### 2022-11-09 manuscript/manuscriptRW.Rmd
manuscript bullet points from results

---
title: "Capturing regeneration in models of forest dynamics"
author:
- Olalla Díaz-Yáñez, Forest Ecology ETH Zurich
- Yannek Käber, Forest Ecology ETH Zurich
- Mats Mahnken, PIK Potsdam
- Samuel Fischer, UFZ
- Rico Fischer, UFZ
- Friedrich Bohn, UFZ
- Xavier Morin, cefe cnrs
- Manfred Lexer, BOKU
- Christian Hochauer, BOKU
- Katarina Merganicova, forim
- Tobias Mette, LWF Bayern
- Werner Rammer, TUM
- Kristin H. Braziunas, TUM
- Paola Mairota, UNIBA
- Josef Bruna, ibot cas
- Heike Lischke, WSL
- Daniel Scherrer, nWSL
- Simon Scheiter, senckenberg
- Thomas Hickler, senckenberg
- Tim Anders, senckenberg
- Jessica Hetzer, senckenberg
- Harald Bugmann, Forest Ecology, ETH Zurich
date: "`r Sys.Date()`"
output:
  word_document:
    reference_docx: template_paper.docx
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

```{r, echo=FALSE}
  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries=FALSE)
  library(bibtex)
  #knitr::opts_chunk$set(echo = TRUE, fig.pos = "h", out.extra = "")

```

# Abstract

*Regeneration is a key process in forest dynamics. Models of forest
dynamics have been long lagging an assessment in their capacities to
properly represent this process. Using 13 models of forest dynamics at
the stand, landscape and global levels, we simulate the regeneration
across 200 sites representing environmental gradients across Europe. The
outputs were compared with regeneration data collected from forest
reserves across Europe. In this study we wanted to asses how well are
models of forest dynamics capturing regeneration levels and what can be
changed to improve this. The results show that most of the models
over-estimate regeneration values. However most of them got right trends
in regeneration across temperature and precipitation gradients. Stand
species richness was better capture than recruitment species richness.
Mortality*

\newpage

# Introduction

Forest recruitment is fundamental for forest dynamics, in particular
with increasing disturbance under climate change and it is also crucial
understand the forest resilience under climate change.

In comparison to recruitment, growth and mortality responses to climate
change are much more studied

Tree recruitment depends on several processes (reproduction, fruit
maturation, seed production, dispersal, germination, juvenile growth and
survival) that are all difficult to assess

Regeneration processes in dynamic forest models

Consequences of over recruit

Species richness in regeneration, mortality and regeneration levels
impacts on forest dynamics

This is the first time, a large number of forest models commonly used to
assess forest dynamics under climate change is evaluated against a
continental-scale multi-species harmonized data set on tree recruitment.

In this study wanted to answer the following questions:

1)  Are models of forests dynamics capturing correctly regeneration
    levels across environmental gradients

2)  Are models of forest dynamics representing correctly initial
    composition of tree populations

3)  Can models of forest dynamics correctly capture the initial
    mortality in tree in-growth between 7 and 10 cm

# Material and methods

A total of 15 models and model variants of forest dynamics at the stand, landscape and global level were used to simulate stand dynamics and sample regeneration in 200 sites selected across Europe.

About the models The models ( Table @ref(tab:tableModels)) were created with different philosophies to represent regeneration. Three models, LPJ-GUESS, the two variants of ForClim and Treemig simulated more species than the one asked because XXXX , these species were assigned to "other" category in the case of TreeMig and in the case of LPJ-GUESS they were removed from the outputs as they are shrubs species (Boreal evergreen shrub, Juniperus oxycedrus and quercus coccifera) LPJ-GUESS also simulate Corylus avellana but only had 3 trees across all samples and they were removed because due to the low degrees of freedom it was not possible to create a model for the category "others" for this model). Forclim simulated a total of 17 species more than the species asked in the protocol . Observed data also have a category others.

About the sites The sites represent the available range of environmental gradients in temperature and precipitation (Figure using the variables degree day and water balance. The sites were selected across a network of plots that collect information on forest reserves [@kaber_sheltered_nodate]

Each modeler group was provided with detailed information on how to perform the simulations including the input variables such as soil or climate, simulation setup and expected output variables. The overarching goal of this simulations was the assessment of the regeneration niche as it arises from empirical data analysis vs. its representation in a wide range of models of forest dynamics.

A GAM model based on XXX was used to compare the trend on the observed data and the simulated outputs. DESCRIBE GAM

Species simulated:  4C only simulated 4 species. Treemig. ForClim and Empirical data have more species assigned all together into category "others". LPJ-GUESS simulated bushes that were not consider in the output of speces composition neither recruitment.

In table @ref(tab:tableModels)

````{r tableModels, echo=FALSE,  message=FALSE}
library(knitr)

modelDF <- readr::read_csv(here::here("data", "modelsDescription.csv"))

modelDF$Type[modelDF$Type ==  "Stand"] <- "S"
modelDF$Type[modelDF$Type ==  "Landscape"] <- "L"
modelDF$Type[modelDF$Type ==  "Global"] <- "G"

modelDF$`Species simulated`[modelDF$`Species simulated` == "F. sylbatica, P. sylvestris, P. abies, A. alba"] <- "F. sylbatica,  P. sylvestris, P. abies,  A. alba"
modelDF$Sampling <- NULL

modelDF$Reference <- "reference"

kable(modelDF,  format = "markdown", caption = "List of models included.")
```

ToDo add a map with the species composition of the empirical data

ToDo check in Haralds paper which models are in which category in the classification of establishment and link this with the results we see here in structure and complexity. For example iLand and Picus have a similar structure but results are very different. Are we meeting the expectations in behaviour fromthe different models?

Species composition in regeneration

We evaluated the species compositions across models and in relation to the observed data by calculating the Shannon index that defines the relative frequency of species referring to basal area (BA) (Equation @ref(eq:shannon))

\begin{equation}
H_n = - \sum_{i = 1}^{s}{p_{i}ln(p_{i})}(\#eq:shannon)
\end{equation}

\begin{equation}
p_{n,i} = \frac{BA_{r_{n,i}}}{BA_{total_{n}}}(\#eq:piShannon)
\end{equation}

where, $s$ is the total number of species in the sample $n$ (11 for the models that simulated all the species and a varied number for those models that did not consider all the species (see Table @ref(tab:tableModels) for more details in the simulated species); $p_{i}$ is the proportion of the species $i$ calculated as the recruited basal area ($BA_{r_{n,i}}$) for that species $i$ over the total basal area ($BA_{total_{n}}$) of that sample $n$.

Main hypothesis:

The recruitment species richness is driven by the simulated adult trees species richness and this is defined on how models represent individual species recruitment per site with respect of adult proportion

Environmental gradients have an impact on the recruitment of the 5 main tree species but extremes are not well captured

Models can capture the relative abundance of recruitment of the 5 main European tree species but they are not doing a good job for minor species

We evaluate the species composition in the recruitment and for all the trees present on the site using the relative frequency of species referring to basal area with the Shannon index calculated per sample, site and model. The higher the value, the higher the diversity of species in a particular site and sample.

Mortality in recruitment

Main hypothesis:

the higher regeneration at threshold of 7cm the stronger is the regeneration mortality, so we will see a larger decease in stem number between recruitment at threshold of 7 cm and 10 cm (as ratio nn7/nn10 (colname nn710))  In othwr words the number of individuals would decrease (Reineke self-thinning rule: This is usually calculated from even age, single species stands. The models could also include the effects of canopy shading and species competition)

Species composition it is also reduced between 7 and 10 cm

Comments on the Reineke rule

the number of individuals would decrease (Reineke self-thinning rule) from 7 to 10cm

assumes a relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands.

This rule is extensively used for management purposes, but it has been initially calibrated for pure, even-aged stands for relatively few species.

This is usually calculated from even age, single species stands. The models could also include the effects of canopy shading and species competition.

Explanations:

$$Reineke exp rk=-1.605: (7/10)^{(-1.605)} = 1.77$$

# Results

## Regeneration levels

-   Recruitment was overestimated across models both for 7 and 10 cm
    diameter thresholds, with the exception of SYBILA and Landis II that
    underestimated recruitment (Figure \@ref(fig:over)). The models with
    higher overestimation are PICUS and TreeMig. The variability across
    sites in recruitment levels is similar or smaller to the observed
    variability for most of the models, with the exception of PICUS and
    TreeMig where the mean recruitment variability is larger.

    ```{r over, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites. The red dotted lines shows the 2.5 and 97.5 percentiles for the 7cm diameter threshold", out.width='100%'}
        knitr::include_graphics(here::here("figures", "recOverAll710.png"))
    ```

-   Models reproduce better the dominating effect of basal area than
    effect of environmental gradients where they have varying effects
    (Figure \@ref(fig:nvTrend710BA), \@ref(fig:nvTrend710BA500) and
    FigureS 1-4). In the observed data the pattern shows that with
    increasing stand basal area the recruitment decreases, this trend
    was captured by Landis II but showing different values of basal area
    for the decreasing. Other models showed different trends such as (1)
    an increase in regeneration with increasing total basal area
    (ForCEEPS, ForCEEPS(f), FORMIND and LPJ-GUESS), (2) a rather
    constant regeneration levels with changing stand basal area (ForClim
    11, xComp, LandClim and aDGVM2) or (3) an increase of recruitment
    until certain value of stand basal area and after that a decrease in
    recruitment, but with different stand basal area thresholds (4C,
    ForClim 1, PICUS, SYBILA and iLand).

-   There are 35 sites that had a diameter threshold of 10cm and
    therefore we do not know what was the recruitment for those sites at
    7cm in the observed data, however the models did simulate
    recruitment for the 7cm threshold on those sites. The Figures
    \@ref(fig:nvTrend710BA), \@ref(fig:nvTrend710BA500) and FigureS 1-4
    show the recruitment from the diameter threshold of 7cm and those
    sites are represented in the empirical data as zeros; this would
    explain why the empirical data has zero recruitment for the lowest
    climatic water balance and highest degree day values while models do
    have recruitment.

-   The mean stand basal area across samples and per bins did not have
    the same categories in across models and empirical data. The bins
    were created using the whole stand basal area range from all the
    outputs (both from empirical and simulated data for all models). For
    those stand basal area values where the empirical data or observed
    models did not have values we assign a zero value so this stand
    basal area categories would be visible in the plots, but this zeros
    are actually not that the recruitment was zero but that this stand
    basal area category was not present for that model. This would
    explain why for the lowest stand basal areas the empirical data has
    no recruitment of for example why ,

```{r nvTrend710BA, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites. The values have been divided in 10 bins and the red lines represent a GAM model showing the trend from the observed data.", out.width='100%'}
    knitr::include_graphics(here::here("figures", "envTrend7Totba.png"))
```

```{r nvTrend710BA500, echo=FALSE, fig.cap="BA to be join with previous figure", out.width='100%'}
    knitr::include_graphics(here::here("figures", "envTrend7Totba_500.png"))
```

## Species composition in regeneration

-   Most models increase or maintain diversity and the same happens with
    the empirical sites (Figure \@ref(fig:H710))

-   In most of the models there are small differences across sites on
    the species richness of recruitment between 7 and 10 cm (Figure
    \@ref(fig:H710). However three models, PICUS, LPJ-GUESS and Landis
    II had higher recruitment species richness than the empirical. The
    model 4C also has higher recruitment than the observed data, however
    this is a special case as this model only simulated 5 species, so
    the total species richness comparison is lower for this model.

    ```{r H710, echo=FALSE, fig.cap="Shannon index of the recruitment calculated by basal area as the mean value across samples per site for diamter treshold of 7 and 10cm. The red dotted lines represent the 20 percentile of the observed data for the treshold of 7cm and 70 percentile of the observed data for the treshold of 10cm ", out.width='100%'}
    knitr::include_graphics(here::here("figures", "H7_10.png"))
    ```

-   Most of the models are capturing well or over-predicting the species
    richness at the stand level (Figure \@ref(fig:combinedRichness) A
    and FigureS 8).

-   The models over predicting species richness at the stand level are
    all the "ForClim family" and FORMIND ] (FigureS 8)

-   Only few models under-predict the species richness at the stand
    level. LPJ-GUESS this is a global model that might not be design to
    capture well this, ForCEEPS (f), xComp, iLand and 4C that did
    simulate less species than the ones available (Figure
    \@ref(fig:combinedRichness), and FigureS 8).

-   Species composition is lower in the recruitment compared to the
    stand level, both for the observed data and simulated data and both
    for recruitment thresholds of 7 and 10 cm. (Figure
    \@ref(fig:combinedRichness)B, and FigureS 9)

-   Most of the models capture correctly the recruited species richness
    at lower values (Figure \@ref(fig:combinedRichness)B, and FigureS 9)

-   Many of the models underestimate species richness in recruitment
    (Figure \@ref(fig:combinedRichness)B, and FigureS 9)

-   Models simulating higher diversity at the stand do not have higher
    diversity in recruitment (FigureS 6 an 7).

-   Both models with feedback and without feedback from the adult trees
    underestimate species richness in recruitment (which ones have
    feedback! enumerate them)) (Appendix figure?)).

    ```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannonn index across all samples per site and for observed and simulated values. A) Shows the three trends present across models for the species richness at the stand level. B) Shows the three trends present across models for recruitment at 7cm treshold. The selection representative model per trend group was based on the Figure XX and tables available in the supplementary information", out.width='75%'}
    knitr::include_graphics(here::here("figures", "combined_richness.png"))
    ```

### **Looking at the recruitment share of the 5 main tree species we found that**

-   Most of the models capture right the share of basal area recruitment
    of the 5 main tree species compared to the observed data in the
    environmental gradient (Figure \@ref(fig:diff)).

-   The models differ in the way they represent the climate niche of
    individual main species (Figure \@ref(fig:diff)).

-   Some models overestimated across the whole gradient one of these 5
    main tree species (for example 4C ForCEEPS, ForCEEPS(f), LandClim,
    FORMIND) while other models overestimate more than one species such
    as Landis II, xComp, PICUS or SYBILA (Figure \@ref(fig:diff)).

-   *Abies alba* was well captured across the gradient for most of the
    models except for xComp, Picus, LanClim and LandisII that
    over-estimated the recruitment share difference. There are also some
    particularities from selected models such as the case of 4C that did
    not simulate *Abies alba* (Figure \@ref(fig:diff)).

-   *Fagus sylvatica* was mainly underestimated cross the environmental
    gradient, with the exception of 4C that overestimate this species
    across the whole environmental gradient (Figure \@ref(fig:diff)).

-   Most of the models over-estimated *Picea abies* recruitment share on
    colder sites (Figure \@ref(fig:diff)).

-   *Pinus sylvestris* and *Quercus* spp. are well capture for those
    models that have recruitment in these species, but the recruitment
    values are always lower so the range of variation is also always
    around 0 (to be confirmed!). However some models lack of recruitment
    for these species (Figure \@ref(fig:diff)).

-   The values that we are seeing are means of sample per site and mean
    across the sites included in each environmental bin. It is important
    to noticed that those groups of sites in the empirical data where
    the recruited basal area of the species was 0 and the stand
    recruited basal area for all species was also 0, the cells color is
    represented as grey (Figure \@ref(fig:diff)). The lack of value in
    the empirical data can be originated due to not recruitment in the
    group of sites or that the sites did not have recruitment values for
    7cm threshold.

    Shall we plot (Figure \@ref(fig:diff)) for 10cm in the appendix and
    remove FiguresS 10-12?

```{r diff, echo=FALSE, fig.cap="Departure of each model in the representation of the main species recruitment relative abundance from the observations across the environmental gradients for a 7cm treshold. The values shown are the mean of samples per site and across sites in bins, where the range of the climatic water balance (cm) and seasonal degree-day sum (dºC) was divided in 10 bins. The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species. The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equalt to zero.", out.width='100%'}
knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles.png"))
```

## Mortality in regeneration

**Main outcomes mortality in number of trees:**

-   There are differences across models on the mortality rate between 7
    and 10 cm (Figure \@ref(fig:mortality11)).

-   Most of the models have a mortality rate larger than those present
    in the empirical data (Figure \@ref(fig:mortality11)).

-   Several models simulate a mortality rate in recruitment that could
    be explained by the self thinning rule (Figure
    \@ref(fig:mortality11)).

-   Models below the Reineke line have lower mortality than the expected
    by self thinning rule, which might point to facilitation or higher
    mortality before the species reached 7 cm diameter (Figure
    \@ref(fig:mortality11)).

-   Values below 1 line show sites where there is a higher number of 10
    cm trees than in 7 cm- this is an artifact of the method - we expect
    space for time substitution, but in reality we just have one
    measurement (Figure \@ref(fig:mortality11)).

-   Models such as SYBILA, xComp, PICUS, LandClim amd TreeMig are above
    the self thinning line (Reineke) indicating that they have higher
    mortality in recruitment of mixed species forest than the expected
    in an even aged single species forests (Figure
    \@ref(fig:mortality11)).

-   Most of the models with lower variation in mortality between 7 and
    10cm also have lower overestimation in recruitment at 7cm (Figure
    \@ref(fig:mortality13) and FigureS 13).

-   Models with higher mortality also show higher tree recruitment at
    7cm (Figure \@ref(fig:mortality13)).

-   There might be a difference between models including stochastic
    processes in recruitment and those which don't, but perhaps the
    reality implies higher mortality rates in the adult component
    (enhance the share of less shade tolerant species?) [to the
    discussion] ((Figure \@ref(fig:mortality11) and
    \@ref(fig:mortality13)).

-   Some of the models that start at a lower number of individuals (7cm)
    have a lower decrease between 7 and 10cm dbh (ratio 7-10cm). The
    species competition and self-thinning is either not that large or it
    happened before the species reached 7cm. This pattern it is not
    consistent across models (FigureS 14-15)

-   There is an effect of the number of individuals at 7cm on the
    mortality rate. Most of the models show a positive effect
    (increasing recruitment at 7cm shows increases the mortality between
    7 and 10cm), but some models show a negative effect. The cloud point
    per models shows the differences across sites, the shape point cloud
    per model shows if the models take into account regeneration density
    when applying mortality to regeneration (FigureS 14-15) This is not
    that clear anymore when we calculate the slope of the linear trends
    per model.

```{r  mortality11, echo=FALSE, fig.cap="Ratio of tree recruitment between 7 and 10 cm recruitment treshold. Dashed blue lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions", out.width='75%'}
knitr::include_graphics(here::here("figures", "ratio7_10.png"))
```

```{r mortality13, echo=FALSE, fig.cap="Ratio of tree recruitment between 7 and 10 cm recruitment treshold and the overestimation proportion (simulated - empirical / empirical) for the mean recruitment trees per model across sites and samples. The horizontal dotted lines show a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions. The vertical line representes the 0 overestimation proportion", out.width='100%'}
knitr::include_graphics(here::here("figures", "mort7_10_medianOver.png"))
```

# Discussion

-   Implication of models that produce very high initial densities

-   Implications of models that have lower mortality

-   The problem of correctly capturing post-disturbance recovery, how
    seed availability and climate conditions define the 15-20 years
    status of a forest. If this is not well capture by the models we can
    not really assess the impacts of disturbances on forests.

-   Range shifts retractions in species expansions because of climate
    change might not also be properly represented if we do not capture
    properly recruitment rates in the environmental gradient .

-   If the composition is not well captured then we can not assess the
    functional composition of the forest appropriately, e.g. their
    sensitive to drought or resilience to new disturbances events.

-   Species richness variations might be explained by the way each model
    considers the recruitment (e.g. in the absence of adults recruitment
    can be linked to either seed rain form elsewhere -dispersal, or to
    sees/seedling bank) Differences might also be expected between
    cohort and individual trees based models.

-   Why models with feedback also underestimate species richness in
    recruitment?

-   Is the adult forest the right forest, how is this relevant and is
    this the consequence for the correct recruitment composition? This
    is highly dependent on the model structure:

    -   Background seed Yes/no

    -   Stochastic vs deterministic

-   We do not know what happen before 7cm in the observed data and does
    not make sense to look at that in the simulated data. 7cm is a large
    sapling, consequences of not having that information to asses forest
    dynamics. Recommendations on data collection.

-   Methodological aspects: difference in observations number between
    observed data (2-4 samples per site) and the simulated data (200
    samples per site) and the implication on variability. Assumption
    equilibrium it is expected overestimation because we do not know the
    status of the observed data more certainly not in equilibrium etc

-   Observed data has much more sites without any recruitment (zero
    inflated data). Many models seem to predict recruitment in all
    sites. The zeros in the emprical data come from no recruitment and
    from tresholds larger than 10cm for some sites. Is this a
    consequence of assuming that the observed data is also in
    equilibrium. Lack of recruitment in sites an indication of
    "non-equilibrium" due to management history?

-   Why the models over predicting species richness at the stand level
    are all the "ForClim family" and FORMIND ?

# References

````{=html}
<!--The relationship was first described by
[@key] for single citation
[@key1; @key2] multiple citation can be separated by semi-colon
[-@key] in order to suppress author name, and just display the year
[see @key1 p 12; also this ref @key2] is alos a valid syntax
```
-->
````

# Supplementary

### 2022-11-09 code/figures.R
Add figures for regeneration levels 7/10cm and VS. Env. Gradients

@@ -1,0 +1,321 @@
+################################################################################
+# Title: Figures
+# Aim: Create figures and analysis for the manuscript
+# Author: Olalla Díaz-Yáñez | @olalla | olalladiaz.net &
+#         some part of figures were inspired in the work of the regeneration
+#         workshop attendees
+################################################################################
+
+# Load plotting styles 
+
+    source("code/plottingManuscript.R")
+
+# Load data ----
+
+    outputsDF <- data.table::fread("data/dataOutputs.csv")
+
+# Assign models and dbh to levels 
+
+    outputsDF$model <- factor(outputsDF$model, levels = modelsOrder)
+    outputsDF$dbh <- factor(outputsDF$dbh, levels = c("10", "7"))
+
+# Calculate Shannon index by basal area ----    
+    
+    ## Recruitment Shannon index----
+    ShannonDF <- outputsDF 
+    ShannonDF$giDIVG <- ShannonDF$r.ba / ShannonDF$Totba
+    ShannonDF$lngiDIVG <- log(ShannonDF$r.ba / ShannonDF$Totba)
+    ShannonDF$mult <- round(ShannonDF$giDIVG * ShannonDF$lngiDIVG, 2)
+    
+    # Sum across species in each observation
+    ShannonIndex <- data.frame(cbind(aggregate(mult ~ site + sample + model + dbh,
+                                               sum, data =  ShannonDF)))
+    
+    ShannonIndex$mult <- (-1) * (ShannonIndex$mult)
+    colnames(ShannonIndex) <- c("site", "sample", "model",
+                                "dbh", "ShannonIndexRecruit")
+    
+    ## Stand Shannon index ----
+    ShannonDFAll <- outputsDF 
+    ShannonDFAll$giDIVG <- ShannonDFAll$ba / ShannonDFAll$Totba
+    ShannonDFAll$lngiDIVG <- log(ShannonDFAll$ba / ShannonDFAll$Totba)
+    ShannonDFAll$mult <- round(ShannonDFAll$giDIVG *  ShannonDFAll$lngiDIVG, 2)
+    
+    # Sum across secies per site and sample
+    ShannonIndexAll <- data.frame(cbind(aggregate(mult ~ site + sample + model + dbh,
+                                                  sum, data = ShannonDFAll)))
+    
+    ShannonIndexAll$mult <- (-1) * (ShannonIndexAll$mult)
+    colnames(ShannonIndexAll) <- c("site", "sample", "model",
+                                   "dbh", "ShannonIndexAllAges")
+    
[…]

### 2022-11-09 code/plottingManuscript.R
addition on plot styles for the manuscript

@@ -7,7 +7,7 @@
 
 # Order of models with names as in the dataset----
 
-    modelsOrder <- c("Empirical" ,"4C", 
+    modelsOrder <- modelsSel <-  c("Empirical" ,"4C", 
                      "ForCEEPS", "ForCEEPS(f)",
                      "FORMIND", "ForClim 1", "ForClim 11", 
                      "SIBYLA", "xComp", "PICUS",

### 2022-11-09 manuscript/manuscriptRW.Rmd
added figures of species richness and corrected some details in the other figures

@@
 output:
-  word_document:
-    reference_docx: template_paper.docx
   bookdown::pdf_document2:
@@
     toc: no
+  word_document:
+    reference_docx: template_paper.docx
 bibliography: Regeneration.bib
@@
 
-Forest recruitment is fundamental for forest dynamics, in particular
-with increasing disturbance under climate change and it is also crucial
+Tree recruitment is fundamental for forest dynamics, in particular with
+increasing disturbance under climate change and it is also crucial
 understand the forest resilience under climate change.
@@
 
-Tree recruitment depends on several processes (reproduction, fruit
+Tree recruitment depends on multiple processes (reproduction, fruit
 maturation, seed production, dispersal, germination, juvenile growth and
@@
 
-Regeneration processes in dynamic forest models
+Models are needed to (1) synthesize existing empirical data and explore
+their consequences, and (2) assess future tree recruitment e.g. in the
+context of global change scenarios
 
-Consequences of over recruit
+Regeneration processes in dynamic forest models and how they have been
+handled to date [@konig_tree_2022; @bugmann2022].
 
-Species richness in regeneration, mortality and regeneration levels
-impacts on forest dynamics
+Tendency of models to be too liberal with recruitment; consequences of
+over-recruitment Species richness in regeneration, mortality and
+regeneration levels may have strong impacts on forest dynamics
 
-This is the first time, a large number of forest models commonly used to
-assess forest dynamics under climate change is evaluated against a
-continental-scale multi-species harmonized data set on tree recruitment.
+An analysis how well models of forest dynamics perform re. recruitment
+is lacking. This is the first time, a large number of forest models
+commonly used to assess forest dynamics under climate change is
+evaluated against a continental-scale multi-species harmonized data set
+on tree recruitment. Briefly praise the dataset [@käber].
 
-In this study wanted to answer the following questions:
-
-1)  Are models of forests dynamics capturing correctly regeneration
-    levels across environmental gradients
-
-2)  Are models of forest dynamics representing correctly initial
-    composition of tree populations
-
-3)  Can models of forest dynamics correctly capture the initial
-    mortality in tree in-growth between 7 and 10 cm
+In this study wanted to answer the following questions: 1) Are models of
+forests dynamics capturing correctly regeneration levels across
+environmental gradients 2) Are models of forest dynamics representing
+correctly initial composition of tree populations 3) Can models of
+forest dynamics correctly capture the initial mortality in tree
+in-growth between 7 and 10 cm
 
@@
 
-A total of 15 models and model variants of forest dynamics at the stand, landscape and global level were used to simulate stand dynamics and sample regeneration in 200 sites selected across Europe.
+A total of 15 models and model variants of forest dynamics at the stand,
+landscape and global level were used to simulate stand dynamics and
+sample regeneration in 200 sites selected across Europe.
 
-About the models The models ( Table @ref(tab:tableModels)) were created with different philosophies to represent regeneration.
-Three models, LPJ-GUESS, the two variants of ForClim and Treemig simulated more species than the one asked because XXXX , these species were assigned to "other" category in the case of TreeMig and in the case of LPJ-GUESS they were removed from the outputs as they are shrubs species (Boreal evergreen shrub, Juniperus oxycedrus and quercus coccifera) LPJ-GUESS also simulate Corylus avellana but only had 3 trees across all samples and they were removed because due to the low degrees of freedom it was not possible to create a model for the category "others" for this model).
-Forclim simulated a total of 17 species more than the species asked in the protocol .
-Observed data also have a category others.
+About the models The models ( Table @ref(tab:tableModels)) were created
+with different philosophies to represent regeneration. Three models,
+LPJ-GUESS, the two variants of ForClim and Treemig simulated more
+species than the one asked because XXXX , these species were assigned to
+"other" category in the case of TreeMig and in the case of LPJ-GUESS
+they were removed from the outputs as they are shrubs species (Boreal
+evergreen shrub, Juniperus oxycedrus and quercus coccifera) LPJ-GUESS
+also simulate Corylus avellana but only had 3 trees across all samples
+and they were removed because due to the low degrees of freedom it was
+not possible to create a model for the category "others" for this
+model). Forclim simulated a total of 17 species more than the species
+asked in the protocol . Observed data also have a category others.
 
-About the sites The sites represent the available range of environmental gradients in temperature and precipitation (Figure using the variables degree day and water balance.
-The sites were selected across a network of plots that collect information on forest reserves [@kaber_sheltered_nodate]
+About the sites The sites represent the available range of environmental
+gradients in temperature and precipitation (Figure using the variables
+degree day and water balance. The sites were selected across a network
+of plots that collect information on forest reserves
+[@kaber_sheltered_nodate]
 
-Each modeler group was provided with detailed information on how to perform the simulations including the input variables such as soil or climate, simulation setup and expected output variables.
-The overarching goal of this simulations was the assessment of the regeneration niche as it arises from empirical data analysis vs.
-its representation in a wide range of models of forest dynamics.
+Each modeler group was provided with detailed information on how to
+perform the simulations including the input variables such as soil or
+climate, simulation setup and expected output variables. The overarching
+goal of this simulations was the assessment of the regeneration niche as
+it arises from empirical data analysis vs. its representation in a wide
+range of models of forest dynamics.
 
-A GAM model based on XXX was used to compare the trend on the observed data and the simulated outputs. DESCRIBE GAM
+A GAM model based on XXX was used to compare the trend on the observed
+data and the simulated outputs. DESCRIBE GAM
 
-Species simulated:  4C only simulated 4 species.
-Treemig.
-ForClim and Empirical data have more species assigned all together into category "others".
-LPJ-GUESS simulated bushes that were not consider in the output of speces composition neither recruitment.
+Species simulated: 4C only simulated 4 species. Treemig. ForClim and
+Empirical data have more species assigned all together into category
+"others". LPJ-GUESS simulated bushes that were not consider in the
+output of speces composition neither recruitment.
 
@@
 
-````{r tableModels, echo=FALSE,  message=FALSE}
+```{r tableModels, echo=FALSE,  message=FALSE}
 library(knitr)
@@
 
-ToDo check in Haralds paper which models are in which category in the classification of establishment and link this with the results we see here in structure and complexity.
-For example iLand and Picus have a similar structure but results are very different.
-Are we meeting the expectations in behaviour fromthe different models?
+ToDo check in Haralds paper which models are in which category in the
+classification of establishment and link this with the results we see
+here in structure and complexity. For example iLand and Picus have a
+similar structure but results are very different. Are we meeting the
+expectations in behaviour fromthe different models?
 
@@
 
-We evaluated the species compositions across models and in relation to the observed data by calculating the Shannon index that defines the relative frequency of species referring to basal area (BA) (Equation @ref(eq:shannon))
+We evaluated the species compositions across models and in relation to
+the observed data by calculating the Shannon index that defines the
+relative frequency of species referring to basal area (BA) (Equation
+@ref(eq:shannon))
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
+```
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
-where, $s$ is the total number of species in the sample $n$ (11 for the models that simulated all the species and a varied number for those models that did not consider all the species (see Table @ref(tab:tableModels) for more details in the simulated species); $p_{i}$ is the proportion of the species $i$ calculated as the recruited basal area ($BA_{r_{n,i}}$) for that species $i$ over the total basal area ($BA_{total_{n}}$) of that sample $n$.
+```
+where, $s$ is the total number of species in the sample $n$ (11 for the
+models that simulated all the species and a varied number for those
+models that did not consider all the species (see Table
+@ref(tab:tableModels) for more details in the simulated species);
+$p_{i}$ is the proportion of the species $i$ calculated as the recruited
+basal area ($BA_{r_{n,i}}$) for that species $i$ over the total basal
+area ($BA_{total_{n}}$) of that sample $n$.
 
@@
 
-The recruitment species richness is driven by the simulated adult trees species richness and this is defined on how models represent individual species recruitment per site with respect of adult proportion
+The recruitment species richness is driven by the simulated adult trees
+species richness and this is defined on how models represent individual
+species recruitment per site with respect of adult proportion
 
-Environmental gradients have an impact on the recruitment of the 5 main tree species but extremes are not well captured
+Environmental gradients have an impact on the recruitment of the 5 main
+tree species but extremes are not well captured
 
-Models can capture the relative abundance of recruitment of the 5 main European tree species but they are not doing a good job for minor species
+Models can capture the relative abundance of recruitment of the 5 main
+European tree species but they are not doing a good job for minor
+species
 
-We evaluate the species composition in the recruitment and for all the trees present on the site using the relative frequency of species referring to basal area with the Shannon index calculated per sample, site and model.
-The higher the value, the higher the diversity of species in a particular site and sample.
+We evaluate the species composition in the recruitment and for all the
+trees present on the site using the relative frequency of species
+referring to basal area with the Shannon index calculated per sample,
+site and model. The higher the value, the higher the diversity of
+species in a particular site and sample.
 
@@
 
-the higher regeneration at threshold of 7cm the stronger is the regeneration mortality, so we will see a larger decease in stem number between recruitment at threshold of 7 cm and 10 cm (as ratio nn7/nn10 (colname nn710))  In othwr words the number of individuals would decrease (Reineke self-thinning rule: This is usually calculated from even age, single species stands.
-The models could also include the effects of canopy shading and species competition)
+the higher regeneration at threshold of 7cm the stronger is the
+regeneration mortality, so we will see a larger decease in stem number
+between recruitment at threshold of 7 cm and 10 cm (as ratio nn7/nn10
+(colname nn710))  In othwr words the number of individuals would
+decrease (Reineke self-thinning rule: This is usually calculated from
+even age, single species stands. The models could also include the
+effects of canopy shading and species competition)
 
@@
 
-the number of individuals would decrease (Reineke self-thinning rule) from 7 to 10cm
+the number of individuals would decrease (Reineke self-thinning rule)
+from 7 to 10cm
 
-assumes a relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands.
+assumes a relationship between the number of stems and the quadratic
+mean diameter in fully stocked pure stands.
 
-This rule is extensively used for management purposes, but it has been initially calibrated for pure, even-aged stands for relatively few species.
+This rule is extensively used for management purposes, but it has been
+initially calibrated for pure, even-aged stands for relatively few
+species.
 
-This is usually calculated from even age, single species stands. The models could also include the effects of canopy shading and species competition.
+This is usually calculated from even age, single species stands. The
+models could also include the effects of canopy shading and species
+competition.
 
@@
 
-## Regeneration levels
+Regeneration levels, species composition and mortality between
+recruitment diameters thresholds of 7 and 10 cm varied across model
+simulation results and in comparison with the observations. As expected,
+recruitment was overestimated across most models for both the 7 and 10
+cm diameter thresholds, with the exception of SYBILA, Landis II and
+aDGVM2 that estimated recruitment rates at the lower end of the
+plausibility interval of the empirical data (Figure \@ref(fig:over)).
+The models with highest overestimation were PICUS and TreeMig. For most
+of the models, the variability across sites in recruitment levels
+(visible from the IQR
+in the box plots of Figure \@ref(fig:over)) is similar to or smaller
+than the observed
+variability for most of the models, with the exception of PICUS, where
+the mean
+recruitment variability is much larger.
 
--   Recruitment was overestimated across models both for 7 and 10 cm
-    diameter thresholds, with the exception of SYBILA and Landis II that
-    underestimated recruitment (Figure \@ref(fig:over)). The models with
-    higher overestimation are PICUS and TreeMig. The variability across
-    sites in recruitment levels is similar or smaller to the observed
-    variability for most of the models, with the exception of PICUS and
-    TreeMig where the mean recruitment variability is larger.
-
-    ```{r over, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites.
-The red dotted lines shows the 2.5 and 97.5 percentiles for the 7cm diameter threshold", out.width='100%'}
+```{r over, echo=FALSE, fig.cap="Mean recruitment values across all samples for the 200 sites.
+The red dotted lines shows the 2.5 and 97.5 percentiles for the 7 cm diameter threshold.", out.width='100%'}
         knitr::include_graphics(here::here("figures", "recOverAll710.png"))
@@
 ```{r nvTrend710BA, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites.
-The values have been divided in 10 bins and the red lines represent a GAM model showing the trend from the observed data.", out.width='100%'}
-    knitr::include_graphics(here::here("figures", "envTrend7Totba.png"))
-```
-
-```{r nvTrend710BA500, echo=FALSE, fig.cap="BA to be join with previous figure", out.width='100%'}
-    knitr::include_graphics(here::here("figures", "envTrend7Totba_500.png"))
+The values have been divided in 10 bins and the red lines represent a GAM model showing the trend from the observed data.", fig.show = 'hold', out.width = '50%'}
+    knitr::include_graphics(c(here::here("figures", "envTrend7Totba.png"),
+                              here::here("figures", "envTrend7Totba_500.png")))
 ```
@@
 
-````{=html}
+```{=html}
 <!--The relationship was first described by
@@
 [see @key1 p 12; also this ref @key2] is alos a valid syntax
+-->
 ```
--->
-````
+\`\`\`
 

### 2022-11-09 code/figures.R

@@ -18,10 +18,16 @@
 
     outputsDF$model <- factor(outputsDF$model, levels = modelsOrder)
     outputsDF$dbh <- factor(outputsDF$dbh, levels = c("10", "7"))
-
-# Calculate Shannon index by basal area ----    
     
-    ## Recruitment Shannon index----
+# Remove the 35 sites from the models data for threshold 7cm as they do not have 
+# observations for those sites
+
+    sitesWoObs <- outputsDF[is.na(outputsDF$r.trees), ]
+    outputsDF <- outputsDF[!(outputsDF$dbh == 7 & outputsDF$site %in% unique(sitesWoObs$site)), ]
+
+# Data preparation ----    
+    
+    ## Recruitment Shannon index  by basal area ----
     ShannonDF <- outputsDF 
     ShannonDF$giDIVG <- ShannonDF$r.ba / ShannonDF$Totba
     ShannonDF$lngiDIVG <- log(ShannonDF$r.ba / ShannonDF$Totba)
@@ -140,11 +146,12 @@
                          wb  = unique(wb))
     
     envTrend <-  envTrend[ envTrend$dbh == "7", ]
-    envTrend <- as.data.table(envTrend) 
+    #envTrend <- as.data.table(envTrend) 
 
     
     # Mean across sites 
     empDataQuartile <- c(0.025, 0.975)
+    
     overDTEmp <- overDTEmp2 |>
         dplyr::group_by(dbh) |>
         dplyr::summarise(r.trees_low = quantile(r.trees, empDataQuartile[1],
@@ -202,11 +209,16 @@
     
     # Plot the boxplot with the observations GAM
     for (modelSel in unique(envTrend$model)) {
-        ienvTrendModel <- envTrend[envTrend$model == modelSel]
+        ienvTrendModel <- envTrend[envTrend$model == modelSel,]
+        
+        max_r <- as.numeric(quantile(ienvTrendModel$r.trees,.99, na.rm = TRUE))
+        ienvTrendModel$r.trees[ienvTrendModel$r.trees > max_r] <- max_r
+        
+        
         axisPlot <- ifelse(modelSel %in% c("Landis II",  "TreeMig", "LPJ-GUESS",
                                             "aDGVM2"), "s", "n")
         boxplot(r.trees ~ ba_cut, 
-                data = ienvTrendModel[dbh == 7 & model == modelSel], 
[…]

### 2022-11-09 manuscript/manuscriptRW.Rmd
add Harald inputs on the result section till species diversity per main species

@@
 
+There are 35 sites that had a diameter threshold of 10 cm and therefore
+we do not know the actual recruitment rates for those sites at 7 cm; yet
+the models simulated recruitment for the 7 cm threshold on those sites.
+The 35 sites were removed from the simulated data as well to make a fair
+comparison.
+
+How many species were simulated per model? why?
+
+Need to say something about zero inflation in the observations vs. the
+simulations (current Figure S5)
+
 A GAM model based on XXX was used to compare the trend on the observed
@@
 Regeneration levels, species composition and mortality between
-recruitment diameters thresholds of 7 and 10 cm varied across model
+recruitment diameters thresholds of 7 and 10cm varied across model
 simulation results and in comparison with the observations. As expected,
@@
 of the models, the variability across sites in recruitment levels
-(visible from the IQR
-in the box plots of Figure \@ref(fig:over)) is similar to or smaller
-than the observed
-variability for most of the models, with the exception of PICUS, where
-the mean
-recruitment variability is much larger.
+(visible from the IQR in the box plots of Figure \@ref(fig:over)) is
+similar to or smaller than the observed variability for most of the
+models, with the exception of PICUS, where the mean recruitment
+variability is much larger.
 
@@
 
--   Models reproduce better the dominating effect of basal area than
-    effect of environmental gradients where they have varying effects
-    (Figure \@ref(fig:nvTrend710BA), \@ref(fig:nvTrend710BA500) and
-    FigureS 1-4). In the observed data the pattern shows that with
-    increasing stand basal area the recruitment decreases, this trend
-    was captured by Landis II but showing different values of basal area
-    for the decreasing. Other models showed different trends such as (1)
-    an increase in regeneration with increasing total basal area
-    (ForCEEPS, ForCEEPS(f), FORMIND and LPJ-GUESS), (2) a rather
-    constant regeneration levels with changing stand basal area (ForClim
-    11, xComp, LandClim and aDGVM2) or (3) an increase of recruitment
-    until certain value of stand basal area and after that a decrease in
-    recruitment, but with different stand basal area thresholds (4C,
-    ForClim 1, PICUS, SYBILA and iLand).
+When looking at the representation of recruitment rates along key
+gradients of light availability (basal area), temperature (degree-days)
+and soil moisture (water balance), distinct features emerge: the models
+reproduce the effect of basal area much better than the effects of
+climatic gradients, where they feature varying effects (Figure
+\@ref(fig:nvTrend710BA), \@ref(fig:nvTrend710BA500) and FigureS 1-4).
 
--   There are 35 sites that had a diameter threshold of 10cm and
-    therefore we do not know what was the recruitment for those sites at
-    7cm in the observed data, however the models did simulate
-    recruitment for the 7cm threshold on those sites. The Figures
-    \@ref(fig:nvTrend710BA), \@ref(fig:nvTrend710BA500) and FigureS 1-4
-    show the recruitment from the diameter threshold of 7cm and those
-    sites are represented in the empirical data as zeros; this would
-    explain why the empirical data has zero recruitment for the lowest
-    climatic water balance and highest degree day values while models do
-    have recruitment.
+In the observed data the pattern shows that with increasing stand basal
+area recruitment decreases. However, this trend was captured by Landis
+II only, albeit at much lower values of basal area than in the
+observations. Other models featured distinctly different trends, such as
+(1) an increase in regeneration with increasing total basal area
+followed of no recruitment at high basal area (ForCEEPS, ForCEEPS(f),
+FORMIND, TreeMig and LPJ-GUESS), (2) almost constant regeneration levels
+with basal area (ForClim 11, xComp, LandClim and aDGVM2) or (3) an
+increase of recruitment up to a certain value of stand basal area
+followed by a decrease of recruitment at even higher stand basal area
+values at model-specific thresholds (4C, ForClim 1, PICUS, SYBILA and
+iLand).
 
--   The mean stand basal area across samples and per bins did not have
-    the same categories in across models and empirical data. The bins
-    were created using the whole stand basal area range from all the
-    outputs (both from empirical and simulated data for all models). For
-    those stand basal area values where the empirical data or observed
-    models did not have values we assign a zero value so this stand
-    basal area categories would be visible in the plots, but this zeros
-    are actually not that the recruitment was zero but that this stand
-    basal area category was not present for that model. This would
-    explain why for the lowest stand basal areas the empirical data has
-    no recruitment of for example why ,
+Models did not match the observed patterns in recruitment across soil
+moisture and temperature gradients and there was either a clear common
+pattern across models. Looking at the soil moisture gradients (FigureS
+2), we can see that a group of models decreases recruitment with
+increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS
+and LandClim), while few models have the opposite trend (ForCEEPs(f),
+FORMIND and LPJ-GUESS), and the rest of do not have a clear pattern
+(ForCEEPS, iLand, Landis II, TreeMig and aDGVM2). Across temperature
+gradients all the models that had a decrease in recruitment with
+increasing soil moisture showed the opposite trend (more recruitment
+with increasing temperature) except for 4C (FigureS 3). The rest of the
+models did not have a clear patterns across the temperature gradient.
 
-```{r nvTrend710BA, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites.
-The values have been divided in 10 bins and the red lines represent a GAM model showing the trend from the observed data.", fig.show = 'hold', out.width = '50%'}
+```{r nvTrend710BA, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the stand basal area.
+(A) y-axis scaled by model; (B) same y-axis across all models.
+The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
     knitr::include_graphics(c(here::here("figures", "envTrend7Totba.png"),
@@
 
-## Species composition in regeneration
+Most models match the level of diversity (as measured by the Shannon
+index) of the observed data quite well (Figure \@ref(fig:H710)),
+although they consistently underestimate its variability across sites.
+Four models overestimate recruitment diversity (4C, PICUS, LANDIS-II and
+LPJ-GUESS, the latter mostly for the 10 cm diameter threshold. The model
+4C is a special case as this model simulated 5 species only. In most
+models there are small differences across sites regarding the species
+richness of recruitment between 7 and 10 cm. Most models maintain or
+even increase in recruitment diversity between the 7 and 10 cm
+thresholds, and the same is visible from the empirical data, however the
+differences are not significant EXPLAIN WHY  "Can we understand this (I
+know this is for the Discussion, but still...)?? This cannot be species
+richness (as there are no new species popping between 7 and 10 cm DBH).
+I'd also expect the number of species to be lower at 10 cm than at 7 cm
+(because some will die off due to insufficient light etc.), but this is
+not reflected in H' because species with zero abundance are ignored
+(ln[0] is not defined). So a strongly increasing evenness must be
+driving this increase of H'."
 
--   Most models increase or maintain diversity and the same happens with
-    the empirical sites (Figure \@ref(fig:H710))
-
--   In most of the models there are small differences across sites on
-    the species richness of recruitment between 7 and 10 cm (Figure
-    \@ref(fig:H710). However three models, PICUS, LPJ-GUESS and Landis
-    II had higher recruitment species richness than the empirical. The
-    model 4C also has higher recruitment than the observed data, however
-    this is a special case as this model only simulated 5 species, so
-    the total species richness comparison is lower for this model.
-
-```{r H710, echo=FALSE, fig.cap="Shannon index of the recruitment calculated by basal area as the mean value across samples per site for diamter treshold of 7 and 10cm.
-The red dotted lines represent the 20 percentile of the observed data for the treshold of 7cm and 70 percentile of the observed data for the treshold of 10cm ", out.width='100%'}
+```{r H710, echo=FALSE, fig.cap="Shannon index of the recruitment, calculated by basal area as the mean value across samples per site for the diameter thresholds of 7 and 10 cm, respectively.
+The red dotted lines represent the 20th percentile of the observed data for the threshold of 7 cm and the 70th percentile of the observed data for the threshold of 10 cm.", out.width='100%'}
     knitr::include_graphics(here::here("figures", "H7_10.png"))
@@
 
--   Most of the models are capturing well or over-predicting the species
-    richness at the stand level (Figure \@ref(fig:combinedRichness) A
-    and FigureS 8).
+Most of the models are capturing well or over-predicting the species
+richness at the stand level (Figure \@ref(fig:combinedRichness) A2 and
+FigureS 6). The models over predicting species richness at the stand
+level are ForClim1, ForClim 11, FORMIND and TreeMig (Figure
+\@ref(fig:combinedRichness) A1). However only few models under-predict
+the species richness at the stand level (Figure
+\@ref(fig:combinedRichness) A3) LPJ-GUESS, ForCEEPS (f), xComp, iLand
+and 4C. Species diversity is lower in the recruitment compared to the
+stand level, for both observed and simulated data, and for both
+recruitment thresholds, i.e. 7 and 10 cm (Figure
+\@ref(fig:combinedRichness)B, and FigureS 7). The majority of the models
+capture reasonably well the recruitment diversity at lower values of
+diversity, however most of them strongly underestimate species diversity
+in recruitment for higher observed diversity levels. Models simulating
+higher diversity at the stand level do not feature higher diversity in
+recruitment
+(FigureS 6 and 7). Both models with feedback and without feedback from
+the adult trees (via seed production) underestimate species richness in
+recruitment.
 
--   The models over predicting species richness at the stand level are
-    all the "ForClim family" and FORMIND ] (FigureS 8)
-
--   Only few models under-predict the species richness at the stand
-    level. LPJ-GUESS this is a global model that might not be design to
-    capture well this, ForCEEPS (f), xComp, iLand and 4C that did
-    simulate less species than the ones available (Figure
-    \@ref(fig:combinedRichness), and FigureS 8).
-
--   Species composition is lower in the recruitment compared to the
-    stand level, both for the observed data and simulated data and both
-    for recruitment thresholds of 7 and 10 cm. (Figure
-    \@ref(fig:combinedRichness)B, and FigureS 9)
-
--   Most of the models capture correctly the recruited species richness
-    at lower values (Figure \@ref(fig:combinedRichness)B, and FigureS 9)
-
--   Many of the models underestimate species richness in recruitment
-    (Figure \@ref(fig:combinedRichness)B, and FigureS 9)
-
--   Models simulating higher diversity at the stand do not have higher
-    diversity in recruitment (FigureS 6 an 7).
-
--   Both models with feedback and without feedback from the adult trees
-    underestimate species richness in recruitment (which ones have
-    feedback! enumerate them)) (Appendix figure?)).
-
-```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannonn index across all samples per site and for observed and simulated values.
+```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site and for observed and simulated values.
 A) Shows the three trends present across models for the species richness at the stand level.
-B) Shows the three trends present across models for recruitment at 7cm treshold.
-The selection representative model per trend group was based on the Figure XX and tables available in the supplementary information", out.width='75%'}
+B) Shows the three trends present across models for recruitment at 7cm threshold.
+The selected representative model per trend group was based on Figure XX and tables available in the supplementary information", out.width='75%'}
     knitr::include_graphics(here::here("figures", "combined_richness.png"))
@@
 
+-   Also BTW, the reason why some models have zero regeneration at very
+    low BA values may be that they never actually simulate such low BA
+    values at the one ha level (which was the sampling unit if I recall
+    correctly). If so, this should be pointed out as well, I think.
+
 # References

### 2022-11-09 code/dataPreparation.R
correct landis input data (changes before email to Josef)

@@ -176,6 +176,7 @@
     landis$model <- "Landis II"
     landis$dbh[landis$dbh == "7 cm" ] <- 7
     landis$dbh[landis$dbh == "10 cm" ] <- 10
+    landis$dbh[landis$dbh == "0" ] <- 0
     landis$dbh <- as.numeric(landis$dbh)
     landis$V1 <- NULL
 
@@ -210,7 +211,7 @@
      aa2 <- rbind(aa7, aa10)
 
      landis <- landis[!landis$dbh == 0,] 
-     sppSimLandis <- rbind(landis, aa2)
+     landis <- rbind(landis, aa2)
      
      
      # Add 0 values to those sites where there is 0 recruitment and 0 BA for a spp
@@ -227,7 +228,18 @@
                           r.ba = 0,
                           r.trees = 0,
                           ba = 0)
-     landis <- rbind(landis, missing)                                                    
+     landis <- rbind(landis, missing) 
+     
+     # Check that same sample do not have ba value for one species and 0 for the same species 
+     # for the two diameter thresholds
+     # e.g. View(landis[landis$site == 3 & landis$sample == 1,])
+     
+     landis <- landis |> 
+         dplyr::group_by(site, sample, model, species) |>
+         dplyr::mutate(type = if(dplyr::n_distinct(ba) > 1) 'multiple values' else 'OK') |>
+         dplyr::mutate(ba = max(ba)) |>
+         dplyr::ungroup()
+     landis$type <- NULL
 
     ### Treemig ----
 

### 2022-11-16 code/figures.R
updates on figures of main trees species and environmental gradient

@@ -63,6 +63,41 @@
     
     ShannonIndex$relShannon <- ShannonIndex$ShannonIndexRecruit / ShannonIndex$ShannonIndexAllAges
     
+# Zero inflated data in recruitment 
+      
+    zeroInflSim <- outputsDF[!outputsDF$model == "Empirical",]
+    zeroInflEmp <- outputsDF[outputsDF$model == "Empirical",]
+    
+    zeroInflSim <-  zeroInflSim[, c("site", "sample", "model", "Totr.ba")]
+    zeroInflSim <- zeroInflSim[!duplicated(zeroInflSim)]
+    zeroInflEmp <-  zeroInflEmp[, c("site", "sample", "Totr.ba")]
+    zeroInflEmp <- zeroInflEmp[!duplicated(zeroInflEmp)]
+    
+    
+    #Number of sites and samples with recruitment equal to 0 in observations
+    zeroInflatedAll <- outputsDF[outputsDF$Totr.ba == 0, ]
+    zeroInflatedAll <- zeroInflatedAll[, c("site", "sample",  "model", "Totr.ba", "dbh")]
+    zeroInflatedAll <-  zeroInflatedAll[!duplicated(zeroInflatedAll)]
+    
+    # Zero data
+    zerroAll <- as.array(table(zeroInflatedAll$model, zeroInflatedAll$dbh))
+    colnames(zerroAll) <- c("zero10", "zero7")
+    #Total observations
+    allObs <- as.array(table(outputsDF$model,outputsDF$dbh ))
+    colnames(allObs) <- c("allObs10", "allObs7")
+    
+    inflatedZeroDF <- as.data.frame(cbind(allObs, zerroAll))
+    
+    inflatedZeroDF$share10 <- round((inflatedZeroDF$zero10  / inflatedZeroDF$allObs10), 4) * 100
+    inflatedZeroDF$share7 <- round((inflatedZeroDF$zero7  / inflatedZeroDF$allObs7) , 4) * 100
+    
+    colnames(inflatedZeroDF) <- c("Total (10 cm)", "Total (7cm)",
+                                  "No recruitment (10 cm)", "No recruitment (7cm)", 
+                                  "Percentage of no recruitment (10 cm)", "Percentage of no recruitment (7 cm)")
+    inflatedZeroDF <- tibble::rownames_to_column(inflatedZeroDF, "Model")
+    
[…]

### 2022-11-16 code/dataPreparation.R
data preparation, forclim joint other category, remove extra samples sybila and picus, more corrections for Landis

@@ -54,6 +54,15 @@
 
     # Assign species simulated outside the ones asked in the protocol to category others
     forclim1$species[!forclim1$species %in% simSppAsked] <- "othr"
+    
+    # Joint values from other category 
+    forclim12 <- forclim1 |>
+      dplyr::group_by(sample, site, species, dbh, model) |>  # for each row
+      dplyr::summarise(r.trees = sum(r.trees),
+                       r.ba = sum(r.ba),
+                       ba = sum(ba)) # calculate the sum of those columns
+    
+    forclim1 <- forclim12
 
     ### ForClim 11 ----
     forclim11 <- data.table::fread("data/simulation_results/forclim/forclim_variant11.csv")
@@ -67,6 +76,14 @@
 
     # Assign species simulated outside the ones asked in the protocol to category others
     forclim11$species[!forclim11$species %in% simSppAsked] <- "othr"
+    # Joint values from other category 
+    forclim112 <- forclim11 |>
+      dplyr::group_by(sample, site, species, dbh, model) |>  # for each row
+      dplyr::summarise(r.trees = sum(r.trees),
+                       r.ba = sum(r.ba),
+                       ba = sum(ba)) # calculate the sum of those columns
+    
+    forclim11 <- forclim112
 
     ### ForCEEPS ----
     forceeps_7cm <- data.table::fread("data/simulation_results/forceeps/Results_Forceeps_7cm.csv")
@@ -87,6 +104,11 @@
     ### SIBYLA ----
     sibyla <- data.table::fread("data/simulation_results/sibyla/TabTreesSibyla0_KJMerganic.csv")
     sibyla$model <- "SIBYLA"
+    
+    # Some sites have more than 200 samples, 
+    # remove the extra ones to have 200 samples per site
+    sibyla <- sibyla[!sibyla$sample > 200,]
+    
 
     ### xComp----
     load("data/simulation_results/xcomp/xcomp2022_simulation_v01_summary07cm.RData")
@@ -119,11 +141,14 @@
     formind$dbh <- formind$dbh * 100
 
 
-    ### Picus ----
+    ### PICUS ----
     picus <- data.table::fread("data/simulation_results/picus/PICUS_OutputData_2022-05-27.csv")
     picus$model <- "PICUS"
+    
+    # It has 201 sample per site, remove one sample
 
-
+    picus <- picus[!picus$sample > 200,]
+    
[…]

### 2022-11-16 manuscript/manuscriptRW.Rmd
manuscript: added most of the coments from Harald on the results bullet points&further changes in the results section

@@
 
-Regeneration levels, species composition and mortality between
-recruitment diameters thresholds of 7 and 10cm varied across model
-simulation results and in comparison with the observations. As expected,
-recruitment was overestimated across most models for both the 7 and 10
-cm diameter thresholds, with the exception of SYBILA, Landis II and
-aDGVM2 that estimated recruitment rates at the lower end of the
-plausibility interval of the empirical data (Figure \@ref(fig:over)).
-The models with highest overestimation were PICUS and TreeMig. For most
-of the models, the variability across sites in recruitment levels
-(visible from the IQR in the box plots of Figure \@ref(fig:over)) is
-similar to or smaller than the observed variability for most of the
-models, with the exception of PICUS, where the mean recruitment
-variability is much larger.
+## Recruitment levels
+
+Results related to regeneration levels, species composition and
+mortality in recruitment varied across models and in comparison with the
+observations. As expected, recruitment was overestimated across most
+models for both the 7 and 10 cm diameter thresholds, with the exception
+of SYBILA, Landis II and aDGVM2 that estimated recruitment rates at the
+lower end of the plausibility interval of the empirical data (Figure
+\@ref(fig:over)). The models with highest overestimation were PICUS and
+TreeMig. For most of the models, the variability across sites in
+recruitment levels (visible from the interquartile range in the box
+plots of Figure \@ref(fig:over)) is similar to or smaller than the
+observed variability for most of the models, with the exception of
+PICUS, where the mean recruitment variability is much larger.
 
@@
 increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS
-and LandClim), while few models have the opposite trend (ForCEEPs(f),
-FORMIND and LPJ-GUESS), and the rest of do not have a clear pattern
-(ForCEEPS, iLand, Landis II, TreeMig and aDGVM2). Across temperature
-gradients all the models that had a decrease in recruitment with
-increasing soil moisture showed the opposite trend (more recruitment
-with increasing temperature) except for 4C (FigureS 3). The rest of the
-models did not have a clear patterns across the temperature gradient.
+and LandClim), while few models had the opposite trend (ForCEEPs(f),
+FORMIND and LPJ-GUESS). Across temperature gradients all the models that
+had a decrease in recruitment with increasing soil moisture showed the
+opposite trend (more recruitment with increasing temperature) except for
+4C (FigureS 3). The rest of the models did not have a clear pattern
+across the temperature gradient.
 
@@
 
-Most models match the level of diversity (as measured by the Shannon
-index) of the observed data quite well (Figure \@ref(fig:H710)),
+## Species composition in recruitment
+
+Most of the models match the level of diversity (as measured by the
+Shannon index) of the observed data quite well (Figure \@ref(fig:H710)),
 although they consistently underestimate its variability across sites.
@@
 thresholds, and the same is visible from the empirical data, however the
-differences are not significant EXPLAIN WHY  "Can we understand this (I
-know this is for the Discussion, but still...)?? This cannot be species
-richness (as there are no new species popping between 7 and 10 cm DBH).
-I'd also expect the number of species to be lower at 10 cm than at 7 cm
-(because some will die off due to insufficient light etc.), but this is
-not reflected in H' because species with zero abundance are ignored
-(ln[0] is not defined). So a strongly increasing evenness must be
-driving this increase of H'."
+differences are not significant. **EXPLAIN WHY  "Can we understand this
+(I know this is for the Discussion, but still...)?? This cannot be
+species richness (as there are no new species popping between 7 and 10
+cm DBH). I'd also expect the number of species to be lower at 10 cm than
+at 7 cm (because some will die off due to insufficient light etc.), but
+this is not reflected in H' because species with zero abundance are
+ignored (ln[0] is not defined). So a strongly increasing evenness must
+be driving this increase of H'."**
 
@@
 higher diversity at the stand level do not feature higher diversity in
-recruitment
-(FigureS 6 and 7). Both models with feedback and without feedback from
-the adult trees (via seed production) underestimate species richness in
-recruitment.
+recruitment (FigureS 6 and 7). Both models with feedback and without
+feedback from the adult trees (via seed production) underestimate
+species richness in recruitment.
 
@@
 
-### **Looking at the recruitment share of the 5 main tree species we found that**
+Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
+sylvestris*) and oaks (*Quercus* spp) were the most frequent species in
+the observed data. Most models capture the share of basal area of
+recruitment of the five main tree species compared to the observed data
+along the two major gradients of climate (Figure \@ref(fig:diff) and
+FigureS XX). However, the models differ in the way they represent the
+climate niche of individual main species. First, some models
+overestimated recruitment across the whole gradient for one of the 5
+main tree species (for example, ForCEEPS, ForCEEPS(f),LandClim or
+LPJ-GUESS), while other models overestimated recruitment for more than
+one species for larger parts of the climate gradient, such as xComp,
+PICUS , SYBILA or LPJ-GUESS (cf. red colors in Figure \@ref(fig:diff)).
+This trends were consistent with those when the regeneration thresholds
+was at 7 and 10cm although the exact changes in recruited basal area
+share changed (FigureS XX).
 
--   Most of the models capture right the share of basal area recruitment
-    of the 5 main tree species compared to the observed data in the
-    environmental gradient (Figure \@ref(fig:diff)).
+When looking individually to each of the five main tree species it is
+possible to observe differences in the simulated recruitment results.
+Silver fir (*Abies alba*) was well captured across the gradient by most
+models except for xComp, Picus, LandClim and Landis-II, which
+overestimated its recruitment share. There are also some particularities
+evident for some models, such as 4C that did not simulate *Abies alba*.
+*Fagus sylvatica* was underestimated by many models across the
+environmental gradient, including ForCEEPS(f) and FORMIND that did not
+have regeneration in most of the sites for this species. Most of the
+models over-estimated *Picea abies* recruitment share on colder sites.
+*Pinus sylvestris* and *Quercus* spp. represented a smaller share of the
+recruited species in the observed data and localized on certain areas of
+the environmental gradient, however many models did have recruitment for
+these two species in most of the climatic gradient although at basal
+area share. Five models (4C, FORMIND, SYBILA, xComp, LandClim) had
+almost no generation or did not regenerate *Pinus sylvestris* and
+*Quercus* spp.
 
--   The models differ in the way they represent the climate niche of
-    individual main species (Figure \@ref(fig:diff)).
-
--   Some models overestimated across the whole gradient one of these 5
-    main tree species (for example 4C ForCEEPS, ForCEEPS(f), LandClim,
-    FORMIND) while other models overestimate more than one species such
-    as Landis II, xComp, PICUS or SYBILA (Figure \@ref(fig:diff)).
-
--   *Abies alba* was well captured across the gradient for most of the
-    models except for xComp, Picus, LanClim and LandisII that
-    over-estimated the recruitment share difference. There are also some
-    particularities from selected models such as the case of 4C that did
-    not simulate *Abies alba* (Figure \@ref(fig:diff)).
-
--   *Fagus sylvatica* was mainly underestimated cross the environmental
-    gradient, with the exception of 4C that overestimate this species
-    across the whole environmental gradient (Figure \@ref(fig:diff)).
-
--   Most of the models over-estimated *Picea abies* recruitment share on
-    colder sites (Figure \@ref(fig:diff)).
-
--   *Pinus sylvestris* and *Quercus* spp. are well capture for those
-    models that have recruitment in these species, but the recruitment
-    values are always lower so the range of variation is also always
-    around 0 (to be confirmed!). However some models lack of recruitment
-    for these species (Figure \@ref(fig:diff)).
-
--   The values that we are seeing are means of sample per site and mean
-    across the sites included in each environmental bin. It is important
-    to noticed that those groups of sites in the empirical data where
-    the recruited basal area of the species was 0 and the stand
-    recruited basal area for all species was also 0, the cells color is
-    represented as grey (Figure \@ref(fig:diff)). The lack of value in
-    the empirical data can be originated due to not recruitment in the
-    group of sites or that the sites did not have recruitment values for
-    7cm threshold.
-
-    Shall we plot (Figure \@ref(fig:diff)) for 10cm in the appendix and
-    remove FiguresS 10-12?
+It is important to keep in mind that the values shown in Figure
+\@ref(fig:diff) are the means of all samples per site, and the mean
+across the sites in each environmental bin. In this figure we are seeing
+together two types of zeros (represented with a lack of dots in the
+corresponding tile), the first type of zero are those sites where the
+recruited basal area of the species was zero and the stand recruited
+basal area for all species was also zero, those were considered as zero
+BA share; the second type of zeros are those sites where there was no
+recruitment for that species but there was recruitment for other
+species.
 
@@
 The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species.
-The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equalt to zero.", out.width='100%'}
-knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles.png"))
+The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equal to zero.", out.width='100%'}
+knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
 ```

### 2022-11-17 code/plottingManuscript.R
change empirical with observed label

@@ -7,7 +7,7 @@
 
 # Order of models with names as in the dataset----
 
-    modelsOrder <- modelsSel <-  c("Empirical" ,"4C", 
+    modelsOrder <- modelsSel <-  c("Observed" ,"4C", 
                      "ForCEEPS", "ForCEEPS(f)",
                      "FORMIND", "ForClim 1", "ForClim 11", 
                      "SIBYLA", "xComp", "PICUS",
@@ -26,7 +26,7 @@
 
 # Colors per model -----
 
-    values_color <- c("Empirical" = "#FF0000",   #red
+    values_color <- c("Observed" = "#FF0000",   #red
                       "4C" = "#FF7F00",          #orange
                       "ForCEEPS" = "#A6CEE3",    #blue light 
                       "ForCEEPS(f)" = "#1F78B4", #blue darker
@@ -46,7 +46,7 @@
 
 # Labels per model -----
 
-    labels <- c("Empirical" = "Empirical", 
+    labels <- c("Observed" = "Observed", 
                 "4C" = "4C", 
                 "Forceeps" = "ForCEEPS",
                 "Forceeps_f" = "ForCEEPS(f)",

### 2022-11-17 manuscript/manuscriptRW.Rmd
add mortality figures and change label "empirical" with "observed" in figures script and manuscript

@@
 Species simulated: 4C only simulated 4 species. Treemig. ForClim and
-Empirical data have more species assigned all together into category
+Observed data have more species assigned all together into category
 "others". LPJ-GUESS simulated bushes that were not consider in the
@@
 
-ToDo add a map with the species composition of the empirical data
+ToDo add a map with the species composition of the observed data
 
@@
 of SYBILA, Landis II and aDGVM2 that estimated recruitment rates at the
-lower end of the plausibility interval of the empirical data (Figure
+lower end of the plausibility interval of the observed data (Figure
 \@ref(fig:over)). The models with highest overestimation were PICUS and
@@
 climatic gradients, where they feature varying effects (Figure
-\@ref(fig:nvTrend710BA), \@ref(fig:nvTrend710BA500) and FigureS 1-4).
+\@ref(fig:nvTrend710BA) and FigureS 1-4).
 
@@
 although they consistently underestimate its variability across sites.
-Four models overestimate recruitment diversity (4C, PICUS, LANDIS-II and
+Four models overestimate recruitment diversity:4C, PICUS, LANDIS-II and
 LPJ-GUESS, the latter mostly for the 10 cm diameter threshold. The model
@@
 models there are small differences across sites regarding the species
-richness of recruitment between 7 and 10 cm. Most models maintain or
-even increase in recruitment diversity between the 7 and 10 cm
-thresholds, and the same is visible from the empirical data, however the
-differences are not significant. **EXPLAIN WHY  "Can we understand this
-(I know this is for the Discussion, but still...)?? This cannot be
-species richness (as there are no new species popping between 7 and 10
-cm DBH). I'd also expect the number of species to be lower at 10 cm than
-at 7 cm (because some will die off due to insufficient light etc.), but
-this is not reflected in H' because species with zero abundance are
-ignored (ln[0] is not defined). So a strongly increasing evenness must
-be driving this increase of H'."**
+richness of recruitment between 7 and 10 cm. Most models maintain the
+recruitment diversity between the 7 and 10 cm thresholds, and the same
+is visible from the observed data, there are only few models where the
+recruitment diversity increases between 7 and 10 cm which might be an
+artifact of the sample strategy and calculating the Shannon index based
+in BA as for a same sample recruitment at 10cm will include larger trees
+with larger BA and if those are from different species, the Shannon
+index will be larger.
 
@@
 The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species.
-The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equal to zero.", out.width='100%'}
+The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the observed data did not becuase both recruitment and stand level recruitments were equal to zero.", out.width='100%'}
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
@@
 
-**Main outcomes mortality in number of trees:**
+There are strong differences across models on the mortality rate between
+7 and 10 cm (Figure \@ref(fig:mortality11)). All models except for the
+two DGVMs LPJ-GUESS and aDGVM2 feature a mortality rate larger than the
+observed data, thus at least partially compensating for the general
+overestimation of recruitment rates (Figure \@ref(fig:over)). Some of
+the models showed strong compensation for the initial overestimation of
+recruitment with larger mortality of recruitment between 7 and 10 cm
+thresholds but not in others. On the one hand, models below the Reineke
+line have lower mortality than expected by the self-thinning rule, which
+may point to facilitation, or simply a higher mortality rate before the
+trees reached 7 cm diameter. On the other hand, models such as xComp,
+PICUS and TreeMig are above the self-thinning line (Reineke), indicating
+that they have higher mortality in recruitment of mixed species forest
+than the expected in an even-aged single species forests, which
+indicates strong compensation for the initial overestimation of
+recruitment in some models. However, some of the models simulate a
+mortality rate in recruitment that is broadly compatible with the
+self-thinning rule, i.e. 4C, SIBYLA , LandClim and LANDIS II.
 
--   There are differences across models on the mortality rate between 7
-    and 10 cm (Figure \@ref(fig:mortality11)).
+Models with higher mortality also show higher tree recruitment at 7cm
+(Figure \@ref(fig:)\@ref(fig:mortality13)). On the other hand, most of
+the models with a low variation in mortality between 7 and 10 cm
+(ForCEEPS, ForCLim 1, ForClim 11 and LandClim) also have lower
+overestimation in recruitment at 7 cm (Figure \@ref(fig:mortality11) and
+FigureS XX). Some of the models that start at a lower number of
+individuals (7cm thershold) have a lower decrease between 7 and 10cm dbh
+(ratio 7-10cm); in these models the species competition and
+self-thinning is either not that large or it happened before the species
+reached 7cm. However, this pattern it is not consistent across models
+(FigureS XX). There is an effect of the number of individuals at 7cm on
+the mortality rate. Most of the models show a positive effect
+(increasing recruitment at 7cm shows increases the mortality between 7
+and 10cm), but some models show a negative effect, however the trend it
+is not very high and not always significant when we look at the slope of
+the linear trends per model (FigureS XX and tableS XX).
 
--   Most of the models have a mortality rate larger than those present
-    in the empirical data (Figure \@ref(fig:mortality11)).
-
--   Several models simulate a mortality rate in recruitment that could
-    be explained by the self thinning rule (Figure
-    \@ref(fig:mortality11)).
-
--   Models below the Reineke line have lower mortality than the expected
-    by self thinning rule, which might point to facilitation or higher
-    mortality before the species reached 7 cm diameter (Figure
-    \@ref(fig:mortality11)).
-
--   Values below 1 line show sites where there is a higher number of 10
-    cm trees than in 7 cm- this is an artifact of the method - we expect
-    space for time substitution, but in reality we just have one
-    measurement (Figure \@ref(fig:mortality11)).
-
--   Models such as SYBILA, xComp, PICUS, LandClim amd TreeMig are above
-    the self thinning line (Reineke) indicating that they have higher
-    mortality in recruitment of mixed species forest than the expected
-    in an even aged single species forests (Figure
-    \@ref(fig:mortality11)).
-
--   Most of the models with lower variation in mortality between 7 and
-    10cm also have lower overestimation in recruitment at 7cm (Figure
-    \@ref(fig:mortality13) and FigureS 13).
-
--   Models with higher mortality also show higher tree recruitment at
-    7cm (Figure \@ref(fig:mortality13)).
-
--   There might be a difference between models including stochastic
-    processes in recruitment and those which don't, but perhaps the
-    reality implies higher mortality rates in the adult component
-    (enhance the share of less shade tolerant species?) [to the
-    discussion] ((Figure \@ref(fig:mortality11) and
-    \@ref(fig:mortality13)).
-
--   Some of the models that start at a lower number of individuals (7cm)
-    have a lower decrease between 7 and 10cm dbh (ratio 7-10cm). The
-    species competition and self-thinning is either not that large or it
-    happened before the species reached 7cm. This pattern it is not
-    consistent across models (FigureS 14-15)
-
--   There is an effect of the number of individuals at 7cm on the
-    mortality rate. Most of the models show a positive effect
-    (increasing recruitment at 7cm shows increases the mortality between
-    7 and 10cm), but some models show a negative effect. The cloud point
-    per models shows the differences across sites, the shape point cloud
-    per model shows if the models take into account regeneration density
-    when applying mortality to regeneration (FigureS 14-15) This is not
-    that clear anymore when we calculate the slope of the linear trends
-    per model.
-
-```{r  mortality11, echo=FALSE, fig.cap="Ratio of tree recruitment between 7 and 10 cm recruitment treshold.
-Dashed blue lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions", out.width='75%'}
+```{r  mortality11, echo=FALSE, fig.cap="Ratio of tree recruitment rates between the 7 and 10 cm thresholds.
+Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree recruitment between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
 knitr::include_graphics(here::here("figures", "ratio7_10.png"))
@@
 
-```{r mortality13, echo=FALSE, fig.cap="Ratio of tree recruitment between 7 and 10 cm recruitment treshold and the overestimation proportion (simulated - empirical / empirical) for the mean recruitment trees per model across sites and samples.
+```{r mortality13, echo=FALSE, fig.cap="Ratio of tree recruitment between 7 and 10 cm recruitment treshold and the overestimation proportion (simulated - observed / observed) for the mean recruitment trees per model across sites and samples.
 The horizontal dotted lines show a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions.
-The vertical line representes the 0 overestimation proportion", out.width='100%'}
+The vertical line represents the 0 overestimation proportion", out.width='100%'}
 knitr::include_graphics(here::here("figures", "mort7_10_medianOver.png"))
@@
 
+-   \# H1: the higher regen7 the stronger the regen mortality
+
+    \# decrease in stem number between regen7 and regen10
+
+    \# as ratio nn7/nn10 (colname nn710)
+
+    \# reference lines
+
+    \# -ratio 1 = no decrease from 7 -\> 10
+
+    \# -ratio for selfthinning under even aged conditions
+
+    \# with Reineke exp rk=-1.605: (7/10)\^(-1.605) = 1.77
+
+-   Models below the Reineke line have lower mortality than expected by
+    the self-thinning
+    rule, which may point to facilitation, or simply a higher mortality
+    rate before the trees reached 7 cm diameter. But then these models
+    would need to underestimate recruitment levels (Fig. 1), no? But
+    this is not what we see... not sure how to explain.
+
+-   methodological limitations compare mortality with the self thinning
+    line etc Values below the 1.0 line show sites where there is a
+    higher number of 10 cm trees than in 7 cm - this is an
+    artifact of the method - we expect space for time substitution, but
+    in reality
+    we just have one measurement (Figure 7).
+
+-   There might be a difference between models including stochastic
+    processes in recruitment and those which do not, but perhaps the
+    reality implies higher mortality rates in the adult component
+    (enhance the share of less shade tolerant species?) [to the
+    discussion] ((Figure \@ref(fig:mortality11) and
+    \@ref(fig:mortality13)).
+
+-    Most models maintain or even increase in recruitment diversity
+    between the 7 and 10 cm thresholds, and the same is visible from the
+    observed data, however the differences are not significant.
+    **EXPLAIN WHY  "Can we understand this (I know this is for the
+    Discussion, but still...)?? This cannot be species richness (as
+    there are no new species popping between 7 and 10 cm DBH). I'd also
+    expect the number of species to be lower at 10 cm than at 7 cm
+    (because some will die off due to insufficient light etc.), but this
+    is not reflected in H' because species with zero abundance are
+    ignored (ln[0] is not defined). So a strongly increasing evenness
+    must be driving this increase of H'."** This could be because we are
+    comparing 7 and 10cm from same samples, and at 10cm you might
+    include more trees and at 7cm This calls for further research and
+    anotehr paper :D
+
 # References

### 2022-11-17 code/figures.R

@@ -13,8 +13,10 @@
 # Load data ----
 
     outputsDF <- data.table::fread("data/dataOutputs.csv")
+    outputsDF$model[outputsDF$model == "Empirical"] <- "Observed"
 
-# Assign models and dbh to levels 
+
+    # Assign models and dbh to levels 
 
     outputsDF$model <- factor(outputsDF$model, levels = modelsOrder)
     outputsDF$dbh <- factor(outputsDF$dbh, levels = c("10", "7"))
@@ -65,8 +67,8 @@
     
 # Zero inflated data in recruitment 
       
-    zeroInflSim <- outputsDF[!outputsDF$model == "Empirical",]
-    zeroInflEmp <- outputsDF[outputsDF$model == "Empirical",]
+    zeroInflSim <- outputsDF[!outputsDF$model == "Observed",]
+    zeroInflEmp <- outputsDF[outputsDF$model == "Observed",]
     
     zeroInflSim <-  zeroInflSim[, c("site", "sample", "model", "Totr.ba")]
     zeroInflSim <- zeroInflSim[!duplicated(zeroInflSim)]
@@ -136,11 +138,11 @@
                                                                              fill = "red")),
                                     name = "Diameter threshold",
                                     labels = c('7 cm', '10 cm')) +
-        ggplot2::geom_hline(ggplot2::aes(yintercept = quantile(overDTsim$r.trees[overDTsim$dbh == "7" & overDTsim$model == "Empirical"],
+        ggplot2::geom_hline(ggplot2::aes(yintercept = quantile(overDTsim$r.trees[overDTsim$dbh == "7" & overDTsim$model == "Observed"],
                                                                c(0.25),
                                                                na.rm = TRUE)), 
                             colour = "#990000", linetype = "dashed") +
-        ggplot2::geom_hline(ggplot2::aes(yintercept = quantile(overDTsim$r.trees[overDTsim$dbh == "10" & overDTsim$model == "Empirical"],
+        ggplot2::geom_hline(ggplot2::aes(yintercept = quantile(overDTsim$r.trees[overDTsim$dbh == "10" & overDTsim$model == "Observed"],
                                                                c(0.70),
                                                                na.rm = TRUE)),
                             colour = "#990000", linetype = "dashed") 
@@ -154,7 +156,7 @@
 ## R VS. env. trends-----
 
[…]

### 2022-11-17 manuscript/manuscriptRW.Rmd
added protocol information and model section in manuscript material and methods

@@
 ---
-title: "Capturing regeneration in models of forest dynamics"
+title: "Drivers of tree regeneration remain challenging to capture for models of forest dynamics"
 author:
@@
 
-In comparison to recruitment, growth and mortality responses to climate
-change are much more studied
+In comparison to recruitment, growth and mortality [@bugmann2019]
+responses to climate change are much more studied
 
@@
 
-A total of 15 models and model variants of forest dynamics at the stand,
-landscape and global level were used to simulate stand dynamics and
-sample regeneration in 200 sites selected across Europe.
+## The models
 
-About the models The models ( Table @ref(tab:tableModels)) were created
-with different philosophies to represent regeneration. Three models,
-LPJ-GUESS, the two variants of ForClim and Treemig simulated more
-species than the one asked because XXXX , these species were assigned to
-"other" category in the case of TreeMig and in the case of LPJ-GUESS
-they were removed from the outputs as they are shrubs species (Boreal
-evergreen shrub, Juniperus oxycedrus and quercus coccifera) LPJ-GUESS
-also simulate Corylus avellana but only had 3 trees across all samples
-and they were removed because due to the low degrees of freedom it was
-not possible to create a model for the category "others" for this
-model). Forclim simulated a total of 17 species more than the species
-asked in the protocol . Observed data also have a category others.
+Fifteen models of forest dynamics (including two model variants) were
+used to assess tree regeneration. The models varied on their scale
+including models at the stand, landscape and global scale (see Table
+\@ref(tab:tableModels)). Their model construction philosophies and
+origin also differs with two models where their formulations were
+derived using statistical analyses based on inventory data (SYBILA and
+xComp) and others with largely process based approach. Some of the
+models largely rely on the original approach underlying forest gap
+models (ForClim 1, ForClim 11, LandClim, Landis II ); and some include
+plant ecophysiology formulations (FORMIND, PICUS, 4C, iLand). This
+different approaches in the models construction also has influence how
+regeneration is defined.
 
-About the sites The sites represent the available range of environmental
-gradients in temperature and precipitation (Figure using the variables
-degree day and water balance. The sites were selected across a network
-of plots that collect information on forest reserves
-[@kaber_sheltered_nodate]
+The models can be categorized as regeneration models and recruitment (or
+ingrowth) models [@vanclay1994, @konig_tree_2022]. Regeneration models
+include in their regeneration modules processes such as flowering and
+pollination, seed production, seed dispersal, germination or seedling
+establishment that lead to the number of seedlings or saplings.
+Recruitment models introduce a number of seedlings or saplings with
+certain characteristic (biomass or dbh) without considering earlier
+development processes. We can also distinguish models that have feedback
+in the regeneration as those in which the level and species composition
+in regeneration is influenced by the existing adult tree community, this
+can be done via the production of seed, seedlings or saplings. The
+starting point for regeneration it was also different per model varying
+from seedbank, seed, sapling and ingrowth. The overall complexity in the
+regeneration modules also varied, following @bugmann2022 we can see that
+most of the models have rather low/high? complexity while few of them
+have high/low?. Further details on the models specific formulation
+related to their regeneration modules and an overall summary across
+models is presented in the supplementary materials.
 
-Each modeler group was provided with detailed information on how to
-perform the simulations including the input variables such as soil or
-climate, simulation setup and expected output variables. The overarching
-goal of this simulations was the assessment of the regeneration niche as
-it arises from empirical data analysis vs. its representation in a wide
+e used repeated tree census data collected in forest reserves and
+unmanaged (or protected) forests across Europe, 127 which were compiled
+in the framework of the EuFoRIa network.
+
+The sampled forests have not been managed for 128 at least one decade,
+but most for much longer, and some are primary forests.
+
+The final data set contained 701 units, each covering on average 1 ha of
+forest area with 142 observations of tree recruitment at a DBH of 7 cm
+(see
+
+aggregated
+
+## The observed data
+
+Regeneration observations for a wide range of environmental gradients
+are hard to obtain and this is one of the reasons why most of the models
+of forest dynamics have never been confronted with a data-set covering a
+wide range of regeneration sites across climate gradients to evaluate
+how well are they capturing regeneration.
+
+The observed data represent the available range of environmental
+gradients in temperature and precipitation from 200 sites of strict
+forest reserves in Europe compiled in the framework of the EuFoRIa
+network. The sample forest have not been managed in at least 10 years,
+and each of the sites contains at least two consecutive measurements.
+Regeneration was measured at 7 and 10 cm threshold, however 35 sites
+only contain information for 10 cm diameter. The 200 sites represent the
+available climatic gradients of temperature and precipitation across
+Europe and they varied in species composition (Figure \@ref(fig:map)).
+For more information on this dataset and their regeneration please refer
+to @kaber_sheltered_nodate.
+
+## Simulation protocol
+
+Each modeler group was provided with detailed protocol (available in the
+supplementary material) on how to perform the simulations regarding
+input variables such as climate or soil, simulation setup and expected
+output variables. No information was provided on the regeneration on
+these sites before the models performed were done. The overarching goal
+of this simulations was the assessment of the regeneration niche as it
+arises from empirical data analysis against its representation in a wide
 range of models of forest dynamics.
+
+The input variables were collected and aggregated to be adapted to the
+different model demands. The time series of climatic variables from 1981
+to 2018 were provided in hourly (Era5-land data @muñozsabater ), daily
+and monthly resolution (CHELSA data, @karger ). Some variables required
+for some models such as relative humidity or vapor pressure deficit had
+to be calculated from these variables. The final use of the climate data
+was different in each model (Table \@ref(tab:tableModels))The soil
+quality data was provided as a continuous value between 1 and 5
+(Soilgrids dataset, @hengl2017, @poggio2021 ). Elevation, slope and
+aspect was also provided for each of the 200 sites, however no other
+spatial information such as coordinates were provided (@astergd ).
+
+The simulations had to be run in the absence of natural disturbances,
+however one the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE
+[ONE SENTENCE]{.underline} OF WHY YOU INCLUDED DISTURBANCES AND WHAT
+TYPE. The simulations had to be set up in a way that they will obtain
+samples on species specific regeneration rates per decade and per ha.
+The modeling teams decided on the simulated area and how they derive
+these samples (further details on how each modeling team prepared the
+simulations are available in the supplementary material).
+
+The simulations were run from bare ground in the absence of management
+to a simulated equilibrium (\"potential natural vegetation\") with the
+current climate. The goal was to derive the regeneration niche.The
+regeneration niche is defined as the passing of a diameter threshold of
+7 and 10 cm (ingrowth; cf. below). Both the width of the regeneration
+niche (i.e., in environmental space) as well as the intensity of the
+recruitment process (i.e., the number of ingrowth trees per area and per
+unit of time) are of interest. The exact length of the simulation should
+was decided by the modeling teams (Table \@ref(tab:tableModels)).
+
+The simulations were run for mixed species forests (not multiple
+single-species simulations) using mixtures of the following
+species/genera: *Fagus sylvatica, Picea abies, Abies alba, Carpinus
+betulus, Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus
+excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris.* Three
+models included additional species in the simulations (ForClim 1,
+ForClim 11 and TreeMig) and three could not simulate all the species
+(4C, xComp, TreeMig, LPJ-GUESS). HERE **FORCLIM** PROVIDES [ONE
+SENTENCE]{.underline} WHY IT SIMULATED MORE SPECIES. HERE **TREEMIG**
+PROVIDES [ONE SENTENCE]{.underline} WHY IT SIMULATED MORE SPECIES AND
+MISSED ONE. HERE **4C** PROVIDES [ONE SENTENCE]{.underline} WHY IT
+SIMULATED LESS SPECIES. HERE **XCOMP** PROVIDES [ONE
+SENTENCE]{.underline} WHY IT SIMULATED LESS SPECIES. The model aDGVM2
+did not have species level option and provided the default approach with
+
+The simualtion have to be provided for two diameter thresholds: 7 and 10
+cm.
 
@@
 simulations (current Figure S5)
+
+```{r map, echo=FALSE, fig.cap="Map with the site observations and species presnet in regeneration.", out.width='80%'}
+knitr::include_graphics(here::here("figures", "theMap.png"))
+```
+
+## Analysis
 
@@
 
-modelDF$Type[modelDF$Type ==  "Stand"] <- "S"
-modelDF$Type[modelDF$Type ==  "Landscape"] <- "L"
-modelDF$Type[modelDF$Type ==  "Global"] <- "G"
+# modelDF$Type[modelDF$Type ==  "Stand"] <- "S"
+# modelDF$Type[modelDF$Type ==  "Landscape"] <- "L"
+# modelDF$Type[modelDF$Type ==  "Global"] <- "G"
 
@@
 
-modelDF$Reference <- "reference"
-
-kable(modelDF,  format = "markdown", caption = "List of models included.")
+kable(modelDF, 
+      #format = "markdown",
+      format = "latex", booktabs = TRUE,
+      caption = "List of models included.") |>
+        kableExtra::landscape() |>
+        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
+    # fix width by column
+      kableExtra::column_spec(column = 1, width = "5em") |> 
+      kableExtra::column_spec(column = 2, width = "5em") |> 
+      kableExtra::column_spec(column = 3, width = "5em") |> 
+      kableExtra::column_spec(column = 4, width = "5em") |>
+      kableExtra::column_spec(column = 5, width = "5em") |>
+      kableExtra::column_spec(column = 6, width = "5em") |>
+      kableExtra::column_spec(column = 7, width = "8em") |>
+      kableExtra::column_spec(column = 8, width = "5em") |>
+      kableExtra::column_spec(column = 9, width = "5em") |>
+      kableExtra::column_spec(column = 10, width = "5em") |>
+      kableExtra::column_spec(column = 11, width = "5em") |>
+      kableExtra::column_spec(column = 12, width = "5em") 
 ```
@@
 
+-    Generally, the approach was appreciated a lot, i.e. the idea of
+    confronting the simulated recruitment along environmental gradients
+    with data from such gradients. With EuFoRIa, we seem to be on a very
+    good track in this regard.\
+    \
+    • Critical comments were given to the choice of an ingrowth limit of
+    10 cm DBH. Many participants said, in one way or the other, that the
+    good match is not surprising because if the model gets overall
+    species composition right, it must also get ingrowth of these rather
+    large trees right. Thus, a better test would be for the ingrowth
+    limit to be lower. My explanation that we wanted to maximize the
+    dataset available and thus chose the highest callipering limit was
+    accepted, but I think it is definitely so that we need to provide
+    data with a lower limit. How much data do we lose if we go down to 4
+    cm? I am afraid that going down to 1 cm is not realistic because
+    we\'d lose most of the data. So the bottomline is that 10 cm are
+    fine, but a very soft test, and hence we need to provide material
+    for a harder test. Can you tell me (1) how much data we\'d lose with
+    4 cm, and (2) what the patterns along the two gradients (CWB and DD)
+    are looking like when using a 4 cm threshold?\
+    \
+    • Many participants also felt that we should not miss out on a third
+    very important axis, i.e. light availability. Obviously this is more
+    tricky because we do not have such measurements, but I think we can
+    work with proxies such as basal area (although it\'s difficult for
+    the small sampling plots, as light availability may be poorly
+    related to BA on a 500 m2 plot...). Have you looked into this
+    already? It would be really nice if we could provide a pattern that
+    shows how recruitment of the different species is becoming sparser
+    and drops out completely as a function of a proxy of light
+    availability.\
+    \
+    Beyond that, some people expressed their desire to calibrate a
+    recruitment model based on the EuFoRIa data, and I tried to explain
+    why I think it is more insightful to use these data for model
+    benchmarking: we can learn more... :-)\
+
 # References

### 2022-11-19 manuscript/manuscriptRW.Rmd
lots of writing in the manuscript and bullet poits in discussion

--- 2022-11-17 manuscript/manuscriptRW.Rmd
@@
 ---
-title: "Capturing regeneration in models of forest dynamics"
+title: "Drivers of tree regeneration remain challenging to be captured by models of forest dynamics"
 author: 
-- Olalla Díaz-Yáñez, Forest Ecology ETH Zurich
-- Yannek Käber, Forest Ecology ETH Zurich
-- Mats Mahnken, PIK Potsdam
-- Samuel Fischer, UFZ
-- Rico Fischer, UFZ
-- Friedrich Bohn, UFZ
-- Xavier Morin, cefe cnrs
-- Manfred Lexer, BOKU
-- Christian Hochauer, BOKU
-- Katarina Merganicova, forim
-- Tobias Mette, LWF Bayern
-- Werner Rammer, TUM
-- Kristin H. Braziunas, TUM
-- Paola Mairota, UNIBA
-- Josef Bruna, ibot cas
-- Heike Lischke, WSL
-- Daniel Scherrer, nWSL
-- Simon Scheiter, senckenberg
-- Thomas Hickler, senckenberg
-- Tim Anders, senckenberg
-- Jessica Hetzer, senckenberg
-- Harald Bugmann, Forest Ecology, ETH Zurich
+- Olalla Díaz-Yáñez, 
+- Yannek Käber, 
+- Mats Mahnken, 
+- Samuel Fischer, 
+- Xavier Morin, 
+- Manfred Lexer, 
+- Christian Hochauer, 
+- Katarina Merganicova, 
+- Tobias Mette, 
+- Werner Rammer, 
+- Kristin H. Braziunas, 
+- Paola Mairota, 
+- Josef Bruna,
+- Heike Lischke, 
+- Daniel Scherrer, 
+- Simon Scheiter, 
+- Thomas Hickler, 
+- Tim Anders,
+- Jessica Hetzer, 
+- Harald Bugmann
 date: "`r Sys.Date()`"
@@
     reference_docx: template_paper.docx
+mainfont: Arial
+fontsize: 12pt
+linestretch: 1.5
 bibliography: Regeneration.bib
@@
   library(knitcitations); cleanbib()
-  cite_options(citation_format = "pandoc", check.entries=FALSE)
+  cite_options(citation_format = "pandoc", check.entries = FALSE)
   library(bibtex)
@@
 
+    Word count: `r as.integer(sub("(\\d+).+$", "\\1", system(sprintf("wc -w %s", knitr::current_input()), intern = TRUE))) - 20`
+
 # Abstract
 
-*Regeneration is a key process in forest dynamics. Models of forest
+Regeneration is a key process in forest dynamics. Models of forest
 dynamics have been long lagging an assessment in their capacities to
-properly represent this process. Using 13 models of forest dynamics at
-the stand, landscape and global levels, we simulate the regeneration
-across 200 sites representing environmental gradients across Europe. The
-outputs were compared with regeneration data collected from forest
-reserves across Europe. In this study we wanted to asses how well are
-models of forest dynamics capturing regeneration levels and what can be
-changed to improve this. The results show that most of the models
-over-estimate regeneration values. However most of them got right trends
-in regeneration across temperature and precipitation gradients. Stand
-species richness was better capture than recruitment species richness.
-Mortality*
+properly represent this process. Using 15 models and variants of forest
+dynamics at the stand, landscape and global levels, we simulate the
+regeneration across 200 sites representing environmental gradients
+across Europe. The simulation results were compared with observed data
+collected from forest reserves across Europe. In this study we want to
+asses how well are models of forest dynamics capturing regeneration
+levels, species composition and mortality in regeneration. The results
+show that .
 
@@
 
-Tree recruitment is fundamental for forest dynamics, in particular with
-increasing disturbance under climate change and it is also crucial
-understand the forest resilience under climate change.
+Forests provide multiple products and services of vital importance.
+Under the pressure of climate change and increasing disturbances, it is
+even more important that we understand how forests will change and
+evaluate their capacity to adapt or be resilient to these changes
+[@lindner2010]. A wide range of models of forest dynamics were developed
+over the past decades considering the impacts of climate in forest
+dynamics [@bugmann2022]. From these studies, it is becoming increasingly
+evident that while we have a reasonably good understanding how to model
+tree growth [@vanclay1997; @bugmann1996], and substantial efforts have
+been dedicated to improving the depiction of tree mortality processes
+[@bugmann2019], tree regeneration remains a less studied process in most
+of these models, while becoming ever more important in the context of
+forest disturbances and forest resilience.
 
-In comparison to recruitment, growth and mortality responses to climate
-change are much more studied
+Tree recruitment depends on multiple processes such as reproduction,
+fruit maturation, seed production, dispersal, germination, juvenile
+growth and survival; these processes are all difficult to assess and
+some of them are scarcely understood and thus appear highly stochastic.
+Finding the right level of complexity in the definition of the factors
+leading to tree recruitment is challenging and models of forest dynamics
+have used different approaches [@bugmann2022]. Overall, models are
+needed to (1) synthesize existing empirical data and explore their
+consequences, and (2) assess future tree recruitment e.g. in the context
+of global change scenarios. Currently, regeneration processes in dynamic
+forest models are handled in a multitude of ways [@konig_tree_2022;
+@bugmann2022], from entirely ignoring tree regeneration (as done in
+classical forest growth models, e.g. @pretzsch2002), across the use of a
+few simple environmental filters, as done in most forest \"gap\" models
+(e.g., @shuman2011) and Dynamic Global Vegetation Models (e.g.,
+@smith2001; @hickler2012), to complex approaches that incorporate local
+feedback from the canopy, multiple ecological processes and often also
+small time steps (e.g., @wehrli2006; @seidl2012), or field-based
+statistical parameterizations, which however are not easy to extrapolate
+in space and time (e.g., @ribbens1994).
 
-Tree recruitment depends on multiple processes (reproduction, fruit
-maturation, seed production, dispersal, germination, juvenile growth and
-survival) that are all difficult to assess
+With the current strategies that models of forest dynamics are using to
+represent tree recruitment we are obtaining results that faced several
+problems related to over-recruitment, needed excess of mortality at
+early stages and inadecuate species composition in regeneration which
+have strong impacts on forest dynamics. Some model outputs are
+over-recruiting which subsequently must die primarily by competition in
+later stages. This is characteristic of \'classical\' forest gap models
+[@gutiérrez2016], at least as long as the simulations comprise a
+multitude of species. Other models use a model output to correct
+recruitment levels for example by routinely calibrate them against the
+predicted canopy species composition affecting as well the species
+dominance.
 
-Models are needed to (1) synthesize existing empirical data and explore
-their consequences, and (2) assess future tree recruitment e.g. in the
-context of global change scenarios
+Correctly capturing the species composition in recruitment is often also
+a challenge, thus leading to unrealistic successional drift in the model
+REFERENCE?, for example by the change of the dominant species via
+stochastic processes that continuously bring seeds of other species that
+the come to dominance, with this strategy a new species might start to
+rise eliminating the former dominant species. A related issues is the
+rarification of species richness due to positive feedback effects such
+that eventually just single-species forest stands remain. This is
+sometimes handled by the incorporation of a low level of seed influx of
+all species at all times (e.g., @lischke2006 ). However, simulated
+species composition is usually exceedingly sensitive to that level
+whereas the parameters of such functions are poorly constrained by field
+data. There is also often a problem with the insuficient constraint of
+parameter values for models that start from very small tree sizes (e.g.,
+10 cm of height or even seeds) and track tree development in a
+process-oriented manner by considering a multitude of ecological
+influences, rather than being statistically driven and emphasizing
+recruitment into a larger size class (e.g., @wehrli2006 ). Most of these
+undesired behaviors in tree recruitment are often not published as such,
+but in the vast majority of studies they are corrected (e.g.,
+@rasche2011 , where for most study sites regeneration was simply
+switched off). Thus, a strong focus on the modeling of tree regeneration
+processes is highly important and long overdue [@price] if we are to
+make reliable projections of future forest dynamics, i.e. when the
+models need to be operated in extrapolation mode, but also from a
+fundamental ecological point of view (increasing systems understanding).
 
-Regeneration processes in dynamic forest models and how they have been
-handled to date [@konig_tree_2022; @bugmann2022].
-
-Tendency of models to be too liberal with recruitment; consequences of
-over-recruitment Species richness in regeneration, mortality and
-regeneration levels may have strong impacts on forest dynamics
-
-An analysis how well models of forest dynamics perform re. recruitment
-is lacking. This is the first time, a large number of forest models
-commonly used to assess forest dynamics under climate change is
-evaluated against a continental-scale multi-species harmonized data set
-on tree recruitment. Briefly praise the dataset [@käber].
-
-In this study wanted to answer the following questions: 1) Are models of
-forests dynamics capturing correctly regeneration levels across
-environmental gradients 2) Are models of forest dynamics representing
-correctly initial composition of tree populations 3) Can models of
-forest dynamics correctly capture the initial mortality in tree
-in-growth between 7 and 10 cm
+This is the first time, a large number of forest models commonly used to
+assess forest dynamics under climate change are evaluated against a
+continental-scale multi-species harmonized data set on tree recruitment
+[@käber2022]. Tree recruitment is understudied in field studies and
+modelling [@hanburybrown_forest_2022], because little data has been
+available and data was very fragmented, i.e., not in one harmonized data
+set. In this study we relied on models with various 'philosophies',
+including models operating at the stand to the global scale. Because of
+the large variability in recruitment patterns in nature and the large
+amount of factor driving recruitment, including some which are not
+included in most models, such as deer browsing, we do not aim at a
+detailed statistical evaluation of each model. Instead, we aim at
+evaluating general recruitment patterns, magnitudes and benchmark the
+regeneration niche of multiple species based on empirical data along a
+wide environmental gradient (temperature, moisture, and light
+availability). More specifically in this study we wanted to answer the
+following questions: 1) are models of forests dynamics capturing
+correctly regeneration levels across environmental gradients; 2) are
+models of forest dynamics representing correctly initial composition of
+tree populations;and, 3) can models of forest dynamics correctly capture
+the initial mortality in tree in-growth?.
 
@@
 
-A total of 15 models and model variants of forest dynamics at the stand,
-landscape and global level were used to simulate stand dynamics and
-sample regeneration in 200 sites selected across Europe.
+## Models
 
-About the models The models ( Table @ref(tab:tableModels)) were created
-with different philosophies to represent regeneration. Three models,
-LPJ-GUESS, the two variants of ForClim and Treemig simulated more
-species than the one asked because XXXX , these species were assigned to
-"other" category in the case of TreeMig and in the case of LPJ-GUESS
-they were removed from the outputs as they are shrubs species (Boreal
-evergreen shrub, Juniperus oxycedrus and quercus coccifera) LPJ-GUESS
-also simulate Corylus avellana but only had 3 trees across all samples
-and they were removed because due to the low degrees of freedom it was
-not possible to create a model for the category "others" for this
-model). Forclim simulated a total of 17 species more than the species
-asked in the protocol . Observed data also have a category others.
+Fifteen models of forest dynamics (including two model variants) were
+used to assess tree regeneration. The models selected simulate forest
+dynamics at the stand, landscape and global scale (Table
+\@ref(tab:tableModels)). Models construction philosophies and origin
+differs, most of the models have a largely process based approach and
+two models have formulations derived from statistical analyses based on
+inventory data (SYBILA and xComp). Some of the models largely rely on
+the original approach underlying forest gap models (ForClim 1, ForClim
+11, LandClim, Landis II ); and some include plant ecophysiology
+formulations (FORMIND, PICUS, 4C, iLand). This different approaches in
+the models construction also has influenced how tree establishment is
+defined.
 
-About the sites The sites represent the available range of environmental
-gradients in temperature and precipitation (Figure using the variables
-degree day and water balance. The sites were selected across a network
-of plots that collect information on forest reserves
-[@kaber_sheltered_nodate]
+Based on their establishment modules the models can be categorized as
+regeneration or recruitment (or ingrowth) models [@vanclay1994;
+@konig_tree_2022] (see Table \@ref(tab:tableModels)). Regeneration
+models include processes such as flowering and pollination, seed
+production, seed dispersal, germination or seedling establishment that
+lead to the number of seedlings or saplings. Recruitment models
+introduce a number of seedlings or saplings with certain characteristic
+such as biomass or dbh, without considering earlier development
+processes. We can also distinguish models that have feedback in the
+regeneration as those in which the level and species composition in
+regeneration is influenced by the existing adult tree community via the
+production of seeds, seedlings or saplings. The starting point for
+regeneration it is also different per model varying from seedbank, seed,
+sapling or ingrowth. The overall complexity in the establishment modules
+also varies, following @bugmann2022 we can see that most of the models
+included in this study can be classified as having rather low complexity
+in their tree recruitment modules while few of them have high complexity
+specifically when defining processes such as XXXX, XXXX (enumerate
+models). Further details on the models specific formulation related to
+their regeneration modules and an overall summary across models is
+presented in the supplementary materials (SHALL WE DO THIS?).
 
-Each modeler group was provided with detailed information on how to
-perform the simulations including the input variables such as soil or
-climate, simulation setup and expected output variables. The overarching
-goal of this simulations was the assessment of the regeneration niche as
-it arises from empirical data analysis vs. its representation in a wide
-range of models of forest dynamics.
-
-There are 35 sites that had a diameter threshold of 10 cm and therefore
-we do not know the actual recruitment rates for those sites at 7 cm; yet
-the models simulated recruitment for the 7 cm threshold on those sites.
-The 35 sites were removed from the simulated data as well to make a fair
-comparison.
-
-How many species were simulated per model? why?
-
-Need to say something about zero inflation in the observations vs. the
-simulations (current Figure S5)
-
-A GAM model based on XXX was used to compare the trend on the observed
-data and the simulated outputs. DESCRIBE GAM
-
-Species simulated: 4C only simulated 4 species. Treemig. ForClim and
-Observed data have more species assigned all together into category
-"others". LPJ-GUESS simulated bushes that were not consider in the
-output of speces composition neither recruitment.
-
-In table @ref(tab:tableModels)
-
-```{r tableModels, echo=FALSE,  message=FALSE}
+```{r tableModels, echo=FALSE,  message=FALSE, warning=FALSE}
 library(knitr)
@@
 modelDF <- readr::read_csv(here::here("data", "modelsDescription.csv"))
-
-modelDF$Type[modelDF$Type ==  "Stand"] <- "S"
-modelDF$Type[modelDF$Type ==  "Landscape"] <- "L"
-modelDF$Type[modelDF$Type ==  "Global"] <- "G"
-
 modelDF$`Species simulated`[modelDF$`Species simulated` == "F. sylbatica, P. sylvestris, P. abies, A. alba"] <- "F. sylbatica,  P. sylvestris, P. abies,  A. alba"
@@
 
-modelDF$Reference <- "reference"
-
-kable(modelDF,  format = "markdown", caption = "List of models included.")
+kable(modelDF, 
+      #format = "markdown",
+      format = "latex", booktabs = TRUE,
+      caption = "List of models included.") |>
+        kableExtra::landscape() |>
+        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
+    # fix width by column
+      kableExtra::column_spec(column = 1, width = "5em") |> 
+      kableExtra::column_spec(column = 2, width = "5em") |> 
+      kableExtra::column_spec(column = 3, width = "5em") |> 
+      kableExtra::column_spec(column = 4, width = "5em") |>
+      kableExtra::column_spec(column = 5, width = "5em") |>
+      kableExtra::column_spec(column = 6, width = "5em") |>
+      kableExtra::column_spec(column = 7, width = "8em") |>
+      kableExtra::column_spec(column = 8, width = "5em") |>
+      kableExtra::column_spec(column = 9, width = "5em") |>
+      kableExtra::column_spec(column = 10, width = "5em") |>
+      kableExtra::column_spec(column = 11, width = "5em") |>
+      kableExtra::column_spec(column = 12, width = "5em") 
 ```
 
-ToDo add a map with the species composition of the observed data
+## Observed data
 
-ToDo check in Haralds paper which models are in which category in the
-classification of establishment and link this with the results we see
-here in structure and complexity. For example iLand and Picus have a
-similar structure but results are very different. Are we meeting the
-expectations in behaviour fromthe different models?
+Regeneration data covering a wide range of environmental gradients is
+hard to obtain and this is one of the reasons why most of the models of
+forest dynamics have never been confronted with a data-set covering a
+wide range of climate gradients in a large number sites to evaluate how
+well are they capturing regeneration. The observed regeneration data
+used in this study represent the available range of environmental
+gradients in temperature and precipitation in Europe using 200 sites of
+strict forest reserves compiled in the framework of the EuFoRIa network
+(Figure \@ref(fig:map)). The sample forests have not been managed in at
+least 10 years (most of them in close to 100 years?), and each of the
+sites contains at least two consecutive measurements. Regeneration was
+measured at 7 and 10 cm threshold in all the sites except for 35 sites
+where only regeneration at 10 cm diameter threshold was available. For
+more details on this dataset and their regeneration information please
+refer to @käber2022.
 
-Species composition in regeneration
+## Simulation protocol
 
-We evaluated the species compositions across models and in relation to
+Each modeler group was provided with detailed protocol (available in the
+supplementary material) with instruction on how to perform the
+simulations regarding input variables such as climate or soil,
+simulation setup and expected output variables. No information was
+provided on the regeneration prior to the simulation. The overarching
+goal of this experiment was the assessment of the regeneration niche as
+it arises from empirical data analysis against its representation in a
+wide range of models of forest dynamics.
+
+The input variables were collected and aggregated to be adapted to the
+different model demands. The time series of climatic variables from 1981
+to 2018 were provided in hourly (Era5-land data, @muñozsabater2019),
+daily and monthly resolution (CHELSA data, @karger2021). Some variables
+required for some models such as relative humidity or vapor pressure
+deficit had to be calculated from these variables. The final strategy on
+how to use the climate data was different in each model based on the
+approach the better fitted the model (Table \@ref(tab:tableModels)). The
+soil quality data was provided as a continuous value between 1 and 5
+(Soilgrids dataset, @hengl2017, @poggio2021). The protocol also provided
+the elevation, slope and aspect for each of the 200 sites
+[@asterscienceteam2019], no other spatial information such as
+coordinates were given with the exception of iLand that used the blurred
+coordinates to obtain detailed soil information values for the different
+sites.
+
+The simulations had to be run in the absence of natural disturbances,
+one the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE [ONE
+SENTENCE]{.underline} OF WHY YOU INCLUDED DISTURBANCES AND WHAT TYPE.
+The simulations had to be set up in a way that they will obtain samples
+on species specific regeneration rates per decade and per ha. The
+modeling teams decided on the simulated area and how they derive these
+samples. Further details on how each modeling team prepared the
+simulations and the outputs are available in the supplementary material.
+
+The simulations were run from bare ground in the absence of management
+to a simulated equilibrium ("potential natural vegetation") with the
+current climate as the ultimate goal was to derive the regeneration
+niche. The regeneration niche is defined as the passing of a diameter
+threshold of 7 and 10 cm (ingrowth). Both the width of the regeneration
+niche (i.e., in environmental space) as well as the intensity of the
+recruitment process (i.e., the number of ingrowth trees per area and per
+unit of time) were of interest. The exact length of the simulation was
+decided by the modeling teams (Table \@ref(tab:tableModels)).
+
+The simulations were run for mixed species forests (not multiple
+single-species simulations) using mixtures of the following species or
+genera: *Fagus sylvatica, Picea abies, Abies alba, Carpinus betulus,
+Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus excelsior,
+Quercus spp., Alnus glutinosa, Pinus sylvestris.* Three models included
+additional species in the simulations (ForClim 1, ForClim 11 and
+TreeMig) and three could not simulate all the species (4C, xComp,
+TreeMig, LPJ-GUESS) (Table \@ref(tab:tableModels)). HERE **FORCLIM**
+PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED MORE SPECIES.
+HERE **TREEMIG** PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED
+MORE SPECIES AND MISSED ONE. HERE **4C** PROVIDES [ONE
+SENTENCE]{.underline} WHY THEY SIMULATED LESS SPECIES. HERE **XCOMP**
+PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED LESS SPECIES. The
+model aDGVM2 did not have species level option and provided the default
+approach with HERE **aDVGM2** PROVIDES [ONE SENTENCE]{.underline} WITH
+THE INFORMATION ON THE DEFAULT SET UP FOR THE SPECIES.
+
+Each of the models reported the number of recruits over a 10 year
+interval per species and per ha for each of the sites. To better
+understand the variation per site each modeling team decided on how to
+sample 200 multiple values per site to show the variation in recruitment
+for spatial units of 1 ha during the simulated phase of potential
+natural vegetation but without providing average of multiple 1-ha
+samples. This was done with different strategies such as sampling
+simulated data from the same 1-ha plot in the equilibrium over time, or
+by sampling several 1-ha plots from the simulated forest at one specific
+point in time (in the equilibrium). This resulted in 880000 observation
+per model that simulated the 11 species indicated in the protocol (200
+sites, 200 samples per site, 11 species and two diameter thresholds).
+Those models that simulated extra species, the recruitment of those
+extra species were aggregated in a category others. The simulated data
+per model and the observations used in this study are provided as a
+supplementary information OR IN A REPOSITORY? available here(XXX).
+
+```{r map, echo=FALSE, fig.cap="The 200 sites with the dominant tree species in the observed recruitment at 10 cm threshold.", out.width='80%'}
+knitr::include_graphics(here::here("figures", "theMap.png"))
+```
+
+## Analysis
+
+The simulation results were analyzed in regard of their regeneration
+levels, species composition and mortality in regeneration. The observed
+data was modeled using a generalized additive model with a negative
+binomial distribution to better understand the trend and regeneration
+levels of the observed regeneration and the simulation results per
+model. The regeneration levels were evaluated against the stand basal
+area (the basal area of all the trees in each sample) as a proxy of
+light availability for regeneration, the degree-days as a proxy for
+temperature and the water balance as a proxy for soil moisture. The mean
+recruited values were calculated as the mean across samples per site and
+when the regeneration levels were evaluated across the total basal area
+and climate gradients the sites.
+
+We evaluated the species composition across models and in relation to
 the observed data by calculating the Shannon index that defines the
-relative frequency of species referring to basal area (BA) (Equation
-@ref(eq:shannon))
+relative frequency of species referring to basal area where the higher
+the value, the higher the diversity of species in a particular site and
+sample (Equation \ref{eqpiShannon}).
 
@@
 \begin{equation}
-H_n = - \sum_{i = 1}^{s}{p_{i}ln(p_{i})}(\#eq:shannon)
+H_n = - \sum_{i = 1}^{s}{p_{i}ln(p_{i})}\label{eqpiShannon}
 \end{equation}
@@
 \begin{equation}
-p_{n,i} = \frac{BA_{r_{n,i}}}{BA_{total_{n}}}(\#eq:piShannon)
+p_{n,i} = \frac{r_{BA_{n,i}}}{Total_{BA_{n}}} \label{eqpiShannon2}
 \end{equation}
 ```
-where, $s$ is the total number of species in the sample $n$ (11 for the
+where, $s$ is the total number of species simulated $n$ (11 for the
 models that simulated all the species and a varied number for those
-models that did not consider all the species (see Table
-@ref(tab:tableModels) for more details in the simulated species);
+models that did not consider all the species, see Table
+\@ref(tab:tableModels) for more details in the simulated species);
 $p_{i}$ is the proportion of the species $i$ calculated as the recruited
-basal area ($BA_{r_{n,i}}$) for that species $i$ over the total basal
-area ($BA_{total_{n}}$) of that sample $n$.
+basal area ($r_{BA_{n,i}}$) for that species $i$ over the stand basal
+area ($Total_{BA_{n}}$) of the sample $n$.
 
-Main hypothesis:
+In order to understand the impact of adult species composition on
+recruitment across climatic gradients we focused on the five main tree
+species: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus
+sylvestris*, and *Quercus* spp. For these species we calculated the mean
+recruited basal area share per site as the mean value across the
+available samples per site:
 
-The recruitment species richness is driven by the simulated adult trees
-species richness and this is defined on how models represent individual
-species recruitment per site with respect of adult proportion
+```{=tex}
+\begin{equation}
+\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{t,i}}{\sum_{i = 1}^{s}{\bar{r}_{t,i}}} \label{eq:BAshare}
+\end{equation}
+```
+where, $s$ is the total number of species simulated; $\bar{r}_{n,i}$ is
+the mean recruitment of the species $i$ per site $t$ across the
+available samples on site.
 
-Environmental gradients have an impact on the recruitment of the 5 main
-tree species but extremes are not well captured
+There were 35 sites that had a diameter threshold of 10 cm and therefore
+we do not know the actual recruitment rates for those sites at 7 cm; yet
+the models simulated recruitment for the 7 cm threshold on those sites.
+The 35 sites were removed from the simulated data to make a fair
+comparison. Furthermore, when looking at the recruited basal area share
+per species ($\bar{R}_{BA_{share_{i}}}$) we categorized the basal area
+share as zero if the total recruited basal area (
+$\sum_{i = 1}^{s}{\bar{r}_{t,i}}$ ) equal zero and the recruited basal
+area of that species ($\bar{r}_{t,i}$) was also zero.
 
-Models can capture the relative abundance of recruitment of the 5 main
-European tree species but they are not doing a good job for minor
-species
+Mortality at regeneration was assessed based on the difference on
+recruitment between 7 and 10 cm diameter thresholds. We used the Reineke
+self-thinning rule [@reineke1933; @pretzsch2003] as a reference to
+estimate if the different recruitment values were above or bellow the
+expected change between these two diameters. The Reineke self-thinning
+rule is usually calculated from even-aged, single species stands and
+assumes a relationship between the number of stems and the quadratic
+mean diameter in fully stocked pure stands. The value used in our
+comparisons was 1.77 and was calculated following the Equation
+\ref{eqReineke}.
 
-We evaluate the species composition in the recruitment and for all the
-trees present on the site using the relative frequency of species
-referring to basal area with the Shannon index calculated per sample,
-site and model. The higher the value, the higher the diversity of
-species in a particular site and sample.
-
-Mortality in recruitment
-
-Main hypothesis:
-
-the higher regeneration at threshold of 7cm the stronger is the
-regeneration mortality, so we will see a larger decease in stem number
-between recruitment at threshold of 7 cm and 10 cm (as ratio nn7/nn10
-(colname nn710))  In othwr words the number of individuals would
-decrease (Reineke self-thinning rule: This is usually calculated from
-even age, single species stands. The models could also include the
-effects of canopy shading and species competition)
-
-Species composition it is also reduced between 7 and 10 cm
-
-Comments on the Reineke rule
-
-the number of individuals would decrease (Reineke self-thinning rule)
-from 7 to 10cm
-
-assumes a relationship between the number of stems and the quadratic
-mean diameter in fully stocked pure stands.
-
-This rule is extensively used for management purposes, but it has been
-initially calibrated for pure, even-aged stands for relatively few
-species.
-
-This is usually calculated from even age, single species stands. The
-models could also include the effects of canopy shading and species
-competition.
-
-Explanations:
-
-$$Reineke exp rk=-1.605: (7/10)^{(-1.605)} = 1.77$$
-
+```{=tex}
+\begin{equation}
+Reineke = (7/10)^{(-1.605)} \label{eqReineke}
+\end{equation}
+```
 # Results
@@
 ```{r over, echo=FALSE, fig.cap="Mean recruitment values across all samples for the 200 sites.
-The red dotted lines shows the 2.5 and 97.5 percentiles for the 7 cm diameter threshold.", out.width='100%'}
+The red dotted lines shows the 2.5 and 97.5 percentiles for the 7 cm diameter threshold in the observed data.", out.width='100%'}
         knitr::include_graphics(here::here("figures", "recOverAll710.png"))
@@
 climatic gradients, where they feature varying effects (Figure
-\@ref(fig:nvTrend710BA) and FigureS 1-4).
+\@ref(fig:nvTrend710BA) and supplementary figures 1 and 2).
 
@@
 moisture and temperature gradients and there was either a clear common
-pattern across models. Looking at the soil moisture gradients (FigureS
-2), we can see that a group of models decreases recruitment with
-increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS
-and LandClim), while few models had the opposite trend (ForCEEPs(f),
-FORMIND and LPJ-GUESS). Across temperature gradients all the models that
-had a decrease in recruitment with increasing soil moisture showed the
-opposite trend (more recruitment with increasing temperature) except for
-4C (FigureS 3). The rest of the models did not have a clear pattern
-across the temperature gradient.
+pattern across models. Looking at the soil moisture gradients
+(supplementary figure 1), we can see that a group of models decreases
+recruitment with increasing soil moisture (4C, ForClim 1, ForClim11,
+SIBYLA, xComp, PICUS and LandClim), while few models had the opposite
+trend (ForCEEPs(f), FORMIND and LPJ-GUESS). Across temperature gradients
+all the models that had a decrease in recruitment with increasing soil
+moisture showed the opposite trend (more recruitment with increasing
+temperature) except for 4C (supplementary figure 2). The rest of the
+models did not have a clear pattern across the temperature gradient.
+
+Both observations and simulated data had no recruitment in some samples
+and sites (supplementary table 1). The observed data had over 3% of the
+samples with no recruitment. Only three models had larger proportion of
+no recruitment (4C, Landis II and aDVM2). Three models always simulate
+recruitment for both 7 and 10 cm threshold (TreeMig and LPJ-GUESS), an
+the rest of the models have a very low percentage of samples with no
+recruitment (0.01% - 1.31%).
 
@@
 although they consistently underestimate its variability across sites.
-Four models overestimate recruitment diversity:4C, PICUS, LANDIS-II and
+Four models overestimate recruitment diversity: 4C, PICUS, LANDIS-II and
 LPJ-GUESS, the latter mostly for the 10 cm diameter threshold. The model
@@
 artifact of the sample strategy and calculating the Shannon index based
-in BA as for a same sample recruitment at 10cm will include larger trees
-with larger BA and if those are from different species, the Shannon
-index will be larger.
+in basal area as for a same sample recruitment at 10cm will include
+larger trees with larger basal area and if those are from different
+species, the Shannon index will be larger.
 
@@
 
-Most of the models are capturing well or over-predicting the species
-richness at the stand level (Figure \@ref(fig:combinedRichness) A2 and
-FigureS 6). The models over predicting species richness at the stand
-level are ForClim1, ForClim 11, FORMIND and TreeMig (Figure
-\@ref(fig:combinedRichness) A1). However only few models under-predict
-the species richness at the stand level (Figure
-\@ref(fig:combinedRichness) A3) LPJ-GUESS, ForCEEPS (f), xComp, iLand
-and 4C. Species diversity is lower in the recruitment compared to the
-stand level, for both observed and simulated data, and for both
-recruitment thresholds, i.e. 7 and 10 cm (Figure
-\@ref(fig:combinedRichness)B, and FigureS 7). The majority of the models
+Species diversity is lower in the recruitment compared to the stand
+level, for both observed and simulated data, and for both recruitment
+thresholds, i.e. 7 and 10 cm (supplementary figure 3 and 4). Models
+simulating higher diversity at the stand level do not feature higher
+diversity in recruitment (supplementary figure 3 and 4). Most of the
+models are capturing well or over-predicting the species richness at the
+stand level (Figure \@ref(fig:combinedRichness)-A-centered plot and
+supplementary figure 5). The models over predicting species richness at
+the stand level are ForClim1, ForClim 11, FORMIND and TreeMig (Figure
+\@ref(fig:combinedRichness)-A-left plot). However only few models
+under-predict the species richness at the stand level LPJ-GUESS,
+ForCEEPS (f), xComp, iLand and 4C. (Figure
+\@ref(fig:combinedRichness)-A-right plot). The majority of the models
 capture reasonably well the recruitment diversity at lower values of
 diversity, however most of them strongly underestimate species diversity
-in recruitment for higher observed diversity levels. Models simulating
-higher diversity at the stand level do not feature higher diversity in
-recruitment (FigureS 6 and 7). Both models with feedback and without
-feedback from the adult trees (via seed production) underestimate
-species richness in recruitment.
+in recruitment for higher observed diversity levels(Figure
+\@ref(fig:combinedRichness)-B and supplementary figure 6). Both models
+with feedback and without feedback from the adult trees (via seed
+production) underestimate species richness in recruitment compared to
+the observed data.
 
@@
 B) Shows the three trends present across models for recruitment at 7cm threshold.
-The selected representative model per trend group was based on Figure XX and tables available in the supplementary information", out.width='75%'}
+The selected representative model per trend group was based on supplementary figure 5 and 6", out.width='75%'}
 knitr::include_graphics(here::here("figures", "combined_richness.png"))
@@
 along the two major gradients of climate (Figure \@ref(fig:diff) and
-FigureS XX). However, the models differ in the way they represent the
-climate niche of individual main species. First, some models
-overestimated recruitment across the whole gradient for one of the 5
-main tree species (for example, ForCEEPS, ForCEEPS(f),LandClim or
+supplementary figure 7). However, the models differ in the way they
+represent the climate niche of individual main species. First, some
+models overestimated recruitment across the whole gradient for one of
+the 5 main tree species (for example, ForCEEPS, ForCEEPS(f),LandClim or
 LPJ-GUESS), while other models overestimated recruitment for more than
 one species for larger parts of the climate gradient, such as xComp,
-PICUS , SYBILA or LPJ-GUESS (cf. red colors in Figure \@ref(fig:diff)).
+PICUS, SYBILA or LPJ-GUESS (cf. red colors in Figure \@ref(fig:diff)).
 This trends were consistent with those when the regeneration thresholds
-was at 7 and 10cm although the exact changes in recruited basal area
-share changed (FigureS XX).
+was at 7 and 10 cm although the exact changes in recruited basal area
+share changed (supplementary figure 7).
 
@@
 environmental gradient, including ForCEEPS(f) and FORMIND that did not
-have regeneration in most of the sites for this species. Most of the
+have regeneration in most of the sites for these species. Most of the
 models over-estimated *Picea abies* recruitment share on colder sites.
@@
 these two species in most of the climatic gradient although at basal
-area share. Five models (4C, FORMIND, SYBILA, xComp, LandClim) had
+area share. Five models (4C, FORMIND, SYBILA, xComp, and LandClim) had
 almost no generation or did not regenerate *Pinus sylvestris* and
@@
 
-It is important to keep in mind that the values shown in Figure
-\@ref(fig:diff) are the means of all samples per site, and the mean
-across the sites in each environmental bin. In this figure we are seeing
-together two types of zeros (represented with a lack of dots in the
-corresponding tile), the first type of zero are those sites where the
-recruited basal area of the species was zero and the stand recruited
-basal area for all species was also zero, those were considered as zero
-BA share; the second type of zeros are those sites where there was no
-recruitment for that species but there was recruitment for other
-species.
+It is important to keep in mind two aspects from Figure \@ref(fig:diff),
+first the values shown are the means of all samples per site, and the
+mean across the sites in each environmental bin. Second, in this figure
+we are seeing together two types of zeros (represented with a lack of
+dots in the corresponding tile), the first type of zero are those sites
+where the recruited basal area of the species was zero and the stand
+recruited basal area for all species was also zero, those were
+considered as zero basal area share; the second type of zeros are those
+sites where there was no recruitment for that species but there was
+recruitment for other species.
 
@@
 The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species.
-The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the observed data did not becuase both recruitment and stand level recruitments were equal to zero.", out.width='100%'}
+The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area.", out.width='100%', out.height= '90%'}
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
@@
 indicates strong compensation for the initial overestimation of
-recruitment in some models. However, some of the models simulate a
+recruitment in some models. However, some of the models simulated a
 mortality rate in recruitment that is broadly compatible with the
@@
 
-Models with higher mortality also show higher tree recruitment at 7cm
-(Figure \@ref(fig:)\@ref(fig:mortality13)). On the other hand, most of
-the models with a low variation in mortality between 7 and 10 cm
-(ForCEEPS, ForCLim 1, ForClim 11 and LandClim) also have lower
-overestimation in recruitment at 7 cm (Figure \@ref(fig:mortality11) and
-FigureS XX). Some of the models that start at a lower number of
-individuals (7cm thershold) have a lower decrease between 7 and 10cm dbh
-(ratio 7-10cm); in these models the species competition and
+Models with higher mortality also showed higher tree recruitment at 7cm
+(Figure \@ref(fig:mortality13)). On the other hand, most of the models
+with a low variation in mortality between 7 and 10 cm (ForCEEPS, ForClim
+1, ForClim 11 and LandClim) also have lower overestimation in
+recruitment at 7 cm (Figure \@ref(fig:mortality11) and supplementary
+figure 8). Some of the models that start at a lower number of
+individuals (7 cm threshold) have a lower decrease between 7 and 10 cm
+dbh (ratio 7-10 cm); in these models the species competition and
 self-thinning is either not that large or it happened before the species
-reached 7cm. However, this pattern it is not consistent across models
-(FigureS XX). There is an effect of the number of individuals at 7cm on
-the mortality rate. Most of the models show a positive effect
-(increasing recruitment at 7cm shows increases the mortality between 7
-and 10cm), but some models show a negative effect, however the trend it
-is not very high and not always significant when we look at the slope of
-the linear trends per model (FigureS XX and tableS XX).
+reached 7 cm. However, this pattern it is not consistent across models
+(supplementary figure 9). There is an effect of the number of
+individuals at 7cm on the mortality rate. Most of the models show a
+positive effect (increasing recruitment at 7cm shows increases the
+mortality between 7 and 10cm), but some models show a negative effect,
+however the trend it is not very high and not always significant when we
+look at the slope of the linear trends per model (supplementary figure 9
+and table 2).
 
@@
 ```{r mortality13, echo=FALSE, fig.cap="Ratio of tree recruitment between 7 and 10 cm recruitment treshold and the overestimation proportion (simulated - observed / observed) for the mean recruitment trees per model across sites and samples.
-The horizontal dotted lines show a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions.
+The horizontal dotted lines show a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke self-thinning ratio under even-aged conditions.
 The vertical line represents the 0 overestimation proportion", out.width='100%'}
@@
 
--   Implication of models that produce very high initial densities
+**General**
 
--   Implications of models that have lower mortality
+Tree recruitment is a fundamental process in forest dynamics, in
+particular with increasing disturbance under climate change.
 
@@
 
+```{=html}
+<!-- -->
+```
+-   the approach: the idea of confronting the simulated recruitment
+    along environmental gradients with data from such gradients.
+
+-   the idea of not doing a calibration a recruitment model based on the
+    EuFoRIa data, and I explain why it is more insightful to use these
+    data for model benchmarking: we can learn more
+
+**Regeneration levels**
+
+-   Implication of models that produce very high initial densities
+
 -   Range shifts retractions in species expansions because of climate
@@
     properly recruitment rates in the environmental gradient .
+
+-   importance of light availability for regeneration and the proxy of
+    stand basal area to define that . Can we see a expected pattern that
+    shows how recruitment of the different species is becoming sparser
+    and drops out completely as a function of a proxy of light
+    availability.
+
+**Regeneration species composition**
 
@@
 
+-   Why the models over predicting species richness at the stand level
+    are all the "ForClim family" and FORMIND ?
+
+-   Most models maintain or even increase in recruitment diversity
+    between the 7 and 10 cm thresholds, and the same is visible from the
+    observed data, however the differences are not significant. EXPLAIN
+    WHY This cannot be species richness (as there are no new species
+    popping between 7 and 10 cm DBH). I'd also expect the number of
+    species to be lower at 10 cm than at 7 cm (because some will die off
+    due to insufficient light etc.), but this is not reflected in H'
+    because species with zero abundance are ignored (ln[0] is not
+    defined). So a strongly increasing evenness must be driving this
+    increase of H'." This could be because we are comparing 7 and 10cm
+    from same samples, and at 10cm you might include more trees and at
+    7cm This calls for further research and another paper?
+
+**Regeneration mortality**
+
+-   Implications of models that have lower mortality in regeneration but
+    large recruitment
+
+-   We expected that the higher regeneration at 7 cm the stronger the
+    regeneration mortality (decrease in stem number between regeneration
+    at 7cm and regeneration at 10cm) . Not the case, why, implications
+
+-   Models below the Reineke line have lower mortality than expected by
+    the self-thinning rule, which may point to facilitation, or simply a
+    higher mortality rate before the trees reached 7 cm diameter. But
+    then these models would need to underestimate recruitment levels
+    (Fig. 1), no? But this is not what we see... not sure how to
+    explain.
+
+-   There might be a difference between models including stochastic
+    processes in recruitment and those which do not, but perhaps the
+    reality implies higher mortality rates in the adult component
+    (enhance the share of less shade tolerant species?)
+
+-   Comments on the Reineke rule
+
+**Methodological aspects**
+
+-   Difference in observations number between observed data (2-4 samples
+    per site) and the simulated data (200 samples per site) and the
+    implication on variability.
+
+-   Assumption equilibrium it is expected overestimation because we do
+    not know the status of the observed data more certainly not in
+    equilibrium etc
+
+-   methodological limitations compare mortality with the self thinning
+    line etc Values below the 1.0 line show sites where there is a
+    higher number of 10 cm trees than in 7 cm - this is an artifact of
+    the method - we expect space for time substitution, but in reality
+    we just have one measurement.
+
+-    choice of an ingrowth limit of 10 /7cm DBH. if the model gets
+    overall species composition right, it must also get ingrowth of
+    these species rather large trees right. a better test would be for
+    the ingrowth limit to be lower. we chose the highest calipering
+    limit of 10 to maximize data availability
+
+-   methodology number of samples in observations vs simulations per
+    site
+
+**Zero inflated**
+
+-   Observed data has much more sites without any recruitment (zero
+    inflated data). Many models seem to predict recruitment in all
+    sites. The zeros in the empirical data come from no recruitment and
+    from thresholds larger than 10cm for some sites. Is this a
+    consequence of assuming that the observed data is also in
+    equilibrium. Lack of recruitment in sites an indication of
+    "non-equilibrium" due to management history?
+
+-   the reason why some models have zero regeneration at very low BA
+    values may be that they never actually simulate such low BA values
+    at the one ha level (which was the sampling unit if I recall
+    correctly). If so, this should be pointed out as well, I think.
+
+**Other aspects**
+
 -   We do not know what happen before 7cm in the observed data and does
@@
 
--   Methodological aspects: difference in observations number between
-    observed data (2-4 samples per site) and the simulated data (200
-    samples per site) and the implication on variability. Assumption
-    equilibrium it is expected overestimation because we do not know the
-    status of the observed data more certainly not in equilibrium etc
+**Models structure and recomendations**
 
--   Observed data has much more sites without any recruitment (zero
-    inflated data). Many models seem to predict recruitment in all
-    sites. The zeros in the emprical data come from no recruitment and
-    from tresholds larger than 10cm for some sites. Is this a
-    consequence of assuming that the observed data is also in
-    equilibrium. Lack of recruitment in sites an indication of
-    "non-equilibrium" due to management history?
+-   how it relates to the results
 
--   Why the models over predicting species richness at the stand level
-    are all the "ForClim family" and FORMIND ?
-
--   Also BTW, the reason why some models have zero regeneration at very
-    low BA values may be that they never actually simulate such low BA
-    values at the one ha level (which was the sampling unit if I recall
-    correctly). If so, this should be pointed out as well, I think.
-
--   \# H1: the higher regen7 the stronger the regen mortality
-
-    \# decrease in stem number between regen7 and regen10
-
-    \# as ratio nn7/nn10 (colname nn710)
-
-    \# reference lines
-
-    \# -ratio 1 = no decrease from 7 -\> 10
-
-    \# -ratio for selfthinning under even aged conditions
-
-    \# with Reineke exp rk=-1.605: (7/10)\^(-1.605) = 1.77
-
--   Models below the Reineke line have lower mortality than expected by
-    the self-thinning
-    rule, which may point to facilitation, or simply a higher mortality
-    rate before the trees reached 7 cm diameter. But then these models
-    would need to underestimate recruitment levels (Fig. 1), no? But
-    this is not what we see... not sure how to explain.
-
--   methodological limitations compare mortality with the self thinning
-    line etc Values below the 1.0 line show sites where there is a
-    higher number of 10 cm trees than in 7 cm - this is an
-    artifact of the method - we expect space for time substitution, but
-    in reality
-    we just have one measurement (Figure 7).
-
--   There might be a difference between models including stochastic
-    processes in recruitment and those which do not, but perhaps the
-    reality implies higher mortality rates in the adult component
-    (enhance the share of less shade tolerant species?) [to the
-    discussion] ((Figure \@ref(fig:mortality11) and
-    \@ref(fig:mortality13)).
-
--    Most models maintain or even increase in recruitment diversity
-    between the 7 and 10 cm thresholds, and the same is visible from the
-    observed data, however the differences are not significant.
-    **EXPLAIN WHY  "Can we understand this (I know this is for the
-    Discussion, but still...)?? This cannot be species richness (as
-    there are no new species popping between 7 and 10 cm DBH). I'd also
-    expect the number of species to be lower at 10 cm than at 7 cm
-    (because some will die off due to insufficient light etc.), but this
-    is not reflected in H' because species with zero abundance are
-    ignored (ln[0] is not defined). So a strongly increasing evenness
-    must be driving this increase of H'."** This could be because we are
-    comparing 7 and 10cm from same samples, and at 10cm you might
-    include more trees and at 7cm This calls for further research and
-    anotehr paper :D
+-   future model development recommendations
 
@@
 ```
-\`\`\`
-
-# Supplementary

### 2022-11-19 code/figures.R
added figure with the map

@@ -1234,8 +1234,8 @@
       ggplot2::geom_point(alpha = .8) +
       ggplot2::geom_hline(ggplot2::aes(yintercept = 1), colour = "#990000", linetype = "dashed") +
       ggplot2::geom_hline(ggplot2::aes(yintercept = 1.77), colour = "#990000", linetype = "dashed") +
-      ggplot2::xlim(c(-5, 15)) +
-      ggplot2::geom_vline(xintercept = 0, linetype = "dotted", color = "red") +
+      ggplot2::xlim(c(-5, 10)) +
+      ggplot2::geom_vline(xintercept = 0, colour = "#990000", linetype = "dashed") +
       ggplot2::scale_colour_manual(values = values_color) +
       ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 0),
                      strip.text.y = ggplot2::element_text(angle = 0),
@@ -1428,8 +1428,70 @@
                     dpi = 300, units = "cm", device = 'png') 
     
     
+# Map -----
+
+    obsSpp <- outputsDF[outputsDF$model == "Observed", ]
     
-   
+    standSpp <- obsSpp |> dplyr::group_by(site) |> dplyr::top_n(1, ba)
+    standSpp <- standSpp[, c("site", "species")]
+    colnames(standSpp) <- c("site", "stand")
+    standSpp <- standSpp |> 
+                 dplyr::group_by(site) |>  
+                 dplyr::filter(dplyr::row_number() == 1)
     
     
-   
+    recruitSpp <- obsSpp |> dplyr::group_by(site, dbh) |> dplyr::top_n(1, r.ba)
+    recruitSpp <- recruitSpp[, c("site", "species", "dbh")]
+    colnames(recruitSpp) <- c("site", "recruitment", "dbh")
+    recruitSpp7 <- recruitSpp[recruitSpp$dbh == 7,]
+    recruitSpp10 <- recruitSpp[recruitSpp$dbh == 10,]
+    
+    recruitSpp7 <- recruitSpp7 |> 
+                  dplyr::group_by(site) |>  
+                  dplyr::filter(dplyr::row_number() == 1)
+    recruitSpp10 <- recruitSpp10 |> 
+      dplyr::group_by(site) |>  
+      dplyr::filter(dplyr::row_number() == 1)
+    
+    # Add location 
+    location <- data.table::fread("data/coords_blurred_dt4326.csv")
+    
[…]

### 2022-11-21 manuscript/manuscriptRW.Rmd
updates in dicusssion (still very messy)

@@
 output:
+  word_document:
+    reference_docx: template_paper.docx
   bookdown::pdf_document2:
@@
     toc: no
-  word_document:
-    reference_docx: template_paper.docx
 mainfont: Arial
@@
   library(bibtex)
+  options(tinytex.verbose = TRUE)
   #knitr::opts_chunk$set(echo = TRUE, fig.pos = "h", out.extra = "")
@@
 growth and survival; these processes are all difficult to assess and
-some of them are scarcely understood and thus appear highly stochastic.
-Finding the right level of complexity in the definition of the factors
-leading to tree recruitment is challenging and models of forest dynamics
-have used different approaches [@bugmann2022]. Overall, models are
-needed to (1) synthesize existing empirical data and explore their
-consequences, and (2) assess future tree recruitment e.g. in the context
-of global change scenarios. Currently, regeneration processes in dynamic
-forest models are handled in a multitude of ways [@konig_tree_2022;
-@bugmann2022], from entirely ignoring tree regeneration (as done in
-classical forest growth models, e.g. @pretzsch2002), across the use of a
-few simple environmental filters, as done in most forest \"gap\" models
-(e.g., @shuman2011) and Dynamic Global Vegetation Models (e.g.,
-@smith2001; @hickler2012), to complex approaches that incorporate local
-feedback from the canopy, multiple ecological processes and often also
-small time steps (e.g., @wehrli2006; @seidl2012), or field-based
-statistical parameterizations, which however are not easy to extrapolate
-in space and time (e.g., @ribbens1994).
+some of them are scarcely understood and thus appear highly stochastic
+[@bogdziewicz2021]. Finding the right level of complexity in the
+definition of the factors leading to tree recruitment is challenging and
+models of forest dynamics have used different approaches [@bugmann2022].
+Overall, models are needed to (1) synthesize existing empirical data and
+explore their consequences, and (2) assess future tree recruitment e.g.
+in the context of global change scenarios. Currently, regeneration
+processes in dynamic forest models are handled in a multitude of ways
+[@konig_tree_2022; @bugmann2022], from entirely ignoring tree
+regeneration (as done in classical forest growth models, e.g.
+@pretzsch2002), across the use of a few simple environmental filters, as
+done in most forest \"gap\" models (e.g., @shuman2011) and Dynamic
+Global Vegetation Models (e.g., @smith2001; @hickler2012), to complex
+approaches that incorporate local feedback from the canopy, multiple
+ecological processes and often also small time steps (e.g., @wehrli2006;
+@seidl2012), or field-based statistical parameterizations, which however
+are not easy to extrapolate in space and time (e.g., @ribbens1994).
 
@@
 represent tree recruitment we are obtaining results that faced several
-problems related to over-recruitment, needed excess of mortality at
-early stages and inadecuate species composition in regeneration which
-have strong impacts on forest dynamics. Some model outputs are
+problems and have strong impacts on forest dynamic such as
+over-recruitment, needed excess of mortality at early stages and
+inadecuate species composition in regeneration. Some model outputs are
 over-recruiting which subsequently must die primarily by competition in
@@
 recruitment into a larger size class (e.g., @wehrli2006 ). Most of these
-undesired behaviors in tree recruitment are often not published as such,
-but in the vast majority of studies they are corrected (e.g.,
-@rasche2011 , where for most study sites regeneration was simply
-switched off). Thus, a strong focus on the modeling of tree regeneration
-processes is highly important and long overdue [@price] if we are to
-make reliable projections of future forest dynamics, i.e. when the
-models need to be operated in extrapolation mode, but also from a
-fundamental ecological point of view (increasing systems understanding).
+undesired behaviors in tree recruitment are often not published, but in
+the vast majority of studies they are corrected (e.g., @rasche2011,
+where for most study sites regeneration was simply switched off). Thus,
+a strong focus on the modeling of tree regeneration processes is highly
+important and long overdue [@price] if we are to make reliable
+projections of future forest dynamics, i.e. when the models need to be
+operated in extrapolation mode, but also from a fundamental ecological
+point of view (increasing systems understanding).
+
+Tree recruitment is understudied in field studies and modelling
+[@hanburybrown_forest_2022], as little data has been available and it is
+often very fragmented, i.e., not in one harmonized data set. On the one
+hand, monitoring on permanent sample plots (such as national forest
+inventories) often measure tree recruitment [@zell_modeling_2019]
+however often with a design that captures the regeneration intensity
+after the stand initiation phase on productive forests [@hallsby2015].
+Data from permanent plots, although available across regions, it is also
+highly heterogeneous in design making their usability more difficult
+[@käber2021]. On the other hand, targeted studies to measure tree
+recruitment on experimental sites are often limited in space, species or
+the relevant processes leading to regeneration that are targeted
+[@berdanier2016; @collet2006].
 
@@
 continental-scale multi-species harmonized data set on tree recruitment
-[@käber2022]. Tree recruitment is understudied in field studies and
-modelling [@hanburybrown_forest_2022], because little data has been
-available and data was very fragmented, i.e., not in one harmonized data
-set. In this study we relied on models with various 'philosophies',
-including models operating at the stand to the global scale. Because of
-the large variability in recruitment patterns in nature and the large
-amount of factor driving recruitment, including some which are not
-included in most models, such as deer browsing, we do not aim at a
-detailed statistical evaluation of each model. Instead, we aim at
-evaluating general recruitment patterns, magnitudes and benchmark the
-regeneration niche of multiple species based on empirical data along a
-wide environmental gradient (temperature, moisture, and light
+[@käber2022]. In this study we relied on models with various
+'philosophies', including models operating at the stand to the global
+scale. Because of the large variability in recruitment patterns in
+nature and the large amount of factor driving recruitment, including
+some which are not included in most models, such as deer browsing, we do
+not aim at a detailed statistical evaluation of each model. Instead, we
+aim at evaluating general recruitment patterns, magnitudes and benchmark
+the regeneration niche of multiple species based on empirical data along
+a wide environmental gradient (temperature, moisture, and light
 availability). More specifically in this study we wanted to answer the
@@
 presented in the supplementary materials (SHALL WE DO THIS?).
+
+Two models, ForClim and ForCEEPs used two alternative model versions.
+HERE **FORCLIM** WRITE 2 SENTENCES ON THE DIFFERENCES BETWEEN FORCLIM 1
+AND 11 AND WHY IT IS RELEVANT TO CONSIDER THE TWO MODEL VERSIONS. HERE
+**FORCEEPS** WRITE 2 SENTENCES ON THE DIFFERENCES BETWEEN FORCEEPS AND
+FORCEEPS(F) AND WHY IT IS RELEVANT TO CONSIDER THE TWO MODEL VERSIONS.
 
@@
 
-**General**
+Tree recruitment is a fundamental process in forest dynamics, in
+particular with increasing disturbance under climate change . Correctly
+capturing regeneration in forest dynamics is fundamental to evaluate for
+example post-disturbance dynamics as this will define the status of the
+forests in the first 15-20 years [@seidl_post-disturbance_2022]. If this
+initial state it is not well captured then we can not properly assess
+aspects such as the forest resilience, or changes in carbon capture. In
+this study, for the first time, projections of tree recruitment done by
+multitude models of forest dynamics were confronted with observed data
+across a large environmental gradients. The observed data that we used
+[@käber2022] is exceptional in this regard, and that's why it was used
+for model benchmarking rather than calibration.
 
-Tree recruitment is a fundamental process in forest dynamics, in
-particular with increasing disturbance under climate change.
+Adequate tree recruitment rates define the structure and composition of
+forests. Most of the models of forest dynamics tended to over-recruit
+trees in all the sites. Excess in recruitment has to lead to a large
+mortality in later stages. Capture properly recruitment rates in the
+environmental gradient might impact future forests. If not capture well
+it is going to be difficult to for models to correctly project range
+shifts retractions in species expansions because of climate change might
+not also be properly represented if we do not capture properly
+recruitment rates in the environmental gradient.
 
--   The problem of correctly capturing post-disturbance recovery, how
-    seed availability and climate conditions define the 15-20 years
-    status of a forest. If this is not well capture by the models we can
-    not really assess the impacts of disturbances on forests.
+Regeneration levels define the success of future forests, but it is
+equally important to correctly identify un-stocked forest areas
+[@rammer2021]. Tree recruitment is often characterized by an excess
+number of zero counts [@fortin2007]. In this study some of the models
+did not have lack of recruitment or their proportion was very low. This
+substantial difference with the observed data might be a consequence of
+assuming that the observed data is also in equilibrium, where lack of
+recruitment in sites could also be an indication of "non-equilibrium"
+due to management history.
 
-```{=html}
-<!-- -->
-```
--   the approach: the idea of confronting the simulated recruitment
-    along environmental gradients with data from such gradients.
-
--   the idea of not doing a calibration a recruitment model based on the
-    EuFoRIa data, and I explain why it is more insightful to use these
-    data for model benchmarking: we can learn more
-
-**Regeneration levels**
-
--   Implication of models that produce very high initial densities
-
--   Range shifts retractions in species expansions because of climate
-    change might not also be properly represented if we do not capture
-    properly recruitment rates in the environmental gradient .
-
--   importance of light availability for regeneration and the proxy of
-    stand basal area to define that . Can we see a expected pattern that
-    shows how recruitment of the different species is becoming sparser
-    and drops out completely as a function of a proxy of light
+Light availability is an important process defining tree recruitment
+[@collet2006; @berdanier2016]. However it is difficult to have measured
+variables defining light availability for the regeneration. In this
+study we used the total stand basal area as a proxy for light
+availability. There was differences in the basal area ranges, some of
+the models had stand basal area lower than the observe data but there
+was also models that did have more stocked stands that the observed data
+and therefore it was not possible to evaluate the recruitment for lower
+density stands. It was also not possible to observed the expected
+pattern showing how recruitment of the different species is becoming
+sparser and drops out completely as a function of a proxy of light
 availability.
 
-**Regeneration species composition**
-
--   If the composition is not well captured then we can not assess the
-    functional composition of the forest appropriately, e.g. their
-    sensitive to drought or resilience to new disturbances events.
-
--   Species richness variations might be explained by the way each model
-    considers the recruitment (e.g. in the absence of adults recruitment
-    can be linked to either seed rain form elsewhere -dispersal, or to
-    sees/seedling bank) Differences might also be expected between
+If the recruitment tree composition is not well captured then we can not
+assess the functional composition of the forest appropriately, e.g.
+their sensitive to drought or resilience to new disturbances events. In
+the models simulation we saw variations on how the main tree species
+were represented in the environmental gradient but also in the in
+overall species richness levels. Most models maintain or even increase
+in recruitment diversity between the 7 and 10 cm thresholds, and the
+same is visible from the observed data, however the differences are not
+significant. Species richness variations might be explained by the way
+each model considers recruitment, e.g. in the absence of adults
+recruitment can be linked to either seed rain from elsewhere -dispersal,
+or to seed/seedling bank). Differences might also be expected between
 cohort and individual trees based models.
@@
     this the consequence for the correct recruitment composition? This
-    is highly dependent on the model structure:
-
-    -   Background seed Yes/no
-
-    -   Stochastic vs deterministic
+    is highly dependent on the model structure:Background seed Yes/no /
+    Stochastic vs deterministic
 
@@
 
--   Most models maintain or even increase in recruitment diversity
-    between the 7 and 10 cm thresholds, and the same is visible from the
-    observed data, however the differences are not significant. EXPLAIN
-    WHY This cannot be species richness (as there are no new species
-    popping between 7 and 10 cm DBH). I'd also expect the number of
-    species to be lower at 10 cm than at 7 cm (because some will die off
-    due to insufficient light etc.), but this is not reflected in H'
-    because species with zero abundance are ignored (ln[0] is not
-    defined). So a strongly increasing evenness must be driving this
-    increase of H'." This could be because we are comparing 7 and 10cm
-    from same samples, and at 10cm you might include more trees and at
-    7cm This calls for further research and another paper?
+We could expect that models that over-recruited trees. will also have a
+large mortality between 7 and 10 cm thresholds. This implies that this
+models have to have the excess of mortality in later stages of the
+forest dynamics. Models below the Reineke line have lower mortality than
+expected by the self-thinning rule, which may point to facilitation, or
+simply a higher mortality rate before the trees reached 7 cm diameter,
+however one would also assume that these models will underestimate
+recruitment which was not the case. Why?. There might be a difference
+between models including stochastic processes in recruitment and those
+which do not, but perhaps the reality implies higher mortality rates in
+the adult component (enhance the share of less shade tolerant species).
 
-**Regeneration mortality**
+There is also the question of how much level of detail one have to give
+to define regeneration processes [@konig_tree_2022; @bugmann2022].
+Models with more detailed and complex recruitment modules, will have
+also increase parametrisation which will lead to lower generalization as
+this values will have to be collected from specific geographical
+regions.
 
--   Implications of models that have lower mortality in regeneration but
-    large recruitment
+-   how it relates to the results
 
--   We expected that the higher regeneration at 7 cm the stronger the
-    regeneration mortality (decrease in stem number between regeneration
-    at 7cm and regeneration at 10cm) . Not the case, why, implications
-
--   Models below the Reineke line have lower mortality than expected by
-    the self-thinning rule, which may point to facilitation, or simply a
-    higher mortality rate before the trees reached 7 cm diameter. But
-    then these models would need to underestimate recruitment levels
-    (Fig. 1), no? But this is not what we see... not sure how to
-    explain.
-
--   There might be a difference between models including stochastic
-    processes in recruitment and those which do not, but perhaps the
-    reality implies higher mortality rates in the adult component
-    (enhance the share of less shade tolerant species?)
-
--   Comments on the Reineke rule
+-   future model development recommendations
 
@@
 
-**Zero inflated**
+-   We use all the models to rpoduce values per 1ha, stand ang global
+    models potentially disregarding the advantages that landscape models
+    present with spatial seed dispersal (TreMig, iLand and LAndClim
+    references). Without spatial context we are limited to the seed
+    influx from the stand itself if the model does not have a background
+    seed input. But even global vegetation models are lacking this
+    dispersal between cells [@hanburybrown_forest_2022] abstraction of
+    space or even the trees themselves. It is important to include
+    spatial considerations on tree recruitmen [@beckage2003]
 
--   Observed data has much more sites without any recruitment (zero
-    inflated data). Many models seem to predict recruitment in all
-    sites. The zeros in the empirical data come from no recruitment and
-    from thresholds larger than 10cm for some sites. Is this a
-    consequence of assuming that the observed data is also in
-    equilibrium. Lack of recruitment in sites an indication of
-    "non-equilibrium" due to management history?
+-   Comparing PFTs with species. We lose the species specific
+    information on with substantial drawbacks [@purves2008]
 
--   the reason why some models have zero regeneration at very low BA
-    values may be that they never actually simulate such low BA values
-    at the one ha level (which was the sampling unit if I recall
-    correctly). If so, this should be pointed out as well, I think.
+-   The loss of species-specific environmental responses through the
+    collation of species communities into PFTs has been widely accepted,
+    as parameterization efforts would otherwise exceed the available
+    means. However, it is doubtful whether the definition of PFTs
+    provides sufficient flexibility for an adequate representation of
+    the ecological processes and the differences between species
 
-**Other aspects**
+-   Recruitment of trees into size classes of ≥4 cm DBH, as in our four
+    data sets, represents a relatively late stage of forest
+    regeneration. It integrates over sometimes long periods since
+    germination and multiple environmental constraints act upon young
+    trees of widely different size. kaber 2021
 
@@
 
-**Models structure and recomendations**
+# Data availability statement
 
--   how it relates to the results
-
--   future model development recommendations
+The code and data that support the findings of this study are openly
+available at zenodo and gitlab (links to be added)
 

### 2022-11-23 manuscript/manuscriptRW.Rmd
keep all the methodological limitation in the discussion with an expansion in other parts

@@
     toc: no
+  word_document:
+    reference_docx: template_paper.docx
 mainfont: Arial
@@
 levels, species composition and mortality in regeneration. The results
-show that .
+show that most of the models of forest dynamics tend to over-estimate
+recruitment rates although this is not always connected to large
+mortality rates in early dynamics stages. The effect of light
+availability on recruitment rates is better captured than temperature or
+soil moisture, however the patterns are not consistent across models.
+Species richness matches quite well compared to the ranges observed in
+the data. Models simulating higher species diversity at the stand level
+do not feature higher diversity in recruitment. *Abies alba* was well
+captured across temperature and soil moisture gradients by most models
+and *Fagus sylvatica* was generally underestimated. *Picea abies*
+recruitment was generally over-estimated on colder sites. HERE TWO
+SENCENTECES DISCUSSION/CONCLUSION. HERE A SO WHAT SENTENCE
 
@@
 @pretzsch2002), across the use of a few simple environmental filters, as
-done in most forest \"gap\" models (e.g., @shuman2011) and Dynamic
-Global Vegetation Models (e.g., @smith2001; @hickler2012), to complex
+done in most forest "gap" models (e.g., @shuman2011) and Dynamic Global
+Vegetation Models (e.g., @smith2001; @hickler2012), to complex
 approaches that incorporate local feedback from the canopy, multiple
@@
 over-recruiting which subsequently must die primarily by competition in
-later stages. This is characteristic of \'classical\' forest gap models
+later stages. This is characteristic of 'classical' forest gap models
 [@gutiérrez2016], at least as long as the simulations comprise a
@@
 specifically when defining processes such as XXXX, XXXX (enumerate
-models). Further details on the models specific formulation related to
-their regeneration modules and an overall summary across models is
-presented in the supplementary materials (SHALL WE DO THIS?).
+models).
 
@@
       kableExtra::column_spec(column = 6, width = "5em") |>
-      kableExtra::column_spec(column = 7, width = "8em") |>
-      kableExtra::column_spec(column = 8, width = "5em") |>
-      kableExtra::column_spec(column = 9, width = "5em") |>
+      kableExtra::column_spec(column = 7, width = "5em") |>
+      kableExtra::column_spec(column = 8, width = "8em") |>
+      kableExtra::column_spec(column = 9, width = "8em") |>
       kableExtra::column_spec(column = 10, width = "5em") |>
@@
 
-Each modeler group was provided with detailed protocol (available in the
-supplementary material) with instruction on how to perform the
+Each modeler group was provided with a detailed protocol (available in
+the supplementary material) with instruction on how to perform the
 simulations regarding input variables such as climate or soil,
@@
 per model and the observations used in this study are provided as a
-supplementary information OR IN A REPOSITORY? available here(XXX).
+supplementary information / repository.
 
@@
 ```{r nvTrend710BA, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the stand basal area.
-(A) y-axis scaled by model; (B) same y-axis across all models.
+(A)same y-axis across all models; (B)y-axis scaled by model.
 The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
-    knitr::include_graphics(c(here::here("figures", "envTrend7Totba.png"),
-                              here::here("figures", "envTrend7Totba_500.png")))
+    knitr::include_graphics(c(here::here("figures", "envTrend7Totba_500.png"),
+                              here::here("figures", "envTrend7Totba.png")))
 ```
@@
 artifact of the sample strategy and calculating the Shannon index based
-in basal area as for a same sample recruitment at 10cm will include
+in basal area as for a same sample recruitment at 10 cm will include
 larger trees with larger basal area and if those are from different
@@
 A) Shows the three trends present across models for the species richness at the stand level.
-B) Shows the three trends present across models for recruitment at 7cm threshold.
-The selected representative model per trend group was based on supplementary figure 5 and 6", out.width='75%'}
+B) Shows the three trends present across models for recruitment at 7 cm threshold.
+The selected representative model per trend group was based on the supplementary figure 5 and 6.", out.width='75%'}
 knitr::include_graphics(here::here("figures", "combined_richness.png"))
@@
 The values shown are the mean of samples per site and across sites in bins, where the range of the climatic water balance (cm) and seasonal degree-day sum (dºC) was divided in 10 bins.
-The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species.
-The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area.", out.width='100%', out.height= '90%'}
+The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species, where a lack of dot represents zero basal area recruitment share.
+The color gradient shows the difference between the simulated and observed ratio of recruited basal area of the species and the total recruited basal area for all the species.", out.width='100%', out.height= '90%'}
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
@@
 mortality rate in recruitment that is broadly compatible with the
-self-thinning rule, i.e. 4C, SIBYLA , LandClim and LANDIS II.
+self-thinning rule, i.e. 4C, SIBYLA, LandClim and LANDIS II.
 
@@
 Tree recruitment is a fundamental process in forest dynamics, in
-particular with increasing disturbance under climate change . Correctly
-capturing regeneration in forest dynamics is fundamental to evaluate for
-example post-disturbance dynamics as this will define the status of the
+particular with increasing disturbance under climate change. Correctly
+capturing regeneration in forest dynamics is fundamental to for example
+evaluate post-disturbance dynamics as this will define the status of the
 forests in the first 15-20 years [@seidl_post-disturbance_2022]. If this
@@
 this study, for the first time, projections of tree recruitment done by
-multitude models of forest dynamics were confronted with observed data
+several models of forest dynamics were confronted with observed data
 across a large environmental gradients. The observed data that we used
-[@käber2022] is exceptional in this regard, and that's why it was used
-for model benchmarking rather than calibration.
+[@käber2022] is exceptional in this regard, and to use it for model
+benchmarking has gave us much more insights than if we have used it for
+calibration.
 
@@
 forests. Most of the models of forest dynamics tended to over-recruit
-trees in all the sites. Excess in recruitment has to lead to a large
-mortality in later stages. Capture properly recruitment rates in the
-environmental gradient might impact future forests. If not capture well
-it is going to be difficult to for models to correctly project range
-shifts retractions in species expansions because of climate change might
-not also be properly represented if we do not capture properly
-recruitment rates in the environmental gradient.
+trees in all the sites. Over estimating recruitment might have
+fundamental implications on the role of regrowth and forest carbon, with
+an overestimation in their capacity for carbon capture for example after
+the impact of disturbances [@pugh2019]. Excess in recruitment also
+requires of larger mortality rates in later stages, changing the
+expected forest densities at early stages and forcing the models to have
+stronger filters. Regeneration levels define the success of future
+forests, but it is equally important to correctly identify un-stocked
+forest areas [@rammer2021]. Tree recruitment is often characterized by
+an excess number of zero counts [@fortin2007]. In this study some of the
+models did not have lack of recruitment or their proportion was very
+low. This substantial difference with the observed data might be a
+consequence of assuming that the observed data is also in equilibrium,
+where lack of recruitment in sites could also be an indication of
+"non-equilibrium" due to management history.
 
-Regeneration levels define the success of future forests, but it is
-equally important to correctly identify un-stocked forest areas
-[@rammer2021]. Tree recruitment is often characterized by an excess
-number of zero counts [@fortin2007]. In this study some of the models
-did not have lack of recruitment or their proportion was very low. This
-substantial difference with the observed data might be a consequence of
-assuming that the observed data is also in equilibrium, where lack of
-recruitment in sites could also be an indication of "non-equilibrium"
-due to management history.
+Light availability effect on tree regeneration have been widely
+documented [@messier1999; @collet2006; @berdanier2016]. However it is
+difficult to have variables defining light availability for the
+regeneration. In this study, we used the total stand basal area as a
+proxy for light availability. There was differences in the basal area
+ranges, some of the models had stand basal areas lower than the observed
+data but there was also models that did have more stocked stands than
+the observed data, therefore it was not possible to evaluate the
+recruitment for lower density stands. Regeneration at low stand
+densities is an interesting aspect to evaluate how well forests are
+recovering e.g. after the creation of gaps due to disturbances. We would
+have expected that the recruitment of the different species would become
+sparser and droped out completely as a function of a proxy of light
+availability, however few models showed this trend indicating that the
+relation between recruitment and light availability it is not yet
+correctly defined.
 
-Light availability is an important process defining tree recruitment
-[@collet2006; @berdanier2016]. However it is difficult to have measured
-variables defining light availability for the regeneration. In this
-study we used the total stand basal area as a proxy for light
-availability. There was differences in the basal area ranges, some of
-the models had stand basal area lower than the observe data but there
-was also models that did have more stocked stands that the observed data
-and therefore it was not possible to evaluate the recruitment for lower
-density stands. It was also not possible to observed the expected
-pattern showing how recruitment of the different species is becoming
-sparser and drops out completely as a function of a proxy of light
-availability.
+When tree composition in recruitment is not well captured, then we can
+not assess the functional composition of the forest appropriately, e.g.
+their sensitive to drought or resilience to new disturbances events
+[@redmond2015; @seidl_post-disturbance_2022]. In the models simulation
+outputs overall species richness levels in recruitment were inside the
+observed ranges, however there were variations on how the main tree
+species were represented in the environmental gradient. Most models
+maintain or even increase in recruitment diversity between the 7 and 10
+cm thresholds, and the same is visible from the observed data, however
+the differences are not significant. Species richness variations might
+be explained by the way each model considers recruitment, e.g. in the
+absence of adults, recruitment can be linked to either seed
+rain-dispersal, or to seed/seedling bank, but also on what processes are
+formulated with stochastic or deterministic approach. Differences might
+also be expected between cohort and individual trees based models
+**how?**. However, our results do not show differences between models
+with an without feedback in recruitment, **why?** The adult forests
+species richness was better captured than the recruitment species
+richness, and only few models over-predicted stand level species
+richness **why these ones?**.
 
-If the recruitment tree composition is not well captured then we can not
-assess the functional composition of the forest appropriately, e.g.
-their sensitive to drought or resilience to new disturbances events. In
-the models simulation we saw variations on how the main tree species
-were represented in the environmental gradient but also in the in
-overall species richness levels. Most models maintain or even increase
-in recruitment diversity between the 7 and 10 cm thresholds, and the
-same is visible from the observed data, however the differences are not
-significant. Species richness variations might be explained by the way
-each model considers recruitment, e.g. in the absence of adults
-recruitment can be linked to either seed rain from elsewhere -dispersal,
-or to seed/seedling bank). Differences might also be expected between
-cohort and individual trees based models.
-
--   Why models with feedback also underestimate species richness in
-    recruitment?
-
--   Is the adult forest the right forest, how is this relevant and is
-    this the consequence for the correct recruitment composition? This
-    is highly dependent on the model structure:Background seed Yes/no /
-    Stochastic vs deterministic
-
--   Why the models over predicting species richness at the stand level
-    are all the "ForClim family" and FORMIND ?
-
-We could expect that models that over-recruited trees. will also have a
-large mortality between 7 and 10 cm thresholds. This implies that this
-models have to have the excess of mortality in later stages of the
-forest dynamics. Models below the Reineke line have lower mortality than
-expected by the self-thinning rule, which may point to facilitation, or
-simply a higher mortality rate before the trees reached 7 cm diameter,
-however one would also assume that these models will underestimate
-recruitment which was not the case. Why?. There might be a difference
-between models including stochastic processes in recruitment and those
-which do not, but perhaps the reality implies higher mortality rates in
-the adult component (enhance the share of less shade tolerant species).
+We could expect that models that over-recruit trees will also have a
+large mortality between 7 and 10 cm thresholds but this was not shown.
+This implies that these models must have the excess of mortality in
+later stages. Higher mortality in the adult component might enhance the
+share of less shade tolerant species for those models. There might be a
+difference between models including stochastic processes in recruitment
+and those which do not ??. There was also a group of models below the
+Reineke line indicating a lower mortality than expected by the
+self-thinning rule, which may point to facilitation, or simply a higher
+mortality rate before the trees reached 7 cm diameter, another
+possibility would have been that these models underestimate recruitment
+which was not the case.
 
@@
 this values will have to be collected from specific geographical
-regions.
-
--   how it relates to the results
-
--   future model development recommendations
-
-**Methodological aspects**
+regions. This is significant relevant when looking at mixed forests and
+the species level parametrization. **How it relates to the results?
+Future model development recommendations?**
 
@@
 
--   Recruitment of trees into size classes of ≥4 cm DBH, as in our four
-    data sets, represents a relatively late stage of forest
+-   Recruitment of trees into size classes of larger 4 cm DBH, as in our
+    four data sets, represents a relatively late stage of forest
     regeneration. It integrates over sometimes long periods since
@@
     dynamics. Recommendations on data collection.
+
+**Conclusion paragraph**
+
+Capture properly recruitment rates in the environmental gradient might
+impact future forests.
 

### 2022-11-23 code/figures.R
zoom map figure 1 and change scale colors

@@ -1475,11 +1475,12 @@
     plotsMap <- ggplot(Europe) +
       ggthemes::theme_map() +
       geom_sf(fill = "transparent" , colour = "lightgrey",  lwd = 0.25) +
-      coord_sf(xlim = c(-25,45), ylim = c(35,60), expand = FALSE) +
+      coord_sf(xlim = c(-25,45), ylim = c(40,55), expand = FALSE) +
       geom_point(data = domSpp10, shape = 20, stroke = FALSE, 
                  mapping = aes(x = X, y = Y, color = recruitment), 
                  alpha = 1, size = 2) +
-      scale_color_manual(values = c("Abies alba" =  "#1B9E77",        
+      scale_color_manual("Dominant species in recruitment",
+                         values = c("Abies alba" =  "#1B9E77",        
                                     "Picea abies" ="#D95F02" ,         
                                     "Fagus sylvatica" = "#7570B3",  
                                     "Pinus sylvestris" = "#E7298A" , 
@@ -1489,7 +1490,8 @@
                                     "Fraxinus excelsior" = "#666666", 
                                     "Tilia cordata" = "#D8B365",       
                                     "Betula spp." = "#F5F5F5",
-                                    "Other spp." = "grey" ) )
+                                    "Other spp." = "grey" ) ) 
+    
     ggplot2::ggsave("figures/theMap.png",
                     plot =   plotsMap,
                     width = 21, height = 12, scale = 0.9,

### 2022-11-23 manuscript/manuscriptRW.Rmd
mainly more changes in the manuscript  discussion section

@@
 ---
-title: "Drivers of tree regeneration remain challenging to be captured by models of forest dynamics"
+title: Drivers of tree regeneration remain challenging to be captured by models of
+  forest dynamics
 author:
@@
 regeneration across 200 sites representing environmental gradients
-across Europe. The simulation results were compared with observed data
+across Europe. The simulation results are compared with observed data
 collected from forest reserves across Europe. In this study we want to
 asses how well are models of forest dynamics capturing regeneration
-levels, species composition and mortality in regeneration. The results
-show that most of the models of forest dynamics tend to over-estimate
-recruitment rates although this is not always connected to large
-mortality rates in early dynamics stages. The effect of light
+levels, species composition and mortality in tree recruitment. The
+results show that most of the models of forest dynamics tend to
+over-estimate recruitment rates although this is not always connected to
+large mortality rates in early dynamics stages. The effect of light
 availability on recruitment rates is better captured than temperature or
@@
 several models of forest dynamics were confronted with observed data
-across a large environmental gradients. The observed data that we used
-[@käber2022] is exceptional in this regard, and to use it for model
-benchmarking has gave us much more insights than if we have used it for
-calibration.
+across a large environmental gradients.
+
+The observed data that we used [@käber2022] is exceptional in this
+regard, and to use it for model benchmarking has gave us insights on
+recruitment levels, species composition and mortality and overall a
+broader understanding than if we have used these data for a model
+calibration. The lower available threshold in the observed data of 7 cm
+provides a limit from which we can understand tree recruitment, which
+might be a rather long period since germination as relevant
+environmental constraint might have already act on those young trees
+[@käber2021]. However recruitment data with lower thresholds are not
+available in a harmonized manner from large environmental gradients. The
+observed data was collected from forest reserves where there has not
+been management for long periods of time, this might help to properly
+assume that these structures are in an equilibrium state. This
+assumption might also justify some of the rather large recruitment rates
+that we are seeing in the models projections compare to the observed
+ones.
 
@@
 models did not have lack of recruitment or their proportion was very
-low. This substantial difference with the observed data might be a
+low. This substantial difference with the observed data might also be a
 consequence of assuming that the observed data is also in equilibrium,
@@
 documented [@messier1999; @collet2006; @berdanier2016]. However it is
-difficult to have variables defining light availability for the
+difficult to collect variables defining light availability on
 regeneration. In this study, we used the total stand basal area as a
@@
 the observed data, therefore it was not possible to evaluate the
-recruitment for lower density stands. Regeneration at low stand
-densities is an interesting aspect to evaluate how well forests are
-recovering e.g. after the creation of gaps due to disturbances. We would
-have expected that the recruitment of the different species would become
-sparser and droped out completely as a function of a proxy of light
-availability, however few models showed this trend indicating that the
-relation between recruitment and light availability it is not yet
-correctly defined.
+recruitment for the extremes of the density ranges in some models.
+Regeneration at low stand densities is interesting to assess how well
+forests are recovering e.g. after the creation of gaps due to
+disturbances. We would have expected that the recruitment of the
+different species would become sparser and droped out completely as a
+function of a proxy of light availability, however few models showed
+this trend indicating that the relation between recruitment and light
+availability it is not yet correctly captured.
 
-When tree composition in recruitment is not well captured, then we can
-not assess the functional composition of the forest appropriately, e.g.
-their sensitive to drought or resilience to new disturbances events
+Correctly capturing tree composition in recruitment help us to assess
+the functional composition of the forest appropriately, e.g. their
+sensitive to drought or resilience to new disturbances events
 [@redmond2015; @seidl_post-disturbance_2022]. In the models simulation
-outputs overall species richness levels in recruitment were inside the
-observed ranges, however there were variations on how the main tree
-species were represented in the environmental gradient. Most models
-maintain or even increase in recruitment diversity between the 7 and 10
-cm thresholds, and the same is visible from the observed data, however
-the differences are not significant. Species richness variations might
-be explained by the way each model considers recruitment, e.g. in the
+outputs overall species richness levels in recruitment was
+underestimated but inside the observed ranges, however there were
+variations on how the main tree species were represented in the
+environmental gradient. The lower species richness in recruitment might
+be a consequence of models running without disturbances, disturbances
+would typically allowed for larger gaps sizes allowing less shade
+tolerant species to grow in larger proportions. Most models maintain or
+even increase in recruitment diversity between the 7 and 10 cm
+thresholds, and the same is visible from the observed data, however the
+differences are not significant. Species richness variations might be
+explained by the way each model considers recruitment, e.g. in the
 absence of adults, recruitment can be linked to either seed
@@
 **how?**. However, our results do not show differences between models
-with an without feedback in recruitment, **why?** The adult forests
+with and without feedback in recruitment, **why?** The adult forests
 species richness was better captured than the recruitment species
@@
 difference between models including stochastic processes in recruitment
-and those which do not ??. There was also a group of models below the
-Reineke line indicating a lower mortality than expected by the
+and those which do not **??**. There was also a group of models below
+the Reineke line indicating a lower mortality than expected by the
 self-thinning rule, which may point to facilitation, or simply a higher
@@
 possibility would have been that these models underestimate recruitment
-which was not the case.
+which was not the case. This might also be partly an artifact of the
+method, as we are expecting space for time substitution, but in reality
+we just have one measurement.
 
@@
 regions. This is significant relevant when looking at mixed forests and
-the species level parametrization. **How it relates to the results?
+the species level parametrization. **How this relates to the results?
 Future model development recommendations?**
 
--   Difference in observations number between observed data (2-4 samples
-    per site) and the simulated data (200 samples per site) and the
-    implication on variability.
-
--   Assumption equilibrium it is expected overestimation because we do
-    not know the status of the observed data more certainly not in
-    equilibrium etc
-
--   methodological limitations compare mortality with the self thinning
-    line etc Values below the 1.0 line show sites where there is a
-    higher number of 10 cm trees than in 7 cm - this is an artifact of
-    the method - we expect space for time substitution, but in reality
-    we just have one measurement.
-
--   choice of an ingrowth limit of 10 /7cm DBH. if the model gets
-    overall species composition right, it must also get ingrowth of
-    these species rather large trees right. a better test would be for
-    the ingrowth limit to be lower. we chose the highest calipering
-    limit of 10 to maximize data availability
-
--   methodology number of samples in observations vs simulations per
-    site
-
--   We use all the models to rpoduce values per 1ha, stand ang global
-    models potentially disregarding the advantages that landscape models
-    present with spatial seed dispersal (TreMig, iLand and LAndClim
-    references). Without spatial context we are limited to the seed
+In these study we considered a wide range of models of forest dynamics,
+from stand to global scale. However, the simulation set up was limited
+to 1 ha recruitment values. It is important to include spatial
+considerations on tree recruitment [@beckage2003] and the lack of it
+might represent a more important disadvantage for landscape level
+models. Without spatial context we are limiting recruitment to the seed
 influx from the stand itself if the model does not have a background
-    seed input. But even global vegetation models are lacking this
-    dispersal between cells [@hanburybrown_forest_2022] abstraction of
-    space or even the trees themselves. It is important to include
-    spatial considerations on tree recruitmen [@beckage2003]
-
--   Comparing PFTs with species. We lose the species specific
-    information on with substantial drawbacks [@purves2008]
-
--   The loss of species-specific environmental responses through the
-    collation of species communities into PFTs has been widely accepted,
-    as parameterization efforts would otherwise exceed the available
-    means. However, it is doubtful whether the definition of PFTs
-    provides sufficient flexibility for an adequate representation of
-    the ecological processes and the differences between species
-
--   Recruitment of trees into size classes of larger 4 cm DBH, as in our
-    four data sets, represents a relatively late stage of forest
-    regeneration. It integrates over sometimes long periods since
-    germination and multiple environmental constraints act upon young
-    trees of widely different size. kaber 2021
-
--   We do not know what happen before 7cm in the observed data and does
-    not make sense to look at that in the simulated data. 7cm is a large
-    sapling, consequences of not having that information to asses forest
-    dynamics. Recommendations on data collection.
+seed input. But even global vegetation models are lacking this dispersal
+between cells where there is an abstraction of the space or even the
+trees themselves [@hanburybrown_forest_2022]. The selected models also
+had limitations on species representation, some models where not able to
+simulate certain species due to lack of species parametrization. We also
+had one model that simulate PFTs providing a less flexible comparison
+the with species values; with the lose of the species specific
+information might also represent substantial ecological
+representativeness drawbacks [@purves2008].
 
@@
 
-Capture properly recruitment rates in the environmental gradient might
-impact future forests.
+Capture properly recruitment rates, species composition and mortality in
+the environmental gradient might impact future forests...
 

### 2022-12-02 manuscript/manuscriptRW.Rmd
additions on the manuscript based on feedback from Harald on 24th November

@@
     reference_docx: template_paper.docx
+  bookdown::word_document2: default
 mainfont: Arial
@@
 
-Regeneration is a key process in forest dynamics. Models of forest
-dynamics have been long lagging an assessment in their capacities to
-properly represent this process. Using 15 models and variants of forest
-dynamics at the stand, landscape and global levels, we simulate the
-regeneration across 200 sites representing environmental gradients
-across Europe. The simulation results are compared with observed data
-collected from forest reserves across Europe. In this study we want to
-asses how well are models of forest dynamics capturing regeneration
-levels, species composition and mortality in tree recruitment. The
-results show that most of the models of forest dynamics tend to
+Tree regeneration is a key process in forest dynamics, particularly in
+the context of forest resilience after disturbance and climate change.
+Models of forest dynamics have been in use since more than 50 years, but
+an assessment of their capacity to properly represent regeneration is
+still lacking. Using 15 models and model variants that were build to
+capture long term (decadal to centennial) forest dynamics at the stand,
+landscape and global levels, we simulate tree regeneration across at 200
+sites representing environmental gradients across Europe. The simulation
+results are benchmarked against data from forest reserves collected at
+these 200 sites. We aim to assess how well the models of forest dynamics
+are capturing the overall abundance, species composition and mortality
+in tree recruitment. The results show that most of the models tend to
 over-estimate recruitment rates although this is not always connected to
-large mortality rates in early dynamics stages. The effect of light
-availability on recruitment rates is better captured than temperature or
-soil moisture, however the patterns are not consistent across models.
-Species richness matches quite well compared to the ranges observed in
-the data. Models simulating higher species diversity at the stand level
-do not feature higher diversity in recruitment. *Abies alba* was well
-captured across temperature and soil moisture gradients by most models
-and *Fagus sylvatica* was generally underestimated. *Picea abies*
-recruitment was generally over-estimated on colder sites. HERE TWO
+large mortality rates in the early stages of population dynamics. In the
+models, the effect of light availability on recruitment rates is
+captured better than the effect of temperature and soil moisture, but
+the patterns are not consistent across models. Species richness matches
+quite well compared to the ranges observed in the data. Models
+simulating higher species diversity at the stand level do not feature
+higher diversity in recruitment. *Abies alba* was well captured across
+temperature and soil moisture gradients by most models, whereas *Fagus
+sylvatica* was generally underestimated. *Picea abies* recruitment was
+generally over-estimated, particularly at colder sites. HERE TWO
 SENCENTECES DISCUSSION/CONCLUSION. HERE A SO WHAT SENTENCE
@@
 
-Forests provide multiple products and services of vital importance.
-Under the pressure of climate change and increasing disturbances, it is
-even more important that we understand how forests will change and
-evaluate their capacity to adapt or be resilient to these changes
+Forests provide a wide range of products and services of vital
+importance. Under the pressure of climate change and increasing
+disturbances, it is becoming ever more important to understand how
+forests structure, composition and functions will change, and evaluate
+forest capacity to adapt or be resilient after disturbance
 [@lindner2010]. A wide range of models of forest dynamics were developed
-over the past decades considering the impacts of climate in forest
+over the past decades considering the impacts of climate on forest
 dynamics [@bugmann2022]. From these studies, it is becoming increasingly
@@
 been dedicated to improving the depiction of tree mortality processes
-[@bugmann2019], tree regeneration remains a less studied process in most
-of these models, while becoming ever more important in the context of
-forest disturbances and forest resilience.
+[@bugmann2019], tree regeneration remains a much less studied process
+that is often represented rather coarsely in these models, which is a
+clear research gap in the context of climate-induced forest disturbances
+and forest resilience.
 
-Tree recruitment depends on multiple processes such as reproduction,
+Tree regeneration depends on multiple processes such as reproduction,
 fruit maturation, seed production, dispersal, germination, juvenile
-growth and survival; these processes are all difficult to assess and
-some of them are scarcely understood and thus appear highly stochastic
-[@bogdziewicz2021]. Finding the right level of complexity in the
-definition of the factors leading to tree recruitment is challenging and
-models of forest dynamics have used different approaches [@bugmann2022].
-Overall, models are needed to (1) synthesize existing empirical data and
-explore their consequences, and (2) assess future tree recruitment e.g.
-in the context of global change scenarios. Currently, regeneration
-processes in dynamic forest models are handled in a multitude of ways
-[@konig_tree_2022; @bugmann2022], from entirely ignoring tree
-regeneration (as done in classical forest growth models, e.g.
-@pretzsch2002), across the use of a few simple environmental filters, as
-done in most forest "gap" models (e.g., @shuman2011) and Dynamic Global
-Vegetation Models (e.g., @smith2001; @hickler2012), to complex
-approaches that incorporate local feedback from the canopy, multiple
-ecological processes and often also small time steps (e.g., @wehrli2006;
-@seidl2012), or field-based statistical parameterizations, which however
-are not easy to extrapolate in space and time (e.g., @ribbens1994).
+growth and survival [@price2001]. All these processes are all difficult
+to assess, and some of them are scarcely understood and thus appear
+highly stochastic [@bogdziewicz2021]. Identifying the appropiate level
+of complexity for the formulation of the key factors that are leading to
+successful tree recruitment is challenging. By tree recruitment, we
+refer to the passing of trees to selected diameter threshold (ingrowth)
+as a definition of successful regeneration [@käber2022]. Overall, models
+are needed to (1) synthesize existing empirical data and explore their
+consequences, and (2) assess future tree recruitment e.g. in the context
+of global change scenarios. Currently, regeneration processes in dynamic
+forest models are handled in a multitude of ways [@konig_tree_2022;
+@bugmann2022], from entirely ignoring tree regeneration (as done in
+classical forest growth models, e.g. @pretzsch2002), across the use of a
+few simple environmental filters, as done in most forest "gap" models
+(e.g., @shuman2011) and Dynamic Global Vegetation Models (e.g.,
+@smith2001; @hickler2012), to complex approaches that incorporate local
+feedback from the canopy, multiple ecological processes and often also
+small time steps (e.g., @wehrli2006; @seidl2012), or field-based
+statistical parameterizations, which however are not easy to extrapolate
+in space and time (e.g., @ribbens1994).
 
-With the current strategies that models of forest dynamics are using to
-represent tree recruitment we are obtaining results that faced several
-problems and have strong impacts on forest dynamic such as
-over-recruitment, needed excess of mortality at early stages and
-inadecuate species composition in regeneration. Some model outputs are
-over-recruiting which subsequently must die primarily by competition in
-later stages. This is characteristic of 'classical' forest gap models
-[@gutiérrez2016], at least as long as the simulations comprise a
-multitude of species. Other models use a model output to correct
-recruitment levels for example by routinely calibrate them against the
-predicted canopy species composition affecting as well the species
-dominance.
+Given the current strategies that are used in models of forest dynamics
+to represent tree recruitment, model behavior is facing several problems
+and that can have strong impacts on forest dynamic, such as very high
+levels of recruitment, which then necessitate excess mortality to early
+stages to arrive at correct stand structure and composition. Also, often
+inadequately high species diversity in regeneration is simulated. This
+is characteristic of \'classical\' forest gap models [@gutiérrez2016],
+at least as long as the simulations comprise a multitude of species.
+Other models use calibration against local canopy-level data to define
+appropriate recruitment levels, which is likely to strongly hamper the
+general applicability of these models e.g. under scenarios of climate
+change.
 
-Correctly capturing the species composition in recruitment is often also
-a challenge, thus leading to unrealistic successional drift in the model
-REFERENCE?, for example by the change of the dominant species via
-stochastic processes that continuously bring seeds of other species that
-the come to dominance, with this strategy a new species might start to
-rise eliminating the former dominant species. A related issues is the
-rarification of species richness due to positive feedback effects such
-that eventually just single-species forest stands remain. This is
-sometimes handled by the incorporation of a low level of seed influx of
-all species at all times (e.g., @lischke2006 ). However, simulated
-species composition is usually exceedingly sensitive to that level
-whereas the parameters of such functions are poorly constrained by field
-data. There is also often a problem with the insuficient constraint of
+Correctly capturing the species composition in recruitment as a function
+of the presence of seed trees in the canopy is often a particular
+challenge, potentially leading to unrealistic successional drift in the
+model, which must be corrected by factors that are hard to parameterize
+[@lischke2006]. A related issues is the excessive reduction of species
+richness due to positive feedback effects such that eventually just
+single-species forest stands remain. This is sometimes handled by the
+incorporation of a low level of seed influx of all species at all times
+(e.g., @lischke2006 ). However, simulated species composition is usually
+exceedingly sensitive to that level, whereas the parameters of such
+functions are poorly constrained by field data. Furthermore, there is
+also often a problem with the insufficient data-based constraint on
 parameter values for models that start from very small tree sizes (e.g.,
@@
 influences, rather than being statistically driven and emphasizing
-recruitment into a larger size class (e.g., @wehrli2006 ). Most of these
-undesired behaviors in tree recruitment are often not published, but in
-the vast majority of studies they are corrected (e.g., @rasche2011,
-where for most study sites regeneration was simply switched off). Thus,
-a strong focus on the modeling of tree regeneration processes is highly
-important and long overdue [@price] if we are to make reliable
+recruitment into a larger size class (e.g., @wehrli2006 ).
+
+Thus, a focus on the modeling of tree regeneration processes is sorely
+needed and long overdue (cf. @price2001) if we are to make reliable
 projections of future forest dynamics, i.e. when the models need to be
@@
 
-Tree recruitment is understudied in field studies and modelling
-[@hanburybrown_forest_2022], as little data has been available and it is
-often very fragmented, i.e., not in one harmonized data set. On the one
-hand, monitoring on permanent sample plots (such as national forest
-inventories) often measure tree recruitment [@zell_modeling_2019]
+Also in the field of ecology, tree recruitment has not received a lot of
+attention [@hanburybrown_forest_2022], although many sub-processes have
+been studied in detail. Yet, little data is available that covers all
+regeneration processes within one species along environmental gradients,
+let alone for a vast suite of species. Thus, the available data are
+fragmented, which constitutes a major problem for model building. On the
+one hand, monitoring on permanent sample plots (such as National Forest
+Inventories) often measure tree recruitment [@zell_modeling_2019],
 however often with a design that captures the regeneration intensity
-after the stand initiation phase on productive forests [@hallsby2015].
-Data from permanent plots, although available across regions, it is also
-highly heterogeneous in design making their usability more difficult
-[@käber2021]. On the other hand, targeted studies to measure tree
-recruitment on experimental sites are often limited in space, species or
-the relevant processes leading to regeneration that are targeted
-[@berdanier2016; @collet2006].
+after the stand initiation phase in productive forests [@hallsby2015].
+Data from permanent plots, although available across regions, are also
+highly heterogeneous, rendering their utility difficult in a model
+context (cf. @käber2021). On the other hand, targeted studies to measure
+tree recruitment on experimental sites are often limited in spatial
+extent, species studied, or sub-set of regeneration processes that are
+targeted [@berdanier2016; @collet2006].
 
-This is the first time, a large number of forest models commonly used to
-assess forest dynamics under climate change are evaluated against a
-continental-scale multi-species harmonized data set on tree recruitment
-[@käber2022]. In this study we relied on models with various
-'philosophies', including models operating at the stand to the global
-scale. Because of the large variability in recruitment patterns in
-nature and the large amount of factor driving recruitment, including
-some which are not included in most models, such as deer browsing, we do
+In the present study, for the first time, a large number of forest
+models commonly used to assess forest dynamics under climate change are
+evaluated against a continental-scale multi-species harmonized data set
+on tree recruitment [@käber2022]. We included models that are based on a
+range of 'philosophies', including models operating at the stand to the
+global scale as well as the range of models from empirically derived to
+"process-based" (e.g. @fabrika2012; @mette2014; @lexer2001;
+@bugmann1996). Because of the large variability in recruitment patterns
+in nature and the large amount of factors driving recruitment, including
+some that are not included in most models, such as deer browsing, we do
 not aim at a detailed statistical evaluation of each model. Instead, we
-aim at evaluating general recruitment patterns, magnitudes and benchmark
-the regeneration niche of multiple species based on empirical data along
-a wide environmental gradient (temperature, moisture, and light
-availability). More specifically in this study we wanted to answer the
-following questions: 1) are models of forests dynamics capturing
-correctly regeneration levels across environmental gradients; 2) are
-models of forest dynamics representing correctly initial composition of
-tree populations;and, 3) can models of forest dynamics correctly capture
-the initial mortality in tree in-growth?.
+aim at evaluating the general recruitment patterns and magnitudes
+simulated by the models, and benchmark the simulated regeneration niche
+of multiple species against empirical data along a wide environmental
+gradient (temperature, moisture, and light availability).
+
+More specifically, we aimed to answer the following questions. 1) Are
+models of forests dynamics capturing correctly tree recruitment levels
+across environmental gradients? 2) Are models of forest dynamics
+representing correctly the initial composition of tree populations? And
+3), do the models of forest dynamics correctly capture the mortality in
+tree in-growth?.
 
@@
 
-Fifteen models of forest dynamics (including two model variants) were
-used to assess tree regeneration. The models selected simulate forest
-dynamics at the stand, landscape and global scale (Table
-\@ref(tab:tableModels)). Models construction philosophies and origin
-differs, most of the models have a largely process based approach and
-two models have formulations derived from statistical analyses based on
-inventory data (SYBILA and xComp). Some of the models largely rely on
-the original approach underlying forest gap models (ForClim 1, ForClim
-11, LandClim, Landis II ); and some include plant ecophysiology
-formulations (FORMIND, PICUS, 4C, iLand). This different approaches in
-the models construction also has influenced how tree establishment is
-defined.
+Fifteen models of forest dynamics (including two model featuring
+variants) were used to assess tree regeneration. The selected models
+simulate forest dynamics at the stand, landscape and global scale (Table
+\@ref(tab:tableModels)). The approaches used for model construction and
+their origin differ, with most of the models featuring a largely
+"process-based" approach, whereas two models based on formulations
+derived from the statistical analysis of inventory data (SYBILA and
+xComp). Some of the models largely rely on the original approach
+underlying forest gap models (ForClim 1, ForClim 11, LandClim, Landis-II
+); whereas some include plant ecophysiological processes (FORMIND,
+PICUS, 4C, iLand). These different approaches underlying the models also
+has strongly influenced the formulation of tree recruitment.
 
-Based on their establishment modules the models can be categorized as
+Based on their establishment modules, the models can be categorized as
 regeneration or recruitment (or ingrowth) models [@vanclay1994;
@@
 production, seed dispersal, germination or seedling establishment that
-lead to the number of seedlings or saplings. Recruitment models
-introduce a number of seedlings or saplings with certain characteristic
-such as biomass or dbh, without considering earlier development
-processes. We can also distinguish models that have feedback in the
-regeneration as those in which the level and species composition in
-regeneration is influenced by the existing adult tree community via the
-production of seeds, seedlings or saplings. The starting point for
-regeneration it is also different per model varying from seedbank, seed,
-sapling or ingrowth. The overall complexity in the establishment modules
-also varies, following @bugmann2022 we can see that most of the models
-included in this study can be classified as having rather low complexity
-in their tree recruitment modules while few of them have high complexity
-specifically when defining processes such as XXXX, XXXX (enumerate
-models).
+lead to the simulated number of seedlings or saplings. Recruitment
+models, in contrast, introduce a certain number of seedlings or saplings
+with certain characteristics such as biomass or diameter, without
+considering earlier development processes. We can further distinguish
+models that feature a feedback in the regeneration, i.e. where the level
+and species composition of regeneration is influenced by the existing
+adult tree community via the production of seeds, seedlings or saplings.
+
+The starting point for regeneration in the models also differs, ranging
+from seedbank, seed, seedling to sapling (i.e., trees much larger than
+10 cm height, often at least 2-3 m tall), which implies that the larger
+size threshold, the larger the number of ecological processes that must
+be aggregated via parameterizations because they cannot be treated
+explicitly. This reduces model complexity but comes at the cost of
+blurring process representation.
+
+The overall complexity in the establishment modules also varies
+considerably. Following @bugmann2022, we can classified seven models
+(iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForCLim and TreeMig) as
+having rather high complexity in their tree recruitment modules (mean
+complexity across all the recruitment process with value above zero in
+Table \@ref(tab:tableModels)). The highest complexity is when defining
+processes such as the process used for modeling establishment or how is
+frost occurrence taken into account in establishment,
 
 Two models, ForClim and ForCEEPs used two alternative model versions.
-HERE **FORCLIM** WRITE 2 SENTENCES ON THE DIFFERENCES BETWEEN FORCLIM 1
-AND 11 AND WHY IT IS RELEVANT TO CONSIDER THE TWO MODEL VERSIONS. HERE
-**FORCEEPS** WRITE 2 SENTENCES ON THE DIFFERENCES BETWEEN FORCEEPS AND
-FORCEEPS(F) AND WHY IT IS RELEVANT TO CONSIDER THE TWO MODEL VERSIONS.
+ForClim variant 1 [@bugmann1996] is based on an establishment module
+that adheres closely to the concept introduced by @botkin1972, whereas
+ForClim variant 11 is adopting a rather different approach for
+establishment model that considers
+space filling along with different \'points of entry\' for the action of
+specific
+environmental factors @huber2020. These two model variants allow us to
+evaluate a more process-based establishment module (variant 11) against
+a simpler module (variant 1) while the rest of the model structure
+remains the same. Similarly,
+the two ForCEEPS variants used here allow us to isolate the importance
+of the
+canopy feedback via seed trees for the quantity and quality of simulated
+regeneration, as one variants includes this feedback whereas the other
+one does
+not.
 
@@
 
-Regeneration data covering a wide range of environmental gradients is
-hard to obtain and this is one of the reasons why most of the models of
-forest dynamics have never been confronted with a data-set covering a
-wide range of climate gradients in a large number sites to evaluate how
-well are they capturing regeneration. The observed regeneration data
-used in this study represent the available range of environmental
-gradients in temperature and precipitation in Europe using 200 sites of
-strict forest reserves compiled in the framework of the EuFoRIa network
-(Figure \@ref(fig:map)). The sample forests have not been managed in at
-least 10 years (most of them in close to 100 years?), and each of the
-sites contains at least two consecutive measurements. Regeneration was
-measured at 7 and 10 cm threshold in all the sites except for 35 sites
-where only regeneration at 10 cm diameter threshold was available. For
-more details on this dataset and their regeneration information please
-refer to @käber2022.
+Recruitment data covering a wide range of environmental conditions are
+hard to obtain, and this is one of the reasons why most of the models of
+forest dynamics have never been confronted with a data-set covering such
+gradients over a large number sites, to evaluate how well they are
+capturing recruitment. The recruitment observations used here are from a
+novel and unprecedented network of sites in forest reserves that
+represent the range of environmental gradients in temperature and
+precipitation in Europe compiled in the framework of the EuFoRIa network
+[@euforia2019] (Figure \@ref(fig:map)). All these forests have not been
+managed for at least 10 years prior to their designation as reserves,
+and most of them provide time series of natural forest dynamics over
+multiple decades and up to ca. 80 years. The censuses periods range from
+37 to 155 years, with most of sites with an average of 14 years. The
+plot sized ranged from 0.02 ha to 5.52 ha, the recruited values were
+further processed and aggregated to provide recruited values per ha
+following @käber2022. The data provide information at the tree level,
+thus providing information for sequential comparisons such as
+recruitment and death between measurements.
+
+We selected 200 sites from this network as the benchmarking data set for
+the simulation exercise. Each of these sites features at least two
+consecutive measurements. Regeneration was measured at threshold of 7 or
+10 cm of stem diameter at all the sites, except for 35 sites where only
+regeneration at the 10 cm diameter threshold was available. The
+observations used in this study featured 30.900 recruited trees, and
+recruitment values per site and sample ranging from zero to 1246 trees
+per ha and a mean of 56 trees per ha recruited. For more details on this
+unique dataset and the detailed recruitment information please refer to
+@käber2022.
 
@@
 
-Each modeler group was provided with a detailed protocol (available in
-the supplementary material) with instruction on how to perform the
-simulations regarding input variables such as climate or soil,
-simulation setup and expected output variables. No information was
-provided on the regeneration prior to the simulation. The overarching
-goal of this experiment was the assessment of the regeneration niche as
-it arises from empirical data analysis against its representation in a
-wide range of models of forest dynamics.
+The overarching goal of this experiment was to assess the regeneration
+niche as it arises from empirical data against its representation in a
+wide range of models of forest dynamics. To this end, each moding group
+was provided with a detailed protocol (available in the Supplementary
+Material X, henceforth "SM") with instructions how to perform the
+simulations, providing input variables on climate and soil conditions,
+the specification of the simulation setup and the expected output
+variables. Neither site information (except for the data specified
+below) nor any data on tree regeneration were provided prior to the
+simulation. That is, the models were run in "blind flight" mode.
 
@@
 how to use the climate data was different in each model based on the
-approach the better fitted the model (Table \@ref(tab:tableModels)). The
-soil quality data was provided as a continuous value between 1 and 5
+approach that best suited the model (Table \@ref(tab:tableModels)). Soil
+quality data was provided as a continuous value between 1 and 5
 (Soilgrids dataset, @hengl2017, @poggio2021). The protocol also provided
@@
 [@asterscienceteam2019], no other spatial information such as
-coordinates were given with the exception of iLand that used the blurred
-coordinates to obtain detailed soil information values for the different
-sites.
+coordinates were provided with the exception of iLand which used the
+blurred coordinates to derive highly detailed soil information values
+for the 200 sites.
 
-The simulations had to be run in the absence of natural disturbances,
-one the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE [ONE
+The simulations were run in the absence of natural disturbances, only
+the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE [ONE
 SENTENCE]{.underline} OF WHY YOU INCLUDED DISTURBANCES AND WHAT TYPE.
-The simulations had to be set up in a way that they will obtain samples
-on species specific regeneration rates per decade and per ha. The
-modeling teams decided on the simulated area and how they derive these
-samples. Further details on how each modeling team prepared the
-simulations and the outputs are available in the supplementary material.
+The simulations were set up to sample species-specific regeneration
+rates per decade and per ha in the equilibrium state of the model,
+typically entailing a \"spin-up\" run (as we did not provide any forest
+data). The modeling teams decided on the simulated area and how they
+derived these samples. The exact length of the simulation was also
+decided by the modeling teams (Table \@ref(tab:tableModels)). Further
+details on how each modeling team prepared the simulations and the
+outputs are available in the SM 2.
 
-The simulations were run from bare ground in the absence of management
-to a simulated equilibrium ("potential natural vegetation") with the
-current climate as the ultimate goal was to derive the regeneration
-niche. The regeneration niche is defined as the passing of a diameter
-threshold of 7 and 10 cm (ingrowth). Both the width of the regeneration
-niche (i.e., in environmental space) as well as the intensity of the
-recruitment process (i.e., the number of ingrowth trees per area and per
-unit of time) were of interest. The exact length of the simulation was
-decided by the modeling teams (Table \@ref(tab:tableModels)).
+The simulations were run in the absence of management to a
+simulated equilibrium ("Potential Natural Vegetation") with the current
+climate, as the ultimate goal was to derive the regeneration niche. This
+entails the assumption that (1) the observations from the forest
+reserves reflect no traces of forest management and (2) an equilibrium
+between forest dynamics and climate. While the former is likely to be
+met at least in many of the EuFoRIA reserves, the latter may be more
+debatable. However, in the absence of detailed data on the history of
+each plot in the EuFoRIa network, some assumptions had to be made. We
+define the regeneration niche as the passing of a diameter threshold of
+7 or 10 cm respectively (synonym: ingrowth). Both the width of the
+regeneration niche (i.e., in environmental space) as well as the
+intensity of the recruitment process (i.e., the number of ingrowth trees
+per area and per unit of time) were of interest.
 
 The simulations were run for mixed species forests (not multiple
-single-species simulations) using mixtures of the following species or
-genera: *Fagus sylvatica, Picea abies, Abies alba, Carpinus betulus,
+single-species simulations) using mixtures of eleven species or genera
+for which ingrowth data of sufficient quality were available from
+EuFoRIa: *Fagus sylvatica, Picea abies, Abies alba, Carpinus betulus,
 Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus excelsior,
-Quercus spp., Alnus glutinosa, Pinus sylvestris.* Three models included
-additional species in the simulations (ForClim 1, ForClim 11 and
-TreeMig) and three could not simulate all the species (4C, xComp,
-TreeMig, LPJ-GUESS) (Table \@ref(tab:tableModels)). HERE **FORCLIM**
-PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED MORE SPECIES.
-HERE **TREEMIG** PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED
-MORE SPECIES AND MISSED ONE. HERE **4C** PROVIDES [ONE
-SENTENCE]{.underline} WHY THEY SIMULATED LESS SPECIES. HERE **XCOMP**
-PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED LESS SPECIES. The
-model aDGVM2 did not have species level option and provided the default
-approach with HERE **aDVGM2** PROVIDES [ONE SENTENCE]{.underline} WITH
-THE INFORMATION ON THE DEFAULT SET UP FOR THE SPECIES.
+Quercus spp., Alnus glutinosa, Pinus sylvestris.* Two models included
+their standard set of species for the simulations which is much larger
+(i.e. ForClim 1, ForClim 11). This also applies to TreeMig, but in this
+model *Betula pendula* was not included because HERE **TREEMIG**
+PROVIDES THE END OF THIS SENTENCE SAYING WHY B pendula WAS NOT INCLUDED.
+In three models, fewer than these elveven species were simulated (4C,
+xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). HERE **4C** PROVIDES
+[ONE SENTENCE]{.underline} WHY THEY SIMULATED LESS SPECIES. HERE
+**XCOMP** PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED LESS
+SPECIES. HERE **LPJ-GUESS** PROVIDES [ONE SENTENCE]{.underline} WHY THEY
+SIMULATED LESS SPECIES. Finally, the
+model aDGVM2 is based on a broad suite of Plant Functional Types rather
+than
+specific species, and hence this model used its default approach of PFTs
+for
+the simulations.
 
-Each of the models reported the number of recruits over a 10 year
-interval per species and per ha for each of the sites. To better
-understand the variation per site each modeling team decided on how to
-sample 200 multiple values per site to show the variation in recruitment
-for spatial units of 1 ha during the simulated phase of potential
-natural vegetation but without providing average of multiple 1-ha
-samples. This was done with different strategies such as sampling
+Each of the models reported the number of recruits by sampling 200 times
+over a 10-year interval for each species and per ha for each of the 200
+sites. Both the 200 samples per site as well as its average were
+reported. The multiple samples per site were used to better understand
+the variation within each site. This was done using different
+strategies, depending on the model. The strategies included (1) sampling
 simulated data from the same 1-ha plot in the equilibrium over time, or
-by sampling several 1-ha plots from the simulated forest at one specific
-point in time (in the equilibrium). This resulted in 880000 observation
-per model that simulated the 11 species indicated in the protocol (200
-sites, 200 samples per site, 11 species and two diameter thresholds).
-Those models that simulated extra species, the recruitment of those
-extra species were aggregated in a category others. The simulated data
-per model and the observations used in this study are provided as a
-supplementary information / repository.
+(2) sampling several 1-ha plots from the simulated forest at one
+specific point in time (in the equilibrium). This resulted in 880.000
+observations per model that simulated the 11 species included in the
+protocol (200 sites, 200 samples per site, 11 species and two diameter
+thresholds). For those models that simulated extra species, the
+recruitment of the extra species was aggregated into a category
+"others". The simulated data per model and the observations used in this
+study are available in a repository (REF HERE when available).
 
-```{r map, echo=FALSE, fig.cap="The 200 sites with the dominant tree species in the observed recruitment at 10 cm threshold.", out.width='80%'}
+Some models did not report all the simulated samples or sites to avoid
+unrealistic results. The model 4C set up a threshold of 100 m2/ha to
+avoid unrealistic stand basal area values, and therefore not all the
+samples and sites were reported, this is the consequence of 4C not being
+able to do long term simulations without management PLEASE **4C** VERIFY
+THIS IS WHAT YOU DID. LPJ-GUESS had 2% of the samples outputs as
+grasslands that were considered in the data as zero stand basal area and
+zero recruitment for all the tree species simulated. Landis II also
+missed X% of the samples HERE **LANDIS-II** PROVIDES [ONE
+SENTENCE]{.underline} WHY THEY DID NOT PROVIDE ALL THE SAMPLES.
+
+```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network that were used in the present study, showing the dominant tree species per site in the observed recruitment at the 10 cm threshold.", out.width='80%'}
 knitr::include_graphics(here::here("figures", "theMap.png"))
@@
 
-The simulation results were analyzed in regard of their regeneration
-levels, species composition and mortality in regeneration. The observed
-data was modeled using a generalized additive model with a negative
-binomial distribution to better understand the trend and regeneration
-levels of the observed regeneration and the simulation results per
-model. The regeneration levels were evaluated against the stand basal
-area (the basal area of all the trees in each sample) as a proxy of
-light availability for regeneration, the degree-days as a proxy for
-temperature and the water balance as a proxy for soil moisture. The mean
-recruited values were calculated as the mean across samples per site and
-when the regeneration levels were evaluated across the total basal area
-and climate gradients the sites.
+The simulation results were analyzed in regarding the recruitment
+levels(i.e., number of recruits per unit time and space), species
+composition of the recruitment, and mortality in the recruitment. The
+observed data were modeled using a Generalized Additive Model with a
+negative binomial distribution to better understand the trend and
+regeneration levels of the observed regeneration, relative to the
+simulation results per model.
+
+Regeneration levels were evaluated along three key environmental
+gradients, i.e. stand basal area (the basal area of all the trees in
+each sample) as a proxy of light availability at the forest floor, the
+annual degree-day sum (DD) [@allen1976; @fischlin1995] as a proxy for
+growing season warmth, and the climatic water balance (CWB) as a proxy
+for soil moisture [@speich2019]. The mean values of recruitment were
+calculated as the mean across the 200 samples per site.
 
 We evaluated the species composition across models and in relation to
-the observed data by calculating the Shannon index that defines the
-relative frequency of species referring to basal area where the higher
-the value, the higher the diversity of species in a particular site and
+the observed data by calculating the Shannon index based on the relative
+frequency of the species in terms of basal area. The higher the value of
+the index, the higher the species diversity at a particular site and
 sample (Equation \ref{eqpiShannon}).
@@
 ```
-where, $s$ is the total number of species simulated $n$ (11 for the
-models that simulated all the species and a varied number for those
-models that did not consider all the species, see Table
-\@ref(tab:tableModels) for more details in the simulated species);
-$p_{i}$ is the proportion of the species $i$ calculated as the recruited
-basal area ($r_{BA_{n,i}}$) for that species $i$ over the stand basal
-area ($Total_{BA_{n}}$) of the sample $n$.
+where, $s$ is the total number of species simulated that have recruited
+basal area above zero in sample $n$; $p_{i}$ is the proportion of the
+species $i$ calculated as the recruited basal area ($r_{BA_{n,i}}$) for
+that species $i$ over the stand basal area ($Total_{BA_{n}}$) of the
+sample $n$.
 
-In order to understand the impact of adult species composition on
-recruitment across climatic gradients we focused on the five main tree
-species: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus
+In order to analyze the impact of adult species composition on
+recruitment across the climatic gradients, we focused on the five main
+tree species: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus
 sylvestris*, and *Quercus* spp. For these species we calculated the mean
@@
 ```
-where, $s$ is the total number of species simulated; $\bar{r}_{n,i}$ is
+where $s$ is the total number of species simulated; $\bar{r}_{n,i}$ is
 the mean recruitment of the species $i$ per site $t$ across the
-available samples on site.
+available samples at that site.
 
-There were 35 sites that had a diameter threshold of 10 cm and therefore
-we do not know the actual recruitment rates for those sites at 7 cm; yet
-the models simulated recruitment for the 7 cm threshold on those sites.
-The 35 sites were removed from the simulated data to make a fair
-comparison. Furthermore, when looking at the recruited basal area share
-per species ($\bar{R}_{BA_{share_{i}}}$) we categorized the basal area
-share as zero if the total recruited basal area (
-$\sum_{i = 1}^{s}{\bar{r}_{t,i}}$ ) equal zero and the recruited basal
-area of that species ($\bar{r}_{t,i}$) was also zero.
+There were 35 sites that had a diameter threshold of 10 cm only and
+therefore we do not know the actual recruitment rates for those sites at
+7 cm; yet the models simulated recruitment for the 7 cm threshold at
+those sites. The 35 sites were removed from the simulated data to make a
+fair comparison regarding the recruitment of the five main species.
+Furthermore, when looking at the recruited basal area share per species
+($\bar{R}_{BA_{share_{i}}}$), we categorized the basal area share as
+zero when the total recruited basal area (
+$\sum_{i = 1}^{s}{\bar{r}_{t,i}}$ ) equaled zero, and also when the
+recruited basal area of that species ($\bar{r}_{t,i}$) was also zero.
 
-Mortality at regeneration was assessed based on the difference on
-recruitment between 7 and 10 cm diameter thresholds. We used the Reineke
-self-thinning rule [@reineke1933; @pretzsch2003] as a reference to
-estimate if the different recruitment values were above or bellow the
-expected change between these two diameters. The Reineke self-thinning
-rule is usually calculated from even-aged, single species stands and
-assumes a relationship between the number of stems and the quadratic
-mean diameter in fully stocked pure stands. The value used in our
-comparisons was 1.77 and was calculated following the Equation
+Mortality in the regeneration was assessed based on the difference in
+recruitment between the 7 and 10 cm diameter thresholds. We used the
+Reineke self-thinning rule [@reineke1933; @pretzsch2003] as a reference
+to estimate whether the different recruitment values were above or
+bellow the expected theoretical rate between these two diameters. The
+Reineke self-thinning rule is usually calculated from even-aged, single
+species stands and assumes a relationship between the number of stems
+and the quadratic mean diameter in fully stocked pure stands. The value
+used in our comparisons was 1.77, calculated using Equation
 \ref{eqReineke}.
@@
 
-Results related to regeneration levels, species composition and
-mortality in recruitment varied across models and in comparison with the
-observations. As expected, recruitment was overestimated across most
-models for both the 7 and 10 cm diameter thresholds, with the exception
-of SYBILA, Landis II and aDGVM2 that estimated recruitment rates at the
-lower end of the plausibility interval of the observed data (Figure
-\@ref(fig:over)). The models with highest overestimation were PICUS and
-TreeMig. For most of the models, the variability across sites in
-recruitment levels (visible from the interquartile range in the box
-plots of Figure \@ref(fig:over)) is similar to or smaller than the
-observed variability for most of the models, with the exception of
-PICUS, where the mean recruitment variability is much larger.
+Simulation results regarding regeneration levels varied strongly across
+the 15 models and typically did not match the levels found in the forest
+reserve data. Recruitment was overestimated in most models for both the
+7 and 10 cm diameter thresholds, with the exception of the empirical
+stand model SYBILA, the landscape model Landis II and the global model
+aDGVM2, which estimated recruitment rates at the lower end of the
+plausibility interval of the observed data (Figure \@ref(fig:over)). The
+models with the highest overestimation were the stand model PICUS and
+the landscape model TreeMig. For most of the models, the variability of
+simulated recruitment levels across the 200 sites (visible from the
+interquartile range in the box plots of Figure \@ref(fig:over)) was
+similar to or smaller than the observed, with the notable exception of
+PICUS, where the mean recruitment variability was much larger than any
+other model.
 
-```{r over, echo=FALSE, fig.cap="Mean recruitment values across all samples for the 200 sites.
+```{r over, echo=FALSE, fig.cap="Mean recruitment rates across all samples per site, plotted for the 200 sites and for each model.
 The red dotted lines shows the 2.5 and 97.5 percentiles for the 7 cm diameter threshold in the observed data.", out.width='100%'}
@@
 
-When looking at the representation of recruitment rates along key
+When evaluating the representation of recruitment rates along key
 gradients of light availability (basal area), temperature (degree-days)
-and soil moisture (water balance), distinct features emerge: the models
-reproduce the effect of basal area much better than the effects of
-climatic gradients, where they feature varying effects (Figure
-\@ref(fig:nvTrend710BA) and supplementary figures 1 and 2).
+and soil moisture (climatic water balance), distinct features emerge:
+the models reproduce the effect of basal area (Figure
+\@ref(fig:nvTrend710BA)) much better than the effects of the climatic
+gradients, where they feature varying effects (supplementary figures S1
+and S2).
 
-In the observed data the pattern shows that with increasing stand basal
-area recruitment decreases. However, this trend was captured by Landis
-II only, albeit at much lower values of basal area than in the
-observations. Other models featured distinctly different trends, such as
-(1) an increase in regeneration with increasing total basal area
-followed of no recruitment at high basal area (ForCEEPS, ForCEEPS(f),
-FORMIND, TreeMig and LPJ-GUESS), (2) almost constant regeneration levels
-with basal area (ForClim 11, xComp, LandClim and aDGVM2) or (3) an
-increase of recruitment up to a certain value of stand basal area
-followed by a decrease of recruitment at even higher stand basal area
-values at model-specific thresholds (4C, ForClim 1, PICUS, SYBILA and
-iLand).
+In the observed data, recruitment rates decreased clearly with
+increasing stand basal area (Figure \@ref(fig:nvTrend710BA)). This trend
+was captured by the Landis II model only, albeit at much lower values of
+basal area than in the observations. Other models featured distinctly
+different trends, such as (1) an increase in regeneration with
+increasing total basal area followed by the absence of recruitment at
+high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and
+LPJ-GUESS), (2) almost constant regeneration levels with basal area
+(ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of
+recruitment up to a certain value of stand basal area followed by a
+decrease of recruitment at even higher values of stand basal area, with
+model-specific thresholds (4C, ForClim 1, PICUS, SYBILA and iLand).
 
-Models did not match the observed patterns in recruitment across soil
-moisture and temperature gradients and there was either a clear common
-pattern across models. Looking at the soil moisture gradients
-(supplementary figure 1), we can see that a group of models decreases
+The models did not match the observed patterns in recruitment across
+gradients of soil moisture and temperature (Figures S1 and S2), and
+there was neither a clear pattern across models. Regarding the soil
+moisture gradient (Figure S1), a group of models features decreasing
 recruitment with increasing soil moisture (4C, ForClim 1, ForClim11,
-SIBYLA, xComp, PICUS and LandClim), while few models had the opposite
-trend (ForCEEPs(f), FORMIND and LPJ-GUESS). Across temperature gradients
-all the models that had a decrease in recruitment with increasing soil
-moisture showed the opposite trend (more recruitment with increasing
-temperature) except for 4C (supplementary figure 2). The rest of the
-models did not have a clear pattern across the temperature gradient.
+SIBYLA, xComp, PICUS and LandClim), while a few models showed the
+opposite trend (ForCEEPs(f), FORMIND and LPJ-GUESS). Across the
+temperature gradient (Figure S2), all the models that featured a
+decrease in recruitment rates with increasing soil moisture showed the
+opposite trend (i.e., more recruitment with increasing temperature),
+except for 4C. The other models did not have a clear pattern across the
+temperature gradient.
 
 Both observations and simulated data had no recruitment in some samples
-and sites (supplementary table 1). The observed data had over 3% of the
-samples with no recruitment. Only three models had larger proportion of
-no recruitment (4C, Landis II and aDVM2). Three models always simulate
-recruitment for both 7 and 10 cm threshold (TreeMig and LPJ-GUESS), an
-the rest of the models have a very low percentage of samples with no
-recruitment (0.01% - 1.31%).
+and at some sites (Table S1). The observed data had slightly more than
+over 3% of the samples with no recruitment. Only three models had a
+larger proportion of no recruitment (4C, Landis II and aDVM2). Three
+models always simulated recruitment for both 7 and 10 cm threshold
+(TreeMig and LPJ-GUESS), i.e., they did not featured zero values. The
+rest of the models had a very low percentage of samples with no
+recruitment (0.01% - 1.31%), i.e., they had distinctly fewer occurrences
+of zeros compared to the observations.
 
-```{r nvTrend710BA, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the stand basal area.
+```{r nvTrend710BA, echo=FALSE, fig.cap="Mean recruitment values across the 200 samples per site, for the 200 sites against the stand basal area.
 (A)same y-axis across all models; (B)y-axis scaled by model.
@@
 
-## Species composition in recruitment
+## Species composition of recruitment
 
-Most of the models match the level of diversity (as measured by the
+Most of the models matched the level of diversity (as measured by the
 Shannon index) of the observed data quite well (Figure \@ref(fig:H710)),
-although they consistently underestimate its variability across sites.
-Four models overestimate recruitment diversity: 4C, PICUS, LANDIS-II and
-LPJ-GUESS, the latter mostly for the 10 cm diameter threshold. The model
-4C is a special case as this model simulated 5 species only. In most
-models there are small differences across sites regarding the species
-richness of recruitment between 7 and 10 cm. Most models maintain the
-recruitment diversity between the 7 and 10 cm thresholds, and the same
-is visible from the observed data, there are only few models where the
-recruitment diversity increases between 7 and 10 cm which might be an
-artifact of the sample strategy and calculating the Shannon index based
-in basal area as for a same sample recruitment at 10 cm will include
-larger trees with larger basal area and if those are from different
-species, the Shannon index will be larger.
+although they consistently underestimated its variability across sites.
+Four models overestimated recruitment diversity: 4C, PICUS, LANDIS-II
+and LPJ-GUESS, the latter mostly for the 10 cm diameter threshold. The
+model 4C is a special case as it simulated 5 species only, i.e., its
+diversity values are not directly comparable to those of the other
+models, nor the observations. Only one model, SIBYLA, consistently
+underestimated recruitment diversity.
 
-```{r H710, echo=FALSE, fig.cap="Shannon index of the recruitment, calculated by basal area as the mean value across samples per site for the diameter thresholds of 7 and 10 cm, respectively.
+In most models, there were only small but significant differences in the
+species diversity of recruitment between 7 and 10 cm across sites (Table
+S2-S4). Furthermore, the observed data and three models (4C, SIBYLA,
+Landis II) maintained the recruitment diversity between the 7 and 10 cm
+thresholds (they were not significant differences between them, see
+Table S2), and the same was visible from the observed data. There are
+only few models with large recruitment diversity increasing between 7
+and 10 cm (ForCEEPS, TreeMig, LPJ-GUESS).
+
+```{r H710, echo=FALSE, fig.cap="Shannon index of tree recruitment, calculated by basal area as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
 The red dotted lines represent the 20th percentile of the observed data for the threshold of 7 cm and the 70th percentile of the observed data for the threshold of 10 cm.", out.width='100%'}
@@
 
-Species diversity is lower in the recruitment compared to the stand
-level, for both observed and simulated data, and for both recruitment
-thresholds, i.e. 7 and 10 cm (supplementary figure 3 and 4). Models
-simulating higher diversity at the stand level do not feature higher
-diversity in recruitment (supplementary figure 3 and 4). Most of the
-models are capturing well or over-predicting the species richness at the
-stand level (Figure \@ref(fig:combinedRichness)-A-centered plot and
-supplementary figure 5). The models over predicting species richness at
-the stand level are ForClim1, ForClim 11, FORMIND and TreeMig (Figure
-\@ref(fig:combinedRichness)-A-left plot). However only few models
-under-predict the species richness at the stand level LPJ-GUESS,
-ForCEEPS (f), xComp, iLand and 4C. (Figure
-\@ref(fig:combinedRichness)-A-right plot). The majority of the models
-capture reasonably well the recruitment diversity at lower values of
-diversity, however most of them strongly underestimate species diversity
-in recruitment for higher observed diversity levels(Figure
-\@ref(fig:combinedRichness)-B and supplementary figure 6). Both models
-with feedback and without feedback from the adult trees (via seed
-production) underestimate species richness in recruitment compared to
-the observed data.
+Species diversity was lower in the recruitment compared to stand-level
+diversity, for both observed and simulated data, and for both
+recruitment thresholds (Figure S3 and S4). Models simulating higher
+diversity at the stand level did feature higher diversity in recruitment
+(Figure S3 and S4). Most of the models were capturing reasonably well or
+were overpredicting the species diversity at the stand level (Figure
+\@ref(fig:combinedRichness)A, center; cf. Figure S5). The models over
+predicting species richness at the stand level were ForClim1, ForClim
+11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A, left).
+Only a few models underpredicted species diversity at the stand level,
+i.e., LPJ-GUESS, ForCEEPS(f), xComp, iLand and 4C. (Figure
+\@ref(fig:combinedRichness)A, right). The majority of the models
+captured reasonably well the recruitment diversity at lower values of
+diversity, but most of them strongly underestimated species diversity in
+recruitment for higher observed diversity levels (Figure
+\@ref(fig:combinedRichness)B, cf. Figure S6). Both the models with
+feedback and those without feedback from the adult trees to recruitment
+(via seed production) underestimated species diversity in the
+recruitment compared to the observations.
 
 ```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site and for observed and simulated values.
-A) Shows the three trends present across models for the species richness at the stand level.
-B) Shows the three trends present across models for recruitment at 7 cm threshold.
-The selected representative model per trend group was based on the supplementary figure 5 and 6.", out.width='75%'}
+n indicates the number of models falling in each group A) The three trends present across models for the species diversity at the stand level.
+B) The three trends present across models for recruitment at the 7 cm threshold.
+The representative model selected here per trend group was based on the full data shown in Figure S5 and S6.", out.width='75%'}
 knitr::include_graphics(here::here("figures", "combined_richness.png"))
@@
 Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
-sylvestris*) and oaks (*Quercus* spp) were the most frequent species in
-the observed data. Most models capture the share of basal area of
-recruitment of the five main tree species compared to the observed data
-along the two major gradients of climate (Figure \@ref(fig:diff) and
-supplementary figure 7). However, the models differ in the way they
-represent the climate niche of individual main species. First, some
-models overestimated recruitment across the whole gradient for one of
-the 5 main tree species (for example, ForCEEPS, ForCEEPS(f),LandClim or
-LPJ-GUESS), while other models overestimated recruitment for more than
-one species for larger parts of the climate gradient, such as xComp,
-PICUS, SYBILA or LPJ-GUESS (cf. red colors in Figure \@ref(fig:diff)).
-This trends were consistent with those when the regeneration thresholds
-was at 7 and 10 cm although the exact changes in recruited basal area
-share changed (supplementary figure 7).
+sylvestris*) and oak (*Quercus* spp.) were the species observed most
+frequently in the EuFoRIa data. Most models captured well the share of
+basal area in the recruitment of these five tree species compared to the
+observed data along the two major gradients of climate (Figure
+\@ref(fig:diff), cf. Figure S7). However, the models differed in the way
+in which they simulated the climate niche of these main species. First,
+some models overestimated recruitment across the entire gradients for at
+least one of the five main tree species (e.g., ForCEEPS, ForCEEPS(f),
+LandClim or LPJ-GUESS), while other models overestimated recruitment for
+more than one species for a large part of the climate gradients, such as
+xComp, PICUS, SYBILA or LPJ-GUESS (cf. red colors in Figure
+\@ref(fig:diff)). These trends were consistent for both regeneration
+thresholds, although the exact changes in the share of recruited basal
+area were somewhat different (Figure S7).
 
-When looking individually to each of the five main tree species it is
-possible to observe differences in the simulated recruitment results.
-Silver fir (*Abies alba*) was well captured across the gradient by most
-models except for xComp, Picus, LandClim and Landis-II, which
-overestimated its recruitment share. There are also some particularities
+Second, there was some congruence in the simulation results by species
+across the models. The share in the recruitment by Silver fir (*Abies
+alba*) was captured well across the gradient by most models except for
+xComp, PICUS, LandClim and Landis-II, which overestimated the
+recruitment share of this species. There are also some particularities
 evident for some models, such as 4C that did not simulate *Abies alba*.
-*Fagus sylvatica* was underestimated by many models across the
-environmental gradient, including ForCEEPS(f) and FORMIND that did not
-have regeneration in most of the sites for these species. Most of the
-models over-estimated *Picea abies* recruitment share on colder sites.
-*Pinus sylvestris* and *Quercus* spp. represented a smaller share of the
-recruited species in the observed data and localized on certain areas of
-the environmental gradient, however many models did have recruitment for
-these two species in most of the climatic gradient although at basal
-area share. Five models (4C, FORMIND, SYBILA, xComp, and LandClim) had
-almost no generation or did not regenerate *Pinus sylvestris* and
-*Quercus* spp.
+*Fagus sylvatica* recruitment was underestimated by many models across
+the environmental gradients, including ForCEEPS(f) and FORMIND, which
+featured negligible regeneration at most sites for this species. Most of
+the models overestimated the *Picea abies* recruitment share in the
+cold-wet part of the gradients. *Pinus sylvestris* and *Quercus* spp.
+represented a smaller share of the recruited species in the observed
+data mostly at warm-dry sites. Many models, however, erroneously
+featured recruitment for these two species in most of the climatic
+gradient although at a low basal area share. Five models (4C, FORMIND,
+SYBILA, xComp, and LandClim) had almost no recruitment or did not
+feature any recruitment of *Pinus sylvestris* and *Quercus* spp. at all.
 
-It is important to keep in mind two aspects from Figure \@ref(fig:diff),
-first the values shown are the means of all samples per site, and the
-mean across the sites in each environmental bin. Second, in this figure
-we are seeing together two types of zeros (represented with a lack of
-dots in the corresponding tile), the first type of zero are those sites
-where the recruited basal area of the species was zero and the stand
-recruited basal area for all species was also zero, those were
-considered as zero basal area share; the second type of zeros are those
-sites where there was no recruitment for that species but there was
-recruitment for other species.
+Lastly, there is no model that performed well across the five species
+regarding the species-specific recruitment rates in the environmental
+space of Figure 6. Some models (e.g., ForClim11, PICUS) tended to
+consistently overestimate the recruitment share of all five species, but
+most models overestimated the recruitment share of some species while
+underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
 
-```{r diff, echo=FALSE, fig.cap="Departure of each model in the representation of the main species recruitment relative abundance from the observations across the environmental gradients for a 7cm treshold.
-The values shown are the mean of samples per site and across sites in bins, where the range of the climatic water balance (cm) and seasonal degree-day sum (dºC) was divided in 10 bins.
-The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species, where a lack of dot represents zero basal area recruitment share.
-The color gradient shows the difference between the simulated and observed ratio of recruited basal area of the species and the total recruited basal area for all the species.", out.width='100%', out.height= '90%'}
+```{r diff, echo=FALSE, fig.cap="Recruitment rates for the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm recruitment threshold.
+The values shown are the mean of the 200 samples per site and across the sites in each bins (tile), with 10 bins per gradient.
+The sizes of the circles represent the ratio between recruited basal area of the species and the total recruited basal area for all species.
+The absence of a circle indicates a zero basal area share in the recruitment, or the absence of recruitment altogether.
+The color gradient (for the models only) shows the difference between the simulated and observed ratio of recruited basal area of the species and the total recruited basal area for all the species.", out.width='100%', out.height= '90%'}
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
@@
 
-There are strong differences across models on the mortality rate between
-7 and 10 cm (Figure \@ref(fig:mortality11)). All models except for the
-two DGVMs LPJ-GUESS and aDGVM2 feature a mortality rate larger than the
-observed data, thus at least partially compensating for the general
-overestimation of recruitment rates (Figure \@ref(fig:over)). Some of
-the models showed strong compensation for the initial overestimation of
-recruitment with larger mortality of recruitment between 7 and 10 cm
-thresholds but not in others. On the one hand, models below the Reineke
-line have lower mortality than expected by the self-thinning rule, which
-may point to facilitation, or simply a higher mortality rate before the
-trees reached 7 cm diameter. On the other hand, models such as xComp,
-PICUS and TreeMig are above the self-thinning line (Reineke), indicating
-that they have higher mortality in recruitment of mixed species forest
-than the expected in an even-aged single species forests, which
-indicates strong compensation for the initial overestimation of
-recruitment in some models. However, some of the models simulated a
-mortality rate in recruitment that is broadly compatible with the
-self-thinning rule, i.e. 4C, SIBYLA, LandClim and LANDIS II.
+There were strong differences among the models regarding the mortality
+rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the
+models featured a mortality rate significantly larger than the observed
+data, thus at least partially compensating for the general
+overestimation of recruitment rates (Figure \@ref(fig:over) and Table
+S5); six models did not have a significantly difference compared to the
+observed data (ForCEEPS, FORMIND, ForClim 1 and 11, and aDGVM2), and
+only the model LPJ-GUESS had a significantly lower mortality rate. Four
+models (xComp, PICUS, LandClim and TreeMig) featured very high mortality
+rates (i.e., well above the Reineke self-thinning line), which
+compensated for the strong initial overestimation of recruitment (cf.
+Figure 2). Interestingly, two models that underestimated overall
+recruitment levels (SIBYLA and Landis-II; cf. Figure \@ref(fig:over))
+still featured mortality rates in the recruitment that were close to but
+still above the self-thinning line (Figure \@ref(fig:mortality11)).
 
-Models with higher mortality also showed higher tree recruitment at 7cm
-(Figure \@ref(fig:mortality13)). On the other hand, most of the models
-with a low variation in mortality between 7 and 10 cm (ForCEEPS, ForClim
-1, ForClim 11 and LandClim) also have lower overestimation in
-recruitment at 7 cm (Figure \@ref(fig:mortality11) and supplementary
-figure 8). Some of the models that start at a lower number of
-individuals (7 cm threshold) have a lower decrease between 7 and 10 cm
-dbh (ratio 7-10 cm); in these models the species competition and
-self-thinning is either not that large or it happened before the species
-reached 7 cm. However, this pattern it is not consistent across models
-(supplementary figure 9). There is an effect of the number of
-individuals at 7cm on the mortality rate. Most of the models show a
-positive effect (increasing recruitment at 7cm shows increases the
-mortality between 7 and 10cm), but some models show a negative effect,
-however the trend it is not very high and not always significant when we
-look at the slope of the linear trends per model (supplementary figure 9
-and table 2).
+On the one hand, the fact that some models feature lower mortality than
+expected by Reineke\'s self-thinning rule may point to facilitation, or
+simply a higher mortality rate
+before the trees had reached 7 cm diameter. On the other hand, the case
+of models
+such as xComp, PICUS and TreeMig, which are well above the self-thinning
+line, may
+indicate that they feature higher recruitment mortality in mixed-species
+forests
+compared to even-aged, single-species forests. However, some models
+yielded a mortality rate of recruitment that is broadly compatible with
+the self-thinning rule, i.e. 4C and LandClim.
+
+The three models with largest overestimation proportion of recruitment
+at 7cm threshold showed had the largest mortality between 7 and 10 cm
+thershold (Figure \@ref(fig:mortality13)A). The rest of the models that
+overestimated recruitment had mortalities lower than the expected shelf
+thinning ratio (Figure \@ref(fig:mortality13)B). However, most of the
+models with a low variation in mortality ratio between 7 and 10 cm
+recruitment (ForCEEPS, ForClim 1, ForClim 11 and LandClim) (Figure S8)
+feature a small overestimation of recruitment at the 7 cm threshold
+(Figure \@ref(fig:mortality11); cf. Figure S8). Some of the models that
+start at a lower number of individuals (7 cm threshold) have a lower
+decrease between 7 and 10 cm dbh (ratio 7-10 cm) (Figure S10); in these
+models the species competition and self-thinning is either not that
+large or it happened before the species reached 7 cm. However, this
+pattern it is not consistent across models (Figure S9). There is an
+effect of the number of individuals at 7cm on the mortality rate. Most
+of the models show a positive effect (increasing recruitment at 7cm
+shows increases the mortality between 7 and 10cm). Some models show a
+negative effect, however the trend it is not very pronounced and not
+always significant regarding the slope of the linear trends per model
+(Figure S9 and Table S2).
 
@@
 particular with increasing disturbance under climate change. Correctly
-capturing regeneration in forest dynamics is fundamental to for example
-evaluate post-disturbance dynamics as this will define the status of the
-forests in the first 15-20 years [@seidl_post-disturbance_2022]. If this
-initial state it is not well captured then we can not properly assess
-aspects such as the forest resilience, or changes in carbon capture. In
-this study, for the first time, projections of tree recruitment done by
-several models of forest dynamics were confronted with observed data
-across a large environmental gradients.
+capturing forest regeneration in dynamic models is fundamental to, for
+example, evaluate post-disturbance dynamics as this will define the
+status of the forests in the first 15-20 years
+[@seidl_post-disturbance_2022]. If this initial state is not captured
+well, we can not properly assess aspects such as the forest resilience,
+or the timing, magnitude and progression of carbon capture. In this
+study, for the first time, projections of tree recruitment from multiple
+models of forest dynamics were confronted with a unique dataset from
+unmanaged forest reserves across a large environmental gradient in
+Europe.
 
-The observed data that we used [@käber2022] is exceptional in this
-regard, and to use it for model benchmarking has gave us insights on
-recruitment levels, species composition and mortality and overall a
-broader understanding than if we have used these data for a model
-calibration. The lower available threshold in the observed data of 7 cm
-provides a limit from which we can understand tree recruitment, which
-might be a rather long period since germination as relevant
-environmental constraint might have already act on those young trees
-[@käber2021]. However recruitment data with lower thresholds are not
-available in a harmonized manner from large environmental gradients. The
-observed data was collected from forest reserves where there has not
-been management for long periods of time, this might help to properly
-assume that these structures are in an equilibrium state. This
-assumption might also justify some of the rather large recruitment rates
-that we are seeing in the models projections compare to the observed
-ones.
+The observed data that we used [@käber2022] are also exceptional with
+respect to the number of recruitment events captured, which is essential
+for a highly stochastic process such as tree recruitment. The use of
+this dataset for model benchmarking provided novel insights on the
+capability of state-of-the art models in simulating recruitment levels,
+the species composition of recruitment, and mortality in an early stage,
+i.e. between a tree diameter of 7 and 10 cm. Overall, when adopting this
+approach a much broader understanding resulted than if we had used these
+data for model calibration: it is primarily from the failure of the
+models that we can gain ecological insights [@trugman2022].
 
-Adequate tree recruitment rates define the structure and composition of
-forests. Most of the models of forest dynamics tended to over-recruit
-trees in all the sites. Over estimating recruitment might have
-fundamental implications on the role of regrowth and forest carbon, with
-an overestimation in their capacity for carbon capture for example after
-the impact of disturbances [@pugh2019]. Excess in recruitment also
-requires of larger mortality rates in later stages, changing the
-expected forest densities at early stages and forcing the models to have
-stronger filters. Regeneration levels define the success of future
-forests, but it is equally important to correctly identify un-stocked
-forest areas [@rammer2021]. Tree recruitment is often characterized by
-an excess number of zero counts [@fortin2007]. In this study some of the
-models did not have lack of recruitment or their proportion was very
-low. This substantial difference with the observed data might also be a
-consequence of assuming that the observed data is also in equilibrium,
-where lack of recruitment in sites could also be an indication of
-"non-equilibrium" due to management history.
+## Recruitment levels
 
-Light availability effect on tree regeneration have been widely
-documented [@messier1999; @collet2006; @berdanier2016]. However it is
-difficult to collect variables defining light availability on
-regeneration. In this study, we used the total stand basal area as a
-proxy for light availability. There was differences in the basal area
-ranges, some of the models had stand basal areas lower than the observed
-data but there was also models that did have more stocked stands than
-the observed data, therefore it was not possible to evaluate the
-recruitment for the extremes of the density ranges in some models.
-Regeneration at low stand densities is interesting to assess how well
-forests are recovering e.g. after the creation of gaps due to
-disturbances. We would have expected that the recruitment of the
-different species would become sparser and droped out completely as a
-function of a proxy of light availability, however few models showed
-this trend indicating that the relation between recruitment and light
-availability it is not yet correctly captured.
+Accurate simulation of tree recruitment rates is pivotal for projecting
+the dynamics of forest stand structure and composition. Most of the
+models of forest dynamics evaluated here tended to overestimate tree
+recruitment rates across the 200 sites. Overestimating has far-reaching
+implications regarding biomass (and carbon) turnover, with a potential
+overestimation in their capacity for carbon capture [@pugh2019].
+Importantly, excess recruitment requires higher mortality rates in later
+stages, either -- as observed for
+some models -- between the 7 and 10 cm diameter thresholds, or either
+before or
+shortly after this stage. Correcting for the expected forest densities
+at early
+stages is thus equivalent to compensating for a first error (excess
+recruitment) by committing a second error (excess mortality). Thus, many
+models appear to get the right \'answer\', i.e., correct stand structure
+and composition, for the wrong reasons. This is problematic particularly
+if the models are to be used under novel conditions (e.g., climate
+change; [@huber2021]) or in a decision support context (e.g.,
+[@thrippleton2021]).
 
-Correctly capturing tree composition in recruitment help us to assess
-the functional composition of the forest appropriately, e.g. their
-sensitive to drought or resilience to new disturbances events
-[@redmond2015; @seidl_post-disturbance_2022]. In the models simulation
-outputs overall species richness levels in recruitment was
-underestimated but inside the observed ranges, however there were
-variations on how the main tree species were represented in the
-environmental gradient. The lower species richness in recruitment might
-be a consequence of models running without disturbances, disturbances
-would typically allowed for larger gaps sizes allowing less shade
-tolerant species to grow in larger proportions. Most models maintain or
-even increase in recruitment diversity between the 7 and 10 cm
-thresholds, and the same is visible from the observed data, however the
-differences are not significant. Species richness variations might be
-explained by the way each model considers recruitment, e.g. in the
-absence of adults, recruitment can be linked to either seed
-rain-dispersal, or to seed/seedling bank, but also on what processes are
-formulated with stochastic or deterministic approach. Differences might
-also be expected between cohort and individual trees based models
-**how?**. However, our results do not show differences between models
-with and without feedback in recruitment, **why?** The adult forests
-species richness was better captured than the recruitment species
-richness, and only few models over-predicted stand level species
-richness **why these ones?**.
+In spite of this criticism, it remains remarkable that most of the
+models were not overly off
+with respect to simulated regeneration levels -- after all, few if any
+of the models\' establishment routines are sufficiently constrained by
+data, and the resulting degrees of freedom that are available for
+modeling regeneration (cf. [@konig_tree_2022]) could have implied that
+model performance would be much worse.
 
-We could expect that models that over-recruit trees will also have a
-large mortality between 7 and 10 cm thresholds but this was not shown.
-This implies that these models must have the excess of mortality in
-later stages. Higher mortality in the adult component might enhance the
-share of less shade tolerant species for those models. There might be a
-difference between models including stochastic processes in recruitment
-and those which do not **??**. There was also a group of models below
-the Reineke line indicating a lower mortality than expected by the
-self-thinning rule, which may point to facilitation, or simply a higher
-mortality rate before the trees reached 7 cm diameter, another
-possibility would have been that these models underestimate recruitment
-which was not the case. This might also be partly an artifact of the
-method, as we are expecting space for time substitution, but in reality
-we just have one measurement.
+Regeneration levels define the success of future forests, but it is
+equally important to correctly identify unstocked forest areas
+[@rammer2021]. Tree recruitment data are characterized by zero inflation
+[@fortin2007]. This was clearly evident from the EuFoRIa dataset, but
+some of the models did not produce zeros at all, or featured a very low
+proportion of zero data. This substantial difference may be due to the
+assumption that the observed data is also in equilibrium situation, and
+the lack of recruitment in the data set could be an indication of a
+non-equilibrium e.g., due to management history.
 
-There is also the question of how much level of detail one have to give
+Light availability as a strong filter for tree regeneration has been
+documented widely [@messier1999; @collet2006; @berdanier2016]. However,
+it is difficult to collect variables defining light availability on
+regeneration at a large spatial and temporal scales. In this study, we
+used the total stand basal area as a proxy for light availability. We
+found pronounced differences in the stand basal area ranges across
+models, but also between models and the observation. This made
+impossible to evaluate the recruitment for the extremes of the density
+ranges in some models. For example, regeneration rates at low stand
+densities are relevant to assess how well forests are recovering e.g.,
+after gap creation due to disturbance[@seidl_post-disturbance_2022;
+@grubb1977]. We expected that with increasing stand basal area, the
+recruitment of the different species would become sparser and drop out
+entirely under low-light condition [@klopcic2012; @zell2019]. However,
+few models showed this trend, thus indicating that the relation between
+recruitment and light availability it is not yet captured correctly in
+the models.
+
+## Species composition of recruitment
+
+Correctly capturing tree species composition of recruitment is important
+to assess the future functional diversity of a forest, e.g., its
+sensitivity to drought or resilience to disturbances events
+[@redmond2015; @seidl_post-disturbance_2022]. In the model simulations,
+overall species diversity levels in recruitment were within the observed
+ranges. Yet, there were variations on how the main tree species were
+represented in the environmental gradient. Most models maintain or even
+increase in recruitment diversity between the 7 and 10 cm thresholds,
+and the same is visible from the observed data, however the differences
+are not significant for all the models. Species richness variations
+might be explained by the way each model considers recruitment
+[@konig_tree_2022], e.g., in the absence of adults, recruitment can be
+linked to either seed rain-dispersal, or to seed/seedling bank, but also
+on what processes are formulated with stochastic or deterministic
+approach. Differences might also be expected between cohort and
+individual trees based models. MAKE REFERENCES TO LITERATURE
+
+Our results do not show differences between models with and without
+feedback in recruitment. The adult forests species richness was better
+captured than the recruitment species richness, and only few models
+overpredicted stand level species richness. Most of the models have
+feedback on recruitment from the adult trees, therefore EXPAND ON
+FEEDBACK ISSUE
+
+## Mortality in regeneration
+
+As mentioned above, it is reasonable to expect that models that
+over-recruit trees will also have a couting-acting high mortality
+between the 7 and 10cm thresholds, but this was nit cinsistently evident
+from the simulations. This implies that these models must have the
+excess of mortality in earlier or later stages. Higher mortality in the
+adult component might enhance the share of less shade tolerant species
+for those models. There might be a difference between models including
+stochastic processes in recruitment and those which do not. There was
+also a group of models below the Reineke line indicating a lower
+mortality than expected by the self-thinning rule, which may point to
+facilitation, or simply a higher mortality rate before the trees reached
+7 cm diameter, another possibility would have been that these models
+underestimate recruitment which was not the case. This might also be
+partly an artifact of the method, as we are expecting space for time
+substitution, but in reality we just have one measurement. NEED MORE
+REFERENCES AND DISCUSSION POINTS
+
+## Research recomendations
+
+There is also the question of how much level of detail one has to give
 to define regeneration processes [@konig_tree_2022; @bugmann2022].
@@
 
-In these study we considered a wide range of models of forest dynamics,
+In this study, we considered a wide range of models of forest dynamics,
 from stand to global scale. However, the simulation set up was limited
@@
 considerations on tree recruitment [@beckage2003] and the lack of it
-might represent a more important disadvantage for landscape level
+might represent an more important disadvantage for landscape level
 models. Without spatial context we are limiting recruitment to the seed
@@
 between cells where there is an abstraction of the space or even the
-trees themselves [@hanburybrown_forest_2022]. The selected models also
-had limitations on species representation, some models where not able to
-simulate certain species due to lack of species parametrization. We also
-had one model that simulate PFTs providing a less flexible comparison
-the with species values; with the lose of the species specific
-information might also represent substantial ecological
+trees themselves [@hanburybrown_forest_2022]. Landscape level models
+have been build to be very accurately at the landscape level. EXPAND ON
+THE DIFFERENT SCALE LEVELS AND IMPLICATIONS
+
+Modeling approach seems more important than scale and accuracy of the
+design. Model behavior (accuracy) is completely unrelated to model scale
+(stand, landscape, global) and approach (empirical, process-based).
+Having both empricial and process based models, the expectation was that
+the empirically based models (SIBYLA, xComp) would have a better
+performance as they are calibrated with inventory data, however this was
+not the case. EXPAND ON
+
+The selected models also had limitations on species representation, some
+models where not able to simulate certain species due to lack of species
+parametrization. We also had one model that simulate PFTs providing a
+less flexible comparison the with species values; with the lose of the
+species specific information might also represent substantial ecological
 representativeness drawbacks [@purves2008].
 
-**Conclusion paragraph**
+**METHODS PARAGRAPH**
+
+Yet, in this study we were not able to assess tree regeneration in a
+holistic sense. The
+lower size threshold of 7 cm in the observed data constituted a hard
+limit based
+on which we can understand some aspects of tree recruitment, which in
+its
+entirety often comprises a rather long period since germination
+[@price2001], and many relevant environmental constraints are acting on
+these young trees [@käber2021] that we were unable to assess. However,
+recruitment data with lower thresholds are simply not available in a
+harmonized manner across large environmental gradients. The EuFoRIa data
+used here were collected in forest reserves where no management has
+taken place for long periods of time. This makes our assumption of an
+equilibrium between forest properties and environmental drivers less
+haphazard than it might appear at first sight. However this assumption
+might also justify some of the rather large recruitment rates that we
+are seeing in the models projections compare to the observed ones, as if
+the observed data it is not in equilibrium the observed recruitment
+would be lower as they are not close to the optimum.
+
+RECOMENDATION PARAGRAPH
+
+Cautious when using this models to evaluate regeneration post
+disturbances we might be overestimation their capacity for carbon
+capture
+
+## Conclusions
 
@@
 the environmental gradient might impact future forests...
+
+Provide succinct answers to the three questions/objectives stated at the
+end of the
+Introduction. Perhaps end with a concluding (\"more positive\") final
+paragraph.
+
+Models eventually get the right stand structure and composition, most of
+them do that
+by compensating for one error (overestimation of recruitment) by another
+error
+(excessive subsequent mortality, if not between 7 and 10 cm, then
+shortly
+thereafter): \"getting the right answer for the wrong reason\". And this
+is
+worrysome in general, but particularly (and very acutely) if the models
+are to
+be used e.g. under changing climatic conditions.
 

### 2022-12-02 code/figures.R
adding some further details into the figures like A, B,C sectionc etc

@@ -15,7 +15,18 @@
     outputsDF <- data.table::fread("data/dataOutputs.csv")
     outputsDF$model[outputsDF$model == "Empirical"] <- "Observed"
 
-
+    
+    # Number of trees in the observed data 
+    obsNum <- outputsDF [outputsDF$model == "Observed",]
+    round( sum(obsNum$r.trees, na.rm = TRUE), 0)
+    
+    obsNumAgg <- obsNum |> 
+                   dplyr::group_by(site, sample) |>
+                   dplyr::summarise( r.ha= sum(r.trees))
+    
+    hist(obsNumAgg$r.ha)
+    range (obsNumAgg$r.ha, na.rm = TRUE)
+    
     # Assign models and dbh to levels 
 
     outputsDF$model <- factor(outputsDF$model, levels = modelsOrder)
@@ -228,7 +239,7 @@
     
 
     #### Fig. R VS BA----
-  png(file = "figures/envTrend7Totba.png", width = 16, height = 12, 
+  png(file = "figures/envTrend7Totba_500.png", width = 16, height = 12, 
         units = "cm", res = 300)
   par(mfrow = c(4, 4), mar = c(2, 3, 1, 1), oma = c(2, 2, 1, 1))
     
@@ -540,10 +551,31 @@
     ggplot2::ggsave("figures/H7_10.png",
                     plot =   H7_10,
                     width = 21, height = 12, scale = 0.9,
-                    dpi = 300, units = "cm", device = 'png') 
+                    dpi = 300, units = "cm", device = 'png')
     
     
+    # Significant difference per model?
     
+    sigH7_10 <- ggpubr::compare_means(ShannonIndexRecruit ~ dbh, 
+                                      data = dat710[!dat710$model == "aDGVM2", ],
+                          group.by = "model", method = "t.test")
+    
+    write.csv(sigH7_10,"figures/sigH7_10.csv", row.names = FALSE)
+  
+    
+    # Significant difference compared to observed?
+    
+    sigH7_Obs <- ggpubr::compare_means(ShannonIndexRecruit ~ model,  
+                                         data = dat710[!dat710$model == "aDGVM2" & dat710$dbh == "7", ],
+                                       ref.group = "Observed", method = "t.test")
+    write.csv( sigH7_Obs,"figures/sigH7_Obs.csv", row.names = FALSE)
+    
+    sigH10_Obs <- ggpubr::compare_means(ShannonIndexRecruit ~ model,  
[…]

### 2022-12-02 manuscript/manuscriptRW.Rmd
a failed attempt: trying to write conclusion in the manuscript

@@
     reference_docx: template_paper.docx
-  bookdown::word_document2: default
 mainfont: Arial
@@
 results are benchmarked against data from forest reserves collected at
-these 200 sites. We aim to assess how well the models of forest dynamics
-are capturing the overall abundance, species composition and mortality
-in tree recruitment. The results show that most of the models tend to
+these sites. We aim to assess how well the models of forest dynamics are
+capturing the overall abundance, species composition and mortality in
+tree recruitment. The results show that most of the models tend to
 over-estimate recruitment rates although this is not always connected to
@@
 simulating higher species diversity at the stand level do not feature
-higher diversity in recruitment. *Abies alba* was well captured across
-temperature and soil moisture gradients by most models, whereas *Fagus
-sylvatica* was generally underestimated. *Picea abies* recruitment was
-generally over-estimated, particularly at colder sites. HERE TWO
-SENCENTECES DISCUSSION/CONCLUSION. HERE A SO WHAT SENTENCE
+higher diversity in recruitment. Looking at the individual species,
+*Abies alba* was well captured across temperature and soil moisture
+gradients by most models, whereas *Fagus sylvatica* was generally
+underestimated. *Picea abies* recruitment was generally over-estimated,
+particularly at colder sites. Increasing complexity in the models
+regeneration modules did not represent and advantage to better capture
+recruitment. Model behavior (accuracy) is completely unrelated to model
+scale (stand, landscape, global) and approach (empirical,
+process-based). We conclude that even-though the overestimation of
+recruitment might represent a challenge for the usability of these
+models for projections under uncertain conditions such a climate change
+and some model formulations should be reconsidered to correct excessive
+recruitment levels, it is still remarkable that all these models can
+capture such a complex process like regeneration in relatively correct
+ranges while never been parametrized with such a data.
 
@@
 xComp). Some of the models largely rely on the original approach
-underlying forest gap models (ForClim 1, ForClim 11, LandClim, Landis-II
-); whereas some include plant ecophysiological processes (FORMIND,
+underlying forest gap models (ForClim 1, ForClim 11, LandClim, Landis
+II); whereas some include plant ecophysiological processes (FORMIND,
 PICUS, 4C, iLand). These different approaches underlying the models also
@@
 thresholds (they were not significant differences between them, see
-Table S2), and the same was visible from the observed data. There are
-only few models with large recruitment diversity increasing between 7
-and 10 cm (ForCEEPS, TreeMig, LPJ-GUESS).
+Table S2), the same was visible from the observed data. There are only
+few models with large recruitment diversity increasing between 7 and 10
+cm (ForCEEPS, TreeMig, LPJ-GUESS).
 
@@
 The absence of a circle indicates a zero basal area share in the recruitment, or the absence of recruitment altogether.
-The color gradient (for the models only) shows the difference between the simulated and observed ratio of recruited basal area of the species and the total recruited basal area for all the species.", out.width='100%', out.height= '90%'}
+The color gradient (for the models only) shows the difference between the simulated and observed ratio of recruited basal area of the species and the total recruited basal area for all the species.", out.width='100%', out.height= '85%'}
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
@@
 
-## Mortality in regeneration
+## Mortality in recruitment
 
@@
 pattern it is not consistent across models (Figure S9). There is an
-effect of the number of individuals at 7cm on the mortality rate. Most
-of the models show a positive effect (increasing recruitment at 7cm
-shows increases the mortality between 7 and 10cm). Some models show a
+effect of the number of individuals at 7 cm on the mortality rate. Most
+of the models show a positive effect (increasing recruitment at 7 cm
+shows increases the mortality between 7 and 10 cm). Some models show a
 negative effect, however the trend it is not very pronounced and not
@@
 if the models are to be used under novel conditions (e.g., climate
-change; [@huber2021]) or in a decision support context (e.g.,
-[@thrippleton2021]).
+change; @huber2021) or in a decision support context (e.g.,
+@thrippleton2021).
 
@@
 data, and the resulting degrees of freedom that are available for
-modeling regeneration (cf. [@konig_tree_2022]) could have implied that
+modeling regeneration (cf. @konig_tree_2022) could have implied that
 model performance would be much worse.
@@
 densities are relevant to assess how well forests are recovering e.g.,
-after gap creation due to disturbance[@seidl_post-disturbance_2022;
+after gap creation due to disturbance [@seidl_post-disturbance_2022;
 @grubb1977]. We expected that with increasing stand basal area, the
@@
 on what processes are formulated with stochastic or deterministic
-approach. Differences might also be expected between cohort and
-individual trees based models. MAKE REFERENCES TO LITERATURE
+approach. The recruitment rates were sampled from simulations in the
+equilibrium, and in this state it is expected that no climax species
+such as *Pinus sylvestris* or *Quercus* would be less represented
+[@klopcic2015], and most of the models did capture this difference also
+present in the observed data.
 
-Our results do not show differences between models with and without
-feedback in recruitment. The adult forests species richness was better
-captured than the recruitment species richness, and only few models
-overpredicted stand level species richness. Most of the models have
-feedback on recruitment from the adult trees, therefore EXPAND ON
-FEEDBACK ISSUE
+The adult forests species richness was better captured than the
+recruitment species richness, and only few models overpredicted stand
+level species richness. Recruitment composition defined as a
+proportional to adult abundance regardless of productivity might be
+consider a cautious assumption to safely maintain relative species
+abundances [@hanbury-brown2022]. Most of the models have feedback on
+recruitment from the adult trees, however we did not see improvements on
+the species richness representation compared to those models without
+feedback. This might have more to do with the models requirements for a
+successfull regeneration than with the habitat niche of the adult trees
+[@grubb1977].
 
-## Mortality in regeneration
+## Mortality in recruitment
 
 As mentioned above, it is reasonable to expect that models that
-over-recruit trees will also have a couting-acting high mortality
-between the 7 and 10cm thresholds, but this was nit cinsistently evident
-from the simulations. This implies that these models must have the
-excess of mortality in earlier or later stages. Higher mortality in the
-adult component might enhance the share of less shade tolerant species
-for those models. There might be a difference between models including
-stochastic processes in recruitment and those which do not. There was
-also a group of models below the Reineke line indicating a lower
-mortality than expected by the self-thinning rule, which may point to
-facilitation, or simply a higher mortality rate before the trees reached
-7 cm diameter, another possibility would have been that these models
-underestimate recruitment which was not the case. This might also be
-partly an artifact of the method, as we are expecting space for time
-substitution, but in reality we just have one measurement. NEED MORE
-REFERENCES AND DISCUSSION POINTS
+over-recruit trees will also have a cauting-acting high mortality
+between the 7 and 10 cm thresholds, but this was not consistently
+evident from the simulations. This implies that these models must have
+the excess of mortality in earlier or later stages. Higher mortality in
+the adult component do have implications on forest dynamics and the
+expected products forests provide. On the one hand, mortality in later
+stages might enhance the share of less shade tolerant species for those
+models [@klopcic2015] and therefore shifting the species composition. On
+the other hand high densities for a large period of time might
+overestimate the potential role of trees regrowth in carbon sink
+dynamics [@pugh2019]. There are several factors that lead to mortality
+in recruitment such as competition [@casper1997], but there might be
+also a differences in mortality between how models define the process
+such as the inclusion of stochasticity. These observations indicate that
+better reconcile differences between recruitment levels and larger size
+classes densities and better representations of early stage mortality
+processes are needed.
 
-## Research recomendations
+## Research limitations and recommendations
 
@@
 regions. This is significant relevant when looking at mixed forests and
-the species level parametrization. **How this relates to the results?
-Future model development recommendations?**
+the species level parametrization. Based in our results there was not a
+significant improvement on how recruitment was captured with increasing
+complexity in the recruitment modules and even some of the more complex
+models had more dificulties correctly capturing recruitment.
 
-In this study, we considered a wide range of models of forest dynamics,
-from stand to global scale. However, the simulation set up was limited
-to 1 ha recruitment values. It is important to include spatial
-considerations on tree recruitment [@beckage2003] and the lack of it
-might represent an more important disadvantage for landscape level
-models. Without spatial context we are limiting recruitment to the seed
-influx from the stand itself if the model does not have a background
-seed input. But even global vegetation models are lacking this dispersal
-between cells where there is an abstraction of the space or even the
-trees themselves [@hanburybrown_forest_2022]. Landscape level models
-have been build to be very accurately at the landscape level. EXPAND ON
-THE DIFFERENT SCALE LEVELS AND IMPLICATIONS
-
-Modeling approach seems more important than scale and accuracy of the
-design. Model behavior (accuracy) is completely unrelated to model scale
-(stand, landscape, global) and approach (empirical, process-based).
-Having both empricial and process based models, the expectation was that
-the empirically based models (SIBYLA, xComp) would have a better
-performance as they are calibrated with inventory data, however this was
-not the case. EXPAND ON
-
-The selected models also had limitations on species representation, some
-models where not able to simulate certain species due to lack of species
-parametrization. We also had one model that simulate PFTs providing a
-less flexible comparison the with species values; with the lose of the
-species specific information might also represent substantial ecological
-representativeness drawbacks [@purves2008].
-
-**METHODS PARAGRAPH**
+It is important to include spatial considerations on tree recruitment
+[@beckage2003]. In this study, we considered a wide range of models of
+forest dynamics, from stand to global scale. However, the simulation set
+up was limited to 1 ha recruitment values. The lack of large spatial
+considerations might represent a more important disadvantage for
+landscape level models as they have been built to be very accurately at
+the landscape level. Without spatial context we are limiting recruitment
+to the seed influx from the stand itself if the model does not have a
+background seed input. But even global vegetation models are lacking
+this dispersal between cells where there is an abstraction of the space
+or even the trees themselves [@hanburybrown_forest_2022]. However, for
+capturing recruitment modeling design seems more important than scale
+and accuracy. Model behavior (accuracy) is completely unrelated to model
+scale (stand, landscape, global) and approach (empirical,
+process-based). Having both empirical and process based models, the
+expectation was that the empirically based models (SIBYLA, xComp) would
+have a better performance as they are calibrated with inventory data,
+however this was not the case.
 
@@
 
-RECOMENDATION PARAGRAPH
-
-Cautious when using this models to evaluate regeneration post
-disturbances we might be overestimation their capacity for carbon
-capture
-
 ## Conclusions
@@
 Capture properly recruitment rates, species composition and mortality in
-the environmental gradient might impact future forests...
+the environmental gradient might impact future forest. Models of forest
+dynamics are important tools to represent forest dynamics. The 15 models
+and variants used in this study represent a wide range of approaches in
+simulating forest dynamics but also face similar challenges when
+representing tree recruitment. Tree recruitment levels were generally
+overestimated, and recruitment niche was not always correctly captured
+as a function of the biotic (light) and abiotic (temperature and
+moisture) [@bugmann2001]. The initial composition of tree populations
+simulated by the different models was well represented. Surprisingly
+changes in the way trees species are represented in recruitment did not
+have a large impact on the recruitment composition , such as the
+inclusion of feedback from adults trees. Related to mortality in tree
+in-growth, the models with higher over estimation of recruitment also
+had a larger tree mortality however this was not enough to reduce the
+high recruitment densities.
 
-Provide succinct answers to the three questions/objectives stated at the
-end of the
-Introduction. Perhaps end with a concluding (\"more positive\") final
-paragraph.
-
-Models eventually get the right stand structure and composition, most of
-them do that
-by compensating for one error (overestimation of recruitment) by another
-error
-(excessive subsequent mortality, if not between 7 and 10 cm, then
-shortly
-thereafter): \"getting the right answer for the wrong reason\". And this
-is
-worrysome in general, but particularly (and very acutely) if the models
-are to
-be used e.g. under changing climatic conditions.
+A concluding (\"more positive\") final paragraph.
 

### 2022-12-02 manuscript/supplementaryRW.Rmd
supplementary material

---
title: "Supplementary figures\nCapturing regeneration in models of forest dynamics"
author: 
- Olalla Díaz-Yáñez, 
- Yannek Käber, 
- Mats Mahnken, 
- Samuel Fischer, 
- Xavier Morin, 
- Manfred Lexer, 
- Christian Hochauer, 
- Katarina Merganicova, 
- Tobias Mette, 
- Werner Rammer, 
- Kristin H. Braziunas, 
- Paola Mairota, 
- Josef Bruna,
- Heike Lischke, 
- Daniel Scherrer, 
- Simon Scheiter, 
- Thomas Hickler, 
- Tim Anders,
- Jessica Hetzer, 
- Harald Bugmann
date: "`r Sys.Date()`"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    latex_engine: xelatex
    extra_dependencies: ["flafter", "float"]
    toc: yes
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options: 
  markdown: 
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H", out.extra = "")

#knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")
```

# Supplementary figures and tables

## Recruitment levels

```{r envTrend7wb, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the sites water balance. (A) same y-axis across all models; (B) y-axis scaled by model. The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data..", fig.show = 'hold', out.width = '80%'}
    knitr::include_graphics(c(here::here("figures", "envTrend7WB_500.png"),
                              here::here("figures", "envTrend7WB.png")))
```

```{r envTrend7dds, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the site seasonal degree-day sum. (A)same y-axis across all models; (B) y-axis scaled by model. The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
    knitr::include_graphics(c(here::here("figures", "envTrend7DDS_500.png"),
                              here::here("figures", "envTrend7DDS.png")))
```

```{r zeroInflated, echo=FALSE,  message=FALSE}
library(knitr)

zeroInflated <- readr::read_csv(here::here("data", "noRecruitment.csv"))
zeroInflated$Model[zeroInflated$Model == "Empirical"] <- "Observations"

kable(zeroInflated,  format = "latex", caption = "Number of total samples, samples with zero recruitment and the proportion of zero recruitment per model and observed data.")|>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
```

## Recruitment species composition

```{r sigDiff, echo=FALSE,  message=FALSE}
library(knitr)

sigDiff <- readr::read_csv(here::here("figures", "sigH7_10.csv"))

kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of Shannon index in each model for 7 and 10 cm threshold.") |>
        #kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
```

```{r sigH7_Obs, echo=FALSE,  message=FALSE}
library(knitr)

sigDiff7 <- readr::read_csv(here::here("figures", "sigH7_Obs.csv"))

kable(sigDiff7,  format = "latex", caption = "Test of the significance of the differences of Shannon index between the observed data and each model for 7cm threshold.") |>
        #kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
```

```{r sigH10_Obs, echo=FALSE,  message=FALSE}
library(knitr)

sigDiff10 <- readr::read_csv(here::here("figures", "sigH10_Obs.csv"))

kable(sigDiff10,  format = "latex", caption = "Test of the significance of the differences of Shannon index between the observed data and each model for 10cm threshold.") |>
        #kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
```

```{r recrStand7, echo=FALSE, fig.cap="Mean Shannon index per site for the recruitment  and at the stand level both for observed and simulated values. NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
knitr::include_graphics(here::here("figures", "recruitmentAdultRichness7.png"))
```

```{r recrStand10, echo=FALSE, fig.cap="Mean Shannon index per site for the recruitment  and at the stand level both for observed and simulated values. NOTE: The recruitment treshold recruitment for treshold 10 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
knitr::include_graphics(here::here("figures", "recruitmentAdultRichness10.png"))
```

```{r simObadult , echo=FALSE, fig.cap="Mean species richness at the stand level per site in the simulated versus observed species richness. The species richness is estimaed using the Shannon index  calculated by basal area. A1) groups models that Simulated species composition in adults is higher than the observed, A2) groups models that Simulated species composition in adults is similar to the observed, A3) groups models where the imulated species composition in adults is lower than the observed. NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
knitr::include_graphics(here::here("figures", "allAgesRichness.png"))
```

```{r simObr, echo=FALSE, fig.cap="Mean species richness per site in the simulated recruitment versus observed species richness. The species richness is estimaed using the Shannon index is calculated by basal area.B1) groups models where the simulated species composition is closer than the other models to observed values, B2) groups models where the sites with lower species composition are underestimated, B3) groups models where the simulated species composition recruitment is lower in the whole range compare to the observe. NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
knitr::include_graphics(here::here("figures", "recruitmentRichness.png"))
```

```{r diff, echo=FALSE, fig.cap="Departure of each model in the representation of the main species recruitment relative abundance from the observations across the environmental gradients for a 10cm treshold. The values shown are the mean of samples per site and across sites in bins, where the range of the climatic water balance (cm) and seasonal degree-day sum (dºC) was divided in 10 bins. The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species. The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equalt to zero.", out.width='100%'}
knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_10.png"))
```

## Mortality

```{r mortTable, echo=FALSE,  message=FALSE}
library(knitr)

sigDiff <- readr::read_csv(here::here("figures", "mortRatio7_10.csv"))

kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold over 10 cm threshold.") |>
#kableExtra::landscape() 
kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
```

```{r mortality12 , echo=FALSE, fig.cap="Interquartile range of the ratio between 7 and 10cm recruitment and the overestimation proportion (sim-obs/obs) for the mean recruitment trees per model across sites and samples", out.width='100%'}
knitr::include_graphics(here::here("figures", "mort7_10_iqrOver.png"))
```

```{r mortality22 , echo=FALSE, fig.cap="Recruitment ration between 7 and 10cm and the initial recruitment tree number at 7cm. Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions. The cloud point per models shows the differences across sites, the shape point cloud per model shows if the models take into account regeneration density when applying mortality to regeneration ", out.width='60%'}
knitr::include_graphics(here::here("figures", "mort7_10.png"))
```

```{r tableModels, echo=FALSE,  message=FALSE}
library(knitr)

tabletrends <- readr::read_csv(here::here("figures", "mort7_10_trends.csv"))
kable(tabletrends,  format = "markdown", caption = "Models trends in mortality between and 10cm.")
```

```{r mortality3 , echo=FALSE, fig.cap="Recruitment ration between 7 and 10cm and the initial recruitment tree number at 7cm. Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions", out.width='60%'}
knitr::include_graphics(here::here("figures", "mort7_10_trends.png"))
```

```{r mortalityDiv3 , echo=FALSE, fig.cap=" Shannon index mena value across samples per site and in BA difference between 10 and 7 cm ", out.width='100%'}
knitr::include_graphics(here::here("figures", "H710.png"))
```

# Model reports

## 4C

## ForCEEPS & ForCEEPS(f)

### Author

Xavier Morin

### Model references

*Please provide here the references to this model where it is possible
to understand the model structure specially for the regeneration module,
this could be a published paper, a website with the model description, a
manual or some other document that the reader can access.*

### Climate

ForCEEPS requires monthly data, I have thus just randomized the years
given by CHELSA database to obotain a 4000-yr long time-series for each
site, with mean monthly temperature and monthly sum of precipitations.

### Soil

Relying on given information, I used the following relationship for
assessing Soil Field Capacity (SFC) : SFC = 8+(X-1)\*(25-8)/(5-1), in
which X is the 'soil_qual' value in the 'soil_quality_200.csv' file.

### Topography

Not considered.

### Simulations

I have run two kinds of simulations : - Simulations with seed rain from
a potential species list, thus independent from the actual forest
composition ("Forclim-like"), for both thresholds (7 and 10cm) -
Simulations with seed rain depending on the actual forest composition
for both thresholds (7 and 10cm) In each site, the simulation consists
of 4000 yr-long simulation of 1ha of forest (ie. 10 patches of 1000 m²),
starting from bare-ground. There was no dispersal between patches, but
the effect of the abundance of adult trees on the composition of
regeneration was considered in half of the simulations as mentioned.

### Outputs

As requested, I have output species specific regeneration rates per
decade and per ha. More precisely, I have considered only the last 2000
years of simulations (= to ensure that pseudo-equilibrium state is
reached). I have exported the results every 10 years for the last 2000
years at the 1ha-level. My outputs thus rely on temporal variability.
Variables: - Site: numbers corresponding to the climate data file -
Sample: one per decade, for a total of 2000 years = identified from 1 to
200 - Species: name specified as requested (eg. 'fasy') - r.trees:
number of trees outcrossing the threshold (7 or 10 cm) during the
decade, for the 10 patches simulated (= 1ha), per species - r.ba: summed
basal area of the r.trees, per species - ba: summed basal area of all
trees per species at the end of the decade - dbh: threshold used And
there are thus 4 output files: Results_Forceeps_7cm.csv
Results_Forceeps_10cm.csv Results_Forceeps_feedback_7cm.csv (=
simulations with feedback of the actual species composition on the seed
rain) Results_Forceeps_feedback 10cm.csv (= simulations with feedback of
the actual species composition on the seed rain)

## FORMIND

## ForClim 1 & ForClim 11

### Authors

### Model references

*Please provide here the references to this model where it is possible
to understand the model structure specially for the regeneration module,
this could be a published paper, a website with the model description, a
manual or some other document that the reader can access.*

### Climate

### Soil

### Topography

### Tree species

### Simulations

### Outputs

## SIBYLA

### Authors

Ján Merganič1, Katarína Merganičová2,3, Marek Fabrika1, Peter Valent1

1Technical University in Zvolen, Faculty of Forestry, T. G. Masaryka 24,
960 53 Zvolen, Slovak Republic

2Department of Biodiversity of Ecosystems and Landscape, Slovak Academy
of Sciences, Štefánikova 3, P.O.Box 25, 81499 Bratislava, Slovakia

3Czech University of Life Sciences Prague, Faculty of Forestry and Wood
Sciences, Kamýcká 129, 16500 Praha 6 -- Suchdol, Czech Republic

### Model references

*Please provide here the references to this model where it is possible
to understand the model structure specially for the regeneration module,
this could be a published paper, a website with the model description, a
manual or some other document that the reader can access.*

SIBYLA is the simulator of forest biodynamics. It belongs to the
category of semi-empirical tree growth simulators. It consists of the
set of mathematical models and algorithms that are transformed into an
integrated software package SIBYLA Suite. The model has been developed
at the Department of Forest Management and Geodesy, Technical University
in Zvolen, Slovakia. The basis of the model was the modelling principle
and algorithms implemented in SILVA 2.2 (Pretzsch 1992, Kahn 1994). The
model is sensitive to climatic factors (length of growing season, mean
temperature during growing season, annual temperature amplitude, and
total precipitation during growing season). The climatic factors modify
height and diameter growth potential, and consequently tree increment.
The climatic factors also influence tree regeneration model. More
detailed information about Sibyla can be found in Fabrika (2005),
<http://etools.tuzvo.sk/sibyla/english/model.htm>.

### Climate

We used CHELSEA monthly and daily time series from 1981 to 2018 to
derive climate variables necessary for simulations with SIBYLA as
follows: Monthly average temperature (tas) values were used to derive
the SIBYLA site-specific characteristic called the „annual temperature
amplitude" (labelled as TAMPL or s5 in the model) that was calculated as
a difference between the maximum and minimum monthly average
temperatures. Monthly average temperature (tas) values for months April
to September were used to derive the SIBYLA site-specific characteristic
called the „average temperature during the growing season" (labelled as
TEMP or s6 in the model). Monthly precipitation sums (pr) for months
April to September were summed up to derive the SIBYLA site-specific
characteristic called the „precipitation total during the growing
season" (labelled as PRECIP or s8 in the model). Daily average
temperature (tas) values were used to derive the SIBYLA site-specific
characteristic called the „length of the growing season" (labelled as
DAYS or s4 in the model).

### Soil

The information on site-specific soil quality (file name:
soil_quality_200.csv) was used to define two SIBYLA site-specific
characteristic called the „soil moisture" (labelled as MOIST or s7 in
the model) and the „soil nutrient supply" (labelled as NUTR or s3).
Since these two characteristics can obtain values in the range from 0 to
1, the provided values on soil quality were converted to this range by
dividing each value with 5 (the maximum). We used the long-term average
of CO2 concentrations in air that represented the period 1981 -- 2018.
The long-term average of CO2 was calculated from the annual CO2 data
obtained from Mauna Loa observations (Keeling et al. 1976). Similarly,
we used the long-term NOx concentration representing the time period
1981-2018. This was obtained by averaging annual values of NOx
concentration in air calculated using the equation by Kahn (1994), where
the year is the driving variable.

### Topograhpy

The growth simulator SIBYLA uses the model of ecological classification
applied in the growth simulator SILVA 2.2, which was derived by Kahn
(1994). Site is specified using ecological site characteristics
describing climate, air, and soil, also called site variables: • s1
(N2O) ... NOx concentration in air (ppb) • s2 (CO2) ... CO2
concentration in air (ppm) • s3 (NUTR) ... soil nutrient supply
(relative value in the range from 0 to 1) • s4 (DAYS) ... number of days
in the vegetation period (days with daily mean temperature above 10°C) •
s5 (TAMPL) ... annual temperature amplitude (the difference between
minimum and maximum monthly temperature in °C) • s6 (TEMP) ... daily
mean temperature in the vegetation period in °C (from April to
September) • s7 (MOIST) ...soil moisture (relative value in the range
from 0 to 1) • s8 (PRECIP) ... precipitation amount in the vegetation
period in mm (from April to September) • s9 (ARID)... aridity index
according to de Martone in mm.°C-1 They directly influence the
production capacity of a stand (tree height and diameter increment).

### Tree species

At the beginning of the simulations we determined balanced tree species
composition of all 12 tree species pre-defined in the „Regeneration
workshop protocol". It means that at the beginning each species was
represented by the same volume, mean tree diameter, mean tree height,
and age.

### Simulations

The simulations were performed following the "Regeneration workshop
protocol". Hence, we simulated the pre-defined 200 sites with the model
SIBYLA under current climate. The simulations were performed with the
help of two software tools for automatised handling of multiple sites to
speed up the process called DBCreator and Automat (Valent et al. 2018).

#### Simulation area and site

We set the simulation area to 1ha, i.e. each site was represented by an
area of 1 ha, to enable a sufficiently large area for species
interactions. Larger areas were not possible due to the restrictions in
the number of trees per plot (the maximum number of trees per plot is
6,000). Due to the restriction in the maximum number of trees per plot
we established a 40-year-old forest stand at each site with the balanced
tree species composition as defined above. We used site-specific
elevation and aspect obtained from the information on topography and
elevation (file name: topo_data_200.csv).

#### Simulation length

Each site was simulated for 3,000 years. Based on the analysis of the
temporal development of multiple stand variables including number of
trees, mean tree and stand characteristics, the simulation time until
the equilibrium was set to 500 years. Hence, the first 500 simulated
years were excluded from the results. The submitted output includes the
required information per decade during the 2,500 years.

#### Dispersal

Relevant aspects of the dispersal (if any) and seed input parameters in
your model The regeneration model operates if there are mature trees
able to produce seeds. Hence, the initial forest stand had to be
established to allow the reproduction and regeneration module. The seed
yield of individual trees is predicted in relation to species-specific
characteristics including tree age interval of seed production and
fertility (age of start, optimum and end), seed production per adult
tree, transmission distance, seed purity, seed quality, germination
capacity, absolute weight of seeds. Seed tree production is further
modified with the regulators and reduction factors, which describe other
conditions necessary for the germination, survival, and the
establishment of the next generation, e.g. stand canopy closure (Fabrika
2005, Merganič and Fabrika 2009, Fabrika et. al. 2009, 2011).

### Outputs

To derive the required information, we worked with the model output for
individual trees per decade. First, we calculated basal area of
individual trees from the simulated diameter at breast height using the
equation for the area of a circle. Total basal area (ba) was obtained by
summing tree basal areas per species and decade. The number or recruits
(r.trees) per species was derived by counting the trees that exceeded
the diameter threshold (7 or 10 cm) in the specific time. The basal area
of recruits (r.ba) per species was calculated by summing tree basal
areas of recruits per species. NA was assigned if no trees of a specific
species occurred in the period.

### References

Fabrika M., Merganič J., Merganičová K., 2009: Natural regeneration
density model developed for the purposes of the individual-tree growth
simulator. Acta Facultatis Forestalis Zvolen, 51(3), p. 125-137, ISSN:
0231-5785 Fabrika, M., (2005): Simulátor biodynamiky lesa SIBYLA,
koncepcia, konštrukcia a programové riešenie. Habilitačná práca.
Technická univerzita vo Zvolene, 238 p. Kahn, M., (1994): Modellierung
der Höhenentwicklung ausgewählter Baumarten in Abhängigkeit vom
Standort. Forstliche Forschungsber. München, Vol. 141, 221 p. Keeling C.
D., R. B. Bacastow, A. E. Bainbridge, C. A. Ekdahl Jr., P. R. Guenther,
L. S. Waterman, J. F. S. Chin 1976: Atmospheric carbon dioxide
variations at Mauna Loa Observatory, Hawaii. Tellus. XXVIII, 6: 538-551,
<https://doi.org/10.1111/j.2153-3490.1976.tb00701.x> Merganič J.,
Fabrika M., 2009: Determination of climatic-site strata based on
rectified climatic rasters for the purposes of natural regeneration
density model. Acta Facultatis Forestalis Zvolen, 51(3), p. 113-124,
ISSN: 0231-5785 Merganič, J., Fabrika, M., Merganičová, K., 2011:
Submodel of height-diameter function for climatic-site strata of the
model of natural regeneration in the growth simulator SIBYLA. Acta
Facultatis Forestalis Zvolen, 53(1), p.155-168, ISSN: 0231-5785
Pretzsch, H., (1992): Konzeption und Konstruktion von Wuchsmodellen für
Rein- und Mischbestände. Forstliche Forschungsberichte München, Nr.115,
358 p. Valent, P., Merganič J., Výbošťok J. 2018: Optimus -- Software
for Multi-criteria analysis. TU Zvolen.

## xComp

### Authors

Tobias Mette, LWF, 24.05.2022

### Model summary

*Please provide here the references to this model where it is possible
to understand the model structure specially for the regeneration module,
this could be a published paper, a website with the model description, a
manual or some other document that the reader can access.*

The forest growth model xcomp was developed to simulate gap dynamics of
monospecific stands with a strong emphasis on position-dependent
individual tree competition. The model consists of three units: (1) the
main growth phase which determines the growth and mortality of each
individual tree above 5 cm dbh, (2) the old-growth phase where
individual trees exceeding a certain dbh-limit decline in vitality and
eventually die, and (3) the regeneration phase where seedling growth and
mortality is modelled depending on light availability. The primary focus
of the model was to establish a set of robust algorithms that can be
flexibly calibrated for different species attributes. The model was used
to simulate single-species stand development cycles for Nothofagus
pumilio in Patagonia (Mette 2014), for European beech in NE-Bavaria
(Mette et al. 2016) and a Norway spruce stand with different mortality
algorithms (Bugmann et al. 2019). For the current article the model was
(4) connected to a climate sensitive species distribution model to
modify growth and mortality in dependence of the site conditions, and
(5) extended from single to multi-species simulations.

### Climate

### Soil

### Topography

### Tree species

### Simulations

The simulation was carried out according to the protocol by Bugmann et
al. (2022). I.e. the simulation is initiated for mixed species stands of
1 ha size at 200 sites of with given climate and soil conditions. The
initial species mix consists of up to 10 species of equal shares with an
initial root mean square diameter dq of 8 cm. From the suggested species
spectrum Alnus glutinosa and Fraxinus excelsior were not included in the
species mix. Quercus spec was split into Q. petraea and Q. pubescens,
but united in the validation to meet the protocol. Xcomp simulates
forest growth in 5 year intervals. Data are validated after having
reached a climate dependent equilibrium -- which is typically the case
after 750 years (= 150 simulation periods). Starting with year 760, data
are read out for 200 10 year intervals (until year 2750, simulation
period 550).

#### Model details

(1) The main growth phase is determined by the processes growth and
    mortality. Mortality is determined by the Reineke exponent rke and
    the stand density sdi and calibrated over a range of 8 cm to 50 cm
    dq (= root mean square diameter) so that stem number N equals: N =
    sdi*(dq/25)\^rke. A tree's resource acquisition area res (m2) is
    approximated as res = 1e4/ (sdi*(dbh/25)\^rke). As the Reineke
    exponent is fixed to 1.6, species with higher/ lower sdi have a
    lower/ higher resource acquisition area for a given dbh. Potential
    dbh growth dd is calculated by multiplying a reference dbh growth
    with a site specific temperature dependent modifier. The individual
    dbh growth in the stand is further modified by the neighbourhood
    competition.

(2) The old growth phase sets in when a tree passes a certain old growth
    dbh (ogDbh) which depends on the site conditions. Having passed this
    threshold dbh a tree is assigned a mortality period that assumes a
    value not greater the a site independent maximum ogPer. The maximum
    old growth time span ranges from 100 years (birch) to 350 years
    (fir). During the old growth phase dbh growth declines and more
    light becomes available for regeneration.

(3) The regeneration phase allows seedling development on a 1x1 m grid
    over the entire stand. First, a light availability matrix (1x1 m) is
    calculated from a cumulative crown shadow which decreases with
    increasing distance to the tree center. Next, regeneration matrices
    (1x1 m) are created for each of the stand's species. In the initial
    period, the regeneration matrix of each species is empty, but at the
    end of each period, a light dependent dbh growth and mortality is
    calculated for each 1m raster cell. The dbh growth is scaled so that
    under full light 5 cm dbh is reached in a defined regeneration
    period (regnPer) that depends on the site conditions. The default
    value is set to 15 years and equal for all species. Dbh growth is
    modified by the light availability and the species' regeneration
    light demand. This is a decisive difference between the species and
    favours shadow-tolerant trees like fir and beech in the absence of
    large gaps. However, each raster cell is afflicted with a species
    bonus or malus that provides local variation to the global
    regeneration parameters. Mortality in the regeneration raster occurs
    as a random process, the likelyhood of which depends on the climate.

(4) The climate sensitivity was calculated by methods of species
    distribution modelling. Effectively species abundance in European
    NFI-plots (Mauri et al. 2017) was related to climate of the plots
    (Karger et al. 2017) using generalized additive models as in Thurm
    et al. (2018). Input variables were summer temperature mean
    (Jun-Aug), winter temperature mean (Dec-Feb) and summer
    precipitation sum (Jun-Aug). Abundance was transformed into
    prevalence () so that the values for each species can assume a range
    from 0 to 1. The prevalence values modify dbh growth dd, old growth
    dbh ogDbh, the regeneration period regnPer, and the regeneration
    mortality regnMort.

(5) The extension from single to multiple species simulation introduced
    no technical changes to the single-species growth and mortality
    behaviour. Competition as the main interaction between trees depends
    on the crown size and density of the neighbouring trees and can turn
    out an dis/advantage for one or the other species. But species
    calibration parameters were not changed due to mixture-specific
    interaction.

### Outputs

The processes that determine the species composition and turnover in
xcomp can be explained best by an example. In Fig. 1, I displayed the
basal area shares resulting from the simulation of site 434 over a time
span of 2750 years. Tab. 1 shows the species calibration parameters for
the site. It is a very favourable site with modelled species prevalences
of 0.93 (beech), 0.86/7 (spruce, maple), 0.81/2 (fir, ped. oak), 0.76/4
(hornbeam, lim, birch), and 0.70 (pine). Starting with more or less
equal basal area shares and 8 cm dq (root mean square dbh), all species
increase their basal area in the first 50 years. Thereafter increased
competition and starting old growth leads to a decline -- first, in all
of the species but spruce, fir and beech, but after 100 years also in
spruce. Only in fir and beech regeneration balances the loss of older
trees. The result is an equilibrium of a dense fir and beech forest with
high basal area where more light demanding species regenerate too little
to contribute significantly.

#### Discussion

xcomp is a very deterministic model where except for mortality and a
random species regeneration malus/ bonus on the regeneration raster.
This leads to equilibria of one or two species. What species depends
strongly on the light demand of the regeneration which favours fir and
beech in the absence of large gaps creating disturbances. The second
most important parameter is the site suitability which is estimated by
the modelled species prevalence and affects growth and thereby
competitivity. As it is the first time that xcomp is used in a multiple
species application there is room for improvements. The automation of a
robust calibration of the main growth phase is a precondition to
investigate how for instance larger disturbances, a weighted
regeneration potential oriented at the main stand, higher or selective
mortality of the regeneration affect the species composition and
turnover.

### references

1.  Bugmann, H., Seidl, R., Hartig, F., Bohn, F., Brůna, J., Cailleret,
    M., ... & Reyer, C. P. (2019). Tree mortality submodels drive
    simulated long‐term forest dynamics: Assessing 15 models from the
    stand to global scale. Ecosphere, 10(2), e02616.
2.  Karger, D. N., Conrad, O., Böhner, J., Kawohl, T., Kreft, H.,
    Soria-Auza, R. W., ... & Kessler, M. (2017). Climatologies at high
    resolution for the earth's land surface areas. Scientific data,
    4(1), 1-20.
3.  Mauri, A., Strona, G., & San-Miguel-Ayanz, J. (2017). EU-Forest, a
    high-resolution tree occurrence dataset for Europe. Scientific data,
    4(1), 1-8.
4.  Mette, T. (2014) Using Reineke to self-calibrate mortality in
    individual tree based forest growth modelling. Beiträge zur
    Forstwissenschaftlichen Tagung, Dresden, 17.-20.09.2014
5.  Mette, T., Falk, W., Blaschke, M., Förster, B., Walentowski, H.
    (2016). Modelling natural forest dynamics in individual tree based
    forest growth simulators. Beiträge zur Forstwissenschaftlichen
    Tagung, Freiburg, 26.-29.09.2016
6.  Thurm, E. A., Hernandez, L., Baltensweiler, A., Ayan, S.,
    Rasztovits, E., Bielak, K., ... & Falk, W. (2018). Alternative tree
    species under climate warming in managed European forests. Forest
    Ecology and Management, 430, 485-497.

Tab. 1: Species calibration parameters for site 434. Spec = species
code, rke = Reineke Exponent, Sdi = stand density index, ogPer =
(maximum) old growth time span (years), regnLgh = regeneration light
demand (canopy shadow value of xcomp), plot_id = site (434), tjja =
summer temperature (Jun-Aug, °C), sitePred = modelled species prevalence
for site (0...1 = min...max), dd = upper diameter growth (cm/ 5 years,
maximum dbh growth up to 85 % higher), ogDbh = old growth dbh (cm, dbh
at which old growth phase begins), regnPer = (minimum) regeneration time
span (year, until seedling reaches 5 cm dbh), regeneration mortality
(climate dependent) Spec Rke Sdi ogPer regnLgh Plot_id Tjja sitePred Dd
ogDbh regnPer regnMort Abal -1.6 1000 350 4 434 0.81 2.8 32.4 17 0 Acps
-1.6 800 150 3 434 0.86 1.8 24.6 17 0 Betu -1.6 700 100 2 434 0.76 1.7
12.8 17 0.001 Cabe -1.6 700 125 3 434 0.76 1.4 16.7 17 0.001 Fasy -1.6
800 150 4 434 0.93 1.9 24.6 17 0 Piab -1.6 1000 350 3 434 0.87 2.9 32.4
17 0 Pisy -1.6 800 250 2 434 0.70 2.0 24.6 17 0.01 Qupe -1.6 700 300 2
434 0.82 1.4 24.6 17 0 Qupu -1.6 700 300 2 434 - - - 0 Tico -1.6 700 150
2.5 434 0.74 24.6 17 0

Fig. 1: Basal area shares in xcomp simulation of site 434 over a time
span of 2750 years.

Fig. 2: xcomp canopy shadow at site 434 after 150 simulation periods
(750 years).

Fig. 3: Histogram of xcomp canopy shadow (vj_lgh) at site 434 after 150
simulation periods (750 years). Species can only establish at sites
where the canopy shadow is lower than their regeneration light demand
(regnLgh in Tab. 1)

## PICUS

### Authors

Christian Hochauer, Manfred J. Lexer

### Model summary

*Please provide here the references to this model where it is possible
to understand the model structure specially for the regeneration module,
this could be a published paper, a website with the model description, a
manual or some other document that the reader can access.*

### Climate

We used the provided monthly climate data (monthly_era5_chelsa_200.csv).
The provided solar radiation rsds was converted to MJ per m².
Additionally, vpd was calculated.

### Soil

We used soil parameters as suggested by the protocol. Between given
class specific values linear interpolation was sued to find
between-class values. Soil Class N WHC pH 1 40 80 4.2 2 - - 4.2 3 - -
4.6 4 - - 5.6 5 100 250 5.6

### Topography

### Tree species

The species composition was defined by the given species from the
protocol: Fagus sylvatica, Picea abies, Abies alba, Carpinus betulus,
Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus excelsior,
Quercus spp., Alnus glutinosa and Pinus sylvestris. For Betula spp.
Betula pendula was simulated, for Quercus spp. the species Quercus
petraea and Quercus robur where used. The establishment from bare ground
was achieved via external seed input (see also in section seed
dispersal), so depending on the environment and random components all
the given species had a chance to grow into the respective simulation.
In addition, on site production of seed is considered depending on age
and size of potential parent trees.

### Simulations

#### Simulation Area

In PICUS the area is defined via the number of 10 m by 10 m patches. For
this simulation we used an area of 12 by 12 patches resulting in an area
of 1.44 ha. This size of the simulated forest provides relatively stable
species composition & structure and reasonable simulation time.

#### Seed dispersal

For the experiment seed input to the simulated forest happens via 2
pathways: (1) external seed input to every patch in any simulated year
(fixed number of seed per patch per species). (2) locally produced seed
by parent trees growing in the simulated forest. Seed production depends
on species, and tree size and a random component accounting for mast
years. Seed dispersal depends on whether a species is anemochorous or
zoochorous. If anemochorous tree height and seed characteristics
determine the maximum dispersal range and the shape of the dispersal
function. Based on germination rate, the environmental factors
(temperature, soil moisture, Nitrogen, light) and the amount of seeds in
combination with a random component seedling establishment is calculated
per patch in any given year. Seedling cohorts grow through 4 height
classes and are then eventually recruited into the tree population.

#### Simulation length

We decided to use the temporal sampling. We first run the model into a
quasi equilibrium and continue afterwards with another 2000 years. The
simulation time until equilibrium was chosen with 600 years from
experience in other projects and a few simulations with test sites where
600 years was a sufficiently long simulation time to reach an
equilibrium state.

### Output

r.trees Counted the number of trees which were not present in the stand
from the previous time period (10 years before) and where above the dbh
threshold per species. For the oaks the value of the two species was
summed into one. The result was then divided by 1.44 to get a per ha
value of r.trees. r.ba Sum of the basal area (circle area calculated
with the dbh) of trees which were not present in the stand from the
previous time period (10 years before) and where above the dbh
threshold. This was then divided by 1.44 to get a per ha value. ba Sum
of the basal area (circle area calculated with the dbh) of all trees
above the dbh threshold. This was then divided by 1.44 to get a per ha
value.

## iLand

### Authors

Werner Rammer Kristin Braziunas

### Model summary

*Please provide here the references to this model where it is possible
to understand the model structure specially for the regeneration module,
this could be a published paper, a website with the model description, a
manual or some other document that the reader can access.*

### Climate

• What climate data have you used, and how have you created your model
climate input parameters?

Climate inputs for iLand are at daily temporal resolution. We used daily
climate from the CHELSA dataset for maximum and minimum temperature,
precipitation, and surface shortwave radiation. We calculated daily
vapor pressure deficit (VPD) from these inputs assuming daily minimum
temperature is dewpoint temperature, a reasonable assumption except
under very arid conditions (Campbell and Norman 2000). We calculated
vapor pressure from temperature using the Tetens formula with different
coefficients for temperatures above or below 0° C (Murray 1967). We then
followed methods outlined by the Food and Agriculture Organization of
the United Nations (Allen et al. 1998) to calculate vapor pressure
deficit as the difference between average daily saturation vapor
pressure (derived from minimum and maximum temperature) and ambient
vapor pressure derived from dewpoint temperature. Some daily VPD values
were negative because there were days in the CHELSA dataset where
minimum temperature exceeded maximum temperature; negative VPD values
were set to 0.

As recommended by workshop organizers, we filled in daily values for two
missing days in 2013 (Jan 2 and Jan 3) with daily values for Jan 1 and
4, respectively.

Atmospheric CO2 concentration was set to 380 ppm and latitude was set to
50° N for all sites. Annual climate year was drawn randomly with
replacement from the full dataset (1981-2018) for each site for the
duration of the simulation.

### Soil

### Topography

### Tree species

• How have you defined the species composition in your model simulation?
(e.g., species levels, PFTs, etc.)? Please describe the species
categories used.

Tree composition is defined at the species level. All individually
requested species (Fagus sylvatica, Picea abies, Abies alba, Carpinus
betulus, Tilia cordata, Acer pseudoplatanus, Fraxinus excelsior, Alnus
glutinosa, Pinus sylvestris) were included. For the two taxa where
species was not defined, we identified a representative species (Betula
pendula for Betula spp. and Quercus robur for Quercus spp.). No other
species were included in model simulations.

### Simulations

#### Simulation area

• How have you defined the simulation area for your model (e.g., size
and other relevant site aspects)?

The simulation extent for each site was 1000 m x 1000 m (total area =
100 ha). Environmental conditions were homogeneous for each site, and we
did not use topographic information. Inputs for soil depth, texture, and
fertility (plant available nitrogen) were extracted from a previously
compiled Europe-wide dataset. We averaged values in a 1 km buffer around
blurred site coordinates provided by workshop organizers. Relationships
between ranked site quality provided by workshop organizers and soils
data from our Europe-wide dataset were weak (Pearson's \|r\| ranged from
0.04 to 0.34), but followed expected trends with lower sand content,
higher water holding capacity, and higher fertility associated with
higher site quality. Albeit the weak correlation, we chose to use our
available data set because we could avoid additional assumptions as it
included all site variables needed for the model.

#### Dispersal

• What are the relevant aspects of the dispersal (if any) and seed input
parameters in your model?

Dispersal parameters are species-specific and include maturity age,
fecundity (seedling potential per m2 crown area), masting (lower seed
production in non-mast years), and dispersal distance (probabilistic
dispersal kernel). A small fraction of seeds are dispersed over longer
distances (long distance dispersal algorithm). Dispersal is calculated
at 20 m spatial resolution. Seedling establishment is further modified
by other environmental constraints and is stochastically spatially
distributed at finer (2 m) resolution. The simulation extent is treated
as a spatially explicit landscape, so neighboring cells can serve as
seed sources for each other.

We initialized our simulation experiments from bare ground with uniform
probability of seed availability set to a low level (0.001) for all
species. Once trees reached maturity, they also contributed to seed
availability and dispersal.

#### Simulation length

• What is the simulation length, and how have you determined the
simulation time until the equilibrium?

Simulations were run for 2010 years, and equilibrium was reached at year
1500. We assessed temporal patterns of species composition and basal
area and defined equilibrium as when forest composition and structure
stabilized across all sites ("potential natural vegetation"). Some sites
had high interannual variation, especially in basal area, after
equilibrium; this variation did not represent directional change and
appeared instead to be related to sites with extreme growth conditions
and variation in randomly selected annual climate.

#### Simulation approaches

• If you have run more than one simulation per site, please describe the
differences of the approaches used.

Only one simulation was run per site.

### Output

• Please describe how have you calculated each of the variables in the
output data

For each site, we generated decadal outputs for the full simulation
extent every 100 years after equilibrium was reached (i.e., 1500-1509,
1600-1609, ..., 2000-2009). This resulted in outputs for a total of 600
ha (100 ha simulation extent x 6 decades). From this list, we randomly
sampled 200 ha without replacement.

We derived recruit density and basal area annually by species. Trees \>
4 m in height are represented as individuals in iLand, so we were able
to identify individual trees that crossed the DBH threshold of either 7
or 10 cm each year. We summed the number of annual recruits and their
basal area in the year they crossed the recruitment threshold for each
10-year output period.

We averaged total density and basal area by species for each decade. We
further included 0 values (complete cases) for species that were missing
from a given sample and site.

### References

Allen, R. G., L. S. Pereira, D. Raes, and M. Smith. 1998. Crop
evapotranspiration - Guidelines for computing crop water requirements.
FAO - Food and Agriculture Organization of the United Nations,
Irrigation and drainage paper 56. Campbell, G. S., and J. M. Norman.
2000. An Introduction to Environmental Biophysics. Springer Science &
Business Media. Murray, F. W. 1967. On the computation of saturation
vapor pressure. Journal of Applied Meteorology 6:203--204.

## LandClim

### Authors

### Model

*Please provide here the references to this model where it is possible
to understand the model structure specially for the regeneration module,
this could be a published paper, a website with the model description, a
manual or some other document that the reader can access.*

### Climate

### Soil

### Topography

### Tree species

### Simulations

### Outputs

## Landis II

### Authors

Josef Brůna, Paola Mairota, Marco Mina, Giorgio Vacchiano

### Model

*Please provide here the references to this model where it is possible
to understand the model structure specially for the regeneration module,
this could be a published paper, a website with the model description, a
manual or some other document that the reader can access.*

Landscape scale forest simulation model LANDIS-II (Scheller et al. 2007)
with PnET-Succession extension (version 4.1). PnET-Succession is based
on the Biomass Succession extension of Scheller and Mladenoff (2004),
embedding elements of the PnET-II ecophysiology model of Aber et al
(1995) to mechanistically simulate growth as a competition for available
light and water (De Bruijn et al. 2014). PnET-Succession simulates the
competition among cohorts for water and light at a monthly time-step as
a function of photosynthetic processes and maintenance respiration that
are explicitly linked to environmental drivers such as temperature,
precipitation, photosynthetic active radiation (PAR), and CO2
concentration. Regeneration depends on distance from a seed source, soil
water, and subcanopy light, while competition is modeled by partitioning
incoming light through multiple canopy layers. Growth of specific cohort
biomass components is allocated in the following order: non structural
carbon, first, then foliage then root and stem. Biomass growth decreases
as cohorts approach their longevity age but mortality can occur at any
time when carbon reserves production is insufficient to support growth
due to shading, water competition, and/or drought (Gustafson et al.
2015), or by any disturbance or management extension used. For output,
we selected Biomass Community Output Extension version 2.0.1, which
generates output in the format: species age biomass (g / m2) at
specified time steps.

### Climate

Climate input was generated by randomly selecting entire years from the
monthly data from the CHELSA dataset and collating them one after the
other to create a 2000 years series. The sampling was repeated 10 times
with different random order of years to create the 10 climate
replicates. Selecting a year at a time eliminates inconsistent seasons
(especially summer). We used monthly mean, minimum and maximum
temperatures and precipitation as input for PnET succession extension.
PnET-Succession also requires mean monthly values of PAR during daylight
hours, which we calculated by dividing the monthly rsds data by the
number of daylight seconds in each month for latitude 47 to get W/m2. We
then multiplied this by 4.6 to get μmol/m2/s and multiplied by 0.5 to
get PAR part of the spectrum, based on the approximation by Tsubo et al.
(2005). Leap years were not considered. The PAR distribution across all
sites and years is in range with values that are typical for temperate
regions (Fig. 1).

Figure 1 Histogram of monthly PAR values for all sites and all available
time

CO2 concentration with yearly timestep was added from the concentrations
table available from EEA (2019). For years 0-1750 CO2 concentration was
maintained fixed to 278 ppm. As for the period 1750-1975 there was only
one value each 5 years, we used a constant value for each 5-year period
(Fig. 2). CO2 concentration was added to the climate input file after
randomization of temperature, precipitation and PAR (see above), so it
reflects realistic values for the simulated years. Simulations were run
from year 0 to 1990 but sampled outputs did not exceed year 1910 (see
Simulation length and descriptions further below). Thus, the drastic
increase in CO2 concentration during the second half of the 1900s did
not affect the delivered results.

Figure 2 CO2 concentration values added to the climate data

### Soil

### Topograhpy

### Tree species

As LANDIS-II cannot start from bareground, initial communities were set
to include all species defined in the protocol at all sites starting
with 10-years old cohorts of the following species: fasy (Fagus
sylvatica), pabi (Picea abies), abal (Abies alba), cabe (Carpinus
betulus), tico (Tilia cordata), acps (Acer pseudoplatanus), betu (Betula
spp.), frex (Fraxinus excelsior), quer (Quercus spp.), algl (Alnus
glutinosa), pisy (Pinus sylvestris). The species-specific parameter
denoting sexual maturity was set to 10 years equally for all species, so
that seed source of all species is available from the onset of the
simulations at all sites and to maintainseed in the ladscape pool.
Vegetative regeneration (resprouting) was prevented, since this would
have an effect on allometry and regeneration. Some species may be under
or overrepresented because their occurrence in LANDIS-II simulations is
typically due to small-scale disturbances or management which were not
included in the simulations. We have started with a set of parameters we
have used for Petter et al. (2020) which were further refined and
modified for PnET v 4.1. Describe species parameters here...

### Simulation

Simulation area was defined as a rectangle of 10×20 dimension with sites
(i.e., cells) of 1-ha. Each site is an individual ecoregion (i.e., one
cell - as in our case - or group of cells sharing similar climate and
soil conditions, as a way in LANDIS-II to define differences in
biophysical conditions across landscapes) with specific values for soil
quality, topography and climate, based on the data provided. Soil type
was set to LOAM at all sites. Rooting depth was calculated from soil
quality codes (1-5, original values ranged from 1.13 to 4.87)
multiplicated by 205. This produces a soil depth in the range 231 - 998
mm, that are within the typical range for PnET-Succession (maximum 1000
mm). A clear trend emerged - deeper soils resulted in higher biomass,
although with some exceptions (Fig. 3).

Figure 3 Biomass of all species on each site ordered by soil quality
(data from replicate 1). Notice a gap after first longevity period, an
artifact of spin-up from "bare ground"

The ecoregion-specific parameter Precipitation Loss Fraction denotes the
water lost to runoff induced by topographic slope. We estimated this
parameter for each site from topo_data_200 dataset as slope \* 0.01
(e.g., 12% slope = 0.12 PrecLossFrac). We acknowledge that this is a
simplification and that better parameterization would be possible if
soil types were available for each site. Tests showed that sites with
extreme slope tended to have lower biomass (Fig. 4), but the signal was
not that clear. The correlation between slope and elevation was 0.6
which could obscure the results.

Figure 4 Biomass of all species on each site sorted by slope, which was
used to modify the fraction of precipitation lost to run off (data from
replicate 1)

Some sites at very high elevation had, correspondingly, very low
temperature (Fig. 5). At these sites, the model showed a visible trend
in producing lower biomass (down to zero at a few sites see Fig. 6
below). Yet this was probably not apparent in the regeneration results,
as extremely low biomass could still support regeneration, even if
cohorts do not live long, due to our modification of sexual maturity
age.

Figure 5 Histogram of elevation of simulated sites.

Figure 6 Biomass of all species on each site ordered by elevation (data
from replicate 1)

#### Dispersal

We have used the non-spatial seeding algorithm "universal dispersal" for
all the model runs. This way, every species can seed any forest site in
the landscape (across all the 200 sites). This mode does not take into
account the spatial arrangement and contiguity of cells (i.e., no active
seed dispersal), which is proper for the sake of the model comparison
exercise. Yet in real landscape seed limitation can also play a role in
species composition. The seed number is affected by species-specific age
of maturity, which we have modified (see Species section). This setting
ensures an influx of seeds from outside consistent with the original
settings. Simulation length We started each model run with initial
communities of all species at age 10 at all sites. Each run was
simulated for 1980 years. Sampling of model outputs was executed in two
time windows starting 50 years after the first cycle following typical
species longevity (see Species parameters). After the year 700, biomass
of most species was relatively stable and we sampled every 50 years,
starting at 750 until the year 1200. The next sampling period started
from 1450 and continued until 1900, i.e., after the next longevity peak
(Fig. 7), again every 50 years (see also Table 1). At each selected
sampling time, we extracted cohorts of each species according to the age
related to the two DBH thresholds (7cm/10cm) (see age-DBH relationships
in Table 2). Additionally, we extracted cohorts that were up to 9 years
older than these thresholds, to get the expected results for a 10-year
interval. For example, Fagus sylvatica reaches DBH 7cm at 17 years,
therefore we have selected all cohorts with age 17-26.

Figure 7 Biomass of each species in time across all sites (data from
replicate 1)

Figure 8 Sum of biomass of all species on each site in time (data from
replicate 1). Notice again the biomass decrease before approximately the
year 700 due to the effect of species longevity. The productivity of the
sites seem to be stable.

#### Sampling design

We have sampled each site 20 times along the simulation run as described
in the simulation length chapter. To deliver more than one simulation
per site, we created 10 replicates with different climate data and
random number generator seed number to account for stochastic processes
in PnET-Succession. This resulted in the expected 200 samples for each
site. Output variables

sample Sample number code (1-200) was defined as sample in time +
(replicate-1)\*20, with sample in time (1-20) according to the sample
order. Replicate (1-10) combines random number and climate sequence (see
Table 1. For example sample number 25 was taken from the climate
replicate 2 and the fifth sample in time (year 950).

Table 1 Sample numbers and their corresponding settings

r.trees As LANDIS-II outputs are expressed in terms of cohort age and
mean biomass per site, we have calculated the mean DBH and biomass of
each species based on age. We estimated the age when each species
reaches 7 cm and 10 cm DBH and the number of individuals using
allometric models from the Italian forest inventory collected in R
package ForIT (Puletti et al. 2014), and unpublished data from the study
by Mina et al. (2021) (Table 2). Please note that the equations used
predict DBH \>0 already for a tree with age 1, this is because it is
based on forestry data and tree cores. The real age when reaching 1.3 m
height is species specific and we did not implement this offset. For
each cohort, we divided the cohort biomass reported by LANDIS-II,
recalculated to kg/ha by the mean biomass of a tree of the same species
and age to obtain the number of trees of each species per site. We
rounded all results up to the nearest integer to avoid any fraction of
trees. The number of individuals was aggregated for all the cohorts of
the age of reaching 7 cm or 10 cm DBH and all cohorts of the same
species that were up to 9 years older (Fig. 9). This approach does not
reflect influence of site conditions or stand densities on Age-DBH and
Age-Biomass. Relationships between age and DBH were derived by fitting
asymptotic regressions using individual tree measurements from an
ensemble of forest inventory plots distributed across south-eastern
Canada (see Supporting information of Mina et al. 2021). This was the
only public dataset for which we could relate tree age, derived from
stem coring and tree ring analysis, to measured DBH. Although tree age
measurements were obtained with a robust methodology, these functions do
not disentangle the influence of site conditions or stand densities.
Additionally, they reflect growing conditions of tree genera in Eastern
North America.

Figure 9 Number of trees in regeneration cohorts (both 7cm and 10 cm
groups) according to species. (data from all replicates and samples)

Table 2 Age at which each simulated species reaches DBH 7cm and 10 cm
Age at DBH 7cm Age at DBH 10cm Comment fasy (for Fagus sylvatica) 17 25\
piab (for Picea abies) 8 12\
abal (for Abies alba) 10 15\
cabe (for Carpinus betulus) 15 23 same equation as betu tico (for Tilia
cordata) 15 23 equation taken from quer acps (for Acer pseudoplatanus)
16 24\
betu (for Betula spp.) 15 23\
frex (for Fraxinus excelsior) 10 15\
quer (for Quercus spp.) 15 23\
algl (for Alnus glutinosa) 15 23 same equation as betu pisy (for Pinus
sylvestris) 9 13

r.ba basal area of regeneration Calculated as a sum of r.ba of all the
cohorts selected in r.trees using the following formula: DBH2 \* π/4 \*
r.tree (Fig. 10)..

Figure 10 Basal area of trees in regeneration cohorts (both 7cm and 10
cm groups) according to species. (data from all replicates and samples)

ba: total basal area We have not calculated this since our allometric
equations do not cover older ages. Therefore, we produced the following
two biomass variables as a surrogate.

sum.bio\
Biomass of all cohorts of the species on site including the regeneration
and even younger regeneration (units kg/ha).

r.bio.sum Biomass of the regeneration cohorts corresponding to the r. ba
and r.trees on site (units kg/ha).

#### Runtime

Simulations on a 16 core virtual computer (each core 2.10 GHz) with 10
replicates running in parallel took about 3 hours 15 minutes. Processing
the results including allometry for selected cohorts took about an hour.
The model runs with a yearly timestep using monthly climate.

### Outputs

### References

Aber J. D. & Federer C. A. (1992). A generalized, lumped-parameter model
of photosynthesis, evapotranspiration and net primary production in
temperate and boreal forest ecosystems. Oecologia 92(4): 463-474. Aber
J.D., Ollinger S.V., Federer C.A., Reich P.B., Goulden M.L., Kicklighter
D.W., Melillo J.M. & Lathrop R.G. Jr. (1995) Predicting the effects of
climate change on water yield and forest production in the northeastern
United States. Climate Research 5(3): 207--22. de Bruijn A., Gustafson
E.J. , Sturtevant B.R., Foster J.R., Miranda B.R., Lichti N.I. & Jacobs
D.F. (2014) Toward more robust projections of forest landscape dynamics
under novel environmental conditions: Embedding PnET within LANDIS-II.
Ecological Modelling (287): 44--57 EEA (2019) Trends in atmospheric
concentrations of CO2 (ppm), CH4 (ppb) and N2O (ppb), between 1800 and
2017 available at
<https://www.eea.europa.eu/ds_resolveuid/217c026ca03649398aadf39d87623e57>
Gustafson, E. J., A. M. G. De Bruijn, R. E. Pangle, J.-M. Limousin, N.
G. McDowell, W. T. Pockman, B. R. Sturtevant, J. D. Muss, and M. E.
Kubiske. 2015. Integrating ecophysiology and forest landscape models to
improve projections of drought effects under climate change. Global
Change Biology 21:843-856. Mina, M., Messier, C., Duveneck, M., Fortin,
M. J., & Aquilué, N. 2021. Network analysis can guide resilience-based
management in forest landscapes under global change. Ecological
Applications, 31(1), 1--18. Petter G., Mairota P., Albrich K., Bebi P.,
Brůna J., Bugmann H., Haffenden A., Scheller R.M., Schmatz D.R., Seidl
R., Speich M., Vacchiano G. & Lischke H. (2020) How robust are future
projections of forest landscape dynamics? Insights from a systematic
comparison of four forest landscape models. Environmental Modelling &
Software 134: 104844. <https://doi.org/10.1016/j.envsoft.2020.104844>
Puletti N., Mura M., Castaldi C., Marchi N., Chiavetta U. and Scotti R.
(2014). ForIT: Functions from the 2nd Italian Forest Inventory (INFC). R
package version 1.0. <https://CRAN.R-project.org/package=ForIT>
Scheller, R. M., Domingo J. B. , Sturtevant B. R. , Williams J. S. ,
Rudy A. , Gustafson E. J. & Mladenoff D. J. (2007) Design, development,
and application of LANDIS-II, a spatial landscape simulation model with
flexible temporal and spatial resolution. Ecological Modelling 201(3-4):
409--419. <https://doi.org/10.1016/j.ecolmodel.2006.10.009> Tsubo M. &
Walker S. (2005) Relationships between photosynthetically active
radiation and clearness index at Bloemfontein, South Africa. Theoretical
and Applied Climatology 80: 17--25.
<https://doi.org/10.1007/s00704-004-0080-5>

## TreeMig

### Authors

### Model

*Please provide here the references to this model where it is possible
to understand the model structure specially for the regeneration module,
this could be a published paper, a website with the model description, a
manual or some other document that the reader can access.*

### Climate

What climate data have you used, and how have you created your model
climate input parameters? I used the monthly time series of tas and
precipitation,additionally slope and aspect from the topo- data set, and
soil quality from the quality data set. slasp= 2*cos(aspect* pi/180)
\*min(1, slope/60)

### Soil

The soil quality I translated into AWC (of  upmost m soil, "bucketsize")
by Bucketsize = 10 + (soil_quality[-1) /4  \* 25 , to covert he range
from 10 cm to 35cm   With these values I calculated the yearly time
series of the bioclimatic variables DDsum (day degree sum above 5.5°C) ,
minWinterTemp (lowest mean monthly temperature) and Drought stress
according to a bucket model, that estimates a monthly PET following
[Thorntwaithe, 1957],  and calculates the monthly soil water content
   based on soil water content in the previous month, precipitation,
interception, and bucketsize   [Bugmann , 1994] . Drought stress is then
given by 1- the ratio between demand (PET-interception) and the supply
(a function of the soil water content). Extending the climate data I
used the monthly data as given, calculated the bioclimate, and extended
since the end of the climate data to get 1000 years, by sampling from
the given data. I used always the same sequence of sampling years in all
simulations and sites. (-\> better comparability of experiments)

### Soil

### Topography

### Tree species

How have you defined the species composition in your model simulation?
(e.g., species levels, PFTs, etc.)? Please describe the species
categories used. Species level, 30 species (see list of species I sent).

### Simulation

How have you defined the simulation area for your model (e.g., size and
other relevant site aspects)? A grid of 15 \* 15 grid cells, each grid
cell with 200m side length.

#### Dispersal

What are the relevant aspects of the dispersal (if any) and seed input
parameters in your model? Seed dispersal is described by species
specific, height dependent seed production. The seeds are then
transported according to a double negative exponential, species specific
kernel, and enter the seed bank in the sink cells, from which seedlings
germinate depending on light and other environmental conditions. At the
start of the simulation, for 10 years seeds of all species are available
in all grid cells, from then on, the normal seed dispersal starts.

There are 5 different experiments related to dispersal NoDisp:  the
dispersal kernels are so narrow, that the huge majority of seeds stays
in the source grid cell. I.e. the spatial variability (see below) only
arises from the sampling of the light classes. But there is still the
feedback between the mother trees and the seedlings. StochDisp:
stochastic dispersal. The seeds are distributed randomly, whereby the
kernel gives the probability of the seeds  landing in a certain distance
DetDisp: deterministic dispersal. The kernel determines the share of the
seeds in the source cell landing in a sink cell in a given distance to
the source cell. FBirthAll: a fixed ingrowth above 1.37m of 20
trees/833m\^2  and year, same for all species, but reduced by
environmental drivers and shade.  This is analogous to the approach of
ForClim,  i.e. ignores the feedback  between mother trees and young
trees and the youth development FBirthIndiv: as FBirthAll, but the fixed
ingrowth is species-specific  (between 4 and 54 trees/833m\^2 )

#### Simulation length

What is the simulation length, and how have you determined the
simulation time until the equilibrium? 1000 years, experience and check
by eye-ball of the simulated basal areas. Variability: I have employed
three different aspects of the model  to account for spatial and
temporal variability and to get the required 200 patches of 1ha.

1)  Basic spatial variability due to the combination of demographic
    stochasticity and short-range spatial interactions. In contrast to
    gap models, TreeMig is based on the concept of frequency
    distributions of tree densities classes in different tree heights on
    ca. 1/12 ha (833m\^2) patches, and the resulting frequency
    distributions in discrete light. During the simulation, in each
    height class the frequencies of the different light classes are
    recorded, as well as the ingrowth of each species in the
    height-light class. By sampling 12 times from the height-light class
    ingrowths according to the height-light class frequency
    distribution, and summing up, the stochastic ingrowth for one ha is
    calculated, for each grid cell.  Then the height class boundaries
    are translated into DBH and the ingrowths in the height classes
    linearly interpolated to 7cm and 10cm.

2)  Spatial smoothing and variability by the formulation of dispersal.
    The different dispersal strategies, see above. With NoDisp, solely
    the basic spatial variability is accounted for. With DetDisp, I
    expect the dispersal to smooth out the variability. StochDisp, in
    turn is expected to create additional variability by stochastic
    founder processes.

3)  Temporal variability by sampling from different decades of the last
    100 years of the  simulation, i.e. years 900-1000

#### Regeneration parameters

How does environment affect regeneration in the model? Seed production:
depends on tree size and thus indirectly on the growth, that  depends on
DDsum,  drought stress, nutrients and light and on adult tree number and
thus on survival  that depends on DDsum, drought stress, nutrients and
light

Seed dispersal:  independent on environment Seed bank dynamics : 
independent on environment Germination and seedling survival: Depends on
drought, browsing, winter temperature and DDSum and light (last 3 are
step functions) Sapling dynamics (up to 1.37): like adult growth and
survival, i.e. depends on DDsum, drought stress and light but light
dependence with a specific parameter values, namely the klights
parameter instead of klighta

#### more simulations per site

If you have run more than one simulation per site, please describe the
differences of the approaches used. I used the following 15
combinationsExperiments (exp_txpe) NoDisp, StochDisp, DetDisp,
FBirthAll, FBirthIndiv), each with the variability setups (comb_type)
200space1time: 200 grid cells at a single randomly chosen decade (
50space4time: 50 grid cells at 4 randomly chosen decades, or
20space10time: 20 grid cells at all 10 decades Always the same random
(nested) sets of decades and grid cells were used Fixed ingrowth values
(for the FBirthAll and FBirthIndiv experiments ) were calculated
beforehand in a  set of simulations with single species under optimal
conditions, without interspecific, but under intraspecific competition.
The ingrowth into the height class above 1.37m per species was averaged
over the last 100  years of the 1000 year simulation and used in the
FBirthIndiv simulations. For the FBirthAll simulations the mean of the
single species ingrowths' (20) was used.

The "standard" combination is DetDisp with 20 grid cells at all 10
decades Please describe how have you calculated each of the variables in
the output data. During the simulation,  in each height class the
frequencies of the different light classes are recorded, as well as the
ingrowth numbers of each species in the height-light class. By sampling
12 times from the height-light class ingrowth numbers according to the
height-light class  frequency distribution, and summing up, the
stochastic ingrowth number for one ha is calculated, for each grid
cell.  Then the height class boundaries are translated into DBH and the
ingrowths in the height classes linearly interpolated to 7cm and 10cm
based.  The ingrowth basal areas into the height classes are obtained by
multiplying the interpolated ingrowth numbers with the basal areas at 
7cm and 10cm, i.e. r.trees \* Pi *(0.07/2)\^2  and r.trees* Pi\* 
(0.1/2)\^2 .

The species specific basal area (at 1.37m) sums are a standard output of
TreeMig, given by multiplying the state variables (number of trees per
grid cell , height class, and species) with  the species specific basal
area of each height class. Files TreeMigRes_all_data.RData: R RData-file
with all experiments (exp_type) and variability combinations (comb_type)
TreeMigRes_Std_data.RData: R RData-file with   exp_type=detDisp) and
variability combinations (comb_type=20space10time)

### Outputs

## LPJ-GUESS

### Authors

### Model

*Please provide here the references to this model where it is possible
to understand the model structure specially for the regeneration module,
this could be a published paper, a website with the model description, a
manual or some other document that the reader can access.*

We used the process-based mechanistic model LPJ-GUESS to simulate
vegetation dynamics considering tree species, age cohorts, gap dynamics
and biogeochemical cycles. Parameters of the model reflect an updated
version of the most common European tree species as well as typical
shrub PFTs as described by Hickler et al. (2012). A detailed description
of LPJ-GUESS is available in Smith et al. 2001 and Smith et al. 2014.

### Climate

What climate data have you used, and how have you created your model
climate input parameters? We used the provided daily time series of
climatic variables from 1981 to 2018. The climatic variables base on the
CHELSA data set with a spatial resolution of 0.008° x 0.008° (\~ 1 km)
(Karger et al., 2021). Daily average temperature (tas) [°C], daily
average maximum temperature (tasmax) [°C], daily average minimum
temperature (tasmin) [°C], precipitation sum (pr) [mm] and surface solar
(shortwave) radiation (rsds) were taken to set up the LPJ-GUESS
simulation. Used climatic variables were converted to netCDF files. As
LPJ-GUESS requires the climatic variables in specific units, temperature
variables (tas, tasmax, tasmin) [°C] were converted to Kelvin and
surface solar radiation (rsds) [J/m2] was converted to W/m2 by dividing
rsds by 86400 s (3600 s/h \* 24 h). Nitrogen deposition was considered
to be constant at a level of 10 kgN/ha/year. Atmospheric carbon dioxide
concentration data from 1900 to 2018 was taken from the global carbon
project (Quéré et al. 2018). We used the same default LPJ-GUESS soil
code for a medium textured soil type (LPJ-GUESS soil class 2) for each
grid cell (Sitch et al. 2003).

### Soil

### Topography

### Tree species

How have you defined the species composition in your model simulation?
(e.g., species levels, PFTs, etc.)? Please describe the species
categories used. LPJ-GUESS can be run in different modes, with different
levels of abstraction of the population and community processes. For
this study, the more detailed 'cohort' mode was used, in which
individuals, patches and vertical canopy structure are represented
explicitly, but living individuals within a cohort (age class) of a
given tree species in a given patch are assumed to be identical (in
terms of all state variables, such as height and stem diameter).
Simulations consider the main tree species as well as other PFTs
yielding an appropriate representation of European vegetation, including
only tree and shrub species that are widely distributed across Europe
and can become dominant in some areas (Hickler et al., 2012).
Mediterranean rain-green small shrubs with shallow roots (e.g. Lavendula
spp., Rosmarinus spp.) and alpine/arctic shrubs (e.g. Vaccinium spp.)
were represented as PFTs in the model. As in earlier applications of
LPJ-GUESS, herbaceous vegetation was represented by two 'generic herb'
PFTs, with C3 and C4 photosynthesis, respectively. The final set
included 16 tree species, one Mediterranean shrub PFT, one boreal/alpine
shrub PFT, and the two herbaceous PFTs described earlier (Table 2). Acer
pseudoplatanus and Alnus glutinosa are not parameterized in LPJ-GUESS
and therefore we did not simulate these species.

### Simulation

How have you defined the simulation area for your model (e.g., size and
other relevant site aspects)? In this study, vegetation is simulated in
so called patches that represent forest areas of fixed size. Patches can
be regarded as sample stands in a grid cell, each of which represents an
0.1 ha (comparable to figure 1 of the tree regeneration workshop
protocol). In this study, vegetation dynamics were simulated by 2000
replicate patches each 0.1 ha in size. To match the required sample
size, we processed the data so that ten 0.1 ha patches were merged to
one 1 ha sample.

#### Dispersal

What are the relevant aspects of the dispersal (if any) and seed input
parameters in your model? Model formulations of establishment are based
on those employed within the 'forest gap' model FORSKA (Leemans &
Prentice, 1989; Prentice et al., 1993). The number of new saplings of
each woody PFT/species and in each patch each year is drawn at random
from the Poisson distribution, with an expectation influenced by a
PFT/species-specific maximum establishment rate and by the 'propagule
pool', i.e. the amount of carbon allocated to reproduction by all
individuals of the PFT/species at all patches in the previous year. No
saplings are established in a given patch if the minimum PAR level at
the forest floor is below a PFT/species-specific threshold, which is
higher for more light-demanding PFTs/species.

Table 1: Parameter settings associated with vegetation establishment.
Parameter Description Parameter setting in this study ifbgestab Whether
background establishment enabled (0,1) 1 ifsme Whether spatial mass
effect enabled for establishment (0,1) 1 ifstochmort Whether stochastic
mortality is enabled (0,1) 1 ifstochestab Whether stochastic
establishment is enabled (0,1) 1 ifdroughtlimitedestab Whether
establishment is limited by growing season drought (0,1) 1 tcmin_est Min
20-year coldest month mean temp for establishment (deg C) PFT/species
specific (see table 2 and 3) tcmax_est Max 20-year coldest month mean
temp for establishment (deg C) PFT/species specific (see table 2 and 3)
twmin_est Min warmest month mean temp for establishment (deg C)
PFT/species specific (see table 2 and 3) twminusc Continentality
parameter for boreal summer-green trees 0 gdd5min_est Min GDD on 5 deg C
base for establishment PFT/species specific (see table 2 and 3) est_max
Max sapling establishment rate (indiv/m2/year) PFT/species specific (see
table 2 and 3) gdd0_min min GDD0 for survival/establishment reestab Pft
reestablishment ("NONE", "RESTRICTED" or "ALL") ALL

#### Simulation length

What is the simulation length, and how have you determined the
simulation time until the equilibrium? The simulations were initialized
from bare ground (no biomass) and the model was spun up for 500 years
until the modelled vegetation was in approximate equilibrium with the
climate and CO2 around the year 1900 (Smith et al., 2001; Hickler et
al., 2012).

If you have run more than one simulation per site, please describe the
differences of the approaches used. We performed one simulation per
site.

### Output v

Please describe how have you calculated each of the variables in the
output data. The generated output file is called
Output_SGN_LPJGUESS.txt. Each row of the output dataset contains
variables regarding the decade 2008-2018. The first two columns describe
the study site (column "site", values reflect site ID) and the 1 ha
samples of that specific site (column "sample", values range from 1 to
200). Output variables are species-specific. Although we simulated a
total of 20 European tree species and PFTs, our output file considers
only the species required by the protocol (column "species", values are
the abbreviations for the species name ("abal" for Abies alba, "cabe"
for Carpinus betulus, "fasy" for Fagus sylvatica, "frex" for Fraxinus
excelsior, "pabi" for Picea abies, "pisy" for Pinus sylvestris , "tico"
for Tilia cordata , "betu_pen" for Betula pendula, "betu_pen" for Betula
pubescens, "quer_coc" for Quercus coccifera, "quer_il" for Quercus ilex,
"quer_pub" for Quercus pubescens, and "quer_rob" for Quercus robur).
Since Acer pseudoplatanus ("acps") and Alnus glutinosa ("algl") are not
parameterized for LPJ-GUESS so far, we have not included these species
in the output. The first variable is the number of 7 cm recruits
"r_7cm.trees" calculated as the sum of all trees of the same species
that cross the threshold of 7 cm DHB between 2008 and 2018 in that
specific sample of that specific site. The variable "r_10cm.trees"
refers to 10 cm threshold respectively. The 7 cm recruits' basal area
„r_7cm.ba" is calculated as a weighted sum of r_7cm.trees multiplied by
recruits basal area at that time (with ba=(dbh/2)∙π). Similarly,
r_10cm.ba was calculated for 10 cm recruits. The last two columns
describes the mean annual basal area and the mean number of trees per
site, sample, and species overall years between 2008 and 2018 (column
"ba").

Figure 1: Exemplary overview over output file.

Table 2: Species/PFT characteristics and parameter of Hickler et al.
2012. r base respiration rate, kallm1 constant in allometry equations,
Tc,min minimum winter temperature for establishment, Tc,max maximum
coldest-month temperature for establishment, GDD5 minimum degree-day sum
above 5 °C for establishment, fAWC minimum growing-season fraction of
available soil water holding capacity in the first soil layer, rfire
fraction of individuals surviving fire, kla:sa leaf longevity, CAmax
maximum woody crown area. Parameter that are associated with
establishment processes are marked in bold.

 Table 3: Parameters associated with different shade tolerance classes
(updates compared to Hickler et al. 2012). Parameter that are associated
with establishment processes are marked in bold.

### References

Hickler, T., Vohland, K., Feehan, J., Miller, P. A., Smith, B., Costa,
L., ... & Sykes, M. T. (2012). Projecting the future distribution of
European potential natural vegetation zones with a generalized, tree
species‐based dynamic vegetation model. Global Ecology and Biogeography,
21(1), 50-63.

Karger, D. N., Lange, S., Hari, C., Reyer, C. P., & Zimmermann, N. E.
(2021). CHELSA-W5E5 v1. 1: W5E5 v1. 0 downscaled with CHELSA v2. 0.

Leemans, R., & Prentice, I. C. (1989). FORSKA-a general forest
succession model. Meddelanden Från Växtbiologiska Institutionen.Prentice
et al., 1993

Quéré, C., Andrew, R. M., Friedlingstein, P., Sitch, S., Hauck, J.,
Pongratz, J., ... & Zheng, B. (2018). Global carbon budget 2018. Earth
System Science Data.

Sitch, S., Smith, B., Prentice, I. C., Arneth, A., Bondeau, A., Cramer,
W., ... & Venevsky, S. (2003). Evaluation of ecosystem dynamics, plant
geography and terrestrial carbon cycling in the LPJ dynamic global
vegetation model. Global change biology, 9(2), 161-185.

Smith, B. (2001). LPJ-GUESS-an ecosystem modelling framework. Department
of Physical Geography and Ecosystems Analysis, INES, Sölvegatan, 12,
22362.

Smith, B., Wårlind, D., Arneth, A., Hickler, T., Leadley, P., Siltberg,
J., & Zaehle, S. (2014). Implications of incorporating N cycling and N
limitations on primary production in an individual-based dynamic
vegetation model. Biogeosciences, 11(7), 2027-2054.

## aDGVM2

### Authors

### Model

*Please provide here the references to this model where it is possible
to understand the model structure specially for the regeneration module,
this could be a published paper, a website with the model description, a
manual or some other document that the reader can access.*

### Climate

### Soil

### Topography

### Tree species

### Simulations

### Outputs

# Protocol

## General steps

Each modeling team will run their simulations prior to the workshop, and
will report on the results before ([short written report](#report)) and
during the workshop. The general approach for the workshop is:

1.  Each team sets up their model to simulate the regeneration niche of
    [eleven major European tree species](#speciescomposition) within the
    known, defined range of environmental conditions of 200 sites. They
    will run at least one simulation per site. It is important to
    conduct mixed-species simulations, not series of monospecific
    stands.

2.  The model is set up for each site based on the [input variable
    information](#inputvariables) provided in this protocol, and the
    simulations need to be run until the [equilibrium is
    reached](#equilibrium).

3.  Each team reports on the results of these exercises in terms of a
    [standardized output dataset](#outputdata) (including density
    variables per 10-yr interval and per ha, for each simulated site and
    species) and in the form of a brief [working report](#report) (\~4
    pages).

4.  At the workshop in June 2022 in Davos, Switzerland, we will evaluate
    the regeneration niche as it results from each model and compare it
    to the empirical estimates of the environmental limits of the
    regeneration niche that will be made available to the participants
    shortly before the workshop. As a group effort, the projections of
    the niche properties arising from the different models will be
    compared. The exact details on how this will be done will be
    provided once we know how many modeling teams will contribute
    results.

## Input variables {#inputvariables}

In this section, we are providing a description of all the necessary
driver information to run the simulations (e.g., location, climate,
soil).

### Coordinates

We intend to provide the site information via unique site IDs. If actual
coordinates are needed, we can provide blurred site coordinates upon
reasonable request.

### Climate

We provide time series of climatic variables from 1981 to 2018 in
hourly, daily and monthly resolution. Based on our survey the climate
data should be sufficient to run all models. Some variables such as
relative humidity or vapor pressure deficit must be calculated from the
variables we provide.

Note that the hourly time series has the lowest spatial resolution, thus
the daily time series should be used for spatial accuracy. If you need
hourly time series we recommend to use the daily time series and only
use the hourly time series to correct for hourly variation. For this
purpose we also provide hourly averages over the period 1981 to 2018 for
each hour of each day of the year. Only for calculating these averages
the 29th of February in leap years was removed from the time series to
obtain values for 365 days.

#### Data sources {#climsources}

The hourly time series are based on the Era5-land data set. Daily and
monthly time series are mostly based on CHELSA data which has a higher
spatial resolution. More details are given below or in the documentation
of the original data source.

The variable names are kept exactly like in the documentation of the
Era5 and CHELSA data sets to facilitate the use of the original data
documentation. Please don't be confused by the ending `*_hourly` in
`surface_thermal_radiation_downwards_hourly` and the two different
variable names for shortwave radiation: `rsds` for the values from
CHELSA and `surface_solar_radiation_downwards_hourly` for the values
from Era5.

All changes that were made to the original data are documented here:

-   The wind variable `wind_speed` is based on the u and v component of
    the wind and calculated with $\sqrt{u^2+v^2}$ where $u$ is the
    variable `u_component_of_wind_10m` and $v$ is
    `v_component_of_wind_10m` of the Era5 data set.

-   The units of temperature variables `temperature_2m` (Era5), `tas`,
    `tasmin`, `tasmax`(all CHELSA) were changed from Kelvin to Degree
    Celsius

-   The units of precipitation variables `total_precipitation_hourly`
    (Era5), `pr` (CHELSA) were changed from m to mm and kG to mm
    respectively.

-   The unit of short wave radiation `rsds` (CHELSA) was changed from kJ
    to J.

We aggregated variables from hourly to daily to monthly values by using
means for temperature, surface pressure and wind speed and sums for
radiation and precipitation.

##### Era5-Land {#era5}

-   Muñoz Sabater, J., (2019): ERA5-Land hourly data from 1981 to
    present. Copernicus Climate Change Service (C3S) Climate Data Store
    (CDS). (23.01.2022),
    [doi:10.24381/cds.e2161bac](https://doi.org/10.24381/cds.e2161bac)
-   [Documentation of the hourly ERA5 data from the Google Earth
    Engine](https://developers.google.com/earth-engine/datasets/catalog/ECMWF_ERA5_LAND_HOURLY#description)
-   Further information is available from the [official copernicus
    website](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-land?tab=overview).

##### CHELSA {#chelsa}

-   Dirk N. Karger, Stefan Lange, Chantal Hari, Christopher P. O. Reyer,
    Niklaus E. Zimmermann (2021): CHELSA-W5E5 v1.0: W5E5 v1.0 downscaled
    with CHELSA v2.0. ISIMIP Repository.
    <https://doi.org/10.48364/ISIMIP.836809>

-   [Official CHELSA website](https://chelsa-climate.org)

#### Monthly time series 1981-2018

file name: monthly_era5_chelsa_200.zip

**Data description**

Variables from the [Era5 data set](#era5) with a spatial resolution of
0.1° x 0.1° (\~ 10 km):

-   average wind speed (`wind_speed`) [m/s]

-   average surface pressure (`surface_pressure`) [Pa]

-   surface thermal (longwave) radiation
    (`surface_thermal_radiation_downwards_hourly`) [J/m2]

Variables from the [CHELSA data set](#chelsa) with a spatial resolution
of 0.008° x 0.008° (\~ 1 km):

-   average temperature (`tas`) [°C]

-   average max temperature (`tasmax`) [°C]

-   average min temperature (`tasmin`) [°C]

-   precipitation sum (`pr`) [mm]

-   surface solar (shortwave) radiation (`rsds`) [J/m2]

#### Daily time series 1981-2018

file name: daily_era5_chelsa_200.zip

**Data description**

Variables from the [Era5 data set](#era5) with a spatial resolution of
0.1° x 0.1° (\~ 10 km):

-   average wind speed (`wind_speed`) [m/s]

-   average surface pressure (`surface_pressure`) [Pa]

-   surface thermal (longwave) radiation
    (`surface_thermal_radiation_downwards_hourly`) [J/m2]

Variables from the [CHELSA data set](#chelsa) with a spatial resolution
of 0.008° x 0.008° (\~ 1 km):

-   average temperature (`tas`) [°C]

-   average max temperature (`tasmax`) [°C]

-   average min temperature (`tasmin`) [°C]

-   precipitation sum (`pr`) [mm]

-   surface solar (shortwave) radiation (`rsds`) [J/m2]

#### Hourly time series 1981-2018

file name: hour_era5_200.zip

file name: yhourMean_era5_200.zip

**Data description**

Variables from the [Era5 data set](#era5) with a spatial resolution of
0.1° x 0.1° (\~ 10 km):

-   average temperature (temperature_2m) [°C]

-   average wind speed (`wind_speed`) [m/s]

-   average surface pressure (`surface_pressure`) [Pa]

-   precipitation sum (total_precipitation_hourly) [mm]

-   surface solar (shortwave) radiation
    (surface_solar_radiation_downwards_hourly) [J/m2]

-   surface thermal (longwave) radiation
    (surface_solar_radiation_downwards_hourly) [J/m2]

### Soil

file name: soil_quality_200.csv

**Data description**

The following variables are available:

-   Soil quality is given as a continuous value between 1 and 5.

**Data source**

-   The soil quality value is based on a random forest that was trained
    to predict expert knowledge based soil quality on a scale from 1 to
    5 from the [SoilGrids250 data
    set](https://www.isric.org/explore/soilgrids) and the [WISE data
    set](https://www.isric.org/explore/wise-databases). The most
    influential variables in the random forest were total depth to
    bedrock and sand content. Thus the variable is a good predictor for
    plant available water storage capacity.

## Topography and elevation

There was an error in a earlier version of this file. Please use the
link above to download the corrected file.

This is the old file and should not be used because the aspect column
contains elevation instead of the aspect.

file name: topo_data_200.csv

**Data description**

The following variables are available:

-   elevation [m.a.s.l]

-   slope [Degree]

-   aspect [Degree]

**Data source**

-   Topographic information was derived from the [ASTER DEM
    V3](https://lpdaac.usgs.gov/products/astgtmv003/).

-   Slope and aspect were calculated with QGIS. The specific tiles that
    were used are documented
    [here](https://ites-fe.ethz.ch/rws/downloads/15122021/aster_dem_files.txt),
    and the processing steps in QGIS are documented
    [here](https://ites-fe.ethz.ch/rws/downloads/15122021/QGIS_command_log.txt).

### Natural disturbances

The simulations must be run in the absence of natural disturbances.

## Management

The simulations must be run in the absence of management.

### Area simulated

From your model outputs, we would like to obtain samples on species
specific regeneration rates per decade and per ha. The modeling teams
decide on the simulated area and how they derive these samples. Please
describe the methodology used in the [output report](#report).

### Simulation duration and equilibrium {#equilibrium}

The simulation should be run from bare ground in the absence of
management to a simulated [equilibrium](#defequilibrium) ("potential
natural vegetation") with current climate. The goal is to derive the
[regeneration niche](#regenniche). The exact length of the simulation
should be decided by the modeling teams. Please describe the simulation
length used in your simulations in the [output report](#report).

[**Regeneration niche:**]{#regenniche} The regeneration niche is defined
as the passing of a diameter threshold of 7 and 10 cm (ingrowth; cf.
below). Both the width of the regeneration niche (i.e., in environmental
space) as well as the intensity of the recruitment process (i.e., the
number of ingrowth trees per area and per unit of time) are of interest.

[**Simulated equilibrium**]{#defequilibrium}: Simulated equilibrium is
the "Potential Natural Vegetation".

### Species composition {#speciescomposition}

**Mixed species simulations** must be run (not multiple single-species
simulations).

**Option A: Species level**

The simulation should be run for **mixtures** of the following
species/genera: *Fagus sylvatica, Picea abies, Abies alba, Carpinus
betulus, Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus
excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris.*

Additional species can be included in the simulation if modelers think
this is important, but they should report only on these 11
species/genera because the benchmarking data refer to them only. Please
report the methodology used in the [output report](#report).

**Option B: PFTs**

If species-level simulation is not possible in your model, substitute
the species by PFTs and use the model's default approach. Please report
the methodology used in the [output report](#report). Explain as well
which PFTs represent which species for your model.

## Diameter threshold

The [output data](#outputdata) have to be provided for two diameter
thresholds: 7 and 10 cm.

## Output

We are expecting to get the number of recruits over a 10-yr interval per
species and per ha for each of the sites. However, we are also
interested in the **variation per site** (i.e., we would like to see the
range of estimates, not only the mean value per site). Each modeling
team will decide how to sample multiple values per site that will show
the variation in recruitment.

Specifically, you should provide us with multiple samples (n = 200; this
number is the same as the number of sites by sheer coincidence) of the
regeneration rate for spatial units of 1 ha (please do [not]{.underline}
provide average of multiple 1-ha samples). This can be done in various
ways, e.g., by sampling simulated data from the same 1-ha plot in the
equilibrium over time, or by sampling several 1-ha plots from the
simulated forest at one specific point in time (in the equilibrium); for
clarification, cf. three sample strategies proposed in [figures
1](#sample1), [2](#sample2) and [3](#sample3).

If, for some reason, you cannot submit **200 samples** of **1 ha**,
please **let us know** in the [workshop chat](#chat) so we can discuss
alternatives. The sampling from the simulations must happen **within**
the simulated phase of Potential Natural Vegetation (for more
information, see how [equilibrium](#defequilibrium), Potential [Natural
Vegetation, and regeneration niche](#regenniche) are defined in this
protocol).

![Figure 1. In this example there are 200 samples collected at year i at
equilibrium differentiated spatially with each square representing 1
ha.](samplingfigure/Slide2.jpeg){#sample1 width="500"}

![Figure 2. In this example, there are 100 spatially differentiated
samples taken in year i at equilibrium and 100 in year i+10. Each square
represents 1 ha.](samplingfigure/Slide3.jpeg){#sample2 width="500"}

![Figure 3. In this example there is 1 spatially differentiated sample
taken in year i at equilibrium and remeasured every 10 years for 2000
years until i+2000. The square represents 1
ha.](samplingfigure/Slide4.jpg){#sample3 width="500"}

### Output data {#outputdata}

Please provide the simulation output data using the following structure:

-   Each row of your output dataset contains the recruitment of 10-year
    interval per species and ha for a unique simulation site.

-   Each column of your output dataset contains one of these variables:

    -   `sample`: this will identify the different samples per site
        (1:n, with n $\ge$ 200)

    -   `site`: site identifier (1:200)

    -   `species`: species simulated with the following categories:

        -   fasy (for *Fagus sylvatica*)
        -   pabi (for *Picea abies*)
        -   abal (for *Abies alba*)
        -   cabe (for *Carpinus betulus*)
        -   tico (for *Tilia cordata*)
        -   acps (for *Acer pseudoplatanus*)
        -   betu (for *Betula* spp.)
        -   frex (for *Fraxinus excelsior*)
        -   quer (for *Quercus* spp.)
        -   algl (for *Alnus glutinosa*)
        -   pisy (for *Pinus sylvestris*)
        -   If you have some other categories please create your own
            species codes and provide the legend in the output report.
            We prefer a lumped category "othr".

    -   `r.trees`: the recruits (trees $ha^{-1}$) per species and
        decade.

    -   `r.ba`: the recruits' basal area ($m^{2}$ $ha^{-1}$) per species
        and decade.

    -   `ba`: total basal area ($m^{2}$ $ha^{-1}$) per species and
        decade

    -   `dbh`: diameter threshold (cm) used to define the recruitment (7
        or 10 cm).

### Output report {#report}

To guide the writing of the output report, please refer to the following
questions:

-   What climate data have you used, and how have you created your model
    climate input parameters?

-   How have you defined the species composition in your model
    simulation? (e.g., species levels, PFTs, etc.)? Please describe the
    species categories used.

-   How have you defined the simulation area for your model (e.g., size
    and other relevant site aspects)?

-   What are the relevant aspects of the dispersal (if any) and seed
    input parameters in your model?

-   What is the simulation length, and how have you determined the
    simulation time until the equilibrium?

-   If you have run more than one simulation per site, please describe
    the differences of the approaches used.

-   Please describe how have you calculated each of the variables in the
    [output data](#outputdata).

### Uploading data and report

Please upload your data and output report as a **zip file** with the
name of your model and any other word that might help us to understand
the files, for example: *thenameofyourmodel_datareport.zip* or
*thenameofyourmodel_data.zip* and *thenameofyourmodel_report.zip*.

You can upload the files here: LINK

### 2022-12-06 manuscript/manuscriptRW.Rmd
add co-author in alphabetical order

@@
 - Yannek Käber,
+- Tim Anders,
+- Kristin H. Braziunas,
+- Josef	Bruna,
+- Samuel Fischer,
+- Jessica Hetzer,
+- Thomas Hickler,
+- Christian	Hochauer,
+- Manfred Lexer,
+- Heike	Lischke,
 - Mats Mahnken,
-- Samuel Fischer,
-- Xavier Morin,
-- Manfred Lexer,
-- Christian Hochauer,
+- Paola	Mairota,
 - Katarina Merganicova,
 - Tobias Mette,
+- Xavier Morin,
 - Werner Rammer,
-- Kristin H. Braziunas,
-- Paola Mairota,
-- Josef Bruna,
-- Heike Lischke,
+- Simon	Scheiter,
 - Daniel Scherrer,
-- Simon Scheiter,
-- Thomas Hickler,
-- Tim Anders,
-- Jessica Hetzer,
 - Harald Bugmann
@@
 inadequately high species diversity in regeneration is simulated. This
-is characteristic of \'classical\' forest gap models [@gutiérrez2016],
-at least as long as the simulations comprise a multitude of species.
-Other models use calibration against local canopy-level data to define
+is characteristic of 'classical' forest gap models [@gutiérrez2016], at
+least as long as the simulations comprise a multitude of species. Other
+models use calibration against local canopy-level data to define
 appropriate recruitment levels, which is likely to strongly hamper the
@@
 ForClim variant 11 is adopting a rather different approach for
-establishment model that considers
-space filling along with different \'points of entry\' for the action of
-specific
-environmental factors @huber2020. These two model variants allow us to
-evaluate a more process-based establishment module (variant 11) against
-a simpler module (variant 1) while the rest of the model structure
-remains the same. Similarly,
-the two ForCEEPS variants used here allow us to isolate the importance
-of the
-canopy feedback via seed trees for the quantity and quality of simulated
-regeneration, as one variants includes this feedback whereas the other
-one does
-not.
+establishment model that considers space filling along with different
+'points of entry' for the action of specific environmental factors
+@huber2020. These two model variants allow us to evaluate a more
+process-based establishment module (variant 11) against a simpler module
+(variant 1) while the rest of the model structure remains the same.
+Similarly, the two ForCEEPS variants used here allow us to isolate the
+importance of the canopy feedback via seed trees for the quantity and
+quality of simulated regeneration, as one variants includes this
+feedback whereas the other one does not.
 
@@
 rates per decade and per ha in the equilibrium state of the model,
-typically entailing a \"spin-up\" run (as we did not provide any forest
+typically entailing a "spin-up" run (as we did not provide any forest
 data). The modeling teams decided on the simulated area and how they
@@
 SPECIES. HERE **LPJ-GUESS** PROVIDES [ONE SENTENCE]{.underline} WHY THEY
-SIMULATED LESS SPECIES. Finally, the
-model aDGVM2 is based on a broad suite of Plant Functional Types rather
-than
-specific species, and hence this model used its default approach of PFTs
-for
-the simulations.
+SIMULATED LESS SPECIES. Finally, the model aDGVM2 is based on a broad
+suite of Plant Functional Types rather than specific species, and hence
+this model used its default approach of PFTs for the simulations.
 
@@
 On the one hand, the fact that some models feature lower mortality than
-expected by Reineke\'s self-thinning rule may point to facilitation, or
-simply a higher mortality rate
-before the trees had reached 7 cm diameter. On the other hand, the case
-of models
-such as xComp, PICUS and TreeMig, which are well above the self-thinning
-line, may
-indicate that they feature higher recruitment mortality in mixed-species
-forests
+expected by Reineke's self-thinning rule may point to facilitation, or
+simply a higher mortality rate before the trees had reached 7 cm
+diameter. On the other hand, the case of models such as xComp, PICUS and
+TreeMig, which are well above the self-thinning line, may indicate that
+they feature higher recruitment mortality in mixed-species forests
 compared to even-aged, single-species forests. However, some models
@@
 Importantly, excess recruitment requires higher mortality rates in later
-stages, either -- as observed for
-some models -- between the 7 and 10 cm diameter thresholds, or either
-before or
-shortly after this stage. Correcting for the expected forest densities
-at early
-stages is thus equivalent to compensating for a first error (excess
-recruitment) by committing a second error (excess mortality). Thus, many
-models appear to get the right \'answer\', i.e., correct stand structure
-and composition, for the wrong reasons. This is problematic particularly
-if the models are to be used under novel conditions (e.g., climate
-change; @huber2021) or in a decision support context (e.g.,
-@thrippleton2021).
+stages, either -- as observed for some models -- between the 7 and 10 cm
+diameter thresholds, or either before or shortly after this stage.
+Correcting for the expected forest densities at early stages is thus
+equivalent to compensating for a first error (excess recruitment) by
+committing a second error (excess mortality). Thus, many models appear
+to get the right 'answer', i.e., correct stand structure and
+composition, for the wrong reasons. This is problematic particularly if
+the models are to be used under novel conditions (e.g., climate change;
+@huber2021) or in a decision support context (e.g., @thrippleton2021).
 
 In spite of this criticism, it remains remarkable that most of the
-models were not overly off
-with respect to simulated regeneration levels -- after all, few if any
-of the models\' establishment routines are sufficiently constrained by
-data, and the resulting degrees of freedom that are available for
-modeling regeneration (cf. @konig_tree_2022) could have implied that
-model performance would be much worse.
+models were not overly off with respect to simulated regeneration levels
+-- after all, few if any of the models' establishment routines are
+sufficiently constrained by data, and the resulting degrees of freedom
+that are available for modeling regeneration (cf. @konig_tree_2022)
+could have implied that model performance would be much worse.
 
@@
 Yet, in this study we were not able to assess tree regeneration in a
-holistic sense. The
-lower size threshold of 7 cm in the observed data constituted a hard
-limit based
-on which we can understand some aspects of tree recruitment, which in
-its
-entirety often comprises a rather long period since germination
-[@price2001], and many relevant environmental constraints are acting on
-these young trees [@käber2021] that we were unable to assess. However,
-recruitment data with lower thresholds are simply not available in a
-harmonized manner across large environmental gradients. The EuFoRIa data
-used here were collected in forest reserves where no management has
-taken place for long periods of time. This makes our assumption of an
-equilibrium between forest properties and environmental drivers less
-haphazard than it might appear at first sight. However this assumption
-might also justify some of the rather large recruitment rates that we
-are seeing in the models projections compare to the observed ones, as if
-the observed data it is not in equilibrium the observed recruitment
-would be lower as they are not close to the optimum.
+holistic sense. The lower size threshold of 7 cm in the observed data
+constituted a hard limit based on which we can understand some aspects
+of tree recruitment, which in its entirety often comprises a rather long
+period since germination [@price2001], and many relevant environmental
+constraints are acting on these young trees [@käber2021] that we were
+unable to assess. However, recruitment data with lower thresholds are
+simply not available in a harmonized manner across large environmental
+gradients. The EuFoRIa data used here were collected in forest reserves
+where no management has taken place for long periods of time. This makes
+our assumption of an equilibrium between forest properties and
+environmental drivers less haphazard than it might appear at first
+sight. However this assumption might also justify some of the rather
+large recruitment rates that we are seeing in the models projections
+compare to the observed ones, as if the observed data it is not in
+equilibrium the observed recruitment would be lower as they are not
+close to the optimum.
 
@@
 
-A concluding (\"more positive\") final paragraph.
+A concluding ("more positive") final paragraph.
 

### 2022-12-07 code/dataPreparation.R
correcting LPJ-GUESS data- missing samples

@@ -106,7 +106,7 @@
     sibyla$model <- "SIBYLA"
     
     # Some sites have more than 200 samples, 
-    # remove the extra ones to have 200 samples per site
+    # remove the extra samples reported to have 200 samples per site
     sibyla <- sibyla[!sibyla$sample > 200,]
     
 
@@ -119,12 +119,12 @@
     xcomp$model <- "xComp"
 
     # Add 0 values to those sites where there is 0 recruitment and 0 BA for a spp
-    # that was simulated for the model
+    # that were simulated by the model
     sppSimxcomp <- unique(xcomp$species)
 
     missingxcomp <- xcomp |> 
                      dplyr::group_by(site, sample, dbh, model) |>
-                     dplyr::summarize(species = sppSimxcomp[!sppSimxcomp  %in% unique(species)],
+                     dplyr::summarize(species = sppSimxcomp[!sppSimxcomp %in% unique(species)],
                                       sample = unique(sample),
                                       site = unique(site),
                                       dbh  = unique(dbh), 
@@ -431,29 +431,86 @@
                        ba = sum(ba)) # calculate the sum of those columns
     
     lpjguess <- lpjguess2 
-
     
-    # Missing samples in some sites
+    
+    
+  # Missing samples in some sites
+  # IMPORTANT! Missing samples were grasslands (authors personal com.)
+  # and will be added here as no recruitment 
     
     # 7cm
     nosamples7 <- as.data.frame(table(lpjguess$site[lpjguess$dbh == 7], 
                                       lpjguess$sample[lpjguess$dbh == 7]))
-    missingSamples7 <- nosamples[nosamples$Freq == 0,]
-    # Sites with missing samples 
-    table(missingSamples7$Var1) 
+    missingSamples7 <- nosamples7[nosamples7$Freq == 0, ]
+    colnames(missingSamples7) <- c("site", "sample", "freq")
+    missingSamples7$sample <- as.numeric(levels(missingSamples7$sample))[missingSamples7$sample] 
+    missingSamples7$site <- as.numeric(levels(missingSamples7$site))[missingSamples7$site] 
+    missingSamples7$freq <- NULL
+    missingSamples7$model <- "LPJ-GUESS" 
+    missingSamples7$dbh <- 7
+    missingSamples7$r.trees  <- 0  
+    missingSamples7$r.ba <- 0  
+    missingSamples7$ba <- 0
+    missingSamples7$species <- "pabi"
+    
[…]

### 2022-12-14 manuscript/manuscriptRW.Rmd
changes based on feedback and discussion with Harald on 12-12-2022

--- 2022-12-02 manuscript/manuscriptRW.Rmd
@@
 ---
-title: Drivers of tree regeneration remain challenging to be captured by models of
-  forest dynamics
+title: "(WT1) Tree establishment in models of forest dynamics: a key priority for further research  \n(WT2) Tree establishment processes in models of forest dynamics: capturing the drivers remains challenging"
 author:
@@
 linkcolor: blue
-csl: draft.csl
+csl: global-change-biology.csl
 editor_options:
@@
 
-Tree regeneration is a key process in forest dynamics, particularly in
+Tree establishment is a key process in forest dynamics, particularly in
 the context of forest resilience after disturbance and climate change.
 Models of forest dynamics have been in use since more than 50 years, but
-an assessment of their capacity to properly represent regeneration is
-still lacking. Using 15 models and model variants that were build to
-capture long term (decadal to centennial) forest dynamics at the stand,
-landscape and global levels, we simulate tree regeneration across at 200
-sites representing environmental gradients across Europe. The simulation
-results are benchmarked against data from forest reserves collected at
-these 200 sites. We aim to assess how well the models of forest dynamics
-are capturing the overall abundance, species composition and mortality
-in tree recruitment. The results show that most of the models tend to
-over-estimate recruitment rates although this is not always connected to
-large mortality rates in the early stages of population dynamics. In the
-models, the effect of light availability on recruitment rates is
-captured better than the effect of temperature and soil moisture, but
-the patterns are not consistent across models. Species richness matches
-quite well compared to the ranges observed in the data. Models
-simulating higher species diversity at the stand level do not feature
-higher diversity in recruitment. *Abies alba* was well captured across
-temperature and soil moisture gradients by most models, whereas *Fagus
-sylvatica* was generally underestimated. *Picea abies* recruitment was
-generally over-estimated, particularly at colder sites. HERE TWO
-SENCENTECES DISCUSSION/CONCLUSION. HERE A SO WHAT SENTENCE
+an assessment of their capacity to represent tree establishment is
+lacking. We assess how well current models are capturing the overall
+abundance, species composition and mortality of tree establishment.
+
+Using 15 models and model variants buil to capture longterm (i.e.,
+decadal to centennial) forest dynamics at the stand, landscape and
+global levels, we simulate tree establishment at 200 sites representing
+environmental gradients across Europe. The results are benchmarked
+against comprehensive data from forest reserves.
+
+Most of the models overestimate ingrowth rates, but this is not always
+compensated by high simulated mortality rates in the early stages of
+population dynamics. In the models, the effect of light availability on
+ingrowth rates is captured better than the effect of temperature and
+soil moisture, but the patterns are not consistent across models.
+Simulated species diversity matches the observed ranges quite well.
+Models simulating higher species diversity at the stand level do not
+feature higher ingrowth diversity. The ingrowth of *Abies alba* is
+captured across temperature and soil moisture gradients, whereas *Fagus
+sylvatica* is generally underestimated. *Picea abies* ingrowth is
+generally overestimated, particularly at colder sites. Increasing
+complexity in the tree establishment modules of the models is not
+related to higher accuracy of simulated tree establishment. Furthermore,
+individual model design is more important than scale (stand, landscape,
+global) and approach (empirical, process-based) to accurately capture
+tree establishment.
+
+Even though the overestimation of tree establishment represent a
+challenge
+for the usefulness of model~~s~~ projections under uncertain conditions
+such a climate change, and although some formulations should be
+reconsidered to correct excessive ingrowth levels, it is remarkable that
+most of the~~se~~ models capture the essential features of the highly
+complex process of tree establishment, while not having been
+parameterized with such a data. We conclude that in spite of all the
+uncertainties, many models hold promise for assessing longterm forest
+dynamics under changing environmental conditions.
 
@@
 Forests provide a wide range of products and services of vital
-importance. Under the pressure of climate change and increasing
-disturbances, it is becoming ever more important to understand how
-forests structure, composition and functions will change, and evaluate
-forest capacity to adapt or be resilient after disturbance
-[@lindner2010]. A wide range of models of forest dynamics were developed
-over the past decades considering the impacts of climate on forest
-dynamics [@bugmann2022]. From these studies, it is becoming increasingly
-evident that while we have a reasonably good understanding how to model
-tree growth [@vanclay1997; @bugmann1996], and substantial efforts have
-been dedicated to improving the depiction of tree mortality processes
-[@bugmann2019], tree regeneration remains a much less studied process
-that is often represented rather coarsely in these models, which is a
-clear research gap in the context of climate-induced forest disturbances
-and forest resilience.
-
-Tree regeneration depends on multiple processes such as reproduction,
+importance to humankind.
+Under the pressure of climate change and increasing disturbances, it is
+becoming ever more important to understand how forest structure,
+composition and function will change, and to evaluate forest capacity to
+adapt or be resilient after disturbance [@lindner2010]. A wide range of
+models of forest dynamics were developed over the past decades
+considering the impacts of climate on forest dynamics [@bugmann2022].
+From these studies, it is evident that while we have a reasonably good
+understanding how to model tree growth [@vanclay1997; @bugmann1996], and
+substantial efforts have been dedicated to improving the depiction of
+tree mortality [@bugmann2019], tree establishment remains a much less
+studied process that is often represented rather coarsely in these
+models, which is a clear research gap in the context of climate-induced
+forest disturbances and forest resilience.
+
+Tree establishment arises from multiple processes such as reproduction,
 fruit maturation, seed production, dispersal, germination, juvenile
-growth and survival [@price2001]. All these processes are all difficult
-to assess, and some of them are scarcely understood and thus appear
-highly stochastic [@bogdziewicz2021]. Identifying the appropiate level
-of complexity for the formulation of the key factors that are leading to
-successful tree recruitment is challenging. By tree recruitment, we
-refer to the passing of trees to selected diameter threshold (ingrowth)
-as a definition of successful regeneration [@käber2022]. Overall, models
-are needed to (1) synthesize existing empirical data and explore their
-consequences, and (2) assess future tree recruitment e.g. in the context
-of global change scenarios. Currently, regeneration processes in dynamic
-forest models are handled in a multitude of ways [@konig_tree_2022;
-@bugmann2022], from entirely ignoring tree regeneration (as done in
-classical forest growth models, e.g. @pretzsch2002), across the use of a
-few simple environmental filters, as done in most forest "gap" models
-(e.g., @shuman2011) and Dynamic Global Vegetation Models (e.g.,
-@smith2001; @hickler2012), to complex approaches that incorporate local
-feedback from the canopy, multiple ecological processes and often also
-small time steps (e.g., @wehrli2006; @seidl2012), or field-based
-statistical parameterizations, which however are not easy to extrapolate
-in space and time (e.g., @ribbens1994).
-
-Given the current strategies that are used in models of forest dynamics
-to represent tree recruitment, model behavior is facing several problems
-and that can have strong impacts on forest dynamic, such as very high
-levels of recruitment, which then necessitate excess mortality to early
-stages to arrive at correct stand structure and composition. Also, often
-inadequately high species diversity in regeneration is simulated. This
-is characteristic of \'classical\' forest gap models [@gutiérrez2016],
-at least as long as the simulations comprise a multitude of species.
-Other models use calibration against local canopy-level data to define
-appropriate recruitment levels, which is likely to strongly hamper the
-general applicability of these models e.g. under scenarios of climate
-change.
-
-Correctly capturing the species composition in recruitment as a function
-of the presence of seed trees in the canopy is often a particular
-challenge, potentially leading to unrealistic successional drift in the
-model, which must be corrected by factors that are hard to parameterize
-[@lischke2006]. A related issues is the excessive reduction of species
-richness due to positive feedback effects such that eventually just
-single-species forest stands remain. This is sometimes handled by the
-incorporation of a low level of seed influx of all species at all times
-(e.g., @lischke2006 ). However, simulated species composition is usually
-exceedingly sensitive to that level, whereas the parameters of such
-functions are poorly constrained by field data. Furthermore, there is
-also often a problem with the insufficient data-based constraint on
-parameter values for models that start from very small tree sizes (e.g.,
-10 cm of height or even seeds) and track tree development in a
-process-oriented manner by considering a multitude of ecological
-influences, rather than being statistically driven and emphasizing
-recruitment into a larger size class (e.g., @wehrli2006 ).
-
-Thus, a focus on the modeling of tree regeneration processes is sorely
-needed and long overdue (cf. @price2001) if we are to make reliable
-projections of future forest dynamics, i.e. when the models need to be
-operated in extrapolation mode, but also from a fundamental ecological
-point of view (increasing systems understanding).
-
-Also in the field of ecology, tree recruitment has not received a lot of
-attention [@hanburybrown_forest_2022], although many sub-processes have
-been studied in detail. Yet, little data is available that covers all
-regeneration processes within one species along environmental gradients,
-let alone for a vast suite of species. Thus, the available data are
-fragmented, which constitutes a major problem for model building. On the
-one hand, monitoring on permanent sample plots (such as National Forest
-Inventories) often measure tree recruitment [@zell_modeling_2019],
-however often with a design that captures the regeneration intensity
-after the stand initiation phase in productive forests [@hallsby2015].
-Data from permanent plots, although available across regions, are also
-highly heterogeneous, rendering their utility difficult in a model
-context (cf. @käber2021). On the other hand, targeted studies to measure
-tree recruitment on experimental sites are often limited in spatial
-extent, species studied, or sub-set of regeneration processes that are
-targeted [@berdanier2016; @collet2006].
+growth and survival [@price2001]. All these processes are difficult to
+assess, and some of them
+are scarcely understood and thus appear highly stochastic
+[@bogdziewicz2021]. Identifying the appropriate level of complexity for
+the formulation of the key factors that are leading to successful tree
+establishment is challenging. Currently, tree
+establishment processes in dynamic forest models are handled in a
+multitude of
+ways [@konig_tree_2022; @bugmann2022], from entirely ignoring it [as
+done in classical forest growth models, e.g., Pretzsch et al.
+-@pretzsch2002], across the use of a few simple environmental filters,
+as done in most forest \"gap\" models [@shuman2011] and Dynamic Global
+Vegetation Models [e.g., @hickler2012; @smith2001], to complex
+approaches that incorporate local feedback from the canopy, multiple
+ecological processes and often also small time steps [e.g., @seidl2012;
+@wehrli2006], or field-based statistical parameterizations, which
+however are not easy to extrapolate in
+space and time [e.g., @ribbens1994].
+
+Overall, models are needed to (1) synthesize existing empirical data and
+explore their consequences, and (2) assess future tree establishment
+e.g., in the context of global change scenarios. Given the current
+strategies that are used in models of forest
+dynamics to represent tree establishment, model behavior is often facing
+problems that can have strong impacts on simulated forest dynamics, such
+as very high levels of tree establishment, which then necessitate excess
+mortality at early stages of tree life to arrive at correct stand
+structure and composition. Also, inadequately high species diversity in
+tree establishment may be simulated. This is characteristic of
+\"classical\" forest gap models [@gutiérrez2016], at least as long as
+the simulations comprise a multitude of species. Other models use
+calibration against local canopy-level data to define appropriate
+ingrowth levels, which is likely to strongly hamper the general
+applicability of these models e.g., under scenarios of climate change.
+Furthermore, correctly capturing the species composition in tree
+establishment as a function of the presence of seed trees in the canopy
+is often a particular challenge, potentially leading to
+unrealistic successional drift in the model, which must be corrected by
+factors that are hard to parameterize [e.g., @lischke2006]. A related
+issue is the excessive reduction of species richness due to positive
+feedback effects such that eventually just single-species ~~forest~~
+stands remain. This is sometimes corrected by the incorporation of a low
+level of seed influx of all species at all times [@lischke2006].
+However, simulated species composition is usually exceedingly sensitive
+to that level, whereas the parameters of such functions are poorly
+constrained by field data. Lastly, there is often a problem with
+insufficient observational constraint on parameter values for models
+that start from very small tree sizes (e.g., 10 cm of height or even
+seeds) and track tree development in a process-oriented manner by
+considering a multitude of ecological influences, rather than
+emphasizing tree establishment into a larger size class [e.g.,
+@wehrli2006].
+
+Thus, a focus on the modeling of tree establishment processes is sorely
+needed and overdue [@price2001] if we are to make reliable projections
+of future forest dynamics, i.e., when the models need to be operated in
+extrapolation mode, but also from a fundamental ecological point of view
+(increasing systems understanding). Also in the field of general
+ecology, tree establishment has not received a lot of attention
+[@hanburybrown_forest_2022], although many sub-processes have been
+studied in detail. Yet, little data is available that covers all the
+processes within one species along environmental gradients, let alone
+for a vast suite of species. Thus, the available data are fragmented,
+which constitutes a major problem for model building. On the one hand,
+monitoring on permanent sample plots (such as National Forest
+Inventories) often measure tree establishment [@zell2019], however with
+a design that captures the rate after the stand initiation
+phase[@hallsby2015]. Data from permanent plots, although available
+across regions, are highly heterogeneous, rendering their utility
+difficult in a modeling context[@käber2021]. On the other hand, targeted
+studies to measure tree establishment on experimental sites are often
+limited in spatial extent, species studied, or the sub-set of processes
+that are targeted [@berdanier2016; @collet2006].
 
@@
 models commonly used to assess forest dynamics under climate change are
-evaluated against a continental-scale multi-species harmonized data set
-on tree recruitment [@käber2022]. We included models that are based on a
-range of 'philosophies', including models operating at the stand to the
-global scale as well as the range of models from empirically derived to
-"process-based" (e.g. @fabrika2012; @mette2014; @lexer2001;
-@bugmann1996). Because of the large variability in recruitment patterns
-in nature and the large amount of factors driving recruitment, including
-some that are not included in most models, such as deer browsing, we do
-not aim at a detailed statistical evaluation of each model. Instead, we
-aim at evaluating the general recruitment patterns and magnitudes
-simulated by the models, and benchmark the simulated regeneration niche
-of multiple species against empirical data along a wide environmental
-gradient (temperature, moisture, and light availability).
+evaluated against a continental-scale, multi-species harmonized data set
+on tree establishment [@käber2022]. By tree establishment, we refer to
+the passing of trees across a specific diameter threshold (\"ingrowth\")
+and thus as successful tree establishment. We included models that are
+based on a range of \"philosophies\", from models operating at the stand
+to the global scale as well as the range of models from empirically
+derived to \"process-based\" [e.g., @bugmann1996; @fabrika2012;
+@mette2014; @lexer2001]. Because of the large variability in tree
+establishment patterns in nature and the large number of factors driving
+this process --- including some that are not included in most models
+such as deer browsing --- we do not aim at a detailed statistical
+evaluation of each model. Instead, we aim at evaluating the general
+ingrowth patterns and magnitudes simulated by the models, and benchmark
+the simulated regeneration niche of multiple species against empirical
+data along a wide environmental gradients of temperature, moisture, and
+light availability.
 
 More specifically, we aimed to answer the following questions. 1) Are
-models of forests dynamics capturing correctly tree recruitment levels
-across environmental gradients? 2) Are models of forest dynamics
-representing correctly the initial composition of tree populations? And
-3), do the models of forest dynamics correctly capture the mortality in
-tree in-growth?.
+models of forest dynamics capturing accurately tree ingrowth levels
+across environmental gradients? 2) Are they matching the initial species
+composition of the tree ingrowth layer? And 3), do they correctly
+represent the mortality in tree establishment?.
 
@@
 Fifteen models of forest dynamics (including two model featuring
-variants) were used to assess tree regeneration. The selected models
-simulate forest dynamics at the stand, landscape and global scale (Table
+variants) were used to assess tree establishment. They simulate forest
+dynamics at the stand, landscape and global scale (Table
 \@ref(tab:tableModels)). The approaches used for model construction and
-their origin differ, with most of the models featuring a largely
-"process-based" approach, whereas two models based on formulations
-derived from the statistical analysis of inventory data (SYBILA and
-xComp). Some of the models largely rely on the original approach
-underlying forest gap models (ForClim 1, ForClim 11, LandClim, Landis-II
-); whereas some include plant ecophysiological processes (FORMIND,
-PICUS, 4C, iLand). These different approaches underlying the models also
-has strongly influenced the formulation of tree recruitment.
-
-Based on their establishment modules, the models can be categorized as
-regeneration or recruitment (or ingrowth) models [@vanclay1994;
-@konig_tree_2022] (see Table \@ref(tab:tableModels)). Regeneration
-models include processes such as flowering and pollination, seed
-production, seed dispersal, germination or seedling establishment that
-lead to the simulated number of seedlings or saplings. Recruitment
-models, in contrast, introduce a certain number of seedlings or saplings
-with certain characteristics such as biomass or diameter, without
-considering earlier development processes. We can further distinguish
-models that feature a feedback in the regeneration, i.e. where the level
-and species composition of regeneration is influenced by the existing
-adult tree community via the production of seeds, seedlings or saplings.
-
-The starting point for regeneration in the models also differs, ranging
-from seedbank, seed, seedling to sapling (i.e., trees much larger than
-10 cm height, often at least 2-3 m tall), which implies that the larger
+their origin differ strongly, with most of the models featuring a
+largely \"process-based\" approach, whereas two models are based on
+formulations derived from the statistical analysis of inventory data
+(SYBILA and xComp). Some of the models largely rely on the original
+approach underlying forest gap models (e.g., ForClim 1 and
+LandClim)[,]{.underline} whereas some [a]{.underline}re based on plant
+ecophysiological processes (e.g., FORMIND and iLand). The different
+approaches underlying the models have strongly influenced the
+formulation of tree establishment.
+
+The establishment modules can be categorized as \'regeneration\' or
+\'recruitment\'
+models [@vanclay1994; @konig_tree_2022] (see Table
+\@ref(tab:tableModels)). Regeneration models include processes such as
+flowering and pollination, seed production, seed dispersal, germination
+seedling [g]{.underline}rowth, which ultimately lead to the simulated
+number of trees. Recruitment models, in contrast, introduce a certain
+number of new trees with certain characteristics such as biomass or
+diameter, without explicitly considering earlier development processes.
+We can further distinguish models that feature a feedback in the
+establishment, i.e., where the level and species composition
+of establishment is influenced by the existing adult tree community via
+the
+production of seeds, seedlings or saplings, from models that do not
+contain such feedback.
+
+The starting point for in the models differs as well, ranging from
+seedbank, seed or seedling to sapling (i.e., trees much larger than 10
+cm height, often at least 2-3 m tall).This implies that the larger the
 size threshold, the larger the number of ecological processes that must
@@
 The overall complexity in the establishment modules also varies
-considerably. Following @bugmann2022, we can classified seven models
-(iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForCLim and TreeMig) as
-having rather high complexity in their tree recruitment modules (mean
-complexity across all the recruitment process with value above zero in
-Table \@ref(tab:tableModels)). The highest complexity is when defining
-processes such as the process used for modeling establishment or how is
-frost occurrence taken into account in establishment,
-
-Two models, ForClim and ForCEEPs used two alternative model versions.
-ForClim variant 1 [@bugmann1996] is based on an establishment module
-that adheres closely to the concept introduced by @botkin1972, whereas
-ForClim variant 11 is adopting a rather different approach for
-establishment model that considers
-space filling along with different \'points of entry\' for the action of
-specific
-environmental factors @huber2020. These two model variants allow us to
-evaluate a more process-based establishment module (variant 11) against
-a simpler module (variant 1) while the rest of the model structure
-remains the same. Similarly,
-the two ForCEEPS variants used here allow us to isolate the importance
-of the
-canopy feedback via seed trees for the quantity and quality of simulated
-regeneration, as one variants includes this feedback whereas the other
-one does
-not.
+considerably. Following @bugmann2022, we can classify seven models
+(iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and TreeMig) as
+having rather high complexity in their establishment modules (average
+complexity across all processes \>0, Table \@ref(tab:tableModels)). In
+most models[,]{.underline} the two processes in the establishment
+modules with the highest complexity are the statistical approach used to
+define the successful number of new treesand how frost occurrence is
+taken into account.
+
+Two models, ForClim and ForCEEPs, were used here with two alternative
+variants.
+ForClim variant 1 [@bugmann1996] is based on a recruitment module that
+adheres closely to the concept introduced by @botkin1972, whereas
+ForClim variant 11 is adopting a rather different approach that
+considers space filling along with different \'points of entry\' for the
+action of specific environmental factors [@huber2020]. These two model
+variants allow us to evaluate a more process-based module (variant 11)
+against a simple module (variant 1), while the rest of the model
+structure is identical. Similarly, the two ForCEEPS variants allow us to
+isolate the importance of the canopy feedback via seed trees for the
+quantity and quality of simulated ingrowth[,]{.underline} as one variant
+includes this feedback whereas the other does not.
 
@@
 modelDF$Sampling <- NULL
+
+#there is a better way of doing footnotes, have a look here https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf and change
+names(modelDF)[2] <- paste0(names(modelDF)[2], "(a)")
+names(modelDF)[3] <- paste0(names(modelDF)[3], "(b)")
+names(modelDF)[4] <- paste0(names(modelDF)[4], "(c)")
+names(modelDF)[5] <- paste0(names(modelDF)[5], "(d)")
+names(modelDF)[6] <- paste0(names(modelDF)[6],"(e)")
+names(modelDF)[11] <- paste0(names(modelDF)[11],"(f)")
 
@@
       kableExtra::column_spec(column = 1, width = "5em") |> 
-      kableExtra::column_spec(column = 2, width = "5em") |> 
-      kableExtra::column_spec(column = 3, width = "5em") |> 
-      kableExtra::column_spec(column = 4, width = "5em") |>
+      kableExtra::column_spec(column = 2, width = "3em") |> 
+      kableExtra::column_spec(column = 3, width = "3em") |> 
+      kableExtra::column_spec(column = 4, width = "4em") |>
       kableExtra::column_spec(column = 5, width = "5em") |>
@@
       kableExtra::column_spec(column = 11, width = "5em") |>
-      kableExtra::column_spec(column = 12, width = "5em") 
+      kableExtra::column_spec(column = 12, width = "5em") |>
+      kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
+                                 "PB-Process based, EM-Empirical", 
+                                 "C-Cohort, I-Individual",
+                                 "RG-Regeneration, RC-Recruitment",
+                                 "Y-Yes, N-No",
+                                 "D-Daily, M-Monthly"),
+
+                 notation = "alphabet") 
 ```
@@
 
-Recruitment data covering a wide range of environmental conditions are
-hard to obtain, and this is one of the reasons why most of the models of
-forest dynamics have never been confronted with a data-set covering such
-gradients over a large number sites, to evaluate how well they are
-capturing recruitment. The recruitment observations used here are from a
-novel and unprecedented network of sites in forest reserves that
-represent the range of environmental gradients in temperature and
-precipitation in Europe compiled in the framework of the EuFoRIa network
-[@euforia2019] (Figure \@ref(fig:map)). All these forests have not been
-managed for at least 10 years prior to their designation as reserves,
-and most of them provide time series of natural forest dynamics over
-multiple decades and up to ca. 80 years. The censuses periods range from
-37 to 155 years, with most of sites with an average of 14 years. The
-plot sized ranged from 0.02 ha to 5.52 ha, the recruited values were
-further processed and aggregated to provide recruited values per ha
-following @käber2022. The data provide information at the tree level,
-thus providing information for sequential comparisons such as
-recruitment and death between measurements.
+Ingrowth data covering a wide range of environmental conditions are hard
+to obtain, and this is one of the reasons why most models of forest
+dynamics have never been confronted with a dataset covering such
+gradients over a large number sites, to evaluate how well ingrowth is
+captured. The observations used here are from a novel and unprecedented
+network of sites in forest reserves that represent the range of
+environmental gradients in temperature and precipitation in central
+Europe as compiled in the framework of the EuFoRIa network[@euforia2019]
+(Figure \@ref(fig:map)). All these forests have not been managed for at
+least 10 years prior to their designation as reserves, and most of them
+provide time series of natural forest dynamics over multiple decades and
+up to ca. 80 years. The census periods range from 3 to 37 years, with an
+average of 14 years. The data provide information at the tree level,
+thus allowing for sequential comparison such as individual-tree ingrowth
+and death between the measurements.
 
 We selected 200 sites from this network as the benchmarking data set for
-the simulation exercise. Each of these sites features at least two
-consecutive measurements. Regeneration was measured at threshold of 7 or
-10 cm of stem diameter at all the sites, except for 35 sites where only
-regeneration at the 10 cm diameter threshold was available. The
-observations used in this study featured 30.900 recruited trees, and
-recruitment values per site and sample ranging from zero to 1246 trees
-per ha and a mean of 56 trees per ha recruited. For more details on this
-unique dataset and the detailed recruitment information please refer to
-@käber2022.
+the simulation. The selection of sites is representative for the
+environmental variation of the data. This was achieved by applying
+kmeans clustering to define 200 clusters of plots from the
+original set of 869 plots along the environmental dimensions of
+temperature, water balance, soil quality and exposition (i.e., slope and
+aspect). Each of these sites
+featured at least two consecutive measurements. Ingrowth was measured at
+threshold of 7 or 10 cm of stem diameter at all sites except for 35
+sites where only ingrowth at the 10 cm diameter threshold was available.
+Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
+processed and aggregated following @käber2022 to provide ingrowth rates
+per ha and per decade. The observations used in this study featured
+30.900 established trees, and ingrowth rates per site, sample and
+decade ranged from zero to 1246 trees per ha, with a mean of 56 trees
+per ha. For more details on this unique dataset and the detailed
+establishment information, cf. @käber2022.
 
@@
 
-The overarching goal of this experiment was to assess the regeneration
+The overarching goal of the experiments was to assess the regeneration
 niche as it arises from empirical data against its representation in a
-wide range of models of forest dynamics. To this end, each moding group
-was provided with a detailed protocol (available in the Supplementary
-Material X, henceforth "SM") with instructions how to perform the
+wide range of models of forest dynamics. We define the regeneration
+niche as the passing of a diameter threshold of 7 or 10
+cm[,]{.underline} respectively (synonym: ingrowth). To this end, each
+modeling group was provided with a detailed protocol (cf. Supplementary
+Material 1, henceforth \"SM\") with instructions how to perform the
 simulations, providing input variables on climate and soil conditions,
-the specification of the simulation setup and the expected output
-variables. Neither site information (except for the data specified
-below) nor any data on tree regeneration were provided prior to the
-simulation. That is, the models were run in "blind flight" mode.
+and the expected output variables. Neither were further site information
+(except for the data specified below) nor any data on tree ingrowth
+provided prior to the simulation. That is, the models were run in
+\"blind flight\" mode.
 
 The input variables were collected and aggregated to be adapted to the
-different model demands. The time series of climatic variables from 1981
-to 2018 were provided in hourly (Era5-land data, @muñozsabater2019),
-daily and monthly resolution (CHELSA data, @karger2021). Some variables
-required for some models such as relative humidity or vapor pressure
-deficit had to be calculated from these variables. The final strategy on
-how to use the climate data was different in each model based on the
-approach that best suited the model (Table \@ref(tab:tableModels)). Soil
-quality data was provided as a continuous value between 1 and 5
-(Soilgrids dataset, @hengl2017, @poggio2021). The protocol also provided
-the elevation, slope and aspect for each of the 200 sites
-[@asterscienceteam2019], no other spatial information such as
-coordinates were provided with the exception of iLand which used the
-blurred coordinates to derive highly detailed soil information values
-for the 200 sites.
-
-The simulations were run in the absence of natural disturbances, only
+needs of the different models. Time series of climatic variables from
+1981 to 2018 were provided in hourly [Era5-land data, Muñoz Sabater
+-@muñozsabater2019], daily and monthly resolution [CHELSA data, Karger
+et al. -@karger2021]. Some variables required for some models such as
+relative humidity or vapor pressure deficit had to be calculated from
+these variables. The final strategy how to use the climate data was
+different in each model, based on the approach that best suited the
+model (Table \@ref(tab:tableModels)). Soil quality data was provided as
+a continuous value between 1 and 5 [Soilgrids dataset, Hengl et al.
+-@hengl2017, Poggio et al. -@poggio2021]. The protocol also provided the
+elevation, slope and aspect for each of the 200 sites
+[@asterscienceteam2019], but no other spatial information such as
+coordinates was provided, with the exception of iLand[,]{.underline}
+which used the blurred coordinates to derive highly detailed soil data.
+
+The simulations were run in the absence of natural disturbances. Only
 the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE [ONE
 SENTENCE]{.underline} OF WHY YOU INCLUDED DISTURBANCES AND WHAT TYPE.
-The simulations were set up to sample species-specific regeneration
-rates per decade and per ha in the equilibrium state of the model,
-typically entailing a \"spin-up\" run (as we did not provide any forest
-data). The modeling teams decided on the simulated area and how they
-derived these samples. The exact length of the simulation was also
-decided by the modeling teams (Table \@ref(tab:tableModels)). Further
-details on how each modeling team prepared the simulations and the
-outputs are available in the SM 2.
+The simulations were set up to sample species-specific ingrowth rates
+per decade and per ha in the equilibrium state of the model, typically
+entailing a "spin-up" run (as we did not provide any forest data). The
+modeling teams decided on the simulated area and how they derived these
+samples. The exact length of the simulation was also decided by the
+modeling teams (Table \@ref(tab:tableModels)). Further details on how
+each modeling team prepared the simulations and the outputs are
+available in SM 2.
 
@@
 entails the assumption that (1) the observations from the forest
-reserves reflect no traces of forest management and (2) an equilibrium
-between forest dynamics and climate. While the former is likely to be
-met at least in many of the EuFoRIA reserves, the latter may be more
-debatable. However, in the absence of detailed data on the history of
-each plot in the EuFoRIa network, some assumptions had to be made. We
-define the regeneration niche as the passing of a diameter threshold of
-7 or 10 cm respectively (synonym: ingrowth). Both the width of the
-regeneration niche (i.e., in environmental space) as well as the
-intensity of the recruitment process (i.e., the number of ingrowth trees
-per area and per unit of time) were of interest.
+reserves reflect no traces of forest management and (2) there is an
+equilibrium between forest dynamics and climate. While the former is
+likely to be met at least in many of the EuFoRIA reserves, the latter
+may be more debatable. However, in the absence of detailed data on the
+history of each plot in the EuFoRIa network, some broad assumptions had
+to be made. Both the width of the regeneration niche (i.e., in
+environmental space) as well as the intensity of the ingrowth process
+(i.e., the number of ingrowth trees per area and per unit of time) were
+of interest.
 
@@
 their standard set of species for the simulations which is much larger
-(i.e. ForClim 1, ForClim 11). This also applies to TreeMig, but in this
+(i.e., ForClim 1, ForClim 11). This also applies to TreeMig, but in this
 model *Betula pendula* was not included because HERE **TREEMIG**
@@
 SPECIES. HERE **LPJ-GUESS** PROVIDES [ONE SENTENCE]{.underline} WHY THEY
-SIMULATED LESS SPECIES. Finally, the
-model aDGVM2 is based on a broad suite of Plant Functional Types rather
-than
-specific species, and hence this model used its default approach of PFTs
-for
-the simulations.
-
-Each of the models reported the number of recruits by sampling 200 times
+SIMULATED LESS SPECIES. Finally, the model aDGVM2 is based on a suite of
+broad Plant Functional Types rather than individual species, and hence
+this model used its default approach of PFTs for the simulations.
+
+Each of the models reported the ingrowth number by sampling 200 times
 over a 10-year interval for each species and per ha for each of the 200
-sites. Both the 200 samples per site as well as its average were
-reported. The multiple samples per site were used to better understand
-the variation within each site. This was done using different
-strategies, depending on the model. The strategies included (1) sampling
-simulated data from the same 1-ha plot in the equilibrium over time, or
-(2) sampling several 1-ha plots from the simulated forest at one
-specific point in time (in the equilibrium). This resulted in 880.000
+sites. Multiple samples per site were used to better understand the
+simulated variation within each site. This was done using different
+strategies, depending on the model, including (1) sampling simulated
+data from the same 1-ha plot in the equilibrium over time, or (2)
+sampling several 1-ha plots from the simulated forest at one specific
+point in time (in the equilibrium). This resulted in 880.000
 observations per model that simulated the 11 species included in the
 protocol (200 sites, 200 samples per site, 11 species and two diameter
-thresholds). For those models that simulated extra species, the
-recruitment of the extra species was aggregated into a category
-"others". The simulated data per model and the observations used in this
-study are available in a repository (REF HERE when available).
-
-Some models did not report all the simulated samples or sites to avoid
-unrealistic results. The model 4C set up a threshold of 100 m2/ha to
-avoid unrealistic stand basal area values, and therefore not all the
-samples and sites were reported, this is the consequence of 4C not being
-able to do long term simulations without management PLEASE **4C** VERIFY
-THIS IS WHAT YOU DID. LPJ-GUESS had 2% of the samples outputs as
-grasslands that were considered in the data as zero stand basal area and
-zero recruitment for all the tree species simulated. Landis II also
-missed X% of the samples HERE **LANDIS-II** PROVIDES [ONE
-SENTENCE]{.underline} WHY THEY DID NOT PROVIDE ALL THE SAMPLES.
-
-```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network that were used in the present study, showing the dominant tree species per site in the observed recruitment at the 10 cm threshold.", out.width='80%'}
+thresholds). For the models that simulated additional species, their
+ingrowth rates were aggregated as \"others\".
+
+Three models did not provide results from all the simulated samples or
+sites to avoid unrealistic results, as follows. In the model 4C, a
+threshold of a maximum basal area of 100 $m^2 ha^{-1}$ was used to avoid
+unrealistic stand basal area data, and therefore not all the samples and
+sites were reported. The reason for this is that 4C is not suitable for
+long-term simulations without management PLEASE **4C** VERIFY THIS IS
+WHAT YOU DID. LPJ-GUESS had 2% of the sample outputs as grasslands that
+were considered in the data as zero stand basal area and zero ingrowth
+for all the tree species simulated. LPJ-GUESS could not simulate one of
+the sites HERE **LPJ-GUESS** PROVIDES [ONE SENTENCE]{.underline} WHY
+THEY DID NOT PROVIDE ALL THE SITES. Landis-II also missed X% of the
+samples HERE **LANDIS-II** PROVIDES [ONE SENTENCE]{.underline} WHY THEY
+DID NOT PROVIDE ALL THE SAMPLES.
+
+```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network that were used in the present study, showing the dominant tree species per site in the observed ingrowth at the 10 cm threshold.", out.width='80%'}
 knitr::include_graphics(here::here("figures", "theMap.png"))
@@
 
-## Analysis
-
-The simulation results were analyzed in regarding the recruitment
-levels(i.e., number of recruits per unit time and space), species
-composition of the recruitment, and mortality in the recruitment. The
-observed data were modeled using a Generalized Additive Model with a
-negative binomial distribution to better understand the trend and
-regeneration levels of the observed regeneration, relative to the
-simulation results per model.
-
-Regeneration levels were evaluated along three key environmental
-gradients, i.e. stand basal area (the basal area of all the trees in
-each sample) as a proxy of light availability at the forest floor, the
-annual degree-day sum (DD) [@allen1976; @fischlin1995] as a proxy for
-growing season warmth, and the climatic water balance (CWB) as a proxy
-for soil moisture [@speich2019]. The mean values of recruitment were
-calculated as the mean across the 200 samples per site.
+## Data analysis
+
+The simulation results were analyzed regarding the ingrowth levels
+(i.e., ingrowth number per unit time and space), species composition of
+the ingrowth, and ingrowth mortality. The observed data were modeled
+using a Generalized Additive Model with a negative binomial distribution
+to better understand the trend and levels of the observed ingrowth,
+relative to the simulation results per model.
+
+Ingrowth levels were evaluated along three key environmental gradients,
+i.e., stand basal area (the basal area of all the trees in each sample)
+as a proxy of light availability at the forest floor, the annual
+degree-day sum (DD) [@allen1976; @fischlin1995] as a proxy for growing
+season warmth, and the climatic water balance (CWB) as a proxy for soil
+moisture [@speich2019]. The mean ingrowth values were calculated as the
+mean across the 200 samples per site.
 
@@
 \begin{equation}
-H_n = - \sum_{i = 1}^{s}{p_{i}ln(p_{i})}\label{eqpiShannon}
+H_n = - \sum_{i = 1}^{s}{p_{i}ln(p_{n,i})}\label{eqpiShannon}
 \end{equation}
@@
 ```
-where, $s$ is the total number of species simulated that have recruited
-basal area above zero in sample $n$; $p_{i}$ is the proportion of the
-species $i$ calculated as the recruited basal area ($r_{BA_{n,i}}$) for
-that species $i$ over the stand basal area ($Total_{BA_{n}}$) of the
-sample $n$.
-
-In order to analyze the impact of adult species composition on
-recruitment across the climatic gradients, we focused on the five main
-tree species: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus
-sylvestris*, and *Quercus* spp. For these species we calculated the mean
-recruited basal area share per site as the mean value across the
-available samples per site:
+where $s$ is the total number of species simulated that have ingrowth
+basal area above zero in sample $n$; $p_{n,i}$ is the proportion of the
+species $i$ calculated as the ingrowth basal area ($r_{BA_{n,i}}$) for
+that species $i$ relative to the stand basal area ($Total_{BA_{n}}$) of
+the sample $n$.
+
+In order to analyze the relationship between the adult species
+composition and ingrowth across the climatic gradients, we focused on
+the five main tree species or genera: *Abies alba*, *Fagus sylvatica*,
+*Picea abies*, *Pinus sylvestris*, and *Quercus* spp. For these species,
+we calculated the mean ingrowth basal area share per site as the mean
+value across the available samples per site:
 
@@
 ```
-where $s$ is the total number of species simulated; $\bar{r}_{n,i}$ is
-the mean recruitment of the species $i$ per site $t$ across the
-available samples at that site.
-
-There were 35 sites that had a diameter threshold of 10 cm only and
-therefore we do not know the actual recruitment rates for those sites at
-7 cm; yet the models simulated recruitment for the 7 cm threshold at
-those sites. The 35 sites were removed from the simulated data to make a
-fair comparison regarding the recruitment of the five main species.
-Furthermore, when looking at the recruited basal area share per species
-($\bar{R}_{BA_{share_{i}}}$), we categorized the basal area share as
-zero when the total recruited basal area (
+where $s$ is the total number of species simulated; $\bar{r}_{t,i}$ is
+the mean ingrowth of the species $i$ per site $t$ across the available
+samples at that site.
+
+The data from 35 sites that had a diameter threshold of 10 cm only, and
+therefore ingrowth rates for these sites at 7 cm are not known. The 35
+sites were removed from the simulated data to make a fair comparison
+regarding the ingrowth of the five main species. Furthermore, regarding
+the ingrowth basal area share per species ($\bar{R}_{BA_{share_{i}}}$),
+we categorized this as zero when the total ingrowth basal area (
 $\sum_{i = 1}^{s}{\bar{r}_{t,i}}$ ) equaled zero, and also when the
-recruited basal area of that species ($\bar{r}_{t,i}$) was also zero.
-
-Mortality in the regeneration was assessed based on the difference in
-recruitment between the 7 and 10 cm diameter thresholds. We used the
+ingrowth basal area of that species ($\bar{r}_{t,i}$) was zero.
+
+Mortality in tree establishment was assessed based on the ratio of
+ingrowth between the 7 and 10 cm diameter thresholds. We used the
 Reineke self-thinning rule [@reineke1933; @pretzsch2003] as a reference
-to estimate whether the different recruitment values were above or
-bellow the expected theoretical rate between these two diameters. The
-Reineke self-thinning rule is usually calculated from even-aged, single
-species stands and assumes a relationship between the number of stems
-and the quadratic mean diameter in fully stocked pure stands. The value
-used in our comparisons was 1.77, calculated using Equation
-\ref{eqReineke}.
+to estimate whether the different ingrowth values were above or below
+the expected theoretical rate between these two diameters. The Reineke
+self-thinning rule is usually calculated from even-aged, single species
+stands and assumes a relationship between the number of stems and the
+quadratic mean diameter in fully stocked pure stands. The value used in
+our comparisons was 1.77, calculated using Equation \ref{eqReineke}.
 
@@
 
-## Recruitment levels
-
-Simulation results regarding regeneration levels varied strongly across
-the 15 models and typically did not match the levels found in the forest
-reserve data. Recruitment was overestimated in most models for both the
+## Ingrowth levels
+
+The ingrowth levels varied strongly across the 15 models and typically
+did not match the levels found in the forest reserve data (Figure
+\@ref(fig:over)). Ingrowth was overestimated in most models for both the
 7 and 10 cm diameter thresholds, with the exception of the empirical
-stand model SYBILA, the landscape model Landis II and the global model
-aDGVM2, which estimated recruitment rates at the lower end of the
+stand model SYBILA, the landscape model Landis-II and the global model
+aDGVM2, which estimated ingrowth rates at the lower end of the
 plausibility interval of the observed data (Figure \@ref(fig:over)). The
-models with the highest overestimation were the stand model PICUS and
-the landscape model TreeMig. For most of the models, the variability of
-simulated recruitment levels across the 200 sites (visible from the
+models with the largest overestimation were the stand model PICUS and
+the landscape model TreeMig. For most models, the variability of
+simulated ingrowth levels across the 200 sites (visible from the
 interquartile range in the box plots of Figure \@ref(fig:over)) was
-similar to or smaller than the observed, with the notable exception of
-PICUS, where the mean recruitment variability was much larger than any
-other model.
-
-```{r over, echo=FALSE, fig.cap="Mean recruitment rates across all samples per site, plotted for the 200 sites and for each model.
+similar to or smaller than observed, with the notable exception of
+PICUS, where simulated ingrowth variability was much larger than that of
+any other model.
+
+```{r over, echo=FALSE, fig.cap="Mean ingrowth rates across all samples per site, plotted for the 200 sites and for each model.
 The red dotted lines shows the 2.5 and 97.5 percentiles for the 7 cm diameter threshold in the observed data.", out.width='100%'}
@@
 
-When evaluating the representation of recruitment rates along key
-gradients of light availability (basal area), temperature (degree-days)
-and soil moisture (climatic water balance), distinct features emerge:
-the models reproduce the effect of basal area (Figure
-\@ref(fig:nvTrend710BA)) much better than the effects of the climatic
-gradients, where they feature varying effects (supplementary figures S1
-and S2).
-
-In the observed data, recruitment rates decreased clearly with
-increasing stand basal area (Figure \@ref(fig:nvTrend710BA)). This trend
-was captured by the Landis II model only, albeit at much lower values of
-basal area than in the observations. Other models featured distinctly
-different trends, such as (1) an increase in regeneration with
-increasing total basal area followed by the absence of recruitment at
-high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and
-LPJ-GUESS), (2) almost constant regeneration levels with basal area
-(ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of
-recruitment up to a certain value of stand basal area followed by a
-decrease of recruitment at even higher values of stand basal area, with
-model-specific thresholds (4C, ForClim 1, PICUS, SYBILA and iLand).
-
-The models did not match the observed patterns in recruitment across
-gradients of soil moisture and temperature (Figures S1 and S2), and
-there was neither a clear pattern across models. Regarding the soil
-moisture gradient (Figure S1), a group of models features decreasing
-recruitment with increasing soil moisture (4C, ForClim 1, ForClim11,
-SIBYLA, xComp, PICUS and LandClim), while a few models showed the
-opposite trend (ForCEEPs(f), FORMIND and LPJ-GUESS). Across the
-temperature gradient (Figure S2), all the models that featured a
-decrease in recruitment rates with increasing soil moisture showed the
-opposite trend (i.e., more recruitment with increasing temperature),
-except for 4C. The other models did not have a clear pattern across the
-temperature gradient.
-
-Both observations and simulated data had no recruitment in some samples
-and at some sites (Table S1). The observed data had slightly more than
-over 3% of the samples with no recruitment. Only three models had a
-larger proportion of no recruitment (4C, Landis II and aDVM2). Three
-models always simulated recruitment for both 7 and 10 cm threshold
-(TreeMig and LPJ-GUESS), i.e., they did not featured zero values. The
-rest of the models had a very low percentage of samples with no
-recruitment (0.01% - 1.31%), i.e., they had distinctly fewer occurrences
-of zeros compared to the observations.
-
-```{r nvTrend710BA, echo=FALSE, fig.cap="Mean recruitment values across the 200 samples per site, for the 200 sites against the stand basal area.
-(A)same y-axis across all models; (B)y-axis scaled by model.
-The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
+When evaluating the representation of ingrowth rates along key gradients
+of light availability (basal area), temperature (degree-days) and soil
+moisture (climatic water balance), distinct features emerge: the models
+reproduced the effect of basal area (Figure \@ref(fig:nvTrend710BA))
+much better than the effects of the climatic gradients, where they
+featured varying effects (Figures S1 and S2).
+
+In the observed data, ingrowth rates decreased clearly with increasing
+stand basal area (Figure \@ref(fig:nvTrend710BA)). This trend was
+captured by the Landis-II model only, albeit at much lower values of
+basal area than in the observations. The other models featured
+distinctly different trends, such as (1) an increase in ingrowth with
+increasing total basal area followed by the absence of ingrowth at high
+values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and
+LPJ-GUESS), (2) almost constant ingrowth levels with basal area (ForClim
+11, xComp, LandClim and aDGVM2), or (3) an increase of ingrowth up to a
+certain value of stand basal area followed by a decrease at even higher
+values of stand basal area, with model-specific thresholds (4C, ForClim
+1, PICUS, SYBILA and iLand).
+
+The models did not match the observed ingrowth patterns across gradients
+of soil moisture and temperature (Figures S1 and S2), and neither was
+there a clear pattern across models. Regarding the soil moisture
+gradient (Figure S1), a group of models featured decreasing ingrowth
+with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp,
+PICUS and LandClim), while a few models showed the opposite trend
+(ForCEEPs(f), FORMIND and LPJ-GUESS). Across the temperature gradient
+(Figure S2), all the models that featured a decrease in ingrowth rates
+with increasing soil moisture showed the opposite trend (i.e., more
+ingrowth with increasing temperature), except for 4C. The other models
+did not feature a clear pattern.
+
+Both observations and simulated data had no ingrowth in some samples and
+at some sites (Table S1). The observed data had slightly more than 3% of
+the samples with no ingrowth. Only three models had a larger proportion
+of no ingrowth (4C, Landis-II and aDVM2). Three models always simulated
+ingrowth for both the 7 and 10 cm threshold (TreeMig and LPJ-GUESS),
+i.e., they did not feature any zero values. The other models had a very
+low percentage of samples with no ingrowth (0.01% - 1.31%), i.e., they
+had distinctly fewer occurrences of zeros compared to the observations.
+There was no significant relationship between the mean complexity of the
+establishment module (Table \@ref(tab:tableModels)) and the
+overestimation proportion at a diameter threshold of 7 cm (Table S2).
+
+```{r nvTrend710BA, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site, for the 200 sites against the stand basal area.
+(A) same y-axis across all models; (B) y-axis scaled by model.
+The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
     knitr::include_graphics(c(here::here("figures", "envTrend7Totba_500.png"),
@@
 
-## Species composition of recruitment
-
-Most of the models matched the level of diversity (as measured by the
-Shannon index) of the observed data quite well (Figure \@ref(fig:H710)),
-although they consistently underestimated its variability across sites.
-Four models overestimated recruitment diversity: 4C, PICUS, LANDIS-II
-and LPJ-GUESS, the latter mostly for the 10 cm diameter threshold. The
-model 4C is a special case as it simulated 5 species only, i.e., its
-diversity values are not directly comparable to those of the other
-models, nor the observations. Only one model, SIBYLA, consistently
-underestimated recruitment diversity.
+## Species composition of ingrowth
+
+Most models matched the level of diversity of the observed data quite
+well (Figure \@ref(fig:H710)), although they consistently underestimated
+its variability across sites. Four models overestimated ingrowth
+diversity: 4C, PICUS, Landis-II and LPJ-GUESS, the latter mostly for the
+10 cm diameter threshold. The model 4C is a special case as it simulated
+five species only, i.e., its diversity values are not directly
+comparable to those of the other models, nor to the observations. Only
+one model, SIBYLA, consistently underestimated ingrowth diversity.
 
 In most models, there were only small but significant differences in the
-species diversity of recruitment between 7 and 10 cm across sites (Table
-S2-S4). Furthermore, the observed data and three models (4C, SIBYLA,
-Landis II) maintained the recruitment diversity between the 7 and 10 cm
-thresholds (they were not significant differences between them, see
-Table S2), and the same was visible from the observed data. There are
-only few models with large recruitment diversity increasing between 7
-and 10 cm (ForCEEPS, TreeMig, LPJ-GUESS).
-
-```{r H710, echo=FALSE, fig.cap="Shannon index of tree recruitment, calculated by basal area as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
+species diversity of ingrowth between 7 and 10 cm across sites (Tables
+S2-S4). Furthermore, three models (4C, SIBYLA, Landis-II) maintained the
+ingrowth diversity between the 7 and 10 cm thresholds (i.e., the
+differences between them were not significant, see Table S3), and the
+same was evident from the observed data. There were a few models with a
+strong increase of ingrowth diversity between 7 and 10 cm (ForCEEPS,
+TreeMig, LPJ-GUESS, Figure \@ref(fig:H710)).
+
+```{r H710, echo=FALSE, fig.cap="Shannon index of tree ingrowth, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
 The red dotted lines represent the 20th percentile of the observed data for the threshold of 7 cm and the 70th percentile of the observed data for the threshold of 10 cm.", out.width='100%'}
@@
 
-Species diversity was lower in the recruitment compared to stand-level
-diversity, for both observed and simulated data, and for both
-recruitment thresholds (Figure S3 and S4). Models simulating higher
-diversity at the stand level did feature higher diversity in recruitment
-(Figure S3 and S4). Most of the models were capturing reasonably well or
-were overpredicting the species diversity at the stand level (Figure
-\@ref(fig:combinedRichness)A, center; cf. Figure S5). The models over
-predicting species richness at the stand level were ForClim1, ForClim
-11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A, left).
-Only a few models underpredicted species diversity at the stand level,
-i.e., LPJ-GUESS, ForCEEPS(f), xComp, iLand and 4C. (Figure
-\@ref(fig:combinedRichness)A, right). The majority of the models
-captured reasonably well the recruitment diversity at lower values of
-diversity, but most of them strongly underestimated species diversity in
-recruitment for higher observed diversity levels (Figure
+For both the observed and the simulated data, and both ingrowth
+thresholds, species diversity was lower in the ingrowth compared to
+stand-level diversity (Figures S3 and S4). Yet, models simulating higher
+diversity at the stand level featured higher diversity in ingrowth
+(Figure S3 and S4). Most models captured reasonably well or
+overpredicted species diversity at the stand level (Figure
+\@ref(fig:combinedRichness)A, center; cf. Figure S5)[,]{.underline} but
+a characteristic feature was that simulations had a much lower
+variability of diversity compared to observations. Models overpredicting
+species richness included ForClim1, ForClim 11, FORMIND and TreeMig
+(Figure \@ref(fig:combinedRichness)A, left). Several models
+underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C
+(Figure \@ref(fig:combinedRichness)A, right). The majority of the models
+captured reasonably well the ingrowth diversity at lower values of
+diversity, but most of them strongly underestimated ingrowth diversity
+for higher observed diversity levels (Figure
 \@ref(fig:combinedRichness)B, cf. Figure S6). Both the models with
-feedback and those without feedback from the adult trees to recruitment
-(via seed production) underestimated species diversity in the
-recruitment compared to the observations.
+feedback and those without feedback from the adult trees to ingrowth
+(via seed production) underestimated ingrowth species diversity compared
+to the observations. Again, simulated diversity was considerably less
+variable than observed diversity (Fig. \@ref(fig:combinedRichness)B).
 
 ```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site and for observed and simulated values.
-n indicates the number of models falling in each group A) The three trends present across models for the species diversity at the stand level.
-B) The three trends present across models for recruitment at the 7 cm threshold.
-The representative model selected here per trend group was based on the full data shown in Figure S5 and S6.", out.width='75%'}
+n indicates the number of models falling in each group.
+A) Examples for the three trends across models for species diversity at the stand level.
+B) Examples for the three trends across models for ingrowth at the 7 cm threshold.
+The selection of the representative model for each trend group was based on the full data shown in Figures S5 and S6.", out.width='75%'}
 knitr::include_graphics(here::here("figures", "combined_richness.png"))
@@
 frequently in the EuFoRIa data. Most models captured well the share of
-basal area in the recruitment of these five tree species compared to the
+basal area in the ingrowth of these five species compared to the
 observed data along the two major gradients of climate (Figure
-\@ref(fig:diff), cf. Figure S7). However, the models differed in the way
-in which they simulated the climate niche of these main species. First,
-some models overestimated recruitment across the entire gradients for at
-least one of the five main tree species (e.g., ForCEEPS, ForCEEPS(f),
-LandClim or LPJ-GUESS), while other models overestimated recruitment for
-more than one species for a large part of the climate gradients, such as
-xComp, PICUS, SYBILA or LPJ-GUESS (cf. red colors in Figure
-\@ref(fig:diff)). These trends were consistent for both regeneration
-thresholds, although the exact changes in the share of recruited basal
-area were somewhat different (Figure S7).
+\@ref(fig:diff), cf. Figure S7). However, the models differed
+considerably in the way in which they simulated their climatic niche.
+
+First, some models overestimated ingrowth across the entire gradients
+for at least one of the five main tree species (Figure \@ref(fig:diff),
+e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models
+overestimated ingrowth for more than one species for a large part of the
+climate gradients, such as xComp, PICUS, SYBILA or LPJ-GUESS (cf. red
+colors in Figure \@ref(fig:diff)). These trends were consistent for both
+ingrowth thresholds, although the exact changes in the share of the
+ingrowth basal area were somewhat different (Figure S7).
 
 Second, there was some congruence in the simulation results by species
-across the models. The share in the recruitment by Silver fir (*Abies
-alba*) was captured well across the gradient by most models except for
-xComp, PICUS, LandClim and Landis-II, which overestimated the
-recruitment share of this species. There are also some particularities
-evident for some models, such as 4C that did not simulate *Abies alba*.
-*Fagus sylvatica* recruitment was underestimated by many models across
-the environmental gradients, including ForCEEPS(f) and FORMIND, which
-featured negligible regeneration at most sites for this species. Most of
-the models overestimated the *Picea abies* recruitment share in the
-cold-wet part of the gradients. *Pinus sylvestris* and *Quercus* spp.
-represented a smaller share of the recruited species in the observed
-data mostly at warm-dry sites. Many models, however, erroneously
-featured recruitment for these two species in most of the climatic
-gradient although at a low basal area share. Five models (4C, FORMIND,
-SYBILA, xComp, and LandClim) had almost no recruitment or did not
-feature any recruitment of *Pinus sylvestris* and *Quercus* spp. at all.
+across the models. The share in the ingrowth by Silver fir (*Abies
+alba*) was captured well across the gradient by most
+models except for xComp, PICUS, LandClim and Landis-II, which
+overestimated the ingrowth share of this species. There were also some
+peculiarities evident for some models, such as 4C that did not simulate
+*Abies alba*. *Fagus sylvatica* ingrowth was underestimated by many
+models across the environmental gradients, including ForCEEPS(f) and
+FORMIND, which featured
+negligible ingrowth at most sites for this species. Most of the models
+overestimated the share of *Picea abies* ingrowth in the cold-wet part
+of the gradients. *Pinus sylvestris* and *Quercus*
+spp. represented a smaller share of the ingrowth in the observed data
+mostly at warm-dry sites. Many models, however, erroneously featured
+ingrowth for these two species along most of the climatic gradients
+although at a low basal area share. Five models (4C, FORMIND, SYBILA,
+xComp, and LandClim) had almost no ingrowth of *Pinus sylvestris* and
+*Quercus*
+spp., or did not feature any ingrowth of of these species at all.
 
 Lastly, there is no model that performed well across the five species
-regarding the species-specific recruitment rates in the environmental
-space of Figure 6. Some models (e.g., ForClim11, PICUS) tended to
-consistently overestimate the recruitment share of all five species, but
-most models overestimated the recruitment share of some species while
-underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
-
-```{r diff, echo=FALSE, fig.cap="Recruitment rates for the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm recruitment threshold.
-The values shown are the mean of the 200 samples per site and across the sites in each bins (tile), with 10 bins per gradient.
-The sizes of the circles represent the ratio between recruited basal area of the species and the total recruited basal area for all species.
-The absence of a circle indicates a zero basal area share in the recruitment, or the absence of recruitment altogether.
-The color gradient (for the models only) shows the difference between the simulated and observed ratio of recruited basal area of the species and the total recruited basal area for all the species.", out.width='100%', out.height= '90%'}
+regarding the species-specific ingrowth rates in the environmental space
+of Figure 6. Some models (e.g., ForClim11, PICUS) tended to consistently
+overestimate the ingrowth share of all five species, but most models
+overestimated the ingrowth share of some species while underestimating
+the share of others (e.g., ForCEEPS, LPJ-GUESS).
+
+```{r diff, echo=FALSE, fig.cap="Ingrowth rates of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm ingrowth threshold.
+The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient.
+The sizes of the circles represent the ratio between the ingrowth basal area of the species and the total ingrowth basal area for all species.
+The absence of a circle indicates a zero basal area share in the ingrowth, or the absence of ingrowth altogether.
+The color gradient (for the models only) shows the difference between the simulated and observed ratio of ingrowth basal area of the species and the total ingrowth basal area for all the species.", out.width='100%', out.height= '85%'}
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
@@
 
-## Mortality in regeneration
+## Ingrowth mortality
 
@@
 data, thus at least partially compensating for the general
-overestimation of recruitment rates (Figure \@ref(fig:over) and Table
-S5); six models did not have a significantly difference compared to the
+overestimation of ingrowth rates (Figure \@ref(fig:over) and Table S5);
+six models did not feature a significant difference compared to the
 observed data (ForCEEPS, FORMIND, ForClim 1 and 11, and aDGVM2), and
@@
 rates (i.e., well above the Reineke self-thinning line), which
-compensated for the strong initial overestimation of recruitment (cf.
+compensated for the strong initial overestimation of ingrowth (cf.
 Figure 2). Interestingly, two models that underestimated overall
-recruitment levels (SIBYLA and Landis-II; cf. Figure \@ref(fig:over))
-still featured mortality rates in the recruitment that were close to but
-still above the self-thinning line (Figure \@ref(fig:mortality11)).
-
-On the one hand, the fact that some models feature lower mortality than
-expected by Reineke\'s self-thinning rule may point to facilitation, or
-simply a higher mortality rate
-before the trees had reached 7 cm diameter. On the other hand, the case
-of models
-such as xComp, PICUS and TreeMig, which are well above the self-thinning
-line, may
-indicate that they feature higher recruitment mortality in mixed-species
-forests
-compared to even-aged, single-species forests. However, some models
-yielded a mortality rate of recruitment that is broadly compatible with
-the self-thinning rule, i.e. 4C and LandClim.
-
-The three models with largest overestimation proportion of recruitment
-at 7cm threshold showed had the largest mortality between 7 and 10 cm
-thershold (Figure \@ref(fig:mortality13)A). The rest of the models that
-overestimated recruitment had mortalities lower than the expected shelf
-thinning ratio (Figure \@ref(fig:mortality13)B). However, most of the
-models with a low variation in mortality ratio between 7 and 10 cm
-recruitment (ForCEEPS, ForClim 1, ForClim 11 and LandClim) (Figure S8)
-feature a small overestimation of recruitment at the 7 cm threshold
-(Figure \@ref(fig:mortality11); cf. Figure S8). Some of the models that
-start at a lower number of individuals (7 cm threshold) have a lower
-decrease between 7 and 10 cm dbh (ratio 7-10 cm) (Figure S10); in these
-models the species competition and self-thinning is either not that
-large or it happened before the species reached 7 cm. However, this
-pattern it is not consistent across models (Figure S9). There is an
-effect of the number of individuals at 7cm on the mortality rate. Most
-of the models show a positive effect (increasing recruitment at 7cm
-shows increases the mortality between 7 and 10cm). Some models show a
-negative effect, however the trend it is not very pronounced and not
-always significant regarding the slope of the linear trends per model
-(Figure S9 and Table S2).
-
-```{r mortality11, echo=FALSE, fig.cap="Ratio of tree recruitment rates between the 7 and 10 cm thresholds.
-Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree recruitment between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
+ingrowth levels (SIBYLA and Landis-II; cf. Figure \@ref(fig:over))
+featured mortality rates that were close to but still above the
+self-thinning line (Figure \@ref(fig:mortality11)).
+
+The fact that some models featured lower mortality than expected by
+Reineke\'s self-thinning rule may point to facilitation, or simply a
+higher mortality rate before the trees had reached 7
+cm diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
+whose mortality was well above the self-thinning line, indicates that
+they feature higher ingrowth mortality in mixed-species forests compared
+to even-aged, single-species forests. Some models yielded a mortality
+rate of ingrowth that is broadly compatible with the self-thinning rule
+(i.e., 4C and LandClim).
+
+The three models with the largest overestimation of the proportion of
+ingrowth at the 7 cm threshold featured the highest mortality between
+the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The other
+models that overestimated ingrowth had mortalities lower than the
+expected self-thinning ratio (Figure \@ref(fig:mortality13)B). It is
+noteworthy that most of the models that featured a low ratio of ingrowth
+between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11 and
+LandClim) (Figure S8) had only a small overestimation of ingrowth at the
+7 cm threshold (Figure \@ref(fig:mortality11) cf. Figure S8). aDGVM2 was
+the only model that underestimated ingrowth at the 7 cm threshold while
+featuring a low mortality between the two thresholds (Figure
+\@ref(fig:mortality13)C), and only a few models were close to the
+correct ingrowth rate at 7 cm while also being close to the Reineke line
+(SIBYLA, Landis-II; cf. Figure \@ref(fig:mortality13)D; and 4C as well
+as LandClim).
+
+Lastly, some models that started at a low number of individuals (7 cm
+threshold) had a lower decrease of ingrowth between 7 and 10 cm dbh
+(ratio 7-10 cm) (Figure S10). In these models, competition and
+self-thinning are either not pronounced or must have occurred before the
+species had reached 7 cm. However, this pattern was not consistent
+across models (Figure S9). It is noteworthy that there is a relationship
+between the ingrowth rate at 7 cm and the mortality rate: most of the
+models showed a positive effect (higher ingrowth at 7 cm is coupled to
+an increase of the mortality between 7 and 10 cm). While some models
+showed a negative effect, this linear trend was not pronounced and not
+always significant regarding the slope (Figure S9 and Table S2).
+
+```{r mortality11, echo=FALSE, fig.cap="Ratio of tree ingrowth rates between the 7 and 10 cm thresholds.
+Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree ingrowth between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
 knitr::include_graphics(here::here("figures", "ratio7_10.png"))
@@
 
-```{r mortality13, echo=FALSE, fig.cap="Ratio of tree recruitment between 7 and 10 cm recruitment treshold and the overestimation proportion (simulated - observed / observed) for the mean recruitment trees per model across sites and samples.
-The horizontal dotted lines show a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke self-thinning ratio under even-aged conditions.
-The vertical line represents the 0 overestimation proportion", out.width='100%'}
+```{r mortality13, echo=FALSE, fig.cap="Ratio of tree ingrowth between the 7 and 10 cm thresholds (ingrowth 7 cm / ingrowth 10 cm) and the overestimation proportion at 7 cm ([simulated -observed] / observed) for the mean ingrowth per model across sites and samples.
+The horizontal dotted lines show a ratio equal to 1, indicating no decrease of ingrowth between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
+The vertical line indicates zero overestimation.", out.width='80%'}
 knitr::include_graphics(here::here("figures", "mort7_10_medianOver.png"))
@@
 
-Tree recruitment is a fundamental process in forest dynamics, in
+Tree establishment is a fundamental process in forest dynamics, in
 particular with increasing disturbance under climate change. Correctly
-capturing forest regeneration in dynamic models is fundamental to, for
-example, evaluate post-disturbance dynamics as this will define the
-status of the forests in the first 15-20 years
-[@seidl_post-disturbance_2022]. If this initial state is not captured
-well, we can not properly assess aspects such as the forest resilience,
-or the timing, magnitude and progression of carbon capture. In this
-study, for the first time, projections of tree recruitment from multiple
-models of forest dynamics were confronted with a unique dataset from
-unmanaged forest reserves across a large environmental gradient in
-Europe.
-
-The observed data that we used [@käber2022] are also exceptional with
-respect to the number of recruitment events captured, which is essential
-for a highly stochastic process such as tree recruitment. The use of
-this dataset for model benchmarking provided novel insights on the
-capability of state-of-the art models in simulating recruitment levels,
-the species composition of recruitment, and mortality in an early stage,
-i.e. between a tree diameter of 7 and 10 cm. Overall, when adopting this
+capturing tree ingrowth in dynamic models is
+fundamental to, for example, evaluate post-disturbance dynamics as it
+will define forest state in the first decades [@seidl2022]. If this
+initial phase is not captured well, we can not properly assess aspects
+such as forest resilience or the timing, magnitude and progression of
+carbon sequestration. In this study, for the first time, projections of
+tree establishment from multiple models of forest dynamics were
+confronted with a unique dataset from unmanaged forest reserves across a
+large environmental gradient in Europe.
+
+The observed data that we used [@käber2022] are also exceptional,
+particularly with respect to the number of events captured, which is
+essential for a highly \'noisy\' process such as tree establishment. The
+use of this dataset for model benchmarking provided novel insights on
+the capability of state-of-the art models to simulate ingrowth levels,
+species composition, and mortality in an early stage of tree life, i.e.,
+between a tree diameter of 7 and 10 cm. Overall, by adopting this
 approach a much broader understanding resulted than if we had used these
 data for model calibration: it is primarily from the failure of the
-models that we can gain ecological insights [@trugman2022].
-
-## Recruitment levels
-
-Accurate simulation of tree recruitment rates is pivotal for projecting
-the dynamics of forest stand structure and composition. Most of the
-models of forest dynamics evaluated here tended to overestimate tree
-recruitment rates across the 200 sites. Overestimating has far-reaching
-implications regarding biomass (and carbon) turnover, with a potential
-overestimation in their capacity for carbon capture [@pugh2019].
-Importantly, excess recruitment requires higher mortality rates in later
-stages, either -- as observed for
-some models -- between the 7 and 10 cm diameter thresholds, or either
-before or
-shortly after this stage. Correcting for the expected forest densities
-at early
-stages is thus equivalent to compensating for a first error (excess
-recruitment) by committing a second error (excess mortality). Thus, many
-models appear to get the right \'answer\', i.e., correct stand structure
-and composition, for the wrong reasons. This is problematic particularly
-if the models are to be used under novel conditions (e.g., climate
-change; [@huber2021]) or in a decision support context (e.g.,
-[@thrippleton2021]).
-
-In spite of this criticism, it remains remarkable that most of the
-models were not overly off
-with respect to simulated regeneration levels -- after all, few if any
-of the models\' establishment routines are sufficiently constrained by
-data, and the resulting degrees of freedom that are available for
-modeling regeneration (cf. [@konig_tree_2022]) could have implied that
-model performance would be much worse.
-
-Regeneration levels define the success of future forests, but it is
-equally important to correctly identify unstocked forest areas
-[@rammer2021]. Tree recruitment data are characterized by zero inflation
-[@fortin2007]. This was clearly evident from the EuFoRIa dataset, but
-some of the models did not produce zeros at all, or featured a very low
-proportion of zero data. This substantial difference may be due to the
-assumption that the observed data is also in equilibrium situation, and
-the lack of recruitment in the data set could be an indication of a
-non-equilibrium e.g., due to management history.
-
-Light availability as a strong filter for tree regeneration has been
-documented widely [@messier1999; @collet2006; @berdanier2016]. However,
-it is difficult to collect variables defining light availability on
-regeneration at a large spatial and temporal scales. In this study, we
-used the total stand basal area as a proxy for light availability. We
-found pronounced differences in the stand basal area ranges across
-models, but also between models and the observation. This made
-impossible to evaluate the recruitment for the extremes of the density
-ranges in some models. For example, regeneration rates at low stand
-densities are relevant to assess how well forests are recovering e.g.,
-after gap creation due to disturbance[@seidl_post-disturbance_2022;
-@grubb1977]. We expected that with increasing stand basal area, the
-recruitment of the different species would become sparser and drop out
-entirely under low-light condition [@klopcic2012; @zell2019]. However,
+models that we can gain ecological insights [e.g., @trugman2022].
+
+## Ingrowth levels
+
+Most of the models tended to overestimate tree ingrowth rates. This has
+potentially far-reaching implications, e.g., regarding biomass (and thus
+carbon) turnover, with a potential overestimation of the capacity for
+forest carbon sequestration [@pugh2019; @pan2011]. Yet, we focused on
+one specific stage, i.e., ingrowth into the 7 and 10 cm diameter
+classes. Trees at this size contribute little to carbon sequestration,
+and if the excess ingrowth rate at this stage is compensated -- in the
+models -- soon thereafter by higher mortality, overall simulated
+ingrowth may still yield trustworthy results at the level of stand
+structure and related forest attributes such as basal area or biomass.
+
+From a modeling point of view, excess ingrowth inevitable requires
+excess mortality rates
+in a later stage, either -- as observed for some models -- between the 7
+and 10 cm diameter thresholds, before the 7 cm are reached, or soon
+after the 10 cm threshold has been crossed. In any case, correcting at
+early stages for the expected forest densities at later stages is
+equivalent to compensating for a first error (excess ingrowth) by a
+second error (excess mortality). Thus, many models appear to get the
+\'right answer\', i.e., correct stand structure and composition, for the
+wrong reasons. It is highly likely that biased projections will result,
+because the two errors are unlikely to be perfectly linked and thus will
+not always compensate each other. Hence, this structural problem of most
+models investigated here is problematic particularly if the models are
+to be used under novel conditions such as under climate change [e.g.,
+@huber2021] or in a decision support context [e.g., @thrippleton2021].
+
+In spite of this criticism, it is remarkable that most models were not
+overly off with respect to simulated ingrowth levels -- after all, few
+if any of the models' establishment routines are sufficiently
+constrained by data, with the exception of the empirical models xComp
+and SIBYLA. It is noteworthy that in spite of their empirical basis,
+both models were off relative to empirical ingrowth levels, in similar
+magnitude as the other models and even in opposite directions (SIBYLA:
+underestimation; xComp: overestimation). It appears that using empirical
+data of limited geographical scope, as done in either of these models,
+leads to extrapolation problems already under current climatic
+conditions. For the other models, whose establishment module is not
+strongly constrained by empirical data, the degrees of freedom that are
+available for modeling ingrowth [@konig_tree_2022] could have implied
+that model performance would be much worse.
+
+Ingrowth levels define the structure and composition of future forests,
+but it is equally important to correctly identify areas where ingrowth
+is lacking [@rammer2021]. There are multiple constraints to the
+regeneration niche of tree species [@price2001], and therefore the
+absence of ingrowth is likely to be common [@fortin2007], even over
+larger areas such as the one-hectare plots used here. Tree ingrowth data
+are characterized by zero-inflation. This was clearly evident from the
+EuFoRIa dataset, but some of the models did not produce zeros at all, or
+featured a very low proportion of zero data. This substantial difference
+may be due to the fact that the simulation results were drawn from
+equilibrium forests, whereas in reality many of the forest reserves are
+recovering from past management activities and have become denser and
+darker over the past decades [e.g., @heiri2009], leading to less
+ingrowth than in an equilibrium situation.
+
+Competition for light as a strong filter for tree establishment has been
+documented widely
+[@messier1999; @collet2006; @berdanier2016]. However, it is difficult to
+measure light availability at large spatial and temporal scales. We used
+total stand basal area as a proxy for light availability [cf.
+@schmid2021]. We found pronounced differences in the stand basal area
+ranges simulated by the models, but also between the models and the
+observations. This made it impossible to evaluate the establishment for
+the extremes of the density ranges in some models. For example, ingrowth
+rates at low stand densities are relevant to assess how well forests are
+recovering e.g., after gap creation due to disturbance [@seidl2022;
+@grubb1977]. At the level of the simulated one-hectare samples, average
+basal area is typically not very low as long as no larger disturbances
+occur, which was explicitly excluded in our simulation protocol. Towards
+the other end of the spectrum, i.e., with increasing basal area, it
+would be reasonable to expect that with increasing stand basal area, the
+ingrowth of the different species would become sparser and drop out
+entirely under low-light conditions [@klopcic2012; @zell2019]. However,
 few models showed this trend, thus indicating that the relation between
-recruitment and light availability it is not yet captured correctly in
-the models.
-
-## Species composition of recruitment
-
-Correctly capturing tree species composition of recruitment is important
-to assess the future functional diversity of a forest, e.g., its
-sensitivity to drought or resilience to disturbances events
-[@redmond2015; @seidl_post-disturbance_2022]. In the model simulations,
-overall species diversity levels in recruitment were within the observed
-ranges. Yet, there were variations on how the main tree species were
-represented in the environmental gradient. Most models maintain or even
-increase in recruitment diversity between the 7 and 10 cm thresholds,
-and the same is visible from the observed data, however the differences
-are not significant for all the models. Species richness variations
-might be explained by the way each model considers recruitment
-[@konig_tree_2022], e.g., in the absence of adults, recruitment can be
-linked to either seed rain-dispersal, or to seed/seedling bank, but also
-on what processes are formulated with stochastic or deterministic
-approach. Differences might also be expected between cohort and
-individual trees based models. MAKE REFERENCES TO LITERATURE
-
-Our results do not show differences between models with and without
-feedback in recruitment. The adult forests species richness was better
-captured than the recruitment species richness, and only few models
-overpredicted stand level species richness. Most of the models have
-feedback on recruitment from the adult trees, therefore EXPAND ON
-FEEDBACK ISSUE
-
-## Mortality in regeneration
-
-As mentioned above, it is reasonable to expect that models that
-over-recruit trees will also have a couting-acting high mortality
-between the 7 and 10cm thresholds, but this was nit cinsistently evident
-from the simulations. This implies that these models must have the
-excess of mortality in earlier or later stages. Higher mortality in the
-adult component might enhance the share of less shade tolerant species
-for those models. There might be a difference between models including
-stochastic processes in recruitment and those which do not. There was
-also a group of models below the Reineke line indicating a lower
-mortality than expected by the self-thinning rule, which may point to
-facilitation, or simply a higher mortality rate before the trees reached
-7 cm diameter, another possibility would have been that these models
-underestimate recruitment which was not the case. This might also be
-partly an artifact of the method, as we are expecting space for time
-substitution, but in reality we just have one measurement. NEED MORE
-REFERENCES AND DISCUSSION POINTS
-
-## Research recomendations
-
-There is also the question of how much level of detail one has to give
-to define regeneration processes [@konig_tree_2022; @bugmann2022].
-Models with more detailed and complex recruitment modules, will have
-also increase parametrisation which will lead to lower generalization as
-this values will have to be collected from specific geographical
-regions. This is significant relevant when looking at mixed forests and
-the species level parametrization. **How this relates to the results?
-Future model development recommendations?**
-
-In this study, we considered a wide range of models of forest dynamics,
-from stand to global scale. However, the simulation set up was limited
-to 1 ha recruitment values. It is important to include spatial
-considerations on tree recruitment [@beckage2003] and the lack of it
-might represent an more important disadvantage for landscape level
-models. Without spatial context we are limiting recruitment to the seed
-influx from the stand itself if the model does not have a background
-seed input. But even global vegetation models are lacking this dispersal
-between cells where there is an abstraction of the space or even the
-trees themselves [@hanburybrown_forest_2022]. Landscape level models
-have been build to be very accurately at the landscape level. EXPAND ON
-THE DIFFERENT SCALE LEVELS AND IMPLICATIONS
-
-Modeling approach seems more important than scale and accuracy of the
-design. Model behavior (accuracy) is completely unrelated to model scale
-(stand, landscape, global) and approach (empirical, process-based).
-Having both empricial and process based models, the expectation was that
-the empirically based models (SIBYLA, xComp) would have a better
-performance as they are calibrated with inventory data, however this was
-not the case. EXPAND ON
-
-The selected models also had limitations on species representation, some
-models where not able to simulate certain species due to lack of species
-parametrization. We also had one model that simulate PFTs providing a
-less flexible comparison the with species values; with the lose of the
-species specific information might also represent substantial ecological
-representativeness drawbacks [@purves2008].
-
-**METHODS PARAGRAPH**
-
-Yet, in this study we were not able to assess tree regeneration in a
-holistic sense. The
-lower size threshold of 7 cm in the observed data constituted a hard
-limit based
-on which we can understand some aspects of tree recruitment, which in
-its
-entirety often comprises a rather long period since germination
-[@price2001], and many relevant environmental constraints are acting on
-these young trees [@käber2021] that we were unable to assess. However,
-recruitment data with lower thresholds are simply not available in a
-harmonized manner across large environmental gradients. The EuFoRIa data
-used here were collected in forest reserves where no management has
-taken place for long periods of time. This makes our assumption of an
-equilibrium between forest properties and environmental drivers less
-haphazard than it might appear at first sight. However this assumption
-might also justify some of the rather large recruitment rates that we
-are seeing in the models projections compare to the observed ones, as if
-the observed data it is not in equilibrium the observed recruitment
-would be lower as they are not close to the optimum.
-
-RECOMENDATION PARAGRAPH
-
-Cautious when using this models to evaluate regeneration post
-disturbances we might be overestimation their capacity for carbon
-capture
+ingrowth and light availability is not yet captured correctly in most
+models.
+
+## Species composition of ingrowth
+
+Correctly capturing species composition in tree establishment is
+important to assess the future functional diversity of a forest, e.g.,
+its sensitivity to drought or resilience to disturbances [@redmond2015;
+@seidl2022]. In the simulations, overall ingrowth species diversity
+levels were well within the observed range for two thirds of the models
+(10 out of 15). Thus, while most of the models are facing difficulties
+to quantitatively match ingrowth levels (cf. above), their performance
+is much better when it comes to matching the quality, i.e., the
+composition of ingrowth as a function of abiotic and biotic conditions.
+
+Yet, there were pronounced variations how the main tree species were
+represented along the environmental gradients. Most models maintained or
+even increased ingrowth diversity between the 7 and 10 cm thresholds,
+and the same is visible from the empirical data, but the differences are
+not significant. Diversity variations may be explained by the way the
+individual models are considering establishment processes
+[@konig_tree_2022], e.g., in the absence of adults, ingrowth can be
+linked to either seed rain-dispersal, or to a seed/seedling bank. The
+ingrowth rates were sampled from simulations in the equilibrium, and in
+this state it is expected that non-climax species such as *Pinus
+sylvestris* or *Quercus* would be of minor importance, or absent
+entirely [@klopcic2015]. Most models captured this difference, which is
+also found in the empirical data. Thus, although the broad patterns are
+matched by many models, improvements in the quantification of the
+regeneration niche of the species are needed, but this cannot be done in
+the absence of robust datasets.
+
+The species diversity of the entire stand was better captured by the
+models than the diversity of tree ingrowth, and only a few models
+overpredicted stand level diversity. Ingrowth composition defined as
+being proportional to adult abundance, regardless of productivity, might
+be a simple and cautious assumption to safely maintain relative species
+abundances [@hanbury-brown2022]. However, based on the simulation
+results, we did not see that models with feedback captured better the
+ingrowth species diversity than those without feedback. The similar
+performance between models with and without feedback might be because
+the models are giving more importance to define the regeneration niche
+via abiotic and biotic filters than with the habitat niche of the adult
+trees [@grubb1977].
+
+## Ingrowth mortality
+
+There are several factors that lead to mortality in establishment such
+as competition [@casper1997] in addition to multiple abiotic factors
+[@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
+expect that models that overestimate ingrowth of small trees will have a
+particularly high mortality between the 7 and 10 cm thresholds. However,
+this was not consistently evident from the simulations. This implies
+that these models must have the excess of mortality in earlier or later
+stages, if we assume that all models would be able to capture the
+structure and composition of adult stands along the EuFoRIa gradient,
+which however was not tested here.
+
+Higher mortality towards the adult stage has important implications for
+forest dynamics and the goods and services provided by forests. On the
+one hand, mortality in later stages may erroneously enhance the share of
+less shade tolerant species in the models [@klopcic2015], therefore
+shifting the species composition. Unrealistic high stem densities for a
+longer period of time may overestimate the role of tree regrowth in
+carbon sink dynamics [@pugh2019]. Given our set of simulations and
+analyses, we cannot conclusively assess what is going on in the models,
+and further studies are required that focus on a wider range of tree
+sizes and the fate of tree establishment along such a size continuum.
+
+## Methodological considerations
+
+The EuFoRIa data as used in this study is unique and highly suited to
+better understand tree establishment. However, three aspects of this
+data may also represent considerable limitations.
+First, we made a comparison of tree ingrowth in a posited equilibrium
+state, but we cannot assess how close the empirical data were to such a
+state. The data were collected in forest reserves where no management
+has taken place for long periods of time. This makes our assumption of
+an equilibrium between forest properties and environmental drivers less
+haphazard than it might appear at first sight. However, it may explain
+why simulated ingrowth rates were higher than the observed ones. Second,
+the data were collected for ingrowth above a 7 cm threshold, thus
+limiting the assessment of tree establishment in a holistic sense. This
+constituted a hard limit based on which we can understand some aspects
+of tree establishment, which in its entirety often comprises a rather
+long period since germination [@price2001]. However, many environmental
+constraints are acting on young trees [@käber2021] that we were unable
+to assess. Yet, ingrowth data with lower thresholds are simply not
+available in a harmonized manner across large environmental gradients.
+Third, the empirical data were
+collected from rather small plots that ranged from 0.2 to up to 1 ha,
+while we sampled simulated ingrowth rates from 1 ha areas, which may
+lead to an incorrect representation of space. Even though this strategy
+it is not ideal, it represents a common challenge when harmonizing
+diverse data sources originating from varied sampling strategies
+[@portier2022]. It would have been extremely challenging for such a
+variety of models to follow a protocol where the spatial sampling size
+was different at each of the 200 sites.
+
+The design of our sampling protocol did not include spatial aspects. We
+considered a wide range of models of forest dynamics, from stand to
+global scale. However, the simulation setup was limited to 1 ha ingrowth
+values. The lack of spatial considerations potentially puts landscape
+models at a disadvantage, as they have been built to be accurate~~l~~ at
+the landscape level. Without spatial context, we are limiting ingrowth
+to the seed influx from the stand itself, if the model does not have a
+background seed input. However, Global Dynamic Vegetation Models should
+not be in disadvantage due to the limited spatial considerations [but
+cf. Snell et al. -@snell2014] as they are usually lacking dispersal
+between cells as they are based on an abstraction of horizontal space,
+or even the trees themselves [@hanburybrown_forest_2022].
+
+## Research recommendations
+
+Increasing complexity in establishment modules has been motivated by
+better process understanding and enhancing of model accuracy
+[@bugmann2022; @bugmann2019]. However, more complex models do not
+necessary lead to better projections and rather to a reduced
+transparency and predictive power [@franklin2020]. The results from this
+study showed that increasing complexity in the establishment modules did
+not represent an improvement in the projections of ingrowth levels,
+species composition or mortality at early tree stages. Thus, the
+question what level of detail one has to use to model establishment
+processes remains open and must be addressed in future research
+[@konig_tree_2022; @bugmann2022].
+
+If it should be adequate to include more detail in the establishment
+models, this will come with higher parameterisation efforts, which will
+most likely lead to lower generalization as the required data will have
+to be collected from specific locations, as there is no general
+establishment dataset available. We recommend on the one hand that the
+further development of the establishment modules is implemented as
+additional features that can be traced back, as done in this study for
+the model variants of ForClim and ForCEEPs, and that model complexity
+and structure must always be connected with model objectives
+[@albrich2020]. On the other hand, we emphasize that datasets like
+EuFoRIa used here are invaluable and should be further expanded in both
+their spatial extent (e.g., towards boreal and Mediterranean conditions)
+as well as in time (e.g., continuing the monitoring into the future).
+
+Based on the results of this study, it is clear that models of forest
+dynamics need a focus on their establishment modules. However, here we
+have only consider tree establishment in the equilibrium state. It is
+equally important to understand how these models project tree
+establishment after changes in forest structure like those brought about
+by disturbances [@seidl2022], or under different management strategies
+[@lindner2000]. However, this will require an entirely different set of
+observed data, and potentially not all models of forest dynamics would
+be able to assess the relationship of these aspects on tree
+establishment, e.g., due to the lack of disturbance or appropiate
+management modules.
+
+We recommend to further investigate the implications of the current
+modeling strategies for tree establishment, particularly those related
+to excess tree establishment and probably late excess mortality, by
+focusing on a wider range of tree sizes and the related ingrowth
+dynamics. We also recommend, especially for the landscape-level models,
+the inclusion of explicit spatial considerations regarding tree ingrowth
+[@beckage2003]; this, however, is facing a serious challenge regarding
+appropriate data sets, as inventory data have wide coverage, but do not
+allow for the assessment of spatial interaction effects.
 
@@
 
-Capture properly recruitment rates, species composition and mortality in
-the environmental gradient might impact future forests...
-
-Provide succinct answers to the three questions/objectives stated at the
-end of the
-Introduction. Perhaps end with a concluding (\"more positive\") final
-paragraph.
-
-Models eventually get the right stand structure and composition, most of
-them do that
-by compensating for one error (overestimation of recruitment) by another
-error
-(excessive subsequent mortality, if not between 7 and 10 cm, then
-shortly
-thereafter): \"getting the right answer for the wrong reason\". And this
-is
-worrysome in general, but particularly (and very acutely) if the models
-are to
-be used e.g. under changing climatic conditions.
+Models of forest dynamics are important tools in science and decision
+support, and how tree establishment is defined has strong implications
+for simulated forest properties. The 15 models and variants used here
+are facing similar challenges when representing tree establishment. They
+generally overestimate tree establishment rates, and the simulated
+establishment niche is not always captured accurately as a function of
+the biotic (light) and abiotic (temperature and moisture). However, most
+models properly capture the diversity of the initial tree populations,
+although changes in their formulations including feedback from adult
+trees did not have an effect on better capturing the species composition
+of ingrowth.
+
+Regarding mortality in the early phase of tree life, many models that
+feature a particularly high overestimation of ingrowth levels are
+compensating this by a larger tree mortality. Often, however, this
+compensation is not sufficient to reduce the high ingrowth levels to
+realistic values. Overall, however, there is no clear pattern across all
+models.
+
+When capturing tree establishment, the specific design decisions taken
+in the development of any model are more important for its behavior
+(accuracy) than scale (stand, landscape, global), background (empirical
+vs. process-based), and complexity. Having both empirical and
+process-based models in our set, the empirically-based models could have
+been expected to have a better performance, as they were calibrated with
+inventory data, but this was not the case. Similarly, higher model
+complexity does not represent an improvement for capturing tree
+establishment.
+
+Even though the establishment routines of most of the models
+investigated here have never been sufficiently constrained by robust
+data, the ingrowth projections are not overly off. This indicates that a
+lot can be gained by a focus on the modeling of establishment processes.
+The representation of forest dynamics in these models would become much
+more robust particularly in the face of climate change and
+post-disturbance dynamics, thus strongly reducing the uncertainty in
+long-term projections of future forest dynamics.
 
@@
 available at zenodo and gitlab (links to be added)
+
+# Acknowledgements
+
+Funding of the workshop by COST (Action No CA19139) and ETH Zurich
+(Grant ETH-35 18-1).
 

### 2022-12-14 code/dataPreparation.R
add count of landis missing samples

@@ -196,7 +196,21 @@
 
 
     ### Landis II ----
-    
+    landisSPP <- data.table::fread("data/simulation_results/landisII/landis_biomass_sum/landis_biomass_sum_site_species_sample.csv")
+    #landis_biomass_sum_site_species_sample.csv: species biomass per site and sample
+    # - all cohorts including regeneration - 142 425 -
+    # this should represent species present, the rest should be zero,
+    # the species were not there at the sampling time
+
+    landisSITE <- data.table::fread("data/simulation_results/landisII/landis_biomass_sum/landis_biomass_sum_site_sample.csv")
+    #landis_biomass_sum_site_sample: total biomass per site and sample -
+    #this has 39916 rows, almost the 40 000 required,
+    #the rest should be zeros, I am sure there were sites where nothing can grow
+    #(extreme climate) which will make the 84 totally missing.
+    #You can once again do your biomass->ba and the magic where you select
+    #the higher biomass for the ones where we have only regeneration.
+
+
     landis <- data.table::fread("data/simulation_results/landisII/LANDIS-II_data_version_03.csv")
     landis$model <- "Landis II"
     landis$dbh[landis$dbh == "7 cm" ] <- 7
@@ -265,19 +279,8 @@
      
      #Correct that there are certain samples from certain sites only present in 
      #one dbh threshold but not in the other
-     #add them assuming 0 recruitment for those sites and samples IS THIS OK?
+     #add them assuming 0 recruitment for those sites and samples 
     
-     dd <- as.data.frame(table(landis$site))
-     sitesWLessObs <- dd$Var1 [dd$Freq < 4400]
-     
-     length(unique(landis$sample[landis$site == 33]))
-     test <- landis[landis$site == 33,]
-     #table(test$species)
-     #table(test$sample)
-     
-     #View(test[test$sample == 117,])
-     # View(landis[landis$site == 33 & landis$sample == 117, ])
-   
      missing7 <- as.data.frame(table(landis$sample[landis$dbh == 7], landis$site[landis$dbh == 7]))
      missing10 <- as.data.frame(table(landis$sample[landis$dbh == 10], landis$site[landis$dbh == 10]))
      
@@ -299,16 +302,77 @@
      landis <- rbind(landis,  mis7,  mis10) 
      
 
-     # Check that there are 200 samples per site NO! WHAT TO DO WITH THIS
-     # landis <- landis |> 
-     #   dplyr::group_by(site, dbh) |>
[…]

### 2022-12-14 code/dataPreparation.R
4C correction data on the BA error estimation

@@ -98,9 +98,9 @@
     forceepsfeedback$model <- "ForCEEPS(f)"
 
     ### 4C ----
-    fourC <- data.table::fread("data/simulation_results/4C/recruitment_data_4c.csv")
+    fourC <- data.table::fread("data/simulation_results/4C/recruitment_data_4c_all_221213.csv")
     fourC$model <- "4C"
-
+    
     ### SIBYLA ----
     sibyla <- data.table::fread("data/simulation_results/sibyla/TabTreesSibyla0_KJMerganic.csv")
     sibyla$model <- "SIBYLA"
@@ -580,6 +580,8 @@
                                        picus, xcomp, sibyla, fourC, formind,
                                        iland, landis, landclim,
                                        treemig, lpjguess, adgvm2)
+    
+    simulationData$end_year <- NULL 
 
     # Add environmental variables to simulated data 
     

### 2022-12-14 code/figures.R
cleaning of figures code

@@ -24,7 +24,7 @@
                    dplyr::group_by(site, sample) |>
                    dplyr::summarise( r.ha= sum(r.trees))
     
-    hist(obsNumAgg$r.ha)
+    #hist(obsNumAgg$r.ha)
     range (obsNumAgg$r.ha, na.rm = TRUE)
     
     # Assign models and dbh to levels 
@@ -375,8 +375,7 @@
     mtext(labEnvVarDDS, side = 1, line = 1, outer = TRUE)
     
     dev.off()
-    
-    
+  
 # Mortality ----    
   ### Fig. R ratio 7/10 ----     
     # Data preparation    
@@ -492,8 +491,8 @@
                        legend.title = ggplot2::element_blank(),
                        strip.background =  ggplot2::element_blank(),
                        legend.position = "none") +
-        ggplot2::xlab(label = "Proportion of overestimation in tree recruiment at 7cm diameter threshold") +
-        ggplot2::ylab(label = "Ratio of recruitment at 7 and 10cm diameter threshold") +
+        ggplot2::xlab(label = "Proportion of overestimation (7 cm)") +
+        ggplot2::ylab(label = "Ratio of recruitment (7 and 10 cm)") +
         
         ggrepel::geom_label_repel(ggplot2::aes(label = model, size = NULL, 
                                                color = NULL),
@@ -502,6 +501,7 @@
     
     ggplot2::ggsave("figures/mort7_10_medianOver.png",
                     plot =  overMortmedian,
+                    width = 16, height = 12, scale = 0.9,
                     dpi = 300, units = "cm", device = 'png')         
     
 # Diversity in recruitment ----
@@ -750,13 +750,14 @@
     ### Prepare data to have a set of Observed and simulated values per model
     empShannon <- ShannonIndex[ShannonIndex$model == "Observed",]
     plotShannonDF <- ShannonIndex[!ShannonIndex$model == "Observed", ]
+
     # Repeat All Rows (Multiple)
     empShannon <- empShannon |>
         dplyr::slice(rep(1:dplyr::n(), length(unique(plotShannonDF$model))))
     empShannon$model <- rep(unique(plotShannonDF$model), each = 579)
     empShannon$type <- "Observed"
     plotShannonDF$type <- "model"
-    
+
     dataPlot <- rbind(empShannon, plotShannonDF)
     dataPlot$type <- factor(dataPlot$type, levels = c("Observed",  "model"))
[…]

### 2022-12-14 manuscript/manuscriptRW.Rmd
updates in the manuscript

@@
 ---
-title: "(WT1) Tree establishment in models of forest dynamics: a key priority for further research  \n(WT2) Tree establishment processes in models of forest dynamics: capturing the drivers remains challenging"
+title: "Tree establishment in models of forest dynamics: a key priority for further research"
 author:
@@
 Even though the overestimation of tree establishment represent a
-challenge
-for the usefulness of model~~s~~ projections under uncertain conditions
-such a climate change, and although some formulations should be
-reconsidered to correct excessive ingrowth levels, it is remarkable that
-most of the~~se~~ models capture the essential features of the highly
-complex process of tree establishment, while not having been
+challenge for the usefulness of model~~s~~ projections under uncertain
+conditions such a climate change, and although some formulations should
+be reconsidered to correct excessive ingrowth levels, it is remarkable
+that most of the~~se~~ models capture the essential features of the
+highly complex process of tree establishment, while not having been
 parameterized with such a data. We conclude that in spite of all the
@@
 Forests provide a wide range of products and services of vital
-importance to humankind.
-Under the pressure of climate change and increasing disturbances, it is
-becoming ever more important to understand how forest structure,
-composition and function will change, and to evaluate forest capacity to
-adapt or be resilient after disturbance [@lindner2010]. A wide range of
-models of forest dynamics were developed over the past decades
-considering the impacts of climate on forest dynamics [@bugmann2022].
-From these studies, it is evident that while we have a reasonably good
-understanding how to model tree growth [@vanclay1997; @bugmann1996], and
-substantial efforts have been dedicated to improving the depiction of
-tree mortality [@bugmann2019], tree establishment remains a much less
-studied process that is often represented rather coarsely in these
-models, which is a clear research gap in the context of climate-induced
-forest disturbances and forest resilience.
+importance to humankind. Under the pressure of climate change and
+increasing disturbances, it is becoming ever more important to
+understand how forest structure, composition and function will change,
+and to evaluate forest capacity to adapt or be resilient after
+disturbance [@lindner2010]. A wide range of models of forest dynamics
+were developed over the past decades considering the impacts of climate
+on forest dynamics [@bugmann2022]. From these studies, it is evident
+that while we have a reasonably good understanding how to model tree
+growth [@vanclay1997; @bugmann1996], and substantial efforts have been
+dedicated to improving the depiction of tree mortality [@bugmann2019],
+tree establishment remains a much less studied process that is often
+represented rather coarsely in these models, which is a clear research
+gap in the context of climate-induced forest disturbances and forest
+resilience.
 
@@
 growth and survival [@price2001]. All these processes are difficult to
-assess, and some of them
-are scarcely understood and thus appear highly stochastic
-[@bogdziewicz2021]. Identifying the appropriate level of complexity for
-the formulation of the key factors that are leading to successful tree
-establishment is challenging. Currently, tree
+assess, and some of them are scarcely understood and thus appear highly
+stochastic [@bogdziewicz2021]. Identifying the appropriate level of
+complexity for the formulation of the key factors that are leading to
+successful tree establishment is challenging. Currently, tree
 establishment processes in dynamic forest models are handled in a
-multitude of
-ways [@konig_tree_2022; @bugmann2022], from entirely ignoring it [as
-done in classical forest growth models, e.g., Pretzsch et al.
--@pretzsch2002], across the use of a few simple environmental filters,
-as done in most forest \"gap\" models [@shuman2011] and Dynamic Global
-Vegetation Models [e.g., @hickler2012; @smith2001], to complex
+multitude of ways [@konig_tree_2022; @bugmann2022], from entirely
+ignoring it [as done in classical forest growth models, e.g., Pretzsch
+et al. -@pretzsch2002], across the use of a few simple environmental
+filters, as done in most forest "gap" models [@shuman2011] and Dynamic
+Global Vegetation Models [e.g., @hickler2012; @smith2001], to complex
 approaches that incorporate local feedback from the canopy, multiple
@@
 @wehrli2006], or field-based statistical parameterizations, which
-however are not easy to extrapolate in
-space and time [e.g., @ribbens1994].
+however are not easy to extrapolate in space and time [e.g.,
+@ribbens1994].
 
@@
 e.g., in the context of global change scenarios. Given the current
-strategies that are used in models of forest
-dynamics to represent tree establishment, model behavior is often facing
-problems that can have strong impacts on simulated forest dynamics, such
-as very high levels of tree establishment, which then necessitate excess
-mortality at early stages of tree life to arrive at correct stand
-structure and composition. Also, inadequately high species diversity in
-tree establishment may be simulated. This is characteristic of
-\"classical\" forest gap models [@gutiérrez2016], at least as long as
-the simulations comprise a multitude of species. Other models use
-calibration against local canopy-level data to define appropriate
-ingrowth levels, which is likely to strongly hamper the general
-applicability of these models e.g., under scenarios of climate change.
-Furthermore, correctly capturing the species composition in tree
-establishment as a function of the presence of seed trees in the canopy
-is often a particular challenge, potentially leading to
-unrealistic successional drift in the model, which must be corrected by
-factors that are hard to parameterize [e.g., @lischke2006]. A related
-issue is the excessive reduction of species richness due to positive
-feedback effects such that eventually just single-species ~~forest~~
-stands remain. This is sometimes corrected by the incorporation of a low
-level of seed influx of all species at all times [@lischke2006].
-However, simulated species composition is usually exceedingly sensitive
-to that level, whereas the parameters of such functions are poorly
-constrained by field data. Lastly, there is often a problem with
-insufficient observational constraint on parameter values for models
-that start from very small tree sizes (e.g., 10 cm of height or even
-seeds) and track tree development in a process-oriented manner by
-considering a multitude of ecological influences, rather than
-emphasizing tree establishment into a larger size class [e.g.,
-@wehrli2006].
+strategies that are used in models of forest dynamics to represent tree
+establishment, model behavior is often facing problems that can have
+strong impacts on simulated forest dynamics, such as very high levels of
+tree establishment, which then necessitate excess mortality at early
+stages of tree life to arrive at correct stand structure and
+composition. Also, inadequately high species diversity in tree
+establishment may be simulated. This is characteristic of "classical"
+forest gap models [@gutiérrez2016], at least as long as the simulations
+comprise a multitude of species. Other models use calibration against
+local canopy-level data to define appropriate ingrowth levels, which is
+likely to strongly hamper the general applicability of these models
+e.g., under scenarios of climate change. Furthermore, correctly
+capturing the species composition in tree establishment as a function of
+the presence of seed trees in the canopy is often a particular
+challenge, potentially leading to unrealistic successional drift in the
+model, which must be corrected by factors that are hard to parameterize
+[e.g., @lischke2006]. A related issue is the excessive reduction of
+species richness due to positive feedback effects such that eventually
+just single-species ~~forest~~ stands remain. This is sometimes
+corrected by the incorporation of a low level of seed influx of all
+species at all times [@lischke2006]. However, simulated species
+composition is usually exceedingly sensitive to that level, whereas the
+parameters of such functions are poorly constrained by field data.
+Lastly, there is often a problem with insufficient observational
+constraint on parameter values for models that start from very small
+tree sizes (e.g., 10 cm of height or even seeds) and track tree
+development in a process-oriented manner by considering a multitude of
+ecological influences, rather than emphasizing tree establishment into a
+larger size class [e.g., @wehrli2006].
 
@@
 on tree establishment [@käber2022]. By tree establishment, we refer to
-the passing of trees across a specific diameter threshold (\"ingrowth\")
+the passing of trees across a specific diameter threshold ("ingrowth")
 and thus as successful tree establishment. We included models that are
-based on a range of \"philosophies\", from models operating at the stand
+based on a range of "philosophies", from models operating at the stand
 to the global scale as well as the range of models from empirically
-derived to \"process-based\" [e.g., @bugmann1996; @fabrika2012;
+derived to "process-based" [e.g., @bugmann1996; @fabrika2012;
 @mette2014; @lexer2001]. Because of the large variability in tree
@@
 their origin differ strongly, with most of the models featuring a
-largely \"process-based\" approach, whereas two models are based on
+largely "process-based" approach, whereas two models are based on
 formulations derived from the statistical analysis of inventory data
@@
 
-The establishment modules can be categorized as \'regeneration\' or
-\'recruitment\'
-models [@vanclay1994; @konig_tree_2022] (see Table
+The establishment modules can be categorized as 'regeneration' or
+'recruitment' models [@vanclay1994; @konig_tree_2022] (see Table
 \@ref(tab:tableModels)). Regeneration models include processes such as
@@
 We can further distinguish models that feature a feedback in the
-establishment, i.e., where the level and species composition
-of establishment is influenced by the existing adult tree community via
-the
+establishment, i.e., where the level and species composition of
+establishment is influenced by the existing adult tree community via the
 production of seeds, seedlings or saplings, from models that do not
@@
 Two models, ForClim and ForCEEPs, were used here with two alternative
-variants.
-ForClim variant 1 [@bugmann1996] is based on a recruitment module that
-adheres closely to the concept introduced by @botkin1972, whereas
-ForClim variant 11 is adopting a rather different approach that
-considers space filling along with different \'points of entry\' for the
+variants. ForClim variant 1 [@bugmann1996] is based on a recruitment
+module that adheres closely to the concept introduced by @botkin1972,
+whereas ForClim variant 11 is adopting a rather different approach that
+considers space filling along with different 'points of entry' for the
 action of specific environmental factors [@huber2020]. These two model
@@
 environmental variation of the data. This was achieved by applying
-kmeans clustering to define 200 clusters of plots from the
-original set of 869 plots along the environmental dimensions of
-temperature, water balance, soil quality and exposition (i.e., slope and
-aspect). Each of these sites
-featured at least two consecutive measurements. Ingrowth was measured at
-threshold of 7 or 10 cm of stem diameter at all sites except for 35
-sites where only ingrowth at the 10 cm diameter threshold was available.
-Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
-processed and aggregated following @käber2022 to provide ingrowth rates
-per ha and per decade. The observations used in this study featured
-30.900 established trees, and ingrowth rates per site, sample and
-decade ranged from zero to 1246 trees per ha, with a mean of 56 trees
-per ha. For more details on this unique dataset and the detailed
-establishment information, cf. @käber2022.
+kmeans clustering to define 200 clusters of plots from the original set
+of 869 plots along the environmental dimensions of temperature, water
+balance, soil quality and exposition (i.e., slope and aspect). Each of
+these sites featured at least two consecutive measurements. Ingrowth was
+measured at threshold of 7 or 10 cm of stem diameter at all sites except
+for 35 sites where only ingrowth at the 10 cm diameter threshold was
+available. Plot size ranged from 0.02 ha to 5.52 ha, and the data were
+further processed and aggregated following @käber2022 to provide
+ingrowth rates per ha and per decade. The observations used in this
+study featured 30.900 established trees, and ingrowth rates per site,
+sample and decade ranged from zero to 1246 trees per ha, with a mean of
+56 trees per ha. For more details on this unique dataset and the
+detailed establishment information, cf. @käber2022.
 
@@
 modeling group was provided with a detailed protocol (cf. Supplementary
-Material 1, henceforth \"SM\") with instructions how to perform the
+Material 1, henceforth "SM") with instructions how to perform the
 simulations, providing input variables on climate and soil conditions,
@@
 (except for the data specified below) nor any data on tree ingrowth
-provided prior to the simulation. That is, the models were run in
-\"blind flight\" mode.
+provided prior to the simulation. That is, the models were run in "blind
+flight" mode.
 
@@
 thresholds). For the models that simulated additional species, their
-ingrowth rates were aggregated as \"others\".
+ingrowth rates were aggregated as "others".
 
@@
 the sites HERE **LPJ-GUESS** PROVIDES [ONE SENTENCE]{.underline} WHY
-THEY DID NOT PROVIDE ALL THE SITES. Landis-II also missed X% of the
-samples HERE **LANDIS-II** PROVIDES [ONE SENTENCE]{.underline} WHY THEY
-DID NOT PROVIDE ALL THE SAMPLES.
+THEY DID NOT PROVIDE ALL THE SITES. Landis-II missed 0.2% of the
+observations HERE **LANDIS-II** PROVIDES [ONE SENTENCE]{.underline} WHY
+THEY DID NOT PROVIDE ALL THE SAMPLES/observations.
 
@@
 across the models. The share in the ingrowth by Silver fir (*Abies
-alba*) was captured well across the gradient by most
-models except for xComp, PICUS, LandClim and Landis-II, which
-overestimated the ingrowth share of this species. There were also some
-peculiarities evident for some models, such as 4C that did not simulate
-*Abies alba*. *Fagus sylvatica* ingrowth was underestimated by many
-models across the environmental gradients, including ForCEEPS(f) and
-FORMIND, which featured
-negligible ingrowth at most sites for this species. Most of the models
-overestimated the share of *Picea abies* ingrowth in the cold-wet part
-of the gradients. *Pinus sylvestris* and *Quercus*
-spp. represented a smaller share of the ingrowth in the observed data
-mostly at warm-dry sites. Many models, however, erroneously featured
-ingrowth for these two species along most of the climatic gradients
-although at a low basal area share. Five models (4C, FORMIND, SYBILA,
-xComp, and LandClim) had almost no ingrowth of *Pinus sylvestris* and
-*Quercus*
-spp., or did not feature any ingrowth of of these species at all.
+alba*) was captured well across the gradient by most models except for
+xComp, PICUS, LandClim and Landis-II, which overestimated the ingrowth
+share of this species. There were also some peculiarities evident for
+some models, such as 4C that did not simulate *Abies alba*. *Fagus
+sylvatica* ingrowth was underestimated by many models across the
+environmental gradients, including ForCEEPS(f) and FORMIND, which
+featured negligible ingrowth at most sites for this species. Most of the
+models overestimated the share of *Picea abies* ingrowth in the cold-wet
+part of the gradients. *Pinus sylvestris* and *Quercus* spp. represented
+a smaller share of the ingrowth in the observed data mostly at warm-dry
+sites. Many models, however, erroneously featured ingrowth for these two
+species along most of the climatic gradients although at a low basal
+area share. Five models (4C, FORMIND, SYBILA, xComp, and LandClim) had
+almost no ingrowth of *Pinus sylvestris* and *Quercus* spp., or did not
+feature any ingrowth of of these species at all.
 
@@
 The fact that some models featured lower mortality than expected by
-Reineke\'s self-thinning rule may point to facilitation, or simply a
-higher mortality rate before the trees had reached 7
-cm diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
-whose mortality was well above the self-thinning line, indicates that
-they feature higher ingrowth mortality in mixed-species forests compared
-to even-aged, single-species forests. Some models yielded a mortality
-rate of ingrowth that is broadly compatible with the self-thinning rule
-(i.e., 4C and LandClim).
+Reineke's self-thinning rule may point to facilitation, or simply a
+higher mortality rate before the trees had reached 7 cm diameter. Yet,
+the case of models such as xComp, PICUS and TreeMig, whose mortality was
+well above the self-thinning line, indicates that they feature higher
+ingrowth mortality in mixed-species forests compared to even-aged,
+single-species forests. Some models yielded a mortality rate of ingrowth
+that is broadly compatible with the self-thinning rule (i.e., 4C and
+LandClim).
 
@@
 particular with increasing disturbance under climate change. Correctly
-capturing tree ingrowth in dynamic models is
-fundamental to, for example, evaluate post-disturbance dynamics as it
-will define forest state in the first decades [@seidl2022]. If this
-initial phase is not captured well, we can not properly assess aspects
-such as forest resilience or the timing, magnitude and progression of
-carbon sequestration. In this study, for the first time, projections of
-tree establishment from multiple models of forest dynamics were
-confronted with a unique dataset from unmanaged forest reserves across a
-large environmental gradient in Europe.
+capturing tree ingrowth in dynamic models is fundamental to, for
+example, evaluate post-disturbance dynamics as it will define forest
+state in the first decades [@seidl2022]. If this initial phase is not
+captured well, we can not properly assess aspects such as forest
+resilience or the timing, magnitude and progression of carbon
+sequestration. In this study, for the first time, projections of tree
+establishment from multiple models of forest dynamics were confronted
+with a unique dataset from unmanaged forest reserves across a large
+environmental gradient in Europe.
 
@@
 particularly with respect to the number of events captured, which is
-essential for a highly \'noisy\' process such as tree establishment. The
+essential for a highly 'noisy' process such as tree establishment. The
 use of this dataset for model benchmarking provided novel insights on
@@
 From a modeling point of view, excess ingrowth inevitable requires
-excess mortality rates
-in a later stage, either -- as observed for some models -- between the 7
-and 10 cm diameter thresholds, before the 7 cm are reached, or soon
-after the 10 cm threshold has been crossed. In any case, correcting at
-early stages for the expected forest densities at later stages is
-equivalent to compensating for a first error (excess ingrowth) by a
-second error (excess mortality). Thus, many models appear to get the
-\'right answer\', i.e., correct stand structure and composition, for the
-wrong reasons. It is highly likely that biased projections will result,
-because the two errors are unlikely to be perfectly linked and thus will
-not always compensate each other. Hence, this structural problem of most
-models investigated here is problematic particularly if the models are
-to be used under novel conditions such as under climate change [e.g.,
-@huber2021] or in a decision support context [e.g., @thrippleton2021].
+excess mortality rates in a later stage, either -- as observed for some
+models -- between the 7 and 10 cm diameter thresholds, before the 7 cm
+are reached, or soon after the 10 cm threshold has been crossed. In any
+case, correcting at early stages for the expected forest densities at
+later stages is equivalent to compensating for a first error (excess
+ingrowth) by a second error (excess mortality). Thus, many models appear
+to get the 'right answer', i.e., correct stand structure and
+composition, for the wrong reasons. It is highly likely that biased
+projections will result, because the two errors are unlikely to be
+perfectly linked and thus will not always compensate each other. Hence,
+this structural problem of most models investigated here is problematic
+particularly if the models are to be used under novel conditions such as
+under climate change [e.g., @huber2021] or in a decision support context
+[e.g., @thrippleton2021].
 
@@
 Competition for light as a strong filter for tree establishment has been
-documented widely
-[@messier1999; @collet2006; @berdanier2016]. However, it is difficult to
-measure light availability at large spatial and temporal scales. We used
-total stand basal area as a proxy for light availability [cf.
-@schmid2021]. We found pronounced differences in the stand basal area
-ranges simulated by the models, but also between the models and the
-observations. This made it impossible to evaluate the establishment for
-the extremes of the density ranges in some models. For example, ingrowth
-rates at low stand densities are relevant to assess how well forests are
-recovering e.g., after gap creation due to disturbance [@seidl2022;
-@grubb1977]. At the level of the simulated one-hectare samples, average
-basal area is typically not very low as long as no larger disturbances
-occur, which was explicitly excluded in our simulation protocol. Towards
-the other end of the spectrum, i.e., with increasing basal area, it
-would be reasonable to expect that with increasing stand basal area, the
-ingrowth of the different species would become sparser and drop out
-entirely under low-light conditions [@klopcic2012; @zell2019]. However,
-few models showed this trend, thus indicating that the relation between
-ingrowth and light availability is not yet captured correctly in most
-models.
+documented widely [@messier1999; @collet2006; @berdanier2016]. However,
+it is difficult to measure light availability at large spatial and
+temporal scales. We used total stand basal area as a proxy for light
+availability [cf. @schmid2021]. We found pronounced differences in the
+stand basal area ranges simulated by the models, but also between the
+models and the observations. This made it impossible to evaluate the
+establishment for the extremes of the density ranges in some models. For
+example, ingrowth rates at low stand densities are relevant to assess
+how well forests are recovering e.g., after gap creation due to
+disturbance [@seidl2022; @grubb1977]. At the level of the simulated
+one-hectare samples, average basal area is typically not very low as
+long as no larger disturbances occur, which was explicitly excluded in
+our simulation protocol. Towards the other end of the spectrum, i.e.,
+with increasing basal area, it would be reasonable to expect that with
+increasing stand basal area, the ingrowth of the different species would
+become sparser and drop out entirely under low-light conditions
+[@klopcic2012; @zell2019]. However, few models showed this trend, thus
+indicating that the relation between ingrowth and light availability is
+not yet captured correctly in most models.
 
@@
 better understand tree establishment. However, three aspects of this
-data may also represent considerable limitations.
-First, we made a comparison of tree ingrowth in a posited equilibrium
-state, but we cannot assess how close the empirical data were to such a
-state. The data were collected in forest reserves where no management
-has taken place for long periods of time. This makes our assumption of
-an equilibrium between forest properties and environmental drivers less
+data may also represent considerable limitations. First, we made a
+comparison of tree ingrowth in a posited equilibrium state, but we
+cannot assess how close the empirical data were to such a state. The
+data were collected in forest reserves where no management has taken
+place for long periods of time. This makes our assumption of an
+equilibrium between forest properties and environmental drivers less
 haphazard than it might appear at first sight. However, it may explain
@@
 available in a harmonized manner across large environmental gradients.
-Third, the empirical data were
-collected from rather small plots that ranged from 0.2 to up to 1 ha,
-while we sampled simulated ingrowth rates from 1 ha areas, which may
-lead to an incorrect representation of space. Even though this strategy
-it is not ideal, it represents a common challenge when harmonizing
-diverse data sources originating from varied sampling strategies
-[@portier2022]. It would have been extremely challenging for such a
-variety of models to follow a protocol where the spatial sampling size
-was different at each of the 200 sites.
+Third, the empirical data were collected from rather small plots that
+ranged from 0.2 to up to 1 ha, while we sampled simulated ingrowth rates
+from 1 ha areas, which may lead to an incorrect representation of space.
+Even though this strategy it is not ideal, it represents a common
+challenge when harmonizing diverse data sources originating from varied
+sampling strategies [@portier2022]. It would have been extremely
+challenging for such a variety of models to follow a protocol where the
+spatial sampling size was different at each of the 200 sites.
 

### 2022-12-14 manuscript/supplementaryRW.Rmd
add correc numbering of figures in the supplementary material

@@
 - \linenumbers
+- \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+
 ---
@@
 
-# Supplementary figures and tables
-
-## Recruitment levels
-
-```{r envTrend7wb, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the sites water balance.
-(A) same y-axis across all models; (B) y-axis scaled by model.
-The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data..", fig.show = 'hold', out.width = '80%'}
-    knitr::include_graphics(c(here::here("figures", "envTrend7WB_500.png"),
-                              here::here("figures", "envTrend7WB.png")))
-```
-
-```{r envTrend7dds, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the site seasonal degree-day sum.
-(A)same y-axis across all models; (B) y-axis scaled by model.
-The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
-    knitr::include_graphics(c(here::here("figures", "envTrend7DDS_500.png"),
-                              here::here("figures", "envTrend7DDS.png")))
-```
-
-```{r zeroInflated, echo=FALSE,  message=FALSE}
-library(knitr)
-
-zeroInflated <- readr::read_csv(here::here("data", "noRecruitment.csv"))
-zeroInflated$Model[zeroInflated$Model == "Empirical"] <- "Observations"
-
-kable(zeroInflated,  format = "latex", caption = "Number of total samples, samples with zero recruitment and the proportion of zero recruitment per model and observed data.")|>
-        kableExtra::landscape() |>
-        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
-```
-
-## Recruitment species composition
-
-```{r sigDiff, echo=FALSE,  message=FALSE}
-library(knitr)
-
-sigDiff <- readr::read_csv(here::here("figures", "sigH7_10.csv"))
-
-kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of Shannon index in each model for 7 and 10 cm threshold.") |>
-        #kableExtra::landscape() |>
-        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
-```
-
-```{r sigH7_Obs, echo=FALSE,  message=FALSE}
-library(knitr)
-
-sigDiff7 <- readr::read_csv(here::here("figures", "sigH7_Obs.csv"))
-
-kable(sigDiff7,  format = "latex", caption = "Test of the significance of the differences of Shannon index between the observed data and each model for 7cm threshold.") |>
-        #kableExtra::landscape() |>
-        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
-```
-
-```{r sigH10_Obs, echo=FALSE,  message=FALSE}
-library(knitr)
-
-sigDiff10 <- readr::read_csv(here::here("figures", "sigH10_Obs.csv"))
-
-kable(sigDiff10,  format = "latex", caption = "Test of the significance of the differences of Shannon index between the observed data and each model for 10cm threshold.") |>
-        #kableExtra::landscape() |>
-        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
-```
-
-```{r recrStand7, echo=FALSE, fig.cap="Mean Shannon index per site for the recruitment  and at the stand level both for observed and simulated values.
-NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
-knitr::include_graphics(here::here("figures", "recruitmentAdultRichness7.png"))
-```
-
-```{r recrStand10, echo=FALSE, fig.cap="Mean Shannon index per site for the recruitment  and at the stand level both for observed and simulated values.
-NOTE: The recruitment treshold recruitment for treshold 10 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
-knitr::include_graphics(here::here("figures", "recruitmentAdultRichness10.png"))
-```
-
-```{r simObadult , echo=FALSE, fig.cap="Mean species richness at the stand level per site in the simulated versus observed species richness.
-The species richness is estimaed using the Shannon index  calculated by basal area.
-A1) groups models that Simulated species composition in adults is higher than the observed, A2) groups models that Simulated species composition in adults is similar to the observed, A3) groups models where the imulated species composition in adults is lower than the observed.
-NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
-knitr::include_graphics(here::here("figures", "allAgesRichness.png"))
-```
-
-```{r simObr, echo=FALSE, fig.cap="Mean species richness per site in the simulated recruitment versus observed species richness.
-The species richness is estimaed using the Shannon index is calculated by basal area.B1) groups models where the simulated species composition is closer than the other models to observed values, B2) groups models where the sites with lower species composition are underestimated, B3) groups models where the simulated species composition recruitment is lower in the whole range compare to the observe.
-NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
-knitr::include_graphics(here::here("figures", "recruitmentRichness.png"))
-```
-
-```{r diff, echo=FALSE, fig.cap="Departure of each model in the representation of the main species recruitment relative abundance from the observations across the environmental gradients for a 10cm treshold.
-The values shown are the mean of samples per site and across sites in bins, where the range of the climatic water balance (cm) and seasonal degree-day sum (dºC) was divided in 10 bins.
-The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species.
-The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equalt to zero.", out.width='100%'}
-knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_10.png"))
-```
-
-## Mortality
-
-```{r mortTable, echo=FALSE,  message=FALSE}
-library(knitr)
-
-sigDiff <- readr::read_csv(here::here("figures", "mortRatio7_10.csv"))
-
-kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold over 10 cm threshold.") |>
-#kableExtra::landscape() 
-kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
-```
-
-```{r mortality12 , echo=FALSE, fig.cap="Interquartile range of the ratio between 7 and 10cm recruitment and the overestimation proportion (sim-obs/obs) for the mean recruitment trees per model across sites and samples", out.width='100%'}
-knitr::include_graphics(here::here("figures", "mort7_10_iqrOver.png"))
-```
-
-```{r mortality22 , echo=FALSE, fig.cap="Recruitment ration between 7 and 10cm and the initial recruitment tree number at 7cm.
-Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions.
-The cloud point per models shows the differences across sites, the shape point cloud per model shows if the models take into account regeneration density when applying mortality to regeneration ", out.width='60%'}
-knitr::include_graphics(here::here("figures", "mort7_10.png"))
-```
-
-```{r tableModels, echo=FALSE,  message=FALSE}
-library(knitr)
-
-tabletrends <- readr::read_csv(here::here("figures", "mort7_10_trends.csv"))
-kable(tabletrends,  format = "markdown", caption = "Models trends in mortality between and 10cm.")
-```
-
-```{r mortality3 , echo=FALSE, fig.cap="Recruitment ration between 7 and 10cm and the initial recruitment tree number at 7cm.
-Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions", out.width='60%'}
-knitr::include_graphics(here::here("figures", "mort7_10_trends.png"))
-```
-
-```{r mortalityDiv3 , echo=FALSE, fig.cap=" Shannon index mena value across samples per site and in BA difference between 10 and 7 cm ", out.width='100%'}
-knitr::include_graphics(here::here("figures", "H710.png"))
-```
-
-# Model reports
+# Supplementary Material 1: Protocol
+
+## General steps
+
+Each modeling team will run their simulations prior to the workshop, and
+will report on the results before ([short written report](#report)) and
+during the workshop. The general approach for the workshop is:
+
+1.  Each team sets up their model to simulate the regeneration niche of
+    [eleven major European tree species](#speciescomposition) within the
+    known, defined range of environmental conditions of 200 sites. They
+    will run at least one simulation per site. It is important to
+    conduct mixed-species simulations, not series of monospecific
+    stands.
+
+2.  The model is set up for each site based on the [input variable
+    information](#inputvariables) provided in this protocol, and the
+    simulations need to be run until the [equilibrium is
+    reached](#equilibrium).
+
+3.  Each team reports on the results of these exercises in terms of a
+    [standardized output dataset](#outputdata) (including density
+    variables per 10-yr interval and per ha, for each simulated site and
+    species) and in the form of a brief [working report](#report) (\~4
+    pages).
+
+4.  At the workshop in June 2022 in Davos, Switzerland, we will evaluate
+    the regeneration niche as it results from each model and compare it
+    to the empirical estimates of the environmental limits of the
+    regeneration niche that will be made available to the participants
+    shortly before the workshop. As a group effort, the projections of
+    the niche properties arising from the different models will be
+    compared. The exact details on how this will be done will be
+    provided once we know how many modeling teams will contribute
+    results.
+
+## Input variables {#inputvariables}
+
+In this section, we are providing a description of all the necessary
+driver information to run the simulations (e.g., location, climate,
+soil).
+
+### Coordinates
+
+We intend to provide the site information via unique site IDs. If actual
+coordinates are needed, we can provide blurred site coordinates upon
+reasonable request.
+
+### Climate
+
+We provide time series of climatic variables from 1981 to 2018 in
+hourly, daily and monthly resolution. Based on our survey the climate
+data should be sufficient to run all models. Some variables such as
+relative humidity or vapor pressure deficit must be calculated from the
+variables we provide.
+
+Note that the hourly time series has the lowest spatial resolution, thus
+the daily time series should be used for spatial accuracy. If you need
+hourly time series we recommend to use the daily time series and only
+use the hourly time series to correct for hourly variation. For this
+purpose we also provide hourly averages over the period 1981 to 2018 for
+each hour of each day of the year. Only for calculating these averages
+the 29th of February in leap years was removed from the time series to
+obtain values for 365 days.
+
+#### Data sources {#climsources}
+
+The hourly time series are based on the Era5-land data set. Daily and
+monthly time series are mostly based on CHELSA data which has a higher
+spatial resolution. More details are given below or in the documentation
+of the original data source.
+
+The variable names are kept exactly like in the documentation of the
+Era5 and CHELSA data sets to facilitate the use of the original data
+documentation. Please don't be confused by the ending `*_hourly` in
+`surface_thermal_radiation_downwards_hourly` and the two different
+variable names for shortwave radiation: `rsds` for the values from
+CHELSA and `surface_solar_radiation_downwards_hourly` for the values
+from Era5.
+
+All changes that were made to the original data are documented here:
+
+-   The wind variable `wind_speed` is based on the u and v component of
+    the wind and calculated with $\sqrt{u^2+v^2}$ where $u$ is the
+    variable `u_component_of_wind_10m` and $v$ is
+    `v_component_of_wind_10m` of the Era5 data set.
+
+-   The units of temperature variables `temperature_2m` (Era5), `tas`,
+    `tasmin`, `tasmax`(all CHELSA) were changed from Kelvin to Degree
+    Celsius
+
+-   The units of precipitation variables `total_precipitation_hourly`
+    (Era5), `pr` (CHELSA) were changed from m to mm and kG to mm
+    respectively.
+
+-   The unit of short wave radiation `rsds` (CHELSA) was changed from kJ
+    to J.
+
+We aggregated variables from hourly to daily to monthly values by using
+means for temperature, surface pressure and wind speed and sums for
+radiation and precipitation.
+
+##### Era5-Land {#era5}
+
+-   Muñoz Sabater, J., (2019): ERA5-Land hourly data from 1981 to
+    present. Copernicus Climate Change Service (C3S) Climate Data Store
+    (CDS). (23.01.2022),
+    [doi:10.24381/cds.e2161bac](https://doi.org/10.24381/cds.e2161bac)
+-   [Documentation of the hourly ERA5 data from the Google Earth
+    Engine](https://developers.google.com/earth-engine/datasets/catalog/ECMWF_ERA5_LAND_HOURLY#description)
+-   Further information is available from the [official copernicus
+    website](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-land?tab=overview).
+
+##### CHELSA {#chelsa}
+
+-   Dirk N. Karger, Stefan Lange, Chantal Hari, Christopher P. O. Reyer,
+    Niklaus E. Zimmermann (2021): CHELSA-W5E5 v1.0: W5E5 v1.0 downscaled
+    with CHELSA v2.0. ISIMIP Repository.
+    <https://doi.org/10.48364/ISIMIP.836809>
+
+-   [Official CHELSA website](https://chelsa-climate.org)
+
+#### Monthly time series 1981-2018
+
+file name: monthly_era5_chelsa_200.zip
+
+**Data description**
+
+Variables from the [Era5 data set](#era5) with a spatial resolution of
+0.1° x 0.1° (\~ 10 km):
+
+-   average wind speed (`wind_speed`) [m/s]
+
+-   average surface pressure (`surface_pressure`) [Pa]
+
+-   surface thermal (longwave) radiation
+    (`surface_thermal_radiation_downwards_hourly`) [J/m2]
+
+Variables from the [CHELSA data set](#chelsa) with a spatial resolution
+of 0.008° x 0.008° (\~ 1 km):
+
+-   average temperature (`tas`) [°C]
+
+-   average max temperature (`tasmax`) [°C]
+
+-   average min temperature (`tasmin`) [°C]
+
+-   precipitation sum (`pr`) [mm]
+
+-   surface solar (shortwave) radiation (`rsds`) [J/m2]
+
+#### Daily time series 1981-2018
+
+file name: daily_era5_chelsa_200.zip
+
+**Data description**
+
+Variables from the [Era5 data set](#era5) with a spatial resolution of
+0.1° x 0.1° (\~ 10 km):
+
+-   average wind speed (`wind_speed`) [m/s]
+
+-   average surface pressure (`surface_pressure`) [Pa]
+
+-   surface thermal (longwave) radiation
+    (`surface_thermal_radiation_downwards_hourly`) [J/m2]
+
+Variables from the [CHELSA data set](#chelsa) with a spatial resolution
+of 0.008° x 0.008° (\~ 1 km):
+
+-   average temperature (`tas`) [°C]
+
+-   average max temperature (`tasmax`) [°C]
+
+-   average min temperature (`tasmin`) [°C]
+
+-   precipitation sum (`pr`) [mm]
+
+-   surface solar (shortwave) radiation (`rsds`) [J/m2]
+
+#### Hourly time series 1981-2018
+
+file name: hour_era5_200.zip
+
+file name: yhourMean_era5_200.zip
+
+**Data description**
+
+Variables from the [Era5 data set](#era5) with a spatial resolution of
+0.1° x 0.1° (\~ 10 km):
+
+-   average temperature (temperature_2m) [°C]
+
+-   average wind speed (`wind_speed`) [m/s]
+
+-   average surface pressure (`surface_pressure`) [Pa]
+
+-   precipitation sum (total_precipitation_hourly) [mm]
+
+-   surface solar (shortwave) radiation
+    (surface_solar_radiation_downwards_hourly) [J/m2]
+
+-   surface thermal (longwave) radiation
+    (surface_solar_radiation_downwards_hourly) [J/m2]
+
+### Soil
+
+file name: soil_quality_200.csv
+
+**Data description**
+
+The following variables are available:
+
+-   Soil quality is given as a continuous value between 1 and 5.
+
+**Data source**
+
+-   The soil quality value is based on a random forest that was trained
+    to predict expert knowledge based soil quality on a scale from 1 to
+    5 from the [SoilGrids250 data
+    set](https://www.isric.org/explore/soilgrids) and the [WISE data
+    set](https://www.isric.org/explore/wise-databases). The most
+    influential variables in the random forest were total depth to
+    bedrock and sand content. Thus the variable is a good predictor for
+    plant available water storage capacity.
+
+## Topography and elevation
+
+There was an error in a earlier version of this file. Please use the
+link above to download the corrected file.
+
+This is the old file and should not be used because the aspect column
+contains elevation instead of the aspect.
+
+file name: topo_data_200.csv
+
+**Data description**
+
+The following variables are available:
+
+-   elevation [m.a.s.l]
+
+-   slope [Degree]
+
+-   aspect [Degree]
+
+**Data source**
+
+-   Topographic information was derived from the [ASTER DEM
+    V3](https://lpdaac.usgs.gov/products/astgtmv003/).
+
+-   Slope and aspect were calculated with QGIS. The specific tiles that
+    were used are documented
+    [here](https://ites-fe.ethz.ch/rws/downloads/15122021/aster_dem_files.txt),
+    and the processing steps in QGIS are documented
+    [here](https://ites-fe.ethz.ch/rws/downloads/15122021/QGIS_command_log.txt).
+
+### Natural disturbances
+
+The simulations must be run in the absence of natural disturbances.
+
+## Management
+
+The simulations must be run in the absence of management.
+
+### Area simulated
+
+From your model outputs, we would like to obtain samples on species
+specific regeneration rates per decade and per ha. The modeling teams
+decide on the simulated area and how they derive these samples. Please
+describe the methodology used in the [output report](#report).
+
+### Simulation duration and equilibrium {#equilibrium}
+
+The simulation should be run from bare ground in the absence of
+management to a simulated [equilibrium](#defequilibrium) ("potential
+natural vegetation") with current climate. The goal is to derive the
+[regeneration niche](#regenniche). The exact length of the simulation
+should be decided by the modeling teams. Please describe the simulation
+length used in your simulations in the [output report](#report).
+
+[**Regeneration niche:**]{#regenniche} The regeneration niche is defined
+as the passing of a diameter threshold of 7 and 10 cm (ingrowth; cf.
+below). Both the width of the regeneration niche (i.e., in environmental
+space) as well as the intensity of the recruitment process (i.e., the
+number of ingrowth trees per area and per unit of time) are of interest.
+
+[**Simulated equilibrium**]{#defequilibrium}: Simulated equilibrium is
+the "Potential Natural Vegetation".
+
+### Species composition {#speciescomposition}
+
+**Mixed species simulations** must be run (not multiple single-species
+simulations).
+
+**Option A: Species level**
+
+The simulation should be run for **mixtures** of the following
+species/genera: *Fagus sylvatica, Picea abies, Abies alba, Carpinus
+betulus, Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus
+excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris.*
+
+Additional species can be included in the simulation if modelers think
+this is important, but they should report only on these 11
+species/genera because the benchmarking data refer to them only. Please
+report the methodology used in the [output report](#report).
+
+**Option B: PFTs**
+
+If species-level simulation is not possible in your model, substitute
+the species by PFTs and use the model's default approach. Please report
+the methodology used in the [output report](#report). Explain as well
+which PFTs represent which species for your model.
+
+## Diameter threshold
+
+The [output data](#outputdata) have to be provided for two diameter
+thresholds: 7 and 10 cm.
+
+## Output
+
+We are expecting to get the number of recruits over a 10-yr interval per
+species and per ha for each of the sites. However, we are also
+interested in the **variation per site** (i.e., we would like to see the
+range of estimates, not only the mean value per site). Each modeling
+team will decide how to sample multiple values per site that will show
+the variation in recruitment.
+
+Specifically, you should provide us with multiple samples (n = 200; this
+number is the same as the number of sites by sheer coincidence) of the
+regeneration rate for spatial units of 1 ha (please do [not]{.underline}
+provide average of multiple 1-ha samples). This can be done in various
+ways, e.g., by sampling simulated data from the same 1-ha plot in the
+equilibrium over time, or by sampling several 1-ha plots from the
+simulated forest at one specific point in time (in the equilibrium); for
+clarification, cf. three sample strategies proposed in [figures
+1](#sample1), [2](#sample2) and [3](#sample3).
+
+If, for some reason, you cannot submit **200 samples** of **1 ha**,
+please **let us know** in the [workshop chat](#chat) so we can discuss
+alternatives. The sampling from the simulations must happen **within**
+the simulated phase of Potential Natural Vegetation (for more
+information, see how [equilibrium](#defequilibrium), Potential [Natural
+Vegetation, and regeneration niche](#regenniche) are defined in this
+protocol).
+
+![Figure 1. In this example there are 200 samples collected at year i at
+equilibrium differentiated spatially with each square representing 1
+ha.](samplingfigure/Slide2.jpeg){#sample1 width="500"}
+
+![Figure 2. In this example, there are 100 spatially differentiated
+samples taken in year i at equilibrium and 100 in year i+10. Each square
+represents 1 ha.](samplingfigure/Slide3.jpeg){#sample2 width="500"}
+
+![Figure 3. In this example there is 1 spatially differentiated sample
+taken in year i at equilibrium and remeasured every 10 years for 2000
+years until i+2000. The square represents 1
+ha.](samplingfigure/Slide4.jpg){#sample3 width="500"}
+
+### Output data {#outputdata}
+
+Please provide the simulation output data using the following structure:
+
+-   Each row of your output dataset contains the recruitment of 10-year
+    interval per species and ha for a unique simulation site.
+
+-   Each column of your output dataset contains one of these variables:
+
+    -   `sample`: this will identify the different samples per site
+        (1:n, with n $\ge$ 200)
+
+    -   `site`: site identifier (1:200)
+
+    -   `species`: species simulated with the following categories:
+
+        -   fasy (for *Fagus sylvatica*)
+        -   pabi (for *Picea abies*)
+        -   abal (for *Abies alba*)
+        -   cabe (for *Carpinus betulus*)
+        -   tico (for *Tilia cordata*)
+        -   acps (for *Acer pseudoplatanus*)
+        -   betu (for *Betula* spp.)
+        -   frex (for *Fraxinus excelsior*)
+        -   quer (for *Quercus* spp.)
+        -   algl (for *Alnus glutinosa*)
+        -   pisy (for *Pinus sylvestris*)
+        -   If you have some other categories please create your own
+            species codes and provide the legend in the output report.
+            We prefer a lumped category "othr".
+
+    -   `r.trees`: the recruits (trees $ha^{-1}$) per species and
+        decade.
+
+    -   `r.ba`: the recruits' basal area ($m^{2}$ $ha^{-1}$) per species
+        and decade.
+
+    -   `ba`: total basal area ($m^{2}$ $ha^{-1}$) per species and
+        decade
+
+    -   `dbh`: diameter threshold (cm) used to define the recruitment (7
+        or 10 cm).
+
+### Output report {#report}
+
+To guide the writing of the output report, please refer to the following
+questions:
+
+-   What climate data have you used, and how have you created your model
+    climate input parameters?
+
+-   How have you defined the species composition in your model
+    simulation? (e.g., species levels, PFTs, etc.)? Please describe the
+    species categories used.
+
+-   How have you defined the simulation area for your model (e.g., size
+    and other relevant site aspects)?
+
+-   What are the relevant aspects of the dispersal (if any) and seed
+    input parameters in your model?
+
+-   What is the simulation length, and how have you determined the
+    simulation time until the equilibrium?
+
+-   If you have run more than one simulation per site, please describe
+    the differences of the approaches used.
+
+-   Please describe how have you calculated each of the variables in the
+    [output data](#outputdata).
+
+### Uploading data and report
+
+Please upload your data and output report as a **zip file** with the
+name of your model and any other word that might help us to understand
+the files, for example: *thenameofyourmodel_datareport.zip* or
+*thenameofyourmodel_data.zip* and *thenameofyourmodel_report.zip*.
+
+You can upload the files here: LINK
+
+# Supplementary Material 2: Model reports
 
 ## 4C
+### Authors
+
+### Model references
+
+*Please provide here the references to this model where it is possible
+to understand the model structure specially for the regeneration module,
+this could be a published paper, a website with the model description, a
+manual or some other document that the reader can access.*
+
+### Climate
+
+### Soil
+
+### Topography
+
+### Tree species
+
+### Simulations
+
+### Outputs
 
@@
 
-• How have you defined the species composition in your model simulation?
-(e.g., species levels, PFTs, etc.)? Please describe the species
-categories used.
-
 Tree composition is defined at the species level. All individually
@@
 #### Simulation area
-
-• How have you defined the simulation area for your model (e.g., size
-and other relevant site aspects)?
 
@@
 
-• What are the relevant aspects of the dispersal (if any) and seed input
-parameters in your model?
-
 Dispersal parameters are species-specific and include maturity age,
@@
 
-• What is the simulation length, and how have you determined the
-simulation time until the equilibrium?
-
 Simulations were run for 2010 years, and equilibrium was reached at year
@@
 #### Simulation approaches
-
-• If you have run more than one simulation per site, please describe the
-differences of the approaches used.
 
@@
 
-# Protocol
-
-## General steps
-
-Each modeling team will run their simulations prior to the workshop, and
-will report on the results before ([short written report](#report)) and
-during the workshop. The general approach for the workshop is:
-
-1.  Each team sets up their model to simulate the regeneration niche of
-    [eleven major European tree species](#speciescomposition) within the
-    known, defined range of environmental conditions of 200 sites. They
-    will run at least one simulation per site. It is important to
-    conduct mixed-species simulations, not series of monospecific
-    stands.
-
-2.  The model is set up for each site based on the [input variable
-    information](#inputvariables) provided in this protocol, and the
-    simulations need to be run until the [equilibrium is
-    reached](#equilibrium).
-
-3.  Each team reports on the results of these exercises in terms of a
-    [standardized output dataset](#outputdata) (including density
-    variables per 10-yr interval and per ha, for each simulated site and
-    species) and in the form of a brief [working report](#report) (\~4
-    pages).
-
-4.  At the workshop in June 2022 in Davos, Switzerland, we will evaluate
-    the regeneration niche as it results from each model and compare it
-    to the empirical estimates of the environmental limits of the
-    regeneration niche that will be made available to the participants
-    shortly before the workshop. As a group effort, the projections of
-    the niche properties arising from the different models will be
-    compared. The exact details on how this will be done will be
-    provided once we know how many modeling teams will contribute
-    results.
-
-## Input variables {#inputvariables}
-
-In this section, we are providing a description of all the necessary
-driver information to run the simulations (e.g., location, climate,
-soil).
-
-### Coordinates
-
-We intend to provide the site information via unique site IDs. If actual
-coordinates are needed, we can provide blurred site coordinates upon
-reasonable request.
-
-### Climate
-
-We provide time series of climatic variables from 1981 to 2018 in
-hourly, daily and monthly resolution. Based on our survey the climate
-data should be sufficient to run all models. Some variables such as
-relative humidity or vapor pressure deficit must be calculated from the
-variables we provide.
-
-Note that the hourly time series has the lowest spatial resolution, thus
-the daily time series should be used for spatial accuracy. If you need
-hourly time series we recommend to use the daily time series and only
-use the hourly time series to correct for hourly variation. For this
-purpose we also provide hourly averages over the period 1981 to 2018 for
-each hour of each day of the year. Only for calculating these averages
-the 29th of February in leap years was removed from the time series to
-obtain values for 365 days.
-
-#### Data sources {#climsources}
-
-The hourly time series are based on the Era5-land data set. Daily and
-monthly time series are mostly based on CHELSA data which has a higher
-spatial resolution. More details are given below or in the documentation
-of the original data source.
-
-The variable names are kept exactly like in the documentation of the
-Era5 and CHELSA data sets to facilitate the use of the original data
-documentation. Please don't be confused by the ending `*_hourly` in
-`surface_thermal_radiation_downwards_hourly` and the two different
-variable names for shortwave radiation: `rsds` for the values from
-CHELSA and `surface_solar_radiation_downwards_hourly` for the values
-from Era5.
-
-All changes that were made to the original data are documented here:
-
--   The wind variable `wind_speed` is based on the u and v component of
-    the wind and calculated with $\sqrt{u^2+v^2}$ where $u$ is the
-    variable `u_component_of_wind_10m` and $v$ is
-    `v_component_of_wind_10m` of the Era5 data set.
-
--   The units of temperature variables `temperature_2m` (Era5), `tas`,
-    `tasmin`, `tasmax`(all CHELSA) were changed from Kelvin to Degree
-    Celsius
-
--   The units of precipitation variables `total_precipitation_hourly`
-    (Era5), `pr` (CHELSA) were changed from m to mm and kG to mm
-    respectively.
-
--   The unit of short wave radiation `rsds` (CHELSA) was changed from kJ
-    to J.
-
-We aggregated variables from hourly to daily to monthly values by using
-means for temperature, surface pressure and wind speed and sums for
-radiation and precipitation.
-
-##### Era5-Land {#era5}
-
--   Muñoz Sabater, J., (2019): ERA5-Land hourly data from 1981 to
-    present. Copernicus Climate Change Service (C3S) Climate Data Store
-    (CDS). (23.01.2022),
-    [doi:10.24381/cds.e2161bac](https://doi.org/10.24381/cds.e2161bac)
--   [Documentation of the hourly ERA5 data from the Google Earth
-    Engine](https://developers.google.com/earth-engine/datasets/catalog/ECMWF_ERA5_LAND_HOURLY#description)
--   Further information is available from the [official copernicus
-    website](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-land?tab=overview).
-
-##### CHELSA {#chelsa}
-
--   Dirk N. Karger, Stefan Lange, Chantal Hari, Christopher P. O. Reyer,
-    Niklaus E. Zimmermann (2021): CHELSA-W5E5 v1.0: W5E5 v1.0 downscaled
-    with CHELSA v2.0. ISIMIP Repository.
-    <https://doi.org/10.48364/ISIMIP.836809>
-
--   [Official CHELSA website](https://chelsa-climate.org)
-
-#### Monthly time series 1981-2018
-
-file name: monthly_era5_chelsa_200.zip
-
-**Data description**
-
-Variables from the [Era5 data set](#era5) with a spatial resolution of
-0.1° x 0.1° (\~ 10 km):
-
--   average wind speed (`wind_speed`) [m/s]
-
--   average surface pressure (`surface_pressure`) [Pa]
-
--   surface thermal (longwave) radiation
-    (`surface_thermal_radiation_downwards_hourly`) [J/m2]
-
-Variables from the [CHELSA data set](#chelsa) with a spatial resolution
-of 0.008° x 0.008° (\~ 1 km):
-
--   average temperature (`tas`) [°C]
-
--   average max temperature (`tasmax`) [°C]
-
--   average min temperature (`tasmin`) [°C]
-
--   precipitation sum (`pr`) [mm]
-
--   surface solar (shortwave) radiation (`rsds`) [J/m2]
-
-#### Daily time series 1981-2018
-
-file name: daily_era5_chelsa_200.zip
-
-**Data description**
-
-Variables from the [Era5 data set](#era5) with a spatial resolution of
-0.1° x 0.1° (\~ 10 km):
-
--   average wind speed (`wind_speed`) [m/s]
-
--   average surface pressure (`surface_pressure`) [Pa]
-
--   surface thermal (longwave) radiation
-    (`surface_thermal_radiation_downwards_hourly`) [J/m2]
-
-Variables from the [CHELSA data set](#chelsa) with a spatial resolution
-of 0.008° x 0.008° (\~ 1 km):
-
--   average temperature (`tas`) [°C]
-
--   average max temperature (`tasmax`) [°C]
-
--   average min temperature (`tasmin`) [°C]
-
--   precipitation sum (`pr`) [mm]
-
--   surface solar (shortwave) radiation (`rsds`) [J/m2]
-
-#### Hourly time series 1981-2018
-
-file name: hour_era5_200.zip
-
-file name: yhourMean_era5_200.zip
-
-**Data description**
-
-Variables from the [Era5 data set](#era5) with a spatial resolution of
-0.1° x 0.1° (\~ 10 km):
-
--   average temperature (temperature_2m) [°C]
-
--   average wind speed (`wind_speed`) [m/s]
-
--   average surface pressure (`surface_pressure`) [Pa]
-
--   precipitation sum (total_precipitation_hourly) [mm]
-
--   surface solar (shortwave) radiation
-    (surface_solar_radiation_downwards_hourly) [J/m2]
-
--   surface thermal (longwave) radiation
-    (surface_solar_radiation_downwards_hourly) [J/m2]
-
-### Soil
-
-file name: soil_quality_200.csv
-
-**Data description**
-
-The following variables are available:
-
--   Soil quality is given as a continuous value between 1 and 5.
-
-**Data source**
-
--   The soil quality value is based on a random forest that was trained
-    to predict expert knowledge based soil quality on a scale from 1 to
-    5 from the [SoilGrids250 data
-    set](https://www.isric.org/explore/soilgrids) and the [WISE data
-    set](https://www.isric.org/explore/wise-databases). The most
-    influential variables in the random forest were total depth to
-    bedrock and sand content. Thus the variable is a good predictor for
-    plant available water storage capacity.
-
-## Topography and elevation
-
-There was an error in a earlier version of this file. Please use the
-link above to download the corrected file.
-
-This is the old file and should not be used because the aspect column
-contains elevation instead of the aspect.
-
-file name: topo_data_200.csv
-
-**Data description**
-
-The following variables are available:
-
--   elevation [m.a.s.l]
-
--   slope [Degree]
-
--   aspect [Degree]
-
-**Data source**
-
--   Topographic information was derived from the [ASTER DEM
-    V3](https://lpdaac.usgs.gov/products/astgtmv003/).
-
--   Slope and aspect were calculated with QGIS. The specific tiles that
-    were used are documented
-    [here](https://ites-fe.ethz.ch/rws/downloads/15122021/aster_dem_files.txt),
-    and the processing steps in QGIS are documented
-    [here](https://ites-fe.ethz.ch/rws/downloads/15122021/QGIS_command_log.txt).
-
-### Natural disturbances
-
-The simulations must be run in the absence of natural disturbances.
-
-## Management
-
-The simulations must be run in the absence of management.
-
-### Area simulated
-
-From your model outputs, we would like to obtain samples on species
-specific regeneration rates per decade and per ha. The modeling teams
-decide on the simulated area and how they derive these samples. Please
-describe the methodology used in the [output report](#report).
-
-### Simulation duration and equilibrium {#equilibrium}
-
-The simulation should be run from bare ground in the absence of
-management to a simulated [equilibrium](#defequilibrium) ("potential
-natural vegetation") with current climate. The goal is to derive the
-[regeneration niche](#regenniche). The exact length of the simulation
-should be decided by the modeling teams. Please describe the simulation
-length used in your simulations in the [output report](#report).
-
-[**Regeneration niche:**]{#regenniche} The regeneration niche is defined
-as the passing of a diameter threshold of 7 and 10 cm (ingrowth; cf.
-below). Both the width of the regeneration niche (i.e., in environmental
-space) as well as the intensity of the recruitment process (i.e., the
-number of ingrowth trees per area and per unit of time) are of interest.
-
-[**Simulated equilibrium**]{#defequilibrium}: Simulated equilibrium is
-the "Potential Natural Vegetation".
-
-### Species composition {#speciescomposition}
-
-**Mixed species simulations** must be run (not multiple single-species
-simulations).
-
-**Option A: Species level**
-
-The simulation should be run for **mixtures** of the following
-species/genera: *Fagus sylvatica, Picea abies, Abies alba, Carpinus
-betulus, Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus
-excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris.*
-
-Additional species can be included in the simulation if modelers think
-this is important, but they should report only on these 11
-species/genera because the benchmarking data refer to them only. Please
-report the methodology used in the [output report](#report).
-
-**Option B: PFTs**
-
-If species-level simulation is not possible in your model, substitute
-the species by PFTs and use the model's default approach. Please report
-the methodology used in the [output report](#report). Explain as well
-which PFTs represent which species for your model.
-
-## Diameter threshold
-
-The [output data](#outputdata) have to be provided for two diameter
-thresholds: 7 and 10 cm.
-
-## Output
-
-We are expecting to get the number of recruits over a 10-yr interval per
-species and per ha for each of the sites. However, we are also
-interested in the **variation per site** (i.e., we would like to see the
-range of estimates, not only the mean value per site). Each modeling
-team will decide how to sample multiple values per site that will show
-the variation in recruitment.
-
-Specifically, you should provide us with multiple samples (n = 200; this
-number is the same as the number of sites by sheer coincidence) of the
-regeneration rate for spatial units of 1 ha (please do [not]{.underline}
-provide average of multiple 1-ha samples). This can be done in various
-ways, e.g., by sampling simulated data from the same 1-ha plot in the
-equilibrium over time, or by sampling several 1-ha plots from the
-simulated forest at one specific point in time (in the equilibrium); for
-clarification, cf. three sample strategies proposed in [figures
-1](#sample1), [2](#sample2) and [3](#sample3).
-
-If, for some reason, you cannot submit **200 samples** of **1 ha**,
-please **let us know** in the [workshop chat](#chat) so we can discuss
-alternatives. The sampling from the simulations must happen **within**
-the simulated phase of Potential Natural Vegetation (for more
-information, see how [equilibrium](#defequilibrium), Potential [Natural
-Vegetation, and regeneration niche](#regenniche) are defined in this
-protocol).
-
-![Figure 1. In this example there are 200 samples collected at year i at
-equilibrium differentiated spatially with each square representing 1
-ha.](samplingfigure/Slide2.jpeg){#sample1 width="500"}
-
-![Figure 2. In this example, there are 100 spatially differentiated
-samples taken in year i at equilibrium and 100 in year i+10. Each square
-represents 1 ha.](samplingfigure/Slide3.jpeg){#sample2 width="500"}
-
-![Figure 3. In this example there is 1 spatially differentiated sample
-taken in year i at equilibrium and remeasured every 10 years for 2000
-years until i+2000. The square represents 1
-ha.](samplingfigure/Slide4.jpg){#sample3 width="500"}
-
-### Output data {#outputdata}
-
-Please provide the simulation output data using the following structure:
-
--   Each row of your output dataset contains the recruitment of 10-year
-    interval per species and ha for a unique simulation site.
-
--   Each column of your output dataset contains one of these variables:
-
-    -   `sample`: this will identify the different samples per site
-        (1:n, with n $\ge$ 200)
-
-    -   `site`: site identifier (1:200)
-
-    -   `species`: species simulated with the following categories:
-
-        -   fasy (for *Fagus sylvatica*)
-        -   pabi (for *Picea abies*)
-        -   abal (for *Abies alba*)
-        -   cabe (for *Carpinus betulus*)
-        -   tico (for *Tilia cordata*)
-        -   acps (for *Acer pseudoplatanus*)
-        -   betu (for *Betula* spp.)
-        -   frex (for *Fraxinus excelsior*)
-        -   quer (for *Quercus* spp.)
-        -   algl (for *Alnus glutinosa*)
-        -   pisy (for *Pinus sylvestris*)
-        -   If you have some other categories please create your own
-            species codes and provide the legend in the output report.
-            We prefer a lumped category "othr".
-
-    -   `r.trees`: the recruits (trees $ha^{-1}$) per species and
-        decade.
-
-    -   `r.ba`: the recruits' basal area ($m^{2}$ $ha^{-1}$) per species
-        and decade.
-
-    -   `ba`: total basal area ($m^{2}$ $ha^{-1}$) per species and
-        decade
-
-    -   `dbh`: diameter threshold (cm) used to define the recruitment (7
-        or 10 cm).
-
-### Output report {#report}
-
-To guide the writing of the output report, please refer to the following
-questions:
-
--   What climate data have you used, and how have you created your model
-    climate input parameters?
-
--   How have you defined the species composition in your model
-    simulation? (e.g., species levels, PFTs, etc.)? Please describe the
-    species categories used.
-
--   How have you defined the simulation area for your model (e.g., size
-    and other relevant site aspects)?
-
--   What are the relevant aspects of the dispersal (if any) and seed
-    input parameters in your model?
-
--   What is the simulation length, and how have you determined the
-    simulation time until the equilibrium?
-
--   If you have run more than one simulation per site, please describe
-    the differences of the approaches used.
-
--   Please describe how have you calculated each of the variables in the
-    [output data](#outputdata).
-
-### Uploading data and report
-
-Please upload your data and output report as a **zip file** with the
-name of your model and any other word that might help us to understand
-the files, for example: *thenameofyourmodel_datareport.zip* or
-*thenameofyourmodel_data.zip* and *thenameofyourmodel_report.zip*.
-
-You can upload the files here: LINK
+\newpage
+\beginsupplement
+
+#  Supplementary Material 3: Supplementary figures and tables
+
+## Recruitment levels
+
+```{r envTrend7wb, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the sites water balance.
+(A) same y-axis across all models; (B) y-axis scaled by model.
+The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data..", fig.show = 'hold', out.width = '80%'}
+    knitr::include_graphics(c(here::here("figures", "envTrend7WB_500.png"),
+                              here::here("figures", "envTrend7WB.png")))
+```
+
+```{r envTrend7dds, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the site seasonal degree-day sum.
+(A)same y-axis across all models; (B) y-axis scaled by model.
+The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
+    knitr::include_graphics(c(here::here("figures", "envTrend7DDS_500.png"),
+                              here::here("figures", "envTrend7DDS.png")))
+```
+
+```{r zeroInflated, echo=FALSE,  message=FALSE}
+library(knitr)
+
+zeroInflated <- readr::read_csv(here::here("data", "noRecruitment.csv"))
+zeroInflated$Model[zeroInflated$Model == "Empirical"] <- "Observations"
+
+kable(zeroInflated,  format = "latex", caption = "Number of total samples, samples with zero recruitment and the proportion of zero recruitment per model and observed data.")|>
+        kableExtra::landscape() |>
+        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+```
+
+## Model complexity
+
+Summary of the linear model of the overestimation proportions versus the mean complexity value on the establishment module.
+```{r tableComp, echo=FALSE,  message=FALSE, render = 'normal_print'}
+library(knitr)
+
+tableCompOver <- readr::read_csv(here::here("figures", "tableCompOver.csv"))
+jtools::summ(lm(overestimation ~ complexity, data = tableCompOver), model.info = FALSE, digits = 5)
+
+```
+
+## Recruitment species composition
+
+```{r sigDiff, echo=FALSE,  message=FALSE}
+library(knitr)
+
+sigDiff <- readr::read_csv(here::here("figures", "sigH7_10.csv"))
+
+kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of Shannon index in each model for 7 and 10 cm threshold.") |>
+        #kableExtra::landscape() |>
+        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+```
+
+```{r sigH7_Obs, echo=FALSE,  message=FALSE}
+library(knitr)
+
+sigDiff7 <- readr::read_csv(here::here("figures", "sigH7_Obs.csv"))
+
+kable(sigDiff7,  format = "latex", caption = "Test of the significance of the differences of Shannon index between the observed data and each model for 7cm threshold.") |>
+        #kableExtra::landscape() |>
+        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+```
+
+```{r sigH10_Obs, echo=FALSE,  message=FALSE}
+library(knitr)
+
+sigDiff10 <- readr::read_csv(here::here("figures", "sigH10_Obs.csv"))
+
+kable(sigDiff10,  format = "latex", caption = "Test of the significance of the differences of Shannon index between the observed data and each model for 10cm threshold.") |>
+        #kableExtra::landscape() |>
+        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+```
+
+```{r recrStand7, echo=FALSE, fig.cap="Mean Shannon index per site for the recruitment  and at the stand level both for observed and simulated values.
+NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
+knitr::include_graphics(here::here("figures", "recruitmentAdultRichness7.png"))
+```
+
+```{r recrStand10, echo=FALSE, fig.cap="Mean Shannon index per site for the recruitment  and at the stand level both for observed and simulated values.
+NOTE: The recruitment treshold recruitment for treshold 10 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
+knitr::include_graphics(here::here("figures", "recruitmentAdultRichness10.png"))
+```
+
+```{r simObadult , echo=FALSE, fig.cap="Mean species richness at the stand level per site in the simulated versus observed species richness.
+The species richness is estimaed using the Shannon index  calculated by basal area.
+A1) groups models that Simulated species composition in adults is higher than the observed, A2) groups models that Simulated species composition in adults is similar to the observed, A3) groups models where the imulated species composition in adults is lower than the observed.
+NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
+knitr::include_graphics(here::here("figures", "allAgesRichness.png"))
+```
+
+```{r simObr, echo=FALSE, fig.cap="Mean species richness per site in the simulated recruitment versus observed species richness.
+The species richness is estimaed using the Shannon index is calculated by basal area.B1) groups models where the simulated species composition is closer than the other models to observed values, B2) groups models where the sites with lower species composition are underestimated, B3) groups models where the simulated species composition recruitment is lower in the whole range compare to the observe.
+NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
+knitr::include_graphics(here::here("figures", "recruitmentRichness.png"))
+```
+
+```{r diff, echo=FALSE, fig.cap="Departure of each model in the representation of the main species recruitment relative abundance from the observations across the environmental gradients for a 10cm treshold.
+The values shown are the mean of samples per site and across sites in bins, where the range of the climatic water balance (cm) and seasonal degree-day sum (dºC) was divided in 10 bins.
+The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species.
+The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equalt to zero.", out.width='100%'}
+knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_10.png"))
+```
+
+## Mortality
+
+```{r mortTable, echo=FALSE,  message=FALSE}
+# library(knitr)
+# 
+# sigDiff <- readr::read_csv(here::here("figures", "mortRatio7_10.csv"))
+# 
+# kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold over 10 cm threshold.") |>
+# #kableExtra::landscape() 
+# kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+```
+
+```{r mortality12 , echo=FALSE, fig.cap="Interquartile range of the ratio between 7 and 10cm recruitment and the overestimation proportion (sim-obs/obs) for the mean recruitment trees per model across sites and samples", out.width='100%'}
+knitr::include_graphics(here::here("figures", "mort7_10_iqrOver.png"))
+```
+
+```{r mortality22 , echo=FALSE, fig.cap="Recruitment ration between 7 and 10cm and the initial recruitment tree number at 7cm.
+Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions.
+The cloud point per models shows the differences across sites, the shape point cloud per model shows if the models take into account regeneration density when applying mortality to regeneration ", out.width='60%'}
+knitr::include_graphics(here::here("figures", "mort7_10.png"))
+```
+
+```{r tableModels, echo=FALSE,  message=FALSE}
+library(knitr)
+
+tabletrends <- readr::read_csv(here::here("figures", "mort7_10_trends.csv"))
+kable(tabletrends,  format = "markdown", caption = "Models trends in mortality between and 10cm.")
+```
+
+```{r mortality3 , echo=FALSE, fig.cap="Recruitment ration between 7 and 10cm and the initial recruitment tree number at 7cm.
+Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions", out.width='60%'}
+knitr::include_graphics(here::here("figures", "mort7_10_trends.png"))
+```
+
+```{r mortalityDiv3 , echo=FALSE, fig.cap=" Shannon index mena value across samples per site and in BA difference between 10 and 7 cm ", out.width='100%'}
+knitr::include_graphics(here::here("figures", "H710.png"))
+```

### 2022-12-15 code/figures.R
correct name of figure envTrend7Totba

@@ -239,7 +239,7 @@
     
 
     #### Fig. R VS BA----
-  png(file = "figures/envTrend7Totba_500.png", width = 16, height = 12, 
+  png(file = "figures/envTrend7Totba.png", width = 16, height = 12, 
         units = "cm", res = 300)
   par(mfrow = c(4, 4), mar = c(2, 3, 1, 1), oma = c(2, 2, 1, 1))
     

### 2022-12-15 manuscript/manuscriptRW.Rmd
corrections in abstract and introduction

@@
 
-Using 15 models and model variants buil to capture longterm (i.e.,
+Using 15 models and model variants built to capture long-term (i.e.,
 decadal to centennial) forest dynamics at the stand, landscape and
@@
 Even though the overestimation of tree establishment represent a
-challenge for the usefulness of model~~s~~ projections under uncertain
+challenge for the usefulness of model projections under uncertain
 conditions such a climate change, and although some formulations should
 be reconsidered to correct excessive ingrowth levels, it is remarkable
-that most of the~~se~~ models capture the essential features of the
-highly complex process of tree establishment, while not having been
+that most of the models capture the essential features of the highly
+complex process of tree establishment, while not having been
 parameterized with such a data. We conclude that in spite of all the
-uncertainties, many models hold promise for assessing longterm forest
+uncertainties, many models hold promise for assessing long-term forest
 dynamics under changing environmental conditions.
@@
 species richness due to positive feedback effects such that eventually
-just single-species ~~forest~~ stands remain. This is sometimes
-corrected by the incorporation of a low level of seed influx of all
-species at all times [@lischke2006]. However, simulated species
-composition is usually exceedingly sensitive to that level, whereas the
-parameters of such functions are poorly constrained by field data.
-Lastly, there is often a problem with insufficient observational
-constraint on parameter values for models that start from very small
-tree sizes (e.g., 10 cm of height or even seeds) and track tree
-development in a process-oriented manner by considering a multitude of
-ecological influences, rather than emphasizing tree establishment into a
-larger size class [e.g., @wehrli2006].
+just single-species stands remain. This is sometimes corrected by the
+incorporation of a low level of seed influx of all species at all times
+[@lischke2006]. However, simulated species composition is usually
+exceedingly sensitive to that level, whereas the parameters of such
+functions are poorly constrained by field data. Lastly, there is often a
+problem with insufficient observational constraint on parameter values
+for models that start from very small tree sizes (e.g., 10 cm of height
+or even seeds) and track tree development in a process-oriented manner
+by considering a multitude of ecological influences, rather than
+emphasizing tree establishment into a larger size class [e.g.,
+@wehrli2006].
 
@@
 Inventories) often measure tree establishment [@zell2019], however with
-a design that captures the rate after the stand initiation
-phase[@hallsby2015]. Data from permanent plots, although available
-across regions, are highly heterogeneous, rendering their utility
-difficult in a modeling context[@käber2021]. On the other hand, targeted
-studies to measure tree establishment on experimental sites are often
-limited in spatial extent, species studied, or the sub-set of processes
-that are targeted [@berdanier2016; @collet2006].
+a design that captures the rate after the stand initiation phase
+[@hallsby2015]. Data from permanent plots, although available across
+regions, are highly heterogeneous, rendering their utility difficult in
+a modeling context[@käber2021]. On the other hand, targeted studies to
+measure tree establishment on experimental sites are often limited in
+spatial extent, species studied, or the sub-set of processes that are
+targeted [@berdanier2016; @collet2006].
 
@@
 establishment patterns in nature and the large number of factors driving
-this process --- including some that are not included in most models
+this process --- including some that are not incorporated in most models
 such as deer browsing --- we do not aim at a detailed statistical

### 2022-12-15 code/figures.R
clean of figures code

@@ -15,17 +15,16 @@
     outputsDF <- data.table::fread("data/dataOutputs.csv")
     outputsDF$model[outputsDF$model == "Empirical"] <- "Observed"
 
-    
     # Number of trees in the observed data 
-    obsNum <- outputsDF [outputsDF$model == "Observed",]
+    obsNum <- outputsDF[outputsDF$model == "Observed", ]
     round( sum(obsNum$r.trees, na.rm = TRUE), 0)
     
     obsNumAgg <- obsNum |> 
                    dplyr::group_by(site, sample) |>
-                   dplyr::summarise( r.ha= sum(r.trees))
+                   dplyr::summarise(r.ha = sum(r.trees))
     
-    #hist(obsNumAgg$r.ha)
-    range (obsNumAgg$r.ha, na.rm = TRUE)
+    sum(obsNumAgg$r.ha, na.rm =  T)
+    range(obsNumAgg$r.ha, na.rm = TRUE)
     
     # Assign models and dbh to levels 
 
@@ -48,19 +47,19 @@
     
     # Sum across species in each observation
     ShannonIndex <- data.frame(cbind(aggregate(mult ~ site + sample + model + dbh,
-                                               sum, data =  ShannonDF)))
+                                               sum, data = ShannonDF)))
     
     ShannonIndex$mult <- (-1) * (ShannonIndex$mult)
-    colnames(ShannonIndex) <- c("site", "sample", "model",
-                                "dbh", "ShannonIndexRecruit")
+    colnames(ShannonIndex) <- c("site", "sample", "model", "dbh",
+                                "ShannonIndexRecruit")
     
     ## Stand Shannon index ----
     ShannonDFAll <- outputsDF 
     ShannonDFAll$giDIVG <- ShannonDFAll$ba / ShannonDFAll$Totba
     ShannonDFAll$lngiDIVG <- log(ShannonDFAll$ba / ShannonDFAll$Totba)
-    ShannonDFAll$mult <- round(ShannonDFAll$giDIVG *  ShannonDFAll$lngiDIVG, 2)
+    ShannonDFAll$mult <- round(ShannonDFAll$giDIVG * ShannonDFAll$lngiDIVG, 2)
     
-    # Sum across secies per site and sample
+    # Sum across species per site and sample
     ShannonIndexAll <- data.frame(cbind(aggregate(mult ~ site + sample + model + dbh,
[…]

### 2022-12-15 manuscript/supplementaryRW.Rmd
figure caption update in manuscript

@@
 
-```{r envTrend7wb, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the sites water balance.
+```{r envTrend7wb, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site, for the 200 sites against the sites water balance.
 (A) same y-axis across all models; (B) y-axis scaled by model.
-The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data..", fig.show = 'hold', out.width = '80%'}
+The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
     knitr::include_graphics(c(here::here("figures", "envTrend7WB_500.png"),
@@
 
-```{r envTrend7dds, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the site seasonal degree-day sum.
-(A)same y-axis across all models; (B) y-axis scaled by model.
-The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
+```{r envTrend7dds, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site, for the 200 sites against the sites seasonal degree-day sum.
+(A) same y-axis across all models; (B) y-axis scaled by model.
+The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
     knitr::include_graphics(c(here::here("figures", "envTrend7DDS_500.png"),
@@
 ```{r mortTable, echo=FALSE,  message=FALSE}
-# library(knitr)
-# 
-# sigDiff <- readr::read_csv(here::here("figures", "mortRatio7_10.csv"))
-# 
-# kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold over 10 cm threshold.") |>
-# #kableExtra::landscape() 
-# kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+library(knitr)
+
+sigDiff <- readr::read_csv(here::here("figures", "mortRatio7_10.csv"))
+
+kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold over 10 cm threshold.") |>
+#kableExtra::landscape()
+kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
 ```
 
-```{r mortality12 , echo=FALSE, fig.cap="Interquartile range of the ratio between 7 and 10cm recruitment and the overestimation proportion (sim-obs/obs) for the mean recruitment trees per model across sites and samples", out.width='100%'}
+```{r mortality12 , echo=FALSE, fig.cap="Interquartile range of the ratio between 7 and 10cm recruitment and the overestimation proportion (sim-obs/obs) for the mean recruitment trees per model across sites and samples", out.width='70%'}
 knitr::include_graphics(here::here("figures", "mort7_10_iqrOver.png"))
@@
 Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions.
-The cloud point per models shows the differences across sites, the shape point cloud per model shows if the models take into account regeneration density when applying mortality to regeneration ", out.width='60%'}
+The cloud point per models shows the differences across sites, the shape point cloud per model shows if the models take into account regeneration density when applying mortality to regeneration ", out.width='70%'}
 knitr::include_graphics(here::here("figures", "mort7_10.png"))
@@
 ```{r mortality3 , echo=FALSE, fig.cap="Recruitment ration between 7 and 10cm and the initial recruitment tree number at 7cm.
-Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions", out.width='60%'}
+Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions", out.width='70%'}
 knitr::include_graphics(here::here("figures", "mort7_10_trends.png"))
@@
 
-```{r mortalityDiv3 , echo=FALSE, fig.cap=" Shannon index mena value across samples per site and in BA difference between 10 and 7 cm ", out.width='100%'}
+```{r mortalityDiv3 , echo=FALSE, fig.cap=" Shannon index mean value across samples per site and in BA difference between 10 and 7 cm ", out.width='70%'}
 knitr::include_graphics(here::here("figures", "H710.png"))

### 2022-12-15 manuscript/manuscriptRW.Rmd
couple of changes in the discussion section

@@
 link-citations: yes
+nocite: |
+  @reyer, @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @seidl2012a, @lischke2006a, @schumacher2006, @scheiter2009, @scheller2004, @morin2021
 linkcolor: blue
@@
 (SYBILA and xComp). Some of the models largely rely on the original
-approach underlying forest gap models (e.g., ForClim 1 and
-LandClim)[,]{.underline} whereas some [a]{.underline}re based on plant
-ecophysiological processes (e.g., FORMIND and iLand). The different
-approaches underlying the models have strongly influenced the
-formulation of tree establishment.
+approach underlying forest gap models (e.g., ForClim 1 and LandClim),
+whereas some are based on plant ecophysiological processes (e.g.,
+FORMIND and iLand). The different approaches underlying the models have
+strongly influenced the formulation of tree establishment.
 
@@
 flowering and pollination, seed production, seed dispersal, germination
-seedling [g]{.underline}rowth, which ultimately lead to the simulated
-number of trees. Recruitment models, in contrast, introduce a certain
-number of new trees with certain characteristics such as biomass or
-diameter, without explicitly considering earlier development processes.
-We can further distinguish models that feature a feedback in the
-establishment, i.e., where the level and species composition of
-establishment is influenced by the existing adult tree community via the
-production of seeds, seedlings or saplings, from models that do not
-contain such feedback.
+seedling growth, which ultimately lead to the simulated number of trees.
+Recruitment models, in contrast, introduce a certain number of new trees
+with certain characteristics such as biomass or diameter, without
+explicitly considering earlier development processes. We can further
+distinguish models that feature a feedback in the establishment, i.e.,
+where the level and species composition of establishment is influenced
+by the existing adult tree community via the production of seeds,
+seedlings or saplings, from models that do not contain such feedback.
 
@@
 seedbank, seed or seedling to sapling (i.e., trees much larger than 10
-cm height, often at least 2-3 m tall).This implies that the larger the
+cm height, often at least 2-3 m tall). This implies that the larger the
 size threshold, the larger the number of ecological processes that must
@@
 complexity across all processes \>0, Table \@ref(tab:tableModels)). In
-most models[,]{.underline} the two processes in the establishment
-modules with the highest complexity are the statistical approach used to
-define the successful number of new treesand how frost occurrence is
-taken into account.
+most models, the two processes in the establishment modules with the
+highest complexity are the statistical approach used to define the
+successful number of new trees and how frost occurrence is taken into
+account.
 
@@
 isolate the importance of the canopy feedback via seed trees for the
-quantity and quality of simulated ingrowth[,]{.underline} as one variant
-includes this feedback whereas the other does not.
+quantity and quality of simulated ingrowth, as one variant includes this
+feedback whereas the other does not.
 
@@
       kableExtra::column_spec(column = 11, width = "5em") |>
-      kableExtra::column_spec(column = 12, width = "5em") |>
+      kableExtra::column_spec(column = 12, width = "8em") |>
       kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
@@
 environmental gradients in temperature and precipitation in central
-Europe as compiled in the framework of the EuFoRIa network[@euforia2019]
-(Figure \@ref(fig:map)). All these forests have not been managed for at
-least 10 years prior to their designation as reserves, and most of them
-provide time series of natural forest dynamics over multiple decades and
-up to ca. 80 years. The census periods range from 3 to 37 years, with an
-average of 14 years. The data provide information at the tree level,
-thus allowing for sequential comparison such as individual-tree ingrowth
-and death between the measurements.
+Europe as compiled in the framework of the EuFoRIa network
+[@euforia2019] (Figure \@ref(fig:map)). All these forests have not been
+managed for at least 10 years prior to their designation as reserves,
+and most of them provide time series of natural forest dynamics over
+multiple decades and up to ca. 80 years. The census periods range from 3
+to 37 years, with an average of 14 years. The data provide information
+at the tree level, thus allowing for sequential comparison such as
+individual-tree ingrowth and death between the measurements.
 
@@
 wide range of models of forest dynamics. We define the regeneration
-niche as the passing of a diameter threshold of 7 or 10
-cm[,]{.underline} respectively (synonym: ingrowth). To this end, each
-modeling group was provided with a detailed protocol (cf. Supplementary
-Material 1, henceforth "SM") with instructions how to perform the
-simulations, providing input variables on climate and soil conditions,
-and the expected output variables. Neither were further site information
-(except for the data specified below) nor any data on tree ingrowth
-provided prior to the simulation. That is, the models were run in "blind
-flight" mode.
+niche as the passing of a diameter threshold of 7 or 10 cm, respectively
+(synonym: ingrowth). To this end, each modeling group was provided with
+a detailed protocol (cf. Supplementary Material 1, henceforth "SM") with
+instructions how to perform the simulations, providing input variables
+on climate and soil conditions, and the expected output variables.
+Neither were further site information (except for the data specified
+below) nor any data on tree ingrowth provided prior to the simulation.
+That is, the models were run in "blind flight" mode.
 
@@
 [@asterscienceteam2019], but no other spatial information such as
-coordinates was provided, with the exception of iLand[,]{.underline}
-which used the blurred coordinates to derive highly detailed soil data.
+coordinates was provided, with the exception of iLand, which used the
+blurred coordinates to derive highly detailed soil data.
 
 The simulations were run in the absence of natural disturbances. Only
-the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE [ONE
-SENTENCE]{.underline} OF WHY YOU INCLUDED DISTURBANCES AND WHAT TYPE.
-The simulations were set up to sample species-specific ingrowth rates
-per decade and per ha in the equilibrium state of the model, typically
-entailing a "spin-up" run (as we did not provide any forest data). The
-modeling teams decided on the simulated area and how they derived these
-samples. The exact length of the simulation was also decided by the
-modeling teams (Table \@ref(tab:tableModels)). Further details on how
-each modeling team prepared the simulations and the outputs are
-available in SM 2.
+the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE ONE SENTENCE
+OF WHY YOU INCLUDED DISTURBANCES AND WHAT TYPE. The simulations were set
+up to sample species-specific ingrowth rates per decade and per ha in
+the equilibrium state of the model, typically entailing a "spin-up" run
+(as we did not provide any forest data). The modeling teams decided on
+the simulated area and how they derived these samples. The exact length
+of the simulation was also decided by the modeling teams (Table
+\@ref(tab:tableModels)). Further details on how each modeling team
+prepared the simulations and the outputs are available in SM 2.
 
@@
 xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). HERE **4C** PROVIDES
-[ONE SENTENCE]{.underline} WHY THEY SIMULATED LESS SPECIES. HERE
-**XCOMP** PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED LESS
-SPECIES. HERE **LPJ-GUESS** PROVIDES [ONE SENTENCE]{.underline} WHY THEY
-SIMULATED LESS SPECIES. Finally, the model aDGVM2 is based on a suite of
-broad Plant Functional Types rather than individual species, and hence
-this model used its default approach of PFTs for the simulations.
+ONE SENTENCE WHY THEY SIMULATED LESS SPECIES. HERE **XCOMP** PROVIDES
+ONE SENTENCE WHY THEY SIMULATED LESS SPECIES. HERE **LPJ-GUESS**
+PROVIDES ONE SENTENCE WHY THEY SIMULATED LESS SPECIES. Finally, the
+model aDGVM2 is based on a suite of broad Plant Functional Types rather
+than individual species, and hence this model used its default approach
+of PFTs for the simulations.
 
@@
 for all the tree species simulated. LPJ-GUESS could not simulate one of
-the sites HERE **LPJ-GUESS** PROVIDES [ONE SENTENCE]{.underline} WHY
-THEY DID NOT PROVIDE ALL THE SITES. Landis-II missed 0.2% of the
-observations HERE **LANDIS-II** PROVIDES [ONE SENTENCE]{.underline} WHY
-THEY DID NOT PROVIDE ALL THE SAMPLES/observations.
+the sites HERE **LPJ-GUESS** PROVIDES ONE SENTENCE WHY THEY DID NOT
+PROVIDE ALL THE SITES. Landis-II missed 0.2% of the observations HERE
+**LANDIS-II** PROVIDES ONE SENTENCE WHY THEY DID NOT PROVIDE ALL THE
+SAMPLES/observations.
 
@@
 establishment module (Table \@ref(tab:tableModels)) and the
-overestimation proportion at a diameter threshold of 7 cm (Table S2).
+overestimation proportion at a diameter threshold of 7 cm (SM3).
 
@@
 overpredicted species diversity at the stand level (Figure
-\@ref(fig:combinedRichness)A, center; cf. Figure S5)[,]{.underline} but
-a characteristic feature was that simulations had a much lower
-variability of diversity compared to observations. Models overpredicting
-species richness included ForClim1, ForClim 11, FORMIND and TreeMig
-(Figure \@ref(fig:combinedRichness)A, left). Several models
-underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C
-(Figure \@ref(fig:combinedRichness)A, right). The majority of the models
+\@ref(fig:combinedRichness)A, center; cf. Figure S5), but a
+characteristic feature was that simulations had a much lower variability
+of diversity compared to observations. Models overpredicting species
+richness included ForClim1, ForClim 11, FORMIND and TreeMig (Figure
+\@ref(fig:combinedRichness)A, left). Several models underpredicted
+species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure
+\@ref(fig:combinedRichness)A, right). The majority of the models
 captured reasonably well the ingrowth diversity at lower values of
@@
 almost no ingrowth of *Pinus sylvestris* and *Quercus* spp., or did not
-feature any ingrowth of of these species at all.
+feature any ingrowth of these species at all.
 
@@
 The species diversity of the entire stand was better captured by the
-models than the diversity of tree ingrowth, and only a few models
-overpredicted stand level diversity. Ingrowth composition defined as
-being proportional to adult abundance, regardless of productivity, might
-be a simple and cautious assumption to safely maintain relative species
-abundances [@hanbury-brown2022]. However, based on the simulation
-results, we did not see that models with feedback captured better the
-ingrowth species diversity than those without feedback. The similar
-performance between models with and without feedback might be because
-the models are giving more importance to define the regeneration niche
-via abiotic and biotic filters than with the habitat niche of the adult
-trees [@grubb1977].
+models than the diversity of
+tree ingrowth, and only a few models overpredicted stand-level
+diversity. Defining ingrowth composition as being proportional to adult
+abundance, regardless of productivity, might be a
+simple and conservative assumption to safely maintain relative species
+abundances [@hanbury-brown2022], although this may be simplistic.
+Furthermore, based on the simulation results there is no evidence that
+models with feedback captured better the ingrowth species diversity than
+those without feedback. The similar performance of models with and
+without feedback is likely because the models are giving more weight to
+the regeneration niche arising abiotic and biotic filters than from the
+habitat niche of the adult trees [@grubb1977].
 
@@
 available in a harmonized manner across large environmental gradients.
-Third, the empirical data were collected from rather small plots that
-ranged from 0.2 to up to 1 ha, while we sampled simulated ingrowth rates
-from 1 ha areas, which may lead to an incorrect representation of space.
-Even though this strategy it is not ideal, it represents a common
-challenge when harmonizing diverse data sources originating from varied
-sampling strategies [@portier2022]. It would have been extremely
-challenging for such a variety of models to follow a protocol where the
-spatial sampling size was different at each of the 200 sites.
+Third, the empirical data were collected from rather small plots, while
+we sampled simulated ingrowth rates from 1 ha areas, which may lead to
+an incorrect representation of space. Even though this strategy it is
+not ideal, it represents a common challenge when harmonizing diverse
+data sources originating from varied sampling strategies [@portier2022].
+It would have been extremely challenging for such a variety of models to
+follow a protocol where the spatial sampling size was different at each
+of the 200 sites.
 
@@
 values. The lack of spatial considerations potentially puts landscape
-models at a disadvantage, as they have been built to be accurate~~l~~ at
-the landscape level. Without spatial context, we are limiting ingrowth
-to the seed influx from the stand itself, if the model does not have a
+models at a disadvantage, as they have been built to be accurate at the
+landscape level. Without spatial context, we are limiting ingrowth to
+the seed influx from the stand itself, if the model does not have a
 background seed input. However, Global Dynamic Vegetation Models should
@@
 establishment niche is not always captured accurately as a function of
-the biotic (light) and abiotic (temperature and moisture). However, most
-models properly capture the diversity of the initial tree populations,
-although changes in their formulations including feedback from adult
-trees did not have an effect on better capturing the species composition
-of ingrowth.
+the biotic (light) and abiotic (temperature and moisture).
+
+However, most models properly capture the diversity of the initial tree
+community, but variations in model formulations, e.g., the presence or
+absence of feedback from the adult trees, did not have a strong effect
+for capturing the species composition of ingrowth.
 

### 2022-12-15 manuscript/manuscriptRW.Rmd
references updates

@@
 nocite: |
-  @reyer, @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @seidl2012a, @lischke2006a, @schumacher2006, @scheiter2009, @scheller2004, @morin2021
+  @reyer, @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2004, @morin2021
 linkcolor: blue
@@
 ecology, tree establishment has not received a lot of attention
-[@hanburybrown_forest_2022], although many sub-processes have been
-studied in detail. Yet, little data is available that covers all the
-processes within one species along environmental gradients, let alone
-for a vast suite of species. Thus, the available data are fragmented,
-which constitutes a major problem for model building. On the one hand,
+[@hanbury-brown2022], although many sub-processes have been studied in
+detail. Yet, little data is available that covers all the processes
+within one species along environmental gradients, let alone for a vast
+suite of species. Thus, the available data are fragmented, which
+constitutes a major problem for model building. On the one hand,
 monitoring on permanent sample plots (such as National Forest
@@
 regions, are highly heterogeneous, rendering their utility difficult in
-a modeling context[@käber2021]. On the other hand, targeted studies to
+a modeling context [@käber2021]. On the other hand, targeted studies to
 measure tree establishment on experimental sites are often limited in
@@
 between cells as they are based on an abstraction of horizontal space,
-or even the trees themselves [@hanburybrown_forest_2022].
+or even the trees themselves [@hanbury-brown2022].
 

### 2022-12-15 code/dataPreparation.R
remove data checks and old notes in dataPreparation

@@ -4,21 +4,6 @@
 # Authors: Olalla Díaz-Yáñez and Yannek Käber 
 ################################################################################ 
 
-# NOTES
-# 0 values in recruitment (r.trees and r.ba) indicate that the site was simulated/observed 
-# for that diameter threshold but there was not recruitment for that species.
-# This also include species were the total basal area in the site was also 0.
-# 
-# NA values in recruitment (r.trees and r.ba) indicate that this site was not observed 
-# this is the case for 35 sites for diameter threshold of 7cm in the observations 
-# data set
-# 
-# Models where not all the species were simulated (4C, LPJ-GUESS and xComp) 
-# do not have values for those species in the dataset
-
-
-
-
 # Load observations -----
 
     observationsData <- data.table::fread("data/observations/observations.csv")
@@ -630,22 +615,6 @@
     colnames(BATot)[5] <- "Totr.ba"
     dataOutputs <- merge(dataOutputs, BATot, by = c("site", "sample", "model"))
 
-    
-# Data checks -----
-   # number of observation for 200 samples per 200 sites for 11 species
-   # (200*200)*11 = 440000 
-    table(dataOutputs$model, dataOutputs$dbh ) # OK LandClim, iLand
-   # models with 12 species 
-   #  (200*200)*12 species = 480000
-   #  (200*200)*10 species  = 400000
-   #  (200*200)*9 species = 360000
-    
-   #  Species per site and model and sample
-    
-    table(dataOutputs$model, dataOutputs$dbh, dataOutputs$site, dataOutputs$sample ) 
-    
-    View (dataOutputs[dataOutputs$model == "ForClim 1",])
-    table(dataOutputs$model)
     
 # Save the data -----
 

### 2022-12-18 code/figures.R
correct figure packaged was not called

@@ -1637,9 +1637,10 @@
     write.csv(tableCompOver,"figures/tableCompOver.csv", row.names = FALSE)
     
     
-    ggplot(tableCompOver, aes(x = complexity, y = overestimation )) +
-      geom_bar(stat = "identity") +
-      stat_smooth(method = "lm", se = TRUE)
+    ggplot2::ggplot(tableCompOver, ggplot2::aes(x = complexity, 
+                                                y = overestimation )) +
+      ggplot2::geom_bar(stat = "identity") +
+      ggplot2::stat_smooth(method = "lm", se = TRUE)
     
     summary(lm(overestimation ~ complexity, data = tableCompOver))
     sjPlot::tab_model(lm(overestimation ~ complexity, data = tableCompOver))

### 2022-12-18 manuscript/supplementaryRW.Rmd
change csl in supplementary

@@
 linkcolor: blue
-csl: draft.csl
+csl: global-change-biology.csl
 editor_options: 

### 2022-12-19 code/dataPreparation.R
commented map figure in the code

@@ -10,7 +10,7 @@
 
 # Load model simulations ----
     
-    # Species asked to be simulated in the procotol
+    # Species asked to be simulated in the protocol
     # Assign other species to other category
     simSppAsked <- c("fasy", "Fagus sylvatica",
                      "pabi", "Picea abies", "piab",
@@ -133,7 +133,8 @@
     # It has 201 sample per site, remove one sample
 
     picus <- picus[!picus$sample > 200,]
-    
+ 
+       
  ## Landscape ----
 
     ### iLand ----

### 2022-12-19 code/figures.R

@@ -1681,48 +1681,50 @@
       dplyr::filter(dplyr::row_number() == 1)
     
     
-    # DO NOT RUN
+    ###### DO NOT RUN #########
   
     # Add location 
-    location <- data.table::fread("data/coords_blurred_dt4326.csv")
+    # location <- data.table::fread("data/coords_blurred_dt4326.csv")
+    # 
+    # domSpp7 <- merge(standSpp, recruitSpp7, by = "site")
+    # domSpp7 <- merge(domSpp7, location, by.x = "site", by.y = "cluster_plot_id200")
+    # 
+    # domSpp10 <- merge(standSpp, recruitSpp10, by = "site")
+    # domSpp10 <- merge(domSpp10, location, by.x = "site", by.y = "cluster_plot_id200")
+    # 
+    # 
+    # library(tidyverse)
+    # library(sf)
+    # library(rnaturalearth)
+    # library(rnaturalearthdata)
+    # library(rgeos)
+    # library(rgdal)
+    # library(raster)
+    # world <- ne_countries(scale = "medium", returnclass = "sf")
+    # Europe <- world[which(world$continent == "Europe"), ]
+    # plotsMap <- ggplot(Europe) +
+    #   ggthemes::theme_map() +
+    #   geom_sf(fill = "transparent" , colour = "lightgrey",  lwd = 0.25) +
+    #   coord_sf(xlim = c(-25,45), ylim = c(40,55), expand = FALSE) +
+    #   geom_point(data = domSpp10, shape = 20, stroke = FALSE,
+    #              mapping = aes(x = X, y = Y, color = recruitment),
+    #              alpha = 1, size = 2) +
+    #   scale_color_manual("Dominant species of ingrowth",
+    #                      values = c("Abies alba" =  "#1B9E77",
+    #                                 "Picea abies" ="#D95F02" ,
+    #                                 "Fagus sylvatica" = "#7570B3",
+    #                                 "Pinus sylvestris" = "#E7298A" ,
+    #                                 "Quercus spp." = "#66A61E",
+    #                                 "Acer pseudoplatanus" =  "#E6AB02",
+    #                                 "Carpinus betulus" = "#A6761D",
+    #                                 "Fraxinus excelsior" = "#666666",
+    #                                 "Tilia cordata" = "#D8B365",
+    #                                 "Betula spp." = "#F5F5F5",
+    #                                 "Other spp." = "grey" ) )
+    # 
+    # ggplot2::ggsave("figures/theMap.png",
[…]

### 2023-01-15 manuscript/supplementaryRW.Rmd
Add modelling protocol for Formind

@@
 
+## FORMIND
+
+### Authors
+
+Samuel M. Fischer, Friedrich Bohn, Rico Fischer
+
+### Model references
+
+<!--
+@article{fischer_lessons_2016,
+	title = {Lessons learned from applying a forest gap model to understand ecosystem and carbon dynamics of complex tropical forests},
+	volume = {326},
+	issn = {03043800},
+	url = {https://linkinghub.elsevier.com/retrieve/pii/S0304380015005505},
+	doi = {10.1016/j.ecolmodel.2015.11.018},
+	language = {en},
+	urldate = {2022-09-15},
+	journal = {Ecological Modelling},
+	author = {Fischer, Rico and Bohn, Friedrich and Dantas de Paula, Mateus and Dislich, Claudia and Groeneveld, J{\"u}rgen and Guti{\'e}rrez, Alvaro G.
+and Kazmierczak, Martin and Knapp, Nikolai and Lehmann, Sebastian and Paulick, Sebastian and P{\"u}tz, Sandro and R{\"o}dig, Edna and Taubert, Franziska and K{\"o}hler, Peter and Huth, Andreas},
+	month = apr,
+	year = {2016},
+	pages = {124--133},
+	file = {Volltext:C\:\\Users\\fischsam\\Zotero\\storage\\EISNTZF6\\Fischer et al. - 2016 - Lessons learned from applying a forest gap model t.pdf:application/pdf},
+}
+-->
+
+A full description of the model can be found in @fischer_lessons_2016. 
+The basis for the parameterization we applied is described in @bohn2014.
+
+### Climate
+
+To incorporate climate, we used the data set `daily_era5_chelsa_200.csv',
+which contains climate data on a daily resolution. We considered the
+covariates precipitation, temperature, and irradiance, the latter
+of which we converted to units used in Formind. Adding on to these
+data, we used a constant and site-independent potential evapotranspiration
+(PET) value of $4\frac{\text{mm}}{\text{d}}$ and assumed that the mean day
+length was $12\text{h}$ at all sites.
+
+Since we simulated longer time periods than the $38$ years for which
+we had climate data, we used the provided data in a randomized fashion.
+For each simulation year, we applied the climate data of a randomly
+selected year. We neglected the existence of leap years, which led
+to a minor time shift in climate data of later years.
+
+### Soil
+
+We modelled soil conditions by using the provided soil quality values
+as proxy for soil depth. To that end, we scaled the provided values
+to the typical range of soil depth values used in Formind for temperate
+forests. Specifically, we multiplied the soil quality values by factor 
+$0.5$ to obtain the modelled soil depth in meters.
+
+### Topography
+
+We did not use any topograpy data.
+
+### Tree species
+
+We considered each of the $11$ species in the protocol explicitly.
+We chose the model parameters for the individual species based on
+the work by @bohn2014. For species that were not
+covered by @bohn2014, we used parameter values of
+similar species with known parameters. The species composition was
+not predefined and evolved solely from the seed input
+and the model dynamics.
+
+### Simulations
+
+#### Simulation area
+
+We simulated a forest area of $1\text{ha}$, which was divided into
+$25$ patches of size $20\text{m}\times20\text{m}$. No interactions
+between trees from different patches were considered in the model.
+That is, the $25$ patches were independent replicates of one another.
+
+#### Seed generation and dispersal
+
+We modelled seed generation as global external seed rain. Specifically, we 
+assumed that for each species, there is a constant seed inflow of 
+$250\frac{\text{seeds}}{\text{ha}\cdot \text{yr}}$, which is distributed 
+evenly among the $25$ individual patches. We supposed that all 
+seeds promptly establish to seedlings with initial stem diameters at 
+breast height (DBH) of 5\text{cm}.
+
+#### Simulation length and sampling approach
+
+We started the simulation at each site with a burn-in phase of $1000$
+years and collected samples of the output variables in subsequent
+years. The length of the burn-in phase was chosen based on earlier
+experiences with the model. To confirm that the model reached its
+limiting behaviour, we plotted the evolution of the output variables
+(e.g. the basal area) at selected sites over time and verified that
+no trends were visible after the burn-in phase.
+
+We computed the output variables in equidistant time intervals of
+$60$ years. Before taking a sample, we saved parts of the model's
+state. Then, we simulated the forest for $10$ years and computed
+the output variables, comparing the model's current state with the
+state saved earlier (see below).
+After taking the sample, we simulated the forest for another $50$
+years to reduce correlations between samples. Then, we repeated the
+procedure until the desired number of $200$ samples was collected.
+
+### Outputs
+
+To determine the number and basal area of recruits, we compared the
+set of large trees before and after the regeneration time period of
+$10$ years. Prior to simulating the forest for the regeneration time
+period, we saved the internal IDs of all trees with DBHs above the
+thresholds of $0.07\text{m}$ or $0.1\text{m}$, respectively. We
+stored these IDs in set data structures $S_{0.07}^{\text{prior}}$ and
+$S_{0.1}^{\text{prior}}$, corresponding to one threshold value each.
+After the regeneration time, we repeated the procedure, obtaining
+tree ID sets $S_{0.07}^{\text{posterior}}$ and $S_{0.1}^{\text{posterior}}$.
+We then computed the set differences 
+$S_{0.07}^{\text{recruits}}=S_{0.07}^{\text{posterior}}\backslash S_{0.07}^{\text{prior}}$
+and $S_{0.1}^{\text{recruits}}=S_{0.1}^{\text{posterior}}\backslash S_{0.1}^{\text{prior}}$
+and determined both the number and cumulative basal area of the trees
+in $S_{0.07}^{\text{recruits}}$ and $S_{0.1}^{\text{recruits}}$ by considered
+species, respectively.
+
+Finally, we computed the total basal area for each considered species
+by adding up the basal areas of all corresponding individuals. Note
+that the initial DBH of seedlings in our model was $0.05\text{m}$.
+Consequently, smaller trees were not included in the basal area computation.
+
 ## ForClim 1 & ForClim 11

### 2023-01-23 code/dataPreparation.R
add dbh as a filter to calculate the total BA and total recruited BA

@@ -603,7 +603,7 @@
 # Calculate r.Totba and Totba per site and sample-----
     # Calculate the total basal area and total recruited basal area 
     simAgg <- dataOutputs |>
-                 dplyr::group_by(site, sample, model) |> 
+                 dplyr::group_by(site, sample, model, dbh) |> 
                  dplyr::summarise(r.trees = round(sum(r.trees), 2), 
                                   r.ba = round(sum(r.ba), 4),
                                   ba = round(sum(ba), 2), # This is total BA in the site, value per ha!
@@ -611,7 +611,7 @@
                                   wb = unique(wb))
 
     # Add the total ba for all the species 
-    BATot <- simAgg[, c("site", "sample", "model", "ba", "r.ba")]
+    BATot <- simAgg[, c("site", "sample", "model", "dbh", "ba", "r.ba")]
     colnames(BATot)[4] <- "Totba"
     colnames(BATot)[5] <- "Totr.ba"
     dataOutputs <- merge(dataOutputs, BATot, by = c("site", "sample", "model"))

### 2023-01-23 code/figures.R
removes a section of the code in figures that was duplicated (related to mortality figures)

@@ -2,8 +2,8 @@
 # Title: Figures
 # Aim: Create figures and analysis for the manuscript
 # Author: Olalla Díaz-Yáñez | @olalla | olalladiaz.net &
-#         some part of figures were inspired in the work of the regeneration
-#         workshop attendees
+#         some part of the figures code were based in the work of the 
+#         regeneration workshop attendees
 ################################################################################
 
 # Load plotting styles 
@@ -506,130 +506,7 @@
     
     dev.off()
     
-# Mortality ----  
-  
-  ### Fig. R ratio 7/10 ----
-     
-    # Data preparation    
-    ddata <- outputsDF[outputsDF$species %in% c(selSpecies, aDVGMSpecies), ]
-    
-    treesTot <- ddata |>
-        dplyr::group_by(model, site, sample, dbh) |>
-        dplyr::summarise(Tot.rtrees = sum(r.trees),
-                         dds = unique(dds),      
-                         wb = unique(wb))
-    meanTrees <- treesTot |>
-        dplyr::group_by(model, site, dbh) |>
-        dplyr::summarise(mean.rtrees = mean(Tot.rtrees),
-                         dds = unique(dds),  
-                         wb = unique(wb))
-    
-    meanTrees7 <- meanTrees[meanTrees$dbh == 7,]
-    colnames(meanTrees7) <- c("model", "site", "dbh", "nn7", "dds", "wb")
-    meanTrees7 <- meanTrees7[, c("model", "site", "nn7")]
-    meanTrees10 <- meanTrees[meanTrees$dbh == 10,]
-    colnames(meanTrees10) <- c("model", "site", "dbh", "nn10", "dds", "wb")
-    meanTrees10 <- meanTrees10[, c("model", "site", "nn10",  "dds", "wb")]
-    
-    mortAll <- merge(meanTrees7, meanTrees10, by = c("model","site"))
-    mortAll$nn710 <- mortAll$nn7 / mortAll$nn10
-    
-    # the plot 
-    ratio7_10 <- ggplot2::ggplot(mortAll, ggplot2::aes(y = nn710, x = model, 
-                                                     fill = model)) + 
-        ggplot2::geom_boxplot() +
-        ggplot2::geom_hline(ggplot2::aes(yintercept = 1), colour = "darkblue", 
-                            linetype = "dashed") +
-        ggplot2::geom_hline(ggplot2::aes(yintercept = 1.77), colour = "darkblue",
[…]

### 2023-01-23 code/dataPreparation.R
an addition to fix in data Preparation the estimation to TotBA

@@ -611,10 +611,11 @@
                                   wb = unique(wb))
 
     # Add the total ba for all the species 
-    BATot <- simAgg[, c("site", "sample", "model", "dbh", "ba", "r.ba")]
+    BATot <- simAgg[, c("site", "sample", "model", "ba", "r.ba", "dbh")]
     colnames(BATot)[4] <- "Totba"
     colnames(BATot)[5] <- "Totr.ba"
-    dataOutputs <- merge(dataOutputs, BATot, by = c("site", "sample", "model"))
+    dataOutputs <- merge(dataOutputs, BATot, 
+                         by = c("site", "sample", "dbh", "model"))
 
     
 # Save the data -----

### 2023-01-23 code/figures.R
changes the axis of Fig5 and FigS 5 (combined_richness.png/allAgesRichness.png) to match Hstand range (changed caused by the TotBA re-estimation)

@@ -792,7 +792,7 @@
                        strip.background =  ggplot2::element_blank(),
                        axis.line = ggplot2::element_line(colour = "black"),
                        legend.position = "right") +
-        ggplot2::coord_cartesian(xlim = c(0, 1.5), ylim = c(0, 1.5)) +  
+        ggplot2::coord_cartesian(xlim = c(0, 2), ylim = c(0, 2)) +  
         ggplot2::labs(y = bquote(bar(H) * " stand"),
                       x = bquote(bar(H) * " stand")) +
         ggplot2::facet_wrap(~model) +
@@ -935,7 +935,7 @@
                                       color = dbh),
                          fill = "white") + 
         hrbrthemes::theme_ipsum() +
-        ggplot2::coord_cartesian(xlim = c(0, 2), ylim = c(0, 2)) +  
+        ggplot2::coord_cartesian(xlim = c(0, 2.5), ylim = c(0, 2.5)) +  
         ggplot2::geom_point(size = 0.6, alpha = 0.7 ) + 
         ggplot2::labs(y = bquote(bar(H) * " simulated stand"),
                       x = bquote(bar(H) * " observed stand")) +

### 2023-01-23 code/figures.R
change the legend scales in HeatEnvGrad_ / Figure 6 and FigS 7 to match the different TotBA values

@@ -998,7 +998,8 @@
                            show.legend = FALSE) +
         ggplot2::scale_color_manual(values = c("7" = "#FDAE61", "10" = "grey"),
                                     name = "Diameter threshold",
-                                    labels = c("7" = '7 cm', "10" = '10 cm'))
+                                    labels = c("7" = '7 cm', "10" = '10 cm')) 
+
     
     ggplot2::ggsave("figures/recruitmentRichness.png",
                     plot = d, width = 22, height = 18, scale = 0.9,
@@ -1083,7 +1084,7 @@
                                                          y = wb_bin,
                                                          fill = Diff)) +
       ggplot2::scale_fill_gradient2(low = "blue", mid = 'white', high = "red",
-                                    breaks = c(-0.6, -0.3, 0, 0.3, 0.6),
+                                    breaks = c(-1, -0.5, 0, 0.5, 1),
                                     midpoint = 0,
                                     name = expression(bar(R) * " BA share difference")) +
  
@@ -1123,7 +1124,7 @@
                                            "(1.11e+03,1.25e+03]" = "")) +
       ggplot2::geom_point(ggplot2::aes(x = dds_bin, y = wb_bin, 
                                        size = r.baTot.r.ShareMean), shape = 1) +
-      ggplot2::scale_size_area(breaks = c(0.1, 0.3, 0.5, 0.7), limits = c(0.0001, 0.8)) + 
+      ggplot2::scale_size_area(breaks = c(0.1, 0.3, 0.6, 0.9), limits = c(0.0001, 1)) + 
       ggplot2::labs(size = expression(bar(R) * " BA share")) 
     
     

### 2023-01-24 manuscript/supplementaryRW.Rmd
Remove duplicate heading for Formind
correct a small spelling error in Formind suplementary report
update report from TreeMig in supplementary

@@
 
-## FORMIND
-
 ### Authors
@@
 
-We did not use any topograpy data.
+We did not use any topography data.
 
@@
 ### Authors
+Heike Lischke
 
@@
 
-*Please provide here the references to this model where it is possible
-to understand the model structure specially for the regeneration module,
-this could be a published paper, a website with the model description, a
-manual or some other document that the reader can access.*
+The model TreeMig (see treemig.wsl.ch) was originally based on the ForClim model, then aggregated by a distribution based approach (Lischke et al.
+1998), and extended to a forest landscape model by including seed production, seed dispersal, seed bank dynamics and seedling establishment (Lischke and Loffler 2006, Lischke et al.
+2006).
 
@@
 
-What climate data have you used, and how have you created your model
-climate input parameters? I used the monthly time series of tas and
-precipitation,additionally slope and aspect from the topo- data set, and
-soil quality from the quality data set. slasp= 2*cos(aspect* pi/180)
-\*min(1, slope/60)
+As climatic input, the monthly time series of mean temperature and precipitation sum were used.
+With these climatic variables and additional data about slope, aspect and soil field capacity, yearly time series of the following bioclimatic variables were calculated:  (1) DDSum, the yearly day degree sum above 5.5°C , (2) MinWiT, the lowest mean monthly temperature, (3) DrStr, drought stress according to a bucket model  that estimates a monthly PET following (Thornthwaite and Mather 1957), using temperature, latitude, slope and aspect as input,  and calculates the monthly soil water content  based on soil water content in the previous month, precipitation, interception, and field capacity (“bucketsize”)   (Bugmann 1994, Fischlin et al.
+1995).
+Drought stress is then given by 1- the ratio between demand (PET-interception) and the supply, a function of the soil water content.
+To get climate data also for the spin-up, bioclimate was calculated from the monthly data   as given, and then extended by sampling from the given data from the end of the climate data to get 1000 years.
+The same sequence of sampling years was used in all simulations and sites.
 
@@
 
-The soil quality I translated into AWC (of  upmost m soil, "bucketsize")
-by Bucketsize = 10 + (soil_quality[-1) /4  \* 25 , to covert he range
-from 10 cm to 35cm   With these values I calculated the yearly time
-series of the bioclimatic variables DDsum (day degree sum above 5.5°C) ,
-minWinterTemp (lowest mean monthly temperature) and Drought stress
-according to a bucket model, that estimates a monthly PET following
-[Thorntwaithe, 1957],  and calculates the monthly soil water content
-   based on soil water content in the previous month, precipitation,
-interception, and bucketsize   [Bugmann , 1994] . Drought stress is then
-given by 1- the ratio between demand (PET-interception) and the supply
-(a function of the soil water content). Extending the climate data I
-used the monthly data as given, calculated the bioclimate, and extended
-since the end of the climate data to get 1000 years, by sampling from
-the given data. I used always the same sequence of sampling years in all
-simulations and sites. (-\> better comparability of experiments)
-
-### Soil
+The soil bucketsize (available water capacity of the upmost 1m soil layer) required for the drought stress calculation, was derived from the soil quality data set, by
+Bucketsize = 10 + (soil_quality-1) /4  * 25 , to covert the range from 10 cm to 35cm
 
@@
 
+Slope and aspect were taken from the topographic data set and converted to a common variable entering the PET calculation, by
+slasp= 2*cos(aspect *pi/180) *min(1, slope/60)
+
 ### Tree species
 
-How have you defined the species composition in your model simulation?
-(e.g., species levels, PFTs, etc.)? Please describe the species
-categories used. Species level, 30 species (see list of species I sent).
+Thirty Central European tree species were simulated, including the standard species of the project .
+Only Betula pubescens was not simulated, because we did not have a parametrization for this species at hand.
+Species: Abies alba, Larix decidua, Picea abies, Pinus cembra, Pinus montana, Pinus sylvestris, Taxus baccata, Acer campestre, Acer platanoides, Acer pseudoplatanus, Alnus glutinosa, Alnus incana, Alnus viridis, Betula pendula, Carpinus betulus, Castanea sativa, Corylus avellana, Fagus sylvatica, Fraxinus excelsior, Populus nigra, Populus tremula, Quercus petraea, Quercus pubescens, Quercus robur, Salix alba, Sorbus aria, Sorbus aucuparia, Tilia cordata, Tilia platyphyllos, Ulmus glabra
 
@@
 
-How have you defined the simulation area for your model (e.g., size and
-other relevant site aspects)? A grid of 15 \* 15 grid cells, each grid
-cell with 200m side length.
+The simulations were carried out on a grid of 15 x 15 cells, each with 200 m side length.
 
@@
 
-What are the relevant aspects of the dispersal (if any) and seed input
-parameters in your model? Seed dispersal is described by species
-specific, height dependent seed production. The seeds are then
-transported according to a double negative exponential, species specific
-kernel, and enter the seed bank in the sink cells, from which seedlings
-germinate depending on light and other environmental conditions. At the
-start of the simulation, for 10 years seeds of all species are available
-in all grid cells, from then on, the normal seed dispersal starts.
-
-There are 5 different experiments related to dispersal NoDisp:  the
-dispersal kernels are so narrow, that the huge majority of seeds stays
-in the source grid cell. I.e. the spatial variability (see below) only
-arises from the sampling of the light classes. But there is still the
-feedback between the mother trees and the seedlings. StochDisp:
-stochastic dispersal. The seeds are distributed randomly, whereby the
-kernel gives the probability of the seeds  landing in a certain distance
-DetDisp: deterministic dispersal. The kernel determines the share of the
-seeds in the source cell landing in a sink cell in a given distance to
-the source cell. FBirthAll: a fixed ingrowth above 1.37m of 20
-trees/833m\^2  and year, same for all species, but reduced by
-environmental drivers and shade.  This is analogous to the approach of
-ForClim,  i.e. ignores the feedback  between mother trees and young
-trees and the youth development FBirthIndiv: as FBirthAll, but the fixed
-ingrowth is species-specific  (between 4 and 54 trees/833m\^2 )
+The seeds produced by the mother trees are transported from the source cell according to a double negative exponential, species specific that determines the share of the seeds in the source cell landing in a sink cell in a given distance to the source cell.
+In the sink cells, the seeds enter the seed bank .
+At the start of the simulation for 10 years seeds of all species are available in all grid cells, from then on, the normal seed dispersal starts.
+
+#### Regeneration processes
+
+TreeMig simulates the full feedback from seed production to ingrowth. Some of the involved processes depend on environment.
+(1) Overall seed production in TreeMig is described by multiplying a species specific seed production per tree that depends on tree size and thus indirectly on the growth, which depends on DDsum, drought stress, nutrients and light and the adult tree number that depends on survival  that depends on DDsum, drought stress, nutrients and light.
+(2) Seed dispersal is independent of environment. 
+(3) Seed bank dynamics – also environment independent -   is given by seed input, seed mortality, and germination.
+Additionally, the number of seeds S of a species in the seed bank of this species (SB) is limited to a carrying capacity C of 1000, by
+SB = C + (SB - C) * Exp(-S/C).  
+(4) Germination and seedling survival depend on browsing, drought, winter temperature, DDSum and light. 
+(5) The saplings up to 1.37 die and grow like adult trees, depending on DDsum, drought stress and light. Only the parameter value of the light dependence differs from that of the adult trees.  
+The regeneration processes were simulated annually in the order 1.
+germination, 2.
+germinated seeds leave the seed bank, 3.
+mortality of seeds in the seed bank, 4.
+seed entry by seed dispersal.
+The ingrowth numbers were very sensitive to this order.
+Simulations with another sequence, in which the seeds in the seed bank died first and then  the  remaining ones germinated,  resulted in about half the ingrowth.
 
@@
 
-What is the simulation length, and how have you determined the
-simulation time until the equilibrium? 1000 years, experience and check
-by eye-ball of the simulated basal areas. Variability: I have employed
-three different aspects of the model  to account for spatial and
-temporal variability and to get the required 200 patches of 1ha.
-
-1)  Basic spatial variability due to the combination of demographic
-    stochasticity and short-range spatial interactions. In contrast to
-    gap models, TreeMig is based on the concept of frequency
-    distributions of tree densities classes in different tree heights on
-    ca. 1/12 ha (833m\^2) patches, and the resulting frequency
-    distributions in discrete light. During the simulation, in each
-    height class the frequencies of the different light classes are
-    recorded, as well as the ingrowth of each species in the
-    height-light class. By sampling 12 times from the height-light class
-    ingrowths according to the height-light class frequency
-    distribution, and summing up, the stochastic ingrowth for one ha is
-    calculated, for each grid cell.  Then the height class boundaries
-    are translated into DBH and the ingrowths in the height classes
-    linearly interpolated to 7cm and 10cm.
-
-2)  Spatial smoothing and variability by the formulation of dispersal.
-    The different dispersal strategies, see above. With NoDisp, solely
-    the basic spatial variability is accounted for. With DetDisp, I
-    expect the dispersal to smooth out the variability. StochDisp, in
-    turn is expected to create additional variability by stochastic
-    founder processes.
-
-3)  Temporal variability by sampling from different decades of the last
-    100 years of the  simulation, i.e. years 900-1000
-
-#### Regeneration parameters
-
-How does environment affect regeneration in the model? Seed production:
-depends on tree size and thus indirectly on the growth, that  depends on
-DDsum,  drought stress, nutrients and light and on adult tree number and
-thus on survival  that depends on DDsum, drought stress, nutrients and
-light
-
-Seed dispersal:  independent on environment Seed bank dynamics : 
-independent on environment Germination and seedling survival: Depends on
-drought, browsing, winter temperature and DDSum and light (last 3 are
-step functions) Sapling dynamics (up to 1.37): like adult growth and
-survival, i.e. depends on DDsum, drought stress and light but light
-dependence with a specific parameter values, namely the klights
-parameter instead of klighta
-
-#### more simulations per site
-
-If you have run more than one simulation per site, please describe the
-differences of the approaches used. I used the following 15
-combinationsExperiments (exp_txpe) NoDisp, StochDisp, DetDisp,
-FBirthAll, FBirthIndiv), each with the variability setups (comb_type)
-200space1time: 200 grid cells at a single randomly chosen decade (
-50space4time: 50 grid cells at 4 randomly chosen decades, or
-20space10time: 20 grid cells at all 10 decades Always the same random
-(nested) sets of decades and grid cells were used Fixed ingrowth values
-(for the FBirthAll and FBirthIndiv experiments ) were calculated
-beforehand in a  set of simulations with single species under optimal
-conditions, without interspecific, but under intraspecific competition.
-The ingrowth into the height class above 1.37m per species was averaged
-over the last 100  years of the 1000 year simulation and used in the
-FBirthIndiv simulations. For the FBirthAll simulations the mean of the
-single species ingrowths' (20) was used.
-
-The "standard" combination is DetDisp with 20 grid cells at all 10
-decades Please describe how have you calculated each of the variables in
-the output data. During the simulation,  in each height class the
-frequencies of the different light classes are recorded, as well as the
-ingrowth numbers of each species in the height-light class. By sampling
-12 times from the height-light class ingrowth numbers according to the
-height-light class  frequency distribution, and summing up, the
-stochastic ingrowth number for one ha is calculated, for each grid
-cell.  Then the height class boundaries are translated into DBH and the
-ingrowths in the height classes linearly interpolated to 7cm and 10cm
-based.  The ingrowth basal areas into the height classes are obtained by
-multiplying the interpolated ingrowth numbers with the basal areas at 
-7cm and 10cm, i.e. r.trees \* Pi *(0.07/2)\^2  and r.trees* Pi\* 
-(0.1/2)\^2 .
-
-The species specific basal area (at 1.37m) sums are a standard output of
-TreeMig, given by multiplying the state variables (number of trees per
-grid cell , height class, and species) with  the species specific basal
-area of each height class. Files TreeMigRes_all_data.RData: R RData-file
-with all experiments (exp_type) and variability combinations (comb_type)
-TreeMigRes_Std_data.RData: R RData-file with   exp_type=detDisp) and
-variability combinations (comb_type=20space10time)
+The simulations were started on bare ground and run for 1000 years, when at all sites the equilibrium was reached.
+
+#### More simulations per site
+
+TreeMig simulates directly the mean dynamics of entire stands, but based on the  variability of light conditions within the stand   [distribution based approach, \Lischke, 1998, aggregation.
+To come up with the required 200 random patches of 1 ha, temporal and spatial variability intrinsic to TreeMig were combined.
+Temporal variability was given by sampling from each decade of the last 100 years of the simulation, i.e.
+years 900-1000.
+For each of theses 10 decades, 20 grid cells were sampled from the simulated grid, reflecting spatial variability, results from demographic stochasticity and short-range spatial interactions
 
 ### Outputs
+
+In contrast to gap models, TreeMig is based on the concept of frequency distributions of tree densities classes in different tree heights on ca.
+1/12 ha (833m^2) patches, and the resulting frequency distributions in discrete light.
+During the simulation, in each height class the frequencies of the different light classes were recorded, as well as the ingrowth of each species in the height-light class.
+By sampling 12 times from the height-light class ingrowths according to the height-light class frequency distribution and summing up, the stochastic ingrowths into all TreeMig height classes for one ha were calculated for each grid cell.
+Then the height class boundaries were translated into DBH and the ingrowths into the height classes linearly interpolated to 7cm and 10cm.
+The ingrowth basal areas into the height classes were obtained by multiplying the interpolated ingrowth numbers with the basal areas at  7cm and 10cm, i.e.
+r.trees * Pi*  (0.07/2)2  and r.trees * Pi*  (0.1/2)2 .
+The species specific basal area (at 1.37m) sums are a standard output of TreeMig, given by multiplying the state variables (number of trees per grid cell, height class, and species) with  the species specific basal area of each height class.
+To account for the ingrown trees dying within a decade, as an approximation the survival in each year to the power of 5 was used.
+Thereby, it was assumed that the trees have to survive in average 5 years in each decade, and that the current mortality is valid for these 5 years.
+
+### References
+
+•	Bugmann, H. 1994. On the ecology of mountainous forests in a changing climate: A simulation study. Dissertation. Swiss Federal Institute of Technology, Zurich.
+•	Fischlin, A., H. Bugmann, and D. Gyalistras. 1995. Sensitivity of a forest ecosystem model to climate parametrization schemes. Env. Poll. 87:267-282.
+•	Lischke, H., and T.
+J.
+Loffler.
+2006.
+Intra-specific density dependence is required to maintain species diversity in spatio-temporal forest simulations with reproduction.
+Ecological Modelling 198:341-361.
+•	Lischke, H., T.
+J.
+Loffler, and A.
+Fischlin.
+1998.
+Aggregation of individual trees and patches in forest succession models: Capturing variability with height structured, random, spatial distributions.
+Theor Popul Biol 54:213-226.
+•	Lischke, H., N.
+E.
+Zimmermann, J.
+Bolliger, S.
+Rickebusch, and T.
+J.
+Loffler.
+2006.
+TreeMig: A forest-landscape model for simulating spatio-temporal patterns from stand to landscape scale.
+Ecological Modelling 199:409-420.
+•	Thornthwaite, C. W., and J. R. Mather. 1957. The water balance. Publications in Climatology VIII:1-69.
 

### 2023-01-24 manuscript/supplementaryRW.Rmd
updates report from ForClim in supplementary

@@
 
+Yannek Käber
+Harald Bugmann
+
 ### Model references
 
-*Please provide here the references to this model where it is possible
-to understand the model structure specially for the regeneration module,
-this could be a published paper, a website with the model description, a
-manual or some other document that the reader can access.*
+Bugmann, H. (1994). On the ecology of mountainous forests in a changing climate: A simulation study.
+Huber, N., Bugmann, H., & Lafond, V.
+(2020).
+Capturing ecological processes in dynamic forest models: Why there is no silver bullet to cope with complexity.
+Ecosphere, 11(5).
+https://doi.org/10.1002/ecs2.3109
 
@@
 
+Temperature and precipitation were aggregated to monthly means and sums, respectively, including their standard deviation.
+These values served as input for the weather generator within the model.
+The weather generator simulates wheather conditions in an annual time step based on the monthly average temperatures and precipitations sums.
+
 ### Soil
 
+Soil data were translated to bucket sizes values based on the assumpation that the lowest soil quality value of 1 refers to a bucket size of 10 cm and the highest soil quality of 5 refers to a bucket size of 35 cm.
+Specifically we used the following formula to calculate the bucket size bucket_size = 10 + (soil_quality-1) /4  * 25.
+
 ### Topography
 
+Topographic information was included by specifying the kSlAsp parameter.
+This parameter defines a value between -2 and 2 which affects the calculated Potential Evapotranspiration within the model.
+For details on the calculation of this value see Käber et.
+al.
+2021.
+
 ### Tree species
 
+All species required were simulated:
+Fagus sylvatica, Picea abies, Abies alba, Carpinus betulus, Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris
+
+In addition the following species were simulated:
+Acer campestre, Acer platanoides, Alnus incana, Alnus viridis, Castanea sativa, Corylus avellana, Larix decidua, Pinus montana, Populus nigra, Populus tremula, Salix alba, Sorbus aria, Sorbus aucuparia, Taxus baccata, Tilia platyphyllos, Ulmus glabra, Pinus cembra
+
 ### Simulations
 
+All simulations were run according to the protocoll.
+For each site we simulated on 16 ha (i.e., 100 patches of 0.08 ha).
+We defined a spin up phase of 1000 years.
+After the spin up we simulated 200 years of forest dynamics.
+
 ### Outputs
+
+We sampled in 10 year intervals between the simulation years 1010 to 1200, which resulted in 16 times 20 intervals each with one ha.
 

### 2023-01-24 manuscript/supplementaryRW.Rmd
updates report from Landis in supplementary

@@
 
-*Please provide here the references to this model where it is possible
-to understand the model structure specially for the regeneration module,
-this could be a published paper, a website with the model description, a
-manual or some other document that the reader can access.*
-
-Landscape scale forest simulation model LANDIS-II (Scheller et al. 2007)
-with PnET-Succession extension (version 4.1). PnET-Succession is based
-on the Biomass Succession extension of Scheller and Mladenoff (2004),
-embedding elements of the PnET-II ecophysiology model of Aber et al
-(1995) to mechanistically simulate growth as a competition for available
-light and water (De Bruijn et al. 2014). PnET-Succession simulates the
-competition among cohorts for water and light at a monthly time-step as
-a function of photosynthetic processes and maintenance respiration that
-are explicitly linked to environmental drivers such as temperature,
-precipitation, photosynthetic active radiation (PAR), and CO2
-concentration. Regeneration depends on distance from a seed source, soil
-water, and subcanopy light, while competition is modeled by partitioning
-incoming light through multiple canopy layers. Growth of specific cohort
-biomass components is allocated in the following order: non structural
-carbon, first, then foliage then root and stem. Biomass growth decreases
-as cohorts approach their longevity age but mortality can occur at any
-time when carbon reserves production is insufficient to support growth
-due to shading, water competition, and/or drought (Gustafson et al.
-2015), or by any disturbance or management extension used. For output,
-we selected Biomass Community Output Extension version 2.0.1, which
-generates output in the format: species age biomass (g / m2) at
-specified time steps.
+*Please provide here the references to this model where it is possible to understand the model structure specially for the regeneration module, this could be a published paper, a website with the model description, a manual or some other document that the reader can access.*
+
+Landscape scale forest simulation model LANDIS-II (Scheller et al.
+2007) with PnET-Succession extension (version 4.1).
+PnET-Succession is based on the Biomass Succession extension of Scheller and Mladenoff (2004), embedding elements of the PnET-II ecophysiology model of Aber et al.
+(1995) to mechanistically simulate growth as a competition for available light and water (De Bruijn et al.
+2014).
+PnET-Succession simulates the competition among cohorts for water and light at a monthly time-step as a function of photosynthetic processes and maintenance respiration that are explicitly linked to environmental drivers such as temperature, precipitation, photosynthetic active radiation (PAR), and CO~2~ concentration.
+Regeneration depends on distance from a seed source, soil water, and subcanopy light, while competition is modeled by partitioning incoming light through multiple canopy layers.
+Growth of specific cohort biomass components is allocated in the following order: non structural carbon, first, then foliage then root and stem.
+Biomass growth decreases as cohorts approach their longevity age but mortality can occur at any time when carbon reserves production is insufficient to support growth due to shading, water competition, and/or drought (Gustafson et al.
+2015), or by any disturbance or management extension used.
+For output, we selected Biomass Community Output Extension version 2.0.1, which generates output in the format: species age biomass (g / m^2^) at specified time steps.
 
@@
 
-Climate input was generated by randomly selecting entire years from the
-monthly data from the CHELSA dataset and collating them one after the
-other to create a 2000 years series. The sampling was repeated 10 times
-with different random order of years to create the 10 climate
-replicates. Selecting a year at a time eliminates inconsistent seasons
-(especially summer). We used monthly mean, minimum and maximum
-temperatures and precipitation as input for PnET succession extension.
-PnET-Succession also requires mean monthly values of PAR during daylight
-hours, which we calculated by dividing the monthly rsds data by the
-number of daylight seconds in each month for latitude 47 to get W/m2. We
-then multiplied this by 4.6 to get μmol/m2/s and multiplied by 0.5 to
-get PAR part of the spectrum, based on the approximation by Tsubo et al.
-(2005). Leap years were not considered. The PAR distribution across all
-sites and years is in range with values that are typical for temperate
-regions (Fig. 1).
-
-Figure 1 Histogram of monthly PAR values for all sites and all available
-time
-
-CO2 concentration with yearly timestep was added from the concentrations
-table available from EEA (2019). For years 0-1750 CO2 concentration was
-maintained fixed to 278 ppm. As for the period 1750-1975 there was only
-one value each 5 years, we used a constant value for each 5-year period
-(Fig. 2). CO2 concentration was added to the climate input file after
-randomization of temperature, precipitation and PAR (see above), so it
-reflects realistic values for the simulated years. Simulations were run
-from year 0 to 1990 but sampled outputs did not exceed year 1910 (see
-Simulation length and descriptions further below). Thus, the drastic
-increase in CO2 concentration during the second half of the 1900s did
-not affect the delivered results.
-
-Figure 2 CO2 concentration values added to the climate data
+Climate input was generated by randomly selecting entire years from the monthly data from the CHELSA dataset and collating them one after the other to create a 2000 years series.
+The sampling was repeated 10 times with different random order of years to create the 10 climate replicates.
+Selecting a year at a time eliminates inconsistent seasons (especially summer).
+We used monthly mean, minimum and maximum temperatures and precipitation as input for PnET succession extension.
+PnET-Succession also requires mean monthly values of PAR during daylight hours, which we calculated by dividing the monthly rsds data by the number of daylight seconds in each month for latitude 47 to get W/m2.
+We then multiplied this by 4.6 to get μmol/m^2^/s and multiplied by 0.5 to get PAR part of the spectrum, based on the approximation by Tsubo et al.
+(2005).
+Leap years were not considered.
+The PAR distribution across all sites and years is in range with values that are typical for temperate regions.
+
+CO~2~ concentration with yearly timestep was added from the concentrations table available from EEA (2019).
+For years 0-1750 CO~2~ concentration was maintained fixed to 278 ppm.
+As for the period 1750-1975 there was only one value each 5 years, we used a constant value for each 5-year period.
+CO~2~ concentration was added to the climate input file after randomization of temperature, precipitation and PAR (see above), so it reflects realistic values for the simulated years.
+Simulations were run from year 0 to 1990 but sampled outputs did not exceed year 1910.
+Thus, the drastic increase in CO~2~ concentration during the second half of the 1900s did not affect the delivered results.
 
@@
 
+Soil type was set to LOAM at all sites.
+Rooting depth was calculated from soil quality codes (1-5, original values ranged from 1.13 to 4.87) multiplicated by 205.
+This produces a soil depth in the range 231 - 998 mm, that are within the typical range for PnET-Succession (maximum 1000 mm).
+A clear trend emerged - deeper soils resulted in higher biomass, although with some exceptions.
+
 ### Topograhpy
 
+The ecoregion-specific parameter Precipitation Loss Fraction denotes the water lost to runoff induced by topographic slope.
+We estimated this parameter for each site from topo_data_200 dataset as slope \* 0.01 (e.g., 12% slope = 0.12 PrecLossFrac).
+We acknowledge that this is a simplification and that better parameterization would be possible if soil types were available for each site.
+Tests showed that sites with extreme slope tended to have lower biomass, but the signal was not that clear.
+The correlation between slope and elevation was 0.6 which could obscure the results.
+
+Some sites at very high elevation had, correspondingly, very low temperature.
+At these sites, the model showed a visible trend in producing lower biomass (down to zero at a few sites).
+Yet this was probably not apparent in the regeneration results, as extremely low biomass could still support regeneration, even if cohorts do not live long, due to our modification of sexual maturity age.
+
 ### Tree species
 
-As LANDIS-II cannot start from bareground, initial communities were set
-to include all species defined in the protocol at all sites starting
-with 10-years old cohorts of the following species: fasy (Fagus
-sylvatica), pabi (Picea abies), abal (Abies alba), cabe (Carpinus
-betulus), tico (Tilia cordata), acps (Acer pseudoplatanus), betu (Betula
-spp.), frex (Fraxinus excelsior), quer (Quercus spp.), algl (Alnus
-glutinosa), pisy (Pinus sylvestris). The species-specific parameter
-denoting sexual maturity was set to 10 years equally for all species, so
-that seed source of all species is available from the onset of the
-simulations at all sites and to maintainseed in the ladscape pool.
-Vegetative regeneration (resprouting) was prevented, since this would
-have an effect on allometry and regeneration. Some species may be under
-or overrepresented because their occurrence in LANDIS-II simulations is
-typically due to small-scale disturbances or management which were not
-included in the simulations. We have started with a set of parameters we
-have used for Petter et al. (2020) which were further refined and
-modified for PnET v 4.1. Describe species parameters here...
+As LANDIS-II cannot start from bare ground, initial communities were set to include all species defined in the protocol at all sites starting with 10-years old cohorts of the following species: fasy (*Fagus sylvatica*), pabi (*Picea abies*), abal (*Abies alba*), cabe (*Carpinus betulus*), tico (*Tilia cordata*), acps (*Acer pseudoplatanus*), betu (*Betula spp.*), frex (*Fraxinus excelsior*), quer (*Quercus spp.*), algl (*Alnus glutinosa*), pisy (*Pinus sylvestris*).
+The species-specific parameter denoting sexual maturity was set to 10 years equally for all species, so that seed source of all species is available from the onset of the simulations at all sites and to maintain seed in the ladscape pool.
+Vegetative regeneration (resprouting) was prevented, since this would have an effect on allometry and regeneration.
+
+Some species may be under or overrepresented because their occurrence in LANDIS-II simulations is typically due to small-scale disturbances or management which were not included in the simulations.
+
+The model requires two sets of parameters representing species-specific life traits.
+The first set is used by the LANDIS-II core module.
+The species-specific parameter denoting "longevity" and "sexual maturity" were the only parameters of this set considered in the current exercise.
+The "sexual maturity" was set to 10 years equally for all species, so that seed source of all species is available from the onset of the simulations at all sites and to maintain seed in the landscape pool.
+Vegetative regeneration (resprouting) was prevented since this would influence allometry and regeneration.
+
+The second set is used by the PnET-Succession extension and includes several ecophysiological parameters.
+For this second set we started with the values already applied in Petter et al.
+(2020) with a previous version of PnET.
+The values of some of the parameters to which PnET is most sensitive and/or for which the range of variation had been modified for PnET4.1 were then iteratively adjusted (individually and then in combination) to meet the requirements of PnET version 4.1 used in this exercise.
+The changes were implemented one by one in simulations with the exercise data and those leading to an output deemed realistic and to a relative stabilization of the biomass trend through the simulation were retained for the implementation of the full set of replicates.
+The retained changes include: 1) decreased percentage of foliar nitrogen for algl (*Alnus glutinosa*); 2) rescaling to the new ranges for all species of the values of the Halfsat and the FracBelowG.
+The first of these parameters represents shade tolerance and its value affects the probability of establishment of a species (Half saturation light level for photosynthesis.); the latter regulates allocation between above and below ground pool (Fraction of non-foliar biomass that is belowground (root pool)) ; 3) a slight decrease of the value of the parameter (Fracfol) affecting the yearly allocation to foliage of the active total woody biomass for the three conifers abal (*Abies alba*), piab (*Picea abies*), pisy (*Pinus sylvestris*).
 
@@
 
-Simulation area was defined as a rectangle of 10×20 dimension with sites
-(i.e., cells) of 1-ha. Each site is an individual ecoregion (i.e., one
-cell - as in our case - or group of cells sharing similar climate and
-soil conditions, as a way in LANDIS-II to define differences in
-biophysical conditions across landscapes) with specific values for soil
-quality, topography and climate, based on the data provided. Soil type
-was set to LOAM at all sites. Rooting depth was calculated from soil
-quality codes (1-5, original values ranged from 1.13 to 4.87)
-multiplicated by 205. This produces a soil depth in the range 231 - 998
-mm, that are within the typical range for PnET-Succession (maximum 1000
-mm). A clear trend emerged - deeper soils resulted in higher biomass,
-although with some exceptions (Fig. 3).
-
-Figure 3 Biomass of all species on each site ordered by soil quality
-(data from replicate 1). Notice a gap after first longevity period, an
-artifact of spin-up from "bare ground"
-
-The ecoregion-specific parameter Precipitation Loss Fraction denotes the
-water lost to runoff induced by topographic slope. We estimated this
-parameter for each site from topo_data_200 dataset as slope \* 0.01
-(e.g., 12% slope = 0.12 PrecLossFrac). We acknowledge that this is a
-simplification and that better parameterization would be possible if
-soil types were available for each site. Tests showed that sites with
-extreme slope tended to have lower biomass (Fig. 4), but the signal was
-not that clear. The correlation between slope and elevation was 0.6
-which could obscure the results.
-
-Figure 4 Biomass of all species on each site sorted by slope, which was
-used to modify the fraction of precipitation lost to run off (data from
-replicate 1)
-
-Some sites at very high elevation had, correspondingly, very low
-temperature (Fig. 5). At these sites, the model showed a visible trend
-in producing lower biomass (down to zero at a few sites see Fig. 6
-below). Yet this was probably not apparent in the regeneration results,
-as extremely low biomass could still support regeneration, even if
-cohorts do not live long, due to our modification of sexual maturity
-age.
-
-Figure 5 Histogram of elevation of simulated sites.
-
-Figure 6 Biomass of all species on each site ordered by elevation (data
-from replicate 1)
+Simulation area was defined as a rectangle of 10×20 dimension with sites (i.e., cells) of 1-ha.
+Each site is an individual ecoregion (i.e., one cell - as in our case - or group of cells sharing similar climate and soil conditions, as a way in LANDIS-II to define differences in biophysical conditions across landscapes) with specific values for soil quality, topography and climate, based on the data provided.
+The model runs with a yearly timestep using monthly climate.
+We started each model run with initial communities of all species at age 10 at all sites.
+Each run was simulated for 1980 years.
 
@@
 
-We have used the non-spatial seeding algorithm "universal dispersal" for
-all the model runs. This way, every species can seed any forest site in
-the landscape (across all the 200 sites). This mode does not take into
-account the spatial arrangement and contiguity of cells (i.e., no active
-seed dispersal), which is proper for the sake of the model comparison
-exercise. Yet in real landscape seed limitation can also play a role in
-species composition. The seed number is affected by species-specific age
-of maturity, which we have modified (see Species section). This setting
-ensures an influx of seeds from outside consistent with the original
-settings. Simulation length We started each model run with initial
-communities of all species at age 10 at all sites. Each run was
-simulated for 1980 years. Sampling of model outputs was executed in two
-time windows starting 50 years after the first cycle following typical
-species longevity (see Species parameters). After the year 700, biomass
-of most species was relatively stable and we sampled every 50 years,
-starting at 750 until the year 1200. The next sampling period started
-from 1450 and continued until 1900, i.e., after the next longevity peak
-(Fig. 7), again every 50 years (see also Table 1). At each selected
-sampling time, we extracted cohorts of each species according to the age
-related to the two DBH thresholds (7cm/10cm) (see age-DBH relationships
-in Table 2). Additionally, we extracted cohorts that were up to 9 years
-older than these thresholds, to get the expected results for a 10-year
-interval. For example, Fagus sylvatica reaches DBH 7cm at 17 years,
-therefore we have selected all cohorts with age 17-26.
-
-Figure 7 Biomass of each species in time across all sites (data from
-replicate 1)
-
-Figure 8 Sum of biomass of all species on each site in time (data from
-replicate 1). Notice again the biomass decrease before approximately the
-year 700 due to the effect of species longevity. The productivity of the
-sites seem to be stable.
+We have used the non-spatial seeding algorithm "universal dispersal" for all the model runs.
+This way, every species can seed any forest site in the landscape (across all the 200 sites).
+This mode does not take into account the spatial arrangement and contiguity of cells (i.e., no active seed dispersal), which is proper for the sake of the model comparison exercise.
+Yet in real landscape seed limitation can also play a role in species composition.
+The seed number is affected by species-specific age of maturity, which we have modified.
+This setting ensures an influx of seeds from outside consistent with the original settings.
 
@@
 
-We have sampled each site 20 times along the simulation run as described
-in the simulation length chapter. To deliver more than one simulation
-per site, we created 10 replicates with different climate data and
-random number generator seed number to account for stochastic processes
-in PnET-Succession. This resulted in the expected 200 samples for each
-site. Output variables
-
-sample Sample number code (1-200) was defined as sample in time +
-(replicate-1)\*20, with sample in time (1-20) according to the sample
-order. Replicate (1-10) combines random number and climate sequence (see
-Table 1. For example sample number 25 was taken from the climate
-replicate 2 and the fifth sample in time (year 950).
-
-Table 1 Sample numbers and their corresponding settings
-
-r.trees As LANDIS-II outputs are expressed in terms of cohort age and
-mean biomass per site, we have calculated the mean DBH and biomass of
-each species based on age. We estimated the age when each species
-reaches 7 cm and 10 cm DBH and the number of individuals using
-allometric models from the Italian forest inventory collected in R
-package ForIT (Puletti et al. 2014), and unpublished data from the study
-by Mina et al. (2021) (Table 2). Please note that the equations used
-predict DBH \>0 already for a tree with age 1, this is because it is
-based on forestry data and tree cores. The real age when reaching 1.3 m
-height is species specific and we did not implement this offset. For
-each cohort, we divided the cohort biomass reported by LANDIS-II,
-recalculated to kg/ha by the mean biomass of a tree of the same species
-and age to obtain the number of trees of each species per site. We
-rounded all results up to the nearest integer to avoid any fraction of
-trees. The number of individuals was aggregated for all the cohorts of
-the age of reaching 7 cm or 10 cm DBH and all cohorts of the same
-species that were up to 9 years older (Fig. 9). This approach does not
-reflect influence of site conditions or stand densities on Age-DBH and
-Age-Biomass. Relationships between age and DBH were derived by fitting
-asymptotic regressions using individual tree measurements from an
-ensemble of forest inventory plots distributed across south-eastern
-Canada (see Supporting information of Mina et al. 2021). This was the
-only public dataset for which we could relate tree age, derived from
-stem coring and tree ring analysis, to measured DBH. Although tree age
-measurements were obtained with a robust methodology, these functions do
-not disentangle the influence of site conditions or stand densities.
-Additionally, they reflect growing conditions of tree genera in Eastern
-North America.
-
-Figure 9 Number of trees in regeneration cohorts (both 7cm and 10 cm
-groups) according to species. (data from all replicates and samples)
-
-Table 2 Age at which each simulated species reaches DBH 7cm and 10 cm
-Age at DBH 7cm Age at DBH 10cm Comment fasy (for Fagus sylvatica) 17 25\
-piab (for Picea abies) 8 12\
-abal (for Abies alba) 10 15\
-cabe (for Carpinus betulus) 15 23 same equation as betu tico (for Tilia
-cordata) 15 23 equation taken from quer acps (for Acer pseudoplatanus)
-16 24\
-betu (for Betula spp.) 15 23\
-frex (for Fraxinus excelsior) 10 15\
-quer (for Quercus spp.) 15 23\
-algl (for Alnus glutinosa) 15 23 same equation as betu pisy (for Pinus
-sylvestris) 9 13
-
-r.ba basal area of regeneration Calculated as a sum of r.ba of all the
-cohorts selected in r.trees using the following formula: DBH2 \* π/4 \*
-r.tree (Fig. 10)..
-
-Figure 10 Basal area of trees in regeneration cohorts (both 7cm and 10
-cm groups) according to species. (data from all replicates and samples)
-
-ba: total basal area We have not calculated this since our allometric
-equations do not cover older ages. Therefore, we produced the following
-two biomass variables as a surrogate.
+We have sampled each site 20 times along the simulation run in two time windows starting 50 years after the first cycle following typical species longevity.
+After the year 700, biomass of most species was relatively stable and we sampled every 50 years, starting at 750 until the year 1200.
+The next sampling period started from 1450 and continued until 1900, i.e., after the next longevity peak, again every 50 years.
+At each selected sampling time, we extracted cohorts of each species according to the age related to the two DBH thresholds (7cm/10cm).
+Additionally, we extracted cohorts that were up to 9 years older than these thresholds, to get the expected results for a 10-year interval.
+For example, *Fagus sylvatica* reaches DBH 7cm at 17 years, therefore we have selected all cohorts with age 17-26.
+
+To deliver more than one simulation per site, we created 10 replicates with different climate replicate and random number generator seed number to account for stochastic processes in PnET-Succession.
+This resulted in the expected 200 samples for each site.
+
+### Outputs
+
+*Please describe how have you calculated each of the variables in the output data.*
+
+Sample number code
+
+Number (1-200) was defined as sample in time + (replicate-1)\*20, with sample in time (1-20) according to the sample order.
+Replicate (1-10) combines random number and climate sequence.
+For example sample number 25 was taken from the climate replicate 2 and the fifth sample in time (year 950).
+
+r.trees
+
+As LANDIS-II outputs are expressed in terms of cohort age and mean biomass per site, we have calculated the mean DBH and biomass of each species based on age.
+We estimated the age when each species reaches 7 cm and 10 cm DBH and the number of individuals using allometric models from the Italian forest inventory collected in R package ForIT (Puletti et al.
+2014), and unpublished data from the study by Mina et al.
+(2021).
+Please note that the equations used predict DBH \>0 already for a tree with age 1, this is because it is based on forestry data and tree cores.
+The real age when reaching 1.3 m height is species specific and we did not implement this offset.
+For each cohort, we divided the cohort biomass reported by LANDIS-II, recalculated to kg/ha by the mean biomass of a tree of the same species and age to obtain the number of trees of each species per site.
+We rounded all results up to the nearest integer to avoid any fraction of trees.
+The number of individuals was aggregated for all the cohorts of the age of reaching 7 cm or 10 cm DBH and all cohorts of the same species that were up to 9 years older.
+This approach does not reflect influence of site conditions or stand densities on Age-DBH and Age-Biomass.
+Relationships between age and DBH were derived by fitting asymptotic regressions using individual tree measurements from an ensemble of forest inventory plots distributed across south-eastern Canada (see Supporting information of Mina et al.
+2021).
+This was the only public dataset for which we could relate tree age, derived from stem coring and tree ring analysis, to measured DBH.
+Although tree age measurements were obtained with a robust methodology, these functions do not disentangle the influence of site conditions or stand densities.
+Additionally, they reflect growing conditions of tree genera in Eastern North America.
+
+r.ba basal area of regeneration Calculated as a sum of r.ba of all the cohorts selected in r.trees using the following formula: DBH2 \* π/4 \* r.tree.
+
+ba: total basal area We have not calculated this since our allometric equations do not cover older ages.
+Therefore, we produced the following two biomass variables as a surrogate.
+The basal area was later calculated as a sum of r.ba and estimated ba based on biomass.
+The estimation includes transformation of kg into tonnes / ha and expansion factor 12.5 to transform ba m^2^/ha.
+
+((( landis\$sum.bio - landis\$r.bio.sum )\* 0.001) / 12.5). This includes
 
 sum.bio\
-Biomass of all cohorts of the species on site including the regeneration
-and even younger regeneration (units kg/ha).
-
-r.bio.sum Biomass of the regeneration cohorts corresponding to the r. ba
-and r.trees on site (units kg/ha).
-
-#### Runtime
-
-Simulations on a 16 core virtual computer (each core 2.10 GHz) with 10
-replicates running in parallel took about 3 hours 15 minutes. Processing
-the results including allometry for selected cohorts took about an hour.
-The model runs with a yearly timestep using monthly climate.
-
-### Outputs
+Biomass of all cohorts of the species on site including the regeneration and even younger regeneration (units kg/ha).
+
+r.bio.sum Biomass of the regeneration cohorts corresponding to the r. ba and r.trees on site (units kg/ha).
 
@@
 
-Aber J. D. & Federer C. A. (1992). A generalized, lumped-parameter model
-of photosynthesis, evapotranspiration and net primary production in
-temperate and boreal forest ecosystems. Oecologia 92(4): 463-474. Aber
-J.D., Ollinger S.V., Federer C.A., Reich P.B., Goulden M.L., Kicklighter
-D.W., Melillo J.M. & Lathrop R.G. Jr. (1995) Predicting the effects of
-climate change on water yield and forest production in the northeastern
-United States. Climate Research 5(3): 207--22. de Bruijn A., Gustafson
-E.J. , Sturtevant B.R., Foster J.R., Miranda B.R., Lichti N.I. & Jacobs
-D.F. (2014) Toward more robust projections of forest landscape dynamics
-under novel environmental conditions: Embedding PnET within LANDIS-II.
-Ecological Modelling (287): 44--57 EEA (2019) Trends in atmospheric
-concentrations of CO2 (ppm), CH4 (ppb) and N2O (ppb), between 1800 and
-2017 available at
-<https://www.eea.europa.eu/ds_resolveuid/217c026ca03649398aadf39d87623e57>
-Gustafson, E. J., A. M. G. De Bruijn, R. E. Pangle, J.-M. Limousin, N.
-G. McDowell, W. T. Pockman, B. R. Sturtevant, J. D. Muss, and M. E.
-Kubiske. 2015. Integrating ecophysiology and forest landscape models to
-improve projections of drought effects under climate change. Global
-Change Biology 21:843-856. Mina, M., Messier, C., Duveneck, M., Fortin,
-M. J., & Aquilué, N. 2021. Network analysis can guide resilience-based
-management in forest landscapes under global change. Ecological
-Applications, 31(1), 1--18. Petter G., Mairota P., Albrich K., Bebi P.,
-Brůna J., Bugmann H., Haffenden A., Scheller R.M., Schmatz D.R., Seidl
-R., Speich M., Vacchiano G. & Lischke H. (2020) How robust are future
-projections of forest landscape dynamics? Insights from a systematic
-comparison of four forest landscape models. Environmental Modelling &
-Software 134: 104844. <https://doi.org/10.1016/j.envsoft.2020.104844>
-Puletti N., Mura M., Castaldi C., Marchi N., Chiavetta U. and Scotti R.
-(2014). ForIT: Functions from the 2nd Italian Forest Inventory (INFC). R
-package version 1.0. <https://CRAN.R-project.org/package=ForIT>
-Scheller, R. M., Domingo J. B. , Sturtevant B. R. , Williams J. S. ,
-Rudy A. , Gustafson E. J. & Mladenoff D. J. (2007) Design, development,
-and application of LANDIS-II, a spatial landscape simulation model with
-flexible temporal and spatial resolution. Ecological Modelling 201(3-4):
-409--419. <https://doi.org/10.1016/j.ecolmodel.2006.10.009> Tsubo M. &
-Walker S. (2005) Relationships between photosynthetically active
-radiation and clearness index at Bloemfontein, South Africa. Theoretical
-and Applied Climatology 80: 17--25.
+Aber J.
+D.
+& Federer C.
+A.
+(1992).
+A generalized, lumped-parameter model of photosynthesis, evapotranspiration and net primary production in temperate and boreal forest ecosystems.
+Oecologia 92(4): 463-474.
+
+Aber J.D., Ollinger S.V., Federer C.A., Reich P.B., Goulden M.L., Kicklighter D.W., Melillo J.M.
+& Lathrop R.G.
+Jr.
+(1995) Predicting the effects of climate change on water yield and forest production in the northeastern United States.
+Climate Research 5(3): 207--22.
+
+de Bruijn A., Gustafson E.J.
+, Sturtevant B.R., Foster J.R., Miranda B.R., Lichti N.I.
+& Jacobs D.F.
+(2014) Toward more robust projections of forest landscape dynamics under novel environmental conditions: Embedding PnET within LANDIS-II.
+Ecological Modelling (287): 44--57
+
+EEA (2019) Trends in atmospheric concentrations of CO2 (ppm), CH4 (ppb) and N2O (ppb), between 1800 and 2017 available at <https://www.eea.europa.eu/ds_resolveuid/217c026ca03649398aadf39d87623e57>
+
+Gustafson, E.
+J., A.
+M.
+G.
+De Bruijn, R.
+E.
+Pangle, J.-M.
+Limousin, N.
+G.
+McDowell, W.
+T.
+Pockman, B.
+R.
+Sturtevant, J.
+D.
+Muss, and M.
+E.
+Kubiske.
+2015.
+Integrating ecophysiology and forest landscape models to improve projections of drought effects under climate change.
+Global Change Biology 21:843-856.
+
+Mina, M., Messier, C., Duveneck, M., Fortin, M.
+J., & Aquilué, N.
+2021.
+Network analysis can guide resilience-based management in forest landscapes under global change.
+Ecological Applications, 31(1), 1--18.
+
+Petter G., Mairota P., Albrich K., Bebi P., Brůna J., Bugmann H., Haffenden A., Scheller R.M., Schmatz D.R., Seidl R., Speich M., Vacchiano G.
+& Lischke H.
+(2020) How robust are future projections of forest landscape dynamics?
+Insights from a systematic comparison of four forest landscape models.
+Environmental Modelling & Software 134: 104844.
+<https://doi.org/10.1016/j.envsoft.2020.104844>
+
+Puletti N., Mura M., Castaldi C., Marchi N., Chiavetta U.
+and Scotti R.
+(2014).
+ForIT: Functions from the 2nd Italian Forest Inventory (INFC).
+R package version 1.0.
+<https://CRAN.R-project.org/package=ForIT>
+
+Scheller, R.
+M., Domingo J.
+B.
+, Sturtevant B.
+R.
+, Williams J.
+S.
+, Rudy A.
+, Gustafson E.
+J.
+& Mladenoff D.
+J.
+(2007) Design, development, and application of LANDIS-II, a spatial landscape simulation model with flexible temporal and spatial resolution.
+Ecological Modelling 201(3-4): 409--419.
+<https://doi.org/10.1016/j.ecolmodel.2006.10.009>
+
+Tsubo M.
+& Walker S.
+(2005) Relationships between photosynthetically active radiation and clearness index at Bloemfontein, South Africa.
+Theoretical and Applied Climatology 80: 17--25.
 <https://doi.org/10.1007/s00704-004-0080-5>

### 2023-01-24 manuscript/supplementaryRW.Rmd
updates report from aDGVM2 in supplementary

@@
 
+Simon Scheiter
+Senckenberg Biodiversity and Climate Research Centre, Senckenberganlage 25, 60325 Frankfurt am Main, Germany
+
 ### Model
 
-*Please provide here the references to this model where it is possible
-to understand the model structure specially for the regeneration module,
-this could be a published paper, a website with the model description, a
-manual or some other document that the reader can access.*
+We used aDGVM2, an individual-based dynamic vegetation model that is based on concepts from community assembly theory and uses a functional trait approach (Langan et al., 2017; Scheiter et al., 2013).
+The model simulates growth, reproduction, and mortality of individual plants while keeping track of state variables, such as biomass, height, and leaf area.
+Each plant in aDGVM2 is characterized by a plant-specific set of trait values.
+Traits describe growth form, leaf characteristics, hydraulic characteristics, resource allocation, architecture, reproduction, mortality, and response to disturbance.
+Most plant traits are linked by trade-offs to constrain possible trait combinations.
+Selection and trait inheritance assemble plant communities that are adapted to biotic and abiotic conditions.
+Plants with trait combinations that allow sufficient growth and reproduction rates, and that allow plants to cope with competition and disturbances can contribute their trait values to the community trait pool.
+Trait mutation and recombination may alter trait values in the community trait pool.
+Randomly drawn seeds from the trait pool are added to the plant population and new plants start growing from seeds.
+Plants that are not adapted to the prevailing disturbance regimes, biotic and abiotic conditions, or that do not allocate enough carbon to reproduction disappear from the population.
+Therefore, successful ecological strategies emerge dynamically from these community assembly and reproduction process.
+Due to the trait-based approach, species or plant functional types (PFTs) are not pre-defined and hard-coded in aDGVM2.
+Rather, different PFTs or ecological strategies that are adjusted to the prevailing biotic and abiotic conditions emerge from community assembly processes (Scheiter et al., 2013), and plants can be classified into PFTs or ecological strategies based on their trait values in a post-processing step.
+The aDGVM2 has primarily been developed for tropical and sub-tropical ecosystems and it has not been tested and benchmarked for Europe.
+We applied the model in the original version.
+Data-model comparisons with model results were not done and the model was not changed to improve agreement with any data for Europe.
 
@@
 
+The aDGVM2 simulates vegetation at daily temporal resolution. We therefore used daily climate input data for the period 1981-2018 according to the modeling protocol.
+
 ### Soil
 
+Soil quality data was converted to maximum rooting depth.
+Specifically, the soil quality value multiplied by 2 and the result was used as maximum rooting depth.
+We used the multiplication, as we typically use maximum soil depth and maximum rooting depth of 10m.
+Previous model simulations showed that rooting depth and soil dept have strong impacts on vegetation dynamics by influencing water availability to plants (Langan et al., 2017).
+We therefore considered it as adequate representation of soil quality.
+Other soil characteristics including volumetric water-holding capacity, soil hydraulic conductivity, soil bulk density, soil depth, soil texture, soil carbon content, soil wilting point and field capacity were taken from the FAO (http://www.fao.org, Nachtergaele et al., 2009).
+
 ### Topography
 
+We used elevation according to the modeling protocol.
+In aDGVM2, elevation is used to calculate atmospheric pressure which in turn influences ecophysiological processes related to water and carbon fluxes at leaf and canopy level.
+Information on slope and aspect was not used, as related processes (e.g., impacts on runoff or radiation balance) are not implemented in aDGVM2.
+
 ### Tree species
 
+The aDGVM2 does not simulate pre-defined species or PFTs.
+The 11 tree species included in the modeling protocol cannot be implemented and parametrized with reasonable effort.
+We therefore classified simulated trees into different ecological strategies as a post-processing step.
+We used all combinations of evergreen/deciduous, light/water triggered phenology, low/high SLA.
+Evergreen/deciduous phenology and light/water triggered phenology are traits implemented in aDGVM2.
+The threshold to distinguish high and low SLA was defined as 20 m2/kg.
+We used these traits as previous simulation experiments showed that patterns of these traits emerge along environmental gradients while patterns of other traits implemented in aDGVM2 are often less clear.
+Ecological strategies considered are: low SLA, deciduous, rain-triggered; low SLA, deciduous, light-triggered; low SLA, evergreen, rain-triggered; low SLA, evergreen, light-triggered; high SLA, deciduous, rain-triggered; high SLA, deciduous, light-triggered; high SLA, evergreen, rain-triggered; high SLA, evergreen, light-triggered.
+
 ### Simulations
 
+Simulations followed the modeling protocol.
+Simulations were conducted for all sites using the provided daily climate and soil data.
+By default, aDGVM2 simulates 1ha stands.
+To implement the modeling protocol, adjustments to scale model results to 1 ha level were therefore not required.
+The model was initialized by default initialization routines, i.e., traits of individual plants were drawn randomly from a pre-defined range of values for each trait.
+Ranges of trait values are provided in Langan et al.
+(2017).
+The model was initialized with bare ground and 2880 tree seedlings with 100g each.
+The aDGVM2 simulates both trees and grasses.
+As the simulation experiment aims at tree/forest regeneration, simulations were conducted only with trees while grasses were removed.
+
+We conducted 5 replicates for each of the 200 sites to account for stochastic processes in aDGVM2.
+Stochastic processes include random initialization of trait values of initial plant populations, demographic processes such as plant mortality or selection of seeds from seed bank, or the community assembly processes including mutation and cross-over.
+Climate time series used for different replicates were identical such that variation between replicates can be attributed by stochastic processes in aDGVM2.
+
+Although aDGVM2 includes routines to simulate fire, simulations were conducted in the absence of fire. Management was not simulated.
+
+### Dispersal 
+The aDGVM2 does not simulate seed dispersal.
+
+### Simulation length
+
+We first conducted a 600-year spin-up with a randomized order of years 1981-2018 provided in the CHELSA daily climate data.
+Spin-up is required to ensure that both state variables such as biomass and tree cover and community trait composition have stabilized and are in a dynamic equilibrium with climate.
+Previous model simulations showed that a 600-year spin-up is sufficient to reach such an equilibrium state (Langan et al., 2017).
+Following the spin-up, we simulated transient vegetation dynamics with daily CHELSA climate data for the period 1981-2018.
+
 ### Outputs
+
+To create output variables, we randomly selected 10-year periods within the transient phase (i.e., period 1981-2018 after spin-up phase) and calculated averages of all required output variables following the modeling protocol.
+The selected 10-year periods differed between sites and replicates.
+
+The aDGVM2 does not simulate pre-defined species or PFTs and trees were classified into different ecological strategies (see section Tree species).
+Codes for the ecological strategies in the species column of the output files are:
+1 - low SLA, deciduous, rain-triggered
+2 - low SLA, deciduous, light-triggered
+3 - low SLA, evergreen, rain-triggered
+4 - low SLA, evergreen, light-triggered
+5 - high SLA, deciduous, rain-triggered
+6 - high SLA, deciduous, light-triggered
+7 - high SLA, evergreen, rain-triggered
+8 - high SLA, evergreen, light-triggered
+
+Simulations were conducted for all 200 sites. We conducted 5 replicates per site to account for stochastic processes in aDGVM. In the output files, replicates are given as follows: 
+Replicate 1 – sample 1:200
+Replicate 2 – sample 201:400
+Replicate 3 – sample 401:600
+Replicate 4 – sample 601:800
+Replicate 5 – sample 801:1000
+
+### References
+
+Langan L; Higgins SI; Scheiter S (2017) Climate-biomes, pedo-biomes or pyro-biomes: which world view explains the tropical forest - savanna boundary in South America?
+JOURNAL OF BIOGEOGRAPHY, 44, 2319-2330.
+
+Nachtergaele, F., van Velthuizen, H., Verelst, L., Batjes, N., Dijk- shoorn, K., Van Engelen, V., Fischer, G., Jones, A., Montanarella, L., and Petri, M.: Harmonized world soil database (version 1.1), FAO, Rome, Italy, IIASA, Laxenburg, Austria, 2009.
+
+Scheiter S, Langan L, Higgins SI (2013) Next generation dynamic global vegetation models: learning from community ecology. NEW PHYTOLOGIST, 198, 957-969.
 

### 2023-01-24 manuscript/supplementaryRW.Rmd
updates report from ForCEEPS in supplementary

@@
 
-Xavier Morin
+Xavier Morin & François de Coligny
 
@@
 
-*Please provide here the references to this model where it is possible
-to understand the model structure specially for the regeneration module,
-this could be a published paper, a website with the model description, a
-manual or some other document that the reader can access.*
+Morin X., de Coligny F., Bugmann H., Limousin J.-M., Ourcival J.-M., Martin-StPaul N., Simioni G., Cailleret M., Prevosto B., Toïgo M., Vennetier M., Cateau E., Guillemot J.
+2021.
+Beyond forest succession: a gap model to study ecosystem functioning and tree community composition under climate change.
+Functional Ecology.
+35, 955-975.
+
+Jourdan M., Dreyfus, P., Riond C., Cordonnier T., Cornet B., de Coligny F., Morin X.
+2021.
+Managing mixed stands can mitigate severe climate change impacts on French alpine forests.
+Regional Environmental Change.
+21, 78.
+
+Morin X., Damestoy T., Toïgo M., Jactel H., Castagneyrol B.
+Meredieu C.
+2020.
+Using forest gap models and experimental data to explore long-term effects of tree diversity on the productivity of mixed planted forests.
+Annals of Forest Science.
+77 https://doi.org/10.1007/s13595-020-00954-0.
 
@@
 
-ForCEEPS requires monthly data, I have thus just randomized the years
-given by CHELSA database to obotain a 4000-yr long time-series for each
-site, with mean monthly temperature and monthly sum of precipitations.
+ForCEEPS requires monthly data.
+A 4000-yr long time-series was obtained for each site by randomizing available years from the CHELSA database, with mean monthly temperature and monthly sum of precipitations (see 1.2.2).
 
@@
 
-Relying on given information, I used the following relationship for
-assessing Soil Field Capacity (SFC) : SFC = 8+(X-1)\*(25-8)/(5-1), in
-which X is the 'soil_qual' value in the 'soil_quality_200.csv' file.
+Soil quality is represented by a continuous value between 1 and 5, as provided (see 1.2.3) Soil Field Capacity (SFC) was assessed as follows:
+SFC = 8+(X-1)*(25-8)/(5-1), in which X is the ‘soil_qual’ value in the provided ‘soil_quality_200.csv’ file.
 
@@
 
-I have run two kinds of simulations : - Simulations with seed rain from
-a potential species list, thus independent from the actual forest
-composition ("Forclim-like"), for both thresholds (7 and 10cm) -
-Simulations with seed rain depending on the actual forest composition
-for both thresholds (7 and 10cm) In each site, the simulation consists
-of 4000 yr-long simulation of 1ha of forest (ie. 10 patches of 1000 m²),
-starting from bare-ground. There was no dispersal between patches, but
-the effect of the abundance of adult trees on the composition of
-regeneration was considered in half of the simulations as mentioned.
+In each site, the simulation consists of 4000 yr-long simulation of 1ha of forest (ie. 10 patches of 1000 m²), starting from bare-ground. There was no dispersal between patches.
+Two kinds of simulations have been carried out in each of the 200 sites. 
+In a first set of simulations, regeneration dynamics was simulated with a seed rain from a potential species list, thus independent from the actual forest composition, as classically done in gap models.
+Under this design, a number of seedling is randomly chosen for each patch, and a species identity is randomly sorted from a potential species list (defined at the start of simulation) assigned to each seedling.
+Then each seedling tries to colonize the site depending on whether the suitability between species requirements and site conditions (climate, light, soil quality, browsing index).
+In a second set of simulations, regeneration dynamics was simulated with a seed rain whose composition depended on the actual forest composition.
+Under this design, a number of seedling is randomly chosen for each patch, but the species identity is assigned according to species relative abundance (calculated across all 10 patches according to species biomass).
+For instance, if at year i the adult trees of species j represent 60% of the total biomass across the 10 patches, then 60% of the seedlings that try to colonize the patches at year i will belong to species j.
+Then, similarly to the first design, each seedling tries to colonize the site depending on whether the suitability between species requirements and site conditions.
+In the two sets of simulations, simulations have been run for both thresholds for regeneration (7 and 10 cm).
 
@@
 
-As requested, I have output species specific regeneration rates per
-decade and per ha. More precisely, I have considered only the last 2000
-years of simulations (= to ensure that pseudo-equilibrium state is
-reached). I have exported the results every 10 years for the last 2000
-years at the 1ha-level. My outputs thus rely on temporal variability.
-Variables: - Site: numbers corresponding to the climate data file -
-Sample: one per decade, for a total of 2000 years = identified from 1 to
-200 - Species: name specified as requested (eg. 'fasy') - r.trees:
-number of trees outcrossing the threshold (7 or 10 cm) during the
-decade, for the 10 patches simulated (= 1ha), per species - r.ba: summed
-basal area of the r.trees, per species - ba: summed basal area of all
-trees per species at the end of the decade - dbh: threshold used And
-there are thus 4 output files: Results_Forceeps_7cm.csv
-Results_Forceeps_10cm.csv Results_Forceeps_feedback_7cm.csv (=
-simulations with feedback of the actual species composition on the seed
-rain) Results_Forceeps_feedback 10cm.csv (= simulations with feedback of
-the actual species composition on the seed rain)
+Outputs consisted in species specific regeneration rates per decade and per ha, by considering the last 2000 years of simulations to ensure that pseudo-equilibrium state is reached.
+Results were aggregated for every 10 years for the last 2000 years at the 1ha-level.
+Variables:
+-	Site: numbers corresponding to the climate data file
+-	Sample: one per decade, for a total of 2000 years = identified from 1 to 200
+-	Species: name specified as requested (eg. ‘fasy’)
+-	r.trees: number of trees outcrossing the threshold (7 or 10 cm) during the decade, for the 10 patches simulated (= 1ha), per species
+-	r.ba: summed basal area of the r.trees, per species
+-	ba: summed basal area of all trees per species at the end of the decade
+-	dbh: threshold used
 

### 2023-01-24 code/dataPreparation.R
add Landis correction in data processing by Josef

@@ -263,6 +263,59 @@
          dplyr::ungroup()
      landis$type <- NULL
      
+     
+     # Corrections done by Josef----
+     emp_site <- unique(landis$site)
+     emp_sample <- unique(landis$sample)
+     emp_dbh <- c(7, 10)
+     emp_species <- unique(landis$species)
+     require(utils)
+     lanempty <- data.frame(expand.grid(emp_site, emp_sample, emp_dbh, emp_species))
+     names(lanempty) <- (c("site", "sample","dbh", "species"))
+     lanempty$r.trees <- 0
+     lanempty$r.ba <- 0
+     lanempty$ba <- 0
+     lanempty$model <- "Landis II"
+     lanempty$ID <- paste0(lanempty$site,"_", lanempty$sample,"_", 
+                           lanempty$dbh,"_", lanempty$species)
+     
+     
+     landis$ID <- paste0(landis$site,"_", landis$sample,"_", 
+                         landis$dbh,"_", landis$species)
+     lanempty$ba <- landis$ba[match(lanempty$ID, landis$ID)]
+     lanempty$r.trees <- landis$r.trees[match(lanempty$ID, landis$ID)]
+     lanempty$r.ba <- landis$r.ba[match(lanempty$ID, landis$ID)]
+     
+     
+     #### check consistency ####
+     lanempty$r.ba[is.na(lanempty$r.ba)] <- 0
+     lanempty$r.trees[is.na(lanempty$r.trees)] <- 0
+     
+     lanempty$ba[is.na(lanempty$ba)] <- 0
+     summary(lanempty$ba)
+     summary(landis$ba)
+     sum(lanempty$ba)
+     sum(landis$ba)
+     
+     summary(lanempty$r.ba)
+     summary(landis$r.ba)
+     sum(lanempty$r.ba)
+     sum(landis$r.ba)
+     
+     summary(lanempty$r.trees)
+     summary(landis$r.trees)
+     sum(lanempty$r.trees)
+     sum(landis$r.trees)
+     
+     
+     
+     landis <- rbind(landis,  lanempty) 
+ 
+     
+     
+     
+     
+     
      #Correct that there are certain samples from certain sites only present in 
      #one dbh threshold but not in the other
      #add them assuming 0 recruitment for those sites and samples 

### 2023-01-24 code/dataPreparation.R
deletes all the Landis postprocessing code, they have submitted a corrected data version

@@ -182,237 +182,13 @@
 
 
     ### Landis II ----
-    landisSPP <- data.table::fread("data/simulation_results/landisII/landis_biomass_sum/landis_biomass_sum_site_species_sample.csv")
-    #landis_biomass_sum_site_species_sample.csv: species biomass per site and sample
-    # - all cohorts including regeneration - 142 425 -
-    # this should represent species present, the rest should be zero,
-    # the species were not there at the sampling time
-
-    landisSITE <- data.table::fread("data/simulation_results/landisII/landis_biomass_sum/landis_biomass_sum_site_sample.csv")
-    #landis_biomass_sum_site_sample: total biomass per site and sample -
-    #this has 39916 rows, almost the 40 000 required,
-    #the rest should be zeros, I am sure there were sites where nothing can grow
-    #(extreme climate) which will make the 84 totally missing.
-    #You can once again do your biomass->ba and the magic where you select
-    #the higher biomass for the ones where we have only regeneration.
-
-
-    landis <- data.table::fread("data/simulation_results/landisII/LANDIS-II_data_version_03.csv")
-    landis$model <- "Landis II"
-    landis$dbh[landis$dbh == "7 cm" ] <- 7
-    landis$dbh[landis$dbh == "10 cm" ] <- 10
-    landis$dbh[landis$dbh == "0" ] <- 0
-    landis$dbh <- as.numeric(landis$dbh)
-    landis$V1 <- NULL
-
-    landis$otherBA <- landis$sum.bio - landis$r.bio.sum 
     
-    #Calculate the BA from (sum.bio - r.bio.sum)
-          #sum.bio: Above ground living biomass of all cohorts 
-          #   of the species on site including the regeneration 
-          #   and even younger regeneration (units kg/ha). 
-          #r.bio.sum: Above ground living biomass of the 
-          #  regeneration cohorts corresponding to the r. ba 
-          #  and r.trees on site (units kg/ha).  
-
-    landis$ba <- ((landis$otherBA * 0.001) / 12.5) + landis$r.ba
-          #1) transform kg into tonnes / ha
-          #2) expansion factor between  10 to 15 (12.5) to transform BA m2/ha
-          #3) Add the r.ba because it is calculated based on the
-          #   allometries for regeneration (there is not allometries 
[…]

### 2023-01-24 manuscript/manuscriptRW.Rmd
added changes based on the comments in abstract and beginning of introduction

@@
 Tree establishment is a key process in forest dynamics, particularly in
-the context of forest resilience after disturbance and climate change.
-Models of forest dynamics have been in use since more than 50 years, but
-an assessment of their capacity to represent tree establishment is
-lacking. We assess how well current models are capturing the overall
-abundance, species composition and mortality of tree establishment.
+the context of forest resilience and climate change. Models of forest
+dynamics have been in use for more than 50 years, but an assessment of
+their capacity to represent tree establishment is lacking. We assess how
+well current models capture the overall abundance, species composition
+and mortality of tree establishment.
 
@@
 environmental gradients across Europe. The results are benchmarked
-against comprehensive data from forest reserves.
+against comprehensive data from unmanaged forests.
 
@@
 compensated by high simulated mortality rates in the early stages of
-population dynamics. In the models, the effect of light availability on
-ingrowth rates is captured better than the effect of temperature and
-soil moisture, but the patterns are not consistent across models.
-Simulated species diversity matches the observed ranges quite well.
-Models simulating higher species diversity at the stand level do not
-feature higher ingrowth diversity. The ingrowth of *Abies alba* is
-captured across temperature and soil moisture gradients, whereas *Fagus
+population dynamics. The effect of light availability on ingrowth rates
+is captured better than the effect of temperature and soil moisture, but
+patterns are not consistent across models. Simulated species diversity
+of ingrowth matches the observed ranges quite well. Models simulating
+higher species diversity at the stand level do not feature higher
+ingrowth diversity. The ingrowth of *Abies alba* is captured across
+temperature and soil moisture gradients, whereas ingrowth of *Fagus
 sylvatica* is generally underestimated. *Picea abies* ingrowth is
@@
 
-Even though the overestimation of tree establishment represent a
+Even though the overestimation of tree establishment represents a
 challenge for the usefulness of model projections under uncertain
-conditions such a climate change, and although some formulations should
+conditions such as climate change, and although some formulations should
 be reconsidered to correct excessive ingrowth levels, it is remarkable
@@
 complex process of tree establishment, while not having been
-parameterized with such a data. We conclude that in spite of all the
-uncertainties, many models hold promise for assessing long-term forest
-dynamics under changing environmental conditions.
+parameterized with such data. We conclude that in spite of all the
+uncertainties, a lot can be gained by a focus on the modeling of
+establishment processes in these model, this will make their long-term
+forest dynamics projections under changing environmental conditions much
+more robust.
 
@@
 Forests provide a wide range of products and services of vital
-importance to humankind. Under the pressure of climate change and
-increasing disturbances, it is becoming ever more important to
-understand how forest structure, composition and function will change,
-and to evaluate forest capacity to adapt or be resilient after
-disturbance [@lindner2010]. A wide range of models of forest dynamics
-were developed over the past decades considering the impacts of climate
-on forest dynamics [@bugmann2022]. From these studies, it is evident
-that while we have a reasonably good understanding how to model tree
-growth [@vanclay1997; @bugmann1996], and substantial efforts have been
-dedicated to improving the depiction of tree mortality [@bugmann2019],
-tree establishment remains a much less studied process that is often
-represented rather coarsely in these models, which is a clear research
-gap in the context of climate-induced forest disturbances and forest
+importance to humankind [@foresteurope2020]. Under the pressure of
+climate change and increasing disturbances, it is becoming ever more
+important to understand how forest structure, composition and function
+will change, and to evaluate forest capacity to adapt or be resilient
+after disturbance [@lindner2010; @seidl2022]. A wide range of models of
+forest dynamics were developed over the past decades considering the
+impacts of climate on forest dynamics [@bugmann2022]. From these
+studies, it is evident that while we have a reasonably good
+understanding how to model tree growth [@vanclay1997; @bugmann1996], and
+substantial efforts have been dedicated to improving the depiction of
+tree mortality [@bugmann2019]. However, tree establishment remains a
+much less studied process that is often represented rather coarsely in
+these models [@walck2011; @leishman1992], which is a clear research gap
+in the context of climate-induced forest disturbances and forest
 resilience.
@@
 The species diversity of the entire stand was better captured by the
-models than the diversity of
-tree ingrowth, and only a few models overpredicted stand-level
-diversity. Defining ingrowth composition as being proportional to adult
-abundance, regardless of productivity, might be a
-simple and conservative assumption to safely maintain relative species
-abundances [@hanbury-brown2022], although this may be simplistic.
-Furthermore, based on the simulation results there is no evidence that
-models with feedback captured better the ingrowth species diversity than
-those without feedback. The similar performance of models with and
-without feedback is likely because the models are giving more weight to
-the regeneration niche arising abiotic and biotic filters than from the
-habitat niche of the adult trees [@grubb1977].
+models than the diversity of tree ingrowth, and only a few models
+overpredicted stand-level diversity. Defining ingrowth composition as
+being proportional to adult abundance, regardless of productivity, might
+be a simple and conservative assumption to safely maintain relative
+species abundances [@hanbury-brown2022], although this may be
+simplistic. Furthermore, based on the simulation results there is no
+evidence that models with feedback captured better the ingrowth species
+diversity than those without feedback. The similar performance of models
+with and without feedback is likely because the models are giving more
+weight to the regeneration niche arising abiotic and biotic filters than
+from the habitat niche of the adult trees [@grubb1977].
 

### 2023-01-25 manuscript/manuscriptRW.Rmd
review of introduction based on co-authors feedback

@@
 fruit maturation, seed production, dispersal, germination, juvenile
-growth and survival [@price2001]. All these processes are difficult to
-assess, and some of them are scarcely understood and thus appear highly
-stochastic [@bogdziewicz2021]. Identifying the appropriate level of
-complexity for the formulation of the key factors that are leading to
-successful tree establishment is challenging. Currently, tree
-establishment processes in dynamic forest models are handled in a
-multitude of ways [@konig_tree_2022; @bugmann2022], from entirely
-ignoring it [as done in classical forest growth models, e.g., Pretzsch
-et al. -@pretzsch2002], across the use of a few simple environmental
-filters, as done in most forest "gap" models [@shuman2011] and Dynamic
-Global Vegetation Models [e.g., @hickler2012; @smith2001], to complex
-approaches that incorporate local feedback from the canopy, multiple
-ecological processes and often also small time steps [e.g., @seidl2012;
-@wehrli2006], or field-based statistical parameterizations, which
-however are not easy to extrapolate in space and time [e.g.,
-@ribbens1994].
+growth and survival [@price2001; @vacchiano2018]. All these processes
+are difficult to assess, and some of them are scarcely understood and
+thus appear highly stochastic [@bogdziewicz2021]. Identifying the
+appropriate level of complexity for the formulation of the key factors
+that are leading to successful tree establishment is challenging.
+Currently, tree establishment processes in dynamic forest models are
+handled in a multitude of ways [@konig_tree_2022; @bugmann2022]: a) from
+entirely ignoring it [as done in classical forest growth models, e.g.,
+Pretzsch et al. -@pretzsch2002], b) across the use of a few simple
+environmental filters, as done in most forest "gap" models [@shuman2011]
+and Dynamic Global Vegetation Models [e.g., @hickler2012; @smith2001],
+c) to complex approaches that incorporate local feedback from the
+canopy, multiple ecological processes and often also small time steps
+[e.g., @seidl2012; @wehrli2006], or d) field-based statistical
+parameterizations, which however are not easy to extrapolate in space
+and time [e.g., @ribbens1994].
 
 Overall, models are needed to (1) synthesize existing empirical data and
-explore their consequences, and (2) assess future tree establishment
-e.g., in the context of global change scenarios. Given the current
-strategies that are used in models of forest dynamics to represent tree
-establishment, model behavior is often facing problems that can have
-strong impacts on simulated forest dynamics, such as very high levels of
-tree establishment, which then necessitate excess mortality at early
-stages of tree life to arrive at correct stand structure and
-composition. Also, inadequately high species diversity in tree
-establishment may be simulated. This is characteristic of "classical"
-forest gap models [@gutiérrez2016], at least as long as the simulations
-comprise a multitude of species. Other models use calibration against
-local canopy-level data to define appropriate ingrowth levels, which is
-likely to strongly hamper the general applicability of these models
-e.g., under scenarios of climate change. Furthermore, correctly
-capturing the species composition in tree establishment as a function of
-the presence of seed trees in the canopy is often a particular
-challenge, potentially leading to unrealistic successional drift in the
-model, which must be corrected by factors that are hard to parameterize
-[e.g., @lischke2006]. A related issue is the excessive reduction of
-species richness due to positive feedback effects such that eventually
-just single-species stands remain. This is sometimes corrected by the
-incorporation of a low level of seed influx of all species at all times
-[@lischke2006]. However, simulated species composition is usually
-exceedingly sensitive to that level, whereas the parameters of such
-functions are poorly constrained by field data. Lastly, there is often a
-problem with insufficient observational constraint on parameter values
-for models that start from very small tree sizes (e.g., 10 cm of height
-or even seeds) and track tree development in a process-oriented manner
-by considering a multitude of ecological influences, rather than
-emphasizing tree establishment into a larger size class [e.g.,
-@wehrli2006].
+explore their consequences, (2) assess future tree establishment e.g.,
+in the context of global change scenarios and, (3) identify most
+important processes needed to define relevant ecological processes.
+Given the current strategies that are used in models of forest dynamics
+to represent tree establishment, model behavior often faces problems
+that can have strong impacts on simulated forest dynamics, such as very
+high levels of tree establishment, which then necessitate excess
+mortality at early stages of tree life to simulate correct stand
+structure and composition. Also, inadequately high species diversity in
+tree establishment may be simulated. This is characteristic of
+"classical" forest gap models [@gutiérrez2016], at least as long as the
+simulations comprise a multitude of species. Other models use
+calibration against local canopy-level data to define appropriate
+ingrowth levels, which is likely to strongly hamper the general
+applicability of these models e.g., under scenarios of climate change.
+Furthermore, correctly capturing the species composition in tree
+establishment as a function of the presence of seed trees in the canopy
+is often a particular challenge, potentially leading to unrealistic
+successional drift in the model, which must be corrected by factors that
+are hard to parameterize [e.g., @lischke2006]. A related issue is the
+excessive reduction of species richness due to positive feedback effects
+such that eventually just single-species stands remain. This is
+sometimes corrected by the incorporation of a low level of seed influx
+of all species at all times [@schumacher2006] or by restricting the
+number of seeds per species in the seedbank [@lischke2006]. However,
+simulated species composition is usually exceedingly sensitive to that
+level, whereas the parameters of such functions are poorly constrained
+by field data. Lastly, there is often a problem with insufficient
+observational constraints on parameter values for models that start from
+very small tree sizes (e.g., 10 cm of height or even seeds) and track
+tree development in a process-oriented manner by considering a multitude
+of ecological influences, rather than emphasizing tree establishment
+into a larger size class [e.g., @wehrli2006].
 
 Thus, a focus on the modeling of tree establishment processes is sorely
-needed and overdue [@price2001] if we are to make reliable projections
-of future forest dynamics, i.e., when the models need to be operated in
-extrapolation mode, but also from a fundamental ecological point of view
-(increasing systems understanding). Also in the field of general
+needed and overdue [@price2001; @walck2011] if we are to make reliable
+projections of future forest dynamics, i.e., when the models need to be
+operated in extrapolation mode, but also from a fundamental ecological
+point of view increasing systems understanding. Also in the field-based
 ecology, tree establishment has not received a lot of attention
@@
 suite of species. Thus, the available data are fragmented, which
-constitutes a major problem for model building. On the one hand,
-monitoring on permanent sample plots (such as National Forest
-Inventories) often measure tree establishment [@zell2019], however with
+constitutes a major problem for model building [@clark1999]. On the one
+hand, monitoring on permanent sample plots (such as National Forest
+Inventories) often measures tree establishment [@zell2019], however with
 a design that captures the rate after the stand initiation phase
 [@hallsby2015]. Data from permanent plots, although available across
-regions, are highly heterogeneous, rendering their utility difficult in
-a modeling context [@käber2021]. On the other hand, targeted studies to
+regions, are highly heterogeneous, rendering their use difficult in a
+modeling context [@käber2021]. On the other hand, targeted studies to
 measure tree establishment on experimental sites are often limited in
@@
 evaluated against a continental-scale, multi-species harmonized data set
-on tree establishment [@käber2022]. By tree establishment, we refer to
-the passing of trees across a specific diameter threshold ("ingrowth")
-and thus as successful tree establishment. We included models that are
-based on a range of "philosophies", from models operating at the stand
-to the global scale as well as the range of models from empirically
-derived to "process-based" [e.g., @bugmann1996; @fabrika2012;
-@mette2014; @lexer2001]. Because of the large variability in tree
-establishment patterns in nature and the large number of factors driving
-this process --- including some that are not incorporated in most models
-such as deer browsing --- we do not aim at a detailed statistical
-evaluation of each model. Instead, we aim at evaluating the general
-ingrowth patterns and magnitudes simulated by the models, and benchmark
-the simulated regeneration niche of multiple species against empirical
-data along a wide environmental gradients of temperature, moisture, and
-light availability.
+on tree establishment [@käber2022]. This is a challenging experiment for
+these models as they do also have to estimate the forest stand above the
+tree establishment. By tree establishment, we refer to the passing of
+trees across a specific diameter threshold ("ingrowth"). We included
+models that are based on a range of "philosophies", from models
+operating at the stand to the global scale as well as the range of
+models from empirically derived to "process-based" [e.g., @bugmann1996;
+@fabrika2012; @mette2014; @lexer2001]. Because of the large variability
+in tree establishment patterns in nature and the large number of factors
+driving this process --- including some that are not incorporated in
+most models such as deer browsing --- we do not aim at a detailed
+statistical evaluation of each model. Instead, we aim at evaluating the
+general ingrowth patterns and magnitudes simulated by the models, and
+benchmark the simulated regeneration niche of multiple species against
+empirical data along a wide environmental gradient of temperature,
+moisture, and light availability.
 
@@
 composition of the tree ingrowth layer? And 3), do they correctly
-represent the mortality in tree establishment?.
+represent the mortality in tree establishment?
 

### 2023-01-27 manuscript/supplementaryRW.Rmd
adds Landis comment in the report about adding 0 values
added LPJ-GUESS report in the supplementary

@@
 - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
-
 ---
@@
 ## 4C
+
 ### Authors
@@
 
-Morin X., de Coligny F., Bugmann H., Limousin J.-M., Ourcival J.-M., Martin-StPaul N., Simioni G., Cailleret M., Prevosto B., Toïgo M., Vennetier M., Cateau E., Guillemot J.
-2021.
-Beyond forest succession: a gap model to study ecosystem functioning and tree community composition under climate change.
-Functional Ecology.
-35, 955-975.
-
-Jourdan M., Dreyfus, P., Riond C., Cordonnier T., Cornet B., de Coligny F., Morin X.
-2021.
-Managing mixed stands can mitigate severe climate change impacts on French alpine forests.
-Regional Environmental Change.
+Morin X., de Coligny F., Bugmann H., Limousin J.-M., Ourcival J.-M.,
+Martin-StPaul N., Simioni G., Cailleret M., Prevosto B., Toïgo M.,
+Vennetier M., Cateau E., Guillemot J. 2021. Beyond forest succession: a
+gap model to study ecosystem functioning and tree community composition
+under climate change. Functional Ecology. 35, 955-975.
+
+Jourdan M., Dreyfus, P., Riond C., Cordonnier T., Cornet B., de Coligny
+F., Morin X. 2021. Managing mixed stands can mitigate severe climate
+change impacts on French alpine forests. Regional Environmental Change.
 21, 78.
 
-Morin X., Damestoy T., Toïgo M., Jactel H., Castagneyrol B.
-Meredieu C.
-2020.
-Using forest gap models and experimental data to explore long-term effects of tree diversity on the productivity of mixed planted forests.
-Annals of Forest Science.
-77 https://doi.org/10.1007/s13595-020-00954-0.
+Morin X., Damestoy T., Toïgo M., Jactel H., Castagneyrol B. Meredieu C.
+2020. Using forest gap models and experimental data to explore long-term
+effects of tree diversity on the productivity of mixed planted forests.
+Annals of Forest Science. 77
+<https://doi.org/10.1007/s13595-020-00954-0>.
 
@@
 
-ForCEEPS requires monthly data.
-A 4000-yr long time-series was obtained for each site by randomizing available years from the CHELSA database, with mean monthly temperature and monthly sum of precipitations (see 1.2.2).
+ForCEEPS requires monthly data. A 4000-yr long time-series was obtained
+for each site by randomizing available years from the CHELSA database,
+with mean monthly temperature and monthly sum of precipitations (see
+1.2.2).
 
@@
 
-Soil quality is represented by a continuous value between 1 and 5, as provided (see 1.2.3) Soil Field Capacity (SFC) was assessed as follows:
-SFC = 8+(X-1)*(25-8)/(5-1), in which X is the ‘soil_qual’ value in the provided ‘soil_quality_200.csv’ file.
+Soil quality is represented by a continuous value between 1 and 5, as
+provided (see 1.2.3) Soil Field Capacity (SFC) was assessed as follows:
+SFC = 8+(X-1)\*(25-8)/(5-1), in which X is the 'soil_qual' value in the
+provided 'soil_quality_200.csv' file.
 
@@
 
-In each site, the simulation consists of 4000 yr-long simulation of 1ha of forest (ie. 10 patches of 1000 m²), starting from bare-ground. There was no dispersal between patches.
-Two kinds of simulations have been carried out in each of the 200 sites. 
-In a first set of simulations, regeneration dynamics was simulated with a seed rain from a potential species list, thus independent from the actual forest composition, as classically done in gap models.
-Under this design, a number of seedling is randomly chosen for each patch, and a species identity is randomly sorted from a potential species list (defined at the start of simulation) assigned to each seedling.
-Then each seedling tries to colonize the site depending on whether the suitability between species requirements and site conditions (climate, light, soil quality, browsing index).
-In a second set of simulations, regeneration dynamics was simulated with a seed rain whose composition depended on the actual forest composition.
-Under this design, a number of seedling is randomly chosen for each patch, but the species identity is assigned according to species relative abundance (calculated across all 10 patches according to species biomass).
-For instance, if at year i the adult trees of species j represent 60% of the total biomass across the 10 patches, then 60% of the seedlings that try to colonize the patches at year i will belong to species j.
-Then, similarly to the first design, each seedling tries to colonize the site depending on whether the suitability between species requirements and site conditions.
-In the two sets of simulations, simulations have been run for both thresholds for regeneration (7 and 10 cm).
+In each site, the simulation consists of 4000 yr-long simulation of 1ha
+of forest (ie. 10 patches of 1000 m²), starting from bare-ground. There
+was no dispersal between patches. Two kinds of simulations have been
+carried out in each of the 200 sites. In a first set of simulations,
+regeneration dynamics was simulated with a seed rain from a potential
+species list, thus independent from the actual forest composition, as
+classically done in gap models. Under this design, a number of seedling
+is randomly chosen for each patch, and a species identity is randomly
+sorted from a potential species list (defined at the start of
+simulation) assigned to each seedling. Then each seedling tries to
+colonize the site depending on whether the suitability between species
+requirements and site conditions (climate, light, soil quality, browsing
+index). In a second set of simulations, regeneration dynamics was
+simulated with a seed rain whose composition depended on the actual
+forest composition. Under this design, a number of seedling is randomly
+chosen for each patch, but the species identity is assigned according to
+species relative abundance (calculated across all 10 patches according
+to species biomass). For instance, if at year i the adult trees of
+species j represent 60% of the total biomass across the 10 patches, then
+60% of the seedlings that try to colonize the patches at year i will
+belong to species j. Then, similarly to the first design, each seedling
+tries to colonize the site depending on whether the suitability between
+species requirements and site conditions. In the two sets of
+simulations, simulations have been run for both thresholds for
+regeneration (7 and 10 cm).
 
@@
 
-Outputs consisted in species specific regeneration rates per decade and per ha, by considering the last 2000 years of simulations to ensure that pseudo-equilibrium state is reached.
-Results were aggregated for every 10 years for the last 2000 years at the 1ha-level.
-Variables:
--	Site: numbers corresponding to the climate data file
--	Sample: one per decade, for a total of 2000 years = identified from 1 to 200
--	Species: name specified as requested (eg. ‘fasy’)
--	r.trees: number of trees outcrossing the threshold (7 or 10 cm) during the decade, for the 10 patches simulated (= 1ha), per species
--	r.ba: summed basal area of the r.trees, per species
--	ba: summed basal area of all trees per species at the end of the decade
--	dbh: threshold used
+Outputs consisted in species specific regeneration rates per decade and
+per ha, by considering the last 2000 years of simulations to ensure that
+pseudo-equilibrium state is reached. Results were aggregated for every
+10 years for the last 2000 years at the 1ha-level. Variables: - Site:
+numbers corresponding to the climate data file - Sample: one per decade,
+for a total of 2000 years = identified from 1 to 200 - Species: name
+specified as requested (eg. 'fasy') - r.trees: number of trees
+outcrossing the threshold (7 or 10 cm) during the decade, for the 10
+patches simulated (= 1ha), per species - r.ba: summed basal area of the
+r.trees, per species - ba: summed basal area of all trees per species at
+the end of the decade - dbh: threshold used
 
@@
 
+```{=html}
 <!--
@@
 -->
-
+```
 A full description of the model can be found in @fischer_lessons_2016.
@@
 
-To incorporate climate, we used the data set `daily_era5_chelsa_200.csv',
-which contains climate data on a daily resolution. We considered the
-covariates precipitation, temperature, and irradiance, the latter
-of which we converted to units used in Formind. Adding on to these
-data, we used a constant and site-independent potential evapotranspiration
-(PET) value of $4\frac{\text{mm}}{\text{d}}$ and assumed that the mean day
-length was $12\text{h}$ at all sites.
-
-Since we simulated longer time periods than the $38$ years for which
-we had climate data, we used the provided data in a randomized fashion.
-For each simulation year, we applied the climate data of a randomly
-selected year. We neglected the existence of leap years, which led
-to a minor time shift in climate data of later years.
+To incorporate climate, we used the data set
+\`daily_era5_chelsa_200.csv', which contains climate data on a daily
+resolution. We considered the covariates precipitation, temperature, and
+irradiance, the latter of which we converted to units used in Formind.
+Adding on to these data, we used a constant and site-independent
+potential evapotranspiration (PET) value of
+$4\frac{\text{mm}}{\text{d}}$ and assumed that the mean day length was
+$12\text{h}$ at all sites.
+
+Since we simulated longer time periods than the $38$ years for which we
+had climate data, we used the provided data in a randomized fashion. For
+each simulation year, we applied the climate data of a randomly selected
+year. We neglected the existence of leap years, which led to a minor
+time shift in climate data of later years.
 
@@
 
-We modelled soil conditions by using the provided soil quality values
-as proxy for soil depth. To that end, we scaled the provided values
-to the typical range of soil depth values used in Formind for temperate
+We modelled soil conditions by using the provided soil quality values as
+proxy for soil depth. To that end, we scaled the provided values to the
+typical range of soil depth values used in Formind for temperate
 forests. Specifically, we multiplied the soil quality values by factor
@@
 
-We considered each of the $11$ species in the protocol explicitly.
-We chose the model parameters for the individual species based on
-the work by @bohn2014. For species that were not
-covered by @bohn2014, we used parameter values of
-similar species with known parameters. The species composition was
-not predefined and evolved solely from the seed input
+We considered each of the $11$ species in the protocol explicitly. We
+chose the model parameters for the individual species based on the work
+by @bohn2014. For species that were not covered by @bohn2014, we used
+parameter values of similar species with known parameters. The species
+composition was not predefined and evolved solely from the seed input
 and the model dynamics.
@@
 
-We simulated a forest area of $1\text{ha}$, which was divided into
-$25$ patches of size $20\text{m}\times20\text{m}$. No interactions
-between trees from different patches were considered in the model.
-That is, the $25$ patches were independent replicates of one another.
+We simulated a forest area of $1\text{ha}$, which was divided into $25$
+patches of size $20\text{m}\times20\text{m}$. No interactions between
+trees from different patches were considered in the model. That is, the
+$25$ patches were independent replicates of one another.
 
@@
 
-We modelled seed generation as global external seed rain. Specifically, we 
-assumed that for each species, there is a constant seed inflow of 
+We modelled seed generation as global external seed rain. Specifically,
+we assumed that for each species, there is a constant seed inflow of
 $250\frac{\text{seeds}}{\text{ha}\cdot \text{yr}}$, which is distributed
-evenly among the $25$ individual patches. We supposed that all 
-seeds promptly establish to seedlings with initial stem diameters at 
-breast height (DBH) of 5\text{cm}.
+evenly among the $25$ individual patches. We supposed that all seeds
+promptly establish to seedlings with initial stem diameters at breast
+height (DBH) of 5\text{cm}.
 
@@
 We started the simulation at each site with a burn-in phase of $1000$
-years and collected samples of the output variables in subsequent
-years. The length of the burn-in phase was chosen based on earlier
-experiences with the model. To confirm that the model reached its
-limiting behaviour, we plotted the evolution of the output variables
-(e.g. the basal area) at selected sites over time and verified that
-no trends were visible after the burn-in phase.
-
-We computed the output variables in equidistant time intervals of
-$60$ years. Before taking a sample, we saved parts of the model's
-state. Then, we simulated the forest for $10$ years and computed
-the output variables, comparing the model's current state with the
-state saved earlier (see below).
-After taking the sample, we simulated the forest for another $50$
-years to reduce correlations between samples. Then, we repeated the
-procedure until the desired number of $200$ samples was collected.
+years and collected samples of the output variables in subsequent years.
+The length of the burn-in phase was chosen based on earlier experiences
+with the model. To confirm that the model reached its limiting
+behaviour, we plotted the evolution of the output variables (e.g. the
+basal area) at selected sites over time and verified that no trends were
+visible after the burn-in phase.
+
+We computed the output variables in equidistant time intervals of $60$
+years. Before taking a sample, we saved parts of the model's state.
+Then, we simulated the forest for $10$ years and computed the output
+variables, comparing the model's current state with the state saved
+earlier (see below). After taking the sample, we simulated the forest
+for another $50$ years to reduce correlations between samples. Then, we
+repeated the procedure until the desired number of $200$ samples was
+collected.
 
@@
 
-To determine the number and basal area of recruits, we compared the
-set of large trees before and after the regeneration time period of
-$10$ years. Prior to simulating the forest for the regeneration time
-period, we saved the internal IDs of all trees with DBHs above the
-thresholds of $0.07\text{m}$ or $0.1\text{m}$, respectively. We
-stored these IDs in set data structures $S_{0.07}^{\text{prior}}$ and
+To determine the number and basal area of recruits, we compared the set
+of large trees before and after the regeneration time period of $10$
+years. Prior to simulating the forest for the regeneration time period,
+we saved the internal IDs of all trees with DBHs above the thresholds of
+$0.07\text{m}$ or $0.1\text{m}$, respectively. We stored these IDs in
+set data structures $S_{0.07}^{\text{prior}}$ and
 $S_{0.1}^{\text{prior}}$, corresponding to one threshold value each.
-After the regeneration time, we repeated the procedure, obtaining
-tree ID sets $S_{0.07}^{\text{posterior}}$ and $S_{0.1}^{\text{posterior}}$.
+After the regeneration time, we repeated the procedure, obtaining tree
+ID sets $S_{0.07}^{\text{posterior}}$ and $S_{0.1}^{\text{posterior}}$.
 We then computed the set differences
 $S_{0.07}^{\text{recruits}}=S_{0.07}^{\text{posterior}}\backslash S_{0.07}^{\text{prior}}$
-and $S_{0.1}^{\text{recruits}}=S_{0.1}^{\text{posterior}}\backslash S_{0.1}^{\text{prior}}$
-and determined both the number and cumulative basal area of the trees
-in $S_{0.07}^{\text{recruits}}$ and $S_{0.1}^{\text{recruits}}$ by considered
-species, respectively.
-
-Finally, we computed the total basal area for each considered species
-by adding up the basal areas of all corresponding individuals. Note
-that the initial DBH of seedlings in our model was $0.05\text{m}$.
-Consequently, smaller trees were not included in the basal area computation.
+and
+$S_{0.1}^{\text{recruits}}=S_{0.1}^{\text{posterior}}\backslash S_{0.1}^{\text{prior}}$
+and determined both the number and cumulative basal area of the trees in
+$S_{0.07}^{\text{recruits}}$ and $S_{0.1}^{\text{recruits}}$ by
+considered species, respectively.
+
+Finally, we computed the total basal area for each considered species by
+adding up the basal areas of all corresponding individuals. Note that
+the initial DBH of seedlings in our model was $0.05\text{m}$.
+Consequently, smaller trees were not included in the basal area
+computation.
 
@@
 
-Yannek Käber
-Harald Bugmann
+Yannek Käber Harald Bugmann
 
@@
 
-Bugmann, H. (1994). On the ecology of mountainous forests in a changing climate: A simulation study.
-Huber, N., Bugmann, H., & Lafond, V.
-(2020).
-Capturing ecological processes in dynamic forest models: Why there is no silver bullet to cope with complexity.
-Ecosphere, 11(5).
-https://doi.org/10.1002/ecs2.3109
+Bugmann, H. (1994). On the ecology of mountainous forests in a changing
+climate: A simulation study. Huber, N., Bugmann, H., & Lafond, V.
+(2020). Capturing ecological processes in dynamic forest models: Why
+there is no silver bullet to cope with complexity. Ecosphere, 11(5).
+<https://doi.org/10.1002/ecs2.3109>
 
@@
 
-Temperature and precipitation were aggregated to monthly means and sums, respectively, including their standard deviation.
-These values served as input for the weather generator within the model.
-The weather generator simulates wheather conditions in an annual time step based on the monthly average temperatures and precipitations sums.
+Temperature and precipitation were aggregated to monthly means and sums,
+respectively, including their standard deviation. These values served as
+input for the weather generator within the model. The weather generator
+simulates wheather conditions in an annual time step based on the
+monthly average temperatures and precipitations sums.
 
@@
 
-Soil data were translated to bucket sizes values based on the assumpation that the lowest soil quality value of 1 refers to a bucket size of 10 cm and the highest soil quality of 5 refers to a bucket size of 35 cm.
-Specifically we used the following formula to calculate the bucket size bucket_size = 10 + (soil_quality-1) /4  * 25.
+Soil data were translated to bucket sizes values based on the
+assumpation that the lowest soil quality value of 1 refers to a bucket
+size of 10 cm and the highest soil quality of 5 refers to a bucket size
+of 35 cm. Specifically we used the following formula to calculate the
+bucket size bucket_size = 10 + (soil_quality-1) /4  \* 25.
 
@@
 Topographic information was included by specifying the kSlAsp parameter.
-This parameter defines a value between -2 and 2 which affects the calculated Potential Evapotranspiration within the model.
-For details on the calculation of this value see Käber et.
-al.
-2021.
+This parameter defines a value between -2 and 2 which affects the
+calculated Potential Evapotranspiration within the model. For details on
+the calculation of this value see Käber et. al. 2021.
 
@@
 
-All species required were simulated:
-Fagus sylvatica, Picea abies, Abies alba, Carpinus betulus, Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris
-
-In addition the following species were simulated:
-Acer campestre, Acer platanoides, Alnus incana, Alnus viridis, Castanea sativa, Corylus avellana, Larix decidua, Pinus montana, Populus nigra, Populus tremula, Salix alba, Sorbus aria, Sorbus aucuparia, Taxus baccata, Tilia platyphyllos, Ulmus glabra, Pinus cembra
+All species required were simulated: Fagus sylvatica, Picea abies, Abies
+alba, Carpinus betulus, Tilia cordata, Acer pseudoplatanus, Betula spp.,
+Fraxinus excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris
+
+In addition the following species were simulated: Acer campestre, Acer
+platanoides, Alnus incana, Alnus viridis, Castanea sativa, Corylus
+avellana, Larix decidua, Pinus montana, Populus nigra, Populus tremula,
+Salix alba, Sorbus aria, Sorbus aucuparia, Taxus baccata, Tilia
+platyphyllos, Ulmus glabra, Pinus cembra
 
@@
 
-All simulations were run according to the protocoll.
-For each site we simulated on 16 ha (i.e., 100 patches of 0.08 ha).
-We defined a spin up phase of 1000 years.
-After the spin up we simulated 200 years of forest dynamics.
+All simulations were run according to the protocoll. For each site we
+simulated on 16 ha (i.e., 100 patches of 0.08 ha). We defined a spin up
+phase of 1000 years. After the spin up we simulated 200 years of forest
+dynamics.
 
@@
 
-We sampled in 10 year intervals between the simulation years 1010 to 1200, which resulted in 16 times 20 intervals each with one ha.
+We sampled in 10 year intervals between the simulation years 1010 to
+1200, which resulted in 16 times 20 intervals each with one ha.
 
@@
 
-*Please provide here the references to this model where it is possible to understand the model structure specially for the regeneration module, this could be a published paper, a website with the model description, a manual or some other document that the reader can access.*
-
-Landscape scale forest simulation model LANDIS-II (Scheller et al.
-2007) with PnET-Succession extension (version 4.1).
-PnET-Succession is based on the Biomass Succession extension of Scheller and Mladenoff (2004), embedding elements of the PnET-II ecophysiology model of Aber et al.
-(1995) to mechanistically simulate growth as a competition for available light and water (De Bruijn et al.
-2014).
-PnET-Succession simulates the competition among cohorts for water and light at a monthly time-step as a function of photosynthetic processes and maintenance respiration that are explicitly linked to environmental drivers such as temperature, precipitation, photosynthetic active radiation (PAR), and CO~2~ concentration.
-Regeneration depends on distance from a seed source, soil water, and subcanopy light, while competition is modeled by partitioning incoming light through multiple canopy layers.
-Growth of specific cohort biomass components is allocated in the following order: non structural carbon, first, then foliage then root and stem.
-Biomass growth decreases as cohorts approach their longevity age but mortality can occur at any time when carbon reserves production is insufficient to support growth due to shading, water competition, and/or drought (Gustafson et al.
-2015), or by any disturbance or management extension used.
-For output, we selected Biomass Community Output Extension version 2.0.1, which generates output in the format: species age biomass (g / m^2^) at specified time steps.
-
-### Climate
-
-Climate input was generated by randomly selecting entire years from the monthly data from the CHELSA dataset and collating them one after the other to create a 2000 years series.
-The sampling was repeated 10 times with different random order of years to create the 10 climate replicates.
-Selecting a year at a time eliminates inconsistent seasons (especially summer).
-We used monthly mean, minimum and maximum temperatures and precipitation as input for PnET succession extension.
-PnET-Succession also requires mean monthly values of PAR during daylight hours, which we calculated by dividing the monthly rsds data by the number of daylight seconds in each month for latitude 47 to get W/m2.
-We then multiplied this by 4.6 to get μmol/m^2^/s and multiplied by 0.5 to get PAR part of the spectrum, based on the approximation by Tsubo et al.
-(2005).
-Leap years were not considered.
-The PAR distribution across all sites and years is in range with values that are typical for temperate regions.
-
-CO~2~ concentration with yearly timestep was added from the concentrations table available from EEA (2019).
-For years 0-1750 CO~2~ concentration was maintained fixed to 278 ppm.
-As for the period 1750-1975 there was only one value each 5 years, we used a constant value for each 5-year period.
-CO~2~ concentration was added to the climate input file after randomization of temperature, precipitation and PAR (see above), so it reflects realistic values for the simulated years.
-Simulations were run from year 0 to 1990 but sampled outputs did not exceed year 1910.
-Thus, the drastic increase in CO~2~ concentration during the second half of the 1900s did not affect the delivered results.
-
-### Soil
-
-Soil type was set to LOAM at all sites.
-Rooting depth was calculated from soil quality codes (1-5, original values ranged from 1.13 to 4.87) multiplicated by 205.
-This produces a soil depth in the range 231 - 998 mm, that are within the typical range for PnET-Succession (maximum 1000 mm).
-A clear trend emerged - deeper soils resulted in higher biomass, although with some exceptions.
-
-### Topograhpy
-
-The ecoregion-specific parameter Precipitation Loss Fraction denotes the water lost to runoff induced by topographic slope.
-We estimated this parameter for each site from topo_data_200 dataset as slope \* 0.01 (e.g., 12% slope = 0.12 PrecLossFrac).
-We acknowledge that this is a simplification and that better parameterization would be possible if soil types were available for each site.
-Tests showed that sites with extreme slope tended to have lower biomass, but the signal was not that clear.
-The correlation between slope and elevation was 0.6 which could obscure the results.
-
-Some sites at very high elevation had, correspondingly, very low temperature.
-At these sites, the model showed a visible trend in producing lower biomass (down to zero at a few sites).
-Yet this was probably not apparent in the regeneration results, as extremely low biomass could still support regeneration, even if cohorts do not live long, due to our modification of sexual maturity age.
-
-### Tree species
-
-As LANDIS-II cannot start from bare ground, initial communities were set to include all species defined in the protocol at all sites starting with 10-years old cohorts of the following species: fasy (*Fagus sylvatica*), pabi (*Picea abies*), abal (*Abies alba*), cabe (*Carpinus betulus*), tico (*Tilia cordata*), acps (*Acer pseudoplatanus*), betu (*Betula spp.*), frex (*Fraxinus excelsior*), quer (*Quercus spp.*), algl (*Alnus glutinosa*), pisy (*Pinus sylvestris*).
-The species-specific parameter denoting sexual maturity was set to 10 years equally for all species, so that seed source of all species is available from the onset of the simulations at all sites and to maintain seed in the ladscape pool.
-Vegetative regeneration (resprouting) was prevented, since this would have an effect on allometry and regeneration.
-
-Some species may be under or overrepresented because their occurrence in LANDIS-II simulations is typically due to small-scale disturbances or management which were not included in the simulations.
-
-The model requires two sets of parameters representing species-specific life traits.
-The first set is used by the LANDIS-II core module.
-The species-specific parameter denoting "longevity" and "sexual maturity" were the only parameters of this set considered in the current exercise.
-The "sexual maturity" was set to 10 years equally for all species, so that seed source of all species is available from the onset of the simulations at all sites and to maintain seed in the landscape pool.
-Vegetative regeneration (resprouting) was prevented since this would influence allometry and regeneration.
-
-The second set is used by the PnET-Succession extension and includes several ecophysiological parameters.
-For this second set we started with the values already applied in Petter et al.
-(2020) with a previous version of PnET.
-The values of some of the parameters to which PnET is most sensitive and/or for which the range of variation had been modified for PnET4.1 were then iteratively adjusted (individually and then in combination) to meet the requirements of PnET version 4.1 used in this exercise.
-The changes were implemented one by one in simulations with the exercise data and those leading to an output deemed realistic and to a relative stabilization of the biomass trend through the simulation were retained for the implementation of the full set of replicates.
-The retained changes include: 1) decreased percentage of foliar nitrogen for algl (*Alnus glutinosa*); 2) rescaling to the new ranges for all species of the values of the Halfsat and the FracBelowG.
-The first of these parameters represents shade tolerance and its value affects the probability of establishment of a species (Half saturation light level for photosynthesis.); the latter regulates allocation between above and below ground pool (Fraction of non-foliar biomass that is belowground (root pool)) ; 3) a slight decrease of the value of the parameter (Fracfol) affecting the yearly allocation to foliage of the active total woody biomass for the three conifers abal (*Abies alba*), piab (*Picea abies*), pisy (*Pinus sylvestris*).
-
-### Simulation
-
-Simulation area was defined as a rectangle of 10×20 dimension with sites (i.e., cells) of 1-ha.
-Each site is an individual ecoregion (i.e., one cell - as in our case - or group of cells sharing similar climate and soil conditions, as a way in LANDIS-II to define differences in biophysical conditions across landscapes) with specific values for soil quality, topography and climate, based on the data provided.
-The model runs with a yearly timestep using monthly climate.
-We started each model run with initial communities of all species at age 10 at all sites.
-Each run was simulated for 1980 years.
-
-#### Dispersal
-
-We have used the non-spatial seeding algorithm "universal dispersal" for all the model runs.
-This way, every species can seed any forest site in the landscape (across all the 200 sites).
-This mode does not take into account the spatial arrangement and contiguity of cells (i.e., no active seed dispersal), which is proper for the sake of the model comparison exercise.
-Yet in real landscape seed limitation can also play a role in species composition.
-The seed number is affected by species-specific age of maturity, which we have modified.
-This setting ensures an influx of seeds from outside consistent with the original settings.
-
-#### Sampling design
-
-We have sampled each site 20 times along the simulation run in two time windows starting 50 years after the first cycle following typical species longevity.
-After the year 700, biomass of most species was relatively stable and we sampled every 50 years, starting at 750 until the year 1200.
-The next sampling period started from 1450 and continued until 1900, i.e., after the next longevity peak, again every 50 years.
-At each selected sampling time, we extracted cohorts of each species according to the age related to the two DBH thresholds (7cm/10cm).
-Additionally, we extracted cohorts that were up to 9 years older than these thresholds, to get the expected results for a 10-year interval.
-For example, *Fagus sylvatica* reaches DBH 7cm at 17 years, therefore we have selected all cohorts with age 17-26.
-
-To deliver more than one simulation per site, we created 10 replicates with different climate replicate and random number generator seed number to account for stochastic processes in PnET-Succession.
-This resulted in the expected 200 samples for each site.
-
-### Outputs
-
-*Please describe how have you calculated each of the variables in the output data.*
-
-Sample number code
-
-Number (1-200) was defined as sample in time + (replicate-1)\*20, with sample in time (1-20) according to the sample order.
-Replicate (1-10) combines random number and climate sequence.
-For example sample number 25 was taken from the climate replicate 2 and the fifth sample in time (year 950).
-
-r.trees
-
-As LANDIS-II outputs are expressed in terms of cohort age and mean biomass per site, we have calculated the mean DBH and biomass of each species based on age.
-We estimated the age when each species reaches 7 cm and 10 cm DBH and the number of individuals using allometric models from the Italian forest inventory collected in R package ForIT (Puletti et al.
-2014), and unpublished data from the study by Mina et al.
-(2021).
-Please note that the equations used predict DBH \>0 already for a tree with age 1, this is because it is based on forestry data and tree cores.
-The real age when reaching 1.3 m height is species specific and we did not implement this offset.
-For each cohort, we divided the cohort biomass reported by LANDIS-II, recalculated to kg/ha by the mean biomass of a tree of the same species and age to obtain the number of trees of each species per site.
-We rounded all results up to the nearest integer to avoid any fraction of trees.
-The number of individuals was aggregated for all the cohorts of the age of reaching 7 cm or 10 cm DBH and all cohorts of the same species that were up to 9 years older.
-This approach does not reflect influence of site conditions or stand densities on Age-DBH and Age-Biomass.
-Relationships between age and DBH were derived by fitting asymptotic regressions using individual tree measurements from an ensemble of forest inventory plots distributed across south-eastern Canada (see Supporting information of Mina et al.
-2021).
-This was the only public dataset for which we could relate tree age, derived from stem coring and tree ring analysis, to measured DBH.
-Although tree age measurements were obtained with a robust methodology, these functions do not disentangle the influence of site conditions or stand densities.
-Additionally, they reflect growing conditions of tree genera in Eastern North America.
-
-r.ba basal area of regeneration Calculated as a sum of r.ba of all the cohorts selected in r.trees using the following formula: DBH2 \* π/4 \* r.tree.
-
-ba: total basal area We have not calculated this since our allometric equations do not cover older ages.
-Therefore, we produced the following two biomass variables as a surrogate.
-The basal area was later calculated as a sum of r.ba and estimated ba based on biomass.
-The estimation includes transformation of kg into tonnes / ha and expansion factor 12.5 to transform ba m^2^/ha.
-
-((( landis\$sum.bio - landis\$r.bio.sum )\* 0.001) / 12.5). This includes
-
-sum.bio\
-Biomass of all cohorts of the species on site including the regeneration and even younger regeneration (units kg/ha).
-
-r.bio.sum Biomass of the regeneration cohorts corresponding to the r. ba and r.trees on site (units kg/ha).
-
-### References
-
-Aber J.
-D.
-& Federer C.
-A.
-(1992).
-A generalized, lumped-parameter model of photosynthesis, evapotranspiration and net primary production in temperate and boreal forest ecosystems.
-Oecologia 92(4): 463-474.
-
-Aber J.D., Ollinger S.V., Federer C.A., Reich P.B., Goulden M.L., Kicklighter D.W., Melillo J.M.
-& Lathrop R.G.
-Jr.
-(1995) Predicting the effects of climate change on water yield and forest production in the northeastern United States.
-Climate Research 5(3): 207--22.
-
-de Bruijn A., Gustafson E.J.
-, Sturtevant B.R., Foster J.R., Miranda B.R., Lichti N.I.
-& Jacobs D.F.
-(2014) Toward more robust projections of forest landscape dynamics under novel environmental conditions: Embedding PnET within LANDIS-II.
-Ecological Modelling (287): 44--57
-
-EEA (2019) Trends in atmospheric concentrations of CO2 (ppm), CH4 (ppb) and N2O (ppb), between 1800 and 2017 available at <https://www.eea.europa.eu/ds_resolveuid/217c026ca03649398aadf39d87623e57>
-
-Gustafson, E.
-J., A.
-M.
-G.
-De Bruijn, R.
-E.
-Pangle, J.-M.
-Limousin, N.
-G.
-McDowell, W.
-T.
-Pockman, B.
-R.
-Sturtevant, J.
-D.
-Muss, and M.
-E.
-Kubiske.
-2015.
-Integrating ecophysiology and forest landscape models to improve projections of drought effects under climate change.
-Global Change Biology 21:843-856.
-
-Mina, M., Messier, C., Duveneck, M., Fortin, M.
-J., & Aquilué, N.
-2021.
-Network analysis can guide resilience-based management in forest landscapes under global change.
-Ecological Applications, 31(1), 1--18.
-
-Petter G., Mairota P., Albrich K., Bebi P., Brůna J., Bugmann H., Haffenden A., Scheller R.M., Schmatz D.R., Seidl R., Speich M., Vacchiano G.
-& Lischke H.
-(2020) How robust are future projections of forest landscape dynamics?
-Insights from a systematic comparison of four forest landscape models.
-Environmental Modelling & Software 134: 104844.
-<https://doi.org/10.1016/j.envsoft.2020.104844>
-
-Puletti N., Mura M., Castaldi C., Marchi N., Chiavetta U.
-and Scotti R.
-(2014).
-ForIT: Functions from the 2nd Italian Forest Inventory (INFC).
-R package version 1.0.
-<https://CRAN.R-project.org/package=ForIT>
-
-Scheller, R.
-M., Domingo J.
-B.
-, Sturtevant B.
-R.
-, Williams J.
-S.
-, Rudy A.
-, Gustafson E.
-J.
-& Mladenoff D.
-J.
-(2007) Design, development, and application of LANDIS-II, a spatial landscape simulation model with flexible temporal and spatial resolution.
-Ecological Modelling 201(3-4): 409--419.
-<https://doi.org/10.1016/j.ecolmodel.2006.10.009>
-
-Tsubo M.
-& Walker S.
-(2005) Relationships between photosynthetically active radiation and clearness index at Bloemfontein, South Africa.
-Theoretical and Applied Climatology 80: 17--25.
-<https://doi.org/10.1007/s00704-004-0080-5>
-
-## TreeMig
-
-### Authors
-Heike Lischke
-
-### Model
-
-The model TreeMig (see treemig.wsl.ch) was originally based on the ForClim model, then aggregated by a distribution based approach (Lischke et al.
-1998), and extended to a forest landscape model by including seed production, seed dispersal, seed bank dynamics and seedling establishment (Lischke and Loffler 2006, Lischke et al.
-2006).
-
-### Climate
-
-As climatic input, the monthly time series of mean temperature and precipitation sum were used.
-With these climatic variables and additional data about slope, aspect and soil field capacity, yearly time series of the following bioclimatic variables were calculated:  (1) DDSum, the yearly day degree sum above 5.5°C , (2) MinWiT, the lowest mean monthly temperature, (3) DrStr, drought stress according to a bucket model  that estimates a monthly PET following (Thornthwaite and Mather 1957), using temperature, latitude, slope and aspect as input,  and calculates the monthly soil water content  based on soil water content in the previous month, precipitation, interception, and field capacity (“bucketsize”)   (Bugmann 1994, Fischlin et al.
-1995).
-Drought stress is then given by 1- the ratio between demand (PET-interception) and the supply, a function of the soil water content.
-To get climate data also for the spin-up, bioclimate was calculated from the monthly data   as given, and then extended by sampling from the given data from the end of the climate data to get 1000 years.
-The same sequence of sampling years was used in all simulations and sites.
-
-### Soil
-
-The soil bucketsize (available water capacity of the upmost 1m soil layer) required for the drought stress calculation, was derived from the soil quality data set, by
-Bucketsize = 10 + (soil_quality-1) /4  * 25 , to covert the range from 10 cm to 35cm
-
-### Topography
-
-Slope and aspect were taken from the topographic data set and converted to a common variable entering the PET calculation, by
-slasp= 2*cos(aspect *pi/180) *min(1, slope/60)
-
-### Tree species
-
-Thirty Central European tree species were simulated, including the standard species of the project .
-Only Betula pubescens was not simulated, because we did not have a parametrization for this species at hand.
-Species: Abies alba, Larix decidua, Picea abies, Pinus cembra, Pinus montana, Pinus sylvestris, Taxus baccata, Acer campestre, Acer platanoides, Acer pseudoplatanus, Alnus glutinosa, Alnus incana, Alnus viridis, Betula pendula, Carpinus betulus, Castanea sativa, Corylus avellana, Fagus sylvatica, Fraxinus excelsior, Populus nigra, Populus tremula, Quercus petraea, Quercus pubescens, Quercus robur, Salix alba, Sorbus aria, Sorbus aucuparia, Tilia cordata, Tilia platyphyllos, Ulmus glabra
-
-### Simulation
-
-The simulations were carried out on a grid of 15 x 15 cells, each with 200 m side length.
-
-#### Dispersal
-
-The seeds produced by the mother trees are transported from the source cell according to a double negative exponential, species specific that determines the share of the seeds in the source cell landing in a sink cell in a given distance to the source cell.
-In the sink cells, the seeds enter the seed bank .
-At the start of the simulation for 10 years seeds of all species are available in all grid cells, from then on, the normal seed dispersal starts.
-
-#### Regeneration processes
-
-TreeMig simulates the full feedback from seed production to ingrowth. Some of the involved processes depend on environment.
-(1) Overall seed production in TreeMig is described by multiplying a species specific seed production per tree that depends on tree size and thus indirectly on the growth, which depends on DDsum, drought stress, nutrients and light and the adult tree number that depends on survival  that depends on DDsum, drought stress, nutrients and light.
-(2) Seed dispersal is independent of environment. 
-(3) Seed bank dynamics – also environment independent -   is given by seed input, seed mortality, and germination.
-Additionally, the number of seeds S of a species in the seed bank of this species (SB) is limited to a carrying capacity C of 1000, by
-SB = C + (SB - C) * Exp(-S/C).  
-(4) Germination and seedling survival depend on browsing, drought, winter temperature, DDSum and light. 
-(5) The saplings up to 1.37 die and grow like adult trees, depending on DDsum, drought stress and light. Only the parameter value of the light dependence differs from that of the adult trees.  
-The regeneration processes were simulated annually in the order 1.
-germination, 2.
-germinated seeds leave the seed bank, 3.
-mortality of seeds in the seed bank, 4.
-seed entry by seed dispersal.
-The ingrowth numbers were very sensitive to this order.
-Simulations with another sequence, in which the seeds in the seed bank died first and then  the  remaining ones germinated,  resulted in about half the ingrowth.
-
-#### Simulation length
-
-The simulations were started on bare ground and run for 1000 years, when at all sites the equilibrium was reached.
-
-#### More simulations per site
-
-TreeMig simulates directly the mean dynamics of entire stands, but based on the  variability of light conditions within the stand   [distribution based approach, \Lischke, 1998, aggregation.
-To come up with the required 200 random patches of 1 ha, temporal and spatial variability intrinsic to TreeMig were combined.
-Temporal variability was given by sampling from each decade of the last 100 years of the simulation, i.e.
-years 900-1000.
-For each of theses 10 decades, 20 grid cells were sampled from the simulated grid, reflecting spatial variability, results from demographic stochasticity and short-range spatial interactions
-
-### Outputs
-
-In contrast to gap models, TreeMig is based on the concept of frequency distributions of tree densities classes in different tree heights on ca.
-1/12 ha (833m^2) patches, and the resulting frequency distributions in discrete light.
-During the simulation, in each height class the frequencies of the different light classes were recorded, as well as the ingrowth of each species in the height-light class.
-By sampling 12 times from the height-light class ingrowths according to the height-light class frequency distribution and summing up, the stochastic ingrowths into all TreeMig height classes for one ha were calculated for each grid cell.
-Then the height class boundaries were translated into DBH and the ingrowths into the height classes linearly interpolated to 7cm and 10cm.
-The ingrowth basal areas into the height classes were obtained by multiplying the interpolated ingrowth numbers with the basal areas at  7cm and 10cm, i.e.
-r.trees * Pi*  (0.07/2)2  and r.trees * Pi*  (0.1/2)2 .
-The species specific basal area (at 1.37m) sums are a standard output of TreeMig, given by multiplying the state variables (number of trees per grid cell, height class, and species) with  the species specific basal area of each height class.
-To account for the ingrown trees dying within a decade, as an approximation the survival in each year to the power of 5 was used.
-Thereby, it was assumed that the trees have to survive in average 5 years in each decade, and that the current mortality is valid for these 5 years.
-
-### References
-
-•	Bugmann, H. 1994. On the ecology of mountainous forests in a changing climate: A simulation study. Dissertation. Swiss Federal Institute of Technology, Zurich.
-•	Fischlin, A., H. Bugmann, and D. Gyalistras. 1995. Sensitivity of a forest ecosystem model to climate parametrization schemes. Env. Poll. 87:267-282.
-•	Lischke, H., and T.
-J.
-Loffler.
-2006.
-Intra-specific density dependence is required to maintain species diversity in spatio-temporal forest simulations with reproduction.
-Ecological Modelling 198:341-361.
-•	Lischke, H., T.
-J.
-Loffler, and A.
-Fischlin.
-1998.
-Aggregation of individual trees and patches in forest succession models: Capturing variability with height structured, random, spatial distributions.
-Theor Popul Biol 54:213-226.
-•	Lischke, H., N.
-E.
-Zimmermann, J.
-Bolliger, S.
-Rickebusch, and T.
-J.
-Loffler.
-2006.
-TreeMig: A forest-landscape model for simulating spatio-temporal patterns from stand to landscape scale.
-Ecological Modelling 199:409-420.
-•	Thornthwaite, C. W., and J. R. Mather. 1957. The water balance. Publications in Climatology VIII:1-69.
-
-## LPJ-GUESS
-
-### Authors
-
-### Model
-
 *Please provide here the references to this model where it is possible
@@
 manual or some other document that the reader can access.*
+
+Landscape scale forest simulation model LANDIS-II (Scheller et al. 2007)
+with PnET-Succession extension (version 4.1). PnET-Succession is based
+on the Biomass Succession extension of Scheller and Mladenoff (2004),
+embedding elements of the PnET-II ecophysiology model of Aber et al.
+(1995) to mechanistically simulate growth as a competition for available
+light and water (De Bruijn et al. 2014). PnET-Succession simulates the
+competition among cohorts for water and light at a monthly time-step as
+a function of photosynthetic processes and maintenance respiration that
+are explicitly linked to environmental drivers such as temperature,
+precipitation, photosynthetic active radiation (PAR), and CO~2~
+concentration. Regeneration depends on distance from a seed source, soil
+water, and subcanopy light, while competition is modeled by partitioning
+incoming light through multiple canopy layers. Growth of specific cohort
+biomass components is allocated in the following order: non structural
+carbon, first, then foliage then root and stem. Biomass growth decreases
+as cohorts approach their longevity age but mortality can occur at any
+time when carbon reserves production is insufficient to support growth
+due to shading, water competition, and/or drought (Gustafson et al.
+2015), or by any disturbance or management extension used. For output,
+we selected Biomass Community Output Extension version 2.0.1, which
+generates output in the format: species age biomass (g / m^2^) at
+specified time steps.
+
+### Climate
+
+Climate input was generated by randomly selecting entire years from the
+monthly data from the CHELSA dataset and collating them one after the
+other to create a 2000 years series. The sampling was repeated 10 times
+with different random order of years to create the 10 climate
+replicates. Selecting a year at a time eliminates inconsistent seasons
+(especially summer). We used monthly mean, minimum and maximum
+temperatures and precipitation as input for PnET succession extension.
+PnET-Succession also requires mean monthly values of PAR during daylight
+hours, which we calculated by dividing the monthly rsds data by the
+number of daylight seconds in each month for latitude 47 to get W/m2. We
+then multiplied this by 4.6 to get μmol/m^2^/s and multiplied by 0.5 to
+get PAR part of the spectrum, based on the approximation by Tsubo et al.
+(2005). Leap years were not considered. The PAR distribution across all
+sites and years is in range with values that are typical for temperate
+regions.
+
+CO~2~ concentration with yearly timestep was added from the
+concentrations table available from EEA (2019). For years 0-1750 CO~2~
+concentration was maintained fixed to 278 ppm. As for the period
+1750-1975 there was only one value each 5 years, we used a constant
+value for each 5-year period. CO~2~ concentration was added to the
+climate input file after randomization of temperature, precipitation and
+PAR (see above), so it reflects realistic values for the simulated
+years. Simulations were run from year 0 to 1990 but sampled outputs did
+not exceed year 1910. Thus, the drastic increase in CO~2~ concentration
+during the second half of the 1900s did not affect the delivered
+results.
+
+### Soil
+
+Soil type was set to LOAM at all sites. Rooting depth was calculated
+from soil quality codes (1-5, original values ranged from 1.13 to 4.87)
+multiplicated by 205. This produces a soil depth in the range 231 - 998
+mm, that are within the typical range for PnET-Succession (maximum 1000
+mm). A clear trend emerged - deeper soils resulted in higher biomass,
+although with some exceptions.
+
+### Topograhpy
+
+The ecoregion-specific parameter Precipitation Loss Fraction denotes the
+water lost to runoff induced by topographic slope. We estimated this
+parameter for each site from topo_data_200 dataset as slope \* 0.01
+(e.g., 12% slope = 0.12 PrecLossFrac). We acknowledge that this is a
+simplification and that better parameterization would be possible if
+soil types were available for each site. Tests showed that sites with
+extreme slope tended to have lower biomass, but the signal was not that
+clear. The correlation between slope and elevation was 0.6 which could
+obscure the results.
+
+Some sites at very high elevation had, correspondingly, very low
+temperature. At these sites, the model showed a visible trend in
+producing lower biomass (down to zero at a few sites). Yet this was
+probably not apparent in the regeneration results, as extremely low
+biomass could still support regeneration, even if cohorts do not live
+long, due to our modification of sexual maturity age.
+
+### Tree species
+
+As LANDIS-II cannot start from bare ground, initial communities were set
+to include all species defined in the protocol at all sites starting
+with 10-years old cohorts of the following species: fasy (*Fagus
+sylvatica*), pabi (*Picea abies*), abal (*Abies alba*), cabe (*Carpinus
+betulus*), tico (*Tilia cordata*), acps (*Acer pseudoplatanus*), betu
+(*Betula spp.*), frex (*Fraxinus excelsior*), quer (*Quercus spp.*),
+algl (*Alnus glutinosa*), pisy (*Pinus sylvestris*). The
+species-specific parameter denoting sexual maturity was set to 10 years
+equally for all species, so that seed source of all species is available
+from the onset of the simulations at all sites and to maintain seed in
+the ladscape pool. Vegetative regeneration (resprouting) was prevented,
+since this would have an effect on allometry and regeneration.
+
+Some species may be under or overrepresented because their occurrence in
+LANDIS-II simulations is typically due to small-scale disturbances or
+management which were not included in the simulations.
+
+The model requires two sets of parameters representing species-specific
+life traits. The first set is used by the LANDIS-II core module. The
+species-specific parameter denoting "longevity" and "sexual maturity"
+were the only parameters of this set considered in the current exercise.
+The "sexual maturity" was set to 10 years equally for all species, so
+that seed source of all species is available from the onset of the
+simulations at all sites and to maintain seed in the landscape pool.
+Vegetative regeneration (resprouting) was prevented since this would
+influence allometry and regeneration.
+
+The second set is used by the PnET-Succession extension and includes
+several ecophysiological parameters. For this second set we started with
+the values already applied in Petter et al. (2020) with a previous
+version of PnET. The values of some of the parameters to which PnET is
+most sensitive and/or for which the range of variation had been modified
+for PnET4.1 were then iteratively adjusted (individually and then in
+combination) to meet the requirements of PnET version 4.1 used in this
+exercise. The changes were implemented one by one in simulations with
+the exercise data and those leading to an output deemed realistic and to
+a relative stabilization of the biomass trend through the simulation
+were retained for the implementation of the full set of replicates. The
+retained changes include: 1) decreased percentage of foliar nitrogen for
+algl (*Alnus glutinosa*); 2) rescaling to the new ranges for all species
+of the values of the Halfsat and the FracBelowG. The first of these
+parameters represents shade tolerance and its value affects the
+probability of establishment of a species (Half saturation light level
+for photosynthesis.); the latter regulates allocation between above and
+below ground pool (Fraction of non-foliar biomass that is belowground
+(root pool)) ; 3) a slight decrease of the value of the parameter
+(Fracfol) affecting the yearly allocation to foliage of the active total
+woody biomass for the three conifers abal (*Abies alba*), piab (*Picea
+abies*), pisy (*Pinus sylvestris*).
+
+### Simulation
+
+Simulation area was defined as a rectangle of 10×20 dimension with sites
+(i.e., cells) of 1-ha. Each site is an individual ecoregion (i.e., one
+cell - as in our case - or group of cells sharing similar climate and
+soil conditions, as a way in LANDIS-II to define differences in
+biophysical conditions across landscapes) with specific values for soil
+quality, topography and climate, based on the data provided. The model
+runs with a yearly timestep using monthly climate. We started each model
+run with initial communities of all species at age 10 at all sites. Each
+run was simulated for 1980 years.
+
+#### Dispersal
+
+We have used the non-spatial seeding algorithm "universal dispersal" for
+all the model runs. This way, every species can seed any forest site in
+the landscape (across all the 200 sites). This mode does not take into
+account the spatial arrangement and contiguity of cells (i.e., no active
+seed dispersal), which is proper for the sake of the model comparison
+exercise. Yet in real landscape seed limitation can also play a role in
+species composition. The seed number is affected by species-specific age
+of maturity, which we have modified. This setting ensures an influx of
+seeds from outside consistent with the original settings.
+
+#### Sampling design
+
+We have sampled each site 20 times along the simulation run in two time
+windows starting 50 years after the first cycle following typical
+species longevity. After the year 700, biomass of most species was
+relatively stable and we sampled every 50 years, starting at 750 until
+the year 1200. The next sampling period started from 1450 and continued
+until 1900, i.e., after the next longevity peak, again every 50 years.
+At each selected sampling time, we extracted cohorts of each species
+according to the age related to the two DBH thresholds (7cm/10cm).
+Additionally, we extracted cohorts that were up to 9 years older than
+these thresholds, to get the expected results for a 10-year interval.
+For example, *Fagus sylvatica* reaches DBH 7cm at 17 years, therefore we
+have selected all cohorts with age 17-26.
+
+To deliver more than one simulation per site, we created 10 replicates
+with different climate replicate and random number generator seed number
+to account for stochastic processes in PnET-Succession. This resulted in
+the expected 200 samples for each site.
+
+### Outputs
+
+Sample number code
+
+Number (1-200) was defined as sample in time + (replicate-1)\*20, with
+sample in time (1-20) according to the sample order. Replicate (1-10)
+combines random number and climate sequence. For example sample number
+25 was taken from the climate replicate 2 and the fifth sample in time
+(year 950).
+
+r.trees
+
+As LANDIS-II outputs are expressed in terms of cohort age and mean
+biomass per site, we have calculated the mean DBH and biomass of each
+species based on age. We estimated the age when each species reaches 7
+cm and 10 cm DBH and the number of individuals using allometric models
+from the Italian forest inventory collected in R package ForIT (Puletti
+et al. 2014), and unpublished data from the study by Mina et al. (2021).
+Please note that the equations used predict DBH \>0 already for a tree
+with age 1, this is because it is based on forestry data and tree cores.
+The real age when reaching 1.3 m height is species specific and we did
+not implement this offset. For each cohort, we divided the cohort
+biomass reported by LANDIS-II, recalculated to kg/ha by the mean biomass
+of a tree of the same species and age to obtain the number of trees of
+each species per site. We rounded all results up to the nearest integer
+to avoid any fraction of trees. The number of individuals was aggregated
+for all the cohorts of the age of reaching 7 cm or 10 cm DBH and all
+cohorts of the same species that were up to 9 years older. This approach
+does not reflect influence of site conditions or stand densities on
+Age-DBH and Age-Biomass. Relationships between age and DBH were derived
+by fitting asymptotic regressions using individual tree measurements
+from an ensemble of forest inventory plots distributed across
+south-eastern Canada (see Supporting information of Mina et al. 2021).
+This was the only public dataset for which we could relate tree age,
+derived from stem coring and tree ring analysis, to measured DBH.
+Although tree age measurements were obtained with a robust methodology,
+these functions do not disentangle the influence of site conditions or
+stand densities. Additionally, they reflect growing conditions of tree
+genera in Eastern North America.
+
+r.ba basal area of regeneration Calculated as a sum of r.ba of all the
+cohorts selected in r.trees using the following formula: DBH2 \* π/4 \*
+r.tree.
+
+ba: total basal area We have not calculated this since our allometric
+equations do not cover older ages. Therefore, we produced the following
+two biomass variables as a surrogate. The basal area was later
+calculated as a sum of r.ba and estimated ba based on biomass. The
+estimation includes transformation of kg into tonnes / ha and expansion
+factor 12.5 to transform ba m^2^/ha.
+
+((( landis\$sum.bio - landis\$r.bio.sum )\* 0.001) / 12.5). This
+includes
+
+sum.bio\
+Biomass of all cohorts of the species on site including the regeneration
+and even younger regeneration (units kg/ha).
+
+r.bio.sum Biomass of the regeneration cohorts corresponding to the r. ba
+and r.trees on site (units kg/ha).
+
+Please noticed that we included 0 values for species missing in a given
+sample and for samples in sites where Landis did not simulate any tree.
+
+### References
+
+Aber J. D. & Federer C. A. (1992). A generalized, lumped-parameter model
+of photosynthesis, evapotranspiration and net primary production in
+temperate and boreal forest ecosystems. Oecologia 92(4): 463-474.
+
+Aber J.D., Ollinger S.V., Federer C.A., Reich P.B., Goulden M.L.,
+Kicklighter D.W., Melillo J.M. & Lathrop R.G. Jr. (1995) Predicting the
+effects of climate change on water yield and forest production in the
+northeastern United States. Climate Research 5(3): 207--22.
+
+de Bruijn A., Gustafson E.J. , Sturtevant B.R., Foster J.R., Miranda
+B.R., Lichti N.I. & Jacobs D.F. (2014) Toward more robust projections of
+forest landscape dynamics under novel environmental conditions:
+Embedding PnET within LANDIS-II. Ecological Modelling (287): 44--57
+
+EEA (2019) Trends in atmospheric concentrations of CO2 (ppm), CH4 (ppb)
+and N2O (ppb), between 1800 and 2017 available at
+<https://www.eea.europa.eu/ds_resolveuid/217c026ca03649398aadf39d87623e57>
+
+Gustafson, E. J., A. M. G. De Bruijn, R. E. Pangle, J.-M. Limousin, N.
+G. McDowell, W. T. Pockman, B. R. Sturtevant, J. D. Muss, and M. E.
+Kubiske. 2015. Integrating ecophysiology and forest landscape models to
+improve projections of drought effects under climate change. Global
+Change Biology 21:843-856.
+
+Mina, M., Messier, C., Duveneck, M., Fortin, M. J., & Aquilué, N. 2021.
+Network analysis can guide resilience-based management in forest
+landscapes under global change. Ecological Applications, 31(1), 1--18.
+
+Petter G., Mairota P., Albrich K., Bebi P., Brůna J., Bugmann H.,
+Haffenden A., Scheller R.M., Schmatz D.R., Seidl R., Speich M.,
+Vacchiano G. & Lischke H. (2020) How robust are future projections of
+forest landscape dynamics? Insights from a systematic comparison of four
+forest landscape models. Environmental Modelling & Software 134: 104844.
+<https://doi.org/10.1016/j.envsoft.2020.104844>
+
+Puletti N., Mura M., Castaldi C., Marchi N., Chiavetta U. and Scotti R.
+(2014). ForIT: Functions from the 2nd Italian Forest Inventory (INFC). R
+package version 1.0. <https://CRAN.R-project.org/package=ForIT>
+
+Scheller, R. M., Domingo J. B. , Sturtevant B. R. , Williams J. S. ,
+Rudy A. , Gustafson E. J. & Mladenoff D. J. (2007) Design, development,
+and application of LANDIS-II, a spatial landscape simulation model with
+flexible temporal and spatial resolution. Ecological Modelling 201(3-4):
+409--419. <https://doi.org/10.1016/j.ecolmodel.2006.10.009>
+
+Tsubo M. & Walker S. (2005) Relationships between photosynthetically
+active radiation and clearness index at Bloemfontein, South Africa.
+Theoretical and Applied Climatology 80: 17--25.
+<https://doi.org/10.1007/s00704-004-0080-5>
+
+## TreeMig
+
+### Authors
+
+Heike Lischke
+
+### Model
+
+The model TreeMig (see treemig.wsl.ch) was originally based on the
+ForClim model, then aggregated by a distribution based approach (Lischke
+et al. 1998), and extended to a forest landscape model by including seed
+production, seed dispersal, seed bank dynamics and seedling
+establishment (Lischke and Loffler 2006, Lischke et al. 2006).
+
+### Climate
+
+As climatic input, the monthly time series of mean temperature and
+precipitation sum were used. With these climatic variables and
+additional data about slope, aspect and soil field capacity, yearly time
+series of the following bioclimatic variables were calculated: (1)
+DDSum, the yearly day degree sum above 5.5°C , (2) MinWiT, the lowest
+mean monthly temperature, (3) DrStr, drought stress according to a
+bucket model that estimates a monthly PET following (Thornthwaite and
+Mather 1957), using temperature, latitude, slope and aspect as input,
+and calculates the monthly soil water content based on soil water
+content in the previous month, precipitation, interception, and field
+capacity ("bucketsize") (Bugmann 1994, Fischlin et al. 1995). Drought
+stress is then given by 1- the ratio between demand (PET-interception)
+and the supply, a function of the soil water content. To get climate
+data also for the spin-up, bioclimate was calculated from the monthly
+data as given, and then extended by sampling from the given data from
+the end of the climate data to get 1000 years. The same sequence of
+sampling years was used in all simulations and sites.
+
+### Soil
+
+The soil bucketsize (available water capacity of the upmost 1m soil
+layer) required for the drought stress calculation, was derived from the
+soil quality data set, by Bucketsize = 10 + (soil_quality-1) /4 \* 25 ,
+to covert the range from 10 cm to 35cm
+
+### Topography
+
+Slope and aspect were taken from the topographic data set and converted
+to a common variable entering the PET calculation, by slasp=
+2*cos(aspect* pi/180) \*min(1, slope/60)
+
+### Tree species
+
+Thirty Central European tree species were simulated, including the
+standard species of the project . Only Betula pubescens was not
+simulated, because we did not have a parametrization for this species at
+hand. Species: Abies alba, Larix decidua, Picea abies, Pinus cembra,
+Pinus montana, Pinus sylvestris, Taxus baccata, Acer campestre, Acer
+platanoides, Acer pseudoplatanus, Alnus glutinosa, Alnus incana, Alnus
+viridis, Betula pendula, Carpinus betulus, Castanea sativa, Corylus
+avellana, Fagus sylvatica, Fraxinus excelsior, Populus nigra, Populus
+tremula, Quercus petraea, Quercus pubescens, Quercus robur, Salix alba,
+Sorbus aria, Sorbus aucuparia, Tilia cordata, Tilia platyphyllos, Ulmus
+glabra
+
+### Simulation
+
+The simulations were carried out on a grid of 15 x 15 cells, each with
+200 m side length.
+
+#### Dispersal
+
+The seeds produced by the mother trees are transported from the source
+cell according to a double negative exponential, species specific that
+determines the share of the seeds in the source cell landing in a sink
+cell in a given distance to the source cell. In the sink cells, the
+seeds enter the seed bank . At the start of the simulation for 10 years
+seeds of all species are available in all grid cells, from then on, the
+normal seed dispersal starts.
+
+#### Regeneration processes
+
+TreeMig simulates the full feedback from seed production to ingrowth.
+Some of the involved processes depend on environment. (1) Overall seed
+production in TreeMig is described by multiplying a species specific
+seed production per tree that depends on tree size and thus indirectly
+on the growth, which depends on DDsum, drought stress, nutrients and
+light and the adult tree number that depends on survival that depends on
+DDsum, drought stress, nutrients and light. (2) Seed dispersal is
+independent of environment. (3) Seed bank dynamics -- also environment
+independent - is given by seed input, seed mortality, and germination.
+Additionally, the number of seeds S of a species in the seed bank of
+this species (SB) is limited to a carrying capacity C of 1000, by\
+SB = C + (SB - C) \* Exp(-S/C).\
+(4) Germination and seedling survival depend on browsing, drought,
+winter temperature, DDSum and light. (5) The saplings up to 1.37 die and
+grow like adult trees, depending on DDsum, drought stress and light.
+Only the parameter value of the light dependence differs from that of
+the adult trees.\
+The regeneration processes were simulated annually in the order 1.
+germination, 2. germinated seeds leave the seed bank, 3. mortality of
+seeds in the seed bank, 4. seed entry by seed dispersal. The ingrowth
+numbers were very sensitive to this order. Simulations with another
+sequence, in which the seeds in the seed bank died first and then the
+remaining ones germinated, resulted in about half the ingrowth.
+
+#### Simulation length
+
+The simulations were started on bare ground and run for 1000 years, when
+at all sites the equilibrium was reached.
+
+#### More simulations per site
+
+TreeMig simulates directly the mean dynamics of entire stands, but based
+on the variability of light conditions within the stand [distribution
+based approach, Lischke, 1998, aggregation. To come up with the required
+200 random patches of 1 ha, temporal and spatial variability intrinsic
+to TreeMig were combined. Temporal variability was given by sampling
+from each decade of the last 100 years of the simulation, i.e. years
+900-1000. For each of theses 10 decades, 20 grid cells were sampled from
+the simulated grid, reflecting spatial variability, results from
+demographic stochasticity and short-range spatial interactions
+
+### Outputs
+
+In contrast to gap models, TreeMig is based on the concept of frequency
+distributions of tree densities classes in different tree heights on ca.
+1/12 ha (833m\^2) patches, and the resulting frequency distributions in
+discrete light. During the simulation, in each height class the
+frequencies of the different light classes were recorded, as well as the
+ingrowth of each species in the height-light class. By sampling 12 times
+from the height-light class ingrowths according to the height-light
+class frequency distribution and summing up, the stochastic ingrowths
+into all TreeMig height classes for one ha were calculated for each grid
+cell. Then the height class boundaries were translated into DBH and the
+ingrowths into the height classes linearly interpolated to 7cm and 10cm.
+The ingrowth basal areas into the height classes were obtained by
+multiplying the interpolated ingrowth numbers with the basal areas at
+7cm and 10cm, i.e. r.trees \* Pi\* (0.07/2)2 and r.trees \* Pi\*
+(0.1/2)2 . The species specific basal area (at 1.37m) sums are a
+standard output of TreeMig, given by multiplying the state variables
+(number of trees per grid cell, height class, and species) with the
+species specific basal area of each height class. To account for the
+ingrown trees dying within a decade, as an approximation the survival in
+each year to the power of 5 was used. Thereby, it was assumed that the
+trees have to survive in average 5 years in each decade, and that the
+current mortality is valid for these 5 years.
+
+### References
+
+• Bugmann, H. 1994. On the ecology of mountainous forests in a changing
+climate: A simulation study. Dissertation. Swiss Federal Institute of
+Technology, Zurich. • Fischlin, A., H. Bugmann, and D. Gyalistras. 1995.
+Sensitivity of a forest ecosystem model to climate parametrization
+schemes. Env. Poll. 87:267-282. • Lischke, H., and T. J. Loffler. 2006.
+Intra-specific density dependence is required to maintain species
+diversity in spatio-temporal forest simulations with reproduction.
+Ecological Modelling 198:341-361. • Lischke, H., T. J. Loffler, and A.
+Fischlin. 1998. Aggregation of individual trees and patches in forest
+succession models: Capturing variability with height structured, random,
+spatial distributions. Theor Popul Biol 54:213-226. • Lischke, H., N. E.
+Zimmermann, J. Bolliger, S. Rickebusch, and T. J. Loffler. 2006.
+TreeMig: A forest-landscape model for simulating spatio-temporal
+patterns from stand to landscape scale. Ecological Modelling
+199:409-420. • Thornthwaite, C. W., and J. R. Mather. 1957. The water
+balance. Publications in Climatology VIII:1-69.
+
+## LPJ-GUESS
+
+### Authors
+
+Jessica Hetzer, Thoms Hickler, Tim Anders
+
+### Model
+
+The model was originally developed by Ben Smith of Lund University in a
+collaboration also involving the Potsdam Institute for Climate Impact
+Research and the Max-Planck Institute for Biogeochemistry. Over the
+years, many people from institutes around the world have contributed to
+the refinement and further development of the model.
 
@@
 of LPJ-GUESS is available in Smith et al. 2001 and Smith et al. 2014.
+Source code is available on demand (see
+<https://web.nateko.lu.se/lpj-guess/>).
 
@@
 
-What climate data have you used, and how have you created your model
-climate input parameters? We used the provided daily time series of
-climatic variables from 1981 to 2018. The climatic variables base on the
-CHELSA data set with a spatial resolution of 0.008° x 0.008° (\~ 1 km)
-(Karger et al., 2021). Daily average temperature (tas) [°C], daily
-average maximum temperature (tasmax) [°C], daily average minimum
-temperature (tasmin) [°C], precipitation sum (pr) [mm] and surface solar
-(shortwave) radiation (rsds) were taken to set up the LPJ-GUESS
-simulation. Used climatic variables were converted to netCDF files. As
-LPJ-GUESS requires the climatic variables in specific units, temperature
-variables (tas, tasmax, tasmin) [°C] were converted to Kelvin and
-surface solar radiation (rsds) [J/m2] was converted to W/m2 by dividing
-rsds by 86400 s (3600 s/h \* 24 h). Nitrogen deposition was considered
-to be constant at a level of 10 kgN/ha/year. Atmospheric carbon dioxide
-concentration data from 1900 to 2018 was taken from the global carbon
-project (Quéré et al. 2018). We used the same default LPJ-GUESS soil
-code for a medium textured soil type (LPJ-GUESS soil class 2) for each
-grid cell (Sitch et al. 2003).
+We used the provided daily time series of climatic variables from 1981
+to 2018. The climatic variables base on the CHELSA data set with a
+spatial resolution of 0.008° x 0.008° (\~ 1 km) (Karger et al., 2021).
+Daily average temperature (tas) [°C], daily average maximum temperature
+(tasmax) [°C], daily average minimum temperature (tasmin) [°C],
+precipitation sum (pr) [mm] and surface solar (shortwave) radiation
+(rsds) were taken to set up the LPJ-GUESS simulation. Used climatic
+variables were converted to netCDF files. As LPJ-GUESS requires the
+climatic variables in specific units, temperature variables (tas,
+tasmax, tasmin) [°C] were converted to Kelvin and surface solar
+radiation (rsds) [J/m2] was converted to W/m2 by dividing rsds by 86400
+s (3600 s/h \* 24 h). Nitrogen deposition was considered to be constant
+at a level of 10 kgN/ha/year. Atmospheric carbon dioxide concentration
+data from 1900 to 2018 was taken from the global carbon project (Quéré
+et al. 2018).
 
@@
 
+All sites We used were simulated withdefault values for a medium
+textured soil type (sand: 0.35, clay: 0.15, silt: 0.5) in LPJ-GUESS.
+
 ### Topography
 
+Topography was not considered.
+
 ### Tree species
 
-How have you defined the species composition in your model simulation?
-(e.g., species levels, PFTs, etc.)? Please describe the species
-categories used. LPJ-GUESS can be run in different modes, with different
-levels of abstraction of the population and community processes. For
-this study, the more detailed 'cohort' mode was used, in which
-individuals, patches and vertical canopy structure are represented
-explicitly, but living individuals within a cohort (age class) of a
-given tree species in a given patch are assumed to be identical (in
-terms of all state variables, such as height and stem diameter).
-Simulations consider the main tree species as well as other PFTs
-yielding an appropriate representation of European vegetation, including
-only tree and shrub species that are widely distributed across Europe
-and can become dominant in some areas (Hickler et al., 2012).
-Mediterranean rain-green small shrubs with shallow roots (e.g. Lavendula
-spp., Rosmarinus spp.) and alpine/arctic shrubs (e.g. Vaccinium spp.)
-were represented as PFTs in the model. As in earlier applications of
-LPJ-GUESS, herbaceous vegetation was represented by two 'generic herb'
-PFTs, with C3 and C4 photosynthesis, respectively. The final set
-included 16 tree species, one Mediterranean shrub PFT, one boreal/alpine
-shrub PFT, and the two herbaceous PFTs described earlier (Table 2). Acer
+LPJ-GUESS can be run in different modes, with different levels of
+abstraction of the population and community processes. For this study,
+the more detailed 'cohort' mode was used, in which individuals, patches
+and vertical canopy structure are represented explicitly, but living
+individuals within a cohort (age class) of a given tree species in a
+given patch are assumed to be identical (in terms of all state
+variables, such as height and stem diameter). Simulations consider the
+main tree species as well as other PFTs yielding an appropriate
+representation of European vegetation, including only tree and shrub
+species that are widely distributed across Europe and can become
+dominant in some areas (Hickler et al., 2012). Mediterranean rain-green
+small shrubs with shallow roots (e.g. Lavendula spp., Rosmarinus spp.)
+and alpine/arctic shrubs (e.g. Vaccinium spp.) were represented as PFTs
+in the model. As in earlier applications of LPJ-GUESS, herbaceous
+vegetation was represented by two 'generic herb' PFTs, with C3 and C4
+photosynthesis, respectively. The final set included 16 tree species,
+one Mediterranean shrub PFT, one boreal/alpine shrub PFT, and the two
+herbaceous PFTs described earlier (Table 2). To increase the chances of
+shade-intolerant species becoming established, patch-destroying
+disturbances were allowed as described in Hickler et al. 2012. Acer
 pseudoplatanus and Alnus glutinosa are not parameterized in LPJ-GUESS
-and therefore we did not simulate these species.
+and therefore not considered.we did not simulate these species.
 
@@
 
-How have you defined the simulation area for your model (e.g., size and
-other relevant site aspects)? In this study, vegetation is simulated in
-so called patches that represent forest areas of fixed size. Patches can
-be regarded as sample stands in a grid cell, each of which represents an
-0.1 ha (comparable to figure 1 of the tree regeneration workshop
-protocol). In this study, vegetation dynamics were simulated by 2000
-replicate patches each 0.1 ha in size. To match the required sample
-size, we processed the data so that ten 0.1 ha patches were merged to
-one 1 ha sample.
+In this study, vegetation is simulated in so called patches that
+represent forest areas of fixed size. Patches can be regarded as sample
+stands in a grid cell, each of which represents an 0.1 ha (comparable to
+figure 1 of the tree regeneration workshop protocol). In this study,
+vegetation dynamics were simulated by 2000 replicate patches each 0.1 ha
+in size. To match the required sample size, we processed the data so
+that ten 0.1 ha patches were merged to one 1 ha sample.
 
@@
 
-What are the relevant aspects of the dispersal (if any) and seed input
-parameters in your model? Model formulations of establishment are based
-on those employed within the 'forest gap' model FORSKA (Leemans &
-Prentice, 1989; Prentice et al., 1993). The number of new saplings of
-each woody PFT/species and in each patch each year is drawn at random
-from the Poisson distribution, with an expectation influenced by a
-PFT/species-specific maximum establishment rate and by the 'propagule
-pool', i.e. the amount of carbon allocated to reproduction by all
-individuals of the PFT/species at all patches in the previous year. No
-saplings are established in a given patch if the minimum PAR level at
-the forest floor is below a PFT/species-specific threshold, which is
-higher for more light-demanding PFTs/species.
+Model formulations of establishment are based on those employed within
+the 'forest gap' model FORSKA (Leemans & Prentice, 1989; Prentice et
+al., 1993). The number of new saplings of each woody PFT/species and in
+each patch each year is drawn at random from the Poisson distribution,
+with an expectation influenced by a PFT/species-specific maximum
+establishment rate and by the 'propagule pool', i.e. the amount of
+carbon allocated to reproduction by all individuals of the PFT/species
+at all patches in the previous year. No saplings are established in a
+given patch if the minimum PAR level at the forest floor is below a
+PFT/species-specific threshold, which is higher for more light-demanding
+PFTs/species.
 
 Table 1: Parameter settings associated with vegetation establishment.
-Parameter Description Parameter setting in this study ifbgestab Whether
-background establishment enabled (0,1) 1 ifsme Whether spatial mass
-effect enabled for establishment (0,1) 1 ifstochmort Whether stochastic
-mortality is enabled (0,1) 1 ifstochestab Whether stochastic
-establishment is enabled (0,1) 1 ifdroughtlimitedestab Whether
-establishment is limited by growing season drought (0,1) 1 tcmin_est Min
-20-year coldest month mean temp for establishment (deg C) PFT/species
-specific (see table 2 and 3) tcmax_est Max 20-year coldest month mean
-temp for establishment (deg C) PFT/species specific (see table 2 and 3)
-twmin_est Min warmest month mean temp for establishment (deg C)
-PFT/species specific (see table 2 and 3) twminusc Continentality
-parameter for boreal summer-green trees 0 gdd5min_est Min GDD on 5 deg C
-base for establishment PFT/species specific (see table 2 and 3) est_max
-Max sapling establishment rate (indiv/m2/year) PFT/species specific (see
-table 2 and 3) gdd0_min min GDD0 for survival/establishment reestab Pft
-reestablishment ("NONE", "RESTRICTED" or "ALL") ALL
+
+![](images/image-1265081375.png)
 
@@
 What is the simulation length, and how have you determined the
-simulation time until the equilibrium? The simulations were initialized
-from bare ground (no biomass) and the model was spun up for 500 years
-until the modelled vegetation was in approximate equilibrium with the
-climate and CO2 around the year 1900 (Smith et al., 2001; Hickler et
-al., 2012).
+simulation time until the equilibrium?
+
+The simulations were initialized from bare ground (no biomass) and the
+model was spun up for 500 years until the modelled vegetation was in
+approximate equilibrium with the climate and CO2 around the year 1900
+(Smith et al., 2001; Hickler et al., 2012).
 
@@
 
-### Output v
+### Output
 
 Please describe how have you calculated each of the variables in the
-output data. The generated output file is called
-Output_SGN_LPJGUESS.txt. Each row of the output dataset contains
-variables regarding the decade 2008-2018. The first two columns describe
-the study site (column "site", values reflect site ID) and the 1 ha
-samples of that specific site (column "sample", values range from 1 to
-200). Output variables are species-specific. Although we simulated a
-total of 20 European tree species and PFTs, our output file considers
-only the species required by the protocol (column "species", values are
-the abbreviations for the species name ("abal" for Abies alba, "cabe"
-for Carpinus betulus, "fasy" for Fagus sylvatica, "frex" for Fraxinus
-excelsior, "pabi" for Picea abies, "pisy" for Pinus sylvestris , "tico"
-for Tilia cordata , "betu_pen" for Betula pendula, "betu_pen" for Betula
-pubescens, "quer_coc" for Quercus coccifera, "quer_il" for Quercus ilex,
-"quer_pub" for Quercus pubescens, and "quer_rob" for Quercus robur).
-Since Acer pseudoplatanus ("acps") and Alnus glutinosa ("algl") are not
-parameterized for LPJ-GUESS so far, we have not included these species
-in the output. The first variable is the number of 7 cm recruits
+output data.
+
+The generated output file is called Output_SGN_LPJGUESS.txt. Each row of
+the output dataset contains variables regarding the decade 2008-2018.
+The first two columns describe the study site (column "site", values
+reflect site ID) and the 1 ha samples of that specific site (column
+"sample", values range from 1 to 200). Output variables are
+species-specific. Although we simulated a total of 20 European tree
+species and PFTs, our output file considers only the species required by
+the protocol (column "species", values are the abbreviations for the
+species name ("abal" for Abies alba, "cabe" for Carpinus betulus, "fasy"
+for Fagus sylvatica, "frex" for Fraxinus excelsior, "pabi" for Picea
+abies, "pisy" for Pinus sylvestris , "tico" for Tilia cordata ,
+"betu_pen" for Betula pendula, "betu_pen" for Betula pubescens,
+"quer_il" for Quercus ilex, "quer_pub" for Quercus pubescens, and
+"quer_rob" for Quercus robur). Since Acer pseudoplatanus ("acps") and
+Alnus glutinosa ("algl") are not parameterized for LPJ-GUESS so far, we
+have not included these species in the output. Further we excluded
+shrubs (alpine/arctic shrubs and Quercus coccifera) and herbaceous
+vegetation from the final output as this study focusses on tree
+regeneration only. The first variable is the number of 7 cm recruits
 "r_7cm.trees" calculated as the sum of all trees of the same species
@@
 maximum woody crown area. Parameter that are associated with
-establishment processes are marked in bold.
-
- Table 3: Parameters associated with different shade tolerance classes
-(updates compared to Hickler et al. 2012). Parameter that are associated
-with establishment processes are marked in bold.
+establishment processes are marked in bold. Species/PFT
+
+![](images/image-2144349832.png)
 
@@
 
-Sitch, S., Smith, B., Prentice, I. C., Arneth, A., Bondeau, A., Cramer,
-W., ... & Venevsky, S. (2003). Evaluation of ecosystem dynamics, plant
-geography and terrestrial carbon cycling in the LPJ dynamic global
-vegetation model. Global change biology, 9(2), 161-185.
-
 Smith, B. (2001). LPJ-GUESS-an ecosystem modelling framework. Department
@@
 
-Simon Scheiter
-Senckenberg Biodiversity and Climate Research Centre, Senckenberganlage 25, 60325 Frankfurt am Main, Germany
+Simon Scheiter Senckenberg Biodiversity and Climate Research Centre,
+Senckenberganlage 25, 60325 Frankfurt am Main, Germany
 
@@
 
-We used aDGVM2, an individual-based dynamic vegetation model that is based on concepts from community assembly theory and uses a functional trait approach (Langan et al., 2017; Scheiter et al., 2013).
-The model simulates growth, reproduction, and mortality of individual plants while keeping track of state variables, such as biomass, height, and leaf area.
-Each plant in aDGVM2 is characterized by a plant-specific set of trait values.
-Traits describe growth form, leaf characteristics, hydraulic characteristics, resource allocation, architecture, reproduction, mortality, and response to disturbance.
-Most plant traits are linked by trade-offs to constrain possible trait combinations.
-Selection and trait inheritance assemble plant communities that are adapted to biotic and abiotic conditions.
-Plants with trait combinations that allow sufficient growth and reproduction rates, and that allow plants to cope with competition and disturbances can contribute their trait values to the community trait pool.
-Trait mutation and recombination may alter trait values in the community trait pool.
-Randomly drawn seeds from the trait pool are added to the plant population and new plants start growing from seeds.
-Plants that are not adapted to the prevailing disturbance regimes, biotic and abiotic conditions, or that do not allocate enough carbon to reproduction disappear from the population.
-Therefore, successful ecological strategies emerge dynamically from these community assembly and reproduction process.
-Due to the trait-based approach, species or plant functional types (PFTs) are not pre-defined and hard-coded in aDGVM2.
-Rather, different PFTs or ecological strategies that are adjusted to the prevailing biotic and abiotic conditions emerge from community assembly processes (Scheiter et al., 2013), and plants can be classified into PFTs or ecological strategies based on their trait values in a post-processing step.
-The aDGVM2 has primarily been developed for tropical and sub-tropical ecosystems and it has not been tested and benchmarked for Europe.
-We applied the model in the original version.
-Data-model comparisons with model results were not done and the model was not changed to improve agreement with any data for Europe.
+We used aDGVM2, an individual-based dynamic vegetation model that is
+based on concepts from community assembly theory and uses a functional
+trait approach (Langan et al., 2017; Scheiter et al., 2013). The model
+simulates growth, reproduction, and mortality of individual plants while
+keeping track of state variables, such as biomass, height, and leaf
+area. Each plant in aDGVM2 is characterized by a plant-specific set of
+trait values. Traits describe growth form, leaf characteristics,
+hydraulic characteristics, resource allocation, architecture,
+reproduction, mortality, and response to disturbance. Most plant traits
+are linked by trade-offs to constrain possible trait combinations.
+Selection and trait inheritance assemble plant communities that are
+adapted to biotic and abiotic conditions. Plants with trait combinations
+that allow sufficient growth and reproduction rates, and that allow
+plants to cope with competition and disturbances can contribute their
+trait values to the community trait pool. Trait mutation and
+recombination may alter trait values in the community trait pool.
+Randomly drawn seeds from the trait pool are added to the plant
+population and new plants start growing from seeds. Plants that are not
+adapted to the prevailing disturbance regimes, biotic and abiotic
+conditions, or that do not allocate enough carbon to reproduction
+disappear from the population. Therefore, successful ecological
+strategies emerge dynamically from these community assembly and
+reproduction process. Due to the trait-based approach, species or plant
+functional types (PFTs) are not pre-defined and hard-coded in aDGVM2.
+Rather, different PFTs or ecological strategies that are adjusted to the
+prevailing biotic and abiotic conditions emerge from community assembly
+processes (Scheiter et al., 2013), and plants can be classified into
+PFTs or ecological strategies based on their trait values in a
+post-processing step. The aDGVM2 has primarily been developed for
+tropical and sub-tropical ecosystems and it has not been tested and
+benchmarked for Europe. We applied the model in the original version.
+Data-model comparisons with model results were not done and the model
+was not changed to improve agreement with any data for Europe.
 
@@
 
-The aDGVM2 simulates vegetation at daily temporal resolution. We therefore used daily climate input data for the period 1981-2018 according to the modeling protocol.
+The aDGVM2 simulates vegetation at daily temporal resolution. We
+therefore used daily climate input data for the period 1981-2018
+according to the modeling protocol.
 
@@
 
-Soil quality data was converted to maximum rooting depth.
-Specifically, the soil quality value multiplied by 2 and the result was used as maximum rooting depth.
-We used the multiplication, as we typically use maximum soil depth and maximum rooting depth of 10m.
-Previous model simulations showed that rooting depth and soil dept have strong impacts on vegetation dynamics by influencing water availability to plants (Langan et al., 2017).
-We therefore considered it as adequate representation of soil quality.
-Other soil characteristics including volumetric water-holding capacity, soil hydraulic conductivity, soil bulk density, soil depth, soil texture, soil carbon content, soil wilting point and field capacity were taken from the FAO (http://www.fao.org, Nachtergaele et al., 2009).
+Soil quality data was converted to maximum rooting depth. Specifically,
+the soil quality value multiplied by 2 and the result was used as
+maximum rooting depth. We used the multiplication, as we typically use
+maximum soil depth and maximum rooting depth of 10m. Previous model
+simulations showed that rooting depth and soil dept have strong impacts
+on vegetation dynamics by influencing water availability to plants
+(Langan et al., 2017). We therefore considered it as adequate
+representation of soil quality. Other soil characteristics including
+volumetric water-holding capacity, soil hydraulic conductivity, soil
+bulk density, soil depth, soil texture, soil carbon content, soil
+wilting point and field capacity were taken from the FAO
+(<http://www.fao.org>, Nachtergaele et al., 2009).
 
@@
 
-We used elevation according to the modeling protocol.
-In aDGVM2, elevation is used to calculate atmospheric pressure which in turn influences ecophysiological processes related to water and carbon fluxes at leaf and canopy level.
-Information on slope and aspect was not used, as related processes (e.g., impacts on runoff or radiation balance) are not implemented in aDGVM2.
+We used elevation according to the modeling protocol. In aDGVM2,
+elevation is used to calculate atmospheric pressure which in turn
+influences ecophysiological processes related to water and carbon fluxes
+at leaf and canopy level. Information on slope and aspect was not used,
+as related processes (e.g., impacts on runoff or radiation balance) are
+not implemented in aDGVM2.
 
@@
 
-The aDGVM2 does not simulate pre-defined species or PFTs.
-The 11 tree species included in the modeling protocol cannot be implemented and parametrized with reasonable effort.
-We therefore classified simulated trees into different ecological strategies as a post-processing step.
-We used all combinations of evergreen/deciduous, light/water triggered phenology, low/high SLA.
-Evergreen/deciduous phenology and light/water triggered phenology are traits implemented in aDGVM2.
-The threshold to distinguish high and low SLA was defined as 20 m2/kg.
-We used these traits as previous simulation experiments showed that patterns of these traits emerge along environmental gradients while patterns of other traits implemented in aDGVM2 are often less clear.
-Ecological strategies considered are: low SLA, deciduous, rain-triggered; low SLA, deciduous, light-triggered; low SLA, evergreen, rain-triggered; low SLA, evergreen, light-triggered; high SLA, deciduous, rain-triggered; high SLA, deciduous, light-triggered; high SLA, evergreen, rain-triggered; high SLA, evergreen, light-triggered.
+The aDGVM2 does not simulate pre-defined species or PFTs. The 11 tree
+species included in the modeling protocol cannot be implemented and
+parametrized with reasonable effort. We therefore classified simulated
+trees into different ecological strategies as a post-processing step. We
+used all combinations of evergreen/deciduous, light/water triggered
+phenology, low/high SLA. Evergreen/deciduous phenology and light/water
+triggered phenology are traits implemented in aDGVM2. The threshold to
+distinguish high and low SLA was defined as 20 m2/kg. We used these
+traits as previous simulation experiments showed that patterns of these
+traits emerge along environmental gradients while patterns of other
+traits implemented in aDGVM2 are often less clear. Ecological strategies
+considered are: low SLA, deciduous, rain-triggered; low SLA, deciduous,
+light-triggered; low SLA, evergreen, rain-triggered; low SLA, evergreen,
+light-triggered; high SLA, deciduous, rain-triggered; high SLA,
+deciduous, light-triggered; high SLA, evergreen, rain-triggered; high
+SLA, evergreen, light-triggered.
 
@@
 
-Simulations followed the modeling protocol.
-Simulations were conducted for all sites using the provided daily climate and soil data.
-By default, aDGVM2 simulates 1ha stands.
-To implement the modeling protocol, adjustments to scale model results to 1 ha level were therefore not required.
-The model was initialized by default initialization routines, i.e., traits of individual plants were drawn randomly from a pre-defined range of values for each trait.
-Ranges of trait values are provided in Langan et al.
-(2017).
-The model was initialized with bare ground and 2880 tree seedlings with 100g each.
-The aDGVM2 simulates both trees and grasses.
-As the simulation experiment aims at tree/forest regeneration, simulations were conducted only with trees while grasses were removed.
-
-We conducted 5 replicates for each of the 200 sites to account for stochastic processes in aDGVM2.
-Stochastic processes include random initialization of trait values of initial plant populations, demographic processes such as plant mortality or selection of seeds from seed bank, or the community assembly processes including mutation and cross-over.
-Climate time series used for different replicates were identical such that variation between replicates can be attributed by stochastic processes in aDGVM2.
-
-Although aDGVM2 includes routines to simulate fire, simulations were conducted in the absence of fire. Management was not simulated.
+Simulations followed the modeling protocol. Simulations were conducted
+for all sites using the provided daily climate and soil data. By
+default, aDGVM2 simulates 1ha stands. To implement the modeling
+protocol, adjustments to scale model results to 1 ha level were
+therefore not required. The model was initialized by default
+initialization routines, i.e., traits of individual plants were drawn
+randomly from a pre-defined range of values for each trait. Ranges of
+trait values are provided in Langan et al. (2017). The model was
+initialized with bare ground and 2880 tree seedlings with 100g each. The
+aDGVM2 simulates both trees and grasses. As the simulation experiment
+aims at tree/forest regeneration, simulations were conducted only with
+trees while grasses were removed.
+
+We conducted 5 replicates for each of the 200 sites to account for
+stochastic processes in aDGVM2. Stochastic processes include random
+initialization of trait values of initial plant populations, demographic
+processes such as plant mortality or selection of seeds from seed bank,
+or the community assembly processes including mutation and cross-over.
+Climate time series used for different replicates were identical such
+that variation between replicates can be attributed by stochastic
+processes in aDGVM2.
+
+Although aDGVM2 includes routines to simulate fire, simulations were
+conducted in the absence of fire. Management was not simulated.
 
 ### Dispersal
+
 The aDGVM2 does not simulate seed dispersal.
@@
 
-We first conducted a 600-year spin-up with a randomized order of years 1981-2018 provided in the CHELSA daily climate data.
-Spin-up is required to ensure that both state variables such as biomass and tree cover and community trait composition have stabilized and are in a dynamic equilibrium with climate.
-Previous model simulations showed that a 600-year spin-up is sufficient to reach such an equilibrium state (Langan et al., 2017).
-Following the spin-up, we simulated transient vegetation dynamics with daily CHELSA climate data for the period 1981-2018.
+We first conducted a 600-year spin-up with a randomized order of years
+1981-2018 provided in the CHELSA daily climate data. Spin-up is required
+to ensure that both state variables such as biomass and tree cover and
+community trait composition have stabilized and are in a dynamic
+equilibrium with climate. Previous model simulations showed that a
+600-year spin-up is sufficient to reach such an equilibrium state
+(Langan et al., 2017). Following the spin-up, we simulated transient
+vegetation dynamics with daily CHELSA climate data for the period
+1981-2018.
 
@@
 
-To create output variables, we randomly selected 10-year periods within the transient phase (i.e., period 1981-2018 after spin-up phase) and calculated averages of all required output variables following the modeling protocol.
-The selected 10-year periods differed between sites and replicates.
-
-The aDGVM2 does not simulate pre-defined species or PFTs and trees were classified into different ecological strategies (see section Tree species).
-Codes for the ecological strategies in the species column of the output files are:
-1 - low SLA, deciduous, rain-triggered
-2 - low SLA, deciduous, light-triggered
-3 - low SLA, evergreen, rain-triggered
-4 - low SLA, evergreen, light-triggered
-5 - high SLA, deciduous, rain-triggered
-6 - high SLA, deciduous, light-triggered
-7 - high SLA, evergreen, rain-triggered
-8 - high SLA, evergreen, light-triggered
-
-Simulations were conducted for all 200 sites. We conducted 5 replicates per site to account for stochastic processes in aDGVM. In the output files, replicates are given as follows: 
-Replicate 1 – sample 1:200
-Replicate 2 – sample 201:400
-Replicate 3 – sample 401:600
-Replicate 4 – sample 601:800
-Replicate 5 – sample 801:1000
+To create output variables, we randomly selected 10-year periods within
+the transient phase (i.e., period 1981-2018 after spin-up phase) and
+calculated averages of all required output variables following the
+modeling protocol. The selected 10-year periods differed between sites
+and replicates.
+
+The aDGVM2 does not simulate pre-defined species or PFTs and trees were
+classified into different ecological strategies (see section Tree
+species). Codes for the ecological strategies in the species column of
+the output files are: 1 - low SLA, deciduous, rain-triggered 2 - low
+SLA, deciduous, light-triggered 3 - low SLA, evergreen, rain-triggered
+4 - low SLA, evergreen, light-triggered 5 - high SLA, deciduous,
+rain-triggered 6 - high SLA, deciduous, light-triggered 7 - high SLA,
+evergreen, rain-triggered 8 - high SLA, evergreen, light-triggered
+
+Simulations were conducted for all 200 sites. We conducted 5 replicates
+per site to account for stochastic processes in aDGVM. In the output
+files, replicates are given as follows: Replicate 1 -- sample 1:200
+Replicate 2 -- sample 201:400 Replicate 3 -- sample 401:600 Replicate 4
+-- sample 601:800 Replicate 5 -- sample 801:1000
 
@@
 
-Langan L; Higgins SI; Scheiter S (2017) Climate-biomes, pedo-biomes or pyro-biomes: which world view explains the tropical forest - savanna boundary in South America?
-JOURNAL OF BIOGEOGRAPHY, 44, 2319-2330.
-
-Nachtergaele, F., van Velthuizen, H., Verelst, L., Batjes, N., Dijk- shoorn, K., Van Engelen, V., Fischer, G., Jones, A., Montanarella, L., and Petri, M.: Harmonized world soil database (version 1.1), FAO, Rome, Italy, IIASA, Laxenburg, Austria, 2009.
-
-Scheiter S, Langan L, Higgins SI (2013) Next generation dynamic global vegetation models: learning from community ecology. NEW PHYTOLOGIST, 198, 957-969.
-
+Langan L; Higgins SI; Scheiter S (2017) Climate-biomes, pedo-biomes or
+pyro-biomes: which world view explains the tropical forest - savanna
+boundary in South America? JOURNAL OF BIOGEOGRAPHY, 44, 2319-2330.
+
+Nachtergaele, F., van Velthuizen, H., Verelst, L., Batjes, N., Dijk-
+shoorn, K., Van Engelen, V., Fischer, G., Jones, A., Montanarella, L.,
+and Petri, M.: Harmonized world soil database (version 1.1), FAO, Rome,
+Italy, IIASA, Laxenburg, Austria, 2009.
+
+Scheiter S, Langan L, Higgins SI (2013) Next generation dynamic global
+vegetation models: learning from community ecology. NEW PHYTOLOGIST,
+198, 957-969.
+
+```{=tex}
 \newpage
 \beginsupplement
-
+```
 # Supplementary Material 3: Supplementary figures and tables
@@
 
-Summary of the linear model of the overestimation proportions versus the mean complexity value on the establishment module.
+Summary of the linear model of the overestimation proportions versus the
+mean complexity value on the establishment module.
+
 ```{r tableComp, echo=FALSE,  message=FALSE, render = 'normal_print'}

### 2023-01-27 manuscript/manuscriptRW.Rmd
added changes based on co-authors comments in the M&M section (not finnished yet)

@@
 
-Most of the models overestimate ingrowth rates, but this is not always
-compensated by high simulated mortality rates in the early stages of
-population dynamics. The effect of light availability on ingrowth rates
-is captured better than the effect of temperature and soil moisture, but
-patterns are not consistent across models. Simulated species diversity
-of ingrowth matches the observed ranges quite well. Models simulating
-higher species diversity at the stand level do not feature higher
-ingrowth diversity. The ingrowth of *Abies alba* is captured across
-temperature and soil moisture gradients, whereas ingrowth of *Fagus
-sylvatica* is generally underestimated. *Picea abies* ingrowth is
-generally overestimated, particularly at colder sites. Increasing
-complexity in the tree establishment modules of the models is not
-related to higher accuracy of simulated tree establishment. Furthermore,
-individual model design is more important than scale (stand, landscape,
-global) and approach (empirical, process-based) to accurately capture
-tree establishment.
+Most of the models overestimate ingrowth rates, which is only
+compensated by some models with high simulated mortality rates in the
+early stages of population dynamics. The effect of light availability on
+ingrowth rates is captured better than the effect of temperature and
+soil moisture, but patterns are not consistent across models. Simulated
+species diversity of ingrowth matches the observed ranges quite well.
+Models simulating higher species diversity at the stand level do not
+feature higher ingrowth diversity. The ingrowth of *Abies alba* is
+captured across temperature and soil moisture gradients, whereas
+ingrowth of *Fagus sylvatica* is generally underestimated. *Picea abies*
+ingrowth is generally overestimated, particularly at colder sites.
+Increasing complexity in the tree establishment modules of the models is
+not related to higher accuracy of simulated tree establishment.
+Furthermore, individual model design is more important than scale
+(stand, landscape, global) and approach (empirical, process-based) to
+accurately capture tree establishment.
 
@@
 models commonly used to assess forest dynamics under climate change are
-evaluated against a continental-scale, multi-species harmonized data set
+evaluated against a continental-scale, multi-species harmonized dataset
 on tree establishment [@käber2022]. This is a challenging experiment for
@@
 
-Fifteen models of forest dynamics (including two model featuring
-variants) were used to assess tree establishment. They simulate forest
-dynamics at the stand, landscape and global scale (Table
+Fifteen models of forest dynamics (including two models featuring
+variants) were used to assess simulated tree establishment. They
+simulate forest dynamics at the stand, landscape or global scale (Table
 \@ref(tab:tableModels)). The approaches used for model construction and
@@
 formulations derived from the statistical analysis of inventory data
-(SYBILA and xComp). Some of the models largely rely on the original
+(SIBYLA and xComp). Some of the models largely rely on the original
 approach underlying forest gap models (e.g., ForClim 1 and LandClim),
@@
 \@ref(tab:tableModels)). Regeneration models include processes such as
-flowering and pollination, seed production, seed dispersal, germination
-seedling growth, which ultimately lead to the simulated number of trees.
-Recruitment models, in contrast, introduce a certain number of new trees
-with certain characteristics such as biomass or diameter, without
-explicitly considering earlier development processes. We can further
-distinguish models that feature a feedback in the establishment, i.e.,
-where the level and species composition of establishment is influenced
-by the existing adult tree community via the production of seeds,
-seedlings or saplings, from models that do not contain such feedback.
-
-The starting point for in the models differs as well, ranging from
-seedbank, seed or seedling to sapling (i.e., trees much larger than 10
-cm height, often at least 2-3 m tall). This implies that the larger the
-size threshold, the larger the number of ecological processes that must
-be aggregated via parameterizations because they cannot be treated
-explicitly. This reduces model complexity but comes at the cost of
-blurring process representation.
-
-The overall complexity in the establishment modules also varies
-considerably. Following @bugmann2022, we can classify seven models
-(iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and TreeMig) as
-having rather high complexity in their establishment modules (average
-complexity across all processes \>0, Table \@ref(tab:tableModels)). In
-most models, the two processes in the establishment modules with the
-highest complexity are the statistical approach used to define the
-successful number of new trees and how frost occurrence is taken into
-account.
+flowering and pollination, seed production, seed dispersal, germination,
+and seedling growth, which ultimately lead to the simulated number of
+established trees. Recruitment models, in contrast, introduce a certain
+number of new trees with certain characteristics such as biomass or
+diameter, without explicitly considering earlier development processes.
+We can further distinguish models that feature a feedback in the
+establishment, i.e., where the level and species composition of
+establishment is influenced by the existing adult tree community via the
+production of seeds, seedlings or saplings, from models that do not
+contain such a feedback.
+
+The starting point for the models establishment differs as well, ranging
+from seedbank, seed or seedling to sapling (i.e., trees much larger than
+10 cm height, often at least 2-3 m tall). Models that start from seed
+need to include a larger number of ecological processes such as
+germination and survival of young seedlings, whereas models that started
+from saplings have to aggregate via parameterizations several ecological
+processes because they are not treated explicitly. This reduces model
+complexity but comes at the cost of blurring process representation. In
+the models used in this study the overall complexity in the
+establishment modules varies considerably. Following @bugmann2022, we
+can classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS,
+ForClim and TreeMig) as having rather high complexity in their
+establishment modules (mean establishment formulation complexity
+complexity across all processes \>0, Table
+\@ref(tab:)\@ref(tab:tableModels)).
 
@@
 structure is identical. Similarly, the two ForCEEPS variants allow us to
-isolate the importance of the canopy feedback via seed trees for the
-quantity and quality of simulated ingrowth, as one variant includes this
-feedback whereas the other does not.
+isolate the importance of the canopy feedback (i.e. simulated actual
+composition and relative abundance of species in the plot)via seed trees
+for the quantity and quality of simulated ingrowth, as one variant
+includes this feedback whereas the other does not.
 
@@
 dynamics have never been confronted with a dataset covering such
-gradients over a large number sites, to evaluate how well ingrowth is
-captured. The observations used here are from a novel and unprecedented
-network of sites in forest reserves that represent the range of
-environmental gradients in temperature and precipitation in central
-Europe as compiled in the framework of the EuFoRIa network
-[@euforia2019] (Figure \@ref(fig:map)). All these forests have not been
-managed for at least 10 years prior to their designation as reserves,
-and most of them provide time series of natural forest dynamics over
-multiple decades and up to ca. 80 years. The census periods range from 3
-to 37 years, with an average of 14 years. The data provide information
-at the tree level, thus allowing for sequential comparison such as
+gradients over a large number of sites, to evaluate how well ingrowth is
+captured. The observations used here are derived from a novel and
+unprecedented network of sites in forest reserves that represent the
+range of environmental gradients in temperature and precipitation in
+central Europe as compiled in the framework of the EuFoRIa network
+[@euforia2019] (Figure \@ref(fig:map)). These forests are unmanged for
+at least 10 years prior to their designation as reserves, and most of
+them provide time series of natural forest dynamics over multiple
+decades and up to ca. 80 years. The census periods range from 3 to 37
+years, with an average of 14 years. The data provide information at the
+tree level, thus allowing for sequential comparison such as
 individual-tree ingrowth and death between the measurements.
 
-We selected 200 sites from this network as the benchmarking data set for
+We selected 200 sites from this network as the benchmarking dataset for
 the simulation. The selection of sites is representative for the
 environmental variation of the data. This was achieved by applying
-kmeans clustering to define 200 clusters of plots from the original set
+k-means clustering to define 200 clusters of plots from the original set
 of 869 plots along the environmental dimensions of temperature, water
@@
 these sites featured at least two consecutive measurements. Ingrowth was
-measured at threshold of 7 or 10 cm of stem diameter at all sites except
-for 35 sites where only ingrowth at the 10 cm diameter threshold was
-available. Plot size ranged from 0.02 ha to 5.52 ha, and the data were
-further processed and aggregated following @käber2022 to provide
-ingrowth rates per ha and per decade. The observations used in this
-study featured 30.900 established trees, and ingrowth rates per site,
-sample and decade ranged from zero to 1246 trees per ha, with a mean of
-56 trees per ha. For more details on this unique dataset and the
-detailed establishment information, cf. @käber2022.
+measured at threshold of 7 and 10 cm of stem diameter at breast height
+in all sites except for 35 sites where only ingrowth at the 10 cm
+diameter threshold was available. Plot size ranged from 0.02 ha to 5.52
+ha, and the data were further processed and aggregated following
+@käber2022 to provide ingrowth rates per ha and per decade. The
+observations used in this study featured 30,900 established trees, and
+ingrowth rates per site, sample and decade ranged from zero to 1246
+trees per ha, with a mean of 56 trees per ha. For more details on this
+unique dataset and the detailed establishment information, cf.
+@käber2022.
 
@@
 wide range of models of forest dynamics. We define the regeneration
-niche as the passing of a diameter threshold of 7 or 10 cm, respectively
-(synonym: ingrowth). To this end, each modeling group was provided with
-a detailed protocol (cf. Supplementary Material 1, henceforth "SM") with
-instructions how to perform the simulations, providing input variables
-on climate and soil conditions, and the expected output variables.
-Neither were further site information (except for the data specified
-below) nor any data on tree ingrowth provided prior to the simulation.
-That is, the models were run in "blind flight" mode.
+niche as the passing of a breast height diameter threshold of 7 or 10
+cm, respectively (synonym: ingrowth). To this end, each modeling group
+was provided with a detailed protocol (cf. Supplementary Material 1,
+henceforth "SM") with instructions how to perform the simulations,
+providing input variables on climate and soil conditions, and the
+expected output variables. Neither were further site information (except
+for the data specified below) nor any data on tree ingrowth provided
+prior to the simulation. That is, the models were run in "blind flight"
+mode.
 
@@
 a continuous value between 1 and 5 [Soilgrids dataset, Hengl et al.
--@hengl2017, Poggio et al. -@poggio2021]. The protocol also provided the
-elevation, slope and aspect for each of the 200 sites
-[@asterscienceteam2019], but no other spatial information such as
-coordinates was provided, with the exception of iLand, which used the
-blurred coordinates to derive highly detailed soil data.
+-@hengl2017]. The protocol also provided the elevation, slope and aspect
+for each of the 200 sites [@asterscienceteam2019], but no other spatial
+information such as coordinates was provided, with the exception of
+iLand and aDGVM2, which used blurred coordinates to derive highly
+detailed soil data.
 
 The simulations were run in the absence of natural disturbances. Only
-the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE ONE SENTENCE
-OF WHY YOU INCLUDED DISTURBANCES AND WHAT TYPE. The simulations were set
-up to sample species-specific ingrowth rates per decade and per ha in
-the equilibrium state of the model, typically entailing a "spin-up" run
-(as we did not provide any forest data). The modeling teams decided on
-the simulated area and how they derived these samples. The exact length
-of the simulation was also decided by the modeling teams (Table
+the model LPJ-GUESS had to include a background disturbance to increase
+chances of shade-intolerant species to establish. The simulations were
+set up to sample species-specific ingrowth rates per decade and per ha
+in the equilibrium state of the model, typically entailing a "spin-up"
+run (as we did not provide any forest data). The modeling teams decided
+on the simulated area and how they derived these samples. The exact
+length of the simulation was also decided by the modeling teams (Table
 \@ref(tab:tableModels)). Further details on how each modeling team
@@
 simulated equilibrium ("Potential Natural Vegetation") with the current
-climate, as the ultimate goal was to derive the regeneration niche. This
-entails the assumption that (1) the observations from the forest
-reserves reflect no traces of forest management and (2) there is an
-equilibrium between forest dynamics and climate. While the former is
-likely to be met at least in many of the EuFoRIA reserves, the latter
-may be more debatable. However, in the absence of detailed data on the
-history of each plot in the EuFoRIa network, some broad assumptions had
-to be made. Both the width of the regeneration niche (i.e., in
-environmental space) as well as the intensity of the ingrowth process
-(i.e., the number of ingrowth trees per area and per unit of time) were
-of interest.
+climate, as the ultimate goal was to evaluate tree establishment under
+comparable and near equilibrium conditions. This entails the assumption
+that the observations from the forest reserves (1) reflect no traces of
+forest management and (2) there is an equilibrium between forest
+dynamics and climate. While the former might be starting to be visible
+in many of the EuFoRIA reserves, the latter may be more debatable.
+However, in the absence of detailed data on the history of each plot in
+the EuFoRIa network, some broad assumptions had to be made. Both the
+width of the regeneration niche (i.e., in environmental space) as well
+as the intensity of the ingrowth process (i.e., the number of ingrowth
+trees per area and per unit of time) were of interest.
 
@@
 Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus excelsior,
-Quercus spp., Alnus glutinosa, Pinus sylvestris.* Two models included
-their standard set of species for the simulations which is much larger
-(i.e., ForClim 1, ForClim 11). This also applies to TreeMig, but in this
-model *Betula pendula* was not included because HERE **TREEMIG**
-PROVIDES THE END OF THIS SENTENCE SAYING WHY B pendula WAS NOT INCLUDED.
-In three models, fewer than these elveven species were simulated (4C,
-xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). HERE **4C** PROVIDES
-ONE SENTENCE WHY THEY SIMULATED LESS SPECIES. HERE **XCOMP** PROVIDES
-ONE SENTENCE WHY THEY SIMULATED LESS SPECIES. HERE **LPJ-GUESS**
-PROVIDES ONE SENTENCE WHY THEY SIMULATED LESS SPECIES. Finally, the
-model aDGVM2 is based on a suite of broad Plant Functional Types rather
-than individual species, and hence this model used its default approach
-of PFTs for the simulations.
+Quercus spp., Alnus glutinosa, Pinus sylvestris.* The same set of
+species were used in all the 200 sites. Two models included their
+standard set of species for the simulations which is much larger (i.e.,
+ForClim 1, ForClim 11, TreeMig). In three models, fewer than these
+eleven species were simulated (4C, xComp, LPJ-GUESS) (Table
+\@ref(tab:tableModels)). In 4C onl*y Fagus sylvatica*, *Picea abies*,
+*Betula* spp., *Quercus* spp. and *Pinus sylvestris* are parameterised.
+XComp simulations did not consider *Fraxinus excelsior* due to a
+depreciated species parameterisation. In LPJ-GUESS, *Acer
+pseudoplatanus* and *Alnus glutinosa* are not parameterized and
+therefore these species could not be included in the simulation.
+Finally, the model aDGVM2 does not simulate individual species. Rather,
+community assembly processes and trait filtering generate plant
+communities that are adjusted to the biotic and abiotic conditions and
+plants can be classified into different ecological strategies based on
+their trait values in a post-processing step.
 
@@
 strategies, depending on the model, including (1) sampling simulated
-data from the same 1-ha plot in the equilibrium over time, or (2)
-sampling several 1-ha plots from the simulated forest at one specific
-point in time (in the equilibrium). This resulted in 880.000
-observations per model that simulated the 11 species included in the
-protocol (200 sites, 200 samples per site, 11 species and two diameter
-thresholds). For the models that simulated additional species, their
-ingrowth rates were aggregated as "others".
+data from the same 1 ha plot in the equilibrium over time, (2) sampling
+several 1 ha plots from the simulated forest at one specific point in
+time (in the equilibrium) or (3) a combination of (1) and (2) . This
+resulted in 880,000 observations per model that simulated the 11 species
+included in the protocol (200 sites, 200 samples per site, 11 species
+and two diameter thresholds). For the models that simulated additional
+species, their ingrowth rates were aggregated as "others".
 
@@
 sites to avoid unrealistic results, as follows. In the model 4C, a
-threshold of a maximum basal area of 100 $m^2 ha^{-1}$ was used to avoid
+threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid
 unrealistic stand basal area data, and therefore not all the samples and
 sites were reported. The reason for this is that 4C is not suitable for
-long-term simulations without management PLEASE **4C** VERIFY THIS IS
-WHAT YOU DID. LPJ-GUESS had 2% of the sample outputs as grasslands that
-were considered in the data as zero stand basal area and zero ingrowth
-for all the tree species simulated. LPJ-GUESS could not simulate one of
-the sites HERE **LPJ-GUESS** PROVIDES ONE SENTENCE WHY THEY DID NOT
-PROVIDE ALL THE SITES. Landis-II missed 0.2% of the observations HERE
-**LANDIS-II** PROVIDES ONE SENTENCE WHY THEY DID NOT PROVIDE ALL THE
-SAMPLES/observations.
+long-term simulations without management, due to misrepresentations in
+density-dependent mortality processes in long-term simulations and
+assumptions of tree geometry that lead to unrealistic single tree
+dimensions of very old individuals. LPJ-GUESS had 2% of the sample
+outputs as grasslands that were considered in the data as zero stand
+basal area and zero ingrowth for all the tree species simulated and one
+site was not simulated.
 
@@
 the ingrowth, and ingrowth mortality. The observed data were modeled
-using a Generalized Additive Model with a negative binomial distribution
-to better understand the trend and levels of the observed ingrowth,
-relative to the simulation results per model.
+using a Generalized Additive Model (GAM) with a negative binomial
+distribution to better understand the trend and levels of the observed
+ingrowth, relative to the simulation results per model. To avoid
+overfitting we used the restricted maximum likelihood
 
@@
 7 and 10 cm diameter thresholds, with the exception of the empirical
-stand model SYBILA, the landscape model Landis-II and the global model
+stand model SIBYLA, the landscape model Landis-II and the global model
 aDGVM2, which estimated ingrowth rates at the lower end of the
@@
 values of stand basal area, with model-specific thresholds (4C, ForClim
-1, PICUS, SYBILA and iLand).
+1, PICUS, SIBYLA and iLand).
 
@@
 overestimated ingrowth for more than one species for a large part of the
-climate gradients, such as xComp, PICUS, SYBILA or LPJ-GUESS (cf. red
+climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf. red
 colors in Figure \@ref(fig:diff)). These trends were consistent for both
@@
 species along most of the climatic gradients although at a low basal
-area share. Five models (4C, FORMIND, SYBILA, xComp, and LandClim) had
+area share. Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had
 almost no ingrowth of *Pinus sylvestris* and *Quercus* spp., or did not
@@
 [@beckage2003]; this, however, is facing a serious challenge regarding
-appropriate data sets, as inventory data have wide coverage, but do not
+appropriate datasets, as inventory data have wide coverage, but do not
 allow for the assessment of spatial interaction effects.

### 2023-01-27 code/figures.R
restrict the GAMs creation and plotting to the observation ranges

@@ -25,6 +25,8 @@
     
     sum(obsNumAgg$r.ha, na.rm =  T)
     range(obsNumAgg$r.ha, na.rm = TRUE)
+    mean(obsNumAgg$r.ha, na.rm = TRUE)
+    median(obsNumAgg$r.ha, na.rm = TRUE)
     
     # Assign models and dbh to levels 
 
@@ -243,12 +245,15 @@
     # Prepare Observed GAM model data for all the plots
     max_Totba = as.numeric(quantile(envTrend$Totba, .99, na.rm = TRUE))
     min_Totba = as.numeric(quantile(envTrend$Totba, .01, na.rm = TRUE))
-    empModel <- envTrend[envTrend$model == "Observed",]
+    empModel <- envTrend[envTrend$model == "Observed", ]
     fm.sim <- mgcv::gam(r.trees ~ s(Totba, k = 3), data = empModel,
                         family = mgcv::nb())
     new_dat <- data.frame(
         Totba = seq(min_Totba, max_Totba, length.out = 10 * bins),
         fct_name = seq(1, bins + 1, length.out = 10 * bins))
+    
+    new_dat <- new_dat[new_dat$Totba < max(empModel$Totba) & new_dat$Totba > min(empModel$Totba), ]
+    
     pred.sim <- predict(fm.sim, new_dat, type = "link", se.fit = TRUE)
     invLink <- family(fm.sim)$linkinv
     
@@ -262,8 +267,10 @@
         
         axisPlot <- ifelse(modelSel %in% c("Landis II",  "TreeMig", "LPJ-GUESS",
                                             "aDGVM2"), "s", "n")
+        
+        dataSel <- ienvTrendModel[ienvTrendModel$dbh == 7 & ienvTrendModel$model == modelSel,]
         boxplot(r.trees ~ ba_cut, 
-                data = ienvTrendModel[ienvTrendModel$dbh == 7 & ienvTrendModel$model == modelSel,], 
+                data = dataSel, 
                 xlab = "",
                 ylab = "",
                 #ylim = c(0, 500),
@@ -271,7 +278,16 @@
                 xaxt = axisPlot,
                 col = values_color[modelSel],
                 frame = F)
+        
+        means <- tapply(dataSel$r.trees, INDEX = dataSel$ba_cut, FUN = mean) # calculate mean
+        points(means, pch = 20, cex = 0.5, col = "lightgrey") #add means as circles to each boxplot
         box(bty = "l")
+        
[…]

### 2023-02-02 code/figures.R
adds table for gam models output

@@ -17,7 +17,7 @@
 
     # Number of trees in the observed data 
     obsNum <- outputsDF[outputsDF$model == "Observed", ]
-    round( sum(obsNum$r.trees, na.rm = TRUE), 0)
+    round(sum(obsNum$r.trees, na.rm = TRUE), 0)
     
     obsNumAgg <- obsNum |> 
                    dplyr::group_by(site, sample) |>
@@ -307,12 +307,13 @@
     # Prepare Observed GAM model data for all the plots
     max_Totba = as.numeric(quantile(envTrend$Totba, .99, na.rm = TRUE))
     min_Totba = as.numeric(quantile(envTrend$Totba, .01, na.rm = TRUE))
-    empModel <- envTrend[envTrend$model == "Observed",]
+    empModel <- envTrend[envTrend$model == "Observed", ]
     fm.sim <- mgcv::gam(r.trees ~ s(Totba, k = 3), data = empModel,
                         family = mgcv::nb())
     new_dat <- data.frame(
-      Totba = seq(min_TotbaEMP, max_TotbaEMP, length.out = 10 * bins),
+      Totba = seq(min_Totba, max_Totba, length.out = 10 * bins),
       fct_name = seq(1, bins + 1, length.out = 10 * bins))
+    
     new_dat <- new_dat[new_dat$Totba < max(empModel$Totba) & new_dat$Totba > min(empModel$Totba), ]
     
     pred.sim <- predict(fm.sim, new_dat, type = "link", se.fit = TRUE)
@@ -328,6 +329,7 @@
       
       axisPlot <- ifelse(modelSel %in% c("Landis II",  "TreeMig", "LPJ-GUESS",
                                          "aDGVM2"), "s", "n")
+      
       dataSel <- ienvTrendModel[ienvTrendModel$dbh == 7 & ienvTrendModel$model == modelSel,]
       boxplot(r.trees ~ ba_cut, 
               data = dataSel, 
@@ -338,9 +340,16 @@
               xaxt = axisPlot,
               col = values_color[modelSel],
               frame = F)
+      
       means <- tapply(dataSel$r.trees, INDEX = dataSel$ba_cut, FUN = mean) # calculate mean
       points(means, pch = 20, cex = 0.5, col = "lightgrey") #add means as circles to each boxplot
       box(bty = "l")
+      
+      linesData <- data.frame(TotBA = new_dat$fct_name, 
+                              r.trees = invLink(pred.sim$fit),
[…]

### 2023-02-02 manuscript/manuscriptRW.Rmd
changes in the M&M of the manuscript based on co-authors feedback

@@
 the ingrowth, and ingrowth mortality. The observed data were modeled
-using a Generalized Additive Model (GAM) with a negative binomial
-distribution to better understand the trend and levels of the observed
-ingrowth, relative to the simulation results per model. To avoid
-overfitting we used the restricted maximum likelihood
+using a Generalized Additive Model [@wood2011] with a negative binomial
+distribution and restricted maximum likelihood to better understand the
+trend and levels of the observed ingrowth, relative to the simulation
+results per model.
 
@@
 the observed data by calculating the Shannon index based on the relative
-frequency of the species in terms of basal area. The higher the value of
-the index, the higher the species diversity at a particular site and
+proportion of the species in terms of basal area. The higher the value
+of the index, the higher the species diversity at a particular site and
 sample (Equation \ref{eqpiShannon}).
@@
 \begin{equation}
-H_n = - \sum_{i = 1}^{s}{p_{i}ln(p_{n,i})}\label{eqpiShannon}
+H_n = - \sum_{i = 1}^{s}{p_{n,i}ln(p_{n,i})}\label{eqpiShannon}
 \end{equation}

### 2023-02-03 code/figures.R
adds divided by total recruited BA to calculate recruited Shannon instead of total basal area

@@ -43,8 +43,8 @@
     
     ## Recruitment Shannon index  by basal area ----
     ShannonDF <- outputsDF 
-    ShannonDF$giDIVG <- ShannonDF$r.ba / ShannonDF$Totba
-    ShannonDF$lngiDIVG <- log(ShannonDF$r.ba / ShannonDF$Totba)
+    ShannonDF$giDIVG <- ShannonDF$r.ba / ShannonDF$Totr.ba
+    ShannonDF$lngiDIVG <- log(ShannonDF$r.ba / ShannonDF$Totr.ba)
     ShannonDF$mult <- round(ShannonDF$giDIVG * ShannonDF$lngiDIVG, 2)
     
     # Sum across species in each observation

### 2023-02-16 manuscript/manuscriptRW.Rmd
firs attempt to re-define paper questions

@@
 
+NEW
+
+1) Are models of forest dynamics capturing accurately tree ingrowth
+levels, initial tree diversity and mortality in tree establishment?
+
+2) Species specific recruitment values in ennvironmental gradients?
+
+3) Are the models traits linked to the tree establishment values?
+
+environmental gradients?
+
+2) Are they matching the initial species composition of the tree
+ingrowth layer? And
+
+3), do they correctly represent the mortality in tree establishment?
+
 # Material and methods

### 2023-02-16 manuscript/manuscriptRW.Rmd
added information on the missing site from LPJ-guess

@@
 dimensions of very old individuals. LPJ-GUESS had 2% of the sample
-outputs as grasslands that were considered in the data as zero stand
-basal area and zero ingrowth for all the tree species simulated and one
-site was not simulated.
+outputs without tree establishment and were considered in the data as
+zero stand basal area and zero ingrowth for all the tree species
+simulated and one site produced grasslands and was not included in the
+results.
 

### 2023-02-21 manuscript/supplementaryRW.Rmd
removed in appendix figure S9

@@
 
-```{r mortality22 , echo=FALSE, fig.cap="Recruitment ration between 7 and 10cm and the initial recruitment tree number at 7cm.
-Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions.
-The cloud point per models shows the differences across sites, the shape point cloud per model shows if the models take into account regeneration density when applying mortality to regeneration ", out.width='70%'}
-knitr::include_graphics(here::here("figures", "mort7_10.png"))
-```
-
 ```{r tableModels, echo=FALSE,  message=FALSE}

### 2023-02-21 manuscript/manuscriptRW.Rmd
comments and changes until discussion

@@
 
-Two models, ForClim and ForCEEPs, were used here with two alternative
+Two models, ForClim and ForCEEPS, were used here with two alternative
 variants. ForClim variant 1 [@bugmann1996] is based on a recruitment
@@
 eleven species were simulated (4C, xComp, LPJ-GUESS) (Table
-\@ref(tab:tableModels)). In 4C onl*y Fagus sylvatica*, *Picea abies*,
+\@ref(tab:tableModels)). In 4C only *Fagus sylvatica*, *Picea abies*,
 *Betula* spp., *Quercus* spp. and *Pinus sylvestris* are parameterised.
@@
 distribution and restricted maximum likelihood to better understand the
-trend and levels of the observed ingrowth, relative to the simulation
-results per model.
+environmental gradients trend in the levels of the observed ingrowth,
+relative to the simulation results per model.
 
 Ingrowth levels were evaluated along three key environmental gradients,
-i.e., stand basal area (the basal area of all the trees in each sample)
-as a proxy of light availability at the forest floor, the annual
-degree-day sum (DD) [@allen1976; @fischlin1995] as a proxy for growing
-season warmth, and the climatic water balance (CWB) as a proxy for soil
-moisture [@speich2019]. The mean ingrowth values were calculated as the
-mean across the 200 samples per site.
-
-We evaluated the species composition across models and in relation to
-the observed data by calculating the Shannon index based on the relative
-proportion of the species in terms of basal area. The higher the value
-of the index, the higher the species diversity at a particular site and
-sample (Equation \ref{eqpiShannon}).
+i.e., stand basal area (the basal area of all the trees in each sample
+including tree establishment) as a proxy of light availability at the
+forest floor, the annual degree-day sum (DD) [@allen1976; @fischlin1995]
+as a proxy for growing season warmth, and the climatic water balance
+(CWB) as a proxy for soil moisture [@speich2019]. The mean ingrowth
+values were calculated as the mean across the 200 samples per site.
+
+We evaluated the species diversity across models and in relation to the
+observed data by calculating the Shannon index based on the relative
+proportion of the species in terms of basal area. The Shannon index was
+calculated for the establishment ($H_{n,r}$) (Equation
+\ref{eqpiShannonR}) and at the stand level ($H_{n,s}$)(Equation
+\ref{eqpiShannon}). The higher the value of the index, the higher the
+species diversity at a particular site and sample.
+
+We also tested for significant trends across XXXXX !!!!!
 
@@
 \begin{equation}
-H_n = - \sum_{i = 1}^{s}{p_{n,i}ln(p_{n,i})}\label{eqpiShannon}
+H_{n,r} = - \sum_{i = 1}^{s}{p_{n,i}ln(p_{n,r,i})}\label{eqpiShannonR}
 \end{equation}
@@
 \begin{equation}
-p_{n,i} = \frac{r_{BA_{n,i}}}{Total_{BA_{n}}} \label{eqpiShannon2}
+p_{n,r,i} = \frac{r_{BA_{n,i}}}{TotalR_{BA_{n}}} \label{eqpiShannonR2}
 \end{equation}
 ```
-where $s$ is the total number of species simulated that have ingrowth
-basal area above zero in sample $n$; $p_{n,i}$ is the proportion of the
-species $i$ calculated as the ingrowth basal area ($r_{BA_{n,i}}$) for
-that species $i$ relative to the stand basal area ($Total_{BA_{n}}$) of
-the sample $n$.
+```{=tex}
+\begin{equation}
+H_{n,s} = - \sum_{i = 1}^{s}{p_{n,i}ln(p_{n,i})}\label{eqpiShannon}
+\end{equation}
+```
+```{=tex}
+\begin{equation}
+p_{n,s,i} = \frac{BA_{n,i}}{Total_{BA_{n}}} \label{eqpiShannon2}
+\end{equation}
+```
+where $s$ is the total number of species present and have basal area
+above zero in sample $n$; $p_{n,r,i}$ is the proportion of the species
+$i$ calculated as the ingrowth basal area ($r_{BA_{n,i}}$) for that
+species $i$ relative to the recruited basal area ($TotalR_{BA_{n}}$) of
+the sample $n$; $p_{n,s,i}$ is the proportion of the species $i$
+calculated as the basal area ($BA_{n,i}$) for that species $i$ relative
+to the total basal area ($TotalR_{BA_{n}}$) of the sample $n$.
 
@@
 \begin{equation}
-\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{t,i}}{\sum_{i = 1}^{s}{\bar{r}_{t,i}}} \label{eq:BAshare}
+\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{BA_{t,i}}}{\sum_{i = 1}^{s}{\bar{r}_{{BA_{t,i}}}}} \label{eq:BAshare}
 \end{equation}
@@
 where $s$ is the total number of species simulated; $\bar{r}_{t,i}$ is
-the mean ingrowth of the species $i$ per site $t$ across the available
-samples at that site.
-
-The data from 35 sites that had a diameter threshold of 10 cm only, and
-therefore ingrowth rates for these sites at 7 cm are not known. The 35
-sites were removed from the simulated data to make a fair comparison
-regarding the ingrowth of the five main species. Furthermore, regarding
-the ingrowth basal area share per species ($\bar{R}_{BA_{share_{i}}}$),
-we categorized this as zero when the total ingrowth basal area (
-$\sum_{i = 1}^{s}{\bar{r}_{t,i}}$ ) equaled zero, and also when the
-ingrowth basal area of that species ($\bar{r}_{t,i}$) was zero.
+the mean basal area ingrowth of the species $i$ per site $t$ across the
+available samples at that site.
+
+We removed 35 sites from the comparisons at 7 cm ingrowth, because field
+data did not include these measurements. Furthermore, regarding the
+ingrowth basal area share per species ($\bar{R}_{BA_{share_{i}}}$), we
+categorized this as zero when the ingrowth basal area of that species
+($\bar{r}_{BA_{t,i}}$) was zero but also when both the ingrowth basal
+area of that species ($\bar{r}_{BA_{t,i}}$) and the total ingrowth basal
+area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.
 
@@
 quadratic mean diameter in fully stocked pure stands. The value used in
-our comparisons was 1.77, calculated using Equation \ref{eqReineke}.
+our comparisons was 1.77 (i.e., we expect stem numbers at 7cm to be 77%
+higher than at 10cm), calculated using Equation \ref{eqReineke}.
 
@@
 similar to or smaller than observed, with the notable exception of
-PICUS, where simulated ingrowth variability was much larger than that of
-any other model.
+PICUS, where simulated ingrowth variability was much larger.
 
 ```{r over, echo=FALSE, fig.cap="Mean ingrowth rates across all samples per site, plotted for the 200 sites and for each model.
-The red dotted lines shows the 2.5 and 97.5 percentiles for the 7 cm diameter threshold in the observed data.", out.width='100%'}
+The red dotted lines shows the 2.5 and 97.5 percentiles for the 7 cm diameter threshold in the observed data.
+There are two boxplots for each model where the left lighted boxplot corresponds to 7cm and the right darker boxplot to 10cm", out.width='100%'}
         knitr::include_graphics(here::here("figures", "recOverAll710.png"))
@@
 
-When evaluating the representation of ingrowth rates along key gradients
-of light availability (basal area), temperature (degree-days) and soil
-moisture (climatic water balance), distinct features emerge: the models
-reproduced the effect of basal area (Figure \@ref(fig:nvTrend710BA))
-much better than the effects of the climatic gradients, where they
-featured varying effects (Figures S1 and S2).
-
-In the observed data, ingrowth rates decreased clearly with increasing
+When evaluating the representation of ingrowth levels along key
+gradients of light availability (basal area), temperature (degree-days)
+and soil moisture (climatic water balance), distinct features emerge:
+the models reproduced the effect of basal area in magnitude and patterns
+(Figure \@ref(fig:nvTrend710BA)) better than the effects of the climatic
+gradients, where they featured varying effects (Figures S1 and S2).
+
+In the observed data, ingrowth levels decreased clearly with increasing
 stand basal area (Figure \@ref(fig:nvTrend710BA)). This trend was
@@
 basal area than in the observations. The other models featured
-distinctly different trends, such as (1) an increase in ingrowth with
-increasing total basal area followed by the absence of ingrowth at high
-values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and
+distinctly different trends, such as (1) an increase in ingrowth levels
+with increasing total basal area followed by the absence of ingrowth at
+high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and
 LPJ-GUESS), (2) almost constant ingrowth levels with basal area (ForClim
@@
 with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp,
-PICUS and LandClim), while a few models showed the opposite trend
-(ForCEEPs(f), FORMIND and LPJ-GUESS). Across the temperature gradient
-(Figure S2), all the models that featured a decrease in ingrowth rates
-with increasing soil moisture showed the opposite trend (i.e., more
-ingrowth with increasing temperature), except for 4C. The other models
-did not feature a clear pattern.
+PICUS, LandClim, TreeMig and iLand), while a few models showed the
+opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS). Across the
+temperature gradient (Figure S2), all the models that featured a
+decrease in ingrowth rates with increasing soil moisture showed the
+opposite trend (i.e., more ingrowth with increasing temperature), except
+for 4C. The other models did not feature a clear pattern.
 
 Both observations and simulated data had no ingrowth in some samples and
-at some sites (Table S1). The observed data had slightly more than 3% of
-the samples with no ingrowth. Only three models had a larger proportion
-of no ingrowth (4C, Landis-II and aDVM2). Three models always simulated
-ingrowth for both the 7 and 10 cm threshold (TreeMig and LPJ-GUESS),
-i.e., they did not feature any zero values. The other models had a very
-low percentage of samples with no ingrowth (0.01% - 1.31%), i.e., they
-had distinctly fewer occurrences of zeros compared to the observations.
-There was no significant relationship between the mean complexity of the
-establishment module (Table \@ref(tab:tableModels)) and the
-overestimation proportion at a diameter threshold of 7 cm (SM3).
+at some sites (Table S1). The observed data had 3% of the samples with
+no ingrowth. Only three models had a larger proportion of no ingrowth
+(4C, Landis-II and aDVM2). CHECK! No LPJ-GUESS anymore with the new data
+submission!!!! Two models always simulated ingrowth for both the 7 and
+10 cm threshold (TreeMig and LPJ-GUESS), i.e., they did not feature any
+zero values. The other models had a very low percentage of samples with
+no ingrowth (0.01% - 1.31%), i.e., they had distinctly fewer occurrences
+of zeros compared to the observations. There was no significant
+relationship between the mean complexity of the establishment module
+(Table \@ref(tab:tableModels)) and the overestimation proportion at a
+diameter threshold of 7 cm (SM3).
 
@@
 Most models matched the level of diversity of the observed data quite
-well (Figure \@ref(fig:H710)), although they consistently underestimated
-its variability across sites. Four models overestimated ingrowth
-diversity: 4C, PICUS, Landis-II and LPJ-GUESS, the latter mostly for the
-10 cm diameter threshold. The model 4C is a special case as it simulated
-five species only, i.e., its diversity values are not directly
-comparable to those of the other models, nor to the observations. Only
-one model, SIBYLA, consistently underestimated ingrowth diversity.
+well as quantified by the Shannon index (Figure \@ref(fig:H710)),
+although they consistently underestimated its variability across sites.
+Four models overestimated ingrowth diversity: 4C, PICUS, Landis-II and
+LPJ-GUESS, the latter mostly for the 10 cm diameter threshold. The model
+4C is a special case, as it simulated five species only, i.e., its
+diversity values are not directly comparable to those of the other
+models, nor to the observations. Only one model, SIBYLA, consistently
+underestimated ingrowth diversity.
 
@@
 species diversity of ingrowth between 7 and 10 cm across sites (Tables
-S2-S4). Furthermore, three models (4C, SIBYLA, Landis-II) maintained the
-ingrowth diversity between the 7 and 10 cm thresholds (i.e., the
-differences between them were not significant, see Table S3), and the
-same was evident from the observed data. There were a few models with a
-strong increase of ingrowth diversity between 7 and 10 cm (ForCEEPS,
-TreeMig, LPJ-GUESS, Figure \@ref(fig:H710)).
+S2-S4). Three models (4C, SIBYLA, Landis-II) maintained the ingrowth
+diversity between the 7 and 10 cm thresholds (i.e., the differences
+between them were not significant, see Table S3), and the same was
+evident from the observed data. There were a few models with a strong
+increase of ingrowth diversity between 7 and 10 cm (ForCEEPS, TreeMig,
+LPJ-GUESS, FORMIND, Figure \@ref(fig:H710)).
 
@@
 of diversity compared to observations. Models overpredicting species
-richness included ForClim1, ForClim 11, FORMIND and TreeMig (Figure
+diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure
 \@ref(fig:combinedRichness)A, left). Several models underpredicted
@@
 ingrowth thresholds, although the exact changes in the share of the
-ingrowth basal area were somewhat different (Figure S7).
+ingrowth basal area were different (Figure S7).
 
 Second, there was some congruence in the simulation results by species
-across the models. The share in the ingrowth by Silver fir (*Abies
+across the models. The share in the ingrowth of Silver fir (*Abies
 alba*) was captured well across the gradient by most models except for
@@
 share of this species. There were also some peculiarities evident for
-some models, such as 4C that did not simulate *Abies alba*. *Fagus
-sylvatica* ingrowth was underestimated by many models across the
-environmental gradients, including ForCEEPS(f) and FORMIND, which
-featured negligible ingrowth at most sites for this species. Most of the
-models overestimated the share of *Picea abies* ingrowth in the cold-wet
-part of the gradients. *Pinus sylvestris* and *Quercus* spp. represented
-a smaller share of the ingrowth in the observed data mostly at warm-dry
-sites. Many models, however, erroneously featured ingrowth for these two
-species along most of the climatic gradients although at a low basal
-area share. Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had
-almost no ingrowth of *Pinus sylvestris* and *Quercus* spp., or did not
-feature any ingrowth of these species at all.
+some models, such as 4C that did not simulate *Abies alba*. The patterns
+across models for *Fagus sylvatica* ingrowth were more complex, the
+ingrowth was underestimated by many models across the environmental
+gradients featuring negligible ingrowth at most sites, while others
+overestimate it in different parts of the environmental space. Most of
+the models overestimated the share of *Picea abies* ingrowth in the
+cold-wet part of the gradients. *Pinus sylvestris* and *Quercus* spp.
+represented a smaller share of the ingrowth in the observed data mostly
+at warm-dry sites. Many models, however, erroneously featured ingrowth
+for these two species along most of the climatic gradients although at a
+low basal area share. Five models (4C, FORMIND, SIBYLA, xComp, and
+LandClim) had almost no ingrowth of *Pinus sylvestris* and *Quercus*
+spp., or did not feature any ingrowth of these species at all, close to
+the patterns from the observed data.
 
@@
 
-The fact that some models featured lower mortality than expected by
-Reineke's self-thinning rule may point to facilitation, or simply a
-higher mortality rate before the trees had reached 7 cm diameter. Yet,
-the case of models such as xComp, PICUS and TreeMig, whose mortality was
-well above the self-thinning line, indicates that they feature higher
-ingrowth mortality in mixed-species forests compared to even-aged,
-single-species forests. Some models yielded a mortality rate of ingrowth
-that is broadly compatible with the self-thinning rule (i.e., 4C and
-LandClim).
+The fact that some models (and observed data) featured lower mortality
+than expected by Reineke's self-thinning rule may point to facilitation,
+or simply a higher mortality rate before the trees had reached 7 cm
+diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
+whose mortality was well above the self-thinning line, indicates that
+they feature higher ingrowth mortality in mixed-species forests compared
+to even-aged, single-species forests. Some models yielded a mortality
+rate of ingrowth that is broadly compatible with the self-thinning rule
+(e.g., 4C and LandClim).
 
@@
 models that overestimated ingrowth had mortalities lower than the
-expected self-thinning ratio (Figure \@ref(fig:mortality13)B). It is
-noteworthy that most of the models that featured a low ratio of ingrowth
-between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11 and
-LandClim) (Figure S8) had only a small overestimation of ingrowth at the
-7 cm threshold (Figure \@ref(fig:mortality11) cf. Figure S8). aDGVM2 was
-the only model that underestimated ingrowth at the 7 cm threshold while
-featuring a low mortality between the two thresholds (Figure
+expected self-thinning ratio with the exception of LandClim (Figure
+\@ref(fig:mortality13)B). It is noteworthy that most of the models that
+featured a low ratio of ingrowth between 7 and 10 cm (i.e., ForCEEPS,
+ForClim 1, ForClim 11 and iLand) had only a small overestimation of
+ingrowth at the 7 cm threshold (Figure \@ref(fig:mortality11)). aDGVM2
+was the only model that underestimated ingrowth at the 7 cm threshold
+while featuring a low mortality between the two thresholds (Figure
 \@ref(fig:mortality13)C), and only a few models were close to the
@@
 
-Lastly, some models that started at a low number of individuals (7 cm
-threshold) had a lower decrease of ingrowth between 7 and 10 cm dbh
+Lastly, some models that had a low number of individuals at 7 cm
+threshold also had a smaller decrease of ingrowth between 7 and 10 cm
 (ratio 7-10 cm) (Figure S10). In these models, competition and
@@
 species had reached 7 cm. However, this pattern was not consistent
-across models (Figure S9). It is noteworthy that there is a relationship
-between the ingrowth rate at 7 cm and the mortality rate: most of the
-models showed a positive effect (higher ingrowth at 7 cm is coupled to
-an increase of the mortality between 7 and 10 cm). While some models
-showed a negative effect, this linear trend was not pronounced and not
-always significant regarding the slope (Figure S9 and Table S2).
+across models (Figure S10). It is noteworthy that there is a
+relationship between the ingrowth rate at 7 cm and the mortality rate:
+most of the models showed a positive effect (higher ingrowth at 7 cm is
+coupled to an increase of the mortality between 7 and 10 cm). While some
+models showed a negative effect, this linear trend was not pronounced
+and not always significant regarding the slope (Figure S9 and Table S6).
 
@@
 Tree establishment is a fundamental process in forest dynamics, in
-particular with increasing disturbance under climate change. Correctly
-capturing tree ingrowth in dynamic models is fundamental to, for
-example, evaluate post-disturbance dynamics as it will define forest
-state in the first decades [@seidl2022]. If this initial phase is not
-captured well, we can not properly assess aspects such as forest
-resilience or the timing, magnitude and progression of carbon
-sequestration. In this study, for the first time, projections of tree
-establishment from multiple models of forest dynamics were confronted
-with a unique dataset from unmanaged forest reserves across a large
-environmental gradient in Europe.
-
-The observed data that we used [@käber2022] are also exceptional,
-particularly with respect to the number of events captured, which is
-essential for a highly 'noisy' process such as tree establishment. The
-use of this dataset for model benchmarking provided novel insights on
-the capability of state-of-the art models to simulate ingrowth levels,
-species composition, and mortality in an early stage of tree life, i.e.,
-between a tree diameter of 7 and 10 cm. Overall, by adopting this
-approach a much broader understanding resulted than if we had used these
-data for model calibration: it is primarily from the failure of the
-models that we can gain ecological insights [e.g., @trugman2022].
+particular under changing conditions such as those caused by climate
+change. Correctly capturing tree ingrowth in dynamic models is
+fundamental to, for example, evaluate post-disturbance dynamics as it
+will define forest state in the first decades [@seidl2022]. If this
+initial phase is not captured well, we can not properly assess aspects
+such as forest resilience or the timing, magnitude and progression of
+carbon sequestration. In this study, for the first time, projections of
+tree establishment from multiple models of forest dynamics were
+confronted with a unique dataset from unmanaged forest reserves across a
+large environmental gradient in Europe.
+
+The observed data that we used [@käber2022] are exceptional,
+particularly with respect to the number of events captured (number of
+sites and repetitive measurements), which is essential for a highly
+'noisy' process such as tree establishment. The use of this dataset for
+model benchmarking provided novel insights on the capability of
+state-of-the-art models to simulate ingrowth levels, species
+composition, and mortality in an early stage of tree life, i.e., between
+a tree diameter of 7 and 10 cm. Overall, by adopting this approach a
+much broader understanding resulted than if we had used these data for
+model calibration: it is primarily from the shortcomings of the models
+that we can gain ecological insights [e.g., @trugman2022].
 
@@
 
-Most of the models tended to overestimate tree ingrowth rates. This has
+Most of the models overestimated tree ingrowth rates. This has
 potentially far-reaching implications, e.g., regarding biomass (and thus
@@
 
-From a modeling point of view, excess ingrowth inevitable requires
+From a modeling point of view, excess ingrowth inevitably requires
 excess mortality rates in a later stage, either -- as observed for some
-models -- between the 7 and 10 cm diameter thresholds, before the 7 cm
-are reached, or soon after the 10 cm threshold has been crossed. In any
-case, correcting at early stages for the expected forest densities at
-later stages is equivalent to compensating for a first error (excess
-ingrowth) by a second error (excess mortality). Thus, many models appear
-to get the 'right answer', i.e., correct stand structure and
-composition, for the wrong reasons. It is highly likely that biased
-projections will result, because the two errors are unlikely to be
-perfectly linked and thus will not always compensate each other. Hence,
-this structural problem of most models investigated here is problematic
-particularly if the models are to be used under novel conditions such as
-under climate change [e.g., @huber2021] or in a decision support context
-[e.g., @thrippleton2021].
+models -- between the 7 and 10 cm diameter thresholds, or soon after the
+10 cm threshold has been crossed. In any case, correcting at early
+stages for the expected forest densities at later stages is equivalent
+to compensating for a first error (excess ingrowth) by a second "error"
+(excess mortality). Thus, many models appear to get the 'right answer',
+i.e., correct stand structure and composition, for the wrong reasons. It
+is highly likely that biased projections will result, because the two
+errors are unlikely to be perfectly linked and thus will not always
+compensate each other. Hence, this structural problem of most models
+investigated here is problematic particularly if the models are to be
+used under novel conditions such as under climate change [e.g.,
+@huber2021] or in a decision support context [e.g., @thrippleton2021].
 
@@
 additional features that can be traced back, as done in this study for
-the model variants of ForClim and ForCEEPs, and that model complexity
+the model variants of ForClim and ForCEEPS, and that model complexity
 and structure must always be connected with model objectives

### 2023-02-24 manuscript/manuscriptRW.Rmd
revision until research recomendations

@@
 
-In spite of this criticism, it is remarkable that most models were not
-overly off with respect to simulated ingrowth levels -- after all, few
-if any of the models' establishment routines are sufficiently
+## Model formulation did not affect model performance
+
+ADD THESE POINTS HERE complexity, empirical v. PB, etc.
+
+In spite of this criticism, it is remarkable that most models did not
+exceedingly deviate from respect to simulated ingrowth levels -- after
+all, few if any of the models' establishment routines are sufficiently
 constrained by data, with the exception of the empirical models xComp
 and SIBYLA. It is noteworthy that in spite of their empirical basis,
-both models were off relative to empirical ingrowth levels, in similar
+both models did not match empirical ingrowth levels, in similar
 magnitude as the other models and even in opposite directions (SIBYLA:
 underestimation; xComp: overestimation). It appears that using empirical
-data of limited geographical scope, as done in either of these models,
-leads to extrapolation problems already under current climatic
-conditions. For the other models, whose establishment module is not
-strongly constrained by empirical data, the degrees of freedom that are
-available for modeling ingrowth [@konig_tree_2022] could have implied
-that model performance would be much worse.
+data of limited geographical scope and constrained to managed forests,
+as done in either of these models, leads to extrapolation problems
+already under current climatic conditions. For the other models, whose
+establishment module is not strongly constrained by empirical data, the
+degrees of freedom that are available for modeling ingrowth
+[@konig_tree_2022] could have implied that model performance would be
+much worse.
 
@@
 equilibrium forests, whereas in reality many of the forest reserves are
-recovering from past management activities and have become denser and
-darker over the past decades [e.g., @heiri2009], leading to less
-ingrowth than in an equilibrium situation.
+recovering from past management activities and have become denser over
+the past decades [e.g., @heiri2009], leading to less ingrowth than in an
+equilibrium situation.
 
 Competition for light as a strong filter for tree establishment has been
-documented widely [@messier1999; @collet2006; @berdanier2016]. However,
-it is difficult to measure light availability at large spatial and
-temporal scales. We used total stand basal area as a proxy for light
-availability [cf. @schmid2021]. We found pronounced differences in the
-stand basal area ranges simulated by the models, but also between the
-models and the observations. This made it impossible to evaluate the
-establishment for the extremes of the density ranges in some models. For
-example, ingrowth rates at low stand densities are relevant to assess
-how well forests are recovering e.g., after gap creation due to
-disturbance [@seidl2022; @grubb1977]. At the level of the simulated
-one-hectare samples, average basal area is typically not very low as
-long as no larger disturbances occur, which was explicitly excluded in
-our simulation protocol. Towards the other end of the spectrum, i.e.,
-with increasing basal area, it would be reasonable to expect that with
+widely documented [@messier1999; @collet2006; @berdanier2016], but
+models did not reproduce this expectation. However, it is difficult to
+measure light availability at large spatial and temporal scales. We used
+total stand basal area as a proxy for light availability [cf.
+@schmid2021]. We found pronounced differences in the stand basal area
+ranges simulated by the models, but also between the models and the
+observations. This made not possible to evaluate the establishment for
+the extremes of the stand density ranges in some models. For example,
+ingrowth rates at low stand densities are relevant to assess how well
+forests are recovering e.g., after gap creation due to disturbance
+[@seidl2022; @grubb1977]. At the level of the simulated one-hectare
+samples, average basal area is typically not very low as long as no
+larger disturbances occur, which was explicitly excluded in our
+simulation protocol. Towards the other end of the spectrum, i.e., with
+increasing basal area, it would be reasonable to expect that with
 increasing stand basal area, the ingrowth of the different species would
@@
 to quantitatively match ingrowth levels (cf. above), their performance
-is much better when it comes to matching the quality, i.e., the
-composition of ingrowth as a function of abiotic and biotic conditions.
-
-Yet, there were pronounced variations how the main tree species were
-represented along the environmental gradients. Most models maintained or
+is much better when it comes to matching the composition of ingrowth as
+a function of abiotic and biotic conditions. Most models maintained or
 even increased ingrowth diversity between the 7 and 10 cm thresholds,
@@
 not significant. Diversity variations may be explained by the way the
-individual models are considering establishment processes
-[@konig_tree_2022], e.g., in the absence of adults, ingrowth can be
-linked to either seed rain-dispersal, or to a seed/seedling bank. The
+individual models consider establishment processes [@konig_tree_2022],
+e.g., in the absence of adults, ingrowth can be linked to either seed
+rain-dispersal, or to a seed/seedling bank.
+
+Yet, there were pronounced variations how the main tree species were
+represented along the environmental gradients, in particular the
+dominant species Abies alba, Fagus sylvatica and Picea abies. The
 ingrowth rates were sampled from simulations in the equilibrium, and in
@@
 sylvestris* or *Quercus* would be of minor importance, or absent
-entirely [@klopcic2015]. Most models captured this difference, which is
-also found in the empirical data. Thus, although the broad patterns are
+[@klopcic2015]. Most models captured this low abundance, which is also
+found in the empirical data. Thus, although the broad patterns are
 matched by many models, improvements in the quantification of the
@@
 species abundances [@hanbury-brown2022], although this may be
-simplistic. Furthermore, based on the simulation results there is no
+simplistic. Furthermore, based on the simulation results, there is no
 evidence that models with feedback captured better the ingrowth species
 diversity than those without feedback. The similar performance of models
-with and without feedback is likely because the models are giving more
-weight to the regeneration niche arising abiotic and biotic filters than
+with and without feedback is likely because the models put more weight
+to the regeneration niche arising from abiotic and biotic filters, than
 from the habitat niche of the adult trees [@grubb1977].
@@
 
-There are several factors that lead to mortality in establishment such
-as competition [@casper1997] in addition to multiple abiotic factors
-[@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
-expect that models that overestimate ingrowth of small trees will have a
-particularly high mortality between the 7 and 10 cm thresholds. However,
-this was not consistently evident from the simulations. This implies
-that these models must have the excess of mortality in earlier or later
-stages, if we assume that all models would be able to capture the
-structure and composition of adult stands along the EuFoRIa gradient,
-which however was not tested here.
-
-Higher mortality towards the adult stage has important implications for
+There are several factors that lead to mortality during the
+establishment phase such as competition [@casper1997] and multiple
+abiotic factors [@schmid2021; @cunningham2006]. As mentioned above, it
+is reasonable to expect that models that overestimate ingrowth of small
+trees will already have a particularly high mortality between the 7 and
+10 cm thresholds. However, this was not consistently evident from the
+simulations. This implies that these models must have the excess of
+mortality in earlier or later stages, if we assume that all models would
+be able to capture the structure and composition of adult stands along
+the EuFoRIa gradient, which however was not directly tested here.
+
+Higher mortality towards the adult phase has important implications for
 forest dynamics and the goods and services provided by forests. On the
 one hand, mortality in later stages may erroneously enhance the share of
-less shade tolerant species in the models [@klopcic2015], therefore
-shifting the species composition. Unrealistic high stem densities for a
-longer period of time may overestimate the role of tree regrowth in
-carbon sink dynamics [@pugh2019]. Given our set of simulations and
-analyses, we cannot conclusively assess what is going on in the models,
-and further studies are required that focus on a wider range of tree
-sizes and the fate of tree establishment along such a size continuum.
+less shade tolerant species in the models [@klopcic2015], if available,
+therefore shifting the species composition. Unrealistic high stem
+densities for a longer period of time may overestimate the role of tree
+regrowth in carbon sink dynamics [@pugh2019]. Given our set of
+simulations and analyses, we cannot conclusively assess what is
+happening in the models, and further studies are required that focus on
+a wider range of tree sizes and the fate of tree establishment along
+such a size continuum.
 
@@
 
-The EuFoRIa data as used in this study is unique and highly suited to
-better understand tree establishment. However, three aspects of this
+The EuFoRIa data as used in this study are unique and well suited to
+better understand tree establishment. However, three aspects of these
 data may also represent considerable limitations. First, we made a
-comparison of tree ingrowth in a posited equilibrium state, but we
-cannot assess how close the empirical data were to such a state. The
-data were collected in forest reserves where no management has taken
-place for long periods of time. This makes our assumption of an
-equilibrium between forest properties and environmental drivers less
-haphazard than it might appear at first sight. However, it may explain
-why simulated ingrowth rates were higher than the observed ones. Second,
-the data were collected for ingrowth above a 7 cm threshold, thus
-limiting the assessment of tree establishment in a holistic sense. This
-constituted a hard limit based on which we can understand some aspects
-of tree establishment, which in its entirety often comprises a rather
-long period since germination [@price2001]. However, many environmental
-constraints are acting on young trees [@käber2021] that we were unable
-to assess. Yet, ingrowth data with lower thresholds are simply not
-available in a harmonized manner across large environmental gradients.
-Third, the empirical data were collected from rather small plots, while
-we sampled simulated ingrowth rates from 1 ha areas, which may lead to
-an incorrect representation of space. Even though this strategy it is
-not ideal, it represents a common challenge when harmonizing diverse
-data sources originating from varied sampling strategies [@portier2022].
-It would have been extremely challenging for such a variety of models to
-follow a protocol where the spatial sampling size was different at each
-of the 200 sites.
-
-The design of our sampling protocol did not include spatial aspects. We
-considered a wide range of models of forest dynamics, from stand to
-global scale. However, the simulation setup was limited to 1 ha ingrowth
-values. The lack of spatial considerations potentially puts landscape
-models at a disadvantage, as they have been built to be accurate at the
-landscape level. Without spatial context, we are limiting ingrowth to
-the seed influx from the stand itself, if the model does not have a
-background seed input. However, Global Dynamic Vegetation Models should
-not be in disadvantage due to the limited spatial considerations [but
-cf. Snell et al. -@snell2014] as they are usually lacking dispersal
-between cells as they are based on an abstraction of horizontal space,
-or even the trees themselves [@hanbury-brown2022].
+comparison of tree ingrowth in a equilibrium state, but we cannot assess
+how close the forests included in the EuFoRIa dataset were to such a
+state. The data were collected in forest reserves where no management
+has taken place for long periods of time. This makes our assumption of
+an equilibrium between forest properties and environmental drivers more
+reasonable than it might appear at first sight. However, the equilibrium
+assumption may explain why simulated ingrowth rates were higher than the
+observed ones. Second, the data were collected for ingrowth above a 7 cm
+threshold, thus limiting the assessment of tree establishment to a
+holistic viewpoint. This constituted a hard limit based on which we can
+understand some aspects of tree establishment, which in its entirety
+often comprises a rather long period since seed production [@price2001].
+However, many environmental constraints are acting on young trees
+[@käber2021] that we were unable to assess. Yet, ingrowth data with
+lower thresholds are simply not available in a harmonized manner across
+large environmental gradients. Third, the empirical data were collected
+from rather small plots, while we sampled simulated ingrowth rates from
+1 ha areas, which may lead to an incorrect representation of space. Even
+though this strategy is not ideal, it represents a common challenge when
+harmonizing diverse data sources originating from varied sampling
+strategies [@portier2022]. It would have been extremely challenging for
+such a variety of models to follow a protocol where the spatial sampling
+size was different at each of the 200 sites and it would have introduce
+additional uncertainty into the results.
+
+The design of our sampling protocol did not include spatial aspects such
+as seed migration nor detailed soil data. We considered a wide range of
+models of forest dynamics, from stand to global scale. However, the
+simulation setup was limited to 1 ha ingrowth values. The lack of
+spatial considerations potentially puts landscape models at a
+disadvantage, as they have been built to be accurate at the landscape
+level. Without spatial context, we are limiting ingrowth to the seed
+influx from the stand itself, if the model does not have a background
+seed input. However, Global Dynamic Vegetation Models should not be in
+disadvantage due to the limited spatial considerations [but cf. Snell et
+al. -@snell2014] as they are usually lacking dispersal between cells as
+they are based on an abstraction of horizontal space, or even the trees
+themselves [@hanbury-brown2022]. Detailed data on soil conditions were
+not available from the observed data and indepentdent data for soil and
+climatic water balance were used instead. Several models, however, used
+more detailed drought stress indicators based e.g., on soil water
+holding capacity that had to be derived from a rough soil quality
+measure which could explain the poor performance of the models against
+climatic water balance.
 
@@
 
-Increasing complexity in establishment modules has been motivated by
-better process understanding and enhancing of model accuracy
-[@bugmann2022; @bugmann2019]. However, more complex models do not
-necessary lead to better projections and rather to a reduced
-transparency and predictive power [@franklin2020]. The results from this
-study showed that increasing complexity in the establishment modules did
-not represent an improvement in the projections of ingrowth levels,
-species composition or mortality at early tree stages. Thus, the
-question what level of detail one has to use to model establishment
-processes remains open and must be addressed in future research
-[@konig_tree_2022; @bugmann2022].
-
-If it should be adequate to include more detail in the establishment
-models, this will come with higher parameterisation efforts, which will
-most likely lead to lower generalization as the required data will have
-to be collected from specific locations, as there is no general
-establishment dataset available. We recommend on the one hand that the
-further development of the establishment modules is implemented as
-additional features that can be traced back, as done in this study for
-the model variants of ForClim and ForCEEPS, and that model complexity
-and structure must always be connected with model objectives
-[@albrich2020]. On the other hand, we emphasize that datasets like
-EuFoRIa used here are invaluable and should be further expanded in both
-their spatial extent (e.g., towards boreal and Mediterranean conditions)
-as well as in time (e.g., continuing the monitoring into the future).
-
 Based on the results of this study, it is clear that models of forest
-dynamics need a focus on their establishment modules. However, here we
-have only consider tree establishment in the equilibrium state. It is
-equally important to understand how these models project tree
+dynamics need a focus on their establishment modules. Increasing
+complexity in establishment modules has been motivated by better process
+understanding and enhancing of model accuracy [@bugmann2022;
+@bugmann2019]. However, more complex models do not necessarily lead to
+better projections and rather to a reduced transparency and predictive
+power [@franklin2020]. The results from this study showed that
+increasing complexity in the establishment modules did not represent an
+improvement in the projections of ingrowth levels, species composition
+or mortality at early tree stages. Thus, the question what level of
+detail one has to use to model establishment processes remains open and
+must be addressed in future research [@konig_tree_2022; @bugmann2022].
+
+We recommend that the further development of the establishment modules
+is implemented as additional features that can be traced back, as done
+in this study for the model variants of ForClim and ForCEEPS, and that
+model complexity and structure must always be connected with model
+objectives [@albrich2020]. If it should be necessary to include more
+detail in the establishment models, this will come with higher
+parameterisation efforts, which will most likely lead to lower
+generalization as the required data will have to be collected from
+specific locations, as there is no general establishment dataset
+available.
+
+More effort should be invested into collecting harmonised datasets on
+tree establishment. We emphasize that datasets like EuFoRIa used here
+are invaluable and should be further expanded in both their spatial
+extent (e.g., towards boreal and Mediterranean conditions) as well as in
+time (e.g., continuing the monitoring into the future). Such a data will
+allow a better evaluation of forest models and help to reduce the
+uncertainty in their projections, which is crucial when they are used as
+tools for predicting potential impacts of climate change.
+
+Here we have only consider tree establishment in the equilibrium state.
+It is equally important to understand how these models project tree
 establishment after changes in forest structure like those brought about
@@
 be able to assess the relationship of these aspects on tree
-establishment, e.g., due to the lack of disturbance or appropiate
+establishment, e.g., due to the lack of disturbance or appropriate
 management modules.

### 2023-03-09 manuscript/manuscriptRW.Rmd
add changes till conclussions

@@
 generalization as the required data will have to be collected from
-specific locations, as there is no general establishment dataset
-available.
+specific locations, as currently there is no general establishment
+dataset available.
 
@@
 
-Here we have only consider tree establishment in the equilibrium state.
-It is equally important to understand how these models project tree
-establishment after changes in forest structure like those brought about
-by disturbances [@seidl2022], or under different management strategies
-[@lindner2000]. However, this will require an entirely different set of
-observed data, and potentially not all models of forest dynamics would
-be able to assess the relationship of these aspects on tree
-establishment, e.g., due to the lack of disturbance or appropriate
-management modules.
+Here we have only considered tree establishment in the equilibrium
+state. It is equally important to understand how these models project
+tree establishment after changes in forest structure by disturbances
+[@seidl2022], or under different management strategies [@lindner2000].
+However, this will require an entirely different set of observed data,
+and potentially not all models of forest dynamics would be able to
+assess the relationship of these aspects on tree establishment, e.g.,
+due to the lack of disturbance or appropriate management modules.
 
@@
 modeling strategies for tree establishment, particularly those related
-to excess tree establishment and probably late excess mortality, by
-focusing on a wider range of tree sizes and the related ingrowth
-dynamics. We also recommend, especially for the landscape-level models,
-the inclusion of explicit spatial considerations regarding tree ingrowth
-[@beckage2003]; this, however, is facing a serious challenge regarding
-appropriate datasets, as inventory data have wide coverage, but do not
-allow for the assessment of spatial interaction effects.
+to excess tree establishment and late excess mortality, by focusing on a
+wider range of tree sizes and the related ingrowth dynamics. We also
+recommend, especially for the landscape-level models, the inclusion of
+explicit spatial considerations regarding tree ingrowth [@beckage2003];
+this, however, is facing a serious challenge regarding appropriate
+datasets, as inventory data have wide coverage, but do not allow for the
+assessment of spatial interaction effects.
 
@@
 establishment niche is not always captured accurately as a function of
-the biotic (light) and abiotic (temperature and moisture).
+the biotic (light) and abiotic (temperature and moisture) factors.
 
@@
 feature a particularly high overestimation of ingrowth levels are
-compensating this by a larger tree mortality. Often, however, this
+compensating for this by a larger tree mortality. Often, this
 compensation is not sufficient to reduce the high ingrowth levels to
-realistic values. Overall, however, there is no clear pattern across all
-models.
+realistic values. Overall, there is no clear mortality pattern across
+all models.
 
@@
 in the development of any model are more important for its behavior
-(accuracy) than scale (stand, landscape, global), background (empirical
-vs. process-based), and complexity. Having both empirical and
+(accuracy) than scale (stand, landscape, global), modelling approach
+(empirical vs. process-based), and complexity. Having both empirical and
 process-based models in our set, the empirically-based models could have

### 2023-03-11 manuscript/manuscriptRW.Rmd
re-define questions of the paper in the introduction

@@
 nocite: |
-  @reyer, @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2004, @morin2021
+  @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @langan2017, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2004, @morin2021, @lasch-born2020, @hickler2012, @fischer_lessons_2016
 linkcolor: blue
@@
 
-More specifically, we aimed to answer the following questions. 1) Are
-models of forest dynamics capturing accurately tree ingrowth levels
-across environmental gradients? 2) Are they matching the initial species
-composition of the tree ingrowth layer? And 3), do they correctly
-represent the mortality in tree establishment?
-
-NEW
+More specifically, we aimed to answer the following questions:
 
@@
 
-2)  Species specific recruitment values in ennvironmental gradients?
-
-3)  Are the models traits linked to the tree establishment values?
-
-environmental gradients?
-
-2)  Are they matching the initial species composition of the tree
-    ingrowth layer? And
-
-3), do they correctly represent the mortality in tree establishment?
+2)  Do model traits explain differences in model performance?
+
+3)  How well do models capture total ingrowth and individual species
+    regeneration niches across environmental gradients of light
+    availability, temperature, and soil moisture?
 
@@
 modelDF <- readr::read_csv(here::here("data", "modelsDescription.csv"))
-modelDF$`Species simulated`[modelDF$`Species simulated` == "F. sylbatica, P. sylvestris, P. abies, A. alba"] <- "F. sylbatica,  P. sylvestris, P. abies,  A. alba"
 modelDF$Sampling <- NULL
 
-#there is a better way of doing footnotes, have a look here https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf and change
+#footnotes
 names(modelDF)[2] <- paste0(names(modelDF)[2], "(a)")
@@
 names(modelDF)[5] <- paste0(names(modelDF)[5], "(d)")
-names(modelDF)[6] <- paste0(names(modelDF)[6],"(e)")
+names(modelDF)[6] <- paste0(names(modelDF)[6], "(e)")
 names(modelDF)[11] <- paste0(names(modelDF)[11],"(f)")
@@
 kable(modelDF, 
-      #format = "markdown",
       format = "latex", booktabs = TRUE,
@@
 
-```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network that were used in the present study, showing the dominant tree species per site in the observed ingrowth at the 10 cm threshold.", out.width='80%'}
+```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network used in the present study, the colors show the dominant tree species per site in the observed ingrowth at the 10 cm threshold.", out.width='80%'}
 knitr::include_graphics(here::here("figures", "theMap.png"))
@@
 \begin{equation}
-H_{n,r} = - \sum_{i = 1}^{s}{p_{n,i}ln(p_{n,r,i})}\label{eqpiShannonR}
+H_{n,r} = - \sum_{i = 1}^{s}{p_{n,i}\ln(p_{n,r,i})}\label{eqpiShannonR}
 \end{equation}
@@
 \begin{equation}
-H_{n,s} = - \sum_{i = 1}^{s}{p_{n,i}ln(p_{n,i})}\label{eqpiShannon}
+H_{n,s} = - \sum_{i = 1}^{s}{p_{n,i}\ln(p_{n,i})}\label{eqpiShannon}
 \end{equation}
@@
 
-```{r over, echo=FALSE, fig.cap="Mean ingrowth rates across all samples per site, plotted for the 200 sites and for each model.
-The red dotted lines shows the 2.5 and 97.5 percentiles for the 7 cm diameter threshold in the observed data.
+```{r over, echo=FALSE, fig.cap="Mean ingrowth levels across all samples per site, plotted for the 200 sites and for each model.
+The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
 There are two boxplots for each model where the left lighted boxplot corresponds to 7cm and the right darker boxplot to 10cm", out.width='100%'}
@@
 temperature gradient (Figure S2), all the models that featured a
-decrease in ingrowth rates with increasing soil moisture showed the
+decrease in ingrowth levels with increasing soil moisture showed the
 opposite trend (i.e., more ingrowth with increasing temperature), except
@@
 ```{r H710, echo=FALSE, fig.cap="Shannon index of tree ingrowth, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
-The red dotted lines represent the 20th percentile of the observed data for the threshold of 7 cm and the 70th percentile of the observed data for the threshold of 10 cm.", out.width='100%'}
+The red dashed lines represent the 25th percentile of the observed data for the threshold of 7 cm and the 75th percentile of the observed data for the threshold of 10 cm.", out.width='100%'}
     knitr::include_graphics(here::here("figures", "H7_10.png"))
@@
 Lastly, there is no model that performed well across the five species
-regarding the species-specific ingrowth rates in the environmental space
-of Figure 6. Some models (e.g., ForClim11, PICUS) tended to consistently
-overestimate the ingrowth share of all five species, but most models
-overestimated the ingrowth share of some species while underestimating
-the share of others (e.g., ForCEEPS, LPJ-GUESS).
-
-```{r diff, echo=FALSE, fig.cap="Ingrowth rates of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm ingrowth threshold.
+regarding the species-specific ingrowth levels in the environmental
+space of Figure 6. Some models (e.g., ForClim11, PICUS) tended to
+consistently overestimate the ingrowth share of all five species, but
+most models overestimated the ingrowth share of some species while
+underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
+
+```{r diff, echo=FALSE, fig.cap="Ingrowth levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm ingrowth threshold.
 The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient.
@@
 data, thus at least partially compensating for the general
-overestimation of ingrowth rates (Figure \@ref(fig:over) and Table S5);
+overestimation of ingrowth levels (Figure \@ref(fig:over) and Table S5);
 six models did not feature a significant difference compared to the
@@
 ```{r mortality13, echo=FALSE, fig.cap="Ratio of tree ingrowth between the 7 and 10 cm thresholds (ingrowth 7 cm / ingrowth 10 cm) and the overestimation proportion at 7 cm ([simulated -observed] / observed) for the mean ingrowth per model across sites and samples.
-The horizontal dotted lines show a ratio equal to 1, indicating no decrease of ingrowth between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
+The horizontal dashed lines show a ratio equal to 1, indicating no decrease of ingrowth between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
 The vertical line indicates zero overestimation.", out.width='80%'}

### 2023-03-11 code/figures.R
crappy tests to check the 7 / 10 cm relation in figure

@@ -125,6 +125,44 @@
                          r.ba = mean(r.ba),
                          Totba = mean(Totba))
     
+    # tests to see that is ok 
+    sumFOR <- overDTsim[ overDTsim$model == "ForCEEPS",]    
+    is7largetthan10 <- list()
+    # 200 samples per site select 
+    for (i in 1:length(unique(sumFOR$site))) {
+      siteVal <- unique(sumFOR$site)
+      
+      value7 <- (sumFOR[sumFOR$dbh == "7" & sumFOR$site == siteVal[1], ])
+      value10 <- (sumFOR[sumFOR$dbh == "10" & sumFOR$site == siteVal[1], ])
+
+      is7smallerthan10[[i]] <- ifelse(as.numeric(value7$r.trees)>as.numeric(value10$r.trees), TRUE, FALSE)
+      
+    }
+    table( unlist(is7smallerthan10))
+    
+    dataTF <- data.frame(unique(sumFOR$site),unlist(is7smallerthan10) )
+    
+    
+    forceepssite3 <-  forceeps[ forceeps$site == 3, ]
+    
+    forceepssite37 <-  forceepssite3[ forceepssite3$dbh == 7,]
+    forceepssite310 <-  forceepssite3[ forceepssite3$dbh == 10,]
+    
+    
+a <- forceepssite37 |>
+      dplyr::group_by(sample) |> 
+      dplyr::summarise(Total=sum(r.trees))
+mean(a$Total)
+
+b<- forceepssite310 |>
+  dplyr::group_by(sample) |> 
+  dplyr::summarise(Total=sum(r.trees))
+mean(b$Total)
+    
+    
+    mean(forceepssite37$r.trees)
+    mean(forceepssite310$r.trees)
+    
   ### Fig. overestimation 7&10 ----
 
     recOverAll <- ggplot2::ggplot() +
@@ -155,7 +193,7 @@
                                                                na.rm = TRUE)), 
                             colour = "#990000", linetype = "dashed") +
         ggplot2::geom_hline(ggplot2::aes(yintercept = quantile(overDTsim$r.trees[overDTsim$dbh == "10" & overDTsim$model == "Observed"],
-                                                               c(0.70),
+                                                               c(0.75),
                                                                na.rm = TRUE)),
                             colour = "#990000", linetype = "dashed") 
     
@@ -605,7 +643,7 @@
                                                                c(0.25))), 
[…]

### 2023-03-11 code/figures.R
correct 7-10 dbh levles in figures

@@ -31,7 +31,7 @@
     # Assign models and dbh to levels 
 
     outputsDF$model <- factor(outputsDF$model, levels = modelsOrder)
-    outputsDF$dbh <- factor(outputsDF$dbh, levels = c("10", "7"))
+    outputsDF$dbh <- factor(outputsDF$dbh, levels = c("7", "10"))
     
 # Remove the 35 sites from the models data for threshold 7cm as they do not have 
 # observations for those sites
@@ -125,44 +125,6 @@
                          r.ba = mean(r.ba),
                          Totba = mean(Totba))
     
-    # tests to see that is ok 
-    sumFOR <- overDTsim[ overDTsim$model == "ForCEEPS",]    
-    is7largetthan10 <- list()
-    # 200 samples per site select 
-    for (i in 1:length(unique(sumFOR$site))) {
-      siteVal <- unique(sumFOR$site)
-      
-      value7 <- (sumFOR[sumFOR$dbh == "7" & sumFOR$site == siteVal[1], ])
-      value10 <- (sumFOR[sumFOR$dbh == "10" & sumFOR$site == siteVal[1], ])
-
-      is7smallerthan10[[i]] <- ifelse(as.numeric(value7$r.trees)>as.numeric(value10$r.trees), TRUE, FALSE)
-      
-    }
-    table( unlist(is7smallerthan10))
-    
-    dataTF <- data.frame(unique(sumFOR$site),unlist(is7smallerthan10) )
-    
-    
-    forceepssite3 <-  forceeps[ forceeps$site == 3, ]
-    
-    forceepssite37 <-  forceepssite3[ forceepssite3$dbh == 7,]
-    forceepssite310 <-  forceepssite3[ forceepssite3$dbh == 10,]
-    
-    
-a <- forceepssite37 |>
-      dplyr::group_by(sample) |> 
-      dplyr::summarise(Total=sum(r.trees))
-mean(a$Total)
-
-b<- forceepssite310 |>
-  dplyr::group_by(sample) |> 
-  dplyr::summarise(Total=sum(r.trees))
-mean(b$Total)
-    
-    
-    mean(forceepssite37$r.trees)
-    mean(forceepssite310$r.trees)
-    
   ### Fig. overestimation 7&10 ----
 
     recOverAll <- ggplot2::ggplot() +
@@ -192,7 +154,7 @@ mean(b$Total)
                                                                c(0.25),
                                                                na.rm = TRUE)), 
                             colour = "#990000", linetype = "dashed") +
[…]

### 2023-03-11 manuscript/supplementaryRW.Rmd
update results of lm between complexity and overestimation

@@
 
-Table 1: Parameter settings associated with vegetation establishment.
-
-![](images/image-1265081375.png)
-
 #### Simulation length
@@
 
-## Recruitment levels
+## Ingrowth levels, tree diversity and mortality in tree establishment
+
+```{r zeroInflated, echo=FALSE,  message=FALSE}
+library(knitr)
+
+zeroInflated <- readr::read_csv(here::here("data", "noRecruitment.csv"))
+zeroInflated$Model[zeroInflated$Model == "Empirical"] <- "Observations"
+
+kable(zeroInflated,  format = "latex", caption = "Number of total samples, samples with zero recruitment and the proportion of zero recruitment per model and observed data.")|>
+        #kableExtra::landscape() |>
+        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+```
+## Model traits and model performance
+
+### Model complexity VS overestimation
+
+```{r plotComp, echo=FALSE, warning=FALSE, fig.cap="Plot and summary atatiatics of the linear model between the overestimation proportion in ingrowth levels and the mean complexity value of the establisment modules", fig.show = 'hold', out.width = '70%'}
+tableCompOver <- readr::read_csv(here::here("figures", "tableCompOver.csv"))
+
+ggplotRegression <- function(fit){
+require(ggplot2)
+ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
+  geom_point() +
+  stat_smooth(method = "lm", col = "blue") +
+  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
+                     "Intercept =",signif(fit$coef[[1]],5 ),
+                     " Slope =",signif(fit$coef[[2]], 5),
+                     " P =",signif(summary(fit)$coef[2,4], 5)))
+}
+
+fit1 <- lm(overestimation ~ complexity, data = tableCompOver)
+ggplotRegression(fit1)
+
+```
+
+## Total ingrowth and individual species regeneration niches
 
@@
 
-```{r zeroInflated, echo=FALSE,  message=FALSE}
-library(knitr)
-
-zeroInflated <- readr::read_csv(here::here("data", "noRecruitment.csv"))
-zeroInflated$Model[zeroInflated$Model == "Empirical"] <- "Observations"
-
-kable(zeroInflated,  format = "latex", caption = "Number of total samples, samples with zero recruitment and the proportion of zero recruitment per model and observed data.")|>
-        kableExtra::landscape() |>
-        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
-```
-
-## Model complexity
-
-Summary of the linear model of the overestimation proportions versus the
-mean complexity value on the establishment module.
-
-```{r tableComp, echo=FALSE,  message=FALSE, render = 'normal_print'}
-library(knitr)
-
-tableCompOver <- readr::read_csv(here::here("figures", "tableCompOver.csv"))
-jtools::summ(lm(overestimation ~ complexity, data = tableCompOver), model.info = FALSE, digits = 5)
-
-```
-
 ## Recruitment species composition

### 2023-03-11 code/figures.R

@@ -94,26 +94,26 @@
     
     # Zero data
     zerroAll <- as.array(table(zeroInflatedAll$model, zeroInflatedAll$dbh))
-    colnames(zerroAll) <- c("zero10", "zero7")
+    colnames(zerroAll) <- c("zero7", "zero10")
     #Total observations
-    allObs <- as.array(table(outputsDF$model,outputsDF$dbh ))
-    colnames(allObs) <- c("allObs10", "allObs7")
+    allObs <- as.array(table(outputsDF$model, outputsDF$dbh ))
+    colnames(allObs) <- c("allObs7", "allObs10")
     
     inflatedZeroDF <- as.data.frame(cbind(allObs, zerroAll))
     
     inflatedZeroDF$share10 <- round((inflatedZeroDF$zero10 / inflatedZeroDF$allObs10), 4) * 100
     inflatedZeroDF$share7 <- round((inflatedZeroDF$zero7 / inflatedZeroDF$allObs7), 4) * 100
     
-    colnames(inflatedZeroDF) <- c("Total (10 cm)", "Total (7cm)",
-                                  "No recruitment (10 cm)", 
-                                  "No recruitment (7cm)", 
-                                  "Percentage of no recruitment (10 cm)", 
-                                  "Percentage of no recruitment (7 cm)")
+    colnames(inflatedZeroDF) <- c("Total (7 cm)", "Total (10cm)",
+                                  "No recruitment (7 cm)", 
+                                  "No recruitment (10cm)", 
+                                  "Percentage of no recruitment (7 cm)", 
+                                  "Percentage of no recruitment (10 cm)")
     inflatedZeroDF <- tibble::rownames_to_column(inflatedZeroDF, "Model")
     
     data.table::fwrite(inflatedZeroDF, file = "data/noRecruitment.csv")
     
-# Levels regeneration ----
+# Ingrowth levels ----
 
     overDTsim <- outputsDF |> 
         dplyr::group_by(site, sample, dbh, model) |>
@@ -165,404 +165,6 @@
                     dpi = 300, units = "cm", 
                     device = 'png') 
     
-## R VS. env. trends-----
-
-    #Total values across all species per site and sample
-    overDTEmp1 <- outputsDF[outputsDF$model == "Observed", ] |>
-        dplyr::group_by(site, sample, dbh) |>
[…]

### 2023-03-11 manuscript/supplementaryRW.Rmd
deletes significanve obs and H7cm and 10cm tables in suplmentary

--- 2023-02-21 manuscript/supplementaryRW.Rmd
@@
 
-## Recruitment levels
+## Ingrowth levels, tree diversity and mortality in tree establishment
+
+```{r zeroInflated, echo=FALSE,  message=FALSE}
+library(knitr)
+
+zeroInflated <- readr::read_csv(here::here("data", "noRecruitment.csv"))
+zeroInflated$Model[zeroInflated$Model == "Empirical"] <- "Observations"
+
+kable(zeroInflated,  format = "latex", caption = "Number of total samples, samples with zero recruitment and the proportion of zero recruitment per model and observed data.")|>
+        #kableExtra::landscape() |>
+        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+```
+
+```{r sigDiff, echo=FALSE,  message=FALSE}
+library(knitr)
+
+sigDiff <- readr::read_csv(here::here("figures", "sigH7_10.csv"))
+
+kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of Shannon index in each model for 7 and 10 cm threshold.") |>
+        #kableExtra::landscape() |>
+        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+```
+
+## Model traits and model performance
+
+### Model complexity VS overestimation
+
+```{r plotComp, echo=FALSE, warning=FALSE, fig.cap="Plot and summary statistics of the linear model between the overestimation proportion in ingrowth levels and the mean complexity value of the establisment modules", fig.show = 'hold', out.width = '70%'}
+
+knitr::include_graphics(here::here("figures", "compOver.png"))
+
+```
+
+## Total ingrowth and individual species regeneration niches
 
@@
 
-```{r zeroInflated, echo=FALSE,  message=FALSE}
-library(knitr)
-
-zeroInflated <- readr::read_csv(here::here("data", "noRecruitment.csv"))
-zeroInflated$Model[zeroInflated$Model == "Empirical"] <- "Observations"
-
-kable(zeroInflated,  format = "latex", caption = "Number of total samples, samples with zero recruitment and the proportion of zero recruitment per model and observed data.")|>
-        kableExtra::landscape() |>
-        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
-```
-
-## Model complexity
-
-Summary of the linear model of the overestimation proportions versus the
-mean complexity value on the establishment module.
-
-```{r tableComp, echo=FALSE,  message=FALSE, render = 'normal_print'}
-library(knitr)
-
-tableCompOver <- readr::read_csv(here::here("figures", "tableCompOver.csv"))
-jtools::summ(lm(overestimation ~ complexity, data = tableCompOver), model.info = FALSE, digits = 5)
-
-```
-
 ## Recruitment species composition
-
-```{r sigDiff, echo=FALSE,  message=FALSE}
-library(knitr)
-
-sigDiff <- readr::read_csv(here::here("figures", "sigH7_10.csv"))
-
-kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of Shannon index in each model for 7 and 10 cm threshold.") |>
-        #kableExtra::landscape() |>
-        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
-```
-
-```{r sigH7_Obs, echo=FALSE,  message=FALSE}
-library(knitr)
-
-sigDiff7 <- readr::read_csv(here::here("figures", "sigH7_Obs.csv"))
-
-kable(sigDiff7,  format = "latex", caption = "Test of the significance of the differences of Shannon index between the observed data and each model for 7cm threshold.") |>
-        #kableExtra::landscape() |>
-        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
-```
-
-```{r sigH10_Obs, echo=FALSE,  message=FALSE}
-library(knitr)
-
-sigDiff10 <- readr::read_csv(here::here("figures", "sigH10_Obs.csv"))
-
-kable(sigDiff10,  format = "latex", caption = "Test of the significance of the differences of Shannon index between the observed data and each model for 10cm threshold.") |>
-        #kableExtra::landscape() |>
-        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
-```
 

### 2023-03-12 manuscript/supplementaryRW.Rmd
remove figure mort7_10_iqrOver.png

@@
 
+```{r recrStand7, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 7cm", out.width='100%'}
+knitr::include_graphics(here::here("figures", "recruitmentAdultRichness7.png"))
+```
+
+```{r recrStand10, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 10cm", out.width='100%'}
+knitr::include_graphics(here::here("figures", "recruitmentAdultRichness10.png"))
+```
+
+```{r simObadult , echo=FALSE, fig.cap="Mean species richness at the stand level per site in the simulated versus observed species richness.
+The species richness is estimaed using the Shannon index  calculated by basal area.
+A1) groups models that Simulated species composition in adults is higher than the observed, A2) groups models that Simulated species composition in adults is similar to the observed, A3) groups models where the imulated species composition in adults is lower than the observed.
+NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
+knitr::include_graphics(here::here("figures", "allAgesRichness.png"))
+```
+
+```{r simObr, echo=FALSE, fig.cap="Mean species richness per site in the simulated recruitment versus observed species richness.
+The species richness is estimaed using the Shannon index is calculated by basal area.B1) groups models where the simulated species composition is closer than the other models to observed values, B2) groups models where the sites with lower species composition are underestimated, B3) groups models where the simulated species composition recruitment is lower in the whole range compare to the observe.
+NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
+knitr::include_graphics(here::here("figures", "recruitmentRichness.png"))
+```
+
+### Mortality
+
+```{r mortTable, echo=FALSE,  message=FALSE}
+library(knitr)
+
+sigDiff <- readr::read_csv(here::here("figures", "mortRatio7_10.csv"))
+
+kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold over 10 cm threshold.") |>
+#kableExtra::landscape()
+kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+```
+
+```{r tableModels, echo=FALSE,  message=FALSE}
+library(knitr)
+
+tabletrends <- readr::read_csv(here::here("figures", "mort7_10_trends.csv"))
+kable(tabletrends,  format = "markdown", caption = "Models trends in mortality between 7 and 10cm, and the ingrowth at 7cm, based in the mean oer site .")
+```
+
+```{r mortality3 , echo=FALSE, fig.cap="Recruitment ration between 7 and 10cm and the initial recruitment tree number at 7cm.
+Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions", out.width='70%'}
+knitr::include_graphics(here::here("figures", "mort7_10_trends.png"))
+```
+
+```{r mortalityDiv3 , echo=FALSE, fig.cap=" Shannon index mean value across samples per site and in BA difference between 10 and 7 cm ", out.width='70%'}
+knitr::include_graphics(here::here("figures", "H710.png"))
+```
+
+\newpage
+
 ## Model traits and model performance
+
+\newpage
 
@@
 
-## Recruitment species composition
-
-```{r recrStand7, echo=FALSE, fig.cap="Mean Shannon index per site for the recruitment  and at the stand level both for observed and simulated values.
-NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
-knitr::include_graphics(here::here("figures", "recruitmentAdultRichness7.png"))
-```
-
-```{r recrStand10, echo=FALSE, fig.cap="Mean Shannon index per site for the recruitment  and at the stand level both for observed and simulated values.
-NOTE: The recruitment treshold recruitment for treshold 10 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
-knitr::include_graphics(here::here("figures", "recruitmentAdultRichness10.png"))
-```
-
-```{r simObadult , echo=FALSE, fig.cap="Mean species richness at the stand level per site in the simulated versus observed species richness.
-The species richness is estimaed using the Shannon index  calculated by basal area.
-A1) groups models that Simulated species composition in adults is higher than the observed, A2) groups models that Simulated species composition in adults is similar to the observed, A3) groups models where the imulated species composition in adults is lower than the observed.
-NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
-knitr::include_graphics(here::here("figures", "allAgesRichness.png"))
-```
-
-```{r simObr, echo=FALSE, fig.cap="Mean species richness per site in the simulated recruitment versus observed species richness.
-The species richness is estimaed using the Shannon index is calculated by basal area.B1) groups models where the simulated species composition is closer than the other models to observed values, B2) groups models where the sites with lower species composition are underestimated, B3) groups models where the simulated species composition recruitment is lower in the whole range compare to the observe.
-NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
-knitr::include_graphics(here::here("figures", "recruitmentRichness.png"))
-```
-
 ```{r diff, echo=FALSE, fig.cap="Departure of each model in the representation of the main species recruitment relative abundance from the observations across the environmental gradients for a 10cm treshold.
@@
 ```
-
-## Mortality
-
-```{r mortTable, echo=FALSE,  message=FALSE}
-library(knitr)
-
-sigDiff <- readr::read_csv(here::here("figures", "mortRatio7_10.csv"))
-
-kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold over 10 cm threshold.") |>
-#kableExtra::landscape()
-kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
-```
-
-```{r mortality12 , echo=FALSE, fig.cap="Interquartile range of the ratio between 7 and 10cm recruitment and the overestimation proportion (sim-obs/obs) for the mean recruitment trees per model across sites and samples", out.width='70%'}
-knitr::include_graphics(here::here("figures", "mort7_10_iqrOver.png"))
-```
-
-```{r tableModels, echo=FALSE,  message=FALSE}
-library(knitr)
-
-tabletrends <- readr::read_csv(here::here("figures", "mort7_10_trends.csv"))
-kable(tabletrends,  format = "markdown", caption = "Models trends in mortality between and 10cm.")
-```
-
-```{r mortality3 , echo=FALSE, fig.cap="Recruitment ration between 7 and 10cm and the initial recruitment tree number at 7cm.
-Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions", out.width='70%'}
-knitr::include_graphics(here::here("figures", "mort7_10_trends.png"))
-```
-
-```{r mortalityDiv3 , echo=FALSE, fig.cap=" Shannon index mean value across samples per site and in BA difference between 10 and 7 cm ", out.width='70%'}
-knitr::include_graphics(here::here("figures", "H710.png"))
-```

### 2023-03-12 code/figures.R

@@ -463,9 +463,9 @@
     plotData2 <- plotShannonDF2[plotShannonDF2$model %in% selModels, ]
     plotData2 <- plotData2[plotData2$dbh == "7", ]
     
-    labelsMod <- c(paste0("Overpredict", "\n(n=2)"),
-                   paste0("", "\n(n=4)"),
-                   paste0("Underpredict", "\n(n=8)"))
+    labelsMod <- c(paste0("Overpredict", "\n(n=7)"),
+                   paste0("", "\n(n=5)"),
+                   paste0("Underpredict", "\n(n=2)"))
     
     plotData2$model <- factor(plotData2$model, levels = selModels, labels = labelsMod)
     
@@ -484,7 +484,7 @@
                        strip.background =  ggplot2::element_blank(),
                        axis.line = ggplot2::element_line(colour = "black"),
                        legend.position = "right") +
-        #ggplot2::coord_cartesian(xlim = c(0, 0.65), ylim = c(0, 0.65)) +  
+        ggplot2::coord_cartesian(xlim = c(0, 2), ylim = c(0, 2)) +  
         ggplot2::labs(y = bquote(bar(H) * " recruitment"),
                       x = bquote(bar(H) * " recruitment")) +
         ggplot2::facet_wrap(~model) +
@@ -523,7 +523,7 @@
         ggplot2::theme(legend.title = ggplot2::element_blank()) +
         ggplot2::labs(y = bquote(bar(H) * " recruitment"),
                       x = bquote(bar(H) * " stand")) +
-        ggplot2::ggtitle("7 cm recruitment") +
+        ggplot2::ggtitle("7 cm diamter threshold") +
         ggplot2::facet_wrap(~model) +
         ggplot2::geom_abline() +
         ggplot2::scale_color_manual(labels = c("Observed" = "Observed", 
@@ -550,7 +550,7 @@
         ggplot2::theme(legend.title = ggplot2::element_blank()) +
         ggplot2::labs(y = bquote(bar(H) * " recruitment"),
                       x = bquote(bar(H) * " stand")) +
-        ggplot2::ggtitle("10 cm recruitment") +
+        ggplot2::ggtitle("10 cm diameter threshold") +
         ggplot2::facet_wrap(~model) +
         ggplot2::geom_abline() +
         ggplot2::scale_color_manual(labels = c("Observed" = "Observed", 
@@ -615,19 +615,19 @@
     
[…]

### 2023-03-12 code/figures.R
remove figure mort7_10_VarmeannOver.png

@@ -1060,40 +1060,7 @@
     
     ggplot2::ggsave("figures/mort7_10_trends.png",
                     plot = mort7_10_trends,
-                    dpi = 300, units = "cm", device = 'png') 
-
-    
-    ### variability 7-10 vs overestimate per model ----
-    VarmeannOver <- ggplot2::ggplot(modelMean2,
-                                    ggplot2::aes(x = diff, y = median,
-                                                 ymin = mean - sd, ymax = mean + sd,
-                                                 label = model, 
-                                                 color = model)) +
-      ggplot2::geom_point(alpha = .8) +
-      ggplot2::geom_text(fontface = "bold",
-                         position = ggplot2::position_jitter(width = 1,
-                                                             height = 1)) +
-      ggplot2::geom_hline(ggplot2::aes(yintercept = 1), colour = "#990000", linetype = "dashed") +
-      ggplot2::geom_hline(ggplot2::aes(yintercept = 1.77), colour = "#990000", linetype = "dashed") +
-      ggplot2::geom_pointrange() +
-      ggplot2::xlim(c(-1, 15)) +
-      ggplot2::ylim(c(0, 4)) +
-      ggplot2::geom_vline(xintercept = 0, linetype = "dotted", color = "red") +
-      ggplot2::scale_colour_manual(values = values_color) +
-      ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 0),
-                     strip.text.y = ggplot2::element_text(angle = 0),
-                     panel.background = ggplot2::element_blank(), 
-                     axis.line = ggplot2::element_line(colour = "black"),
-                     legend.key =  ggplot2::element_blank(), 
-                     legend.title = ggplot2::element_blank(),
-                     strip.background =  ggplot2::element_blank(),
-                     legend.position = "right") +
-      ggplot2::xlab(label = "Proportion of overestimation in tree recruiment") +
-      ggplot2::ylab(label = "Median and SD of tree share between 10 and 7cm") 
-    
-    ggplot2::ggsave("figures/mort7_10_VarmeannOver.png",
-                    plot =  VarmeannOver,
-                    dpi = 300, units = "cm", device = 'png') 
+                    dpi = 300, units = "cm", device = 'png')
     
 # Complexity model structure-----  
 

### 2023-03-12 code/figures.R
remove inter quantil range estimation in figures script

@@ -870,16 +870,6 @@
     
   ### Fig. median vs overestimate per model ----
     # Prepare the data
-    ## interquartile range (IQR) 
-    ##calculated as the difference between the 3rd quartile and the 1st quartile
-    ##in rate of ntrees between 10 and 7 cm
-    iqrDT <- mortAll[is.finite(mortAll$nn710),] |> 
-      dplyr::group_by(model) |> 
-      dplyr::summarise(iqr7_10 = IQR(nn710, na.rm = TRUE),
-                       median = median(nn710, na.rm = TRUE),
-                       mean = mean(nn710, na.rm = TRUE),
-                       sd = sd(nn710, na.rm = TRUE) )
-    
     # Overestimation at 7cm   
     dbhSel <- 7
     simResdbh2 <- outputsDF |> dplyr::filter(dbh %in% dbhSel) # this is only for dbh 7cm

### 2023-03-12 code/figures.R
add back iqrDT in figures

@@ -870,8 +870,17 @@
     
   ### Fig. median vs overestimate per model ----
     # Prepare the data
+    ##calculated as the difference between the 3rd quartile and the 1st quartile
+    ##in rate of ntrees between 10 and 7 cm
+    iqrDT <- mortAll[is.finite(mortAll$nn710),] |>
+                dplyr::group_by(model) |> 
+                dplyr::summarise(iqr7_10 = IQR(nn710, na.rm = TRUE),
+                                 median = median(nn710, na.rm = TRUE),
+                                 mean = mean(nn710, na.rm = TRUE),
+                                 sd = sd(nn710, na.rm = TRUE) )
+    
     # Overestimation at 7cm   
-    dbhSel <- 7
+    dbhSel <- "7"
     simResdbh2 <- outputsDF |> dplyr::filter(dbh %in% dbhSel) # this is only for dbh 7cm
     aggNrecr <- simResdbh2 |>  
       dplyr::group_by(model, site, sample) |> 
@@ -991,7 +1000,7 @@
     
     prep2 <- mortAllnoInf  |>
       dplyr::group_by(model) |> 
-      dplyr::do(mod = lm(nn710 ~ nn7, na.action=na.omit, data = .)) |>
+      dplyr::do(mod = lm(nn710 ~ nn7, na.action = na.omit, data = .)) |>
       dplyr::mutate(Slope = summary(mod)$coeff[2],
                     Significance = summary(mod)$coeff[2, 4]) |> 
       dplyr::ungroup() |>
@@ -1003,7 +1012,6 @@
     # Categorize the trends 
     
     trends <- prep2[, c("model", "Slope", "Significance")]
-    
     trends$trend[trends$Slope < 0] <- "negative"
     postive <- trends[trends$Slope > 0,]
     median(postive$Slope)
@@ -1023,7 +1031,7 @@
     
     # Plot the trends 
     mort7_10_trends <- 
-      ggplot2::ggplot(mortAllnoInf, ggplot2::aes(x = nn7, y = nn710, color = model)) +
+      ggplot2::ggplot(mortAllnoInf , ggplot2::aes(x = nn7, y = nn710, color = model)) +
       ggplot2::geom_point(alpha = .2) +
       ggplot2::ylim(c(0, 3.5)) +
       ggplot2::xlim(c(.2, 100)) +
@@ -1041,7 +1049,7 @@
       ggplot2::xlab(label = expression("R (trees ha"^-1 * "10yr"^-1 * ")(7cm) (mean per site)")) +
[…]

### 2023-03-12 code/figures.R
remove old code for mortality trends in appendix

@@ -937,56 +937,7 @@
     
     
     #### FigS. R change from 7 ----   
-    
-    medianPoints <- mortAll |>
-      dplyr::group_by(model) |> 
-      dplyr::summarise(nn7 = median(nn7, na.rm = TRUE),
-                       nn10 = median(nn10, na.rm = TRUE),
-                       nn710 = median(nn710, na.rm = TRUE))  
-    
-    
-    mort7_10_a <- ggplot2::ggplot(mortAll, ggplot2::aes(x = nn7, y = nn710, color = model)) +
-      ggplot2::geom_point(alpha = .2) +
-      ggplot2::ylim(c(0, 3.5)) +
-      ggplot2::xlim(c(.2, 100)) +
-      ggplot2::geom_hline(ggplot2::aes(yintercept = 1), colour = "#990000", linetype = "dashed") +
-      ggplot2::geom_hline(ggplot2::aes(yintercept = 1.77), colour = "#990000", linetype = "dashed") +
-      ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 0),
-                     strip.text.y = ggplot2::element_text(angle = 0),
-                     panel.background = ggplot2::element_blank(), 
-                     axis.line = ggplot2::element_line(colour = "black"),
-                     legend.key =  ggplot2::element_blank(), 
-                     legend.title = ggplot2::element_blank(),
-                     strip.background =  ggplot2::element_blank()) +
-      ggplot2::scale_colour_manual(values = values_color) +
-      ggplot2::scale_fill_manual(values = values_color) +
-      ggplot2::xlab(label = expression("R (trees ha"^-1 * "10yr"^-1 * ")(7cm) (mean per site)")) +
-      ggplot2::ylab(label = "R (7cm) / R (10cm) (mean per site)") 
-    
-    mort7_10 <- mort7_10_a + ggplot2::geom_point(data = medianPoints, size = 6, alpha = .9, 
-                                                 pch = 21, ggplot2::aes(fill = model))  
-    
-    ggplot2::ggsave("figures/mort7_10.png",
-                    plot = mort7_10,
-                    dpi = 300, units = "cm", device = 'png') 
-    
-    mort7_10_trends <- mort7_10_a +
-      ggplot2::geom_smooth(method = lm, level = 0.99, se = F,
-                           ggplot2::aes(group = model)) +
-      ggplot2::geom_smooth(mortAll, 
-                           mapping = ggplot2::aes(group = 1,
[…]

### 2023-03-12 code/figures.R
remove the categorization of the mortality trends in figures script

@@ -959,27 +959,7 @@
 
     write.csv(prep2[, c("model", "Slope", "Significance")],
               "figures/mort7_10_trends.csv", row.names = FALSE)
-    
-    # Categorize the trends 
-    
-    trends <- prep2[, c("model", "Slope", "Significance")]
-    trends$trend[trends$Slope < 0] <- "negative"
-    postive <- trends[trends$Slope > 0,]
-    median(postive$Slope)
-    trends$trend[trends$Slope <  median(postive$Slope) & trends$Slope > 0] <- "positve"
-    trends$trend[trends$Slope >=  median(postive$Slope)] <- "strong positive"
-    
-    negative <- paste0(trends$model[trends$trend == "negative"],sep = ", ", collapse = "")
-    positive <- paste0(trends$model[trends$trend == "positve"],sep = ", ", collapse = "")
-    positvestrong <- paste0(trends$model[trends$trend == "strong positive"],
-                            sep = ", ", collapse = "")
-    
-    tabletrend <- data.frame(negative =  negative, positive  = positive , 
-                             positvestrong = positvestrong)
-    colnames(tabletrend) <- c("Negative", "Positive", "Strong positive")
-    write.csv(tabletrend,
-              "figures/summaryTrend_mort7_10_trends.csv", row.names = FALSE)
-    
+
     # Plot the trends 
     mort7_10_trends <- 
       ggplot2::ggplot(mortAllnoInf , ggplot2::aes(x = nn7, y = nn710, color = model)) +

### 2023-03-12 manuscript/supplementaryRW.Rmd
remove H710.png from figures script and appendix document

@@
 
-```{r zeroInflated, echo=FALSE,  message=FALSE}
+```{r zeroInflated, echo=FALSE,  message=FALSE, fig.pos='H'}
 library(knitr)
@@
 kable(zeroInflated,  format = "latex", caption = "Number of total samples, samples with zero recruitment and the proportion of zero recruitment per model and observed data.")|>
+        kableExtra::kable_styling(latex_options = "HOLD_position")|>
         #kableExtra::landscape() |>
@@
 
-```{r sigDiff, echo=FALSE,  message=FALSE}
+```{r sigDiff, echo=FALSE,  message=FALSE, fig.pos='H'}
 library(knitr)
@@
 kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of Shannon index in each model for 7 and 10 cm threshold.") |>
+        kableExtra::kable_styling(latex_options = "HOLD_position")|>
         #kableExtra::landscape() |>
@@
 
-```{r recrStand7, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 7cm", out.width='100%'}
+```{r recrStand7, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 7cm", out.width='100%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "recruitmentAdultRichness7.png"))
@@
 
-```{r recrStand10, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 10cm", out.width='100%'}
+```{r recrStand10, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 10cm", out.width='100%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "recruitmentAdultRichness10.png"))
@@
 A1) groups models that Simulated species composition in adults is higher than the observed, A2) groups models that Simulated species composition in adults is similar to the observed, A3) groups models where the imulated species composition in adults is lower than the observed.
-NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
+NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "allAgesRichness.png"))
@@
 The species richness is estimaed using the Shannon index is calculated by basal area.B1) groups models where the simulated species composition is closer than the other models to observed values, B2) groups models where the sites with lower species composition are underestimated, B3) groups models where the simulated species composition recruitment is lower in the whole range compare to the observe.
-NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%'}
+NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "recruitmentRichness.png"))
@@
 
-```{r mortTable, echo=FALSE,  message=FALSE}
+```{r mortTable, echo=FALSE,  message=FALSE, fig.pos='H'}
 library(knitr)
@@
 
-kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold over 10 cm threshold.") |>
+kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold over 10 cm threshold between the observed values and the simulated values per model.") |>
+    kableExtra::kable_styling(latex_options = "HOLD_position")|>
 #kableExtra::landscape()
@@
 
-```{r tableModels, echo=FALSE,  message=FALSE}
+```{r tableModels, echo=FALSE,  message=FALSE, fig.pos='H'}
 library(knitr)
@@
 tabletrends <- readr::read_csv(here::here("figures", "mort7_10_trends.csv"))
-kable(tabletrends,  format = "markdown", caption = "Models trends in mortality between 7 and 10cm, and the ingrowth at 7cm, based in the mean oer site .")
+kable(tabletrends,  format = "latex", caption = "Models trends in mortality between 7 and 10cm, and the ingrowth at 7cm, based in the mean oer site .")|>
+kableExtra::kable_styling(latex_options = "HOLD_position")
 ```
@@
 ```{r mortality3 , echo=FALSE, fig.cap="Recruitment ration between 7 and 10cm and the initial recruitment tree number at 7cm.
-Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions", out.width='70%'}
+Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions", out.width='70%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "mort7_10_trends.png"))
@@
 
-```{r mortalityDiv3 , echo=FALSE, fig.cap=" Shannon index mean value across samples per site and in BA difference between 10 and 7 cm ", out.width='70%'}
-knitr::include_graphics(here::here("figures", "H710.png"))
-```
-
 \newpage
@@
 
-```{r plotComp, echo=FALSE, warning=FALSE, fig.cap="Plot and summary statistics of the linear model between the overestimation proportion in ingrowth levels and the mean complexity value of the establisment modules", fig.show = 'hold', out.width = '70%'}
+```{r plotComp, echo=FALSE, warning=FALSE, fig.cap="Plot and summary statistics of the linear model between the overestimation proportion in ingrowth levels and the mean complexity value of the establisment modules", fig.show = 'hold', out.width = '70%', fig.pos='H'}
 
@@
 (A) same y-axis across all models; (B) y-axis scaled by model.
-The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
+The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%', fig.pos='H'}
     knitr::include_graphics(c(here::here("figures", "envTrend7WB_500.png"),
@@
 (A) same y-axis across all models; (B) y-axis scaled by model.
-The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
+The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%', fig.pos='H'}
     knitr::include_graphics(c(here::here("figures", "envTrend7DDS_500.png"),
@@
 The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species.
-The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equalt to zero.", out.width='100%'}
+The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equalt to zero.", out.width='100%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_10.png"))

### 2023-03-12 code/figures.R

@@ -297,52 +297,6 @@
                     #width = 55, height = 18, scale = 0.9,
                     dpi = 300, units = "cm", device = 'png') 
     
-    #### FigS. H10-H7 ----
-    
-    ShannonIndexM <- ShannonIndex |>  dplyr::group_by(model, site, dbh) |> 
-        dplyr::summarise(ShannonIndexRecruit = mean(ShannonIndexRecruit),
-                         ShannonIndexAllAges = mean(ShannonIndexAllAges),
-                         relShannon = mean(relShannon))
-    
-    S10 <- ShannonIndexM[ShannonIndexM$dbh == "10",]
-    colnames(S10) <- c("model", "site", "dbh", "ShannonIndexRecruit10", 
-                       "ShannonIndexAllAges10", "relShannon10")
-    S7 <- ShannonIndexM[ShannonIndexM$dbh == "7",]
-    colnames(S7) <- c("model", "site", "dbh", "ShannonIndexRecruit7", 
-                      "ShannonIndexAllAges7", "relShannon7")
-    
-    S107 <- merge(S10[, c("model", "site", "ShannonIndexRecruit10", 
-                          "ShannonIndexAllAges10", "relShannon10")], 
-                  S7[,  c("model", "site", "ShannonIndexRecruit7", 
-                          "ShannonIndexAllAges7", "relShannon7")],
-                  by = c("site",  "model"))
-    
-    S107$H710 <- S107$ShannonIndexRecruit10 - S107$ShannonIndexRecruit7
-    
-    H710 <- ggplot2::ggplot(S107,
-                            ggplot2::aes(y = H710, 
-                                         x = model, 
-                                         fill = model)) + 
-        ggplot2::geom_boxplot(notch = F) +
-        #ggplot2::ylim(c(-0.5, 0.5)) +
-        ggplot2::xlab(label = "") +
-        ggplot2::ylab(label = "H10 - H7") +
-        ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90),
-                       strip.text.y = ggplot2::element_text(angle = 0),
-                       legend.position = "none",
-                       panel.background = ggplot2::element_blank(), 
-                       axis.line = ggplot2::element_line(colour = "black"),
-                       legend.key =  ggplot2::element_blank(), 
-                       legend.title = ggplot2::element_blank(),
-                       strip.background =  ggplot2::element_blank()) +
[…]

### 2023-03-12 manuscript/manuscriptRW.Rmd
combine the 3 gradients figure in one figure and add the zoomed ones into the appendix

@@
     toc: no
+    extra_dependencies: "subfig"
   word_document:
@@
 isolate the importance of the canopy feedback (i.e. simulated actual
-composition and relative abundance of species in the plot)via seed trees
-for the quantity and quality of simulated ingrowth, as one variant
+composition and relative abundance of species in the plot) via seed
+trees for the quantity and quality of simulated ingrowth, as one variant
 includes this feedback whereas the other does not.
@@
 The simulation results were analyzed regarding the ingrowth levels
-(i.e., ingrowth number per unit time and space), species composition of
-the ingrowth, and ingrowth mortality. The observed data were modeled
-using a Generalized Additive Model [@wood2011] with a negative binomial
-distribution and restricted maximum likelihood to better understand the
-environmental gradients trend in the levels of the observed ingrowth,
-relative to the simulation results per model.
-
-Ingrowth levels were evaluated along three key environmental gradients,
-i.e., stand basal area (the basal area of all the trees in each sample
-including tree establishment) as a proxy of light availability at the
-forest floor, the annual degree-day sum (DD) [@allen1976; @fischlin1995]
-as a proxy for growing season warmth, and the climatic water balance
-(CWB) as a proxy for soil moisture [@speich2019]. The mean ingrowth
-values were calculated as the mean across the 200 samples per site.
-
-We evaluated the species diversity across models and in relation to the
-observed data by calculating the Shannon index based on the relative
-proportion of the species in terms of basal area. The Shannon index was
-calculated for the establishment ($H_{n,r}$) (Equation
-\ref{eqpiShannonR}) and at the stand level ($H_{n,s}$)(Equation
-\ref{eqpiShannon}). The higher the value of the index, the higher the
-species diversity at a particular site and sample.
-
-We also tested for significant trends across XXXXX !!!!!
+(i.e., ingrowth number per unit time and space), species diversity of
+the ingrowth, and ingrowth mortality. We evaluated the species diversity
+across models and in relation to the observed data by calculating the
+Shannon index based on the relative proportion of the species in terms
+of basal area. The Shannon index was calculated for the establishment
+($H_{n,r}$) (Equation \ref{eqpiShannonR}) and at the stand level
+($H_{n,s}$)(Equation \ref{eqpiShannon}). The higher the value of the
+index, the higher the species diversity at a particular site and sample.
 
@@
 to the total basal area ($TotalR_{BA_{n}}$) of the sample $n$.
-
-In order to analyze the relationship between the adult species
-composition and ingrowth across the climatic gradients, we focused on
-the five main tree species or genera: *Abies alba*, *Fagus sylvatica*,
-*Picea abies*, *Pinus sylvestris*, and *Quercus* spp. For these species,
-we calculated the mean ingrowth basal area share per site as the mean
-value across the available samples per site:
-
-```{=tex}
-\begin{equation}
-\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{BA_{t,i}}}{\sum_{i = 1}^{s}{\bar{r}_{{BA_{t,i}}}}} \label{eq:BAshare}
-\end{equation}
-```
-where $s$ is the total number of species simulated; $\bar{r}_{t,i}$ is
-the mean basal area ingrowth of the species $i$ per site $t$ across the
-available samples at that site.
-
-We removed 35 sites from the comparisons at 7 cm ingrowth, because field
-data did not include these measurements. Furthermore, regarding the
-ingrowth basal area share per species ($\bar{R}_{BA_{share_{i}}}$), we
-categorized this as zero when the ingrowth basal area of that species
-($\bar{r}_{BA_{t,i}}$) was zero but also when both the ingrowth basal
-area of that species ($\bar{r}_{BA_{t,i}}$) and the total ingrowth basal
-area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.
 
@@
 ```
+We assessed model traits in relation to model performance focusing on
+features such as model complexity as defined in @bugmann2022, model type
+(empirical or process based), models with and without canopy feedback
+(the level and species composition of the ingrowth is influenced by the
+existing adult tree community) and the models scale (stand, landscape or
+global). We tested for significant differences considering each of these
+model features and between observed and simulated ingrowth
+overestimation, ingrowth levels and species diversity respectively.
+
+The total ingrowth and individual species regeneration niches were
+evaluated across environmental gradients of light availability,
+temperature, and soil moisture. Ingrowth levels were evaluated along
+three key environmental gradients, i.e., stand basal area (the basal
+area of all the trees in each sample including tree establishment) as a
+proxy of light availability at the forest floor, the annual degree-day
+sum (DD) [@allen1976; @fischlin1995] as a proxy for growing season
+warmth, and the climatic water balance (CWB) as a proxy for soil
+moisture [@speich2019]. The mean ingrowth values were calculated as the
+mean across the 200 samples per site. The observed data were modeled
+using a Generalized Additive Model [@wood2011] with a negative binomial
+distribution and restricted maximum likelihood to better understand the
+environmental gradients trend in the levels of the observed ingrowth,
+relative to the simulation results per model.
+
+In order to analyze the individual species regeneration niches across
+the climatic gradients, we focused on the five main tree species or
+genera: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus
+sylvestris*, and *Quercus* spp. For these species, we calculated the
+mean ingrowth basal area share per site as the mean value across the
+available samples per site:
+
+```{=tex}
+\begin{equation}
+\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{BA_{t,i}}}{\sum_{i = 1}^{s}{\bar{r}_{{BA_{t,i}}}}} \label{eq:BAshare}
+\end{equation}
+```
+where $s$ is the total number of species simulated; $\bar{r}_{t,i}$ is
+the mean basal area ingrowth of the species $i$ per site $t$ across the
+available samples at that site.
+
+We removed 35 sites from the comparisons at 7 cm ingrowth, because field
+data did not include these measurements. Furthermore, regarding the
+ingrowth basal area share per species ($\bar{R}_{BA_{share_{i}}}$), we
+categorized this as zero when the ingrowth basal area of that species
+($\bar{r}_{BA_{t,i}}$) was zero but also when both the ingrowth basal
+area of that species ($\bar{r}_{BA_{t,i}}$) and the total ingrowth basal
+area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.
+
 # Results
 
-## Ingrowth levels
+## Ingrowth levels, tree diversity and mortality in tree establishment
 
@@
 stand model SIBYLA, the landscape model Landis-II and the global model
-aDGVM2, which estimated ingrowth rates at the lower end of the
+aDGVM2, which estimated ingrowth levels at the lower end of the
 plausibility interval of the observed data (Figure \@ref(fig:over)). The
@@
 ```{r over, echo=FALSE, fig.cap="Mean ingrowth levels across all samples per site, plotted for the 200 sites and for each model.
-The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
+The red dashed lines shows the 25th and 75th percentiles for the 7cm diameter threshold in the observed data.
 There are two boxplots for each model where the left lighted boxplot corresponds to 7cm and the right darker boxplot to 10cm", out.width='100%'}
@@
 ```
+
+Both observations and simulated data had no ingrowth in some samples and
+at some sites (Table S1). The observed data had 4% of the samples with
+no ingrowth. Only three models had a larger proportion of no ingrowth
+(4C, Landis-II and aDVM2). Two models always simulated ingrowth for both
+the 7 and 10 cm threshold (xComp and TreeMig), i.e., they did not
+feature any zero values. The other models had a very low percentage of
+samples with no ingrowth (0.01% - 2.39%), i.e., they had distinctly
+fewer occurrences of zeros compared to the observations.
+
+Most models matched the level of diversity of the observed data quite
+well as quantified by the Shannon index (Figure \@ref(fig:H710)). Five
+models overestimated ingrowth diversity: ForCEEPS, ForClim 11, PICUS,
+TreeMig and LandClim, the latter mostly for the 7 cm diameter threshold.
+The model 4C is a special case, as it simulated five species only, i.e.,
+its diversity values are not directly comparable to those of the other
+models, nor to the observations. Only one model, Landis-II, consistently
+underestimated ingrowth diversity.
+
+```{r H710, echo=FALSE, fig.cap="Shannon index of tree ingrowth, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
+The red dashed lines represent the 25th percentile of the observed data for the threshold of 7 cm and the 75th percentile of the observed data for the threshold of 10 cm.", out.width='100%'}
+    knitr::include_graphics(here::here("figures", "H7_10.png"))
+```
+
+In most models, there were only small but significant differences in the
+species diversity of ingrowth between 7 and 10 cm across sites (Tables
+S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained
+the ingrowth diversity between the 7 and 10 cm thresholds (i.e., the
+differences between them were not significant, see Table S2), and the
+same was evident from the observed data.
+
+For both the observed and the simulated data, and both ingrowth
+thresholds, species diversity in the ingrowth compared to stand-level
+diversity did not have clear patterns across all the models (Figures S1
+and S2). Most models captured reasonably well or overpredicted species
+diversity at the stand level (Figure \@ref(fig:combinedRichness)A,
+center; cf. Figure S3), but a characteristic feature was that
+simulations had a much lower variability of diversity compared to
+observations. Models overpredicting species diversity included ForClim1,
+ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A,
+left). Several models underpredicted species diversity, i.e., LPJ-GUESS,
+ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).
+
+The majority of the models overpredicted ingrowth diversity and only two
+of them underestimated ingrowth diversity (4C and Landis-II) (Figure
+\@ref(fig:combinedRichness)A and C, cf. Figure S4). Both the models with
+feedback and those without feedback from the adult trees to ingrowth
+(via seed production) overestimate, underestimated and captured
+reasonably well ingrowth species diversity compared to the observations.
+
+```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site and for observed and simulated values.
+n indicates the number of models falling in each group.
+A) Examples for the three trends across models for species diversity at the stand level.
+B) Examples for the three trends across models for ingrowth at the 7 cm threshold.
+The selection of the representative model for each trend group was based on the full data shown in Figures S5 and S6.", out.width='75%'}
+knitr::include_graphics(here::here("figures", "combined_richness.png"))
+```
+
+There were strong differences among the models regarding the mortality
+rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the
+models featured a mortality rate significantly larger than the observed
+data, thus at least partially compensating for the general
+overestimation of ingrowth levels (Figure \@ref(fig:over) and Table S5);
+six models did not feature a significant difference compared to the
+observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
+Four models (xComp, PICUS, LandClim and TreeMig) featured very high
+mortality rates (i.e., well above the Reineke self-thinning line), which
+compensated for the strong initial overestimation of ingrowth (cf.
+Figure \@ref(fig:over)). Interestingly, two models that underestimated
+overall ingrowth levels (SIBYLA and Landis-II; cf. Figure
+\@ref(fig:over)) featured mortality rates that were close to but still
+above the self-thinning line (Figure \@ref(fig:mortality11)).
+
+```{r mortality11, echo=FALSE, fig.cap="Ratio of tree ingrowth rates between the 7 and 10 cm thresholds.
+Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree ingrowth between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
+knitr::include_graphics(here::here("figures", "ratio7_10.png"))
+```
+
+The fact that some models (and observed data) featured lower mortality
+than expected by Reineke's self-thinning rule may point to facilitation,
+or simply a higher mortality rate before the trees had reached 7 cm
+diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
+whose mortality was well above the self-thinning line, indicates that
+they feature higher ingrowth mortality in mixed-species forests compared
+to even-aged, single-species forests. Some models yielded a mortality
+rate of ingrowth that is broadly compatible with the self-thinning rule
+(e.g., 4C and LandClim).
+
+The three models with the largest overestimation of the proportion of
+ingrowth at the 7 cm threshold featured the highest mortality between
+the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The other
+models that overestimated ingrowth had mortalities lower than the
+expected self-thinning ratio with the exception of LandClim (Figure
+\@ref(fig:mortality13)B). It is noteworthy that most of the models that
+featured a low ratio of ingrowth between 7 and 10 cm (i.e., ForCEEPS,
+ForClim 1, ForClim 11 and iLand) had only a small overestimation of
+ingrowth at the 7 cm threshold (Figure \@ref(fig:mortality11)). aDGVM2
+was the only model that underestimated ingrowth at the 7 cm threshold
+while featuring a low mortality between the two thresholds (Figure
+\@ref(fig:mortality13)C), and only a few models were close to the
+correct ingrowth rate at 7 cm while also being close to the Reineke line
+(SIBYLA, Landis-II; cf. Figure \@ref(fig:mortality13)D; and 4C as well
+as LandClim).
+
+```{r mortality13, echo=FALSE, fig.cap="Ratio of tree ingrowth between the 7 and 10 cm thresholds (ingrowth 7 cm / ingrowth 10 cm) and the overestimation proportion at 7 cm ([simulated -observed] / observed) for the mean ingrowth per model across sites and samples.
+The horizontal dashed lines show a ratio equal to 1, indicating no decrease of ingrowth between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
+The vertical line indicates zero overestimation.", out.width='80%'}
+knitr::include_graphics(here::here("figures", "mort7_10_medianOver.png"))
+```
+
+Lastly, some models that had a low number of individuals at 7 cm
+threshold also had a smaller decrease of ingrowth between 7 and 10 cm
+(ratio 7-10 cm) (Figure S5). In these models, competition and
+self-thinning are either not pronounced or must have occurred before the
+species had reached 7 cm. However, this pattern was not consistent
+across models (Figure S5). It is noteworthy that there is a relationship
+between the ingrowth rate at 7 cm and the mortality rate (Table S4):
+most of the models showed a positive effect (higher ingrowth at 7 cm is
+coupled to an increase of the mortality between 7 and 10 cm). While some
+models showed a negative effect, this linear trend was not pronounced
+and not always significant regarding the slope.
+
+## Model traits and model performance
+
+There was no significant relationship between the mean complexity of the
+establishment module (Table \@ref(tab:tableModels)) and the
+overestimation proportion at a diameter threshold of 7 cm (Figure S7).
+All the models except 4C had significant differences between the
+observed mean ingrowth values and the simulated, therefore it was not
+possible to differentiate in terms of performance between models forest
+type (empirical or process based) and across models scales (stand,
+landscape or global) (Table S5). The ingrowth species diversity values
+simulated per models were also assessed againts the observed data to see
+if there was a difference between models that include feedback compare
+to those that did not include feedback, all the models except
+ForCEEPs(f) and FORMIND had significant differences (Table S6). It is
+noteworthy the change from overpredicted species diversity in the
+ingrowth for the model ForCEEPS to a closer diversity levels when the
+recruitment module included feedback (ForCEEPs(f)). Again, simulated
+diversity was considerably less variable than observed diversity.
+
+## Total ingrowth and individual species regeneration niches
 
@@
 
-Both observations and simulated data had no ingrowth in some samples and
-at some sites (Table S1). The observed data had 3% of the samples with
-no ingrowth. Only three models had a larger proportion of no ingrowth
-(4C, Landis-II and aDVM2). CHECK! No LPJ-GUESS anymore with the new data
-submission!!!! Two models always simulated ingrowth for both the 7 and
-10 cm threshold (TreeMig and LPJ-GUESS), i.e., they did not feature any
-zero values. The other models had a very low percentage of samples with
-no ingrowth (0.01% - 1.31%), i.e., they had distinctly fewer occurrences
-of zeros compared to the observations. There was no significant
-relationship between the mean complexity of the establishment module
-(Table \@ref(tab:tableModels)) and the overestimation proportion at a
-diameter threshold of 7 cm (SM3).
-
-```{r nvTrend710BA, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site, for the 200 sites against the stand basal area.
-(A) same y-axis across all models; (B) y-axis scaled by model.
-The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
+```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree sum.
+The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
+The same graphs with y-axis scaled by model as availalbe in Figure S", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
+
    knitr::include_graphics(c(here::here("figures", "envTrend7Totba_500.png"),
-                              here::here("figures", "envTrend7Totba.png")))
-```
-
-## Species composition of ingrowth
-
-Most models matched the level of diversity of the observed data quite
-well as quantified by the Shannon index (Figure \@ref(fig:H710)),
-although they consistently underestimated its variability across sites.
-Four models overestimated ingrowth diversity: 4C, PICUS, Landis-II and
-LPJ-GUESS, the latter mostly for the 10 cm diameter threshold. The model
-4C is a special case, as it simulated five species only, i.e., its
-diversity values are not directly comparable to those of the other
-models, nor to the observations. Only one model, SIBYLA, consistently
-underestimated ingrowth diversity.
-
-In most models, there were only small but significant differences in the
-species diversity of ingrowth between 7 and 10 cm across sites (Tables
-S2-S4). Three models (4C, SIBYLA, Landis-II) maintained the ingrowth
-diversity between the 7 and 10 cm thresholds (i.e., the differences
-between them were not significant, see Table S3), and the same was
-evident from the observed data. There were a few models with a strong
-increase of ingrowth diversity between 7 and 10 cm (ForCEEPS, TreeMig,
-LPJ-GUESS, FORMIND, Figure \@ref(fig:H710)).
-
-```{r H710, echo=FALSE, fig.cap="Shannon index of tree ingrowth, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
-The red dashed lines represent the 25th percentile of the observed data for the threshold of 7 cm and the 75th percentile of the observed data for the threshold of 10 cm.", out.width='100%'}
-    knitr::include_graphics(here::here("figures", "H7_10.png"))
-```
-
-For both the observed and the simulated data, and both ingrowth
-thresholds, species diversity was lower in the ingrowth compared to
-stand-level diversity (Figures S3 and S4). Yet, models simulating higher
-diversity at the stand level featured higher diversity in ingrowth
-(Figure S3 and S4). Most models captured reasonably well or
-overpredicted species diversity at the stand level (Figure
-\@ref(fig:combinedRichness)A, center; cf. Figure S5), but a
-characteristic feature was that simulations had a much lower variability
-of diversity compared to observations. Models overpredicting species
-diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure
-\@ref(fig:combinedRichness)A, left). Several models underpredicted
-species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure
-\@ref(fig:combinedRichness)A, right). The majority of the models
-captured reasonably well the ingrowth diversity at lower values of
-diversity, but most of them strongly underestimated ingrowth diversity
-for higher observed diversity levels (Figure
-\@ref(fig:combinedRichness)B, cf. Figure S6). Both the models with
-feedback and those without feedback from the adult trees to ingrowth
-(via seed production) underestimated ingrowth species diversity compared
-to the observations. Again, simulated diversity was considerably less
-variable than observed diversity (Fig. \@ref(fig:combinedRichness)B).
-
-```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site and for observed and simulated values.
-n indicates the number of models falling in each group.
-A) Examples for the three trends across models for species diversity at the stand level.
-B) Examples for the three trends across models for ingrowth at the 7 cm threshold.
-The selection of the representative model for each trend group was based on the full data shown in Figures S5 and S6.", out.width='75%'}
-knitr::include_graphics(here::here("figures", "combined_richness.png"))
+                              here::here("figures", "envTrend7WB_500.png"),
+                              here::here("figures", "envTrend7DDS_500.png")))
 ```
@@
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
-```
-
-## Ingrowth mortality
-
-There were strong differences among the models regarding the mortality
-rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the
-models featured a mortality rate significantly larger than the observed
-data, thus at least partially compensating for the general
-overestimation of ingrowth levels (Figure \@ref(fig:over) and Table S5);
-six models did not feature a significant difference compared to the
-observed data (ForCEEPS, FORMIND, ForClim 1 and 11, and aDGVM2), and
-only the model LPJ-GUESS had a significantly lower mortality rate. Four
-models (xComp, PICUS, LandClim and TreeMig) featured very high mortality
-rates (i.e., well above the Reineke self-thinning line), which
-compensated for the strong initial overestimation of ingrowth (cf.
-Figure 2). Interestingly, two models that underestimated overall
-ingrowth levels (SIBYLA and Landis-II; cf. Figure \@ref(fig:over))
-featured mortality rates that were close to but still above the
-self-thinning line (Figure \@ref(fig:mortality11)).
-
-The fact that some models (and observed data) featured lower mortality
-than expected by Reineke's self-thinning rule may point to facilitation,
-or simply a higher mortality rate before the trees had reached 7 cm
-diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
-whose mortality was well above the self-thinning line, indicates that
-they feature higher ingrowth mortality in mixed-species forests compared
-to even-aged, single-species forests. Some models yielded a mortality
-rate of ingrowth that is broadly compatible with the self-thinning rule
-(e.g., 4C and LandClim).
-
-The three models with the largest overestimation of the proportion of
-ingrowth at the 7 cm threshold featured the highest mortality between
-the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The other
-models that overestimated ingrowth had mortalities lower than the
-expected self-thinning ratio with the exception of LandClim (Figure
-\@ref(fig:mortality13)B). It is noteworthy that most of the models that
-featured a low ratio of ingrowth between 7 and 10 cm (i.e., ForCEEPS,
-ForClim 1, ForClim 11 and iLand) had only a small overestimation of
-ingrowth at the 7 cm threshold (Figure \@ref(fig:mortality11)). aDGVM2
-was the only model that underestimated ingrowth at the 7 cm threshold
-while featuring a low mortality between the two thresholds (Figure
-\@ref(fig:mortality13)C), and only a few models were close to the
-correct ingrowth rate at 7 cm while also being close to the Reineke line
-(SIBYLA, Landis-II; cf. Figure \@ref(fig:mortality13)D; and 4C as well
-as LandClim).
-
-Lastly, some models that had a low number of individuals at 7 cm
-threshold also had a smaller decrease of ingrowth between 7 and 10 cm
-(ratio 7-10 cm) (Figure S10). In these models, competition and
-self-thinning are either not pronounced or must have occurred before the
-species had reached 7 cm. However, this pattern was not consistent
-across models (Figure S10). It is noteworthy that there is a
-relationship between the ingrowth rate at 7 cm and the mortality rate:
-most of the models showed a positive effect (higher ingrowth at 7 cm is
-coupled to an increase of the mortality between 7 and 10 cm). While some
-models showed a negative effect, this linear trend was not pronounced
-and not always significant regarding the slope (Figure S9 and Table S6).
-
-```{r mortality11, echo=FALSE, fig.cap="Ratio of tree ingrowth rates between the 7 and 10 cm thresholds.
-Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree ingrowth between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
-knitr::include_graphics(here::here("figures", "ratio7_10.png"))
-```
-
-```{r mortality13, echo=FALSE, fig.cap="Ratio of tree ingrowth between the 7 and 10 cm thresholds (ingrowth 7 cm / ingrowth 10 cm) and the overestimation proportion at 7 cm ([simulated -observed] / observed) for the mean ingrowth per model across sites and samples.
-The horizontal dashed lines show a ratio equal to 1, indicating no decrease of ingrowth between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
-The vertical line indicates zero overestimation.", out.width='80%'}
-knitr::include_graphics(here::here("figures", "mort7_10_medianOver.png"))
 ```

### 2023-03-12 manuscript/supplementaryRW.Rmd
combine the 3 gradients figure in one figure and add the zoomed ones into the appendix

@@
 
-### Mortality
-
 ```{r mortTable, echo=FALSE,  message=FALSE, fig.pos='H'}
@@
 
+```{r plotComp, echo=FALSE, warning=FALSE, fig.cap="Plot and summary statistics of the linear model between the overestimation proportion in ingrowth levels and the mean complexity value of the establisment modules", fig.show = 'hold', out.width = '70%', fig.pos='H'}
+
+knitr::include_graphics(here::here("figures", "compOver.png"))
+
+```
+
+```{r rtreesDiff, echo=FALSE,  message=FALSE, fig.pos='H'}
+library(knitr)
+
+rtreesDiff <- readr::read_csv(here::here("figures", "rtreesDiff.csv"))
+
+kable(rtreesDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold of the observed values and the simulated values per model.") |>
+    kableExtra::kable_styling(latex_options = "HOLD_position") |>
+    kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+```
+
+```{r sigH7_SimObs, echo=FALSE,  message=FALSE, fig.pos='H'}
+library(knitr)
+
+rtreesDiff <- readr::read_csv(here::here("figures", "sigH7_SimObs.csv"))
+
+kable(rtreesDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold of the observed values and the simulated values per model.") |>
+    kableExtra::kable_styling(latex_options = "HOLD_position") |>
+    kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+```
+
 \newpage
-
-### Model complexity VS overestimation
-
-```{r plotComp, echo=FALSE, warning=FALSE, fig.cap="Plot and summary statistics of the linear model between the overestimation proportion in ingrowth levels and the mean complexity value of the establisment modules", fig.show = 'hold', out.width = '70%', fig.pos='H'}
-
-knitr::include_graphics(here::here("figures", "compOver.png"))
-
-```
-
 ## Total ingrowth and individual species regeneration niches
 
-```{r envTrend7wb, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site, for the 200 sites against the sites water balance.
-(A) same y-axis across all models; (B) y-axis scaled by model.
-The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%', fig.pos='H'}
-    knitr::include_graphics(c(here::here("figures", "envTrend7WB_500.png"),
-                              here::here("figures", "envTrend7WB.png")))
-```
-
-```{r envTrend7dds, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site, for the 200 sites against the sites seasonal degree-day sum.
-(A) same y-axis across all models; (B) y-axis scaled by model.
-The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%', fig.pos='H'}
-    knitr::include_graphics(c(here::here("figures", "envTrend7DDS_500.png"),
+```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree sum.
+The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
+The same graphs with y-axis scaled by model as availalbe in Figure S", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
+     knitr::include_graphics(c(here::here("figures", "envTrend7Totba.png"),
+                              here::here("figures", "envTrend7WB.png"),
                               here::here("figures", "envTrend7DDS.png")))

### 2023-03-12 code/figures.R
correct the dots size in the HeatEnvGrad figure

@@ -216,7 +216,7 @@
                     dpi = 300, units = "cm", device = 'png')
     
     
-    # Significant difference per model?
+    # Significant difference per model between 7 and 10 diameter threshold
     
     sigH7_10 <- ggpubr::compare_means(ShannonIndexRecruit ~ dbh, 
                                       data = dat710[!dat710$model == "aDGVM2", ],
@@ -617,144 +617,6 @@
                     plot = d, width = 22, height = 18, scale = 0.9,
                     dpi = 300, units = "cm", device = 'png')
     
-    
-    ### Fig. r.ba/Totr.ba & env gradient -----
-    
-    # Data preparation
-  
-  for (dbhSel in c(7, 10)) {
-    dbhSel <- dbhSel
-    simResdbh <- outputsDF |> dplyr::filter(dbh %in% dbhSel)
-    #simResdbh <- outputsDF 
-    simResdbh$r.baShare <- simResdbh$r.ba / simResdbh$Totba
-    simResdbh$baShare <- simResdbh$ba / simResdbh$Totba
-    
-    simResdbh$r.baTotRecShare <- round(simResdbh$r.ba, 4) / round(simResdbh$Totr.ba, 4)
-    
-    # For this figure we consider that values with 0 r.ba and 0 Tot.r.ba have a 
-    # recruitment share of 0 instead of NaN
-    simResdbh$r.baTotRecShare [simResdbh$r.ba == 0 & simResdbh$Totr.ba == 0] <- 0
-    
-    # Aggregate values from simulations to get total recruitment per site and sample
-    simResAgg <- simResdbh |>
-      dplyr::group_by(site, species, model) |> 
-      dplyr::summarise(r.baShareMean = round(mean(r.baShare), 2), 
-                       baShareMean = round(mean(baShare), 2), 
-                       r.baTot.r.ShareMean = round(mean(r.baTotRecShare),
-                                                   2), 
-                       dds = unique(dds), 
-                       wb = unique(wb)) 
-    
-    mainSppBAshare <-  simResAgg[simResAgg$species %in% c("Abies alba", 
-                                                          "Fagus sylvatica",
-                                                          "Picea abies", 
-                                                          "Pinus sylvestris",
-                                                          "Quercus spp."), ]
-    mainSppBAshare$model <- factor(mainSppBAshare$model, levels = modelsOrder)
-    mainSppBAshare$dds <- round(mainSppBAshare$dds, 0)
-    
-    #Divide dds and wb in bins based on all the possible values for all sites
[…]

### 2023-03-12 manuscript/manuscriptRW.Rmd
added proper PROCLIAS funding text in the manuscript

@@
 the models reproduced the effect of basal area in magnitude and patterns
-(Figure \@ref(fig:nvTrend710BA)) better than the effects of the climatic
-gradients, where they featured varying effects (Figures S1 and S2).
+(Figure \@ref(fig:nvTrend710BAlightTemp) and S7) better than the effects
+of the climatic gradients, where they featured varying effects.
 
 In the observed data, ingrowth levels decreased clearly with increasing
-stand basal area (Figure \@ref(fig:nvTrend710BA)). This trend was
-captured by the Landis-II model only, albeit at much lower values of
+stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)). This trend
+was captured by the Landis-II model only, albeit at much lower values of
 basal area than in the observations. The other models featured
@@
 
-The models did not match the observed ingrowth patterns across gradients
-of soil moisture and temperature (Figures S1 and S2), and neither was
+The observed ingrowth did not change much across the climatic water
+balance gradient, and showed a slight increase with temperature. The
+models did not match the observed ingrowth patterns across gradients of
+soil moisture and temperature (Figure
+\@ref(fig:)\@ref(fig:nvTrend710BAlightTemp) and S7), and neither was
 there a clear pattern across models. Regarding the soil moisture
-gradient (Figure S1), a group of models featured decreasing ingrowth
-with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp,
-PICUS, LandClim, TreeMig and iLand), while a few models showed the
-opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS). Across the
-temperature gradient (Figure S2), all the models that featured a
-decrease in ingrowth levels with increasing soil moisture showed the
-opposite trend (i.e., more ingrowth with increasing temperature), except
-for 4C. The other models did not feature a clear pattern.
-
-```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree sum.
+gradient, a group of models featured decreasing ingrowth with increasing
+soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim,
+TreeMig and iLand), while a few models showed the opposite trend
+(ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient,
+all the models that featured a decrease in ingrowth levels with
+increasing soil moisture showed the opposite trend (i.e., more ingrowth
+with increasing temperature), except for 4C. The other models did not
+feature a clear pattern.
+
+```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site with y-axis scaled by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree sum.
 The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
-The same graphs with y-axis scaled by model as availalbe in Figure S", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
-
-   knitr::include_graphics(c(here::here("figures", "envTrend7Totba_500.png"),
-                              here::here("figures", "envTrend7WB_500.png"),
-                              here::here("figures", "envTrend7DDS_500.png")))
+For comparing the magnitude of recruitment across models the same graphs with equal y-axis are availalbe in Figure S7", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
+
+   knitr::include_graphics(c(here::here("figures", "envTrend7Totba.png"),
+                              here::here("figures", "envTrend7WB.png"),
+                              here::here("figures", "envTrend7DDS.png")))
 ```
@@
 observed data along the two major gradients of climate (Figure
-\@ref(fig:diff), cf. Figure S7). However, the models differed
+\@ref(fig:diff), cf. Figure S8). However, the models differed
 considerably in the way in which they simulated their climatic niche.
@@
 
-Funding of the workshop by COST (Action No CA19139) and ETH Zurich
+Funding of the workshop by COST Action CA19139 PROCLIAS (PROcess-based
+models for CLimate Impact Attribution across Sectors), supported by COST
+(European Cooperation in Science and
+Technology; [https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich
 (Grant ETH-35 18-1).

### 2023-03-12 manuscript/manuscriptRW.Rmd
re-tructure of discussion following the new flow of ideas

@@
 compensated by some models with high simulated mortality rates in the
-early stages of population dynamics. The effect of light availability on
-ingrowth rates is captured better than the effect of temperature and
-soil moisture, but patterns are not consistent across models. Simulated
-species diversity of ingrowth matches the observed ranges quite well.
-Models simulating higher species diversity at the stand level do not
-feature higher ingrowth diversity. The ingrowth of *Abies alba* is
-captured across temperature and soil moisture gradients, whereas
+early stages of population dynamics. Simulated species diversity of
+ingrowth matches the observed ranges quite well. Models simulating
+higher species diversity at the stand level do not feature higher
+ingrowth diversity. The effect of light availability on ingrowth rates
+is captured better than the effect of temperature and soil moisture, but
+patterns are not consistent across models. The ingrowth of *Abies alba*
+is captured across temperature and soil moisture gradients, whereas
 ingrowth of *Fagus sylvatica* is generally underestimated. *Picea abies*
@@
 
-More specifically, we aimed to answer the following questions:
-
-1)  Are models of forest dynamics capturing accurately tree ingrowth
-    levels, initial tree diversity and mortality in tree establishment?
-
-2)  Do model traits explain differences in model performance?
-
-3)  How well do models capture total ingrowth and individual species
-    regeneration niches across environmental gradients of light
-    availability, temperature, and soil moisture?
+More specifically, we aimed to answer the following questions: (1) Are
+models of forest dynamics capturing accurately tree ingrowth levels,
+initial tree diversity and mortality in tree establishment? (2)Do model
+traits explain differences in model performance? (3) How well do models
+capture total ingrowth and individual species regeneration niches across
+environmental gradients of light availability, temperature, and soil
+moisture?
 
@@
 ingrowth thresholds, although the exact changes in the share of the
-ingrowth basal area were different (Figure S7).
+ingrowth basal area were different (Figure S8).
 
@@
 alba*) was captured well across the gradient by most models except for
-xComp, PICUS, LandClim and Landis-II, which overestimated the ingrowth
-share of this species. There were also some peculiarities evident for
-some models, such as 4C that did not simulate *Abies alba*. The patterns
+xComp, PICUS, and LandClim, which overestimated the ingrowth share of
+this species. There were also some peculiarities evident for some
+models, such as 4C that did not simulate *Abies alba*. The patterns
 across models for *Fagus sylvatica* ingrowth were more complex, the
@@
 regarding the species-specific ingrowth levels in the environmental
-space of Figure 6. Some models (e.g., ForClim11, PICUS) tended to
-consistently overestimate the ingrowth share of all five species, but
-most models overestimated the ingrowth share of some species while
-underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
+space of Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
+tended to consistently overestimate the ingrowth share of all five
+species, but most models overestimated the ingrowth share of some
+species while underestimating the share of others (e.g., ForCEEPS,
+LPJ-GUESS).
 
@@
 
-## Ingrowth levels
+## Most of the models overestimated tree ingrowth rates
 
@@
 
-## Model formulation did not affect model performance
-
-ADD THESE POINTS HERE complexity, empirical v. PB, etc.
+Ingrowth levels define the structure and composition of future forests,
+but it is equally important to correctly identify areas where ingrowth
+is lacking [@rammer2021]. There are multiple constraints to the
+regeneration niche of tree species [@price2001], and therefore the
+absence of ingrowth is likely to be common [@fortin2007], even over
+larger areas such as the one-hectare plots used here. Tree ingrowth data
+are characterized by zero-inflation. This was clearly evident from the
+EuFoRIa dataset, but some of the models did not produce zeros at all, or
+featured a very low proportion of zero data. This substantial difference
+may be due to the fact that the simulation results were drawn from
+equilibrium forests, whereas in reality many of the forest reserves are
+recovering from past management activities and have become denser over
+the past decades [e.g., @heiri2009], leading to less ingrowth than in an
+equilibrium situation.
+
+Correctly capturing species composition in tree establishment is
+important to assess the future functional diversity of a forest, e.g.,
+its sensitivity to drought or resilience to disturbances [@redmond2015;
+@seidl2022]. In the simulations, overall ingrowth species diversity
+levels were well within the observed range for two thirds of the models
+(10 out of 15). Thus, while most of the models are facing difficulties
+to quantitatively match ingrowth levels (cf. above), their performance
+is much better when it comes to matching the composition of ingrowth as
+a function of abiotic and biotic conditions. Most models maintained or
+even increased ingrowth diversity between the 7 and 10 cm thresholds,
+and the same is visible from the empirical data, but the differences are
+not significant. Diversity variations may be explained by the way the
+individual models consider establishment processes [@konig_tree_2022],
+e.g., in the absence of adults, ingrowth can be linked to either seed
+rain-dispersal, or to a seed/seedling bank.
+
+The species diversity of the entire stand was better captured by the
+models than the diversity of tree ingrowth, and only a few models
+overpredicted stand-level diversity. Defining ingrowth composition as
+being proportional to adult abundance, regardless of productivity, might
+be a simple and conservative assumption to safely maintain relative
+species abundances [@hanbury-brown2022], although this may be
+simplistic. Furthermore, based on the simulation results, there is no
+evidence that models with feedback captured better the ingrowth species
+diversity than those without feedback. The similar performance of models
+with and without feedback is likely because the models put more weight
+to the regeneration niche arising from abiotic and biotic filters, than
+from the habitat niche of the adult trees [@grubb1977].
+
+There are several factors that lead to mortality during the
+establishment phase such as competition [@casper1997] and multiple
+abiotic factors [@schmid2021; @cunningham2006]. As mentioned above, it
+is reasonable to expect that models that overestimate ingrowth of small
+trees will already have a particularly high mortality between the 7 and
+10 cm thresholds. However, this was not consistently evident from the
+simulations. This implies that these models must have the excess of
+mortality in earlier or later stages, if we assume that all models would
+be able to capture the structure and composition of adult stands along
+the EuFoRIa gradient, which however was not directly tested here.
+
+Higher mortality towards the adult phase has important implications for
+forest dynamics and the goods and services provided by forests. On the
+one hand, mortality in later stages may erroneously enhance the share of
+less shade tolerant species in the models [@klopcic2015], if available,
+therefore shifting the species composition. Unrealistic high stem
+densities for a longer period of time may overestimate the role of tree
+regrowth in carbon sink dynamics [@pugh2019]. Given our set of
+simulations and analyses, we cannot conclusively assess what is
+happening in the models, and further studies are required that focus on
+a wider range of tree sizes and the fate of tree establishment along
+such a size continuum.
+
+## Model traits did not affect model performance
 
@@
 
-Ingrowth levels define the structure and composition of future forests,
-but it is equally important to correctly identify areas where ingrowth
-is lacking [@rammer2021]. There are multiple constraints to the
-regeneration niche of tree species [@price2001], and therefore the
-absence of ingrowth is likely to be common [@fortin2007], even over
-larger areas such as the one-hectare plots used here. Tree ingrowth data
-are characterized by zero-inflation. This was clearly evident from the
-EuFoRIa dataset, but some of the models did not produce zeros at all, or
-featured a very low proportion of zero data. This substantial difference
-may be due to the fact that the simulation results were drawn from
-equilibrium forests, whereas in reality many of the forest reserves are
-recovering from past management activities and have become denser over
-the past decades [e.g., @heiri2009], leading to less ingrowth than in an
-equilibrium situation.
+Increasing complexity in establishment modules has been motivated by
+better process understanding and enhancing of model accuracy
+[@bugmann2022; @bugmann2019]. However, more complex models do not
+necessarily lead to better projections and rather to a reduced
+transparency and predictive power [@franklin2020]. The results from this
+study showed that increasing complexity in the establishment modules did
+not represent an improvement in the projections of ingrowth levels,
+species composition or mortality at early tree stages. Thus, the
+question what level of detail one has to use to model establishment
+processes remains open [@konig_tree_2022; @bugmann2022].
+
+## Regeneration niche
 
@@
 
-## Species composition of ingrowth
-
-Correctly capturing species composition in tree establishment is
-important to assess the future functional diversity of a forest, e.g.,
-its sensitivity to drought or resilience to disturbances [@redmond2015;
-@seidl2022]. In the simulations, overall ingrowth species diversity
-levels were well within the observed range for two thirds of the models
-(10 out of 15). Thus, while most of the models are facing difficulties
-to quantitatively match ingrowth levels (cf. above), their performance
-is much better when it comes to matching the composition of ingrowth as
-a function of abiotic and biotic conditions. Most models maintained or
-even increased ingrowth diversity between the 7 and 10 cm thresholds,
-and the same is visible from the empirical data, but the differences are
-not significant. Diversity variations may be explained by the way the
-individual models consider establishment processes [@konig_tree_2022],
-e.g., in the absence of adults, ingrowth can be linked to either seed
-rain-dispersal, or to a seed/seedling bank.
-
-Yet, there were pronounced variations how the main tree species were
+There were pronounced variations how the main tree species were
 represented along the environmental gradients, in particular the
@@
 the absence of robust datasets.
-
-The species diversity of the entire stand was better captured by the
-models than the diversity of tree ingrowth, and only a few models
-overpredicted stand-level diversity. Defining ingrowth composition as
-being proportional to adult abundance, regardless of productivity, might
-be a simple and conservative assumption to safely maintain relative
-species abundances [@hanbury-brown2022], although this may be
-simplistic. Furthermore, based on the simulation results, there is no
-evidence that models with feedback captured better the ingrowth species
-diversity than those without feedback. The similar performance of models
-with and without feedback is likely because the models put more weight
-to the regeneration niche arising from abiotic and biotic filters, than
-from the habitat niche of the adult trees [@grubb1977].
-
-## Ingrowth mortality
-
-There are several factors that lead to mortality during the
-establishment phase such as competition [@casper1997] and multiple
-abiotic factors [@schmid2021; @cunningham2006]. As mentioned above, it
-is reasonable to expect that models that overestimate ingrowth of small
-trees will already have a particularly high mortality between the 7 and
-10 cm thresholds. However, this was not consistently evident from the
-simulations. This implies that these models must have the excess of
-mortality in earlier or later stages, if we assume that all models would
-be able to capture the structure and composition of adult stands along
-the EuFoRIa gradient, which however was not directly tested here.
-
-Higher mortality towards the adult phase has important implications for
-forest dynamics and the goods and services provided by forests. On the
-one hand, mortality in later stages may erroneously enhance the share of
-less shade tolerant species in the models [@klopcic2015], if available,
-therefore shifting the species composition. Unrealistic high stem
-densities for a longer period of time may overestimate the role of tree
-regrowth in carbon sink dynamics [@pugh2019]. Given our set of
-simulations and analyses, we cannot conclusively assess what is
-happening in the models, and further studies are required that focus on
-a wider range of tree sizes and the fate of tree establishment along
-such a size continuum.
 
@@
 Based on the results of this study, it is clear that models of forest
-dynamics need a focus on their establishment modules. Increasing
-complexity in establishment modules has been motivated by better process
-understanding and enhancing of model accuracy [@bugmann2022;
-@bugmann2019]. However, more complex models do not necessarily lead to
-better projections and rather to a reduced transparency and predictive
-power [@franklin2020]. The results from this study showed that
-increasing complexity in the establishment modules did not represent an
-improvement in the projections of ingrowth levels, species composition
-or mortality at early tree stages. Thus, the question what level of
-detail one has to use to model establishment processes remains open and
-must be addressed in future research [@konig_tree_2022; @bugmann2022].
-
-We recommend that the further development of the establishment modules
-is implemented as additional features that can be traced back, as done
-in this study for the model variants of ForClim and ForCEEPS, and that
-model complexity and structure must always be connected with model
-objectives [@albrich2020]. If it should be necessary to include more
-detail in the establishment models, this will come with higher
+dynamics need a focus on their establishment modules. As shown in the
+results of the study it is still uncertain what level of detail one has
+to use to model establishment processes and must be addressed in future
+research. We recommend that the further development of the establishment
+modules is implemented as additional features that can be traced back,
+as done in this study for the model variants of ForClim and ForCEEPS,
+and that model complexity and structure must always be connected with
+model objectives [@albrich2020]. If it should be necessary to include
+more detail in the establishment models, this will come with higher
 parameterisation efforts, which will most likely lead to lower
@@
 
-# Data availability statement
-
-The code and data that support the findings of this study are openly
-available at zenodo and gitlab (links to be added)
+# Reproducibility and data availability statement
+
+This study compiles with the basic rules of reproducibility and has make
+available the protocol of the study, the data and all the software that
+leads of the findings presented in this study. The protocol is available
+here XXX. The code and data that support the findings of this study are
+openly available at zenodo and gihub (links to be added)
 
@@
 
-Funding of the workshop by COST Action CA19139 PROCLIAS (PROcess-based
-models for CLimate Impact Attribution across Sectors), supported by COST
-(European Cooperation in Science and
+Funding of the workshop was provided by the COST Action CA19139 PROCLIAS
+(PROcess-based models for CLimate Impact Attribution across Sectors),
+supported by COST (European Cooperation in Science and
 Technology; [https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich
-(Grant ETH-35 18-1).
+(Grant ETH-35 18-1). ODY was funded by SNF project XXXX
 

### 2023-03-12 manuscript/manuscriptRW.Rmd
changes in lats paragraph abstract

@@
 global levels, we simulate tree establishment at 200 sites representing
-environmental gradients across Europe. The results are benchmarked
-against comprehensive data from unmanaged forests.
+environmental gradients across Europe. The results are evaluated against
+comprehensive data from unmanaged forests.
 
@@
 
-Even though the overestimation of tree establishment represents a
-challenge for the usefulness of model projections under uncertain
-conditions such as climate change, and although some formulations should
-be reconsidered to correct excessive ingrowth levels, it is remarkable
-that most of the models capture the essential features of the highly
-complex process of tree establishment, while not having been
-parameterized with such data. We conclude that in spite of all the
-uncertainties, a lot can be gained by a focus on the modeling of
+In spite of more or less substantial mismatches between model results
+and data, it is remarkable that most of the models capture the essential
+features of the highly complex process of tree establishment, while not
+having been parameterized with such data. We conclude that in spite of
+all the uncertainties, a lot can be gained by a focus on the modeling of
 establishment processes in these model, this will make their long-term

### 2023-03-12 code/figures.R

@@ -1473,11 +1473,7 @@
       
     }
     
-    
-    
-    
-    
-    
+   
       
 # Map -----
     # IMPORTANT: This is the only figure that is not reproducible

### 2023-03-12 manuscript/manuscriptRW.Rmd
change deffinition for forclim 11

@@
 models from empirically derived to "process-based" [e.g., @bugmann1996;
-@fabrika2012; @mette2014; @lexer2001]. Because of the large variability
+@fabrika2012; @lexer2001; @reyer2014]. Because of the large variability
 in tree establishment patterns in nature and the large number of factors
@@
 module that adheres closely to the concept introduced by @botkin1972,
-whereas ForClim variant 11 is adopting a rather different approach that
-considers space filling along with different 'points of entry' for the
-action of specific environmental factors [@huber2020]. These two model
-variants allow us to evaluate a more process-based module (variant 11)
-against a simple module (variant 1), while the rest of the model
-structure is identical. Similarly, the two ForCEEPS variants allow us to
-isolate the importance of the canopy feedback (i.e. simulated actual
-composition and relative abundance of species in the plot) via seed
-trees for the quantity and quality of simulated ingrowth, as one variant
-includes this feedback whereas the other does not.
+whereas ForClim variant 11 is adopting a different approach where
+individual species properties and its relation to the environment are
+more important [@huber2020]. These two model variants allow us to
+evaluate a more process-based and complex module (variant 11) against a
+simple module (variant 1), while the rest of the model structure is
+identical. Similarly, the two ForCEEPS variants allow us to isolate the
+importance of the canopy feedback (i.e. simulated actual composition and
+relative abundance of species in the plot) via seed trees for the
+quantity and quality of simulated ingrowth, as one variant includes this
+feedback whereas the other does not.
 

### 2023-03-12 manuscript/manuscriptRW.Rmd
redefine thresholds in the observed data

@@
 establishment modules (mean establishment formulation complexity
-complexity across all processes \>0, Table
-\@ref(tab:)\@ref(tab:tableModels)).
+complexity across all processes \>0, Table \@ref(tab:tableModels)).
 
@@
 balance, soil quality and exposition (i.e., slope and aspect). Each of
-these sites featured at least two consecutive measurements. Ingrowth was
-measured at threshold of 7 and 10 cm of stem diameter at breast height
-in all sites except for 35 sites where only ingrowth at the 10 cm
-diameter threshold was available. Plot size ranged from 0.02 ha to 5.52
-ha, and the data were further processed and aggregated following
-@käber2022 to provide ingrowth rates per ha and per decade. The
-observations used in this study featured 30,900 established trees, and
-ingrowth rates per site, sample and decade ranged from zero to 1246
-trees per ha, with a mean of 56 trees per ha. For more details on this
-unique dataset and the detailed establishment information, cf.
-@käber2022.
+these sites featured at least two consecutive measurements. Ingrowth
+thresholds for the sites differed between 0 and 10 cm, we defined two
+data sets where one included all sites with a dbh threshold of 7cm or
+lower and the other included another 35 sites with dbh thresholds
+between 7 and 10 cm. Thus we focused on two data sets with the
+harmonized dbh thresholds of 7 cm and 10 cm.
+
+Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
+processed and aggregated following @käber2022 to provide ingrowth rates
+per ha and per decade. The observations used in this study featured
+30,900 established trees, and ingrowth rates per site, sample and
+decade ranged from zero to 1246 trees per ha, with a mean of 56 trees
+per ha. Adult species composition was also available for the simulated
+species. For more details on this unique dataset and the detailed
+establishment information, cf. @käber2022.
 

### 2023-03-13 manuscript/manuscriptRW.Rmd
I remove the right answer wrong reason sentend in discussion

@@
 are hard to parameterize [e.g., @lischke2006]. A related issue is the
-excessive reduction of species richness due to positive feedback effects
-such that eventually just single-species stands remain. This is
+excessive reduction of species diversity due to positive feedback
+effects such that eventually just single-species stands remain. This is
 sometimes corrected by the incorporation of a low level of seed influx
@@
 
-The overarching goal of the experiments was to assess the regeneration
-niche as it arises from empirical data against its representation in a
-wide range of models of forest dynamics. We define the regeneration
-niche as the passing of a breast height diameter threshold of 7 or 10
-cm, respectively (synonym: ingrowth). To this end, each modeling group
-was provided with a detailed protocol (cf. Supplementary Material 1,
-henceforth "SM") with instructions how to perform the simulations,
-providing input variables on climate and soil conditions, and the
-expected output variables. Neither were further site information (except
-for the data specified below) nor any data on tree ingrowth provided
-prior to the simulation. That is, the models were run in "blind flight"
-mode.
+The overarching goal of the experiments was to assess the tree
+establishment as it arises from empirical data against its
+representation in a wide range of models of forest dynamics. We define
+the tree establishment as the passing of a breast height diameter
+threshold of 7 or 10 cm, respectively (synonym: ingrowth). To this end,
+each modeling group was provided with a detailed protocol (cf.
+Supplementary Material 1, henceforth "SM") with instructions how to
+perform the simulations, providing input variables on climate and soil
+conditions, and the expected output variables. Neither were further site
+information (except for the data specified below) nor any data on tree
+ingrowth provided prior to the simulation. That is, the models were run
+in "blind flight" mode.
 
@@
 to compensating for a first error (excess ingrowth) by a second "error"
-(excess mortality). Thus, many models appear to get the 'right answer',
-i.e., correct stand structure and composition, for the wrong reasons. It
-is highly likely that biased projections will result, because the two
-errors are unlikely to be perfectly linked and thus will not always
-compensate each other. Hence, this structural problem of most models
-investigated here is problematic particularly if the models are to be
-used under novel conditions such as under climate change [e.g.,
-@huber2021] or in a decision support context [e.g., @thrippleton2021].
+(excess mortality). It is highly likely that biased projections will
+result, because the two errors are unlikely to be perfectly linked and
+thus will not always compensate each other. Hence, this structural
+problem of most models investigated here is problematic particularly if
+the models are to be used under novel conditions such as under climate
+change [e.g., @huber2021] or in a decision support context [e.g.,
+@thrippleton2021].
 
@@
 size was different at each of the 200 sites and it would have introduce
-additional uncertainty into the results.
+additional uncertainty into the results. We can not forget here that the
+observed data is just a snapshot of an extremely complex and stochastic
+reality, which might be the explanation of e.g. lack of clear patterns
+when confronted with gradients of climater water balance and degree
+days, while the models outputs are the realizations of the definitions
+of the regeneration niche.
 

### 2023-03-13 manuscript/manuscriptRW.Rmd
substitude rates per levles with ingrowth and added a new paragraph in research recomendations

@@
 
-Most of the models overestimate ingrowth rates, which is only
+Most of the models overestimate ingrowth levels, which is only
 compensated by some models with high simulated mortality rates in the
@@
 higher species diversity at the stand level do not feature higher
-ingrowth diversity. The effect of light availability on ingrowth rates
+ingrowth diversity. The effect of light availability on ingrowth levels
 is captured better than the effect of temperature and soil moisture, but
@@
 \@ref(fig:mortality13)C), and only a few models were close to the
-correct ingrowth rate at 7 cm while also being close to the Reineke line
-(SIBYLA, Landis-II; cf. Figure \@ref(fig:mortality13)D; and 4C as well
-as LandClim).
+correct ingrowth levels at 7 cm while also being close to the Reineke
+line (SIBYLA, Landis-II; cf. Figure \@ref(fig:mortality13)D; and 4C as
+well as LandClim).
 
@@
 across models (Figure S5). It is noteworthy that there is a relationship
-between the ingrowth rate at 7 cm and the mortality rate (Table S4):
+between the ingrowth levels at 7 cm and the mortality rate (Table S4):
 most of the models showed a positive effect (higher ingrowth at 7 cm is
@@
 
-## Most of the models overestimated tree ingrowth rates
-
-Most of the models overestimated tree ingrowth rates. This has
+## Most of the models overestimated tree ingrowth levels
+
+Most of the models overestimated tree ingrowth levels. This has
 potentially far-reaching implications, e.g., regarding biomass (and thus
@@
 classes. Trees at this size contribute little to carbon sequestration,
-and if the excess ingrowth rate at this stage is compensated -- in the
-models -- soon thereafter by higher mortality, overall simulated
-ingrowth may still yield trustworthy results at the level of stand
-structure and related forest attributes such as basal area or biomass.
+and if the excess ingrowth at this stage is compensated -- in the models
+-- soon thereafter by higher mortality, overall simulated ingrowth may
+still yield trustworthy results at the level of stand structure and
+related forest attributes such as basal area or biomass.
 
@@
 
-Increasing complexity in establishment modules has been motivated by
-better process understanding and enhancing of model accuracy
-[@bugmann2022; @bugmann2019]. However, more complex models do not
-necessarily lead to better projections and rather to a reduced
-transparency and predictive power [@franklin2020]. The results from this
-study showed that increasing complexity in the establishment modules did
-not represent an improvement in the projections of ingrowth levels,
-species composition or mortality at early tree stages. Thus, the
-question what level of detail one has to use to model establishment
-processes remains open [@konig_tree_2022; @bugmann2022].
+The results from this study showed that increasing complexity in the
+establishment modules did not represent an improvement in the
+projections of ingrowth levels, species composition or mortality at
+early tree stages. Increasing complexity in establishment modules has
+been motivated by better process understanding and enhancing of model
+accuracy [@bugmann2022; @bugmann2019]. However, more complex models do
+not necessarily lead to better projections and rather to a reduced
+transparency and predictive power [@franklin2020]. Thus, the question
+what level of detail one has to use to model establishment processes
+remains open.[@konig_tree_2022; @bugmann2022].
 
@@
 the extremes of the stand density ranges in some models. For example,
-ingrowth rates at low stand densities are relevant to assess how well
+ingrowth levels at low stand densities are relevant to assess how well
 forests are recovering e.g., after gap creation due to disturbance
@@
 indicating that the relation between ingrowth and light availability is
-not yet captured correctly in most models.
+not yet captured correctly in most models. On the other hand several of
+the models with an increase of ingrowth with increasing basal area also
+include the feedback between seed production of mature trees and
+ingrowth. Thus it seems that the light competition does not sufficiently
+compensate the increased seed availability with higher basal area.
 
@@
 dominant species Abies alba, Fagus sylvatica and Picea abies. The
-ingrowth rates were sampled from simulations in the equilibrium, and in
+ingrowth levels were sampled from simulations in the equilibrium, and in
 this state it is expected that non-climax species such as *Pinus
@@
 reasonable than it might appear at first sight. However, the equilibrium
-assumption may explain why simulated ingrowth rates were higher than the
-observed ones. Second, the data were collected for ingrowth above a 7 cm
-threshold, thus limiting the assessment of tree establishment to a
+assumption may explain why simulated ingrowth levels were higher than
+the observed ones. Second, the data were collected for ingrowth above a
+7 cm threshold, thus limiting the assessment of tree establishment to a
 holistic viewpoint. This constituted a hard limit based on which we can
@@
 large environmental gradients. Third, the empirical data were collected
-from rather small plots, while we sampled simulated ingrowth rates from
+from rather small plots, while we sampled simulated ingrowth levels from
 1 ha areas, which may lead to an incorrect representation of space. Even
@@
 reality, which might be the explanation of e.g. lack of clear patterns
-when confronted with gradients of climater water balance and degree
-days, while the models outputs are the realizations of the definitions
-of the regeneration niche.
+when confronted with gradients of climate water balance and degree days,
+while the models outputs are the realizations of the definitions of the
+regeneration niche.
 
@@
 
+Exercises like the one presented here where the models are run in "blind
+flight" without the possibility of being tuned towards capturing the
+observed values should be repeated. Such a benchmarking exercises should
+next focus in different aspects such as specific model traits and
+ecological formulations to better understand the implications of the
+model assumptions. Furthermore, the inclusion in this exercises with
+different scales, type, populations structures, tree establishment
+modules and complexity ensure a wider benefit to the whole modeling
+community.
+
 ## Conclusions
@@
 are facing similar challenges when representing tree establishment. They
-generally overestimate tree establishment rates, and the simulated
+generally overestimate tree establishment levels, and the simulated
 establishment niche is not always captured accurately as a function of

### 2023-03-13 manuscript/supplementaryRW.Rmd
added the env gradient with same y axis

@@
     latex_engine: xelatex
-    extra_dependencies: ["flafter", "float"]
+    extra_dependencies: ["flafter", "float", "subfig"]
     toc: yes
@@
 
-```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree sum.
-The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
-The same graphs with y-axis scaled by model as availalbe in Figure S", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
-     knitr::include_graphics(c(here::here("figures", "envTrend7Totba.png"),
-                              here::here("figures", "envTrend7WB.png"),
-                              here::here("figures", "envTrend7DDS.png")))
+```{r nvTrend710, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree sum.
+The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
+   knitr::include_graphics(c(here::here("figures", "envTrend7Totba_500.png"),
+                              here::here("figures", "envTrend7WB_500.png"),
+                              here::here("figures", "envTrend7DDS_500.png")))
 ```

### 2023-03-13 code/figures.R
added text on the figure of overprediction

@@ -744,11 +744,14 @@
       ggplot2::annotate("text", x = 10, y = 2.6, label = "A") + 
       ggplot2::annotate("text", x = 10, y = 1.6, label = "B") + 
       ggplot2::annotate("text", x = -4.6, y = 1.6, label = "C") + 
-      ggplot2::annotate("text", x = -4.6, y = 2.6, label = "D") 
+      ggplot2::annotate("text", x = -4.6, y = 2.6, label = "D") + 
+    
+      ggplot2::annotate("text", x = 8.5, y = 1.81, label = "Reineke self-thinning") + 
+      ggplot2::annotate("text", x = 7.9, y = 1.03, label = "Ratio equal to 1")
     
     ggplot2::ggsave("figures/mort7_10_medianOver.png",
                     plot =  overMortmedian,
-                    width = 16, height = 12, scale = 0.9,
+                    width = 18, height = 12, scale = 0.9,
                     dpi = 300, units = "cm", device = 'png')  
     
     

### 2023-03-13 code/plottingManuscript.R
change landsi name to not capitals in the plotting parameters

@@ -58,7 +58,7 @@
                 "xcomp" = "xComp",
                 "iLand" = "iLand", 
                 "LandClim" = "LandClim",  
-                "Landis II" = "LANDIS-II",  
+                "Landis II" = "Landis-II",  
                 "Treemig" = "TreeMig",
                 "aDGVM2" = "aDGVM2", 
                 "LPJ-GUESS" = "LPJ-GUESS")

### 2023-03-13 manuscript/supplementaryRW.Rmd
remove figure captions from xcomp

@@
 
-Tab. 1: Species calibration parameters for site 434. Spec = species
-code, rke = Reineke Exponent, Sdi = stand density index, ogPer =
-(maximum) old growth time span (years), regnLgh = regeneration light
-demand (canopy shadow value of xcomp), plot_id = site (434), tjja =
-summer temperature (Jun-Aug, °C), sitePred = modelled species prevalence
-for site (0...1 = min...max), dd = upper diameter growth (cm/ 5 years,
-maximum dbh growth up to 85 % higher), ogDbh = old growth dbh (cm, dbh
-at which old growth phase begins), regnPer = (minimum) regeneration time
-span (year, until seedling reaches 5 cm dbh), regeneration mortality
-(climate dependent) Spec Rke Sdi ogPer regnLgh Plot_id Tjja sitePred Dd
-ogDbh regnPer regnMort Abal -1.6 1000 350 4 434 0.81 2.8 32.4 17 0 Acps
--1.6 800 150 3 434 0.86 1.8 24.6 17 0 Betu -1.6 700 100 2 434 0.76 1.7
-12.8 17 0.001 Cabe -1.6 700 125 3 434 0.76 1.4 16.7 17 0.001 Fasy -1.6
-800 150 4 434 0.93 1.9 24.6 17 0 Piab -1.6 1000 350 3 434 0.87 2.9 32.4
-17 0 Pisy -1.6 800 250 2 434 0.70 2.0 24.6 17 0.01 Qupe -1.6 700 300 2
-434 0.82 1.4 24.6 17 0 Qupu -1.6 700 300 2 434 - - - 0 Tico -1.6 700 150
-2.5 434 0.74 24.6 17 0
-
-Fig. 1: Basal area shares in xcomp simulation of site 434 over a time
-span of 2750 years.
-
-Fig. 2: xcomp canopy shadow at site 434 after 150 simulation periods
-(750 years).
-
-Fig. 3: Histogram of xcomp canopy shadow (vj_lgh) at site 434 after 150
-simulation periods (750 years). Species can only establish at sites
-where the canopy shadow is lower than their regeneration light demand
-(regnLgh in Tab. 1)
-
 ## PICUS
@@
 
-Werner Rammer Kristin Braziunas
+Werner Rammer, Kristin Braziunas
 
@@
 
-*Please provide here the references to this model where it is possible
-to understand the model structure specially for the regeneration module,
-this could be a published paper, a website with the model description, a
-manual or some other document that the reader can access.*
+The most complete reference resource for the iLand regeneration module is:
+Seidl, R., Spies, T.
+a.
+T.
+A.
+T.
+a., Rammer, W., Steel, E.
+A.
+A., Pabst, R.
+J.
+R.
+J.
+R.
+J., & Olsen, K.
+(2012).
+Multi-scale Drivers of Spatial Variation in Old-Growth Forest Carbon Density Disentangled with Lidar and an Individual-Based Landscape Model.
+Ecosystems, 15(8), 1321-1335.
+https://doi.org/10.1007/s10021-012-9587-2
 
@@
 ```{r simObadult , echo=FALSE, fig.cap="Mean species richness at the stand level per site in the simulated versus observed species richness.
-The species richness is estimaed using the Shannon index  calculated by basal area.
+The species richness is estimated using the Shannon index  calculated by basal area.
 A1) groups models that Simulated species composition in adults is higher than the observed, A2) groups models that Simulated species composition in adults is similar to the observed, A3) groups models where the imulated species composition in adults is lower than the observed.
@@
 ```{r simObr, echo=FALSE, fig.cap="Mean species richness per site in the simulated recruitment versus observed species richness.
-The species richness is estimaed using the Shannon index is calculated by basal area.B1) groups models where the simulated species composition is closer than the other models to observed values, B2) groups models where the sites with lower species composition are underestimated, B3) groups models where the simulated species composition recruitment is lower in the whole range compare to the observe.
+The species richness is estimated using the Shannon index is calculated by basal area.B1) groups models where the simulated species composition is closer than the other models to observed values, B2) groups models where the sites with lower species composition are underestimated, B3) groups models where the simulated species composition recruitment is lower in the whole range compare to the observe.
 NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%', fig.pos='H'}
@@
 The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species.
-The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equalt to zero.", out.width='100%', fig.pos='H'}
+The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equalt to zero.", out.width='100%',out.height= "80%", fig.pos='H'}
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_10.png"))

### 2023-03-15 code/figures.R
correct quantile lines in H recruitment figure

@@ -206,8 +206,8 @@
         ggplot2::geom_hline(ggplot2::aes(yintercept = quantile(dat710$ShannonIndexRecruit[dat710$dbh == "7" & dat710$model == "Observed"],
                                                                c(0.25))), 
                             colour = "#990000", linetype = "dashed") +
-        ggplot2::geom_hline(ggplot2::aes(yintercept = quantile(dat710$ShannonIndexRecruit[dat710$dbh == "10" & dat710$model == "Observed"],
-                                                               c(0.77))),
+        ggplot2::geom_hline(ggplot2::aes(yintercept = quantile(dat710$ShannonIndexRecruit[dat710$dbh == "7" & dat710$model == "Observed"],
+                                                               c(0.75))),
                             colour = "#990000", linetype = "dashed") 
     
     ggplot2::ggsave("figures/H7_10.png",

### 2023-03-16 manuscript/manuscriptRW.Rmd
no idea what has changed here but I save it just in case
small additions manuscript

@@
     extra_dependencies: "subfig"
-  word_document:
-    reference_docx: template_paper.docx
 mainfont: Arial
@@
 all the uncertainties, a lot can be gained by a focus on the modeling of
-establishment processes in these model, this will make their long-term
+establishment processes in these models, this will make their long-term
 forest dynamics projections under changing environmental conditions much
@@
 models of forest dynamics capturing accurately tree ingrowth levels,
-initial tree diversity and mortality in tree establishment? (2)Do model
+initial tree diversity and mortality in tree establishment? (2) Do model
 traits explain differences in model performance? (3) How well do models
@@
 need to include a larger number of ecological processes such as
-germination and survival of young seedlings, whereas models that started
+germination and survival of young seedlings, whereas models that start
 from saplings have to aggregate via parameterizations several ecological
@@
 identical. Similarly, the two ForCEEPS variants allow us to isolate the
-importance of the canopy feedback (i.e. simulated actual composition and
-relative abundance of species in the plot) via seed trees for the
+importance of the canopy feedback (i.e., simulated actual composition
+and relative abundance of species in the plot) via seed trees for the
 quantity and quality of simulated ingrowth, as one variant includes this
@@
       format = "latex", booktabs = TRUE,
-      caption = "List of models included.") |>
+      caption = "Models included and their characteristics.") |>
         kableExtra::landscape() |>
@@
 thresholds for the sites differed between 0 and 10 cm, we defined two
-data sets where one included all sites with a dbh threshold of 7cm or
-lower and the other included another 35 sites with dbh thresholds
-between 7 and 10 cm. Thus we focused on two data sets with the
-harmonized dbh thresholds of 7 cm and 10 cm.
+data sets where one included all sites with a diameter threshold of 7 cm
+or lower and the other included another 35 sites with diameter
+thresholds between 7 and 10 cm. Thus we focused on two data sets with
+the harmonized diameter thresholds of 7 cm and 10 cm.
 
@@
 of basal area. The Shannon index was calculated for the establishment
-($H_{n,r}$) (Equation \ref{eqpiShannonR}) and at the stand level
-($H_{n,s}$)(Equation \ref{eqpiShannon}). The higher the value of the
+($H_{R_{n}}$) (Equation \ref{eqpiShannonR}) and at the stand level
+($H_{S_{n,s}}$)(Equation \ref{eqpiShannon}). The higher the value of the
 index, the higher the species diversity at a particular site and sample.
@@
 \begin{equation}
-H_{n,r} = - \sum_{i = 1}^{s}{p_{n,i}\ln(p_{n,r,i})}\label{eqpiShannonR}
+H_{R_{n}} = - \sum_{i = 1}^{s}{p_{R_{n,i}}\ln(p_{R_{n,i}})}\label{eqpiShannonR}
 \end{equation}
@@
 \begin{equation}
-p_{n,r,i} = \frac{r_{BA_{n,i}}}{TotalR_{BA_{n}}} \label{eqpiShannonR2}
+p_{R_{n,i}} = \frac{r_{BA_{n,i}}}{TotalR_{BA_{n}}} \label{eqpiShannonR2}
 \end{equation}
@@
 \begin{equation}
-H_{n,s} = - \sum_{i = 1}^{s}{p_{n,i}\ln(p_{n,i})}\label{eqpiShannon}
+H_{S_{n}} = - \sum_{i = 1}^{s}{p_{S_{n,i}}\ln(p_{S_{n,i}})}\label{eqpiShannon}
 \end{equation}
@@
 \begin{equation}
-p_{n,s,i} = \frac{BA_{n,i}}{Total_{BA_{n}}} \label{eqpiShannon2}
+p_{S_{n,i}} = \frac{BA_{n,i}}{Total_{BA_{n}}} \label{eqpiShannon2}
 \end{equation}
@@
 above zero in sample $n$; $p_{n,r,i}$ is the proportion of the species
-$i$ calculated as the ingrowth basal area ($r_{BA_{n,i}}$) for that
-species $i$ relative to the recruited basal area ($TotalR_{BA_{n}}$) of
-the sample $n$; $p_{n,s,i}$ is the proportion of the species $i$
-calculated as the basal area ($BA_{n,i}$) for that species $i$ relative
-to the total basal area ($TotalR_{BA_{n}}$) of the sample $n$.
+$i$ in sample $n$ calculated as the ingrowth basal area ($r_{BA_{n,i}}$)
+for that species $i$ relative to the recruited basal area
+($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the proportion
+of the species $i$ calculated as the basal area of all trees
+($BA_{n,i}$) of that species $i$ relative to the total basal area
+($TotalR_{BA_{n}}$) of the sample $n$.
 
@@
 quadratic mean diameter in fully stocked pure stands. The value used in
-our comparisons was 1.77 (i.e., we expect stem numbers at 7cm to be 77%
-higher than at 10cm), calculated using Equation \ref{eqReineke}.
+our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm to be 77%
+higher than at 10 cm), calculated using Equation \ref{eqReineke}.
 
@@
 features such as model complexity as defined in @bugmann2022, model type
-(empirical or process based), models with and without canopy feedback
-(the level and species composition of the ingrowth is influenced by the
-existing adult tree community) and the models scale (stand, landscape or
-global). We tested for significant differences considering each of these
-model features and between observed and simulated ingrowth
+(empirical or process based), models with and without canopy feedback,
+and the models scale (stand, landscape or global). We tested for
+significant differences using two sample t-tests by considering each of
+these model features considering observed and simulated ingrowth
 overestimation, ingrowth levels and species diversity respectively.
@@
 proxy of light availability at the forest floor, the annual degree-day
-sum (DD) [@allen1976; @fischlin1995] as a proxy for growing season
-warmth, and the climatic water balance (CWB) as a proxy for soil
-moisture [@speich2019]. The mean ingrowth values were calculated as the
-mean across the 200 samples per site. The observed data were modeled
-using a Generalized Additive Model [@wood2011] with a negative binomial
+sum [@allen1976; @fischlin1995] as a proxy for growing season warmth,
+and the climatic water balance as a proxy for soil moisture
+[@speich2019]. The mean ingrowth values were calculated as the mean
+across the 200 samples per site. The observed data were modeled using a
+Generalized Additive Model [@wood2011] with a negative binomial
 distribution and restricted maximum likelihood to better understand the
@@
 ```{r over, echo=FALSE, fig.cap="Mean ingrowth levels across all samples per site, plotted for the 200 sites and for each model.
-The red dashed lines shows the 25th and 75th percentiles for the 7cm diameter threshold in the observed data.
-There are two boxplots for each model where the left lighted boxplot corresponds to 7cm and the right darker boxplot to 10cm", out.width='100%'}
+The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
+There are two boxplots for each model where the left lighted boxplot corresponds to 7 cm and the right darker boxplot to 10 cm", out.width='100%'}
         knitr::include_graphics(here::here("figures", "recOverAll710.png"))
@@
 ```{r H710, echo=FALSE, fig.cap="Shannon index of tree ingrowth, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
-The red dashed lines represent the 25th percentile of the observed data for the threshold of 7 cm and the 75th percentile of the observed data for the threshold of 10 cm.", out.width='100%'}
+The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.", out.width='100%'}
     knitr::include_graphics(here::here("figures", "H7_10.png"))
@@
 and S2). Most models captured reasonably well or overpredicted species
-diversity at the stand level (Figure \@ref(fig:combinedRichness)A,
-center; cf. Figure S3), but a characteristic feature was that
+diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left
+and center; cf. Figure S3), but a characteristic feature was that
 simulations had a much lower variability of diversity compared to
@@
 of them underestimated ingrowth diversity (4C and Landis-II) (Figure
-\@ref(fig:combinedRichness)A and C, cf. Figure S4). Both the models with
+\@ref(fig:combinedRichness)B and Figure S4). Both the models with
 feedback and those without feedback from the adult trees to ingrowth
@@
 reasonably well ingrowth species diversity compared to the observations.
+Again, simulated diversity was considerably less variable than observed
+diversity in the ingrowth.
 
@@
 data, thus at least partially compensating for the general
-overestimation of ingrowth levels (Figure \@ref(fig:over) and Table S5);
+overestimation of ingrowth levels (Figure \@ref(fig:over) and Table S3);
 six models did not feature a significant difference compared to the
@@
 establishment module (Table \@ref(tab:tableModels)) and the
-overestimation proportion at a diameter threshold of 7 cm (Figure S7).
+overestimation proportion at a diameter threshold of 7 cm (Figure S6).
 All the models except 4C had significant differences between the
@@
 ingrowth for the model ForCEEPS to a closer diversity levels when the
-recruitment module included feedback (ForCEEPs(f)). Again, simulated
-diversity was considerably less variable than observed diversity.
+recruitment module included feedback (ForCEEPs(f)).
 
@@
 models did not match the observed ingrowth patterns across gradients of
-soil moisture and temperature (Figure
-\@ref(fig:)\@ref(fig:nvTrend710BAlightTemp) and S7), and neither was
-there a clear pattern across models. Regarding the soil moisture
-gradient, a group of models featured decreasing ingrowth with increasing
-soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim,
-TreeMig and iLand), while a few models showed the opposite trend
-(ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient,
-all the models that featured a decrease in ingrowth levels with
-increasing soil moisture showed the opposite trend (i.e., more ingrowth
-with increasing temperature), except for 4C. The other models did not
-feature a clear pattern.
+soil moisture and temperature (Figure \@ref(fig:nvTrend710BAlightTemp)
+and S7), and neither was there a clear pattern across models. Regarding
+the soil moisture gradient, a group of models featured decreasing
+ingrowth with increasing soil moisture (4C, ForClim 1, ForClim11,
+SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models
+showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS). Across
+the temperature gradient, all the models that featured a decrease in
+ingrowth levels with increasing soil moisture showed the opposite trend
+(i.e., more ingrowth with increasing temperature), except for 4C. The
+other models did not feature a clear pattern.
 
@@
 @seidl2022]. In the simulations, overall ingrowth species diversity
-levels were well within the observed range for two thirds of the models
-(10 out of 15). Thus, while most of the models are facing difficulties
-to quantitatively match ingrowth levels (cf. above), their performance
-is much better when it comes to matching the composition of ingrowth as
-a function of abiotic and biotic conditions. Most models maintained or
-even increased ingrowth diversity between the 7 and 10 cm thresholds,
+levels were well within the observed range for half of the models (7 out
+of 14). Thus, while most of the models are facing difficulties to
+quantitatively match ingrowth levels (cf. above), their performance is
+better when it comes to matching the composition of ingrowth as a
+function of abiotic and biotic conditions. Most models maintained or
+even decreased ingrowth diversity between the 7 and 10 cm thresholds,
 and the same is visible from the empirical data, but the differences are
@@
 evidence that models with feedback captured better the ingrowth species
-diversity than those without feedback. The similar performance of models
-with and without feedback is likely because the models put more weight
-to the regeneration niche arising from abiotic and biotic filters, than
-from the habitat niche of the adult trees [@grubb1977].
+diversity than those without feedback with the exception of ForCEEPS
+that significantly improved the ingrowth species diversity with the
+model version that included feedback (ForCEEPS(f)). The similar
+performance of models with and without feedback is likely because the
+models put more weight to the regeneration niche arising from abiotic
+and biotic filters, than from the habitat niche of the adult trees
+[@grubb1977].
 
@@
 Technology; [https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich
-(Grant ETH-35 18-1). ODY was funded by SNF project XXXX
+(Grant ETH-35 18-1). ODY was funded by SNF project XXXX. KB received
+support from the European Research Council under the European Union's
+Horizon 2020 research and innovation program (Grant Agreement
+101001905). JB acknowledges funding from the long-term research
+development project no. RVO 67985939 from the Czech Academy of Sciences.
+MM acknowledges financial support from I-Maestro (Innovative forest
+management strategies for a resilient bioeconomy under climate change
+and disturbances, grant no. 22035418, 2019--2022) funded by the ERA-NET
+Cofund ForestValue. KM received funding from the Project No.
+CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant no.
+ITMS2014+313011W580 supported by the Integrated Infrastructure
+Operational Programme funded by the ERDF
 

### 2023-03-17 manuscript/manuscriptRW.Rmd
changes in manuscript second revision internal from Harald

@@
     extra_dependencies: "subfig"
+  word_document:
+    reference_docx: template_paper.docx
 mainfont: Arial
@@
 Tree establishment is a key process in forest dynamics, particularly in
-the context of forest resilience and climate change. Models of forest
-dynamics have been in use for more than 50 years, but an assessment of
-their capacity to represent tree establishment is lacking. We assess how
-well current models capture the overall abundance, species composition
-and mortality of tree establishment.
-
-Using 15 models and model variants built to capture long-term (i.e.,
-decadal to centennial) forest dynamics at the stand, landscape and
-global levels, we simulate tree establishment at 200 sites representing
-environmental gradients across Europe. The results are evaluated against
-comprehensive data from unmanaged forests.
+the context of forest resilience and climate change. Models are pivotal
+for assessing long term (i.e., decadal to centennial) forest dynamics,
+and they have been in use for more than 50 years. However, an assessment
+of their ability to accurately represent tree establishment is lacking.
+We assess how well current models capture the overall abundance, species
+composition and mortality of tree establishment.
+
+Using 15 models and model variants built to capture long-term forest
+dynamics at the stand, landscape and global levels, we simulate tree
+establishment at 200 sites representing large environmental gradients
+across Europe. The results are evaluated against comprehensive data from
+unmanaged forests.
 
@@
 patterns are not consistent across models. The ingrowth of *Abies alba*
-is captured across temperature and soil moisture gradients, whereas
-ingrowth of *Fagus sylvatica* is generally underestimated. *Picea abies*
-ingrowth is generally overestimated, particularly at colder sites.
-Increasing complexity in the tree establishment modules of the models is
-not related to higher accuracy of simulated tree establishment.
-Furthermore, individual model design is more important than scale
-(stand, landscape, global) and approach (empirical, process-based) to
-accurately capture tree establishment.
-
-In spite of more or less substantial mismatches between model results
-and data, it is remarkable that most of the models capture the essential
+is captured well by most models across temperature and soil moisture
+gradients, whereas ingrowth of *Fagus sylvatica* is generally
+underestimated. *Picea abies* ingrowth is generally overestimated,
+particularly at colder sites. Increasing complexity in the tree
+establishment modules of the models is not related to higher accuracy of
+simulated tree establishment. Furthermore, individual model design is
+more important than scale (stand, landscape, global) and approach
+(empirical, process-based) to accurately capture tree establishment.
+
+Despite sometimes considerable mismatches between simulation results and
+data, it is remarkable that most of the models capture the essential
 features of the highly complex process of tree establishment, while not
@@
 all the uncertainties, a lot can be gained by a focus on the modeling of
-establishment processes in these models, this will make their long-term
-forest dynamics projections under changing environmental conditions much
-more robust.
+establishment processes in these models. This has the potential to
+render long-term projections of forest dynamics under changing
+environmental conditions much more robust.
 
@@
 important to understand how forest structure, composition and function
-will change, and to evaluate forest capacity to adapt or be resilient
+will change, and to evaluate forest capacity to adapt to or be resilient
 after disturbance [@lindner2010; @seidl2022]. A wide range of models of
@@
 impacts of climate on forest dynamics [@bugmann2022]. From these
-studies, it is evident that while we have a reasonably good
-understanding how to model tree growth [@vanclay1997; @bugmann1996], and
-substantial efforts have been dedicated to improving the depiction of
-tree mortality [@bugmann2019]. However, tree establishment remains a
-much less studied process that is often represented rather coarsely in
-these models [@walck2011; @leishman1992], which is a clear research gap
-in the context of climate-induced forest disturbances and forest
-resilience.
-
-Tree establishment arises from multiple processes such as reproduction,
+studies, it is evident that we have a reasonably good understanding how
+to model tree growth [@vanclay1997; @bugmann1996], and substantial
+efforts have been dedicated to improving the representation of tree
+mortality [@bugmann2019; @cailleret2017]. In contrast, tree
+establishment is a much less studied process that is often represented
+rather coarsely in these models [@walck2011; @leishman1992; @price2001],
+which is a clear research gap in the context of climate-induced forest
+disturbances and forest resilience.
+
+Tree establishment arises from multiple processes such as pollination,
 fruit maturation, seed production, dispersal, germination, juvenile
@@
 thus appear highly stochastic [@bogdziewicz2021]. Identifying the
-appropriate level of complexity for the formulation of the key factors
-that are leading to successful tree establishment is challenging.
-Currently, tree establishment processes in dynamic forest models are
-handled in a multitude of ways [@konig_tree_2022; @bugmann2022]: a) from
-entirely ignoring it [as done in classical forest growth models, e.g.,
-Pretzsch et al. -@pretzsch2002], b) across the use of a few simple
-environmental filters, as done in most forest "gap" models [@shuman2011]
-and Dynamic Global Vegetation Models [e.g., @hickler2012; @smith2001],
-c) to complex approaches that incorporate local feedback from the
-canopy, multiple ecological processes and often also small time steps
-[e.g., @seidl2012; @wehrli2006], or d) field-based statistical
-parameterizations, which however are not easy to extrapolate in space
-and time [e.g., @ribbens1994].
+appropriate level of complexity for the mathematical formulation of the
+key factors that are leading to successful tree establishment is
+challenging. Currently, tree establishment processes in dynamic forest
+models are handled in a multitude of ways [@konig_tree_2022;
+@bugmann2022]: from a) entirely ignoring it [as done in classical forest
+growth models, e.g., Pretzsch et al. -@pretzsch2002], across b) the use
+of a few simple environmental filters, as done in most forest "gap"
+models [@shuman2011] and Dynamic Global Vegetation Models [e.g.,
+@hickler2012; @smith2001], to c) complex approaches that incorporate
+local feedback from the canopy, multiple ecological processes and often
+also small time steps [e.g., @seidl2012; @wehrli2006], or d) field-based
+statistical parameterizations, which however are not easy to extrapolate
+in space and time [e.g., @ribbens1994].
 
 Overall, models are needed to (1) synthesize existing empirical data and
-explore their consequences, (2) assess future tree establishment e.g.,
-in the context of global change scenarios and, (3) identify most
-important processes needed to define relevant ecological processes.
-Given the current strategies that are used in models of forest dynamics
-to represent tree establishment, model behavior often faces problems
-that can have strong impacts on simulated forest dynamics, such as very
-high levels of tree establishment, which then necessitate excess
+explore their consequences, (2) assess future tree establishment e.g. in
+the context of global change scenarios, and (3) identify the most
+important processes that are shaping ecological patterns. Given the
+current strategies that are used in models of forest dynamics to
+represent tree establishment, model behavior often is prone to problems,
+such as very high levels of tree establishment that necessitate excess
 mortality at early stages of tree life to simulate correct stand
 structure and composition. Also, inadequately high species diversity in
-tree establishment may be simulated. This is characteristic of
+tree establishment may be simulated, which is characteristic of
 "classical" forest gap models [@gutiérrez2016], at least as long as the
-simulations comprise a multitude of species. Other models use
-calibration against local canopy-level data to define appropriate
-ingrowth levels, which is likely to strongly hamper the general
-applicability of these models e.g., under scenarios of climate change.
-Furthermore, correctly capturing the species composition in tree
-establishment as a function of the presence of seed trees in the canopy
-is often a particular challenge, potentially leading to unrealistic
-successional drift in the model, which must be corrected by factors that
-are hard to parameterize [e.g., @lischke2006]. A related issue is the
-excessive reduction of species diversity due to positive feedback
-effects such that eventually just single-species stands remain. This is
-sometimes corrected by the incorporation of a low level of seed influx
-of all species at all times [@schumacher2006] or by restricting the
-number of seeds per species in the seedbank [@lischke2006]. However,
-simulated species composition is usually exceedingly sensitive to that
-level, whereas the parameters of such functions are poorly constrained
-by field data. Lastly, there is often a problem with insufficient
-observational constraints on parameter values for models that start from
-very small tree sizes (e.g., 10 cm of height or even seeds) and track
-tree development in a process-oriented manner by considering a multitude
-of ecological influences, rather than emphasizing tree establishment
-into a larger size class [e.g., @wehrli2006].
+simulations comprise a multitude of species. Some models use calibration
+against local canopy-level data to define appropriate ingrowth levels,
+which is likely to strongly hamper the general applicability of these
+models e.g. under scenarios of climate change. Furthermore, correctly
+capturing the species composition in tree establishment as a function of
+the presence of seed trees in the canopy is often a particular
+challenge, potentially leading to unrealistic successional drift in the
+model, which must be corrected by factors that are hard to parameterize
+[e.g., @lischke2006]. A related issue is the excessive reduction of
+species diversity due to positive feedback effects such that eventually
+just single-species stands remain. This is sometimes corrected by the
+incorporation of a low level of seed influx of all species at all times
+[@schumacher2006] or by restricting the number of seeds per species in
+the seedbank [@lischke2006]. However, simulated species composition is
+usually exceedingly sensitive to that level, whereas the parameters of
+such functions are poorly constrained by field data. Lastly, there is
+often a problem with insufficient observational constraints on parameter
+values for models that start from very small tree sizes (e.g., 10 cm of
+height or even seeds) and track tree development in a process-oriented
+manner by considering a multitude of ecological influences, rather than
+emphasizing tree establishment into a larger size class [e.g.,
+@wehrli2006].
 
@@
 projections of future forest dynamics, i.e., when the models need to be
-operated in extrapolation mode, but also from a fundamental ecological
-point of view increasing systems understanding. Also in the field-based
+operated in extrapolation mode, as well as from a fundamental ecological
+point of view for increasing systems understanding. Also in empirical
 ecology, tree establishment has not received a lot of attention
@@
 evaluated against a continental-scale, multi-species harmonized dataset
-on tree establishment [@käber2022]. This is a challenging experiment for
-these models as they do also have to estimate the forest stand above the
-tree establishment. By tree establishment, we refer to the passing of
-trees across a specific diameter threshold ("ingrowth"). We included
-models that are based on a range of "philosophies", from models
-operating at the stand to the global scale as well as the range of
-models from empirically derived to "process-based" [e.g., @bugmann1996;
-@fabrika2012; @lexer2001; @reyer2014]. Because of the large variability
-in tree establishment patterns in nature and the large number of factors
-driving this process --- including some that are not incorporated in
-most models such as deer browsing --- we do not aim at a detailed
-statistical evaluation of each model. Instead, we aim at evaluating the
-general ingrowth patterns and magnitudes simulated by the models, and
-benchmark the simulated regeneration niche of multiple species against
-empirical data along a wide environmental gradient of temperature,
-moisture, and light availability.
-
-More specifically, we aimed to answer the following questions: (1) Are
+on tree establishment [@käber2022]. By tree establishment, we refer to
+the passing of trees across a specific diameter threshold ("ingrowth").
+We included models that are based on a range of "philosophies", from
+models operating at the stand to the global scale as well as the range
+of models from empirically derived to "process-based" [e.g.,
+@bugmann1996; @fabrika2012; @lexer2001; @reyer2014]. Because of the
+large variability in tree establishment patterns in nature and the large
+number of factors driving this process --- including some that are not
+incorporated in most models such as deer browsing --- we do not aim at a
+detailed statistical evaluation of each model. Instead, we aim at
+evaluating the general ingrowth patterns and magnitudes simulated by the
+models, and benchmark the simulated regeneration niche of multiple
+species against empirical data along a wide environmental gradient of
+temperature, moisture, and light availability.
+
+More specifically, we aim to answer the following questions: (1) Are
 models of forest dynamics capturing accurately tree ingrowth levels,
-initial tree diversity and mortality in tree establishment? (2) Do model
-traits explain differences in model performance? (3) How well do models
-capture total ingrowth and individual species regeneration niches across
-environmental gradients of light availability, temperature, and soil
-moisture?
+initial tree species diversity and mortality in the establishment? (2)
+Do model traits explain differences in model performance? (3) How well
+do the models capture total ingrowth and individual species regeneration
+niches across environmental gradients of light availability,
+temperature, and soil moisture?
 
@@
 FORMIND and iLand). The different approaches underlying the models have
-strongly influenced the formulation of tree establishment.
+strongly influenced the formulation of tree establishment processes.
 
@@
 
-The starting point for the models establishment differs as well, ranging
-from seedbank, seed or seedling to sapling (i.e., trees much larger than
-10 cm height, often at least 2-3 m tall). Models that start from seed
-need to include a larger number of ecological processes such as
+The starting point for tree establishment in the models differs as well,
+ranging from seedbank, seed or seedling to sapling (i.e., trees much
+larger than 10 cm height, often ca. 2-3 m tall). Models that start from
+seed need to include a larger number of ecological processes such as
 germination and survival of young seedlings, whereas models that start
 from saplings have to aggregate via parameterizations several ecological
-processes because they are not treated explicitly. This reduces model
-complexity but comes at the cost of blurring process representation. In
-the models used in this study the overall complexity in the
+processes because they are not treated explicitly. This later approach
+reduces model complexity but comes at the cost of blurring process
+representation.
+
+In the models used in this study, the overall complexity in the
 establishment modules varies considerably. Following @bugmann2022, we
@@
 establishment modules (mean establishment formulation complexity
-complexity across all processes \>0, Table \@ref(tab:tableModels)).
-
-Two models, ForClim and ForCEEPS, were used here with two alternative
-variants. ForClim variant 1 [@bugmann1996] is based on a recruitment
-module that adheres closely to the concept introduced by @botkin1972,
-whereas ForClim variant 11 is adopting a different approach where
-individual species properties and its relation to the environment are
-more important [@huber2020]. These two model variants allow us to
-evaluate a more process-based and complex module (variant 11) against a
-simple module (variant 1), while the rest of the model structure is
-identical. Similarly, the two ForCEEPS variants allow us to isolate the
-importance of the canopy feedback (i.e., simulated actual composition
-and relative abundance of species in the plot) via seed trees for the
-quantity and quality of simulated ingrowth, as one variant includes this
-feedback whereas the other does not.
+complexity across all processes \>0, Table \@ref(tab:tableModels)). Two
+models, ForClim and ForCEEPS, were used here with two alternative
+variants of establishment. ForClim variant 1 [@bugmann1996] is based on
+a recruitment module that adheres closely to the concept introduced by
+@botkin1972, whereas ForClim variant 11 is adopting a different approach
+where individual species properties and their relation to the
+environment are important [@huber2020]. These two model variants allow
+us to evaluate a more process-based and complex module (variant 11)
+against a simple module (variant 1), while the rest of the model
+structure is identical. Similarly, the two ForCEEPS variants allow us to
+isolate the importance of the canopy feedback (i.e., simulated actual
+composition and relative abundance of species in the plot) via seed
+trees for the quantity and quality of simulated ingrowth, as one variant
+includes this feedback whereas the other does not.
 
@@
 central Europe as compiled in the framework of the EuFoRIa network
-[@euforia2019] (Figure \@ref(fig:map)). These forests are unmanged for
-at least 10 years prior to their designation as reserves, and most of
-them provide time series of natural forest dynamics over multiple
-decades and up to ca. 80 years. The census periods range from 3 to 37
-years, with an average of 14 years. The data provide information at the
-tree level, thus allowing for sequential comparison such as
+[@euforia2019] (Figure \@ref(fig:map)). These forests have been
+unmanaged for at least 10 years prior to their designation as reserves,
+and most of them provide time series of natural forest dynamics over
+multiple decades and up to ca. 80 years. The census periods range from 3
+to 37 years, with an average of 14 years. The data provide information
+at the tree level, thus allowing for sequential comparison such as
 individual-tree ingrowth and death between the measurements.
@@
 these sites featured at least two consecutive measurements. Ingrowth
-thresholds for the sites differed between 0 and 10 cm, we defined two
-data sets where one included all sites with a diameter threshold of 7 cm
-or lower and the other included another 35 sites with diameter
-thresholds between 7 and 10 cm. Thus we focused on two data sets with
-the harmonized diameter thresholds of 7 cm and 10 cm.
+thresholds for the sites differed between 0 and 10 cm. For our study, we
+defined two data sets where one included 165 sites with a diameter
+threshold of 7 cm or lower, and the other included another 35 sites with
+diameter thresholds between 7 and 10 cm. Thus, we focused on two data
+sets with harmonized diameter thresholds of 7 cm and 10 cm,
+respectively.
 
@@
 per ha and per decade. The observations used in this study featured
-30,900 established trees, and ingrowth rates per site, sample and
+30,900 established trees. Ingrowth rates per site, sample and
 decade ranged from zero to 1246 trees per ha, with a mean of 56 trees
@@
 information (except for the data specified below) nor any data on tree
-ingrowth provided prior to the simulation. That is, the models were run
-in "blind flight" mode.
+ingrowth or forest stand features provided prior to the simulation. That
+is, the models were run in "blind flight" mode.
 
@@
 -@muñozsabater2019], daily and monthly resolution [CHELSA data, Karger
-et al. -@karger2021]. Some variables required for some models such as
-relative humidity or vapor pressure deficit had to be calculated from
-these variables. The final strategy how to use the climate data was
-different in each model, based on the approach that best suited the
-model (Table \@ref(tab:tableModels)). Soil quality data was provided as
-a continuous value between 1 and 5 [Soilgrids dataset, Hengl et al.
--@hengl2017]. The protocol also provided the elevation, slope and aspect
-for each of the 200 sites [@asterscienceteam2019], but no other spatial
-information such as coordinates was provided, with the exception of
-iLand and aDGVM2, which used blurred coordinates to derive highly
-detailed soil data.
+et al. -@karger2021]. Some variables required for some models, such as
+relative humidity or vapor pressure deficit, were calculated from these
+variables. The final strategy how to use the climate data was different
+in each model, based on the approach that best suited the model (Table
+\@ref(tab:tableModels)). Soil quality data were provided as continuous
+values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
+The protocol also provided the elevation, slope and aspect for each of
+the 200 sites [@asterscienceteam2019], but no other spatial information
+such as coordinates, with the exception of iLand and aDGVM2, which
+required blurred coordinates to derive highly detailed soil data.
 
@@
 Quercus spp., Alnus glutinosa, Pinus sylvestris.* The same set of
-species were used in all the 200 sites. Two models included their
-standard set of species for the simulations which is much larger (i.e.,
-ForClim 1, ForClim 11, TreeMig). In three models, fewer than these
-eleven species were simulated (4C, xComp, LPJ-GUESS) (Table
+species was used at all 200 sites. Two models included their standard
+set of species for the simulations, which is much larger (i.e., ForClim
+1, ForClim 11, TreeMig). In three models, fewer than these eleven
+species were simulated (4C, xComp, LPJ-GUESS) (Table
 \@ref(tab:tableModels)). In 4C only *Fagus sylvatica*, *Picea abies*,
@@
 community assembly processes and trait filtering generate plant
-communities that are adjusted to the biotic and abiotic conditions and
-plants can be classified into different ecological strategies based on
-their trait values in a post-processing step.
+communities that are adjusted to the biotic and abiotic conditions, and
+the simulated plants can be classified into ecological strategies based
+on their trait values in a post-processing step.
 
@@
 dimensions of very old individuals. LPJ-GUESS had 2% of the sample
-outputs without tree establishment and were considered in the data as
-zero stand basal area and zero ingrowth for all the tree species
-simulated and one site produced grasslands and was not included in the
-results.
+outputs without tree establishment, and these were considered as zero
+stand basal area and zero ingrowth for all the tree species simulated;
+one site produced grassland rather than a forest and was not included in
+the results.
 
@@
 
-The simulation results were analyzed regarding the ingrowth levels
-(i.e., ingrowth number per unit time and space), species diversity of
-the ingrowth, and ingrowth mortality. We evaluated the species diversity
-across models and in relation to the observed data by calculating the
-Shannon index based on the relative proportion of the species in terms
-of basal area. The Shannon index was calculated for the establishment
-($H_{R_{n}}$) (Equation \ref{eqpiShannonR}) and at the stand level
-($H_{S_{n,s}}$)(Equation \ref{eqpiShannon}). The higher the value of the
-index, the higher the species diversity at a particular site and sample.
+The simulation results were analyzed regarding (1) ingrowth levels
+(i.e., ingrowth number per unit time and space), (2) species diversity
+of the ingrowth, and (3) ingrowth mortality. We evaluated the species
+diversity across models and in relation to the observed data by
+calculating the Shannon index based on the relative proportion of the
+species in terms of basal area. The Shannon index was calculated for the
+establishment ($H_{R_{n}}$) (Equation \ref{eqpiShannonR}) and at the
+stand level ($H_{S_{n,s}}$)(Equation \ref{eqpiShannon}). The higher the
+value of the index, the higher the species diversity at a particular
+site and sample.
 
@@
 ```
-where $s$ is the total number of species present and have basal area
-above zero in sample $n$; $p_{n,r,i}$ is the proportion of the species
-$i$ in sample $n$ calculated as the ingrowth basal area ($r_{BA_{n,i}}$)
-for that species $i$ relative to the recruited basal area
-($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the proportion
-of the species $i$ calculated as the basal area of all trees
+where $s$ is the total number of species present and that have a basal
+area larger than zero in sample $n$; $p_{n,r,i}$ is the proportion of
+the species $i$ in sample $n$ calculated as the ingrowth basal area
+($r_{BA_{n,i}}$) for that species $i$ relative to the total recruited
+basal area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
+proportion of the species $i$ calculated as the basal area of all trees
 ($BA_{n,i}$) of that species $i$ relative to the total basal area
@@
 Reineke self-thinning rule [@reineke1933; @pretzsch2003] as a reference
-to estimate whether the different ingrowth values were above or below
-the expected theoretical rate between these two diameters. The Reineke
-self-thinning rule is usually calculated from even-aged, single species
-stands and assumes a relationship between the number of stems and the
-quadratic mean diameter in fully stocked pure stands. The value used in
-our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm to be 77%
-higher than at 10 cm), calculated using Equation \ref{eqReineke}.
+to estimate whether the ratio of ingrowth between the 7 and 10 cm
+diameter thresholds were above or below the expected theoretical rate
+between these two diameters. The Reineke self-thinning rule is usually
+calculated from even-aged, single species stands and assumes a fixed
+relationship between the number of stems and the quadratic mean diameter
+in fully stocked pure stands. The value used in our comparisons was 1.77
+(i.e., we expect stem numbers at 7 cm to be 77% higher than at 10 cm),
+calculated using Equation \ref{eqReineke}.
 
@@
 ```
-We assessed model traits in relation to model performance focusing on
-features such as model complexity as defined in @bugmann2022, model type
-(empirical or process based), models with and without canopy feedback,
-and the models scale (stand, landscape or global). We tested for
-significant differences using two sample t-tests by considering each of
-these model features considering observed and simulated ingrowth
-overestimation, ingrowth levels and species diversity respectively.
-
-The total ingrowth and individual species regeneration niches were
-evaluated across environmental gradients of light availability,
-temperature, and soil moisture. Ingrowth levels were evaluated along
-three key environmental gradients, i.e., stand basal area (the basal
-area of all the trees in each sample including tree establishment) as a
+We assessed model preformance in relation to model traits focusing on
+(1) model complexity as defined by @bugmann2022, (2) model type
+(empirical or process based), (3) the presence of absence of a canopy
+feedback for ingrowth, and (4) the scale of application of the model
+(stand, landscape or global). We tested for significant differences
+using two-sided t-tests by considering each of these model features
+regarding the observed and simulated ingrowth levels and species
+diversity, respectively.
+
+Total ingrowth and the regeneration niches of the individual species
+were evaluated across the environmental gradients of light availability,
+temperature, and soil moisture. Stand basal area (the basal area of all
+the trees in each sample including tree establishment) was used as a
 proxy of light availability at the forest floor, the annual degree-day
@@
 and the climatic water balance as a proxy for soil moisture
-[@speich2019]. The mean ingrowth values were calculated as the mean
-across the 200 samples per site. The observed data were modeled using a
-Generalized Additive Model [@wood2011] with a negative binomial
-distribution and restricted maximum likelihood to better understand the
-environmental gradients trend in the levels of the observed ingrowth,
+[@speich2019]. Ingrowth values were calculated as the mean across the
+200 samples per site. The observed data were modeled using a Generalized
+Additive Model [@wood2011] with a negative binomial distribution and
+restricted maximum likelihood to better understand the relationship
+between the environmental gradients and the levels of observed ingrowth,
 relative to the simulation results per model.
 
-In order to analyze the individual species regeneration niches across
-the climatic gradients, we focused on the five main tree species or
-genera: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus
-sylvestris*, and *Quercus* spp. For these species, we calculated the
-mean ingrowth basal area share per site as the mean value across the
-available samples per site:
+In order to analyze the regeneration niches across the climatic
+gradients, we focused on the five main tree species or genera: *Abies
+alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
+*Quercus* spp. For these species, we calculated the share in the
+ingrowth basal area per site as the mean across the available samples
+per site:
 
@@
 ```
-where $s$ is the total number of species simulated; $\bar{r}_{t,i}$ is
-the mean basal area ingrowth of the species $i$ per site $t$ across the
-available samples at that site.
-
-We removed 35 sites from the comparisons at 7 cm ingrowth, because field
-data did not include these measurements. Furthermore, regarding the
-ingrowth basal area share per species ($\bar{R}_{BA_{share_{i}}}$), we
-categorized this as zero when the ingrowth basal area of that species
-($\bar{r}_{BA_{t,i}}$) was zero but also when both the ingrowth basal
-area of that species ($\bar{r}_{BA_{t,i}}$) and the total ingrowth basal
-area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.
+where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
+is the mean basal area ingrowth of the species $i$ per site $t$ across
+the available samples at that site.
+
+We removed 35 sites from the comparisons at a threshold of 7 cm
+ingrowth, because field data did not include these measurements.
+Furthermore, regarding the share in the ingrowth basal area per species
+($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the
+ingrowth basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero but
+also when both the ingrowth basal area of that species
+($\bar{r}_{BA_{t,i}}$) and the total ingrowth basal area (
+$\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.
 
@@
 
-## Ingrowth levels, tree diversity and mortality in tree establishment
-
-The ingrowth levels varied strongly across the 15 models and typically
-did not match the levels found in the forest reserve data (Figure
-\@ref(fig:over)). Ingrowth was overestimated in most models for both the
-7 and 10 cm diameter thresholds, with the exception of the empirical
-stand model SIBYLA, the landscape model Landis-II and the global model
-aDGVM2, which estimated ingrowth levels at the lower end of the
-plausibility interval of the observed data (Figure \@ref(fig:over)). The
-models with the largest overestimation were the stand model PICUS and
-the landscape model TreeMig. For most models, the variability of
+## Ingrowth levels
+
+Simulated ingrowth levels varied strongly across the 15 models and
+typically did not match the levels found in the forest reserve data
+(Figure \@ref(fig:over)). Ingrowth was overestimated in most models for
+both the 7 and 10 cm diameter thresholds, with the exception of the
+empirical stand model SIBYLA, the landscape model Landis-II and the
+global model aDGVM2, which estimated ingrowth levels at the lower end of
+the plausibility interval of the observed data (Figure \@ref(fig:over)).
+The models with the largest overestimation were the stand model PICUS
+and the landscape model TreeMig. For most models, the variability of
 simulated ingrowth levels across the 200 sites (visible from the
@@
 The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
-There are two boxplots for each model where the left lighted boxplot corresponds to 7 cm and the right darker boxplot to 10 cm", out.width='100%'}
+There are two boxplots for each model where the left lighter boxplot corresponds to 7 cm and the right darker boxplot to 10 cm", out.width='100%'}
         knitr::include_graphics(here::here("figures", "recOverAll710.png"))
@@
 
-Both observations and simulated data had no ingrowth in some samples and
-at some sites (Table S1). The observed data had 4% of the samples with
-no ingrowth. Only three models had a larger proportion of no ingrowth
-(4C, Landis-II and aDVM2). Two models always simulated ingrowth for both
-the 7 and 10 cm threshold (xComp and TreeMig), i.e., they did not
-feature any zero values. The other models had a very low percentage of
-samples with no ingrowth (0.01% - 2.39%), i.e., they had distinctly
-fewer occurrences of zeros compared to the observations.
+Both the observations and the simulated data had no ingrowth in some
+samples and at some sites (for details, cf. Table S1). The observed data
+had 4% of the samples with no ingrowth. Only three models had a larger
+proportion of no ingrowth (4C, Landis-II and aDVM2). Two models always
+simulated ingrowth for both the 7 and 10 cm threshold (xComp and
+TreeMig), i.e., they did not feature any zero values. The other ten
+models had a very low percentage of samples with no ingrowth (0.01% -
+2.39%), i.e., they had distinctly fewer occurrences of zeros compared to
+the observations.
+
+## Tree species diversity of ingrowth
 
@@
 well as quantified by the Shannon index (Figure \@ref(fig:H710)). Five
-models overestimated ingrowth diversity: ForCEEPS, ForClim 11, PICUS,
-TreeMig and LandClim, the latter mostly for the 7 cm diameter threshold.
-The model 4C is a special case, as it simulated five species only, i.e.,
-its diversity values are not directly comparable to those of the other
-models, nor to the observations. Only one model, Landis-II, consistently
-underestimated ingrowth diversity.
+models overestimated ingrowth diversity: ForCEEPS, ForClim 1, PICUS,
+TreeMig and LandClim, the latter particularly for the 7 cm diameter
+threshold. The model 4C is a special case, as it simulated five species
+only, i.e., its diversity values are not directly comparable to those of
+the other models, nor to the observations. Only one model, Landis-II,
+consistently underestimated ingrowth diversity.
 
@@
 thresholds, species diversity in the ingrowth compared to stand-level
-diversity did not have clear patterns across all the models (Figures S1
-and S2). Most models captured reasonably well or overpredicted species
+diversity did not have clear patterns across the models (Figures S1 and
+S2). Most models captured reasonably well or overpredicted species
 diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left
@@
 
-The majority of the models overpredicted ingrowth diversity and only two
-of them underestimated ingrowth diversity (4C and Landis-II) (Figure
+The majority of the models overpredicted ingrowth diversity; only two of
+them underestimated ingrowth diversity (4C and Landis-II) (Figure
 \@ref(fig:combinedRichness)B and Figure S4). Both the models with
@@
 
-```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site and for observed and simulated values.
+```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site for observed and vs.simulated data.
 n indicates the number of models falling in each group.
@@
 B) Examples for the three trends across models for ingrowth at the 7 cm threshold.
-The selection of the representative model for each trend group was based on the full data shown in Figures S5 and S6.", out.width='75%'}
+The full data are shown in Figures S5 and S6.", out.width='75%'}
 knitr::include_graphics(here::here("figures", "combined_richness.png"))
 ```
+
+## Mortality in tree establishment
 
@@
 models that overestimated ingrowth had mortalities lower than the
-expected self-thinning ratio with the exception of LandClim (Figure
+expected self-thinning ratio, with the exception of LandClim (Figure
 \@ref(fig:mortality13)B). It is noteworthy that most of the models that
@@
 
-Lastly, some models that had a low number of individuals at 7 cm
+Lastly, some models that had a low number of individuals at the 7 cm
 threshold also had a smaller decrease of ingrowth between 7 and 10 cm
@@
 
-## Model traits and model performance
+## Model performance and model traits
 
@@
 All the models except 4C had significant differences between the
-observed mean ingrowth values and the simulated, therefore it was not
-possible to differentiate in terms of performance between models forest
-type (empirical or process based) and across models scales (stand,
-landscape or global) (Table S5). The ingrowth species diversity values
-simulated per models were also assessed againts the observed data to see
-if there was a difference between models that include feedback compare
-to those that did not include feedback, all the models except
-ForCEEPs(f) and FORMIND had significant differences (Table S6). It is
-noteworthy the change from overpredicted species diversity in the
-ingrowth for the model ForCEEPS to a closer diversity levels when the
-recruitment module included feedback (ForCEEPs(f)).
-
-## Total ingrowth and individual species regeneration niches
-
-When evaluating the representation of ingrowth levels along key
-gradients of light availability (basal area), temperature (degree-days)
-and soil moisture (climatic water balance), distinct features emerge:
-the models reproduced the effect of basal area in magnitude and patterns
-(Figure \@ref(fig:nvTrend710BAlightTemp) and S7) better than the effects
-of the climatic gradients, where they featured varying effects.
+observed and simulated mean ingrowth values. Therefore it was not
+possible to differentiate in terms of performance between models type
+(empirical or process based), or their scales (stand, landscape or
+global) (Table S5). The values of ingrowth species diversity simulated
+by each model were also assessed against the observed data to see if
+there was a difference between models that include a feedback compared
+to those that did not; all the models except ForCEEPs(f) and FORMIND had
+significant differences (Table S6). It is noteworthy that for the model
+ForCEEPS the pattern changed from the overprediction of species
+diversity in the ingrowth to a diversity level that is closer to
+observations when the recruitment module included feedback
+(ForCEEPs(f)).
+
+## Ingrowth gradients on regeneration niches
+
+When evaluating ingrowth levels along key gradients of light
+availability (basal area), temperature (degree-days) and soil moisture
+(climatic water balance), distinct features emerge: the models
+reproduced the effect of basal area in both its magnitude and patterns
+(Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better
+than the effects of the climatic gradients, where they featured varying
+patterns.
 
 In the observed data, ingrowth levels decreased clearly with increasing
-stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)). This trend
+stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a). This trend
 was captured by the Landis-II model only, albeit at much lower values of
@@
 balance gradient, and showed a slight increase with temperature. The
-models did not match the observed ingrowth patterns across gradients of
-soil moisture and temperature (Figure \@ref(fig:nvTrend710BAlightTemp)
-and S7), and neither was there a clear pattern across models. Regarding
-the soil moisture gradient, a group of models featured decreasing
-ingrowth with increasing soil moisture (4C, ForClim 1, ForClim11,
-SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models
-showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS). Across
-the temperature gradient, all the models that featured a decrease in
-ingrowth levels with increasing soil moisture showed the opposite trend
-(i.e., more ingrowth with increasing temperature), except for 4C. The
-other models did not feature a clear pattern.
-
-```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site with y-axis scaled by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree sum.
+models did not match the observed ingrowth patterns across the gradients
+of soil moisture and temperature (Figure
+\@ref(fig:nvTrend710BAlightTemp)b,c and S7), and neither was there a
+clear pattern across models. Regarding the soil moisture gradient, a
+group of models featured decreasing ingrowth with increasing soil
+moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim,
+TreeMig and iLand), while a few models showed the opposite trend
+(ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient,
+all the models that featured a decrease in ingrowth levels with
+increasing soil moisture showed the opposite trend (i.e., more ingrowth
+with increasing temperature), except for 4C. The other models did not
+feature a clear pattern.
+
+```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum.
 The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
-For comparing the magnitude of recruitment across models the same graphs with equal y-axis are availalbe in Figure S7", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
+For comparing the magnitude of recruitment across models the same graphs with equal y-axis cf.
+Figure S7.", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
 
@@
 basal area in the ingrowth of these five species compared to the
-observed data along the two major gradients of climate (Figure
+observed data along the two major climatic gradients (Figure
 \@ref(fig:diff), cf. Figure S8). However, the models differed
-considerably in the way in which they simulated their climatic niche.
+considerably in the way in which they simulated these climatic niches.
 
@@
 models, such as 4C that did not simulate *Abies alba*. The patterns
-across models for *Fagus sylvatica* ingrowth were more complex, the
+across models for *Fagus sylvatica* ingrowth were more complex, as its
 ingrowth was underestimated by many models across the environmental
 gradients featuring negligible ingrowth at most sites, while others
-overestimate it in different parts of the environmental space. Most of
-the models overestimated the share of *Picea abies* ingrowth in the
-cold-wet part of the gradients. *Pinus sylvestris* and *Quercus* spp.
-represented a smaller share of the ingrowth in the observed data mostly
-at warm-dry sites. Many models, however, erroneously featured ingrowth
-for these two species along most of the climatic gradients although at a
-low basal area share. Five models (4C, FORMIND, SIBYLA, xComp, and
-LandClim) had almost no ingrowth of *Pinus sylvestris* and *Quercus*
-spp., or did not feature any ingrowth of these species at all, close to
-the patterns from the observed data.
+overestimate it in different parts of the environmental space. Most
+models overestimated the share of *Picea abies* ingrowth in the cold-wet
+part of the gradients. *Pinus sylvestris* and *Quercus* spp. represented
+a smaller share of the ingrowth in the observed data mostly at warm-dry
+sites. Many models, however, erroneously featured ingrowth for these two
+species along most of the climatic gradients although at a low basal
+area share. Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had
+almost no ingrowth of *Pinus sylvestris* and *Quercus* spp., or did not
+feature any ingrowth of these species at all, close to the patterns in
+the observed data.
 
@@
 fundamental to, for example, evaluate post-disturbance dynamics as it
-will define forest state in the first decades [@seidl2022]. If this
+will define the forest state in the first decades [@seidl2022]. If this
 initial phase is not captured well, we can not properly assess aspects
@@
 particularly with respect to the number of events captured (number of
-sites and repetitive measurements), which is essential for a highly
+sites and repeated measurements), which is essential for a highly
 'noisy' process such as tree establishment. The use of this dataset for
-model benchmarking provided novel insights on the capability of
-state-of-the-art models to simulate ingrowth levels, species
+model benchmarking provided novel insights on the ability of
+state-of-the-art models to accurately simulate ingrowth levels, species
 composition, and mortality in an early stage of tree life, i.e., between
@@
 
-## Most of the models overestimated tree ingrowth levels
+## Ingrowth levels
 
@@
 
+## Species composition of ingrowth
+
 Correctly capturing species composition in tree establishment is
@@
 its sensitivity to drought or resilience to disturbances [@redmond2015;
-@seidl2022]. In the simulations, overall ingrowth species diversity
-levels were well within the observed range for half of the models (7 out
-of 14). Thus, while most of the models are facing difficulties to
+@seidl2022]. In the simulations, overall species diversity levels in the
+ingrowth were well within the observed range for half of the models (7
+out of 14). Thus, while most of the models are facing difficulties to
 quantitatively match ingrowth levels (cf. above), their performance is
@@
 even decreased ingrowth diversity between the 7 and 10 cm thresholds,
-and the same is visible from the empirical data, but the differences are
-not significant. Diversity variations may be explained by the way the
-individual models consider establishment processes [@konig_tree_2022],
-e.g., in the absence of adults, ingrowth can be linked to either seed
-rain-dispersal, or to a seed/seedling bank.
-
-The species diversity of the entire stand was better captured by the
+and the same is visible from the empirical data, but the differences
+were not significant. Diversity variations may be explained by the way
+the individual models consider establishment processes
+[@konig_tree_2022], e.g., in the absence of adults, ingrowth can be
+linked to either seed rain from dispersal or a seed/seedling bank;
+neither process is incororated across many models studied here, however.
+
+The species diversity of the entire stand was captured better by the
 models than the diversity of tree ingrowth, and only a few models
@@
 simplistic. Furthermore, based on the simulation results, there is no
-evidence that models with feedback captured better the ingrowth species
-diversity than those without feedback with the exception of ForCEEPS
-that significantly improved the ingrowth species diversity with the
-model version that included feedback (ForCEEPS(f)). The similar
-performance of models with and without feedback is likely because the
-models put more weight to the regeneration niche arising from abiotic
-and biotic filters, than from the habitat niche of the adult trees
-[@grubb1977].
+evidence that models with feedback from the canopy captured better the
+species diversity of the ingrowth than those without feedback, with the
+exception of ForCEEPS that significantly improved ingrowth species
+diversity with the model version that included feedback (ForCEEPS(f)).
+The similar performance of models with and without feedback is likely
+because the models put more weight on the regeneration niche arising
+from abiotic and biotic filters, than from the habitat niche of the
+adult trees [@grubb1977].
+
+## Ingrowth mortality
 
 There are several factors that lead to mortality during the
-establishment phase such as competition [@casper1997] and multiple
+establishment phase, such as competition [@casper1997] and multiple
 abiotic factors [@schmid2021; @cunningham2006]. As mentioned above, it
-is reasonable to expect that models that overestimate ingrowth of small
-trees will already have a particularly high mortality between the 7 and
-10 cm thresholds. However, this was not consistently evident from the
+is reasonable to expect that models that overestimate ingrowth will
+already have a particularly high mortality between the 7 and 10 cm
+thresholds. However, this was not consistently evident from the
 simulations. This implies that these models must have the excess of
-mortality in earlier or later stages, if we assume that all models would
-be able to capture the structure and composition of adult stands along
-the EuFoRIa gradient, which however was not directly tested here.
+mortality in later stages, if we assume that all models would be able to
+capture the structure and composition of adult stands along the EuFoRIa
+gradient, which however was not directly tested here.
 
@@
 one hand, mortality in later stages may erroneously enhance the share of
-less shade tolerant species in the models [@klopcic2015], if available,
-therefore shifting the species composition. Unrealistic high stem
-densities for a longer period of time may overestimate the role of tree
-regrowth in carbon sink dynamics [@pugh2019]. Given our set of
-simulations and analyses, we cannot conclusively assess what is
-happening in the models, and further studies are required that focus on
-a wider range of tree sizes and the fate of tree establishment along
-such a size continuum.
-
-## Model traits did not affect model performance
-
-In spite of this criticism, it is remarkable that most models did not
-exceedingly deviate from respect to simulated ingrowth levels -- after
-all, few if any of the models' establishment routines are sufficiently
-constrained by data, with the exception of the empirical models xComp
-and SIBYLA. It is noteworthy that in spite of their empirical basis,
-both models did not match empirical ingrowth levels, in similar
-magnitude as the other models and even in opposite directions (SIBYLA:
-underestimation; xComp: overestimation). It appears that using empirical
-data of limited geographical scope and constrained to managed forests,
-as done in either of these models, leads to extrapolation problems
-already under current climatic conditions. For the other models, whose
-establishment module is not strongly constrained by empirical data, the
-degrees of freedom that are available for modeling ingrowth
-[@konig_tree_2022] could have implied that model performance would be
-much worse.
-
-The results from this study showed that increasing complexity in the
-establishment modules did not represent an improvement in the
-projections of ingrowth levels, species composition or mortality at
-early tree stages. Increasing complexity in establishment modules has
-been motivated by better process understanding and enhancing of model
-accuracy [@bugmann2022; @bugmann2019]. However, more complex models do
-not necessarily lead to better projections and rather to a reduced
+less shade tolerant species in the models [@klopcic2015], therefore
+shifting the species composition. Unrealistic high stem densities for a
+longer period of time may overestimate the role of tree regrowth in
+carbon sink dynamics [@pugh2019]. Given our set of simulations and
+analyses, we cannot conclusively assess what is happening in the models,
+and further studies are required that focus on a wider range of tree
+sizes and the fate of tree establishment along such a size continuum.
+
+## Model performance and model traits
+
+In spite of the critical considerations above, it is remarkable that
+most models did not deviate exceedingly with respect to simulated
+ingrowth levels -- after all, few if any of the models' establishment
+routines are sufficiently constrained by data, with the exception of the
+empirical models xComp and SIBYLA. It is noteworthy that in spite of
+their empirical basis, these two models did not match empirical ingrowth
+levels, in a similar magnitude as the other models and even in opposite
+directions (SIBYLA: underestimation; xComp: overestimation). It appears
+that using empirical data of limited geographical scope and constrained
+to managed forests, as done in these models, leads to extrapolation
+problems already under current climatic conditions. For the other
+models, whose establishment module is not strongly constrained by
+empirical data, the degrees of freedom that are available for modeling
+ingrowth [@konig_tree_2022] could have implied that model performance
+would be much worse than what we found.
+
+Our study showed that increasing complexity in the establishment modules
+did not represent an improvement in the projections of ingrowth levels,
+species composition or mortality at early tree stages, as there was no
+significant relationship between model performance and model complexity.
+Increasing complexity in establishment modules has been motivated by
+better process understanding and enhancing of model accuracy
+[@bugmann2022; @bugmann2019]. However, more complex models do not
+necessarily lead to better projections and rather to a reduced
 transparency and predictive power [@franklin2020]. Thus, the question
-what level of detail one has to use to model establishment processes
-remains open.[@konig_tree_2022; @bugmann2022].
+what level of detail is appropriate and parsimonious for modeling tree
+establishment processes remains open [@konig_tree_2022; @bugmann2022].
 
@@
 Competition for light as a strong filter for tree establishment has been
-widely documented [@messier1999; @collet2006; @berdanier2016], but
-models did not reproduce this expectation. However, it is difficult to
-measure light availability at large spatial and temporal scales. We used
-total stand basal area as a proxy for light availability [cf.
-@schmid2021]. We found pronounced differences in the stand basal area
-ranges simulated by the models, but also between the models and the
-observations. This made not possible to evaluate the establishment for
+widely documented [@messier1999; @collet2006; @berdanier2016], but the
+models examined here did not reproduce this expectation. However, it is
+difficult to measure light availability at large spatial and temporal
+scales. We used total stand basal area as a proxy for light availability
+[cf. @schmid2021]. We found pronounced differences in the stand basal
+area ranges simulated by the models, but also between the models and the
+observations. This made it impossible to evaluate the establishment for
 the extremes of the stand density ranges in some models. For example,
 ingrowth levels at low stand densities are relevant to assess how well
-forests are recovering e.g., after gap creation due to disturbance
+forests are recovering e.g. after gap creation due to disturbance
 [@seidl2022; @grubb1977]. At the level of the simulated one-hectare
@@
 simulation protocol. Towards the other end of the spectrum, i.e., with
-increasing basal area, it would be reasonable to expect that with
-increasing stand basal area, the ingrowth of the different species would
-become sparser and drop out entirely under low-light conditions
-[@klopcic2012; @zell2019]. However, few models showed this trend, thus
-indicating that the relation between ingrowth and light availability is
-not yet captured correctly in most models. On the other hand several of
-the models with an increase of ingrowth with increasing basal area also
-include the feedback between seed production of mature trees and
-ingrowth. Thus it seems that the light competition does not sufficiently
-compensate the increased seed availability with higher basal area.
-
-There were pronounced variations how the main tree species were
-represented along the environmental gradients, in particular the
-dominant species Abies alba, Fagus sylvatica and Picea abies. The
-ingrowth levels were sampled from simulations in the equilibrium, and in
-this state it is expected that non-climax species such as *Pinus
-sylvestris* or *Quercus* would be of minor importance, or absent
-[@klopcic2015]. Most models captured this low abundance, which is also
-found in the empirical data. Thus, although the broad patterns are
+increasing stand basal area, it would be reasonable to expect that the
+ingrowth of the different species would become sparser and drop out
+entirely under low-light conditions [@klopcic2012; @zell2019]. However,
+few models showed this trend, thus indicating that the relation between
+ingrowth and light availability is not yet captured correctly in most
+models. Yet, several of the models that did feature an increase of
+ingrowth with increasing basal area include a feedback between seed
+production of mature trees and ingrowth. Thus, it seems that higher
+light competition does not sufficiently compensate for increased seed
+availability with higher basal area due to more mature trees.
+
+There were pronounced differences how the main tree species were
+represented by the models along the environmental gradients, in
+particular the dominant species Abies alba, Fagus sylvatica and Picea
+abies. The ingrowth levels were sampled from simulations in the
+equilibrium, and in this state it is expected that non-climax species
+such as *Pinus sylvestris* or *Quercus* would be of minor importance, or
+absent [@klopcic2015]. Most models captured this low abundance, which is
+also found in the empirical data. Thus, although the broad patterns are
 matched by many models, improvements in the quantification of the
@@
 
-The EuFoRIa data as used in this study are unique and well suited to
-better understand tree establishment. However, three aspects of these
-data may also represent considerable limitations. First, we made a
-comparison of tree ingrowth in a equilibrium state, but we cannot assess
-how close the forests included in the EuFoRIa dataset were to such a
-state. The data were collected in forest reserves where no management
-has taken place for long periods of time. This makes our assumption of
-an equilibrium between forest properties and environmental drivers more
+The EuFoRIa data as used here are unique and well suited to better
+understand tree establishment. However, three aspects of these data may
+also represent considerable limitations. First, we made a comparison of
+tree ingrowth in an equilibrium state, but we cannot assess how close
+the forests included in the EuFoRIa dataset were to such a state. The
+data were collected in forest reserves where no management has taken
+place for long periods of time. This makes our assumption of an
+equilibrium between forest properties and environmental drivers more
 reasonable than it might appear at first sight. However, the equilibrium
 assumption may explain why simulated ingrowth levels were higher than
-the observed ones. Second, the data were collected for ingrowth above a
-7 cm threshold, thus limiting the assessment of tree establishment to a
-holistic viewpoint. This constituted a hard limit based on which we can
-understand some aspects of tree establishment, which in its entirety
-often comprises a rather long period since seed production [@price2001].
-However, many environmental constraints are acting on young trees
-[@käber2021] that we were unable to assess. Yet, ingrowth data with
-lower thresholds are simply not available in a harmonized manner across
-large environmental gradients. Third, the empirical data were collected
-from rather small plots, while we sampled simulated ingrowth levels from
-1 ha areas, which may lead to an incorrect representation of space. Even
-though this strategy is not ideal, it represents a common challenge when
-harmonizing diverse data sources originating from varied sampling
-strategies [@portier2022]. It would have been extremely challenging for
-such a variety of models to follow a protocol where the spatial sampling
-size was different at each of the 200 sites and it would have introduce
-additional uncertainty into the results. We can not forget here that the
-observed data is just a snapshot of an extremely complex and stochastic
-reality, which might be the explanation of e.g. lack of clear patterns
-when confronted with gradients of climate water balance and degree days,
-while the models outputs are the realizations of the definitions of the
-regeneration niche.
+the observed ones [@brzeziecki2021]. Second, the data were collected for
+ingrowth above a 7 cm threshold, thus limiting the assessment of tree
+establishment to a holistic viewpoint. This constituted a hard limit
+based on which we can understand only some aspects of tree
+establishment, which in its entirety often comprises a rather long
+period since seed production [@price2001]. In reality, many
+environmental constraints are acting on young trees [@käber2021] that we
+were unable to assess. Yet, ingrowth data with lower thresholds are
+simply not available in a harmonized manner across large environmental
+gradients. Third, the empirical data were collected from rather small
+plots, while we sampled simulated ingrowth levels from 1 ha areas, which
+may lead to an incorrect representation of space. Even though this
+strategy is not ideal, it represents a common challenge when harmonizing
+diverse data sources originating from varied sampling strategies
+[@portier2022]. It would have been extremely challenging for such a
+variety of models to follow a protocol where the spatial sampling size
+was different at each of the 200 sites, and it would have introduced
+additional uncertainty in the results. We must not ignore that the
+observed data represent just a snapshot of an extremely complex and
+stochastic reality. This, may at least parlty explain e.g. the lack of
+clear patterns in the simulation results along the gradients of climate
+water balance and degree-days.
 
 The design of our sampling protocol did not include spatial aspects such
-as seed migration nor detailed soil data. We considered a wide range of
-models of forest dynamics, from stand to global scale. However, the
-simulation setup was limited to 1 ha ingrowth values. The lack of
-spatial considerations potentially puts landscape models at a
-disadvantage, as they have been built to be accurate at the landscape
-level. Without spatial context, we are limiting ingrowth to the seed
-influx from the stand itself, if the model does not have a background
-seed input. However, Global Dynamic Vegetation Models should not be in
-disadvantage due to the limited spatial considerations [but cf. Snell et
-al. -@snell2014] as they are usually lacking dispersal between cells as
-they are based on an abstraction of horizontal space, or even the trees
-themselves [@hanbury-brown2022]. Detailed data on soil conditions were
-not available from the observed data and indepentdent data for soil and
-climatic water balance were used instead. Several models, however, used
-more detailed drought stress indicators based e.g., on soil water
-holding capacity that had to be derived from a rough soil quality
-measure which could explain the poor performance of the models against
-climatic water balance.
+as seed dispersal or detailed soil data. Although, we considered a wide
+range of models of forest dynamics, from stand to global scales, the
+simulation setup was limited to 1 ha ingrowth values. This lack of
+spatial considerations is appropriate for stand-scale models, but it
+potentially puts landscape models at a disadvantage, as they have been
+built to be accurate at the landscape level. Without spatial context, we
+are limiting ingrowth to the seed influx from the stand itself, unless
+the model has a background seed input. However, global models should not
+be at a disadvantage due to the limited spatial considerations [but cf.
+Snell et al. -@snell2014], as they are usually lacking dispersal between
+cells and are based on a strong abstraction of horizontal space
+[@hanbury-brown2022]. Lastly, detailed data on soil conditions were not
+available from the observed data, and independent, admittedly rather
+coarse data for soil and climatic water balance were used instead. It is
+noteworthy that several models represent drought based on more detailed
+indicators based e.g., on soil water holding capacity, which had to be
+derived from a rough soil quality measure. This may at least partially
+explain the poor performance of the models against the drought axis
+(i.e., climatic water balance).
 
@@
 Based on the results of this study, it is clear that models of forest
-dynamics need a focus on their establishment modules. As shown in the
-results of the study it is still uncertain what level of detail one has
-to use to model establishment processes and must be addressed in future
-research. We recommend that the further development of the establishment
-modules is implemented as additional features that can be traced back,
-as done in this study for the model variants of ForClim and ForCEEPS,
-and that model complexity and structure must always be connected with
-model objectives [@albrich2020]. If it should be necessary to include
-more detail in the establishment models, this will come with higher
-parameterisation efforts, which will most likely lead to lower
-generalization as the required data will have to be collected from
-specific locations, as currently there is no general establishment
-dataset available.
-
-More effort should be invested into collecting harmonised datasets on
-tree establishment. We emphasize that datasets like EuFoRIa used here
-are invaluable and should be further expanded in both their spatial
-extent (e.g., towards boreal and Mediterranean conditions) as well as in
-time (e.g., continuing the monitoring into the future). Such a data will
-allow a better evaluation of forest models and help to reduce the
-uncertainty in their projections, which is crucial when they are used as
-tools for predicting potential impacts of climate change.
-
-Here we have only considered tree establishment in the equilibrium
-state. It is equally important to understand how these models project
-tree establishment after changes in forest structure by disturbances
-[@seidl2022], or under different management strategies [@lindner2000].
-However, this will require an entirely different set of observed data,
-and potentially not all models of forest dynamics would be able to
-assess the relationship of these aspects on tree establishment, e.g.,
-due to the lack of disturbance or appropriate management modules.
+dynamics need a focus on tree establishment modules to complement
+earlier efforts to make them more robust regarding growth and mortality.
+It remain uncertain what level of detail is required to model
+establishment processes, and this must be addressed in future research.
+We recommend that the further development of the establishment modules
+is implemented as additional features that can be traced back, as done
+here for the variants of ForClim and ForCEEPS, and that model complexity
+and structure must always be connected with modeling objectives
+[@albrich2020]. If it should be necessary to include more detail in the
+establishment models, this will come with higher parameterisation
+efforts. This will most likely lead to lower generalization because the
+required data will have to be collected from specific locations, as
+currently there is no general establishment dataset available.
+
+Therefore, we recommend that more effort should be invested into
+collecting harmonised datasets on tree establishment. We emphasize that
+datasets like EuFoRIa used here are invaluable and should be further
+expanded in both their spatial extent (e.g., towards boreal and
+Mediterranean conditions) as well as in time (e.g., continuing the
+monitoring into the future). Such data will allow for a better
+evaluation of forest models and help to reduce the uncertainty in their
+projections, which is crucial when they are used as tools for predicting
+e.g. the impacts of anthropogenic climate change.
+
+In the present study, we have considered tree establishment in the
+equilibrium state only. It is equally important to understand how these
+models project tree establishment after changes in forest structure by
+disturbances [@seidl2022], or under different management strategies
+[@lindner2000]. However, this will require an entirely different set of
+observed data, and potentially not all models of forest dynamics would
+be able to assess the relationship of these aspects on tree
+establishment, e.g., due to the lack of disturbance or appropriate
+management modules.
 
 We recommend to further investigate the implications of the current
-modeling strategies for tree establishment, particularly those related
-to excess tree establishment and late excess mortality, by focusing on a
-wider range of tree sizes and the related ingrowth dynamics. We also
-recommend, especially for the landscape-level models, the inclusion of
-explicit spatial considerations regarding tree ingrowth [@beckage2003];
-this, however, is facing a serious challenge regarding appropriate
-datasets, as inventory data have wide coverage, but do not allow for the
-assessment of spatial interaction effects.
-
-Exercises like the one presented here where the models are run in "blind
-flight" without the possibility of being tuned towards capturing the
-observed values should be repeated. Such a benchmarking exercises should
-next focus in different aspects such as specific model traits and
-ecological formulations to better understand the implications of the
-model assumptions. Furthermore, the inclusion in this exercises with
-different scales, type, populations structures, tree establishment
-modules and complexity ensure a wider benefit to the whole modeling
+modeling strategies for tree establishment and, ultimately, simulated
+forest stand structure. This applies particularly to the erroneous
+patterns of excess tree establishment and later excess mortality, by
+focusing on a wider range of tree sizes and the related ingrowth
+dynamics. We also recommend, especially for the landscape-level models,
+the inclusion of explicit spatial considerations regarding tree ingrowth
+[@beckage2003]; this, however, is facing a serious challenge regarding
+appropriate datasets, as inventory data have wide coverage, but by
+definition do not allow for the assessment of spatial interaction
+effects.
+
+Exercises like the one presented here, where the models are operated in
+"blind flight" mode, i.e. without the possibility of being tuned towards
+capturing the expected patterns, should be repeated. Such benchmarking
+exercises should next focus on aspects such as specific model traits and
+the ecological formulations of particular (sub-)processes, so as to
+better understand the implications of the assumptions on which the
+models are based. Furthermore, the inclusion of a wide range of models
+with different scales, type, populations structures, tree establishment
+modules and complexity will ensure a wider benefit to the whole modeling
 community.
@@
 Models of forest dynamics are important tools in science and decision
-support, and how tree establishment is defined has strong implications
+support. The formulation of tree establishment has strong implications
 for simulated forest properties. The 15 models and variants used here
-are facing similar challenges when representing tree establishment. They
-generally overestimate tree establishment levels, and the simulated
-establishment niche is not always captured accurately as a function of
-the biotic (light) and abiotic (temperature and moisture) factors.
+are facing similar challenges in their representation of tree
+establishment: they generally overestimate tree establishment levels,
+and the simulated establishment niche is not always captured accurately
+as a function of biotic (light) and abiotic (temperature and moisture)
+factors.
 
@@
 
-This study compiles with the basic rules of reproducibility and has make
+This study complies with the basic rules of reproducibility and has made
 available the protocol of the study, the data and all the software that
-leads of the findings presented in this study. The protocol is available
-here XXX. The code and data that support the findings of this study are
-openly available at zenodo and gihub (links to be added)
+has led to the findings presented in this paper. The protocol is
+available here:( link to be added). The code and data that support the
+findings of this study are openly available at zenodo and gihub (links
+to be added).
 
@@
 
-Funding of the workshop was provided by the COST Action CA19139 PROCLIAS
+This paper arises from a workshop held in Davos (Switzerland) in June
+2022, co-led by Olalla Díaz-Yñaez, Yannek Käber and Harald Bugmann.
+Funding for the workshop was provided by COST Action CA19139 PROCLIAS
 (PROcess-based models for CLimate Impact Attribution across Sectors),
@@
 Technology; [https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich
-(Grant ETH-35 18-1). ODY was funded by SNF project XXXX. KB received
-support from the European Research Council under the European Union's
-Horizon 2020 research and innovation program (Grant Agreement
-101001905). JB acknowledges funding from the long-term research
-development project no. RVO 67985939 from the Czech Academy of Sciences.
-MM acknowledges financial support from I-Maestro (Innovative forest
-management strategies for a resilient bioeconomy under climate change
-and disturbances, grant no. 22035418, 2019--2022) funded by the ERA-NET
-Cofund ForestValue. KM received funding from the Project No.
-CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant no.
-ITMS2014+313011W580 supported by the Integrated Infrastructure
-Operational Programme funded by the ERDF
+(Research Grant ETH-35 18-1). ODY was funded by the Swiss National
+Science Foundation project . KB received support from the European
+Research Council under the European Union's Horizon 2020 research and
+innovation program (Grant Agreement 101001905). JB acknowledges funding
+from the long-term research development project no. RVO 67985939 from
+the Czech Academy of Sciences. MM acknowledges financial support from
+I-Maestro (Innovative forest management strategies for a resilient
+bioeconomy under climate change and disturbances, grant no. 22035418,
+2019--2022) funded by the ERA-NET Cofund ForestValue. KM received
+funding from the Project No. CZ.02.1.01/0.0/0.0/16_019/0000803 financed
+by OP RDE, grant no. ITMS2014+313011W580 supported by the Integrated
+Infrastructure Operational Programme funded by the ERDF
 

### 2023-03-17 code/figures.R

@@ -446,7 +446,7 @@
     
     
     xx <- ggpubr::ggarrange(sm, sl,
-                            labels = c("A", "B"),
+                            labels = c("a", "b"),
                             ncol = 1, nrow = 2, heights = c(3, 3), align = "v")
     
     xx <- ggpubr::annotate_figure(xx, 
@@ -741,10 +741,10 @@
                                              color = NULL),
                                 nudge_y = 0.05) +
       ggplot2::scale_fill_manual(values = values_color) +
-      ggplot2::annotate("text", x = 10, y = 2.6, label = "A") + 
-      ggplot2::annotate("text", x = 10, y = 1.6, label = "B") + 
-      ggplot2::annotate("text", x = -4.6, y = 1.6, label = "C") + 
-      ggplot2::annotate("text", x = -4.6, y = 2.6, label = "D") + 
+      ggplot2::annotate("text", x = 10, y = 2.6, label = "a") + 
+      ggplot2::annotate("text", x = 10, y = 1.6, label = "b") + 
+      ggplot2::annotate("text", x = -4.6, y = 1.6, label = "c") + 
+      ggplot2::annotate("text", x = -4.6, y = 2.6, label = "d") + 
     
       ggplot2::annotate("text", x = 8.5, y = 1.81, label = "Reineke self-thinning") + 
       ggplot2::annotate("text", x = 7.9, y = 1.03, label = "Ratio equal to 1")

### 2023-03-30 manuscript/manuscriptRW.Rmd
changes added during the meeting with Harald

@@
 abiotic factors [@schmid2021; @cunningham2006]. As mentioned above, it
-is reasonable to expect that models that overestimate ingrowth will
-already have a particularly high mortality between the 7 and 10 cm
-thresholds. However, this was not consistently evident from the
-simulations. This implies that these models must have the excess of
-mortality in later stages, if we assume that all models would be able to
-capture the structure and composition of adult stands along the EuFoRIa
-gradient, which however was not directly tested here.
+is reasonable to expect that models that overestimate ingrowth may have
+a particularly high mortality between the 7 and 10 cm thresholds.
+However, this was not consistently evident from the simulations. This
+implies that these models must have an excess mortality in later stages,
+if we assume that all models were able to capture the structure and
+composition of the adult stands along the EuFoRIa gradient; this however
+was not tested here.
 
@@
 equilibrium between forest properties and environmental drivers more
-reasonable than it might appear at first sight. However, the equilibrium
-assumption may explain why simulated ingrowth levels were higher than
-the observed ones [@brzeziecki2021]. Second, the data were collected for
-ingrowth above a 7 cm threshold, thus limiting the assessment of tree
-establishment to a holistic viewpoint. This constituted a hard limit
-based on which we can understand only some aspects of tree
-establishment, which in its entirety often comprises a rather long
-period since seed production [@price2001]. In reality, many
-environmental constraints are acting on young trees [@käber2021] that we
-were unable to assess. Yet, ingrowth data with lower thresholds are
-simply not available in a harmonized manner across large environmental
-gradients. Third, the empirical data were collected from rather small
-plots, while we sampled simulated ingrowth levels from 1 ha areas, which
-may lead to an incorrect representation of space. Even though this
-strategy is not ideal, it represents a common challenge when harmonizing
-diverse data sources originating from varied sampling strategies
-[@portier2022]. It would have been extremely challenging for such a
-variety of models to follow a protocol where the spatial sampling size
-was different at each of the 200 sites, and it would have introduced
-additional uncertainty in the results. We must not ignore that the
-observed data represent just a snapshot of an extremely complex and
-stochastic reality. This, may at least parlty explain e.g. the lack of
-clear patterns in the simulation results along the gradients of climate
-water balance and degree-days.
+reasonable than it might appear at first sight. In an analysis of
+primeval forests at demographic equilibrium, @brzeziecki2021, they found
+higher ingrowth rates than the ones observed in the Euforia data. Thus,
+the overestimate of ingrowth rates by the models may not be so
+problematic !!! Second, the data were collected for ingrowth above a 7
+cm threshold, thus limiting the assessment of tree establishment to a
+holistic viewpoint. This constituted a hard limit based on which we can
+understand only some aspects of tree establishment, which in its
+entirety often comprises a rather long period since seed production
+[@price2001]. In reality, many environmental constraints are acting on
+young trees [@käber2021] that we were unable to assess. Yet, ingrowth
+data with lower thresholds are simply not available in a harmonized
+manner across large environmental gradients. Third, the empirical data
+were collected from rather small plots, while we sampled simulated
+ingrowth levels from 1 ha areas, which may lead to an incorrect
+representation of space. Even though this strategy is not ideal, it
+represents a common challenge when harmonizing diverse data sources
+originating from varied sampling strategies [@portier2022]. It would
+have been extremely challenging for such a variety of models to follow a
+protocol where the spatial sampling size was different at each of the
+200 sites, and it would have introduced additional uncertainty in the
+results. We must not ignore that the observed data represent just a
+snapshot of an extremely complex and stochastic reality. This, may at
+least parlty explain e.g. the lack of clear patterns in the simulation
+results along the gradients of climate water balance and degree-days.
 
@@
 This paper arises from a workshop held in Davos (Switzerland) in June
-2022, co-led by Olalla Díaz-Yñaez, Yannek Käber and Harald Bugmann.
+2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
 Funding for the workshop was provided by COST Action CA19139 PROCLIAS
@@
 (Research Grant ETH-35 18-1). ODY was funded by the Swiss National
-Science Foundation project . KB received support from the European
+Science Foundation project XXXX . KB received support from the European
 Research Council under the European Union's Horizon 2020 research and
@@
 by OP RDE, grant no. ITMS2014+313011W580 supported by the Integrated
-Infrastructure Operational Programme funded by the ERDF
+Infrastructure Operational Programme funded by the ERDF.
 

### 2023-03-31 code/figures.R
zoom to map figure

@@ -1530,7 +1530,7 @@
     # plotsMap <- ggplot(Europe) +
     #   ggthemes::theme_map() +
     #   geom_sf(fill = "transparent" , colour = "lightgrey",  lwd = 0.25) +
-    #   coord_sf(xlim = c(-25,45), ylim = c(40,55), expand = FALSE) +
+    #   coord_sf(xlim = c(-15,30), ylim = c(40,55), expand = FALSE) +
     #   geom_point(data = domSpp10, shape = 20, stroke = FALSE,
     #              mapping = aes(x = X, y = Y, color = recruitment),
     #              alpha = 1, size = 2) +
@@ -1552,4 +1552,3 @@
     #                 width = 21, height = 12, scale = 0.9,
     #                 dpi = 300, units = "cm", device = 'png')
 
-    

### 2023-03-31 code/figures.R
map figure zoom updates

@@ -1533,7 +1533,7 @@
     #   coord_sf(xlim = c(-15,30), ylim = c(40,55), expand = FALSE) +
     #   geom_point(data = domSpp10, shape = 20, stroke = FALSE,
     #              mapping = aes(x = X, y = Y, color = recruitment),
-    #              alpha = 1, size = 2) +
+    #              alpha = 1, size = 2.5) +
     #   scale_color_manual("Dominant species of ingrowth",
     #                      values = c("Abies alba" =  "#1B9E77",
     #                                 "Picea abies" ="#D95F02" ,

### 2023-03-31 manuscript/manuscriptRW.Rmd
remove word counting from manuscript
update protocol reference in the manuscript

@@
 
-    Word count: `r as.integer(sub("(\\d+).+$", "\\1", system(sprintf("wc -w %s", knitr::current_input()), intern = TRUE))) - 20`
-
 # Abstract
@@
 threshold of 7 or 10 cm, respectively (synonym: ingrowth). To this end,
-each modeling group was provided with a detailed protocol (cf.
-Supplementary Material 1, henceforth "SM") with instructions how to
-perform the simulations, providing input variables on climate and soil
-conditions, and the expected output variables. Neither were further site
-information (except for the data specified below) nor any data on tree
-ingrowth or forest stand features provided prior to the simulation. That
-is, the models were run in "blind flight" mode.
+each modeling group was provided with a detailed protocol [@díaz-yáñez]
+with instructions how to perform the simulations, providing input
+variables on climate and soil conditions, and the expected output
+variables. Neither were further site information (except for the data
+specified below) nor any data on tree ingrowth or forest stand features
+provided prior to the simulation. That is, the models were run in "blind
+flight" mode.
 
@@
 \@ref(tab:tableModels)). Further details on how each modeling team
-prepared the simulations and the outputs are available in SM 2.
+prepared the simulations and the outputs are available in Supplementary
+Material 1.
 
@@
 (i.e., ingrowth number per unit time and space), (2) species diversity
-of the ingrowth, and (3) ingrowth mortality. We evaluated the species
-diversity across models and in relation to the observed data by
-calculating the Shannon index based on the relative proportion of the
+of the ingrowth, (3) ingrowth mortality, (4) model performance and model
+traits, and (5) ingrowth gradients on regeneration niches. We evaluated
+the species diversity across models and in relation to the observed data
+by calculating the Shannon index based on the relative proportion of the
 species in terms of basal area. The Shannon index was calculated for the
 establishment ($H_{R_{n}}$) (Equation \ref{eqpiShannonR}) and at the
-stand level ($H_{S_{n,s}}$)(Equation \ref{eqpiShannon}). The higher the
+stand level ($H_{S_{n,s}}$) (Equation \ref{eqpiShannon}). The higher the
 value of the index, the higher the species diversity at a particular
@@
 
-## Regeneration niche
+## Ingrowth gradients on regeneration niches
 
@@
 the overestimate of ingrowth rates by the models may not be so
-problematic !!! Second, the data were collected for ingrowth above a 7
-cm threshold, thus limiting the assessment of tree establishment to a
+problematic. Second, the data were collected for ingrowth above a 7 cm
+threshold, thus limiting the assessment of tree establishment to a
 holistic viewpoint. This constituted a hard limit based on which we can
@@
 snapshot of an extremely complex and stochastic reality. This, may at
-least parlty explain e.g. the lack of clear patterns in the simulation
+least partly explain e.g. the lack of clear patterns in the simulation
 results along the gradients of climate water balance and degree-days.
@@
 Therefore, we recommend that more effort should be invested into
-collecting harmonised datasets on tree establishment. We emphasize that
+collecting harmonized datasets on tree establishment. We emphasize that
 datasets like EuFoRIa used here are invaluable and should be further
@@
 has led to the findings presented in this paper. The protocol is
-available here:( link to be added). The code and data that support the
-findings of this study are openly available at zenodo and gihub (links
-to be added).
+available here @díaz-yáñez . The code and data that support the findings
+of this study are openly available at zenodo (*link to be added*) and
+gihub (*link to be added*).
 
@@
 (Research Grant ETH-35 18-1). ODY was funded by the Swiss National
-Science Foundation project XXXX . KB received support from the European
-Research Council under the European Union's Horizon 2020 research and
-innovation program (Grant Agreement 101001905). JB acknowledges funding
-from the long-term research development project no. RVO 67985939 from
-the Czech Academy of Sciences. MM acknowledges financial support from
-I-Maestro (Innovative forest management strategies for a resilient
-bioeconomy under climate change and disturbances, grant no. 22035418,
-2019--2022) funded by the ERA-NET Cofund ForestValue. KM received
-funding from the Project No. CZ.02.1.01/0.0/0.0/16_019/0000803 financed
-by OP RDE, grant no. ITMS2014+313011W580 supported by the Integrated
-Infrastructure Operational Programme funded by the ERDF.
+Science Foundation project 200020_188882 / 1 . KB received support from
+the European Research Council under the European Union's Horizon 2020
+research and innovation program (Grant Agreement 101001905). JB
+acknowledges funding from the long-term research development project no.
+RVO 67985939 from the Czech Academy of Sciences. MM acknowledges
+financial support from I-Maestro (Innovative forest management
+strategies for a resilient bioeconomy under climate change and
+disturbances, grant no. 22035418, 2019--2022) funded by the ERA-NET
+Cofund ForestValue. KM received funding from the Project No.
+CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant no.
+ITMS2014+313011W580 supported by the Integrated Infrastructure
+Operational Programme funded by the ERDF.
 

### 2023-03-31 manuscript/supplementaryRW.Rmd
update supplementary

@@
 
-# Supplementary Material 1: Protocol
-
-## General steps
-
-Each modeling team will run their simulations prior to the workshop, and
-will report on the results before ([short written report](#report)) and
-during the workshop. The general approach for the workshop is:
-
-1.  Each team sets up their model to simulate the regeneration niche of
-    [eleven major European tree species](#speciescomposition) within the
-    known, defined range of environmental conditions of 200 sites. They
-    will run at least one simulation per site. It is important to
-    conduct mixed-species simulations, not series of monospecific
-    stands.
-
-2.  The model is set up for each site based on the [input variable
-    information](#inputvariables) provided in this protocol, and the
-    simulations need to be run until the [equilibrium is
-    reached](#equilibrium).
-
-3.  Each team reports on the results of these exercises in terms of a
-    [standardized output dataset](#outputdata) (including density
-    variables per 10-yr interval and per ha, for each simulated site and
-    species) and in the form of a brief [working report](#report) (\~4
-    pages).
-
-4.  At the workshop in June 2022 in Davos, Switzerland, we will evaluate
-    the regeneration niche as it results from each model and compare it
-    to the empirical estimates of the environmental limits of the
-    regeneration niche that will be made available to the participants
-    shortly before the workshop. As a group effort, the projections of
-    the niche properties arising from the different models will be
-    compared. The exact details on how this will be done will be
-    provided once we know how many modeling teams will contribute
-    results.
-
-## Input variables {#inputvariables}
-
-In this section, we are providing a description of all the necessary
-driver information to run the simulations (e.g., location, climate,
-soil).
-
-### Coordinates
-
-We intend to provide the site information via unique site IDs. If actual
-coordinates are needed, we can provide blurred site coordinates upon
-reasonable request.
-
-### Climate
-
-We provide time series of climatic variables from 1981 to 2018 in
-hourly, daily and monthly resolution. Based on our survey the climate
-data should be sufficient to run all models. Some variables such as
-relative humidity or vapor pressure deficit must be calculated from the
-variables we provide.
-
-Note that the hourly time series has the lowest spatial resolution, thus
-the daily time series should be used for spatial accuracy. If you need
-hourly time series we recommend to use the daily time series and only
-use the hourly time series to correct for hourly variation. For this
-purpose we also provide hourly averages over the period 1981 to 2018 for
-each hour of each day of the year. Only for calculating these averages
-the 29th of February in leap years was removed from the time series to
-obtain values for 365 days.
-
-#### Data sources {#climsources}
-
-The hourly time series are based on the Era5-land data set. Daily and
-monthly time series are mostly based on CHELSA data which has a higher
-spatial resolution. More details are given below or in the documentation
-of the original data source.
-
-The variable names are kept exactly like in the documentation of the
-Era5 and CHELSA data sets to facilitate the use of the original data
-documentation. Please don't be confused by the ending `*_hourly` in
-`surface_thermal_radiation_downwards_hourly` and the two different
-variable names for shortwave radiation: `rsds` for the values from
-CHELSA and `surface_solar_radiation_downwards_hourly` for the values
-from Era5.
-
-All changes that were made to the original data are documented here:
-
--   The wind variable `wind_speed` is based on the u and v component of
-    the wind and calculated with $\sqrt{u^2+v^2}$ where $u$ is the
-    variable `u_component_of_wind_10m` and $v$ is
-    `v_component_of_wind_10m` of the Era5 data set.
-
--   The units of temperature variables `temperature_2m` (Era5), `tas`,
-    `tasmin`, `tasmax`(all CHELSA) were changed from Kelvin to Degree
-    Celsius
-
--   The units of precipitation variables `total_precipitation_hourly`
-    (Era5), `pr` (CHELSA) were changed from m to mm and kG to mm
-    respectively.
-
--   The unit of short wave radiation `rsds` (CHELSA) was changed from kJ
-    to J.
-
-We aggregated variables from hourly to daily to monthly values by using
-means for temperature, surface pressure and wind speed and sums for
-radiation and precipitation.
-
-##### Era5-Land {#era5}
-
--   Muñoz Sabater, J., (2019): ERA5-Land hourly data from 1981 to
-    present. Copernicus Climate Change Service (C3S) Climate Data Store
-    (CDS). (23.01.2022),
-    [doi:10.24381/cds.e2161bac](https://doi.org/10.24381/cds.e2161bac)
--   [Documentation of the hourly ERA5 data from the Google Earth
-    Engine](https://developers.google.com/earth-engine/datasets/catalog/ECMWF_ERA5_LAND_HOURLY#description)
--   Further information is available from the [official copernicus
-    website](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-land?tab=overview).
-
-##### CHELSA {#chelsa}
-
--   Dirk N. Karger, Stefan Lange, Chantal Hari, Christopher P. O. Reyer,
-    Niklaus E. Zimmermann (2021): CHELSA-W5E5 v1.0: W5E5 v1.0 downscaled
-    with CHELSA v2.0. ISIMIP Repository.
-    <https://doi.org/10.48364/ISIMIP.836809>
-
--   [Official CHELSA website](https://chelsa-climate.org)
-
-#### Monthly time series 1981-2018
-
-file name: monthly_era5_chelsa_200.zip
-
-**Data description**
-
-Variables from the [Era5 data set](#era5) with a spatial resolution of
-0.1° x 0.1° (\~ 10 km):
-
--   average wind speed (`wind_speed`) [m/s]
-
--   average surface pressure (`surface_pressure`) [Pa]
-
--   surface thermal (longwave) radiation
-    (`surface_thermal_radiation_downwards_hourly`) [J/m2]
-
-Variables from the [CHELSA data set](#chelsa) with a spatial resolution
-of 0.008° x 0.008° (\~ 1 km):
-
--   average temperature (`tas`) [°C]
-
--   average max temperature (`tasmax`) [°C]
-
--   average min temperature (`tasmin`) [°C]
-
--   precipitation sum (`pr`) [mm]
-
--   surface solar (shortwave) radiation (`rsds`) [J/m2]
-
-#### Daily time series 1981-2018
-
-file name: daily_era5_chelsa_200.zip
-
-**Data description**
-
-Variables from the [Era5 data set](#era5) with a spatial resolution of
-0.1° x 0.1° (\~ 10 km):
-
--   average wind speed (`wind_speed`) [m/s]
-
--   average surface pressure (`surface_pressure`) [Pa]
-
--   surface thermal (longwave) radiation
-    (`surface_thermal_radiation_downwards_hourly`) [J/m2]
-
-Variables from the [CHELSA data set](#chelsa) with a spatial resolution
-of 0.008° x 0.008° (\~ 1 km):
-
--   average temperature (`tas`) [°C]
-
--   average max temperature (`tasmax`) [°C]
-
--   average min temperature (`tasmin`) [°C]
-
--   precipitation sum (`pr`) [mm]
-
--   surface solar (shortwave) radiation (`rsds`) [J/m2]
-
-#### Hourly time series 1981-2018
-
-file name: hour_era5_200.zip
-
-file name: yhourMean_era5_200.zip
-
-**Data description**
-
-Variables from the [Era5 data set](#era5) with a spatial resolution of
-0.1° x 0.1° (\~ 10 km):
-
--   average temperature (temperature_2m) [°C]
-
--   average wind speed (`wind_speed`) [m/s]
-
--   average surface pressure (`surface_pressure`) [Pa]
-
--   precipitation sum (total_precipitation_hourly) [mm]
-
--   surface solar (shortwave) radiation
-    (surface_solar_radiation_downwards_hourly) [J/m2]
-
--   surface thermal (longwave) radiation
-    (surface_solar_radiation_downwards_hourly) [J/m2]
-
-### Soil
-
-file name: soil_quality_200.csv
-
-**Data description**
-
-The following variables are available:
-
--   Soil quality is given as a continuous value between 1 and 5.
-
-**Data source**
-
--   The soil quality value is based on a random forest that was trained
-    to predict expert knowledge based soil quality on a scale from 1 to
-    5 from the [SoilGrids250 data
-    set](https://www.isric.org/explore/soilgrids) and the [WISE data
-    set](https://www.isric.org/explore/wise-databases). The most
-    influential variables in the random forest were total depth to
-    bedrock and sand content. Thus the variable is a good predictor for
-    plant available water storage capacity.
-
-## Topography and elevation
-
-There was an error in a earlier version of this file. Please use the
-link above to download the corrected file.
-
-This is the old file and should not be used because the aspect column
-contains elevation instead of the aspect.
-
-file name: topo_data_200.csv
-
-**Data description**
-
-The following variables are available:
-
--   elevation [m.a.s.l]
-
--   slope [Degree]
-
--   aspect [Degree]
-
-**Data source**
-
--   Topographic information was derived from the [ASTER DEM
-    V3](https://lpdaac.usgs.gov/products/astgtmv003/).
-
--   Slope and aspect were calculated with QGIS. The specific tiles that
-    were used are documented
-    [here](https://ites-fe.ethz.ch/rws/downloads/15122021/aster_dem_files.txt),
-    and the processing steps in QGIS are documented
-    [here](https://ites-fe.ethz.ch/rws/downloads/15122021/QGIS_command_log.txt).
-
-### Natural disturbances
-
-The simulations must be run in the absence of natural disturbances.
-
-## Management
-
-The simulations must be run in the absence of management.
-
-### Area simulated
-
-From your model outputs, we would like to obtain samples on species
-specific regeneration rates per decade and per ha. The modeling teams
-decide on the simulated area and how they derive these samples. Please
-describe the methodology used in the [output report](#report).
-
-### Simulation duration and equilibrium {#equilibrium}
-
-The simulation should be run from bare ground in the absence of
-management to a simulated [equilibrium](#defequilibrium) ("potential
-natural vegetation") with current climate. The goal is to derive the
-[regeneration niche](#regenniche). The exact length of the simulation
-should be decided by the modeling teams. Please describe the simulation
-length used in your simulations in the [output report](#report).
-
-[**Regeneration niche:**]{#regenniche} The regeneration niche is defined
-as the passing of a diameter threshold of 7 and 10 cm (ingrowth; cf.
-below). Both the width of the regeneration niche (i.e., in environmental
-space) as well as the intensity of the recruitment process (i.e., the
-number of ingrowth trees per area and per unit of time) are of interest.
-
-[**Simulated equilibrium**]{#defequilibrium}: Simulated equilibrium is
-the "Potential Natural Vegetation".
-
-### Species composition {#speciescomposition}
-
-**Mixed species simulations** must be run (not multiple single-species
-simulations).
-
-**Option A: Species level**
-
-The simulation should be run for **mixtures** of the following
-species/genera: *Fagus sylvatica, Picea abies, Abies alba, Carpinus
-betulus, Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus
-excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris.*
-
-Additional species can be included in the simulation if modelers think
-this is important, but they should report only on these 11
-species/genera because the benchmarking data refer to them only. Please
-report the methodology used in the [output report](#report).
-
-**Option B: PFTs**
-
-If species-level simulation is not possible in your model, substitute
-the species by PFTs and use the model's default approach. Please report
-the methodology used in the [output report](#report). Explain as well
-which PFTs represent which species for your model.
-
-## Diameter threshold
-
-The [output data](#outputdata) have to be provided for two diameter
-thresholds: 7 and 10 cm.
-
-## Output
-
-We are expecting to get the number of recruits over a 10-yr interval per
-species and per ha for each of the sites. However, we are also
-interested in the **variation per site** (i.e., we would like to see the
-range of estimates, not only the mean value per site). Each modeling
-team will decide how to sample multiple values per site that will show
-the variation in recruitment.
-
-Specifically, you should provide us with multiple samples (n = 200; this
-number is the same as the number of sites by sheer coincidence) of the
-regeneration rate for spatial units of 1 ha (please do [not]{.underline}
-provide average of multiple 1-ha samples). This can be done in various
-ways, e.g., by sampling simulated data from the same 1-ha plot in the
-equilibrium over time, or by sampling several 1-ha plots from the
-simulated forest at one specific point in time (in the equilibrium); for
-clarification, cf. three sample strategies proposed in [figures
-1](#sample1), [2](#sample2) and [3](#sample3).
-
-If, for some reason, you cannot submit **200 samples** of **1 ha**,
-please **let us know** in the [workshop chat](#chat) so we can discuss
-alternatives. The sampling from the simulations must happen **within**
-the simulated phase of Potential Natural Vegetation (for more
-information, see how [equilibrium](#defequilibrium), Potential [Natural
-Vegetation, and regeneration niche](#regenniche) are defined in this
-protocol).
-
-![Figure 1. In this example there are 200 samples collected at year i at
-equilibrium differentiated spatially with each square representing 1
-ha.](samplingfigure/Slide2.jpeg){#sample1 width="500"}
-
-![Figure 2. In this example, there are 100 spatially differentiated
-samples taken in year i at equilibrium and 100 in year i+10. Each square
-represents 1 ha.](samplingfigure/Slide3.jpeg){#sample2 width="500"}
-
-![Figure 3. In this example there is 1 spatially differentiated sample
-taken in year i at equilibrium and remeasured every 10 years for 2000
-years until i+2000. The square represents 1
-ha.](samplingfigure/Slide4.jpg){#sample3 width="500"}
-
-### Output data {#outputdata}
-
-Please provide the simulation output data using the following structure:
-
--   Each row of your output dataset contains the recruitment of 10-year
-    interval per species and ha for a unique simulation site.
-
--   Each column of your output dataset contains one of these variables:
-
-    -   `sample`: this will identify the different samples per site
-        (1:n, with n $\ge$ 200)
-
-    -   `site`: site identifier (1:200)
-
-    -   `species`: species simulated with the following categories:
-
-        -   fasy (for *Fagus sylvatica*)
-        -   pabi (for *Picea abies*)
-        -   abal (for *Abies alba*)
-        -   cabe (for *Carpinus betulus*)
-        -   tico (for *Tilia cordata*)
-        -   acps (for *Acer pseudoplatanus*)
-        -   betu (for *Betula* spp.)
-        -   frex (for *Fraxinus excelsior*)
-        -   quer (for *Quercus* spp.)
-        -   algl (for *Alnus glutinosa*)
-        -   pisy (for *Pinus sylvestris*)
-        -   If you have some other categories please create your own
-            species codes and provide the legend in the output report.
-            We prefer a lumped category "othr".
-
-    -   `r.trees`: the recruits (trees $ha^{-1}$) per species and
-        decade.
-
-    -   `r.ba`: the recruits' basal area ($m^{2}$ $ha^{-1}$) per species
-        and decade.
-
-    -   `ba`: total basal area ($m^{2}$ $ha^{-1}$) per species and
-        decade
-
-    -   `dbh`: diameter threshold (cm) used to define the recruitment (7
-        or 10 cm).
-
-### Output report {#report}
-
-To guide the writing of the output report, please refer to the following
-questions:
-
--   What climate data have you used, and how have you created your model
-    climate input parameters?
-
--   How have you defined the species composition in your model
-    simulation? (e.g., species levels, PFTs, etc.)? Please describe the
-    species categories used.
-
--   How have you defined the simulation area for your model (e.g., size
-    and other relevant site aspects)?
-
--   What are the relevant aspects of the dispersal (if any) and seed
-    input parameters in your model?
-
--   What is the simulation length, and how have you determined the
-    simulation time until the equilibrium?
-
--   If you have run more than one simulation per site, please describe
-    the differences of the approaches used.
-
--   Please describe how have you calculated each of the variables in the
-    [output data](#outputdata).
-
-### Uploading data and report
-
-Please upload your data and output report as a **zip file** with the
-name of your model and any other word that might help us to understand
-the files, for example: *thenameofyourmodel_datareport.zip* or
-*thenameofyourmodel_data.zip* and *thenameofyourmodel_report.zip*.
-
-You can upload the files here: LINK
-
-# Supplementary Material 2: Model reports
+# Supplementary Material 1: Model reports
 
@@
 
-Yannek Käber Harald Bugmann
+Yannek Käber, Harald Bugmann
 
@@
 ### Model references
-
-*Please provide here the references to this model where it is possible
-to understand the model structure specially for the regeneration module,
-this could be a published paper, a website with the model description, a
-manual or some other document that the reader can access.*
 
@@
 height and diameter growth potential, and consequently tree increment.
-The climatic factors also influence tree regeneration model. More
-detailed information about Sibyla can be found in Fabrika (2005),
+The climatic factors also influence tree regeneration model.
+
+More detailed information about Sibyla can be found in Fabrika (2005) and in the link: 
 <http://etools.tuzvo.sk/sibyla/english/model.htm>.
@@
 
-• What climate data have you used, and how have you created your model
-climate input parameters?
-
 Climate inputs for iLand are at daily temporal resolution. We used daily
@@
 ### Output
-
-• Please describe how have you calculated each of the variables in the
-output data
 
@@
 "ba").
-
-Figure 1: Exemplary overview over output file.
 
@@
 ```
-# Supplementary Material 3: Supplementary figures and tables
+# Supplementary Material 2: Supplementary figures and tables
 
@@
 ```
+
+## References

### 2023-03-31 manuscript/protocol.Rmd
added the protocol uploaded in OSF as a different file

---
title: "Supplementary figures\nCapturing regeneration in models of forest dynamics"
author: 
- Olalla Díaz-Yáñez, 
- Yannek Käber, 
- Harald Bugmann
date: "`r Sys.Date()`"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    latex_engine: xelatex
    extra_dependencies: ["flafter", "float", "subfig"]
    toc: yes
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: global-change-biology.csl
editor_options: 
  markdown: 
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
- \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H", out.extra = "")

#knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")
```

# Supplementary Material 1: Protocol

## General steps

Each modeling team will run their simulations prior to the workshop, and
will report on the results before ([short written report](#report)) and
during the workshop. The general approach for the workshop is:

1.  Each team sets up their model to simulate the regeneration niche of
    [eleven major European tree species](#speciescomposition) within the
    known, defined range of environmental conditions of 200 sites. They
    will run at least one simulation per site. It is important to
    conduct mixed-species simulations, not series of monospecific
    stands.

2.  The model is set up for each site based on the [input variable
    information](#inputvariables) provided in this protocol, and the
    simulations need to be run until the [equilibrium is
    reached](#equilibrium).

3.  Each team reports on the results of these exercises in terms of a
    [standardized output dataset](#outputdata) (including density
    variables per 10-yr interval and per ha, for each simulated site and
    species) and in the form of a brief [working report](#report) (\~4
    pages).

4.  At the workshop in June 2022 in Davos, Switzerland, we will evaluate
    the regeneration niche as it results from each model and compare it
    to the empirical estimates of the environmental limits of the
    regeneration niche that will be made available to the participants
    shortly before the workshop. As a group effort, the projections of
    the niche properties arising from the different models will be
    compared. The exact details on how this will be done will be
    provided once we know how many modeling teams will contribute
    results.

## Input variables {#inputvariables}

In this section, we are providing a description of all the necessary
driver information to run the simulations (e.g., location, climate,
soil).

### Coordinates

We intend to provide the site information via unique site IDs. If actual
coordinates are needed, we can provide blurred site coordinates upon
reasonable request.

### Climate

We provide time series of climatic variables from 1981 to 2018 in
hourly, daily and monthly resolution. Based on our survey the climate
data should be sufficient to run all models. Some variables such as
relative humidity or vapor pressure deficit must be calculated from the
variables we provide.

Note that the hourly time series has the lowest spatial resolution, thus
the daily time series should be used for spatial accuracy. If you need
hourly time series we recommend to use the daily time series and only
use the hourly time series to correct for hourly variation. For this
purpose we also provide hourly averages over the period 1981 to 2018 for
each hour of each day of the year. Only for calculating these averages
the 29th of February in leap years was removed from the time series to
obtain values for 365 days.

#### Data sources {#climsources}

The hourly time series are based on the Era5-land data set. Daily and
monthly time series are mostly based on CHELSA data which has a higher
spatial resolution. More details are given below or in the documentation
of the original data source.

The variable names are kept exactly like in the documentation of the
Era5 and CHELSA data sets to facilitate the use of the original data
documentation. Please don't be confused by the ending `*_hourly` in
`surface_thermal_radiation_downwards_hourly` and the two different
variable names for shortwave radiation: `rsds` for the values from
CHELSA and `surface_solar_radiation_downwards_hourly` for the values
from Era5.

All changes that were made to the original data are documented here:

-   The wind variable `wind_speed` is based on the u and v component of
    the wind and calculated with $\sqrt{u^2+v^2}$ where $u$ is the
    variable `u_component_of_wind_10m` and $v$ is
    `v_component_of_wind_10m` of the Era5 data set.

-   The units of temperature variables `temperature_2m` (Era5), `tas`,
    `tasmin`, `tasmax`(all CHELSA) were changed from Kelvin to Degree
    Celsius

-   The units of precipitation variables `total_precipitation_hourly`
    (Era5), `pr` (CHELSA) were changed from m to mm and kG to mm
    respectively.

-   The unit of short wave radiation `rsds` (CHELSA) was changed from kJ
    to J.

We aggregated variables from hourly to daily to monthly values by using
means for temperature, surface pressure and wind speed and sums for
radiation and precipitation.

##### Era5-Land {#era5}

-   Muñoz Sabater, J., (2019): ERA5-Land hourly data from 1981 to
    present. Copernicus Climate Change Service (C3S) Climate Data Store
    (CDS). (23.01.2022),
    [doi:10.24381/cds.e2161bac](https://doi.org/10.24381/cds.e2161bac)
-   [Documentation of the hourly ERA5 data from the Google Earth
    Engine](https://developers.google.com/earth-engine/datasets/catalog/ECMWF_ERA5_LAND_HOURLY#description)
-   Further information is available from the [official copernicus
    website](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-land?tab=overview).

##### CHELSA {#chelsa}

-   Dirk N. Karger, Stefan Lange, Chantal Hari, Christopher P. O. Reyer,
    Niklaus E. Zimmermann (2021): CHELSA-W5E5 v1.0: W5E5 v1.0 downscaled
    with CHELSA v2.0. ISIMIP Repository.
    <https://doi.org/10.48364/ISIMIP.836809>

-   [Official CHELSA website](https://chelsa-climate.org)

#### Monthly time series 1981-2018

file name: monthly_era5_chelsa_200.zip

**Data description**

Variables from the [Era5 data set](#era5) with a spatial resolution of
0.1° x 0.1° (\~ 10 km):

-   average wind speed (`wind_speed`) [m/s]

-   average surface pressure (`surface_pressure`) [Pa]

-   surface thermal (longwave) radiation
    (`surface_thermal_radiation_downwards_hourly`) [J/m2]

Variables from the [CHELSA data set](#chelsa) with a spatial resolution
of 0.008° x 0.008° (\~ 1 km):

-   average temperature (`tas`) [°C]

-   average max temperature (`tasmax`) [°C]

-   average min temperature (`tasmin`) [°C]

-   precipitation sum (`pr`) [mm]

-   surface solar (shortwave) radiation (`rsds`) [J/m2]

#### Daily time series 1981-2018

file name: daily_era5_chelsa_200.zip

**Data description**

Variables from the [Era5 data set](#era5) with a spatial resolution of
0.1° x 0.1° (\~ 10 km):

-   average wind speed (`wind_speed`) [m/s]

-   average surface pressure (`surface_pressure`) [Pa]

-   surface thermal (longwave) radiation
    (`surface_thermal_radiation_downwards_hourly`) [J/m2]

Variables from the [CHELSA data set](#chelsa) with a spatial resolution
of 0.008° x 0.008° (\~ 1 km):

-   average temperature (`tas`) [°C]

-   average max temperature (`tasmax`) [°C]

-   average min temperature (`tasmin`) [°C]

-   precipitation sum (`pr`) [mm]

-   surface solar (shortwave) radiation (`rsds`) [J/m2]

#### Hourly time series 1981-2018

file name: hour_era5_200.zip

file name: yhourMean_era5_200.zip

**Data description**

Variables from the [Era5 data set](#era5) with a spatial resolution of
0.1° x 0.1° (\~ 10 km):

-   average temperature (temperature_2m) [°C]

-   average wind speed (`wind_speed`) [m/s]

-   average surface pressure (`surface_pressure`) [Pa]

-   precipitation sum (total_precipitation_hourly) [mm]

-   surface solar (shortwave) radiation
    (surface_solar_radiation_downwards_hourly) [J/m2]

-   surface thermal (longwave) radiation
    (surface_solar_radiation_downwards_hourly) [J/m2]

### Soil

file name: soil_quality_200.csv

**Data description**

The following variables are available:

-   Soil quality is given as a continuous value between 1 and 5.

**Data source**

-   The soil quality value is based on a random forest that was trained
    to predict expert knowledge based soil quality on a scale from 1 to
    5 from the [SoilGrids250 data
    set](https://www.isric.org/explore/soilgrids) and the [WISE data
    set](https://www.isric.org/explore/wise-databases). The most
    influential variables in the random forest were total depth to
    bedrock and sand content. Thus the variable is a good predictor for
    plant available water storage capacity.

## Topography and elevation

There was an error in a earlier version of this file. Please use the
link above to download the corrected file.

This is the old file and should not be used because the aspect column
contains elevation instead of the aspect.

file name: topo_data_200.csv

**Data description**

The following variables are available:

-   elevation [m.a.s.l]

-   slope [Degree]

-   aspect [Degree]

**Data source**

-   Topographic information was derived from the [ASTER DEM
    V3](https://lpdaac.usgs.gov/products/astgtmv003/).

-   Slope and aspect were calculated with QGIS. The specific tiles that
    were used are documented
    [here](https://ites-fe.ethz.ch/rws/downloads/15122021/aster_dem_files.txt),
    and the processing steps in QGIS are documented
    [here](https://ites-fe.ethz.ch/rws/downloads/15122021/QGIS_command_log.txt).

### Natural disturbances

The simulations must be run in the absence of natural disturbances.

## Management

The simulations must be run in the absence of management.

### Area simulated

From your model outputs, we would like to obtain samples on species
specific regeneration rates per decade and per ha. The modeling teams
decide on the simulated area and how they derive these samples. Please
describe the methodology used in the [output report](#report).

### Simulation duration and equilibrium {#equilibrium}

The simulation should be run from bare ground in the absence of
management to a simulated [equilibrium](#defequilibrium) ("potential
natural vegetation") with current climate. The goal is to derive the
[regeneration niche](#regenniche). The exact length of the simulation
should be decided by the modeling teams. Please describe the simulation
length used in your simulations in the [output report](#report).

[**Regeneration niche:**]{#regenniche} The regeneration niche is defined
as the passing of a diameter threshold of 7 and 10 cm (ingrowth; cf.
below). Both the width of the regeneration niche (i.e., in environmental
space) as well as the intensity of the recruitment process (i.e., the
number of ingrowth trees per area and per unit of time) are of interest.

[**Simulated equilibrium**]{#defequilibrium}: Simulated equilibrium is
the "Potential Natural Vegetation".

### Species composition {#speciescomposition}

**Mixed species simulations** must be run (not multiple single-species
simulations).

**Option A: Species level**

The simulation should be run for **mixtures** of the following
species/genera: *Fagus sylvatica, Picea abies, Abies alba, Carpinus
betulus, Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus
excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris.*

Additional species can be included in the simulation if modelers think
this is important, but they should report only on these 11
species/genera because the benchmarking data refer to them only. Please
report the methodology used in the [output report](#report).

**Option B: PFTs**

If species-level simulation is not possible in your model, substitute
the species by PFTs and use the model's default approach. Please report
the methodology used in the [output report](#report). Explain as well
which PFTs represent which species for your model.

## Diameter threshold

The [output data](#outputdata) have to be provided for two diameter
thresholds: 7 and 10 cm.

## Output

We are expecting to get the number of recruits over a 10-yr interval per
species and per ha for each of the sites. However, we are also
interested in the **variation per site** (i.e., we would like to see the
range of estimates, not only the mean value per site). Each modeling
team will decide how to sample multiple values per site that will show
the variation in recruitment.

Specifically, you should provide us with multiple samples (n = 200; this
number is the same as the number of sites by sheer coincidence) of the
regeneration rate for spatial units of 1 ha (please do [not]{.underline}
provide average of multiple 1-ha samples). This can be done in various
ways, e.g., by sampling simulated data from the same 1-ha plot in the
equilibrium over time, or by sampling several 1-ha plots from the
simulated forest at one specific point in time (in the equilibrium); for
clarification, cf. three sample strategies proposed in [figures
1](#sample1), [2](#sample2) and [3](#sample3).

If, for some reason, you cannot submit **200 samples** of **1 ha**,
please **let us know** in the [workshop chat](#chat) so we can discuss
alternatives. The sampling from the simulations must happen **within**
the simulated phase of Potential Natural Vegetation (for more
information, see how [equilibrium](#defequilibrium), Potential [Natural
Vegetation, and regeneration niche](#regenniche) are defined in this
protocol).

![Figure 1. In this example there are 200 samples collected at year i at
equilibrium differentiated spatially with each square representing 1
ha.](samplingfigure/Slide2.jpeg){#sample1 width="500"}

![Figure 2. In this example, there are 100 spatially differentiated
samples taken in year i at equilibrium and 100 in year i+10. Each square
represents 1 ha.](samplingfigure/Slide3.jpeg){#sample2 width="500"}

![Figure 3. In this example there is 1 spatially differentiated sample
taken in year i at equilibrium and remeasured every 10 years for 2000
years until i+2000. The square represents 1
ha.](samplingfigure/Slide4.jpg){#sample3 width="500"}

### Output data {#outputdata}

Please provide the simulation output data using the following structure:

-   Each row of your output dataset contains the recruitment of 10-year
    interval per species and ha for a unique simulation site.

-   Each column of your output dataset contains one of these variables:

    -   `sample`: this will identify the different samples per site
        (1:n, with n $\ge$ 200)

    -   `site`: site identifier (1:200)

    -   `species`: species simulated with the following categories:

        -   fasy (for *Fagus sylvatica*)
        -   pabi (for *Picea abies*)
        -   abal (for *Abies alba*)
        -   cabe (for *Carpinus betulus*)
        -   tico (for *Tilia cordata*)
        -   acps (for *Acer pseudoplatanus*)
        -   betu (for *Betula* spp.)
        -   frex (for *Fraxinus excelsior*)
        -   quer (for *Quercus* spp.)
        -   algl (for *Alnus glutinosa*)
        -   pisy (for *Pinus sylvestris*)
        -   If you have some other categories please create your own
            species codes and provide the legend in the output report.
            We prefer a lumped category "othr".

    -   `r.trees`: the recruits (trees $ha^{-1}$) per species and
        decade.

    -   `r.ba`: the recruits' basal area ($m^{2}$ $ha^{-1}$) per species
        and decade.

    -   `ba`: total basal area ($m^{2}$ $ha^{-1}$) per species and
        decade

    -   `dbh`: diameter threshold (cm) used to define the recruitment (7
        or 10 cm).

### Output report {#report}

To guide the writing of the output report, please refer to the following
questions:

-   What climate data have you used, and how have you created your model
    climate input parameters?

-   How have you defined the species composition in your model
    simulation? (e.g., species levels, PFTs, etc.)? Please describe the
    species categories used.

-   How have you defined the simulation area for your model (e.g., size
    and other relevant site aspects)?

-   What are the relevant aspects of the dispersal (if any) and seed
    input parameters in your model?

-   What is the simulation length, and how have you determined the
    simulation time until the equilibrium?

-   If you have run more than one simulation per site, please describe
    the differences of the approaches used.

-   Please describe how have you calculated each of the variables in the
    [output data](#outputdata).

### Uploading data and report

Please upload your data and output report as a **zip file** with the
name of your model and any other word that might help us to understand
the files, for example: *thenameofyourmodel_datareport.zip* or
*thenameofyourmodel_data.zip* and *thenameofyourmodel_report.zip*.

You can upload the files here: LINK

### 2023-04-11 manuscript/supplementaryRW.Rmd
added model resources for Landis in the supplementary

@@
 
-More detailed information about Sibyla can be found in Fabrika (2005) and in the link: 
-<http://etools.tuzvo.sk/sibyla/english/model.htm>.
+More detailed information about Sibyla can be found in Fabrika (2005)
+and in the link: <http://etools.tuzvo.sk/sibyla/english/model.htm>.
 
@@
 
-The most complete reference resource for the iLand regeneration module is:
-Seidl, R., Spies, T.
-a.
-T.
-A.
-T.
-a., Rammer, W., Steel, E.
-A.
-A., Pabst, R.
-J.
-R.
-J.
-R.
-J., & Olsen, K.
-(2012).
-Multi-scale Drivers of Spatial Variation in Old-Growth Forest Carbon Density Disentangled with Lidar and an Individual-Based Landscape Model.
-Ecosystems, 15(8), 1321-1335.
-https://doi.org/10.1007/s10021-012-9587-2
+The most complete reference resource for the iLand regeneration module
+is: Seidl, R., Spies, T. a. T. A. T. a., Rammer, W., Steel, E. A. A.,
+Pabst, R. J. R. J. R. J., & Olsen, K. (2012). Multi-scale Drivers of
+Spatial Variation in Old-Growth Forest Carbon Density Disentangled with
+Lidar and an Individual-Based Landscape Model. Ecosystems, 15(8),
+1321-1335. <https://doi.org/10.1007/s10021-012-9587-2>
 
@@
 ### Model
-
-*Please provide here the references to this model where it is possible
-to understand the model structure specially for the regeneration module,
-this could be a published paper, a website with the model description, a
-manual or some other document that the reader can access.*
 
@@
 specified time steps.
+
+The most complete reference resources for the LANDIS-II model and
+regeneration module used here are:
+
+-   Scheller et al. 2007 -
+    <https://doi.org/10.1016/j.ecolmodel.2006.10.00>
+-   De Bruijn et al. 2014 -
+    <https://doi.org/10.1016/j.ecolmodel.2014.05.004>
+-   Petter et al. 2020 - <https://doi.org/10.1016/j.envsoft.2020.104844>
+    (supplementary material)
+-   User guide for PNET 4.1 online -
+    <https://github.com/LANDIS-II-Foundation/Extension-PnET-Succession/blob/master/deploy/docs/LANDIS-II%20PnET-Succession%20v4.1%20User%20Guide.pdf>
 
@@
 \newpage
+
 ## Total ingrowth and individual species regeneration niches

### 2023-04-11 manuscript/supplementaryRW.Rmd
update model references to make it homogeneous across models reports in the suplementary

@@
 
-### Model references
-
-*Please provide here the references to this model where it is possible
-to understand the model structure specially for the regeneration module,
-this could be a published paper, a website with the model description, a
-manual or some other document that the reader can access.*
+Mahnken M., Gutsch M., Reyer C.P.O., Lasch-Born P.
+
+### Model
+
+The most complete reference resource for the 4C (v2.2) model and
+regeneration module is:
+
+Lasch-Bonn et al. 2020. Description and evaluation of the process-based
+forest model 4C v2.2 at four European forest sites. DOI:
+[https://doi.org/10.5194/gmd-13-5311-2020)](https://doi.org/10.5194/gmd-13-5311-2020))
 
@@
 
+We used the provided daily ERA5-CHELSA climate data set and complemented
+the CHELSA time series data regarding relative humidity with the model
+after Eccel et al. (2015) based on minimum and mean daily temperature to
+accomodate data needs of 4C. The missing data for January 2nd and 3rd
+2013 were filled with data from January 1st and 4th 2013 respectively.
+For construction of longer climate time series we looped through the
+provided time series from 1981 to 2018 up to the maximum simulation
+length of 2500 years. We amended situations where leap year
+inconsistencies arised by removing February 29 or duplicating data from
+February 28. Otherwise, the unaltered daily data were used to run 4C on
+a daily basis. We applied a constant atmopsheric CO2 concentration of
+380 ppm.
+
 ### Soil
 
+We used parametrized soils readily available for 4C from the EU soil
+data base. The link from parametrized soil to soil quality was done by
+extracting the plant available water storage capacity as the model
+bucket size of the rooted soil horizons. We excluded bucket size values
+of lower than 10 cm and larger than 35 cm. The bucket size was
+reprojected to the range from 0 to 1. In addition we tested the
+sensitivity of dominant tree height at age 100 years for monospecific
+stands of all five species descibed above to the available soils using
+model simulations with 4C. The dominant tree height at age 100 years was
+then also reprojected to the range from 0 to 1 in order to derive a
+model specific soil quality indicator for all paramterized soils. Then
+we multiplied this relative model specific soil quality with the
+relative soil quality derived from the bucket size to the range from 1
+to 5. The match between paramterized soil and plot was then done by
+finding the pair of soil quality with the minimum difference between
+simuation protocol prescribed soil quality and and the soil quality of
+the paramterized soils as described above.
+
 ### Topography
 
+Topographical information was not considered in the simulations.
+
 ### Tree species
 
+We initialized stands with five out of the eleven proposed
+species/genera: Fagus sylvatica (fasy), Picea abies (piab), Betula spp.
+(betu; 4C parameters for Betula pendula), Quercus spp. (quer; 4C
+parameters for Quercus robur) and Pinus sylvestris (pisy) since only
+these are fully parameterized for an application as proposed in the
+simulation protocol. Each plot was initialized with all five tree
+species with similar distributions of height classes and number of
+individuals: 2000 individuals/ha with height around 0.25 m, 1000
+individuals/ha with height around 0.75 m and 500 individuals/ha with
+height around 1.75 m. In total 17500 individuals/ha.
+Carpinus betulus,
+Tilia cordata, Acer pseudoplatanus, Fraxinus excelsior, Alnus glutinosa
+and Abies alba are not fully parameterized in 4C.
+
 ### Simulations
 
+#### Dispersal and seed input
+
+There is no explicitly modelled seed dispersal in a stand and between
+stands. We applied a constant sapling establishment rate every five
+years of 2000 individuals/ha for all five species; 10000 individuals/ha
+in total every five years. The constant sapling establishment provides
+constant regeneration potential if favorable conditions for sapling
+growth (especially light and water regime) are present. This approach
+enables regeneration of other species than only the one present in the
+stand to establish if competition allows. The resulting recruitment
+patterns are an outcome of the dynamical processes from the sapling
+stage onward. Seed production and germination were not explicitly
+simulated.
+
+#### Simulation length and equilibrium criteria
+
+To reach equilibrium and obtain dynamics in the equilibrium period, we
+simulated a maximum of 2500 years. 4C is usually applied for much
+shorter simulation periods and different technical as well as
+model-theory related issues arise when simulating longer than 100-150
+years. Therefore 2500 years can be seen as the upper limit of continuous
+simulation in the defined simulation setup. The first 500 years were not
+sampled for recruitment as we assume the model needs this time to reach
+equilibrium. In this long simulation period 4C partially projects
+unrealistically large tree heights resulting in premature termination of
+the simulation (36 plots) so that not all plots were simulated for the
+full 2500 years. We started sampling the recruits after the first 500
+simulation years from non-overlapping 10 year periods to obtain 200
+samples. For shorter simulation periods we produced 200 samples from
+overlapping 10 year periods.
+In addition, we excluded those plots in which 4C simulated
+unrealistically high basal area values of over 150 m² (19 plots). We
+used this large threshold to give the model maximum flexibility to
+explore the solution space in these long simulation runs although we
+acknowledge that this is an unrealistic overestimation that is probably
+originating from misrepresented density dependent mortality processes in
+long simulation runs and misrepresented single tree dimensions due to
+wrong assumptions of tree geometry and biomass share of such old-aged
+individuals. Furthermore those plots with less than 1000 simulation
+years (16 plots) before termination were excluded because the sample
+space after the first 500 years was too small. Additionally we included
+another subset criterion removing all samples that have a ba \> 100 to
+exclude unrealistically high samples without the need to exclude the
+full site. This resulted in discarding 0.8 % of the sample
+
 ### Outputs
 
+We provide 200 samples each for 165 of the 200 plots (see reason for
+excluded plots in "Simulation length and equilibrium criteria"). These
+stem from 10 year periods that overlap partially (for \< 2500 simulation
+years) or do not overlap (for 2500 simulation years). The number of
+recruits (r.trees) is the sum of individuals that cross the threshold
+diameter (7/10 cm) in the 10 year sample period per hectare that have
+not already died until the sample year (Fig. 3). The recruit's basal
+area (r.ba) is the sum of the basal area per hectare of the trees that
+cross the diameter threshold that have not already died until the sample
+year. The total stand basal area (ba) is the basal area per hectare of
+all trees in the stand \> 0 cm diameter at breast height (no lower
+boundary) at the sample time per species. We provide the data for both
+diameter thresholds, 7 and 10 cm. In over 90 % of the decades no
+recruitment takes place.
+
+[4C
+(v2.2)]([https://doi.org/10.5194/gmd-13-5311-2020)](https://doi.org/10.5194/gmd-13-5311-2020))
+is usually not used for simulations of forest equillibrium dynamics,
+potential natural vegetation assessment and long simulation periods but
+rather shorter-term simulations (up to 100-150 years). In addition, only
+five species out of the pool of eleven potentially relevant species here
+are parameterized in 4C so that projections of the potential natural
+vegetation done wih 4C need to be considered in the light of model
+uncertainty.
+For example, the spatial homogeneity in the model leads to
+cyclic recruitment patterns (see Fig. 4). When the adult stands dies of,
+young individuals are recruited into the larger diameter classes because
+more light reaches lower canopy layer since larger (adult) trees do not
+anymore prevent the light reaching lower canopy layers. This results in
+increased growth of the regeneration layer during those times. In most
+stands Fagus sylvatica dominates the stand after some time but
+occasionally individuals of other species, like Picea abies or Quercus
+spp. cross the recruitment diameter threshold.
+
 ## ForCEEPS & ForCEEPS(f)
@@
 
-### Model references
-
-Morin X., de Coligny F., Bugmann H., Limousin J.-M., Ourcival J.-M.,
+### Model
+
+The most complete reference resource for the ForCEEPS & ForCEEPS(f)
+model and regeneration module are:
+
+-   Morin X., de Coligny F., Bugmann H., Limousin J.-M., Ourcival J.-M.,
     Martin-StPaul N., Simioni G., Cailleret M., Prevosto B., Toïgo M.,
-Vennetier M., Cateau E., Guillemot J. 2021. Beyond forest succession: a
-gap model to study ecosystem functioning and tree community composition
-under climate change. Functional Ecology. 35, 955-975.
-
-Jourdan M., Dreyfus, P., Riond C., Cordonnier T., Cornet B., de Coligny
-F., Morin X. 2021. Managing mixed stands can mitigate severe climate
-change impacts on French alpine forests. Regional Environmental Change.
-21, 78.
-
-Morin X., Damestoy T., Toïgo M., Jactel H., Castagneyrol B. Meredieu C.
-2020. Using forest gap models and experimental data to explore long-term
-effects of tree diversity on the productivity of mixed planted forests.
-Annals of Forest Science. 77
+    Vennetier M., Cateau E., Guillemot J. 2021. Beyond forest
+    succession: a gap model to study ecosystem functioning and tree
+    community composition under climate change. Functional Ecology. 35,
+    955-975.
+
+-   Jourdan M., Dreyfus, P., Riond C., Cordonnier T., Cornet B., de
+    Coligny F., Morin X. 2021. Managing mixed stands can mitigate severe
+    climate change impacts on French alpine forests. Regional
+    Environmental Change. 21, 78.
+
+-   Morin X., Damestoy T., Toïgo M., Jactel H., Castagneyrol B.
+    Meredieu C. 2020. Using forest gap models and experimental data to
+    explore long-term effects of tree diversity on the productivity of
+    mixed planted forests. Annals of Forest Science. 77
     <https://doi.org/10.1007/s13595-020-00954-0>.
@@
 
-### Model references
+### Model
 
@@
 
-### Model references
-
-Bugmann, H. (1994). On the ecology of mountainous forests in a changing
-climate: A simulation study. Huber, N., Bugmann, H., & Lafond, V.
-(2020). Capturing ecological processes in dynamic forest models: Why
-there is no silver bullet to cope with complexity. Ecosphere, 11(5).
-<https://doi.org/10.1002/ecs2.3109>
+### Model
+
+The most complete reference resource for the ForClim 1 & ForClim 11
+model and regeneration module is:
+
+-   Bugmann, H. (1994). On the ecology of mountainous forests in a
+    changing climate: A simulation study. Huber, N., Bugmann, H., &
+    Lafond, V. (2020). Capturing ecological processes in dynamic forest
+    models: Why there is no silver bullet to cope with complexity.
+    Ecosphere, 11(5). <https://doi.org/10.1002/ecs2.3109>
 
@@
 
-Soil data were translated to bucket sizes values based on the
-assumpation that the lowest soil quality value of 1 refers to a bucket
-size of 10 cm and the highest soil quality of 5 refers to a bucket size
-of 35 cm. Specifically we used the following formula to calculate the
-bucket size bucket_size = 10 + (soil_quality-1) /4  \* 25.
+Soil data were translated to bucket sizes values based on the assumption
+that the lowest soil quality value of 1 refers to a bucket size of 10 cm
+and the highest soil quality of 5 refers to a bucket size of 35 cm.
+Specifically we used the following formula to calculate the bucket size
+bucket_size = 10 + (soil_quality-1) /4  \* 25.
 
@@
 
-All simulations were run according to the protocoll. For each site we
+All simulations were run according to the protocol. For each site we
 simulated on 16 ha (i.e., 100 patches of 0.08 ha). We defined a spin up
@@
 
-More detailed information about Sibyla can be found in Fabrika (2005)
-and in the link: <http://etools.tuzvo.sk/sibyla/english/model.htm>.
+The most complete reference resource for the SIBYLA model and
+regeneration module is:
+
+-    Fabrika (2005) and in the link:
+    <http://etools.tuzvo.sk/sibyla/english/model.htm>.
 
@@
 
-### Model summary
+### Model
 
@@
 (5) extended from single to multi-species simulations.
+
+The most complete reference resource for the xComp model and
+regeneration module is:
+
+-
 
@@
 manual or some other document that the reader can access.*
+
+The most complete reference resource for the PICUS model and
+regeneration module is:
+
+-
 
@@
 
-### Model summary
+### Model
 
 The most complete reference resource for the iLand regeneration module
-is: Seidl, R., Spies, T. a. T. A. T. a., Rammer, W., Steel, E. A. A.,
+is:
+
+-   Seidl, R., Spies, T. a. T. A. T. a., Rammer, W., Steel, E. A. A.,
     Pabst, R. J. R. J. R. J., & Olsen, K. (2012). Multi-scale Drivers of
-Spatial Variation in Old-Growth Forest Carbon Density Disentangled with
-Lidar and an Individual-Based Landscape Model. Ecosystems, 15(8),
-1321-1335. <https://doi.org/10.1007/s10021-012-9587-2>
+    Spatial Variation in Old-Growth Forest Carbon Density Disentangled
+    with Lidar and an Individual-Based Landscape Model. Ecosystems,
+    15(8), 1321-1335. <https://doi.org/10.1007/s10021-012-9587-2>
 
@@
 manual or some other document that the reader can access.*
+
+The most complete reference resource for the LandClim model and
+regeneration module is:
+
+-
 
@@
 The most complete reference resources for the LANDIS-II model and
-regeneration module used here are:
+regeneration module are:
 
@@
 
+The most complete reference resource for the TreeMig model and
+regeneration module are:
+
+-   Lischke, H., T. J. Loffler, and A. Fischlin. 1998. Aggregation of
+    individual trees and patches in forest succession models: Capturing
+    variability with height structured, random, spatial distributions.
+    Theor Popul Biol 54:213-226.
+
+-   Lischke, H., N. E. Zimmermann, J. Bolliger, S. Rickebusch, and T. J.
+    Loffler. 2006. TreeMig: A forest-landscape model for simulating
+    spatio-temporal patterns from stand to landscape scale. Ecological
+    Modelling 199:409-420.
+
+-   Lischke, H., and T. J. Loffler. 2006. Intra-specific density
+    dependence is required to maintain species diversity in
+    spatio-temporal forest simulations with reproduction. Ecological
+    Modelling 198:341-361.
+
+-   The online resources available at treemig.wsl.ch
+
 ### Climate
@@
 
-• Bugmann, H. 1994. On the ecology of mountainous forests in a changing
+Bugmann, H. 1994. On the ecology of mountainous forests in a changing
 climate: A simulation study. Dissertation. Swiss Federal Institute of
-Technology, Zurich. • Fischlin, A., H. Bugmann, and D. Gyalistras. 1995.
-Sensitivity of a forest ecosystem model to climate parametrization
-schemes. Env. Poll. 87:267-282. • Lischke, H., and T. J. Loffler. 2006.
-Intra-specific density dependence is required to maintain species
-diversity in spatio-temporal forest simulations with reproduction.
-Ecological Modelling 198:341-361. • Lischke, H., T. J. Loffler, and A.
-Fischlin. 1998. Aggregation of individual trees and patches in forest
-succession models: Capturing variability with height structured, random,
-spatial distributions. Theor Popul Biol 54:213-226. • Lischke, H., N. E.
-Zimmermann, J. Bolliger, S. Rickebusch, and T. J. Loffler. 2006.
-TreeMig: A forest-landscape model for simulating spatio-temporal
-patterns from stand to landscape scale. Ecological Modelling
-199:409-420. • Thornthwaite, C. W., and J. R. Mather. 1957. The water
-balance. Publications in Climatology VIII:1-69.
+Technology, Zurich.
+
+Fischlin, A., H. Bugmann, and D. Gyalistras. 1995. Sensitivity of a
+forest ecosystem model to climate parametrization schemes. Env. Poll.
+87:267-282.
+
+Lischke, H., and T. J. Loffler. 2006. Intra-specific density dependence
+is required to maintain species diversity in spatio-temporal forest
+simulations with reproduction. Ecological Modelling 198:341-361.
+
+Lischke, H., T. J. Loffler, and A. Fischlin. 1998. Aggregation of
+individual trees and patches in forest succession models: Capturing
+variability with height structured, random, spatial distributions. Theor
+Popul Biol 54:213-226.
+
+Lischke, H., N. E. Zimmermann, J. Bolliger, S. Rickebusch, and T. J.
+Loffler. 2006. TreeMig: A forest-landscape model for simulating
+spatio-temporal patterns from stand to landscape scale. Ecological
+Modelling 199:409-420.
+
+Thornthwaite, C. W., and J. R. Mather. 1957. The water balance.
+Publications in Climatology VIII:1-69.
 
@@
 version of the most common European tree species as well as typical
-shrub PFTs as described by Hickler et al. (2012). A detailed description
-of LPJ-GUESS is available in Smith et al. 2001 and Smith et al. 2014.
-Source code is available on demand (see
+shrub PFTs as described by Hickler et al. (2012).
+
+The most complete reference resource for the LPJ-GUESS model and
+regeneration module is:
+
+-   Smith, B. (2001). LPJ-GUESS-an ecosystem modelling framework.
+    Department of Physical Geography and Ecosystems Analysis, INES,
+    Sölvegatan, 12, 22362.
+
+-   Smith, B., Wårlind, D., Arneth, A., Hickler, T., Leadley, P.,
+    Siltberg, J., & Zaehle, S. (2014). Implications of incorporating N
+    cycling and N limitations on primary production in an
+    individual-based dynamic vegetation model. Biogeosciences, 11(7),
+    2027-2054.
+
+-   Source code is available on demand (see
     <https://web.nateko.lu.se/lpj-guess/>).
@@
 
-All sites We used were simulated withdefault values for a medium
+All sites We used were simulated with default values for a medium
 textured soil type (sand: 0.35, clay: 0.15, silt: 0.5) in LPJ-GUESS.
@@
 
+The most complete reference resource for the aDGVM2 model and
+regeneration module is:
+
+-   Scheiter S, Langan L, Higgins SI (2013) Next generation dynamic
+    global vegetation models: learning from community ecology. NEW
+    PHYTOLOGIST, 198, 957-969.
+
 ### Climate

### 2023-04-14 manuscript/supplementaryRW.Rmd
update report PICUS

@@
 individuals/ha with height around 0.75 m and 500 individuals/ha with
-height around 1.75 m. In total 17500 individuals/ha.
-Carpinus betulus,
+height around 1.75 m. In total 17500 individuals/ha. Carpinus betulus,
 Tilia cordata, Acer pseudoplatanus, Fraxinus excelsior, Alnus glutinosa
@@
 samples. For shorter simulation periods we produced 200 samples from
-overlapping 10 year periods.
-In addition, we excluded those plots in which 4C simulated
-unrealistically high basal area values of over 150 m² (19 plots). We
-used this large threshold to give the model maximum flexibility to
-explore the solution space in these long simulation runs although we
-acknowledge that this is an unrealistic overestimation that is probably
-originating from misrepresented density dependent mortality processes in
-long simulation runs and misrepresented single tree dimensions due to
-wrong assumptions of tree geometry and biomass share of such old-aged
-individuals. Furthermore those plots with less than 1000 simulation
-years (16 plots) before termination were excluded because the sample
-space after the first 500 years was too small. Additionally we included
-another subset criterion removing all samples that have a ba \> 100 to
-exclude unrealistically high samples without the need to exclude the
-full site. This resulted in discarding 0.8 % of the sample
+overlapping 10 year periods. In addition, we excluded those plots in
+which 4C simulated unrealistically high basal area values of over 150 m²
+(19 plots). We used this large threshold to give the model maximum
+flexibility to explore the solution space in these long simulation runs
+although we acknowledge that this is an unrealistic overestimation that
+is probably originating from misrepresented density dependent mortality
+processes in long simulation runs and misrepresented single tree
+dimensions due to wrong assumptions of tree geometry and biomass share
+of such old-aged individuals. Furthermore those plots with less than
+1000 simulation years (16 plots) before termination were excluded
+because the sample space after the first 500 years was too small.
+Additionally we included another subset criterion removing all samples
+that have a ba \> 100 to exclude unrealistically high samples without
+the need to exclude the full site. This resulted in discarding 0.8 % of
+the sample
 
@@
 [4C
-(v2.2)]([https://doi.org/10.5194/gmd-13-5311-2020)](https://doi.org/10.5194/gmd-13-5311-2020))
+(v2.2)](%5Bhttps://doi.org/10.5194/gmd-13-5311-2020)](<https://doi.org/10.5194/gmd-13-5311-2020>))
 is usually not used for simulations of forest equillibrium dynamics,
@@
 vegetation done wih 4C need to be considered in the light of model
-uncertainty.
-For example, the spatial homogeneity in the model leads to
+uncertainty. For example, the spatial homogeneity in the model leads to
 cyclic recruitment patterns (see Fig. 4). When the adult stands dies of,
@@
 
-*Please provide here the references to this model where it is possible
-to understand the model structure specially for the regeneration module,
-this could be a published paper, a website with the model description, a
-manual or some other document that the reader can access.*
-
-The most complete reference resource for the PICUS model and
-regeneration module is:
-
--
+The core elements of PICUS are described in Lexer and Hönninger (2001)
+and Seidl et al. (2005). Specific details regarding the regeneration
+module can be found in Woltjer et al. (2008).
+
+-   Lexer MJ, Hönninger K (2001) A modified 3D-patch model for spatially
+    explicit simulation of vegetation composition in heterogeneous
+    landscapes. For Ecol Manag 144:43--65.
+    doi:10.1016/S0378-1127(00)00386-8
+
+-   Seidl R, Lexer MJ, Jäger D, Hönninger K (2005) Evaluating the
+    accuracy and generality of a hybrid patch model. Tree Physiol
+    25:939--951. doi:10.1093/treephys/25.7.939
+
+-   Woltjer M, Rammer W, Brauner M, Seidl R, Mohren GMJ, Lexer MJ (2008)
+    Coupling a 3D patch model and a rockfall module to assess rockfall
+    protection in mountain forests. J Environ Manag 87:373--388.
+    doi:10.1016/j.jenvman.2007.01.031
 
@@
 
-We used soil parameters as suggested by the protocol. Between given
-class specific values linear interpolation was sued to find
-between-class values. Soil Class N WHC pH 1 40 80 4.2 2 - - 4.2 3 - -
-4.6 4 - - 5.6 5 100 250 5.6
+Based on the provided continuous value for soil quality between 1 and 5
+minimum (soil quality = 1) and maximum (soil quality = 5) values for
+soil water holding capacity (80mm, 200mm), plant available Nitrogen (40
+kg ha^-1^ yr^-1^, 100 kg ha^-1^ yr^-1^) and soil pH (4.2, 5.6) were
+defined. Linear interpolation was used to scale between the minimum and
+maximum values.
 
 ### Topography
+
+Not considered.
 
@@
 
-r.trees Counted the number of trees which were not present in the stand
-from the previous time period (10 years before) and where above the dbh
-threshold per species. For the oaks the value of the two species was
-summed into one. The result was then divided by 1.44 to get a per ha
-value of r.trees. r.ba Sum of the basal area (circle area calculated
-with the dbh) of trees which were not present in the stand from the
-previous time period (10 years before) and where above the dbh
-threshold. This was then divided by 1.44 to get a per ha value. ba Sum
-of the basal area (circle area calculated with the dbh) of all trees
-above the dbh threshold. This was then divided by 1.44 to get a per ha
-value.
+Output consisted of species-specific regeneration rates per decade and
+per ha. Oak species were aggregated. A sample was taken for every 10
+years for a total period of 2000 years once the model had reached
+equilibrium (see 1.7.7.3). Ingrowth was determined by counting the trees
+passing the 7cm and 10cm DBH thresholds every 10 years. Basal area of
+the ingrowth was determined at the end of each decade.
 

### 2023-04-14 manuscript/manuscriptRW.Rmd
Add second name letter for Manfred

@@
 - Christian	Hochauer,
-- Manfred Lexer,
+- Manfred J. Lexer,
 - Heike	Lischke,

### 2023-04-14 manuscript/supplementaryRW.Rmd
Add second name letter for Manfred

@@
 - Xavier Morin, 
-- Manfred Lexer, 
+- Manfred J. Lexer, 
 - Christian Hochauer, 

### 2023-04-14 manuscript/manuscriptRW.Rmd
changes based on PICUS team feedback

@@
 the seedbank [@lischke2006]. However, simulated species composition is
-usually exceedingly sensitive to that level, whereas the parameters of
-such functions are poorly constrained by field data. Lastly, there is
-often a problem with insufficient observational constraints on parameter
-values for models that start from very small tree sizes (e.g., 10 cm of
-height or even seeds) and track tree development in a process-oriented
-manner by considering a multitude of ecological influences, rather than
-emphasizing tree establishment into a larger size class [e.g.,
-@wehrli2006].
+usually exceedingly sensitive to such a formulations, whereas the
+parameters of such functions are poorly constrained by field data.
+Lastly, there is often a problem with insufficient observational
+constraints on parameter values for models that start from very small
+tree sizes (e.g., 10 cm of height or even seeds) and track tree
+development in a process-oriented manner by considering a multitude of
+ecological influences, rather than emphasizing tree establishment into a
+larger size class [e.g., @wehrli2006].
 
@@
 
-Fifteen models of forest dynamics (including two models featuring
+Fifteen models of forest dynamics (including two models featuring model
 variants) were used to assess simulated tree establishment. They
@@
 from saplings have to aggregate via parameterizations several ecological
-processes because they are not treated explicitly. This later approach
-reduces model complexity but comes at the cost of blurring process
+processes which are not treated explicitly. This later approach reduces
+model complexity but comes at the cost of blurring process
 representation.
@@
 ForClim and TreeMig) as having rather high complexity in their
-establishment modules (mean establishment formulation complexity
-complexity across all processes \>0, Table \@ref(tab:tableModels)). Two
-models, ForClim and ForCEEPS, were used here with two alternative
-variants of establishment. ForClim variant 1 [@bugmann1996] is based on
-a recruitment module that adheres closely to the concept introduced by
+establishment modules (mean establishment formulation complexity across
+all processes \>0, Table \@ref(tab:tableModels)). Two models, ForClim
+and ForCEEPS, were used here with two alternative variants of
+establishment. ForClim variant 1 [@bugmann1996] is based on a
+recruitment module that adheres closely to the concept introduced by
 @botkin1972, whereas ForClim variant 11 is adopting a different approach
@@
 composition and relative abundance of species in the plot) via seed
-trees for the quantity and quality of simulated ingrowth, as one variant
-includes this feedback whereas the other does not.
+trees for the quantity and quality (e.g., diversity and composition) of
+simulated ingrowth, as one variant includes this feedback whereas the
+other does not.
 
@@
 k-means clustering to define 200 clusters of plots from the original set
-of 869 plots along the environmental dimensions of temperature, water
-balance, soil quality and exposition (i.e., slope and aspect). Each of
-these sites featured at least two consecutive measurements. Ingrowth
-thresholds for the sites differed between 0 and 10 cm. For our study, we
-defined two data sets where one included 165 sites with a diameter
-threshold of 7 cm or lower, and the other included another 35 sites with
-diameter thresholds between 7 and 10 cm. Thus, we focused on two data
-sets with harmonized diameter thresholds of 7 cm and 10 cm,
-respectively.
+of 869 plots along the environmental dimensions of temperature, climatic
+water balance, soil quality and exposition (i.e., slope and aspect).
+Each of these sites featured at least two consecutive measurements.
+Ingrowth thresholds for the sites differed between diameters of 0 and 10
+cm. For our study, we defined two data sets where one included 165 sites
+with a diameter threshold of 7 cm or lower, and the other included
+another 35 sites with diameter thresholds between 7 and 10 cm.
 
@@
 decade ranged from zero to 1246 trees per ha, with a mean of 56 trees
-per ha. Adult species composition was also available for the simulated
-species. For more details on this unique dataset and the detailed
-establishment information, cf. @käber2022.
+per ha. Adult species composition was also available at the plots. For
+more details on this unique dataset and the detailed establishment
+information, cf. @käber2022.
 
@@
 
-```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site for observed and vs.simulated data.
+```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site for observed and simulated data.
 n indicates the number of models falling in each group.
@@
 B) Examples for the three trends across models for ingrowth at the 7 cm threshold.
-The full data are shown in Figures S5 and S6.", out.width='75%'}
+The full data are shown in Figures S3 and S4.", out.width='75%'}
 knitr::include_graphics(here::here("figures", "combined_richness.png"))
@@
 linked to either seed rain from dispersal or a seed/seedling bank;
-neither process is incororated across many models studied here, however.
+neither process is incorporated across many models studied here,
+however.
 
@@
 scales. We used total stand basal area as a proxy for light availability
-[cf. @schmid2021]. We found pronounced differences in the stand basal
-area ranges simulated by the models, but also between the models and the
-observations. This made it impossible to evaluate the establishment for
-the extremes of the stand density ranges in some models. For example,
-ingrowth levels at low stand densities are relevant to assess how well
-forests are recovering e.g. after gap creation due to disturbance
-[@seidl2022; @grubb1977]. At the level of the simulated one-hectare
-samples, average basal area is typically not very low as long as no
-larger disturbances occur, which was explicitly excluded in our
-simulation protocol. Towards the other end of the spectrum, i.e., with
-increasing stand basal area, it would be reasonable to expect that the
-ingrowth of the different species would become sparser and drop out
+[cf. @schmid2021]. However we could not consider light availability
+restrictions caused by ground vegetation which might be an important
+filter for some models [@woltjer2008]. We found pronounced differences
+in the stand basal area ranges simulated by the models, but also between
+the models and the observations. This made it impossible to evaluate the
+establishment for the extremes of the stand density ranges in some
+models. For example, ingrowth levels at low stand densities are relevant
+to assess how well forests are recovering e.g. after gap creation due to
+disturbance [@seidl2022; @grubb1977]. At the level of the simulated
+one-hectare samples, average basal area is typically not very low as
+long as no larger disturbances occur, which was explicitly excluded in
+our simulation protocol. Towards the other end of the spectrum, i.e.,
+with increasing stand basal area, it would be reasonable to expect that
+the ingrowth of the different species would become sparser and drop out
 entirely under low-light conditions [@klopcic2012; @zell2019]. However,
@@
 reasonable than it might appear at first sight. In an analysis of
-primeval forests at demographic equilibrium, @brzeziecki2021, they found
+primeval forests at demographic equilibrium, @brzeziecki2021 found
 higher ingrowth rates than the ones observed in the Euforia data. Thus,
@@
 results. We must not ignore that the observed data represent just a
-snapshot of an extremely complex and stochastic reality. This, may at
-least partly explain e.g. the lack of clear patterns in the simulation
+snapshot of an extremely complex and stochastic reality. This may at
+least partly explain e.g., the lack of clear patterns in the simulation
 results along the gradients of climate water balance and degree-days.
@@
 earlier efforts to make them more robust regarding growth and mortality.
-It remain uncertain what level of detail is required to model
+It remains uncertain what level of detail is required to model
 establishment processes, and this must be addressed in future research.

### 2023-04-14 manuscript/supplementaryRW.Rmd
changes based on PICUS team feedback

@@
 The species richness is estimated using the Shannon index  calculated by basal area.
-A1) groups models that Simulated species composition in adults is higher than the observed, A2) groups models that Simulated species composition in adults is similar to the observed, A3) groups models where the imulated species composition in adults is lower than the observed.
+A1) groups models where the simulated species composition in adults is higher than the observed, A2) groups models where the simulated species composition in adults is similar to the observed, A3) groups models where the simulated species composition in adults is lower than the observed.
 NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%', fig.pos='H'}
@@
 
-```{r simObr, echo=FALSE, fig.cap="Mean species richness per site in the simulated recruitment versus observed species richness.
-The species richness is estimated using the Shannon index is calculated by basal area.B1) groups models where the simulated species composition is closer than the other models to observed values, B2) groups models where the sites with lower species composition are underestimated, B3) groups models where the simulated species composition recruitment is lower in the whole range compare to the observe.
+```{r simObr, echo=FALSE, fig.cap="Mean species richness per site in the simulated recruitment versus observed species richness in the recruitment.
+The species richness is estimated using the Shannon index using the basal area.
+B1) groups models where the simulated species composition in recruitment is higher than the observed, A2) groups models where the simulated species composition in recruitment is similar to the observed, A3) groups models where the simulated species composition in recruitment is lower than the observed.
 NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%', fig.pos='H'}

### 2023-04-14 manuscript/supplementaryRW.Rmd
updated references for landis model in the appendix

@@
 
--   Scheller et al. 2007 -
-    <https://doi.org/10.1016/j.ecolmodel.2006.10.00>
--   De Bruijn et al. 2014 -
-    <https://doi.org/10.1016/j.ecolmodel.2014.05.004>
--   Petter et al. 2020 - <https://doi.org/10.1016/j.envsoft.2020.104844>
+-   Scheller R.M. & Mladenoff D.J. (2004) A forest growth and biomass
+    module for a landscape simulation model, LANDIS: design, validation,
+    and application. Ecological Modelling 180:
+    211--229. <https://doi.org/10.1016/j.ecolmodel.2004.01.022>
+-   Scheller, R. M., Domingo J. B. , Sturtevant B. R. , Williams J. S. ,
+    Rudy A. , Gustafson E. J. & Mladenoff D. J. (2007) Design,
+    development, and application of LANDIS-II, a spatial landscape
+    simulation model with flexible temporal and spatial resolution.
+    Ecological Modelling 201(3-4): 409--419.
+    <https://doi.org/10.1016/j.ecolmodel.2006.10.009>
+-   de Bruijn A., Gustafson E.J. , Sturtevant B.R., Foster J.R., Miranda
+    B.R., Lichti N.I. & Jacobs D.F. (2014) Toward more robust
+    projections of forest landscape dynamics under novel environmental
+    conditions: Embedding PnET within LANDIS-II. Ecological Modelling
+    (287): 44--57 <https://doi.org/10.1016/j.ecolmodel.2014.05.004>
+-   Petter G., Mairota P., Albrich K., Bebi P., Brůna J., Bugmann H.,
+    Haffenden A., Scheller R.M., Schmatz D.R., Seidl R., Speich M.,
+    Vacchiano G. & Lischke H. (2020) How robust are future projections
+    of forest landscape dynamics? Insights from a systematic comparison
+    of four forest landscape models. Environmental Modelling & Software
+    134: 104844. <https://doi.org/10.1016/j.envsoft.2020.104844>
     (supplementary material)
--   User guide for PNET 4.1 online -
-    <https://github.com/LANDIS-II-Foundation/Extension-PnET-Succession/blob/master/deploy/docs/LANDIS-II%20PnET-Succession%20v4.1%20User%20Guide.pdf>
+-   Gustafson E.J., De Bruijn A.M.G., Pangle R.E., Limousin J.M.,
+    Mcdowell N.G., Pockman W.T., Sturtevant B.R., Muss J.D. & Kubiske
+    M.E. (2015) Integrating ecophysiology and forest landscape models to
+    improve projections of drought effects under climate change. Global
+    Change Biology 21: 843--856. <https://doi.org/10.1111/gcb.12713>\
 

### 2023-04-14 manuscript/supplementaryRW.Rmd
added landclim report to supplementary

@@
     landscapes. For Ecol Manag 144:43--65.
-    doi:10.1016/S0378-1127(00)00386-8
+    [doi:10.1016/S0378-1127(00)00386-8](doi:10.1016/S0378-1127(00)00386-8){.uri}
 
@@
     accuracy and generality of a hybrid patch model. Tree Physiol
-    25:939--951. doi:10.1093/treephys/25.7.939
+    25:939--951. <doi:10.1093/treephys/25.7.939>
 
@@
     protection in mountain forests. J Environ Manag 87:373--388.
-    doi:10.1016/j.jenvman.2007.01.031
+    <doi:10.1016/j.jenvman.2007.01.031>
 
@@
 
+Olalla Díaz-Yáñez, Harald Bugmann
+
 ### Model
 
-*Please provide here the references to this model where it is possible
-to understand the model structure specially for the regeneration module,
-this could be a published paper, a website with the model description, a
-manual or some other document that the reader can access.*
-
-The most complete reference resource for the LandClim model and
-regeneration module is:
-
--
+We use the LandClim version 2.0, which calculates forest dynamics at a
+decadal time step. The complete reference resource for the LandClim
+model and regeneration module are:
+
+-   SCHUMACHER, S., H. BUGMANN, AND D. J. MLADENOFF. (2004). "Improving
+    the formulation of tree growth and succession in a spatially
+    explicit landscape model" Ecological Modelling 180: 175-194.
+
+-   SCHUMACHER,  S. AND H. BUGMANN. (2006). "The relative importance of
+    climatic effects, wildfires and management for future forest
+    landscape dynamics in the Swiss Alps" Global Change Biology 12:
+    1435-1450.
+
+-   SCHUMACHER, S., B. REINEKING, J. SIBOLD, AND H. BUGMANN. (2006).
+    "Modeling the impact of climate and vegetation on fire regimes in
+    mountain landscapes" Landscape Ecology 21: 539-554.
+
+-   Manual available on the model web page:
+    <https://ites-fe.ethz.ch/openaccess/products/landclim>
 
@@
 
+Climate inputs for LandClim are at monthly temporal resolution. We used
+monthly climate from the CHELSA dataset provided in the protocol using
+temperature and precipitation variables. -during the simulations, we
+sampled the climate time series randomly.
+
 ### Soil
 
+We used the soil quality provided to infer the Available Water Holding
+Capacity needed in LandClim. We assigned 8 cm for soil quality one and
+25 cm for soil quality five and interpolated the values for the other
+categories linearly between the two extremes.
+
 ### Topography
 
+We used the altitude and soil information provided in the protocol to
+create the different landscapes, one per site. However, each landscape
+created per site had no slope, altitude, or aspect variation.
+
 ### Tree species
 
+We simulated forest dynamics only considering the 11 species proposed in
+the protocol: *Fagus sylvatica, Picea abies, Abies alba, Carpinus
+betulus, Tilia cordata, Acer pseudoplatanus, Betula* spp.*, Fraxinus
+excelsior, Quercus* spp.*, Alnus glutinosa,* and *Pinus sylvestris.*
+
 ### Simulations
 
+We simulate forest dynamics from bare ground for each of the sites. We
+prepared landscapes of 250x250m without topographical elevation for each
+of the sites. The simulations we run from bare ground to 2000 years and
+the samples were taken from the year 2000. On the regeneration module,
+we use a background probability of 0.001, a dispersal mode based on the
+formulation random asymmetric, and the same establishment probability
+for all the species included in the simulation of 0.9.
+
 ### Outputs
+
+We estimated the outputs based on a sample of 25 random cells every ten
+years from 2000 to 4000 of the simulation for one simulation per site,
+providing 200 samples per site.
 

### 2023-04-21 manuscript/manuscriptRW.Rmd
update acknowlegments with CH information
changes in the manuscript based on the comments from Mats and Christopher

@@
 - Werner Rammer,
+- Christopher P.O. Reyer,
 - Simon	Scheiter,
@@
 
-Despite sometimes considerable mismatches between simulation results and
-data, it is remarkable that most of the models capture the essential
-features of the highly complex process of tree establishment, while not
-having been parameterized with such data. We conclude that in spite of
-all the uncertainties, a lot can be gained by a focus on the modeling of
-establishment processes in these models. This has the potential to
-render long-term projections of forest dynamics under changing
-environmental conditions much more robust.
+Despite considerable mismatches between simulation results and data in
+some instances, it is remarkable that most of the models capture the
+essential features of the highly complex process of tree establishment,
+while not having been parameterized with such data. We conclude that in
+spite of all the uncertainties, a lot can be gained by a focus on the
+modeling of establishment processes in these models. This has the
+potential to render long-term projections of forest dynamics under
+changing environmental conditions much more robust.
 
@@
 importance to humankind [@foresteurope2020]. Under the pressure of
-climate change and increasing disturbances, it is becoming ever more
-important to understand how forest structure, composition and function
-will change, and to evaluate forest capacity to adapt to or be resilient
-after disturbance [@lindner2010; @seidl2022]. A wide range of models of
-forest dynamics were developed over the past decades considering the
-impacts of climate on forest dynamics [@bugmann2022]. From these
-studies, it is evident that we have a reasonably good understanding how
-to model tree growth [@vanclay1997; @bugmann1996], and substantial
-efforts have been dedicated to improving the representation of tree
-mortality [@bugmann2019; @cailleret2017]. In contrast, tree
-establishment is a much less studied process that is often represented
-rather coarsely in these models [@walck2011; @leishman1992; @price2001],
-which is a clear research gap in the context of climate-induced forest
-disturbances and forest resilience.
+climate change, increasing disturbances and changing societal demands on
+forests ecosystems services, it is becoming ever more important to
+understand how forest structure, composition and function will change,
+and to evaluate forest capacity to adapt to or be resilient after
+disturbance [@lindner2010; @seidl2022]. A wide range of models of forest
+dynamics were developed over the past decades considering the impacts of
+climate on forest dynamics [@bugmann2022]. From these studies, it is
+evident that we have a reasonably good understanding how to model tree
+growth [@vanclay1997; @bugmann1996], and substantial efforts have been
+dedicated to improving the representation of tree mortality
+[@bugmann2019; @cailleret2017]. In contrast, tree establishment is a
+much less studied process that is often represented rather coarsely in
+these models [@walck2011; @leishman1992; @price2001], which is a clear
+research gap in the context of climate-induced forest disturbances and
+forest resilience.
 
@@
 Ingrowth thresholds for the sites differed between diameters of 0 and 10
-cm. For our study, we defined two data sets where one included 165 sites
+cm. For our study, we defined two datasets where one included 165 sites
 with a diameter threshold of 7 cm or lower, and the other included
@@
 threshold of 7 or 10 cm, respectively (synonym: ingrowth). To this end,
-each modeling group was provided with a detailed protocol [@díaz-yáñez]
-with instructions how to perform the simulations, providing input
-variables on climate and soil conditions, and the expected output
-variables. Neither were further site information (except for the data
-specified below) nor any data on tree ingrowth or forest stand features
-provided prior to the simulation. That is, the models were run in "blind
-flight" mode.
+each modeling group was provided with a detailed protocol
+[@díaz-yáñez2022] with instructions how to perform the simulations,
+providing input variables on climate and soil conditions, and the
+expected output variables. Neither were further site information (except
+for the data specified below) nor any data on tree ingrowth or forest
+stand features provided prior to the simulation. That is, the models
+were run in "blind flight" mode.
 
@@
 
-Three models did not provide results from all the simulated samples or
+Two models did not provide results from all the simulated samples or
 sites to avoid unrealistic results, as follows. In the model 4C, a
@@
 ($BA_{n,i}$) of that species $i$ relative to the total basal area
-($TotalR_{BA_{n}}$) of the sample $n$.
+($TotalR_{BA_{n}}$) of the sample $n$. The species diversity was not
+assessed for the model aDGVM2 because it does not simulate individual
+species.
 
@@
 ```
-We assessed model preformance in relation to model traits focusing on
+We assessed model performance in relation to model traits focusing on
 (1) model complexity as defined by @bugmann2022, (2) model type
@@
 regarding the species-specific ingrowth levels in the environmental
-space of Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
+space (Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
 tended to consistently overestimate the ingrowth share of all five
@@
 primeval forests at demographic equilibrium, @brzeziecki2021 found
-higher ingrowth rates than the ones observed in the Euforia data. Thus,
-the overestimate of ingrowth rates by the models may not be so
+higher ingrowth rates than the ones observed in the EuFoRia data. Thus,
+the overestimation of ingrowth rates by the models may not be so
 problematic. Second, the data were collected for ingrowth above a 7 cm
@@
 has led to the findings presented in this paper. The protocol is
-available here @díaz-yáñez . The code and data that support the findings
-of this study are openly available at zenodo (*link to be added*) and
-gihub (*link to be added*).
+available here @díaz-yáñez2022 . The code and data that support the
+findings of this study are openly available at zenodo (*link to be
+added*) and gihub (*link to be added*).
 
@@
 ITMS2014+313011W580 supported by the Integrated Infrastructure
-Operational Programme funded by the ERDF.
+Operational Programme funded by the ERDF. CH acknowledges financial
+support from Waldfonds WASIM2100 project.
 

### 2023-04-25 code/figures.R
changes in the map figure such as italic kegend and lss spp in legend

@@ -446,7 +446,7 @@
     
     
     xx <- ggpubr::ggarrange(sm, sl,
-                            labels = c("a", "b"),
+                            labels = c("A", "B"),
                             ncol = 1, nrow = 2, heights = c(3, 3), align = "v")
     
     xx <- ggpubr::annotate_figure(xx, 
@@ -741,10 +741,10 @@
                                              color = NULL),
                                 nudge_y = 0.05) +
       ggplot2::scale_fill_manual(values = values_color) +
-      ggplot2::annotate("text", x = 10, y = 2.6, label = "a") + 
-      ggplot2::annotate("text", x = 10, y = 1.6, label = "b") + 
-      ggplot2::annotate("text", x = -4.6, y = 1.6, label = "c") + 
-      ggplot2::annotate("text", x = -4.6, y = 2.6, label = "d") + 
+      ggplot2::annotate("text", x = 10, y = 2.6, label = "A") + 
+      ggplot2::annotate("text", x = 10, y = 1.6, label = "B") + 
+      ggplot2::annotate("text", x = -4.6, y = 1.6, label = "C") + 
+      ggplot2::annotate("text", x = -4.6, y = 2.6, label = "D") + 
     
       ggplot2::annotate("text", x = 8.5, y = 1.81, label = "Reineke self-thinning") + 
       ggplot2::annotate("text", x = 7.9, y = 1.03, label = "Ratio equal to 1")
@@ -1491,7 +1491,6 @@
                  dplyr::group_by(site) |>  
                  dplyr::filter(dplyr::row_number() == 1)
     
-    
     recruitSpp <- obsSpp |> dplyr::group_by(site, dbh) |> dplyr::top_n(1, r.ba)
     recruitSpp <- recruitSpp[, c("site", "species", "dbh")]
     colnames(recruitSpp) <- c("site", "recruitment", "dbh")
@@ -1505,50 +1504,70 @@
       dplyr::group_by(site) |>  
       dplyr::filter(dplyr::row_number() == 1)
     
+    # Check if sites that had multiple samples
+    # had different dominant spp in recruitment
+    recruitSpp1 <- obsSpp |>  
+      dplyr::filter(r.ba > 0) |> # do not consider no recruitment 
+      dplyr::group_by(site, dbh, sample) |> 
+      dplyr::top_n(1, r.ba) #select with higher r.ba
+    # Select only dbh 10
[…]

### 2023-04-25 code/figures.R
dashed line to figure 4

@@ -148,7 +148,7 @@
         ggplot2::scale_alpha_manual(values = c(0.3, 1),
                                     guide = ggplot2::guide_legend(override.aes = list(alpha = c(0.4, 1),
                                                                              fill = "red")),
-                                    name = "Diameter threshold",
+                                    #name = "Diameter threshold",
                                     labels = c('7 cm', '10 cm')) +
         ggplot2::geom_hline(ggplot2::aes(yintercept = quantile(overDTsim$r.trees[overDTsim$dbh == "7" & overDTsim$model == "Observed"],
                                                                c(0.25),
@@ -161,7 +161,7 @@
     
     ggplot2::ggsave("figures/recOverAll710.png",
                     plot =  recOverAll,
-                    width = 21, height = 12,
+                    width = 21, height = 12, scale = 0.9,
                     dpi = 300, units = "cm", 
                     device = 'png') 
     
@@ -201,7 +201,7 @@
         ggplot2::scale_alpha_manual(values = c(0.3, 1),
                                     guide = ggplot2::guide_legend(override.aes = list(alpha = c(0.4, 1),
                                                                              fill = "red")),
-                                    name = "Diameter threshold",
+                                    #name = "Diameter threshold",
                                     labels = c('7 cm', '10 cm')) +
         ggplot2::geom_hline(ggplot2::aes(yintercept = quantile(dat710$ShannonIndexRecruit[dat710$dbh == "7" & dat710$model == "Observed"],
                                                                c(0.25))), 
@@ -382,10 +382,10 @@
     # Plot stand observed VS. simulated
     selModels2 <-  c("ForClim 11", "iLand", "4C")
     plotData3 <- plotShannonDF2[plotShannonDF2$model %in% selModels2 ,]
-    plotData3 <-  plotData3[plotData3$dbh == "7", ]
+    plotData3 <- plotData3[plotData3$dbh == "7", ]
     
     labels2 <- c(paste0("Overpredict", "\n(n=4)"),
-                 paste0("", "\n(n=7)"),
+                 paste0("Intermediate", "\n(n=7)"),
                  paste0("Underpredict", "\n(n=3)"))
     
     plotData3$model <- factor(plotData3$model, levels = selModels2, labels = labels2)
@@ -408,7 +408,7 @@
         ggplot2::labs(y = bquote(bar(H) * " stand"),
                       x = bquote(bar(H) * " stand")) +
         ggplot2::facet_wrap(~model) +
[…]

### 2023-04-25 code/figures.R
change ylab in ratio7_10 to match next figure ylab that is the same

@@ -655,7 +655,8 @@
                           linetype = "dashed") +
       ggplot2::ylim(c(0,4)) +
       ggplot2::xlab(label = "") +
-      ggplot2::ylab(label = bquote(bar(R) * "(7cm) / " * bar(R) * "(10cm)")) +
+      ggplot2::ylab(label = "Ratio of recruitment (7 and 10 cm)") +
+                    # bquote(bar(R) * "(7cm) / " * bar(R) * "(10cm)")) +
       ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90),
                      strip.text.y = ggplot2::element_text(angle = 0),
                      legend.position = "none",

### 2023-04-25 manuscript/manuscriptRW.Rmd
update references table 1

@@
 nocite: |
-  @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @langan2017, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2004, @morin2021, @lasch-born2020, @hickler2012, @fischer_lessons_2016
+  @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @langan2017, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2007, @morin2021, @lasch-born2020, @hickler2012, @fischer_lessons_2016, @langan2017, @scheiter2013
 linkcolor: blue
@@
 compensated by some models with high simulated mortality rates in the
-early stages of population dynamics. Simulated species diversity of
-ingrowth matches the observed ranges quite well. Models simulating
+early stages of individual trees dynamics. Simulated species diversity
+of ingrowth matches the observed ranges quite well. Models simulating
 higher species diversity at the stand level do not feature higher
@@
 more important than scale (stand, landscape, global) and approach
-(empirical, process-based) to accurately capture tree establishment.
+(empirical, process-based) for accurately capturing tree establishment.
 
@@
 while not having been parameterized with such data. We conclude that in
-spite of all the uncertainties, a lot can be gained by a focus on the
-modeling of establishment processes in these models. This has the
-potential to render long-term projections of forest dynamics under
-changing environmental conditions much more robust.
+spite of all the uncertainties, much can be gained by evaluating and
+refining model establishment processes. This has the potential to render
+long-term projections of forest dynamics under changing environmental
+conditions much more robust.
 
@@
 importance to humankind [@foresteurope2020]. Under the pressure of
-climate change, increasing disturbances and changing societal demands on
-forests ecosystems services, it is becoming ever more important to
-understand how forest structure, composition and function will change,
-and to evaluate forest capacity to adapt to or be resilient after
-disturbance [@lindner2010; @seidl2022]. A wide range of models of forest
-dynamics were developed over the past decades considering the impacts of
-climate on forest dynamics [@bugmann2022]. From these studies, it is
-evident that we have a reasonably good understanding how to model tree
-growth [@vanclay1997; @bugmann1996], and substantial efforts have been
-dedicated to improving the representation of tree mortality
-[@bugmann2019; @cailleret2017]. In contrast, tree establishment is a
-much less studied process that is often represented rather coarsely in
-these models [@walck2011; @leishman1992; @price2001], which is a clear
-research gap in the context of climate-induced forest disturbances and
-forest resilience.
+climate change, increasing disturbances impacts and changing societal
+demands on forests ecosystems services, it is becoming ever more
+important to understand how forest structure, composition and function
+will change, and to evaluate forest capacity to adapt to or be resilient
+after disturbance [@lindner2010; @seidl2022]. A wide range of models of
+forest dynamics were developed over the past decades considering the
+impacts of climate on forest dynamics [@bugmann2022]. From these
+studies, it is evident that we have a reasonably good understanding how
+to model tree growth [@vanclay1997; @bugmann1996], and substantial
+efforts have been dedicated to improving the representation of tree
+mortality [@bugmann2019; @cailleret2017]. In contrast, tree
+establishment is a much less studied process that is often represented
+rather coarsely in these models [@walck2011; @leishman1992; @price2001],
+which is a clear research gap in the context of climate-induced forest
+disturbances and forest resilience.
 
@@
 models are handled in a multitude of ways [@konig_tree_2022;
-@bugmann2022]: from a) entirely ignoring it [as done in classical forest
-growth models, e.g., Pretzsch et al. -@pretzsch2002], across b) the use
+@bugmann2022]: from 1) entirely ignoring it [as done in classical forest
+growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2) the use
 of a few simple environmental filters, as done in most forest "gap"
 models [@shuman2011] and Dynamic Global Vegetation Models [e.g.,
-@hickler2012; @smith2001], to c) complex approaches that incorporate
+@hickler2012; @smith2001], to 3) complex approaches that incorporate
 local feedback from the canopy, multiple ecological processes and often
-also small time steps [e.g., @seidl2012; @wehrli2006], or d) field-based
+also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
 statistical parameterizations, which however are not easy to extrapolate
@@
 Overall, models are needed to (1) synthesize existing empirical data and
-explore their consequences, (2) assess future tree establishment e.g. in
-the context of global change scenarios, and (3) identify the most
+explore their relationships, (2) assess future tree establishment e.g.
+in the context of global change scenarios, and (3) identify the most
 important processes that are shaping ecological patterns. Given the
@@
 "classical" forest gap models [@gutiérrez2016], at least as long as the
-simulations comprise a multitude of species. Some models use calibration
-against local canopy-level data to define appropriate ingrowth levels,
-which is likely to strongly hamper the general applicability of these
-models e.g. under scenarios of climate change. Furthermore, correctly
-capturing the species composition in tree establishment as a function of
-the presence of seed trees in the canopy is often a particular
-challenge, potentially leading to unrealistic successional drift in the
-model, which must be corrected by factors that are hard to parameterize
-[e.g., @lischke2006]. A related issue is the excessive reduction of
-species diversity due to positive feedback effects such that eventually
-just single-species stands remain. This is sometimes corrected by the
-incorporation of a low level of seed influx of all species at all times
-[@schumacher2006] or by restricting the number of seeds per species in
-the seedbank [@lischke2006]. However, simulated species composition is
-usually exceedingly sensitive to such a formulations, whereas the
-parameters of such functions are poorly constrained by field data.
-Lastly, there is often a problem with insufficient observational
-constraints on parameter values for models that start from very small
-tree sizes (e.g., 10 cm of height or even seeds) and track tree
-development in a process-oriented manner by considering a multitude of
-ecological influences, rather than emphasizing tree establishment into a
-larger size class [e.g., @wehrli2006].
+simulations set-up comprise a multitude of species. Some models use
+calibration against local canopy-level data to define appropriate
+ingrowth levels, which is likely to strongly hamper the general
+applicability of these models e.g. under scenarios of climate change.
+Furthermore, correctly capturing the species composition in tree
+establishment as a function of the presence of seed trees in the canopy
+is often a particular challenge, potentially leading to unrealistic
+successional drift in the model, which must be corrected by factors that
+are hard to parameterize [e.g., @lischke2006]. A related issue is the
+excessive reduction of species diversity due to positive feedback
+effects such that eventually just single-species stands remain. This is
+sometimes corrected by the incorporation of a low level of seed influx
+of all species at all times [@schumacher2006] or by restricting the
+number of seeds per species in the seedbank [@lischke2006]. However,
+simulated species composition is usually exceedingly sensitive to
+assumptions about seed availability, whereas the parameters of such
+functions are poorly constrained by field data. Lastly, there is often a
+problem with insufficient observational constraints on parameter values
+for models that start from very small tree sizes (e.g., 10 cm of height
+or even seeds) and track tree development in a process-oriented manner
+by considering a multitude of ecological influences, rather than
+emphasizing tree establishment into a larger size class [e.g.,
+@wehrli2006].
 
@@
 [@hanbury-brown2022], although many sub-processes have been studied in
-detail. Yet, little data is available that covers all the processes
-within one species along environmental gradients, let alone for a vast
-suite of species. Thus, the available data are fragmented, which
-constitutes a major problem for model building [@clark1999]. On the one
-hand, monitoring on permanent sample plots (such as National Forest
-Inventories) often measures tree establishment [@zell2019], however with
-a design that captures the rate after the stand initiation phase
-[@hallsby2015]. Data from permanent plots, although available across
-regions, are highly heterogeneous, rendering their use difficult in a
-modeling context [@käber2021]. On the other hand, targeted studies to
-measure tree establishment on experimental sites are often limited in
+detail [@miina2006]. Yet, little data is available that covers all the
+processes within one species along environmental gradients, let alone
+for a vast suite of species. Thus, the available data are fragmented,
+which constitutes a major problem for model building [@clark1999]. On
+the one hand, monitoring on permanent sample plots (such as National
+Forest Inventories) often measures tree establishment [@zell2019],
+however with a design that captures the rate after the stand initiation
+phase [@hallsby2015]. Data from permanent plots, although available
+across regions, are highly heterogeneous, rendering their use difficult
+in a modeling context [@käber2021]. On the other hand, targeted studies
+to measure tree establishment on experimental sites are often limited in
 spatial extent, species studied, or the sub-set of processes that are
@@
 species against empirical data along a wide environmental gradient of
-temperature, moisture, and light availability.
+temperature, moisture, and light availability [@grubb1977].
 
@@
 and seedling growth, which ultimately lead to the simulated number of
-established trees. Recruitment models, in contrast, introduce a certain
-number of new trees with certain characteristics such as biomass or
-diameter, without explicitly considering earlier development processes.
-We can further distinguish models that feature a feedback in the
-establishment, i.e., where the level and species composition of
-establishment is influenced by the existing adult tree community via the
-production of seeds, seedlings or saplings, from models that do not
-contain such a feedback.
+established trees. Recruitment models, in contrast, introduce a number
+of new trees with certain characteristics such as biomass or diameter,
+without explicitly considering earlier development processes. We can
+further distinguish models that feature a feedback in the establishment,
+i.e., where the level and species composition of establishment is
+influenced by the existing adult tree community via the production of
+seeds, seedlings or saplings, from models that do not contain such a
+feedback.
 
@@
 range of environmental gradients in temperature and precipitation in
-central Europe as compiled in the framework of the EuFoRIa network
+Central Europe as compiled in the framework of the EuFoRIa network
 [@euforia2019] (Figure \@ref(fig:map)). These forests have been
@@
 We selected 200 sites from this network as the benchmarking dataset for
-the simulation. The selection of sites is representative for the
+the simulation. The selection of sites is representative of the
 environmental variation of the data. This was achieved by applying
@@
 each modeling group was provided with a detailed protocol
-[@díaz-yáñez2022] with instructions how to perform the simulations,
-providing input variables on climate and soil conditions, and the
-expected output variables. Neither were further site information (except
-for the data specified below) nor any data on tree ingrowth or forest
-stand features provided prior to the simulation. That is, the models
-were run in "blind flight" mode.
-
-The input variables were collected and aggregated to be adapted to the
-needs of the different models. Time series of climatic variables from
-1981 to 2018 were provided in hourly [Era5-land data, Muñoz Sabater
--@muñozsabater2019], daily and monthly resolution [CHELSA data, Karger
-et al. -@karger2021]. Some variables required for some models, such as
-relative humidity or vapor pressure deficit, were calculated from these
-variables. The final strategy how to use the climate data was different
-in each model, based on the approach that best suited the model (Table
-\@ref(tab:tableModels)). Soil quality data were provided as continuous
-values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
-The protocol also provided the elevation, slope and aspect for each of
-the 200 sites [@asterscienceteam2019], but no other spatial information
-such as coordinates, with the exception of iLand and aDGVM2, which
-required blurred coordinates to derive highly detailed soil data.
+[@díaz-yáñez2022] with instructions for how to perform the simulations,
+providing input variables on climate and soil conditions, and the list
+of expected output variables. Neither were further site information
+(except for the data specified below) nor any data on tree ingrowth or
+forest stand features provided prior to the simulation. That is, the
+models were run in "blind flight" mode.
+
+The input variables were collected from different data sources and
+aggregated to be adapted to the needs of the different models. Time
+series of climatic variables from 1981 to 2018 were provided in hourly
+[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
+resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
+required for some models, such as relative humidity or vapor pressure
+deficit, were calculated from these variables. The final instructions
+for using the climate data was different in each model, based on the
+approach that best suited the model (Table \@ref(tab:tableModels)). Soil
+quality data were provided as continuous values between 1 and 5
+[Soilgrids dataset, Hengl et al. -@hengl2017]. The protocol also
+provided the elevation, slope and aspect for each of the 200 sites
+[@asterscienceteam2019], but no other spatial information such as
+coordinates, with the exception of iLand and aDGVM2, which required
+blurred coordinates to derive highly detailed soil data.
 
@@
 for which ingrowth data of sufficient quality were available from
-EuFoRIa: *Fagus sylvatica, Picea abies, Abies alba, Carpinus betulus,
-Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus excelsior,
-Quercus spp., Alnus glutinosa, Pinus sylvestris.* The same set of
-species was used at all 200 sites. Two models included their standard
-set of species for the simulations, which is much larger (i.e., ForClim
-1, ForClim 11, TreeMig). In three models, fewer than these eleven
-species were simulated (4C, xComp, LPJ-GUESS) (Table
-\@ref(tab:tableModels)). In 4C only *Fagus sylvatica*, *Picea abies*,
-*Betula* spp., *Quercus* spp. and *Pinus sylvestris* are parameterised.
-XComp simulations did not consider *Fraxinus excelsior* due to a
-depreciated species parameterisation. In LPJ-GUESS, *Acer
+EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
+betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
+Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus
+sylvestris L.* The same set of species was used at all 200 sites. Two
+models included their standard set of species for the simulations, which
+is much larger (i.e., ForClim 1, ForClim 11, TreeMig). In three models,
+fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS)
+(Table \@ref(tab:tableModels)). In 4C only *Fagus sylvatica*, *Picea
+abies*, *Betula* spp., *Quercus* spp. and *Pinus sylvestris* are
+parameterised. XComp simulations did not consider *Fraxinus excelsior*
+due to a depreciated species parameterisation. In LPJ-GUESS, *Acer
 pseudoplatanus* and *Alnus glutinosa* are not parameterized and
@@
 
-Each of the models reported the ingrowth number by sampling 200 times
-over a 10-year interval for each species and per ha for each of the 200
+Each of the models reported the ingrowth number by sampling 200 times in
+a 10-year interval for each species and per ha for each of the 200
 sites. Multiple samples per site were used to better understand the
@@
 
-```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network used in the present study, the colors show the dominant tree species per site in the observed ingrowth at the 10 cm threshold.", out.width='80%'}
+```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed ingrowth at the 10 cm threshold.", out.width='80%'}
 knitr::include_graphics(here::here("figures", "theMap.png"))
@@
 of the ingrowth, (3) ingrowth mortality, (4) model performance and model
-traits, and (5) ingrowth gradients on regeneration niches. We evaluated
-the species diversity across models and in relation to the observed data
-by calculating the Shannon index based on the relative proportion of the
-species in terms of basal area. The Shannon index was calculated for the
-establishment ($H_{R_{n}}$) (Equation \ref{eqpiShannonR}) and at the
-stand level ($H_{S_{n,s}}$) (Equation \ref{eqpiShannon}). The higher the
-value of the index, the higher the species diversity at a particular
-site and sample.
+traits, and (5) ingrowth gradients along regeneration niches. We
+evaluated the species diversity across models and in relation to the
+observed data by calculating the Shannon index based on the relative
+proportion of the species in terms of basal area. The Shannon index was
+calculated for the establishment ($H_{R_{n}}$) (Equation
+\ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation
+\ref{eqpiShannon}). The higher the value of the index, the higher the
+species diversity at a particular site and sample.
 
@@
 where $s$ is the total number of species present and that have a basal
-area larger than zero in sample $n$; $p_{n,r,i}$ is the proportion of
+area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
 the species $i$ in sample $n$ calculated as the ingrowth basal area
@@
 \begin{equation}
-Reineke = (7/10)^{(-1.605)} \label{eqReineke}
+Reineke = \frac{7}{10}^{-1.605} \label{eqReineke}
 \end{equation}
@@
 (1) model complexity as defined by @bugmann2022, (2) model type
-(empirical or process based), (3) the presence of absence of a canopy
+(empirical or process based), (3) the presence or absence of a canopy
 feedback for ingrowth, and (4) the scale of application of the model
@@
 is the mean basal area ingrowth of the species $i$ per site $t$ across
-the available samples at that site.
-
-We removed 35 sites from the comparisons at a threshold of 7 cm
-ingrowth, because field data did not include these measurements.
-Furthermore, regarding the share in the ingrowth basal area per species
-($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the
-ingrowth basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero but
-also when both the ingrowth basal area of that species
-($\bar{r}_{BA_{t,i}}$) and the total ingrowth basal area (
-$\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.
+the available samples at that site. Furthermore, regarding the share in
+the ingrowth basal area per species ($\bar{R}_{BA_{share_{i}}}$), we
+categorized this as zero when the ingrowth basal area of that species
+($\bar{r}_{BA_{t,i}}$) was zero but also when both the ingrowth basal
+area of that species ($\bar{r}_{BA_{t,i}}$) and the total ingrowth basal
+area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.
 
@@
 
-The majority of the models overpredicted ingrowth diversity; only two of
-them underestimated ingrowth diversity (4C and Landis-II) (Figure
+The majority of the models overpredicted ingrowth diversity; as already
+observed in figure \@ref(fig:H710), only two of them underestimated
+ingrowth diversity (4C and Landis-II) (Figure
 \@ref(fig:combinedRichness)B and Figure S4). Both the models with
@@
 ```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site for observed and simulated data.
+Each plot shows one pattern represented for one examplary model of each category (overperdicted, intermediate and underpredicted).
+The full data with the models grouping are shown in Figures S3 and S4.
 n indicates the number of models falling in each group.
 A) Examples for the three trends across models for species diversity at the stand level.
-B) Examples for the three trends across models for ingrowth at the 7 cm threshold.
-The full data are shown in Figures S3 and S4.", out.width='75%'}
+B) Examples for the three trends across models for ingrowth at the 7 cm threshold.", out.width='75%'}
 knitr::include_graphics(here::here("figures", "combined_richness.png"))
@@
 coupled to an increase of the mortality between 7 and 10 cm). While some
-models showed a negative effect, this linear trend was not pronounced
+models showed a negative effect, their linear trend was not pronounced
 and not always significant regarding the slope.
@@
 change. Correctly capturing tree ingrowth in dynamic models is
-fundamental to, for example, evaluate post-disturbance dynamics as it
-will define the forest state in the first decades [@seidl2022]. If this
-initial phase is not captured well, we can not properly assess aspects
-such as forest resilience or the timing, magnitude and progression of
-carbon sequestration. In this study, for the first time, projections of
-tree establishment from multiple models of forest dynamics were
-confronted with a unique dataset from unmanaged forest reserves across a
-large environmental gradient in Europe.
+fundamental to, for example, evaluate post-disturbance dynamics and
+potential long-term recovery trajectories as it will define the forest
+state in the first decades [@seidl2022]. If this initial phase is not
+captured well, we can not properly assess aspects such as forest
+resilience or the timing, magnitude and progression of carbon
+sequestration. In this study, for the first time, projections of tree
+establishment from multiple models of forest dynamics were confronted
+with a unique dataset from unmanaged forest reserves across a large
+environmental gradient in Europe.
 
 The observed data that we used [@käber2022] are exceptional,
-particularly with respect to the number of events captured (number of
+particularly with respect to the number of records captured (number of
 sites and repeated measurements), which is essential for a highly
@@
 the individual models consider establishment processes
-[@konig_tree_2022], e.g., in the absence of adults, ingrowth can be
-linked to either seed rain from dispersal or a seed/seedling bank;
-neither process is incorporated across many models studied here,
-however.
+[@konig_tree_2022], e.g., without considering the feedback from adults,
+ingrowth can be linked to either seed rain from dispersal or a
+seed/seedling bank; neither process is incorporated across many models
+studied here, however.
 
@@
 problems already under current climatic conditions. For the other
-models, whose establishment module is not strongly constrained by
+models, whose establishment modules are not strongly constrained by
 empirical data, the degrees of freedom that are available for modeling
@@
 necessarily lead to better projections and rather to a reduced
-transparency and predictive power [@franklin2020]. Thus, the question
+transparency and predictive power [@franklin2020]. Thus, the question of
 what level of detail is appropriate and parsimonious for modeling tree
@@
 
-There were pronounced differences how the main tree species were
+There were pronounced differences in how the main tree species were
 represented by the models along the environmental gradients, in
-particular the dominant species Abies alba, Fagus sylvatica and Picea
-abies. The ingrowth levels were sampled from simulations in the
+particular the dominant species *Abies alba*, *Fagus sylvatica* and
+*Picea abies*. The ingrowth levels were sampled from simulations in the
 equilibrium, and in this state it is expected that non-climax species
@@
 regeneration niche of the species are needed, but this cannot be done in
-the absence of robust datasets.
+the absence of robust datasets across multiple stand development stages.
 
@@
 be at a disadvantage due to the limited spatial considerations [but cf.
-Snell et al. -@snell2014], as they are usually lacking dispersal between
-cells and are based on a strong abstraction of horizontal space
-[@hanbury-brown2022]. Lastly, detailed data on soil conditions were not
-available from the observed data, and independent, admittedly rather
-coarse data for soil and climatic water balance were used instead. It is
-noteworthy that several models represent drought based on more detailed
-indicators based e.g., on soil water holding capacity, which had to be
-derived from a rough soil quality measure. This may at least partially
-explain the poor performance of the models against the drought axis
-(i.e., climatic water balance).
+Snell et al. -@snell2018; or Lehsten et al. -@lehsten2019], as they are
+usually lacking dispersal between cells and are based on a strong
+abstraction of horizontal space [@hanbury-brown2022]. Lastly, detailed
+data on soil conditions were not available from the observed data, and
+independent, admittedly rather coarse data for soil and climatic water
+balance were used instead. It is noteworthy that several models
+represent drought based on more detailed indicators based e.g., on soil
+water holding capacity, which had to be derived from a rough soil
+quality measure. This may at least partially explain the poor
+performance of the models against the drought axis (i.e., climatic water
+balance).
 
@@
 Based on the results of this study, it is clear that models of forest
-dynamics need a focus on tree establishment modules to complement
-earlier efforts to make them more robust regarding growth and mortality.
-It remains uncertain what level of detail is required to model
-establishment processes, and this must be addressed in future research.
-We recommend that the further development of the establishment modules
-is implemented as additional features that can be traced back, as done
-here for the variants of ForClim and ForCEEPS, and that model complexity
-and structure must always be connected with modeling objectives
-[@albrich2020]. If it should be necessary to include more detail in the
-establishment models, this will come with higher parameterisation
-efforts. This will most likely lead to lower generalization because the
-required data will have to be collected from specific locations, as
-currently there is no general establishment dataset available.
+dynamics need a focus on tree establishment modules to make them more
+robust regarding growth and mortality. It remains uncertain what level
+of detail is required to model establishment processes, and this must be
+addressed in future research. We recommend that the improvement of the
+establishment modules is implemented as additional features that can be
+traced back, as done here for the variants of ForClim and ForCEEPS, and
+that model complexity and structure must always be connected with
+modeling objectives [@albrich2020]. If it should be necessary to include
+more detail in the establishment models, this will come with higher
+parameterisation efforts. This will most likely lead to lower
+generalization because the required data will have to be collected from
+specific locations, as currently there is no general establishment
+dataset available.
 
@@
 exercises should next focus on aspects such as specific model traits and
-the ecological formulations of particular (sub-)processes, so as to
-better understand the implications of the assumptions on which the
-models are based. Furthermore, the inclusion of a wide range of models
-with different scales, type, populations structures, tree establishment
+the ecological formulations of particular (sub-)processes, to better
+understand the implications of the assumptions on which the models are
+based. Furthermore, the inclusion of a wide range of models with
+different scales, type, populations structures, tree establishment
 modules and complexity will ensure a wider benefit to the whole modeling
@@
 However, most models properly capture the diversity of the initial tree
-community, but variations in model formulations, e.g., the presence or
-absence of feedback from the adult trees, did not have a strong effect
-for capturing the species composition of ingrowth.
+community, but differences between model formulations, e.g., the
+presence or absence of feedback from the adult trees, did not have a
+strong effect for capturing the species composition of ingrowth.
 
@@
 findings of this study are openly available at zenodo (*link to be
-added*) and gihub (*link to be added*).
+added*) and github (*link to be added*).
 
@@
 disturbances, grant no. 22035418, 2019--2022) funded by the ERA-NET
-Cofund ForestValue. KM received funding from the Project No.
-CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant no.
-ITMS2014+313011W580 supported by the Integrated Infrastructure
+Cofund ForestValue. MMina acknowledges funding from the H2020-MSCA-IF
+project REINFORCE (Grant No. 891671). KM received funding from the
+Project No. CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant
+no. ITMS2014+313011W580 supported by the Integrated Infrastructure
 Operational Programme funded by the ERDF. CH acknowledges financial

### 2023-04-25 manuscript/supplementaryRW.Rmd
corrections in 4C, authors names and others

@@
 - Yannek Käber,
+- Tim Anders,
+- Kristin H. Braziunas,
+- Josef	Bruna,
+- Samuel Fischer,
+- Jessica Hetzer,
+- Thomas Hickler,
+- Christian	Hochauer,
+- Manfred J. Lexer,
+- Heike	Lischke,
 - Mats Mahnken,
-- Samuel Fischer, 
-- Xavier Morin, 
-- Manfred J. Lexer, 
-- Christian Hochauer, 
+- Paola	Mairota,
 - Katarina Merganicova,
 - Tobias Mette,
+- Xavier Morin,
 - Werner Rammer,
-- Kristin H. Braziunas, 
-- Paola Mairota, 
-- Josef Bruna,
-- Heike Lischke, 
+- Christopher P.O. Reyer,
+- Simon	Scheiter,
 - Daniel Scherrer,
-- Simon Scheiter, 
-- Thomas Hickler, 
-- Tim Anders,
-- Jessica Hetzer, 
 - Harald Bugmann
@@
 the CHELSA time series data regarding relative humidity with the model
-after Eccel et al. (2015) based on minimum and mean daily temperature to
-accomodate data needs of 4C. The missing data for January 2nd and 3rd
-2013 were filled with data from January 1st and 4th 2013 respectively.
-For construction of longer climate time series we looped through the
-provided time series from 1981 to 2018 up to the maximum simulation
-length of 2500 years. We amended situations where leap year
-inconsistencies arised by removing February 29 or duplicating data from
-February 28. Otherwise, the unaltered daily data were used to run 4C on
-a daily basis. We applied a constant atmopsheric CO2 concentration of
-380 ppm.
+after [Eccel et al. (2012)](https://doi.org/10.1002/met.258) based on
+minimum and mean daily temperature to accomodate data needs of 4C. The
+missing data for January 2nd and 3rd 2013 were filled with data from
+January 1st and 4th 2013 respectively. For construction of longer
+climate time series we looped through the provided time series from 1981
+to 2018 up to the maximum simulation length of 2500 years. We amended
+situations where leap year inconsistencies arised by removing February
+29 or duplicating data from February 28. Otherwise, the unaltered daily
+data were used to run 4C on a daily basis. We applied a constant
+atmopsheric CO2 concentration of 380 ppm.
 
@@
 sensitivity of dominant tree height at age 100 years for monospecific
-stands of all five species descibed above to the available soils using
-model simulations with 4C. The dominant tree height at age 100 years was
-then also reprojected to the range from 0 to 1 in order to derive a
-model specific soil quality indicator for all paramterized soils. Then
-we multiplied this relative model specific soil quality with the
-relative soil quality derived from the bucket size to the range from 1
-to 5. The match between paramterized soil and plot was then done by
-finding the pair of soil quality with the minimum difference between
-simuation protocol prescribed soil quality and and the soil quality of
-the paramterized soils as described above.
+stands of all five species simulated to the available soils using model
+simulations with 4C. The dominant tree height at age 100 years was then
+also reprojected to the range from 0 to 1 in order to derive a model
+specific soil quality indicator for all paramterized soils. Then we
+multiplied this relative model specific soil quality with the relative
+soil quality derived from the bucket size to the range from 1 to 5. The
+match between paramterized soil and plot was then done by finding the
+pair of soil quality with the minimum difference between simuation
+protocol prescribed soil quality and the soil quality of the
+paramterized soils as described above.
 
@@
 growth (especially light and water regime) are present. This approach
-enables regeneration of other species than only the one present in the
+enables regeneration of other species than only those present in the
 stand to establish if competition allows. The resulting recruitment
@@
 Additionally we included another subset criterion removing all samples
-that have a ba \> 100 to exclude unrealistically high samples without
-the need to exclude the full site. This resulted in discarding 0.8 % of
-the sample
+that have a ba \> 90 to exclude unrealistically high samples without the
+need to exclude the full site. This resulted in discarding 0.8 % of the
+samples.
 
@@
 
-We provide 200 samples each for 165 of the 200 plots (see reason for
-excluded plots in "Simulation length and equilibrium criteria"). These
-stem from 10 year periods that overlap partially (for \< 2500 simulation
+We provide 200 samples each for 48 plots and at least 190 samples each
+for 117 plots while not simulating 35 plots (see reason for excluded
+plots in \"Simulation length and equilibrium criteria\"). These stem
+from 10 year periods that overlap partially (for \< 2500 simulation
 years) or do not overlap (for 2500 simulation years). The number of
@@
 diameter (7/10 cm) in the 10 year sample period per hectare that have
-not already died until the sample year (Fig. 3). The recruit's basal
-area (r.ba) is the sum of the basal area per hectare of the trees that
-cross the diameter threshold that have not already died until the sample
-year. The total stand basal area (ba) is the basal area per hectare of
-all trees in the stand \> 0 cm diameter at breast height (no lower
-boundary) at the sample time per species. We provide the data for both
-diameter thresholds, 7 and 10 cm. In over 90 % of the decades no
-recruitment takes place.
+not already died until the sample year. The recruit's basal area (r.ba)
+is the sum of the basal area per hectare of the trees that cross the
+diameter threshold that have not already died until the sample year. The
+total stand basal area (ba) is the basal area per hectare of all trees
+in the stand \> 0 cm diameter at breast height (no lower boundary) at
+the sample time per species. We provide the data for both diameter
+thresholds, 7 and 10 cm. In over 90 % of the decades no recruitment
+takes place.
 
@@
 uncertainty. For example, the spatial homogeneity in the model leads to
-cyclic recruitment patterns (see Fig. 4). When the adult stands dies of,
-young individuals are recruited into the larger diameter classes because
-more light reaches lower canopy layer since larger (adult) trees do not
+cyclic recruitment patterns. When the adult stands dies of, young
+individuals are recruited into the larger diameter classes because more
+light reaches lower canopy layer since larger (adult) trees do not
 anymore prevent the light reaching lower canopy layers. This results in
@@
 
-Please noticed that we included 0 values for species missing in a given
+Please notice that we included 0 values for species missing in a given
 sample and for samples in sites where Landis did not simulate any tree.
@@
 
-Jessica Hetzer, Thoms Hickler, Tim Anders
+Tim Anders, Jessica Hetzer, Thomas Hickler
 
@@
 tasmax, tasmin) [°C] were converted to Kelvin and surface solar
-radiation (rsds) [J/m2] was converted to W/m2 by dividing rsds by 86400
+radiation (rsds) [Jm-2] was converted to W/m2 by dividing rsds by 86400
 s (3600 s/h \* 24 h). Nitrogen deposition was considered to be constant
@@
 
-All sites We used were simulated with default values for a medium
-textured soil type (sand: 0.35, clay: 0.15, silt: 0.5) in LPJ-GUESS.
+All sites were simulated with default values for a medium textured soil
+type (sand: 0.35, clay: 0.15, 823 silt: 0.5) in LPJ-GUESS.
 
@@
 dominant in some areas (Hickler et al., 2012). Mediterranean rain-green
-small shrubs with shallow roots (e.g. Lavendula spp., Rosmarinus spp.)
-and alpine/arctic shrubs (e.g. Vaccinium spp.) were represented as PFTs
-in the model. As in earlier applications of LPJ-GUESS, herbaceous
+small shrubs with shallow roots (e.g. *Lavendula* spp., *Rosmarinus*
+spp.) and alpine/arctic shrubs (e.g. Vaccinium spp.) were represented as
+PFTs in the model. As in earlier applications of LPJ-GUESS, herbaceous
 vegetation was represented by two 'generic herb' PFTs, with C3 and C4
@@
 shade-intolerant species becoming established, patch-destroying
-disturbances were allowed as described in Hickler et al. 2012. Acer
-pseudoplatanus and Alnus glutinosa are not parameterized in LPJ-GUESS
-and therefore not considered.we did not simulate these species.
+disturbances were allowed as described in Hickler et al. 2012. *Acer
+pseudoplatanus* and *Alnus glutinosa* are not parameterized in LPJ-GUESS
+and therefore not considered.
 
@@
 ### Output
-
-Please describe how have you calculated each of the variables in the
-output data.
 
@@
 "r_7cm.trees" calculated as the sum of all trees of the same species
-that cross the threshold of 7 cm DHB between 2008 and 2018 in that
+that cross the threshold of 7 cm DBH between 2008 and 2018 in that
 specific sample of that specific site. The variable "r_10cm.trees"
@@
 
-Table 2: Species/PFT characteristics and parameter of Hickler et al.
+Table 1: Species/PFT characteristics and parameter of Hickler et al.
 2012. r base respiration rate, kallm1 constant in allometry equations,
@@
 
-```{r mortality3 , echo=FALSE, fig.cap="Recruitment ration between 7 and 10cm and the initial recruitment tree number at 7cm.
-Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke selfthinning ratio under evenaged conditions", out.width='70%', fig.pos='H'}
+```{r mortality3 , echo=FALSE, fig.cap="Recruitment ratio between 7 and 10cm and the initial recruitment tree number at 7cm.
+Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke self-thinning ratio under evenaged conditions", out.width='70%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "mort7_10_trends.png"))

### 2023-04-25 manuscript/supplementaryRW.Rmd
added corrections in iLand report in the supplementary

@@
 
-### Topography
-
-### Tree species
-
-Tree composition is defined at the species level. All individually
-requested species (Fagus sylvatica, Picea abies, Abies alba, Carpinus
-betulus, Tilia cordata, Acer pseudoplatanus, Fraxinus excelsior, Alnus
-glutinosa, Pinus sylvestris) were included. For the two taxa where
-species was not defined, we identified a representative species (Betula
-pendula for Betula spp. and Quercus robur for Quercus spp.). No other
-species were included in model simulations.
-
-### Simulations
-
-#### Simulation area
-
-The simulation extent for each site was 1000 m x 1000 m (total area =
-100 ha). Environmental conditions were homogeneous for each site, and we
-did not use topographic information. Inputs for soil depth, texture, and
-fertility (plant available nitrogen) were extracted from a previously
-compiled Europe-wide dataset. We averaged values in a 1 km buffer around
+Inputs for soil depth, texture, and fertility (plant available nitrogen)
+were extracted from a previously compiled Europe-wide dataset
+(unpublished dataset). We averaged values in a 1 km buffer around
 blurred site coordinates provided by workshop organizers. Relationships
@@
 included all site variables needed for the model.
+
+### Topography
+
+Topographic information was not used.
+
+### Tree species
+
+Tree composition is defined at the species level. All individually
+requested species (Fagus sylvatica, Picea abies, Abies alba, Carpinus
+betulus, Tilia cordata, Acer pseudoplatanus, Fraxinus excelsior, Alnus
+glutinosa, Pinus sylvestris) were included. For the two taxa where
+species was not defined, we identified a representative species (Betula
+pendula for Betula spp. and Quercus robur for Quercus spp.). No other
+species were included in model simulations.
+
+### Simulations
+
+#### Simulation area
+
+The simulation extent for each site was 1000 m x 1000 m (total area =
+100 ha). Environmental conditions were homogeneous for each site.
 

### 2023-04-26 manuscript/manuscriptRW.Rmd
expading acknowledgement section
small change in the abstract

@@
 establishment at 200 sites representing large environmental gradients
-across Europe. The results are evaluated against comprehensive data from
-unmanaged forests.
+across central Europe. The results are evaluated against comprehensive
+data from unmanaged forests.
 
@@
 2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
-Funding for the workshop was provided by COST Action CA19139 PROCLIAS
-(PROcess-based models for CLimate Impact Attribution across Sectors),
-supported by COST (European Cooperation in Science and
+Several people on top of the co-authors list have been involved in the
+development of this study in several degrees that we would like to thank
+here including Martin Gutsch and Giorgio Vacchiano. Funding for the
+workshop was provided by COST Action CA19139 PROCLIAS (PROcess-based
+models for CLimate Impact Attribution across Sectors), supported by COST
+(European Cooperation in Science and
 Technology; [https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich

### 2023-04-26 manuscript/supplementaryRW.Rmd
updated LandClim report

@@
 for 117 plots while not simulating 35 plots (see reason for excluded
-plots in \"Simulation length and equilibrium criteria\"). These stem
-from 10 year periods that overlap partially (for \< 2500 simulation
-years) or do not overlap (for 2500 simulation years). The number of
-recruits (r.trees) is the sum of individuals that cross the threshold
-diameter (7/10 cm) in the 10 year sample period per hectare that have
-not already died until the sample year. The recruit's basal area (r.ba)
-is the sum of the basal area per hectare of the trees that cross the
-diameter threshold that have not already died until the sample year. The
-total stand basal area (ba) is the basal area per hectare of all trees
-in the stand \> 0 cm diameter at breast height (no lower boundary) at
-the sample time per species. We provide the data for both diameter
+plots in "Simulation length and equilibrium criteria"). These stem from
+10 year periods that overlap partially (for \< 2500 simulation years) or
+do not overlap (for 2500 simulation years). The number of recruits
+(r.trees) is the sum of individuals that cross the threshold diameter
+(7/10 cm) in the 10 year sample period per hectare that have not already
+died until the sample year. The recruit's basal area (r.ba) is the sum
+of the basal area per hectare of the trees that cross the diameter
+threshold that have not already died until the sample year. The total
+stand basal area (ba) is the basal area per hectare of all trees in the
+stand \> 0 cm diameter at breast height (no lower boundary) at the
+sample time per species. We provide the data for both diameter
 thresholds, 7 and 10 cm. In over 90 % of the decades no recruitment
@@
 
-We use the LandClim version 2.0, which calculates forest dynamics at a
-decadal time step. The complete reference resource for the LandClim
-model and regeneration module are:
+We used LandClim version 2.0, which calculates forest dynamics at a
+decadal time step. The complete reference resources for the LandClim
+model and its regeneration module are:
 
@@
 Climate inputs for LandClim are at monthly temporal resolution. We used
-monthly climate from the CHELSA dataset provided in the protocol using
-temperature and precipitation variables. -during the simulations, we
-sampled the climate time series randomly.
+monthly climate from the CHELSA dataset provided in the protocol, using
+temperature and precipitation variables. During the simulations, we
+randomly sampled the annual climate data from the CHELSA time series.
 
@@
 
-We used the soil quality provided to infer the Available Water Holding
-Capacity needed in LandClim. We assigned 8 cm for soil quality one and
-25 cm for soil quality five and interpolated the values for the other
-categories linearly between the two extremes.
+We used the soil quality provided in the protocol to infer the Available
+Water Holding Capacity (WHC) needed in LandClim. We assigned WHC = 8 cm
+for soil quality one and WHC = 25 cm for soil quality five, and
+interpolated the values for the other categories linearly between these
+two extremes.
 
@@
 We used the altitude and soil information provided in the protocol to
-create the different landscapes, one per site. However, each landscape
-created per site had no slope, altitude, or aspect variation.
+create 200 landscapes, one per site. However, each landscape created per
+site had no slope, altitude, or aspect variation.
 
@@
 
-We simulated forest dynamics only considering the 11 species proposed in
+We simulated forest dynamics considering only the 11 species proposed in
 the protocol: *Fagus sylvatica, Picea abies, Abies alba, Carpinus
@@
 
-We simulate forest dynamics from bare ground for each of the sites. We
-prepared landscapes of 250x250m without topographical elevation for each
-of the sites. The simulations we run from bare ground to 2000 years and
-the samples were taken from the year 2000. On the regeneration module,
-we use a background probability of 0.001, a dispersal mode based on the
-formulation random asymmetric, and the same establishment probability
-for all the species included in the simulation of 0.9.
+We simulated forest dynamics from bare ground for each site, using
+patches (pixels) with a size of 20x20 cm^2^. We prepared landscapes of
+250x250 m without topographical variation for each site. The simulations
+we run from bare ground to 2000 years, and the samples were taken
+starting in the year 2000. In the regeneration module, we used (1) a
+base seeding probability of 0.001, which defines if seeds are available
+due to background seed rain; (2) a dispersal mode based on the
+formulation \'random asymmetric\'; and (3) the same establishment
+probability of 0.9 for all species included in the simulation of, which
+sets the same probability for all the species to regenerate
+successfully, provided that the abiotic and light conditions are
+suitable for establishment.
 
@@
 We estimated the outputs based on a sample of 25 random cells every ten
-years from 2000 to 4000 of the simulation for one simulation per site,
-providing 200 samples per site.
+years from the simulation years 2010 to 4000 for one simulation per
+site, thus providing the required 200 samples per site.
 

### 2023-04-26 manuscript/manuscriptRW.Rmd
added authors and corrected accents in surnames

@@
 - Tim Anders,
+- Friedrich Bohn,
 - Kristin H. Braziunas,
-- Josef	Bruna,
-- Samuel Fischer,
+- Josef	Brůna,
+- Rico Fischer,
+- Samuel M. Fischer,
 - Jessica Hetzer,
@@
 - Paola	Mairota,
-- Katarina Merganicova,
+- Ján Merganič,
+- Katarina Merganičová,
 - Tobias Mette,
+- Marco Mina,
 - Xavier Morin,
@@
 
-```{r tableModels, echo=FALSE,  message=FALSE, warning=FALSE}
+```{r tableModels, echo=FALSE,  message=FALSE, warning=TRUE}
 library(knitr)
-
 modelDF <- readr::read_csv(here::here("data", "modelsDescription.csv"))

### 2023-04-26 manuscript/manuscriptRW.Rmd
mainly I added oxford commas in mansucript

@@
 We assess how well current models capture the overall abundance, species
-composition and mortality of tree establishment.
+composition, and mortality of tree establishment.
 
@@
 long-term projections of forest dynamics under changing environmental
-conditions much more robust.
+conditions that are much more robust.
 

### 2023-04-26 manuscript/supplementaryRW.Rmd
remove in-text citations from FORMIND report supplementary materials

@@
 
-```{=html}
-<!--
-@article{fischer_lessons_2016,
-    title = {Lessons learned from applying a forest gap model to understand ecosystem and carbon dynamics of complex tropical forests},
-    volume = {326},
-    issn = {03043800},
-    url = {https://linkinghub.elsevier.com/retrieve/pii/S0304380015005505},
-    doi = {10.1016/j.ecolmodel.2015.11.018},
-    language = {en},
-    urldate = {2022-09-15},
-    journal = {Ecological Modelling},
-    author = {Fischer, Rico and Bohn, Friedrich and Dantas de Paula, Mateus and Dislich, Claudia and Groeneveld, J{\"u}rgen and Guti{\'e}rrez, Alvaro G.
-and Kazmierczak, Martin and Knapp, Nikolai and Lehmann, Sebastian and Paulick, Sebastian and P{\"u}tz, Sandro and R{\"o}dig, Edna and Taubert, Franziska and K{\"o}hler, Peter and Huth, Andreas},
-    month = apr,
-    year = {2016},
-    pages = {124--133},
-    file = {Volltext:C\:\\Users\\fischsam\\Zotero\\storage\\EISNTZF6\\Fischer et al. - 2016 - Lessons learned from applying a forest gap model t.pdf:application/pdf},
-}
--->
-```
-A full description of the model can be found in @fischer_lessons_2016.
-The basis for the parameterization we applied is described in @bohn2014.
+A full description of the model can be found in Fischer et al (2016) .
+The basis for the parameterization we applied is described in Bohn et al
+(2014):
+
+-   Bohn, F. J., Frank, K., & Huth, A. (2014). Of climate and its
+    resulting tree growth: Simulating the productivity of temperate
+    forests. Ecological Modelling, 278, 9--17.
+    <https://doi.org/10.1016/j.ecolmodel.>
+
+-   Fischer, R., Bohn, F., Dantas de Paula, M., Dislich, C., Groeneveld,
+    J., Gutiérrez, A. G., Kazmierczak, M., Knapp, N., Lehmann, S.,
+    Paulick, S., Pütz, S., Rödig, E., Taubert, F., Köhler, P., &
+    Huth, A. (2016).Lessons learned from applying a forest gap model to
+    understand ecosystem and carbon dynamics of complex tropical
+    forests. Ecological Modelling, 326, 124--133.
+    <https://doi.org/10.1016/j.ecolmodel.2015.>
 

### 2023-04-26 manuscript/manuscriptRW.Rmd
reduce abstract to 300 words

@@
 the context of forest resilience and climate change. Models are pivotal
-for assessing long term (i.e., decadal to centennial) forest dynamics,
-and they have been in use for more than 50 years. However, an assessment
-of their ability to accurately represent tree establishment is lacking.
-We assess how well current models capture the overall abundance, species
-composition, and mortality of tree establishment.
-
-Using 15 models and model variants built to capture long-term forest
-dynamics at the stand, landscape and global levels, we simulate tree
-establishment at 200 sites representing large environmental gradients
-across central Europe. The results are evaluated against comprehensive
-data from unmanaged forests.
+for assessing long-term forest dynamics, and they have been in use for
+more than 50 years. However, an assessment of their ability to
+accurately represent tree establishment is lacking. We assess how well
+current models capture the overall abundance, species composition, and
+mortality of tree establishment.
+
+Using 15 models built to capture long-term forest dynamics at the stand,
+landscape, and global levels, we simulate tree establishment at 200
+sites representing large environmental gradients across Central Europe.
+The results are evaluated against comprehensive data from unmanaged
+forests.
 
@@
 early stages of individual trees dynamics. Simulated species diversity
-of ingrowth matches the observed ranges quite well. Models simulating
-higher species diversity at the stand level do not feature higher
-ingrowth diversity. The effect of light availability on ingrowth levels
-is captured better than the effect of temperature and soil moisture, but
-patterns are not consistent across models. The ingrowth of *Abies alba*
-is captured well by most models across temperature and soil moisture
-gradients, whereas ingrowth of *Fagus sylvatica* is generally
-underestimated. *Picea abies* ingrowth is generally overestimated,
-particularly at colder sites. Increasing complexity in the tree
-establishment modules of the models is not related to higher accuracy of
-simulated tree establishment. Furthermore, individual model design is
-more important than scale (stand, landscape, global) and approach
-(empirical, process-based) for accurately capturing tree establishment.
+of ingrowth matches the observed ranges. Models simulating higher
+species diversity at the stand level do not feature higher ingrowth
+diversity. The effect of light availability on ingrowth levels is
+captured better than the effect of temperature and soil moisture, but
+patterns are not consistent across models. Increasing complexity in the
+tree establishment modules of the models is not related to higher
+accuracy of simulated tree establishment. Furthermore, individual model
+design is more important than scale (stand, landscape, global) and
+approach (empirical, process-based) for accurately capturing tree
+establishment.
 
@@
 essential features of the highly complex process of tree establishment,
-while not having been parameterized with such data. We conclude that in
-spite of all the uncertainties, much can be gained by evaluating and
-refining model establishment processes. This has the potential to render
-long-term projections of forest dynamics under changing environmental
-conditions that are much more robust.
+while not having been parameterized with such data. We conclude that
+much can be gained by evaluating and refining model establishment
+processes. This has the potential to render long-term projections of
+forest dynamics under changing environmental conditions that are much
+more robust.
 
@@
 Overall, models are needed to (1) synthesize existing empirical data and
-explore their relationships, (2) assess future tree establishment e.g.
+explore their relationships, (2) assess future tree establishment e.g.,
 in the context of global change scenarios, and (3) identify the most
@@
 per ha and per decade. The observations used in this study featured
-30,900 established trees. Ingrowth rates per site, sample and
-decade ranged from zero to 1246 trees per ha, with a mean of 56 trees
-per ha. Adult species composition was also available at the plots. For
-more details on this unique dataset and the detailed establishment
+30,900 established trees. Ingrowth rates per site, sample and decade
+ranged from zero to 1246 trees per ha, with a mean of 56 trees per ha.
+Adult species composition was also available at the plots. For more
+details on this unique dataset and the detailed establishment
 information, cf. @käber2022.
@@
 
-The simulations were run in the absence of management to a
-simulated equilibrium ("Potential Natural Vegetation") with the current
-climate, as the ultimate goal was to evaluate tree establishment under
-comparable and near equilibrium conditions. This entails the assumption
-that the observations from the forest reserves (1) reflect no traces of
-forest management and (2) there is an equilibrium between forest
-dynamics and climate. While the former might be starting to be visible
-in many of the EuFoRIA reserves, the latter may be more debatable.
-However, in the absence of detailed data on the history of each plot in
-the EuFoRIa network, some broad assumptions had to be made. Both the
-width of the regeneration niche (i.e., in environmental space) as well
-as the intensity of the ingrowth process (i.e., the number of ingrowth
-trees per area and per unit of time) were of interest.
+The simulations were run in the absence of management to a simulated
+equilibrium ("Potential Natural Vegetation") with the current climate,
+as the ultimate goal was to evaluate tree establishment under comparable
+and near equilibrium conditions. This entails the assumption that the
+observations from the forest reserves (1) reflect no traces of forest
+management and (2) there is an equilibrium between forest dynamics and
+climate. While the former might be starting to be visible in many of the
+EuFoRIA reserves, the latter may be more debatable. However, in the
+absence of detailed data on the history of each plot in the EuFoRIa
+network, some broad assumptions had to be made. Both the width of the
+regeneration niche (i.e., in environmental space) as well as the
+intensity of the ingrowth process (i.e., the number of ingrowth trees
+per area and per unit of time) were of interest.
 
@@
 Exercises like the one presented here, where the models are operated in
-"blind flight" mode, i.e. without the possibility of being tuned towards
-capturing the expected patterns, should be repeated. Such benchmarking
-exercises should next focus on aspects such as specific model traits and
-the ecological formulations of particular (sub-)processes, to better
-understand the implications of the assumptions on which the models are
-based. Furthermore, the inclusion of a wide range of models with
-different scales, type, populations structures, tree establishment
-modules and complexity will ensure a wider benefit to the whole modeling
-community.
+"blind flight" mode, i.e., without the possibility of being tuned
+towards capturing the expected patterns, should be repeated. Such
+benchmarking exercises should next focus on aspects such as specific
+model traits and the ecological formulations of particular
+(sub-)processes, to better understand the implications of the
+assumptions on which the models are based. Furthermore, the inclusion of
+a wide range of models with different scales, type, populations
+structures, tree establishment modules and complexity will ensure a
+wider benefit to the whole modeling community.
 
@@
 has led to the findings presented in this paper. The protocol is
-available here @díaz-yáñez2022 . The code and data that support the
+available at @díaz-yáñez2022 . The code and data that support the
 findings of this study are openly available at zenodo (*link to be
-added*) and github (*link to be added*).
+added*) and github (*link to be added*) [for the review process we have
+provided these materials [on this
+link](https://polybox.ethz.ch/index.php/s/q5MOB1CLcjARs3H) (password:
+DownDataCodeRegen1)].
 
@@
 2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
-Several people on top of the co-authors list have been involved in the
-development of this study in several degrees that we would like to thank
-here including Martin Gutsch and Giorgio Vacchiano. Funding for the
-workshop was provided by COST Action CA19139 PROCLIAS (PROcess-based
-models for CLimate Impact Attribution across Sectors), supported by COST
-(European Cooperation in Science and
-Technology; [https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich
-(Research Grant ETH-35 18-1). ODY was funded by the Swiss National
-Science Foundation project 200020_188882 / 1 . KB received support from
-the European Research Council under the European Union's Horizon 2020
+Funding for the workshop was provided by COST Action CA19139 PROCLIAS
+(PROcess-based models for CLimate Impact Attribution across Sectors),
+supported by COST (European Cooperation in Science and Technology;
+[https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research
+Grant ETH-35 18-1). ODY was funded by the Swiss National Science
+Foundation project 200020_188882 / 1. KB received support from the
+European Research Council under the European Union's Horizon 2020
 research and innovation program (Grant Agreement 101001905). JB
@@
 Operational Programme funded by the ERDF. CH acknowledges financial
-support from Waldfonds WASIM2100 project.
+support from Waldfonds WASIM2100 project. We would also like to thank
+all the people that helped to make this study possible with
+contributions at different stages such as Martin Gutsch and Giorgio
+Vacchiano.
 
@@
 ```{=html}
-<!--The relationship was first described by
-[@key] for single citation
-[@key1; @key2] multiple citation can be separated by semi-colon
-[-@key] in order to suppress author name, and just display the year
-[see @key1 p 12; also this ref @key2] is alos a valid syntax
--->
-```
+```

### 2023-04-26 code/figures.R
remove sigH10_Obs.csv

@@ -33,8 +33,8 @@
     outputsDF$model <- factor(outputsDF$model, levels = modelsOrder)
     outputsDF$dbh <- factor(outputsDF$dbh, levels = c("7", "10"))
     
-# Remove the 35 sites from the models data for threshold 7cm as they do not have 
-# observations for those sites
+# Remove the 35 sites from the models data for threshold 7cm 
+# as they do not have observations for those sites
 
     sitesWoObs <- outputsDF[is.na(outputsDF$r.trees), ]
     outputsDF <- outputsDF[!(outputsDF$dbh == 7 & outputsDF$site %in% unique(sitesWoObs$site)), ]
@@ -72,7 +72,6 @@
     ShannonIndex <- merge(ShannonIndex, ShannonIndexAll, by = c("site", "sample",
                                                                 "model","dbh"))
     
-    
     ## Relative Shannon index ----
     
     ShannonIndex$relShannon <- ShannonIndex$ShannonIndexRecruit / ShannonIndex$ShannonIndexAllAges
@@ -84,7 +83,7 @@
     
     zeroInflSim <-  zeroInflSim[, c("site", "sample", "model", "Totr.ba")]
     zeroInflSim <- zeroInflSim[!duplicated(zeroInflSim)]
-    zeroInflEmp <-  zeroInflEmp[, c("site", "sample", "Totr.ba")]
+    zeroInflEmp <- zeroInflEmp[, c("site", "sample", "Totr.ba")]
     zeroInflEmp <- zeroInflEmp[!duplicated(zeroInflEmp)]
     
     #Number of sites and samples with recruitment equal to 0 in observations
@@ -148,7 +147,7 @@
         ggplot2::scale_alpha_manual(values = c(0.3, 1),
                                     guide = ggplot2::guide_legend(override.aes = list(alpha = c(0.4, 1),
                                                                              fill = "red")),
-                                    #name = "Diameter threshold",
+                                    name = "", #"Diameter threshold",
                                     labels = c('7 cm', '10 cm')) +
         ggplot2::geom_hline(ggplot2::aes(yintercept = quantile(overDTsim$r.trees[overDTsim$dbh == "7" & overDTsim$model == "Observed"],
                                                                c(0.25),
@@ -201,7 +200,7 @@
         ggplot2::scale_alpha_manual(values = c(0.3, 1),
                                     guide = ggplot2::guide_legend(override.aes = list(alpha = c(0.4, 1),
                                                                              fill = "red")),
[…]

### 2023-04-26 code/figures.R
remove code from H7.png

@@ -223,35 +223,6 @@
     
     write.csv(sigH7_10,"figures/sigH7_10.csv", row.names = FALSE)
     
-    ####  Figs. H7 per model -----
-    
-    dat <- ShannonIndex[ShannonIndex$dbh == "7", ]
-    dat <- dat |>  dplyr::group_by(model, site) |> 
-        dplyr::summarise(ShannonIndexRecruit = mean(ShannonIndexRecruit))
-    
-    H7 <- ggplot2::ggplot(dat,
-                          ggplot2::aes(y = ShannonIndexRecruit, 
-                                       x = model, 
-                                       fill = model)) + 
-        ggplot2::geom_boxplot(notch = F) +
-        #ggplot2::ylim(c(0,0.3)) +
-        ggplot2::xlab(label = "") +
-        ggplot2::ylab(label = "H (7cm)") +
-        ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90),
-                       strip.text.y = ggplot2::element_text(angle = 0),
-                       legend.position = "none",
-                       panel.background = ggplot2::element_blank(), 
-                       axis.line = ggplot2::element_line(colour = "black"),
-                       legend.key =  ggplot2::element_blank(), 
-                       legend.title = ggplot2::element_blank(),
-                       strip.background =  ggplot2::element_blank()) +
-        ggplot2::scale_fill_manual(values = values_color) 
-    
-    ggplot2::ggsave("figures/H7.png",
-                    plot =  H7,
-                    width = 21, height = 12, scale = 0.9,
-                    dpi = 300, units = "cm", device = 'png') 
-    
     #### FigS. H10 per model -----
     dat10 <- ShannonIndex[ShannonIndex$dbh == "10",]
     dat10 <- dat10 |>  dplyr::group_by(model, site) |> 

### 2023-04-26 code/figures.R
remove code from H10.png

@@ -223,34 +223,6 @@
     
     write.csv(sigH7_10,"figures/sigH7_10.csv", row.names = FALSE)
     
-    #### FigS. H10 per model -----
-    dat10 <- ShannonIndex[ShannonIndex$dbh == "10",]
-    dat10 <- dat10 |>  dplyr::group_by(model, site) |> 
-        dplyr::summarise(ShannonIndexRecruit = mean(ShannonIndexRecruit))
-    
-    H10 <- ggplot2::ggplot(dat10,
-                           ggplot2::aes(y = ShannonIndexRecruit, 
-                                        x = model, 
-                                        fill = model)) + 
-        ggplot2::geom_boxplot(notch = F) +
-        #ggplot2::ylim(c(0,0.3)) +
-        ggplot2::xlab(label = "") +
-        ggplot2::ylab(label = "H (7cm)") +
-        ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90),
-                       strip.text.y = ggplot2::element_text(angle = 0),
-                       legend.position = "none",
-                       panel.background = ggplot2::element_blank(), 
-                       axis.line = ggplot2::element_line(colour = "black"),
-                       legend.key =  ggplot2::element_blank(), 
-                       legend.title = ggplot2::element_blank(),
-                       strip.background =  ggplot2::element_blank()) +
-        ggplot2::scale_fill_manual(values = values_color) 
-    
-    ggplot2::ggsave("figures/H10.png",
-                    plot =  H10,
-                    #width = 55, height = 18, scale = 0.9,
-                    dpi = 300, units = "cm", device = 'png') 
-    
     #### FigS. H10-H7 / H7 ----
     
     medianPointsH <- S107 |>

### 2023-04-26 code/figures.R
remove code from H10H7_H7.png

@@ -223,42 +223,7 @@
     
     write.csv(sigH7_10,"figures/sigH7_10.csv", row.names = FALSE)
     
-    #### FigS. H10-H7 / H7 ----
-    
-    medianPointsH <- S107 |>
-        dplyr::group_by(model) |> 
-        dplyr::summarise(ShannonIndexRecruit7 = median(ShannonIndexRecruit7, na.rm = TRUE),
-                         H710 = median(H710, na.rm = TRUE))  
-    
-    
-    H10H7_H7  <- ggplot2::ggplot(S107,
-                                 ggplot2::aes(x = ShannonIndexRecruit7,
-                                              y = H710, 
-                                              color = model)) +
-        ggplot2::geom_point(alpha = .2) +
-        #ggplot2::xlim(c(0, 0.4)) +
-        ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 0),
-                       strip.text.y = ggplot2::element_text(angle = 0),
-                       panel.background = ggplot2::element_blank(), 
-                       axis.line = ggplot2::element_line(colour = "black"),
-                       legend.key =  ggplot2::element_blank(), 
-                       legend.title = ggplot2::element_blank(),
-                       strip.background =  ggplot2::element_blank()) +
-        ggplot2::scale_colour_manual(values = values_color) +
-        ggplot2::scale_fill_manual(values = values_color) +
-        ggplot2::xlab(label = "H7") +
-        ggplot2::ylab(label = "H10-H7") +
-        ggplot2::geom_point(data = medianPointsH, size = 6, alpha = .9, 
-                            pch = 21, ggplot2::aes(fill = model))  
-    
-    
-    
-    ggplot2::ggsave("figures/H10H7_H7.png",
-                    plot = H10H7_H7,
-                    dpi = 300, units = "cm", device = 'png') 
-    
-    
-    
+   
     ### Fig. H obs. VS sim. in stand and recruitment  ----
     
     ### Data preparation

### 2023-04-26 code/figures.R
final cleaning in fogures script

@@ -577,9 +577,9 @@
     
     
   ### Fig. median vs overestimate per model ----
-    # Prepare the data
-    ##calculated as the difference between the 3rd quartile and the 1st quartile
-    ##in rate of ntrees between 10 and 7 cm
+    ## Prepare the data
+    ## calculated as the difference between the 3rd quartile and the 1st quartile
+    ## in rate of ntrees between 10 and 7 cm
     iqrDT <- mortAll[is.finite(mortAll$nn710),] |>
                 dplyr::group_by(model) |> 
                 dplyr::summarise(iqr7_10 = IQR(nn710, na.rm = TRUE),
@@ -751,7 +751,6 @@
     
     
 ## Model traits-----    
-## 
  ### Complexity VS overestimation proportion-----
 
     # Prepare data for species diversity -----
@@ -781,7 +780,7 @@
     #Figure
     tableCompOver <- readr::read_csv(here::here("figures", "tableCompOver.csv"))
     
-    ggplotRegression <- function(fit){
+    ggplotRegression <- function(fit) {
       require(ggplot2)
       ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
         geom_point() +

### 2023-04-26 code/figures.R
clean figures script

@@ -110,7 +110,7 @@
                                   "Percentage of no recruitment (10 cm)")
     inflatedZeroDF <- tibble::rownames_to_column(inflatedZeroDF, "Model")
     
-    data.table::fwrite(inflatedZeroDF, file = "data/noRecruitment.csv")
+    data.table::fwrite(inflatedZeroDF, file = "figures/noRecruitment.csv")
     
 # Ingrowth levels ----
 

### 2023-04-26 manuscript/manuscriptRW.Rmd
clean code lines in manuscript

@@
 ```{r, echo=FALSE}
-  # devtools::install_github("cboettig/knitcitations@v1")
   library(knitcitations); cleanbib()
@@
   options(tinytex.verbose = TRUE)
-  #knitr::opts_chunk$set(echo = TRUE, fig.pos = "h", out.extra = "")
-
 ```

### 2023-04-26 manuscript/supplementaryRW.Rmd
updates in supplementary

@@
 chose the model parameters for the individual species based on the work
-by @bohn2014. For species that were not covered by @bohn2014, we used
-parameter values of similar species with known parameters. The species
-composition was not predefined and evolved solely from the seed input
-and the model dynamics.
+by Bohn et al (2014). For species that were not covered by Bohn et al
+(2014), we used parameter values of similar species with known
+parameters. The species composition was not predefined and evolved
+solely from the seed input and the model dynamics.
 
@@
 
-zeroInflated <- readr::read_csv(here::here("data", "noRecruitment.csv"))
+zeroInflated <- readr::read_csv(here::here("figures", "noRecruitment.csv"))
 zeroInflated$Model[zeroInflated$Model == "Empirical"] <- "Observations"

### 2023-04-26 manuscript/supplementaryRW.Rmd
details from supplementary

@@
 
-Lasch-Bonn et al. 2020. Description and evaluation of the process-based
-forest model 4C v2.2 at four European forest sites. DOI:
+-   Lasch-Bonn et al. 2020. Description and evaluation of the
+    process-based forest model 4C v2.2 at four European forest sites.
+    DOI:
     [https://doi.org/10.5194/gmd-13-5311-2020)](https://doi.org/10.5194/gmd-13-5311-2020))
@@
 
-[4C
-(v2.2)](%5Bhttps://doi.org/10.5194/gmd-13-5311-2020)](<https://doi.org/10.5194/gmd-13-5311-2020>))
-is usually not used for simulations of forest equillibrium dynamics,
-potential natural vegetation assessment and long simulation periods but
-rather shorter-term simulations (up to 100-150 years). In addition, only
-five species out of the pool of eleven potentially relevant species here
-are parameterized in 4C so that projections of the potential natural
+[4C (v2.2)](https://doi.org/10.5194/gmd-13-5311-2020) is usually not
+used for simulations of forest equillibrium dynamics, potential natural
+vegetation assessment and long simulation periods but rather
+shorter-term simulations (up to 100-150 years). In addition, only five
+species out of the pool of eleven potentially relevant species here are
+parameterized in 4C so that projections of the potential natural
 vegetation done wih 4C need to be considered in the light of model
@@
 
-Ján Merganič1, Katarína Merganičová2,3, Marek Fabrika1, Peter Valent1
-
-1Technical University in Zvolen, Faculty of Forestry, T. G. Masaryka 24,
-960 53 Zvolen, Slovak Republic
-
-2Department of Biodiversity of Ecosystems and Landscape, Slovak Academy
-of Sciences, Štefánikova 3, P.O.Box 25, 81499 Bratislava, Slovakia
-
-3Czech University of Life Sciences Prague, Faculty of Forestry and Wood
-Sciences, Kamýcká 129, 16500 Praha 6 -- Suchdol, Czech Republic
+Ján Merganič, Katarína Merganičová, Marek Fabrika, Peter Valent
 
@@
 
--   Fabrika (2005) and in the link:
+-   Fabrika (2005) "Simulátor biodynamiky lesa SIBYLA." *Koncepcia,
+    konštrukcia a programové riešenie. Habilitačná práca, Technická
+    univerzita vo Zvolene* 
+-   Materials available in
     <http://etools.tuzvo.sk/sibyla/english/model.htm>.
@@
 (1994). Site is specified using ecological site characteristics
-describing climate, air, and soil, also called site variables: • s1
-(N2O) ... NOx concentration in air (ppb) • s2 (CO2) ... CO2
-concentration in air (ppm) • s3 (NUTR) ... soil nutrient supply
-(relative value in the range from 0 to 1) • s4 (DAYS) ... number of days
-in the vegetation period (days with daily mean temperature above 10°C) •
-s5 (TAMPL) ... annual temperature amplitude (the difference between
-minimum and maximum monthly temperature in °C) • s6 (TEMP) ... daily
-mean temperature in the vegetation period in °C (from April to
-September) • s7 (MOIST) ...soil moisture (relative value in the range
-from 0 to 1) • s8 (PRECIP) ... precipitation amount in the vegetation
-period in mm (from April to September) • s9 (ARID)... aridity index
-according to de Martone in mm.°C-1 They directly influence the
-production capacity of a stand (tree height and diameter increment).
+describing climate, air, and soil, also called site variables:
+
+• s1 (N2O) ... NOx concentration in air (ppb)
+
+• s2 (CO2) ... CO2 concentration in air (ppm)
+
+• s3 (NUTR) ... soil nutrient supply (relative value in the range from 0
+to 1)
+
+• s4 (DAYS) ... number of days in the vegetation period (days with daily
+mean temperature above 10°C)
+
+• s5 (TAMPL) ... annual temperature amplitude (the difference between
+minimum and maximum monthly temperature in °C)
+
+• s6 (TEMP) ... daily mean temperature in the vegetation period in °C
+(from April to September)
+
+• s7 (MOIST) ...soil moisture (relative value in the range from 0 to 1)
+
+• s8 (PRECIP) ... precipitation amount in the vegetation period in mm
+(from April to September)
+
+• s9 (ARID)... aridity index according to de Martone in mm.°C-1 They
+directly influence the production capacity of a stand (tree height and
+diameter increment).
 
@@
 koncepcia, konštrukcia a programové riešenie. Habilitačná práca.
-Technická univerzita vo Zvolene, 238 p. Kahn, M., (1994): Modellierung
-der Höhenentwicklung ausgewählter Baumarten in Abhängigkeit vom
-Standort. Forstliche Forschungsber. München, Vol. 141, 221 p. Keeling C.
-D., R. B. Bacastow, A. E. Bainbridge, C. A. Ekdahl Jr., P. R. Guenther,
-L. S. Waterman, J. F. S. Chin 1976: Atmospheric carbon dioxide
-variations at Mauna Loa Observatory, Hawaii. Tellus. XXVIII, 6: 538-551,
-<https://doi.org/10.1111/j.2153-3490.1976.tb00701.x> Merganič J.,
-Fabrika M., 2009: Determination of climatic-site strata based on
-rectified climatic rasters for the purposes of natural regeneration
-density model. Acta Facultatis Forestalis Zvolen, 51(3), p. 113-124,
-ISSN: 0231-5785 Merganič, J., Fabrika, M., Merganičová, K., 2011:
-Submodel of height-diameter function for climatic-site strata of the
-model of natural regeneration in the growth simulator SIBYLA. Acta
-Facultatis Forestalis Zvolen, 53(1), p.155-168, ISSN: 0231-5785
+Technická univerzita vo
+
+Zvolene, 238 p. Kahn, M., (1994): Modellierung der Höhenentwicklung
+ausgewählter Baumarten in Abhängigkeit vom Standort. Forstliche
+Forschungsber. München, Vol. 141, 221 p. Keeling C. D., R. B. Bacastow,
+A. E. Bainbridge, C. A. Ekdahl Jr., P. R. Guenther, L. S. Waterman, J.
+F. S. Chin 1976: Atmospheric carbon dioxide variations at Mauna Loa
+Observatory, Hawaii. Tellus. XXVIII, 6: 538-551,
+<https://doi.org/10.1111/j.2153-3490.1976.tb00701.x>
+
+Merganič J., Fabrika M., 2009: Determination of climatic-site strata
+based on rectified climatic rasters for the purposes of natural
+regeneration density model. Acta Facultatis Forestalis Zvolen, 51(3), p.
+113-124, ISSN: 0231-5785
+
+Merganič, J., Fabrika, M., Merganičová, K., 2011: Submodel of
+height-diameter function for climatic-site strata of the model of
+natural regeneration in the growth simulator SIBYLA. Acta Facultatis
+Forestalis Zvolen, 53(1), p.155-168, ISSN: 0231-5785
+
 Pretzsch, H., (1992): Konzeption und Konstruktion von Wuchsmodellen für

### 2023-04-26 code/dataPreparation.R

@@ -1,7 +1,8 @@
 ################################################################################
 # Title: Data preparation from model simulations and observations
 # Aim: Prepare and merge data from simulations and observations
-# Authors: Olalla Díaz-Yáñez and Yannek Käber 
+# Authors: Olalla Díaz-Yáñez | @olalla | olalladiaz.net
+#          and Yannek Käber 
 ################################################################################ 
 
 # Load observations -----

### 2023-04-26 manuscript/manuscriptRW.Rmd
change establishment and ingrowth to REGENERATION

@@
 ---
-title: "Tree establishment in models of forest dynamics: a key priority for further research"
+title: "Tree regeneration in models of forest dynamics: a key priority for further research"
 author:
@@
 
-Tree establishment is a key process in forest dynamics, particularly in
+Tree regeneration is a key process in forest dynamics, particularly in
 the context of forest resilience and climate change. Models are pivotal
@@
 more than 50 years. However, an assessment of their ability to
-accurately represent tree establishment is lacking. We assess how well
+accurately represent tree regeneration is lacking. We assess how well
 current models capture the overall abundance, species composition, and
-mortality of tree establishment.
+mortality of tree regeneration.
 
 Using 15 models built to capture long-term forest dynamics at the stand,
-landscape, and global levels, we simulate tree establishment at 200
-sites representing large environmental gradients across Central Europe.
-The results are evaluated against comprehensive data from unmanaged
-forests.
-
-Most of the models overestimate ingrowth levels, which is only
+landscape, and global levels, we simulate tree regeneration at 200 sites
+representing large environmental gradients across Central Europe. The
+results are evaluated against comprehensive data from unmanaged forests.
+
+Most of the models overestimate regeneration levels, which is only
 compensated by some models with high simulated mortality rates in the
 early stages of individual trees dynamics. Simulated species diversity
-of ingrowth matches the observed ranges. Models simulating higher
-species diversity at the stand level do not feature higher ingrowth
-diversity. The effect of light availability on ingrowth levels is
+of regeneration matches the observed ranges. Models simulating higher
+species diversity at the stand level do not feature higher regeneration
+diversity. The effect of light availability on regeneration levels is
 captured better than the effect of temperature and soil moisture, but
 patterns are not consistent across models. Increasing complexity in the
-tree establishment modules of the models is not related to higher
-accuracy of simulated tree establishment. Furthermore, individual model
+tree regeneration modules of the models is not related to higher
+accuracy of simulated tree regeneration. Furthermore, individual model
 design is more important than scale (stand, landscape, global) and
 approach (empirical, process-based) for accurately capturing tree
-establishment.
+regeneration.
 
@@
 some instances, it is remarkable that most of the models capture the
-essential features of the highly complex process of tree establishment,
+essential features of the highly complex process of tree regeneration,
 while not having been parameterized with such data. We conclude that
-much can be gained by evaluating and refining model establishment
+much can be gained by evaluating and refining model regeneration
 processes. This has the potential to render long-term projections of
@@
 efforts have been dedicated to improving the representation of tree
-mortality [@bugmann2019; @cailleret2017]. In contrast, tree
-establishment is a much less studied process that is often represented
-rather coarsely in these models [@walck2011; @leishman1992; @price2001],
-which is a clear research gap in the context of climate-induced forest
-disturbances and forest resilience.
-
-Tree establishment arises from multiple processes such as pollination,
+mortality [@bugmann2019; @cailleret2017]. In contrast, tree regeneration
+is a much less studied process that is often represented rather coarsely
+in these models [@walck2011; @leishman1992; @price2001], which is a
+clear research gap in the context of climate-induced forest disturbances
+and forest resilience.
+
+Tree regeneration arises from multiple processes such as pollination,
 fruit maturation, seed production, dispersal, germination, juvenile
@@
 appropriate level of complexity for the mathematical formulation of the
-key factors that are leading to successful tree establishment is
-challenging. Currently, tree establishment processes in dynamic forest
+key factors that are leading to successful tree regeneration is
+challenging. Currently, tree regeneration processes in dynamic forest
 models are handled in a multitude of ways [@konig_tree_2022;
@@
 Overall, models are needed to (1) synthesize existing empirical data and
-explore their relationships, (2) assess future tree establishment e.g.,
+explore their relationships, (2) assess future tree regeneration e.g.,
 in the context of global change scenarios, and (3) identify the most
@@
 current strategies that are used in models of forest dynamics to
-represent tree establishment, model behavior often is prone to problems,
-such as very high levels of tree establishment that necessitate excess
+represent tree regeneration, model behavior often is prone to problems,
+such as very high levels of tree regeneration that necessitate excess
 mortality at early stages of tree life to simulate correct stand
 structure and composition. Also, inadequately high species diversity in
-tree establishment may be simulated, which is characteristic of
+tree regeneration may be simulated, which is characteristic of
 "classical" forest gap models [@gutiérrez2016], at least as long as the
@@
 calibration against local canopy-level data to define appropriate
-ingrowth levels, which is likely to strongly hamper the general
+regeneration levels, which is likely to strongly hamper the general
 applicability of these models e.g. under scenarios of climate change.
 Furthermore, correctly capturing the species composition in tree
-establishment as a function of the presence of seed trees in the canopy
+regeneration as a function of the presence of seed trees in the canopy
 is often a particular challenge, potentially leading to unrealistic
@@
 by considering a multitude of ecological influences, rather than
-emphasizing tree establishment into a larger size class [e.g.,
+emphasizing tree regeneration into a larger size class [e.g.,
 @wehrli2006].
 
-Thus, a focus on the modeling of tree establishment processes is sorely
+Thus, a focus on the modeling of tree regeneration processes is sorely
 needed and overdue [@price2001; @walck2011] if we are to make reliable
@@
 point of view for increasing systems understanding. Also in empirical
-ecology, tree establishment has not received a lot of attention
+ecology, tree regeneration has not received a lot of attention
 [@hanbury-brown2022], although many sub-processes have been studied in
@@
 the one hand, monitoring on permanent sample plots (such as National
-Forest Inventories) often measures tree establishment [@zell2019],
+Forest Inventories) often measures tree regeneration [@zell2019],
 however with a design that captures the rate after the stand initiation
@@
 in a modeling context [@käber2021]. On the other hand, targeted studies
-to measure tree establishment on experimental sites are often limited in
+to measure tree regeneration on experimental sites are often limited in
 spatial extent, species studied, or the sub-set of processes that are
@@
 evaluated against a continental-scale, multi-species harmonized dataset
-on tree establishment [@käber2022]. By tree establishment, we refer to
-the passing of trees across a specific diameter threshold ("ingrowth").
-We included models that are based on a range of "philosophies", from
-models operating at the stand to the global scale as well as the range
-of models from empirically derived to "process-based" [e.g.,
-@bugmann1996; @fabrika2012; @lexer2001; @reyer2014]. Because of the
-large variability in tree establishment patterns in nature and the large
-number of factors driving this process --- including some that are not
-incorporated in most models such as deer browsing --- we do not aim at a
-detailed statistical evaluation of each model. Instead, we aim at
-evaluating the general ingrowth patterns and magnitudes simulated by the
-models, and benchmark the simulated regeneration niche of multiple
-species against empirical data along a wide environmental gradient of
+on tree regeneration [@käber2022]. By tree regeneration, we refer to the
+passing of trees across a specific diameter threshold ("ingrowth"). We
+included models that are based on a range of "philosophies", from models
+operating at the stand to the global scale as well as the range of
+models from empirically derived to "process-based" [e.g., @bugmann1996;
+@fabrika2012; @lexer2001; @reyer2014]. Because of the large variability
+in tree regeneration patterns in nature and the large number of factors
+driving this process --- including some that are not incorporated in
+most models such as deer browsing --- we do not aim at a detailed
+statistical evaluation of each model. Instead, we aim at evaluating the
+general regeneration patterns and magnitudes simulated by the models,
+and benchmark the simulated regeneration niche of multiple species
+against empirical data along a wide environmental gradient of
 temperature, moisture, and light availability [@grubb1977].
@@
 More specifically, we aim to answer the following questions: (1) Are
-models of forest dynamics capturing accurately tree ingrowth levels,
-initial tree species diversity and mortality in the establishment? (2)
-Do model traits explain differences in model performance? (3) How well
-do the models capture total ingrowth and individual species regeneration
-niches across environmental gradients of light availability,
-temperature, and soil moisture?
+models of forest dynamics capturing accurately tree regeneration levels,
+initial tree species diversity and mortality in the regeneration? (2) Do
+model traits explain differences in model performance? (3) How well do
+the models capture total regeneration and individual species
+regeneration niches across environmental gradients of light
+availability, temperature, and soil moisture?
 
@@
 Fifteen models of forest dynamics (including two models featuring model
-variants) were used to assess simulated tree establishment. They
-simulate forest dynamics at the stand, landscape or global scale (Table
+variants) were used to assess simulated tree regeneration. They simulate
+forest dynamics at the stand, landscape or global scale (Table
 \@ref(tab:tableModels)). The approaches used for model construction and
@@
 FORMIND and iLand). The different approaches underlying the models have
-strongly influenced the formulation of tree establishment processes.
-
-The establishment modules can be categorized as 'regeneration' or
-'recruitment' models [@vanclay1994; @konig_tree_2022] (see Table
-\@ref(tab:tableModels)). Regeneration models include processes such as
-flowering and pollination, seed production, seed dispersal, germination,
-and seedling growth, which ultimately lead to the simulated number of
-established trees. Recruitment models, in contrast, introduce a number
-of new trees with certain characteristics such as biomass or diameter,
-without explicitly considering earlier development processes. We can
-further distinguish models that feature a feedback in the establishment,
-i.e., where the level and species composition of establishment is
-influenced by the existing adult tree community via the production of
-seeds, seedlings or saplings, from models that do not contain such a
-feedback.
-
-The starting point for tree establishment in the models differs as well,
+strongly influenced the formulation of tree regeneration processes.
+
+The regeneration modules has been categorized as 'regeneration' or
+'recruitment' models [@vanclay1994; @konig_tree_2022]. *Regeneration
+models* include processes such as flowering and pollination, seed
+production, seed dispersal, germination, and seedling growth, which
+ultimately lead to the simulated number of established trees.
+*Recruitment models*, in contrast, introduce a number of new trees with
+certain characteristics such as biomass or diameter, without explicitly
+considering earlier development processes. We can further distinguish
+models that feature a feedback in the regeneration, i.e., where the
+level and species composition of regeneration is influenced by the
+existing adult tree community via the production of seeds, seedlings or
+saplings, from models that do not contain such a feedback.
+
+The starting point for tree regeneration in the models differs as well,
 ranging from seedbank, seed or seedling to sapling (i.e., trees much
@@
 In the models used in this study, the overall complexity in the
-establishment modules varies considerably. Following @bugmann2022, we
-can classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS,
+regeneration modules varies considerably. Following @bugmann2022, we can
+classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS,
 ForClim and TreeMig) as having rather high complexity in their
-establishment modules (mean establishment formulation complexity across
+regeneration modules (mean regeneration formulation complexity across
 all processes \>0, Table \@ref(tab:tableModels)). Two models, ForClim
 and ForCEEPS, were used here with two alternative variants of
-establishment. ForClim variant 1 [@bugmann1996] is based on a
-recruitment module that adheres closely to the concept introduced by
-@botkin1972, whereas ForClim variant 11 is adopting a different approach
-where individual species properties and their relation to the
-environment are important [@huber2020]. These two model variants allow
-us to evaluate a more process-based and complex module (variant 11)
-against a simple module (variant 1), while the rest of the model
-structure is identical. Similarly, the two ForCEEPS variants allow us to
-isolate the importance of the canopy feedback (i.e., simulated actual
-composition and relative abundance of species in the plot) via seed
-trees for the quantity and quality (e.g., diversity and composition) of
-simulated ingrowth, as one variant includes this feedback whereas the
-other does not.
+regeneration. ForClim variant 1 [@bugmann1996] is based on a recruitment
+module that adheres closely to the concept introduced by @botkin1972,
+whereas ForClim variant 11 is adopting a different approach where
+individual species properties and their relation to the environment are
+important [@huber2020]. These two model variants allow us to evaluate a
+more process-based and complex module (variant 11) against a simple
+module (variant 1), while the rest of the model structure is identical.
+Similarly, the two ForCEEPS variants allow us to isolate the importance
+of the canopy feedback (i.e., simulated actual composition and relative
+abundance of species in the plot) via seed trees for the quantity and
+quality (e.g., diversity and composition) of simulated regeneration, as
+one variant includes this feedback whereas the other does not.
 
@@
 
-Ingrowth data covering a wide range of environmental conditions are hard
-to obtain, and this is one of the reasons why most models of forest
+Regeneration data covering a wide range of environmental conditions are
+hard to obtain, and this is one of the reasons why most models of forest
 dynamics have never been confronted with a dataset covering such
-gradients over a large number of sites, to evaluate how well ingrowth is
-captured. The observations used here are derived from a novel and
-unprecedented network of sites in forest reserves that represent the
-range of environmental gradients in temperature and precipitation in
-Central Europe as compiled in the framework of the EuFoRIa network
-[@euforia2019] (Figure \@ref(fig:map)). These forests have been
-unmanaged for at least 10 years prior to their designation as reserves,
-and most of them provide time series of natural forest dynamics over
-multiple decades and up to ca. 80 years. The census periods range from 3
-to 37 years, with an average of 14 years. The data provide information
-at the tree level, thus allowing for sequential comparison such as
-individual-tree ingrowth and death between the measurements.
+gradients over a large number of sites, to evaluate how well
+regeneration is captured. The observations used here are derived from a
+novel and unprecedented network of sites in forest reserves that
+represent the range of environmental gradients in temperature and
+precipitation in Central Europe as compiled in the framework of the
+EuFoRIa network [@euforia2019] (Figure \@ref(fig:map)). These forests
+have been unmanaged for at least 10 years prior to their designation as
+reserves, and most of them provide time series of natural forest
+dynamics over multiple decades and up to ca. 80 years. The census
+periods range from 3 to 37 years, with an average of 14 years. The data
+provide information at the tree level, thus allowing for sequential
+comparison such as individual-tree regeneration and death between the
+measurements.
 
@@
 Each of these sites featured at least two consecutive measurements.
-Ingrowth thresholds for the sites differed between diameters of 0 and 10
-cm. For our study, we defined two datasets where one included 165 sites
-with a diameter threshold of 7 cm or lower, and the other included
+egeneration thresholds for the sites differed between diameters of 0 and
+10 cm. For our study, we defined two datasets where one included 165
+sites with a diameter threshold of 7 cm or lower, and the other included
 another 35 sites with diameter thresholds between 7 and 10 cm.
@@
 Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
-processed and aggregated following @käber2022 to provide ingrowth rates
-per ha and per decade. The observations used in this study featured
-30,900 established trees. Ingrowth rates per site, sample and decade
-ranged from zero to 1246 trees per ha, with a mean of 56 trees per ha.
-Adult species composition was also available at the plots. For more
-details on this unique dataset and the detailed establishment
+processed and aggregated following @käber2022 to provide regeneration
+rates per ha and per decade. The observations used in this study
+featured 30,900 established trees. Regeneration rates per site, sample
+and decade ranged from zero to 1246 trees per ha, with a mean of 56
+trees per ha. Adult species composition was also available at the plots.
+For more details on this unique dataset and the detailed regeneration
 information, cf. @käber2022.
@@
 The overarching goal of the experiments was to assess the tree
-establishment as it arises from empirical data against its
-representation in a wide range of models of forest dynamics. We define
-the tree establishment as the passing of a breast height diameter
-threshold of 7 or 10 cm, respectively (synonym: ingrowth). To this end,
-each modeling group was provided with a detailed protocol
-[@díaz-yáñez2022] with instructions for how to perform the simulations,
-providing input variables on climate and soil conditions, and the list
-of expected output variables. Neither were further site information
-(except for the data specified below) nor any data on tree ingrowth or
-forest stand features provided prior to the simulation. That is, the
-models were run in "blind flight" mode.
+regeneration as it arises from empirical data against its representation
+in a wide range of models of forest dynamics. We define the tree
+regeneration as the passing of a breast height diameter threshold of 7
+or 10 cm, respectively (synonym: ingrowth). To this end, each modeling
+group was provided with a detailed protocol [@díaz-yáñez2022] with
+instructions for how to perform the simulations, providing input
+variables on climate and soil conditions, and the list of expected
+output variables. Neither were further site information (except for the
+data specified below) nor any data on tree regeneration or forest stand
+features provided prior to the simulation. That is, the models were run
+in "blind flight" mode.
 
@@
 chances of shade-intolerant species to establish. The simulations were
-set up to sample species-specific ingrowth rates per decade and per ha
-in the equilibrium state of the model, typically entailing a "spin-up"
-run (as we did not provide any forest data). The modeling teams decided
-on the simulated area and how they derived these samples. The exact
-length of the simulation was also decided by the modeling teams (Table
-\@ref(tab:tableModels)). Further details on how each modeling team
-prepared the simulations and the outputs are available in Supplementary
-Material 1.
+set up to sample species-specific regeneration rates per decade and per
+ha in the equilibrium state of the model, typically entailing a
+"spin-up" run (as we did not provide any forest data). The modeling
+teams decided on the simulated area and how they derived these samples.
+The exact length of the simulation was also decided by the modeling
+teams (Table \@ref(tab:tableModels)). Further details on how each
+modeling team prepared the simulations and the outputs are available in
+Supplementary Material 1.
 
@@
 equilibrium ("Potential Natural Vegetation") with the current climate,
-as the ultimate goal was to evaluate tree establishment under comparable
+as the ultimate goal was to evaluate tree regeneration under comparable
 and near equilibrium conditions. This entails the assumption that the
@@
 regeneration niche (i.e., in environmental space) as well as the
-intensity of the ingrowth process (i.e., the number of ingrowth trees
-per area and per unit of time) were of interest.
+intensity of the regeneration process (i.e., the number of ingrowth
+trees per area and per unit of time) were of interest.
 
@@
 single-species simulations) using mixtures of eleven species or genera
-for which ingrowth data of sufficient quality were available from
+for which regeneration data of sufficient quality were available from
 EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
@@
 
-Each of the models reported the ingrowth number by sampling 200 times in
-a 10-year interval for each species and per ha for each of the 200
-sites. Multiple samples per site were used to better understand the
+Each of the models reported the regeneration number by sampling 200
+times in a 10-year interval for each species and per ha for each of the
+200 sites. Multiple samples per site were used to better understand the
 simulated variation within each site. This was done using different
@@
 and two diameter thresholds). For the models that simulated additional
-species, their ingrowth rates were aggregated as "others".
+species, their regeneration rates were aggregated as "others".
 
@@
 dimensions of very old individuals. LPJ-GUESS had 2% of the sample
-outputs without tree establishment, and these were considered as zero
-stand basal area and zero ingrowth for all the tree species simulated;
-one site produced grassland rather than a forest and was not included in
-the results.
-
-```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed ingrowth at the 10 cm threshold.", out.width='80%'}
+outputs without tree regeneration, and these were considered as zero
+stand basal area and zero regeneration for all the tree species
+simulated; one site produced grassland rather than a forest and was not
+included in the results.
+
+```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed regeneration at the 10 cm threshold.", out.width='80%'}
 knitr::include_graphics(here::here("figures", "theMap.png"))
@@
 
-The simulation results were analyzed regarding (1) ingrowth levels
-(i.e., ingrowth number per unit time and space), (2) species diversity
-of the ingrowth, (3) ingrowth mortality, (4) model performance and model
-traits, and (5) ingrowth gradients along regeneration niches. We
+The simulation results were analyzed regarding (1) regeneration levels
+(i.e., ingrowth number per unit time and space), (2) regeneration
+species diversity, (3) regeneration mortality, (4) model performance and
+model traits, and (5) ingrowth gradients along regeneration niches. We
 evaluated the species diversity across models and in relation to the
@@
 proportion of the species in terms of basal area. The Shannon index was
-calculated for the establishment ($H_{R_{n}}$) (Equation
+calculated for the regeneration ($H_{R_{n}}$) (Equation
 \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation
@@
 area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
-the species $i$ in sample $n$ calculated as the ingrowth basal area
+the species $i$ in sample $n$ calculated as the regeneration basal area
 ($r_{BA_{n,i}}$) for that species $i$ relative to the total recruited
@@
 
-Mortality in tree establishment was assessed based on the ratio of
-ingrowth between the 7 and 10 cm diameter thresholds. We used the
+Mortality in tree regeneration was assessed based on the ratio of
+regeneration between the 7 and 10 cm diameter thresholds. We used the
 Reineke self-thinning rule [@reineke1933; @pretzsch2003] as a reference
-to estimate whether the ratio of ingrowth between the 7 and 10 cm
+to estimate whether the ratio of regeneration between the 7 and 10 cm
 diameter thresholds were above or below the expected theoretical rate
@@
 (empirical or process based), (3) the presence or absence of a canopy
-feedback for ingrowth, and (4) the scale of application of the model
+feedback for regeneration, and (4) the scale of application of the model
 (stand, landscape or global). We tested for significant differences
 using two-sided t-tests by considering each of these model features
-regarding the observed and simulated ingrowth levels and species
+regarding the observed and simulated regeneration levels and species
 diversity, respectively.
 
-Total ingrowth and the regeneration niches of the individual species
+Total regeneration and the regeneration niches of the individual species
 were evaluated across the environmental gradients of light availability,
 temperature, and soil moisture. Stand basal area (the basal area of all
-the trees in each sample including tree establishment) was used as a
+the trees in each sample including tree regeneration) was used as a
 proxy of light availability at the forest floor, the annual degree-day
@@
 and the climatic water balance as a proxy for soil moisture
-[@speich2019]. Ingrowth values were calculated as the mean across the
-200 samples per site. The observed data were modeled using a Generalized
-Additive Model [@wood2011] with a negative binomial distribution and
-restricted maximum likelihood to better understand the relationship
-between the environmental gradients and the levels of observed ingrowth,
-relative to the simulation results per model.
+[@speich2019]. Regeneration values were calculated as the mean across
+the 200 samples per site. The observed data were modeled using a
+Generalized Additive Model [@wood2011] with a negative binomial
+distribution and restricted maximum likelihood to better understand the
+relationship between the environmental gradients and the levels of
+observed regeneration, relative to the simulation results per model.
 
@@
 *Quercus* spp. For these species, we calculated the share in the
-ingrowth basal area per site as the mean across the available samples
-per site:
+regeneration basal area per site as the mean across the available
+samples per site:
 
@@
 where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
-is the mean basal area ingrowth of the species $i$ per site $t$ across
-the available samples at that site. Furthermore, regarding the share in
-the ingrowth basal area per species ($\bar{R}_{BA_{share_{i}}}$), we
-categorized this as zero when the ingrowth basal area of that species
-($\bar{r}_{BA_{t,i}}$) was zero but also when both the ingrowth basal
-area of that species ($\bar{r}_{BA_{t,i}}$) and the total ingrowth basal
-area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.
+is the mean basal area regeneration of the species $i$ per site $t$
+across the available samples at that site. Furthermore, regarding the
+share in the regeneration basal area per species
+($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the
+regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero
+but also when both the regeneration basal area of that species
+($\bar{r}_{BA_{t,i}}$) and the total regeneration basal area (
+$\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.
 
@@
 
-## Ingrowth levels
-
-Simulated ingrowth levels varied strongly across the 15 models and
+## Regeneration levels
+
+Simulated regeneration levels varied strongly across the 15 models and
 typically did not match the levels found in the forest reserve data
-(Figure \@ref(fig:over)). Ingrowth was overestimated in most models for
-both the 7 and 10 cm diameter thresholds, with the exception of the
+(Figure \@ref(fig:over)). Regeneration was overestimated in most models
+for both the 7 and 10 cm diameter thresholds, with the exception of the
 empirical stand model SIBYLA, the landscape model Landis-II and the
-global model aDGVM2, which estimated ingrowth levels at the lower end of
-the plausibility interval of the observed data (Figure \@ref(fig:over)).
-The models with the largest overestimation were the stand model PICUS
-and the landscape model TreeMig. For most models, the variability of
-simulated ingrowth levels across the 200 sites (visible from the
-interquartile range in the box plots of Figure \@ref(fig:over)) was
-similar to or smaller than observed, with the notable exception of
-PICUS, where simulated ingrowth variability was much larger.
-
-```{r over, echo=FALSE, fig.cap="Mean ingrowth levels across all samples per site, plotted for the 200 sites and for each model.
+global model aDGVM2, which estimated regeneration levels at the lower
+end of the plausibility interval of the observed data (Figure
+\@ref(fig:over)). The models with the largest overestimation were the
+stand model PICUS and the landscape model TreeMig. For most models, the
+variability of simulated regeneration levels across the 200 sites
+(visible from the interquartile range in the box plots of Figure
+\@ref(fig:over)) was similar to or smaller than observed, with the
+notable exception of PICUS, where simulated regeneration variability was
+much larger.
+
+```{r over, echo=FALSE, fig.cap="Mean regeneration levels across all samples per site, plotted for the 200 sites and for each model.
 The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
@@
 
-Both the observations and the simulated data had no ingrowth in some
+Both the observations and the simulated data had no regeneration in some
 samples and at some sites (for details, cf. Table S1). The observed data
-had 4% of the samples with no ingrowth. Only three models had a larger
-proportion of no ingrowth (4C, Landis-II and aDVM2). Two models always
-simulated ingrowth for both the 7 and 10 cm threshold (xComp and
-TreeMig), i.e., they did not feature any zero values. The other ten
-models had a very low percentage of samples with no ingrowth (0.01% -
-2.39%), i.e., they had distinctly fewer occurrences of zeros compared to
-the observations.
-
-## Tree species diversity of ingrowth
+had 4% of the samples with no regeneration. Only three models had a
+larger proportion of no regeneration (4C, Landis-II and aDVM2). Two
+models always simulated regeneration for both the 7 and 10 cm threshold
+(xComp and TreeMig), i.e., they did not feature any zero values. The
+other ten models had a very low percentage of samples with no
+regeneration (0.01% - 2.39%), i.e., they had distinctly fewer
+occurrences of zeros compared to the observations.
+
+## Tree species diversity of regeneration
 
@@
 well as quantified by the Shannon index (Figure \@ref(fig:H710)). Five
-models overestimated ingrowth diversity: ForCEEPS, ForClim 1, PICUS,
+models overestimated regeneration diversity: ForCEEPS, ForClim 1, PICUS,
 TreeMig and LandClim, the latter particularly for the 7 cm diameter
@@
 the other models, nor to the observations. Only one model, Landis-II,
-consistently underestimated ingrowth diversity.
-
-```{r H710, echo=FALSE, fig.cap="Shannon index of tree ingrowth, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
+consistently underestimated regeneration diversity.
+
+```{r H710, echo=FALSE, fig.cap="Shannon index of tree regeneration, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
 The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.", out.width='100%'}
@@
 In most models, there were only small but significant differences in the
-species diversity of ingrowth between 7 and 10 cm across sites (Tables
-S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained
-the ingrowth diversity between the 7 and 10 cm thresholds (i.e., the
-differences between them were not significant, see Table S2), and the
-same was evident from the observed data.
-
-For both the observed and the simulated data, and both ingrowth
-thresholds, species diversity in the ingrowth compared to stand-level
-diversity did not have clear patterns across the models (Figures S1 and
-S2). Most models captured reasonably well or overpredicted species
-diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left
-and center; cf. Figure S3), but a characteristic feature was that
-simulations had a much lower variability of diversity compared to
-observations. Models overpredicting species diversity included ForClim1,
-ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A,
-left). Several models underpredicted species diversity, i.e., LPJ-GUESS,
-ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).
-
-The majority of the models overpredicted ingrowth diversity; as already
-observed in figure \@ref(fig:H710), only two of them underestimated
-ingrowth diversity (4C and Landis-II) (Figure
+species diversity of regeneration between 7 and 10 cm across sites
+(Tables S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS)
+maintained the regeneration diversity between the 7 and 10 cm thresholds
+(i.e., the differences between them were not significant, see Table S2),
+and the same was evident from the observed data.
+
+For both the observed and the simulated data, and both regeneration
+thresholds, species diversity in the regeneration compared to
+stand-level diversity did not have clear patterns across the models
+(Figures S1 and S2). Most models captured reasonably well or
+overpredicted species diversity at the stand level (Figure
+\@ref(fig:combinedRichness)A, left and center; cf. Figure S3), but a
+characteristic feature was that simulations had a much lower variability
+of diversity compared to observations. Models overpredicting species
+diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure
+\@ref(fig:combinedRichness)A, left). Several models underpredicted
+species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure
+\@ref(fig:combinedRichness)A, right).
+
+The majority of the models overpredicted regeneration diversity; as
+already observed in figure \@ref(fig:H710), only two of them
+underestimated regeneration diversity (4C and Landis-II) (Figure
 \@ref(fig:combinedRichness)B and Figure S4). Both the models with
-feedback and those without feedback from the adult trees to ingrowth
+feedback and those without feedback from the adult trees to regeneration
 (via seed production) overestimate, underestimated and captured
-reasonably well ingrowth species diversity compared to the observations.
-Again, simulated diversity was considerably less variable than observed
-diversity in the ingrowth.
+reasonably well regeneration species diversity compared to the
+observations. Again, simulated diversity was considerably less variable
+than observed diversity in the regeneration.
 
@@
 A) Examples for the three trends across models for species diversity at the stand level.
-B) Examples for the three trends across models for ingrowth at the 7 cm threshold.", out.width='75%'}
+B) Examples for the three trends across models for regeneration at the 7 cm threshold.", out.width='75%'}
 knitr::include_graphics(here::here("figures", "combined_richness.png"))
@@
 
-## Mortality in tree establishment
+## Mortality in tree regeneration
 
@@
 data, thus at least partially compensating for the general
-overestimation of ingrowth levels (Figure \@ref(fig:over) and Table S3);
-six models did not feature a significant difference compared to the
+overestimation of regeneration levels (Figure \@ref(fig:over) and Table
+S3); six models did not feature a significant difference compared to the
 observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
@@
 mortality rates (i.e., well above the Reineke self-thinning line), which
-compensated for the strong initial overestimation of ingrowth (cf.
+compensated for the strong initial overestimation of regeneration (cf.
 Figure \@ref(fig:over)). Interestingly, two models that underestimated
-overall ingrowth levels (SIBYLA and Landis-II; cf. Figure
+overall regeneration levels (SIBYLA and Landis-II; cf. Figure
 \@ref(fig:over)) featured mortality rates that were close to but still
@@
 
-```{r mortality11, echo=FALSE, fig.cap="Ratio of tree ingrowth rates between the 7 and 10 cm thresholds.
-Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree ingrowth between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
+```{r mortality11, echo=FALSE, fig.cap="Ratio of tree regeneration rates between the 7 and 10 cm thresholds.
+Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree regeneration between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
 knitr::include_graphics(here::here("figures", "ratio7_10.png"))
@@
 whose mortality was well above the self-thinning line, indicates that
-they feature higher ingrowth mortality in mixed-species forests compared
-to even-aged, single-species forests. Some models yielded a mortality
-rate of ingrowth that is broadly compatible with the self-thinning rule
-(e.g., 4C and LandClim).
+they feature higher regeneration mortality in mixed-species forests
+compared to even-aged, single-species forests. Some models yielded a
+mortality rate of regeneration that is broadly compatible with the
+self-thinning rule (e.g., 4C and LandClim).
 
 The three models with the largest overestimation of the proportion of
-ingrowth at the 7 cm threshold featured the highest mortality between
-the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The other
-models that overestimated ingrowth had mortalities lower than the
-expected self-thinning ratio, with the exception of LandClim (Figure
+regeneration at the 7 cm threshold featured the highest mortality
+between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The
+other models that overestimated regeneration had mortalities lower than
+the expected self-thinning ratio, with the exception of LandClim (Figure
 \@ref(fig:mortality13)B). It is noteworthy that most of the models that
-featured a low ratio of ingrowth between 7 and 10 cm (i.e., ForCEEPS,
-ForClim 1, ForClim 11 and iLand) had only a small overestimation of
-ingrowth at the 7 cm threshold (Figure \@ref(fig:mortality11)). aDGVM2
-was the only model that underestimated ingrowth at the 7 cm threshold
-while featuring a low mortality between the two thresholds (Figure
-\@ref(fig:mortality13)C), and only a few models were close to the
-correct ingrowth levels at 7 cm while also being close to the Reineke
-line (SIBYLA, Landis-II; cf. Figure \@ref(fig:mortality13)D; and 4C as
-well as LandClim).
-
-```{r mortality13, echo=FALSE, fig.cap="Ratio of tree ingrowth between the 7 and 10 cm thresholds (ingrowth 7 cm / ingrowth 10 cm) and the overestimation proportion at 7 cm ([simulated -observed] / observed) for the mean ingrowth per model across sites and samples.
-The horizontal dashed lines show a ratio equal to 1, indicating no decrease of ingrowth between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
+featured a low ratio of regeneration between 7 and 10 cm (i.e.,
+ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small
+overestimation of regeneration at the 7 cm threshold (Figure
+\@ref(fig:mortality11)). aDGVM2 was the only model that underestimated
+regeneration at the 7 cm threshold while featuring a low mortality
+between the two thresholds (Figure \@ref(fig:mortality13)C), and only a
+few models were close to the correct regeneration levels at 7 cm while
+also being close to the Reineke line (SIBYLA, Landis-II; cf. Figure
+\@ref(fig:mortality13)D; and 4C as well as LandClim).
+
+```{r mortality13, echo=FALSE, fig.cap="Ratio of tree regeneration between the 7 and 10 cm thresholds (regeneration 7 cm / regeneration 10 cm) and the overestimation proportion at 7 cm ([simulated -observed] / observed) for the mean regeneration per model across sites and samples.
+The horizontal dashed lines show a ratio equal to 1, indicating no decrease of regeneration between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
 The vertical line indicates zero overestimation.", out.width='80%'}
@@
 Lastly, some models that had a low number of individuals at the 7 cm
-threshold also had a smaller decrease of ingrowth between 7 and 10 cm
-(ratio 7-10 cm) (Figure S5). In these models, competition and
+threshold also had a smaller decrease of regeneration between 7 and 10
+cm (ratio 7-10 cm) (Figure S5). In these models, competition and
 self-thinning are either not pronounced or must have occurred before the
@@
 across models (Figure S5). It is noteworthy that there is a relationship
-between the ingrowth levels at 7 cm and the mortality rate (Table S4):
-most of the models showed a positive effect (higher ingrowth at 7 cm is
-coupled to an increase of the mortality between 7 and 10 cm). While some
-models showed a negative effect, their linear trend was not pronounced
-and not always significant regarding the slope.
+between the regeneration levels at 7 cm and the mortality rate (Table
+S4): most of the models showed a positive effect (higher regeneration at
+7 cm is coupled to an increase of the mortality between 7 and 10 cm).
+While some models showed a negative effect, their linear trend was not
+pronounced and not always significant regarding the slope.
 
@@
 There was no significant relationship between the mean complexity of the
-establishment module (Table \@ref(tab:tableModels)) and the
+regeneration module (Table \@ref(tab:tableModels)) and the
 overestimation proportion at a diameter threshold of 7 cm (Figure S6).
 All the models except 4C had significant differences between the
-observed and simulated mean ingrowth values. Therefore it was not
+observed and simulated mean regeneration values. Therefore it was not
 possible to differentiate in terms of performance between models type
 (empirical or process based), or their scales (stand, landscape or
-global) (Table S5). The values of ingrowth species diversity simulated
-by each model were also assessed against the observed data to see if
-there was a difference between models that include a feedback compared
-to those that did not; all the models except ForCEEPs(f) and FORMIND had
-significant differences (Table S6). It is noteworthy that for the model
-ForCEEPS the pattern changed from the overprediction of species
-diversity in the ingrowth to a diversity level that is closer to
-observations when the recruitment module included feedback
+global) (Table S5). The values of regeneration species diversity
+simulated by each model were also assessed against the observed data to
+see if there was a difference between models that include a feedback
+compared to those that did not; all the models except ForCEEPs(f) and
+FORMIND had significant differences (Table S6). It is noteworthy that
+for the model ForCEEPS the pattern changed from the overprediction of
+species diversity in the regeneration to a diversity level that is
+closer to observations when the recruitment module included feedback
 (ForCEEPs(f)).
 
-## Ingrowth gradients on regeneration niches
-
-When evaluating ingrowth levels along key gradients of light
+## Regeneration gradients on regeneration niches
+
+When evaluating regeneration levels along key gradients of light
 availability (basal area), temperature (degree-days) and soil moisture
@@
 
-In the observed data, ingrowth levels decreased clearly with increasing
-stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a). This trend
-was captured by the Landis-II model only, albeit at much lower values of
-basal area than in the observations. The other models featured
-distinctly different trends, such as (1) an increase in ingrowth levels
-with increasing total basal area followed by the absence of ingrowth at
-high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and
-LPJ-GUESS), (2) almost constant ingrowth levels with basal area (ForClim
-11, xComp, LandClim and aDGVM2), or (3) an increase of ingrowth up to a
-certain value of stand basal area followed by a decrease at even higher
-values of stand basal area, with model-specific thresholds (4C, ForClim
-1, PICUS, SIBYLA and iLand).
-
-The observed ingrowth did not change much across the climatic water
+In the observed data, regeneration levels decreased clearly with
+increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a).
+This trend was captured by the Landis-II model only, albeit at much
+lower values of basal area than in the observations. The other models
+featured distinctly different trends, such as (1) an increase in
+regeneration levels with increasing total basal area followed by the
+absence of regeneration at high values of basal area (ForCEEPS,
+ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant
+regeneration levels with basal area (ForClim 11, xComp, LandClim and
+aDGVM2), or (3) an increase of regeneration up to a certain value of
+stand basal area followed by a decrease at even higher values of stand
+basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA
+and iLand).
+
+The observed regeneration did not change much across the climatic water
 balance gradient, and showed a slight increase with temperature. The
-models did not match the observed ingrowth patterns across the gradients
-of soil moisture and temperature (Figure
+models did not match the observed regeneration patterns across the
+gradients of soil moisture and temperature (Figure
 \@ref(fig:nvTrend710BAlightTemp)b,c and S7), and neither was there a
 clear pattern across models. Regarding the soil moisture gradient, a
-group of models featured decreasing ingrowth with increasing soil
+group of models featured decreasing regeneration with increasing soil
 moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim,
@@
 (ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient,
-all the models that featured a decrease in ingrowth levels with
-increasing soil moisture showed the opposite trend (i.e., more ingrowth
-with increasing temperature), except for 4C. The other models did not
-feature a clear pattern.
-
-```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum.
+all the models that featured a decrease in regeneration levels with
+increasing soil moisture showed the opposite trend (i.e., more
+regeneration with increasing temperature), except for 4C. The other
+models did not feature a clear pattern.
+
+```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean regeneration values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum.
 The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
@@
 frequently in the EuFoRIa data. Most models captured well the share of
-basal area in the ingrowth of these five species compared to the
+basal area in the regeneration of these five species compared to the
 observed data along the two major climatic gradients (Figure
@@
 
-First, some models overestimated ingrowth across the entire gradients
-for at least one of the five main tree species (Figure \@ref(fig:diff),
-e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models
-overestimated ingrowth for more than one species for a large part of the
-climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf. red
-colors in Figure \@ref(fig:diff)). These trends were consistent for both
-ingrowth thresholds, although the exact changes in the share of the
-ingrowth basal area were different (Figure S8).
+First, some models overestimated regeneration across the entire
+gradients for at least one of the five main tree species (Figure
+\@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS),
+while other models overestimated regeneration for more than one species
+for a large part of the climate gradients, such as xComp, PICUS, SIBYLA
+or LPJ-GUESS (cf. red colors in Figure \@ref(fig:diff)). These trends
+were consistent for both regeneration thresholds, although the exact
+changes in the share of the regeneration basal area were different
+(Figure S8).
 
 Second, there was some congruence in the simulation results by species
-across the models. The share in the ingrowth of Silver fir (*Abies
+across the models. The share in the regeneration of Silver fir (*Abies
 alba*) was captured well across the gradient by most models except for
-xComp, PICUS, and LandClim, which overestimated the ingrowth share of
-this species. There were also some peculiarities evident for some
+xComp, PICUS, and LandClim, which overestimated the regeneration share
+of this species. There were also some peculiarities evident for some
 models, such as 4C that did not simulate *Abies alba*. The patterns
-across models for *Fagus sylvatica* ingrowth were more complex, as its
-ingrowth was underestimated by many models across the environmental
-gradients featuring negligible ingrowth at most sites, while others
-overestimate it in different parts of the environmental space. Most
-models overestimated the share of *Picea abies* ingrowth in the cold-wet
-part of the gradients. *Pinus sylvestris* and *Quercus* spp. represented
-a smaller share of the ingrowth in the observed data mostly at warm-dry
-sites. Many models, however, erroneously featured ingrowth for these two
-species along most of the climatic gradients although at a low basal
-area share. Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had
-almost no ingrowth of *Pinus sylvestris* and *Quercus* spp., or did not
-feature any ingrowth of these species at all, close to the patterns in
-the observed data.
+across models for *Fagus sylvatica* regeneration were more complex, as
+its regeneration was underestimated by many models across the
+environmental gradients featuring negligible regeneration at most sites,
+while others overestimate it in different parts of the environmental
+space. Most models overestimated the share of *Picea abies* regeneration
+in the cold-wet part of the gradients. *Pinus sylvestris* and *Quercus*
+spp. represented a smaller share of the regeneration in the observed
+data mostly at warm-dry sites. Many models, however, erroneously
+featured regeneration for these two species along most of the climatic
+gradients although at a low basal area share. Five models (4C, FORMIND,
+SIBYLA, xComp, and LandClim) had almost no regeneration of *Pinus
+sylvestris* and *Quercus* spp., or did not feature any regeneration of
+these species at all, close to the patterns in the observed data.
 
 Lastly, there is no model that performed well across the five species
-regarding the species-specific ingrowth levels in the environmental
+regarding the species-specific regeneration levels in the environmental
 space (Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
-tended to consistently overestimate the ingrowth share of all five
-species, but most models overestimated the ingrowth share of some
+tended to consistently overestimate the regeneration share of all five
+species, but most models overestimated the regeneration share of some
 species while underestimating the share of others (e.g., ForCEEPS,
@@
 
-```{r diff, echo=FALSE, fig.cap="Ingrowth levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm ingrowth threshold.
+```{r diff, echo=FALSE, fig.cap="Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold.
 The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient.
-The sizes of the circles represent the ratio between the ingrowth basal area of the species and the total ingrowth basal area for all species.
-The absence of a circle indicates a zero basal area share in the ingrowth, or the absence of ingrowth altogether.
-The color gradient (for the models only) shows the difference between the simulated and observed ratio of ingrowth basal area of the species and the total ingrowth basal area for all the species.", out.width='100%', out.height= '85%'}
+The sizes of the circles represent the ratio between the regeneration basal area of the species and the total regeneration basal area for all species.
+The absence of a circle indicates a zero basal area share in the regeneration, or the absence of regeneration altogether.
+The color gradient (for the models only) shows the difference between the simulated and observed ratio of regeneration basal area of the species and the total regeneration basal area for all the species.", out.width='100%', out.height= '85%'}
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
@@
 
-Tree establishment is a fundamental process in forest dynamics, in
+Tree regeneration is a fundamental process in forest dynamics, in
 particular under changing conditions such as those caused by climate
-change. Correctly capturing tree ingrowth in dynamic models is
+change. Correctly capturing tree regeneration in dynamic models is
 fundamental to, for example, evaluate post-disturbance dynamics and
@@
 sequestration. In this study, for the first time, projections of tree
-establishment from multiple models of forest dynamics were confronted
+regeneration from multiple models of forest dynamics were confronted
 with a unique dataset from unmanaged forest reserves across a large
@@
 sites and repeated measurements), which is essential for a highly
-'noisy' process such as tree establishment. The use of this dataset for
+'noisy' process such as tree regeneration. The use of this dataset for
 model benchmarking provided novel insights on the ability of
-state-of-the-art models to accurately simulate ingrowth levels, species
-composition, and mortality in an early stage of tree life, i.e., between
-a tree diameter of 7 and 10 cm. Overall, by adopting this approach a
-much broader understanding resulted than if we had used these data for
-model calibration: it is primarily from the shortcomings of the models
-that we can gain ecological insights [e.g., @trugman2022].
-
-## Ingrowth levels
-
-Most of the models overestimated tree ingrowth levels. This has
+state-of-the-art models to accurately simulate regeneration levels,
+species composition, and mortality in an early stage of tree life, i.e.,
+between a tree diameter of 7 and 10 cm. Overall, by adopting this
+approach a much broader understanding resulted than if we had used these
+data for model calibration: it is primarily from the shortcomings of the
+models that we can gain ecological insights [e.g., @trugman2022].
+
+## Regeneration levels
+
+Most of the models overestimated tree regeneration levels. This has
 potentially far-reaching implications, e.g., regarding biomass (and thus
@@
 forest carbon sequestration [@pugh2019; @pan2011]. Yet, we focused on
-one specific stage, i.e., ingrowth into the 7 and 10 cm diameter
+one specific stage, i.e., regeneration into the 7 and 10 cm diameter
 classes. Trees at this size contribute little to carbon sequestration,
-and if the excess ingrowth at this stage is compensated -- in the models
--- soon thereafter by higher mortality, overall simulated ingrowth may
-still yield trustworthy results at the level of stand structure and
-related forest attributes such as basal area or biomass.
-
-From a modeling point of view, excess ingrowth inevitably requires
+and if the excess regeneration at this stage is compensated -- in the
+models -- soon thereafter by higher mortality, overall simulated
+regeneration may still yield trustworthy results at the level of stand
+structure and related forest attributes such as basal area or biomass.
+
+From a modeling point of view, excess regeneration inevitably requires
 excess mortality rates in a later stage, either -- as observed for some
@@
 stages for the expected forest densities at later stages is equivalent
-to compensating for a first error (excess ingrowth) by a second "error"
-(excess mortality). It is highly likely that biased projections will
-result, because the two errors are unlikely to be perfectly linked and
-thus will not always compensate each other. Hence, this structural
+to compensating for a first error (excess regeneration) by a second
+"error" (excess mortality). It is highly likely that biased projections
+will result, because the two errors are unlikely to be perfectly linked
+and thus will not always compensate each other. Hence, this structural
 problem of most models investigated here is problematic particularly if
@@
 
-Ingrowth levels define the structure and composition of future forests,
-but it is equally important to correctly identify areas where ingrowth
-is lacking [@rammer2021]. There are multiple constraints to the
-regeneration niche of tree species [@price2001], and therefore the
-absence of ingrowth is likely to be common [@fortin2007], even over
-larger areas such as the one-hectare plots used here. Tree ingrowth data
-are characterized by zero-inflation. This was clearly evident from the
-EuFoRIa dataset, but some of the models did not produce zeros at all, or
-featured a very low proportion of zero data. This substantial difference
-may be due to the fact that the simulation results were drawn from
-equilibrium forests, whereas in reality many of the forest reserves are
-recovering from past management activities and have become denser over
-the past decades [e.g., @heiri2009], leading to less ingrowth than in an
-equilibrium situation.
-
-## Species composition of ingrowth
-
-Correctly capturing species composition in tree establishment is
+Regeneration levels define the structure and composition of future
+forests, but it is equally important to correctly identify areas where
+regeneration is lacking [@rammer2021]. There are multiple constraints to
+the regeneration niche of tree species [@price2001], and therefore the
+absence of regeneration is likely to be common [@fortin2007], even over
+larger areas such as the one-hectare plots used here. Tree regeneration
+data are characterized by zero-inflation. This was clearly evident from
+the EuFoRIa dataset, but some of the models did not produce zeros at
+all, or featured a very low proportion of zero data. This substantial
+difference may be due to the fact that the simulation results were drawn
+from equilibrium forests, whereas in reality many of the forest reserves
+are recovering from past management activities and have become denser
+over the past decades [e.g., @heiri2009], leading to less regeneration
+than in an equilibrium situation.
+
+## Species composition of regeneration
+
+Correctly capturing species composition in tree regeneration is
 important to assess the future functional diversity of a forest, e.g.,
@@
 @seidl2022]. In the simulations, overall species diversity levels in the
-ingrowth were well within the observed range for half of the models (7
-out of 14). Thus, while most of the models are facing difficulties to
-quantitatively match ingrowth levels (cf. above), their performance is
-better when it comes to matching the composition of ingrowth as a
+regeneration were well within the observed range for half of the models
+(7 out of 14). Thus, while most of the models are facing difficulties to
+quantitatively match regeneration levels (cf. above), their performance
+is better when it comes to matching the composition of regeneration as a
 function of abiotic and biotic conditions. Most models maintained or
-even decreased ingrowth diversity between the 7 and 10 cm thresholds,
-and the same is visible from the empirical data, but the differences
-were not significant. Diversity variations may be explained by the way
-the individual models consider establishment processes
+even decreased regeneration diversity between the 7 and 10 cm
+thresholds, and the same is visible from the empirical data, but the
+differences were not significant. Diversity variations may be explained
+by the way the individual models consider regeneration processes
 [@konig_tree_2022], e.g., without considering the feedback from adults,
-ingrowth can be linked to either seed rain from dispersal or a
+regeneration can be linked to either seed rain from dispersal or a
 seed/seedling bank; neither process is incorporated across many models
@@
 The species diversity of the entire stand was captured better by the
-models than the diversity of tree ingrowth, and only a few models
-overpredicted stand-level diversity. Defining ingrowth composition as
-being proportional to adult abundance, regardless of productivity, might
-be a simple and conservative assumption to safely maintain relative
-species abundances [@hanbury-brown2022], although this may be
+models than the diversity of tree regeneration, and only a few models
+overpredicted stand-level diversity. Defining regeneration composition
+as being proportional to adult abundance, regardless of productivity,
+might be a simple and conservative assumption to safely maintain
+relative species abundances [@hanbury-brown2022], although this may be
 simplistic. Furthermore, based on the simulation results, there is no
 evidence that models with feedback from the canopy captured better the
-species diversity of the ingrowth than those without feedback, with the
-exception of ForCEEPS that significantly improved ingrowth species
-diversity with the model version that included feedback (ForCEEPS(f)).
-The similar performance of models with and without feedback is likely
-because the models put more weight on the regeneration niche arising
-from abiotic and biotic filters, than from the habitat niche of the
-adult trees [@grubb1977].
-
-## Ingrowth mortality
-
-There are several factors that lead to mortality during the
-establishment phase, such as competition [@casper1997] and multiple
-abiotic factors [@schmid2021; @cunningham2006]. As mentioned above, it
-is reasonable to expect that models that overestimate ingrowth may have
-a particularly high mortality between the 7 and 10 cm thresholds.
-However, this was not consistently evident from the simulations. This
-implies that these models must have an excess mortality in later stages,
-if we assume that all models were able to capture the structure and
+species diversity of the regeneration than those without feedback, with
+the exception of ForCEEPS that significantly improved regeneration
+species diversity with the model version that included feedback
+(ForCEEPS(f)). The similar performance of models with and without
+feedback is likely because the models put more weight on the
+regeneration niche arising from abiotic and biotic filters, than from
+the habitat niche of the adult trees [@grubb1977].
+
+## Regeneration mortality
+
+There are several factors that lead to mortality during the regeneration
+phase, such as competition [@casper1997] and multiple abiotic factors
+[@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
+expect that models that overestimate regeneration may have a
+particularly high mortality between the 7 and 10 cm thresholds. However,
+this was not consistently evident from the simulations. This implies
+that these models must have an excess mortality in later stages, if we
+assume that all models were able to capture the structure and
 composition of the adult stands along the EuFoRIa gradient; this however
@@
 and further studies are required that focus on a wider range of tree
-sizes and the fate of tree establishment along such a size continuum.
+sizes and the fate of tree regeneration along such a size continuum.
 
@@
 most models did not deviate exceedingly with respect to simulated
-ingrowth levels -- after all, few if any of the models' establishment
+regeneration levels -- after all, few if any of the models' regeneration
 routines are sufficiently constrained by data, with the exception of the
 empirical models xComp and SIBYLA. It is noteworthy that in spite of
-their empirical basis, these two models did not match empirical ingrowth
-levels, in a similar magnitude as the other models and even in opposite
-directions (SIBYLA: underestimation; xComp: overestimation). It appears
-that using empirical data of limited geographical scope and constrained
-to managed forests, as done in these models, leads to extrapolation
-problems already under current climatic conditions. For the other
-models, whose establishment modules are not strongly constrained by
-empirical data, the degrees of freedom that are available for modeling
-ingrowth [@konig_tree_2022] could have implied that model performance
-would be much worse than what we found.
-
-Our study showed that increasing complexity in the establishment modules
-did not represent an improvement in the projections of ingrowth levels,
-species composition or mortality at early tree stages, as there was no
-significant relationship between model performance and model complexity.
-Increasing complexity in establishment modules has been motivated by
-better process understanding and enhancing of model accuracy
-[@bugmann2022; @bugmann2019]. However, more complex models do not
-necessarily lead to better projections and rather to a reduced
+their empirical basis, these two models did not match empirical
+regeneration levels, in a similar magnitude as the other models and even
+in opposite directions (SIBYLA: underestimation; xComp: overestimation).
+It appears that using empirical data of limited geographical scope and
+constrained to managed forests, as done in these models, leads to
+extrapolation problems already under current climatic conditions. For
+the other models, whose regeneration modules are not strongly
+constrained by empirical data, the degrees of freedom that are available
+for modeling regeneration [@konig_tree_2022] could have implied that
+model performance would be much worse than what we found.
+
+Our study showed that increasing complexity in the regeneration modules
+did not represent an improvement in the projections of regeneration
+levels, species composition or mortality at early tree stages, as there
+was no significant relationship between model performance and model
+complexity. Increasing complexity in regeneration modules has been
+motivated by better process understanding and enhancing of model
+accuracy [@bugmann2022; @bugmann2019]. However, more complex models do
+not necessarily lead to better projections and rather to a reduced
 transparency and predictive power [@franklin2020]. Thus, the question of
 what level of detail is appropriate and parsimonious for modeling tree
-establishment processes remains open [@konig_tree_2022; @bugmann2022].
-
-## Ingrowth gradients on regeneration niches
-
-Competition for light as a strong filter for tree establishment has been
+regeneration processes remains open [@konig_tree_2022; @bugmann2022].
+
+## Regeneration gradients on regeneration niches
+
+Competition for light as a strong filter for tree regeneration has been
 widely documented [@messier1999; @collet2006; @berdanier2016], but the
@@
 the models and the observations. This made it impossible to evaluate the
-establishment for the extremes of the stand density ranges in some
-models. For example, ingrowth levels at low stand densities are relevant
-to assess how well forests are recovering e.g. after gap creation due to
-disturbance [@seidl2022; @grubb1977]. At the level of the simulated
-one-hectare samples, average basal area is typically not very low as
-long as no larger disturbances occur, which was explicitly excluded in
-our simulation protocol. Towards the other end of the spectrum, i.e.,
-with increasing stand basal area, it would be reasonable to expect that
-the ingrowth of the different species would become sparser and drop out
-entirely under low-light conditions [@klopcic2012; @zell2019]. However,
-few models showed this trend, thus indicating that the relation between
-ingrowth and light availability is not yet captured correctly in most
-models. Yet, several of the models that did feature an increase of
-ingrowth with increasing basal area include a feedback between seed
-production of mature trees and ingrowth. Thus, it seems that higher
-light competition does not sufficiently compensate for increased seed
-availability with higher basal area due to more mature trees.
+regeneration for the extremes of the stand density ranges in some
+models. For example, regeneration levels at low stand densities are
+relevant to assess how well forests are recovering e.g. after gap
+creation due to disturbance [@seidl2022; @grubb1977]. At the level of
+the simulated one-hectare samples, average basal area is typically not
+very low as long as no larger disturbances occur, which was explicitly
+excluded in our simulation protocol. Towards the other end of the
+spectrum, i.e., with increasing stand basal area, it would be reasonable
+to expect that the regeneration of the different species would become
+sparser and drop out entirely under low-light conditions [@klopcic2012;
+@zell2019]. However, few models showed this trend, thus indicating that
+the relation between regeneration and light availability is not yet
+captured correctly in most models. Yet, several of the models that did
+feature an increase of regeneration with increasing basal area include a
+feedback between seed production of mature trees and regeneration. Thus,
+it seems that higher light competition does not sufficiently compensate
+for increased seed availability with higher basal area due to more
+mature trees.
 
@@
 particular the dominant species *Abies alba*, *Fagus sylvatica* and
-*Picea abies*. The ingrowth levels were sampled from simulations in the
-equilibrium, and in this state it is expected that non-climax species
-such as *Pinus sylvestris* or *Quercus* would be of minor importance, or
-absent [@klopcic2015]. Most models captured this low abundance, which is
-also found in the empirical data. Thus, although the broad patterns are
-matched by many models, improvements in the quantification of the
-regeneration niche of the species are needed, but this cannot be done in
-the absence of robust datasets across multiple stand development stages.
+*Picea abies*. The regeneration levels were sampled from simulations in
+the equilibrium, and in this state it is expected that non-climax
+species such as *Pinus sylvestris* or *Quercus* would be of minor
+importance, or absent [@klopcic2015]. Most models captured this low
+abundance, which is also found in the empirical data. Thus, although the
+broad patterns are matched by many models, improvements in the
+quantification of the regeneration niche of the species are needed, but
+this cannot be done in the absence of robust datasets across multiple
+stand development stages.
 
@@
 The EuFoRIa data as used here are unique and well suited to better
-understand tree establishment. However, three aspects of these data may
+understand tree regeneration. However, three aspects of these data may
 also represent considerable limitations. First, we made a comparison of
-tree ingrowth in an equilibrium state, but we cannot assess how close
-the forests included in the EuFoRIa dataset were to such a state. The
-data were collected in forest reserves where no management has taken
+tree regeneration in an equilibrium state, but we cannot assess how
+close the forests included in the EuFoRIa dataset were to such a state.
+The data were collected in forest reserves where no management has taken
 place for long periods of time. This makes our assumption of an
@@
 primeval forests at demographic equilibrium, @brzeziecki2021 found
-higher ingrowth rates than the ones observed in the EuFoRia data. Thus,
-the overestimation of ingrowth rates by the models may not be so
-problematic. Second, the data were collected for ingrowth above a 7 cm
-threshold, thus limiting the assessment of tree establishment to a
+higher regeneration rates than the ones observed in the EuFoRia data.
+Thus, the overestimation of regeneration rates by the models may not be
+so problematic. Second, the data were collected for regeneration above a
+7 cm threshold, thus limiting the assessment of tree regeneration to a
 holistic viewpoint. This constituted a hard limit based on which we can
-understand only some aspects of tree establishment, which in its
-entirety often comprises a rather long period since seed production
-[@price2001]. In reality, many environmental constraints are acting on
-young trees [@käber2021] that we were unable to assess. Yet, ingrowth
-data with lower thresholds are simply not available in a harmonized
-manner across large environmental gradients. Third, the empirical data
-were collected from rather small plots, while we sampled simulated
-ingrowth levels from 1 ha areas, which may lead to an incorrect
-representation of space. Even though this strategy is not ideal, it
-represents a common challenge when harmonizing diverse data sources
-originating from varied sampling strategies [@portier2022]. It would
-have been extremely challenging for such a variety of models to follow a
-protocol where the spatial sampling size was different at each of the
-200 sites, and it would have introduced additional uncertainty in the
-results. We must not ignore that the observed data represent just a
-snapshot of an extremely complex and stochastic reality. This may at
-least partly explain e.g., the lack of clear patterns in the simulation
-results along the gradients of climate water balance and degree-days.
+understand only some aspects of tree regeneration, which in its entirety
+often comprises a rather long period since seed production [@price2001].
+In reality, many environmental constraints are acting on young trees
+[@käber2021] that we were unable to assess. Yet, regeneration data with
+lower thresholds are simply not available in a harmonized manner across
+large environmental gradients. Third, the empirical data were collected
+from rather small plots, while we sampled simulated regeneration levels
+from 1 ha areas, which may lead to an incorrect representation of space.
+Even though this strategy is not ideal, it represents a common challenge
+when harmonizing diverse data sources originating from varied sampling
+strategies [@portier2022]. It would have been extremely challenging for
+such a variety of models to follow a protocol where the spatial sampling
+size was different at each of the 200 sites, and it would have
+introduced additional uncertainty in the results. We must not ignore
+that the observed data represent just a snapshot of an extremely complex
+and stochastic reality. This may at least partly explain e.g., the lack
+of clear patterns in the simulation results along the gradients of
+climate water balance and degree-days.
 
@@
 range of models of forest dynamics, from stand to global scales, the
-simulation setup was limited to 1 ha ingrowth values. This lack of
+simulation setup was limited to 1 ha regeneration values. This lack of
 spatial considerations is appropriate for stand-scale models, but it
@@
 built to be accurate at the landscape level. Without spatial context, we
-are limiting ingrowth to the seed influx from the stand itself, unless
-the model has a background seed input. However, global models should not
-be at a disadvantage due to the limited spatial considerations [but cf.
-Snell et al. -@snell2018; or Lehsten et al. -@lehsten2019], as they are
-usually lacking dispersal between cells and are based on a strong
-abstraction of horizontal space [@hanbury-brown2022]. Lastly, detailed
-data on soil conditions were not available from the observed data, and
-independent, admittedly rather coarse data for soil and climatic water
-balance were used instead. It is noteworthy that several models
-represent drought based on more detailed indicators based e.g., on soil
-water holding capacity, which had to be derived from a rough soil
-quality measure. This may at least partially explain the poor
-performance of the models against the drought axis (i.e., climatic water
-balance).
+are limiting regeneration to the seed influx from the stand itself,
+unless the model has a background seed input. However, global models
+should not be at a disadvantage due to the limited spatial
+considerations [but cf. Snell et al. -@snell2018; or Lehsten et al.
+-@lehsten2019], as they are usually lacking dispersal between cells and
+are based on a strong abstraction of horizontal space
+[@hanbury-brown2022]. Lastly, detailed data on soil conditions were not
+available from the observed data, and independent, admittedly rather
+coarse data for soil and climatic water balance were used instead. It is
+noteworthy that several models represent drought based on more detailed
+indicators based e.g., on soil water holding capacity, which had to be
+derived from a rough soil quality measure. This may at least partially
+explain the poor performance of the models against the drought axis
+(i.e., climatic water balance).
 
@@
 Based on the results of this study, it is clear that models of forest
-dynamics need a focus on tree establishment modules to make them more
+dynamics need a focus on tree regeneration modules to make them more
 robust regarding growth and mortality. It remains uncertain what level
-of detail is required to model establishment processes, and this must be
+of detail is required to model regeneration processes, and this must be
 addressed in future research. We recommend that the improvement of the
-establishment modules is implemented as additional features that can be
+regeneration modules is implemented as additional features that can be
 traced back, as done here for the variants of ForClim and ForCEEPS, and
@@
 modeling objectives [@albrich2020]. If it should be necessary to include
-more detail in the establishment models, this will come with higher
+more detail in the regeneration models, this will come with higher
 parameterisation efforts. This will most likely lead to lower
 generalization because the required data will have to be collected from
-specific locations, as currently there is no general establishment
+specific locations, as currently there is no general regeneration
 dataset available.
@@
 Therefore, we recommend that more effort should be invested into
-collecting harmonized datasets on tree establishment. We emphasize that
+collecting harmonized datasets on tree regeneration. We emphasize that
 datasets like EuFoRIa used here are invaluable and should be further
@@
 
-In the present study, we have considered tree establishment in the
+In the present study, we have considered tree regeneration in the
 equilibrium state only. It is equally important to understand how these
-models project tree establishment after changes in forest structure by
+models project tree regeneration after changes in forest structure by
 disturbances [@seidl2022], or under different management strategies
@@
 be able to assess the relationship of these aspects on tree
-establishment, e.g., due to the lack of disturbance or appropriate
+regeneration, e.g., due to the lack of disturbance or appropriate
 management modules.
@@
 We recommend to further investigate the implications of the current
-modeling strategies for tree establishment and, ultimately, simulated
+modeling strategies for tree regeneration and, ultimately, simulated
 forest stand structure. This applies particularly to the erroneous
-patterns of excess tree establishment and later excess mortality, by
-focusing on a wider range of tree sizes and the related ingrowth
+patterns of excess tree regeneration and later excess mortality, by
+focusing on a wider range of tree sizes and the related regeneration
 dynamics. We also recommend, especially for the landscape-level models,
-the inclusion of explicit spatial considerations regarding tree ingrowth
-[@beckage2003]; this, however, is facing a serious challenge regarding
-appropriate datasets, as inventory data have wide coverage, but by
-definition do not allow for the assessment of spatial interaction
-effects.
+the inclusion of explicit spatial considerations regarding tree
+regeneration [@beckage2003]; this, however, is facing a serious
+challenge regarding appropriate datasets, as inventory data have wide
+coverage, but by definition do not allow for the assessment of spatial
+interaction effects.
 
@@
 a wide range of models with different scales, type, populations
-structures, tree establishment modules and complexity will ensure a
-wider benefit to the whole modeling community.
+structures, tree regeneration modules and complexity will ensure a wider
+benefit to the whole modeling community.
 
@@
 Models of forest dynamics are important tools in science and decision
-support. The formulation of tree establishment has strong implications
+support. The formulation of tree regeneration has strong implications
 for simulated forest properties. The 15 models and variants used here
 are facing similar challenges in their representation of tree
-establishment: they generally overestimate tree establishment levels,
-and the simulated establishment niche is not always captured accurately
-as a function of biotic (light) and abiotic (temperature and moisture)
+regeneration: they generally overestimate tree regeneration levels, and
+the simulated regeneration niche is not always captured accurately as a
+function of biotic (light) and abiotic (temperature and moisture)
 factors.
@@
 presence or absence of feedback from the adult trees, did not have a
-strong effect for capturing the species composition of ingrowth.
+strong effect for capturing the species composition of regeneration.
 
 Regarding mortality in the early phase of tree life, many models that
-feature a particularly high overestimation of ingrowth levels are
+feature a particularly high overestimation of regeneration levels are
 compensating for this by a larger tree mortality. Often, this
-compensation is not sufficient to reduce the high ingrowth levels to
+compensation is not sufficient to reduce the high regeneration levels to
 realistic values. Overall, there is no clear mortality pattern across
@@
 
-When capturing tree establishment, the specific design decisions taken
-in the development of any model are more important for its behavior
+When capturing tree regeneration, the specific design decisions taken in
+the development of any model are more important for its behavior
 (accuracy) than scale (stand, landscape, global), modelling approach
@@
 complexity does not represent an improvement for capturing tree
-establishment.
-
-Even though the establishment routines of most of the models
-investigated here have never been sufficiently constrained by robust
-data, the ingrowth projections are not overly off. This indicates that a
-lot can be gained by a focus on the modeling of establishment processes.
-The representation of forest dynamics in these models would become much
-more robust particularly in the face of climate change and
-post-disturbance dynamics, thus strongly reducing the uncertainty in
-long-term projections of future forest dynamics.
+regeneration.
+
+Even though the regeneration routines of most of the models investigated
+here have never been sufficiently constrained by robust data, the
+regeneration projections are not overly off. This indicates that a lot
+can be gained by a focus on the modeling of regeneration processes. The
+representation of forest dynamics in these models would become much more
+robust particularly in the face of climate change and post-disturbance
+dynamics, thus strongly reducing the uncertainty in long-term
+projections of future forest dynamics.
 

### 2023-04-26 manuscript/manuscriptRW.Rmd
changes in table 1 to remove the model classification

@@
 names(modelDF)[4] <- paste0(names(modelDF)[4], "(c)")
-names(modelDF)[5] <- paste0(names(modelDF)[5], "(d)")
-names(modelDF)[6] <- paste0(names(modelDF)[6], "(e)")
-names(modelDF)[11] <- paste0(names(modelDF)[11],"(f)")
+names(modelDF)[6] <- paste0(names(modelDF)[6], "(d)")
+names(modelDF)[11] <- paste0(names(modelDF)[11],"(e)")
 
@@
       kableExtra::column_spec(column = 2, width = "3em") |>
-      kableExtra::column_spec(column = 3, width = "3em") |>
-      kableExtra::column_spec(column = 4, width = "4em") |>
+      kableExtra::column_spec(column = 3, width = "4em") |>
+      kableExtra::column_spec(column = 4, width = "5em") |>
       kableExtra::column_spec(column = 5, width = "5em") |>
       kableExtra::column_spec(column = 6, width = "5em") |>
-      kableExtra::column_spec(column = 7, width = "5em") |>
+      kableExtra::column_spec(column = 7, width = "8em") |>
       kableExtra::column_spec(column = 8, width = "8em") |>
-      kableExtra::column_spec(column = 9, width = "8em") |>
+      kableExtra::column_spec(column = 9, width = "5em") |>
       kableExtra::column_spec(column = 10, width = "5em") |>
-      kableExtra::column_spec(column = 11, width = "5em") |>
-      kableExtra::column_spec(column = 12, width = "8em") |>
+      kableExtra::column_spec(column = 11, width = "8em") |>
       kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
@@
                                  "C-Cohort, I-Individual",
-                                 "RG-Regeneration, RC-Recruitment",
                                  "Y-Yes, N-No",

### 2023-04-28 manuscript/manuscriptRW.Rmd
changes from Harald

@@
 importance to humankind [@foresteurope2020]. Under the pressure of
-climate change, increasing disturbances impacts and changing societal
-demands on forests ecosystems services, it is becoming ever more
-important to understand how forest structure, composition and function
-will change, and to evaluate forest capacity to adapt to or be resilient
+climate change, increasing disturbance impacts and changing societal
+demands on forest ecosystem services, it is becoming ever more important
+to understand how forest structure, composition and function will
+change, and to evaluate forest capacity to adapt to or be resilient
 after disturbance [@lindner2010; @seidl2022]. A wide range of models of
 forest dynamics were developed over the past decades considering the
-impacts of climate on forest dynamics [@bugmann2022]. From these
-studies, it is evident that we have a reasonably good understanding how
-to model tree growth [@vanclay1997; @bugmann1996], and substantial
-efforts have been dedicated to improving the representation of tree
-mortality [@bugmann2019; @cailleret2017]. In contrast, tree regeneration
-is a much less studied process that is often represented rather coarsely
-in these models [@walck2011; @leishman1992; @price2001], which is a
-clear research gap in the context of climate-induced forest disturbances
-and forest resilience.
+impacts of climate [@bugmann2022]. From these studies, it is evident
+that we have a reasonably good understanding how to model tree growth
+[@vanclay1997; @bugmann1996], and substantial efforts have been
+dedicated to improving the representation of tree mortality
+[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is a much
+less studied process that is often represented rather coarsely in models
+[@walck2011; @leishman1992; @price2001]. Which is a clear research gap
+in the context of climate-induced forest disturbances and forest
+resilience.
 
@@
 statistical parameterizations, which however are not easy to extrapolate
-in space and time [e.g., @ribbens1994].
-
-Overall, models are needed to (1) synthesize existing empirical data and
-explore their relationships, (2) assess future tree regeneration e.g.,
-in the context of global change scenarios, and (3) identify the most
-important processes that are shaping ecological patterns. Given the
-current strategies that are used in models of forest dynamics to
-represent tree regeneration, model behavior often is prone to problems,
-such as very high levels of tree regeneration that necessitate excess
-mortality at early stages of tree life to simulate correct stand
-structure and composition. Also, inadequately high species diversity in
-tree regeneration may be simulated, which is characteristic of
-"classical" forest gap models [@gutiérrez2016], at least as long as the
-simulations set-up comprise a multitude of species. Some models use
-calibration against local canopy-level data to define appropriate
-regeneration levels, which is likely to strongly hamper the general
-applicability of these models e.g. under scenarios of climate change.
-Furthermore, correctly capturing the species composition in tree
-regeneration as a function of the presence of seed trees in the canopy
-is often a particular challenge, potentially leading to unrealistic
-successional drift in the model, which must be corrected by factors that
-are hard to parameterize [e.g., @lischke2006]. A related issue is the
-excessive reduction of species diversity due to positive feedback
-effects such that eventually just single-species stands remain. This is
-sometimes corrected by the incorporation of a low level of seed influx
-of all species at all times [@schumacher2006] or by restricting the
-number of seeds per species in the seedbank [@lischke2006]. However,
-simulated species composition is usually exceedingly sensitive to
-assumptions about seed availability, whereas the parameters of such
-functions are poorly constrained by field data. Lastly, there is often a
-problem with insufficient observational constraints on parameter values
-for models that start from very small tree sizes (e.g., 10 cm of height
-or even seeds) and track tree development in a process-oriented manner
-by considering a multitude of ecological influences, rather than
-emphasizing tree regeneration into a larger size class [e.g.,
-@wehrli2006].
+in space and time [e.g., @ribbens1994]. Overall, models are needed to
+(1) synthesize existing empirical data and explore their relationships,
+(2) assess future tree regeneration e.g. in the context of global change
+scenarios, and (3) identify the most important processes that are
+shaping ecological patterns. Given the current strategies that are used
+in models of forest dynamics to represent tree regeneration, model
+behavior often is prone to problems, such as very high levels of tree
+regeneration that necessitate excess mortality at early stages of tree
+life to simulate correct stand structure and composition. Also,
+inadequately high species diversity in tree regeneration may be
+simulated, which is characteristic of "classical" forest gap models
+[@gutiérrez2016], at least as long as the simulation set-up comprises a
+multitude of species. Some models use calibration against local
+canopy-level data to constrain simulated regeneration levels, which is
+likely to hamper the general applicability of these models e.g. under
+scenarios of climate change. Furthermore, correctly capturing the
+species composition in tree regeneration as a function of the presence
+of seed trees in the canopy is often a particular challenge, potentially
+leading to unrealistic successional drift in the model, which must be
+corrected by factors that are hard to parameterize [e.g., @lischke2006].
+A related issue is the excessive reduction of species diversity due to
+positive feedback effects, such that eventually just single-species
+stands remain. This is sometimes corrected by the incorporation of a low
+level of seed influx of all species at all times [@schumacher2006] or by
+restricting the number of seeds per species in the seedbank
+[@lischke2006]. However, simulated species composition is usually
+exceedingly sensitive to assumptions about seed availability, whereas
+the parameters of such functions are poorly constrained by field data.
+Lastly, there is often a problem with insufficient observational
+constraints on parameter values for models that start from very small
+tree sizes (e.g., 10 cm of height or even seeds) and track tree
+development in a process-oriented manner by considering a multitude of
+ecological influences, rather than emphasizing tree regeneration into a
+larger size class [e.g., @wehrli2006].
 
@@
 [@hanbury-brown2022], although many sub-processes have been studied in
-detail [@miina2006]. Yet, little data is available that covers all the
+detail [@miina2006]. Yet, little data are available that covers all the
 processes within one species along environmental gradients, let alone
@@
 which constitutes a major problem for model building [@clark1999]. On
-the one hand, monitoring on permanent sample plots (such as National
-Forest Inventories) often measures tree regeneration [@zell2019],
-however with a design that captures the rate after the stand initiation
-phase [@hallsby2015]. Data from permanent plots, although available
-across regions, are highly heterogeneous, rendering their use difficult
-in a modeling context [@käber2021]. On the other hand, targeted studies
-to measure tree regeneration on experimental sites are often limited in
+the one hand, monitoring on permanent plots (such as National Forest
+Inventories) often measures tree regeneration [@zell2019], however with
+a design that captures the rate after the stand initiation phase
+[@hallsby2015]. Data from permanent plots, although available across
+regions, are highly heterogeneous, rendering their use difficult in a
+modeling context [@käber2021]. On the other hand, targeted studies to
+measure tree regeneration on experimental sites are often limited in
 spatial extent, species studied, or the sub-set of processes that are
@@
 models from empirically derived to "process-based" [e.g., @bugmann1996;
-@fabrika2012; @lexer2001; @reyer2014]. Because of the large variability
-in tree regeneration patterns in nature and the large number of factors
-driving this process --- including some that are not incorporated in
-most models such as deer browsing --- we do not aim at a detailed
-statistical evaluation of each model. Instead, we aim at evaluating the
-general regeneration patterns and magnitudes simulated by the models,
-and benchmark the simulated regeneration niche of multiple species
-against empirical data along a wide environmental gradient of
+@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
+tree regeneration patterns in nature and the large number of factors
+driving this process --- including some that are not incorporated
+explicitly in most models, such as deer browsing --- we do not aim for a
+detailed statistical evaluation of each model. Instead, we aim to
+evaluate the general regeneration patterns and magnitudes simulated by
+the models, and benchmark the simulated regeneration niche of multiple
+species against empirical data along a wide environmental gradient of
 temperature, moisture, and light availability [@grubb1977].
@@
 Fifteen models of forest dynamics (including two models featuring model
-variants) were used to assess simulated tree regeneration. They simulate
-forest dynamics at the stand, landscape or global scale (Table
-\@ref(tab:tableModels)). The approaches used for model construction and
-their origin differ strongly, with most of the models featuring a
-largely "process-based" approach, whereas two models are based on
-formulations derived from the statistical analysis of inventory data
-(SIBYLA and xComp). Some of the models largely rely on the original
-approach underlying forest gap models (e.g., ForClim 1 and LandClim),
+variants) were used to simulate forest dynamics at the stand, landscape
+or global scale (Table \@ref(tab:tableModels)). The approaches used for
+model construction and their origin differ strongly, with most of the
+models featuring a largely "process-based" approach, whereas two models
+are based on formulations derived from the statistical analysis of
+inventory data (SIBYLA and xComp). Some of the models largely rely on
+the original approach underlying forest gap models (e.g., ForClim 1),
 whereas some are based on plant ecophysiological processes (e.g.,
@@
 
-The regeneration modules has been categorized as 'regeneration' or
-'recruitment' models [@vanclay1994; @konig_tree_2022]. *Regeneration
+The regeneration models can be differentiated into 'regeneration' and
+'recruitment' approaches [@vanclay1994; @konig_tree_2022]. *Regeneration
 models* include processes such as flowering and pollination, seed
@@
 from saplings have to aggregate via parameterizations several ecological
-processes which are not treated explicitly. This later approach reduces
+processes that are not treated explicitly. This latter approach reduces
 model complexity but comes at the cost of blurring process
@@
 
-In the models used in this study, the overall complexity in the
-regeneration modules varies considerably. Following @bugmann2022, we can
-classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS,
-ForClim and TreeMig) as having rather high complexity in their
-regeneration modules (mean regeneration formulation complexity across
-all processes \>0, Table \@ref(tab:tableModels)). Two models, ForClim
-and ForCEEPS, were used here with two alternative variants of
-regeneration. ForClim variant 1 [@bugmann1996] is based on a recruitment
-module that adheres closely to the concept introduced by @botkin1972,
-whereas ForClim variant 11 is adopting a different approach where
-individual species properties and their relation to the environment are
-important [@huber2020]. These two model variants allow us to evaluate a
-more process-based and complex module (variant 11) against a simple
-module (variant 1), while the rest of the model structure is identical.
-Similarly, the two ForCEEPS variants allow us to isolate the importance
-of the canopy feedback (i.e., simulated actual composition and relative
-abundance of species in the plot) via seed trees for the quantity and
-quality (e.g., diversity and composition) of simulated regeneration, as
-one variant includes this feedback whereas the other does not.
+In the models used here, the overall complexity in the regeneration
+modules varies considerably. Following @bugmann2022, we can classify
+seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and
+TreeMig) as having rather high complexity in their regeneration modules
+(mean regeneration formulation complexity across all processes \>0,
+Table \@ref(tab:tableModels)). Two models, ForClim and ForCEEPS, were
+used here with two alternative variants of regeneration. ForClim variant
+1 [@bugmann1996] is based on a recruitment module that adheres closely
+to the concept introduced by @botkin1972, whereas ForClim variant 11 is
+adopting a different approach where individual species properties and
+their relation to the environment are important [@huber2020]. These two
+model variants allow us to evaluate a more process-based and complex
+module (variant 11) against a simple module (variant 1), while the rest
+of the model structure is identical. Similarly, the two ForCEEPS
+variants allow us to isolate the importance of the canopy feedback
+(i.e., simulated actual composition and relative abundance of species in
+the plot) via seed trees for the quantity and quality (e.g., diversity
+and composition) of simulated regeneration, as one variant includes this
+feedback whereas the other does not.
 
@@
 periods range from 3 to 37 years, with an average of 14 years. The data
-provide information at the tree level, thus allowing for sequential
-comparison such as individual-tree regeneration and death between the
-measurements.
+provide information at the tree level, thus allowing for the sequential
+comparison of processes such as individual-tree regeneration and death
+between the measurements.
 
 We selected 200 sites from this network as the benchmarking dataset for
-the simulation. The selection of sites is representative of the
-environmental variation of the data. This was achieved by applying
-k-means clustering to define 200 clusters of plots from the original set
-of 869 plots along the environmental dimensions of temperature, climatic
-water balance, soil quality and exposition (i.e., slope and aspect).
-Each of these sites featured at least two consecutive measurements.
-egeneration thresholds for the sites differed between diameters of 0 and
-10 cm. For our study, we defined two datasets where one included 165
-sites with a diameter threshold of 7 cm or lower, and the other included
-another 35 sites with diameter thresholds between 7 and 10 cm.
+the simulation to be representative of the environmental variation
+contained in the data. This was achieved by applying k-means clustering
+to define 200 clusters of plots from the original set of 869 plots along
+the environmental dimensions of temperature, climatic water balance,
+soil quality and exposition (i.e., slope and aspect). Each of these
+sites featured at least two consecutive measurements. Regeneration
+thresholds for these sites differed between diameters of 0 and 10 cm.
+For our study, we defined two datasets where one included 165 sites with
+a diameter threshold of 7 cm or lower, and the other included another 35
+sites with diameter thresholds between 7 and 10 cm.
 
@@
 rates per ha and per decade. The observations used in this study
-featured 30,900 established trees. Regeneration rates per site, sample
-and decade ranged from zero to 1246 trees per ha, with a mean of 56
-trees per ha. Adult species composition was also available at the plots.
-For more details on this unique dataset and the detailed regeneration
+featured 30,900 newly established trees. Regeneration rates per site,
+sample and decade ranged from zero to 1246 trees, with a mean of 56
+trees. Adult species composition was also available for each plot. For
+more details on this unique dataset and the detailed regeneration
 information, cf. @käber2022.
@@
 regeneration as it arises from empirical data against its representation
-in a wide range of models of forest dynamics. We define the tree
+in a wide range of models of forest dynamics. We define tree
 regeneration as the passing of a breast height diameter threshold of 7
@@
 group was provided with a detailed protocol [@díaz-yáñez2022] with
-instructions for how to perform the simulations, providing input
-variables on climate and soil conditions, and the list of expected
-output variables. Neither were further site information (except for the
-data specified below) nor any data on tree regeneration or forest stand
+instructions how to perform the simulations, providing input variables
+on climate and soil conditions, and the list of expected output
+variables. Neither were further site information (except for the data
+specified below) nor any data on tree regeneration or forest stand
 features provided prior to the simulation. That is, the models were run
@@
 deficit, were calculated from these variables. The final instructions
-for using the climate data was different in each model, based on the
-approach that best suited the model (Table \@ref(tab:tableModels)). Soil
-quality data were provided as continuous values between 1 and 5
-[Soilgrids dataset, Hengl et al. -@hengl2017]. The protocol also
-provided the elevation, slope and aspect for each of the 200 sites
-[@asterscienceteam2019], but no other spatial information such as
-coordinates, with the exception of iLand and aDGVM2, which required
-blurred coordinates to derive highly detailed soil data.
+for using the climate data were slightly different in each model, based
+on the approach that best suited the model (Table
+\@ref(tab:tableModels)). Soil quality data were provided as continuous
+values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
+The protocol also provided the elevation, slope and aspect for each of
+the 200 sites [@asterscienceteam2019], but no other spatial information
+such as coordinates, with the exception of iLand and aDGVM2, which
+required blurred coordinates to derive highly detailed soil data.
 
@@
 as the ultimate goal was to evaluate tree regeneration under comparable
-and near equilibrium conditions. This entails the assumption that the
-observations from the forest reserves (1) reflect no traces of forest
-management and (2) there is an equilibrium between forest dynamics and
+and near-equilibrium conditions. This entails the assumption that (1)
+the observations from the forest reserves reflect no traces of forest
+management, and (2) there is an equilibrium between forest dynamics and
 climate. While the former might be starting to be visible in many of the
@@
 
-The simulations were run for mixed species forests (not multiple
+The simulations were run for mixed-species forests (not multiple
 single-species simulations) using mixtures of eleven species or genera
@@
 abies*, *Betula* spp., *Quercus* spp. and *Pinus sylvestris* are
-parameterised. XComp simulations did not consider *Fraxinus excelsior*
+parameterised. xComp simulations did not consider *Fraxinus excelsior*
 due to a depreciated species parameterisation. In LPJ-GUESS, *Acer
@@
 several 1 ha plots from the simulated forest at one specific point in
-time (in the equilibrium) or (3) a combination of (1) and (2) . This
+time (in the equilibrium) or (3) a combination of (1) and (2). This
 resulted in 880,000 observations per model that simulated the 11 species
@@
 (i.e., ingrowth number per unit time and space), (2) regeneration
-species diversity, (3) regeneration mortality, (4) model performance and
-model traits, and (5) ingrowth gradients along regeneration niches. We
-evaluated the species diversity across models and in relation to the
-observed data by calculating the Shannon index based on the relative
-proportion of the species in terms of basal area. The Shannon index was
-calculated for the regeneration ($H_{R_{n}}$) (Equation
-\ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation
-\ref{eqpiShannon}). The higher the value of the index, the higher the
-species diversity at a particular site and sample.
+species diversity, (3) regeneration mortality, (4) the relationship
+between model performance and model traits, and (5) ingrowth gradients
+along the regeneration niches. We evaluated species diversity across the
+models and in relation to the observed data by calculating the Shannon
+index $H$ based on the relative proportion of the species in terms of
+basal area. It was calculated for the regeneration ($H_{R_{n}}$)
+(Equation \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$)
+(Equation \ref{eqpiShannon}). The higher the value of the index, the
+higher the species diversity at a particular site and sample.
 
@@
 area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
-the species $i$ in sample $n$ calculated as the regeneration basal area
-($r_{BA_{n,i}}$) for that species $i$ relative to the total recruited
-basal area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
-proportion of the species $i$ calculated as the basal area of all trees
-($BA_{n,i}$) of that species $i$ relative to the total basal area
-($TotalR_{BA_{n}}$) of the sample $n$. The species diversity was not
-assessed for the model aDGVM2 because it does not simulate individual
-species.
+species $i$ in sample $n$ calculated as the regeneration basal area
+($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal
+area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
+proportion of species $i$ calculated as the basal area of all trees
+($BA_{n,i}$) of that species $i$ relative to total basal area
+($TotalR_{BA_{n}}$) of the sample $n$. Species diversity was not
+assessed for aDGVM2 as this model does not simulate individual species.
 
@@
 to estimate whether the ratio of regeneration between the 7 and 10 cm
-diameter thresholds were above or below the expected theoretical rate
-between these two diameters. The Reineke self-thinning rule is usually
-calculated from even-aged, single species stands and assumes a fixed
-relationship between the number of stems and the quadratic mean diameter
-in fully stocked pure stands. The value used in our comparisons was 1.77
-(i.e., we expect stem numbers at 7 cm to be 77% higher than at 10 cm),
-calculated using Equation \ref{eqReineke}.
+diameter thresholds were above or below the expected theoretical rate.
+The Reineke self-thinning rule is usually calculated for even-aged,
+single species stands and assumes a fixed relationship between the
+number of stems and the quadratic mean diameter in fully stocked pure
+stands. The value used in our comparisons was 1.77 (i.e., we expect stem
+numbers at 7 cm to be 77% higher than at 10 cm), calculated using
+Equation \ref{eqReineke}.
 
@@
 the trees in each sample including tree regeneration) was used as a
-proxy of light availability at the forest floor, the annual degree-day
+proxy for light availability at the forest floor, the annual degree-day
 sum [@allen1976; @fischlin1995] as a proxy for growing season warmth,
@@
 In order to analyze the regeneration niches across the climatic
-gradients, we focused on the five main tree species or genera: *Abies
+gradients, we focused on five common tree species or genera: *Abies
 alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
@@
 where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
-is the mean basal area regeneration of the species $i$ per site $t$
-across the available samples at that site. Furthermore, regarding the
-share in the regeneration basal area per species
-($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the
-regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero
-but also when both the regeneration basal area of that species
-($\bar{r}_{BA_{t,i}}$) and the total regeneration basal area (
-$\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.
+is the mean basal area in the regeneration (subsequently referred to as
+"regeneration basal area") of species $i$ per site $t$ across the
+available samples at that site. Furthermore, regarding the share in the
+regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we
+categorized this as zero when the regeneration basal area of that
+species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the
+regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) and the
+total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$)
+equaled zero.
 
@@
 Most models matched the level of diversity of the observed data quite
-well as quantified by the Shannon index (Figure \@ref(fig:H710)). Five
-models overestimated regeneration diversity: ForCEEPS, ForClim 1, PICUS,
-TreeMig and LandClim, the latter particularly for the 7 cm diameter
-threshold. The model 4C is a special case, as it simulated five species
-only, i.e., its diversity values are not directly comparable to those of
-the other models, nor to the observations. Only one model, Landis-II,
-consistently underestimated regeneration diversity.
+well (Figure \@ref(fig:H710)). Five models overestimated regeneration
+diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
+particularly for the 7 cm diameter threshold. The model 4C is a special
+case, as it simulated five species only, i.e., its diversity values are
+not directly comparable to those of the other models, nor to the
+observations. Only one model, Landis-II, consistently underestimated
+regeneration diversity.
 
@@
 thresholds, species diversity in the regeneration compared to
-stand-level diversity did not have clear patterns across the models
+stand-level diversity did not feature clear patterns across the models
 (Figures S1 and S2). Most models captured reasonably well or
@@
 \@ref(fig:combinedRichness)A, left and center; cf. Figure S3), but a
-characteristic feature was that simulations had a much lower variability
-of diversity compared to observations. Models overpredicting species
-diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure
-\@ref(fig:combinedRichness)A, left). Several models underpredicted
-species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure
-\@ref(fig:combinedRichness)A, right).
+characteristic feature was that the simulations had a much lower
+variability of diversity compared to observations. Models overpredicting
+species diversity included ForClim1, ForClim 11, FORMIND and TreeMig
+(Figure \@ref(fig:combinedRichness)A, left). Several models
+underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C
+(Figure \@ref(fig:combinedRichness)A, right).
 
 The majority of the models overpredicted regeneration diversity; as
-already observed in figure \@ref(fig:H710), only two of them
-underestimated regeneration diversity (4C and Landis-II) (Figure
+already observed in Figure \@ref(fig:H710), only two of them
+underestimated it (4C and Landis-II) (Figure
 \@ref(fig:combinedRichness)B and Figure S4). Both the models with
 feedback and those without feedback from the adult trees to regeneration
-(via seed production) overestimate, underestimated and captured
+(via seed production) overestimated, underestimated or captured
 reasonably well regeneration species diversity compared to the
-observations. Again, simulated diversity was considerably less variable
-than observed diversity in the regeneration.
+observations. Again, simulated diversity had considerably lower
+variability than observed diversity in the regeneration.
 
@@
 self-thinning are either not pronounced or must have occurred before the
-species had reached 7 cm. However, this pattern was not consistent
-across models (Figure S5). It is noteworthy that there is a relationship
+trees had reached 7 cm. However, this pattern was not consistent across
+models (Figure S5). It is noteworthy that there is a relationship
 between the regeneration levels at 7 cm and the mortality rate (Table
@@
 7 cm is coupled to an increase of the mortality between 7 and 10 cm).
-While some models showed a negative effect, their linear trend was not
+While some models showed a negative effect, the linear trend was not
 pronounced and not always significant regarding the slope.
@@
 availability (basal area), temperature (degree-days) and soil moisture
-(climatic water balance), distinct features emerge: the models
+(climatic water balance), distinct features emerged: the models
 reproduced the effect of basal area in both its magnitude and patterns
@@
 First, some models overestimated regeneration across the entire
-gradients for at least one of the five main tree species (Figure
+gradients of at least one of the five main tree species (Figure
 \@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS),
-while other models overestimated regeneration for more than one species
+while other models overestimated regeneration of more than one species
 for a large part of the climate gradients, such as xComp, PICUS, SIBYLA
@@
 environmental gradients featuring negligible regeneration at most sites,
-while others overestimate it in different parts of the environmental
+while others overestimated it in different parts of the environmental
 space. Most models overestimated the share of *Picea abies* regeneration
 in the cold-wet part of the gradients. *Pinus sylvestris* and *Quercus*
-spp. represented a smaller share of the regeneration in the observed
-data mostly at warm-dry sites. Many models, however, erroneously
-featured regeneration for these two species along most of the climatic
-gradients although at a low basal area share. Five models (4C, FORMIND,
+spp. represented a small share of the regeneration in the observed data
+mostly at warm-dry sites. Many models, however, erroneously featured
+regeneration for these two species along most of the climatic gradients,
+although with a low share of basal area. Five models (4C, FORMIND,
 SIBYLA, xComp, and LandClim) had almost no regeneration of *Pinus
 sylvestris* and *Quercus* spp., or did not feature any regeneration of
-these species at all, close to the patterns in the observed data.
+these species at all.
 
@@
 
-Tree regeneration is a fundamental process in forest dynamics, in
-particular under changing conditions such as those caused by climate
-change. Correctly capturing tree regeneration in dynamic models is
-fundamental to, for example, evaluate post-disturbance dynamics and
-potential long-term recovery trajectories as it will define the forest
-state in the first decades [@seidl2022]. If this initial phase is not
-captured well, we can not properly assess aspects such as forest
-resilience or the timing, magnitude and progression of carbon
-sequestration. In this study, for the first time, projections of tree
-regeneration from multiple models of forest dynamics were confronted
-with a unique dataset from unmanaged forest reserves across a large
-environmental gradient in Europe.
-
-The observed data that we used [@käber2022] are exceptional,
-particularly with respect to the number of records captured (number of
-sites and repeated measurements), which is essential for a highly
-'noisy' process such as tree regeneration. The use of this dataset for
-model benchmarking provided novel insights on the ability of
-state-of-the-art models to accurately simulate regeneration levels,
-species composition, and mortality in an early stage of tree life, i.e.,
-between a tree diameter of 7 and 10 cm. Overall, by adopting this
-approach a much broader understanding resulted than if we had used these
-data for model calibration: it is primarily from the shortcomings of the
-models that we can gain ecological insights [e.g., @trugman2022].
+Tree regeneration is a fundamental process in forest dynamics. Correctly
+capturing tree regeneration in dynamic models is fundamental to, for
+example, evaluate post-disturbance dynamics and potential long-term
+recovery trajectories as it will define the forest state in the first
+decades [@seidl2022]. If this initial phase is not captured well, we can
+not properly assess aspects such as forest resilience or the timing,
+magnitude and progression of carbon sequestration. In this study, for
+the first time, projections of tree regeneration from multiple models of
+forest dynamics were confronted with a unique dataset from unmanaged
+forest reserves across a large environmental gradient in Europe.
+
+The EuFoRIa data [@käber2022] are exceptional, particularly with respect
+to the number of records (number of sites and repeated measurements),
+which is essential for capturing a highly 'noisy' process such as tree
+regeneration. The use of this dataset for model benchmarking provided
+novel insights on the ability of state-of-the-art models to accurately
+simulate regeneration levels, species composition, and mortality in an
+early stage of tree life, i.e., between a tree diameter of 7 and 10 cm.
+Overall, by adopting this approach a much broader understanding resulted
+than if we had used these data for model calibration: it is primarily
+from the shortcomings of the models that we can gain ecological insights
+[cf. @trugman2022].
 
@@
 to compensating for a first error (excess regeneration) by a second
-"error" (excess mortality). It is highly likely that biased projections
+error (excess mortality). It is highly likely that biased projections
 will result, because the two errors are unlikely to be perfectly linked
@@
 In spite of the critical considerations above, it is remarkable that
-most models did not deviate exceedingly with respect to simulated
-regeneration levels -- after all, few if any of the models' regeneration
-routines are sufficiently constrained by data, with the exception of the
-empirical models xComp and SIBYLA. It is noteworthy that in spite of
-their empirical basis, these two models did not match empirical
-regeneration levels, in a similar magnitude as the other models and even
-in opposite directions (SIBYLA: underestimation; xComp: overestimation).
-It appears that using empirical data of limited geographical scope and
-constrained to managed forests, as done in these models, leads to
-extrapolation problems already under current climatic conditions. For
-the other models, whose regeneration modules are not strongly
-constrained by empirical data, the degrees of freedom that are available
-for modeling regeneration [@konig_tree_2022] could have implied that
-model performance would be much worse than what we found.
+most models did not deviate exceedingly from the observations with
+respect to simulated regeneration levels -- after all, few if any of the
+models' regeneration routines are sufficiently constrained by data, with
+the exception of the empirical models xComp and SIBYLA. It is noteworthy
+that in spite of their empirical basis, these two models did not match
+empirical regeneration levels, in a similar magnitude as the other
+models and even in opposite directions (SIBYLA: underestimation; xComp:
+overestimation). It appears that using empirical data of limited
+geographical scope and constrained to managed forests, as done in these
+models, leads to extrapolation problems already under current climatic
+conditions. For the other models, whose regeneration modules are not
+strongly constrained by empirical data, the degrees of freedom that are
+available for modeling regeneration [@konig_tree_2022] could have
+implied that model performance would be much worse than what we found.
 
 Our study showed that increasing complexity in the regeneration modules
-did not represent an improvement in the projections of regeneration
+is not linked with a higher accuracy of the projections of regeneration
 levels, species composition or mortality at early tree stages, as there
@@
 motivated by better process understanding and enhancing of model
-accuracy [@bugmann2022; @bugmann2019]. However, more complex models do
-not necessarily lead to better projections and rather to a reduced
-transparency and predictive power [@franklin2020]. Thus, the question of
-what level of detail is appropriate and parsimonious for modeling tree
-regeneration processes remains open [@konig_tree_2022; @bugmann2022].
+accuracy [@bugmann2022]. However, more complex models do not necessarily
+lead to better projections and rather to a reduced transparency and
+lower predictive power [@franklin2020]. Thus, the question of what level
+of detail is appropriate and parsimonious for modeling tree regeneration
+processes remains open [@konig_tree_2022; @bugmann2022].
 
@@
 scales. We used total stand basal area as a proxy for light availability
-[cf. @schmid2021]. However we could not consider light availability
-restrictions caused by ground vegetation which might be an important
-filter for some models [@woltjer2008]. We found pronounced differences
-in the stand basal area ranges simulated by the models, but also between
-the models and the observations. This made it impossible to evaluate the
-regeneration for the extremes of the stand density ranges in some
-models. For example, regeneration levels at low stand densities are
-relevant to assess how well forests are recovering e.g. after gap
-creation due to disturbance [@seidl2022; @grubb1977]. At the level of
-the simulated one-hectare samples, average basal area is typically not
-very low as long as no larger disturbances occur, which was explicitly
-excluded in our simulation protocol. Towards the other end of the
-spectrum, i.e., with increasing stand basal area, it would be reasonable
-to expect that the regeneration of the different species would become
-sparser and drop out entirely under low-light conditions [@klopcic2012;
-@zell2019]. However, few models showed this trend, thus indicating that
-the relation between regeneration and light availability is not yet
-captured correctly in most models. Yet, several of the models that did
-feature an increase of regeneration with increasing basal area include a
-feedback between seed production of mature trees and regeneration. Thus,
-it seems that higher light competition does not sufficiently compensate
-for increased seed availability with higher basal area due to more
-mature trees.
-
-There were pronounced differences in how the main tree species were
+[cf. @schmid2021]. However, we were unable to consider light
+availability restrictions caused by ground vegetation, which may be an
+important filter for some EuForIa stands [@woltjer2008]. We found
+pronounced differences in the stand basal area ranges simulated by the
+models, but also between models and observations. This made it
+impossible to evaluate the regeneration for the extremes of the stand
+density ranges in some models. For example, regeneration levels at low
+stand densities are relevant to assess how well forests are recovering
+e.g. after gap creation due to disturbance [@seidl2022; @grubb1977]. At
+the level of the simulated one-hectare samples, average basal area is
+typically not very low as long as no larger disturbances occur, which
+was explicitly excluded in our simulation protocol. Towards the other
+end of the spectrum, i.e., with increasing stand basal area, it would be
+reasonable to expect that the regeneration of the different species
+would become sparser and drop out entirely under low-light conditions
+[@klopcic2012; @zell2019]. However, few models showed this trend, thus
+indicating that the relation between regeneration and light availability
+is not yet captured correctly in most models. Yet, several of the models
+that did feature an increase of regeneration with increasing basal area
+include a feedback between seed production of mature trees and
+regeneration. Thus, it seems that higher light competition does not
+sufficiently compensate for increased seed availability with higher
+basal area due to a higher abundance of mature trees.
+
+There were pronounced differences how the main tree species were
 represented by the models along the environmental gradients, in
@@
 understand tree regeneration. However, three aspects of these data may
-also represent considerable limitations. First, we made a comparison of
-tree regeneration in an equilibrium state, but we cannot assess how
-close the forests included in the EuFoRIa dataset were to such a state.
-The data were collected in forest reserves where no management has taken
-place for long periods of time. This makes our assumption of an
-equilibrium between forest properties and environmental drivers more
-reasonable than it might appear at first sight. In an analysis of
-primeval forests at demographic equilibrium, @brzeziecki2021 found
-higher regeneration rates than the ones observed in the EuFoRia data.
-Thus, the overestimation of regeneration rates by the models may not be
-so problematic. Second, the data were collected for regeneration above a
-7 cm threshold, thus limiting the assessment of tree regeneration to a
-holistic viewpoint. This constituted a hard limit based on which we can
-understand only some aspects of tree regeneration, which in its entirety
-often comprises a rather long period since seed production [@price2001].
-In reality, many environmental constraints are acting on young trees
-[@käber2021] that we were unable to assess. Yet, regeneration data with
-lower thresholds are simply not available in a harmonized manner across
-large environmental gradients. Third, the empirical data were collected
-from rather small plots, while we sampled simulated regeneration levels
-from 1 ha areas, which may lead to an incorrect representation of space.
-Even though this strategy is not ideal, it represents a common challenge
-when harmonizing diverse data sources originating from varied sampling
+represent considerable limitations. First, we made a comparison of tree
+regeneration in an equilibrium state, but we cannot assess how close the
+forests included in the EuFoRIa dataset were to such a state. The data
+were collected in forest reserves where no management has taken place
+for long periods of time. This makes our assumption of an equilibrium
+between forest properties and environmental drivers more reasonable than
+it might appear at first sight. In an analysis of primeval forests at
+demographic equilibrium, @brzeziecki2021 found higher regeneration rates
+than the ones observed in the EuFoRia data. Thus, the overestimation of
+regeneration rates by the models may not be so problematic. Second, the
+data were collected for regeneration above a 7 cm threshold, thus
+limiting the assessment of tree regeneration to a holistic viewpoint.
+This constituted a hard limit based on which we can understand only some
+aspects of tree regeneration, which in its entirety often comprises a
+rather long period since seed production [@price2001]. In reality, many
+environmental constraints are acting on young trees [@käber2021] that we
+were unable to assess. Yet, regeneration data with lower thresholds are
+simply not available in a harmonized manner across large environmental
+gradients. Third, the empirical data were collected from rather small
+plots, while we sampled simulated regeneration levels from 1 ha areas,
+which may lead to an incorrect representation of space. Even though the
+strategy we adopted is not ideal, it represents a common challenge when
+harmonizing diverse data sources originating from varied sampling
 strategies [@portier2022]. It would have been extremely challenging for
@@
 size was different at each of the 200 sites, and it would have
-introduced additional uncertainty in the results. We must not ignore
-that the observed data represent just a snapshot of an extremely complex
-and stochastic reality. This may at least partly explain e.g., the lack
-of clear patterns in the simulation results along the gradients of
-climate water balance and degree-days.
+introduced additional uncertainty in the results. Ultimately, we need to
+acknowledge that the observed data represent just a snapshot of an
+extremely complex and stochastic reality. This may at least partly
+explain e.g. the lack of clear patterns in the simulation results along
+the gradients of climate water balance and degree-days.
 
 The design of our sampling protocol did not include spatial aspects such
-as seed dispersal or detailed soil data. Although, we considered a wide
+as seed dispersal or detailed soil data. While we considered a wide
 range of models of forest dynamics, from stand to global scales, the
 simulation setup was limited to 1 ha regeneration values. This lack of
-spatial considerations is appropriate for stand-scale models, but it
-potentially puts landscape models at a disadvantage, as they have been
-built to be accurate at the landscape level. Without spatial context, we
-are limiting regeneration to the seed influx from the stand itself,
-unless the model has a background seed input. However, global models
-should not be at a disadvantage due to the limited spatial
+consideration of spatial scale is appropriate for stand-scale models,
+but it potentially puts the landscape models at a disadvantage, as they
+have been built to be accurate at the landscape level. Without spatial
+context, we are limiting regeneration to the seed influx from the stand
+itself, unless the model has a background seed input. Yet, the global
+models should not be at a disadvantage due to the limited spatial
 considerations [but cf. Snell et al. -@snell2018; or Lehsten et al.
@@
 available from the observed data, and independent, admittedly rather
-coarse data for soil and climatic water balance were used instead. It is
-noteworthy that several models represent drought based on more detailed
-indicators based e.g., on soil water holding capacity, which had to be
+coarse data for soil and climatic water balance had to be used instead.
+It is noteworthy that many models represent drought based on detailed
+indicators based e.g. on soil water holding capacity, which had to be
 derived from a rough soil quality measure. This may at least partially
-explain the poor performance of the models against the drought axis
-(i.e., climatic water balance).
+explain the unsatisfactory performance of many models along the drought
+axis (i.e., climatic water balance).
 
@@
 
-Based on the results of this study, it is clear that models of forest
-dynamics need a focus on tree regeneration modules to make them more
-robust regarding growth and mortality. It remains uncertain what level
-of detail is required to model regeneration processes, and this must be
-addressed in future research. We recommend that the improvement of the
-regeneration modules is implemented as additional features that can be
-traced back, as done here for the variants of ForClim and ForCEEPS, and
-that model complexity and structure must always be connected with
-modeling objectives [@albrich2020]. If it should be necessary to include
-more detail in the regeneration models, this will come with higher
-parameterisation efforts. This will most likely lead to lower
-generalization because the required data will have to be collected from
-specific locations, as currently there is no general regeneration
-dataset available.
-
-Therefore, we recommend that more effort should be invested into
+With our study, we have demonstrated that models of forest dynamics need
+a focus on tree regeneration modules to make them more robust. It
+remains uncertain what level of detail is required to model regeneration
+processes, and this must be addressed in future research. We recommend
+that the improvement of the regeneration modules is implemented as
+additional features that can be traced back, as done here for the
+variants of ForClim and ForCEEPS, and that model complexity and
+structure must always be connected with modeling objectives
+[@albrich2020]. If it should be necessary to include more detail in the
+regeneration models, this will come with higher parameterisation
+efforts. This will most likely lead to lower generalization because the
+required data will have to be collected from specific locations, as
+currently there is no general, comprehensive regeneration dataset
+available.
+
+Therefore, we further recommend that more effort should be invested into
 collecting harmonized datasets on tree regeneration. We emphasize that
-datasets like EuFoRIa used here are invaluable and should be further
-expanded in both their spatial extent (e.g., towards boreal and
-Mediterranean conditions) as well as in time (e.g., continuing the
-monitoring into the future). Such data will allow for a better
-evaluation of forest models and help to reduce the uncertainty in their
-projections, which is crucial when they are used as tools for predicting
-e.g. the impacts of anthropogenic climate change.
+datasets such as EuFoRIa are invaluable and should be expanded in both
+their spatial extent (e.g., towards boreal and Mediterranean conditions)
+as well as in time (e.g., continuing the monitoring into the future).
+Such data will allow for a better evaluation of forest models and help
+to reduce the uncertainty in their projections, which is crucial when
+they are used as tools for predicting e.g. the impacts of anthropogenic
+climate change.
 
@@
 be able to assess the relationship of these aspects on tree
-regeneration, e.g., due to the lack of disturbance or appropriate
+regeneration, e.g. due to the lack of disturbance or appropriate
 management modules.
 
-We recommend to further investigate the implications of the current
+We recommend to investigate in detail the implications of the current
 modeling strategies for tree regeneration and, ultimately, simulated
@@
 the inclusion of explicit spatial considerations regarding tree
-regeneration [@beckage2003]; this, however, is facing a serious
-challenge regarding appropriate datasets, as inventory data have wide
-coverage, but by definition do not allow for the assessment of spatial
-interaction effects.
+regeneration [@beckage2003]; this, however, is a serious challenge
+regarding appropriate datasets, as inventory data have wide coverage,
+but by definition do not allow for the assessment of spatial interaction
+effects.
 
@@
 assumptions on which the models are based. Furthermore, the inclusion of
-a wide range of models with different scales, type, populations
+a wide range of models with different scales, type, population
 structures, tree regeneration modules and complexity will ensure a wider
@@
 Models of forest dynamics are important tools in science and decision
-support. The formulation of tree regeneration has strong implications
-for simulated forest properties. The 15 models and variants used here
-are facing similar challenges in their representation of tree
+support, and the formulation of tree regeneration has strong
+implications for simulated forest properties. The 15 models and variants
+used here are facing similar challenges in their representation of tree
 regeneration: they generally overestimate tree regeneration levels, and
@@
 However, most models properly capture the diversity of the initial tree
-community, but differences between model formulations, e.g., the
+community, and differences between model formulations, e.g., the
 presence or absence of feedback from the adult trees, did not have a
@@
 has led to the findings presented in this paper. The protocol is
-available at @díaz-yáñez2022 . The code and data that support the
+available at @díaz-yáñez2022. The code and data that support the
 findings of this study are openly available at zenodo (*link to be

### 2023-04-28 manuscript/manuscriptRW.Rmd
added more changes from Haralds feedback

@@
 - \linenumbers
+- \usepackage{caption}
+- \captionsetup[figure]{font=scriptsize}
+- \captionsetup[table]{font=scriptsize}
 ---
@@
 of a few simple environmental filters, as done in most forest "gap"
-models [@shuman2011] and Dynamic Global Vegetation Models [e.g.,
+models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
 @hickler2012; @smith2001], to 3) complex approaches that incorporate
@@
 evaluated against a continental-scale, multi-species harmonized dataset
-on tree regeneration [@käber2022]. By tree regeneration, we refer to the
+on tree regeneration [@käber2023]. By tree regeneration, we refer to the
 passing of trees across a specific diameter threshold ("ingrowth"). We
@@
       kableExtra::column_spec(column = 10, width = "5em") |>
-      kableExtra::column_spec(column = 11, width = "8em") |>
+      kableExtra::column_spec(column = 11, width = "10em") |>
       kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
@@
 Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
-processed and aggregated following @käber2022 to provide regeneration
+processed and aggregated following @käber2023 to provide regeneration
 rates per ha and per decade. The observations used in this study
@@
 more details on this unique dataset and the detailed regeneration
-information, cf. @käber2022.
+information, cf. @käber2023.
 
@@
 regeneration between the 7 and 10 cm diameter thresholds. We used the
-Reineke self-thinning rule [@reineke1933; @pretzsch2003] as a reference
+Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference
 to estimate whether the ratio of regeneration between the 7 and 10 cm
@@
 ```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site for observed and simulated data.
-Each plot shows one pattern represented for one examplary model of each category (overperdicted, intermediate and underpredicted).
-The full data with the models grouping are shown in Figures S3 and S4.
+Each plot shows one pattern represented by one exemplary model of each category (overperdicted, intermediate and underpredicted).
+The full data with the grouping of the models are shown in Figures S3 and S4.
 n indicates the number of models falling in each group.
@@
 The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
-For comparing the magnitude of recruitment across models the same graphs with equal y-axis cf.
-Figure S7.", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
+For comparing the magnitude of recruitment across models the same graphs with equally scaled y-axis can be found in Figure S7.", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
 
@@
 
-The EuFoRIa data [@käber2022] are exceptional, particularly with respect
+The EuFoRIa data [@käber2023] are exceptional, particularly with respect
 to the number of records (number of sites and repeated measurements),

### 2023-04-28 manuscript/supplementaryRW.Rmd
added more changes from Haralds feedback

@@
 due to background seed rain; (2) a dispersal mode based on the
-formulation \'random asymmetric\'; and (3) the same establishment
+formulation 'random asymmetric'; and (3) the same establishment
 probability of 0.9 for all species included in the simulation of, which

### 2023-04-28 manuscript/manuscriptRW.Rmd
changes in abstract from Harald

--- 2023-04-26 manuscript/manuscriptRW.Rmd
@@
 ---
-title: "Tree regeneration in models of forest dynamics: a key priority for further research"
+title: "Tree establishment in models of forest dynamics: a key priority for further research"
 author:
@@
 
-Tree regeneration is a key process in forest dynamics, particularly in
+Tree establishment is a key process in forest dynamics, particularly in
 the context of forest resilience and climate change. Models are pivotal
@@
 more than 50 years. However, an assessment of their ability to
-accurately represent tree regeneration is lacking. We assess how well
+accurately represent tree establishment is lacking. We assess how well
 current models capture the overall abundance, species composition, and
-mortality of tree regeneration.
+mortality of tree establishment.
 
 Using 15 models built to capture long-term forest dynamics at the stand,
-landscape, and global levels, we simulate tree regeneration at 200 sites
-representing large environmental gradients across Central Europe. The
-results are evaluated against comprehensive data from unmanaged forests.
-
-Most of the models overestimate regeneration levels, which is only
-compensated by some models with high simulated mortality rates in the
+landscape, and global levels, we simulate tree establishment at 200
+sites representing large environmental gradients across Central Europe.
+The results are evaluated against comprehensive data from unmanaged
+forests.
+
+Most of the models overestimate ingrowth levels, which is only
+compensated in some models by high simulated mortality rates in the
 early stages of individual trees dynamics. Simulated species diversity
-of regeneration matches the observed ranges. Models simulating higher
-species diversity at the stand level do not feature higher regeneration
-diversity. The effect of light availability on regeneration levels is
+of ingrowth matches the observed ranges. Models simulating higher
+species diversity at the stand level do not feature higher ingrowth
+diversity. The effect of light availability on ingrowth levels is
 captured better than the effect of temperature and soil moisture, but
 patterns are not consistent across models. Increasing complexity in the
-tree regeneration modules of the models is not related to higher
-accuracy of simulated tree regeneration. Furthermore, individual model
+tree establishment modules of the models is not related to higher
+accuracy of simulated tree establishment. Furthermore, individual model
 design is more important than scale (stand, landscape, global) and
 approach (empirical, process-based) for accurately capturing tree
-regeneration.
-
-Despite considerable mismatches between simulation results and data in
-some instances, it is remarkable that most of the models capture the
-essential features of the highly complex process of tree regeneration,
+establishment.
+
+Despite considerable mismatches between simulation results and data, it is remarkable that most models capture the
+essential features of the highly complex process of tree establishment,
 while not having been parameterized with such data. We conclude that
-much can be gained by evaluating and refining model regeneration
+much can be gained by evaluating and refining the modeling of establishment
 processes. This has the potential to render long-term projections of
@@
 efforts have been dedicated to improving the representation of tree
-mortality [@bugmann2019; @cailleret2017]. In contrast, tree regeneration
-is a much less studied process that is often represented rather coarsely
-in these models [@walck2011; @leishman1992; @price2001], which is a
-clear research gap in the context of climate-induced forest disturbances
-and forest resilience.
-
-Tree regeneration arises from multiple processes such as pollination,
+mortality [@bugmann2019; @cailleret2017]. In contrast, tree
+establishment is a much less studied process that is often represented
+rather coarsely in these models [@walck2011; @leishman1992; @price2001],
+which is a clear research gap in the context of climate-induced forest
+disturbances and forest resilience.
+
+Tree establishment arises from multiple processes such as pollination,
 fruit maturation, seed production, dispersal, germination, juvenile
@@
 appropriate level of complexity for the mathematical formulation of the
-key factors that are leading to successful tree regeneration is
-challenging. Currently, tree regeneration processes in dynamic forest
+key factors that are leading to successful tree establishment is
+challenging. Currently, tree establishment processes in dynamic forest
 models are handled in a multitude of ways [@konig_tree_2022;
@@
 Overall, models are needed to (1) synthesize existing empirical data and
-explore their relationships, (2) assess future tree regeneration e.g.,
+explore their relationships, (2) assess future tree establishment e.g.,
 in the context of global change scenarios, and (3) identify the most
@@
 current strategies that are used in models of forest dynamics to
-represent tree regeneration, model behavior often is prone to problems,
-such as very high levels of tree regeneration that necessitate excess
+represent tree establishment, model behavior often is prone to problems,
+such as very high levels of tree establishment that necessitate excess
 mortality at early stages of tree life to simulate correct stand
 structure and composition. Also, inadequately high species diversity in
-tree regeneration may be simulated, which is characteristic of
+tree establishment may be simulated, which is characteristic of
 "classical" forest gap models [@gutiérrez2016], at least as long as the
@@
 calibration against local canopy-level data to define appropriate
-regeneration levels, which is likely to strongly hamper the general
+ingrowth levels, which is likely to strongly hamper the general
 applicability of these models e.g. under scenarios of climate change.
 Furthermore, correctly capturing the species composition in tree
-regeneration as a function of the presence of seed trees in the canopy
+establishment as a function of the presence of seed trees in the canopy
 is often a particular challenge, potentially leading to unrealistic
@@
 by considering a multitude of ecological influences, rather than
-emphasizing tree regeneration into a larger size class [e.g.,
+emphasizing tree establishment into a larger size class [e.g.,
 @wehrli2006].
 
-Thus, a focus on the modeling of tree regeneration processes is sorely
+Thus, a focus on the modeling of tree establishment processes is sorely
 needed and overdue [@price2001; @walck2011] if we are to make reliable
@@
 point of view for increasing systems understanding. Also in empirical
-ecology, tree regeneration has not received a lot of attention
+ecology, tree establishment has not received a lot of attention
 [@hanbury-brown2022], although many sub-processes have been studied in
@@
 the one hand, monitoring on permanent sample plots (such as National
-Forest Inventories) often measures tree regeneration [@zell2019],
+Forest Inventories) often measures tree establishment [@zell2019],
 however with a design that captures the rate after the stand initiation
@@
 in a modeling context [@käber2021]. On the other hand, targeted studies
-to measure tree regeneration on experimental sites are often limited in
+to measure tree establishment on experimental sites are often limited in
 spatial extent, species studied, or the sub-set of processes that are
@@
 evaluated against a continental-scale, multi-species harmonized dataset
-on tree regeneration [@käber2022]. By tree regeneration, we refer to the
-passing of trees across a specific diameter threshold ("ingrowth"). We
-included models that are based on a range of "philosophies", from models
-operating at the stand to the global scale as well as the range of
-models from empirically derived to "process-based" [e.g., @bugmann1996;
-@fabrika2012; @lexer2001; @reyer2014]. Because of the large variability
-in tree regeneration patterns in nature and the large number of factors
-driving this process --- including some that are not incorporated in
-most models such as deer browsing --- we do not aim at a detailed
-statistical evaluation of each model. Instead, we aim at evaluating the
-general regeneration patterns and magnitudes simulated by the models,
-and benchmark the simulated regeneration niche of multiple species
-against empirical data along a wide environmental gradient of
+on tree establishment [@käber2022]. By tree establishment, we refer to
+the passing of trees across a specific diameter threshold ("ingrowth").
+We included models that are based on a range of "philosophies", from
+models operating at the stand to the global scale as well as the range
+of models from empirically derived to "process-based" [e.g.,
+@bugmann1996; @fabrika2012; @lexer2001; @reyer2014]. Because of the
+large variability in tree establishment patterns in nature and the large
+number of factors driving this process --- including some that are not
+incorporated in most models such as deer browsing --- we do not aim at a
+detailed statistical evaluation of each model. Instead, we aim at
+evaluating the general ingrowth patterns and magnitudes simulated by the
+models, and benchmark the simulated regeneration niche of multiple
+species against empirical data along a wide environmental gradient of
 temperature, moisture, and light availability [@grubb1977].
@@
 More specifically, we aim to answer the following questions: (1) Are
-models of forest dynamics capturing accurately tree regeneration levels,
-initial tree species diversity and mortality in the regeneration? (2) Do
-model traits explain differences in model performance? (3) How well do
-the models capture total regeneration and individual species
-regeneration niches across environmental gradients of light
-availability, temperature, and soil moisture?
+models of forest dynamics capturing accurately tree ingrowth levels,
+initial tree species diversity and mortality in the establishment? (2)
+Do model traits explain differences in model performance? (3) How well
+do the models capture total ingrowth and individual species regeneration
+niches across environmental gradients of light availability,
+temperature, and soil moisture?
 
@@
 Fifteen models of forest dynamics (including two models featuring model
-variants) were used to assess simulated tree regeneration. They simulate
-forest dynamics at the stand, landscape or global scale (Table
+variants) were used to assess simulated tree establishment. They
+simulate forest dynamics at the stand, landscape or global scale (Table
 \@ref(tab:tableModels)). The approaches used for model construction and
@@
 FORMIND and iLand). The different approaches underlying the models have
-strongly influenced the formulation of tree regeneration processes.
-
-The regeneration modules has been categorized as 'regeneration' or
-'recruitment' models [@vanclay1994; @konig_tree_2022]. *Regeneration
-models* include processes such as flowering and pollination, seed
-production, seed dispersal, germination, and seedling growth, which
-ultimately lead to the simulated number of established trees.
-*Recruitment models*, in contrast, introduce a number of new trees with
-certain characteristics such as biomass or diameter, without explicitly
-considering earlier development processes. We can further distinguish
-models that feature a feedback in the regeneration, i.e., where the
-level and species composition of regeneration is influenced by the
-existing adult tree community via the production of seeds, seedlings or
-saplings, from models that do not contain such a feedback.
-
-The starting point for tree regeneration in the models differs as well,
+strongly influenced the formulation of tree establishment processes.
+
+The establishment modules can be categorized as 'regeneration' or
+'recruitment' models [@vanclay1994; @konig_tree_2022] (see Table
+\@ref(tab:tableModels)). Regeneration models include processes such as
+flowering and pollination, seed production, seed dispersal, germination,
+and seedling growth, which ultimately lead to the simulated number of
+established trees. Recruitment models, in contrast, introduce a number
+of new trees with certain characteristics such as biomass or diameter,
+without explicitly considering earlier development processes. We can
+further distinguish models that feature a feedback in the establishment,
+i.e., where the level and species composition of establishment is
+influenced by the existing adult tree community via the production of
+seeds, seedlings or saplings, from models that do not contain such a
+feedback.
+
+The starting point for tree establishment in the models differs as well,
 ranging from seedbank, seed or seedling to sapling (i.e., trees much
@@
 In the models used in this study, the overall complexity in the
-regeneration modules varies considerably. Following @bugmann2022, we can
-classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS,
+establishment modules varies considerably. Following @bugmann2022, we
+can classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS,
 ForClim and TreeMig) as having rather high complexity in their
-regeneration modules (mean regeneration formulation complexity across
+establishment modules (mean establishment formulation complexity across
 all processes \>0, Table \@ref(tab:tableModels)). Two models, ForClim
 and ForCEEPS, were used here with two alternative variants of
-regeneration. ForClim variant 1 [@bugmann1996] is based on a recruitment
-module that adheres closely to the concept introduced by @botkin1972,
-whereas ForClim variant 11 is adopting a different approach where
-individual species properties and their relation to the environment are
-important [@huber2020]. These two model variants allow us to evaluate a
-more process-based and complex module (variant 11) against a simple
-module (variant 1), while the rest of the model structure is identical.
-Similarly, the two ForCEEPS variants allow us to isolate the importance
-of the canopy feedback (i.e., simulated actual composition and relative
-abundance of species in the plot) via seed trees for the quantity and
-quality (e.g., diversity and composition) of simulated regeneration, as
-one variant includes this feedback whereas the other does not.
+establishment. ForClim variant 1 [@bugmann1996] is based on a
+recruitment module that adheres closely to the concept introduced by
+@botkin1972, whereas ForClim variant 11 is adopting a different approach
+where individual species properties and their relation to the
+environment are important [@huber2020]. These two model variants allow
+us to evaluate a more process-based and complex module (variant 11)
+against a simple module (variant 1), while the rest of the model
+structure is identical. Similarly, the two ForCEEPS variants allow us to
+isolate the importance of the canopy feedback (i.e., simulated actual
+composition and relative abundance of species in the plot) via seed
+trees for the quantity and quality (e.g., diversity and composition) of
+simulated ingrowth, as one variant includes this feedback whereas the
+other does not.
 
@@
 names(modelDF)[4] <- paste0(names(modelDF)[4], "(c)")
-names(modelDF)[6] <- paste0(names(modelDF)[6], "(d)")
-names(modelDF)[11] <- paste0(names(modelDF)[11],"(e)")
+names(modelDF)[5] <- paste0(names(modelDF)[5], "(d)")
+names(modelDF)[6] <- paste0(names(modelDF)[6], "(e)")
+names(modelDF)[11] <- paste0(names(modelDF)[11],"(f)")
 
@@
       kableExtra::column_spec(column = 2, width = "3em") |>
-      kableExtra::column_spec(column = 3, width = "4em") |>
-      kableExtra::column_spec(column = 4, width = "5em") |>
+      kableExtra::column_spec(column = 3, width = "3em") |>
+      kableExtra::column_spec(column = 4, width = "4em") |>
       kableExtra::column_spec(column = 5, width = "5em") |>
       kableExtra::column_spec(column = 6, width = "5em") |>
-      kableExtra::column_spec(column = 7, width = "8em") |>
+      kableExtra::column_spec(column = 7, width = "5em") |>
       kableExtra::column_spec(column = 8, width = "8em") |>
-      kableExtra::column_spec(column = 9, width = "5em") |>
+      kableExtra::column_spec(column = 9, width = "8em") |>
       kableExtra::column_spec(column = 10, width = "5em") |>
-      kableExtra::column_spec(column = 11, width = "8em") |>
+      kableExtra::column_spec(column = 11, width = "5em") |>
+      kableExtra::column_spec(column = 12, width = "8em") |>
       kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
@@
                                  "C-Cohort, I-Individual",
+                                 "RG-Regeneration, RC-Recruitment",
                                  "Y-Yes, N-No",
@@
 
-Regeneration data covering a wide range of environmental conditions are
-hard to obtain, and this is one of the reasons why most models of forest
+Ingrowth data covering a wide range of environmental conditions are hard
+to obtain, and this is one of the reasons why most models of forest
 dynamics have never been confronted with a dataset covering such
-gradients over a large number of sites, to evaluate how well
-regeneration is captured. The observations used here are derived from a
-novel and unprecedented network of sites in forest reserves that
-represent the range of environmental gradients in temperature and
-precipitation in Central Europe as compiled in the framework of the
-EuFoRIa network [@euforia2019] (Figure \@ref(fig:map)). These forests
-have been unmanaged for at least 10 years prior to their designation as
-reserves, and most of them provide time series of natural forest
-dynamics over multiple decades and up to ca. 80 years. The census
-periods range from 3 to 37 years, with an average of 14 years. The data
-provide information at the tree level, thus allowing for sequential
-comparison such as individual-tree regeneration and death between the
-measurements.
+gradients over a large number of sites, to evaluate how well ingrowth is
+captured. The observations used here are derived from a novel and
+unprecedented network of sites in forest reserves that represent the
+range of environmental gradients in temperature and precipitation in
+Central Europe as compiled in the framework of the EuFoRIa network
+[@euforia2019] (Figure \@ref(fig:map)). These forests have been
+unmanaged for at least 10 years prior to their designation as reserves,
+and most of them provide time series of natural forest dynamics over
+multiple decades and up to ca. 80 years. The census periods range from 3
+to 37 years, with an average of 14 years. The data provide information
+at the tree level, thus allowing for sequential comparison such as
+individual-tree ingrowth and death between the measurements.
 
@@
 Each of these sites featured at least two consecutive measurements.
-egeneration thresholds for the sites differed between diameters of 0 and
-10 cm. For our study, we defined two datasets where one included 165
-sites with a diameter threshold of 7 cm or lower, and the other included
+Ingrowth thresholds for the sites differed between diameters of 0 and 10
+cm. For our study, we defined two datasets where one included 165 sites
+with a diameter threshold of 7 cm or lower, and the other included
 another 35 sites with diameter thresholds between 7 and 10 cm.
@@
 Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
-processed and aggregated following @käber2022 to provide regeneration
-rates per ha and per decade. The observations used in this study
-featured 30,900 established trees. Regeneration rates per site, sample
-and decade ranged from zero to 1246 trees per ha, with a mean of 56
-trees per ha. Adult species composition was also available at the plots.
-For more details on this unique dataset and the detailed regeneration
+processed and aggregated following @käber2022 to provide ingrowth rates
+per ha and per decade. The observations used in this study featured
+30,900 established trees. Ingrowth rates per site, sample and decade
+ranged from zero to 1246 trees per ha, with a mean of 56 trees per ha.
+Adult species composition was also available at the plots. For more
+details on this unique dataset and the detailed establishment
 information, cf. @käber2022.
@@
 The overarching goal of the experiments was to assess the tree
-regeneration as it arises from empirical data against its representation
-in a wide range of models of forest dynamics. We define the tree
-regeneration as the passing of a breast height diameter threshold of 7
-or 10 cm, respectively (synonym: ingrowth). To this end, each modeling
-group was provided with a detailed protocol [@díaz-yáñez2022] with
-instructions for how to perform the simulations, providing input
-variables on climate and soil conditions, and the list of expected
-output variables. Neither were further site information (except for the
-data specified below) nor any data on tree regeneration or forest stand
-features provided prior to the simulation. That is, the models were run
-in "blind flight" mode.
+establishment as it arises from empirical data against its
+representation in a wide range of models of forest dynamics. We define
+the tree establishment as the passing of a breast height diameter
+threshold of 7 or 10 cm, respectively (synonym: ingrowth). To this end,
+each modeling group was provided with a detailed protocol
+[@díaz-yáñez2022] with instructions for how to perform the simulations,
+providing input variables on climate and soil conditions, and the list
+of expected output variables. Neither were further site information
+(except for the data specified below) nor any data on tree ingrowth or
+forest stand features provided prior to the simulation. That is, the
+models were run in "blind flight" mode.
 
@@
 chances of shade-intolerant species to establish. The simulations were
-set up to sample species-specific regeneration rates per decade and per
-ha in the equilibrium state of the model, typically entailing a
-"spin-up" run (as we did not provide any forest data). The modeling
-teams decided on the simulated area and how they derived these samples.
-The exact length of the simulation was also decided by the modeling
-teams (Table \@ref(tab:tableModels)). Further details on how each
-modeling team prepared the simulations and the outputs are available in
-Supplementary Material 1.
+set up to sample species-specific ingrowth rates per decade and per ha
+in the equilibrium state of the model, typically entailing a "spin-up"
+run (as we did not provide any forest data). The modeling teams decided
+on the simulated area and how they derived these samples. The exact
+length of the simulation was also decided by the modeling teams (Table
+\@ref(tab:tableModels)). Further details on how each modeling team
+prepared the simulations and the outputs are available in Supplementary
+Material 1.
 
@@
 equilibrium ("Potential Natural Vegetation") with the current climate,
-as the ultimate goal was to evaluate tree regeneration under comparable
+as the ultimate goal was to evaluate tree establishment under comparable
 and near equilibrium conditions. This entails the assumption that the
@@
 regeneration niche (i.e., in environmental space) as well as the
-intensity of the regeneration process (i.e., the number of ingrowth
-trees per area and per unit of time) were of interest.
+intensity of the ingrowth process (i.e., the number of ingrowth trees
+per area and per unit of time) were of interest.
 
@@
 single-species simulations) using mixtures of eleven species or genera
-for which regeneration data of sufficient quality were available from
+for which ingrowth data of sufficient quality were available from
 EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
@@
 
-Each of the models reported the regeneration number by sampling 200
-times in a 10-year interval for each species and per ha for each of the
-200 sites. Multiple samples per site were used to better understand the
+Each of the models reported the ingrowth number by sampling 200 times in
+a 10-year interval for each species and per ha for each of the 200
+sites. Multiple samples per site were used to better understand the
 simulated variation within each site. This was done using different
@@
 and two diameter thresholds). For the models that simulated additional
-species, their regeneration rates were aggregated as "others".
+species, their ingrowth rates were aggregated as "others".
 
@@
 dimensions of very old individuals. LPJ-GUESS had 2% of the sample
-outputs without tree regeneration, and these were considered as zero
-stand basal area and zero regeneration for all the tree species
-simulated; one site produced grassland rather than a forest and was not
-included in the results.
-
-```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed regeneration at the 10 cm threshold.", out.width='80%'}
+outputs without tree establishment, and these were considered as zero
+stand basal area and zero ingrowth for all the tree species simulated;
+one site produced grassland rather than a forest and was not included in
+the results.
+
+```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed ingrowth at the 10 cm threshold.", out.width='80%'}
 knitr::include_graphics(here::here("figures", "theMap.png"))
@@
 
-The simulation results were analyzed regarding (1) regeneration levels
-(i.e., ingrowth number per unit time and space), (2) regeneration
-species diversity, (3) regeneration mortality, (4) model performance and
-model traits, and (5) ingrowth gradients along regeneration niches. We
+The simulation results were analyzed regarding (1) ingrowth levels
+(i.e., ingrowth number per unit time and space), (2) species diversity
+of the ingrowth, (3) ingrowth mortality, (4) model performance and model
+traits, and (5) ingrowth gradients along regeneration niches. We
 evaluated the species diversity across models and in relation to the
@@
 proportion of the species in terms of basal area. The Shannon index was
-calculated for the regeneration ($H_{R_{n}}$) (Equation
+calculated for the establishment ($H_{R_{n}}$) (Equation
 \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation
@@
 area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
-the species $i$ in sample $n$ calculated as the regeneration basal area
+the species $i$ in sample $n$ calculated as the ingrowth basal area
 ($r_{BA_{n,i}}$) for that species $i$ relative to the total recruited
@@
 
-Mortality in tree regeneration was assessed based on the ratio of
-regeneration between the 7 and 10 cm diameter thresholds. We used the
+Mortality in tree establishment was assessed based on the ratio of
+ingrowth between the 7 and 10 cm diameter thresholds. We used the
 Reineke self-thinning rule [@reineke1933; @pretzsch2003] as a reference
-to estimate whether the ratio of regeneration between the 7 and 10 cm
+to estimate whether the ratio of ingrowth between the 7 and 10 cm
 diameter thresholds were above or below the expected theoretical rate
@@
 (empirical or process based), (3) the presence or absence of a canopy
-feedback for regeneration, and (4) the scale of application of the model
+feedback for ingrowth, and (4) the scale of application of the model
 (stand, landscape or global). We tested for significant differences
 using two-sided t-tests by considering each of these model features
-regarding the observed and simulated regeneration levels and species
+regarding the observed and simulated ingrowth levels and species
 diversity, respectively.
 
-Total regeneration and the regeneration niches of the individual species
+Total ingrowth and the regeneration niches of the individual species
 were evaluated across the environmental gradients of light availability,
 temperature, and soil moisture. Stand basal area (the basal area of all
-the trees in each sample including tree regeneration) was used as a
+the trees in each sample including tree establishment) was used as a
 proxy of light availability at the forest floor, the annual degree-day
@@
 and the climatic water balance as a proxy for soil moisture
-[@speich2019]. Regeneration values were calculated as the mean across
-the 200 samples per site. The observed data were modeled using a
-Generalized Additive Model [@wood2011] with a negative binomial
-distribution and restricted maximum likelihood to better understand the
-relationship between the environmental gradients and the levels of
-observed regeneration, relative to the simulation results per model.
+[@speich2019]. Ingrowth values were calculated as the mean across the
+200 samples per site. The observed data were modeled using a Generalized
+Additive Model [@wood2011] with a negative binomial distribution and
+restricted maximum likelihood to better understand the relationship
+between the environmental gradients and the levels of observed ingrowth,
+relative to the simulation results per model.
 
@@
 *Quercus* spp. For these species, we calculated the share in the
-regeneration basal area per site as the mean across the available
-samples per site:
+ingrowth basal area per site as the mean across the available samples
+per site:
 
@@
 where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
-is the mean basal area regeneration of the species $i$ per site $t$
-across the available samples at that site. Furthermore, regarding the
-share in the regeneration basal area per species
-($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the
-regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero
-but also when both the regeneration basal area of that species
-($\bar{r}_{BA_{t,i}}$) and the total regeneration basal area (
-$\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.
+is the mean basal area ingrowth of the species $i$ per site $t$ across
+the available samples at that site. Furthermore, regarding the share in
+the ingrowth basal area per species ($\bar{R}_{BA_{share_{i}}}$), we
+categorized this as zero when the ingrowth basal area of that species
+($\bar{r}_{BA_{t,i}}$) was zero but also when both the ingrowth basal
+area of that species ($\bar{r}_{BA_{t,i}}$) and the total ingrowth basal
+area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$ ) equaled zero.
 
@@
 
-## Regeneration levels
-
-Simulated regeneration levels varied strongly across the 15 models and
+## Ingrowth levels
+
+Simulated ingrowth levels varied strongly across the 15 models and
 typically did not match the levels found in the forest reserve data
-(Figure \@ref(fig:over)). Regeneration was overestimated in most models
-for both the 7 and 10 cm diameter thresholds, with the exception of the
+(Figure \@ref(fig:over)). Ingrowth was overestimated in most models for
+both the 7 and 10 cm diameter thresholds, with the exception of the
 empirical stand model SIBYLA, the landscape model Landis-II and the
-global model aDGVM2, which estimated regeneration levels at the lower
-end of the plausibility interval of the observed data (Figure
-\@ref(fig:over)). The models with the largest overestimation were the
-stand model PICUS and the landscape model TreeMig. For most models, the
-variability of simulated regeneration levels across the 200 sites
-(visible from the interquartile range in the box plots of Figure
-\@ref(fig:over)) was similar to or smaller than observed, with the
-notable exception of PICUS, where simulated regeneration variability was
-much larger.
-
-```{r over, echo=FALSE, fig.cap="Mean regeneration levels across all samples per site, plotted for the 200 sites and for each model.
+global model aDGVM2, which estimated ingrowth levels at the lower end of
+the plausibility interval of the observed data (Figure \@ref(fig:over)).
+The models with the largest overestimation were the stand model PICUS
+and the landscape model TreeMig. For most models, the variability of
+simulated ingrowth levels across the 200 sites (visible from the
+interquartile range in the box plots of Figure \@ref(fig:over)) was
+similar to or smaller than observed, with the notable exception of
+PICUS, where simulated ingrowth variability was much larger.
+
+```{r over, echo=FALSE, fig.cap="Mean ingrowth levels across all samples per site, plotted for the 200 sites and for each model.
 The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
@@
 
-Both the observations and the simulated data had no regeneration in some
+Both the observations and the simulated data had no ingrowth in some
 samples and at some sites (for details, cf. Table S1). The observed data
-had 4% of the samples with no regeneration. Only three models had a
-larger proportion of no regeneration (4C, Landis-II and aDVM2). Two
-models always simulated regeneration for both the 7 and 10 cm threshold
-(xComp and TreeMig), i.e., they did not feature any zero values. The
-other ten models had a very low percentage of samples with no
-regeneration (0.01% - 2.39%), i.e., they had distinctly fewer
-occurrences of zeros compared to the observations.
-
-## Tree species diversity of regeneration
+had 4% of the samples with no ingrowth. Only three models had a larger
+proportion of no ingrowth (4C, Landis-II and aDVM2). Two models always
+simulated ingrowth for both the 7 and 10 cm threshold (xComp and
+TreeMig), i.e., they did not feature any zero values. The other ten
+models had a very low percentage of samples with no ingrowth (0.01% -
+2.39%), i.e., they had distinctly fewer occurrences of zeros compared to
+the observations.
+
+## Tree species diversity of ingrowth
 
@@
 well as quantified by the Shannon index (Figure \@ref(fig:H710)). Five
-models overestimated regeneration diversity: ForCEEPS, ForClim 1, PICUS,
+models overestimated ingrowth diversity: ForCEEPS, ForClim 1, PICUS,
 TreeMig and LandClim, the latter particularly for the 7 cm diameter
@@
 the other models, nor to the observations. Only one model, Landis-II,
-consistently underestimated regeneration diversity.
-
-```{r H710, echo=FALSE, fig.cap="Shannon index of tree regeneration, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
+consistently underestimated ingrowth diversity.
+
+```{r H710, echo=FALSE, fig.cap="Shannon index of tree ingrowth, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
 The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.", out.width='100%'}
@@
 In most models, there were only small but significant differences in the
-species diversity of regeneration between 7 and 10 cm across sites
-(Tables S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS)
-maintained the regeneration diversity between the 7 and 10 cm thresholds
-(i.e., the differences between them were not significant, see Table S2),
-and the same was evident from the observed data.
-
-For both the observed and the simulated data, and both regeneration
-thresholds, species diversity in the regeneration compared to
-stand-level diversity did not have clear patterns across the models
-(Figures S1 and S2). Most models captured reasonably well or
-overpredicted species diversity at the stand level (Figure
-\@ref(fig:combinedRichness)A, left and center; cf. Figure S3), but a
-characteristic feature was that simulations had a much lower variability
-of diversity compared to observations. Models overpredicting species
-diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure
-\@ref(fig:combinedRichness)A, left). Several models underpredicted
-species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure
-\@ref(fig:combinedRichness)A, right).
-
-The majority of the models overpredicted regeneration diversity; as
-already observed in figure \@ref(fig:H710), only two of them
-underestimated regeneration diversity (4C and Landis-II) (Figure
+species diversity of ingrowth between 7 and 10 cm across sites (Tables
+S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained
+the ingrowth diversity between the 7 and 10 cm thresholds (i.e., the
+differences between them were not significant, see Table S2), and the
+same was evident from the observed data.
+
+For both the observed and the simulated data, and both ingrowth
+thresholds, species diversity in the ingrowth compared to stand-level
+diversity did not have clear patterns across the models (Figures S1 and
+S2). Most models captured reasonably well or overpredicted species
+diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left
+and center; cf. Figure S3), but a characteristic feature was that
+simulations had a much lower variability of diversity compared to
+observations. Models overpredicting species diversity included ForClim1,
+ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A,
+left). Several models underpredicted species diversity, i.e., LPJ-GUESS,
+ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).
+
+The majority of the models overpredicted ingrowth diversity; as already
+observed in figure \@ref(fig:H710), only two of them underestimated
+ingrowth diversity (4C and Landis-II) (Figure
 \@ref(fig:combinedRichness)B and Figure S4). Both the models with
-feedback and those without feedback from the adult trees to regeneration
+feedback and those without feedback from the adult trees to ingrowth
 (via seed production) overestimate, underestimated and captured
-reasonably well regeneration species diversity compared to the
-observations. Again, simulated diversity was considerably less variable
-than observed diversity in the regeneration.
+reasonably well ingrowth species diversity compared to the observations.
+Again, simulated diversity was considerably less variable than observed
+diversity in the ingrowth.
 
@@
 A) Examples for the three trends across models for species diversity at the stand level.
-B) Examples for the three trends across models for regeneration at the 7 cm threshold.", out.width='75%'}
+B) Examples for the three trends across models for ingrowth at the 7 cm threshold.", out.width='75%'}
 knitr::include_graphics(here::here("figures", "combined_richness.png"))
@@
 
-## Mortality in tree regeneration
+## Mortality in tree establishment
 
@@
 data, thus at least partially compensating for the general
-overestimation of regeneration levels (Figure \@ref(fig:over) and Table
-S3); six models did not feature a significant difference compared to the
+overestimation of ingrowth levels (Figure \@ref(fig:over) and Table S3);
+six models did not feature a significant difference compared to the
 observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
@@
 mortality rates (i.e., well above the Reineke self-thinning line), which
-compensated for the strong initial overestimation of regeneration (cf.
+compensated for the strong initial overestimation of ingrowth (cf.
 Figure \@ref(fig:over)). Interestingly, two models that underestimated
-overall regeneration levels (SIBYLA and Landis-II; cf. Figure
+overall ingrowth levels (SIBYLA and Landis-II; cf. Figure
 \@ref(fig:over)) featured mortality rates that were close to but still
@@
 
-```{r mortality11, echo=FALSE, fig.cap="Ratio of tree regeneration rates between the 7 and 10 cm thresholds.
-Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree regeneration between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
+```{r mortality11, echo=FALSE, fig.cap="Ratio of tree ingrowth rates between the 7 and 10 cm thresholds.
+Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree ingrowth between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
 knitr::include_graphics(here::here("figures", "ratio7_10.png"))
@@
 whose mortality was well above the self-thinning line, indicates that
-they feature higher regeneration mortality in mixed-species forests
-compared to even-aged, single-species forests. Some models yielded a
-mortality rate of regeneration that is broadly compatible with the
-self-thinning rule (e.g., 4C and LandClim).
+they feature higher ingrowth mortality in mixed-species forests compared
+to even-aged, single-species forests. Some models yielded a mortality
+rate of ingrowth that is broadly compatible with the self-thinning rule
+(e.g., 4C and LandClim).
 
 The three models with the largest overestimation of the proportion of
-regeneration at the 7 cm threshold featured the highest mortality
-between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The
-other models that overestimated regeneration had mortalities lower than
-the expected self-thinning ratio, with the exception of LandClim (Figure
+ingrowth at the 7 cm threshold featured the highest mortality between
+the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The other
+models that overestimated ingrowth had mortalities lower than the
+expected self-thinning ratio, with the exception of LandClim (Figure
 \@ref(fig:mortality13)B). It is noteworthy that most of the models that
-featured a low ratio of regeneration between 7 and 10 cm (i.e.,
-ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small
-overestimation of regeneration at the 7 cm threshold (Figure
-\@ref(fig:mortality11)). aDGVM2 was the only model that underestimated
-regeneration at the 7 cm threshold while featuring a low mortality
-between the two thresholds (Figure \@ref(fig:mortality13)C), and only a
-few models were close to the correct regeneration levels at 7 cm while
-also being close to the Reineke line (SIBYLA, Landis-II; cf. Figure
-\@ref(fig:mortality13)D; and 4C as well as LandClim).
-
-```{r mortality13, echo=FALSE, fig.cap="Ratio of tree regeneration between the 7 and 10 cm thresholds (regeneration 7 cm / regeneration 10 cm) and the overestimation proportion at 7 cm ([simulated -observed] / observed) for the mean regeneration per model across sites and samples.
-The horizontal dashed lines show a ratio equal to 1, indicating no decrease of regeneration between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
+featured a low ratio of ingrowth between 7 and 10 cm (i.e., ForCEEPS,
+ForClim 1, ForClim 11 and iLand) had only a small overestimation of
+ingrowth at the 7 cm threshold (Figure \@ref(fig:mortality11)). aDGVM2
+was the only model that underestimated ingrowth at the 7 cm threshold
+while featuring a low mortality between the two thresholds (Figure
+\@ref(fig:mortality13)C), and only a few models were close to the
+correct ingrowth levels at 7 cm while also being close to the Reineke
+line (SIBYLA, Landis-II; cf. Figure \@ref(fig:mortality13)D; and 4C as
+well as LandClim).
+
+```{r mortality13, echo=FALSE, fig.cap="Ratio of tree ingrowth between the 7 and 10 cm thresholds (ingrowth 7 cm / ingrowth 10 cm) and the overestimation proportion at 7 cm ([simulated -observed] / observed) for the mean ingrowth per model across sites and samples.
+The horizontal dashed lines show a ratio equal to 1, indicating no decrease of ingrowth between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
 The vertical line indicates zero overestimation.", out.width='80%'}
@@
 Lastly, some models that had a low number of individuals at the 7 cm
-threshold also had a smaller decrease of regeneration between 7 and 10
-cm (ratio 7-10 cm) (Figure S5). In these models, competition and
+threshold also had a smaller decrease of ingrowth between 7 and 10 cm
+(ratio 7-10 cm) (Figure S5). In these models, competition and
 self-thinning are either not pronounced or must have occurred before the
@@
 across models (Figure S5). It is noteworthy that there is a relationship
-between the regeneration levels at 7 cm and the mortality rate (Table
-S4): most of the models showed a positive effect (higher regeneration at
-7 cm is coupled to an increase of the mortality between 7 and 10 cm).
-While some models showed a negative effect, their linear trend was not
-pronounced and not always significant regarding the slope.
+between the ingrowth levels at 7 cm and the mortality rate (Table S4):
+most of the models showed a positive effect (higher ingrowth at 7 cm is
+coupled to an increase of the mortality between 7 and 10 cm). While some
+models showed a negative effect, their linear trend was not pronounced
+and not always significant regarding the slope.
 
@@
 There was no significant relationship between the mean complexity of the
-regeneration module (Table \@ref(tab:tableModels)) and the
+establishment module (Table \@ref(tab:tableModels)) and the
 overestimation proportion at a diameter threshold of 7 cm (Figure S6).
 All the models except 4C had significant differences between the
-observed and simulated mean regeneration values. Therefore it was not
+observed and simulated mean ingrowth values. Therefore it was not
 possible to differentiate in terms of performance between models type
 (empirical or process based), or their scales (stand, landscape or
-global) (Table S5). The values of regeneration species diversity
-simulated by each model were also assessed against the observed data to
-see if there was a difference between models that include a feedback
-compared to those that did not; all the models except ForCEEPs(f) and
-FORMIND had significant differences (Table S6). It is noteworthy that
-for the model ForCEEPS the pattern changed from the overprediction of
-species diversity in the regeneration to a diversity level that is
-closer to observations when the recruitment module included feedback
+global) (Table S5). The values of ingrowth species diversity simulated
+by each model were also assessed against the observed data to see if
+there was a difference between models that include a feedback compared
+to those that did not; all the models except ForCEEPs(f) and FORMIND had
+significant differences (Table S6). It is noteworthy that for the model
+ForCEEPS the pattern changed from the overprediction of species
+diversity in the ingrowth to a diversity level that is closer to
+observations when the recruitment module included feedback
 (ForCEEPs(f)).
 
-## Regeneration gradients on regeneration niches
-
-When evaluating regeneration levels along key gradients of light
+## Ingrowth gradients on regeneration niches
+
+When evaluating ingrowth levels along key gradients of light
 availability (basal area), temperature (degree-days) and soil moisture
@@
 
-In the observed data, regeneration levels decreased clearly with
-increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a).
-This trend was captured by the Landis-II model only, albeit at much
-lower values of basal area than in the observations. The other models
-featured distinctly different trends, such as (1) an increase in
-regeneration levels with increasing total basal area followed by the
-absence of regeneration at high values of basal area (ForCEEPS,
-ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant
-regeneration levels with basal area (ForClim 11, xComp, LandClim and
-aDGVM2), or (3) an increase of regeneration up to a certain value of
-stand basal area followed by a decrease at even higher values of stand
-basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA
-and iLand).
-
-The observed regeneration did not change much across the climatic water
+In the observed data, ingrowth levels decreased clearly with increasing
+stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a). This trend
+was captured by the Landis-II model only, albeit at much lower values of
+basal area than in the observations. The other models featured
+distinctly different trends, such as (1) an increase in ingrowth levels
+with increasing total basal area followed by the absence of ingrowth at
+high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and
+LPJ-GUESS), (2) almost constant ingrowth levels with basal area (ForClim
+11, xComp, LandClim and aDGVM2), or (3) an increase of ingrowth up to a
+certain value of stand basal area followed by a decrease at even higher
+values of stand basal area, with model-specific thresholds (4C, ForClim
+1, PICUS, SIBYLA and iLand).
+
+The observed ingrowth did not change much across the climatic water
 balance gradient, and showed a slight increase with temperature. The
-models did not match the observed regeneration patterns across the
-gradients of soil moisture and temperature (Figure
+models did not match the observed ingrowth patterns across the gradients
+of soil moisture and temperature (Figure
 \@ref(fig:nvTrend710BAlightTemp)b,c and S7), and neither was there a
 clear pattern across models. Regarding the soil moisture gradient, a
-group of models featured decreasing regeneration with increasing soil
+group of models featured decreasing ingrowth with increasing soil
 moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim,
@@
 (ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient,
-all the models that featured a decrease in regeneration levels with
-increasing soil moisture showed the opposite trend (i.e., more
-regeneration with increasing temperature), except for 4C. The other
-models did not feature a clear pattern.
-
-```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean regeneration values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum.
+all the models that featured a decrease in ingrowth levels with
+increasing soil moisture showed the opposite trend (i.e., more ingrowth
+with increasing temperature), except for 4C. The other models did not
+feature a clear pattern.
+
+```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum.
 The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
@@
 frequently in the EuFoRIa data. Most models captured well the share of
-basal area in the regeneration of these five species compared to the
+basal area in the ingrowth of these five species compared to the
 observed data along the two major climatic gradients (Figure
@@
 
-First, some models overestimated regeneration across the entire
-gradients for at least one of the five main tree species (Figure
-\@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS),
-while other models overestimated regeneration for more than one species
-for a large part of the climate gradients, such as xComp, PICUS, SIBYLA
-or LPJ-GUESS (cf. red colors in Figure \@ref(fig:diff)). These trends
-were consistent for both regeneration thresholds, although the exact
-changes in the share of the regeneration basal area were different
-(Figure S8).
+First, some models overestimated ingrowth across the entire gradients
+for at least one of the five main tree species (Figure \@ref(fig:diff),
+e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models
+overestimated ingrowth for more than one species for a large part of the
+climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf. red
+colors in Figure \@ref(fig:diff)). These trends were consistent for both
+ingrowth thresholds, although the exact changes in the share of the
+ingrowth basal area were different (Figure S8).
 
 Second, there was some congruence in the simulation results by species
-across the models. The share in the regeneration of Silver fir (*Abies
+across the models. The share in the ingrowth of Silver fir (*Abies
 alba*) was captured well across the gradient by most models except for
-xComp, PICUS, and LandClim, which overestimated the regeneration share
-of this species. There were also some peculiarities evident for some
+xComp, PICUS, and LandClim, which overestimated the ingrowth share of
+this species. There were also some peculiarities evident for some
 models, such as 4C that did not simulate *Abies alba*. The patterns
-across models for *Fagus sylvatica* regeneration were more complex, as
-its regeneration was underestimated by many models across the
-environmental gradients featuring negligible regeneration at most sites,
-while others overestimate it in different parts of the environmental
-space. Most models overestimated the share of *Picea abies* regeneration
-in the cold-wet part of the gradients. *Pinus sylvestris* and *Quercus*
-spp. represented a smaller share of the regeneration in the observed
-data mostly at warm-dry sites. Many models, however, erroneously
-featured regeneration for these two species along most of the climatic
-gradients although at a low basal area share. Five models (4C, FORMIND,
-SIBYLA, xComp, and LandClim) had almost no regeneration of *Pinus
-sylvestris* and *Quercus* spp., or did not feature any regeneration of
-these species at all, close to the patterns in the observed data.
+across models for *Fagus sylvatica* ingrowth were more complex, as its
+ingrowth was underestimated by many models across the environmental
+gradients featuring negligible ingrowth at most sites, while others
+overestimate it in different parts of the environmental space. Most
+models overestimated the share of *Picea abies* ingrowth in the cold-wet
+part of the gradients. *Pinus sylvestris* and *Quercus* spp. represented
+a smaller share of the ingrowth in the observed data mostly at warm-dry
+sites. Many models, however, erroneously featured ingrowth for these two
+species along most of the climatic gradients although at a low basal
+area share. Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had
+almost no ingrowth of *Pinus sylvestris* and *Quercus* spp., or did not
+feature any ingrowth of these species at all, close to the patterns in
+the observed data.
 
 Lastly, there is no model that performed well across the five species
-regarding the species-specific regeneration levels in the environmental
+regarding the species-specific ingrowth levels in the environmental
 space (Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
-tended to consistently overestimate the regeneration share of all five
-species, but most models overestimated the regeneration share of some
+tended to consistently overestimate the ingrowth share of all five
+species, but most models overestimated the ingrowth share of some
 species while underestimating the share of others (e.g., ForCEEPS,
@@
 
-```{r diff, echo=FALSE, fig.cap="Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold.
+```{r diff, echo=FALSE, fig.cap="Ingrowth levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm ingrowth threshold.
 The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient.
-The sizes of the circles represent the ratio between the regeneration basal area of the species and the total regeneration basal area for all species.
-The absence of a circle indicates a zero basal area share in the regeneration, or the absence of regeneration altogether.
-The color gradient (for the models only) shows the difference between the simulated and observed ratio of regeneration basal area of the species and the total regeneration basal area for all the species.", out.width='100%', out.height= '85%'}
+The sizes of the circles represent the ratio between the ingrowth basal area of the species and the total ingrowth basal area for all species.
+The absence of a circle indicates a zero basal area share in the ingrowth, or the absence of ingrowth altogether.
+The color gradient (for the models only) shows the difference between the simulated and observed ratio of ingrowth basal area of the species and the total ingrowth basal area for all the species.", out.width='100%', out.height= '85%'}
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
@@
 
-Tree regeneration is a fundamental process in forest dynamics, in
+Tree establishment is a fundamental process in forest dynamics, in
 particular under changing conditions such as those caused by climate
-change. Correctly capturing tree regeneration in dynamic models is
+change. Correctly capturing tree ingrowth in dynamic models is
 fundamental to, for example, evaluate post-disturbance dynamics and
@@
 sequestration. In this study, for the first time, projections of tree
-regeneration from multiple models of forest dynamics were confronted
+establishment from multiple models of forest dynamics were confronted
 with a unique dataset from unmanaged forest reserves across a large
@@
 sites and repeated measurements), which is essential for a highly
-'noisy' process such as tree regeneration. The use of this dataset for
+'noisy' process such as tree establishment. The use of this dataset for
 model benchmarking provided novel insights on the ability of
-state-of-the-art models to accurately simulate regeneration levels,
-species composition, and mortality in an early stage of tree life, i.e.,
-between a tree diameter of 7 and 10 cm. Overall, by adopting this
-approach a much broader understanding resulted than if we had used these
-data for model calibration: it is primarily from the shortcomings of the
-models that we can gain ecological insights [e.g., @trugman2022].
-
-## Regeneration levels
-
-Most of the models overestimated tree regeneration levels. This has
+state-of-the-art models to accurately simulate ingrowth levels, species
+composition, and mortality in an early stage of tree life, i.e., between
+a tree diameter of 7 and 10 cm. Overall, by adopting this approach a
+much broader understanding resulted than if we had used these data for
+model calibration: it is primarily from the shortcomings of the models
+that we can gain ecological insights [e.g., @trugman2022].
+
+## Ingrowth levels
+
+Most of the models overestimated tree ingrowth levels. This has
 potentially far-reaching implications, e.g., regarding biomass (and thus
@@
 forest carbon sequestration [@pugh2019; @pan2011]. Yet, we focused on
-one specific stage, i.e., regeneration into the 7 and 10 cm diameter
+one specific stage, i.e., ingrowth into the 7 and 10 cm diameter
 classes. Trees at this size contribute little to carbon sequestration,
-and if the excess regeneration at this stage is compensated -- in the
-models -- soon thereafter by higher mortality, overall simulated
-regeneration may still yield trustworthy results at the level of stand
-structure and related forest attributes such as basal area or biomass.
-
-From a modeling point of view, excess regeneration inevitably requires
+and if the excess ingrowth at this stage is compensated -- in the models
+-- soon thereafter by higher mortality, overall simulated ingrowth may
+still yield trustworthy results at the level of stand structure and
+related forest attributes such as basal area or biomass.
+
+From a modeling point of view, excess ingrowth inevitably requires
 excess mortality rates in a later stage, either -- as observed for some
@@
 stages for the expected forest densities at later stages is equivalent
-to compensating for a first error (excess regeneration) by a second
-"error" (excess mortality). It is highly likely that biased projections
-will result, because the two errors are unlikely to be perfectly linked
-and thus will not always compensate each other. Hence, this structural
+to compensating for a first error (excess ingrowth) by a second "error"
+(excess mortality). It is highly likely that biased projections will
+result, because the two errors are unlikely to be perfectly linked and
+thus will not always compensate each other. Hence, this structural
 problem of most models investigated here is problematic particularly if
@@
 
-Regeneration levels define the structure and composition of future
-forests, but it is equally important to correctly identify areas where
-regeneration is lacking [@rammer2021]. There are multiple constraints to
-the regeneration niche of tree species [@price2001], and therefore the
-absence of regeneration is likely to be common [@fortin2007], even over
-larger areas such as the one-hectare plots used here. Tree regeneration
-data are characterized by zero-inflation. This was clearly evident from
-the EuFoRIa dataset, but some of the models did not produce zeros at
-all, or featured a very low proportion of zero data. This substantial
-difference may be due to the fact that the simulation results were drawn
-from equilibrium forests, whereas in reality many of the forest reserves
-are recovering from past management activities and have become denser
-over the past decades [e.g., @heiri2009], leading to less regeneration
-than in an equilibrium situation.
-
-## Species composition of regeneration
-
-Correctly capturing species composition in tree regeneration is
+Ingrowth levels define the structure and composition of future forests,
+but it is equally important to correctly identify areas where ingrowth
+is lacking [@rammer2021]. There are multiple constraints to the
+regeneration niche of tree species [@price2001], and therefore the
+absence of ingrowth is likely to be common [@fortin2007], even over
+larger areas such as the one-hectare plots used here. Tree ingrowth data
+are characterized by zero-inflation. This was clearly evident from the
+EuFoRIa dataset, but some of the models did not produce zeros at all, or
+featured a very low proportion of zero data. This substantial difference
+may be due to the fact that the simulation results were drawn from
+equilibrium forests, whereas in reality many of the forest reserves are
+recovering from past management activities and have become denser over
+the past decades [e.g., @heiri2009], leading to less ingrowth than in an
+equilibrium situation.
+
+## Species composition of ingrowth
+
+Correctly capturing species composition in tree establishment is
 important to assess the future functional diversity of a forest, e.g.,
@@
 @seidl2022]. In the simulations, overall species diversity levels in the
-regeneration were well within the observed range for half of the models
-(7 out of 14). Thus, while most of the models are facing difficulties to
-quantitatively match regeneration levels (cf. above), their performance
-is better when it comes to matching the composition of regeneration as a
+ingrowth were well within the observed range for half of the models (7
+out of 14). Thus, while most of the models are facing difficulties to
+quantitatively match ingrowth levels (cf. above), their performance is
+better when it comes to matching the composition of ingrowth as a
 function of abiotic and biotic conditions. Most models maintained or
-even decreased regeneration diversity between the 7 and 10 cm
-thresholds, and the same is visible from the empirical data, but the
-differences were not significant. Diversity variations may be explained
-by the way the individual models consider regeneration processes
+even decreased ingrowth diversity between the 7 and 10 cm thresholds,
+and the same is visible from the empirical data, but the differences
+were not significant. Diversity variations may be explained by the way
+the individual models consider establishment processes
 [@konig_tree_2022], e.g., without considering the feedback from adults,
-regeneration can be linked to either seed rain from dispersal or a
+ingrowth can be linked to either seed rain from dispersal or a
 seed/seedling bank; neither process is incorporated across many models
@@
 The species diversity of the entire stand was captured better by the
-models than the diversity of tree regeneration, and only a few models
-overpredicted stand-level diversity. Defining regeneration composition
-as being proportional to adult abundance, regardless of productivity,
-might be a simple and conservative assumption to safely maintain
-relative species abundances [@hanbury-brown2022], although this may be
+models than the diversity of tree ingrowth, and only a few models
+overpredicted stand-level diversity. Defining ingrowth composition as
+being proportional to adult abundance, regardless of productivity, might
+be a simple and conservative assumption to safely maintain relative
+species abundances [@hanbury-brown2022], although this may be
 simplistic. Furthermore, based on the simulation results, there is no
 evidence that models with feedback from the canopy captured better the
-species diversity of the regeneration than those without feedback, with
-the exception of ForCEEPS that significantly improved regeneration
-species diversity with the model version that included feedback
-(ForCEEPS(f)). The similar performance of models with and without
-feedback is likely because the models put more weight on the
-regeneration niche arising from abiotic and biotic filters, than from
-the habitat niche of the adult trees [@grubb1977].
-
-## Regeneration mortality
-
-There are several factors that lead to mortality during the regeneration
-phase, such as competition [@casper1997] and multiple abiotic factors
-[@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
-expect that models that overestimate regeneration may have a
-particularly high mortality between the 7 and 10 cm thresholds. However,
-this was not consistently evident from the simulations. This implies
-that these models must have an excess mortality in later stages, if we
-assume that all models were able to capture the structure and
+species diversity of the ingrowth than those without feedback, with the
+exception of ForCEEPS that significantly improved ingrowth species
+diversity with the model version that included feedback (ForCEEPS(f)).
+The similar performance of models with and without feedback is likely
+because the models put more weight on the regeneration niche arising
+from abiotic and biotic filters, than from the habitat niche of the
+adult trees [@grubb1977].
+
+## Ingrowth mortality
+
+There are several factors that lead to mortality during the
+establishment phase, such as competition [@casper1997] and multiple
+abiotic factors [@schmid2021; @cunningham2006]. As mentioned above, it
+is reasonable to expect that models that overestimate ingrowth may have
+a particularly high mortality between the 7 and 10 cm thresholds.
+However, this was not consistently evident from the simulations. This
+implies that these models must have an excess mortality in later stages,
+if we assume that all models were able to capture the structure and
 composition of the adult stands along the EuFoRIa gradient; this however
@@
 and further studies are required that focus on a wider range of tree
-sizes and the fate of tree regeneration along such a size continuum.
+sizes and the fate of tree establishment along such a size continuum.
 
@@
 most models did not deviate exceedingly with respect to simulated
-regeneration levels -- after all, few if any of the models' regeneration
+ingrowth levels -- after all, few if any of the models' establishment
 routines are sufficiently constrained by data, with the exception of the
 empirical models xComp and SIBYLA. It is noteworthy that in spite of
-their empirical basis, these two models did not match empirical
-regeneration levels, in a similar magnitude as the other models and even
-in opposite directions (SIBYLA: underestimation; xComp: overestimation).
-It appears that using empirical data of limited geographical scope and
-constrained to managed forests, as done in these models, leads to
-extrapolation problems already under current climatic conditions. For
-the other models, whose regeneration modules are not strongly
-constrained by empirical data, the degrees of freedom that are available
-for modeling regeneration [@konig_tree_2022] could have implied that
-model performance would be much worse than what we found.
-
-Our study showed that increasing complexity in the regeneration modules
-did not represent an improvement in the projections of regeneration
-levels, species composition or mortality at early tree stages, as there
-was no significant relationship between model performance and model
-complexity. Increasing complexity in regeneration modules has been
-motivated by better process understanding and enhancing of model
-accuracy [@bugmann2022; @bugmann2019]. However, more complex models do
-not necessarily lead to better projections and rather to a reduced
+their empirical basis, these two models did not match empirical ingrowth
+levels, in a similar magnitude as the other models and even in opposite
+directions (SIBYLA: underestimation; xComp: overestimation). It appears
+that using empirical data of limited geographical scope and constrained
+to managed forests, as done in these models, leads to extrapolation
+problems already under current climatic conditions. For the other
+models, whose establishment modules are not strongly constrained by
+empirical data, the degrees of freedom that are available for modeling
+ingrowth [@konig_tree_2022] could have implied that model performance
+would be much worse than what we found.
+
+Our study showed that increasing complexity in the establishment modules
+did not represent an improvement in the projections of ingrowth levels,
+species composition or mortality at early tree stages, as there was no
+significant relationship between model performance and model complexity.
+Increasing complexity in establishment modules has been motivated by
+better process understanding and enhancing of model accuracy
+[@bugmann2022; @bugmann2019]. However, more complex models do not
+necessarily lead to better projections and rather to a reduced
 transparency and predictive power [@franklin2020]. Thus, the question of
 what level of detail is appropriate and parsimonious for modeling tree
-regeneration processes remains open [@konig_tree_2022; @bugmann2022].
-
-## Regeneration gradients on regeneration niches
-
-Competition for light as a strong filter for tree regeneration has been
+establishment processes remains open [@konig_tree_2022; @bugmann2022].
+
+## Ingrowth gradients on regeneration niches
+
+Competition for light as a strong filter for tree establishment has been
 widely documented [@messier1999; @collet2006; @berdanier2016], but the
@@
 the models and the observations. This made it impossible to evaluate the
-regeneration for the extremes of the stand density ranges in some
-models. For example, regeneration levels at low stand densities are
-relevant to assess how well forests are recovering e.g. after gap
-creation due to disturbance [@seidl2022; @grubb1977]. At the level of
-the simulated one-hectare samples, average basal area is typically not
-very low as long as no larger disturbances occur, which was explicitly
-excluded in our simulation protocol. Towards the other end of the
-spectrum, i.e., with increasing stand basal area, it would be reasonable
-to expect that the regeneration of the different species would become
-sparser and drop out entirely under low-light conditions [@klopcic2012;
-@zell2019]. However, few models showed this trend, thus indicating that
-the relation between regeneration and light availability is not yet
-captured correctly in most models. Yet, several of the models that did
-feature an increase of regeneration with increasing basal area include a
-feedback between seed production of mature trees and regeneration. Thus,
-it seems that higher light competition does not sufficiently compensate
-for increased seed availability with higher basal area due to more
-mature trees.
+establishment for the extremes of the stand density ranges in some
+models. For example, ingrowth levels at low stand densities are relevant
+to assess how well forests are recovering e.g. after gap creation due to
+disturbance [@seidl2022; @grubb1977]. At the level of the simulated
+one-hectare samples, average basal area is typically not very low as
+long as no larger disturbances occur, which was explicitly excluded in
+our simulation protocol. Towards the other end of the spectrum, i.e.,
+with increasing stand basal area, it would be reasonable to expect that
+the ingrowth of the different species would become sparser and drop out
+entirely under low-light conditions [@klopcic2012; @zell2019]. However,
+few models showed this trend, thus indicating that the relation between
+ingrowth and light availability is not yet captured correctly in most
+models. Yet, several of the models that did feature an increase of
+ingrowth with increasing basal area include a feedback between seed
+production of mature trees and ingrowth. Thus, it seems that higher
+light competition does not sufficiently compensate for increased seed
+availability with higher basal area due to more mature trees.
 
@@
 particular the dominant species *Abies alba*, *Fagus sylvatica* and
-*Picea abies*. The regeneration levels were sampled from simulations in
-the equilibrium, and in this state it is expected that non-climax
-species such as *Pinus sylvestris* or *Quercus* would be of minor
-importance, or absent [@klopcic2015]. Most models captured this low
-abundance, which is also found in the empirical data. Thus, although the
-broad patterns are matched by many models, improvements in the
-quantification of the regeneration niche of the species are needed, but
-this cannot be done in the absence of robust datasets across multiple
-stand development stages.
+*Picea abies*. The ingrowth levels were sampled from simulations in the
+equilibrium, and in this state it is expected that non-climax species
+such as *Pinus sylvestris* or *Quercus* would be of minor importance, or
+absent [@klopcic2015]. Most models captured this low abundance, which is
+also found in the empirical data. Thus, although the broad patterns are
+matched by many models, improvements in the quantification of the
+regeneration niche of the species are needed, but this cannot be done in
+the absence of robust datasets across multiple stand development stages.
 
@@
 The EuFoRIa data as used here are unique and well suited to better
-understand tree regeneration. However, three aspects of these data may
+understand tree establishment. However, three aspects of these data may
 also represent considerable limitations. First, we made a comparison of
-tree regeneration in an equilibrium state, but we cannot assess how
-close the forests included in the EuFoRIa dataset were to such a state.
-The data were collected in forest reserves where no management has taken
+tree ingrowth in an equilibrium state, but we cannot assess how close
+the forests included in the EuFoRIa dataset were to such a state. The
+data were collected in forest reserves where no management has taken
 place for long periods of time. This makes our assumption of an
@@
 primeval forests at demographic equilibrium, @brzeziecki2021 found
-higher regeneration rates than the ones observed in the EuFoRia data.
-Thus, the overestimation of regeneration rates by the models may not be
-so problematic. Second, the data were collected for regeneration above a
-7 cm threshold, thus limiting the assessment of tree regeneration to a
+higher ingrowth rates than the ones observed in the EuFoRia data. Thus,
+the overestimation of ingrowth rates by the models may not be so
+problematic. Second, the data were collected for ingrowth above a 7 cm
+threshold, thus limiting the assessment of tree establishment to a
 holistic viewpoint. This constituted a hard limit based on which we can
-understand only some aspects of tree regeneration, which in its entirety
-often comprises a rather long period since seed production [@price2001].
-In reality, many environmental constraints are acting on young trees
-[@käber2021] that we were unable to assess. Yet, regeneration data with
-lower thresholds are simply not available in a harmonized manner across
-large environmental gradients. Third, the empirical data were collected
-from rather small plots, while we sampled simulated regeneration levels
-from 1 ha areas, which may lead to an incorrect representation of space.
-Even though this strategy is not ideal, it represents a common challenge
-when harmonizing diverse data sources originating from varied sampling
-strategies [@portier2022]. It would have been extremely challenging for
-such a variety of models to follow a protocol where the spatial sampling
-size was different at each of the 200 sites, and it would have
-introduced additional uncertainty in the results. We must not ignore
-that the observed data represent just a snapshot of an extremely complex
-and stochastic reality. This may at least partly explain e.g., the lack
-of clear patterns in the simulation results along the gradients of
-climate water balance and degree-days.
+understand only some aspects of tree establishment, which in its
+entirety often comprises a rather long period since seed production
+[@price2001]. In reality, many environmental constraints are acting on
+young trees [@käber2021] that we were unable to assess. Yet, ingrowth
+data with lower thresholds are simply not available in a harmonized
+manner across large environmental gradients. Third, the empirical data
+were collected from rather small plots, while we sampled simulated
+ingrowth levels from 1 ha areas, which may lead to an incorrect
+representation of space. Even though this strategy is not ideal, it
+represents a common challenge when harmonizing diverse data sources
+originating from varied sampling strategies [@portier2022]. It would
+have been extremely challenging for such a variety of models to follow a
+protocol where the spatial sampling size was different at each of the
+200 sites, and it would have introduced additional uncertainty in the
+results. We must not ignore that the observed data represent just a
+snapshot of an extremely complex and stochastic reality. This may at
+least partly explain e.g., the lack of clear patterns in the simulation
+results along the gradients of climate water balance and degree-days.
 
@@
 range of models of forest dynamics, from stand to global scales, the
-simulation setup was limited to 1 ha regeneration values. This lack of
+simulation setup was limited to 1 ha ingrowth values. This lack of
 spatial considerations is appropriate for stand-scale models, but it
@@
 built to be accurate at the landscape level. Without spatial context, we
-are limiting regeneration to the seed influx from the stand itself,
-unless the model has a background seed input. However, global models
-should not be at a disadvantage due to the limited spatial
-considerations [but cf. Snell et al. -@snell2018; or Lehsten et al.
--@lehsten2019], as they are usually lacking dispersal between cells and
-are based on a strong abstraction of horizontal space
-[@hanbury-brown2022]. Lastly, detailed data on soil conditions were not
-available from the observed data, and independent, admittedly rather
-coarse data for soil and climatic water balance were used instead. It is
-noteworthy that several models represent drought based on more detailed
-indicators based e.g., on soil water holding capacity, which had to be
-derived from a rough soil quality measure. This may at least partially
-explain the poor performance of the models against the drought axis
-(i.e., climatic water balance).
+are limiting ingrowth to the seed influx from the stand itself, unless
+the model has a background seed input. However, global models should not
+be at a disadvantage due to the limited spatial considerations [but cf.
+Snell et al. -@snell2018; or Lehsten et al. -@lehsten2019], as they are
+usually lacking dispersal between cells and are based on a strong
+abstraction of horizontal space [@hanbury-brown2022]. Lastly, detailed
+data on soil conditions were not available from the observed data, and
+independent, admittedly rather coarse data for soil and climatic water
+balance were used instead. It is noteworthy that several models
+represent drought based on more detailed indicators based e.g., on soil
+water holding capacity, which had to be derived from a rough soil
+quality measure. This may at least partially explain the poor
+performance of the models against the drought axis (i.e., climatic water
+balance).
 
@@
 Based on the results of this study, it is clear that models of forest
-dynamics need a focus on tree regeneration modules to make them more
+dynamics need a focus on tree establishment modules to make them more
 robust regarding growth and mortality. It remains uncertain what level
-of detail is required to model regeneration processes, and this must be
+of detail is required to model establishment processes, and this must be
 addressed in future research. We recommend that the improvement of the
-regeneration modules is implemented as additional features that can be
+establishment modules is implemented as additional features that can be
 traced back, as done here for the variants of ForClim and ForCEEPS, and
@@
 modeling objectives [@albrich2020]. If it should be necessary to include
-more detail in the regeneration models, this will come with higher
+more detail in the establishment models, this will come with higher
 parameterisation efforts. This will most likely lead to lower
 generalization because the required data will have to be collected from
-specific locations, as currently there is no general regeneration
+specific locations, as currently there is no general establishment
 dataset available.
@@
 Therefore, we recommend that more effort should be invested into
-collecting harmonized datasets on tree regeneration. We emphasize that
+collecting harmonized datasets on tree establishment. We emphasize that
 datasets like EuFoRIa used here are invaluable and should be further
@@
 
-In the present study, we have considered tree regeneration in the
+In the present study, we have considered tree establishment in the
 equilibrium state only. It is equally important to understand how these
-models project tree regeneration after changes in forest structure by
+models project tree establishment after changes in forest structure by
 disturbances [@seidl2022], or under different management strategies
@@
 be able to assess the relationship of these aspects on tree
-regeneration, e.g., due to the lack of disturbance or appropriate
+establishment, e.g., due to the lack of disturbance or appropriate
 management modules.
@@
 We recommend to further investigate the implications of the current
-modeling strategies for tree regeneration and, ultimately, simulated
+modeling strategies for tree establishment and, ultimately, simulated
 forest stand structure. This applies particularly to the erroneous
-patterns of excess tree regeneration and later excess mortality, by
-focusing on a wider range of tree sizes and the related regeneration
+patterns of excess tree establishment and later excess mortality, by
+focusing on a wider range of tree sizes and the related ingrowth
 dynamics. We also recommend, especially for the landscape-level models,
-the inclusion of explicit spatial considerations regarding tree
-regeneration [@beckage2003]; this, however, is facing a serious
-challenge regarding appropriate datasets, as inventory data have wide
-coverage, but by definition do not allow for the assessment of spatial
-interaction effects.
+the inclusion of explicit spatial considerations regarding tree ingrowth
+[@beckage2003]; this, however, is facing a serious challenge regarding
+appropriate datasets, as inventory data have wide coverage, but by
+definition do not allow for the assessment of spatial interaction
+effects.
 
@@
 a wide range of models with different scales, type, populations
-structures, tree regeneration modules and complexity will ensure a wider
-benefit to the whole modeling community.
+structures, tree establishment modules and complexity will ensure a
+wider benefit to the whole modeling community.
 
@@
 Models of forest dynamics are important tools in science and decision
-support. The formulation of tree regeneration has strong implications
+support. The formulation of tree establishment has strong implications
 for simulated forest properties. The 15 models and variants used here
 are facing similar challenges in their representation of tree
-regeneration: they generally overestimate tree regeneration levels, and
-the simulated regeneration niche is not always captured accurately as a
-function of biotic (light) and abiotic (temperature and moisture)
+establishment: they generally overestimate tree establishment levels,
+and the simulated establishment niche is not always captured accurately
+as a function of biotic (light) and abiotic (temperature and moisture)
 factors.
@@
 presence or absence of feedback from the adult trees, did not have a
-strong effect for capturing the species composition of regeneration.
+strong effect for capturing the species composition of ingrowth.
 
 Regarding mortality in the early phase of tree life, many models that
-feature a particularly high overestimation of regeneration levels are
+feature a particularly high overestimation of ingrowth levels are
 compensating for this by a larger tree mortality. Often, this
-compensation is not sufficient to reduce the high regeneration levels to
+compensation is not sufficient to reduce the high ingrowth levels to
 realistic values. Overall, there is no clear mortality pattern across
@@
 
-When capturing tree regeneration, the specific design decisions taken in
-the development of any model are more important for its behavior
+When capturing tree establishment, the specific design decisions taken
+in the development of any model are more important for its behavior
 (accuracy) than scale (stand, landscape, global), modelling approach
@@
 complexity does not represent an improvement for capturing tree
-regeneration.
-
-Even though the regeneration routines of most of the models investigated
-here have never been sufficiently constrained by robust data, the
-regeneration projections are not overly off. This indicates that a lot
-can be gained by a focus on the modeling of regeneration processes. The
-representation of forest dynamics in these models would become much more
-robust particularly in the face of climate change and post-disturbance
-dynamics, thus strongly reducing the uncertainty in long-term
-projections of future forest dynamics.
+establishment.
+
+Even though the establishment routines of most of the models
+investigated here have never been sufficiently constrained by robust
+data, the ingrowth projections are not overly off. This indicates that a
+lot can be gained by a focus on the modeling of establishment processes.
+The representation of forest dynamics in these models would become much
+more robust particularly in the face of climate change and
+post-disturbance dynamics, thus strongly reducing the uncertainty in
+long-term projections of future forest dynamics.
 

### 2023-04-28 manuscript/manuscriptRW.Rmd
fix merge issues on the abstract

--- 2023-04-28 manuscript/manuscriptRW.Rmd
@@
 Using 15 models built to capture long-term forest dynamics at the stand,
-landscape, and global levels, we simulate tree regeneration at 200 sites
-representing large environmental gradients across Central Europe. The
-results are evaluated against comprehensive data from unmanaged forests.
+landscape, and global levels, we simulate tree regeneration at 200
+sites representing large environmental gradients across Central Europe.
+The results are evaluated against comprehensive data from unmanaged
+forests.
 
 Most of the models overestimate regeneration levels, which is only
-compensated by some models with high simulated mortality rates in the
+compensated in some models by high simulated mortality rates in the
 early stages of individual trees dynamics. Simulated species diversity
@@
 
-Despite considerable mismatches between simulation results and data in
-some instances, it is remarkable that most of the models capture the
+Despite considerable mismatches between simulation results and data, it is remarkable that most models capture the
 essential features of the highly complex process of tree regeneration,
 while not having been parameterized with such data. We conclude that
-much can be gained by evaluating and refining model regeneration
+much can be gained by evaluating and refining the modeling of regeneration
 processes. This has the potential to render long-term projections of

### 2023-04-28 manuscript/supplementaryRW.Rmd
xComp report updates from Tobi

@@
 
-Tobias Mette, LWF, 24.05.2022
+Tobias Mette, LWF.
 
@@
 modify growth and mortality in dependence of the site conditions, and
-(5) extended from single to multi-species simulations.
+(5) extended from single to multi-species simulations (c.f. Section
+1.6.7).
 
 The most complete reference resource for the xComp model and
-regeneration module is:
-
--
+regeneration module are:
+
+-   Mette, T. (2014). xcomp Tutorial (v2014_08). Technical Report, TUM.
+    DOI: 10.13140/RG.2.2.32524.59529
+
+-   Mette, T. (2014) Using Reineke to self-calibrate mortality in
+    individual tree based forest growth modelling. Beiträge zur
+    Forstwissenschaftlichen Tagung, Dresden,
+    17.-20.09.2014. 10.13140/RG.2.2.11553.07527
 
@@
 
+The species parameterisation in xcomp is sensitive to long-term climate,
+i.e. 30-year averages of summer temperature, summer precipitation and
+winter temperature. Reactions to daily, monthly or periodic extremes are
+not implemented as deterministic but as stochastically processes
+functionally dependent on the long-term climate. The climate sensitivity
+of growth, mortality and regeneration is described in Section 1.6.7.
+
 ### Soil
 
+Soil moisture class MOIST was used to modify the climate sensitivity via
+summer precipitation PJJA by PJJA(mod) = PJJA+15\*MOIST-45 (in mm).
+
 ### Topography
 
+Topography is no input parameter in xcomp. It may only effect the
+simulation results indirectly through climate or soil parameters.
+
 ### Tree species
+
+From the suggested species spectrum *Alnus glutinosa* and *Fraxinus
+excelsior* were not included in the species mix. *Quercus spec* was
+split into *Q. petraea* and *Q. pubescens*, but united in the validation
+to meet the protocol.
 
@@
 initial species mix consists of up to 10 species of equal shares with an
-initial root mean square diameter dq of 8 cm. From the suggested species
-spectrum Alnus glutinosa and Fraxinus excelsior were not included in the
-species mix. Quercus spec was split into Q. petraea and Q. pubescens,
-but united in the validation to meet the protocol. Xcomp simulates
-forest growth in 5 year intervals. Data are validated after having
-reached a climate dependent equilibrium -- which is typically the case
-after 750 years (= 150 simulation periods). Starting with year 760, data
-are read out for 200 10 year intervals (until year 2750, simulation
-period 550).
-
-#### Model details
+initial root mean square diameter dq of 8 cm. Xcomp simulates forest
+growth in 5 year intervals. Data are validated after having reached a
+climate dependent equilibrium -- which is typically the case after 750
+years (= 150 simulation periods). Starting with year 760, data are read
+out for 200 10 year intervals (until year 2750, simulation period 550).
+Further simulation relevant model details:
 
 (1)   The main growth phase is determined by the processes growth and
-    mortality. Mortality is determined by the Reineke exponent rke and
-    the stand density sdi and calibrated over a range of 8 cm to 50 cm
-    dq (= root mean square diameter) so that stem number N equals: N =
-    sdi*(dq/25)\^rke. A tree's resource acquisition area res (m2) is
-    approximated as res = 1e4/ (sdi*(dbh/25)\^rke). As the Reineke
-    exponent is fixed to 1.6, species with higher/ lower sdi have a
-    lower/ higher resource acquisition area for a given dbh. Potential
-    dbh growth dd is calculated by multiplying a reference dbh growth
-    with a site specific temperature dependent modifier. The individual
-    dbh growth in the stand is further modified by the neighbourhood
-    competition.
-
-(2) The old growth phase sets in when a tree passes a certain old growth
-    dbh (ogDbh) which depends on the site conditions. Having passed this
-    threshold dbh a tree is assigned a mortality period that assumes a
-    value not greater the a site independent maximum ogPer. The maximum
-    old growth time span ranges from 100 years (birch) to 350 years
-    (fir). During the old growth phase dbh growth declines and more
-    light becomes available for regeneration.
+mortality. Mortality is determined by the Reineke exponent rke and the
+stand density sdi and calibrated over a range of 8 cm to 50 cm dq (=
+root mean square diameter) so that stem number N equals: N =
+sdi\*(dq/25)\^rke. A tree\'s resource acquisition area res (m2) is
+approximated as res = 1e4/ (sdi\*(dbh/25)\^rke). As the Reineke exponent
+is fixed to ‑1.6, species with higher/ lower sdi have a lower/ higher
+resource acquisition area for a given dbh. Potential dbh growth dd is
+calculated by multiplying a reference dbh growth with a site specific
+temperature dependent modifier. The individual dbh growth in the stand
+is further modified by the neighbourhood competition.
+
+(2)   The old growth phase sets in when a tree passes a certain old
+growth dbh (ogDbh) which depends on the site conditions. Having passed
+this threshold dbh a tree is assigned a mortality period that assumes a
+value not greater the a site independent maximum ogPer. The maximum old
+growth time span ranges from 100 years (birch) to 350 years (fir).
+During the old growth phase dbh growth declines and more light becomes
+available for regeneration.
 
@@
 increasing distance to the tree center. Next, regeneration matrices
-    (1x1 m) are created for each of the stand's species. In the initial
-    period, the regeneration matrix of each species is empty, but at the
-    end of each period, a light dependent dbh growth and mortality is
-    calculated for each 1m raster cell. The dbh growth is scaled so that
-    under full light 5 cm dbh is reached in a defined regeneration
-    period (regnPer) that depends on the site conditions. The default
-    value is set to 15 years and equal for all species. Dbh growth is
-    modified by the light availability and the species' regeneration
-    light demand. This is a decisive difference between the species and
-    favours shadow-tolerant trees like fir and beech in the absence of
-    large gaps. However, each raster cell is afflicted with a species
-    bonus or malus that provides local variation to the global
-    regeneration parameters. Mortality in the regeneration raster occurs
-    as a random process, the likelyhood of which depends on the climate.
+(1x1 m) are created for each of the stand\'s species. In the initial
+period, the regeneration matrix of each species is empty, but at the end
+of each period, a light dependent dbh growth and mortality is calculated
+for each 1m raster cell. The dbh growth is scaled so that under full
+light 5 cm dbh is reached in a defined regeneration period (regnPer)
+that depends on the site conditions. The default value is set to 15
+years and equal for all species. Dbh growth is modified by the light
+availability and the species\' regeneration light demand. This is a
+decisive difference between the species and favours shadow-tolerant
+trees like fir and beech in the absence of large gaps. However, each
+raster cell is afflicted with a species bonus or malus that provides
+local variation to the global regeneration parameters. Mortality in the
+regeneration raster occurs as a random process, the likelyhood of which
+depends on the climate.
 
@@
 NFI-plots (Mauri et al. 2017) was related to climate of the plots
-    (Karger et al. 2017) using generalized additive models as in Thurm
-    et al. (2018). Input variables were summer temperature mean
-    (Jun-Aug), winter temperature mean (Dec-Feb) and summer
-    precipitation sum (Jun-Aug). Abundance was transformed into
-    prevalence () so that the values for each species can assume a range
-    from 0 to 1. The prevalence values modify dbh growth dd, old growth
-    dbh ogDbh, the regeneration period regnPer, and the regeneration
-    mortality regnMort.
-
-(5) The extension from single to multiple species simulation introduced
-    no technical changes to the single-species growth and mortality
-    behaviour. Competition as the main interaction between trees depends
-    on the crown size and density of the neighbouring trees and can turn
-    out an dis/advantage for one or the other species. But species
+(Karger et al. 2017) using generalized additive models as in Thurm et
+al. (2018). Input variables were summer temperature mean (Jun-Aug),
+winter temperature mean (Dec-Feb) and summer precipitation sum
+(Jun-Aug). Abundance was transformed into prevalence () so that the
+values for each species can assume a range from 0 to 1. The prevalence
+values modify dbh growth dd, old growth dbh ogDbh, the regeneration
+period regnPer, and the regeneration mortality regnMort.
+
+(5)   The extension from single to multiple species simulation
+introduced no technical changes to the single-species growth and
+mortality behaviour. Competition as the main interaction between trees
+depends on the crown size and density of the neighbouring trees and can
+turn out an dis/advantage for one or the other species. But species
 calibration parameters were not changed due to mixture-specific
@@
 
-The processes that determine the species composition and turnover in
-xcomp can be explained best by an example. In Fig. 1, I displayed the
-basal area shares resulting from the simulation of site 434 over a time
-span of 2750 years. Tab. 1 shows the species calibration parameters for
-the site. It is a very favourable site with modelled species prevalences
-of 0.93 (beech), 0.86/7 (spruce, maple), 0.81/2 (fir, ped. oak), 0.76/4
-(hornbeam, lim, birch), and 0.70 (pine). Starting with more or less
-equal basal area shares and 8 cm dq (root mean square dbh), all species
-increase their basal area in the first 50 years. Thereafter increased
-competition and starting old growth leads to a decline -- first, in all
-of the species but spruce, fir and beech, but after 100 years also in
-spruce. Only in fir and beech regeneration balances the loss of older
-trees. The result is an equilibrium of a dense fir and beech forest with
-high basal area where more light demanding species regenerate too little
-to contribute significantly.
-
-#### Discussion
-
-xcomp is a very deterministic model where except for mortality and a
-random species regeneration malus/ bonus on the regeneration raster.
-This leads to equilibria of one or two species. What species depends
-strongly on the light demand of the regeneration which favours fir and
-beech in the absence of large gaps creating disturbances. The second
-most important parameter is the site suitability which is estimated by
-the modelled species prevalence and affects growth and thereby
-competitivity. As it is the first time that xcomp is used in a multiple
-species application there is room for improvements. The automation of a
-robust calibration of the main growth phase is a precondition to
-investigate how for instance larger disturbances, a weighted
-regeneration potential oriented at the main stand, higher or selective
-mortality of the regeneration affect the species composition and
-turnover.
-
-### references
+Since the simulation output tables contain all biometric variables on an
+individual tree basis, all required outputs can be derived directly from
+the output tables. Being a very deterministic model what typically
+happens in xcomp is that the equilibria consist of one or two species.
+Shade tolerance favours fir and beech in the absence of large gaps
+creating disturbances. The second most important parameter is the site
+suitability which is estimated by the modelled species prevalence and
+affects growth and thereby competitivity. As it is the first time that
+xcomp is used in a multiple species application there is room for
+improvements. The automation of a robust calibration of the main growth
+phase is a precondition to investigate how for instance larger
+disturbances, a weighted regeneration potential oriented at the main
+stand, higher or selective mortality of the regeneration affect the
+species composition and turnover.
+
+### References
 
 1.  Bugmann, H., Seidl, R., Hartig, F., Bohn, F., Brůna, J., Cailleret,
-    M., ... & Reyer, C. P. (2019). Tree mortality submodels drive
+    M., \... & Reyer, C. P. (2019). Tree mortality submodels drive
     simulated long‐term forest dynamics: Assessing 15 models from the
@@
 2.  Karger, D. N., Conrad, O., Böhner, J., Kawohl, T., Kreft, H.,
-    Soria-Auza, R. W., ... & Kessler, M. (2017). Climatologies at high
-    resolution for the earth's land surface areas. Scientific data,
+    Soria-Auza, R. W., \... & Kessler, M. (2017). Climatologies at high
+    resolution for the earth\'s land surface areas. Scientific data,
     4(1), 1-20.
@@
     4(1), 1-8.
-4.  Mette, T. (2014) Using Reineke to self-calibrate mortality in
-    individual tree based forest growth modelling. Beiträge zur
-    Forstwissenschaftlichen Tagung, Dresden, 17.-20.09.2014
-5.  Mette, T., Falk, W., Blaschke, M., Förster, B., Walentowski, H.
+4.  Mette, T., Falk, W., Blaschke, M., Förster, B., Walentowski, H.
     (2016). Modelling natural forest dynamics in individual tree based
     forest growth simulators. Beiträge zur Forstwissenschaftlichen
-    Tagung, Freiburg, 26.-29.09.2016
-6.  Thurm, E. A., Hernandez, L., Baltensweiler, A., Ayan, S.,
-    Rasztovits, E., Bielak, K., ... & Falk, W. (2018). Alternative tree
+    Tagung, Freiburg, 26.-29.09.2016. 10.13140/RG.2.2.29169.15201
+5.  Thurm, E. A., Hernandez, L., Baltensweiler, A., Ayan, S.,
+    Rasztovits, E., Bielak, K., \... & Falk, W. (2018). Alternative tree
     species under climate warming in managed European forests. Forest

### 2023-04-28 manuscript/manuscriptRW.Rmd
change of one word in the manuscript

--- 2023-04-28 manuscript/manuscriptRW.Rmd
@@
 Most of the models overestimate regeneration levels, which is only
-compensated by some models with high simulated mortality rates in the
+compensated in some models by high simulated mortality rates in the
 early stages of individual trees dynamics. Simulated species diversity
@@
 
-Despite considerable mismatches between simulation results and data in
-some instances, it is remarkable that most of the models capture the
-essential features of the highly complex process of tree regeneration,
-while not having been parameterized with such data. We conclude that
-much can be gained by evaluating and refining model regeneration
-processes. This has the potential to render long-term projections of
-forest dynamics under changing environmental conditions that are much
-more robust.
+Despite considerable mismatches between simulation results and data, it
+is remarkable that most models capture the essential features of the
+highly complex process of tree regeneration, while not having been
+parameterized with such data. We conclude that much can be gained by
+evaluating and refining the modeling of regeneration processes. This has
+the potential to render long-term projections of forest dynamics under
+changing environmental conditions that are much more robust.
 
@@
 statistical parameterizations, which however are not easy to extrapolate
-in space and time [e.g., @ribbens1994]. Overall, models are needed to
-(1) synthesize existing empirical data and explore their relationships,
-(2) assess future tree regeneration e.g. in the context of global change
-scenarios, and (3) identify the most important processes that are
-shaping ecological patterns. Given the current strategies that are used
-in models of forest dynamics to represent tree regeneration, model
-behavior often is prone to problems, such as very high levels of tree
-regeneration that necessitate excess mortality at early stages of tree
-life to simulate correct stand structure and composition. Also,
-inadequately high species diversity in tree regeneration may be
-simulated, which is characteristic of "classical" forest gap models
-[@gutiérrez2016], at least as long as the simulation set-up comprises a
-multitude of species. Some models use calibration against local
-canopy-level data to constrain simulated regeneration levels, which is
-likely to hamper the general applicability of these models e.g. under
-scenarios of climate change. Furthermore, correctly capturing the
-species composition in tree regeneration as a function of the presence
-of seed trees in the canopy is often a particular challenge, potentially
-leading to unrealistic successional drift in the model, which must be
-corrected by factors that are hard to parameterize [e.g., @lischke2006].
-A related issue is the excessive reduction of species diversity due to
-positive feedback effects, such that eventually just single-species
-stands remain. This is sometimes corrected by the incorporation of a low
-level of seed influx of all species at all times [@schumacher2006] or by
-restricting the number of seeds per species in the seedbank
-[@lischke2006]. However, simulated species composition is usually
-exceedingly sensitive to assumptions about seed availability, whereas
-the parameters of such functions are poorly constrained by field data.
-Lastly, there is often a problem with insufficient observational
-constraints on parameter values for models that start from very small
-tree sizes (e.g., 10 cm of height or even seeds) and track tree
-development in a process-oriented manner by considering a multitude of
-ecological influences, rather than emphasizing tree regeneration into a
-larger size class [e.g., @wehrli2006].
+in space and time [e.g., @ribbens1994].
+
+Overall, models are needed to (1) synthesize existing empirical data and
+explore their relationships, (2) assess future tree regeneration e.g. in
+the context of global change scenarios, and (3) identify the most
+important processes that are shaping ecological patterns. Given the
+current strategies that are used in models of forest dynamics to
+represent tree regeneration, model behavior often is prone to problems,
+such as very high levels of tree regeneration that necessitate excess
+mortality at early stages of tree life to simulate correct stand
+structure and composition. Also, inadequately high species diversity in
+tree regeneration may be simulated, which is characteristic of
+"classical" forest gap models [@gutiérrez2016], at least as long as the
+simulation set-up comprises a multitude of species. Some models use
+calibration against local canopy-level data to constrain simulated
+regeneration levels, which is likely to hamper the general applicability
+of these models e.g. under scenarios of climate change. Furthermore,
+correctly capturing the species composition in tree regeneration as a
+function of the presence of seed trees in the canopy is often a
+particular challenge, potentially leading to unrealistic successional
+drift in the model, which must be corrected by factors that are hard to
+parameterize [e.g., @lischke2006]. A related issue is the excessive
+reduction of species diversity due to positive feedback effects, such
+that eventually just single-species stands remain. This is sometimes
+corrected by the incorporation of a low level of seed influx of all
+species at all times [@schumacher2006] or by restricting the number of
+seeds per species in the seedbank [@lischke2006]. However, simulated
+species composition is usually exceedingly sensitive to assumptions
+about seed availability, whereas the parameters of such functions are
+poorly constrained by field data. Lastly, there is often a problem with
+insufficient observational constraints on parameter values for models
+that start from very small tree sizes (e.g., 10 cm of height or even
+seeds) and track tree development in a process-oriented manner by
+considering a multitude of ecological influences, rather than
+emphasizing tree regeneration into a larger size class [e.g.,
+@wehrli2006].
 
@@
 
-```{r mortality13, echo=FALSE, fig.cap="Ratio of tree regeneration between the 7 and 10 cm thresholds (regeneration 7 cm / regeneration 10 cm) and the overestimation proportion at 7 cm ([simulated -observed] / observed) for the mean regeneration per model across sites and samples.
+```{r mortality13, echo=FALSE, fig.cap="Ratio of tree regeneration between the 7 and 10 cm thresholds (regeneration 7 cm / regeneration 10 cm) and the overestimation proportion at 7 cm ([simulated - observed] / observed) for the mean regeneration per model across sites and samples.
 The horizontal dashed lines show a ratio equal to 1, indicating no decrease of regeneration between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.

### 2023-04-28 manuscript/supplementaryRW.Rmd
small last changes in supplementary materials

@@
 -   Bugmann, H. (1994). On the ecology of mountainous forests in a
-    changing climate: A simulation study. Huber, N., Bugmann, H., &
-    Lafond, V. (2020). Capturing ecological processes in dynamic forest
-    models: Why there is no silver bullet to cope with complexity.
-    Ecosphere, 11(5). <https://doi.org/10.1002/ecs2.3109>
+    changing climate: A simulation study.
+-   Huber, N., Bugmann, H., & Lafond, V. (2020). Capturing ecological
+    processes in dynamic forest models: Why there is no silver bullet to
+    cope with complexity. Ecosphere, 11(5).
+    <https://doi.org/10.1002/ecs2.3109>
 
@@
 The most complete reference resource for the SIBYLA model and
-regeneration module is:
+regeneration module are:
 
@@
 module can be found in Woltjer et al. (2008).
+
+The most complete reference resource for the SIBYLA model and
+regeneration module are:
 
@@
 The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species.
-The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equalt to zero.", out.width='100%',out.height= "80%", fig.pos='H'}
+The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area: grey cell represent areas of the envirnomental gradient in which the model had R BA share  and the empirical dara did not becuase both recruitment and stand level recruitments were equal to zero.", out.width='100%',out.height= "80%", fig.pos='H'}
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_10.png"))
 ```
-
-## References

### 2023-05-15 manuscript/manuscriptRW.Rmd
add all authors affiliations in the date section and allignt

@@
 ---
+pretitle: "Journal: Ecological Monographs. Manuscript type: Article"
 title: "Tree regeneration in models of forest dynamics: a key priority for further research"
 author: 
-- Olalla Díaz-Yáñez,
-- Yannek Käber,
-- Tim Anders,
-- Friedrich Bohn,
-- Kristin H. Braziunas,
-- Josef	Brůna,
-- Rico Fischer,
-- Samuel M. Fischer,
-- Jessica Hetzer,
-- Thomas Hickler,
-- Christian	Hochauer,
-- Manfred J. Lexer,
-- Heike	Lischke,
-- Mats Mahnken,
-- Paola	Mairota,
-- Ján Merganič,
-- Katarina Merganičová,
-- Tobias Mette,
-- Marco Mina,
-- Xavier Morin,
-- Werner Rammer,
-- Christopher P.O. Reyer,
-- Simon	Scheiter,
-- Daniel Scherrer,
-- Harald Bugmann
-date: "`r Sys.Date()`"
+- Olalla Díaz-Yáñez$^{*1}$, 
+- Yannek Käber$^1$, 
+- Tim Anders$^2$,
+- Friedrich Bohn$^3$,
+- Kristin H. Braziunas$^4$,
+- Josef	Brůna$^5$,
+- Rico Fischer$^3$,
+- Samuel M. Fischer$^3$,
+- Jessica Hetzer$^2$,
+- Thomas Hickler$^2$,
+- Christian	Hochauer$^6$,
+- Manfred J. Lexer$^6$,
+- Heike	Lischke$^7$,
+- Mats Mahnken$^8$,
+- Paola	Mairota$^9$,
+- Ján Merganič$^{10}$,
+- Katarina Merganičová$^{11,12}$,
+- Tobias Mette$^{13}$,
+- Marco Mina$^{14}$,
+- Xavier Morin$^{15}$
+- Werner Rammer$^4$,
+- Christopher P.O. Reyer$^8$,
+- Simon	Scheiter$^2$,
+- Daniel Scherrer$^{16}$,
+- Harald Bugmann$^1$
+date: $^*$Corresponding author (olalla.diaz@usys.ethz.ch) \newline
+      $^1$ETH Zurich. Department of Environmental Systems Science. Institute of Terrestrial Ecosystems. Universitätstrasse 16, 8092 Zürich, Switzerland. \newline
+      $^2$Senckenberg Biodiversity and Climate Research Centre (SBiK-F), 60325 Frankfurt am Main, Germany. \newline
+      $^3$Department of Computational Hydrosystem, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline
+      $^4$Ecosystem Dynamics and Forest Management Group, TUM School of Life Sciences, Technical University of Munich, 85354 Freising, Germany. \newline
+       $^5$Department of Geoecology, Institute of Botany of the Czech Academy of Sciences, Průhonice, Czech Republic. \newline
+       $^6$Institute of Silviculture, Department of Forest and Soil Sciences, University of Natural Resources and Life Sciences, Peter Jordan Strasse 82, A-1190 Vienna, Austria \newline
+       $^7$Dynamic Macroecology, Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland. \newline
+       $^8$Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, 14473 Potsdam, Germany. \newline
+       $^9$Department of Soil Sciences, of Plants and Food, University of Bari Aldo Moro, Italy. \newline
+       $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen \newline
+       $^{11}$Czech University of Life Sciences Prague, Faculty of Forestry and Wood Sciences, Kamýcká 129, 16500 Praha 6 – Suchdol, Czech Republic. \newline
+       $^{12}$Department of Biodiversity of Ecosystems and Landscape, Slovak Academy of Sciences, Štefánikova 3, P.O.Box 25, 81499 Bratislava, Slovakia. \newline 
+       $^{13}$Dpt. of Soil and Climate, Bavarian State Institute of Forestry. \newline 
+       $^{14}$Institute for Alpine Environment, Eurac Research, Bolzano/Bozen, Italy. \newline 
+       $^{15}$CEFE, CNRS Campus du CNRS 1919, route de Mende 34293 Montpellier cedex, France. \newline 
+       $^{16}$Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland
+    \newline  
+    \newline
+    \today 
 output:
@@
 - \captionsetup[table]{font=scriptsize}
+- \usepackage{titling}
+- \predate{\begin{flushleft}}
+- \postdate{\end{flushleft}}
 ---
@@
 
+\newpage
+
 # Abstract
 
-Tree regeneration is a key process in forest dynamics, particularly in
-the context of forest resilience and climate change. Models are pivotal
-for assessing long-term forest dynamics, and they have been in use for
-more than 50 years. However, an assessment of their ability to
-accurately represent tree regeneration is lacking. We assess how well
-current models capture the overall abundance, species composition, and
-mortality of tree regeneration.
+Tree regeneration is a key process in forest dynamics, particularly in the context of forest resilience and climate change.
+Models are pivotal for assessing long-term forest dynamics, and they have been in use for more than 50 years.
+However, an assessment of their ability to accurately represent tree regeneration is lacking.
+We assess how well current models capture the overall abundance, species composition, and mortality of tree regeneration.
 
-Using 15 models built to capture long-term forest dynamics at the stand,
-landscape, and global levels, we simulate tree regeneration at 200 sites
-representing large environmental gradients across Central Europe. The
-results are evaluated against comprehensive data from unmanaged forests.
+Using 15 models built to capture long-term forest dynamics at the stand, landscape, and global levels, we simulate tree regeneration at 200 sites representing large environmental gradients across Central Europe.
+The results are evaluated against comprehensive data from unmanaged forests.
 
-Most of the models overestimate regeneration levels, which is only
-compensated in some models by high simulated mortality rates in the
-early stages of individual trees dynamics. Simulated species diversity
-of regeneration matches the observed ranges. Models simulating higher
-species diversity at the stand level do not feature higher regeneration
-diversity. The effect of light availability on regeneration levels is
-captured better than the effect of temperature and soil moisture, but
-patterns are not consistent across models. Increasing complexity in the
-tree regeneration modules of the models is not related to higher
-accuracy of simulated tree regeneration. Furthermore, individual model
-design is more important than scale (stand, landscape, global) and
-approach (empirical, process-based) for accurately capturing tree
-regeneration.
+Most of the models overestimate regeneration levels, which is only compensated in some models by high simulated mortality rates in the early stages of individual trees dynamics.
+Simulated species diversity of regeneration matches the observed ranges.
+Models simulating higher species diversity at the stand level do not feature higher regeneration diversity.
+The effect of light availability on regeneration levels is captured better than the effect of temperature and soil moisture, but patterns are not consistent across models.
+Increasing complexity in the tree regeneration modules of the models is not related to higher accuracy of simulated tree regeneration.
+Furthermore, individual model design is more important than scale (stand, landscape, global) and approach (empirical, process-based) for accurately capturing tree regeneration.
 
-Despite considerable mismatches between simulation results and data, it
-is remarkable that most models capture the essential features of the
-highly complex process of tree regeneration, while not having been
-parameterized with such data. We conclude that much can be gained by
-evaluating and refining the modeling of regeneration processes. This has
-the potential to render long-term projections of forest dynamics under
-changing environmental conditions that are much more robust.
+Despite considerable mismatches between simulation results and data, it is remarkable that most models capture the essential features of the highly complex process of tree regeneration, while not having been parameterized with such data.
+We conclude that much can be gained by evaluating and refining the modeling of regeneration processes.
+This has the potential to render long-term projections of forest dynamics under changing environmental conditions that are much more robust.
 
@@
 
-Forests provide a wide range of products and services of vital
-importance to humankind [@foresteurope2020]. Under the pressure of
-climate change, increasing disturbance impacts and changing societal
-demands on forest ecosystem services, it is becoming ever more important
-to understand how forest structure, composition and function will
-change, and to evaluate forest capacity to adapt to or be resilient
-after disturbance [@lindner2010; @seidl2022]. A wide range of models of
-forest dynamics were developed over the past decades considering the
-impacts of climate [@bugmann2022]. From these studies, it is evident
-that we have a reasonably good understanding how to model tree growth
-[@vanclay1997; @bugmann1996], and substantial efforts have been
-dedicated to improving the representation of tree mortality
-[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is a much
-less studied process that is often represented rather coarsely in models
-[@walck2011; @leishman1992; @price2001]. Which is a clear research gap
-in the context of climate-induced forest disturbances and forest
-resilience.
+Forests provide a wide range of products and services of vital importance to humankind [@foresteurope2020].
+Under the pressure of climate change, increasing disturbance impacts and changing societal demands on forest ecosystem services, it is becoming ever more important to understand how forest structure, composition and function will change, and to evaluate forest capacity to adapt to or be resilient after disturbance [@lindner2010; @seidl2022].
+A wide range of models of forest dynamics were developed over the past decades considering the impacts of climate [@bugmann2022].
+From these studies, it is evident that we have a reasonably good understanding how to model tree growth [@vanclay1997; @bugmann1996], and substantial efforts have been dedicated to improving the representation of tree mortality [@bugmann2019; @cailleret2017].
+In contrast, tree regeneration is a much less studied process that is often represented rather coarsely in models [@walck2011; @leishman1992; @price2001].
+Which is a clear research gap in the context of climate-induced forest disturbances and forest resilience.
 
-Tree regeneration arises from multiple processes such as pollination,
-fruit maturation, seed production, dispersal, germination, juvenile
-growth and survival [@price2001; @vacchiano2018]. All these processes
-are difficult to assess, and some of them are scarcely understood and
-thus appear highly stochastic [@bogdziewicz2021]. Identifying the
-appropriate level of complexity for the mathematical formulation of the
-key factors that are leading to successful tree regeneration is
-challenging. Currently, tree regeneration processes in dynamic forest
-models are handled in a multitude of ways [@konig_tree_2022;
-@bugmann2022]: from 1) entirely ignoring it [as done in classical forest
-growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2) the use
-of a few simple environmental filters, as done in most forest "gap"
-models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
-@hickler2012; @smith2001], to 3) complex approaches that incorporate
-local feedback from the canopy, multiple ecological processes and often
-also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
-statistical parameterizations, which however are not easy to extrapolate
-in space and time [e.g., @ribbens1994].
+Tree regeneration arises from multiple processes such as pollination, fruit maturation, seed production, dispersal, germination, juvenile growth and survival [@price2001; @vacchiano2018].
+All these processes are difficult to assess, and some of them are scarcely understood and thus appear highly stochastic [@bogdziewicz2021].
+Identifying the appropriate level of complexity for the mathematical formulation of the key factors that are leading to successful tree regeneration is challenging.
+Currently, tree regeneration processes in dynamic forest models are handled in a multitude of ways [@konig_tree_2022; @bugmann2022]: from 1) entirely ignoring it [as done in classical forest growth models, e.g., Pretzsch et al.
+-@pretzsch2002], across 2) the use of a few simple environmental filters, as done in most forest "gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g., @hickler2012; @smith2001], to 3) complex approaches that incorporate local feedback from the canopy, multiple ecological processes and often also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based statistical parameterizations, which however are not easy to extrapolate in space and time [e.g., @ribbens1994].
 
-Overall, models are needed to (1) synthesize existing empirical data and
-explore their relationships, (2) assess future tree regeneration e.g. in
-the context of global change scenarios, and (3) identify the most
-important processes that are shaping ecological patterns. Given the
-current strategies that are used in models of forest dynamics to
-represent tree regeneration, model behavior often is prone to problems,
-such as very high levels of tree regeneration that necessitate excess
-mortality at early stages of tree life to simulate correct stand
-structure and composition. Also, inadequately high species diversity in
-tree regeneration may be simulated, which is characteristic of
-"classical" forest gap models [@gutiérrez2016], at least as long as the
-simulation set-up comprises a multitude of species. Some models use
-calibration against local canopy-level data to constrain simulated
-regeneration levels, which is likely to hamper the general applicability
-of these models e.g. under scenarios of climate change. Furthermore,
-correctly capturing the species composition in tree regeneration as a
-function of the presence of seed trees in the canopy is often a
-particular challenge, potentially leading to unrealistic successional
-drift in the model, which must be corrected by factors that are hard to
-parameterize [e.g., @lischke2006]. A related issue is the excessive
-reduction of species diversity due to positive feedback effects, such
-that eventually just single-species stands remain. This is sometimes
-corrected by the incorporation of a low level of seed influx of all
-species at all times [@schumacher2006] or by restricting the number of
-seeds per species in the seedbank [@lischke2006]. However, simulated
-species composition is usually exceedingly sensitive to assumptions
-about seed availability, whereas the parameters of such functions are
-poorly constrained by field data. Lastly, there is often a problem with
-insufficient observational constraints on parameter values for models
-that start from very small tree sizes (e.g., 10 cm of height or even
-seeds) and track tree development in a process-oriented manner by
-considering a multitude of ecological influences, rather than
-emphasizing tree regeneration into a larger size class [e.g.,
-@wehrli2006].
+Overall, models are needed to (1) synthesize existing empirical data and explore their relationships, (2) assess future tree regeneration e.g.
+in the context of global change scenarios, and (3) identify the most important processes that are shaping ecological patterns.
+Given the current strategies that are used in models of forest dynamics to represent tree regeneration, model behavior often is prone to problems, such as very high levels of tree regeneration that necessitate excess mortality at early stages of tree life to simulate correct stand structure and composition.
+Also, inadequately high species diversity in tree regeneration may be simulated, which is characteristic of "classical" forest gap models [@gutiérrez2016], at least as long as the simulation set-up comprises a multitude of species.
+Some models use calibration against local canopy-level data to constrain simulated regeneration levels, which is likely to hamper the general applicability of these models e.g.
+under scenarios of climate change.
+Furthermore, correctly capturing the species composition in tree regeneration as a function of the presence of seed trees in the canopy is often a particular challenge, potentially leading to unrealistic successional drift in the model, which must be corrected by factors that are hard to parameterize [e.g., @lischke2006].
+A related issue is the excessive reduction of species diversity due to positive feedback effects, such that eventually just single-species stands remain.
+This is sometimes corrected by the incorporation of a low level of seed influx of all species at all times [@schumacher2006] or by restricting the number of seeds per species in the seedbank [@lischke2006].
+However, simulated species composition is usually exceedingly sensitive to assumptions about seed availability, whereas the parameters of such functions are poorly constrained by field data.
+Lastly, there is often a problem with insufficient observational constraints on parameter values for models that start from very small tree sizes (e.g., 10 cm of height or even seeds) and track tree development in a process-oriented manner by considering a multitude of ecological influences, rather than emphasizing tree regeneration into a larger size class [e.g., @wehrli2006].
 
-Thus, a focus on the modeling of tree regeneration processes is sorely
-needed and overdue [@price2001; @walck2011] if we are to make reliable
-projections of future forest dynamics, i.e., when the models need to be
-operated in extrapolation mode, as well as from a fundamental ecological
-point of view for increasing systems understanding. Also in empirical
-ecology, tree regeneration has not received a lot of attention
-[@hanbury-brown2022], although many sub-processes have been studied in
-detail [@miina2006]. Yet, little data are available that covers all the
-processes within one species along environmental gradients, let alone
-for a vast suite of species. Thus, the available data are fragmented,
-which constitutes a major problem for model building [@clark1999]. On
-the one hand, monitoring on permanent plots (such as National Forest
-Inventories) often measures tree regeneration [@zell2019], however with
-a design that captures the rate after the stand initiation phase
-[@hallsby2015]. Data from permanent plots, although available across
-regions, are highly heterogeneous, rendering their use difficult in a
-modeling context [@käber2021]. On the other hand, targeted studies to
-measure tree regeneration on experimental sites are often limited in
-spatial extent, species studied, or the sub-set of processes that are
-targeted [@berdanier2016; @collet2006].
+Thus, a focus on the modeling of tree regeneration processes is sorely needed and overdue [@price2001; @walck2011] if we are to make reliable projections of future forest dynamics, i.e., when the models need to be operated in extrapolation mode, as well as from a fundamental ecological point of view for increasing systems understanding.
+Also in empirical ecology, tree regeneration has not received a lot of attention [@hanbury-brown2022], although many sub-processes have been studied in detail [@miina2006].
+Yet, little data are available that covers all the processes within one species along environmental gradients, let alone for a vast suite of species.
+Thus, the available data are fragmented, which constitutes a major problem for model building [@clark1999].
+On the one hand, monitoring on permanent plots (such as National Forest Inventories) often measures tree regeneration [@zell2019], however with a design that captures the rate after the stand initiation phase [@hallsby2015].
+Data from permanent plots, although available across regions, are highly heterogeneous, rendering their use difficult in a modeling context [@käber2021].
+On the other hand, targeted studies to measure tree regeneration on experimental sites are often limited in spatial extent, species studied, or the sub-set of processes that are targeted [@berdanier2016; @collet2006].
 
-In the present study, for the first time, a large number of forest
-models commonly used to assess forest dynamics under climate change are
-evaluated against a continental-scale, multi-species harmonized dataset
-on tree regeneration [@käber2023]. By tree regeneration, we refer to the
-passing of trees across a specific diameter threshold ("ingrowth"). We
-included models that are based on a range of "philosophies", from models
-operating at the stand to the global scale as well as the range of
-models from empirically derived to "process-based" [e.g., @bugmann1996;
-@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
-tree regeneration patterns in nature and the large number of factors
-driving this process --- including some that are not incorporated
-explicitly in most models, such as deer browsing --- we do not aim for a
-detailed statistical evaluation of each model. Instead, we aim to
-evaluate the general regeneration patterns and magnitudes simulated by
-the models, and benchmark the simulated regeneration niche of multiple
-species against empirical data along a wide environmental gradient of
-temperature, moisture, and light availability [@grubb1977].
+In the present study, for the first time, a large number of forest models commonly used to assess forest dynamics under climate change are evaluated against a continental-scale, multi-species harmonized dataset on tree regeneration [@käber2023].
+By tree regeneration, we refer to the passing of trees across a specific diameter threshold ("ingrowth").
+We included models that are based on a range of "philosophies", from models operating at the stand to the global scale as well as the range of models from empirically derived to "process-based" [e.g., @bugmann1996; @fabrika2012; @lexer2001; @reyer2014].
+Due to the large variability in tree regeneration patterns in nature and the large number of factors driving this process --- including some that are not incorporated explicitly in most models, such as deer browsing --- we do not aim for a detailed statistical evaluation of each model.
+Instead, we aim to evaluate the general regeneration patterns and magnitudes simulated by the models, and benchmark the simulated regeneration niche of multiple species against empirical data along a wide environmental gradient of temperature, moisture, and light availability [@grubb1977].
 
-More specifically, we aim to answer the following questions: (1) Are
-models of forest dynamics capturing accurately tree regeneration levels,
-initial tree species diversity and mortality in the regeneration? (2) Do
-model traits explain differences in model performance? (3) How well do
-the models capture total regeneration and individual species
-regeneration niches across environmental gradients of light
-availability, temperature, and soil moisture?
+More specifically, we aim to answer the following questions: (1) Are models of forest dynamics capturing accurately tree regeneration levels, initial tree species diversity and mortality in the regeneration?
+(2) Do model traits explain differences in model performance?
+(3) How well do the models capture total regeneration and individual species regeneration niches across environmental gradients of light availability, temperature, and soil moisture?
 
@@
 
-Fifteen models of forest dynamics (including two models featuring model
-variants) were used to simulate forest dynamics at the stand, landscape
-or global scale (Table \@ref(tab:tableModels)). The approaches used for
-model construction and their origin differ strongly, with most of the
-models featuring a largely "process-based" approach, whereas two models
-are based on formulations derived from the statistical analysis of
-inventory data (SIBYLA and xComp). Some of the models largely rely on
-the original approach underlying forest gap models (e.g., ForClim 1),
-whereas some are based on plant ecophysiological processes (e.g.,
-FORMIND and iLand). The different approaches underlying the models have
-strongly influenced the formulation of tree regeneration processes.
+Fifteen models of forest dynamics (including two models featuring model variants) were used to simulate forest dynamics at the stand, landscape or global scale (Table \@ref(tab:tableModels)).
+The approaches used for model construction and their origin differ strongly, with most of the models featuring a largely "process-based" approach, whereas two models are based on formulations derived from the statistical analysis of inventory data (SIBYLA and xComp).
+Some of the models largely rely on the original approach underlying forest gap models (e.g., ForClim 1), whereas some are based on plant ecophysiological processes (e.g., FORMIND and iLand).
+The different approaches underlying the models have strongly influenced the formulation of tree regeneration processes.
 
-The regeneration models can be differentiated into 'regeneration' and
-'recruitment' approaches [@vanclay1994; @konig_tree_2022]. *Regeneration
-models* include processes such as flowering and pollination, seed
-production, seed dispersal, germination, and seedling growth, which
-ultimately lead to the simulated number of established trees.
-*Recruitment models*, in contrast, introduce a number of new trees with
-certain characteristics such as biomass or diameter, without explicitly
-considering earlier development processes. We can further distinguish
-models that feature a feedback in the regeneration, i.e., where the
-level and species composition of regeneration is influenced by the
-existing adult tree community via the production of seeds, seedlings or
-saplings, from models that do not contain such a feedback.
+The regeneration models can be differentiated into 'regeneration' and 'recruitment' approaches [@vanclay1994; @konig_tree_2022].
+*Regeneration models* include processes such as flowering and pollination, seed production, seed dispersal, germination, and seedling growth, which ultimately lead to the simulated number of established trees.
+*Recruitment models*, in contrast, introduce a number of new trees with certain characteristics such as biomass or diameter, without explicitly considering earlier development processes.
+We can further distinguish models that feature a feedback in the regeneration, i.e., where the level and species composition of regeneration is influenced by the existing adult tree community via the production of seeds, seedlings or saplings, from models that do not contain such a feedback.
 
-The starting point for tree regeneration in the models differs as well,
-ranging from seedbank, seed or seedling to sapling (i.e., trees much
-larger than 10 cm height, often ca. 2-3 m tall). Models that start from
-seed need to include a larger number of ecological processes such as
-germination and survival of young seedlings, whereas models that start
-from saplings have to aggregate via parameterizations several ecological
-processes that are not treated explicitly. This latter approach reduces
-model complexity but comes at the cost of blurring process
-representation.
+The starting point for tree regeneration in the models differs as well, ranging from seedbank, seed or seedling to sapling (i.e., trees much larger than 10 cm height, often ca.
+2-3 m tall).
+Models that start from seed need to include a larger number of ecological processes such as germination and survival of young seedlings, whereas models that start from saplings have to aggregate via parameterizations several ecological processes that are not treated explicitly.
+This latter approach reduces model complexity but comes at the cost of blurring process representation.
 
-In the models used here, the overall complexity in the regeneration
-modules varies considerably. Following @bugmann2022, we can classify
-seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and
-TreeMig) as having rather high complexity in their regeneration modules
-(mean regeneration formulation complexity across all processes \>0,
-Table \@ref(tab:tableModels)). Two models, ForClim and ForCEEPS, were
-used here with two alternative variants of regeneration. ForClim variant
-1 [@bugmann1996] is based on a recruitment module that adheres closely
-to the concept introduced by @botkin1972, whereas ForClim variant 11 is
-adopting a different approach where individual species properties and
-their relation to the environment are important [@huber2020]. These two
-model variants allow us to evaluate a more process-based and complex
-module (variant 11) against a simple module (variant 1), while the rest
-of the model structure is identical. Similarly, the two ForCEEPS
-variants allow us to isolate the importance of the canopy feedback
-(i.e., simulated actual composition and relative abundance of species in
-the plot) via seed trees for the quantity and quality (e.g., diversity
-and composition) of simulated regeneration, as one variant includes this
-feedback whereas the other does not.
+In the models used here, the overall complexity in the regeneration modules varies considerably.
+Following @bugmann2022, we can classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and TreeMig) as having rather high complexity in their regeneration modules (mean regeneration formulation complexity across all processes \>0, Table \@ref(tab:tableModels)).
+Two models, ForClim and ForCEEPS, were used here with two alternative variants of regeneration.
+ForClim variant 1 [@bugmann1996] is based on a recruitment module that adheres closely to the concept introduced by @botkin1972, whereas ForClim variant 11 is adopting a different approach where individual species properties and their relation to the environment are important [@huber2020].
+These two model variants allow us to evaluate a more process-based and complex module (variant 11) against a simple module (variant 1), while the rest of the model structure is identical.
+Similarly, the two ForCEEPS variants allow us to isolate the importance of the canopy feedback (i.e., simulated actual composition and relative abundance of species in the plot) via seed trees for the quantity and quality (e.g., diversity and composition) of simulated regeneration, as one variant includes this feedback whereas the other does not.
 
@@
 
-Regeneration data covering a wide range of environmental conditions are
-hard to obtain, and this is one of the reasons why most models of forest
-dynamics have never been confronted with a dataset covering such
-gradients over a large number of sites, to evaluate how well
-regeneration is captured. The observations used here are derived from a
-novel and unprecedented network of sites in forest reserves that
-represent the range of environmental gradients in temperature and
-precipitation in Central Europe as compiled in the framework of the
-EuFoRIa network [@euforia2019] (Figure \@ref(fig:map)). These forests
-have been unmanaged for at least 10 years prior to their designation as
-reserves, and most of them provide time series of natural forest
-dynamics over multiple decades and up to ca. 80 years. The census
-periods range from 3 to 37 years, with an average of 14 years. The data
-provide information at the tree level, thus allowing for the sequential
-comparison of processes such as individual-tree regeneration and death
-between the measurements.
+Regeneration data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset covering such gradients over a large number of sites, to evaluate how well regeneration is captured.
+The observations used here are derived from a novel and unprecedented network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network [@euforia2019] (Figure \@ref(fig:map)).
+These forests have been unmanaged for at least 10 years prior to their designation as reserves, and most of them provide time series of natural forest dynamics over multiple decades and up to ca.
+80 years.
+The census periods range from 3 to 37 years, with an average of 14 years.
+The data provide information at the tree level, thus allowing for the sequential comparison of processes such as individual-tree regeneration and death between the measurements.
 
-We selected 200 sites from this network as the benchmarking dataset for
-the simulation to be representative of the environmental variation
-contained in the data. This was achieved by applying k-means clustering
-to define 200 clusters of plots from the original set of 869 plots along
-the environmental dimensions of temperature, climatic water balance,
-soil quality and exposition (i.e., slope and aspect). Each of these
-sites featured at least two consecutive measurements. Regeneration
-thresholds for these sites differed between diameters of 0 and 10 cm.
-For our study, we defined two datasets where one included 165 sites with
-a diameter threshold of 7 cm or lower, and the other included another 35
-sites with diameter thresholds between 7 and 10 cm.
+We selected 200 sites from this network as the benchmarking dataset for the simulation to be representative of the environmental variation contained in the data.
+This was achieved by applying k-means clustering to define 200 clusters of plots from the original set of 869 plots along the environmental dimensions of temperature, climatic water balance, soil quality and exposition (i.e., slope and aspect).
+Each of these sites featured at least two consecutive measurements.
+Regeneration thresholds for these sites differed between diameters of 0 and 10 cm.
+For our study, we defined two datasets where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds between 7 and 10 cm.
 
-Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
-processed and aggregated following @käber2023 to provide regeneration
-rates per ha and per decade. The observations used in this study
-featured 30,900 newly established trees. Regeneration rates per site,
-sample and decade ranged from zero to 1246 trees, with a mean of 56
-trees. Adult species composition was also available for each plot. For
-more details on this unique dataset and the detailed regeneration
-information, cf. @käber2023.
+Plot size ranged from 0.02 ha to 5.52 ha, and the data were further processed and aggregated following @käber2023 to provide regeneration rates per ha and per decade.
+The observations used in this study featured 30,900 newly established trees.
+Regeneration rates per site, sample and decade ranged from zero to 1246 trees, with a mean of 56 trees.
+Adult species composition was also available for each plot.
+For more details on this unique dataset and the detailed regeneration information, cf.
+@käber2023.
 
@@
 
-The overarching goal of the experiments was to assess the tree
-regeneration as it arises from empirical data against its representation
-in a wide range of models of forest dynamics. We define tree
-regeneration as the passing of a breast height diameter threshold of 7
-or 10 cm, respectively (synonym: ingrowth). To this end, each modeling
-group was provided with a detailed protocol [@díaz-yáñez2022] with
-instructions how to perform the simulations, providing input variables
-on climate and soil conditions, and the list of expected output
-variables. Neither were further site information (except for the data
-specified below) nor any data on tree regeneration or forest stand
-features provided prior to the simulation. That is, the models were run
-in "blind flight" mode.
+The overarching goal of the experiments was to assess the tree regeneration as it arises from empirical data against its representation in a wide range of models of forest dynamics.
+We define tree regeneration as the passing of a breast height diameter threshold of 7 or 10 cm, respectively (synonym: ingrowth).
+To this end, each modeling group was provided with a detailed protocol [@díaz-yáñez2022] with instructions how to perform the simulations, providing input variables on climate and soil conditions, and the list of expected output variables.
+Neither were further site information (except for the data specified below) nor any data on tree regeneration or forest stand features provided prior to the simulation.
+That is, the models were run in "blind flight" mode.
 
-The input variables were collected from different data sources and
-aggregated to be adapted to the needs of the different models. Time
-series of climatic variables from 1981 to 2018 were provided in hourly
-[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
-resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
-required for some models, such as relative humidity or vapor pressure
-deficit, were calculated from these variables. The final instructions
-for using the climate data were slightly different in each model, based
-on the approach that best suited the model (Table
-\@ref(tab:tableModels)). Soil quality data were provided as continuous
-values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
-The protocol also provided the elevation, slope and aspect for each of
-the 200 sites [@asterscienceteam2019], but no other spatial information
-such as coordinates, with the exception of iLand and aDGVM2, which
-required blurred coordinates to derive highly detailed soil data.
+The input variables were collected from different data sources and aggregated to be adapted to the needs of the different models.
+Time series of climatic variables from 1981 to 2018 were provided in hourly [Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly resolution [CHELSA data, Karger et al.
+-@karger2021].
+Some variables required for some models, such as relative humidity or vapor pressure deficit, were calculated from these variables.
+The final instructions for using the climate data were slightly different in each model, based on the approach that best suited the model (Table \@ref(tab:tableModels)).
+Soil quality data were provided as continuous values between 1 and 5 [Soilgrids dataset, Hengl et al.
+-@hengl2017].
+The protocol also provided the elevation, slope and aspect for each of the 200 sites [@asterscienceteam2019], but no other spatial information such as coordinates, with the exception of iLand and aDGVM2, which required blurred coordinates to derive highly detailed soil data.
 
-The simulations were run in the absence of natural disturbances. Only
-the model LPJ-GUESS had to include a background disturbance to increase
-chances of shade-intolerant species to establish. The simulations were
-set up to sample species-specific regeneration rates per decade and per
-ha in the equilibrium state of the model, typically entailing a
-"spin-up" run (as we did not provide any forest data). The modeling
-teams decided on the simulated area and how they derived these samples.
-The exact length of the simulation was also decided by the modeling
-teams (Table \@ref(tab:tableModels)). Further details on how each
-modeling team prepared the simulations and the outputs are available in
-Supplementary Material 1.
+The simulations were run in the absence of natural disturbances.
+Only the model LPJ-GUESS had to include a background disturbance to increase chances of shade-intolerant species to establish.
+The simulations were set up to sample species-specific regeneration rates per decade and per ha in the equilibrium state of the model, typically entailing a "spin-up" run (as we did not provide any forest data).
+The modeling teams decided on the simulated area and how they derived these samples.
+The exact length of the simulation was also decided by the modeling teams (Table \@ref(tab:tableModels)).
+Further details on how each modeling team prepared the simulations and the outputs are available in Supplementary Material 1.
 
-The simulations were run in the absence of management to a simulated
-equilibrium ("Potential Natural Vegetation") with the current climate,
-as the ultimate goal was to evaluate tree regeneration under comparable
-and near-equilibrium conditions. This entails the assumption that (1)
-the observations from the forest reserves reflect no traces of forest
-management, and (2) there is an equilibrium between forest dynamics and
-climate. While the former might be starting to be visible in many of the
-EuFoRIA reserves, the latter may be more debatable. However, in the
-absence of detailed data on the history of each plot in the EuFoRIa
-network, some broad assumptions had to be made. Both the width of the
-regeneration niche (i.e., in environmental space) as well as the
-intensity of the regeneration process (i.e., the number of ingrowth
-trees per area and per unit of time) were of interest.
+The simulations were run in the absence of management to a simulated equilibrium ("Potential Natural Vegetation") with the current climate, as the ultimate goal was to evaluate tree regeneration under comparable and near-equilibrium conditions.
+This entails the assumption that (1) the observations from the forest reserves reflect no traces of forest management, and (2) there is an equilibrium between forest dynamics and climate.
+While the former might be starting to be visible in many of the EuFoRIA reserves, the latter may be more debatable.
+However, in the absence of detailed data on the history of each plot in the EuFoRIa network, some broad assumptions had to be made.
+Both the width of the regeneration niche (i.e., in environmental space) as well as the intensity of the regeneration process (i.e., the number of ingrowth trees per area and per unit of time) were of interest.
 
-The simulations were run for mixed-species forests (not multiple
-single-species simulations) using mixtures of eleven species or genera
-for which regeneration data of sufficient quality were available from
-EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
-betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
-Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus
-sylvestris L.* The same set of species was used at all 200 sites. Two
-models included their standard set of species for the simulations, which
-is much larger (i.e., ForClim 1, ForClim 11, TreeMig). In three models,
-fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS)
-(Table \@ref(tab:tableModels)). In 4C only *Fagus sylvatica*, *Picea
-abies*, *Betula* spp., *Quercus* spp. and *Pinus sylvestris* are
-parameterised. xComp simulations did not consider *Fraxinus excelsior*
-due to a depreciated species parameterisation. In LPJ-GUESS, *Acer
-pseudoplatanus* and *Alnus glutinosa* are not parameterized and
-therefore these species could not be included in the simulation.
-Finally, the model aDGVM2 does not simulate individual species. Rather,
-community assembly processes and trait filtering generate plant
-communities that are adjusted to the biotic and abiotic conditions, and
-the simulated plants can be classified into ecological strategies based
-on their trait values in a post-processing step.
+The simulations were run for mixed-species forests (not multiple single-species simulations) using mixtures of eleven species or genera for which regeneration data of sufficient quality were available from EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp.
+L., Fraxinus excelsior L., Quercus spp.
+L., Alnus glutinosa L., Pinus sylvestris L.* The same set of species was used at all 200 sites.
+Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig).
+In three models, fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)).
+In 4C only *Fagus sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp.
+and *Pinus sylvestris* are parameterised.
+xComp simulations did not consider *Fraxinus excelsior* due to a depreciated species parameterisation.
+In LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not parameterized and therefore these species could not be included in the simulation.
+Finally, the model aDGVM2 does not simulate individual species.
+Rather, community assembly processes and trait filtering generate plant communities that are adjusted to the biotic and abiotic conditions, and the simulated plants can be classified into ecological strategies based on their trait values in a post-processing step.
 
-Each of the models reported the regeneration number by sampling 200
-times in a 10-year interval for each species and per ha for each of the
-200 sites. Multiple samples per site were used to better understand the
-simulated variation within each site. This was done using different
-strategies, depending on the model, including (1) sampling simulated
-data from the same 1 ha plot in the equilibrium over time, (2) sampling
-several 1 ha plots from the simulated forest at one specific point in
-time (in the equilibrium) or (3) a combination of (1) and (2). This
-resulted in 880,000 observations per model that simulated the 11 species
-included in the protocol (200 sites, 200 samples per site, 11 species
-and two diameter thresholds). For the models that simulated additional
-species, their regeneration rates were aggregated as "others".
+Each of the models reported the regeneration number by sampling 200 times in a 10-year interval for each species and per ha for each of the 200 sites.
+Multiple samples per site were used to better understand the simulated variation within each site.
+This was done using different strategies, depending on the model, including (1) sampling simulated data from the same 1 ha plot in the equilibrium over time, (2) sampling several 1 ha plots from the simulated forest at one specific point in time (in the equilibrium) or (3) a combination of (1) and (2).
+This resulted in 880,000 observations per model that simulated the 11 species included in the protocol (200 sites, 200 samples per site, 11 species and two diameter thresholds).
+For the models that simulated additional species, their regeneration rates were aggregated as "others".
 
-Two models did not provide results from all the simulated samples or
-sites to avoid unrealistic results, as follows. In the model 4C, a
-threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid
-unrealistic stand basal area data, and therefore not all the samples and
-sites were reported. The reason for this is that 4C is not suitable for
-long-term simulations without management, due to misrepresentations in
-density-dependent mortality processes in long-term simulations and
-assumptions of tree geometry that lead to unrealistic single tree
-dimensions of very old individuals. LPJ-GUESS had 2% of the sample
-outputs without tree regeneration, and these were considered as zero
-stand basal area and zero regeneration for all the tree species
-simulated; one site produced grassland rather than a forest and was not
-included in the results.
+Two models did not provide results from all the simulated samples or sites to avoid unrealistic results, as follows.
+In the model 4C, a threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid unrealistic stand basal area data, and therefore not all the samples and sites were reported.
+The reason for this is that 4C is not suitable for long-term simulations without management, due to misrepresentations in density-dependent mortality processes in long-term simulations and assumptions of tree geometry that lead to unrealistic single tree dimensions of very old individuals.
+LPJ-GUESS had 2% of the sample outputs without tree regeneration, and these were considered as zero stand basal area and zero regeneration for all the tree species simulated; one site produced grassland rather than a forest and was not included in the results.
 
@@
 
-The simulation results were analyzed regarding (1) regeneration levels
-(i.e., ingrowth number per unit time and space), (2) regeneration
-species diversity, (3) regeneration mortality, (4) the relationship
-between model performance and model traits, and (5) ingrowth gradients
-along the regeneration niches. We evaluated species diversity across the
-models and in relation to the observed data by calculating the Shannon
-index $H$ based on the relative proportion of the species in terms of
-basal area. It was calculated for the regeneration ($H_{R_{n}}$)
-(Equation \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$)
-(Equation \ref{eqpiShannon}). The higher the value of the index, the
-higher the species diversity at a particular site and sample.
+The simulation results were analyzed regarding (1) regeneration levels (i.e., ingrowth number per unit time and space), (2) regeneration species diversity, (3) regeneration mortality, (4) the relationship between model performance and model traits, and (5) ingrowth gradients along the regeneration niches.
+We evaluated species diversity across the models and in relation to the observed data by calculating the Shannon index $H$ based on the relative proportion of the species in terms of basal area.
+It was calculated for the regeneration ($H_{R_{n}}$) (Equation \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation \ref{eqpiShannon}).
+The higher the value of the index, the higher the species diversity at a particular site and sample.
 
@@
 ```
-where $s$ is the total number of species present and that have a basal
-area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
-species $i$ in sample $n$ calculated as the regeneration basal area
-($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal
-area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
-proportion of species $i$ calculated as the basal area of all trees
-($BA_{n,i}$) of that species $i$ relative to total basal area
-($TotalR_{BA_{n}}$) of the sample $n$. Species diversity was not
-assessed for aDGVM2 as this model does not simulate individual species.
+where $s$ is the total number of species present and that have a basal area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of species $i$ in sample $n$ calculated as the regeneration basal area ($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the proportion of species $i$ calculated as the basal area of all trees ($BA_{n,i}$) of that species $i$ relative to total basal area ($TotalR_{BA_{n}}$) of the sample $n$.
+Species diversity was not assessed for aDGVM2 as this model does not simulate individual species.
 
-Mortality in tree regeneration was assessed based on the ratio of
-regeneration between the 7 and 10 cm diameter thresholds. We used the
-Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference
-to estimate whether the ratio of regeneration between the 7 and 10 cm
-diameter thresholds were above or below the expected theoretical rate.
-The Reineke self-thinning rule is usually calculated for even-aged,
-single species stands and assumes a fixed relationship between the
-number of stems and the quadratic mean diameter in fully stocked pure
-stands. The value used in our comparisons was 1.77 (i.e., we expect stem
-numbers at 7 cm to be 77% higher than at 10 cm), calculated using
-Equation \ref{eqReineke}.
+Mortality in tree regeneration was assessed based on the ratio of regeneration between the 7 and 10 cm diameter thresholds.
+We used the Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference to estimate whether the ratio of regeneration between the 7 and 10 cm diameter thresholds were above or below the expected theoretical rate.
+The Reineke self-thinning rule is usually calculated for even-aged, single species stands and assumes a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands.
+The value used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm to be 77% higher than at 10 cm), calculated using Equation \ref{eqReineke}.
 
@@
 ```
-We assessed model performance in relation to model traits focusing on
-(1) model complexity as defined by @bugmann2022, (2) model type
-(empirical or process based), (3) the presence or absence of a canopy
-feedback for regeneration, and (4) the scale of application of the model
-(stand, landscape or global). We tested for significant differences
-using two-sided t-tests by considering each of these model features
-regarding the observed and simulated regeneration levels and species
-diversity, respectively.
+We assessed model performance in relation to model traits focusing on (1) model complexity as defined by @bugmann2022, (2) model type (empirical or process based), (3) the presence or absence of a canopy feedback for regeneration, and (4) the scale of application of the model (stand, landscape or global).
+We tested for significant differences using two-sided t-tests by considering each of these model features regarding the observed and simulated regeneration levels and species diversity, respectively.
 
-Total regeneration and the regeneration niches of the individual species
-were evaluated across the environmental gradients of light availability,
-temperature, and soil moisture. Stand basal area (the basal area of all
-the trees in each sample including tree regeneration) was used as a
-proxy for light availability at the forest floor, the annual degree-day
-sum [@allen1976; @fischlin1995] as a proxy for growing season warmth,
-and the climatic water balance as a proxy for soil moisture
-[@speich2019]. Regeneration values were calculated as the mean across
-the 200 samples per site. The observed data were modeled using a
-Generalized Additive Model [@wood2011] with a negative binomial
-distribution and restricted maximum likelihood to better understand the
-relationship between the environmental gradients and the levels of
-observed regeneration, relative to the simulation results per model.
+Total regeneration and the regeneration niches of the individual species were evaluated across the environmental gradients of light availability, temperature, and soil moisture.
+Stand basal area (the basal area of all the trees in each sample including tree regeneration) was used as a proxy for light availability at the forest floor, the annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing season warmth, and the climatic water balance as a proxy for soil moisture [@speich2019].
+Regeneration values were calculated as the mean across the 200 samples per site.
+The observed data were modeled using a Generalized Additive Model [@wood2011] with a negative binomial distribution and restricted maximum likelihood to better understand the relationship between the environmental gradients and the levels of observed regeneration, relative to the simulation results per model.
 
-In order to analyze the regeneration niches across the climatic
-gradients, we focused on five common tree species or genera: *Abies
-alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
-*Quercus* spp. For these species, we calculated the share in the
-regeneration basal area per site as the mean across the available
-samples per site:
+In order to analyze the regeneration niches across the climatic gradients, we focused on five common tree species or genera: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and *Quercus* spp.
+For these species, we calculated the share in the regeneration basal area per site as the mean across the available samples per site:
 
@@
 ```
-where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
-is the mean basal area in the regeneration (subsequently referred to as
-"regeneration basal area") of species $i$ per site $t$ across the
-available samples at that site. Furthermore, regarding the share in the
-regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we
-categorized this as zero when the regeneration basal area of that
-species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the
-regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) and the
-total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$)
-equaled zero.
+where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$ is the mean basal area in the regeneration (subsequently referred to as "regeneration basal area") of species $i$ per site $t$ across the available samples at that site.
+Furthermore, regarding the share in the regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) and the total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) equaled zero.
 
@@
 
-Simulated regeneration levels varied strongly across the 15 models and
-typically did not match the levels found in the forest reserve data
-(Figure \@ref(fig:over)). Regeneration was overestimated in most models
-for both the 7 and 10 cm diameter thresholds, with the exception of the
-empirical stand model SIBYLA, the landscape model Landis-II and the
-global model aDGVM2, which estimated regeneration levels at the lower
-end of the plausibility interval of the observed data (Figure
-\@ref(fig:over)). The models with the largest overestimation were the
-stand model PICUS and the landscape model TreeMig. For most models, the
-variability of simulated regeneration levels across the 200 sites
-(visible from the interquartile range in the box plots of Figure
-\@ref(fig:over)) was similar to or smaller than observed, with the
-notable exception of PICUS, where simulated regeneration variability was
-much larger.
+Simulated regeneration levels varied strongly across the 15 models and typically did not match the levels found in the forest reserve data (Figure \@ref(fig:over)).
+Regeneration was overestimated in most models for both the 7 and 10 cm diameter thresholds, with the exception of the empirical stand model SIBYLA, the landscape model Landis-II and the global model aDGVM2, which estimated regeneration levels at the lower end of the plausibility interval of the observed data (Figure \@ref(fig:over)).
+The models with the largest overestimation were the stand model PICUS and the landscape model TreeMig.
+For most models, the variability of simulated regeneration levels across the 200 sites (visible from the interquartile range in the box plots of Figure \@ref(fig:over)) was similar to or smaller than observed, with the notable exception of PICUS, where simulated regeneration variability was much larger.
 
@@
 
-Both the observations and the simulated data had no regeneration in some
-samples and at some sites (for details, cf. Table S1). The observed data
-had 4% of the samples with no regeneration. Only three models had a
-larger proportion of no regeneration (4C, Landis-II and aDVM2). Two
-models always simulated regeneration for both the 7 and 10 cm threshold
-(xComp and TreeMig), i.e., they did not feature any zero values. The
-other ten models had a very low percentage of samples with no
-regeneration (0.01% - 2.39%), i.e., they had distinctly fewer
-occurrences of zeros compared to the observations.
+Both the observations and the simulated data had no regeneration in some samples and at some sites (for details, cf.
+Table S1).
+The observed data had 4% of the samples with no regeneration.
+Only three models had a larger proportion of no regeneration (4C, Landis-II and aDVM2).
+Two models always simulated regeneration for both the 7 and 10 cm threshold (xComp and TreeMig), i.e., they did not feature any zero values.
+The other ten models had a very low percentage of samples with no regeneration (0.01% - 2.39%), i.e., they had distinctly fewer occurrences of zeros compared to the observations.
 
@@
 
-Most models matched the level of diversity of the observed data quite
-well (Figure \@ref(fig:H710)). Five models overestimated regeneration
-diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
-particularly for the 7 cm diameter threshold. The model 4C is a special
-case, as it simulated five species only, i.e., its diversity values are
-not directly comparable to those of the other models, nor to the
-observations. Only one model, Landis-II, consistently underestimated
-regeneration diversity.
+Most models matched the level of diversity of the observed data quite well (Figure \@ref(fig:H710)).
+Five models overestimated regeneration diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter particularly for the 7 cm diameter threshold.
+The model 4C is a special case, as it simulated five species only, i.e., its diversity values are not directly comparable to those of the other models, nor to the observations.
+Only one model, Landis-II, consistently underestimated regeneration diversity.
 
@@
 
-In most models, there were only small but significant differences in the
-species diversity of regeneration between 7 and 10 cm across sites
-(Tables S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS)
-maintained the regeneration diversity between the 7 and 10 cm thresholds
-(i.e., the differences between them were not significant, see Table S2),
-and the same was evident from the observed data.
+In most models, there were only small but significant differences in the species diversity of regeneration between 7 and 10 cm across sites (Tables S2).
+Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained the regeneration diversity between the 7 and 10 cm thresholds (i.e., the differences between them were not significant, see Table S2), and the same was evident from the observed data.
 
-For both the observed and the simulated data, and both regeneration
-thresholds, species diversity in the regeneration compared to
-stand-level diversity did not feature clear patterns across the models
-(Figures S1 and S2). Most models captured reasonably well or
-overpredicted species diversity at the stand level (Figure
-\@ref(fig:combinedRichness)A, left and center; cf. Figure S3), but a
-characteristic feature was that the simulations had a much lower
-variability of diversity compared to observations. Models overpredicting
-species diversity included ForClim1, ForClim 11, FORMIND and TreeMig
-(Figure \@ref(fig:combinedRichness)A, left). Several models
-underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C
-(Figure \@ref(fig:combinedRichness)A, right).
+For both the observed and the simulated data, and both regeneration thresholds, species diversity in the regeneration compared to stand-level diversity did not feature clear patterns across the models (Figures S1 and S2).
+Most models captured reasonably well or overpredicted species diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left and center; cf.
+Figure S3), but a characteristic feature was that the simulations had a much lower variability of diversity compared to observations.
+Models overpredicting species diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A, left).
+Several models underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).
 
-The majority of the models overpredicted regeneration diversity; as
-already observed in Figure \@ref(fig:H710), only two of them
-underestimated it (4C and Landis-II) (Figure
-\@ref(fig:combinedRichness)B and Figure S4). Both the models with
-feedback and those without feedback from the adult trees to regeneration
-(via seed production) overestimated, underestimated or captured
-reasonably well regeneration species diversity compared to the
-observations. Again, simulated diversity had considerably lower
-variability than observed diversity in the regeneration.
+The majority of the models overpredicted regeneration diversity; as already observed in Figure \@ref(fig:H710), only two of them underestimated it (4C and Landis-II) (Figure \@ref(fig:combinedRichness)B and Figure S4).
+Both the models with feedback and those without feedback from the adult trees to regeneration (via seed production) overestimated, underestimated or captured reasonably well regeneration species diversity compared to the observations.
+Again, simulated diversity had considerably lower variability than observed diversity in the regeneration.
 
@@
 
-There were strong differences among the models regarding the mortality
-rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the
-models featured a mortality rate significantly larger than the observed
-data, thus at least partially compensating for the general
-overestimation of regeneration levels (Figure \@ref(fig:over) and Table
-S3); six models did not feature a significant difference compared to the
-observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
-Four models (xComp, PICUS, LandClim and TreeMig) featured very high
-mortality rates (i.e., well above the Reineke self-thinning line), which
-compensated for the strong initial overestimation of regeneration (cf.
-Figure \@ref(fig:over)). Interestingly, two models that underestimated
-overall regeneration levels (SIBYLA and Landis-II; cf. Figure
-\@ref(fig:over)) featured mortality rates that were close to but still
-above the self-thinning line (Figure \@ref(fig:mortality11)).
+There were strong differences among the models regarding the mortality rate between 7 and 10 cm (Figure \@ref(fig:mortality11)).
+Most of the models featured a mortality rate significantly larger than the observed data, thus at least partially compensating for the general overestimation of regeneration levels (Figure \@ref(fig:over) and Table S3); six models did not feature a significant difference compared to the observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
+Four models (xComp, PICUS, LandClim and TreeMig) featured very high mortality rates (i.e., well above the Reineke self-thinning line), which compensated for the strong initial overestimation of regeneration (cf.
+Figure \@ref(fig:over)).
+Interestingly, two models that underestimated overall regeneration levels (SIBYLA and Landis-II; cf.
+Figure \@ref(fig:over)) featured mortality rates that were close to but still above the self-thinning line (Figure \@ref(fig:mortality11)).
 
@@
 
-The fact that some models (and observed data) featured lower mortality
-than expected by Reineke's self-thinning rule may point to facilitation,
-or simply a higher mortality rate before the trees had reached 7 cm
-diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
-whose mortality was well above the self-thinning line, indicates that
-they feature higher regeneration mortality in mixed-species forests
-compared to even-aged, single-species forests. Some models yielded a
-mortality rate of regeneration that is broadly compatible with the
-self-thinning rule (e.g., 4C and LandClim).
+The fact that some models (and observed data) featured lower mortality than expected by Reineke's self-thinning rule may point to facilitation, or simply a higher mortality rate before the trees had reached 7 cm diameter.
+Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher regeneration mortality in mixed-species forests compared to even-aged, single-species forests.
+Some models yielded a mortality rate of regeneration that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).
 
-The three models with the largest overestimation of the proportion of
-regeneration at the 7 cm threshold featured the highest mortality
-between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The
-other models that overestimated regeneration had mortalities lower than
-the expected self-thinning ratio, with the exception of LandClim (Figure
-\@ref(fig:mortality13)B). It is noteworthy that most of the models that
-featured a low ratio of regeneration between 7 and 10 cm (i.e.,
-ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small
-overestimation of regeneration at the 7 cm threshold (Figure
-\@ref(fig:mortality11)). aDGVM2 was the only model that underestimated
-regeneration at the 7 cm threshold while featuring a low mortality
-between the two thresholds (Figure \@ref(fig:mortality13)C), and only a
-few models were close to the correct regeneration levels at 7 cm while
-also being close to the Reineke line (SIBYLA, Landis-II; cf. Figure
-\@ref(fig:mortality13)D; and 4C as well as LandClim).
+The three models with the largest overestimation of the proportion of regeneration at the 7 cm threshold featured the highest mortality between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A).
+The other models that overestimated regeneration had mortalities lower than the expected self-thinning ratio, with the exception of LandClim (Figure \@ref(fig:mortality13)B).
+It is noteworthy that most of the models that featured a low ratio of regeneration between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small overestimation of regeneration at the 7 cm threshold (Figure \@ref(fig:mortality11)).
+aDGVM2 was the only model that underestimated regeneration at the 7 cm threshold while featuring a low mortality between the two thresholds (Figure \@ref(fig:mortality13)C), and only a few models were close to the correct regeneration levels at 7 cm while also being close to the Reineke line (SIBYLA, Landis-II; cf.
+Figure \@ref(fig:mortality13)D; and 4C as well as LandClim).
 
@@
 
-Lastly, some models that had a low number of individuals at the 7 cm
-threshold also had a smaller decrease of regeneration between 7 and 10
-cm (ratio 7-10 cm) (Figure S5). In these models, competition and
-self-thinning are either not pronounced or must have occurred before the
-trees had reached 7 cm. However, this pattern was not consistent across
-models (Figure S5). It is noteworthy that there is a relationship
-between the regeneration levels at 7 cm and the mortality rate (Table
-S4): most of the models showed a positive effect (higher regeneration at
-7 cm is coupled to an increase of the mortality between 7 and 10 cm).
-While some models showed a negative effect, the linear trend was not
-pronounced and not always significant regarding the slope.
+Lastly, some models that had a low number of individuals at the 7 cm threshold also had a smaller decrease of regeneration between 7 and 10 cm (ratio 7-10 cm) (Figure S5).
+In these models, competition and self-thinning are either not pronounced or must have occurred before the trees had reached 7 cm.
+However, this pattern was not consistent across models (Figure S5).
+It is noteworthy that there is a relationship between the regeneration levels at 7 cm and the mortality rate (Table S4): most of the models showed a positive effect (higher regeneration at 7 cm is coupled to an increase of the mortality between 7 and 10 cm).
+While some models showed a negative effect, the linear trend was not pronounced and not always significant regarding the slope.
 
@@
 
-There was no significant relationship between the mean complexity of the
-regeneration module (Table \@ref(tab:tableModels)) and the
-overestimation proportion at a diameter threshold of 7 cm (Figure S6).
-All the models except 4C had significant differences between the
-observed and simulated mean regeneration values. Therefore it was not
-possible to differentiate in terms of performance between models type
-(empirical or process based), or their scales (stand, landscape or
-global) (Table S5). The values of regeneration species diversity
-simulated by each model were also assessed against the observed data to
-see if there was a difference between models that include a feedback
-compared to those that did not; all the models except ForCEEPs(f) and
-FORMIND had significant differences (Table S6). It is noteworthy that
-for the model ForCEEPS the pattern changed from the overprediction of
-species diversity in the regeneration to a diversity level that is
-closer to observations when the recruitment module included feedback
-(ForCEEPs(f)).
+There was no significant relationship between the mean complexity of the regeneration module (Table \@ref(tab:tableModels)) and the overestimation proportion at a diameter threshold of 7 cm (Figure S6).
+All the models except 4C had significant differences between the observed and simulated mean regeneration values.
+Therefore it was not possible to differentiate in terms of performance between models type (empirical or process based), or their scales (stand, landscape or global) (Table S5).
+The values of regeneration species diversity simulated by each model were also assessed against the observed data to see if there was a difference between models that include a feedback compared to those that did not; all the models except ForCEEPs(f) and FORMIND had significant differences (Table S6).
+It is noteworthy that for the model ForCEEPS the pattern changed from the overprediction of species diversity in the regeneration to a diversity level that is closer to observations when the recruitment module included feedback (ForCEEPs(f)).
 
@@
 
-When evaluating regeneration levels along key gradients of light
-availability (basal area), temperature (degree-days) and soil moisture
-(climatic water balance), distinct features emerged: the models
-reproduced the effect of basal area in both its magnitude and patterns
-(Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better
-than the effects of the climatic gradients, where they featured varying
-patterns.
+When evaluating regeneration levels along key gradients of light availability (basal area), temperature (degree-days) and soil moisture (climatic water balance), distinct features emerged: the models reproduced the effect of basal area in both its magnitude and patterns (Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better than the effects of the climatic gradients, where they featured varying patterns.
 
-In the observed data, regeneration levels decreased clearly with
-increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a).
-This trend was captured by the Landis-II model only, albeit at much
-lower values of basal area than in the observations. The other models
-featured distinctly different trends, such as (1) an increase in
-regeneration levels with increasing total basal area followed by the
-absence of regeneration at high values of basal area (ForCEEPS,
-ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant
-regeneration levels with basal area (ForClim 11, xComp, LandClim and
-aDGVM2), or (3) an increase of regeneration up to a certain value of
-stand basal area followed by a decrease at even higher values of stand
-basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA
-and iLand).
+In the observed data, regeneration levels decreased clearly with increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a).
+This trend was captured by the Landis-II model only, albeit at much lower values of basal area than in the observations.
+The other models featured distinctly different trends, such as (1) an increase in regeneration levels with increasing total basal area followed by the absence of regeneration at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant regeneration levels with basal area (ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of regeneration up to a certain value of stand basal area followed by a decrease at even higher values of stand basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).
 
-The observed regeneration did not change much across the climatic water
-balance gradient, and showed a slight increase with temperature. The
-models did not match the observed regeneration patterns across the
-gradients of soil moisture and temperature (Figure
-\@ref(fig:nvTrend710BAlightTemp)b,c and S7), and neither was there a
-clear pattern across models. Regarding the soil moisture gradient, a
-group of models featured decreasing regeneration with increasing soil
-moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim,
-TreeMig and iLand), while a few models showed the opposite trend
-(ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient,
-all the models that featured a decrease in regeneration levels with
-increasing soil moisture showed the opposite trend (i.e., more
-regeneration with increasing temperature), except for 4C. The other
-models did not feature a clear pattern.
+The observed regeneration did not change much across the climatic water balance gradient, and showed a slight increase with temperature.
+The models did not match the observed regeneration patterns across the gradients of soil moisture and temperature (Figure \@ref(fig:nvTrend710BAlightTemp)b,c and S7), and neither was there a clear pattern across models.
+Regarding the soil moisture gradient, a group of models featured decreasing regeneration with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS).
+Across the temperature gradient, all the models that featured a decrease in regeneration levels with increasing soil moisture showed the opposite trend (i.e., more regeneration with increasing temperature), except for 4C.
+The other models did not feature a clear pattern.
 
@@
 
-Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
-sylvestris*) and oak (*Quercus* spp.) were the species observed most
-frequently in the EuFoRIa data. Most models captured well the share of
-basal area in the regeneration of these five species compared to the
-observed data along the two major climatic gradients (Figure
-\@ref(fig:diff), cf. Figure S8). However, the models differed
-considerably in the way in which they simulated these climatic niches.
+Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus sylvestris*) and oak (*Quercus* spp.) were the species observed most frequently in the EuFoRIa data.
+Most models captured well the share of basal area in the regeneration of these five species compared to the observed data along the two major climatic gradients (Figure \@ref(fig:diff), cf.
+Figure S8).
+However, the models differed considerably in the way in which they simulated these climatic niches.
 
-First, some models overestimated regeneration across the entire
-gradients of at least one of the five main tree species (Figure
-\@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS),
-while other models overestimated regeneration of more than one species
-for a large part of the climate gradients, such as xComp, PICUS, SIBYLA
-or LPJ-GUESS (cf. red colors in Figure \@ref(fig:diff)). These trends
-were consistent for both regeneration thresholds, although the exact
-changes in the share of the regeneration basal area were different
-(Figure S8).
+First, some models overestimated regeneration across the entire gradients of at least one of the five main tree species (Figure \@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated regeneration of more than one species for a large part of the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
+red colors in Figure \@ref(fig:diff)).
+These trends were consistent for both regeneration thresholds, although the exact changes in the share of the regeneration basal area were different (Figure S8).
 
-Second, there was some congruence in the simulation results by species
-across the models. The share in the regeneration of Silver fir (*Abies
-alba*) was captured well across the gradient by most models except for
-xComp, PICUS, and LandClim, which overestimated the regeneration share
-of this species. There were also some peculiarities evident for some
-models, such as 4C that did not simulate *Abies alba*. The patterns
-across models for *Fagus sylvatica* regeneration were more complex, as
-its regeneration was underestimated by many models across the
-environmental gradients featuring negligible regeneration at most sites,
-while others overestimated it in different parts of the environmental
-space. Most models overestimated the share of *Picea abies* regeneration
-in the cold-wet part of the gradients. *Pinus sylvestris* and *Quercus*
-spp. represented a small share of the regeneration in the observed data
-mostly at warm-dry sites. Many models, however, erroneously featured
-regeneration for these two species along most of the climatic gradients,
-although with a low share of basal area. Five models (4C, FORMIND,
-SIBYLA, xComp, and LandClim) had almost no regeneration of *Pinus
-sylvestris* and *Quercus* spp., or did not feature any regeneration of
-these species at all.
+Second, there was some congruence in the simulation results by species across the models.
+The share in the regeneration of Silver fir (*Abies alba*) was captured well across the gradient by most models except for xComp, PICUS, and LandClim, which overestimated the regeneration share of this species.
+There were also some peculiarities evident for some models, such as 4C that did not simulate *Abies alba*.
+The patterns across models for *Fagus sylvatica* regeneration were more complex, as its regeneration was underestimated by many models across the environmental gradients featuring negligible regeneration at most sites, while others overestimated it in different parts of the environmental space.
+Most models overestimated the share of *Picea abies* regeneration in the cold-wet part of the gradients.
+*Pinus sylvestris* and *Quercus* spp.
+represented a small share of the regeneration in the observed data mostly at warm-dry sites.
+Many models, however, erroneously featured regeneration for these two species along most of the climatic gradients, although with a low share of basal area.
+Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had almost no regeneration of *Pinus sylvestris* and *Quercus* spp., or did not feature any regeneration of these species at all.
 
-Lastly, there is no model that performed well across the five species
-regarding the species-specific regeneration levels in the environmental
-space (Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
-tended to consistently overestimate the regeneration share of all five
-species, but most models overestimated the regeneration share of some
-species while underestimating the share of others (e.g., ForCEEPS,
-LPJ-GUESS).
+Lastly, there is no model that performed well across the five species regarding the species-specific regeneration levels in the environmental space (Figure \@ref(fig:diff)).
+Some models (e.g., ForClim11, PICUS) tended to consistently overestimate the regeneration share of all five species, but most models overestimated the regeneration share of some species while underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
 
@@
 
-Tree regeneration is a fundamental process in forest dynamics. Correctly
-capturing tree regeneration in dynamic models is fundamental to, for
-example, evaluate post-disturbance dynamics and potential long-term
-recovery trajectories as it will define the forest state in the first
-decades [@seidl2022]. If this initial phase is not captured well, we can
-not properly assess aspects such as forest resilience or the timing,
-magnitude and progression of carbon sequestration. In this study, for
-the first time, projections of tree regeneration from multiple models of
-forest dynamics were confronted with a unique dataset from unmanaged
-forest reserves across a large environmental gradient in Europe.
+Tree regeneration is a fundamental process in forest dynamics.
+Correctly capturing tree regeneration in dynamic models is fundamental to, for example, evaluate post-disturbance dynamics and potential long-term recovery trajectories as it will define the forest state in the first decades [@seidl2022].
+If this initial phase is not captured well, we can not properly assess aspects such as forest resilience or the timing, magnitude and progression of carbon sequestration.
+In this study, for the first time, projections of tree regeneration from multiple models of forest dynamics were confronted with a unique dataset from unmanaged forest reserves across a large environmental gradient in Europe.
 
-The EuFoRIa data [@käber2023] are exceptional, particularly with respect
-to the number of records (number of sites and repeated measurements),
-which is essential for capturing a highly 'noisy' process such as tree
-regeneration. The use of this dataset for model benchmarking provided
-novel insights on the ability of state-of-the-art models to accurately
-simulate regeneration levels, species composition, and mortality in an
-early stage of tree life, i.e., between a tree diameter of 7 and 10 cm.
-Overall, by adopting this approach a much broader understanding resulted
-than if we had used these data for model calibration: it is primarily
-from the shortcomings of the models that we can gain ecological insights
-[cf. @trugman2022].
+The EuFoRIa data [@käber2023] are exceptional, particularly with respect to the number of records (number of sites and repeated measurements), which is essential for capturing a highly 'noisy' process such as tree regeneration.
+The use of this dataset for model benchmarking provided novel insights on the ability of state-of-the-art models to accurately simulate regeneration levels, species composition, and mortality in an early stage of tree life, i.e., between a tree diameter of 7 and 10 cm.
+Overall, by adopting this approach a much broader understanding resulted than if we had used these data for model calibration: it is primarily from the shortcomings of the models that we can gain ecological insights [cf.
+@trugman2022].
 
@@
 
-Most of the models overestimated tree regeneration levels. This has
-potentially far-reaching implications, e.g., regarding biomass (and thus
-carbon) turnover, with a potential overestimation of the capacity for
-forest carbon sequestration [@pugh2019; @pan2011]. Yet, we focused on
-one specific stage, i.e., regeneration into the 7 and 10 cm diameter
-classes. Trees at this size contribute little to carbon sequestration,
-and if the excess regeneration at this stage is compensated -- in the
-models -- soon thereafter by higher mortality, overall simulated
-regeneration may still yield trustworthy results at the level of stand
-structure and related forest attributes such as basal area or biomass.
+Most of the models overestimated tree regeneration levels.
+This has potentially far-reaching implications, e.g., regarding biomass (and thus carbon) turnover, with a potential overestimation of the capacity for forest carbon sequestration [@pugh2019; @pan2011].
+Yet, we focused on one specific stage, i.e., regeneration into the 7 and 10 cm diameter classes.
+Trees at this size contribute little to carbon sequestration, and if the excess regeneration at this stage is compensated -- in the models -- soon thereafter by higher mortality, overall simulated regeneration may still yield trustworthy results at the level of stand structure and related forest attributes such as basal area or biomass.
 
-From a modeling point of view, excess regeneration inevitably requires
-excess mortality rates in a later stage, either -- as observed for some
-models -- between the 7 and 10 cm diameter thresholds, or soon after the
-10 cm threshold has been crossed. In any case, correcting at early
-stages for the expected forest densities at later stages is equivalent
-to compensating for a first error (excess regeneration) by a second
-error (excess mortality). It is highly likely that biased projections
-will result, because the two errors are unlikely to be perfectly linked
-and thus will not always compensate each other. Hence, this structural
-problem of most models investigated here is problematic particularly if
-the models are to be used under novel conditions such as under climate
-change [e.g., @huber2021] or in a decision support context [e.g.,
-@thrippleton2021].
+From a modeling point of view, excess regeneration inevitably requires excess mortality rates in a later stage, either -- as observed for some models -- between the 7 and 10 cm diameter thresholds, or soon after the 10 cm threshold has been crossed.
+In any case, correcting at early stages for the expected forest densities at later stages is equivalent to compensating for a first error (excess regeneration) by a second error (excess mortality).
+It is highly likely that biased projections will result, because the two errors are unlikely to be perfectly linked and thus will not always compensate each other.
+Hence, this structural problem of most models investigated here is problematic particularly if the models are to be used under novel conditions such as under climate change [e.g., @huber2021] or in a decision support context [e.g., @thrippleton2021].
 
-Regeneration levels define the structure and composition of future
-forests, but it is equally important to correctly identify areas where
-regeneration is lacking [@rammer2021]. There are multiple constraints to
-the regeneration niche of tree species [@price2001], and therefore the
-absence of regeneration is likely to be common [@fortin2007], even over
-larger areas such as the one-hectare plots used here. Tree regeneration
-data are characterized by zero-inflation. This was clearly evident from
-the EuFoRIa dataset, but some of the models did not produce zeros at
-all, or featured a very low proportion of zero data. This substantial
-difference may be due to the fact that the simulation results were drawn
-from equilibrium forests, whereas in reality many of the forest reserves
-are recovering from past management activities and have become denser
-over the past decades [e.g., @heiri2009], leading to less regeneration
-than in an equilibrium situation.
+Regeneration levels define the structure and composition of future forests, but it is equally important to correctly identify areas where regeneration is lacking [@rammer2021].
+There are multiple constraints to the regeneration niche of tree species [@price2001], and therefore the absence of regeneration is likely to be common [@fortin2007], even over larger areas such as the one-hectare plots used here.
+Tree regeneration data are characterized by zero-inflation.
+This was clearly evident from the EuFoRIa dataset, but some of the models did not produce zeros at all, or featured a very low proportion of zero data.
+This substantial difference may be due to the fact that the simulation results were drawn from equilibrium forests, whereas in reality many of the forest reserves are recovering from past management activities and have become denser over the past decades [e.g., @heiri2009], leading to less regeneration than in an equilibrium situation.
 
@@
 
-Correctly capturing species composition in tree regeneration is
-important to assess the future functional diversity of a forest, e.g.,
-its sensitivity to drought or resilience to disturbances [@redmond2015;
-@seidl2022]. In the simulations, overall species diversity levels in the
-regeneration were well within the observed range for half of the models
-(7 out of 14). Thus, while most of the models are facing difficulties to
-quantitatively match regeneration levels (cf. above), their performance
-is better when it comes to matching the composition of regeneration as a
-function of abiotic and biotic conditions. Most models maintained or
-even decreased regeneration diversity between the 7 and 10 cm
-thresholds, and the same is visible from the empirical data, but the
-differences were not significant. Diversity variations may be explained
-by the way the individual models consider regeneration processes
-[@konig_tree_2022], e.g., without considering the feedback from adults,
-regeneration can be linked to either seed rain from dispersal or a
-seed/seedling bank; neither process is incorporated across many models
-studied here, however.
+Correctly capturing species composition in tree regeneration is important to assess the future functional diversity of a forest, e.g., its sensitivity to drought or resilience to disturbances [@redmond2015; @seidl2022].
+In the simulations, overall species diversity levels in the regeneration were well within the observed range for half of the models (7 out of 14).
+Thus, while most of the models are facing difficulties to quantitatively match regeneration levels (cf.
+above), their performance is better when it comes to matching the composition of regeneration as a function of abiotic and biotic conditions.
+Most models maintained or even decreased regeneration diversity between the 7 and 10 cm thresholds, and the same is visible from the empirical data, but the differences were not significant.
+Diversity variations may be explained by the way the individual models consider regeneration processes [@konig_tree_2022], e.g., without considering the feedback from adults, regeneration can be linked to either seed rain from dispersal or a seed/seedling bank; neither process is incorporated across many models studied here, however.
 
-The species diversity of the entire stand was captured better by the
-models than the diversity of tree regeneration, and only a few models
-overpredicted stand-level diversity. Defining regeneration composition
-as being proportional to adult abundance, regardless of productivity,
-might be a simple and conservative assumption to safely maintain
-relative species abundances [@hanbury-brown2022], although this may be
-simplistic. Furthermore, based on the simulation results, there is no
-evidence that models with feedback from the canopy captured better the
-species diversity of the regeneration than those without feedback, with
-the exception of ForCEEPS that significantly improved regeneration
-species diversity with the model version that included feedback
-(ForCEEPS(f)). The similar performance of models with and without
-feedback is likely because the models put more weight on the
-regeneration niche arising from abiotic and biotic filters, than from
-the habitat niche of the adult trees [@grubb1977].
+The species diversity of the entire stand was captured better by the models than the diversity of tree regeneration, and only a few models overpredicted stand-level diversity.
+Defining regeneration composition as being proportional to adult abundance, regardless of productivity, might be a simple and conservative assumption to safely maintain relative species abundances [@hanbury-brown2022], although this may be simplistic.
+Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy captured better the species diversity of the regeneration than those without feedback, with the exception of ForCEEPS that significantly improved regeneration species diversity with the model version that included feedback (ForCEEPS(f)).
+The similar performance of models with and without feedback is likely because the models put more weight on the regeneration niche arising from abiotic and biotic filters, than from the habitat niche of the adult trees [@grubb1977].
 
@@
 
-There are several factors that lead to mortality during the regeneration
-phase, such as competition [@casper1997] and multiple abiotic factors
-[@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
-expect that models that overestimate regeneration may have a
-particularly high mortality between the 7 and 10 cm thresholds. However,
-this was not consistently evident from the simulations. This implies
-that these models must have an excess mortality in later stages, if we
-assume that all models were able to capture the structure and
-composition of the adult stands along the EuFoRIa gradient; this however
-was not tested here.
+There are several factors that lead to mortality during the regeneration phase, such as competition [@casper1997] and multiple abiotic factors [@schmid2021; @cunningham2006].
+As mentioned above, it is reasonable to expect that models that overestimate regeneration may have a particularly high mortality between the 7 and 10 cm thresholds.
+However, this was not consistently evident from the simulations.
+This implies that these models must have an excess mortality in later stages, if we assume that all models were able to capture the structure and composition of the adult stands along the EuFoRIa gradient; this however was not tested here.
 
-Higher mortality towards the adult phase has important implications for
-forest dynamics and the goods and services provided by forests. On the
-one hand, mortality in later stages may erroneously enhance the share of
-less shade tolerant species in the models [@klopcic2015], therefore
-shifting the species composition. Unrealistic high stem densities for a
-longer period of time may overestimate the role of tree regrowth in
-carbon sink dynamics [@pugh2019]. Given our set of simulations and
-analyses, we cannot conclusively assess what is happening in the models,
-and further studies are required that focus on a wider range of tree
-sizes and the fate of tree regeneration along such a size continuum.
+Higher mortality towards the adult phase has important implications for forest dynamics and the goods and services provided by forests.
+On the one hand, mortality in later stages may erroneously enhance the share of less shade tolerant species in the models [@klopcic2015], therefore shifting the species composition.
+Unrealistic high stem densities for a longer period of time may overestimate the role of tree regrowth in carbon sink dynamics [@pugh2019].
+Given our set of simulations and analyses, we cannot conclusively assess what is happening in the models, and further studies are required that focus on a wider range of tree sizes and the fate of tree regeneration along such a size continuum.
 
@@
 
-In spite of the critical considerations above, it is remarkable that
-most models did not deviate exceedingly from the observations with
-respect to simulated regeneration levels -- after all, few if any of the
-models' regeneration routines are sufficiently constrained by data, with
-the exception of the empirical models xComp and SIBYLA. It is noteworthy
-that in spite of their empirical basis, these two models did not match
-empirical regeneration levels, in a similar magnitude as the other
-models and even in opposite directions (SIBYLA: underestimation; xComp:
-overestimation). It appears that using empirical data of limited
-geographical scope and constrained to managed forests, as done in these
-models, leads to extrapolation problems already under current climatic
-conditions. For the other models, whose regeneration modules are not
-strongly constrained by empirical data, the degrees of freedom that are
-available for modeling regeneration [@konig_tree_2022] could have
-implied that model performance would be much worse than what we found.
+In spite of the critical considerations above, it is remarkable that most models did not deviate exceedingly from the observations with respect to simulated regeneration levels -- after all, few if any of the models' regeneration routines are sufficiently constrained by data, with the exception of the empirical models xComp and SIBYLA.
+It is noteworthy that in spite of their empirical basis, these two models did not match empirical regeneration levels, in a similar magnitude as the other models and even in opposite directions (SIBYLA: underestimation; xComp: overestimation).
+It appears that using empirical data of limited geographical scope and constrained to managed forests, as done in these models, leads to extrapolation problems already under current climatic conditions.
+For the other models, whose regeneration modules are not strongly constrained by empirical data, the degrees of freedom that are available for modeling regeneration [@konig_tree_2022] could have implied that model performance would be much worse than what we found.
 
-Our study showed that increasing complexity in the regeneration modules
-is not linked with a higher accuracy of the projections of regeneration
-levels, species composition or mortality at early tree stages, as there
-was no significant relationship between model performance and model
-complexity. Increasing complexity in regeneration modules has been
-motivated by better process understanding and enhancing of model
-accuracy [@bugmann2022]. However, more complex models do not necessarily
-lead to better projections and rather to a reduced transparency and
-lower predictive power [@franklin2020]. Thus, the question of what level
-of detail is appropriate and parsimonious for modeling tree regeneration
-processes remains open [@konig_tree_2022; @bugmann2022].
+Our study showed that increasing complexity in the regeneration modules is not linked with a higher accuracy of the projections of regeneration levels, species composition or mortality at early tree stages, as there was no significant relationship between model performance and model complexity.
+Increasing complexity in regeneration modules has been motivated by better process understanding and enhancing of model accuracy [@bugmann2022].
+However, more complex models do not necessarily lead to better projections and rather to a reduced transparency and lower predictive power [@franklin2020].
+Thus, the question of what level of detail is appropriate and parsimonious for modeling tree regeneration processes remains open [@konig_tree_2022; @bugmann2022].
 
@@
 
-Competition for light as a strong filter for tree regeneration has been
-widely documented [@messier1999; @collet2006; @berdanier2016], but the
-models examined here did not reproduce this expectation. However, it is
-difficult to measure light availability at large spatial and temporal
-scales. We used total stand basal area as a proxy for light availability
-[cf. @schmid2021]. However, we were unable to consider light
-availability restrictions caused by ground vegetation, which may be an
-important filter for some EuForIa stands [@woltjer2008]. We found
-pronounced differences in the stand basal area ranges simulated by the
-models, but also between models and observations. This made it
-impossible to evaluate the regeneration for the extremes of the stand
-density ranges in some models. For example, regeneration levels at low
-stand densities are relevant to assess how well forests are recovering
-e.g. after gap creation due to disturbance [@seidl2022; @grubb1977]. At
-the level of the simulated one-hectare samples, average basal area is
-typically not very low as long as no larger disturbances occur, which
-was explicitly excluded in our simulation protocol. Towards the other
-end of the spectrum, i.e., with increasing stand basal area, it would be
-reasonable to expect that the regeneration of the different species
-would become sparser and drop out entirely under low-light conditions
-[@klopcic2012; @zell2019]. However, few models showed this trend, thus
-indicating that the relation between regeneration and light availability
-is not yet captured correctly in most models. Yet, several of the models
-that did feature an increase of regeneration with increasing basal area
-include a feedback between seed production of mature trees and
-regeneration. Thus, it seems that higher light competition does not
-sufficiently compensate for increased seed availability with higher
-basal area due to a higher abundance of mature trees.
+Competition for light as a strong filter for tree regeneration has been widely documented [@messier1999; @collet2006; @berdanier2016], but the models examined here did not reproduce this expectation.
+However, it is difficult to measure light availability at large spatial and temporal scales.
+We used total stand basal area as a proxy for light availability [cf.
+@schmid2021].
+However, we were unable to consider light availability restrictions caused by ground vegetation, which may be an important filter for some EuForIa stands [@woltjer2008].
+We found pronounced differences in the stand basal area ranges simulated by the models, but also between models and observations.
+This made it impossible to evaluate the regeneration for the extremes of the stand density ranges in some models.
+For example, regeneration levels at low stand densities are relevant to assess how well forests are recovering e.g.
+after gap creation due to disturbance [@seidl2022; @grubb1977].
+At the level of the simulated one-hectare samples, average basal area is typically not very low as long as no larger disturbances occur, which was explicitly excluded in our simulation protocol.
+Towards the other end of the spectrum, i.e., with increasing stand basal area, it would be reasonable to expect that the regeneration of the different species would become sparser and drop out entirely under low-light conditions [@klopcic2012; @zell2019].
+However, few models showed this trend, thus indicating that the relation between regeneration and light availability is not yet captured correctly in most models.
+Yet, several of the models that did feature an increase of regeneration with increasing basal area include a feedback between seed production of mature trees and regeneration.
+Thus, it seems that higher light competition does not sufficiently compensate for increased seed availability with higher basal area due to a higher abundance of mature trees.
 
-There were pronounced differences how the main tree species were
-represented by the models along the environmental gradients, in
-particular the dominant species *Abies alba*, *Fagus sylvatica* and
-*Picea abies*. The regeneration levels were sampled from simulations in
-the equilibrium, and in this state it is expected that non-climax
-species such as *Pinus sylvestris* or *Quercus* would be of minor
-importance, or absent [@klopcic2015]. Most models captured this low
-abundance, which is also found in the empirical data. Thus, although the
-broad patterns are matched by many models, improvements in the
-quantification of the regeneration niche of the species are needed, but
-this cannot be done in the absence of robust datasets across multiple
-stand development stages.
+There were pronounced differences how the main tree species were represented by the models along the environmental gradients, in particular the dominant species *Abies alba*, *Fagus sylvatica* and *Picea abies*.
+The regeneration levels were sampled from simulations in the equilibrium, and in this state it is expected that non-climax species such as *Pinus sylvestris* or *Quercus* would be of minor importance, or absent [@klopcic2015].
+Most models captured this low abundance, which is also found in the empirical data.
+Thus, although the broad patterns are matched by many models, improvements in the quantification of the regeneration niche of the species are needed, but this cannot be done in the absence of robust datasets across multiple stand development stages.
 
@@
 
-The EuFoRIa data as used here are unique and well suited to better
-understand tree regeneration. However, three aspects of these data may
-represent considerable limitations. First, we made a comparison of tree
-regeneration in an equilibrium state, but we cannot assess how close the
-forests included in the EuFoRIa dataset were to such a state. The data
-were collected in forest reserves where no management has taken place
-for long periods of time. This makes our assumption of an equilibrium
-between forest properties and environmental drivers more reasonable than
-it might appear at first sight. In an analysis of primeval forests at
-demographic equilibrium, @brzeziecki2021 found higher regeneration rates
-than the ones observed in the EuFoRia data. Thus, the overestimation of
-regeneration rates by the models may not be so problematic. Second, the
-data were collected for regeneration above a 7 cm threshold, thus
-limiting the assessment of tree regeneration to a holistic viewpoint.
-This constituted a hard limit based on which we can understand only some
-aspects of tree regeneration, which in its entirety often comprises a
-rather long period since seed production [@price2001]. In reality, many
-environmental constraints are acting on young trees [@käber2021] that we
-were unable to assess. Yet, regeneration data with lower thresholds are
-simply not available in a harmonized manner across large environmental
-gradients. Third, the empirical data were collected from rather small
-plots, while we sampled simulated regeneration levels from 1 ha areas,
-which may lead to an incorrect representation of space. Even though the
-strategy we adopted is not ideal, it represents a common challenge when
-harmonizing diverse data sources originating from varied sampling
-strategies [@portier2022]. It would have been extremely challenging for
-such a variety of models to follow a protocol where the spatial sampling
-size was different at each of the 200 sites, and it would have
-introduced additional uncertainty in the results. Ultimately, we need to
-acknowledge that the observed data represent just a snapshot of an
-extremely complex and stochastic reality. This may at least partly
-explain e.g. the lack of clear patterns in the simulation results along
-the gradients of climate water balance and degree-days.
+The EuFoRIa data as used here are unique and well suited to better understand tree regeneration.
+However, three aspects of these data may represent considerable limitations.
+First, we made a comparison of tree regeneration in an equilibrium state, but we cannot assess how close the forests included in the EuFoRIa dataset were to such a state.
+The data were collected in forest reserves where no management has taken place for long periods of time.
+This makes our assumption of an equilibrium between forest properties and environmental drivers more reasonable than it might appear at first sight.
+In an analysis of primeval forests at demographic equilibrium, @brzeziecki2021 found higher regeneration rates than the ones observed in the EuFoRia data.
+Thus, the overestimation of regeneration rates by the models may not be so problematic.
+Second, the data were collected for regeneration above a 7 cm threshold, thus limiting the assessment of tree regeneration to a holistic viewpoint.
+This constituted a hard limit based on which we can understand only some aspects of tree regeneration, which in its entirety often comprises a rather long period since seed production [@price2001].
+In reality, many environmental constraints are acting on young trees [@käber2021] that we were unable to assess.
+Yet, regeneration data with lower thresholds are simply not available in a harmonized manner across large environmental gradients.
+Third, the empirical data were collected from rather small plots, while we sampled simulated regeneration levels from 1 ha areas, which may lead to an incorrect representation of space.
+Even though the strategy we adopted is not ideal, it represents a common challenge when harmonizing diverse data sources originating from varied sampling strategies [@portier2022].
+It would have been extremely challenging for such a variety of models to follow a protocol where the spatial sampling size was different at each of the 200 sites, and it would have introduced additional uncertainty in the results.
+Ultimately, we need to acknowledge that the observed data represent just a snapshot of an extremely complex and stochastic reality.
+This may at least partly explain e.g.
+the lack of clear patterns in the simulation results along the gradients of climate water balance and degree-days.
 
-The design of our sampling protocol did not include spatial aspects such
-as seed dispersal or detailed soil data. While we considered a wide
-range of models of forest dynamics, from stand to global scales, the
-simulation setup was limited to 1 ha regeneration values. This lack of
-consideration of spatial scale is appropriate for stand-scale models,
-but it potentially puts the landscape models at a disadvantage, as they
-have been built to be accurate at the landscape level. Without spatial
-context, we are limiting regeneration to the seed influx from the stand
-itself, unless the model has a background seed input. Yet, the global
-models should not be at a disadvantage due to the limited spatial
-considerations [but cf. Snell et al. -@snell2018; or Lehsten et al.
--@lehsten2019], as they are usually lacking dispersal between cells and
-are based on a strong abstraction of horizontal space
-[@hanbury-brown2022]. Lastly, detailed data on soil conditions were not
-available from the observed data, and independent, admittedly rather
-coarse data for soil and climatic water balance had to be used instead.
-It is noteworthy that many models represent drought based on detailed
-indicators based e.g. on soil water holding capacity, which had to be
-derived from a rough soil quality measure. This may at least partially
-explain the unsatisfactory performance of many models along the drought
-axis (i.e., climatic water balance).
+The design of our sampling protocol did not include spatial aspects such as seed dispersal or detailed soil data.
+While we considered a wide range of models of forest dynamics, from stand to global scales, the simulation setup was limited to 1 ha regeneration values.
+This lack of consideration of spatial scale is appropriate for stand-scale models, but it potentially puts the landscape models at a disadvantage, as they have been built to be accurate at the landscape level.
+Without spatial context, we are limiting regeneration to the seed influx from the stand itself, unless the model has a background seed input.
+Yet, the global models should not be at a disadvantage due to the limited spatial considerations [but cf.
+Snell et al.
+-@snell2018; or Lehsten et al.
+-@lehsten2019], as they are usually lacking dispersal between cells and are based on a strong abstraction of horizontal space [@hanbury-brown2022].
+Lastly, detailed data on soil conditions were not available from the observed data, and independent, admittedly rather coarse data for soil and climatic water balance had to be used instead.
+It is noteworthy that many models represent drought based on detailed indicators based e.g.
+on soil water holding capacity, which had to be derived from a rough soil quality measure.
+This may at least partially explain the unsatisfactory performance of many models along the drought axis (i.e., climatic water balance).
 
@@
 
-With our study, we have demonstrated that models of forest dynamics need
-a focus on tree regeneration modules to make them more robust. It
-remains uncertain what level of detail is required to model regeneration
-processes, and this must be addressed in future research. We recommend
-that the improvement of the regeneration modules is implemented as
-additional features that can be traced back, as done here for the
-variants of ForClim and ForCEEPS, and that model complexity and
-structure must always be connected with modeling objectives
-[@albrich2020]. If it should be necessary to include more detail in the
-regeneration models, this will come with higher parameterisation
-efforts. This will most likely lead to lower generalization because the
-required data will have to be collected from specific locations, as
-currently there is no general, comprehensive regeneration dataset
-available.
+With our study, we have demonstrated that models of forest dynamics need a focus on tree regeneration modules to make them more robust.
+It remains uncertain what level of detail is required to model regeneration processes, and this must be addressed in future research.
+We recommend that the improvement of the regeneration modules is implemented as additional features that can be traced back, as done here for the variants of ForClim and ForCEEPS, and that model complexity and structure must always be connected with modeling objectives [@albrich2020].
+If it should be necessary to include more detail in the regeneration models, this will come with higher parameterisation efforts.
+This will most likely lead to lower generalization because the required data will have to be collected from specific locations, as currently there is no general, comprehensive regeneration dataset available.
 
-Therefore, we further recommend that more effort should be invested into
-collecting harmonized datasets on tree regeneration. We emphasize that
-datasets such as EuFoRIa are invaluable and should be expanded in both
-their spatial extent (e.g., towards boreal and Mediterranean conditions)
-as well as in time (e.g., continuing the monitoring into the future).
-Such data will allow for a better evaluation of forest models and help
-to reduce the uncertainty in their projections, which is crucial when
-they are used as tools for predicting e.g. the impacts of anthropogenic
-climate change.
+Therefore, we further recommend that more effort should be invested into collecting harmonized datasets on tree regeneration.
+We emphasize that datasets such as EuFoRIa are invaluable and should be expanded in both their spatial extent (e.g., towards boreal and Mediterranean conditions) as well as in time (e.g., continuing the monitoring into the future).
+Such data will allow for a better evaluation of forest models and help to reduce the uncertainty in their projections, which is crucial when they are used as tools for predicting e.g.
+the impacts of anthropogenic climate change.
 
-In the present study, we have considered tree regeneration in the
-equilibrium state only. It is equally important to understand how these
-models project tree regeneration after changes in forest structure by
-disturbances [@seidl2022], or under different management strategies
-[@lindner2000]. However, this will require an entirely different set of
-observed data, and potentially not all models of forest dynamics would
-be able to assess the relationship of these aspects on tree
-regeneration, e.g. due to the lack of disturbance or appropriate
-management modules.
+In the present study, we have considered tree regeneration in the equilibrium state only.
+It is equally important to understand how these models project tree regeneration after changes in forest structure by disturbances [@seidl2022], or under different management strategies [@lindner2000].
+However, this will require an entirely different set of observed data, and potentially not all models of forest dynamics would be able to assess the relationship of these aspects on tree regeneration, e.g.
+due to the lack of disturbance or appropriate management modules.
 
-We recommend to investigate in detail the implications of the current
-modeling strategies for tree regeneration and, ultimately, simulated
-forest stand structure. This applies particularly to the erroneous
-patterns of excess tree regeneration and later excess mortality, by
-focusing on a wider range of tree sizes and the related regeneration
-dynamics. We also recommend, especially for the landscape-level models,
-the inclusion of explicit spatial considerations regarding tree
-regeneration [@beckage2003]; this, however, is a serious challenge
-regarding appropriate datasets, as inventory data have wide coverage,
-but by definition do not allow for the assessment of spatial interaction
-effects.
+We recommend to investigate in detail the implications of the current modeling strategies for tree regeneration and, ultimately, simulated forest stand structure.
+This applies particularly to the erroneous patterns of excess tree regeneration and later excess mortality, by focusing on a wider range of tree sizes and the related regeneration dynamics.
+We also recommend, especially for the landscape-level models, the inclusion of explicit spatial considerations regarding tree regeneration [@beckage2003]; this, however, is a serious challenge regarding appropriate datasets, as inventory data have wide coverage, but by definition do not allow for the assessment of spatial interaction effects.
 
-Exercises like the one presented here, where the models are operated in
-"blind flight" mode, i.e., without the possibility of being tuned
-towards capturing the expected patterns, should be repeated. Such
-benchmarking exercises should next focus on aspects such as specific
-model traits and the ecological formulations of particular
-(sub-)processes, to better understand the implications of the
-assumptions on which the models are based. Furthermore, the inclusion of
-a wide range of models with different scales, type, population
-structures, tree regeneration modules and complexity will ensure a wider
-benefit to the whole modeling community.
+Exercises like the one presented here, where the models are operated in "blind flight" mode, i.e., without the possibility of being tuned towards capturing the expected patterns, should be repeated.
+Such benchmarking exercises should next focus on aspects such as specific model traits and the ecological formulations of particular (sub-)processes, to better understand the implications of the assumptions on which the models are based.
+Furthermore, the inclusion of a wide range of models with different scales, type, population structures, tree regeneration modules and complexity will ensure a wider benefit to the whole modeling community.
 
@@
 
-Models of forest dynamics are important tools in science and decision
-support, and the formulation of tree regeneration has strong
-implications for simulated forest properties. The 15 models and variants
-used here are facing similar challenges in their representation of tree
-regeneration: they generally overestimate tree regeneration levels, and
-the simulated regeneration niche is not always captured accurately as a
-function of biotic (light) and abiotic (temperature and moisture)
-factors.
+Models of forest dynamics are important tools in science and decision support, and the formulation of tree regeneration has strong implications for simulated forest properties.
+The 15 models and variants used here are facing similar challenges in their representation of tree regeneration: they generally overestimate tree regeneration levels, and the simulated regeneration niche is not always captured accurately as a function of biotic (light) and abiotic (temperature and moisture) factors.
 
-However, most models properly capture the diversity of the initial tree
-community, and differences between model formulations, e.g., the
-presence or absence of feedback from the adult trees, did not have a
-strong effect for capturing the species composition of regeneration.
+However, most models properly capture the diversity of the initial tree community, and differences between model formulations, e.g., the presence or absence of feedback from the adult trees, did not have a strong effect for capturing the species composition of regeneration.
 
-Regarding mortality in the early phase of tree life, many models that
-feature a particularly high overestimation of regeneration levels are
-compensating for this by a larger tree mortality. Often, this
-compensation is not sufficient to reduce the high regeneration levels to
-realistic values. Overall, there is no clear mortality pattern across
-all models.
+Regarding mortality in the early phase of tree life, many models that feature a particularly high overestimation of regeneration levels are compensating for this by a larger tree mortality.
+Often, this compensation is not sufficient to reduce the high regeneration levels to realistic values.
+Overall, there is no clear mortality pattern across all models.
 
-When capturing tree regeneration, the specific design decisions taken in
-the development of any model are more important for its behavior
-(accuracy) than scale (stand, landscape, global), modelling approach
-(empirical vs. process-based), and complexity. Having both empirical and
-process-based models in our set, the empirically-based models could have
-been expected to have a better performance, as they were calibrated with
-inventory data, but this was not the case. Similarly, higher model
-complexity does not represent an improvement for capturing tree
-regeneration.
+When capturing tree regeneration, the specific design decisions taken in the development of any model are more important for its behavior (accuracy) than scale (stand, landscape, global), modelling approach (empirical vs.
+process-based), and complexity.
+Having both empirical and process-based models in our set, the empirically-based models could have been expected to have a better performance, as they were calibrated with inventory data, but this was not the case.
+Similarly, higher model complexity does not represent an improvement for capturing tree regeneration.
 
-Even though the regeneration routines of most of the models investigated
-here have never been sufficiently constrained by robust data, the
-regeneration projections are not overly off. This indicates that a lot
-can be gained by a focus on the modeling of regeneration processes. The
-representation of forest dynamics in these models would become much more
-robust particularly in the face of climate change and post-disturbance
-dynamics, thus strongly reducing the uncertainty in long-term
-projections of future forest dynamics.
+Even though the regeneration routines of most of the models investigated here have never been sufficiently constrained by robust data, the regeneration projections are not overly off.
+This indicates that a lot can be gained by a focus on the modeling of regeneration processes.
+The representation of forest dynamics in these models would become much more robust particularly in the face of climate change and post-disturbance dynamics, thus strongly reducing the uncertainty in long-term projections of future forest dynamics.
 
@@
 
-This study complies with the basic rules of reproducibility and has made
-available the protocol of the study, the data and all the software that
-has led to the findings presented in this paper. The protocol is
-available at @díaz-yáñez2022. The code and data that support the
-findings of this study are openly available at zenodo (*link to be
-added*) and github (*link to be added*) [for the review process we have
-provided these materials [on this
-link](https://polybox.ethz.ch/index.php/s/q5MOB1CLcjARs3H) (password:
-DownDataCodeRegen1)].
+This study complies with the basic rules of reproducibility and has made available the protocol of the study, the data and all the software that has led to the findings presented in this paper.
+The protocol is available at @díaz-yáñez2022.
+The code and data that support the findings of this study are openly available at zenodo (*link to be added*) and github (*link to be added*) [for the review process we have provided these materials [on this link](https://polybox.ethz.ch/index.php/s/q5MOB1CLcjARs3H) (password: DownDataCodeRegen1)].
 
@@
 
-This paper arises from a workshop held in Davos (Switzerland) in June
-2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
-Funding for the workshop was provided by COST Action CA19139 PROCLIAS
-(PROcess-based models for CLimate Impact Attribution across Sectors),
-supported by COST (European Cooperation in Science and Technology;
-[https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research
-Grant ETH-35 18-1). ODY was funded by the Swiss National Science
-Foundation project 200020_188882 / 1. KB received support from the
-European Research Council under the European Union's Horizon 2020
-research and innovation program (Grant Agreement 101001905). JB
-acknowledges funding from the long-term research development project no.
-RVO 67985939 from the Czech Academy of Sciences. MM acknowledges
-financial support from I-Maestro (Innovative forest management
-strategies for a resilient bioeconomy under climate change and
-disturbances, grant no. 22035418, 2019--2022) funded by the ERA-NET
-Cofund ForestValue. MMina acknowledges funding from the H2020-MSCA-IF
-project REINFORCE (Grant No. 891671). KM received funding from the
-Project No. CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant
-no. ITMS2014+313011W580 supported by the Integrated Infrastructure
-Operational Programme funded by the ERDF. CH acknowledges financial
-support from Waldfonds WASIM2100 project. We would also like to thank
-all the people that helped to make this study possible with
-contributions at different stages such as Martin Gutsch and Giorgio
-Vacchiano.
+This paper arises from a workshop held in Davos (Switzerland) in June 2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
+Funding for the workshop was provided by COST Action CA19139 PROCLIAS (PROcess-based models for CLimate Impact Attribution across Sectors), supported by COST (European Cooperation in Science and Technology; [https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research Grant ETH-35 18-1).
+ODY was funded by the Swiss National Science Foundation project 200020_188882 / 1.
+KB received support from the European Research Council under the European Union's Horizon 2020 research and innovation program (Grant Agreement 101001905).
+JB acknowledges funding from the long-term research development project no.
+RVO 67985939 from the Czech Academy of Sciences.
+MM acknowledges financial support from I-Maestro (Innovative forest management strategies for a resilient bioeconomy under climate change and disturbances, grant no.
+22035418, 2019--2022) funded by the ERA-NET Cofund ForestValue.
+MMina acknowledges funding from the H2020-MSCA-IF project REINFORCE (Grant No.
+891671).
+KM received funding from the Project No.
+CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant no.
+ITMS2014+313011W580 supported by the Integrated Infrastructure Operational Programme funded by the ERDF.
+CH acknowledges financial support from Waldfonds WASIM2100 project.
+We would also like to thank all the people that helped to make this study possible with contributions at different stages such as Martin Gutsch and Giorgio Vacchiano.
 

### 2023-05-15 manuscript/manuscriptRW.Rmd
remove date in manuscript

@@
     \newline
-    \today 
 output:

### 2023-05-15 manuscript/manuscriptRW.Rmd
data statement added

--- 2023-05-15 manuscript/manuscriptRW.Rmd
@@
 ---
-pretitle: "Journal: Ecological Monographs. Manuscript type: Article"
 title: "Tree regeneration in models of forest dynamics: a key priority for further research"
+subtitle: "Submitted to the journal Ecological Monographs. Manuscript type: Article"
 author: 
@@
 - Harald Bugmann$^1$
-date: $^*$Corresponding author (olalla.diaz@usys.ethz.ch) \newline
-      $^1$ETH Zurich. Department of Environmental Systems Science. Institute of Terrestrial Ecosystems. Universitätstrasse 16, 8092 Zürich, Switzerland. \newline
+date: $^1$ETH Zurich. Department of Environmental Systems Science. Institute of Terrestrial Ecosystems. Universitätstrasse 16, 8092 Zürich, Switzerland. \newline
       $^2$Senckenberg Biodiversity and Climate Research Centre (SBiK-F), 60325 Frankfurt am Main, Germany. \newline
@@
        $^{15}$CEFE, CNRS Campus du CNRS 1919, route de Mende 34293 Montpellier cedex, France. \newline 
-       $^{16}$Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland
-    \newline  
-    \newline
-    \today 
+       $^{16}$Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland.  \newline
+       $^*$Corresponding author (olalla.diaz@usys.ethz.ch)
 output:
@@
 ```
+
+OPEN RESEARCH STATEMENT: This submission uses novel code, both the data and code are provided in an external repository to be evaluated during the peer review process.
+The repository can be downloaded from [here](https://polybox.ethz.ch/index.php/s/q5MOB1CLcjARs3H) (password: DownDataCodeRegen1)].
+The protocol is available at @díaz-yáñez2022.
+We will make publicly available the code and data after the revision.
+
+**Keywords:** tree regeneration | tree recruitment | tree establishment | forest dynamics | models | landscape | stand | global
 
@@
 
-# Reproducibility and data availability statement
-
-This study complies with the basic rules of reproducibility and has made available the protocol of the study, the data and all the software that has led to the findings presented in this paper.
-The protocol is available at @díaz-yáñez2022.
-The code and data that support the findings of this study are openly available at zenodo (*link to be added*) and github (*link to be added*) [for the review process we have provided these materials [on this link](https://polybox.ethz.ch/index.php/s/q5MOB1CLcjARs3H) (password: DownDataCodeRegen1)].
+# Author contributions
+O.
+Díaz-Yáñez and Y.
+Käber contributed equally and share the first authorship.
+H.
+Bugmann, Y.
+Käber, and O.
+Díaz-Yáñez developed the study design.
+All authors contributed to preparing the simulations with their respective models.
+O.
+Díaz-Yáñez led the data analysis with contributions during the workshop from Y.Käber, T.
+Anders, K.
+H.
+Braziunas, J.
+Brůna, S.
+M.
+Fischer, J.
+Hetzer, T.
+Hickler, H.
+Lischke, M.Mahnken, P.
+Mairota, K.
+Merganičová, T.
+Mette, X.
+Morin, W.
+Rammer and D.
+Scherrer
+H. Bugmann. O. Díaz-Yáñez and H. Bugmann led the writing of the manuscript. All authors participated in the revision of the manuscript and approved its submission.
 
@@
 
+# Conflict of interest statement
+
+The authors declare no conflicts of interest.
+
 # References

### 2023-05-22 manuscript/supplementaryRW.Rmd

@@
 root mean square diameter) so that stem number N equals: N =
-sdi\*(dq/25)\^rke. A tree\'s resource acquisition area res (m2) is
+sdi\*(dq/25)\^rke. A tree's resource acquisition area res (m2) is
 approximated as res = 1e4/ (sdi\*(dbh/25)\^rke). As the Reineke exponent
@@
 increasing distance to the tree center. Next, regeneration matrices
-(1x1 m) are created for each of the stand\'s species. In the initial
+(1x1 m) are created for each of the stand's species. In the initial
 period, the regeneration matrix of each species is empty, but at the end
@@
 years and equal for all species. Dbh growth is modified by the light
-availability and the species\' regeneration light demand. This is a
+availability and the species' regeneration light demand. This is a
 decisive difference between the species and favours shadow-tolerant
@@
     Soria-Auza, R. W., \... & Kessler, M. (2017). Climatologies at high
-    resolution for the earth\'s land surface areas. Scientific data,
+    resolution for the earth's land surface areas. Scientific data,
     4(1), 1-20.

### 2023-05-22 manuscript/manuscriptRW.Rmd
updates from EcolMod prep

@@
 ---
-title: "Tree regeneration in models of forest dynamics: a key priority for further research"
-subtitle: "Submitted to the journal Ecological Monographs. Manuscript type: Article"
+subtitle: "\\fontsize{16pt}{3pt}\\selectfont \\textbf{\\textit{Tree regeneration in models of forest dynamics: a key priority for further research}} "
+title: "\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to the journal Ecological Monographs. Manuscript type: Article}}"
 author: 
@@
     extra_dependencies: "subfig"
+    latex_engine: xelatex
   word_document:
-    reference_docx: template_paper.docx
-mainfont: Arial
+    reference_docx: template_paper.docxl
 fontsize: 12pt
-linestretch: 1.5
+linestretch: 2
 bibliography: Regeneration.bib
@@
 linkcolor: blue
-csl: global-change-biology.csl
+csl: chicago-author-date.csl
 editor_options:
@@
 - \usepackage{titling}
-- \predate{\begin{flushleft}}
-- \postdate{\end{flushleft}}
+- \predate{ }
+- \postdate{  }
+- \usepackage{fontspec}
+- \setmainfont{Times New Roman}
 ---
@@
 
-OPEN RESEARCH STATEMENT: This submission uses novel code, both the data and code are provided in an external repository to be evaluated during the peer review process.
-The repository can be downloaded from [here](https://polybox.ethz.ch/index.php/s/q5MOB1CLcjARs3H) (password: DownDataCodeRegen1)].
-The protocol is available at @díaz-yáñez2022.
-We will make publicly available the code and data after the revision.
-
-**Keywords:** tree regeneration | tree recruitment | tree establishment | forest dynamics | models | landscape | stand | global
+OPEN RESEARCH STATEMENT: This submission uses novel code, both the data
+and code are provided in an external repository to be evaluated during
+the peer review process. The repository can be downloaded from
+[here](https://polybox.ethz.ch/index.php/s/q5MOB1CLcjARs3H) (password:
+DownDataCodeRegen1). The protocol is available at @díaz-yáñez2022. We
+will make publicly available the code and data after the revision.
+
+**Keywords:** tree regeneration \| tree recruitment \| tree
+establishment \| forest dynamics \| models \| landscape \| stand \|
+global
 
@@
 
-Tree regeneration is a key process in forest dynamics, particularly in the context of forest resilience and climate change.
-Models are pivotal for assessing long-term forest dynamics, and they have been in use for more than 50 years.
-However, an assessment of their ability to accurately represent tree regeneration is lacking.
-We assess how well current models capture the overall abundance, species composition, and mortality of tree regeneration.
-
-Using 15 models built to capture long-term forest dynamics at the stand, landscape, and global levels, we simulate tree regeneration at 200 sites representing large environmental gradients across Central Europe.
-The results are evaluated against comprehensive data from unmanaged forests.
-
-Most of the models overestimate regeneration levels, which is only compensated in some models by high simulated mortality rates in the early stages of individual trees dynamics.
-Simulated species diversity of regeneration matches the observed ranges.
-Models simulating higher species diversity at the stand level do not feature higher regeneration diversity.
-The effect of light availability on regeneration levels is captured better than the effect of temperature and soil moisture, but patterns are not consistent across models.
-Increasing complexity in the tree regeneration modules of the models is not related to higher accuracy of simulated tree regeneration.
-Furthermore, individual model design is more important than scale (stand, landscape, global) and approach (empirical, process-based) for accurately capturing tree regeneration.
-
-Despite considerable mismatches between simulation results and data, it is remarkable that most models capture the essential features of the highly complex process of tree regeneration, while not having been parameterized with such data.
-We conclude that much can be gained by evaluating and refining the modeling of regeneration processes.
-This has the potential to render long-term projections of forest dynamics under changing environmental conditions that are much more robust.
+Tree regeneration is a key process in forest dynamics, particularly in
+the context of forest resilience and climate change. Models are pivotal
+for assessing long-term forest dynamics, and they have been in use for
+more than 50 years. However, an assessment of their ability to
+accurately represent tree regeneration is lacking. We assess how well
+current models capture the overall abundance, species composition, and
+mortality of tree regeneration.
+
+Using 15 models built to capture long-term forest dynamics at the stand,
+landscape, and global levels, we simulate tree regeneration at 200 sites
+representing large environmental gradients across Central Europe. The
+results are evaluated against comprehensive data from unmanaged forests.
+
+Most of the models overestimate regeneration levels, which is only
+compensated in some models by high simulated mortality rates in the
+early stages of individual trees dynamics. Simulated species diversity
+of regeneration matches the observed ranges. Models simulating higher
+species diversity at the stand level do not feature higher regeneration
+diversity. The effect of light availability on regeneration levels is
+captured better than the effect of temperature and soil moisture, but
+patterns are not consistent across models. Increasing complexity in the
+tree regeneration modules of the models is not related to higher
+accuracy of simulated tree regeneration. Furthermore, individual model
+design is more important than scale (stand, landscape, global) and
+approach (empirical, process-based) for accurately capturing tree
+regeneration.
+
+Despite considerable mismatches between simulation results and data, it
+is remarkable that most models capture the essential features of the
+highly complex process of tree regeneration, while not having been
+parameterized with such data. We conclude that much can be gained by
+evaluating and refining the modeling of regeneration processes. This has
+the potential to render long-term projections of forest dynamics under
+changing environmental conditions that are much more robust.
 
@@
 
-Forests provide a wide range of products and services of vital importance to humankind [@foresteurope2020].
-Under the pressure of climate change, increasing disturbance impacts and changing societal demands on forest ecosystem services, it is becoming ever more important to understand how forest structure, composition and function will change, and to evaluate forest capacity to adapt to or be resilient after disturbance [@lindner2010; @seidl2022].
-A wide range of models of forest dynamics were developed over the past decades considering the impacts of climate [@bugmann2022].
-From these studies, it is evident that we have a reasonably good understanding how to model tree growth [@vanclay1997; @bugmann1996], and substantial efforts have been dedicated to improving the representation of tree mortality [@bugmann2019; @cailleret2017].
-In contrast, tree regeneration is a much less studied process that is often represented rather coarsely in models [@walck2011; @leishman1992; @price2001].
-Which is a clear research gap in the context of climate-induced forest disturbances and forest resilience.
-
-Tree regeneration arises from multiple processes such as pollination, fruit maturation, seed production, dispersal, germination, juvenile growth and survival [@price2001; @vacchiano2018].
-All these processes are difficult to assess, and some of them are scarcely understood and thus appear highly stochastic [@bogdziewicz2021].
-Identifying the appropriate level of complexity for the mathematical formulation of the key factors that are leading to successful tree regeneration is challenging.
-Currently, tree regeneration processes in dynamic forest models are handled in a multitude of ways [@konig_tree_2022; @bugmann2022]: from 1) entirely ignoring it [as done in classical forest growth models, e.g., Pretzsch et al.
--@pretzsch2002], across 2) the use of a few simple environmental filters, as done in most forest "gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g., @hickler2012; @smith2001], to 3) complex approaches that incorporate local feedback from the canopy, multiple ecological processes and often also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based statistical parameterizations, which however are not easy to extrapolate in space and time [e.g., @ribbens1994].
-
-Overall, models are needed to (1) synthesize existing empirical data and explore their relationships, (2) assess future tree regeneration e.g.
-in the context of global change scenarios, and (3) identify the most important processes that are shaping ecological patterns.
-Given the current strategies that are used in models of forest dynamics to represent tree regeneration, model behavior often is prone to problems, such as very high levels of tree regeneration that necessitate excess mortality at early stages of tree life to simulate correct stand structure and composition.
-Also, inadequately high species diversity in tree regeneration may be simulated, which is characteristic of "classical" forest gap models [@gutiérrez2016], at least as long as the simulation set-up comprises a multitude of species.
-Some models use calibration against local canopy-level data to constrain simulated regeneration levels, which is likely to hamper the general applicability of these models e.g.
-under scenarios of climate change.
-Furthermore, correctly capturing the species composition in tree regeneration as a function of the presence of seed trees in the canopy is often a particular challenge, potentially leading to unrealistic successional drift in the model, which must be corrected by factors that are hard to parameterize [e.g., @lischke2006].
-A related issue is the excessive reduction of species diversity due to positive feedback effects, such that eventually just single-species stands remain.
-This is sometimes corrected by the incorporation of a low level of seed influx of all species at all times [@schumacher2006] or by restricting the number of seeds per species in the seedbank [@lischke2006].
-However, simulated species composition is usually exceedingly sensitive to assumptions about seed availability, whereas the parameters of such functions are poorly constrained by field data.
-Lastly, there is often a problem with insufficient observational constraints on parameter values for models that start from very small tree sizes (e.g., 10 cm of height or even seeds) and track tree development in a process-oriented manner by considering a multitude of ecological influences, rather than emphasizing tree regeneration into a larger size class [e.g., @wehrli2006].
-
-Thus, a focus on the modeling of tree regeneration processes is sorely needed and overdue [@price2001; @walck2011] if we are to make reliable projections of future forest dynamics, i.e., when the models need to be operated in extrapolation mode, as well as from a fundamental ecological point of view for increasing systems understanding.
-Also in empirical ecology, tree regeneration has not received a lot of attention [@hanbury-brown2022], although many sub-processes have been studied in detail [@miina2006].
-Yet, little data are available that covers all the processes within one species along environmental gradients, let alone for a vast suite of species.
-Thus, the available data are fragmented, which constitutes a major problem for model building [@clark1999].
-On the one hand, monitoring on permanent plots (such as National Forest Inventories) often measures tree regeneration [@zell2019], however with a design that captures the rate after the stand initiation phase [@hallsby2015].
-Data from permanent plots, although available across regions, are highly heterogeneous, rendering their use difficult in a modeling context [@käber2021].
-On the other hand, targeted studies to measure tree regeneration on experimental sites are often limited in spatial extent, species studied, or the sub-set of processes that are targeted [@berdanier2016; @collet2006].
-
-In the present study, for the first time, a large number of forest models commonly used to assess forest dynamics under climate change are evaluated against a continental-scale, multi-species harmonized dataset on tree regeneration [@käber2023].
-By tree regeneration, we refer to the passing of trees across a specific diameter threshold ("ingrowth").
-We included models that are based on a range of "philosophies", from models operating at the stand to the global scale as well as the range of models from empirically derived to "process-based" [e.g., @bugmann1996; @fabrika2012; @lexer2001; @reyer2014].
-Due to the large variability in tree regeneration patterns in nature and the large number of factors driving this process --- including some that are not incorporated explicitly in most models, such as deer browsing --- we do not aim for a detailed statistical evaluation of each model.
-Instead, we aim to evaluate the general regeneration patterns and magnitudes simulated by the models, and benchmark the simulated regeneration niche of multiple species against empirical data along a wide environmental gradient of temperature, moisture, and light availability [@grubb1977].
-
-More specifically, we aim to answer the following questions: (1) Are models of forest dynamics capturing accurately tree regeneration levels, initial tree species diversity and mortality in the regeneration?
-(2) Do model traits explain differences in model performance?
-(3) How well do the models capture total regeneration and individual species regeneration niches across environmental gradients of light availability, temperature, and soil moisture?
+Forests provide a wide range of products and services of vital
+importance to humankind [@foresteurope2020]. Under the pressure of
+climate change, increasing disturbance impacts and changing societal
+demands on forest ecosystem services, it is becoming ever more important
+to understand how forest structure, composition and function will
+change, and to evaluate forest capacity to adapt to or be resilient
+after disturbance [@lindner2010; @seidl2022]. A wide range of models of
+forest dynamics were developed over the past decades considering the
+impacts of climate [@bugmann2022]. From these studies, it is evident
+that we have a reasonably good understanding how to model tree growth
+[@vanclay1997; @bugmann1996], and substantial efforts have been
+dedicated to improving the representation of tree mortality
+[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is a much
+less studied process that is often represented rather coarsely in models
+[@walck2011; @leishman1992; @price2001]. Which is a clear research gap
+in the context of climate-induced forest disturbances and forest
+resilience.
+
+Tree regeneration arises from multiple processes such as pollination,
+fruit maturation, seed production, dispersal, germination, juvenile
+growth and survival [@price2001; @vacchiano2018]. All these processes
+are difficult to assess, and some of them are scarcely understood and
+thus appear highly stochastic [@bogdziewicz2021]. Identifying the
+appropriate level of complexity for the mathematical formulation of the
+key factors that are leading to successful tree regeneration is
+challenging. Currently, tree regeneration processes in dynamic forest
+models are handled in a multitude of ways [@konig_tree_2022;
+@bugmann2022]: from 1) entirely ignoring it [as done in classical forest
+growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2) the use
+of a few simple environmental filters, as done in most forest "gap"
+models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
+@hickler2012; @smith2001], to 3) complex approaches that incorporate
+local feedback from the canopy, multiple ecological processes and often
+also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
+statistical parameterizations, which however are not easy to extrapolate
+in space and time [e.g., @ribbens1994].
+
+Overall, models are needed to (1) synthesize existing empirical data and
+explore their relationships, (2) assess future tree regeneration e.g. in
+the context of global change scenarios, and (3) identify the most
+important processes that are shaping ecological patterns. Given the
+current strategies that are used in models of forest dynamics to
+represent tree regeneration, model behavior often is prone to problems,
+such as very high levels of tree regeneration that necessitate excess
+mortality at early stages of tree life to simulate correct stand
+structure and composition. Also, inadequately high species diversity in
+tree regeneration may be simulated, which is characteristic of
+"classical" forest gap models [@gutiérrez2016], at least as long as the
+simulation set-up comprises a multitude of species. Some models use
+calibration against local canopy-level data to constrain simulated
+regeneration levels, which is likely to hamper the general applicability
+of these models e.g. under scenarios of climate change. Furthermore,
+correctly capturing the species composition in tree regeneration as a
+function of the presence of seed trees in the canopy is often a
+particular challenge, potentially leading to unrealistic successional
+drift in the model, which must be corrected by factors that are hard to
+parameterize [e.g., @lischke2006]. A related issue is the excessive
+reduction of species diversity due to positive feedback effects, such
+that eventually just single-species stands remain. This is sometimes
+corrected by the incorporation of a low level of seed influx of all
+species at all times [@schumacher2006] or by restricting the number of
+seeds per species in the seedbank [@lischke2006]. However, simulated
+species composition is usually exceedingly sensitive to assumptions
+about seed availability, whereas the parameters of such functions are
+poorly constrained by field data. Lastly, there is often a problem with
+insufficient observational constraints on parameter values for models
+that start from very small tree sizes (e.g., 10 cm of height or even
+seeds) and track tree development in a process-oriented manner by
+considering a multitude of ecological influences, rather than
+emphasizing tree regeneration into a larger size class [e.g.,
+@wehrli2006].
+
+Thus, a focus on the modeling of tree regeneration processes is sorely
+needed and overdue [@price2001; @walck2011] if we are to make reliable
+projections of future forest dynamics, i.e., when the models need to be
+operated in extrapolation mode, as well as from a fundamental ecological
+point of view for increasing systems understanding. Also in empirical
+ecology, tree regeneration has not received a lot of attention
+[@hanbury-brown2022], although many sub-processes have been studied in
+detail [@miina2006]. Yet, little data are available that covers all the
+processes within one species along environmental gradients, let alone
+for a vast suite of species. Thus, the available data are fragmented,
+which constitutes a major problem for model building [@clark1999]. On
+the one hand, monitoring on permanent plots (such as National Forest
+Inventories) often measures tree regeneration [@zell2019], however with
+a design that captures the rate after the stand initiation phase
+[@hallsby2015]. Data from permanent plots, although available across
+regions, are highly heterogeneous, rendering their use difficult in a
+modeling context [@käber2021]. On the other hand, targeted studies to
+measure tree regeneration on experimental sites are often limited in
+spatial extent, species studied, or the sub-set of processes that are
+targeted [@berdanier2016; @collet2006].
+
+In the present study, for the first time, a large number of forest
+models commonly used to assess forest dynamics under climate change are
+evaluated against a continental-scale, multi-species harmonized dataset
+on tree regeneration [@käber2023]. By tree regeneration, we refer to the
+passing of trees across a specific diameter threshold ("ingrowth"). We
+included models that are based on a range of "philosophies", from models
+operating at the stand to the global scale as well as the range of
+models from empirically derived to "process-based" [e.g., @bugmann1996;
+@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
+tree regeneration patterns in nature and the large number of factors
+driving this process --- including some that are not incorporated
+explicitly in most models, such as deer browsing --- we do not aim for a
+detailed statistical evaluation of each model. Instead, we aim to
+evaluate the general regeneration patterns and magnitudes simulated by
+the models, and benchmark the simulated regeneration niche of multiple
+species against empirical data along a wide environmental gradient of
+temperature, moisture, and light availability [@grubb1977].
+
+More specifically, we aim to answer the following questions: (1) Are
+models of forest dynamics capturing accurately tree regeneration levels,
+initial tree species diversity and mortality in the regeneration? (2) Do
+model traits explain differences in model performance? (3) How well do
+the models capture total regeneration and individual species
+regeneration niches across environmental gradients of light
+availability, temperature, and soil moisture?
 
@@
 
-Fifteen models of forest dynamics (including two models featuring model variants) were used to simulate forest dynamics at the stand, landscape or global scale (Table \@ref(tab:tableModels)).
-The approaches used for model construction and their origin differ strongly, with most of the models featuring a largely "process-based" approach, whereas two models are based on formulations derived from the statistical analysis of inventory data (SIBYLA and xComp).
-Some of the models largely rely on the original approach underlying forest gap models (e.g., ForClim 1), whereas some are based on plant ecophysiological processes (e.g., FORMIND and iLand).
-The different approaches underlying the models have strongly influenced the formulation of tree regeneration processes.
-
-The regeneration models can be differentiated into 'regeneration' and 'recruitment' approaches [@vanclay1994; @konig_tree_2022].
-*Regeneration models* include processes such as flowering and pollination, seed production, seed dispersal, germination, and seedling growth, which ultimately lead to the simulated number of established trees.
-*Recruitment models*, in contrast, introduce a number of new trees with certain characteristics such as biomass or diameter, without explicitly considering earlier development processes.
-We can further distinguish models that feature a feedback in the regeneration, i.e., where the level and species composition of regeneration is influenced by the existing adult tree community via the production of seeds, seedlings or saplings, from models that do not contain such a feedback.
-
-The starting point for tree regeneration in the models differs as well, ranging from seedbank, seed or seedling to sapling (i.e., trees much larger than 10 cm height, often ca.
-2-3 m tall).
-Models that start from seed need to include a larger number of ecological processes such as germination and survival of young seedlings, whereas models that start from saplings have to aggregate via parameterizations several ecological processes that are not treated explicitly.
-This latter approach reduces model complexity but comes at the cost of blurring process representation.
-
-In the models used here, the overall complexity in the regeneration modules varies considerably.
-Following @bugmann2022, we can classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and TreeMig) as having rather high complexity in their regeneration modules (mean regeneration formulation complexity across all processes \>0, Table \@ref(tab:tableModels)).
-Two models, ForClim and ForCEEPS, were used here with two alternative variants of regeneration.
-ForClim variant 1 [@bugmann1996] is based on a recruitment module that adheres closely to the concept introduced by @botkin1972, whereas ForClim variant 11 is adopting a different approach where individual species properties and their relation to the environment are important [@huber2020].
-These two model variants allow us to evaluate a more process-based and complex module (variant 11) against a simple module (variant 1), while the rest of the model structure is identical.
-Similarly, the two ForCEEPS variants allow us to isolate the importance of the canopy feedback (i.e., simulated actual composition and relative abundance of species in the plot) via seed trees for the quantity and quality (e.g., diversity and composition) of simulated regeneration, as one variant includes this feedback whereas the other does not.
+Fifteen models of forest dynamics (including two models featuring model
+variants) were used to simulate forest dynamics at the stand, landscape
+or global scale (Table \@ref(tab:tableModels)). The approaches used for
+model construction and their origin differ strongly, with most of the
+models featuring a largely "process-based" approach, whereas two models
+are based on formulations derived from the statistical analysis of
+inventory data (SIBYLA and xComp). Some of the models largely rely on
+the original approach underlying forest gap models (e.g., ForClim 1),
+whereas some are based on plant ecophysiological processes (e.g.,
+FORMIND and iLand). The different approaches underlying the models have
+strongly influenced the formulation of tree regeneration processes.
+
+The regeneration models can be differentiated into 'regeneration' and
+'recruitment' approaches [@vanclay1994; @konig_tree_2022]. *Regeneration
+models* include processes such as flowering and pollination, seed
+production, seed dispersal, germination, and seedling growth, which
+ultimately lead to the simulated number of established trees.
+*Recruitment models*, in contrast, introduce a number of new trees with
+certain characteristics such as biomass or diameter, without explicitly
+considering earlier development processes. We can further distinguish
+models that feature a feedback in the regeneration, i.e., where the
+level and species composition of regeneration is influenced by the
+existing adult tree community via the production of seeds, seedlings or
+saplings, from models that do not contain such a feedback.
+
+The starting point for tree regeneration in the models differs as well,
+ranging from seedbank, seed or seedling to sapling (i.e., trees much
+larger than 10 cm height, often ca. 2-3 m tall). Models that start from
+seed need to include a larger number of ecological processes such as
+germination and survival of young seedlings, whereas models that start
+from saplings have to aggregate via parameterizations several ecological
+processes that are not treated explicitly. This latter approach reduces
+model complexity but comes at the cost of blurring process
+representation.
+
+In the models used here, the overall complexity in the regeneration
+modules varies considerably. Following @bugmann2022, we can classify
+seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and
+TreeMig) as having rather high complexity in their regeneration modules
+(mean regeneration formulation complexity across all processes \>0,
+Table \@ref(tab:tableModels)). Two models, ForClim and ForCEEPS, were
+used here with two alternative variants of regeneration. ForClim variant
+1 [@bugmann1996] is based on a recruitment module that adheres closely
+to the concept introduced by @botkin1972, whereas ForClim variant 11 is
+adopting a different approach where individual species properties and
+their relation to the environment are important [@huber2020]. These two
+model variants allow us to evaluate a more process-based and complex
+module (variant 11) against a simple module (variant 1), while the rest
+of the model structure is identical. Similarly, the two ForCEEPS
+variants allow us to isolate the importance of the canopy feedback
+(i.e., simulated actual composition and relative abundance of species in
+the plot) via seed trees for the quantity and quality (e.g., diversity
+and composition) of simulated regeneration, as one variant includes this
+feedback whereas the other does not.
+
+## Observed data
+
+Regeneration data covering a wide range of environmental conditions are
+hard to obtain, and this is one of the reasons why most models of forest
+dynamics have never been confronted with a dataset covering such
+gradients over a large number of sites, to evaluate how well
+regeneration is captured. The observations used here are derived from a
+novel and unprecedented network of sites in forest reserves that
+represent the range of environmental gradients in temperature and
+precipitation in Central Europe as compiled in the framework of the
+EuFoRIa network [@euforia2019] (Figure \@ref(fig:map)). These forests
+have been unmanaged for at least 10 years prior to their designation as
+reserves, and most of them provide time series of natural forest
+dynamics over multiple decades and up to ca. 80 years. The census
+periods range from 3 to 37 years, with an average of 14 years. The data
+provide information at the tree level, thus allowing for the sequential
+comparison of processes such as individual-tree regeneration and death
+between the measurements.
+
+We selected 200 sites from this network as the benchmarking dataset for
+the simulation to be representative of the environmental variation
+contained in the data. This was achieved by applying k-means clustering
+to define 200 clusters of plots from the original set of 869 plots along
+the environmental dimensions of temperature, climatic water balance,
+soil quality and exposition (i.e., slope and aspect). Each of these
+sites featured at least two consecutive measurements. Regeneration
+thresholds for these sites differed between diameters of 0 and 10 cm.
+For our study, we defined two datasets where one included 165 sites with
+a diameter threshold of 7 cm or lower, and the other included another 35
+sites with diameter thresholds between 7 and 10 cm.
+
+Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
+processed and aggregated following @käber2023 to provide regeneration
+rates per ha and per decade. The observations used in this study
+featured 30,900 newly established trees. Regeneration rates per site,
+sample and decade ranged from zero to 1246 trees, with a mean of 56
+trees. Adult species composition was also available for each plot. For
+more details on this unique dataset and the detailed regeneration
+information, cf. @käber2023.
+
+## Simulation protocol
+
+The overarching goal of the experiments was to assess the tree
+regeneration as it arises from empirical data against its representation
+in a wide range of models of forest dynamics. We define tree
+regeneration as the passing of a breast height diameter threshold of 7
+or 10 cm, respectively (synonym: ingrowth). To this end, each modeling
+group was provided with a detailed protocol [@díaz-yáñez2022] with
+instructions how to perform the simulations, providing input variables
+on climate and soil conditions, and the list of expected output
+variables. Neither were further site information (except for the data
+specified below) nor any data on tree regeneration or forest stand
+features provided prior to the simulation. That is, the models were run
+in "blind flight" mode.
+
+The input variables were collected from different data sources and
+aggregated to be adapted to the needs of the different models. Time
+series of climatic variables from 1981 to 2018 were provided in hourly
+[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
+resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
+required for some models, such as relative humidity or vapor pressure
+deficit, were calculated from these variables. The final instructions
+for using the climate data were slightly different in each model, based
+on the approach that best suited the model (Table
+\@ref(tab:tableModels)). Soil quality data were provided as continuous
+values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
+The protocol also provided the elevation, slope and aspect for each of
+the 200 sites [@asterscienceteam2019], but no other spatial information
+such as coordinates, with the exception of iLand and aDGVM2, which
+required blurred coordinates to derive highly detailed soil data.
+
+The simulations were run in the absence of natural disturbances. Only
+the model LPJ-GUESS had to include a background disturbance to increase
+chances of shade-intolerant species to establish. The simulations were
+set up to sample species-specific regeneration rates per decade and per
+ha in the equilibrium state of the model, typically entailing a
+"spin-up" run (as we did not provide any forest data). The modeling
+teams decided on the simulated area and how they derived these samples.
+The exact length of the simulation was also decided by the modeling
+teams (Table \@ref(tab:tableModels)). Further details on how each
+modeling team prepared the simulations and the outputs are available in
+Supplementary Material 1.
+
+The simulations were run in the absence of management to a simulated
+equilibrium ("Potential Natural Vegetation") with the current climate,
+as the ultimate goal was to evaluate tree regeneration under comparable
+and near-equilibrium conditions. This entails the assumption that (1)
+the observations from the forest reserves reflect no traces of forest
+management, and (2) there is an equilibrium between forest dynamics and
+climate. While the former might be starting to be visible in many of the
+EuFoRIA reserves, the latter may be more debatable. However, in the
+absence of detailed data on the history of each plot in the EuFoRIa
+network, some broad assumptions had to be made. Both the width of the
+regeneration niche (i.e., in environmental space) as well as the
+intensity of the regeneration process (i.e., the number of ingrowth
+trees per area and per unit of time) were of interest.
+
+The simulations were run for mixed-species forests (not multiple
+single-species simulations) using mixtures of eleven species or genera
+for which regeneration data of sufficient quality were available from
+EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
+betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
+Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus
+sylvestris L.* The same set of species was used at all 200 sites. Two
+models included their standard set of species for the simulations, which
+is much larger (i.e., ForClim 1, ForClim 11, TreeMig). In three models,
+fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS)
+(Table \@ref(tab:tableModels)). In 4C only *Fagus sylvatica*, *Picea
+abies*, *Betula* spp., *Quercus* spp. and *Pinus sylvestris* are
+parameterised. xComp simulations did not consider *Fraxinus excelsior*
+due to a depreciated species parameterisation. In LPJ-GUESS, *Acer
+pseudoplatanus* and *Alnus glutinosa* are not parameterized and
+therefore these species could not be included in the simulation.
+Finally, the model aDGVM2 does not simulate individual species. Rather,
+community assembly processes and trait filtering generate plant
+communities that are adjusted to the biotic and abiotic conditions, and
+the simulated plants can be classified into ecological strategies based
+on their trait values in a post-processing step.
+
+Each of the models reported the regeneration number by sampling 200
+times in a 10-year interval for each species and per ha for each of the
+200 sites. Multiple samples per site were used to better understand the
+simulated variation within each site. This was done using different
+strategies, depending on the model, including (1) sampling simulated
+data from the same 1 ha plot in the equilibrium over time, (2) sampling
+several 1 ha plots from the simulated forest at one specific point in
+time (in the equilibrium) or (3) a combination of (1) and (2). This
+resulted in 880,000 observations per model that simulated the 11 species
+included in the protocol (200 sites, 200 samples per site, 11 species
+and two diameter thresholds). For the models that simulated additional
+species, their regeneration rates were aggregated as "others".
+
+Two models did not provide results from all the simulated samples or
+sites to avoid unrealistic results, as follows. In the model 4C, a
+threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid
+unrealistic stand basal area data, and therefore not all the samples and
+sites were reported. The reason for this is that 4C is not suitable for
+long-term simulations without management, due to misrepresentations in
+density-dependent mortality processes in long-term simulations and
+assumptions of tree geometry that lead to unrealistic single tree
+dimensions of very old individuals. LPJ-GUESS had 2% of the sample
+outputs without tree regeneration, and these were considered as zero
+stand basal area and zero regeneration for all the tree species
+simulated; one site produced grassland rather than a forest and was not
+included in the results.
+
+## Data analysis
+
+The simulation results were analyzed regarding (1) regeneration levels
+(i.e., ingrowth number per unit time and space), (2) regeneration
+species diversity, (3) regeneration mortality, (4) the relationship
+between model performance and model traits, and (5) ingrowth gradients
+along the regeneration niches. We evaluated species diversity across the
+models and in relation to the observed data by calculating the Shannon
+index $H$ based on the relative proportion of the species in terms of
+basal area. It was calculated for the regeneration ($H_{R_{n}}$)
+(Equation \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$)
+(Equation \ref{eqpiShannon}). The higher the value of the index, the
+higher the species diversity at a particular site and sample.
+
+```{=tex}
+\begin{equation}
+H_{R_{n}} = - \sum_{i = 1}^{s}{p_{R_{n,i}}\ln(p_{R_{n,i}})}\label{eqpiShannonR}
+\end{equation}
+```
+```{=tex}
+\begin{equation}
+p_{R_{n,i}} = \frac{r_{BA_{n,i}}}{TotalR_{BA_{n}}} \label{eqpiShannonR2}
+\end{equation}
+```
+```{=tex}
+\begin{equation}
+H_{S_{n}} = - \sum_{i = 1}^{s}{p_{S_{n,i}}\ln(p_{S_{n,i}})}\label{eqpiShannon}
+\end{equation}
+```
+```{=tex}
+\begin{equation}
+p_{S_{n,i}} = \frac{BA_{n,i}}{Total_{BA_{n}}} \label{eqpiShannon2}
+\end{equation}
+```
+where $s$ is the total number of species present and that have a basal
+area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
+species $i$ in sample $n$ calculated as the regeneration basal area
+($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal
+area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
+proportion of species $i$ calculated as the basal area of all trees
+($BA_{n,i}$) of that species $i$ relative to total basal area
+($TotalR_{BA_{n}}$) of the sample $n$. Species diversity was not
+assessed for aDGVM2 as this model does not simulate individual species.
+
+Mortality in tree regeneration was assessed based on the ratio of
+regeneration between the 7 and 10 cm diameter thresholds. We used the
+Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference
+to estimate whether the ratio of regeneration between the 7 and 10 cm
+diameter thresholds were above or below the expected theoretical rate.
+The Reineke self-thinning rule is usually calculated for even-aged,
+single species stands and assumes a fixed relationship between the
+number of stems and the quadratic mean diameter in fully stocked pure
+stands. The value used in our comparisons was 1.77 (i.e., we expect stem
+numbers at 7 cm to be 77% higher than at 10 cm), calculated using
+Equation \ref{eqReineke}.
+
+```{=tex}
+\begin{equation}
+Reineke = \frac{7}{10}^{-1.605} \label{eqReineke}
+\end{equation}
+```
+We assessed model performance in relation to model traits focusing on
+(1) model complexity as defined by @bugmann2022, (2) model type
+(empirical or process based), (3) the presence or absence of a canopy
+feedback for regeneration, and (4) the scale of application of the model
+(stand, landscape or global). We tested for significant differences
+using two-sided t-tests by considering each of these model features
+regarding the observed and simulated regeneration levels and species
+diversity, respectively.
+
+Total regeneration and the regeneration niches of the individual species
+were evaluated across the environmental gradients of light availability,
+temperature, and soil moisture. Stand basal area (the basal area of all
+the trees in each sample including tree regeneration) was used as a
+proxy for light availability at the forest floor, the annual degree-day
+sum [@allen1976; @fischlin1995] as a proxy for growing season warmth,
+and the climatic water balance as a proxy for soil moisture
+[@speich2019]. Regeneration values were calculated as the mean across
+the 200 samples per site. The observed data were modeled using a
+Generalized Additive Model [@wood2011] with a negative binomial
+distribution and restricted maximum likelihood to better understand the
+relationship between the environmental gradients and the levels of
+observed regeneration, relative to the simulation results per model.
+
+In order to analyze the regeneration niches across the climatic
+gradients, we focused on five common tree species or genera: *Abies
+alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
+*Quercus* spp. For these species, we calculated the share in the
+regeneration basal area per site as the mean across the available
+samples per site:
+
+```{=tex}
+\begin{equation}
+\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{BA_{t,i}}}{\sum_{i = 1}^{s}{\bar{r}_{{BA_{t,i}}}}} \label{eq:BAshare}
+\end{equation}
+```
+where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
+is the mean basal area in the regeneration (subsequently referred to as
+"regeneration basal area") of species $i$ per site $t$ across the
+available samples at that site. Furthermore, regarding the share in the
+regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we
+categorized this as zero when the regeneration basal area of that
+species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the
+regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) and the
+total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$)
+equaled zero.
+
+# Results
+
+## Regeneration levels
+
+Simulated regeneration levels varied strongly across the 15 models and
+typically did not match the levels found in the forest reserve data
+(Figure \@ref(fig:over)). Regeneration was overestimated in most models
+for both the 7 and 10 cm diameter thresholds, with the exception of the
+empirical stand model SIBYLA, the landscape model Landis-II and the
+global model aDGVM2, which estimated regeneration levels at the lower
+end of the plausibility interval of the observed data (Figure
+\@ref(fig:over)). The models with the largest overestimation were the
+stand model PICUS and the landscape model TreeMig. For most models, the
+variability of simulated regeneration levels across the 200 sites
+(visible from the interquartile range in the box plots of Figure
+\@ref(fig:over)) was similar to or smaller than observed, with the
+notable exception of PICUS, where simulated regeneration variability was
+much larger.
+
+Both the observations and the simulated data had no regeneration in some
+samples and at some sites (for details, cf. Table S1). The observed data
+had 4% of the samples with no regeneration. Only three models had a
+larger proportion of no regeneration (4C, Landis-II and aDVM2). Two
+models always simulated regeneration for both the 7 and 10 cm threshold
+(xComp and TreeMig), i.e., they did not feature any zero values. The
+other ten models had a very low percentage of samples with no
+regeneration (0.01% - 2.39%), i.e., they had distinctly fewer
+occurrences of zeros compared to the observations.
+
+## Tree species diversity of regeneration
+
+Most models matched the level of diversity of the observed data quite
+well (Figure \@ref(fig:H710)). Five models overestimated regeneration
+diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
+particularly for the 7 cm diameter threshold. The model 4C is a special
+case, as it simulated five species only, i.e., its diversity values are
+not directly comparable to those of the other models, nor to the
+observations. Only one model, Landis-II, consistently underestimated
+regeneration diversity.
+
+In most models, there were only small but significant differences in the
+species diversity of regeneration between 7 and 10 cm across sites
+(Tables S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS)
+maintained the regeneration diversity between the 7 and 10 cm thresholds
+(i.e., the differences between them were not significant, see Table S2),
+and the same was evident from the observed data.
+
+For both the observed and the simulated data, and both regeneration
+thresholds, species diversity in the regeneration compared to
+stand-level diversity did not feature clear patterns across the models
+(Figures S1 and S2). Most models captured reasonably well or
+overpredicted species diversity at the stand level (Figure
+\@ref(fig:combinedRichness)A, left and center; cf. Figure S3), but a
+characteristic feature was that the simulations had a much lower
+variability of diversity compared to observations. Models overpredicting
+species diversity included ForClim1, ForClim 11, FORMIND and TreeMig
+(Figure \@ref(fig:combinedRichness)A, left). Several models
+underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C
+(Figure \@ref(fig:combinedRichness)A, right).
+
+The majority of the models overpredicted regeneration diversity; as
+already observed in Figure \@ref(fig:H710), only two of them
+underestimated it (4C and Landis-II) (Figure
+\@ref(fig:combinedRichness)B and Figure S4). Both the models with
+feedback and those without feedback from the adult trees to regeneration
+(via seed production) overestimated, underestimated or captured
+reasonably well regeneration species diversity compared to the
+observations. Again, simulated diversity had considerably lower
+variability than observed diversity in the regeneration.
+
+## Mortality in tree regeneration
+
+There were strong differences among the models regarding the mortality
+rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the
+models featured a mortality rate significantly larger than the observed
+data, thus at least partially compensating for the general
+overestimation of regeneration levels (Figure \@ref(fig:over) and Table
+S3); six models did not feature a significant difference compared to the
+observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
+Four models (xComp, PICUS, LandClim and TreeMig) featured very high
+mortality rates (i.e., well above the Reineke self-thinning line), which
+compensated for the strong initial overestimation of regeneration (cf.
+Figure \@ref(fig:over)). Interestingly, two models that underestimated
+overall regeneration levels (SIBYLA and Landis-II; cf. Figure
+\@ref(fig:over)) featured mortality rates that were close to but still
+above the self-thinning line (Figure \@ref(fig:mortality11)).
+
+The fact that some models (and observed data) featured lower mortality
+than expected by Reineke's self-thinning rule may point to facilitation,
+or simply a higher mortality rate before the trees had reached 7 cm
+diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
+whose mortality was well above the self-thinning line, indicates that
+they feature higher regeneration mortality in mixed-species forests
+compared to even-aged, single-species forests. Some models yielded a
+mortality rate of regeneration that is broadly compatible with the
+self-thinning rule (e.g., 4C and LandClim).
+
+The three models with the largest overestimation of the proportion of
+regeneration at the 7 cm threshold featured the highest mortality
+between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The
+other models that overestimated regeneration had mortalities lower than
+the expected self-thinning ratio, with the exception of LandClim (Figure
+\@ref(fig:mortality13)B). It is noteworthy that most of the models that
+featured a low ratio of regeneration between 7 and 10 cm (i.e.,
+ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small
+overestimation of regeneration at the 7 cm threshold (Figure
+\@ref(fig:mortality11)). aDGVM2 was the only model that underestimated
+regeneration at the 7 cm threshold while featuring a low mortality
+between the two thresholds (Figure \@ref(fig:mortality13)C), and only a
+few models were close to the correct regeneration levels at 7 cm while
+also being close to the Reineke line (SIBYLA, Landis-II; cf. Figure
+\@ref(fig:mortality13)D; and 4C as well as LandClim).
+
+Lastly, some models that had a low number of individuals at the 7 cm
+threshold also had a smaller decrease of regeneration between 7 and 10
+cm (ratio 7-10 cm) (Figure S5). In these models, competition and
+self-thinning are either not pronounced or must have occurred before the
+trees had reached 7 cm. However, this pattern was not consistent across
+models (Figure S5). It is noteworthy that there is a relationship
+between the regeneration levels at 7 cm and the mortality rate (Table
+S4): most of the models showed a positive effect (higher regeneration at
+7 cm is coupled to an increase of the mortality between 7 and 10 cm).
+While some models showed a negative effect, the linear trend was not
+pronounced and not always significant regarding the slope.
+
+## Model performance and model traits
+
+There was no significant relationship between the mean complexity of the
+regeneration module (Table \@ref(tab:tableModels)) and the
+overestimation proportion at a diameter threshold of 7 cm (Figure S6).
+All the models except 4C had significant differences between the
+observed and simulated mean regeneration values. Therefore it was not
+possible to differentiate in terms of performance between models type
+(empirical or process based), or their scales (stand, landscape or
+global) (Table S5). The values of regeneration species diversity
+simulated by each model were also assessed against the observed data to
+see if there was a difference between models that include a feedback
+compared to those that did not; all the models except ForCEEPs(f) and
+FORMIND had significant differences (Table S6). It is noteworthy that
+for the model ForCEEPS the pattern changed from the overprediction of
+species diversity in the regeneration to a diversity level that is
+closer to observations when the recruitment module included feedback
+(ForCEEPs(f)).
+
+## Regeneration gradients on regeneration niches
+
+When evaluating regeneration levels along key gradients of light
+availability (basal area), temperature (degree-days) and soil moisture
+(climatic water balance), distinct features emerged: the models
+reproduced the effect of basal area in both its magnitude and patterns
+(Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better
+than the effects of the climatic gradients, where they featured varying
+patterns.
+
+In the observed data, regeneration levels decreased clearly with
+increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a).
+This trend was captured by the Landis-II model only, albeit at much
+lower values of basal area than in the observations. The other models
+featured distinctly different trends, such as (1) an increase in
+regeneration levels with increasing total basal area followed by the
+absence of regeneration at high values of basal area (ForCEEPS,
+ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant
+regeneration levels with basal area (ForClim 11, xComp, LandClim and
+aDGVM2), or (3) an increase of regeneration up to a certain value of
+stand basal area followed by a decrease at even higher values of stand
+basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA
+and iLand).
+
+The observed regeneration did not change much across the climatic water
+balance gradient, and showed a slight increase with temperature. The
+models did not match the observed regeneration patterns across the
+gradients of soil moisture and temperature (Figure
+\@ref(fig:nvTrend710BAlightTemp)b,c and S7), and neither was there a
+clear pattern across models. Regarding the soil moisture gradient, a
+group of models featured decreasing regeneration with increasing soil
+moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim,
+TreeMig and iLand), while a few models showed the opposite trend
+(ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient,
+all the models that featured a decrease in regeneration levels with
+increasing soil moisture showed the opposite trend (i.e., more
+regeneration with increasing temperature), except for 4C. The other
+models did not feature a clear pattern.
+
+Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
+sylvestris*) and oak (*Quercus* spp.) were the species observed most
+frequently in the EuFoRIa data. Most models captured well the share of
+basal area in the regeneration of these five species compared to the
+observed data along the two major climatic gradients (Figure
+\@ref(fig:diff), cf. Figure S8). However, the models differed
+considerably in the way in which they simulated these climatic niches.
+
+First, some models overestimated regeneration across the entire
+gradients of at least one of the five main tree species (Figure
+\@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS),
+while other models overestimated regeneration of more than one species
+for a large part of the climate gradients, such as xComp, PICUS, SIBYLA
+or LPJ-GUESS (cf. red colors in Figure \@ref(fig:diff)). These trends
+were consistent for both regeneration thresholds, although the exact
+changes in the share of the regeneration basal area were different
+(Figure S8).
+
+Second, there was some congruence in the simulation results by species
+across the models. The share in the regeneration of Silver fir (*Abies
+alba*) was captured well across the gradient by most models except for
+xComp, PICUS, and LandClim, which overestimated the regeneration share
+of this species. There were also some peculiarities evident for some
+models, such as 4C that did not simulate *Abies alba*. The patterns
+across models for *Fagus sylvatica* regeneration were more complex, as
+its regeneration was underestimated by many models across the
+environmental gradients featuring negligible regeneration at most sites,
+while others overestimated it in different parts of the environmental
+space. Most models overestimated the share of *Picea abies* regeneration
+in the cold-wet part of the gradients. *Pinus sylvestris* and *Quercus*
+spp. represented a small share of the regeneration in the observed data
+mostly at warm-dry sites. Many models, however, erroneously featured
+regeneration for these two species along most of the climatic gradients,
+although with a low share of basal area. Five models (4C, FORMIND,
+SIBYLA, xComp, and LandClim) had almost no regeneration of *Pinus
+sylvestris* and *Quercus* spp., or did not feature any regeneration of
+these species at all.
+
+Lastly, there is no model that performed well across the five species
+regarding the species-specific regeneration levels in the environmental
+space (Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
+tended to consistently overestimate the regeneration share of all five
+species, but most models overestimated the regeneration share of some
+species while underestimating the share of others (e.g., ForCEEPS,
+LPJ-GUESS).
+
+# Discussion
+
+Tree regeneration is a fundamental process in forest dynamics. Correctly
+capturing tree regeneration in dynamic models is fundamental to, for
+example, evaluate post-disturbance dynamics and potential long-term
+recovery trajectories as it will define the forest state in the first
+decades [@seidl2022]. If this initial phase is not captured well, we can
+not properly assess aspects such as forest resilience or the timing,
+magnitude and progression of carbon sequestration. In this study, for
+the first time, projections of tree regeneration from multiple models of
+forest dynamics were confronted with a unique dataset from unmanaged
+forest reserves across a large environmental gradient in Europe.
+
+The EuFoRIa data [@käber2023] are exceptional, particularly with respect
+to the number of records (number of sites and repeated measurements),
+which is essential for capturing a highly 'noisy' process such as tree
+regeneration. The use of this dataset for model benchmarking provided
+novel insights on the ability of state-of-the-art models to accurately
+simulate regeneration levels, species composition, and mortality in an
+early stage of tree life, i.e., between a tree diameter of 7 and 10 cm.
+Overall, by adopting this approach a much broader understanding resulted
+than if we had used these data for model calibration: it is primarily
+from the shortcomings of the models that we can gain ecological insights
+[cf. @trugman2022].
+
+## Regeneration levels
+
+Most of the models overestimated tree regeneration levels. This has
+potentially far-reaching implications, e.g., regarding biomass (and thus
+carbon) turnover, with a potential overestimation of the capacity for
+forest carbon sequestration [@pugh2019; @pan2011]. Yet, we focused on
+one specific stage, i.e., regeneration into the 7 and 10 cm diameter
+classes. Trees at this size contribute little to carbon sequestration,
+and if the excess regeneration at this stage is compensated -- in the
+models -- soon thereafter by higher mortality, overall simulated
+regeneration may still yield trustworthy results at the level of stand
+structure and related forest attributes such as basal area or biomass.
+
+From a modeling point of view, excess regeneration inevitably requires
+excess mortality rates in a later stage, either -- as observed for some
+models -- between the 7 and 10 cm diameter thresholds, or soon after the
+10 cm threshold has been crossed. In any case, correcting at early
+stages for the expected forest densities at later stages is equivalent
+to compensating for a first error (excess regeneration) by a second
+error (excess mortality). It is highly likely that biased projections
+will result, because the two errors are unlikely to be perfectly linked
+and thus will not always compensate each other. Hence, this structural
+problem of most models investigated here is problematic particularly if
+the models are to be used under novel conditions such as under climate
+change [e.g., @huber2021] or in a decision support context [e.g.,
+@thrippleton2021].
+
+Regeneration levels define the structure and composition of future
+forests, but it is equally important to correctly identify areas where
+regeneration is lacking [@rammer2021]. There are multiple constraints to
+the regeneration niche of tree species [@price2001], and therefore the
+absence of regeneration is likely to be common [@fortin2007], even over
+larger areas such as the one-hectare plots used here. Tree regeneration
+data are characterized by zero-inflation. This was clearly evident from
+the EuFoRIa dataset, but some of the models did not produce zeros at
+all, or featured a very low proportion of zero data. This substantial
+difference may be due to the fact that the simulation results were drawn
+from equilibrium forests, whereas in reality many of the forest reserves
+are recovering from past management activities and have become denser
+over the past decades [e.g., @heiri2009], leading to less regeneration
+than in an equilibrium situation.
+
+## Species composition of regeneration
+
+Correctly capturing species composition in tree regeneration is
+important to assess the future functional diversity of a forest, e.g.,
+its sensitivity to drought or resilience to disturbances [@redmond2015;
+@seidl2022]. In the simulations, overall species diversity levels in the
+regeneration were well within the observed range for half of the models
+(7 out of 14). Thus, while most of the models are facing difficulties to
+quantitatively match regeneration levels (cf. above), their performance
+is better when it comes to matching the composition of regeneration as a
+function of abiotic and biotic conditions. Most models maintained or
+even decreased regeneration diversity between the 7 and 10 cm
+thresholds, and the same is visible from the empirical data, but the
+differences were not significant. Diversity variations may be explained
+by the way the individual models consider regeneration processes
+[@konig_tree_2022], e.g., without considering the feedback from adults,
+regeneration can be linked to either seed rain from dispersal or a
+seed/seedling bank; neither process is incorporated across many models
+studied here, however.
+
+The species diversity of the entire stand was captured better by the
+models than the diversity of tree regeneration, and only a few models
+overpredicted stand-level diversity. Defining regeneration composition
+as being proportional to adult abundance, regardless of productivity,
+might be a simple and conservative assumption to safely maintain
+relative species abundances [@hanbury-brown2022], although this may be
+simplistic. Furthermore, based on the simulation results, there is no
+evidence that models with feedback from the canopy captured better the
+species diversity of the regeneration than those without feedback, with
+the exception of ForCEEPS that significantly improved regeneration
+species diversity with the model version that included feedback
+(ForCEEPS(f)). The similar performance of models with and without
+feedback is likely because the models put more weight on the
+regeneration niche arising from abiotic and biotic filters, than from
+the habitat niche of the adult trees [@grubb1977].
+
+## Regeneration mortality
+
+There are several factors that lead to mortality during the regeneration
+phase, such as competition [@casper1997] and multiple abiotic factors
+[@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
+expect that models that overestimate regeneration may have a
+particularly high mortality between the 7 and 10 cm thresholds. However,
+this was not consistently evident from the simulations. This implies
+that these models must have an excess mortality in later stages, if we
+assume that all models were able to capture the structure and
+composition of the adult stands along the EuFoRIa gradient; this however
+was not tested here.
+
+Higher mortality towards the adult phase has important implications for
+forest dynamics and the goods and services provided by forests. On the
+one hand, mortality in later stages may erroneously enhance the share of
+less shade tolerant species in the models [@klopcic2015], therefore
+shifting the species composition. Unrealistic high stem densities for a
+longer period of time may overestimate the role of tree regrowth in
+carbon sink dynamics [@pugh2019]. Given our set of simulations and
+analyses, we cannot conclusively assess what is happening in the models,
+and further studies are required that focus on a wider range of tree
+sizes and the fate of tree regeneration along such a size continuum.
+
+## Model performance and model traits
+
+In spite of the critical considerations above, it is remarkable that
+most models did not deviate exceedingly from the observations with
+respect to simulated regeneration levels -- after all, few if any of the
+models' regeneration routines are sufficiently constrained by data, with
+the exception of the empirical models xComp and SIBYLA. It is noteworthy
+that in spite of their empirical basis, these two models did not match
+empirical regeneration levels, in a similar magnitude as the other
+models and even in opposite directions (SIBYLA: underestimation; xComp:
+overestimation). It appears that using empirical data of limited
+geographical scope and constrained to managed forests, as done in these
+models, leads to extrapolation problems already under current climatic
+conditions. For the other models, whose regeneration modules are not
+strongly constrained by empirical data, the degrees of freedom that are
+available for modeling regeneration [@konig_tree_2022] could have
+implied that model performance would be much worse than what we found.
+
+Our study showed that increasing complexity in the regeneration modules
+is not linked with a higher accuracy of the projections of regeneration
+levels, species composition or mortality at early tree stages, as there
+was no significant relationship between model performance and model
+complexity. Increasing complexity in regeneration modules has been
+motivated by better process understanding and enhancing of model
+accuracy [@bugmann2022]. However, more complex models do not necessarily
+lead to better projections and rather to a reduced transparency and
+lower predictive power [@franklin2020]. Thus, the question of what level
+of detail is appropriate and parsimonious for modeling tree regeneration
+processes remains open [@konig_tree_2022; @bugmann2022].
+
+## Regeneration gradients on regeneration niches
+
+Competition for light as a strong filter for tree regeneration has been
+widely documented [@messier1999; @collet2006; @berdanier2016], but the
+models examined here did not reproduce this expectation. However, it is
+difficult to measure light availability at large spatial and temporal
+scales. We used total stand basal area as a proxy for light availability
+[cf. @schmid2021]. However, we were unable to consider light
+availability restrictions caused by ground vegetation, which may be an
+important filter for some EuForIa stands [@woltjer2008]. We found
+pronounced differences in the stand basal area ranges simulated by the
+models, but also between models and observations. This made it
+impossible to evaluate the regeneration for the extremes of the stand
+density ranges in some models. For example, regeneration levels at low
+stand densities are relevant to assess how well forests are recovering
+e.g. after gap creation due to disturbance [@seidl2022; @grubb1977]. At
+the level of the simulated one-hectare samples, average basal area is
+typically not very low as long as no larger disturbances occur, which
+was explicitly excluded in our simulation protocol. Towards the other
+end of the spectrum, i.e., with increasing stand basal area, it would be
+reasonable to expect that the regeneration of the different species
+would become sparser and drop out entirely under low-light conditions
+[@klopcic2012; @zell2019]. However, few models showed this trend, thus
+indicating that the relation between regeneration and light availability
+is not yet captured correctly in most models. Yet, several of the models
+that did feature an increase of regeneration with increasing basal area
+include a feedback between seed production of mature trees and
+regeneration. Thus, it seems that higher light competition does not
+sufficiently compensate for increased seed availability with higher
+basal area due to a higher abundance of mature trees.
+
+There were pronounced differences how the main tree species were
+represented by the models along the environmental gradients, in
+particular the dominant species *Abies alba*, *Fagus sylvatica* and
+*Picea abies*. The regeneration levels were sampled from simulations in
+the equilibrium, and in this state it is expected that non-climax
+species such as *Pinus sylvestris* or *Quercus* would be of minor
+importance, or absent [@klopcic2015]. Most models captured this low
+abundance, which is also found in the empirical data. Thus, although the
+broad patterns are matched by many models, improvements in the
+quantification of the regeneration niche of the species are needed, but
+this cannot be done in the absence of robust datasets across multiple
+
+## Methodological considerations
+
+The EuFoRIa data as used here are unique and well suited to better
+understand tree regeneration. However, three aspects of these data may
+represent considerable limitations. First, we made a comparison of tree
+regeneration in an equilibrium state, but we cannot assess how close the
+forests included in the EuFoRIa dataset were to such a state. The data
+were collected in forest reserves where no management has taken place
+for long periods of time. This makes our assumption of an equilibrium
+between forest properties and environmental drivers more reasonable than
+it might appear at first sight. In an analysis of primeval forests at
+demographic equilibrium, @brzeziecki2021 found higher regeneration rates
+than the ones observed in the EuFoRia data. Thus, the overestimation of
+regeneration rates by the models may not be so problematic. Second, the
+data were collected for regeneration above a 7 cm threshold, thus
+limiting the assessment of tree regeneration to a holistic viewpoint.
+This constituted a hard limit based on which we can understand only some
+aspects of tree regeneration, which in its entirety often comprises a
+rather long period since seed production [@price2001]. In reality, many
+environmental constraints are acting on young trees [@käber2021] that we
+were unable to assess. Yet, regeneration data with lower thresholds are
+simply not available in a harmonized manner across large environmental
+gradients. Third, the empirical data were collected from rather small
+plots, while we sampled simulated regeneration levels from 1 ha areas,
+which may lead to an incorrect representation of space. Even though the
+strategy we adopted is not ideal, it represents a common challenge when
+harmonizing diverse data sources originating from varied sampling
+strategies [@portier2022]. It would have been extremely challenging for
+such a variety of models to follow a protocol where the spatial sampling
+size was different at each of the 200 sites, and it would have
+introduced additional uncertainty in the results. Ultimately, we need to
+acknowledge that the observed data represent just a snapshot of an
+extremely complex and stochastic reality. This may at least partly
+explain e.g. the lack of clear patterns in the simulation results along
+the gradients of climate water balance and degree-days.
+
+The design of our sampling protocol did not include spatial aspects such
+as seed dispersal or detailed soil data. While we considered a wide
+range of models of forest dynamics, from stand to global scales, the
+simulation setup was limited to 1 ha regeneration values. This lack of
+consideration of spatial scale is appropriate for stand-scale models,
+but it potentially puts the landscape models at a disadvantage, as they
+have been built to be accurate at the landscape level. Without spatial
+context, we are limiting regeneration to the seed influx from the stand
+itself, unless the model has a background seed input. Yet, the global
+models should not be at a disadvantage due to the limited spatial
+considerations [but cf. Snell et al. -@snell2018; or Lehsten et al.
+-@lehsten2019], as they are usually lacking dispersal between cells and
+are based on a strong abstraction of horizontal space
+[@hanbury-brown2022]. Lastly, detailed data on soil conditions were not
+available from the observed data, and independent, admittedly rather
+coarse data for soil and climatic water balance had to be used instead.
+It is noteworthy that many models represent drought based on detailed
+indicators based e.g. on soil water holding capacity, which had to be
+derived from a rough soil quality measure. This may at least partially
+explain the unsatisfactory performance of many models along the drought
+axis (i.e., climatic water balance).
+
+## Research recommendations
+
+With our study, we have demonstrated that models of forest dynamics need
+a focus on tree regeneration modules to make them more robust. It
+remains uncertain what level of detail is required to model regeneration
+processes, and this must be addressed in future research. We recommend
+that the improvement of the regeneration modules is implemented as
+additional features that can be traced back, as done here for the
+variants of ForClim and ForCEEPS, and that model complexity and
+structure must always be connected with modeling objectives
+[@albrich2020]. If it should be necessary to include more detail in the
+regeneration models, this will come with higher parameterisation
+efforts. This will most likely lead to lower generalization because the
+required data will have to be collected from specific locations, as
+currently there is no general, comprehensive regeneration dataset
+available.
+
+Therefore, we further recommend that more effort should be invested into
+collecting harmonized datasets on tree regeneration. We emphasize that
+datasets such as EuFoRIa are invaluable and should be expanded in both
+their spatial extent (e.g., towards boreal and Mediterranean conditions)
+as well as in time (e.g., continuing the monitoring into the future).
+Such data will allow for a better evaluation of forest models and help
+to reduce the uncertainty in their projections, which is crucial when
+they are used as tools for predicting e.g. the impacts of anthropogenic
+climate change.
+
+In the present study, we have considered tree regeneration in the
+equilibrium state only. It is equally important to understand how these
+models project tree regeneration after changes in forest structure by
+disturbances [@seidl2022], or under different management strategies
+[@lindner2000]. However, this will require an entirely different set of
+observed data, and potentially not all models of forest dynamics would
+be able to assess the relationship of these aspects on tree
+regeneration, e.g. due to the lack of disturbance or appropriate
+management modules.
+
+We recommend to investigate in detail the implications of the current
+modeling strategies for tree regeneration and, ultimately, simulated
+forest stand structure. This applies particularly to the erroneous
+patterns of excess tree regeneration and later excess mortality, by
+focusing on a wider range of tree sizes and the related regeneration
+dynamics. We also recommend, especially for the landscape-level models,
+the inclusion of explicit spatial considerations regarding tree
+regeneration [@beckage2003]; this, however, is a serious challenge
+regarding appropriate datasets, as inventory data have wide coverage,
+but by definition do not allow for the assessment of spatial interaction
+effects.
+
+Exercises like the one presented here, where the models are operated in
+"blind flight" mode, i.e., without the possibility of being tuned
+towards capturing the expected patterns, should be repeated. Such
+benchmarking exercises should next focus on aspects such as specific
+model traits and the ecological formulations of particular
+(sub-)processes, to better understand the implications of the
+assumptions on which the models are based. Furthermore, the inclusion of
+a wide range of models with different scales, type, population
+structures, tree regeneration modules and complexity will ensure a wider
+benefit to the whole modeling community.
+
+## Conclusions
+
+Models of forest dynamics are important tools in science and decision
+support, and the formulation of tree regeneration has strong
+implications for simulated forest properties. The 15 models and variants
+used here are facing similar challenges in their representation of tree
+regeneration: they generally overestimate tree regeneration levels, and
+the simulated regeneration niche is not always captured accurately as a
+function of biotic (light) and abiotic (temperature and moisture)
+factors.
+
+However, most models properly capture the diversity of the initial tree
+community, and differences between model formulations, e.g., the
+presence or absence of feedback from the adult trees, did not have a
+strong effect for capturing the species composition of regeneration.
+
+Regarding mortality in the early phase of tree life, many models that
+feature a particularly high overestimation of regeneration levels are
+compensating for this by a larger tree mortality. Often, this
+compensation is not sufficient to reduce the high regeneration levels to
+realistic values. Overall, there is no clear mortality pattern across
+all models.
+
+When capturing tree regeneration, the specific design decisions taken in
+the development of any model are more important for its behavior
+(accuracy) than scale (stand, landscape, global), modelling approach
+(empirical vs. process-based), and complexity. Having both empirical and
+process-based models in our set, the empirically-based models could have
+been expected to have a better performance, as they were calibrated with
+inventory data, but this was not the case. Similarly, higher model
+complexity does not represent an improvement for capturing tree
+regeneration.
+
+Even though the regeneration routines of most of the models investigated
+here have never been sufficiently constrained by robust data, the
+regeneration projections are not overly off. This indicates that a lot
+can be gained by a focus on the modeling of regeneration processes. The
+representation of forest dynamics in these models would become much more
+robust particularly in the face of climate change and post-disturbance
+dynamics, thus strongly reducing the uncertainty in long-term
+projections of future forest dynamics.
+
+# Author contributions
+
+O. Díaz-Yáñez and Y. Käber contributed equally and share the first
+authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study
+design. All authors contributed to preparing the simulations with their
+respective models. O. Díaz-Yáñez led the data analysis with
+contributions during the workshop from Y.Käber, T. Anders, K. H.
+Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T. Hickler, H. Lischke,
+M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X. Morin, W. Rammer and
+D. Scherrer H. Bugmann. O. Díaz-Yáñez and H. Bugmann led the writing of
+the manuscript. All authors participated in the revision of the
+manuscript and approved its submission.
+
+# Acknowledgements
+
+This paper arises from a workshop held in Davos (Switzerland) in June
+2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
+Funding for the workshop was provided by COST Action CA19139 PROCLIAS
+(PROcess-based models for CLimate Impact Attribution across Sectors),
+supported by COST (European Cooperation in Science and Technology;
+[https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research
+Grant ETH-35 18-1). ODY was funded by the Swiss National Science
+Foundation project 200020_188882 / 1. KB received support from the
+European Research Council under the European Union's Horizon 2020
+research and innovation program (Grant Agreement 101001905). JB
+acknowledges funding from the long-term research development project no.
+RVO 67985939 from the Czech Academy of Sciences. MM acknowledges
+financial support from I-Maestro (Innovative forest management
+strategies for a resilient bioeconomy under climate change and
+disturbances, grant no. 22035418, 2019--2022) funded by the ERA-NET
+Cofund ForestValue. MMina acknowledges funding from the H2020-MSCA-IF
+project REINFORCE (Grant No. 891671). KM received funding from the
+Project No. CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant
+no. ITMS2014+313011W580 supported by the Integrated Infrastructure
+Operational Programme funded by the ERDF. CH acknowledges financial
+support from Waldfonds WASIM2100 project. We would also like to thank
+all the people that helped to make this study possible with
+contributions at different stages such as Martin Gutsch and Giorgio
+Vacchiano.
+
+# Conflict of interest statement
+
+The authors declare no conflicts of interest.
+
+# References
+
+```{=html}
+```
+::: {#refs}
+:::
+
+\newpage
+
+# Table
 
@@
 
-## Observed data
-
-Regeneration data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset covering such gradients over a large number of sites, to evaluate how well regeneration is captured.
-The observations used here are derived from a novel and unprecedented network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network [@euforia2019] (Figure \@ref(fig:map)).
-These forests have been unmanaged for at least 10 years prior to their designation as reserves, and most of them provide time series of natural forest dynamics over multiple decades and up to ca.
-80 years.
-The census periods range from 3 to 37 years, with an average of 14 years.
-The data provide information at the tree level, thus allowing for the sequential comparison of processes such as individual-tree regeneration and death between the measurements.
-
-We selected 200 sites from this network as the benchmarking dataset for the simulation to be representative of the environmental variation contained in the data.
-This was achieved by applying k-means clustering to define 200 clusters of plots from the original set of 869 plots along the environmental dimensions of temperature, climatic water balance, soil quality and exposition (i.e., slope and aspect).
-Each of these sites featured at least two consecutive measurements.
-Regeneration thresholds for these sites differed between diameters of 0 and 10 cm.
-For our study, we defined two datasets where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds between 7 and 10 cm.
-
-Plot size ranged from 0.02 ha to 5.52 ha, and the data were further processed and aggregated following @käber2023 to provide regeneration rates per ha and per decade.
-The observations used in this study featured 30,900 newly established trees.
-Regeneration rates per site, sample and decade ranged from zero to 1246 trees, with a mean of 56 trees.
-Adult species composition was also available for each plot.
-For more details on this unique dataset and the detailed regeneration information, cf.
-@käber2023.
-
-## Simulation protocol
-
-The overarching goal of the experiments was to assess the tree regeneration as it arises from empirical data against its representation in a wide range of models of forest dynamics.
-We define tree regeneration as the passing of a breast height diameter threshold of 7 or 10 cm, respectively (synonym: ingrowth).
-To this end, each modeling group was provided with a detailed protocol [@díaz-yáñez2022] with instructions how to perform the simulations, providing input variables on climate and soil conditions, and the list of expected output variables.
-Neither were further site information (except for the data specified below) nor any data on tree regeneration or forest stand features provided prior to the simulation.
-That is, the models were run in "blind flight" mode.
-
-The input variables were collected from different data sources and aggregated to be adapted to the needs of the different models.
-Time series of climatic variables from 1981 to 2018 were provided in hourly [Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly resolution [CHELSA data, Karger et al.
--@karger2021].
-Some variables required for some models, such as relative humidity or vapor pressure deficit, were calculated from these variables.
-The final instructions for using the climate data were slightly different in each model, based on the approach that best suited the model (Table \@ref(tab:tableModels)).
-Soil quality data were provided as continuous values between 1 and 5 [Soilgrids dataset, Hengl et al.
--@hengl2017].
-The protocol also provided the elevation, slope and aspect for each of the 200 sites [@asterscienceteam2019], but no other spatial information such as coordinates, with the exception of iLand and aDGVM2, which required blurred coordinates to derive highly detailed soil data.
-
-The simulations were run in the absence of natural disturbances.
-Only the model LPJ-GUESS had to include a background disturbance to increase chances of shade-intolerant species to establish.
-The simulations were set up to sample species-specific regeneration rates per decade and per ha in the equilibrium state of the model, typically entailing a "spin-up" run (as we did not provide any forest data).
-The modeling teams decided on the simulated area and how they derived these samples.
-The exact length of the simulation was also decided by the modeling teams (Table \@ref(tab:tableModels)).
-Further details on how each modeling team prepared the simulations and the outputs are available in Supplementary Material 1.
-
-The simulations were run in the absence of management to a simulated equilibrium ("Potential Natural Vegetation") with the current climate, as the ultimate goal was to evaluate tree regeneration under comparable and near-equilibrium conditions.
-This entails the assumption that (1) the observations from the forest reserves reflect no traces of forest management, and (2) there is an equilibrium between forest dynamics and climate.
-While the former might be starting to be visible in many of the EuFoRIA reserves, the latter may be more debatable.
-However, in the absence of detailed data on the history of each plot in the EuFoRIa network, some broad assumptions had to be made.
-Both the width of the regeneration niche (i.e., in environmental space) as well as the intensity of the regeneration process (i.e., the number of ingrowth trees per area and per unit of time) were of interest.
-
-The simulations were run for mixed-species forests (not multiple single-species simulations) using mixtures of eleven species or genera for which regeneration data of sufficient quality were available from EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp.
-L., Fraxinus excelsior L., Quercus spp.
-L., Alnus glutinosa L., Pinus sylvestris L.* The same set of species was used at all 200 sites.
-Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig).
-In three models, fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)).
-In 4C only *Fagus sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp.
-and *Pinus sylvestris* are parameterised.
-xComp simulations did not consider *Fraxinus excelsior* due to a depreciated species parameterisation.
-In LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not parameterized and therefore these species could not be included in the simulation.
-Finally, the model aDGVM2 does not simulate individual species.
-Rather, community assembly processes and trait filtering generate plant communities that are adjusted to the biotic and abiotic conditions, and the simulated plants can be classified into ecological strategies based on their trait values in a post-processing step.
-
-Each of the models reported the regeneration number by sampling 200 times in a 10-year interval for each species and per ha for each of the 200 sites.
-Multiple samples per site were used to better understand the simulated variation within each site.
-This was done using different strategies, depending on the model, including (1) sampling simulated data from the same 1 ha plot in the equilibrium over time, (2) sampling several 1 ha plots from the simulated forest at one specific point in time (in the equilibrium) or (3) a combination of (1) and (2).
-This resulted in 880,000 observations per model that simulated the 11 species included in the protocol (200 sites, 200 samples per site, 11 species and two diameter thresholds).
-For the models that simulated additional species, their regeneration rates were aggregated as "others".
-
-Two models did not provide results from all the simulated samples or sites to avoid unrealistic results, as follows.
-In the model 4C, a threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid unrealistic stand basal area data, and therefore not all the samples and sites were reported.
-The reason for this is that 4C is not suitable for long-term simulations without management, due to misrepresentations in density-dependent mortality processes in long-term simulations and assumptions of tree geometry that lead to unrealistic single tree dimensions of very old individuals.
-LPJ-GUESS had 2% of the sample outputs without tree regeneration, and these were considered as zero stand basal area and zero regeneration for all the tree species simulated; one site produced grassland rather than a forest and was not included in the results.
-
-```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed regeneration at the 10 cm threshold.", out.width='80%'}
+\newpage
+
+# Figure captions
+
+**Figure 1.** The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used
+in the present study, the colors show the dominant tree species per site
+in the observed regeneration at the 10 cm threshold.
+
+**Figure 2.** Mean regeneration levels across all samples per site,
+plotted for the 200 sites and for each model. The red dashed lines shows
+the 25th and 75th percentiles for the 7 cm diameter threshold in the
+observed data. There are two boxplots for each model where the left
+lighter boxplot corresponds to 7 cm and the right darker boxplot to 10
+cm
+
+**Figure 3.** Shannon index of tree regeneration, calculated by basal
+area, as the mean value across the 200 samples per site for the diameter
+thresholds of 7 and 10 cm, respectively. The red dashed lines shows the
+25th and 75th percentiles for the 7 cm diameter threshold in the
+observed data.
+
+**Figure 4.** Mean Shannon index across all samples per site for
+observed and simulated data. Each plot shows one pattern represented by
+one exemplary model of each category (overperdicted, intermediate and
+underpredicted). The full data with the grouping of the models are shown
+in Figures S3 and S4. n indicates the number of models falling in each
+group. A) Examples for the three trends across models for species
+diversity at the stand level. B) Examples for the three trends across
+models for regeneration at the 7 cm threshold.
+
+**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm
+thresholds. Dashed blue lines mark a ratio equal to 1, indicating no
+decrease in tree regeneration between 7 and 10 cm, and a ratio equal to
+1.77, corresponding to the Reineke self-thinning ratio under even-aged
+conditions.
+
+**Figure 6.** Ratio of tree regeneration between the 7 and 10 cm
+thresholds (regeneration 7 cm / regeneration 10 cm) and the
+overestimation proportion at 7 cm ([simulated - observed] / observed)
+for the mean regeneration per model across sites and samples. The
+horizontal dashed lines show a ratio equal to 1, indicating no decrease
+of regeneration between 7 and 10 cm, and a ratio equal to 1.77
+corresponding to the Reineke self-thinning line under even-aged
+conditions. The vertical line indicates zero overestimation.
+
+**Figure 7.** Mean regeneration values across the 200 samples per site
+with the y-axis scaled differently by model, for the 200 sites against
+gradients of (a) total basal area; (b) climatic water balance; (c)
+seasonal degree-day sum. The values were split into ten bins; the red
+lines represent a GAM model showing the trend in the observed data. For
+comparing the magnitude of recruitment across models the same graphs
+with equally scaled y-axis can be found in Figure S7.
+
+**Figure 8.** Regeneration levels of the five main species in terms of
+their basal area share in the observations (top row) and the performance
+of each model across the environmental gradients (other rows) for the 7
+cm regeneration threshold. The values shown are the mean of the 200
+samples per site and across the sites in each bin (tile), with ten bins
+per gradient. The sizes of the circles represent the ratio between the
+regeneration basal area of the species and the total regeneration basal
+area for all species. The absence of a circle indicates a zero basal
+area share in the regeneration, or the absence of regeneration
+altogether. The color gradient (for the models only) shows the
+difference between the simulated and observed ratio of regeneration
+basal area of the species and the total regeneration basal area for all
+the species.
+
+\newpage
+
+# Figures
+
+```{r map, echo=FALSE, fig.cap=".", out.width='80%'}
 knitr::include_graphics(here::here("figures", "theMap.png"))
@@
 
-## Data analysis
-
-The simulation results were analyzed regarding (1) regeneration levels (i.e., ingrowth number per unit time and space), (2) regeneration species diversity, (3) regeneration mortality, (4) the relationship between model performance and model traits, and (5) ingrowth gradients along the regeneration niches.
-We evaluated species diversity across the models and in relation to the observed data by calculating the Shannon index $H$ based on the relative proportion of the species in terms of basal area.
-It was calculated for the regeneration ($H_{R_{n}}$) (Equation \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation \ref{eqpiShannon}).
-The higher the value of the index, the higher the species diversity at a particular site and sample.
-
-```{=tex}
-\begin{equation}
-H_{R_{n}} = - \sum_{i = 1}^{s}{p_{R_{n,i}}\ln(p_{R_{n,i}})}\label{eqpiShannonR}
-\end{equation}
-```
-```{=tex}
-\begin{equation}
-p_{R_{n,i}} = \frac{r_{BA_{n,i}}}{TotalR_{BA_{n}}} \label{eqpiShannonR2}
-\end{equation}
-```
-```{=tex}
-\begin{equation}
-H_{S_{n}} = - \sum_{i = 1}^{s}{p_{S_{n,i}}\ln(p_{S_{n,i}})}\label{eqpiShannon}
-\end{equation}
-```
-```{=tex}
-\begin{equation}
-p_{S_{n,i}} = \frac{BA_{n,i}}{Total_{BA_{n}}} \label{eqpiShannon2}
-\end{equation}
-```
-where $s$ is the total number of species present and that have a basal area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of species $i$ in sample $n$ calculated as the regeneration basal area ($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the proportion of species $i$ calculated as the basal area of all trees ($BA_{n,i}$) of that species $i$ relative to total basal area ($TotalR_{BA_{n}}$) of the sample $n$.
-Species diversity was not assessed for aDGVM2 as this model does not simulate individual species.
-
-Mortality in tree regeneration was assessed based on the ratio of regeneration between the 7 and 10 cm diameter thresholds.
-We used the Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference to estimate whether the ratio of regeneration between the 7 and 10 cm diameter thresholds were above or below the expected theoretical rate.
-The Reineke self-thinning rule is usually calculated for even-aged, single species stands and assumes a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands.
-The value used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm to be 77% higher than at 10 cm), calculated using Equation \ref{eqReineke}.
-
-```{=tex}
-\begin{equation}
-Reineke = \frac{7}{10}^{-1.605} \label{eqReineke}
-\end{equation}
-```
-We assessed model performance in relation to model traits focusing on (1) model complexity as defined by @bugmann2022, (2) model type (empirical or process based), (3) the presence or absence of a canopy feedback for regeneration, and (4) the scale of application of the model (stand, landscape or global).
-We tested for significant differences using two-sided t-tests by considering each of these model features regarding the observed and simulated regeneration levels and species diversity, respectively.
-
-Total regeneration and the regeneration niches of the individual species were evaluated across the environmental gradients of light availability, temperature, and soil moisture.
-Stand basal area (the basal area of all the trees in each sample including tree regeneration) was used as a proxy for light availability at the forest floor, the annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing season warmth, and the climatic water balance as a proxy for soil moisture [@speich2019].
-Regeneration values were calculated as the mean across the 200 samples per site.
-The observed data were modeled using a Generalized Additive Model [@wood2011] with a negative binomial distribution and restricted maximum likelihood to better understand the relationship between the environmental gradients and the levels of observed regeneration, relative to the simulation results per model.
-
-In order to analyze the regeneration niches across the climatic gradients, we focused on five common tree species or genera: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and *Quercus* spp.
-For these species, we calculated the share in the regeneration basal area per site as the mean across the available samples per site:
-
-```{=tex}
-\begin{equation}
-\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{BA_{t,i}}}{\sum_{i = 1}^{s}{\bar{r}_{{BA_{t,i}}}}} \label{eq:BAshare}
-\end{equation}
-```
-where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$ is the mean basal area in the regeneration (subsequently referred to as "regeneration basal area") of species $i$ per site $t$ across the available samples at that site.
-Furthermore, regarding the share in the regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) and the total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) equaled zero.
-
-# Results
-
-## Regeneration levels
-
-Simulated regeneration levels varied strongly across the 15 models and typically did not match the levels found in the forest reserve data (Figure \@ref(fig:over)).
-Regeneration was overestimated in most models for both the 7 and 10 cm diameter thresholds, with the exception of the empirical stand model SIBYLA, the landscape model Landis-II and the global model aDGVM2, which estimated regeneration levels at the lower end of the plausibility interval of the observed data (Figure \@ref(fig:over)).
-The models with the largest overestimation were the stand model PICUS and the landscape model TreeMig.
-For most models, the variability of simulated regeneration levels across the 200 sites (visible from the interquartile range in the box plots of Figure \@ref(fig:over)) was similar to or smaller than observed, with the notable exception of PICUS, where simulated regeneration variability was much larger.
-
-```{r over, echo=FALSE, fig.cap="Mean regeneration levels across all samples per site, plotted for the 200 sites and for each model.
-The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
-There are two boxplots for each model where the left lighter boxplot corresponds to 7 cm and the right darker boxplot to 10 cm", out.width='100%'}
+\newpage
+
+```{r over, echo=FALSE, fig.cap=".", out.width='100%'}
         knitr::include_graphics(here::here("figures", "recOverAll710.png"))
@@
 
-Both the observations and the simulated data had no regeneration in some samples and at some sites (for details, cf.
-Table S1).
-The observed data had 4% of the samples with no regeneration.
-Only three models had a larger proportion of no regeneration (4C, Landis-II and aDVM2).
-Two models always simulated regeneration for both the 7 and 10 cm threshold (xComp and TreeMig), i.e., they did not feature any zero values.
-The other ten models had a very low percentage of samples with no regeneration (0.01% - 2.39%), i.e., they had distinctly fewer occurrences of zeros compared to the observations.
-
-## Tree species diversity of regeneration
-
-Most models matched the level of diversity of the observed data quite well (Figure \@ref(fig:H710)).
-Five models overestimated regeneration diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter particularly for the 7 cm diameter threshold.
-The model 4C is a special case, as it simulated five species only, i.e., its diversity values are not directly comparable to those of the other models, nor to the observations.
-Only one model, Landis-II, consistently underestimated regeneration diversity.
-
-```{r H710, echo=FALSE, fig.cap="Shannon index of tree regeneration, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
-The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.", out.width='100%'}
+\newpage
+
+```{r H710, echo=FALSE, fig.cap=".", out.width='100%'}
     knitr::include_graphics(here::here("figures", "H7_10.png"))
@@
 
-In most models, there were only small but significant differences in the species diversity of regeneration between 7 and 10 cm across sites (Tables S2).
-Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained the regeneration diversity between the 7 and 10 cm thresholds (i.e., the differences between them were not significant, see Table S2), and the same was evident from the observed data.
-
-For both the observed and the simulated data, and both regeneration thresholds, species diversity in the regeneration compared to stand-level diversity did not feature clear patterns across the models (Figures S1 and S2).
-Most models captured reasonably well or overpredicted species diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left and center; cf.
-Figure S3), but a characteristic feature was that the simulations had a much lower variability of diversity compared to observations.
-Models overpredicting species diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A, left).
-Several models underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).
-
-The majority of the models overpredicted regeneration diversity; as already observed in Figure \@ref(fig:H710), only two of them underestimated it (4C and Landis-II) (Figure \@ref(fig:combinedRichness)B and Figure S4).
-Both the models with feedback and those without feedback from the adult trees to regeneration (via seed production) overestimated, underestimated or captured reasonably well regeneration species diversity compared to the observations.
-Again, simulated diversity had considerably lower variability than observed diversity in the regeneration.
-
-```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site for observed and simulated data.
-Each plot shows one pattern represented by one exemplary model of each category (overperdicted, intermediate and underpredicted).
-The full data with the grouping of the models are shown in Figures S3 and S4.
-n indicates the number of models falling in each group.
-A) Examples for the three trends across models for species diversity at the stand level.
-B) Examples for the three trends across models for regeneration at the 7 cm threshold.", out.width='75%'}
+\newpage
+
+```{r combinedRichness, echo=FALSE, fig.cap=".", out.width='75%'}
 knitr::include_graphics(here::here("figures", "combined_richness.png"))
@@
 
-## Mortality in tree regeneration
-
-There were strong differences among the models regarding the mortality rate between 7 and 10 cm (Figure \@ref(fig:mortality11)).
-Most of the models featured a mortality rate significantly larger than the observed data, thus at least partially compensating for the general overestimation of regeneration levels (Figure \@ref(fig:over) and Table S3); six models did not feature a significant difference compared to the observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
-Four models (xComp, PICUS, LandClim and TreeMig) featured very high mortality rates (i.e., well above the Reineke self-thinning line), which compensated for the strong initial overestimation of regeneration (cf.
-Figure \@ref(fig:over)).
-Interestingly, two models that underestimated overall regeneration levels (SIBYLA and Landis-II; cf.
-Figure \@ref(fig:over)) featured mortality rates that were close to but still above the self-thinning line (Figure \@ref(fig:mortality11)).
-
-```{r mortality11, echo=FALSE, fig.cap="Ratio of tree regeneration rates between the 7 and 10 cm thresholds.
-Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree regeneration between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
+\newpage
+
+```{r mortality11, echo=FALSE, fig.cap=".", out.width='75%'}
 knitr::include_graphics(here::here("figures", "ratio7_10.png"))
@@
 
-The fact that some models (and observed data) featured lower mortality than expected by Reineke's self-thinning rule may point to facilitation, or simply a higher mortality rate before the trees had reached 7 cm diameter.
-Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher regeneration mortality in mixed-species forests compared to even-aged, single-species forests.
-Some models yielded a mortality rate of regeneration that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).
-
-The three models with the largest overestimation of the proportion of regeneration at the 7 cm threshold featured the highest mortality between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A).
-The other models that overestimated regeneration had mortalities lower than the expected self-thinning ratio, with the exception of LandClim (Figure \@ref(fig:mortality13)B).
-It is noteworthy that most of the models that featured a low ratio of regeneration between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small overestimation of regeneration at the 7 cm threshold (Figure \@ref(fig:mortality11)).
-aDGVM2 was the only model that underestimated regeneration at the 7 cm threshold while featuring a low mortality between the two thresholds (Figure \@ref(fig:mortality13)C), and only a few models were close to the correct regeneration levels at 7 cm while also being close to the Reineke line (SIBYLA, Landis-II; cf.
-Figure \@ref(fig:mortality13)D; and 4C as well as LandClim).
-
-```{r mortality13, echo=FALSE, fig.cap="Ratio of tree regeneration between the 7 and 10 cm thresholds (regeneration 7 cm / regeneration 10 cm) and the overestimation proportion at 7 cm ([simulated - observed] / observed) for the mean regeneration per model across sites and samples.
-The horizontal dashed lines show a ratio equal to 1, indicating no decrease of regeneration between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
-The vertical line indicates zero overestimation.", out.width='80%'}
+\newpage
+
+```{r mortality13, echo=FALSE, fig.cap=".", out.width='80%'}
 knitr::include_graphics(here::here("figures", "mort7_10_medianOver.png"))
@@
 
-Lastly, some models that had a low number of individuals at the 7 cm threshold also had a smaller decrease of regeneration between 7 and 10 cm (ratio 7-10 cm) (Figure S5).
-In these models, competition and self-thinning are either not pronounced or must have occurred before the trees had reached 7 cm.
-However, this pattern was not consistent across models (Figure S5).
-It is noteworthy that there is a relationship between the regeneration levels at 7 cm and the mortality rate (Table S4): most of the models showed a positive effect (higher regeneration at 7 cm is coupled to an increase of the mortality between 7 and 10 cm).
-While some models showed a negative effect, the linear trend was not pronounced and not always significant regarding the slope.
-
-## Model performance and model traits
-
-There was no significant relationship between the mean complexity of the regeneration module (Table \@ref(tab:tableModels)) and the overestimation proportion at a diameter threshold of 7 cm (Figure S6).
-All the models except 4C had significant differences between the observed and simulated mean regeneration values.
-Therefore it was not possible to differentiate in terms of performance between models type (empirical or process based), or their scales (stand, landscape or global) (Table S5).
-The values of regeneration species diversity simulated by each model were also assessed against the observed data to see if there was a difference between models that include a feedback compared to those that did not; all the models except ForCEEPs(f) and FORMIND had significant differences (Table S6).
-It is noteworthy that for the model ForCEEPS the pattern changed from the overprediction of species diversity in the regeneration to a diversity level that is closer to observations when the recruitment module included feedback (ForCEEPs(f)).
-
-## Regeneration gradients on regeneration niches
-
-When evaluating regeneration levels along key gradients of light availability (basal area), temperature (degree-days) and soil moisture (climatic water balance), distinct features emerged: the models reproduced the effect of basal area in both its magnitude and patterns (Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better than the effects of the climatic gradients, where they featured varying patterns.
-
-In the observed data, regeneration levels decreased clearly with increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a).
-This trend was captured by the Landis-II model only, albeit at much lower values of basal area than in the observations.
-The other models featured distinctly different trends, such as (1) an increase in regeneration levels with increasing total basal area followed by the absence of regeneration at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant regeneration levels with basal area (ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of regeneration up to a certain value of stand basal area followed by a decrease at even higher values of stand basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).
-
-The observed regeneration did not change much across the climatic water balance gradient, and showed a slight increase with temperature.
-The models did not match the observed regeneration patterns across the gradients of soil moisture and temperature (Figure \@ref(fig:nvTrend710BAlightTemp)b,c and S7), and neither was there a clear pattern across models.
-Regarding the soil moisture gradient, a group of models featured decreasing regeneration with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS).
-Across the temperature gradient, all the models that featured a decrease in regeneration levels with increasing soil moisture showed the opposite trend (i.e., more regeneration with increasing temperature), except for 4C.
-The other models did not feature a clear pattern.
-
-```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean regeneration values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum.
-The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
-For comparing the magnitude of recruitment across models the same graphs with equally scaled y-axis can be found in Figure S7.", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
+\newpage
+
+```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap=".", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
 
@@
 
-Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus sylvestris*) and oak (*Quercus* spp.) were the species observed most frequently in the EuFoRIa data.
-Most models captured well the share of basal area in the regeneration of these five species compared to the observed data along the two major climatic gradients (Figure \@ref(fig:diff), cf.
-Figure S8).
-However, the models differed considerably in the way in which they simulated these climatic niches.
-
-First, some models overestimated regeneration across the entire gradients of at least one of the five main tree species (Figure \@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated regeneration of more than one species for a large part of the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
-red colors in Figure \@ref(fig:diff)).
-These trends were consistent for both regeneration thresholds, although the exact changes in the share of the regeneration basal area were different (Figure S8).
-
-Second, there was some congruence in the simulation results by species across the models.
-The share in the regeneration of Silver fir (*Abies alba*) was captured well across the gradient by most models except for xComp, PICUS, and LandClim, which overestimated the regeneration share of this species.
-There were also some peculiarities evident for some models, such as 4C that did not simulate *Abies alba*.
-The patterns across models for *Fagus sylvatica* regeneration were more complex, as its regeneration was underestimated by many models across the environmental gradients featuring negligible regeneration at most sites, while others overestimated it in different parts of the environmental space.
-Most models overestimated the share of *Picea abies* regeneration in the cold-wet part of the gradients.
-*Pinus sylvestris* and *Quercus* spp.
-represented a small share of the regeneration in the observed data mostly at warm-dry sites.
-Many models, however, erroneously featured regeneration for these two species along most of the climatic gradients, although with a low share of basal area.
-Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had almost no regeneration of *Pinus sylvestris* and *Quercus* spp., or did not feature any regeneration of these species at all.
-
-Lastly, there is no model that performed well across the five species regarding the species-specific regeneration levels in the environmental space (Figure \@ref(fig:diff)).
-Some models (e.g., ForClim11, PICUS) tended to consistently overestimate the regeneration share of all five species, but most models overestimated the regeneration share of some species while underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
-
-```{r diff, echo=FALSE, fig.cap="Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold.
-The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient.
-The sizes of the circles represent the ratio between the regeneration basal area of the species and the total regeneration basal area for all species.
-The absence of a circle indicates a zero basal area share in the regeneration, or the absence of regeneration altogether.
-The color gradient (for the models only) shows the difference between the simulated and observed ratio of regeneration basal area of the species and the total regeneration basal area for all the species.", out.width='100%', out.height= '85%'}
+\newpage
+
+```{r diff, echo=FALSE, fig.cap=".", out.width='100%', out.height= '85%'}
 knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
@@
 
-# Discussion
-
-Tree regeneration is a fundamental process in forest dynamics.
-Correctly capturing tree regeneration in dynamic models is fundamental to, for example, evaluate post-disturbance dynamics and potential long-term recovery trajectories as it will define the forest state in the first decades [@seidl2022].
-If this initial phase is not captured well, we can not properly assess aspects such as forest resilience or the timing, magnitude and progression of carbon sequestration.
-In this study, for the first time, projections of tree regeneration from multiple models of forest dynamics were confronted with a unique dataset from unmanaged forest reserves across a large environmental gradient in Europe.
-
-The EuFoRIa data [@käber2023] are exceptional, particularly with respect to the number of records (number of sites and repeated measurements), which is essential for capturing a highly 'noisy' process such as tree regeneration.
-The use of this dataset for model benchmarking provided novel insights on the ability of state-of-the-art models to accurately simulate regeneration levels, species composition, and mortality in an early stage of tree life, i.e., between a tree diameter of 7 and 10 cm.
-Overall, by adopting this approach a much broader understanding resulted than if we had used these data for model calibration: it is primarily from the shortcomings of the models that we can gain ecological insights [cf.
-@trugman2022].
-
-## Regeneration levels
-
-Most of the models overestimated tree regeneration levels.
-This has potentially far-reaching implications, e.g., regarding biomass (and thus carbon) turnover, with a potential overestimation of the capacity for forest carbon sequestration [@pugh2019; @pan2011].
-Yet, we focused on one specific stage, i.e., regeneration into the 7 and 10 cm diameter classes.
-Trees at this size contribute little to carbon sequestration, and if the excess regeneration at this stage is compensated -- in the models -- soon thereafter by higher mortality, overall simulated regeneration may still yield trustworthy results at the level of stand structure and related forest attributes such as basal area or biomass.
-
-From a modeling point of view, excess regeneration inevitably requires excess mortality rates in a later stage, either -- as observed for some models -- between the 7 and 10 cm diameter thresholds, or soon after the 10 cm threshold has been crossed.
-In any case, correcting at early stages for the expected forest densities at later stages is equivalent to compensating for a first error (excess regeneration) by a second error (excess mortality).
-It is highly likely that biased projections will result, because the two errors are unlikely to be perfectly linked and thus will not always compensate each other.
-Hence, this structural problem of most models investigated here is problematic particularly if the models are to be used under novel conditions such as under climate change [e.g., @huber2021] or in a decision support context [e.g., @thrippleton2021].
-
-Regeneration levels define the structure and composition of future forests, but it is equally important to correctly identify areas where regeneration is lacking [@rammer2021].
-There are multiple constraints to the regeneration niche of tree species [@price2001], and therefore the absence of regeneration is likely to be common [@fortin2007], even over larger areas such as the one-hectare plots used here.
-Tree regeneration data are characterized by zero-inflation.
-This was clearly evident from the EuFoRIa dataset, but some of the models did not produce zeros at all, or featured a very low proportion of zero data.
-This substantial difference may be due to the fact that the simulation results were drawn from equilibrium forests, whereas in reality many of the forest reserves are recovering from past management activities and have become denser over the past decades [e.g., @heiri2009], leading to less regeneration than in an equilibrium situation.
-
-## Species composition of regeneration
-
-Correctly capturing species composition in tree regeneration is important to assess the future functional diversity of a forest, e.g., its sensitivity to drought or resilience to disturbances [@redmond2015; @seidl2022].
-In the simulations, overall species diversity levels in the regeneration were well within the observed range for half of the models (7 out of 14).
-Thus, while most of the models are facing difficulties to quantitatively match regeneration levels (cf.
-above), their performance is better when it comes to matching the composition of regeneration as a function of abiotic and biotic conditions.
-Most models maintained or even decreased regeneration diversity between the 7 and 10 cm thresholds, and the same is visible from the empirical data, but the differences were not significant.
-Diversity variations may be explained by the way the individual models consider regeneration processes [@konig_tree_2022], e.g., without considering the feedback from adults, regeneration can be linked to either seed rain from dispersal or a seed/seedling bank; neither process is incorporated across many models studied here, however.
-
-The species diversity of the entire stand was captured better by the models than the diversity of tree regeneration, and only a few models overpredicted stand-level diversity.
-Defining regeneration composition as being proportional to adult abundance, regardless of productivity, might be a simple and conservative assumption to safely maintain relative species abundances [@hanbury-brown2022], although this may be simplistic.
-Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy captured better the species diversity of the regeneration than those without feedback, with the exception of ForCEEPS that significantly improved regeneration species diversity with the model version that included feedback (ForCEEPS(f)).
-The similar performance of models with and without feedback is likely because the models put more weight on the regeneration niche arising from abiotic and biotic filters, than from the habitat niche of the adult trees [@grubb1977].
-
-## Regeneration mortality
-
-There are several factors that lead to mortality during the regeneration phase, such as competition [@casper1997] and multiple abiotic factors [@schmid2021; @cunningham2006].
-As mentioned above, it is reasonable to expect that models that overestimate regeneration may have a particularly high mortality between the 7 and 10 cm thresholds.
-However, this was not consistently evident from the simulations.
-This implies that these models must have an excess mortality in later stages, if we assume that all models were able to capture the structure and composition of the adult stands along the EuFoRIa gradient; this however was not tested here.
-
-Higher mortality towards the adult phase has important implications for forest dynamics and the goods and services provided by forests.
-On the one hand, mortality in later stages may erroneously enhance the share of less shade tolerant species in the models [@klopcic2015], therefore shifting the species composition.
-Unrealistic high stem densities for a longer period of time may overestimate the role of tree regrowth in carbon sink dynamics [@pugh2019].
-Given our set of simulations and analyses, we cannot conclusively assess what is happening in the models, and further studies are required that focus on a wider range of tree sizes and the fate of tree regeneration along such a size continuum.
-
-## Model performance and model traits
-
-In spite of the critical considerations above, it is remarkable that most models did not deviate exceedingly from the observations with respect to simulated regeneration levels -- after all, few if any of the models' regeneration routines are sufficiently constrained by data, with the exception of the empirical models xComp and SIBYLA.
-It is noteworthy that in spite of their empirical basis, these two models did not match empirical regeneration levels, in a similar magnitude as the other models and even in opposite directions (SIBYLA: underestimation; xComp: overestimation).
-It appears that using empirical data of limited geographical scope and constrained to managed forests, as done in these models, leads to extrapolation problems already under current climatic conditions.
-For the other models, whose regeneration modules are not strongly constrained by empirical data, the degrees of freedom that are available for modeling regeneration [@konig_tree_2022] could have implied that model performance would be much worse than what we found.
-
-Our study showed that increasing complexity in the regeneration modules is not linked with a higher accuracy of the projections of regeneration levels, species composition or mortality at early tree stages, as there was no significant relationship between model performance and model complexity.
-Increasing complexity in regeneration modules has been motivated by better process understanding and enhancing of model accuracy [@bugmann2022].
-However, more complex models do not necessarily lead to better projections and rather to a reduced transparency and lower predictive power [@franklin2020].
-Thus, the question of what level of detail is appropriate and parsimonious for modeling tree regeneration processes remains open [@konig_tree_2022; @bugmann2022].
-
-## Regeneration gradients on regeneration niches
-
-Competition for light as a strong filter for tree regeneration has been widely documented [@messier1999; @collet2006; @berdanier2016], but the models examined here did not reproduce this expectation.
-However, it is difficult to measure light availability at large spatial and temporal scales.
-We used total stand basal area as a proxy for light availability [cf.
-@schmid2021].
-However, we were unable to consider light availability restrictions caused by ground vegetation, which may be an important filter for some EuForIa stands [@woltjer2008].
-We found pronounced differences in the stand basal area ranges simulated by the models, but also between models and observations.
-This made it impossible to evaluate the regeneration for the extremes of the stand density ranges in some models.
-For example, regeneration levels at low stand densities are relevant to assess how well forests are recovering e.g.
-after gap creation due to disturbance [@seidl2022; @grubb1977].
-At the level of the simulated one-hectare samples, average basal area is typically not very low as long as no larger disturbances occur, which was explicitly excluded in our simulation protocol.
-Towards the other end of the spectrum, i.e., with increasing stand basal area, it would be reasonable to expect that the regeneration of the different species would become sparser and drop out entirely under low-light conditions [@klopcic2012; @zell2019].
-However, few models showed this trend, thus indicating that the relation between regeneration and light availability is not yet captured correctly in most models.
-Yet, several of the models that did feature an increase of regeneration with increasing basal area include a feedback between seed production of mature trees and regeneration.
-Thus, it seems that higher light competition does not sufficiently compensate for increased seed availability with higher basal area due to a higher abundance of mature trees.
-
-There were pronounced differences how the main tree species were represented by the models along the environmental gradients, in particular the dominant species *Abies alba*, *Fagus sylvatica* and *Picea abies*.
-The regeneration levels were sampled from simulations in the equilibrium, and in this state it is expected that non-climax species such as *Pinus sylvestris* or *Quercus* would be of minor importance, or absent [@klopcic2015].
-Most models captured this low abundance, which is also found in the empirical data.
-Thus, although the broad patterns are matched by many models, improvements in the quantification of the regeneration niche of the species are needed, but this cannot be done in the absence of robust datasets across multiple stand development stages.
-
-## Methodological considerations
-
-The EuFoRIa data as used here are unique and well suited to better understand tree regeneration.
-However, three aspects of these data may represent considerable limitations.
-First, we made a comparison of tree regeneration in an equilibrium state, but we cannot assess how close the forests included in the EuFoRIa dataset were to such a state.
-The data were collected in forest reserves where no management has taken place for long periods of time.
-This makes our assumption of an equilibrium between forest properties and environmental drivers more reasonable than it might appear at first sight.
-In an analysis of primeval forests at demographic equilibrium, @brzeziecki2021 found higher regeneration rates than the ones observed in the EuFoRia data.
-Thus, the overestimation of regeneration rates by the models may not be so problematic.
-Second, the data were collected for regeneration above a 7 cm threshold, thus limiting the assessment of tree regeneration to a holistic viewpoint.
-This constituted a hard limit based on which we can understand only some aspects of tree regeneration, which in its entirety often comprises a rather long period since seed production [@price2001].
-In reality, many environmental constraints are acting on young trees [@käber2021] that we were unable to assess.
-Yet, regeneration data with lower thresholds are simply not available in a harmonized manner across large environmental gradients.
-Third, the empirical data were collected from rather small plots, while we sampled simulated regeneration levels from 1 ha areas, which may lead to an incorrect representation of space.
-Even though the strategy we adopted is not ideal, it represents a common challenge when harmonizing diverse data sources originating from varied sampling strategies [@portier2022].
-It would have been extremely challenging for such a variety of models to follow a protocol where the spatial sampling size was different at each of the 200 sites, and it would have introduced additional uncertainty in the results.
-Ultimately, we need to acknowledge that the observed data represent just a snapshot of an extremely complex and stochastic reality.
-This may at least partly explain e.g.
-the lack of clear patterns in the simulation results along the gradients of climate water balance and degree-days.
-
-The design of our sampling protocol did not include spatial aspects such as seed dispersal or detailed soil data.
-While we considered a wide range of models of forest dynamics, from stand to global scales, the simulation setup was limited to 1 ha regeneration values.
-This lack of consideration of spatial scale is appropriate for stand-scale models, but it potentially puts the landscape models at a disadvantage, as they have been built to be accurate at the landscape level.
-Without spatial context, we are limiting regeneration to the seed influx from the stand itself, unless the model has a background seed input.
-Yet, the global models should not be at a disadvantage due to the limited spatial considerations [but cf.
-Snell et al.
--@snell2018; or Lehsten et al.
--@lehsten2019], as they are usually lacking dispersal between cells and are based on a strong abstraction of horizontal space [@hanbury-brown2022].
-Lastly, detailed data on soil conditions were not available from the observed data, and independent, admittedly rather coarse data for soil and climatic water balance had to be used instead.
-It is noteworthy that many models represent drought based on detailed indicators based e.g.
-on soil water holding capacity, which had to be derived from a rough soil quality measure.
-This may at least partially explain the unsatisfactory performance of many models along the drought axis (i.e., climatic water balance).
-
-## Research recommendations
-
-With our study, we have demonstrated that models of forest dynamics need a focus on tree regeneration modules to make them more robust.
-It remains uncertain what level of detail is required to model regeneration processes, and this must be addressed in future research.
-We recommend that the improvement of the regeneration modules is implemented as additional features that can be traced back, as done here for the variants of ForClim and ForCEEPS, and that model complexity and structure must always be connected with modeling objectives [@albrich2020].
-If it should be necessary to include more detail in the regeneration models, this will come with higher parameterisation efforts.
-This will most likely lead to lower generalization because the required data will have to be collected from specific locations, as currently there is no general, comprehensive regeneration dataset available.
-
-Therefore, we further recommend that more effort should be invested into collecting harmonized datasets on tree regeneration.
-We emphasize that datasets such as EuFoRIa are invaluable and should be expanded in both their spatial extent (e.g., towards boreal and Mediterranean conditions) as well as in time (e.g., continuing the monitoring into the future).
-Such data will allow for a better evaluation of forest models and help to reduce the uncertainty in their projections, which is crucial when they are used as tools for predicting e.g.
-the impacts of anthropogenic climate change.
-
-In the present study, we have considered tree regeneration in the equilibrium state only.
-It is equally important to understand how these models project tree regeneration after changes in forest structure by disturbances [@seidl2022], or under different management strategies [@lindner2000].
-However, this will require an entirely different set of observed data, and potentially not all models of forest dynamics would be able to assess the relationship of these aspects on tree regeneration, e.g.
-due to the lack of disturbance or appropriate management modules.
-
-We recommend to investigate in detail the implications of the current modeling strategies for tree regeneration and, ultimately, simulated forest stand structure.
-This applies particularly to the erroneous patterns of excess tree regeneration and later excess mortality, by focusing on a wider range of tree sizes and the related regeneration dynamics.
-We also recommend, especially for the landscape-level models, the inclusion of explicit spatial considerations regarding tree regeneration [@beckage2003]; this, however, is a serious challenge regarding appropriate datasets, as inventory data have wide coverage, but by definition do not allow for the assessment of spatial interaction effects.
-
-Exercises like the one presented here, where the models are operated in "blind flight" mode, i.e., without the possibility of being tuned towards capturing the expected patterns, should be repeated.
-Such benchmarking exercises should next focus on aspects such as specific model traits and the ecological formulations of particular (sub-)processes, to better understand the implications of the assumptions on which the models are based.
-Furthermore, the inclusion of a wide range of models with different scales, type, population structures, tree regeneration modules and complexity will ensure a wider benefit to the whole modeling community.
-
-## Conclusions
-
-Models of forest dynamics are important tools in science and decision support, and the formulation of tree regeneration has strong implications for simulated forest properties.
-The 15 models and variants used here are facing similar challenges in their representation of tree regeneration: they generally overestimate tree regeneration levels, and the simulated regeneration niche is not always captured accurately as a function of biotic (light) and abiotic (temperature and moisture) factors.
-
-However, most models properly capture the diversity of the initial tree community, and differences between model formulations, e.g., the presence or absence of feedback from the adult trees, did not have a strong effect for capturing the species composition of regeneration.
-
-Regarding mortality in the early phase of tree life, many models that feature a particularly high overestimation of regeneration levels are compensating for this by a larger tree mortality.
-Often, this compensation is not sufficient to reduce the high regeneration levels to realistic values.
-Overall, there is no clear mortality pattern across all models.
-
-When capturing tree regeneration, the specific design decisions taken in the development of any model are more important for its behavior (accuracy) than scale (stand, landscape, global), modelling approach (empirical vs.
-process-based), and complexity.
-Having both empirical and process-based models in our set, the empirically-based models could have been expected to have a better performance, as they were calibrated with inventory data, but this was not the case.
-Similarly, higher model complexity does not represent an improvement for capturing tree regeneration.
-
-Even though the regeneration routines of most of the models investigated here have never been sufficiently constrained by robust data, the regeneration projections are not overly off.
-This indicates that a lot can be gained by a focus on the modeling of regeneration processes.
-The representation of forest dynamics in these models would become much more robust particularly in the face of climate change and post-disturbance dynamics, thus strongly reducing the uncertainty in long-term projections of future forest dynamics.
-
-# Author contributions
-O.
-Díaz-Yáñez and Y.
-Käber contributed equally and share the first authorship.
-H.
-Bugmann, Y.
-Käber, and O.
-Díaz-Yáñez developed the study design.
-All authors contributed to preparing the simulations with their respective models.
-O.
-Díaz-Yáñez led the data analysis with contributions during the workshop from Y.Käber, T.
-Anders, K.
-H.
-Braziunas, J.
-Brůna, S.
-M.
-Fischer, J.
-Hetzer, T.
-Hickler, H.
-Lischke, M.Mahnken, P.
-Mairota, K.
-Merganičová, T.
-Mette, X.
-Morin, W.
-Rammer and D.
-Scherrer
-H. Bugmann. O. Díaz-Yáñez and H. Bugmann led the writing of the manuscript. All authors participated in the revision of the manuscript and approved its submission.
-
-# Acknowledgements
-
-This paper arises from a workshop held in Davos (Switzerland) in June 2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
-Funding for the workshop was provided by COST Action CA19139 PROCLIAS (PROcess-based models for CLimate Impact Attribution across Sectors), supported by COST (European Cooperation in Science and Technology; [https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research Grant ETH-35 18-1).
-ODY was funded by the Swiss National Science Foundation project 200020_188882 / 1.
-KB received support from the European Research Council under the European Union's Horizon 2020 research and innovation program (Grant Agreement 101001905).
-JB acknowledges funding from the long-term research development project no.
-RVO 67985939 from the Czech Academy of Sciences.
-MM acknowledges financial support from I-Maestro (Innovative forest management strategies for a resilient bioeconomy under climate change and disturbances, grant no.
-22035418, 2019--2022) funded by the ERA-NET Cofund ForestValue.
-MMina acknowledges funding from the H2020-MSCA-IF project REINFORCE (Grant No.
-891671).
-KM received funding from the Project No.
-CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant no.
-ITMS2014+313011W580 supported by the Integrated Infrastructure Operational Programme funded by the ERDF.
-CH acknowledges financial support from Waldfonds WASIM2100 project.
-We would also like to thank all the people that helped to make this study possible with contributions at different stages such as Martin Gutsch and Giorgio Vacchiano.
-
-# Conflict of interest statement
-
-The authors declare no conflicts of interest.
-
-# References
-
-```{=html}
-```
+\newpage

### 2023-05-22 manuscript/supplementaryRW.Rmd

@@
 ---
-title: "Supplementary figures\nCapturing regeneration in models of forest dynamics"
+title: "Appendix figures\nTree regeneration in models of forest dynamics: a key priority for further research"
 author: 
@@
 linkcolor: blue
-csl: global-change-biology.csl
+csl: chicago-author-date.csl
 editor_options: 
@@
     wrap: 72
+fontsize: 12pt
+linestretch: 2
 header-includes:
@@
 - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+- \usepackage{fontspec}
+- \setmainfont{Times New Roman}
 ---
@@
 
-# Supplementary Material 1: Model reports
+# Appendix 1: Model reports
 
@@
 ```
-# Supplementary Material 2: Supplementary figures and tables
+# Appendix 2: Supplementary figures and tables
 

### 2023-05-22 manuscript/manuscriptRW.Rmd
add table 1 to manuscript folder and change path in the manuscript

@@
 library(knitr)
-modelDF <- readr::read_csv(here::here("data", "modelsDescription.csv"))
+modelDF <- readr::read_csv(here::here("manuscript", "modelsDescription.csv"))
 modelDF$Sampling <- NULL

### 2023-06-13 manuscript/manuscriptRW.Rmd
no relevant changes

@@
 subtitle: "\\fontsize{16pt}{3pt}\\selectfont \\textbf{\\textit{Tree regeneration in models of forest dynamics: a key priority for further research}} "
-title: "\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to the journal Ecological Monographs. Manuscript type: Article}}"
+title: "\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to Ecosphere. Manuscript type: Article}}"
 author: 
@@
 
-OPEN RESEARCH STATEMENT: This submission uses novel code, both the data
+**OPEN RESEARCH STATEMENT:** This submission uses novel code, both the data
 and code are provided in an external repository to be evaluated during
 the peer review process. The repository can be downloaded from
-[here](https://polybox.ethz.ch/index.php/s/q5MOB1CLcjARs3H) (password:
-DownDataCodeRegen1). The protocol is available at @díaz-yáñez2022. We
-will make publicly available the code and data after the revision.
+[this figshare private link: https://figshare.com/s/100857ba67d2351dcb15](https://figshare.com/s/100857ba67d2351dcb15).
+The protocol is available at @díaz-yáñez2022 ([with link: https://osf.io/czdxp/](https://osf.io/czdxp/ )).
+We
+will make publicly available the code and data if the manuscript is accepted.
 

### 2023-06-13 manuscript/supplementaryRW.Rmd
no relevant changes

@@
 ---
-title: "Appendix figures\nTree regeneration in models of forest dynamics: a key priority for further research"
+title: "Appendix"
+subtitle: "Tree regeneration in models of forest dynamics: a key priority for further research"
 author: 
@@
 - Harald Bugmann
-date: "`r Sys.Date()`"
+date: ""
 output: 

### 2023-06-13 code/figures.R

@@ -1408,13 +1408,13 @@
     ###### DO NOT RUN #########
 
     #Add location
-    location <- data.table::fread("data/coords_blurred_dt4326.csv")
+    #location <- data.table::fread("data/coords_blurred_dt4326.csv")
 
-    domSpp7 <- merge(standSpp, recruitSpp7, by = "site")
-    domSpp7 <- merge(domSpp7, location, by.x = "site", by.y = "cluster_plot_id200")
+    #domSpp7 <- merge(standSpp, recruitSpp7, by = "site")
+    #domSpp7 <- merge(domSpp7, location, by.x = "site", by.y = "cluster_plot_id200")
 
-    domSpp10 <- merge(standSpp, recruitSpp10, by = "site")
-    domSpp10 <- merge(domSpp10, location, by.x = "site", by.y = "cluster_plot_id200")
+    #domSpp10 <- merge(standSpp, recruitSpp10, by = "site")
+    #domSpp10 <- merge(domSpp10, location, by.x = "site", by.y = "cluster_plot_id200")
 
     # Species to be plotted
     #unique(domSpp10$stand)

### 2023-09-18 manuscript/review/authorsResponse.qmd
added basic structure for review

---
title: Authors Response to Reviewer Comments
subtitle: "ECOSPHERE Manuscript ID (ECS23-0342) Tree regeneration in models of forest dynamics: a key priority for further research"

author:
  - name: Olalla Díaz-Yáñez*,
    attributes:
        corresponding: true
  - name: Yannek Käber,
  - name: Tim Anders,
  - name: Friedrich Bohn,
  - name: Kristin H. Braziunas,
  - name: Josef	Brůna,
  - name: Rico Fischer,
  - name: Samuel M. Fischer,
  - name: Jessica Hetzer,
  - name: Thomas Hickler,
  - name: Christian	Hochauer,
  - name: Manfred J. Lexer,
  - name: Heike	Lischke,
  - name: Mats Mahnken,
  - name: Paola	Mairota,
  - name: Ján Merganič,
  - name: Katarina Merganičová,
  - name: Tobias Mette,
  - name: Marco Mina$,
  - name: Xavier Morin,
  - name: Werner Rammer,
  - name: Christopher P.O. Reyer,
  - name: Simon	Scheiter,
  - name: Daniel Scherrer,
  - name: Harald Bugmann

date: last-modified
link-citations: true
linkcolor: blue
format:
   pdf: 
     toc: true
---

## Journal formatting

* Respond to the review comments in the space provided. You can use this space to document any changes you make to the original manuscript. In order to expedite the processing of the resubmitted manuscript, please be as specific as possible in your response to the review comments. If you disagree with a point, please explain why. Please note that the field in ScholarOne does not retain type formats such as italics, boldface, or colors, so please format the responses accordingly. We suggest you upload a separate file for your "Author's Response to Reviewer Comments" document and use our template: <https://www.esa.org/wp-content/uploads/2021/04/Author-Response-to-Reviewers-Template.docx>.

* A version uploaded in the file list should be given the designation "Additional File for Review but NOT for publication".\

* Please also include a "track changes" version of your file, along with a clean copy as your Main Document. This "track changes" version should be uploaded in the file list list and given the designation "Additional File for Review but NOT for publication".

* IMPORTANT: Your original files are available to you when you upload your revised manuscript. Please replace files of the earlier version and delete any redundant files before completing the submission.

* To submit your revised manuscript, log into <https://mc.manuscriptcentral.com/ecosphere> and enter your Author Center. You will find your manuscript title listed under "Manuscripts with Decisions." Under "Actions," click on "Create a Revision." Your manuscript number has been appended to denote a revision. Please DO NOT upload your revised manuscript as a new submission.

* DEADLINE: We encourage you to submit your revised manuscript within three weeks from today's date. The final due date for submitting your manuscript with the same manuscript number is 14-Dec-2023.

* If you are unable to meet the final deadline, you can request an extension by contacting the the journal's editorial staff ([esajournals\@esa.org](mailto:esajournals@esa.org)). Please note that we can only consider extensions of a few weeks. If you cannot submit by the final due date, you must create a new submission for your manuscript. If creating a new submission, be sure to note the previous manuscript ID (ECS23-0342) in the online submission form and include an "Author's Response to Reviewer Comments" document (using the template linked to earlier in this message) in the file list.

* Editorial staff note: In addition to revisions according to the review comments below, please make revisions to your paper to conform to the journal style at this stage, especially the tables, figures, and supporting information, anticipating the possibility that your paper may be accepted for publication.

* Please provide the main manuscript file composed in Word, taking care to use the Table function for the tables. However, if you created your manuscript in LaTex, please provide only the .tex file of the main manuscript, a .pdf file of the main manuscript (that is the exact same in content as the .tex file), and a .bib file.

* Ecosphere page style allows for five levels of headings. Please use different type styles (italics, boldface, boldface italics, etc.) consistently to distinguish between the levels of headings in the article and delete the section numbers in the headings.

* Literature references: Please check that each citation has complete information, and check that all references in text have a matching citation in the References section and vice versa.

* Table and figure references: Please be sure that all tables and all figures have in-text citations in the main body of the manuscript, and that tables and figures are called out in numerical order (i.e., Figure 1 is called out before Figure 2).\

* Figures: Each figure should be provided in an individual, complete file (all panels in one image, identified by figure number in the file name), preferably as a high-resolution (300-600 dpi) image file or high-quality PDF. Preferred file formats include TIF, EPS, PDF, or AI at 600 ppi, while JPEG, PPT/PPTX, or DOC/DOCX are acceptable if the resolution is sufficient. Only one version of each figure should be provided. Please delete figure images from the main manuscript file. Please check for misspellings and be sure that all parts, including all lettering, symbols, labels, and scale numbers, are discernible, clear, and readable when the figures are presented at the maximum allowable size (18 cm wide by 22 cm tall) in the typeset and paged PDF version of the article. See additional instructions for preparation in the Manuscript Preparation Guide.

* Figure captions: Please be sure to include a list of figure captions in the main manuscript file and ensure that any abbreviation that appears in a figure is identified in the caption.

* Supporting Information: Each appendix must be assembled exactly as you intend it to appear online. (Appendices are posted "as is"; ESA and Wiley staff do not copy edit or assemble appendices.) Note also that the Ecosphere naming convention for appendices is of the style "Appendix S1," "Appendix S2," etc., and for tables and figures within appendices is "Table S1," "Table S2," etc. or "Fig. S1," "Fig. S2," etc., and requires that the numbering begins at "S1" in each appendix (and revise references to material in other appendices as necessary). Please be sure to eliminate line numbering and to include the title of the journal, the title of the article, and all of the authors' names on the first page of each appendix. In addition, if there are literature references within an appendix, the appendix should include a list of literature cited.

* If the journal style is not followed at submission of the revision, we may return the manuscript to you for further revision before sending it along to the Subject-matter Editor. Closely following our manuscript preparation guide at this stage would expedite the production of your paper for publication, should it be accepted. You can download our guide here: <https://www.esa.org/wp-content/uploads/2022/05/ESA-Manuscript-Preparation-Guide.pdf>

## Response to the Editor

Dear Dr. Díaz-Yáñez:

Thank you very much for submitting your manuscript "Tree regeneration in models of forest dynamics: a key priority for further research" for review by Ecosphere. The reviewers and I appreciate the work you have accomplished. I am willing to consider a revised version for publication in the journal, assuming that you are able to modify the manuscript according to the reviewers' extensive recommendations. Your revisions should address the specific points made in the review comments.  Please note that Reviewer #2 made extensive comments directly on the manuscript.
\
I should note that personally I can hardly disagree with your basic conclusion that models of forest dynamics need to do a good job of capturing regeneration and recruitment processes.  I have devoted most of my research over the past 40 years to developing the field methods to allow incorporation of accurate seedling, sapling and canopy recruitment in the models SORTIE and SORTIE-ND, in systems ranging from the tropics to boreal forests. The specific ecological processes incorporated in those models depended on the goals of the modeling and the ecology of the system, and included processes such as masting cycles, effects of disturbance on seedbed substrate dynamics, seed and seedling predation by small mammals, browsing by ungulates, Janzen-Connell effects, and, of course, detailed analyses of shading and neighborhood competition. More recently, this includes developing statistical models of seedling recruitment from the US forest inventory network that have been directly incorporated in the model for forests of the entire eastern US.  To my mind, the difficulties the models you reviewed encountered in capturing canopy recruitment speak more to the limitations of the approaches used in model development than in any inherent difficulties in capturing regeneration and recruitment in models of forest dynamics.

Sincerely,

Dr. Charles Canham
Subject-matter Editor, Ecosphere

## Response to reviewer 1

### General Comments

The manuscript compares how regeneration and recruitment is simulated in 15 forest models, using a large network of forest reserves in Europe as the observed data. I want to start by saying there are many strengths of the paper. I was impressed with the comparison of 15 models, that the observed data was not used to parameterize or fine-tune the models, and the high quality of the writing. The introduction and discussion were easy and enjoyable to read, and the topic is critical and relevant. Some of the details about the models and methods were placed in the very substantial appendix, and I agree with their choices. However, the manuscript is still quite long with lots of multi-paneled figures. If the journal does not have a page limit, then this is not an issue. If there is a page limit, then some tough choices will have to be made for which figures to keep. Figures 2, 3, 6 and 7 are the most important (in my opinion).

I have a few major comments. One of which requires a different statistical analysis -- which is why I do not consider this a 'minor revision'. However, there are no fatal flaws, and all of my comments are fixable. I hope they serve to improve the manuscript.

##### Our general comments to reviewer 1:

### Major comments

#### COMMENT 1

'regeneration' versus 'recruitment'. I personally have had this conversation multiple times over the years, so I understand that there is variation in the definitions and uses. However, the authors do define these terms in the methods (L167-171) and I completely agree with them. Regeneration is for flowers, seeds, and seedlings while recruitment is for reaching a certain size threshold. However, the entire manuscript refers to regeneration as "passing of trees across a specific diameter threshold ("ingrowth") (L142), either reaching a size of 7 or 10 cm DBH. This is recruitment. Since this is the response variable for the manuscript, I would suggest replacing 'regeneration' with 'recruitment' throughout the manuscript. \

##### Response to comment 1:

#### COMMENT 2
The difference between 7 cm and 10 cm DBH in the observed data. From the methods, there is one dataset that has sites where recruitment was defined as 7 cm DBH, and then another where recruitment was somewhere between 7 -- 10 cm DBH. It was not clear how these were treated and compared in the results. I assume that the models simulated all 200 sites, but then the recruitment results were compared like-to-like. So, if a site defined recruitment as DBH = 8.0 cm, then all 15 models would compare simulated recruitment at that site only at 8 cm. Since the data are boxplots, I am not able to tell. For example, Figure 2 -- there should be 165 points for all of the 7 cm bars (observed and simulated) and 35 data points for the 10 cm bars (observed and simulated). If true, then please clarify. If not true, then please explain. Also, how was the data analysis performed when looking at the change in mortality and species diversity between 7 to 10 cm? I assume only the 165 stands were used. Please clarify. \

##### Response to comment 2:

#### COMMENT 3
The t-tests. The results in Table S2 and S3 (comparing at 7 cm and 10 cm DBH) should be at least by paired t-tests not two-sided t-tests (L319), where you are comparing the difference within the same stand at 7 and 10 cm. The results Table S5 and S6 should not be a t-test at all. Performing 15 individual t-tests is a classic Type-1 error situation (technically this applies to Tables S2 and S3 as well, a comment on that below). For Tables S5 and S6, I would suggest a general linear mixed effects model. You can take the difference between the observed, and the simulated output and make this the response variable (Y). Then, include model complexity, feedback, model type and scale as fixed explanatory variables. And include the stand ID as the random effect. This will more accurately answer your question about if any of these fixed explanatory variables can predict over or under representation of recruitment. For Tables S2 and S3, you could do the same thing (i.e., calculate the difference between the observed and the simulated output for each model) and make this your response variable. But the only fixed effect would be the actual model, and then you could do an ANOVA (again, stand ID as the random effect) and then a post-hoc test to see which models were or were not different from observed. If you do want to keep the 15 paired t-tests, then you need to account for Type I error rates and adjust your p-values accordingly.\

##### Response to comment 3:

### Minor comments:

#### COMMENT 4
The abstract is not clear if the models do a good or poor job in simulating regeneration (L59, that they match observed ranges, and then L66, considerable mismatch between simulations and observed data, but that they do capture tree regeneration). Please clarify the main message in the abstract.

##### Response to comment 4:

#### COMMENT 5
L123-138: this paragraph has a lot of "on the other hand" type of statements, and it does not lead nicely into the next paragraph about the current study. It was a bit of a surprise (i.e., I did not expect the main purpose of the study when I started to read the following paragraph). It seems like the main issue is that we just don't know how uncertain regeneration and recruitment methods are, because we lack the comparison across models and a comprehensive empirical recruitment data set that we can use to compare. And that is what your study is doing.

##### Response to comment 5:

#### COMMENT 6
Table 1. Please clarify what the letters are for (a, b, c and d in the table headers). For models that start from saplings or ingrowth, please add in the size as well (i.e., start from saplings with DBH = 1.0). Could also add in the classification scheme mentioned in the methods (L168, regeneration or recruitment) to Table 1. \

##### Response to comment 6:

#### COMMENT 7
L218 -- do I understand this correctly, that regeneration rates are a mean of 56 trees per ha per decade? This seems really low (assuming that this is for \>10 cm DBH).  \

##### Response to comment 7:

#### COMMENT 8
L264 -- 274: how will this be accounted for later, that not all models simulated the same set of species?

##### Response to comment 8:

#### COMMENT 9
L340: "regeneration basal area of that species" is mentioned twice, please delete the second one.

##### Response to comment 9:

#### COMMENT 10
L364: so please remove the 4C model from the figure as it is not comparable. Do you need to also remove other models that don't simulate the same number of species too?\

##### Response to comment 10:

#### COMMENT 11
Table S2 and S3: What are the 3 columns for p-values? Please report just the one you used. No need to have a column to state 't-test' -- just move this to the table caption. Then add in a column for df, the test statistic (t) and the effect size (was it higher or lower and by how much?). NOTE, this comment may become irrelevant if you change the statistical analyses.

##### Response to comment 11:

#### COMMENT 12
Figure S1 and S2 -- I can't tell the difference between observed and simulated. Are they different colours? The legend should be larger. What is the black line for?

##### Response to comment 12:

#### COMMENT 13
Figure 4 (and Figure S3) are interesting in that there are no patterns with overstory diversity, and that this was not explained by feedback with seed production. What happens if you colour the points, by those from models with the feedback? Do you address in the discussion why this feedback makes no difference to recruitment? \

##### Response to comment 13:

#### COMMENT 14
All figures in the Appendix, please make the X and Y axis labels larger and centered. I actually thought labels were missing initially and it took a while to find them.

##### Response to comment 14:

#### COMMENT 15
L400 -- 402: where is this conclusion about mixed species or monocultures coming from? I don't see that in the figures or the results.

##### Response to comment 15:

#### COMMENT 16
Figure 6 is excellent.

##### Response to comment 16:

#### COMMENT 17
Table S4: please clean up. No need to report 7 places after the decimal point. Please clarify in the table caption what the slope represents (maybe mention Figure S5 too, as I assume it is illustrating the slopes). Another suggestion, would be to make the slopes solid lines if significant, or dashed lines if not significant. Not sure if that would help, but if it does then it could eliminate the need for this table.

##### Response to comment 17:

#### COMMENT 18
Not sure both Figure 7 and Figure S7 are necessary (I admit I spent a few minutes going back and forth between them trying to understand how they differed). I would just keep Figure 7 and note that the scales differ.

##### Response to comment 18:

#### COMMENT 19
L547-548: predicting stand diversity, did I miss this? I don't believe Figure 4 shows this, as if you put all data points on the same panel it would just be a random scatter of points that covers the entire space (not necessarily closer to the 1:1 line).

##### Response to comment 19:

#### COMMENT 20
L575: that most models did not deviate exceedingly from observations seems in contrast to what the previous section was stating. What is this based on?

##### Response to comment 20:

#### COMMENT 21
L584: What does this mean, about degrees of freedom for modelling regeneration?

##### Response to comment 21:

#### COMMENT 22
L624: Missing the end of the sentence.

##### Response to comment 22:

#### COMMENT 23
L647: Completely agree that observed data is a snapshot of a stochastic and dynamic system and care should always be taken with comparing to simulation data, however I disagree about the following stamen (L648). The models should be able to capture broader patterns along climatic or environmental gradients.

##### Response to comment 23:

#### COMMENT 24
L655: But the landscape models generally did about the same (i.e., not consistently worse) then the rest, correct?

##### Response to comment 24:

#### COMMENT 25
L663: But in general, the models were more sensitive to water balance and the observed data did not have any real pattern. Why would this coarser representation of water balance in the model, cause it to be more sensitive?

##### Response to comment 25:

#### COMMENT 26
L675: what would a 'comprehensive regeneration dataset" look like? There are large data sets for seed production (MASTIF, MASTREE, etc.) and seedling data from national forest inventory programs. Does EuFoRIa (L677) have everything you need? Is it just a matter of expanding the spatial and temporal extent? Please explain what sort of data that the models would need (imagine an ideal world).

##### Response to comment 26:

## Response to reviewer 2

### General Comments

The paper evaluates whether forest dynamics models are successful at capturing the essence of regeneration processes, by comparing outputs from 15 simulation models with tree recruitment data obtained from 300 mature forests spread across central/eastern Europe.  Specifically, the models predicted how many trees of different species recruited each year (i.e. grew larger than 7-cm or 10-cm stem diameter) and indices derived from these numbers were compared with analogous indices derived from repeat-measured forest inventory plots.  Most of the simulation models performed poorly for most of the indices examined.  In particular, they tended to overestimate recruitment rates.

It is valuable to demonstrate that leading forest dynamics models are unable to characterise regeneration processes very well.  I also thought the "blind trial" approach was clever.  Huge effort must have gone into running 15 models under various starting conditions and synthesising the outputs into a coherent paper.  Choices have been made about which details to include in the main body of the text vs supplementary information and I thought a good balance was struck, although I have made several suggestions on the pdf.

##### Our general comments to reviewer 2:

### Major comments

#### COMMENT 1
Where the paper falls short, in my opinion, is in providing a clear path forward.  It shows that the models are too simplistic to simulate regeneration processes accurately but provides few insights into how we can improve their performance.  The paper reports that creating more complex models is not valuable in and of itself, but that's not to say that intelligent improvement of specific subcomponents may hold the key.  For instance, deer browsing and competition with herb layer plants may well contribute to lower recruitment being observed than simulated, but how can we demonstrate that, and if the inadequate representation of these processes in the models is the problem, what should be done to improve the models?  The review by Hanbury-Brown et al. (2022) on "future forests within Earth system models: regeneration processes critical to prediction" provides a good summary of current knowledge and future directions, including thoughts on 1. reproductive allocation and seed production; 2. dispersal;  3. seed survival, germination, and resprouting; and 4. seedling survival and growth. I would encourage the authors to frame their discussion in a similar way.  Advanced statistical approaches that bring together simulation models and Approximate Bayesian Computation to estimate parameters of recruitment submodels may hold the key (e.g. <https://onlinelibrary.wiley.com/doi/full/10.1111/ecog.04824> ).  To me, a key question is how we can make better use of field measurements and statistical analyses to refine forest dynamics models, i.e. not only parameterise them but also to identify which submodels need refinement and the function that need inclusion.

##### Response to comment 1:

#### COMMENT 2
I felt the readability of the paper would be improved by introducing a "road mapping" paragraph at the end of the introduction or start of the methods section that sets out the approach in broad terms.

##### Response to comment 2:

#### COMMENT 3
The words "ingrowth", "regeneration" and "recruitment" are used interchangeably to describe the recruitment of new trees into the 7-cm diameter class.  I would encourage the authors to use the term "ingrowth" or "recruitment" to this to describe this process,  and keep "regeneration" as an overarching term describing the entire process of seed production, dispersal, early establishment, and onward growth.

##### Response to comment 3:

#### COMMENT 4
The article is written entirely by Europeans, unless I'm mistaken, and very much focused on the central/eastern European forestry literature.  I would encourage a deeper dive into the relevant North American and Asian literature when introducing and discussing their work.

##### Response to comment 4:

### Minor comments:
These comment have been extracted from the comments done by reviewer 2 in the PDF document of the manuscript.

#### COMMENT 5

##### Response to comment 5:

## Other changes made in the manuscript

### 2023-09-19 manuscript/review/authorsResponse.qmd
review documents and updates on the original manuscript

@@
 
-* Respond to the review comments in the space provided.
+-   Respond to the review comments in the space provided.
 You can use this space to document any changes you make to the original manuscript.
@@
 
-* A version uploaded in the file list should be given the designation "Additional File for Review but NOT for publication".\
+-   A version uploaded in the file list should be given the designation "Additional File for Review but NOT for publication".
 
-* Please also include a "track changes" version of your file, along with a clean copy as your Main Document.
+-   Please also include a "track changes" version of your file, along with a clean copy as your Main Document.
 This "track changes" version should be uploaded in the file list list and given the designation "Additional File for Review but NOT for publication".
 
-* IMPORTANT: Your original files are available to you when you upload your revised manuscript.
+-   IMPORTANT: Your original files are available to you when you upload your revised manuscript.
 Please replace files of the earlier version and delete any redundant files before completing the submission.
 
-* To submit your revised manuscript, log into <https://mc.manuscriptcentral.com/ecosphere> and enter your Author Center.
+-   To submit your revised manuscript, log into <https://mc.manuscriptcentral.com/ecosphere> and enter your Author Center.
 You will find your manuscript title listed under "Manuscripts with Decisions." Under "Actions," click on "Create a Revision." Your manuscript number has been appended to denote a revision.
@@
 
-* DEADLINE: We encourage you to submit your revised manuscript within three weeks from today's date. The final due date for submitting your manuscript with the same manuscript number is 14-Dec-2023.
+-   DEADLINE: We encourage you to submit your revised manuscript within three weeks from today's date. The final due date for submitting your manuscript with the same manuscript number is 14-Dec-2023.
 
-* If you are unable to meet the final deadline, you can request an extension by contacting the the journal's editorial staff ([esajournals\@esa.org](mailto:esajournals@esa.org)).
+-   If you are unable to meet the final deadline, you can request an extension by contacting the the journal's editorial staff ([esajournals\@esa.org](mailto:esajournals@esa.org)).
 Please note that we can only consider extensions of a few weeks.
@@
 
-* Editorial staff note: In addition to revisions according to the review comments below, please make revisions to your paper to conform to the journal style at this stage, especially the tables, figures, and supporting information, anticipating the possibility that your paper may be accepted for publication.
+-   Editorial staff note: In addition to revisions according to the review comments below, please make revisions to your paper to conform to the journal style at this stage, especially the tables, figures, and supporting information, anticipating the possibility that your paper may be accepted for publication.
 
-* Please provide the main manuscript file composed in Word, taking care to use the Table function for the tables.
+-   Please provide the main manuscript file composed in Word, taking care to use the Table function for the tables.
 However, if you created your manuscript in LaTex, please provide only the .tex file of the main manuscript, a .pdf file of the main manuscript (that is the exact same in content as the .tex file), and a .bib file.
 
-* Ecosphere page style allows for five levels of headings.
+-   Ecosphere page style allows for five levels of headings.
 Please use different type styles (italics, boldface, boldface italics, etc.) consistently to distinguish between the levels of headings in the article and delete the section numbers in the headings.
 
-* Literature references: Please check that each citation has complete information, and check that all references in text have a matching citation in the References section and vice versa.
+-   Literature references: Please check that each citation has complete information, and check that all references in text have a matching citation in the References section and vice versa.
 
-* Table and figure references: Please be sure that all tables and all figures have in-text citations in the main body of the manuscript, and that tables and figures are called out in numerical order (i.e., Figure 1 is called out before Figure 2).\
+-   Table and figure references: Please be sure that all tables and all figures have in-text citations in the main body of the manuscript, and that tables and figures are called out in numerical order (i.e., Figure 1 is called out before Figure 2).
 
-* Figures: Each figure should be provided in an individual, complete file (all panels in one image, identified by figure number in the file name), preferably as a high-resolution (300-600 dpi) image file or high-quality PDF.
+-   Figures: Each figure should be provided in an individual, complete file (all panels in one image, identified by figure number in the file name), preferably as a high-resolution (300-600 dpi) image file or high-quality PDF.
 Preferred file formats include TIF, EPS, PDF, or AI at 600 ppi, while JPEG, PPT/PPTX, or DOC/DOCX are acceptable if the resolution is sufficient.
@@
 
-* Figure captions: Please be sure to include a list of figure captions in the main manuscript file and ensure that any abbreviation that appears in a figure is identified in the caption.
+-   Figure captions: Please be sure to include a list of figure captions in the main manuscript file and ensure that any abbreviation that appears in a figure is identified in the caption.
 
-* Supporting Information: Each appendix must be assembled exactly as you intend it to appear online.
+-   Supporting Information: Each appendix must be assembled exactly as you intend it to appear online.
 (Appendices are posted "as is"; ESA and Wiley staff do not copy edit or assemble appendices.) Note also that the Ecosphere naming convention for appendices is of the style "Appendix S1," "Appendix S2," etc., and for tables and figures within appendices is "Table S1," "Table S2," etc.
@@
 
-* If the journal style is not followed at submission of the revision, we may return the manuscript to you for further revision before sending it along to the Subject-matter Editor.
+-   If the journal style is not followed at submission of the revision, we may return the manuscript to you for further revision before sending it along to the Subject-matter Editor.
 Closely following our manuscript preparation guide at this stage would expedite the production of your paper for publication, should it be accepted.
@@
 ## Response to the Editor
+
+### General Comments
 
@@
 Your revisions should address the specific points made in the review comments.
- Please note that Reviewer #2 made extensive comments directly on the manuscript.
-\
+ Please note that Reviewer #2 made extensive comments directly on the manuscript.\
 I should note that personally I can hardly disagree with your basic conclusion that models of forest dynamics need to do a good job of capturing regeneration and recruitment processes.
@@
 
-Dr. Charles Canham
-Subject-matter Editor, Ecosphere
+Dr. Charles Canham Subject-matter Editor, Ecosphere
+
+### Response to editors comments
 
@@
 This is recruitment.
-Since this is the response variable for the manuscript, I would suggest replacing 'regeneration' with 'recruitment' throughout the manuscript. \
+Since this is the response variable for the manuscript, I would suggest replacing 'regeneration' with 'recruitment' throughout the manuscript.\
 
@@
 
+Thank you for pointing
+
 #### COMMENT 2
+
 The difference between 7 cm and 10 cm DBH in the observed data.
@@
 
+Thank you for highlighting that we didn not managed to appropriately communicate our diameter thresholds approach correctly in the manuscript. The explanation is that the observed data
+
+In this revised version we have revised the explanation in L XX
+
 #### COMMENT 3
+
 The t-tests.
@@
 
+Based on the comments from Reviewer 1, we have re-analyse some of the results. More specifically we
+
 ### Minor comments:
@@
 #### COMMENT 4
+
 The abstract is not clear if the models do a good or poor job in simulating regeneration (L59, that they match observed ranges, and then L66, considerable mismatch between simulations and observed data, but that they do capture tree regeneration).
@@
 #### COMMENT 5
+
 L123-138: this paragraph has a lot of "on the other hand" type of statements, and it does not lead nicely into the next paragraph about the current study.
@@
 #### COMMENT 6
+
 Table 1.
@@
 #### COMMENT 7
+
 L218 -- do I understand this correctly, that regeneration rates are a mean of 56 trees per ha per decade? This seems really low (assuming that this is for \>10 cm DBH).  \
@@
 #### COMMENT 8
+
 L264 -- 274: how will this be accounted for later, that not all models simulated the same set of species?
@@
 #### COMMENT 9
+
 L340: "regeneration basal area of that species" is mentioned twice, please delete the second one.
@@
 #### COMMENT 10
+
 L364: so please remove the 4C model from the figure as it is not comparable. Do you need to also remove other models that don't simulate the same number of species too?\
@@
 #### COMMENT 11
+
 Table S2 and S3: What are the 3 columns for p-values?
@@
 #### COMMENT 12
+
 Figure S1 and S2 -- I can't tell the difference between observed and simulated. Are they different colours? The legend should be larger. What is the black line for?
@@
 #### COMMENT 13
+
 Figure 4 (and Figure S3) are interesting in that there are no patterns with overstory diversity, and that this was not explained by feedback with seed production.
@@
 #### COMMENT 14
+
 All figures in the Appendix, please make the X and Y axis labels larger and centered. I actually thought labels were missing initially and it took a while to find them.
@@
 #### COMMENT 15
+
 L400 -- 402: where is this conclusion about mixed species or monocultures coming from? I don't see that in the figures or the results.
@@
 #### COMMENT 16
+
 Figure 6 is excellent.
@@
 #### COMMENT 17
+
 Table S4: please clean up.
@@
 #### COMMENT 18
+
 Not sure both Figure 7 and Figure S7 are necessary (I admit I spent a few minutes going back and forth between them trying to understand how they differed).
@@
 #### COMMENT 19
+
 L547-548: predicting stand diversity, did I miss this?
@@
 #### COMMENT 20
+
 L575: that most models did not deviate exceedingly from observations seems in contrast to what the previous section was stating. What is this based on?
@@
 #### COMMENT 21
+
 L584: What does this mean, about degrees of freedom for modelling regeneration?
@@
 #### COMMENT 22
+
 L624: Missing the end of the sentence.
@@
 #### COMMENT 23
+
 L647: Completely agree that observed data is a snapshot of a stochastic and dynamic system and care should always be taken with comparing to simulation data, however I disagree about the following stamen (L648).
@@
 #### COMMENT 24
+
 L655: But the landscape models generally did about the same (i.e., not consistently worse) then the rest, correct?
@@
 #### COMMENT 25
+
 L663: But in general, the models were more sensitive to water balance and the observed data did not have any real pattern.
@@
 #### COMMENT 26
+
 L675: what would a 'comprehensive regeneration dataset" look like?
@@
 #### COMMENT 1
+
 Where the paper falls short, in my opinion, is in providing a clear path forward.
@@
 #### COMMENT 2
+
 I felt the readability of the paper would be improved by introducing a "road mapping" paragraph at the end of the introduction or start of the methods section that sets out the approach in broad terms.
@@
 #### COMMENT 3
+
 The words "ingrowth", "regeneration" and "recruitment" are used interchangeably to describe the recruitment of new trees into the 7-cm diameter class.
@@
 #### COMMENT 4
+
 The article is written entirely by Europeans, unless I'm mistaken, and very much focused on the central/eastern European forestry literature.
@@
 ### Minor comments:
-These comment have been extracted from the comments done by reviewer 2 in the PDF document of the manuscript.
+
+These comment have been extracted from the reviewer 2 comments on the PDF document of the manuscript.
 
@@
 
+L51-52 "However, an assessment of their ability to accurately represent tree regeneration is lacking." Not true.
+
 ##### Response to comment 5:
 
+#### COMMENT 6
+
+L56 "The results are evaluated against comprehensive data from unmanaged forests." comprehensive?
+
+##### Response to comment 6:
+
+#### COMMENT 7
+
+L56 "Models simulating higher species diversity at the stand level do not feature higher regeneration diversity." feature?
+
+##### Response to comment 7:
+
 ## Other changes made in the manuscript

### 2023-09-19 manuscript/review/authorsResponse.qmd
first draft letter to editor

@@
 ### Response to editors comments
+
+Dear Charles,
+
+Thank you very much for taking the time to review our manuscript and for your insightful thoughts on our findings.
+
+We totally agree with your comments.
+Our main take away from this manuscript it is not that it would be impossible to calibrate all 15 models against e.g.
+the EuFoRIa data, but the issue is - as you wrote - that this needs to be bound to the goals of the study.
+And when the goal is to provide extrapolation into a future climate, then calibration against data under current climate may be unreliable.
+This is why we belive it is fundamentally valuable to test model performance "out of the box\", as we have done here.
+
+In this revised version we have modify the manuscript according to the reviewers' recommendations and we have addressed each of the specific points made in the review comments.
+We believe the manuscript has significantly improved based on this changes.
 

### 2023-09-19 manuscript/manuscriptRW.Rmd
added basic structure for review
review documents and updates on the original manuscript
comments 5-7 reviewer 2

@@
   word_document:
-    reference_docx: template_paper.docxl
+    reference_docx: template_paper.docx
 fontsize: 12pt
@@
 
-**OPEN RESEARCH STATEMENT:** This submission uses novel code, both the data
-and code are provided in an external repository to be evaluated during
-the peer review process. The repository can be downloaded from
-[this figshare private link: https://figshare.com/s/100857ba67d2351dcb15](https://figshare.com/s/100857ba67d2351dcb15).
-The protocol is available at @díaz-yáñez2022 ([with link: https://osf.io/czdxp/](https://osf.io/czdxp/ )).
-We
-will make publicly available the code and data if the manuscript is accepted.
+**OPEN RESEARCH STATEMENT:** This submission uses novel code, both the
+data and code are provided in an external repository to be evaluated
+during the peer review process. The repository can be downloaded from
+[this figshare private link:
+https://figshare.com/s/100857ba67d2351dcb15](https://figshare.com/s/100857ba67d2351dcb15).
+The protocol is available at @díaz-yáñez2022 ([with link:
+https://osf.io/czdxp/](https://osf.io/czdxp/)). We will make publicly
+available the code and data if the manuscript is accepted.
 
@@
 for assessing long-term forest dynamics, and they have been in use for
-more than 50 years. However, an assessment of their ability to
-accurately represent tree regeneration is lacking. We assess how well
-current models capture the overall abundance, species composition, and
-mortality of tree regeneration.
+more than 50 years. However, there is a need to evaluate their capacity
+to accurately represent tree regeneration. We assess how well current
+models capture the overall abundance, species composition, and mortality
+of tree regeneration.
 
@@
 representing large environmental gradients across Central Europe. The
-results are evaluated against comprehensive data from unmanaged forests.
+results are evaluated against extensive data from unmanaged forests.
 
@@
 of regeneration matches the observed ranges. Models simulating higher
-species diversity at the stand level do not feature higher regeneration
+species diversity at the stand level do not present higher regeneration
 diversity. The effect of light availability on regeneration levels is

### 2023-09-19 manuscript/review/authorsResponse.qmd
comments 5-7 reviewer 2

@@
 
-These comment have been extracted from the reviewer 2 comments on the PDF document of the manuscript.
+These comments have been extracted from the reviewer 2 comments on the PDF document of the manuscript.
 
@@
 
+Based on the reviewers comments we have modified this sentence to: "However, there is a need to evaluate their capacity to accurately represent tree regeneration"
+
 #### COMMENT 6
@@
 ##### Response to comment 6:
+
+Following the reviewers comment we have changed the word "comprehensive" to "extensive" to express that we used a wide-range data that considers many important elements of regeneration.
 
@@
 
+Following the reviewers comments we have changed "feature" with "present".
+
+#### COMMENT 8
+
+##### Response to comment 8:
+
 ## Other changes made in the manuscript

### 2023-09-20 manuscript/manuscriptRW.Rmd
reviewer 2 almost all comments

@@
 after disturbance [@lindner2010; @seidl2022]. A wide range of models of
-forest dynamics were developed over the past decades considering the
-impacts of climate [@bugmann2022]. From these studies, it is evident
+forest dynamics have been developed over the past decades considering
+the impacts of climate [@bugmann2022]. From these studies, it is evident
 that we have a reasonably good understanding how to model tree growth
@@
 less studied process that is often represented rather coarsely in models
-[@walck2011; @leishman1992; @price2001]. Which is a clear research gap
-in the context of climate-induced forest disturbances and forest
-resilience.
-
-Tree regeneration arises from multiple processes such as pollination,
+[@walck2011; @leishman1992; @price2001]. This presents an important
+research gap, particularly in the context of climate-induced forest
+disturbances and forest resilience.
+
+Tree regeneration arises from multiple processes including pollination,
 fruit maturation, seed production, dispersal, germination, juvenile
@@
 models are handled in a multitude of ways [@konig_tree_2022;
-@bugmann2022]: from 1) entirely ignoring it [as done in classical forest
-growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2) the use
-of a few simple environmental filters, as done in most forest "gap"
-models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
+@bugmann2022]: from 1) entirely ignoring them [as done in classical
+forest growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2)
+the use of a few simple environmental filters, as done in most forest
+"gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
 @hickler2012; @smith2001], to 3) complex approaches that incorporate
@@
 tree regeneration patterns in nature and the large number of factors
-driving this process --- including some that are not incorporated
-explicitly in most models, such as deer browsing --- we do not aim for a
-detailed statistical evaluation of each model. Instead, we aim to
-evaluate the general regeneration patterns and magnitudes simulated by
-the models, and benchmark the simulated regeneration niche of multiple
-species against empirical data along a wide environmental gradient of
-temperature, moisture, and light availability [@grubb1977].
+driving and constraining this process --- including some that are not
+incorporated explicitly in most models, such as deer browsing --- we do
+not aim for a detailed statistical evaluation of each model. Instead, we
+aim to evaluate the general regeneration patterns and magnitudes
+simulated by the models, and benchmark the simulated regeneration niche
+of multiple species against empirical data along a wide environmental
+gradient of temperature, moisture, and light availability [@grubb1977].
 
@@
 the environmental dimensions of temperature, climatic water balance,
-soil quality and exposition (i.e., slope and aspect). Each of these
-sites featured at least two consecutive measurements. Regeneration
-thresholds for these sites differed between diameters of 0 and 10 cm.
-For our study, we defined two datasets where one included 165 sites with
-a diameter threshold of 7 cm or lower, and the other included another 35
+soil quality, slope and aspect. Each of these sites featured at least
+two consecutive measurements. Regeneration minimum size thresholds for
+these sites differed between stem diameters of 0 and 10 cm. For our
+study, we defined two datasets where one included 165 sites with a
+diameter threshold of 7 cm or lower, and the other included another 35
 sites with diameter thresholds between 7 and 10 cm.
@@
 
-The overarching goal of the experiments was to assess the tree
+The overarching goal of the simulation experiments was to assess tree
 regeneration as it arises from empirical data against its representation
@@
 regeneration as the passing of a breast height diameter threshold of 7
-or 10 cm, respectively (synonym: ingrowth). To this end, each modeling
-group was provided with a detailed protocol [@díaz-yáñez2022] with
-instructions how to perform the simulations, providing input variables
-on climate and soil conditions, and the list of expected output
-variables. Neither were further site information (except for the data
-specified below) nor any data on tree regeneration or forest stand
-features provided prior to the simulation. That is, the models were run
-in "blind flight" mode.
+and 10 cm (synonym: ingrowth). To this end, each modeling group was
+provided with a detailed protocol [@díaz-yáñez2022] with instructions
+how to perform the simulations, input variables on climate and soil
+conditions, and the list of expected output variables. Neither were
+further site information (except for the data specified below) nor any
+data on tree regeneration or forest stand features provided prior to the
+simulation.
 
@@
 
-Each of the models reported the regeneration number by sampling 200
-times in a 10-year interval for each species and per ha for each of the
-200 sites. Multiple samples per site were used to better understand the
-simulated variation within each site. This was done using different
-strategies, depending on the model, including (1) sampling simulated
-data from the same 1 ha plot in the equilibrium over time, (2) sampling
-several 1 ha plots from the simulated forest at one specific point in
-time (in the equilibrium) or (3) a combination of (1) and (2). This
-resulted in 880,000 observations per model that simulated the 11 species
-included in the protocol (200 sites, 200 samples per site, 11 species
-and two diameter thresholds). For the models that simulated additional
-species, their regeneration rates were aggregated as "others".
+Each of the models reported the number of trees crossing the two
+diameter thresholds by sampling 200 times in a 10-year interval for each
+species and per ha for each of the 200 sites. Multiple samples per site
+were used to better understand the simulated variation within each site.
+This was done using different strategies, depending on the model,
+including (1) sampling simulated data from the same 1 ha plot in the
+equilibrium over time, (2) sampling several 1 ha plots from the
+simulated forest at one specific point in time (in the equilibrium) or
+(3) a combination of (1) and (2). This resulted in 880,000 observations
+per model that simulated the 11 species included in the protocol (200
+sites, 200 samples per site, 11 species and two diameter thresholds).
+For the models that simulated additional species, their regeneration
+rates were aggregated as "others".
 
@@
 \begin{equation}
-Reineke = \frac{7}{10}^{-1.605} \label{eqReineke}
+Reineke = \frac{N7}{N10}^{-1.605} \label{eqReineke}
 \end{equation}
@@
 over the past decades [e.g., @heiri2009], leading to less regeneration
-than in an equilibrium situation.
+than in an equilibrium situation. Another possible reason for this
+difference may be the exclusion of factors like deer browsing or failure
+to accurately measure establishment filters, such as those to site
+limitants or competition with herb layers.
 
@@
 simplistic. Furthermore, based on the simulation results, there is no
-evidence that models with feedback from the canopy captured better the
-species diversity of the regeneration than those without feedback, with
-the exception of ForCEEPS that significantly improved regeneration
-species diversity with the model version that included feedback
-(ForCEEPS(f)). The similar performance of models with and without
-feedback is likely because the models put more weight on the
-regeneration niche arising from abiotic and biotic filters, than from
-the habitat niche of the adult trees [@grubb1977].
+evidence that models with feedback from the canopy (in terms of species
+composition of regeneration) captured better the species diversity of
+the regeneration than those without feedback, with the exception of
+ForCEEPS that significantly improved regeneration species diversity with
+the model version that included feedback (ForCEEPS(f)). The similar
+performance of models with and without feedback is likely because the
+models put more weight on the regeneration niche arising from abiotic
+and biotic filters, than from the habitat niche of the adult trees
+[@grubb1977].
 
@@
 widely documented [@messier1999; @collet2006; @berdanier2016], but the
-models examined here did not reproduce this expectation. However, it is
-difficult to measure light availability at large spatial and temporal
-scales. We used total stand basal area as a proxy for light availability
-[cf. @schmid2021]. However, we were unable to consider light
-availability restrictions caused by ground vegetation, which may be an
-important filter for some EuForIa stands [@woltjer2008]. We found
+models examined here did not reproduce this expectation (i.e.,
+decreasing regeneration levels with increasing stand basal area).
+However, it is difficult to measure light availability at large spatial
+and temporal scales. We used total stand basal area as a proxy for light
+availability [cf. @schmid2021]. However, we were unable to consider
+light availability restrictions caused by ground vegetation, which may
+be an important filter for some EuForIa stands [@woltjer2008]. We found
 pronounced differences in the stand basal area ranges simulated by the

### 2023-09-20 manuscript/review/authorsResponse.qmd
reviewer 2 almost all comments

@@
 linkcolor: blue
+number-sections: true
 format:
@@
 
-## Journal formatting
+## Internal instructions
+
+-   underlined comments are comments that still need attention
+
+-
+
+## Journal formatting (ToDo list)
 
@@
 And when the goal is to provide extrapolation into a future climate, then calibration against data under current climate may be unreliable.
-This is why we belive it is fundamentally valuable to test model performance "out of the box\", as we have done here.
+This is why we belive it is fundamentally valuable to test model performance "out of the box", as we have done here.
 
@@
 
-#### COMMENT 1
+#### [COMMENT 1]{.underline}
 
@@
 
+[I have to think about this again]{.underline}
+
 Thank you for pointing
 
-#### COMMENT 2
+#### [COMMENT 2]{.underline}
 
@@
 
-#### COMMENT 3
+#### [COMMENT 3]{.underline}
 
@@
 
-Based on the comments from Reviewer 1, we have re-analyse some of the results. More specifically we
+[Based on the comments from Reviewer 1, we have re-analyse some of the results. More specifically we]{.underline}
 
@@
 
-#### COMMENT 4
+#### [COMMENT 4]{.underline}
 
@@
 
-#### COMMENT 5
+[check]{.underline}
+
+#### [COMMENT 5]{.underline}
 
@@
 
-#### COMMENT 6
+[check]{.underline}
+
+#### [COMMENT 6]{.underline}
 
@@
 
-#### COMMENT 7
+[change this]{.underline}
+
+#### [COMMENT 7]{.underline}
 
@@
 
-#### COMMENT 8
+[check]{.underline}
+
+#### [COMMENT 8]{.underline}
 
@@
 
-#### COMMENT 9
+#### [COMMENT 9]{.underline}
 
@@
 
-#### COMMENT 2
+#### COMMENT 2 {#sec-R2C2}
 
@@
 
+L76-78 "A wide range of models of forest dynamics were developed over the past decades considering the impacts of climate"
+
 ##### Response to comment 8:
 
+Following the reviewers comment, we have changed the word "were" with "have been".
+
+#### [COMMENT 9]{.underline}
+
+L79 K. Vanclay and Skovsgaard 1997;
+
+##### Response to comment 9:
+
+Following the reviewers comment, we have corrected the reference style.
+
+#### COMMENT 10
+
+L83-84 "Which is a clear research gap in the context of climate-induced forest disturbances and forest resilience." incomplete sentence
+
+##### Response to comment 10:
+
+Following the reviewers comment, we have changed this sentence to: "This presents an important research gap, particularly in the context of climate-induced forest disturbances and forest resilience."
+
+#### COMMENT 11
+
+L85-86 "Tree regeneration arises from multiple processes such as pollination, fruit maturation, seed production, dispersal, germination, juvenile growth and survival" including
+
+##### Response to comment 11:
+
+Following the reviewers comment, we have changed "such as" with "including"
+
+#### COMMENT 12
+
+L85-86 "Tree regeneration arises from multiple processes such as pollination, fruit maturation, seed production, dispersal, germination, juvenile growth and survival" including
+
+##### Response to comment 12:
+
+Following the reviewers comment, we have changed "such as" with "including"
+
+#### COMMENT 13
+
+L90-92 "Currently, tree regeneration processes in dynamic forest models are handled in a multitude of ways (König et al.
+2022; Bugmann and Seidl 2022): from 1) entirely ignoring it (as done in classical forest growth models, e.g., Pretzsch et al.
+2002)," it.
+
+##### Response to comment 13:
+
+Following the reviewer's comment we have changed "it" with "them"
+
+#### [COMMENT 14]{.underline}
+
+L96-98 Particularly given your conclusion that the models are poor at predicting regeneration processes, should we now return to statistical parameterisation afresh and find better ways of extrapolation/interpolation across space/time?
+
+##### Response to comment 14:
+
+It is indeed interesting to account with the increasing data vailablity to better understand tree regeneration processes.
+However, we believe that the main take away it is not that it would be impossible to calibrate all 15 models against e.g.
+the EuFoRIa data to better capture regeneration, but that the focus should be in the study goals.
+And when the goal is to provide extrapolation into a future climate, then calibration against data under current climate may be unreliable.
+With the approach we have followed in this study where we have asked the models to simulate forest dynamics and capture regeneration without statistical predetermination we can better assess their behavior and understand how their current model structures and inclusion of exclusion of certain ecological processes impact the outputs we are seeing.
+
+[Following this and other comments from the editor, we have tried to capture this better in the discussion/conclussions section.]{.underline}
+
+#### [COMMENT 15]{.underline}
+
+L100-101 "Overall, models are needed to...(3) identify the most important processes that are shaping ecological patterns." doesn't that depend on what goes into the model?
+
+##### Response to comment 15:
+
+We are not complete sure of the reviewers point here.
+If the reviewer is referring to the fact that the most important processes will be those included in the model, it is true that you can only asses those processes included or how including or not another processes impact regeneration.
+This was also the idea with considering a large pool of models, that we could assess how their structural complexity, or the inclusion/exclusion of a certain process affects the results we are seeing.
+This also goes very much in line with our research recommendations presented in the discussion where we proposed that improvement of the regeneration modules is implemented as additional features that can be traced back, as done here for the variants of ForClim and ForCEEPS, and that model complexity and structure must always be connected with modeling objectives.
+
+#### [COMMENT 16]{.underline}
+
+L101-106 "Given the current strategies that are used in models of forest dynamics to represent tree regeneration, model behavior often is prone to problems, such as very high levels of tree regeneration that necessitate excess mortality at early stages of tree life to simulate correct stand structure and composition." ref?
+tens of thousands of seeds\...
+
+##### Response to comment 16:
+
+[Following the reviewer's comment we have added a reference]{.underline}
+
+[ToDo find a reference!]{.underline}
+
+#### [COMMENT 17]{.underline}
+
+L112-114 "A related issue is the excessive reduction of species diversity due to positive feedback effects, such that eventually just single-species stands remain." ref
+
+##### Response to comment 17:
+
+[Following the reviewer's comment we have added a reference]{.underline}
+
+[ToDo find a reference!]{.underline}
+
+#### [**COMMENT 18**]{.underline}
+
+L127 I suggest making better use of the hanbury-brown 2022 review.
+
+##### Response to comment 18:
+
+We currently use the exciting Hanbury-brown 2022 review in several places across our manuscripts, following the reviewer's comment here and the general comment 2 in @sec-R2C2 we have also get inspired by their structure [by creating a road map to help the reader to navigate the manuscript content.]{.underline}
+
+#### [**COMMENT 19**]{.underline}
+
+L145-147 "Due to the large variability in tree regeneration patterns in nature and the large number of factors driving this process including some that are not incorporated explicitly in most models, such as deer browsing---"
+
+##### Response to comment 19:
+
+We believe the reviewer highlighted the words "models, such as deer browsing" because they think this is not correct.
+From our sample pool [only 4 models]{.underline} incorporated explicitly deer browsing.
+
+####
+
+[**COMMENT 20**]{.underline}
+
+L150 "regeneration niche of multiple species"
+
+L154 "traits"
+
+##### Response to comment 20:
+
+[We are unsure why the reviewer highlighted those words]{.underline}
+
+#### **COMMENT 21**
+
+L185 "(mean regeneration formulation complexity across all processes \>0, Table 1)." Meaning
+
+##### Response to comment 21:
+
+This statement relates to Bugmann and Seidl (2022) where complexity values were provided.
+The exact details of how we calculated this mean regeneration formulation complexity value are available in the software suplementary materials to this paper: \`\`\`figure.R\`\`\` line 706-750.
+This value represents the mean complexity of regeneration formulations per model.
+
+#### [**COMMENT 22**]{.underline}
+
+L186-188 "ForClim variant 1 (Bugmann et al. 1996) is based on a recruitment module that adheres closely to the concept introduced by Botkin, Janak and Wallis (1972)" which is...
+
+##### Response to comment 22:
+
+[Is there a way to explain this shortly??]{.underline}
+
+#### **COMMENT 23**
+
+L197-198 "Regeneration data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset" not sure that's true.  Some of the large forest inventories include small seedlings and saplings in nested plots, which are a step in the right direction.
+
+##### Response to comment 22:
+
+We agree with the reviewer that permanent plots such as those present in National Forest Inventories could potentially be a good source of information to obtain regeneration data.
+However these often are collected from managed sites, do not capture ingrowth below 10cm, and are hard to harmonized between different inventory strategies.
+
+#### **COMMENT 24**
+
+L200-202 "The observations used here are derived from a novel and unprecedented network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network (EuFoRIa 2019)" I don't feel "novel and unprecedented" adds anything to this sentence.  There are plenty of networks outside Europe providing this sort of data
+
+##### Response to comment 24:
+
+We agree with the reviewer's comment that there are other forest inventory data networks outside Central Europe even on forest reserves that could provide this sort of data.
+However, the EuFoRIa network is indeed novel and unprecedented network in Central Europe.
+
+#### **COMMENT 25**
+
+L200-202 "The observations used here are derived from a novel and unprecedented network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network (EuFoRIa 2019)" I don't feel "novel and unprecedented" adds anything to this sentence.  There are plenty of networks outside Europe providing this sort of data
+
+##### Response to comment 25: {#sec-R2C25r}
+
+We agree with the reviewer's comment that there are other forest inventory data networks outside Central Europe even on forest reserves that could provide this sort of data.
+However, the EuFoRIa network is indeed novel and unprecedented network of forest reserves in Europe.
+
+#### **COMMENT 26**
+
+L208-209 "We selected 200 sites from this network as the benchmarking dataset for the simulation to be representative of the environmental variation contained in the data." COOL
+
+##### Response to comment 26:
+
+We agree :)
+
+#### **COMMENT 27**
+
+L211-212 "exposition (i.e.,slope and aspect)." unfamiliar term to me
+
+##### Response to comment 27:
+
+Following the reviewer's comment we have changed "and exposition (i.e.,slope and aspect)." with ", slope and aspect."
+
+#### **COMMENT 28**
+
+L212-213 "Regeneration thresholds for these sites differed between diameters of 0 and 10 cm" minimum size threshold? and stem diameters
+
+##### Response to comment 28:
+
+Following the reviewer's comment we have changed this sentence to: "Regeneration minimum size thresholds for these sites differed between stem diameters of 0 and 10 cm"
+
+#### **COMMENT 29**
+
+L215 "another 35 sites with diameter thresholds between 7 and 10 cm." OK, so we're talking about v large trees here
+
+##### Response to comment 29:
+
+The minimum threshold that could be standarized across reserves was of at least 7cm.
+
+**COMMENT 30**
+
+L215 "another 35 sites with diameter thresholds between 7 and 10 cm." OK, so we're talking about v large trees here
+
+L218 "featured 30,900 newly established trees" so they are trees....
+
+##### Response to comment 30:
+
+The minimum threshold that could be standardized across reserves was of at least 7cm.
+
+#### **COMMENT 31**
+
+L215 "on this unique dataset" not so unique
+
+##### Response to comment 31:
+
+Please, see our reply on comment 25 in here @sec-R2C25r
+
+#### **COMMENT 32**
+
+L221 "nformation, cf. Käber et al. (2023)." in review and not available
+
+##### Response to comment 32:
+
+This paper have recently being published and the reference has been updated accordingly.
+
+#### **COMMENT 33**
+
+L223 "The overarching goal of the experiments was to assess the tree regeneration as it arises" simulation experiments and remove "the"
+
+Response to comment 33:
+
+Changed as suggested
+
+#### [**COMMENT 34**]{.underline}
+
+L225 "tree regeneration as the passing of a breast height diameter threshold of 7 or 10 cm, respectively" I suggest "tree recruitment" not "tree regeneration" here. respectively??
+
+##### 
+Response to comment 34:
+
+We have removed "respectively".
+
+[We have re-assessed the terms used across this manuscript XXX]{.underline}
+
+####
+
+**COMMENT 35**
+
+L 227-228 "simulations, providing input variables" provided with
+
+##### 
+Response to comment 35:
+
+We have removed "providing".
+
+#### **COMMENT 36**
+
+L229 "Neither were further site information (except for the data specified below) nor any data " These were "blind trials":  modelling group were not provided with site information \...
+
+##### 
+Response to comment 36:
+
+The reviewer is correct that we did not provide e.g.
+site coordinates (with the exception of iLand that needed this for a model specific soil input, this is described in the manuscript), but we did provide topography, climate and soil site specific variables.
+
+#### **COMMENT 37**
+
+L230-231 "That is, the models were run in \"blind flight\" mode." Remove
+
+##### 
+Response to comment 37:
+
+Changed as suggested.
+
+#### **COMMENT 38**
+
+L237-238 "Soil quality data were provided as continuous values between 1 and 5" ?hydrological quality?
+
+##### 
+Response to comment 38:
+
+The soil quality value is based on a random forest that was trained to predict expert knowledge based soil quality on a scale from 1 to 5 from the [SoilGrids250 data set](https://www.isric.org/explore/soilgrids) and the [WISE data set](https://www.isric.org/explore/wise-databases).
+The most influential variables in the random forest were total depth to bedrock and sand content.
+Thus the variable is a good predictor for plant available water storage capacity.
+This information has been provided in the protocol of this study.
+
+####
+
+**COMMENT 39**
+
+L247 "The exact length of the simulation was also decided by the modeling teams" seems odd not to constrain that.
+
+##### 
+Response to comment 39:
+
+We understand the reviewers concern on not constraining the length of the simulations, but we decided to leave it to the assessment of the individual modelers, as we wanted to obtain simulation that were run from from bare ground in the absence of management to a simulated equilibrium (\"potential natural vegetation\") with current climate, and each modeler know best how to obtain this in each model.
+This information is currently available in the manuscript and in the protocol of this study.
+
+#### **COMMENT 40**
+
+L250 "The simulations were run in the absence of management to a simulated equilibrium" The absence of disturbance and running to equilibrium set quite specific conditions; suggest this is emphasised more in abstract / intro / discussion.
+
+##### 
+Response to comment 40:
+
+We agree with the reviewer that the simulations set up regarding absence of management and to a simulated equilibrium are quite specific.
+We found challenging to incorporate this in the introduction as we see this a more of a methodological specification, and in the abstract due to space constraints as we though is more interesting to focus in the results and meaning of our findings than in methodological aspects.
+However we currently have a lengthy discussion on the equilibrium assumption and also about the importance of considering disturbances, which we think it would be a great further research direction to better understand regeneration from models of forest dynamics.
+
+#### **COMMENT 41**
+
+L252-253 "This entails the assumption that (1) the observations from the forest reserves reflect no traces of forest management," That's seldom the case in Europe where the legacies of management are evident centuries on
+
+##### 
+Response to comment 41:
+
+The reviewer is correct and management legacies might be evident for long periods of time.
+The data were collected in forest reserves where no management has taken place for long periods of time which makes our assumption of an equilibrium between forest properties and environmental drivers reasonable.
+This is the best data we have and we have address its limitations, as this one, in the discussion because we are also aware that is not ideal.
+
+#### **COMMENT 42**
+
+L275 "Each of the models reported the regeneration number by sampling 200 times in a 10-year interval" regeneration number = number of "recruits" entering the 7-cm size class?
+
+##### 
+Response to comment 42:
+
+Following the reviewer's comment we have changed "regeneration number" with "number of trees crossing the two diameter thresholds"
+
+#### [**COMMENT 43**]{.underline}
+
+L295 "(2) regeneration species diversity" species diversity of recruitment
+
+L295 "(3) regeneration mortality" mortality of recruits
+
+##### 
+Response to comment 43:
+
+[Change if we change regeneration with recruit]{.underline}
+
+####
+
+**COMMENT 44**
+
+L296 "ingrowth gradients along the regeneration niches." ?? not clear to me. How is regeneration niche defined??
+
+##### 
+Response to comment 44:
+
+We defined the regeneration niche as the passing of a diameter threshold of 7 and 10 cm and we focused both in the width of the regeneration niche (i.e., in environmental space) as well as the intensity of the regeneration process (i.e., the number of ingrowth trees per area and per unit of time) were of interest.
+This is currently defined in the material and methods section and in the protocol of this study.
+
+#### **COMMENT 45**
+
+Equations 1-5.
+not convinced these formulae need to be shown as  the Shannon index is very well known.  Consider converting the Shannon index to the effective number of species,  by taking the negative of the SI and exponentiating it.  This is gaining ground in ecology as it's more intuitive to interpret than SI  (see paper by Jost).
+
+##### 
+Response to comment 45:
+
+We agree with the reviewer that the Shannon index is very well known index and its equation as well.
+However we think is still important that we define how we did calculate the Shannon index based on the available data we had as we had some internal discussion on how to do this and the meaning of this index in this context, we though the reader will benefit from a more detailed explanation.
+In any case the exact details of how we did calculate this index based on the simulation outputs are also available in the software supplementary materials, in the script figures.R L42-78.
+We also thank the reviewer for the suggestion of using the effective number of species instead of the Shannon index, it is indeed a good alternative, however we have decided to leave it with the SI.
+
+#### [**COMMENT 46**]{.underline}
+
+L308 "Mortality in tree regeneration was assessed based on the ratio of regeneration between the 7 and 10" [of recent recruits.]{.underline} Now I understand the 7 and 10 cm thresholds mentioned earlier;  please clarify earlier mention.
+
+##### 
+Response to comment 46:
+
+Following the reviewer's comment on the lack of claritiy of the selected thresholds explanation we have changed at the beginning of the section simulation protocol the first sentence where we explain that we considered tree regeneration as the passing of a breast height diameter threshold of 7 and 10 cm.
+We have also reported later on in this section (but before this line) that each of the models reported the number of trees crossing the two diameter thresholds.
+
+####
+
+**COMMENT 47**
+
+L311-313 "The Reineke self-thinning rule is usually calculated for even-aged, single species stands and assumes a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands." more to the point,  it's always (I think) applied to all stems in the stand,  not just regenerating trees, as it's fundamental concept is that entire canopies are space filling .
+Why would we expect it to apply to a narrow cohort?  I think not.
+
+equation 5 Suggest writing N7 / N10
+
+Given my comment above, is there a strong justification for the -1.605 exponent?  Or just keep it at one?
+
+##### 
+Response to comment 47:
+
+Follwoing the reviewer recomendation we have changed the equation formulation to better represent the Reineke relationship.
+The reviewer is correct that the Reineke is applied to all stems in the stand, in our case the N value is only changing due to the available regeneration at the two diameter thresholds as the adult trees number is the same.
+By comparing our expected relationship of recruited trees at 7 and 10cm with the Reineke value we are just comparing to a theoretical value in fully stocked stands.
+We have use the coefficient -1.605 because Reineke attributed a general validity to this allometric coefficient for fully stocked, even-aged forest stands, regardless of tree species and site.
+
+#### COMMENT 48
+
+equation 5 Suggest writing N7 / N10
+
+Given my comment above, is there a strong justification for the -1.605 exponent?  Or just keep it at one?
+
+##### 
+Response to comment 48:
+
+The Stand Density Index by Reineke (1933) is based on the relationship between quadratic mean diameter and number of stems per unit area in fully stocked stands.
+We have use the coefficient -1.605 because Reineke attributed a general validity to this allometric coefficient for fully stocked, even-aged forest stands, regardless of tree species and site.
+This relationship was obtained by using pairs of variates for quadratic mean diameter and number of stems per unit area.
+In our study we are using the value of 1.77 to define this relationship between 7 and 10cm
+
+#### [COMMENT 49]{.underline}
+
+L316-317 "model complexity as defined by Bugmann and Seidl"
+
+##### 
+Response to comment 49:
+
+[Think how can I add something else without making this much longer!]{.underline}
+
+#### COMMENT 50
+
+L317 "model type (empirical or process based)," if statistical analyses are used to estimate the coefficients of complex functions that underpin a process based model, is that model empirical or process based?
+
+##### 
+Response to comment 50:
+
+The reviewer's comment is a very valid one as it is true that many processes in process based models have been defined based on empirical approaches.
+In this study we have categorized the models in process based and empirical based on their general approaches to define forest dynamics.
+This categorization is available in table 1.
+It is also possible to find further details on each of the models in the supplementary materials.
+
+#### COMMENT 51
+
+L496 "confronted with a unique dataset from unmanaged"
+
+L498 "The EuFoRIa data (Käber et al. 2023) are exceptional, particularly"
+
+I think there are plenty of other datasets that could be used and have better information on smaller trees than EuFoRIa,  so suggest a more nuanced sales pitch here.
+
+##### 
+Response to comment 51:
+
+To our knowledge this is the only network of forest reserves in Europe covering a large environmental gradient and from which a standardization effort has been made toward understanding tree regeneration across different countries.
+We would be happy to learn of any other network of this characteristics in Europe and potentially use it in follow up studies.
+
+####
+
+COMMENT 52
+
+L525-527 "There are multiple constraints to the regeneration niche of tree species (Price et al.
+2001), and therefore the absence of regeneration is likely to be common (Fortin and DeBlois 2007), even over larger areas" YES - good point.  This should be featured more prominent I think
+
+##### 
+Response to comment 52:
+
+Thank you for this remark, we agree with the reviewer that this is an important point and specially how different models are capturing or missing to represent the absence of regeneration.
+We have now captured this aspect also in the introduction in the sentence: "Due to the large variability in tree regeneration patterns in nature and the large number of factors driving and constraining this process" on top of the current whole paragraph on this issue present in the discussion.
+
+#### COMMENT 53
+
+L530-533 "This substantial difference may be due to the fact that the simulation results were drawn from equilibrium forests, whereas in reality many of the forest reserves are recovering from past management activities and have become denser over the past decades (e.g., Heiri et al.
+2009), leading to less regeneration than in an equilibrium situation.
+" But also \-\-- browsing, establishment sites,  competition with herb layer all reduce recruitment
+
+##### 
+Response to comment 53:
+
+Following the reviewer's comment we have included this point by adding into this paragraph: "Another possible reason for this difference may be the exclusion of factors like deer browsing or failure to accurately measure establishment filters, such as those to site limitants or competition with herb layers."
+
+#### COMMENT 54
+
+L552 "there is no evidence that models with feedback from the canopy captured better the species" could you explain what you mean by "feedback" here?
+
+##### 
+Response to comment 54:
+
+Following the reviewers comment we have modified this sentence to clarifu what do we mean here with feedback: " Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the regeneration than those without feedback, "
+
+This had also been defined in the material and methods where we stated that "i.e., where the level and species composition of regeneration is influenced by the existing adult tree community via the production of seeds, seedlings or saplings, from models that do not contain such a feedback."
+
+#### 
+COMMENT 55
+
+L586-587 "Our study showed that increasing complexity in the regenerationmodules is not linkedwith a higher accuracy of the projections of regeneration levels," Surely more complex models WOULD be better if they captured the relevant processes!  Arguing for simple vs complex models is a distraction here\...
+we need models that are as simple as possible but not too simple..
+
+##### 
+Response to comment 55:
+
+We agree with the reviewers statement that we need the "right amount" of complexity which considers the most relevant processes.
+however there are limited studies looking at this aspect in models of forest dynamics and even more limited looking at this aspect in the regeneration related processes in models of forest dynamics.
+We still think it is important to discuss this issue in the context of our findings even though this was not the solely focus of our research.
+
+#### 
+COMMENT 56
+
+L596-598 "Competition for light as a strong filter for tree regeneration has been widely documented (Messier et al.
+1999; Collet and Chenost 2006; Berdanier and Clark 2016), but the models examined here did not reproduce this expectation.
+", because \...
+remind us of the evidence you are putting forward
+
+##### 
+Response to comment 56:
+
+Following the reviewers comment we have added the following sentence in this part of the text:
+
+"(i.e., decreasing regeneration levels with increasing stand basal area)"
+
+#### [COMMENT 57]{.underline}
+
+L603-606 "This made it impossible to evaluate the regeneration for the extremes of the stand density ranges in some models.
+For example, regeneration levels at low stand densities are relevant to assess how well forests are recovering e.g.
+after gap creation due to disturbance" Agreed \...
+and this paper isn't testing performance of models in predicting regeneration in large gaps.  It's possible that the models do a better job at this,  but it's not something you tested.  Suggest making that point e.g.
+in the abstract.
+
+##### 
+Response to comment 57:
+
+[Should we add here:]{.underline}
+
+["Using 15 models built to capture long-term forest dynamics at the stand, landscape, and global levels, we simulate tree regeneration at 200 sites representing large environmental gradients across Central Europe WITHOUT DISTURBANCES CONSIDERATION" ?
+or something like this?]{.underline}
+
+#### 
+[COMMENT 58]{.underline}
+
+L676-677 "Therefore, we further recommend that more effort should be invested into collecting harmonized datasets on tree regeneration." regeneration in what sense?
+  Seedlings/ ground layer/ \....  What is really needed?
+
+##### 
+Response to comment 58:
+
+Following the reviewers comment we have included this text:
+
+####
+
+[COMMENT 59]{.underline}
+
+L685-687 However, this will require an entirely different set of observed data, and potentially not all models of forest dynamics would be able to assess the relationship of these aspects on tree regeneration, e.g.
+due to the lack of disturbance or appropriate management modules.
+
+##### 
+Response to comment 59:
+
+Following the reviewers comment we have included this text:
+
+####
+
+[COMMENT 60]{.underline}
+
+L696 "Exercises like the one presented here, where the models are operated in \"blind flight\" mode." Definition of blind flight "To do something based on guesswork, intuition, or without any help or instructions"\....is that what you mean?
+
+#####
+
+Response to comment 60:
+
+Following the reviewers comment we have included this text:
+
 ## Other changes made in the manuscript

### 2023-09-22 manuscript/review/authorsResponse.qmd
Some suggestions for R Comments 1-5. More to follow on monday.

@@
 
-Thank you for pointing
+Thank you for pointing this out.
+We also discussed terminology several times and we agree.
+We followed your suggestion replaced 'regeneration' with 'recruitment' where the difference is not clear.
+However, we believe that in some places regeneration is more accurate.
+Especially in many parts of the introduction and discussion.
+We argue that regeneration is often only modeled via "recruitment", but with the aim to simulate the whole process of regeneration.
+In this context it makes more sense to use the general term (e.g., for discussion consequences.)
+
+COMMENT from YK: I would argue that recruitment is evidence for successful regeneration.
+This allows us to use recruitment for discussions observation vs.
+simulation in both aspects: technical (recruitment) and general (regeneration).
 
@@
 
+SUGGESTION YK: We understand the necessity of a clear statement in this regard. Reformulate by replacing "capture" with "do a good job in capturing" after models [L67].
+
 #### [COMMENT 5]{.underline}
@@
 [check]{.underline}
+
+SUGGESTION YK: I think the paragraph can be simplified by dissentangling the issues and the corresponding approaches to resolve them.
+I suggest to resolve this comment with changing the text something like:
+
+A focus on the modeling of tree regeneration processes is sorely
+needed and overdue [@price2001; @walck2011] if we are to make reliable
+projections of future forest dynamics, i.e., when the models need to be
+operated in extrapolation mode, as well as from a fundamental ecological
+point of view for increasing systems understanding. 
+Tree regeneration has often been overlooked in previous research [@hanbury-brown2022] because of the challenges induced by trade offs between the spatial and temporal scales on which processes of regeneration act and the availability of data needed to analyse these processes [@clark1999].
+Data on forest regeneration data are often fragmented,
+which constitutes a major problem for model building [@clark1999].
+For example monitoring on permanent plots (such as National Forest
+Inventories) often measures tree regeneration with a design that captures the rate after the stand initiation phase [@hallsby2015; @zell2019]; Data from permanent plots, although available across
+regions, are highly heterogeneous, rendering their use difficult in a
+modeling context [@käber2021]; and targeted studies to
+measure tree regeneration on experimental sites are often limited in
+spatial extent, species studied, or the sub-set of processes that are
+targeted [@berdanier2016; @collet2006].
 

### 2023-09-24 manuscript/manuscriptRW.Rmd
changes in both reviewers.

@@
 
-**OPEN RESEARCH STATEMENT:** This submission uses novel code, both the
-data and code are provided in an external repository to be evaluated
-during the peer review process. The repository can be downloaded from
-[this figshare private link:
-https://figshare.com/s/100857ba67d2351dcb15](https://figshare.com/s/100857ba67d2351dcb15).
-The protocol is available at @díaz-yáñez2022 ([with link:
-https://osf.io/czdxp/](https://osf.io/czdxp/)). We will make publicly
-available the code and data if the manuscript is accepted.
-
-**Keywords:** tree regeneration \| tree recruitment \| tree
-establishment \| forest dynamics \| models \| landscape \| stand \|
-global
+**OPEN RESEARCH STATEMENT:** This submission uses novel code, both the data and code are provided in an external repository to be evaluated during the peer review process.
+The repository can be downloaded from [this figshare private link: https://figshare.com/s/100857ba67d2351dcb15](https://figshare.com/s/100857ba67d2351dcb15).
+The protocol is available at @díaz-yáñez2022 ([with link: https://osf.io/czdxp/](https://osf.io/czdxp/)).
+We will make publicly available the code and data if the manuscript is accepted.
+
+**Keywords:** tree regeneration \| tree recruitment \| tree establishment \| forest dynamics \| models \| landscape \| stand \| global
 
@@
 
-Tree regeneration is a key process in forest dynamics, particularly in
-the context of forest resilience and climate change. Models are pivotal
-for assessing long-term forest dynamics, and they have been in use for
-more than 50 years. However, there is a need to evaluate their capacity
-to accurately represent tree regeneration. We assess how well current
-models capture the overall abundance, species composition, and mortality
-of tree regeneration.
-
-Using 15 models built to capture long-term forest dynamics at the stand,
-landscape, and global levels, we simulate tree regeneration at 200 sites
-representing large environmental gradients across Central Europe. The
-results are evaluated against extensive data from unmanaged forests.
-
-Most of the models overestimate regeneration levels, which is only
-compensated in some models by high simulated mortality rates in the
-early stages of individual trees dynamics. Simulated species diversity
-of regeneration matches the observed ranges. Models simulating higher
-species diversity at the stand level do not present higher regeneration
-diversity. The effect of light availability on regeneration levels is
-captured better than the effect of temperature and soil moisture, but
-patterns are not consistent across models. Increasing complexity in the
-tree regeneration modules of the models is not related to higher
-accuracy of simulated tree regeneration. Furthermore, individual model
-design is more important than scale (stand, landscape, global) and
-approach (empirical, process-based) for accurately capturing tree
-regeneration.
-
-Despite considerable mismatches between simulation results and data, it
-is remarkable that most models capture the essential features of the
-highly complex process of tree regeneration, while not having been
-parameterized with such data. We conclude that much can be gained by
-evaluating and refining the modeling of regeneration processes. This has
-the potential to render long-term projections of forest dynamics under
-changing environmental conditions that are much more robust.
+Tree regeneration is a key process in forest dynamics, particularly in the context of forest resilience and climate change.
+Models are pivotal for assessing long-term forest dynamics, and they have been in use for more than 50 years.
+However, there is a need to evaluate their capacity to accurately represent tree regeneration.
+We assess how well current models capture the overall abundance, species composition, and mortality of tree regeneration.
+
+Using 15 models built to capture long-term forest dynamics at the stand, landscape, and global levels, we simulate tree regeneration at 200 sites representing large environmental gradients across Central Europe.
+The results are evaluated against extensive data from unmanaged forests.
+
+Most of the models overestimate regeneration levels, which is only compensated in some models by high simulated mortality rates in the early stages of individual trees dynamics.
+Simulated species diversity of regeneration matches the observed ranges.
+Models simulating higher species diversity at the stand level do not present higher regeneration diversity.
+The effect of light availability on regeneration levels is captured better than the effect of temperature and soil moisture, but patterns are not consistent across models.
+Increasing complexity in the tree regeneration modules of the models is not related to higher accuracy of simulated tree regeneration.
+Furthermore, individual model design is more important than scale (stand, landscape, global) and approach (empirical, process-based) for accurately capturing tree regeneration.
+
+Despite the mismatches between simulation results and data, it is remarkable that most models capture the essential features of the highly complex process of tree regeneration, while not having been parameterized with such data.
+We conclude that much can be gained by evaluating and refining the modeling of regeneration processes.
+This has the potential to render long-term projections of forest dynamics under changing environmental conditions that are much more robust.
 
@@
 
-Forests provide a wide range of products and services of vital
-importance to humankind [@foresteurope2020]. Under the pressure of
-climate change, increasing disturbance impacts and changing societal
-demands on forest ecosystem services, it is becoming ever more important
-to understand how forest structure, composition and function will
-change, and to evaluate forest capacity to adapt to or be resilient
-after disturbance [@lindner2010; @seidl2022]. A wide range of models of
-forest dynamics have been developed over the past decades considering
-the impacts of climate [@bugmann2022]. From these studies, it is evident
-that we have a reasonably good understanding how to model tree growth
-[@vanclay1997; @bugmann1996], and substantial efforts have been
-dedicated to improving the representation of tree mortality
-[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is a much
-less studied process that is often represented rather coarsely in models
-[@walck2011; @leishman1992; @price2001]. This presents an important
-research gap, particularly in the context of climate-induced forest
-disturbances and forest resilience.
-
-Tree regeneration arises from multiple processes including pollination,
-fruit maturation, seed production, dispersal, germination, juvenile
-growth and survival [@price2001; @vacchiano2018]. All these processes
-are difficult to assess, and some of them are scarcely understood and
-thus appear highly stochastic [@bogdziewicz2021]. Identifying the
-appropriate level of complexity for the mathematical formulation of the
-key factors that are leading to successful tree regeneration is
-challenging. Currently, tree regeneration processes in dynamic forest
-models are handled in a multitude of ways [@konig_tree_2022;
-@bugmann2022]: from 1) entirely ignoring them [as done in classical
-forest growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2)
-the use of a few simple environmental filters, as done in most forest
-"gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
-@hickler2012; @smith2001], to 3) complex approaches that incorporate
-local feedback from the canopy, multiple ecological processes and often
-also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
-statistical parameterizations, which however are not easy to extrapolate
-in space and time [e.g., @ribbens1994].
-
-Overall, models are needed to (1) synthesize existing empirical data and
-explore their relationships, (2) assess future tree regeneration e.g. in
-the context of global change scenarios, and (3) identify the most
-important processes that are shaping ecological patterns. Given the
-current strategies that are used in models of forest dynamics to
-represent tree regeneration, model behavior often is prone to problems,
-such as very high levels of tree regeneration that necessitate excess
-mortality at early stages of tree life to simulate correct stand
-structure and composition. Also, inadequately high species diversity in
-tree regeneration may be simulated, which is characteristic of
-"classical" forest gap models [@gutiérrez2016], at least as long as the
-simulation set-up comprises a multitude of species. Some models use
-calibration against local canopy-level data to constrain simulated
-regeneration levels, which is likely to hamper the general applicability
-of these models e.g. under scenarios of climate change. Furthermore,
-correctly capturing the species composition in tree regeneration as a
-function of the presence of seed trees in the canopy is often a
-particular challenge, potentially leading to unrealistic successional
-drift in the model, which must be corrected by factors that are hard to
-parameterize [e.g., @lischke2006]. A related issue is the excessive
-reduction of species diversity due to positive feedback effects, such
-that eventually just single-species stands remain. This is sometimes
-corrected by the incorporation of a low level of seed influx of all
-species at all times [@schumacher2006] or by restricting the number of
-seeds per species in the seedbank [@lischke2006]. However, simulated
-species composition is usually exceedingly sensitive to assumptions
-about seed availability, whereas the parameters of such functions are
-poorly constrained by field data. Lastly, there is often a problem with
-insufficient observational constraints on parameter values for models
-that start from very small tree sizes (e.g., 10 cm of height or even
-seeds) and track tree development in a process-oriented manner by
-considering a multitude of ecological influences, rather than
-emphasizing tree regeneration into a larger size class [e.g.,
-@wehrli2006].
-
-Thus, a focus on the modeling of tree regeneration processes is sorely
-needed and overdue [@price2001; @walck2011] if we are to make reliable
-projections of future forest dynamics, i.e., when the models need to be
-operated in extrapolation mode, as well as from a fundamental ecological
-point of view for increasing systems understanding. Also in empirical
-ecology, tree regeneration has not received a lot of attention
-[@hanbury-brown2022], although many sub-processes have been studied in
-detail [@miina2006]. Yet, little data are available that covers all the
-processes within one species along environmental gradients, let alone
-for a vast suite of species. Thus, the available data are fragmented,
-which constitutes a major problem for model building [@clark1999]. On
-the one hand, monitoring on permanent plots (such as National Forest
-Inventories) often measures tree regeneration [@zell2019], however with
-a design that captures the rate after the stand initiation phase
-[@hallsby2015]. Data from permanent plots, although available across
-regions, are highly heterogeneous, rendering their use difficult in a
-modeling context [@käber2021]. On the other hand, targeted studies to
-measure tree regeneration on experimental sites are often limited in
-spatial extent, species studied, or the sub-set of processes that are
-targeted [@berdanier2016; @collet2006].
-
-In the present study, for the first time, a large number of forest
-models commonly used to assess forest dynamics under climate change are
-evaluated against a continental-scale, multi-species harmonized dataset
-on tree regeneration [@käber2023]. By tree regeneration, we refer to the
-passing of trees across a specific diameter threshold ("ingrowth"). We
-included models that are based on a range of "philosophies", from models
-operating at the stand to the global scale as well as the range of
-models from empirically derived to "process-based" [e.g., @bugmann1996;
-@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
-tree regeneration patterns in nature and the large number of factors
-driving and constraining this process --- including some that are not
-incorporated explicitly in most models, such as deer browsing --- we do
-not aim for a detailed statistical evaluation of each model. Instead, we
-aim to evaluate the general regeneration patterns and magnitudes
-simulated by the models, and benchmark the simulated regeneration niche
-of multiple species against empirical data along a wide environmental
-gradient of temperature, moisture, and light availability [@grubb1977].
-
-More specifically, we aim to answer the following questions: (1) Are
-models of forest dynamics capturing accurately tree regeneration levels,
-initial tree species diversity and mortality in the regeneration? (2) Do
-model traits explain differences in model performance? (3) How well do
-the models capture total regeneration and individual species
-regeneration niches across environmental gradients of light
-availability, temperature, and soil moisture?
+Forests provide a wide range of products and services of vital importance to humankind [@foresteurope2020].
+Under the pressure of climate change, increasing disturbance impacts and changing societal demands on forest ecosystem services, it is becoming ever more important to understand how forest structure, composition and function will change, and to evaluate forest capacity to adapt to or be resilient after disturbance [@lindner2010; @seidl2022].
+A wide range of models of forest dynamics have been developed over the past decades considering the impacts of climate [@bugmann2022].
+From these studies, it is evident that we have a reasonably good understanding how to model tree growth [@vanclay1997; @bugmann1996], and substantial efforts have been dedicated to improving the representation of tree mortality [@bugmann2019; @cailleret2017].
+In contrast, tree regeneration is a much less studied process that is often represented rather coarsely in models [@walck2011; @leishman1992; @price2001].
+This presents an important research gap, particularly in the context of climate-induced forest disturbances and forest resilience.
+
+Tree regeneration arises from multiple processes including pollination, fruit maturation, seed production, dispersal, germination, juvenile growth and survival [@price2001; @vacchiano2018].
+All these processes are difficult to assess, and some of them are scarcely understood and thus appear highly stochastic [@bogdziewicz2021].
+Identifying the appropriate level of complexity for the mathematical formulation of the key factors that are leading to successful tree regeneration is challenging.
+Currently, tree regeneration processes in dynamic forest models are handled in a multitude of ways [@konig_tree_2022; @bugmann2022]: from 1) entirely ignoring them [as done in classical forest growth models, e.g., Pretzsch et al.
+-@pretzsch2002], across 2) the use of a few simple environmental filters, as done in most forest "gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g., @hickler2012; @smith2001], to 3) complex approaches that incorporate local feedback from the canopy, multiple ecological processes and often also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based statistical parameterizations, which however are not easy to extrapolate in space and time [e.g., @ribbens1994].
+
+Overall, models are needed to (1) synthesize existing empirical data and explore their relationships, (2) assess future tree regeneration e.g.
+in the context of global change scenarios, and (3) identify the most important processes that are shaping ecological patterns.
+Given the current strategies that are used in models of forest dynamics to represent tree regeneration, model behavior often is prone to problems, such as very high levels of tree regeneration that necessitate excess mortality at early stages of tree life to simulate correct stand structure and composition.
+Also, inadequately high species diversity in tree regeneration may be simulated, which is characteristic of "classical" forest gap models [@gutiérrez2016], at least as long as the simulation set-up comprises a multitude of species.
+Some models use calibration against local canopy-level data to constrain simulated regeneration levels, which is likely to hamper the general applicability of these models e.g.
+under scenarios of climate change.
+Furthermore, correctly capturing the species composition in tree regeneration as a function of the presence of seed trees in the canopy is often a particular challenge, potentially leading to unrealistic successional drift in the model, which must be corrected by factors that are hard to parameterize [e.g., @lischke2006].
+A related issue is the excessive reduction of species diversity due to positive feedback effects, such that eventually just single-species stands remain.
+This is sometimes corrected by the incorporation of a low level of seed influx of all species at all times [@schumacher2006] or by restricting the number of seeds per species in the seedbank [@lischke2006].
+However, simulated species composition is usually exceedingly sensitive to assumptions about seed availability, whereas the parameters of such functions are poorly constrained by field data.
+Lastly, there is often a problem with insufficient observational constraints on parameter values for models that start from very small tree sizes (e.g., 10 cm of height or even seeds) and track tree development in a process-oriented manner by considering a multitude of ecological influences, rather than emphasizing tree regeneration into a larger size class [e.g., @wehrli2006].
+
+Modeling tree regeneration processes is challenging, and even in empirical ecology, it has not received much attention [@hanbury-brown2022], although many sub-processes have been studied in detail [@miina2006].
+Yet, little data are available that covers all the processes within one species along environmental gradients, let alone for a vast suite of species.
+Data on forest regeneration data are often fragmented, which constitutes a major problem for model building [@clark1999a].
+For example, monitoring on permanent plots (such as National Forest Inventories) often measures tree regeneration [@zell2019], however with a design that captures the rate after the stand initiation phase [@hallsby2015].
+Data from permanent plots, although available across regions, are also highly heterogeneous, rendering their use difficult in a modeling context [@käber2021], and targeted studies to measure tree regeneration on experimental sites are often limited in spatial extent, species studied, or the sub-set of processes that are targeted [@berdanier2016; @collet2006].
+
+Thus, a focus on the modeling of tree regeneration processes is sorely needed and overdue [@price2001; @walck2011] if we are to make reliable projections of future forest dynamics, i.e., when the models need to be operated in extrapolation mode, as well as from a fundamental ecological point of view for increasing systems understanding.
+In the present study, for the first time, a large number of forest models commonly used to assess forest dynamics under climate change are evaluated against a continental-scale, multi-species harmonized dataset on tree regeneration [@käber2023].
+By tree regeneration, we refer to the passing of trees across a specific diameter threshold ("ingrowth").
+We included models that are based on a range of "philosophies", from models operating at the stand to the global scale as well as the range of models from empirically derived to "process-based" [e.g., @bugmann1996; @fabrika2012; @lexer2001; @reyer2014].
+Due to the large variability in tree regeneration patterns in nature and the large number of factors driving and constraining this process --- including some that are not incorporated explicitly in most models, such as deer browsing --- we do not aim for a detailed statistical evaluation of each model.
+Instead, we aim to evaluate the general regeneration patterns and magnitudes simulated by the models, and benchmark the simulated regeneration niche of multiple species against empirical data along a wide environmental gradient of temperature, moisture, and light availability [@grubb1977].
+
+More specifically, we aim to answer the following questions: (1) Are models of forest dynamics capturing accurately tree regeneration levels, initial tree species diversity and mortality in the regeneration?
+(2) Do model traits explain differences in model performance?
+(3) How well do the models capture total regeneration and individual species regeneration niches across environmental gradients of light availability, temperature, and soil moisture?
+We evaluate the performance of regeneration models by analyzing simulated data alongside observed data.
+Specifically, we examine tree species diversity, mortality rates in tree regeneration, and regeneration levels.
+Then, we link the model's traits and performance to define regeneration and we assess the regeneration gradients on regeneration niches.
+Finally, we provide research recommendations for next steps towards improving tree regeneration modules.
 
@@
 
-Fifteen models of forest dynamics (including two models featuring model
-variants) were used to simulate forest dynamics at the stand, landscape
-or global scale (Table \@ref(tab:tableModels)). The approaches used for
-model construction and their origin differ strongly, with most of the
-models featuring a largely "process-based" approach, whereas two models
-are based on formulations derived from the statistical analysis of
-inventory data (SIBYLA and xComp). Some of the models largely rely on
-the original approach underlying forest gap models (e.g., ForClim 1),
-whereas some are based on plant ecophysiological processes (e.g.,
-FORMIND and iLand). The different approaches underlying the models have
-strongly influenced the formulation of tree regeneration processes.
-
-The regeneration models can be differentiated into 'regeneration' and
-'recruitment' approaches [@vanclay1994; @konig_tree_2022]. *Regeneration
-models* include processes such as flowering and pollination, seed
-production, seed dispersal, germination, and seedling growth, which
-ultimately lead to the simulated number of established trees.
-*Recruitment models*, in contrast, introduce a number of new trees with
-certain characteristics such as biomass or diameter, without explicitly
-considering earlier development processes. We can further distinguish
-models that feature a feedback in the regeneration, i.e., where the
-level and species composition of regeneration is influenced by the
-existing adult tree community via the production of seeds, seedlings or
-saplings, from models that do not contain such a feedback.
-
-The starting point for tree regeneration in the models differs as well,
-ranging from seedbank, seed or seedling to sapling (i.e., trees much
-larger than 10 cm height, often ca. 2-3 m tall). Models that start from
-seed need to include a larger number of ecological processes such as
-germination and survival of young seedlings, whereas models that start
-from saplings have to aggregate via parameterizations several ecological
-processes that are not treated explicitly. This latter approach reduces
-model complexity but comes at the cost of blurring process
-representation.
-
-In the models used here, the overall complexity in the regeneration
-modules varies considerably. Following @bugmann2022, we can classify
-seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and
-TreeMig) as having rather high complexity in their regeneration modules
-(mean regeneration formulation complexity across all processes \>0,
-Table \@ref(tab:tableModels)). Two models, ForClim and ForCEEPS, were
-used here with two alternative variants of regeneration. ForClim variant
-1 [@bugmann1996] is based on a recruitment module that adheres closely
-to the concept introduced by @botkin1972, whereas ForClim variant 11 is
-adopting a different approach where individual species properties and
-their relation to the environment are important [@huber2020]. These two
-model variants allow us to evaluate a more process-based and complex
-module (variant 11) against a simple module (variant 1), while the rest
-of the model structure is identical. Similarly, the two ForCEEPS
-variants allow us to isolate the importance of the canopy feedback
-(i.e., simulated actual composition and relative abundance of species in
-the plot) via seed trees for the quantity and quality (e.g., diversity
-and composition) of simulated regeneration, as one variant includes this
-feedback whereas the other does not.
+Fifteen models of forest dynamics (including two models featuring model variants) were used to simulate forest dynamics at the stand, landscape or global scale (Table \@ref(tab:tableModels)).
+The approaches used for model construction and their origin differ strongly, with most of the models featuring a largely "process-based" approach, whereas two models are based on formulations derived from the statistical analysis of inventory data (SIBYLA and xComp).
+Some of the models largely rely on the original approach underlying forest gap models (e.g., ForClim 1), whereas some are based on plant ecophysiological processes (e.g., FORMIND and iLand).
+The different approaches underlying the models have strongly influenced the formulation of tree regeneration processes.
+
+The regeneration models can be differentiated into 'regeneration' and 'recruitment' approaches [@vanclay1994; @konig_tree_2022].
+*Regeneration models* include processes such as flowering and pollination, seed production, seed dispersal, germination, and seedling growth, which ultimately lead to the simulated number of established trees.
+*Recruitment models*, in contrast, introduce a number of new trees with certain characteristics such as biomass or diameter, without explicitly considering earlier development processes.
+We can further distinguish models that feature a feedback in the regeneration, i.e., where the level and species composition of regeneration is influenced by the existing adult tree community via the production of seeds, seedlings or saplings, from models that do not contain such a feedback.
+
+The starting point for tree regeneration in the models differs as well, ranging from seedbank, seed or seedling to sapling (i.e., trees much larger than 10 cm height, often ca.
+2-3 m tall).
+Models that start from seed need to include a larger number of ecological processes such as germination and survival of young seedlings, whereas models that start from saplings have to aggregate via parameterizations several ecological processes that are not treated explicitly.
+This latter approach reduces model complexity but comes at the cost of blurring process representation.
+
+In the models used here, the overall complexity in the regeneration modules varies considerably.
+Following @bugmann2022, we can classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and TreeMig) as having rather high complexity in their regeneration modules (mean regeneration formulation complexity across all processes \>0, Table \@ref(tab:tableModels)).
+Two models, ForClim and ForCEEPS, were used here with two alternative variants of regeneration.
+ForClim variant 1 [@bugmann1996] is based on a recruitment module that adheres closely to the concept introduced by @botkin1972, whereas ForClim variant 11 is adopting a different approach where individual species properties and their relation to the environment are important [@huber2020].
+These two model variants allow us to evaluate a more process-based and complex module (variant 11) against a simple module (variant 1), while the rest of the model structure is identical.
+Similarly, the two ForCEEPS variants allow us to isolate the importance of the canopy feedback (i.e., simulated actual composition and relative abundance of species in the plot) via seed trees for the quantity and quality (e.g., diversity and composition) of simulated regeneration, as one variant includes this feedback whereas the other does not.
 
@@
 
-Regeneration data covering a wide range of environmental conditions are
-hard to obtain, and this is one of the reasons why most models of forest
-dynamics have never been confronted with a dataset covering such
-gradients over a large number of sites, to evaluate how well
-regeneration is captured. The observations used here are derived from a
-novel and unprecedented network of sites in forest reserves that
-represent the range of environmental gradients in temperature and
-precipitation in Central Europe as compiled in the framework of the
-EuFoRIa network [@euforia2019] (Figure \@ref(fig:map)). These forests
-have been unmanaged for at least 10 years prior to their designation as
-reserves, and most of them provide time series of natural forest
-dynamics over multiple decades and up to ca. 80 years. The census
-periods range from 3 to 37 years, with an average of 14 years. The data
-provide information at the tree level, thus allowing for the sequential
-comparison of processes such as individual-tree regeneration and death
-between the measurements.
-
-We selected 200 sites from this network as the benchmarking dataset for
-the simulation to be representative of the environmental variation
-contained in the data. This was achieved by applying k-means clustering
-to define 200 clusters of plots from the original set of 869 plots along
-the environmental dimensions of temperature, climatic water balance,
-soil quality, slope and aspect. Each of these sites featured at least
-two consecutive measurements. Regeneration minimum size thresholds for
-these sites differed between stem diameters of 0 and 10 cm. For our
-study, we defined two datasets where one included 165 sites with a
-diameter threshold of 7 cm or lower, and the other included another 35
-sites with diameter thresholds between 7 and 10 cm.
-
-Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
-processed and aggregated following @käber2023 to provide regeneration
-rates per ha and per decade. The observations used in this study
-featured 30,900 newly established trees. Regeneration rates per site,
-sample and decade ranged from zero to 1246 trees, with a mean of 56
-trees. Adult species composition was also available for each plot. For
-more details on this unique dataset and the detailed regeneration
-information, cf. @käber2023.
+Regeneration data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset covering such gradients over a large number of sites, to evaluate how well regeneration is captured.
+The observations used here are derived from a novel and unprecedented network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network [@euforia2019] (Figure \@ref(fig:map)).
+These forests have been unmanaged for at least 10 years prior to their designation as reserves, and most of them provide time series of natural forest dynamics over multiple decades and up to ca.
+80 years.
+The census periods range from 3 to 37 years, with an average of 14 years.
+The data provide information at the tree level, thus allowing for the sequential comparison of processes such as individual-tree regeneration and death between the measurements.
+
+We selected 200 sites from this network as the benchmarking dataset for the simulation to be representative of the environmental variation contained in the data.
+This was achieved by applying k-means clustering to define 200 clusters of plots from the original set of 869 plots along the environmental dimensions of temperature, climatic water balance, soil quality, slope and aspect.
+Each of these sites featured at least two consecutive measurements.
+Regeneration minimum size thresholds for these sites differed between stem diameters of 0 and 10 cm.
+For our study, we defined two datasets where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds from 10 cm.
+Therefore we had in total 200 sites with observations of newly established trees above 10 cm and 165 sites above 7cm.
+In the results section any result from observed data at 7 cm threshold is originated from the 165 sites.
+
+Plot size ranged from 0.02 ha to 5.52 ha, and the data were further processed and aggregated following @käber2023 to provide regeneration rates per ha and per decade.
+The observations used in this study featured 30,900 newly established trees.
+Regeneration rates per site, sample and decade ranged from zero to 1246 trees, with a mean of 56 trees.
+Adult species composition was also available for each plot.
+For more details on this unique dataset and the detailed regeneration information, cf.
+@käber2023.
 
@@
 
-The overarching goal of the simulation experiments was to assess tree
-regeneration as it arises from empirical data against its representation
-in a wide range of models of forest dynamics. We define tree
-regeneration as the passing of a breast height diameter threshold of 7
-and 10 cm (synonym: ingrowth). To this end, each modeling group was
-provided with a detailed protocol [@díaz-yáñez2022] with instructions
-how to perform the simulations, input variables on climate and soil
-conditions, and the list of expected output variables. Neither were
-further site information (except for the data specified below) nor any
-data on tree regeneration or forest stand features provided prior to the
-simulation.
-
-The input variables were collected from different data sources and
-aggregated to be adapted to the needs of the different models. Time
-series of climatic variables from 1981 to 2018 were provided in hourly
-[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
-resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
-required for some models, such as relative humidity or vapor pressure
-deficit, were calculated from these variables. The final instructions
-for using the climate data were slightly different in each model, based
-on the approach that best suited the model (Table
-\@ref(tab:tableModels)). Soil quality data were provided as continuous
-values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
-The protocol also provided the elevation, slope and aspect for each of
-the 200 sites [@asterscienceteam2019], but no other spatial information
-such as coordinates, with the exception of iLand and aDGVM2, which
-required blurred coordinates to derive highly detailed soil data.
-
-The simulations were run in the absence of natural disturbances. Only
-the model LPJ-GUESS had to include a background disturbance to increase
-chances of shade-intolerant species to establish. The simulations were
-set up to sample species-specific regeneration rates per decade and per
-ha in the equilibrium state of the model, typically entailing a
-"spin-up" run (as we did not provide any forest data). The modeling
-teams decided on the simulated area and how they derived these samples.
-The exact length of the simulation was also decided by the modeling
-teams (Table \@ref(tab:tableModels)). Further details on how each
-modeling team prepared the simulations and the outputs are available in
-Supplementary Material 1.
-
-The simulations were run in the absence of management to a simulated
-equilibrium ("Potential Natural Vegetation") with the current climate,
-as the ultimate goal was to evaluate tree regeneration under comparable
-and near-equilibrium conditions. This entails the assumption that (1)
-the observations from the forest reserves reflect no traces of forest
-management, and (2) there is an equilibrium between forest dynamics and
-climate. While the former might be starting to be visible in many of the
-EuFoRIA reserves, the latter may be more debatable. However, in the
-absence of detailed data on the history of each plot in the EuFoRIa
-network, some broad assumptions had to be made. Both the width of the
-regeneration niche (i.e., in environmental space) as well as the
-intensity of the regeneration process (i.e., the number of ingrowth
-trees per area and per unit of time) were of interest.
-
-The simulations were run for mixed-species forests (not multiple
-single-species simulations) using mixtures of eleven species or genera
-for which regeneration data of sufficient quality were available from
-EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
-betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
-Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus
-sylvestris L.* The same set of species was used at all 200 sites. Two
-models included their standard set of species for the simulations, which
-is much larger (i.e., ForClim 1, ForClim 11, TreeMig). In three models,
-fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS)
-(Table \@ref(tab:tableModels)). In 4C only *Fagus sylvatica*, *Picea
-abies*, *Betula* spp., *Quercus* spp. and *Pinus sylvestris* are
-parameterised. xComp simulations did not consider *Fraxinus excelsior*
-due to a depreciated species parameterisation. In LPJ-GUESS, *Acer
-pseudoplatanus* and *Alnus glutinosa* are not parameterized and
-therefore these species could not be included in the simulation.
-Finally, the model aDGVM2 does not simulate individual species. Rather,
-community assembly processes and trait filtering generate plant
-communities that are adjusted to the biotic and abiotic conditions, and
-the simulated plants can be classified into ecological strategies based
-on their trait values in a post-processing step.
-
-Each of the models reported the number of trees crossing the two
-diameter thresholds by sampling 200 times in a 10-year interval for each
-species and per ha for each of the 200 sites. Multiple samples per site
-were used to better understand the simulated variation within each site.
-This was done using different strategies, depending on the model,
-including (1) sampling simulated data from the same 1 ha plot in the
-equilibrium over time, (2) sampling several 1 ha plots from the
-simulated forest at one specific point in time (in the equilibrium) or
-(3) a combination of (1) and (2). This resulted in 880,000 observations
-per model that simulated the 11 species included in the protocol (200
-sites, 200 samples per site, 11 species and two diameter thresholds).
-For the models that simulated additional species, their regeneration
-rates were aggregated as "others".
-
-Two models did not provide results from all the simulated samples or
-sites to avoid unrealistic results, as follows. In the model 4C, a
-threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid
-unrealistic stand basal area data, and therefore not all the samples and
-sites were reported. The reason for this is that 4C is not suitable for
-long-term simulations without management, due to misrepresentations in
-density-dependent mortality processes in long-term simulations and
-assumptions of tree geometry that lead to unrealistic single tree
-dimensions of very old individuals. LPJ-GUESS had 2% of the sample
-outputs without tree regeneration, and these were considered as zero
-stand basal area and zero regeneration for all the tree species
-simulated; one site produced grassland rather than a forest and was not
-included in the results.
+The overarching goal of the simulation experiments was to assess tree regeneration as it arises from empirical data against its representation in a wide range of models of forest dynamics.
+We define tree regeneration as the passing of a breast height diameter threshold of 7 and 10 cm (synonym: ingrowth).
+To this end, each modeling group was provided with a detailed protocol [@díaz-yáñez2022] with instructions how to perform the simulations, input variables on climate and soil conditions, and the list of expected output variables.
+Neither were further site information (except for the data specified below) nor any data on tree regeneration or forest stand features provided prior to the simulation.
+
+The input variables were collected from different data sources and aggregated to be adapted to the needs of the different models.
+Time series of climatic variables from 1981 to 2018 were provided in hourly [Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly resolution [CHELSA data, Karger et al.
+-@karger2021].
+Some variables required for some models, such as relative humidity or vapor pressure deficit, were calculated from these variables.
+The final instructions for using the climate data were slightly different in each model, based on the approach that best suited the model (Table \@ref(tab:tableModels)).
+Soil quality data were provided as continuous values between 1 and 5 [Soilgrids dataset, Hengl et al.
+-@hengl2017].
+The protocol also provided the elevation, slope and aspect for each of the 200 sites [@asterscienceteam2019], but no other spatial information such as coordinates, with the exception of iLand and aDGVM2, which required blurred coordinates to derive highly detailed soil data.
+
+The simulations were run in the absence of natural disturbances.
+Only the model LPJ-GUESS had to include a background disturbance to increase chances of shade-intolerant species to establish.
+The simulations were set up to sample species-specific regeneration rates per decade and per ha in the equilibrium state of the model, typically entailing a "spin-up" run (as we did not provide any forest data).
+The modeling teams decided on the simulated area and how they derived these samples.
+The exact length of the simulation was also decided by the modeling teams (Table \@ref(tab:tableModels)).
+Further details on how each modeling team prepared the simulations and the outputs are available in Supplementary Material 1.
+
+The simulations were run in the absence of management to a simulated equilibrium ("Potential Natural Vegetation") with the current climate, as the ultimate goal was to evaluate tree regeneration under comparable and near-equilibrium conditions.
+This entails the assumption that (1) the observations from the forest reserves reflect no traces of forest management, and (2) there is an equilibrium between forest dynamics and climate.
+While the former might be starting to be visible in many of the EuFoRIA reserves, the latter may be more debatable.
+However, in the absence of detailed data on the history of each plot in the EuFoRIa network, some broad assumptions had to be made.
+Both the width of the regeneration niche (i.e., in environmental space) as well as the intensity of the regeneration process (i.e., the number of ingrowth trees per area and per unit of time) were of interest.
+
+The simulations were run for mixed-species forests (not multiple single-species simulations) using mixtures of eleven species or genera for which regeneration data of sufficient quality were available from EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp.
+L., Fraxinus excelsior L., Quercus spp.
+L., Alnus glutinosa L., Pinus sylvestris L.* The same set of species was used at all 200 sites.
+Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species simulated beyond the eleven species were lumped in a category "others".
+In three models, fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)).
+In 4C only *Fagus sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp.
+and *Pinus sylvestris* are parameterised.
+xComp simulations did not consider *Fraxinus excelsior* due to a depreciated species parameterisation.
+In LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not parameterized and therefore these species could not be included in the simulation.
+Finally, the model aDGVM2 does not simulate individual species.
+Rather, community assembly processes and trait filtering generate plant communities that are adjusted to the biotic and abiotic conditions, and the simulated plants can be classified into ecological strategies based on their trait values in a post-processing step.
+
+Each of the models reported the number of trees crossing the two diameter thresholds by sampling 200 times in a 10-year interval for each species and per ha for each of the 200 sites.
+Multiple samples per site were used to better understand the simulated variation within each site.
+This was done using different strategies, depending on the model, including (1) sampling simulated data from the same 1 ha plot in the equilibrium over time, (2) sampling several 1 ha plots from the simulated forest at one specific point in time (in the equilibrium) or (3) a combination of (1) and (2).
+This resulted in 880,000 observations per model that simulated the 11 species included in the protocol (200 sites, 200 samples per site, 11 species and two diameter thresholds).
+For the models that simulated additional species, their regeneration rates were aggregated as "others".
+
+Two models did not provide results from all the simulated samples or sites to avoid unrealistic results, as follows.
+In the model 4C, a threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid unrealistic stand basal area data, and therefore not all the samples and sites were reported.
+The reason for this is that 4C is not suitable for long-term simulations without management, due to misrepresentations in density-dependent mortality processes in long-term simulations and assumptions of tree geometry that lead to unrealistic single tree dimensions of very old individuals.
+LPJ-GUESS had 2% of the sample outputs without tree regeneration, and these were considered as zero stand basal area and zero regeneration for all the tree species simulated; one site produced grassland rather than a forest and was not included in the results.
 
@@
 ```
-where $s$ is the total number of species present and that have a basal
-area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
-species $i$ in sample $n$ calculated as the regeneration basal area
-($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal
-area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
-proportion of species $i$ calculated as the basal area of all trees
-($BA_{n,i}$) of that species $i$ relative to total basal area
-($TotalR_{BA_{n}}$) of the sample $n$. Species diversity was not
-assessed for aDGVM2 as this model does not simulate individual species.
-
-Mortality in tree regeneration was assessed based on the ratio of
-regeneration between the 7 and 10 cm diameter thresholds. We used the
-Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference
-to estimate whether the ratio of regeneration between the 7 and 10 cm
-diameter thresholds were above or below the expected theoretical rate.
-The Reineke self-thinning rule is usually calculated for even-aged,
-single species stands and assumes a fixed relationship between the
-number of stems and the quadratic mean diameter in fully stocked pure
-stands. The value used in our comparisons was 1.77 (i.e., we expect stem
-numbers at 7 cm to be 77% higher than at 10 cm), calculated using
-Equation \ref{eqReineke}.
+where $s$ is the total number of species present and that have a basal area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of species $i$ in sample $n$ calculated as the regeneration basal area ($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the proportion of species $i$ calculated as the basal area of all trees ($BA_{n,i}$) of that species $i$ relative to total basal area ($TotalR_{BA_{n}}$) of the sample $n$.
+Species diversity was not assessed for aDGVM2 as this model does not simulate individual species.
+
+Mortality in tree regeneration was assessed based on the ratio of regeneration between the 7 and 10 cm diameter thresholds.
+We used the Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference to estimate whether the ratio of regeneration between the 7 and 10 cm diameter thresholds were above or below the expected theoretical rate.
+The Reineke self-thinning rule is usually calculated for even-aged, single species stands and assumes a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands.
+The value used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm to be 77% higher than at 10 cm), calculated using Equation \ref{eqReineke}.
 
@@
 ```
-We assessed model performance in relation to model traits focusing on
-(1) model complexity as defined by @bugmann2022, (2) model type
-(empirical or process based), (3) the presence or absence of a canopy
-feedback for regeneration, and (4) the scale of application of the model
-(stand, landscape or global). We tested for significant differences
-using two-sided t-tests by considering each of these model features
-regarding the observed and simulated regeneration levels and species
-diversity, respectively.
-
-Total regeneration and the regeneration niches of the individual species
-were evaluated across the environmental gradients of light availability,
-temperature, and soil moisture. Stand basal area (the basal area of all
-the trees in each sample including tree regeneration) was used as a
-proxy for light availability at the forest floor, the annual degree-day
-sum [@allen1976; @fischlin1995] as a proxy for growing season warmth,
-and the climatic water balance as a proxy for soil moisture
-[@speich2019]. Regeneration values were calculated as the mean across
-the 200 samples per site. The observed data were modeled using a
-Generalized Additive Model [@wood2011] with a negative binomial
-distribution and restricted maximum likelihood to better understand the
-relationship between the environmental gradients and the levels of
-observed regeneration, relative to the simulation results per model.
-
-In order to analyze the regeneration niches across the climatic
-gradients, we focused on five common tree species or genera: *Abies
-alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
-*Quercus* spp. For these species, we calculated the share in the
-regeneration basal area per site as the mean across the available
-samples per site:
+We assessed model performance in relation to model traits focusing on (1) model complexity as defined by @bugmann2022, (2) model type (empirical or process based), (3) the presence or absence of a canopy feedback for regeneration, and (4) the scale of application of the model (stand, landscape or global).
+We tested for significant differences using two-sided t-tests by considering each of these model features regarding the observed and simulated regeneration levels and species diversity, respectively.
+
+Total regeneration and the regeneration niches of the individual species were evaluated across the environmental gradients of light availability, temperature, and soil moisture.
+Stand basal area (the basal area of all the trees in each sample including tree regeneration) was used as a proxy for light availability at the forest floor, the annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing season warmth, and the climatic water balance as a proxy for soil moisture [@speich2019].
+Regeneration values were calculated as the mean across the 200 samples per site.
+The observed data were modeled using a Generalized Additive Model [@wood2011] with a negative binomial distribution and restricted maximum likelihood to better understand the relationship between the environmental gradients and the levels of observed regeneration, relative to the simulation results per model.
+
+In order to analyze the regeneration niches across the climatic gradients, we focused on five common tree species or genera: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and *Quercus* spp.
+For these species, we calculated the share in the regeneration basal area per site as the mean across the available samples per site:
 
@@
 ```
-where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
-is the mean basal area in the regeneration (subsequently referred to as
-"regeneration basal area") of species $i$ per site $t$ across the
-available samples at that site. Furthermore, regarding the share in the
-regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we
-categorized this as zero when the regeneration basal area of that
-species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the
-regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) and the
-total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$)
-equaled zero.
+where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$ is the mean basal area in the regeneration (subsequently referred to as "regeneration basal area") of species $i$ at site $t$ across the available samples at that site.
+Furthermore, regarding the share in the regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero.
 
@@
 
-Simulated regeneration levels varied strongly across the 15 models and
-typically did not match the levels found in the forest reserve data
-(Figure \@ref(fig:over)). Regeneration was overestimated in most models
-for both the 7 and 10 cm diameter thresholds, with the exception of the
-empirical stand model SIBYLA, the landscape model Landis-II and the
-global model aDGVM2, which estimated regeneration levels at the lower
-end of the plausibility interval of the observed data (Figure
-\@ref(fig:over)). The models with the largest overestimation were the
-stand model PICUS and the landscape model TreeMig. For most models, the
-variability of simulated regeneration levels across the 200 sites
-(visible from the interquartile range in the box plots of Figure
-\@ref(fig:over)) was similar to or smaller than observed, with the
-notable exception of PICUS, where simulated regeneration variability was
-much larger.
-
-Both the observations and the simulated data had no regeneration in some
-samples and at some sites (for details, cf. Table S1). The observed data
-had 4% of the samples with no regeneration. Only three models had a
-larger proportion of no regeneration (4C, Landis-II and aDVM2). Two
-models always simulated regeneration for both the 7 and 10 cm threshold
-(xComp and TreeMig), i.e., they did not feature any zero values. The
-other ten models had a very low percentage of samples with no
-regeneration (0.01% - 2.39%), i.e., they had distinctly fewer
-occurrences of zeros compared to the observations.
+Simulated regeneration levels varied strongly across the 15 models and typically did not match the levels found in the forest reserve data (Figure \@ref(fig:over)).
+Regeneration was overestimated in most models for both the 7 and 10 cm diameter thresholds, with the exception of the empirical stand model SIBYLA, the landscape model Landis-II and the global model aDGVM2, which estimated regeneration levels at the lower end of the plausibility interval of the observed data (Figure \@ref(fig:over)).
+The models with the largest overestimation were the stand model PICUS and the landscape model TreeMig.
+For most models, the variability of simulated regeneration levels across the 200 sites (visible from the interquartile range in the box plots of Figure \@ref(fig:over)) was similar to or smaller than observed, with the notable exception of PICUS, where simulated regeneration variability was much larger.
+
+Both the observations and the simulated data had no regeneration in some samples and at some sites (for details, cf.
+Table S1).
+The observed data had 4% of the samples with no regeneration.
+Only three models had a larger proportion of no regeneration (4C, Landis-II and aDVM2).
+Two models always simulated regeneration for both the 7 and 10 cm threshold (xComp and TreeMig), i.e., they did not feature any zero values.
+The other ten models had a very low percentage of samples with no regeneration (0.01% - 2.39%), i.e., they had distinctly fewer occurrences of zeros compared to the observations.
 
@@
 
-Most models matched the level of diversity of the observed data quite
-well (Figure \@ref(fig:H710)). Five models overestimated regeneration
-diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
-particularly for the 7 cm diameter threshold. The model 4C is a special
-case, as it simulated five species only, i.e., its diversity values are
-not directly comparable to those of the other models, nor to the
-observations. Only one model, Landis-II, consistently underestimated
-regeneration diversity.
-
-In most models, there were only small but significant differences in the
-species diversity of regeneration between 7 and 10 cm across sites
-(Tables S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS)
-maintained the regeneration diversity between the 7 and 10 cm thresholds
-(i.e., the differences between them were not significant, see Table S2),
-and the same was evident from the observed data.
-
-For both the observed and the simulated data, and both regeneration
-thresholds, species diversity in the regeneration compared to
-stand-level diversity did not feature clear patterns across the models
-(Figures S1 and S2). Most models captured reasonably well or
-overpredicted species diversity at the stand level (Figure
-\@ref(fig:combinedRichness)A, left and center; cf. Figure S3), but a
-characteristic feature was that the simulations had a much lower
-variability of diversity compared to observations. Models overpredicting
-species diversity included ForClim1, ForClim 11, FORMIND and TreeMig
-(Figure \@ref(fig:combinedRichness)A, left). Several models
-underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C
-(Figure \@ref(fig:combinedRichness)A, right).
-
-The majority of the models overpredicted regeneration diversity; as
-already observed in Figure \@ref(fig:H710), only two of them
-underestimated it (4C and Landis-II) (Figure
-\@ref(fig:combinedRichness)B and Figure S4). Both the models with
-feedback and those without feedback from the adult trees to regeneration
-(via seed production) overestimated, underestimated or captured
-reasonably well regeneration species diversity compared to the
-observations. Again, simulated diversity had considerably lower
-variability than observed diversity in the regeneration.
+Most models matched the level of diversity of the observed data quite well (Figure \@ref(fig:H710)).
+Five models overestimated regeneration diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter particularly for the 7 cm diameter threshold.
+The model 4C is a special case, as it simulated five species only, i.e., its diversity values are not directly comparable to those of the other models, nor to the observations.
+Only one model, Landis-II, consistently underestimated regeneration diversity.
+
+In most models, there were only small but significant differences in the species diversity of regeneration between 7 and 10 cm across sites (Tables S2).
+Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained the regeneration diversity between the 7 and 10 cm thresholds (i.e., the differences between them were not significant, see Table S2), and the same was evident from the observed data.
+
+For both the observed and the simulated data, and both regeneration thresholds, species diversity in the regeneration compared to stand-level diversity did not feature clear patterns across the models (Figures S1 and S2).
+Most models captured reasonably well or overpredicted species diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left and center; cf.
+Figure S3), but a characteristic feature was that the simulations had a much lower variability of diversity compared to observations.
+Models overpredicting species diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A, left).
+Several models underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).
+
+The majority of the models overpredicted regeneration diversity; as already observed in Figure \@ref(fig:H710), only two of them underestimated it (4C and Landis-II) (Figure \@ref(fig:combinedRichness)B and Figure S4).
+Both the models with feedback and those without feedback from the adult trees to regeneration (via seed production) overestimated, underestimated or captured reasonably well regeneration species diversity compared to the observations.
+Again, simulated diversity had considerably lower variability than observed diversity in the regeneration.
 
@@
 
-There were strong differences among the models regarding the mortality
-rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the
-models featured a mortality rate significantly larger than the observed
-data, thus at least partially compensating for the general
-overestimation of regeneration levels (Figure \@ref(fig:over) and Table
-S3); six models did not feature a significant difference compared to the
-observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
-Four models (xComp, PICUS, LandClim and TreeMig) featured very high
-mortality rates (i.e., well above the Reineke self-thinning line), which
-compensated for the strong initial overestimation of regeneration (cf.
-Figure \@ref(fig:over)). Interestingly, two models that underestimated
-overall regeneration levels (SIBYLA and Landis-II; cf. Figure
-\@ref(fig:over)) featured mortality rates that were close to but still
-above the self-thinning line (Figure \@ref(fig:mortality11)).
-
-The fact that some models (and observed data) featured lower mortality
-than expected by Reineke's self-thinning rule may point to facilitation,
-or simply a higher mortality rate before the trees had reached 7 cm
-diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
-whose mortality was well above the self-thinning line, indicates that
-they feature higher regeneration mortality in mixed-species forests
-compared to even-aged, single-species forests. Some models yielded a
-mortality rate of regeneration that is broadly compatible with the
-self-thinning rule (e.g., 4C and LandClim).
-
-The three models with the largest overestimation of the proportion of
-regeneration at the 7 cm threshold featured the highest mortality
-between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The
-other models that overestimated regeneration had mortalities lower than
-the expected self-thinning ratio, with the exception of LandClim (Figure
-\@ref(fig:mortality13)B). It is noteworthy that most of the models that
-featured a low ratio of regeneration between 7 and 10 cm (i.e.,
-ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small
-overestimation of regeneration at the 7 cm threshold (Figure
-\@ref(fig:mortality11)). aDGVM2 was the only model that underestimated
-regeneration at the 7 cm threshold while featuring a low mortality
-between the two thresholds (Figure \@ref(fig:mortality13)C), and only a
-few models were close to the correct regeneration levels at 7 cm while
-also being close to the Reineke line (SIBYLA, Landis-II; cf. Figure
-\@ref(fig:mortality13)D; and 4C as well as LandClim).
-
-Lastly, some models that had a low number of individuals at the 7 cm
-threshold also had a smaller decrease of regeneration between 7 and 10
-cm (ratio 7-10 cm) (Figure S5). In these models, competition and
-self-thinning are either not pronounced or must have occurred before the
-trees had reached 7 cm. However, this pattern was not consistent across
-models (Figure S5). It is noteworthy that there is a relationship
-between the regeneration levels at 7 cm and the mortality rate (Table
-S4): most of the models showed a positive effect (higher regeneration at
-7 cm is coupled to an increase of the mortality between 7 and 10 cm).
-While some models showed a negative effect, the linear trend was not
-pronounced and not always significant regarding the slope.
+There were strong differences among the models regarding the mortality rate between 7 and 10 cm (Figure \@ref(fig:mortality11)).
+Most of the models featured a mortality rate significantly larger than the observed data, thus at least partially compensating for the general overestimation of regeneration levels (Figure \@ref(fig:over) and Table S3); six models did not feature a significant difference compared to the observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
+Four models (xComp, PICUS, LandClim and TreeMig) featured very high mortality rates (i.e., well above the Reineke self-thinning line), which compensated for the strong initial overestimation of regeneration (cf.
+Figure \@ref(fig:over)).
+Interestingly, two models that underestimated overall regeneration levels (SIBYLA and Landis-II; cf.
+Figure \@ref(fig:over)) featured mortality rates that were close to but still above the self-thinning line (Figure \@ref(fig:mortality11)).
+
+The fact that some models (and observed data) featured lower mortality than expected by Reineke's self-thinning rule may point to facilitation, or simply a higher mortality rate before the trees had reached 7 cm diameter.
+Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher regeneration mortality in mixed-species forests compared to even-aged, single-species forests.
+Some models yielded a mortality rate of regeneration that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).
+
+The three models with the largest overestimation of the proportion of regeneration at the 7 cm threshold featured the highest mortality between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A).
+The other models that overestimated regeneration had mortalities lower than the expected self-thinning ratio, with the exception of LandClim (Figure \@ref(fig:mortality13)B).
+It is noteworthy that most of the models that featured a low ratio of regeneration between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small overestimation of regeneration at the 7 cm threshold (Figure \@ref(fig:mortality11)).
+aDGVM2 was the only model that underestimated regeneration at the 7 cm threshold while featuring a low mortality between the two thresholds (Figure \@ref(fig:mortality13)C), and only a few models were close to the correct regeneration levels at 7 cm while also being close to the Reineke line (SIBYLA, Landis-II; cf.
+Figure \@ref(fig:mortality13)D; and 4C as well as LandClim).
+
+Lastly, some models that had a low number of individuals at the 7 cm threshold also had a smaller decrease of regeneration between 7 and 10 cm (ratio 7-10 cm) (Figure S5).
+In these models, competition and self-thinning are either not pronounced or must have occurred before the trees had reached 7 cm.
+However, this pattern was not consistent across models (Figure S5).
+It is noteworthy that there is a relationship between the regeneration levels at 7 cm and the mortality rate (Table S4): most of the models showed a positive effect (higher regeneration at 7 cm is coupled to an increase of the mortality between 7 and 10 cm).
+While some models showed a negative effect, the linear trend was not pronounced and not always significant regarding the slope.
 
@@
 
-When evaluating regeneration levels along key gradients of light
-availability (basal area), temperature (degree-days) and soil moisture
-(climatic water balance), distinct features emerged: the models
-reproduced the effect of basal area in both its magnitude and patterns
-(Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better
-than the effects of the climatic gradients, where they featured varying
-patterns.
-
-In the observed data, regeneration levels decreased clearly with
-increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a).
-This trend was captured by the Landis-II model only, albeit at much
-lower values of basal area than in the observations. The other models
-featured distinctly different trends, such as (1) an increase in
-regeneration levels with increasing total basal area followed by the
-absence of regeneration at high values of basal area (ForCEEPS,
-ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant
-regeneration levels with basal area (ForClim 11, xComp, LandClim and
-aDGVM2), or (3) an increase of regeneration up to a certain value of
-stand basal area followed by a decrease at even higher values of stand
-basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA
-and iLand).
-
-The observed regeneration did not change much across the climatic water
-balance gradient, and showed a slight increase with temperature. The
-models did not match the observed regeneration patterns across the
-gradients of soil moisture and temperature (Figure
-\@ref(fig:nvTrend710BAlightTemp)b,c and S7), and neither was there a
-clear pattern across models. Regarding the soil moisture gradient, a
-group of models featured decreasing regeneration with increasing soil
-moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim,
-TreeMig and iLand), while a few models showed the opposite trend
-(ForCEEPS(f), FORMIND and LPJ-GUESS). Across the temperature gradient,
-all the models that featured a decrease in regeneration levels with
-increasing soil moisture showed the opposite trend (i.e., more
-regeneration with increasing temperature), except for 4C. The other
-models did not feature a clear pattern.
-
-Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
-sylvestris*) and oak (*Quercus* spp.) were the species observed most
-frequently in the EuFoRIa data. Most models captured well the share of
-basal area in the regeneration of these five species compared to the
-observed data along the two major climatic gradients (Figure
-\@ref(fig:diff), cf. Figure S8). However, the models differed
-considerably in the way in which they simulated these climatic niches.
-
-First, some models overestimated regeneration across the entire
-gradients of at least one of the five main tree species (Figure
-\@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS),
-while other models overestimated regeneration of more than one species
-for a large part of the climate gradients, such as xComp, PICUS, SIBYLA
-or LPJ-GUESS (cf. red colors in Figure \@ref(fig:diff)). These trends
-were consistent for both regeneration thresholds, although the exact
-changes in the share of the regeneration basal area were different
-(Figure S8).
-
-Second, there was some congruence in the simulation results by species
-across the models. The share in the regeneration of Silver fir (*Abies
-alba*) was captured well across the gradient by most models except for
-xComp, PICUS, and LandClim, which overestimated the regeneration share
-of this species. There were also some peculiarities evident for some
-models, such as 4C that did not simulate *Abies alba*. The patterns
-across models for *Fagus sylvatica* regeneration were more complex, as
-its regeneration was underestimated by many models across the
-environmental gradients featuring negligible regeneration at most sites,
-while others overestimated it in different parts of the environmental
-space. Most models overestimated the share of *Picea abies* regeneration
-in the cold-wet part of the gradients. *Pinus sylvestris* and *Quercus*
-spp. represented a small share of the regeneration in the observed data
-mostly at warm-dry sites. Many models, however, erroneously featured
-regeneration for these two species along most of the climatic gradients,
-although with a low share of basal area. Five models (4C, FORMIND,
-SIBYLA, xComp, and LandClim) had almost no regeneration of *Pinus
-sylvestris* and *Quercus* spp., or did not feature any regeneration of
-these species at all.
-
-Lastly, there is no model that performed well across the five species
-regarding the species-specific regeneration levels in the environmental
-space (Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
-tended to consistently overestimate the regeneration share of all five
-species, but most models overestimated the regeneration share of some
-species while underestimating the share of others (e.g., ForCEEPS,
-LPJ-GUESS).
+When evaluating regeneration levels along key gradients of light availability (basal area), temperature (degree-days) and soil moisture (climatic water balance), distinct features emerged: the models reproduced the effect of basal area in both its magnitude and patterns (Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better than the effects of the climatic gradients, where they featured varying patterns.
+
+In the observed data, regeneration levels decreased clearly with increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a).
+This trend was captured by the Landis-II model only, albeit at much lower values of basal area than in the observations.
+The other models featured distinctly different trends, such as (1) an increase in regeneration levels with increasing total basal area followed by the absence of regeneration at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant regeneration levels with basal area (ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of regeneration up to a certain value of stand basal area followed by a decrease at even higher values of stand basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).
+
+The observed regeneration did not change much across the climatic water balance gradient, and showed a slight increase with temperature.
+The models did not match the observed regeneration patterns across the gradients of soil moisture and temperature (Figure \@ref(fig:nvTrend710BAlightTemp)b,c), and neither was there a clear pattern across models.
+Regarding the soil moisture gradient, a group of models featured decreasing regeneration with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS).
+Across the temperature gradient, all the models that featured a decrease in regeneration levels with increasing soil moisture showed the opposite trend (i.e., more regeneration with increasing temperature), except for 4C.
+The other models did not feature a clear pattern.
+
+Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus sylvestris*) and oak (*Quercus* spp.) were the species observed most frequently in the EuFoRIa data.
+Most models captured well the share of basal area in the regeneration of these five species compared to the observed data along the two major climatic gradients (Figure \@ref(fig:diff), cf.
+Figure S8).
+However, the models differed considerably in the way in which they simulated these climatic niches.
+
+First, some models overestimated regeneration across the entire gradients of at least one of the five main tree species (Figure \@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated regeneration of more than one species for a large part of the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
+red colors in Figure \@ref(fig:diff)).
+These trends were consistent for both regeneration thresholds, although the exact changes in the share of the regeneration basal area were different (Figure S8).
+
+Second, there was some congruence in the simulation results by species across the models.
+The share in the regeneration of Silver fir (*Abies alba*) was captured well across the gradient by most models except for xComp, PICUS, and LandClim, which overestimated the regeneration share of this species.
+There were also some peculiarities evident for some models, such as 4C that did not simulate *Abies alba*.
+The patterns across models for *Fagus sylvatica* regeneration were more complex, as its regeneration was underestimated by many models across the environmental gradients featuring negligible regeneration at most sites, while others overestimated it in different parts of the environmental space.
+Most models overestimated the share of *Picea abies* regeneration in the cold-wet part of the gradients.
+*Pinus sylvestris* and *Quercus* spp.
+represented a small share of the regeneration in the observed data mostly at warm-dry sites.
+Many models, however, erroneously featured regeneration for these two species along most of the climatic gradients, although with a low share of basal area.
+Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had almost no regeneration of *Pinus sylvestris* and *Quercus* spp., or did not feature any regeneration of these species at all.
+
+Lastly, there is no model that performed well across the five species regarding the species-specific regeneration levels in the environmental space (Figure \@ref(fig:diff)).
+Some models (e.g., ForClim11, PICUS) tended to consistently overestimate the regeneration share of all five species, but most models overestimated the regeneration share of some species while underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
 
@@
 
-Tree regeneration is a fundamental process in forest dynamics. Correctly
-capturing tree regeneration in dynamic models is fundamental to, for
-example, evaluate post-disturbance dynamics and potential long-term
-recovery trajectories as it will define the forest state in the first
-decades [@seidl2022]. If this initial phase is not captured well, we can
-not properly assess aspects such as forest resilience or the timing,
-magnitude and progression of carbon sequestration. In this study, for
-the first time, projections of tree regeneration from multiple models of
-forest dynamics were confronted with a unique dataset from unmanaged
-forest reserves across a large environmental gradient in Europe.
-
-The EuFoRIa data [@käber2023] are exceptional, particularly with respect
-to the number of records (number of sites and repeated measurements),
-which is essential for capturing a highly 'noisy' process such as tree
-regeneration. The use of this dataset for model benchmarking provided
-novel insights on the ability of state-of-the-art models to accurately
-simulate regeneration levels, species composition, and mortality in an
-early stage of tree life, i.e., between a tree diameter of 7 and 10 cm.
-Overall, by adopting this approach a much broader understanding resulted
-than if we had used these data for model calibration: it is primarily
-from the shortcomings of the models that we can gain ecological insights
-[cf. @trugman2022].
+Tree regeneration is a fundamental process in forest dynamics.
+Correctly capturing tree regeneration in dynamic models is fundamental to, for example, evaluate post-disturbance dynamics and potential long-term recovery trajectories as it will define the forest state in the first decades [@seidl2022].
+If this initial phase is not captured well, we can not properly assess aspects such as forest resilience or the timing, magnitude and progression of carbon sequestration.
+In this study, for the first time, projections of tree regeneration from multiple models of forest dynamics were confronted with a unique dataset from unmanaged forest reserves across a large environmental gradient in Europe.
+
+The EuFoRIa data [@käber2023] are exceptional, particularly with respect to the number of records (number of sites and repeated measurements), which is essential for capturing a highly 'noisy' process such as tree regeneration.
+The use of this dataset for model benchmarking provided novel insights on the ability of state-of-the-art models to accurately simulate regeneration levels, species composition, and mortality in an early stage of tree life, i.e., between a tree diameter of 7 and 10 cm.
+Overall, by adopting this approach a much broader understanding resulted than if we had used these data for model calibration: it is primarily from the shortcomings of the models that we can gain ecological insights [cf.
+@trugman2022].
 
@@
 
-Most of the models overestimated tree regeneration levels. This has
-potentially far-reaching implications, e.g., regarding biomass (and thus
-carbon) turnover, with a potential overestimation of the capacity for
-forest carbon sequestration [@pugh2019; @pan2011]. Yet, we focused on
-one specific stage, i.e., regeneration into the 7 and 10 cm diameter
-classes. Trees at this size contribute little to carbon sequestration,
-and if the excess regeneration at this stage is compensated -- in the
-models -- soon thereafter by higher mortality, overall simulated
-regeneration may still yield trustworthy results at the level of stand
-structure and related forest attributes such as basal area or biomass.
-
-From a modeling point of view, excess regeneration inevitably requires
-excess mortality rates in a later stage, either -- as observed for some
-models -- between the 7 and 10 cm diameter thresholds, or soon after the
-10 cm threshold has been crossed. In any case, correcting at early
-stages for the expected forest densities at later stages is equivalent
-to compensating for a first error (excess regeneration) by a second
-error (excess mortality). It is highly likely that biased projections
-will result, because the two errors are unlikely to be perfectly linked
-and thus will not always compensate each other. Hence, this structural
-problem of most models investigated here is problematic particularly if
-the models are to be used under novel conditions such as under climate
-change [e.g., @huber2021] or in a decision support context [e.g.,
-@thrippleton2021].
-
-Regeneration levels define the structure and composition of future
-forests, but it is equally important to correctly identify areas where
-regeneration is lacking [@rammer2021]. There are multiple constraints to
-the regeneration niche of tree species [@price2001], and therefore the
-absence of regeneration is likely to be common [@fortin2007], even over
-larger areas such as the one-hectare plots used here. Tree regeneration
-data are characterized by zero-inflation. This was clearly evident from
-the EuFoRIa dataset, but some of the models did not produce zeros at
-all, or featured a very low proportion of zero data. This substantial
-difference may be due to the fact that the simulation results were drawn
-from equilibrium forests, whereas in reality many of the forest reserves
-are recovering from past management activities and have become denser
-over the past decades [e.g., @heiri2009], leading to less regeneration
-than in an equilibrium situation. Another possible reason for this
-difference may be the exclusion of factors like deer browsing or failure
-to accurately measure establishment filters, such as those to site
-limitants or competition with herb layers.
+Most of the models overestimated tree regeneration levels.
+This has potentially far-reaching implications, e.g., regarding biomass (and thus carbon) turnover, with a potential overestimation of the capacity for forest carbon sequestration [@pugh2019; @pan2011].
+Yet, we focused on one specific stage, i.e., regeneration into the 7 and 10 cm diameter classes.
+Trees at this size contribute little to carbon sequestration, and if the excess regeneration at this stage is compensated -- in the models -- soon thereafter by higher mortality, overall simulated regeneration may still yield trustworthy results at the level of stand structure and related forest attributes such as basal area or biomass.
+
+From a modeling point of view, excess regeneration inevitably requires excess mortality rates in a later stage, either -- as observed for some models -- between the 7 and 10 cm diameter thresholds, or soon after the 10 cm threshold has been crossed.
+In any case, correcting at early stages for the expected forest densities at later stages is equivalent to compensating for a first error (excess regeneration) by a second error (excess mortality).
+It is highly likely that biased projections will result, because the two errors are unlikely to be perfectly linked and thus will not always compensate each other.
+Hence, this structural problem of most models investigated here is problematic particularly if the models are to be used under novel conditions such as under climate change [e.g., @huber2021] or in a decision support context [e.g., @thrippleton2021].
+
+Regeneration levels define the structure and composition of future forests, but it is equally important to correctly identify areas where regeneration is lacking [@rammer2021].
+There are multiple constraints to the regeneration niche of tree species [@price2001], and therefore the absence of regeneration is likely to be common [@fortin2007], even over larger areas such as the one-hectare plots used here.
+Tree regeneration data are characterized by zero-inflation.
+This was clearly evident from the EuFoRIa dataset, but some of the models did not produce zeros at all, or featured a very low proportion of zero data.
+This substantial difference may be due to the fact that the simulation results were drawn from equilibrium forests, whereas in reality many of the forest reserves are recovering from past management activities and have become denser over the past decades [e.g., @heiri2009], leading to less regeneration than in an equilibrium situation.
+Another possible reason for this difference may be the exclusion of factors like deer browsing or failure to accurately measure establishment filters, such as those to site limitants or competition with herb layers.
 
@@
 
-Correctly capturing species composition in tree regeneration is
-important to assess the future functional diversity of a forest, e.g.,
-its sensitivity to drought or resilience to disturbances [@redmond2015;
-@seidl2022]. In the simulations, overall species diversity levels in the
-regeneration were well within the observed range for half of the models
-(7 out of 14). Thus, while most of the models are facing difficulties to
-quantitatively match regeneration levels (cf. above), their performance
-is better when it comes to matching the composition of regeneration as a
-function of abiotic and biotic conditions. Most models maintained or
-even decreased regeneration diversity between the 7 and 10 cm
-thresholds, and the same is visible from the empirical data, but the
-differences were not significant. Diversity variations may be explained
-by the way the individual models consider regeneration processes
-[@konig_tree_2022], e.g., without considering the feedback from adults,
-regeneration can be linked to either seed rain from dispersal or a
-seed/seedling bank; neither process is incorporated across many models
-studied here, however.
-
-The species diversity of the entire stand was captured better by the
-models than the diversity of tree regeneration, and only a few models
-overpredicted stand-level diversity. Defining regeneration composition
-as being proportional to adult abundance, regardless of productivity,
-might be a simple and conservative assumption to safely maintain
-relative species abundances [@hanbury-brown2022], although this may be
-simplistic. Furthermore, based on the simulation results, there is no
-evidence that models with feedback from the canopy (in terms of species
-composition of regeneration) captured better the species diversity of
-the regeneration than those without feedback, with the exception of
-ForCEEPS that significantly improved regeneration species diversity with
-the model version that included feedback (ForCEEPS(f)). The similar
-performance of models with and without feedback is likely because the
-models put more weight on the regeneration niche arising from abiotic
-and biotic filters, than from the habitat niche of the adult trees
-[@grubb1977].
+Correctly capturing species composition in tree regeneration is important to assess the future functional diversity of a forest, e.g., its sensitivity to drought or resilience to disturbances [@redmond2015; @seidl2022].
+In the simulations, overall species diversity levels in the regeneration were well within the observed range for half of the models (7 out of 14).
+Thus, while most of the models are facing difficulties to quantitatively match regeneration levels (cf.
+above), their performance is better when it comes to matching the composition of regeneration as a function of abiotic and biotic conditions.
+Most models maintained or even decreased regeneration diversity between the 7 and 10 cm thresholds, and the same is visible from the empirical data, but the differences were not significant.
+Diversity variations may be explained by the way the individual models consider regeneration processes [@konig_tree_2022], e.g., without considering the feedback from adults, regeneration can be linked to either seed rain from dispersal or a seed/seedling bank; neither process is incorporated across many models studied here, however.
+
+The species diversity of the entire stand was captured better by the models than the diversity of tree regeneration, and only a few models overpredicted stand-level diversity.
+Defining regeneration composition as being proportional to adult abundance, regardless of productivity, might be a simple and conservative assumption to safely maintain relative species abundances [@hanbury-brown2022], although this may be simplistic.
+Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the regeneration than those without feedback, with the exception of ForCEEPS that significantly improved regeneration species diversity with the model version that included feedback (ForCEEPS(f)).
+The similar performance of models with and without feedback is likely because the models put more weight on the regeneration niche arising from abiotic and biotic filters, than from the habitat niche of the adult trees [@grubb1977].
 
@@
 
-There are several factors that lead to mortality during the regeneration
-phase, such as competition [@casper1997] and multiple abiotic factors
-[@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
-expect that models that overestimate regeneration may have a
-particularly high mortality between the 7 and 10 cm thresholds. However,
-this was not consistently evident from the simulations. This implies
-that these models must have an excess mortality in later stages, if we
-assume that all models were able to capture the structure and
-composition of the adult stands along the EuFoRIa gradient; this however
-was not tested here.
-
-Higher mortality towards the adult phase has important implications for
-forest dynamics and the goods and services provided by forests. On the
-one hand, mortality in later stages may erroneously enhance the share of
-less shade tolerant species in the models [@klopcic2015], therefore
-shifting the species composition. Unrealistic high stem densities for a
-longer period of time may overestimate the role of tree regrowth in
-carbon sink dynamics [@pugh2019]. Given our set of simulations and
-analyses, we cannot conclusively assess what is happening in the models,
-and further studies are required that focus on a wider range of tree
-sizes and the fate of tree regeneration along such a size continuum.
+There are several factors that lead to mortality during the regeneration phase, such as competition [@casper1997] and multiple abiotic factors [@schmid2021; @cunningham2006].
+As mentioned above, it is reasonable to expect that models that overestimate regeneration may have a particularly high mortality between the 7 and 10 cm thresholds.
+However, this was not consistently evident from the simulations.
+This implies that these models must have an excess mortality in later stages, if we assume that all models were able to capture the structure and composition of the adult stands along the EuFoRIa gradient; this however was not tested here.
+
+Higher mortality towards the adult phase has important implications for forest dynamics and the goods and services provided by forests.
+On the one hand, mortality in later stages may erroneously enhance the share of less shade tolerant species in the models [@klopcic2015], therefore shifting the species composition.
+Unrealistic high stem densities for a longer period of time may overestimate the role of tree regrowth in carbon sink dynamics [@pugh2019].
+Given our set of simulations and analyses, we cannot conclusively assess what is happening in the models, and further studies are required that focus on a wider range of tree sizes and the fate of tree regeneration along such a size continuum.
 
@@
 
-In spite of the critical considerations above, it is remarkable that
-most models did not deviate exceedingly from the observations with
-respect to simulated regeneration levels -- after all, few if any of the
-models' regeneration routines are sufficiently constrained by data, with
-the exception of the empirical models xComp and SIBYLA. It is noteworthy
-that in spite of their empirical basis, these two models did not match
-empirical regeneration levels, in a similar magnitude as the other
-models and even in opposite directions (SIBYLA: underestimation; xComp:
-overestimation). It appears that using empirical data of limited
-geographical scope and constrained to managed forests, as done in these
-models, leads to extrapolation problems already under current climatic
-conditions. For the other models, whose regeneration modules are not
-strongly constrained by empirical data, the degrees of freedom that are
-available for modeling regeneration [@konig_tree_2022] could have
-implied that model performance would be much worse than what we found.
-
-Our study showed that increasing complexity in the regeneration modules
-is not linked with a higher accuracy of the projections of regeneration
-levels, species composition or mortality at early tree stages, as there
-was no significant relationship between model performance and model
-complexity. Increasing complexity in regeneration modules has been
-motivated by better process understanding and enhancing of model
-accuracy [@bugmann2022]. However, more complex models do not necessarily
-lead to better projections and rather to a reduced transparency and
-lower predictive power [@franklin2020]. Thus, the question of what level
-of detail is appropriate and parsimonious for modeling tree regeneration
-processes remains open [@konig_tree_2022; @bugmann2022].
+In spite of the critical considerations above, it is remarkable that most models did not deviate exceedingly from the observations with respect to simulated regeneration levels -- after all, few if any of the models' regeneration routines are sufficiently constrained by data, with the exception of the empirical models xComp and SIBYLA.
+It is noteworthy that in spite of their empirical basis, these two models did not match empirical regeneration levels, in a similar magnitude as the other models and even in opposite directions (SIBYLA: underestimation; xComp: overestimation).
+It appears that using empirical data of limited geographical scope and constrained to managed forests, as done in these models, leads to extrapolation problems already under current climatic conditions.
+For the other models, whose regeneration modules are not strongly constrained by empirical data, the multiple strategies that are available for modeling regeneration [@konig_tree_2022] could have implied that model performance would be much worse than what we found.
+
+Our study showed that increasing complexity in the regeneration modules is not linked with a higher accuracy of the projections of regeneration levels, species composition or mortality at early tree stages, as there was no significant relationship between model performance and model complexity.
+Increasing complexity in regeneration modules has been motivated by better process understanding and enhancing of model accuracy [@bugmann2022].
+However, more complex models do not necessarily lead to better projections and rather to a reduced transparency and lower predictive power [@franklin2020].
+Thus, the question of what level of detail is appropriate and parsimonious for modeling tree regeneration processes remains open [@konig_tree_2022; @bugmann2022].
 
@@
 
-Competition for light as a strong filter for tree regeneration has been
-widely documented [@messier1999; @collet2006; @berdanier2016], but the
-models examined here did not reproduce this expectation (i.e.,
-decreasing regeneration levels with increasing stand basal area).
-However, it is difficult to measure light availability at large spatial
-and temporal scales. We used total stand basal area as a proxy for light
-availability [cf. @schmid2021]. However, we were unable to consider
-light availability restrictions caused by ground vegetation, which may
-be an important filter for some EuForIa stands [@woltjer2008]. We found
-pronounced differences in the stand basal area ranges simulated by the
-models, but also between models and observations. This made it
-impossible to evaluate the regeneration for the extremes of the stand
-density ranges in some models. For example, regeneration levels at low
-stand densities are relevant to assess how well forests are recovering
-e.g. after gap creation due to disturbance [@seidl2022; @grubb1977]. At
-the level of the simulated one-hectare samples, average basal area is
-typically not very low as long as no larger disturbances occur, which
-was explicitly excluded in our simulation protocol. Towards the other
-end of the spectrum, i.e., with increasing stand basal area, it would be
-reasonable to expect that the regeneration of the different species
-would become sparser and drop out entirely under low-light conditions
-[@klopcic2012; @zell2019]. However, few models showed this trend, thus
-indicating that the relation between regeneration and light availability
-is not yet captured correctly in most models. Yet, several of the models
-that did feature an increase of regeneration with increasing basal area
-include a feedback between seed production of mature trees and
-regeneration. Thus, it seems that higher light competition does not
-sufficiently compensate for increased seed availability with higher
-basal area due to a higher abundance of mature trees.
-
-There were pronounced differences how the main tree species were
-represented by the models along the environmental gradients, in
-particular the dominant species *Abies alba*, *Fagus sylvatica* and
-*Picea abies*. The regeneration levels were sampled from simulations in
-the equilibrium, and in this state it is expected that non-climax
-species such as *Pinus sylvestris* or *Quercus* would be of minor
-importance, or absent [@klopcic2015]. Most models captured this low
-abundance, which is also found in the empirical data. Thus, although the
-broad patterns are matched by many models, improvements in the
-quantification of the regeneration niche of the species are needed, but
-this cannot be done in the absence of robust datasets across multiple
+Competition for light as a strong filter for tree regeneration has been widely documented [@messier1999; @collet2006; @berdanier2016], but the models examined here did not reproduce this expectation (i.e., decreasing regeneration levels with increasing stand basal area).
+However, it is difficult to measure light availability at large spatial and temporal scales.
+We used total stand basal area as a proxy for light availability [cf.
+@schmid2021].
+However, we were unable to consider light availability restrictions caused by ground vegetation, which may be an important filter for some EuForIa stands [@woltjer2008].
+We found pronounced differences in the stand basal area ranges simulated by the models, but also between models and observations.
+This made it impossible to evaluate the regeneration for the extremes of the stand density ranges in some models.
+For example, regeneration levels at low stand densities are relevant to assess how well forests are recovering e.g.
+after gap creation due to disturbance [@seidl2022; @grubb1977].
+At the level of the simulated one-hectare samples, average basal area is typically not very low as long as no larger disturbances occur, which was explicitly excluded in our simulation protocol.
+Towards the other end of the spectrum, i.e., with increasing stand basal area, it would be reasonable to expect that the regeneration of the different species would become sparser and drop out entirely under low-light conditions [@klopcic2012; @zell2019].
+However, few models showed this trend, thus indicating that the relation between regeneration and light availability is not yet captured correctly in most models.
+Yet, several of the models that did feature an increase of regeneration with increasing basal area include a feedback between seed production of mature trees and regeneration.
+Thus, it seems that higher light competition does not sufficiently compensate for increased seed availability with higher basal area due to a higher abundance of mature trees.
+
+There were pronounced differences how the main tree species were represented by the models along the environmental gradients, in particular the dominant species *Abies alba*, *Fagus sylvatica* and *Picea abies*.
+The regeneration levels were sampled from simulations in the equilibrium, and in this state it is expected that non-climax species such as *Pinus sylvestris* or *Quercus* would be of minor importance, or absent [@klopcic2015].
+Most models captured this low abundance, which is also found in the empirical data.
+Thus, although the broad patterns are matched by many models, improvements in the quantification of the regeneration niche of the species are needed, but this cannot be done in the absence of robust datasets across multiple species.
 
@@
 
-The EuFoRIa data as used here are unique and well suited to better
-understand tree regeneration. However, three aspects of these data may
-represent considerable limitations. First, we made a comparison of tree
-regeneration in an equilibrium state, but we cannot assess how close the
-forests included in the EuFoRIa dataset were to such a state. The data
-were collected in forest reserves where no management has taken place
-for long periods of time. This makes our assumption of an equilibrium
-between forest properties and environmental drivers more reasonable than
-it might appear at first sight. In an analysis of primeval forests at
-demographic equilibrium, @brzeziecki2021 found higher regeneration rates
-than the ones observed in the EuFoRia data. Thus, the overestimation of
-regeneration rates by the models may not be so problematic. Second, the
-data were collected for regeneration above a 7 cm threshold, thus
-limiting the assessment of tree regeneration to a holistic viewpoint.
-This constituted a hard limit based on which we can understand only some
-aspects of tree regeneration, which in its entirety often comprises a
-rather long period since seed production [@price2001]. In reality, many
-environmental constraints are acting on young trees [@käber2021] that we
-were unable to assess. Yet, regeneration data with lower thresholds are
-simply not available in a harmonized manner across large environmental
-gradients. Third, the empirical data were collected from rather small
-plots, while we sampled simulated regeneration levels from 1 ha areas,
-which may lead to an incorrect representation of space. Even though the
-strategy we adopted is not ideal, it represents a common challenge when
-harmonizing diverse data sources originating from varied sampling
-strategies [@portier2022]. It would have been extremely challenging for
-such a variety of models to follow a protocol where the spatial sampling
-size was different at each of the 200 sites, and it would have
-introduced additional uncertainty in the results. Ultimately, we need to
-acknowledge that the observed data represent just a snapshot of an
-extremely complex and stochastic reality. This may at least partly
-explain e.g. the lack of clear patterns in the simulation results along
-the gradients of climate water balance and degree-days.
-
-The design of our sampling protocol did not include spatial aspects such
-as seed dispersal or detailed soil data. While we considered a wide
-range of models of forest dynamics, from stand to global scales, the
-simulation setup was limited to 1 ha regeneration values. This lack of
-consideration of spatial scale is appropriate for stand-scale models,
-but it potentially puts the landscape models at a disadvantage, as they
-have been built to be accurate at the landscape level. Without spatial
-context, we are limiting regeneration to the seed influx from the stand
-itself, unless the model has a background seed input. Yet, the global
-models should not be at a disadvantage due to the limited spatial
-considerations [but cf. Snell et al. -@snell2018; or Lehsten et al.
--@lehsten2019], as they are usually lacking dispersal between cells and
-are based on a strong abstraction of horizontal space
-[@hanbury-brown2022]. Lastly, detailed data on soil conditions were not
-available from the observed data, and independent, admittedly rather
-coarse data for soil and climatic water balance had to be used instead.
-It is noteworthy that many models represent drought based on detailed
-indicators based e.g. on soil water holding capacity, which had to be
-derived from a rough soil quality measure. This may at least partially
-explain the unsatisfactory performance of many models along the drought
-axis (i.e., climatic water balance).
+The EuFoRIa data as used here are unique and well suited to better understand tree regeneration.
+However, three aspects of these data may represent considerable limitations.
+First, we made a comparison of tree regeneration in an equilibrium state, but we cannot assess how close the forests included in the EuFoRIa dataset were to such a state.
+The data were collected in forest reserves where no management has taken place for long periods of time.
+This makes our assumption of an equilibrium between forest properties and environmental drivers more reasonable than it might appear at first sight.
+In an analysis of primeval forests at demographic equilibrium, @brzeziecki2021 found higher regeneration rates than the ones observed in the EuFoRia data.
+Thus, the overestimation of regeneration rates by the models may not be so problematic.
+Second, the data were collected for regeneration above a 7 cm threshold, thus limiting the assessment of tree regeneration to a holistic viewpoint.
+This constituted a hard limit based on which we can understand only some aspects of tree regeneration, which in its entirety often comprises a rather long period since seed production [@price2001].
+In reality, many environmental constraints are acting on young trees [@käber2021] that we were unable to assess.
+Yet, regeneration data with lower thresholds are simply not available in a harmonized manner across large environmental gradients.
+Third, the empirical data were collected from rather small plots, while we sampled simulated regeneration levels from 1 ha areas, which may lead to an incorrect representation of space.
+Even though the strategy we adopted is not ideal, it represents a common challenge when harmonizing diverse data sources originating from varied sampling strategies [@portier2022].
+It would have been extremely challenging for such a variety of models to follow a protocol where the spatial sampling size was different at each of the 200 sites, and it would have introduced additional uncertainty in the results.
+Ultimately, we need to acknowledge that the observed data represent just a snapshot of an extremely complex and stochastic reality.
+This may at least partly explain e.g.
+the lack of clear patterns in the simulation results along the gradients of climate water balance and degree-days.
+
+The design of our sampling protocol did not include spatial aspects such as seed dispersal or detailed soil data.
+While we considered a wide range of models of forest dynamics, from stand to global scales, the simulation setup was limited to 1 ha regeneration values.
+This lack of consideration of spatial scale is appropriate for stand-scale models, but it potentially puts the landscape models at a disadvantage, as they have been built to be accurate at the landscape level.
+Without spatial context, we are limiting regeneration to the seed influx from the stand itself, unless the model has a background seed input.
+Yet, the global models should not be at a disadvantage due to the limited spatial considerations [but cf.
+Snell et al.
+-@snell2018; or Lehsten et al.
+-@lehsten2019], as they are usually lacking dispersal between cells and are based on a strong abstraction of horizontal space [@hanbury-brown2022].
+Lastly, detailed data on soil conditions were not available from the observed data, and independent, admittedly rather coarse data for soil and climatic water balance had to be used instead.
+It is noteworthy that many models represent drought based on detailed indicators based e.g.
+on soil water holding capacity, which had to be derived from a rough soil quality measure.
+This may at least partially explain the unsatisfactory performance of many models along the drought axis (i.e., climatic water balance).
 
@@
 
-With our study, we have demonstrated that models of forest dynamics need
-a focus on tree regeneration modules to make them more robust. It
-remains uncertain what level of detail is required to model regeneration
-processes, and this must be addressed in future research. We recommend
-that the improvement of the regeneration modules is implemented as
-additional features that can be traced back, as done here for the
-variants of ForClim and ForCEEPS, and that model complexity and
-structure must always be connected with modeling objectives
-[@albrich2020]. If it should be necessary to include more detail in the
-regeneration models, this will come with higher parameterisation
-efforts. This will most likely lead to lower generalization because the
-required data will have to be collected from specific locations, as
-currently there is no general, comprehensive regeneration dataset
-available.
-
-Therefore, we further recommend that more effort should be invested into
-collecting harmonized datasets on tree regeneration. We emphasize that
-datasets such as EuFoRIa are invaluable and should be expanded in both
-their spatial extent (e.g., towards boreal and Mediterranean conditions)
-as well as in time (e.g., continuing the monitoring into the future).
-Such data will allow for a better evaluation of forest models and help
-to reduce the uncertainty in their projections, which is crucial when
-they are used as tools for predicting e.g. the impacts of anthropogenic
-climate change.
-
-In the present study, we have considered tree regeneration in the
-equilibrium state only. It is equally important to understand how these
-models project tree regeneration after changes in forest structure by
-disturbances [@seidl2022], or under different management strategies
-[@lindner2000]. However, this will require an entirely different set of
-observed data, and potentially not all models of forest dynamics would
-be able to assess the relationship of these aspects on tree
-regeneration, e.g. due to the lack of disturbance or appropriate
-management modules.
-
-We recommend to investigate in detail the implications of the current
-modeling strategies for tree regeneration and, ultimately, simulated
-forest stand structure. This applies particularly to the erroneous
-patterns of excess tree regeneration and later excess mortality, by
-focusing on a wider range of tree sizes and the related regeneration
-dynamics. We also recommend, especially for the landscape-level models,
-the inclusion of explicit spatial considerations regarding tree
-regeneration [@beckage2003]; this, however, is a serious challenge
-regarding appropriate datasets, as inventory data have wide coverage,
-but by definition do not allow for the assessment of spatial interaction
-effects.
-
-Exercises like the one presented here, where the models are operated in
-"blind flight" mode, i.e., without the possibility of being tuned
-towards capturing the expected patterns, should be repeated. Such
-benchmarking exercises should next focus on aspects such as specific
-model traits and the ecological formulations of particular
-(sub-)processes, to better understand the implications of the
-assumptions on which the models are based. Furthermore, the inclusion of
-a wide range of models with different scales, type, population
-structures, tree regeneration modules and complexity will ensure a wider
-benefit to the whole modeling community.
+With our study, we have demonstrated that models of forest dynamics need a focus on tree regeneration modules to make them more robust.
+It remains uncertain what level of detail is required to model regeneration processes, and this must be addressed in future research.
+We recommend that the improvement of the regeneration modules is implemented as additional features that can be traced back, as done here for the variants of ForClim and ForCEEPS, and that model complexity and structure must always be connected with modeling objectives [@albrich2020].
+If it should be necessary to include more detail in the regeneration models, this will come with higher parameterisation efforts.
+This will most likely lead to lower generalization because the required data will have to be collected from specific locations, as currently there is no general, comprehensive regeneration dataset available.
+
+Therefore, we further recommend that more effort should be invested into collecting harmonized datasets on tree regeneration.
+We emphasize that datasets such as EuFoRIa are invaluable and should be expanded in both their spatial extent (e.g., towards boreal and Mediterranean conditions) as well as in time (e.g., continuing the monitoring into the future).
+Such data will allow for a better evaluation of forest models and help to reduce the uncertainty in their projections, which is crucial when they are used as tools for predicting e.g.
+the impacts of anthropogenic climate change.
+
+In the present study, we have considered tree regeneration in the equilibrium state only.
+It is equally important to understand how these models project tree regeneration after changes in forest structure by disturbances [@seidl2022], or under different management strategies [@lindner2000].
+However, this will require an entirely different set of observed data, and potentially not all models of forest dynamics would be able to assess the relationship of these aspects on tree regeneration, e.g.
+due to the lack of disturbance or appropriate management modules.
+
+We recommend to investigate in detail the implications of the current modeling strategies for tree regeneration and, ultimately, simulated forest stand structure.
+This applies particularly to the erroneous patterns of excess tree regeneration and later excess mortality, by focusing on a wider range of tree sizes and the related regeneration dynamics.
+We also recommend, especially for the landscape-level models, the inclusion of explicit spatial considerations regarding tree regeneration [@beckage2003]; this, however, is a serious challenge regarding appropriate datasets, as inventory data have wide coverage, but by definition do not allow for the assessment of spatial interaction effects.
+
+Exercises like the one presented here, where the models are operated in "blind flight" mode, i.e., without the possibility of being tuned towards capturing the expected patterns, should be repeated.
+Such benchmarking exercises should next focus on aspects such as specific model traits and the ecological formulations of particular (sub-)processes, to better understand the implications of the assumptions on which the models are based.
+Furthermore, the inclusion of a wide range of models with different scales, type, population structures, tree regeneration modules and complexity will ensure a wider benefit to the whole modeling community.
 
@@
 
-Models of forest dynamics are important tools in science and decision
-support, and the formulation of tree regeneration has strong
-implications for simulated forest properties. The 15 models and variants
-used here are facing similar challenges in their representation of tree
-regeneration: they generally overestimate tree regeneration levels, and
-the simulated regeneration niche is not always captured accurately as a
-function of biotic (light) and abiotic (temperature and moisture)
-factors.
-
-However, most models properly capture the diversity of the initial tree
-community, and differences between model formulations, e.g., the
-presence or absence of feedback from the adult trees, did not have a
-strong effect for capturing the species composition of regeneration.
-
-Regarding mortality in the early phase of tree life, many models that
-feature a particularly high overestimation of regeneration levels are
-compensating for this by a larger tree mortality. Often, this
-compensation is not sufficient to reduce the high regeneration levels to
-realistic values. Overall, there is no clear mortality pattern across
-all models.
-
-When capturing tree regeneration, the specific design decisions taken in
-the development of any model are more important for its behavior
-(accuracy) than scale (stand, landscape, global), modelling approach
-(empirical vs. process-based), and complexity. Having both empirical and
-process-based models in our set, the empirically-based models could have
-been expected to have a better performance, as they were calibrated with
-inventory data, but this was not the case. Similarly, higher model
-complexity does not represent an improvement for capturing tree
-regeneration.
-
-Even though the regeneration routines of most of the models investigated
-here have never been sufficiently constrained by robust data, the
-regeneration projections are not overly off. This indicates that a lot
-can be gained by a focus on the modeling of regeneration processes. The
-representation of forest dynamics in these models would become much more
-robust particularly in the face of climate change and post-disturbance
-dynamics, thus strongly reducing the uncertainty in long-term
-projections of future forest dynamics.
+Models of forest dynamics are important tools in science and decision support, and the formulation of tree regeneration has strong implications for simulated forest properties.
+The 15 models and variants used here are facing similar challenges in their representation of tree regeneration: they generally overestimate tree regeneration levels, and the simulated regeneration niche is not always captured accurately as a function of biotic (light) and abiotic (temperature and moisture) factors.
+
+However, most models properly capture the diversity of the initial tree community, and differences between model formulations, e.g., the presence or absence of feedback from the adult trees, did not have a strong effect for capturing the species composition of regeneration.
+
+Regarding mortality in the early phase of tree life, many models that feature a particularly high overestimation of regeneration levels are compensating for this by a larger tree mortality.
+Often, this compensation is not sufficient to reduce the high regeneration levels to realistic values.
+Overall, there is no clear mortality pattern across all models.
+
+When capturing tree regeneration, the specific design decisions taken in the development of any model are more important for its behavior (accuracy) than scale (stand, landscape, global), modelling approach (empirical vs.
+process-based), and complexity.
+Having both empirical and process-based models in our set, the empirically-based models could have been expected to have a better performance, as they were calibrated with inventory data, but this was not the case.
+Similarly, higher model complexity does not represent an improvement for capturing tree regeneration.
+
+Even though the regeneration routines of most of the models investigated here have never been sufficiently constrained by robust data, the regeneration projections are not overly off.
+This indicates that a lot can be gained by a focus on the modeling of regeneration processes.
+The representation of forest dynamics in these models would become much more robust particularly in the face of climate change and post-disturbance dynamics, thus strongly reducing the uncertainty in long-term projections of future forest dynamics.
 
@@
 
-O. Díaz-Yáñez and Y. Käber contributed equally and share the first
-authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study
-design. All authors contributed to preparing the simulations with their
-respective models. O. Díaz-Yáñez led the data analysis with
-contributions during the workshop from Y.Käber, T. Anders, K. H.
-Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T. Hickler, H. Lischke,
-M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X. Morin, W. Rammer and
-D. Scherrer H. Bugmann. O. Díaz-Yáñez and H. Bugmann led the writing of
-the manuscript. All authors participated in the revision of the
-manuscript and approved its submission.
+O.
+Díaz-Yáñez and Y.
+Käber contributed equally and share the first authorship.
+H.
+Bugmann, Y.
+Käber, and O.
+Díaz-Yáñez developed the study design.
+All authors contributed to preparing the simulations with their respective models.
+O.
+Díaz-Yáñez led the data analysis with contributions during the workshop from Y.Käber, T.
+Anders, K.
+H.
+Braziunas, J.
+Brůna, S.
+M.
+Fischer, J.
+Hetzer, T.
+Hickler, H.
+Lischke, M.Mahnken, P.
+Mairota, K.
+Merganičová, T.
+Mette, X.
+Morin, W.
+Rammer and D.
+Scherrer H.
+Bugmann.
+O.
+Díaz-Yáñez and H.
+Bugmann led the writing of the manuscript.
+All authors participated in the revision of the manuscript and approved its submission.
 
@@
       format = "latex", booktabs = TRUE,
-      caption = "Models included and their characteristics.") |>
+      caption = "") |>
         kableExtra::landscape() |>
@@
       kableExtra::column_spec(column = 5, width = "5em") |>
-      kableExtra::column_spec(column = 6, width = "5em") |>
-      kableExtra::column_spec(column = 7, width = "8em") |>
+      kableExtra::column_spec(column = 6, width = "6em") |>
+      kableExtra::column_spec(column = 7, width = "5em") |>
       kableExtra::column_spec(column = 8, width = "8em") |>
-      kableExtra::column_spec(column = 9, width = "5em") |>
+      kableExtra::column_spec(column = 9, width = "8em") |>
       kableExtra::column_spec(column = 10, width = "5em") |>
-      kableExtra::column_spec(column = 11, width = "10em") |>
+      kableExtra::column_spec(column = 11, width = "5em") |>
+      kableExtra::column_spec(column = 12, width = "10em") |>
       kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
@@
                                  "Y-Yes, N-No",
+                                 "R-Regeneration, Rc-Recruitment",
                                  "D-Daily, M-Monthly"),
@@
 
-# Figure captions
-
-**Figure 1.** The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used
-in the present study, the colors show the dominant tree species per site
-in the observed regeneration at the 10 cm threshold.
-
-**Figure 2.** Mean regeneration levels across all samples per site,
-plotted for the 200 sites and for each model. The red dashed lines shows
-the 25th and 75th percentiles for the 7 cm diameter threshold in the
-observed data. There are two boxplots for each model where the left
-lighter boxplot corresponds to 7 cm and the right darker boxplot to 10
-cm
-
-**Figure 3.** Shannon index of tree regeneration, calculated by basal
-area, as the mean value across the 200 samples per site for the diameter
-thresholds of 7 and 10 cm, respectively. The red dashed lines shows the
-25th and 75th percentiles for the 7 cm diameter threshold in the
-observed data.
-
-**Figure 4.** Mean Shannon index across all samples per site for
-observed and simulated data. Each plot shows one pattern represented by
-one exemplary model of each category (overperdicted, intermediate and
-underpredicted). The full data with the grouping of the models are shown
-in Figures S3 and S4. n indicates the number of models falling in each
-group. A) Examples for the three trends across models for species
-diversity at the stand level. B) Examples for the three trends across
-models for regeneration at the 7 cm threshold.
-
-**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm
-thresholds. Dashed blue lines mark a ratio equal to 1, indicating no
-decrease in tree regeneration between 7 and 10 cm, and a ratio equal to
-1.77, corresponding to the Reineke self-thinning ratio under even-aged
-conditions.
-
-**Figure 6.** Ratio of tree regeneration between the 7 and 10 cm
-thresholds (regeneration 7 cm / regeneration 10 cm) and the
-overestimation proportion at 7 cm ([simulated - observed] / observed)
-for the mean regeneration per model across sites and samples. The
-horizontal dashed lines show a ratio equal to 1, indicating no decrease
-of regeneration between 7 and 10 cm, and a ratio equal to 1.77
-corresponding to the Reineke self-thinning line under even-aged
-conditions. The vertical line indicates zero overestimation.
-
-**Figure 7.** Mean regeneration values across the 200 samples per site
-with the y-axis scaled differently by model, for the 200 sites against
-gradients of (a) total basal area; (b) climatic water balance; (c)
-seasonal degree-day sum. The values were split into ten bins; the red
-lines represent a GAM model showing the trend in the observed data. For
-comparing the magnitude of recruitment across models the same graphs
-with equally scaled y-axis can be found in Figure S7.
-
-**Figure 8.** Regeneration levels of the five main species in terms of
-their basal area share in the observations (top row) and the performance
-of each model across the environmental gradients (other rows) for the 7
-cm regeneration threshold. The values shown are the mean of the 200
-samples per site and across the sites in each bin (tile), with ten bins
-per gradient. The sizes of the circles represent the ratio between the
-regeneration basal area of the species and the total regeneration basal
-area for all species. The absence of a circle indicates a zero basal
-area share in the regeneration, or the absence of regeneration
-altogether. The color gradient (for the models only) shows the
-difference between the simulated and observed ratio of regeneration
-basal area of the species and the total regeneration basal area for all
-the species.
+# Table and Figure captions
+
+**Table 1.** Models included, their characteristics and simulation strategies used in this study.
+Where Scale refers to the models scale, Type to the models type, Population structure to the model population approach, Feedback to the inclusion or not of adult tree feedback to the level and species composition of regeneration, Approach to the model regeneration module approach, Start from to the stage from which trees are recruited, Complexity, to the mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), Species to which species from the original eleven species were simulated , Runtime (spin up) [years] to the number of years used per model in their simulations for the spin up period and regeneration sampling; Climate data type (use of years) to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details.
+Further information can be found on the individual model reports available in the supplementary materials.
+
+**Figure 1.** The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed regeneration at the 10 cm threshold.
+
+**Figure 2.** Mean regeneration levels across all samples per site, plotted for the 200 sites and for each model.
+The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
+There are two boxplots for each model where the left lighter boxplot corresponds to 7 cm and the right darker boxplot to 10 cm
+
+**Figure 3.** Shannon index of tree regeneration, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
+The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
+
+**Figure 4.** Mean Shannon index across all samples per site for observed and simulated data.
+Each plot shows one pattern represented by one exemplary model of each category (overpredicted, intermediate and underpredicted).
+The full data with the grouping of the models are shown in Figures S3 and S4.
+n indicates the number of models falling in each group.
+A) Examples for the three trends across models for species diversity at the stand level.
+B) Examples for the three trends across models for regeneration at the 7 cm threshold (165 sites).
+
+**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm thresholds.
+Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree regeneration between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.
+
+**Figure 6.** Ratio of tree regeneration between the 7 and 10 cm thresholds (regeneration 7 cm / regeneration 10 cm) and the overestimation proportion at 7 cm ([simulated - observed] / observed) for the mean regeneration per model across sites and samples.
+The horizontal dashed lines show a ratio equal to 1, indicating no decrease of regeneration between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
+The vertical line indicates zero overestimation.
+
+**Figure 7.** Mean regeneration values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum.
+The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
+Y-axis are not equally scaled.
+
+**Figure 8.** Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold.
+The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient.
+The sizes of the circles represent the ratio between the regeneration basal area of the species and the total regeneration basal area for all species.
+The absence of a circle indicates a zero basal area share in the regeneration, or the absence of regeneration altogether.
+The color gradient (for the models only) shows the difference between the simulated and observed ratio of regeneration basal area of the species and the total regeneration basal area for all the species.
 

### 2023-09-24 manuscript/review/authorsResponse.qmd
changes in both reviewers.

--- 2023-09-20 manuscript/review/authorsResponse.qmd
@@
 ---
-title: Authors Response to Reviewer Comments
+title: Authors Response to Reviewer's Comments
 subtitle: "ECOSPHERE Manuscript ID (ECS23-0342) Tree regeneration in models of forest dynamics: a key priority for further research"
@@
 number-sections: true
-format:
-   pdf: 
-     toc: true
+format: docx
 ---
@@
 -   underlined comments are comments that still need attention
-
--
 
@@
 
-Dear Charles,
-
-Thank you very much for taking the time to review our manuscript and for your insightful thoughts on our findings.
-
-We totally agree with your comments.
-Our main take away from this manuscript it is not that it would be impossible to calibrate all 15 models against e.g.
-the EuFoRIa data, but the issue is - as you wrote - that this needs to be bound to the goals of the study.
-And when the goal is to provide extrapolation into a future climate, then calibration against data under current climate may be unreliable.
-This is why we belive it is fundamentally valuable to test model performance "out of the box", as we have done here.
-
-In this revised version we have modify the manuscript according to the reviewers' recommendations and we have addressed each of the specific points made in the review comments.
-We believe the manuscript has significantly improved based on this changes.
+Dear Dr. Charles Canham,
+
+I want to express my gratitude for taking the time to review our manuscript. Your insightful thoughts on our findings were very helpful.
+
+Our main takeaway from this manuscript is that while it may be possible to calibrate all 15 models against the EuFoRIa data, it is important to ensure that this aligns with the study's goals.
+If the goal is to provide extrapolation into a future climate, calibrating against data under the current climate may not be reliable.
+This is why we believe it is valuable to test model performance without any changes or calibration as we have done in this study.
+
+We have carefully reviewed and addressed each of the specific points made in the review comments in this revised version.
+We believe that the changes we made have significantly improved the manuscript.
+Once again, I want to thank you for the time and effort you dedicated to reviewing our work.
+We look forward to hearing from you again.
+
+Best regards,
+
+Olalla Díaz-Yáñez on behalf of all the co-authors
 
@@
 
+Dear Reviewer 1,
+
+Thank you for taking the time to review our manuscript.
+We appreciate your thoughtful comments and have carefully considered each of them.
+We have made changes to the manuscript based on your feedback and believe that these changes have improved the overall quality of the work.
+
+We understand your concern about the length of the manuscript and the number of figures included.
+We have made an effort to reduce the text where possible and have verified that we are following the journal guidelines.
+Despite our best efforts, however, it has been difficult to significantly reduce the length of the manuscript.
+
+Please find below our responses to each of your comments and the concrete actions we have taken in response.
+
+Thank you again for your valuable feedback.
+
 ### Major comments
@@
 
-[I have to think about this again]{.underline}
-
-Thank you for pointing
-
-#### [COMMENT 2]{.underline}
+[We should discuss this point again... I do not agree to simple change regeneration with recruitment. Specially without a detailed definition as we had in the previous version.]{.underline}
+
+[SUGGESTION FROM YANNEK:]{.underline}
+
+[Thank you for pointing this out.
+We also discussed terminology several times and we agree.
+We followed your suggestion replaced 'regeneration' with 'recruitment' where the difference is not clear.
+However, we believe that in some places regeneration is more accurate.
+Especially in many parts of the introduction and discussion.
+We argue that regeneration is often only modeled via "recruitment", but with the aim to simulate the whole process of regeneration.
+In this context it makes more sense to use the general term (e.g., for discussion consequences.)]{.underline}
+
+[COMMENT from YK: I would argue that recruitment is evidence for successful regeneration.
+This allows us to use recruitment for discussions observation vs.
+simulation in both aspects: technical (recruitment) and general (regeneration).]{.underline}
+
+#### COMMENT 2
 
@@
 I assume only the 165 stands were used.
-Please clarify. \
+Please clarify.
 
@@
 
-Thank you for highlighting that we didn not managed to appropriately communicate our diameter thresholds approach correctly in the manuscript. The explanation is that the observed data
-
-In this revised version we have revised the explanation in L XX
+Thank you for bringing to our attention the issue with the way we communicated our diameter thresholds approach in the manuscript.
+The observed data had varying diameter thresholds for reporting tree recruitment, and, for the purposes of this study, we merged the recruited trees values into two thresholds of 7 and 10cm.
+
+In total, we had 165 sites where the records showed the number of recruited trees at a 7cm threshold.
+For these sites, any recruited tree above 7cm was counted.
+We also had 35 sites where the minimum diameter threshold was 10cm, and for these sites, we could only account for trees recruited from 10cm diameter.
+We used the other 165 sites with 7cm diameter threshold in the 10cm database, but only accounting for newly established trees above 10cm.
+There was no like-to-like comparison for any other diameter threshold, only for 7 and 10cm.
+
+The reviewer correctly pointed out that in Figure 2 for the 7cm threshold, there were 165 "dots" in the observed data, while for the 10cm diameter threshold, there were 200 "dots." However, simulated values were obtained for the 200 sites for both thresholds, as explained in the Simulation protocol section.
+Therefore, the simulated values in the figures are from the 200 sites.
+
+When examining the change in mortality between 7 to 10cm or species diversity comparison between observed and simulated data (Figure 4), we could only use the 165 sites that contained both pieces of information at the 7 and 10cm thresholds in the observed data.
+This is reported in the supplementary materials software, script \`\`\`Figures.R\`\`\` line 536 and L284-349 and now better explained in the manuscript.
+
+To better clarify this aspect, we have made the following changes:
+
+1\) We have better explained the difference on the observed data in the section 3.2 Observed data with the following text:
+
+"For our study, we defined two datasets where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds from 10 cm.
+Therefore we had in total 200 sites with observations of newly established trees above 10 cm and 165 sites above 7cm.
+In the results section any result from observed data at 7 cm threshold is originated from the 165 sites."
+
+2\) We have added the following text in the caption of figure 4
+
+"B) Examples for the three trends across models for regeneration at the 7 cm threshold (165 sites)."
 
@@
 
-[Based on the comments from Reviewer 1, we have re-analyse some of the results. More specifically we]{.underline}
+[I am going to do this.]{.underline}
 
@@
 
-#### [COMMENT 4]{.underline}
+#### COMMENT 4
 
@@
 
-[check]{.underline}
-
-#### [COMMENT 5]{.underline}
+We have made some changes to the abstract based on feedback from both reviewers, including clarifying certain parts for better understanding.
+The models were able to capture some aspects of regeneration more effectively than others.
+One example, is the sentence highlighted by the reviewer that was refereing to the diversity of regeneration.
+
+#### COMMENT 5
 
@@
 
-[check]{.underline}
-
-#### [COMMENT 6]{.underline}
+Following the reviewers comments we have made several changes in this paragraph to improve its readability and flow if ideas with the main purpose of the study text.
+
+#### COMMENT 6
 
@@
 
-[change this]{.underline}
-
-#### [COMMENT 7]{.underline}
+Following the reviewers comment we have added a new column classifying the models in recruit or regeneration models; and the letters a,b,c, d,e and f table are now better described.
+Regarding the suggestion on adding into this table the size of the saplings or ingrowth: there are models that start from saplings or ingrowth what without a constant diameter threshold as they are biomass based (e.g.
+LandClim), we find it complicated to add all this information in this table without making it too large and therefore we have decided not to included.
+Furthermore the models reports in the supplementary materials provide more details on how the models simulate regeneration.
+
+ToDo All models check that the new column in the table is correct for your model.
+
+#### COMMENT 7
 
@@
 
-[check]{.underline}
-
-#### [COMMENT 8]{.underline}
+Yes, this is the mean value across all observations and sites, there are two main reasons why this number might seem low, first due to the zero inflated nature of the data, and second because these are forest reserves, corresponding with denser forests in which regeneration would be expected to be lower.
+This value is in agreement with a further observed data from reserved (Käber, Y., et al (2023).
+Sheltered or suppressed?
+Tree regeneration in unmanaged European forests.
+Journal of Ecology, 00, 1--15.
+https://doi.org/10.1111/1365-2745.14181)
+
+#### COMMENT 8
 
@@
 
-#### [COMMENT 9]{.underline}
+For those models that simulated more than the eleven species, we created a different category "others" were we lumped those extra recruited species.
+This were counted when we calculated the diversity index and for total recruitment values.
+Following the reviewer comments we have added this information in the simulation protocol section with the following text:
+
+"Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species simulated beyond the eleven species were lumped in a category "others""
+
+#### COMMENT 9
 
@@
 
+Following the reviewers comment we have modified this text to improve it readability with:
+
+"Furthermore, regarding the share in the regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero."
+
 #### COMMENT 10
@@
 
-#### COMMENT 11
+Even thought it is not directly comparable, we still believe is interesting for the reader that we share the findings related to the 4C model and those that simulate more species that were lumped in the category others.
+
+#### [COMMENT 11]{.underline}
 
@@
 
+[To be addressed in combination with COMMENT 3]{.underline}
+
 #### COMMENT 12
@@
 ##### Response to comment 12:
+
+Following the reviewers comment we have increased the legend size, and now the difference between observed and simulated is visible.
+We have also explained in the caption the meaning of the black line:
+
+"Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 7cm.
+The black line marks where the mean Shannon index per site at recruitment and at the stand level would be equal"
 
@@
 What happens if you colour the points, by those from models with the feedback?
-Do you address in the discussion why this feedback makes no difference to recruitment? \
+Do you address in the discussion why this feedback makes no difference to recruitment?
 
@@
 
+We agree with the reviewer that it was interesting to find that the models with feedback did not captured better the species diversity of the regeneration compared with those without, with the exeption of ForCEEPS.
+The Figure 4 shows three representative patterns observed across models using 6 of them as an example, and the Figure S3 provides the details for each model.
+Models with feedback are defined in table 1 and are ForCEEPS(f) , PICUS, SIBYLA, xComp, iLand, LandClim, Landis-II, TreeMig, LPJ-GUESS, and aDGVM2.
+We have addressed this point in the discussion section with the following text:
+
+"Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the regeneration than those without feedback, with the exception of ForCEEPS that significantly improved regeneration species diversity with the model version that included feedback (ForCEEPS(f)).
+The similar performance of models with and without feedback is likely because the models put more weight on the regeneration niche arising from abiotic and biotic filters, than from the habitat niche of the adult trees [@grubb1977]."
+
 #### COMMENT 14
@@
 
+Following the reviewers recommendation we have changed the axis labels in the appendix figures.
+
 #### COMMENT 15
@@
 
+This paragraph is addressing the comparison with the theoretical Reineke value which is usually calculated for even-aged, single species stands and assumes a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands.
+
 #### COMMENT 16
@@
 
-#### COMMENT 17
+Thank you! We also think is a good summary figure.
+
+#### [COMMENT 17]{.underline}
 
@@
 
+[Clean together with comment 2]{.underline}
+
 #### COMMENT 18
@@
 
+Following the reviewer comment we have removed figure S7.
+
 #### COMMENT 19
@@
 
+Figure 4 shows three examples of the patterns observed across models (overestimation intermediate and underestimation) using one exemplary model of each category , the actual patterns per model can be found in Figure S3.
+In the results we showed that most models captured reasonably well or overpredicted species diversity at the stand level, as most of the models are under the patterns presented in plot 1 and 2 from A on Figure 4.
+We have decided to keep the referenced statement link to this findings in the discussion as it was.
+
 #### COMMENT 20
@@
 
+In this statement we were referring only to regeneration levels.
+Although we agree with the reviewer that we did state that most of the models overestimate regeneration, here our point is that even-thought they did overestimate regeneration they did not deviate exceedingly, specially under the consideration that these models have never been confronted with this type of data.
+
 #### COMMENT 21
@@
 
+With this statement we were referring to the multiple potential combinations of approaches of modeling tree regeneration. Based on the reviewers comment we have modified this for clarity.
+
 #### COMMENT 22
@@
 ##### Response to comment 22:
+
+Thank you for noticing this. We have now corrected this mistake.
 
@@
 
+We agree with the reviewer that ideally models should be able to capture broader patterns.
+With this sentence we just wan to explain that the stochasticity of this process may have partly explained the results on this regard.
+
 #### COMMENT 24
@@
 ##### Response to comment 24:
+
+The reviewer is correct that landscape level models were not significantly performing worst than others.
+Here we just wanted to discuss that maybe another spatial consideration will have allowed landscape level models to better represent regeneration processes as they are designed for that.
 
@@
 
-#### COMMENT 26
+We agree with the reviewer, a coarse representation of water balance would not make it more sensitive to water deficits.
+And this was exactly our point in this part of the discussion where we wrote: "This may at least partially explain the unsatisfactory performance of many models along the drought axis (i.e., climatic water balance)."
+
+#### [COMMENT 26]{.underline}
 
@@
 
+Following the reviewers comment we have expanded this aspect in the following paragraph XXXX
+
+[Both reviewers asked to be more specific on the data we are missing, what vairables would you say without being too detailed? also linked to comment 59 from reviewer 2]{.underline}
+
 ## Response to reviewer 2
@@
 
+Dear Reviewer 2,
+
+We would like to express our sincere gratitude for all the detailed comments they provided on our manuscript.
+We have carefully considered all of their feedback and made the necessary changes to improve the manuscript.
+Thanks to your comments, we have improved our manuscript.
+
+Thank you again for your valuable feedback.
+
 ### Major comments
 
-#### COMMENT 1
+#### [COMMENT 1]{.underline}
 
@@
 
-#### COMMENT 2 {#sec-R2C2}
+[I think we have proposed some ideas of how to do this better in the discussion. What else do you think we should add here?]{.underline}
+
+#### [COMMENT 2]{.underline} {#sec-R2C2}
 
@@
 
-#### COMMENT 3
+Following the reviewer comment we have added a "road map" at the end of the introduction.
+
+#### [COMMENT 3]{.underline}
 
@@
 
+[Decide final terminology]{.underline}
+
 #### COMMENT 4
@@
 ##### Response to comment 4:
+
+The reviewer is correct that most of the people involved in this manuscript have a central European focus in their current work, and so did the study set up.
+We have included some relevant global or with a focus outside Europe references when relevant.
+e would be happy to consider some other relevant suggestions.
 
@@
 
-[Following this and other comments from the editor, we have tried to capture this better in the discussion/conclussions section.]{.underline}
-
-#### [COMMENT 15]{.underline}
+[Following this and other comments from the editor, we have tried to capture this better in the discussion/conclussions section. Should we add something on this?]{.underline}
+
+#### COMMENT 15
 
@@
 
-[ToDo find a reference!]{.underline}
+[ToDo find a reference! Any suggestions?]{.underline}
 
@@
 
-[ToDo find a reference!]{.underline}
-
-#### [**COMMENT 18**]{.underline}
+[ToDo find a reference!Any suggestions?]{.underline}
+
+#### **COMMENT 18**
 
@@
 
-We currently use the exciting Hanbury-brown 2022 review in several places across our manuscripts, following the reviewer's comment here and the general comment 2 in @sec-R2C2 we have also get inspired by their structure [by creating a road map to help the reader to navigate the manuscript content.]{.underline}
-
-#### [**COMMENT 19**]{.underline}
+We currently use the exciting Hanbury-brown 2022 review in several places across our manuscripts, following the reviewer's comment here and the general comment 2 in @sec-R2C2 we have also get inspired by their structure by creating a road map to help the reader to navigate the manuscript content.
+
+#### **COMMENT 19**
 
@@
 
-We believe the reviewer highlighted the words "models, such as deer browsing" because they think this is not correct.
-From our sample pool [only 4 models]{.underline} incorporated explicitly deer browsing.
+We believe the reviewer highlighted the words "models, such as deer browsing" because they think this is not correct. From our sample pool only 4 models incorporated explicitly deer browsing.
 
@@
 
-[**COMMENT 20**]{.underline}
+#### **COMMENT 20**
 
@@
 
-[We are unsure why the reviewer highlighted those words]{.underline}
+We are unsure why the reviewer highlighted those words
 
@@
 
-#### **COMMENT 25**
-
-L200-202 "The observations used here are derived from a novel and unprecedented network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network (EuFoRIa 2019)" I don't feel "novel and unprecedented" adds anything to this sentence.  There are plenty of networks outside Europe providing this sort of data
-
-##### Response to comment 25: {#sec-R2C25r}
-
-We agree with the reviewer's comment that there are other forest inventory data networks outside Central Europe even on forest reserves that could provide this sort of data.
-However, the EuFoRIa network is indeed novel and unprecedented network of forest reserves in Europe.
-
 #### **COMMENT 26**
@@
 
-Response to comment 33:
+##### Response to comment 33:
 
@@
 #####
-Response to comment 34:
+
+##### Response to comment 34:
 
@@
 
-##### 
-Response to comment 35:
+##### Response to comment 35:
 
@@
 
-L229 "Neither were further site information (except for the data specified below) nor any data " These were "blind trials":  modelling group were not provided with site information \...
-
-##### 
-Response to comment 36:
+L229 "Neither were further site information (except for the data specified below) nor any data" These were "blind trials":  modelling group were not provided with site information \...
+
+##### Response to comment 36:
 
@@
 
-L230-231 "That is, the models were run in \"blind flight\" mode." Remove
-
-##### 
-Response to comment 37:
+L230-231 "That is, the models were run in "blind flight" mode." Remove
+
+##### Response to comment 37:
 
@@
 
-##### 
-Response to comment 38:
+##### Response to comment 38:
 
@@
 
-##### 
-Response to comment 39:
-
-We understand the reviewers concern on not constraining the length of the simulations, but we decided to leave it to the assessment of the individual modelers, as we wanted to obtain simulation that were run from from bare ground in the absence of management to a simulated equilibrium (\"potential natural vegetation\") with current climate, and each modeler know best how to obtain this in each model.
+##### Response to comment 39:
+
+We understand the reviewers concern on not constraining the length of the simulations, but we decided to leave it to the assessment of the individual modelers, as we wanted to obtain simulation that were run from from bare ground in the absence of management to a simulated equilibrium ("potential natural vegetation") with current climate, and each modeler know best how to obtain this in each model.
 This information is currently available in the manuscript and in the protocol of this study.
@@
 
-##### 
-Response to comment 40:
+##### Response to comment 40:
 
@@
 
-##### 
-Response to comment 41:
+##### Response to comment 41:
 
@@
 
-##### 
-Response to comment 42:
+##### Response to comment 42:
 
@@
 
-##### 
-Response to comment 43:
+##### Response to comment 43:
 
@@
 
-**COMMENT 44**
+#### **COMMENT 44**
 
@@
 
-##### 
-Response to comment 44:
+##### Response to comment 44:
 
@@
 
-##### 
-Response to comment 45:
+L1046 Such converting to effective no of species using Jost formula
+
+##### Response to comment 45:
 
@@
 
-#### [**COMMENT 46**]{.underline}
+#### **COMMENT 46**
 
@@
 
-##### 
-Response to comment 46:
+##### Response to comment 46:
 
@@
 
-**COMMENT 47**
+#### COMMENT 47
 
@@
 
-##### 
-Response to comment 47:
-
-Follwoing the reviewer recomendation we have changed the equation formulation to better represent the Reineke relationship.
+##### Response to comment 47:
+
+Following the reviewer recomendation we have changed the equation formulation to better represent the Reineke relationship.
 The reviewer is correct that the Reineke is applied to all stems in the stand, in our case the N value is only changing due to the available regeneration at the two diameter thresholds as the adult trees number is the same.
@@
 
-#### COMMENT 48
-
-equation 5 Suggest writing N7 / N10
-
-Given my comment above, is there a strong justification for the -1.605 exponent?  Or just keep it at one?
-
-##### 
-Response to comment 48:
-
-The Stand Density Index by Reineke (1933) is based on the relationship between quadratic mean diameter and number of stems per unit area in fully stocked stands.
-We have use the coefficient -1.605 because Reineke attributed a general validity to this allometric coefficient for fully stocked, even-aged forest stands, regardless of tree species and site.
-This relationship was obtained by using pairs of variates for quadratic mean diameter and number of stems per unit area.
-In our study we are using the value of 1.77 to define this relationship between 7 and 10cm
-
 #### [COMMENT 49]{.underline}
 
-L316-317 "model complexity as defined by Bugmann and Seidl"
-
-##### 
-Response to comment 49:
-
-[Think how can I add something else without making this much longer!]{.underline}
+L316-317 "model complexity as defined by Bugmann and Seidl" say a little more...
+
+##### Response to comment 49:
+
+[Any suggestion of how can I add something else without making this much longer!]{.underline}
 
@@
 
-##### 
-Response to comment 50:
+##### Response to comment 50:
 
@@
 
-##### 
-Response to comment 51:
+##### Response to comment 51:
 
@@
 
-COMMENT 52
+#### COMMENT 52
 
@@
 
-##### 
-Response to comment 52:
+##### Response to comment 52:
 
@@
 L530-533 "This substantial difference may be due to the fact that the simulation results were drawn from equilibrium forests, whereas in reality many of the forest reserves are recovering from past management activities and have become denser over the past decades (e.g., Heiri et al.
-2009), leading to less regeneration than in an equilibrium situation.
-" But also \-\-- browsing, establishment sites,  competition with herb layer all reduce recruitment
-
-##### 
-Response to comment 53:
+2009), leading to less regeneration than in an equilibrium situation." But also \-\-- browsing, establishment sites,  competition with herb layer all reduce recruitment
+
+##### Response to comment 53:
 
@@
 
-##### 
-Response to comment 54:
+##### Response to comment 54:
 
@@
 ####
-COMMENT 55
+
+#### COMMENT 55
 
@@
 
-##### 
-Response to comment 55:
+##### Response to comment 55:
 
@@
 ####
-COMMENT 56
+
+#### COMMENT 56
 
 L596-598 "Competition for light as a strong filter for tree regeneration has been widely documented (Messier et al.
-1999; Collet and Chenost 2006; Berdanier and Clark 2016), but the models examined here did not reproduce this expectation.
-", because \...
+1999; Collet and Chenost 2006; Berdanier and Clark 2016), but the models examined here did not reproduce this expectation.", because \...
 remind us of the evidence you are putting forward
 
-##### 
-Response to comment 56:
+##### Response to comment 56:
 
@@
 
-##### 
-Response to comment 57:
+##### Response to comment 57:
 
@@
 ####
-[COMMENT 58]{.underline}
+
+#### COMMENT 58
+
+L656 "Yet, the global models should not be at a disadvantage due to the limited spatial consideration" Don't quite follow your point here. what does "at a disadvantage" mean?
+
+##### Response to comment 58:
+
+In this sentence we are discussing how the sampling strategy presented in the study protocol might have represented a disadvantage for landscape level models due to the limited spatial scale, but how this might not be the same for Global level models as they usually lack dispersal between cells.
+
+#### [COMMENT 59]{.underline}
 
 L676-677 "Therefore, we further recommend that more effort should be invested into collecting harmonized datasets on tree regeneration." regeneration in what sense?
-  Seedlings/ ground layer/ \....  What is really needed?
-
-##### 
-Response to comment 58:
-
-Following the reviewers comment we have included this text:
+  Seedlings/ ground layer/ ....  What is really needed?
+
+##### Response to comment 58:
+
+[Linked to comment 26 from reviewer 1]{.underline}
+
+#### COMMENT 60
+
+L685-687 "However, this will require an entirely different set of observed data, and potentially not all models of forest dynamics would be able to assess the relationship of these aspects on tree regeneration, e.g.
+due to the lack of disturbance or appropriate management modules." more akin to data already available in national inventories I think
+
+##### Response to comment 60:
+
+We agree with the reviewer that a more systematic kind of inventory could be useful to detect regeneration patterns under management or disturbances impacts.
 
@@
 
-[COMMENT 59]{.underline}
-
-L685-687 However, this will require an entirely different set of observed data, and potentially not all models of forest dynamics would be able to assess the relationship of these aspects on tree regeneration, e.g.
-due to the lack of disturbance or appropriate management modules.
-
-##### 
-Response to comment 59:
-
-Following the reviewers comment we have included this text:
+#### COMMENT 61
+
+L696 "Exercises like the one presented here, where the models are operated in "blind flight" mode." Definition of blind flight "To do something based on guesswork, intuition, or without any help or instructions"....is that what you mean?
+
+##### Response to comment 60:
+
+With blind flight we refer to doing the simulations without having important information about the expected regeneration values.
+
+#### COMMENT 62
+
+Table 1 caption: A useful table.
+I'd encourage the authors to write a lengthier table heading, which briefly summarises the methods section, so that readers can know what "start from", "runtime for sampling" mean without having to delve into the text.
+Also I'd like to see an explanation of "formulation complexity" in the text and here.
+The supplementary information gives more details about model assumptions, and I do wonder whether those could be included here; in particular, seed production and dispersal is clearly important for regeneration, so could columns be introduced to compare assumptions of model with respect to these.
+Also deer browsing, herb-layers and nurse logs all have an influence of regeneration, but not incorporated into these models; I suggest the table heading makes that clear.
+
+##### Response to comment 62:
+
+Following the reviewers comment we have expanded the caption of table 1 with the following text:
+
+"Models included, their characteristics and simulation strategies used in this study.
+Where Scale refers to the models scale, *Type* to the models type, *Population structure* to the model population approach, *Feedback* to the inclusion or not of adult tree feedback to the level and species composition of regeneration, *Approach* to the model regeneration module approach, *Start from* to the stage from which trees are recruited, *Complexity,* to *the* mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), *Species* to which species from the original eleven species were simulated , *Runtime (spin up) \[years\]* to the number of years used per model in their simulations for the spin up period and regeneration sampling; *Climate data type (use of years)* to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details.
+Further information can be found on the individual model reports available in the supplementary materials."
+
+#### COMMENT 63
+
+L831 Add publisher
+
+L939 Correct doi
+
+##### Response to comment 63:
+
+Changed as suggested.
 
 #### 
-
-[COMMENT 60]{.underline}
-
-L696 "Exercises like the one presented here, where the models are operated in \"blind flight\" mode." Definition of blind flight "To do something based on guesswork, intuition, or without any help or instructions"\....is that what you mean?
-
-#####
-
-Response to comment 60:
-
-Following the reviewers comment we have included this text:
+[COMMENT 64]{.underline}
+
+L1041 suggest sticking with "ingrowth" (as used in the figure) or "recruitment"
+
+##### Response to comment 65:
+
+[To follow up after terminology discussion]{.underline}
+
+#### COMMENT 65
+
+L1050 Overperdicted
+
+##### Response to comment 65:
+
+Changed as suggested.
+
+#### 
+COMMENT 66
+
+L1056 no mortality of recruits as they grow from 7 to 10 cm diameter.
+
+##### Response to comment 66:
+
+We use the difference of observed trees recruited between 7 and 10cm as a proxy for mortality.
+
+#### 
+COMMENT 67
+
+L1077 Explain the fitted curves
+
+##### Response to comment 67:
+
+We are not sure if we are correctly understanding the reviewers point here. In figure 8 we do not show any fitted curves.
 

### 2023-09-24 manuscript/supplementaryRW.Rmd
changes in both reviewers.

@@
 
-```{r recrStand7, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 7cm", out.width='100%', fig.pos='H'}
+```{r recrStand7, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 7cm.
+The black line marks where the mean Shannon index per site at recruitment and at the stand level would be equal", out.width='100%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "recruitmentAdultRichness7.png"))
@@
 
-```{r nvTrend710, echo=FALSE, fig.cap="Mean ingrowth values across the 200 samples per site, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree sum.
-The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
-   knitr::include_graphics(c(here::here("figures", "envTrend7Totba_500.png"),
-                              here::here("figures", "envTrend7WB_500.png"),
-                              here::here("figures", "envTrend7DDS_500.png")))
-```
-
 ```{r diff, echo=FALSE, fig.cap="Departure of each model in the representation of the main species recruitment relative abundance from the observations across the environmental gradients for a 10cm treshold.

### 2023-09-24 code/figures.R

@@ -26,6 +26,7 @@
     sum(obsNumAgg$r.ha, na.rm =  T)
     range(obsNumAgg$r.ha, na.rm = TRUE)
     mean(obsNumAgg$r.ha, na.rm = TRUE)
+    mean(obsNumAgg$r.ha[!obsNumAgg$r.ha == 0], na.rm = TRUE)
     median(obsNumAgg$r.ha, na.rm = TRUE)
     
     # Assign models and dbh to levels 
@@ -364,16 +365,17 @@
                          fill = "white") + 
         ggplot2::coord_cartesian(xlim = c(0, 2), ylim = c(0, 2)) +  
         ggplot2::geom_point(size = 0.1, alpha = 0.3 ) + 
-        hrbrthemes::theme_ipsum() +
+        ggplot2::theme_classic() +
         ggplot2::theme(legend.title = ggplot2::element_blank()) +
         ggplot2::labs(y = bquote(bar(H) * " recruitment"),
                       x = bquote(bar(H) * " stand")) +
-        ggplot2::ggtitle("7 cm diamter threshold") +
+        ggplot2::ggtitle("7 cm diameter threshold") +
         ggplot2::facet_wrap(~model) +
         ggplot2::geom_abline() +
         ggplot2::scale_color_manual(labels = c("Observed" = "Observed", 
                                                "model" = "Simulated"),
-                                    values = c("#FDAE61", "grey"))
+                                    values = c("#FDAE61", "grey")) +
+        ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 10)))
     
     
     ggplot2::ggsave("figures/recruitmentAdultRichness7.png",
@@ -391,7 +393,7 @@
                          fill = "white") + 
         ggplot2::coord_cartesian(xlim = c(0, 2), ylim = c(0, 2)) +  
         ggplot2::geom_point(size = 0.1, alpha = 0.3 ) + 
-        hrbrthemes::theme_ipsum() +
+        ggplot2::theme_classic() +
         ggplot2::theme(legend.title = ggplot2::element_blank()) +
         ggplot2::labs(y = bquote(bar(H) * " recruitment"),
                       x = bquote(bar(H) * " stand")) +
@@ -400,7 +402,8 @@
         ggplot2::geom_abline() +
         ggplot2::scale_color_manual(labels = c("Observed" = "Observed", 
                                                "model" = "Simulated"),
-                                    values = c("#FDAE61", "grey"))
+                                    values = c("#FDAE61", "grey")) +
[…]

### 2023-10-01 manuscript/manuscriptRW.Rmd
r2 comment 22/ 24

@@
 Two models, ForClim and ForCEEPS, were used here with two alternative variants of regeneration.
-ForClim variant 1 [@bugmann1996] is based on a recruitment module that adheres closely to the concept introduced by @botkin1972, whereas ForClim variant 11 is adopting a different approach where individual species properties and their relation to the environment are important [@huber2020].
+ForClim variant 1 [@bugmann1996] is based on a recruitment module that adheres closely to the concept introduced by @botkin1972 which is an environmental filter considering species specific thresholds of light availability and climatic variables.
+Whereas ForClim variant 11 is adopting a slightly more complex approach where individual species properties and their relation to the environment are represented gradually along with the relative suitability of different species (i.e., competition among tree species) [@huber2020].
 These two model variants allow us to evaluate a more process-based and complex module (variant 11) against a simple module (variant 1), while the rest of the model structure is identical.

### 2023-10-01 manuscript/review/authorsResponse.qmd
r2 comment 22/ 24

@@
 We agree with the reviewer's comment that there are other forest inventory data networks outside Central Europe even on forest reserves that could provide this sort of data.
-However, the EuFoRIa network is indeed novel and unprecedented network in Central Europe.
+However, within Europe no such data existed before.
+The research network (EuFoRIa) was established in 2019 and therefor is in fact novel, and within Europe such data is also without precedence.
+
+Yet we acknowledge that our statement needs to be put in a European context much clearer. We hope that our adjustments to the text resolve the exaggeration of our initial formulation.
 
@@
 
+​
+
+​
+
+​
+
+​
+
+​
+
+​
+
+​
+
+​
+
+​
+
+​
+
 ## Other changes made in the manuscript

### 2023-10-02 manuscript/manuscriptRW.Rmd
unsaved changes

@@
 Regeneration data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset covering such gradients over a large number of sites, to evaluate how well regeneration is captured.
-The observations used here are derived from a novel and unprecedented network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network [@euforia2019] (Figure \@ref(fig:map)).
+The observations used here are derived from a novel network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network [@euforia2019;TODO add @kaber2023] (Figure \@ref(fig:map)).
+The environmental coverage and the variety of forests represented by this data is unprecedented within Europe.
 These forests have been unmanaged for at least 10 years prior to their designation as reserves, and most of them provide time series of natural forest dynamics over multiple decades and up to ca.

### 2023-10-03 manuscript/manuscriptRW.Rmd
changes following the review

@@
 The fact that some models (and observed data) featured lower mortality than expected by Reineke's self-thinning rule may point to facilitation, or simply a higher mortality rate before the trees had reached 7 cm diameter.
-Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher regeneration mortality in mixed-species forests compared to even-aged, single-species forests.
+Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher regeneration mortality compared to even-aged, single-species forests.
 Some models yielded a mortality rate of regeneration that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).
@@
 It would have been extremely challenging for such a variety of models to follow a protocol where the spatial sampling size was different at each of the 200 sites, and it would have introduced additional uncertainty in the results.
-Ultimately, we need to acknowledge that the observed data represent just a snapshot of an extremely complex and stochastic reality.
-This may at least partly explain e.g.
-the lack of clear patterns in the simulation results along the gradients of climate water balance and degree-days.
 
@@
 
-Therefore, we further recommend that more effort should be invested into collecting harmonized datasets on tree regeneration.
+Therefore, we further recommend that more effort should be invested into collecting harmonized datasets in a site-specific manner covering the different aspects leading to tree regeneration.
 We emphasize that datasets such as EuFoRIa are invaluable and should be expanded in both their spatial extent (e.g., towards boreal and Mediterranean conditions) as well as in time (e.g., continuing the monitoring into the future).

### 2023-10-03 manuscript/review/authorsResponse.qmd
changes following the review

@@
 number-sections: true
-format: docx
+format: pdf
 ---
 
-## Internal instructions
+# Internal instructions
 
@@
 
-## Journal formatting (ToDo list)
+# Journal formatting (ToDo list)
 
@@
 
-## Response to the Editor
-
-### General Comments
-
-Dear Dr. Díaz-Yáñez:
-
-Thank you very much for submitting your manuscript "Tree regeneration in models of forest dynamics: a key priority for further research" for review by Ecosphere.
-The reviewers and I appreciate the work you have accomplished.
-I am willing to consider a revised version for publication in the journal, assuming that you are able to modify the manuscript according to the reviewers' extensive recommendations.
-Your revisions should address the specific points made in the review comments.
- Please note that Reviewer #2 made extensive comments directly on the manuscript.\
+# Response to the Editor
+
+## General Comments
+
+Your revisions should address the specific points made in the review comments.  Please note that Reviewer #2 made extensive comments directly on the manuscript.\
 I should note that personally I can hardly disagree with your basic conclusion that models of forest dynamics need to do a good job of capturing regeneration and recruitment processes.
@@
 
-Sincerely,
-
-Dr. Charles Canham Subject-matter Editor, Ecosphere
-
-### Response to editors comments
-
-Dear Dr. Charles Canham,
-
-I want to express my gratitude for taking the time to review our manuscript. Your insightful thoughts on our findings were very helpful.
-
-Our main takeaway from this manuscript is that while it may be possible to calibrate all 15 models against the EuFoRIa data, it is important to ensure that this aligns with the study's goals.
-If the goal is to provide extrapolation into a future climate, calibrating against data under the current climate may not be reliable.
-This is why we believe it is valuable to test model performance without any changes or calibration as we have done in this study.
-
-We have carefully reviewed and addressed each of the specific points made in the review comments in this revised version.
-We believe that the changes we made have significantly improved the manuscript.
-Once again, I want to thank you for the time and effort you dedicated to reviewing our work.
-We look forward to hearing from you again.
-
-Best regards,
-
-Olalla Díaz-Yáñez on behalf of all the co-authors
-
-## Response to reviewer 1
-
-### General Comments
+**Response to editors comments**
+
+Our main take-away from this manuscript is that while it would certainly be possible to calibrate all 15 models against the EuFoRIa data, it is important to ensure that this aligns with the study's goals.
+If the goal is to provide extrapolation into a future climate, calibrating against data under the current climate may not be reliable, as shown by the two models in our set that *were* calibrated against such data (i.e., SIBYLA and xComp).
+This is why we believe it is valuable to test the performance of the models "as they are", i.e.
+as we have done in this study.
+
+We have carefully reviewed and addressed each of the specific points made in the review comments in this revised version, as explained below.
+
+# Response to reviewer 1
+
+## General Comments
 
@@
 
-##### Our general comments to reviewer 1:
-
-Dear Reviewer 1,
+Our general comments to reviewer 1:
 
@@
 We appreciate your thoughtful comments and have carefully considered each of them.
-We have made changes to the manuscript based on your feedback and believe that these changes have improved the overall quality of the work.
-
-We understand your concern about the length of the manuscript and the number of figures included.
+We have made changes to the manuscript based on your feedback and believe that these changes have improved the overall quality of our manuscript.
+
+We appreciate the concern about the length of the manuscript and the number of figures included.
 We have made an effort to reduce the text where possible and have verified that we are following the journal guidelines.
@@
 
-Please find below our responses to each of your comments and the concrete actions we have taken in response.
-
-Thank you again for your valuable feedback.
-
-### Major comments
-
-#### [COMMENT 1]{.underline}
+Our responses to each of the comments and the concrete actions we have taken are detailed below.
+
+## Major comments
+
+### COMMENT 1
 
@@
 This is recruitment.
-Since this is the response variable for the manuscript, I would suggest replacing 'regeneration' with 'recruitment' throughout the manuscript.\
-
-##### Response to comment 1:
-
-[We should discuss this point again... I do not agree to simple change regeneration with recruitment. Specially without a detailed definition as we had in the previous version.]{.underline}
-
-[SUGGESTION FROM YANNEK:]{.underline}
-
-[Thank you for pointing this out.
-We also discussed terminology several times and we agree.
-We followed your suggestion replaced 'regeneration' with 'recruitment' where the difference is not clear.
-However, we believe that in some places regeneration is more accurate.
-Especially in many parts of the introduction and discussion.
-We argue that regeneration is often only modeled via "recruitment", but with the aim to simulate the whole process of regeneration.
-In this context it makes more sense to use the general term (e.g., for discussion consequences.)]{.underline}
-
-[COMMENT from YK: I would argue that recruitment is evidence for successful regeneration.
-This allows us to use recruitment for discussions observation vs.
-simulation in both aspects: technical (recruitment) and general (regeneration).]{.underline}
-
-#### COMMENT 2
+Since this is the response variable for the manuscript, I would suggest replacing 'regeneration' with 'recruitment' throughout the manuscript.
+
+**Response to comment 1:**
+
+Thank you for pointing this out.
+We discussed terminology several times and basically agree.
+We have followed your suggestion and have replaced the erroneous use of 'regeneration' by 'recruitment' where we refer to ingrowth into a certain DBH class.
+However, in some places the term 'regeneration' is more accurate, especially in many parts of the Introduction and Discussion sections.
+'Regeneration' is a broader term that starts from flowering and ends with ingrowth into a given size class; that is, recruitment is the result of successful regeneration.
+Thus, where we refer to the entire process of regeneration, it would not be appropriate to use 'recruitment', and therefore we maintain the former term where it is fitting.
+
+### COMMENT 2
 
@@
 
-##### Response to comment 2:
-
-Thank you for bringing to our attention the issue with the way we communicated our diameter thresholds approach in the manuscript.
-The observed data had varying diameter thresholds for reporting tree recruitment, and, for the purposes of this study, we merged the recruited trees values into two thresholds of 7 and 10cm.
-
-In total, we had 165 sites where the records showed the number of recruited trees at a 7cm threshold.
-For these sites, any recruited tree above 7cm was counted.
-We also had 35 sites where the minimum diameter threshold was 10cm, and for these sites, we could only account for trees recruited from 10cm diameter.
-We used the other 165 sites with 7cm diameter threshold in the 10cm database, but only accounting for newly established trees above 10cm.
-There was no like-to-like comparison for any other diameter threshold, only for 7 and 10cm.
-
-The reviewer correctly pointed out that in Figure 2 for the 7cm threshold, there were 165 "dots" in the observed data, while for the 10cm diameter threshold, there were 200 "dots." However, simulated values were obtained for the 200 sites for both thresholds, as explained in the Simulation protocol section.
-Therefore, the simulated values in the figures are from the 200 sites.
-
-When examining the change in mortality between 7 to 10cm or species diversity comparison between observed and simulated data (Figure 4), we could only use the 165 sites that contained both pieces of information at the 7 and 10cm thresholds in the observed data.
-This is reported in the supplementary materials software, script \`\`\`Figures.R\`\`\` line 536 and L284-349 and now better explained in the manuscript.
-
-To better clarify this aspect, we have made the following changes:
-
-1\) We have better explained the difference on the observed data in the section 3.2 Observed data with the following text:
+**Response to comment 2:**
+
+We realized that there is an issue with the way we communicated our diameter thresholds approach in the manuscript.
+The observed data had varying diameter thresholds for reporting tree recruitment, and, for the purposes of this study, we merged the recruited trees values into two thresholds of 7 and 10 cm.
+ That is, in all plots (*n* = 165) that had a calipering limit \<7 cm, only the successful recruitment events at 7 cm were considered; and analogously for plots (*n* = 35) with a callipering limit between 7 and 10 cm, where only recruitment at 10 cm was considered.
+Thus, the 7 cm dataset (*n* = 165) actually is a subset of the 10 cm dataset (*n* = 200), as all data from the 7-cm threshold could be re-calculated for the 10-cm threshold.
+Thus, there was no 'like-to-like' comparison for any other diameter threshold, only for 7 and 10cm.
+
+When examining the change in mortality between 7 and 10 cm, or the difference in species diversity between observed and simulated data (Figure 4), we could only use the 165 sites that contained both pieces of information at the 7 and 10 cm thresholds in the observed data.
+This is reported in the supplementary materials software, script \`\`\`Figures.R\`\`\` line 536 and L284-349 and now better explained in the manuscript:
+
+1\) We have better explained the difference on the observed data in the section 3.2 Observed data L\[XX-XX\]
 
@@
 
-2\) We have added the following text in the caption of figure 4
+2\) We have added a new text in the caption of figure 4 L\[XX-XX\]
 
@@
 
-#### [COMMENT 3]{.underline}
+### [COMMENT 3]{.underline}
 
@@
 
-##### Response to comment 3:
-
-[I am going to do this.]{.underline}
-
-### Minor comments:
-
-#### COMMENT 4
+**Response to comment 3:**
+
+[PENDING: I am going to do this.]{.underline}
+
+## Minor comments:
+
+### COMMENT 4
 
@@
 
-##### Response to comment 4:
-
-We have made some changes to the abstract based on feedback from both reviewers, including clarifying certain parts for better understanding.
+Response to comment 4:
+
+We have changed the abstract based on feedback from both reviewers, including clarifying certain parts for better understanding.
 The models were able to capture some aspects of regeneration more effectively than others.
-One example, is the sentence highlighted by the reviewer that was refereing to the diversity of regeneration.
-
-#### COMMENT 5
+One example, is the sentence highlighted by the reviewer that was referring to the diversity of regeneration.
+
+### COMMENT 5
 
@@
 
-##### Response to comment 5:
-
-Following the reviewers comments we have made several changes in this paragraph to improve its readability and flow if ideas with the main purpose of the study text.
-
-#### COMMENT 6
+**Response to comment 5:**
+
+We have made several changes in this paragraph to improve its readability and flow of ideas towards explaining the main purpose of the study.
+
+### COMMENT 6
 
@@
 For models that start from saplings or ingrowth, please add in the size as well (i.e., start from saplings with DBH = 1.0).
-Could also add in the classification scheme mentioned in the methods (L168, regeneration or recruitment) to Table 1. \
-
-##### Response to comment 6:
-
-Following the reviewers comment we have added a new column classifying the models in recruit or regeneration models; and the letters a,b,c, d,e and f table are now better described.
-Regarding the suggestion on adding into this table the size of the saplings or ingrowth: there are models that start from saplings or ingrowth what without a constant diameter threshold as they are biomass based (e.g.
-LandClim), we find it complicated to add all this information in this table without making it too large and therefore we have decided not to included.
-Furthermore the models reports in the supplementary materials provide more details on how the models simulate regeneration.
-
-ToDo All models check that the new column in the table is correct for your model.
-
-#### COMMENT 7
-
-L218 -- do I understand this correctly, that regeneration rates are a mean of 56 trees per ha per decade? This seems really low (assuming that this is for \>10 cm DBH).  \
-
-##### Response to comment 7:
-
-Yes, this is the mean value across all observations and sites, there are two main reasons why this number might seem low, first due to the zero inflated nature of the data, and second because these are forest reserves, corresponding with denser forests in which regeneration would be expected to be lower.
-This value is in agreement with a further observed data from reserved (Käber, Y., et al (2023).
+Could also add in the classification scheme mentioned in the methods (L168, regeneration or recruitment) to Table 1.
+
+**Response to comment 6:**
+
+We have added a new column classifying the models into recruitment or regeneration models; and the letters a,b,c, d,e and f in the table are now better described.
+Regarding the suggestion of adding the size of the saplings or ingrowth: we are concerned that this would be more confusing than helpful.
+For example, there are models that start from saplings without a diameter threshold, as they are biomass-based (e.g. LandClim), and providing all information for each model would make this table rather overloaded, without a clear benefit, as we believe.
+Therefore we have decided not to include this information.
+We would also like to note that the 15 model reports in the Supplementary Materials provide more details how regeneration is simulated.
+
+### COMMENT 7
+
+L218 -- do I understand this correctly, that regeneration rates are a mean of 56 trees per ha per decade? This seems really low (assuming that this is for \>10 cm DBH).
+
+**Response to comment 7:**
+
+Yes, this is the mean value across all observations and sites.
+There are two major reasons why this number appears low.
+First due to the zero-inflated nature of the data, and second because these are forest reserves, typically being characterized by denser forests in which regeneration can be expected to be low.
+This value is in agreement with further observed data from reserves (Käber, Y., et al (2023).
 Sheltered or suppressed?
 Tree regeneration in unmanaged European forests.
-Journal of Ecology, 00, 1--15.
-https://doi.org/10.1111/1365-2745.14181)
-
-#### COMMENT 8
+Journal of Ecology, https://doi.org/10.1111/1365-2745.14181).
+
+### COMMENT 8
 
@@
 
-##### Response to comment 8:
+**Response to comment 8:**
 
 For those models that simulated more than the eleven species, we created a different category "others" were we lumped those extra recruited species.
-This were counted when we calculated the diversity index and for total recruitment values.
-Following the reviewer comments we have added this information in the simulation protocol section with the following text:
-
-"Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species simulated beyond the eleven species were lumped in a category "others""
-
-#### COMMENT 9
+These were counted when we calculated the diversity index as well as for total recruitment values.
+We have added this information in the simulation protocol section L\[XX-XX\].
+
+"Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species simulated beyond the eleven species were lumped in a category"others""
+
+### COMMENT 9
 
@@
 
-##### Response to comment 9:
-
-Following the reviewers comment we have modified this text to improve it readability with:
+**Response to comment 9:**
+
+Following the reviewers comment we have modified the text L\[XX-XX\]
 
@@
 
-#### COMMENT 10
-
-L364: so please remove the 4C model from the figure as it is not comparable. Do you need to also remove other models that don't simulate the same number of species too?\
-
-##### Response to comment 10:
-
-Even thought it is not directly comparable, we still believe is interesting for the reader that we share the findings related to the 4C model and those that simulate more species that were lumped in the category others.
-
-#### [COMMENT 11]{.underline}
+### COMMENT 10
+
+L364: so please remove the 4C model from the figure as it is not comparable. Do you need to also remove other models that don't simulate the same number of species too?
+
+**Response to comment 10:**
+
+We acknowledge the point that the reviewer is making, but even though it is not directly comparable, we still believe to be insightful for the reader that we share the findings related to the 4C model.
+Note that in those models that simulate more species, these additional species were lumped in the category 'others' (as explained further above), thus with little impact on the comparison.
+
+### [COMMENT 11]{.underline}
 
@@
 
-##### Response to comment 11:
+**Response to comment 11:**
 
@@
 
-#### COMMENT 12
+### COMMENT 12
 
@@
 
-##### Response to comment 12:
-
-Following the reviewers comment we have increased the legend size, and now the difference between observed and simulated is visible.
-We have also explained in the caption the meaning of the black line:
+**Response to comment 12:**
+
+We have increased the font size of the legend, and now the difference between observed and simulated is visible. We have also explained the meaning of the black line in the caption L\[XX-XX\].
 
@@
 
-#### COMMENT 13
+### COMMENT 13
 
@@
 
-##### Response to comment 13:
-
-We agree with the reviewer that it was interesting to find that the models with feedback did not captured better the species diversity of the regeneration compared with those without, with the exeption of ForCEEPS.
-The Figure 4 shows three representative patterns observed across models using 6 of them as an example, and the Figure S3 provides the details for each model.
-Models with feedback are defined in table 1 and are ForCEEPS(f) , PICUS, SIBYLA, xComp, iLand, LandClim, Landis-II, TreeMig, LPJ-GUESS, and aDGVM2.
-We have addressed this point in the discussion section with the following text:
+**Response to comment 13:**
+
+We agree that it is interesting that the models with feedback did not capture better the species diversity of the regeneration compared with those without, with the exeption of ForCEEPS.
+Figure 4 shows three representative patterns observed across models using 6 of the models as an example, and Figure S3 provides the details for each model.
+We believe it is not very informative to color the points from the models with feedback in Figure 4 as this will only color all the points from some of the panels and it is possible to identify the models with feedback in table 1 We have addressed this point in the discussion section L\[XX-XX\].
 
@@
 
-#### COMMENT 14
+### COMMENT 14
 
@@
 
-##### Response to comment 14:
-
-Following the reviewers recommendation we have changed the axis labels in the appendix figures.
-
-#### COMMENT 15
+**Response to comment 14:**
+
+Done.
+
+### COMMENT 15
 
@@
 
-##### Response to comment 15:
-
-This paragraph is addressing the comparison with the theoretical Reineke value which is usually calculated for even-aged, single species stands and assumes a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands.
-
-#### COMMENT 16
+**Response to comment 15:**
+
+This paragraph is addressing the comparison with the theoretical Reineke value which is usually calculated for even-aged, single species stands. We have change this text to clarify this aspect.
+
+### COMMENT 16
 
@@
 
-##### Response to comment 16:
-
-Thank you! We also think is a good summary figure.
-
-#### [COMMENT 17]{.underline}
+**Response to comment 16:**
+
+Thank you!
+
+### [COMMENT 17]{.underline}
 
@@
 
-##### Response to comment 17:
+**Response to comment 17:**
 
@@
 
-#### COMMENT 18
+### COMMENT 18
 
@@
 
-##### Response to comment 18:
-
-Following the reviewer comment we have removed figure S7.
-
-#### COMMENT 19
+**Response to comment 18:**
+
+We have removed figure S7.
+
+### COMMENT 19
 
@@
 
-##### Response to comment 19:
-
-Figure 4 shows three examples of the patterns observed across models (overestimation intermediate and underestimation) using one exemplary model of each category , the actual patterns per model can be found in Figure S3.
-In the results we showed that most models captured reasonably well or overpredicted species diversity at the stand level, as most of the models are under the patterns presented in plot 1 and 2 from A on Figure 4.
-We have decided to keep the referenced statement link to this findings in the discussion as it was.
-
-#### COMMENT 20
+**Response to comment 19:**
+
+Figure 4 shows three examples of the patterns observed across models (overestimation intermediate and underestimation) using one exemplary model of each category.
+The actual patterns per model can be found in Figure S3.
+In the results, we showed that most models capture reasonably well or overpredicted species diversity at the stand level, as most of the models are following the patterns presented in plot 1 and 2 (part A of Figure 4).
+We have decided to keep the referenced statement regarding this finding in the discussion.
+
+### COMMENT 20
 
@@
 
-##### Response to comment 20:
+Response to comment 20:
 
 In this statement we were referring only to regeneration levels.
-Although we agree with the reviewer that we did state that most of the models overestimate regeneration, here our point is that even-thought they did overestimate regeneration they did not deviate exceedingly, specially under the consideration that these models have never been confronted with this type of data.
-
-#### COMMENT 21
+Although we agree with the reviewer that we did state that most of the models overestimate regeneration, here our point is that even though they did overestimate regeneration, they did not deviate exceedingly, especially given that they have never been confronted with this type of data.
+
+### COMMENT 21
 
@@
 
-##### Response to comment 21:
-
-With this statement we were referring to the multiple potential combinations of approaches of modeling tree regeneration. Based on the reviewers comment we have modified this for clarity.
-
-#### COMMENT 22
+Response to comment 21:
+
+With this statement, we were referring to the multiple potential combinations of approaches of modeling tree regeneration. We have modified this for clarity L\[XX-XX\]
+
+### COMMENT 22
 
@@
 
-##### Response to comment 22:
-
-Thank you for noticing this. We have now corrected this mistake.
-
-#### COMMENT 23
+**Response to comment 22:**
+
+This was a glitch, which now is corrected L\[ XX-XX\]
+
+### COMMENT 23
 
@@
 
-##### Response to comment 23:
-
-We agree with the reviewer that ideally models should be able to capture broader patterns.
-With this sentence we just wan to explain that the stochasticity of this process may have partly explained the results on this regard.
-
-#### COMMENT 24
+**Response to comment 23:**
+
+We have removed these statements.
+
+### COMMENT 24
 
@@
 
-##### Response to comment 24:
-
-The reviewer is correct that landscape level models were not significantly performing worst than others.
-Here we just wanted to discuss that maybe another spatial consideration will have allowed landscape level models to better represent regeneration processes as they are designed for that.
-
-#### COMMENT 25
+Response to comment 24:
+
+Yes, the landscape level models were not performing significantly worse than the others.
+However, they were not performing better either, potentially because they were used in "single site" rather than "landscape" mode.
+Thus, the explicit consideration of horizontal space would possibly allow the landscape level models to better represent regeneration processes, as they are designed for that.
+This, however, was completely out of the scope of the present analysis, as there are no landscape-level data of the surroundings of the 200 plots.
+
+### COMMENT 25
 
@@
 
-##### Response to comment 25:
-
-We agree with the reviewer, a coarse representation of water balance would not make it more sensitive to water deficits.
-And this was exactly our point in this part of the discussion where we wrote: "This may at least partially explain the unsatisfactory performance of many models along the drought axis (i.e., climatic water balance)."
-
-#### [COMMENT 26]{.underline}
+**Response to comment 25:**
+
+Our point here is that the models, although many of them have detailed schemes for the water balance, may not have performed well because of the absence of site-specific (truly reliable) soil information.
+
+### COMMENT 26
 
@@
 
-##### Response to comment 26:
-
-Following the reviewers comment we have expanded this aspect in the following paragraph XXXX
-
-[Both reviewers asked to be more specific on the data we are missing, what vairables would you say without being too detailed? also linked to comment 59 from reviewer 2]{.underline}
-
-## Response to reviewer 2
-
-### General Comments
-
-The paper evaluates whether forest dynamics models are successful at capturing the essence of regeneration processes, by comparing outputs from 15 simulation models with tree recruitment data obtained from 300 mature forests spread across central/eastern Europe.
- Specifically, the models predicted how many trees of different species recruited each year (i.e.
-grew larger than 7-cm or 10-cm stem diameter) and indices derived from these numbers were compared with analogous indices derived from repeat-measured forest inventory plots.
- Most of the simulation models performed poorly for most of the indices examined.
- In particular, they tended to overestimate recruitment rates.
+**Response to comment 26:**
+
+We have changed this text L\[XX-XX\].
+Although there are multiple data sources (such as MASTIF, MASTREE, National Forest Inventories, Forest Reserve plots, etc.), usually these data do not refer to the same locations and thus cannot be used in a site-specific manner for calibrating or testing models that intend to follow the ‚regeneration' chain of processes.
+
+"Therefore, we further recommend that more effort should be invested into collecting harmonized datasets in a site-specific manner covering the different aspects leading to tree regeneration."
+
+# Response to reviewer 2
+
+## General Comments
 
@@
 
-##### Our general comments to reviewer 2:
-
-Dear Reviewer 2,
-
-We would like to express our sincere gratitude for all the detailed comments they provided on our manuscript.
-We have carefully considered all of their feedback and made the necessary changes to improve the manuscript.
-Thanks to your comments, we have improved our manuscript.
-
-Thank you again for your valuable feedback.
-
-### Major comments
-
-#### [COMMENT 1]{.underline}
+**Our general comments to reviewer 2:**
+
+We would like to thank Reviewer #2 for this positive overall assessment and the level of detail of the review.
+We have taken care to go through all suggestions made in the PDF, and have changed/improved the manuscript in most instances; however these suggestions are too numerous to be dealt with on an individual basis here.
+
+## Major comments
+
+### COMMENT 1
 
@@
 
-##### Response to comment 1:
-
-[I think we have proposed some ideas of how to do this better in the discussion. What else do you think we should add here?]{.underline}
-
-#### [COMMENT 2]{.underline} {#sec-R2C2}
+Response to comment 1:
+
+In this first assessment we were not aiming at identify what specific subcomponents may hold the key and therefore dased on the current approach we could not provide more concrete recommendation for model development which most likely be specific for each of the 15 models.
+Maybe we need more such studies, including calibration as a first step, and then application against a withheld part of the data; or application (and comparison) under climate change scenarios to evaluate the robustness of the signal under future conditions.
+
+### COMMENT 2 {#sec-R2C2}
 
@@
 
-##### Response to comment 2:
-
-Following the reviewer comment we have added a "road map" at the end of the introduction.
-
-#### [COMMENT 3]{.underline}
+**Response to comment 2:**
+
+We have added a "road map" at the end of the introduction.
+
+### [COMMENT 3]{.underline}
 
@@
 
-##### Response to comment 3:
-
-[Decide final terminology]{.underline}
-
-#### COMMENT 4
+**Response to comment 3:**
+
+Thank you for pointing this out.
+We discussed terminology several times and basically agree.
+We have followed your suggestion and have now use recruitment to discuss the process and regeneration as the overarching term.
+
+### COMMENT 4
 
@@
 
-##### Response to comment 4:
-
-The reviewer is correct that most of the people involved in this manuscript have a central European focus in their current work, and so did the study set up.
-We have included some relevant global or with a focus outside Europe references when relevant.
-e would be happy to consider some other relevant suggestions.
-
-### Minor comments:
+**Response to comment 4:**
+
+Indeed, most of the people involved in this manuscript have a central European focus in their current work, and so did the study setup.
+We have included global references or references with a focus outside Europe when they were relevant.
+We have evaluated whether this could be further "internationalized", but felt our choices were appropriate.
+
+## Minor comments:
 
@@
 
-#### COMMENT 5
+### COMMENT 5
 
@@
 
-##### Response to comment 5:
-
-Based on the reviewers comments we have modified this sentence to: "However, there is a need to evaluate their capacity to accurately represent tree regeneration"
-
-#### COMMENT 6
+Response to comment 5:
+
+We have modified this sentence L\[XX-XX\]
+
+"However, there is a need to evaluate their capacity to accurately represent tree regeneration"
+
+### COMMENT 6
 
@@
 
-##### Response to comment 6:
-
-Following the reviewers comment we have changed the word "comprehensive" to "extensive" to express that we used a wide-range data that considers many important elements of regeneration.
-
-#### COMMENT 7
+Response to comment 6:
+
+We have changed the word "comprehensive" to "extensive" to express that we used a wide-range data that considers many important elements of regeneration.
+
+### COMMENT 7
 
@@
 
-##### Response to comment 7:
-
-Following the reviewers comments we have changed "feature" with "present".
-
-#### COMMENT 8
+**Response to comment 7:**
+
+We have changed "feature" with "present".
+
+### COMMENT 8
 
@@
 
-##### Response to comment 8:
-
-Following the reviewers comment, we have changed the word "were" with "have been".
-
-#### [COMMENT 9]{.underline}
+**Response to comment 8:**
+
+We have changed the word "were" with "have been".
+
+### COMMENT 9
 
@@
 
-##### Response to comment 9:
-
-Following the reviewers comment, we have corrected the reference style.
-
-#### COMMENT 10
+**Response to comment 9:**
+
+We have corrected the reference style.
+
+### COMMENT 10
 
@@
 
-##### Response to comment 10:
-
-Following the reviewers comment, we have changed this sentence to: "This presents an important research gap, particularly in the context of climate-induced forest disturbances and forest resilience."
-
-#### COMMENT 11
+**Response to comment 10:**
+
+We have changed this sentence to: "This presents an important research gap, particularly in the context of climate-induced forest disturbances and forest resilience."
+
+### COMMENT 11
 
@@
 
-##### Response to comment 11:
-
-Following the reviewers comment, we have changed "such as" with "including"
-
-#### COMMENT 12
-
-L85-86 "Tree regeneration arises from multiple processes such as pollination, fruit maturation, seed production, dispersal, germination, juvenile growth and survival" including
-
-##### Response to comment 12:
-
-Following the reviewers comment, we have changed "such as" with "including"
-
-#### COMMENT 13
+Response to comment 11:
+
+We have changed "such as" with "including"
+
+### COMMENT 12
 
@@
 
-##### Response to comment 13:
-
-Following the reviewer's comment we have changed "it" with "them"
-
-#### [COMMENT 14]{.underline}
+Response to comment 12:
+
+We have changed "it" with "them"
+
+### [COMMENT 13]{.underline}
 
@@
 
-##### Response to comment 14:
-
-It is indeed interesting to account with the increasing data vailablity to better understand tree regeneration processes.
+Response to comment 13:
+
+It is indeed interesting to account with the increasing data availability to better understand tree regeneration processes.
 However, we believe that the main take away it is not that it would be impossible to calibrate all 15 models against e.g.
@@
 
-[Following this and other comments from the editor, we have tried to capture this better in the discussion/conclussions section. Should we add something on this?]{.underline}
-
-#### COMMENT 15
+### COMMENT 14
 
@@
 
-##### Response to comment 15:
+Response to comment 14:
 
@@
 
-#### [COMMENT 16]{.underline}
+### [COMMENT 15]{.underline}
 
@@
 
-##### Response to comment 16:
-
-[Following the reviewer's comment we have added a reference]{.underline}
+**Response to comment 15:**
 
@@
 
-#### [COMMENT 17]{.underline}
+### [COMMENT 16]{.underline}
 
@@
 
-##### Response to comment 17:
-
-[Following the reviewer's comment we have added a reference]{.underline}
+**Response to comment 16:**
 
@@
 
-#### **COMMENT 18**
+### **COMMENT 17**
 
@@
 
-##### Response to comment 18:
-
-We currently use the exciting Hanbury-brown 2022 review in several places across our manuscripts, following the reviewer's comment here and the general comment 2 in @sec-R2C2 we have also get inspired by their structure by creating a road map to help the reader to navigate the manuscript content.
-
-#### **COMMENT 19**
+**Response to comment 17:**
+
+We currently use the exciting Hanbury-brown 2022 review in several places across our manuscripts, following the reviewer's comment here and the general comment 2 we have also get inspired by their structure by creating a road map to help the reader to navigate the manuscript content.
+
+### **COMMENT 18**
 
@@
 
-##### Response to comment 19:
+**Response to comment 18:**
 
@@
 
-####
-
-#### **COMMENT 20**
-
-L150 "regeneration niche of multiple species"
-
-L154 "traits"
-
-##### Response to comment 20:
-
-We are unsure why the reviewer highlighted those words
-
-#### **COMMENT 21**
+### **COMMENT 19**
 
@@
 
-##### Response to comment 21:
+Response to comment 19:
 
 This statement relates to Bugmann and Seidl (2022) where complexity values were provided.
-The exact details of how we calculated this mean regeneration formulation complexity value are available in the software suplementary materials to this paper: \`\`\`figure.R\`\`\` line 706-750.
+The exact details of how we calculated this mean regeneration formulation complexity value are available in the software supplementary materials to this paper: \`\`\`figure.R\`\`\` line 706-750.
 This value represents the mean complexity of regeneration formulations per model.
 
-#### [**COMMENT 22**]{.underline}
+### [**COMMENT 20**]{.underline}
 
@@
 
-##### Response to comment 22:
-
-[Is there a way to explain this shortly??]{.underline}
-
-#### **COMMENT 23**
+Response to comment 22:
+
+[Is there a way to explain this shortly \@ Yannek ??]{.underline}
+
+### **COMMENT 23**
 
@@
 
-##### Response to comment 22:
+Response to comment 22:
 
@@
 
-#### **COMMENT 24**
+### **COMMENT 24**
 
@@
 
-##### Response to comment 24:
+Response to comment 24:
 
@@
 
-#### **COMMENT 26**
+### **COMMENT 26**
 
@@
 
-##### Response to comment 26:
+Response to comment 26:
 
@@
 
-#### **COMMENT 27**
+### **COMMENT 27**
 
@@
 
-##### Response to comment 27:
-
-Following the reviewer's comment we have changed "and exposition (i.e.,slope and aspect)." with ", slope and aspect."
-
-#### **COMMENT 28**
+Response to comment 27:
+
+We have changed "and exposition (i.e.,slope and aspect)." with ", slope and aspect."
+
+### **COMMENT 28**
 
@@
 
-##### Response to comment 28:
+Response to comment 28:
 
@@
 
-#### **COMMENT 29**
+### **COMMENT 29**
 
@@
 
-##### Response to comment 29:
+Response to comment 29:
 
@@
 
-**COMMENT 30**
+### **COMMENT 30**
 
@@
 
-##### Response to comment 30:
+Response to comment 30:
 
@@
 
-#### **COMMENT 31**
+### **COMMENT 31**
 
@@
 
-##### Response to comment 31:
-
-Please, see our reply on comment 25 in here @sec-R2C25r
-
-#### **COMMENT 32**
+Response to comment 31:
+
+Please, see our reply on comment 25
+
+### **COMMENT 32**
 
@@
 
-##### Response to comment 32:
+Response to comment 32:
 
@@
 
-#### **COMMENT 33**
+### **COMMENT 33**
 
@@
 
-##### Response to comment 33:
+Response to comment 33:
 
@@
 
-#### [**COMMENT 34**]{.underline}
+### [**COMMENT 34**]{.underline}
 
@@
 
-#####
-
-##### Response to comment 34:
+Response to comment 34:
 
@@
 
-[We have re-assessed the terms used across this manuscript XXX]{.underline}
-
-####
-
-**COMMENT 35**
+We have re-assessed the terms used across this manuscript and change it across the text accordingly.
+
+### **COMMENT 35**
 
@@
 
-##### Response to comment 35:
+Response to comment 35:
 
@@
 
-#### **COMMENT 36**
+### **COMMENT 36**
 
@@
 
-##### Response to comment 36:
+Response to comment 36:
 
@@
 
-#### **COMMENT 37**
-
-L230-231 "That is, the models were run in "blind flight" mode." Remove
-
-##### Response to comment 37:
+### **COMMENT 37**
+
+L230-231 "That is, the models were run in"blind flight" mode." Remove
+
+Response to comment 37:
 
@@
 
-#### **COMMENT 38**
+### **COMMENT 38**
 
@@
 
-##### Response to comment 38:
+Response to comment 38:
 
@@
 
-####
-
 **COMMENT 39**
@@
 
-##### Response to comment 39:
+Response to comment 39:
 
@@
 
-#### **COMMENT 40**
+### **COMMENT 40**
 
@@
 
-##### Response to comment 40:
+Response to comment 40:
 
@@
 
-#### **COMMENT 41**
+### **COMMENT 41**
 
@@
 
-##### Response to comment 41:
+Response to comment 41:
 
@@
 
-#### **COMMENT 42**
+### **COMMENT 42**
 
@@
 
-##### Response to comment 42:
+Response to comment 42:
 
@@
 
-#### [**COMMENT 43**]{.underline}
+### **COMMENT 43**
 
@@
 
-##### Response to comment 43:
-
-[Change if we change regeneration with recruit]{.underline}
-
-####
-
-#### **COMMENT 44**
+Response to comment 43:
+
+We have change the terminology.
+
+### **COMMENT 44**
 
@@
 
-##### Response to comment 44:
+Response to comment 44:
 
@@
 
-#### **COMMENT 45**
+### **COMMENT 45**
 
@@
 
-##### Response to comment 45:
+Response to comment 45:
 
@@
 
-#### **COMMENT 46**
+### **COMMENT 46**
 
@@
 
-##### Response to comment 46:
-
-Following the reviewer's comment on the lack of claritiy of the selected thresholds explanation we have changed at the beginning of the section simulation protocol the first sentence where we explain that we considered tree regeneration as the passing of a breast height diameter threshold of 7 and 10 cm.
+Response to comment 46:
+
+Following the reviewer's comment on the lack of clarity of the selected thresholds explanation we have changed at the beginning of the section simulation protocol the first sentence where we explain that we considered tree regeneration as the passing of a breast height diameter threshold of 7 and 10 cm.
 We have also reported later on in this section (but before this line) that each of the models reported the number of trees crossing the two diameter thresholds.
 
-####
-
-#### COMMENT 47
+### COMMENT 47
 
@@
 
-##### Response to comment 47:
+Response to comment 47:
 
@@
 
-#### [COMMENT 49]{.underline}
-
-L316-317 "model complexity as defined by Bugmann and Seidl" say a little more...
-
-##### Response to comment 49:
-
-[Any suggestion of how can I add something else without making this much longer!]{.underline}
-
-#### COMMENT 50
+### COMMENT 50
 
@@
 
-##### Response to comment 50:
+Response to comment 50:
 
@@
 
-#### COMMENT 51
+### COMMENT 51
 
@@
 
-##### Response to comment 51:
-
-To our knowledge this is the only network of forest reserves in Europe covering a large environmental gradient and from which a standardization effort has been made toward understanding tree regeneration across different countries.
-We would be happy to learn of any other network of this characteristics in Europe and potentially use it in follow up studies.
-
-####
-
-#### COMMENT 52
+Response to comment 51:
+
+We have changed the text.
+
+###
+
+### COMMENT 52
 
@@
 
-##### Response to comment 52:
+Response to comment 52:
 
 Thank you for this remark, we agree with the reviewer that this is an important point and specially how different models are capturing or missing to represent the absence of regeneration.
-We have now captured this aspect also in the introduction in the sentence: "Due to the large variability in tree regeneration patterns in nature and the large number of factors driving and constraining this process" on top of the current whole paragraph on this issue present in the discussion.
-
-#### COMMENT 53
+We have now captured this aspect also in the introduction in the sentence L\[XX-XX\]
+
+"Due to the large variability in tree regeneration patterns in nature and the large number of factors driving and constraining this process" on top of the current whole paragraph on this issue present in the discussion.
+
+### COMMENT 53
 
@@
 
-##### Response to comment 53:
-
-Following the reviewer's comment we have included this point by adding into this paragraph: "Another possible reason for this difference may be the exclusion of factors like deer browsing or failure to accurately measure establishment filters, such as those to site limitants or competition with herb layers."
-
-#### COMMENT 54
+Response to comment 53:
+
+We have included this point by adding into this paragraph: "Another possible reason for this difference may be the exclusion of factors like deer browsing or failure to accurately measure establishment filters, such as those to site limitants or competition with herb layers."
+
+### COMMENT 54
 
@@
 
-##### Response to comment 54:
-
-Following the reviewers comment we have modified this sentence to clarifu what do we mean here with feedback: " Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the regeneration than those without feedback, "
+Response to comment 54:
+
+We have modified this sentence to clarify what do we mean here with feedback: " Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the regeneration than those without feedback, "
 
@@
 
-####
-
-#### COMMENT 55
+###
+
+### COMMENT 55
 
@@
 
-##### Response to comment 55:
+Response to comment 55:
 
@@
 
-####
-
-#### COMMENT 56
+###
+
+### COMMENT 56
 
@@
 
-##### Response to comment 56:
+Response to comment 56:
 
@@
 
-#### [COMMENT 57]{.underline}
+### [COMMENT 57]{.underline}
 
@@
 
-##### Response to comment 57:
+Response to comment 57:
 
@@
 
-####
-
-#### COMMENT 58
+###
+
+### COMMENT 58
 
@@
 
-##### Response to comment 58:
+Response to comment 58:
 
@@
 
-#### [COMMENT 59]{.underline}
+### COMMENT 59
 
@@
 
-##### Response to comment 58:
-
-[Linked to comment 26 from reviewer 1]{.underline}
-
-#### COMMENT 60
+Response to comment 58:
+
+We have changed this text L\[XX-XX\]
+
+### COMMENT 60
 
@@
 
-##### Response to comment 60:
+Response to comment 60:
 
@@
 
-####
-
-#### COMMENT 61
-
-L696 "Exercises like the one presented here, where the models are operated in "blind flight" mode." Definition of blind flight "To do something based on guesswork, intuition, or without any help or instructions"....is that what you mean?
-
-##### Response to comment 60:
+###
+
+### COMMENT 61
+
+L696 "Exercises like the one presented here, where the models are operated in"blind flight" mode." Definition of blind flight "To do something based on guesswork, intuition, or without any help or instructions"....is that what you mean?
+
+Response to comment 60:
 
@@
 
-#### COMMENT 62
+### COMMENT 62
 
@@
 
-##### Response to comment 62:
+Response to comment 62:
 
@@
 
-#### COMMENT 63
+### COMMENT 63
 
@@
 
-##### Response to comment 63:
+Response to comment 63:
 
@@
 
-#### 
-[COMMENT 64]{.underline}
+### COMMENT 64
 
@@
 
-##### Response to comment 65:
-
-[To follow up after terminology discussion]{.underline}
-
-#### COMMENT 65
+Response to comment 65:
+
+We have changed the terminology.
+
+### COMMENT 65
 
@@
 
-##### Response to comment 65:
+Response to comment 65:
 
@@
 
-#### 
-COMMENT 66
-
-L1056 no mortality of recruits as they grow from 7 to 10 cm diameter.
-
-##### Response to comment 66:
-
-We use the difference of observed trees recruited between 7 and 10cm as a proxy for mortality.
-
-#### 
-COMMENT 67
-
-L1077 Explain the fitted curves
-
-##### Response to comment 67:
-
-We are not sure if we are correctly understanding the reviewers point here. In figure 8 we do not show any fitted curves.
-
 ​
@@
 ​
-
-​
-
-​
-
-​
-
-​
-
-​
-
-​
-
-​
-
-​
-
-## Other changes made in the manuscript

### 2023-10-03 manuscript/review/editorsLetter.qmd
changes following the review

---
title: "letterEditor_revision"
format: pdf
editor: visual
---

Dear Dr. Charles Canham,

I want to express my gratitude for taking the time to review our manuscript. Your insightful thoughts on our findings were very helpful.

We believe that the changes we made have significantly improved the manuscript.

Once again, I want to thank you for the time and effort you dedicated to reviewing our work. We look forward to hearing from you again.

Best regards,

Olalla Díaz-Yáñez on behalf of all the co-authors

##

## Reviewer 1

Dear Reviewer 1,

Thank you for taking the time to review our manuscript. We appreciate your thoughtful comments and have carefully considered each of them. We have made changes to the manuscript based on your feedback and believe that these changes have improved the overall quality of the work.

Please find below our responses to each of your comments and the concrete actions we have taken in response.

Thank you again for your valuable feedback.

### 2023-10-09 manuscript/review/authorsResponse.qmd
export pdf response to review

@@
 number-sections: true
-format: pdf
+format: 
+  pdf:
+    toc: true
+    number-sections: true
+    colorlinks: true
 ---
@@
 
-### [COMMENT 3]{.underline}
+### COMMENT 3
 
@@
 And when the goal is to provide extrapolation into a future climate, then calibration against data under current climate may be unreliable.
-With the approach we have followed in this study where we have asked the models to simulate forest dynamics and capture regeneration without statistical predetermination we can better assess their behavior and understand how their current model structures and inclusion of exclusion of certain ecological processes impact the outputs we are seeing.
+With the approach we have followed in this study where we have asked the models to simulate forest dynamics and capture regeneration without statistical parametrisation we can better assess their behavior and understand how their current model structures and inclusion of exclusion of certain ecological processes impact the outputs we are seeing.
 
@@
 
-### **COMMENT 17**
+### COMMENT 17
 

### 2023-10-10 manuscript/manuscriptRW.Rmd
more updates in the revision before sending to co-authors

@@
 
-Most of the models overestimate regeneration levels, which is only compensated in some models by high simulated mortality rates in the early stages of individual trees dynamics.
-Simulated species diversity of regeneration matches the observed ranges.
-Models simulating higher species diversity at the stand level do not present higher regeneration diversity.
-The effect of light availability on regeneration levels is captured better than the effect of temperature and soil moisture, but patterns are not consistent across models.
-Increasing complexity in the tree regeneration modules of the models is not related to higher accuracy of simulated tree regeneration.
+Most of the models overestimate recruitment levels, which is only compensated in some models by high simulated mortality rates in the early stages of individual trees dynamics.
+Simulated species diversity of recruitment matches the observed ranges.
+Models simulating higher species diversity at the stand level do not present higher recruitment diversity.
+The effect of light availability on recruitment levels is captured better than the effect of temperature and soil moisture, but patterns are not consistent across models.
+Increasing complexity in the tree regeneration modules of the models is not related to higher accuracy of simulated tree recruitment.
 Furthermore, individual model design is more important than scale (stand, landscape, global) and approach (empirical, process-based) for accurately capturing tree regeneration.
@@
 Thus, a focus on the modeling of tree regeneration processes is sorely needed and overdue [@price2001; @walck2011] if we are to make reliable projections of future forest dynamics, i.e., when the models need to be operated in extrapolation mode, as well as from a fundamental ecological point of view for increasing systems understanding.
-In the present study, for the first time, a large number of forest models commonly used to assess forest dynamics under climate change are evaluated against a continental-scale, multi-species harmonized dataset on tree regeneration [@käber2023].
-By tree regeneration, we refer to the passing of trees across a specific diameter threshold ("ingrowth").
+In the present study, for the first time, a large number of forest models commonly used to assess forest dynamics under climate change are evaluated against a continental-scale, multi-species harmonized dataset on tree recruitment [@käber2023].
+By tree recruitment, we refer to the passing of trees across a specific diameter threshold ("ingrowth").
 We included models that are based on a range of "philosophies", from models operating at the stand to the global scale as well as the range of models from empirically derived to "process-based" [e.g., @bugmann1996; @fabrika2012; @lexer2001; @reyer2014].
@@
 
-More specifically, we aim to answer the following questions: (1) Are models of forest dynamics capturing accurately tree regeneration levels, initial tree species diversity and mortality in the regeneration?
+More specifically, we aim to answer the following questions: (1) Are models of forest dynamics capturing accurately tree recruitment levels, initial tree species diversity and mortality in the recruitment?
 (2) Do model traits explain differences in model performance?
-(3) How well do the models capture total regeneration and individual species regeneration niches across environmental gradients of light availability, temperature, and soil moisture?
+(3) How well do the models capture total recruitment and individual species regeneration niches across environmental gradients of light availability, temperature, and soil moisture?
 We evaluate the performance of regeneration models by analyzing simulated data alongside observed data.
-Specifically, we examine tree species diversity, mortality rates in tree regeneration, and regeneration levels.
-Then, we link the model's traits and performance to define regeneration and we assess the regeneration gradients on regeneration niches.
+Specifically, we examine tree species diversity, mortality rates in tree recruitment, and recruitment levels.
+Then, we link the model's traits and performance to define regeneration and we assess the recruitment gradients on regeneration niches.
 Finally, we provide research recommendations for next steps towards improving tree regeneration modules.
@@
 Two models, ForClim and ForCEEPS, were used here with two alternative variants of regeneration.
-ForClim variant 1 [@bugmann1996] is based on a recruitment module that adheres closely to the concept introduced by @botkin1972 which is an environmental filter considering species specific thresholds of light availability and climatic variables.
+ForClim variant 1 [@bugmann1996] is based on a regeneration module that adheres closely to the concept introduced by @botkin1972 which is an environmental filter considering species specific thresholds of light availability and climatic variables.
 Whereas ForClim variant 11 is adopting a slightly more complex approach where individual species properties and their relation to the environment are represented gradually along with the relative suitability of different species (i.e., competition among tree species) [@huber2020].
@@
 
-Regeneration data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset covering such gradients over a large number of sites, to evaluate how well regeneration is captured.
-The observations used here are derived from a novel network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network [@euforia2019;TODO add @kaber2023] (Figure \@ref(fig:map)).
+Recruitment data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset covering such gradients over a large number of sites, to evaluate how well regeneration is captured.
+The observations used here are derived from a novel network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network [@euforia2019;@käber2023] (Figure \@ref(fig:map)).
 The environmental coverage and the variety of forests represented by this data is unprecedented within Europe.
@@
 Each of these sites featured at least two consecutive measurements.
-Regeneration minimum size thresholds for these sites differed between stem diameters of 0 and 10 cm.
+Recruitment minimum size thresholds for these sites differed between stem diameters of 0 and 10 cm.
 For our study, we defined two datasets where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds from 10 cm.
 Therefore we had in total 200 sites with observations of newly established trees above 10 cm and 165 sites above 7cm.
-In the results section any result from observed data at 7 cm threshold is originated from the 165 sites.
+In the results section, any result from observed data at 7 cm threshold is originated from the 165 sites.
 
-Plot size ranged from 0.02 ha to 5.52 ha, and the data were further processed and aggregated following @käber2023 to provide regeneration rates per ha and per decade.
+Plot size ranged from 0.02 ha to 5.52 ha, and the data were further processed and aggregated following @käber2023 to provide recruitment rates per ha and per decade.
 The observations used in this study featured 30,900 newly established trees.
-Regeneration rates per site, sample and decade ranged from zero to 1246 trees, with a mean of 56 trees.
+Recruitment rates per site, sample and decade ranged from zero to 1246 trees, with a mean of 56 trees.
 Adult species composition was also available for each plot.
-For more details on this unique dataset and the detailed regeneration information, cf.
+For more details on this dataset and the detailed regeneration information, cf.
 @käber2023.
@@
 
-The overarching goal of the simulation experiments was to assess tree regeneration as it arises from empirical data against its representation in a wide range of models of forest dynamics.
-We define tree regeneration as the passing of a breast height diameter threshold of 7 and 10 cm (synonym: ingrowth).
+The overarching goal of the simulation experiments was to assess tree recruitment as it arises from empirical data against its representation in a wide range of models of forest dynamics.
+We define tree recruitment as the passing of a breast height diameter threshold of 7 and 10 cm (synonym: ingrowth).
 To this end, each modeling group was provided with a detailed protocol [@díaz-yáñez2022] with instructions how to perform the simulations, input variables on climate and soil conditions, and the list of expected output variables.
-Neither were further site information (except for the data specified below) nor any data on tree regeneration or forest stand features provided prior to the simulation.
+Neither were further site information (except for the data specified below) nor any data on tree recruitment or forest stand features provided prior to the simulation.
 
@@
 Only the model LPJ-GUESS had to include a background disturbance to increase chances of shade-intolerant species to establish.
-The simulations were set up to sample species-specific regeneration rates per decade and per ha in the equilibrium state of the model, typically entailing a "spin-up" run (as we did not provide any forest data).
+The simulations were set up to sample species-specific recruitment rates per decade and per ha in the equilibrium state of the model, typically entailing a "spin-up" run (as we did not provide any forest data).
 The modeling teams decided on the simulated area and how they derived these samples.
@@
 
-The simulations were run in the absence of management to a simulated equilibrium ("Potential Natural Vegetation") with the current climate, as the ultimate goal was to evaluate tree regeneration under comparable and near-equilibrium conditions.
+The simulations were run in the absence of management to a simulated equilibrium ("Potential Natural Vegetation") with the current climate, as the ultimate goal was to evaluate tree recruitment under comparable and near-equilibrium conditions.
 This entails the assumption that (1) the observations from the forest reserves reflect no traces of forest management, and (2) there is an equilibrium between forest dynamics and climate.
@@
 However, in the absence of detailed data on the history of each plot in the EuFoRIa network, some broad assumptions had to be made.
-Both the width of the regeneration niche (i.e., in environmental space) as well as the intensity of the regeneration process (i.e., the number of ingrowth trees per area and per unit of time) were of interest.
+Both the width of the regeneration niche (i.e., in environmental space) as well as the intensity of the regeneration process (i.e., the number of recruited trees per area and per unit of time) were of interest.
 
@@
 The reason for this is that 4C is not suitable for long-term simulations without management, due to misrepresentations in density-dependent mortality processes in long-term simulations and assumptions of tree geometry that lead to unrealistic single tree dimensions of very old individuals.
-LPJ-GUESS had 2% of the sample outputs without tree regeneration, and these were considered as zero stand basal area and zero regeneration for all the tree species simulated; one site produced grassland rather than a forest and was not included in the results.
+LPJ-GUESS had 2% of the sample outputs without tree recruitment, and these were considered as zero stand basal area and zero regeneration for all the tree species simulated; one site produced grassland rather than a forest and was not included in the results.
 
@@
 
-The simulation results were analyzed regarding (1) regeneration levels (i.e., ingrowth number per unit time and space), (2) regeneration species diversity, (3) regeneration mortality, (4) the relationship between model performance and model traits, and (5) ingrowth gradients along the regeneration niches.
+The simulation results were analyzed regarding (1) recruitment levels (i.e., ingrowth number per unit time and space), (2) recruitment species diversity, (3) recruitment mortality, (4) the relationship between model performance and model traits, and (5) ingrowth gradients along the regeneration niches.
 We evaluated species diversity across the models and in relation to the observed data by calculating the Shannon index $H$ based on the relative proportion of the species in terms of basal area.
@@
 
-Mortality in tree regeneration was assessed based on the ratio of regeneration between the 7 and 10 cm diameter thresholds.
-We used the Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference to estimate whether the ratio of regeneration between the 7 and 10 cm diameter thresholds were above or below the expected theoretical rate.
+Mortality in tree recruitment was assessed based on the ratio of recruitment between the 7 and 10 cm diameter thresholds.
+We used the Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference to estimate whether the ratio of recruitment between the 7 and 10 cm diameter thresholds were above or below the expected theoretical rate.
 The Reineke self-thinning rule is usually calculated for even-aged, single species stands and assumes a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands.
@@
 We assessed model performance in relation to model traits focusing on (1) model complexity as defined by @bugmann2022, (2) model type (empirical or process based), (3) the presence or absence of a canopy feedback for regeneration, and (4) the scale of application of the model (stand, landscape or global).
-We tested for significant differences using two-sided t-tests by considering each of these model features regarding the observed and simulated regeneration levels and species diversity, respectively.
+We tested for significant differences between simulated and observed values of the recruitment levels and species diversity by using paired t-tests.
 
-Total regeneration and the regeneration niches of the individual species were evaluated across the environmental gradients of light availability, temperature, and soil moisture.
+Total recruitment and the regeneration niches of the individual species were evaluated across the environmental gradients of light availability, temperature, and soil moisture.
 Stand basal area (the basal area of all the trees in each sample including tree regeneration) was used as a proxy for light availability at the forest floor, the annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing season warmth, and the climatic water balance as a proxy for soil moisture [@speich2019].
-Regeneration values were calculated as the mean across the 200 samples per site.
+Recruitment values of each species were calculated as the mean across the 200 samples per site.
 The observed data were modeled using a Generalized Additive Model [@wood2011] with a negative binomial distribution and restricted maximum likelihood to better understand the relationship between the environmental gradients and the levels of observed regeneration, relative to the simulation results per model.
@@
 In order to analyze the regeneration niches across the climatic gradients, we focused on five common tree species or genera: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and *Quercus* spp.
-For these species, we calculated the share in the regeneration basal area per site as the mean across the available samples per site:
+For these species, we calculated the share in the recruited basal area per site, where each site represents a part of the environmental space, as the mean across the available samples per site:
 
@@
 ```
-where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$ is the mean basal area in the regeneration (subsequently referred to as "regeneration basal area") of species $i$ at site $t$ across the available samples at that site.
-Furthermore, regarding the share in the regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero.
+where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$ is the mean basal area in the recruitment (subsequently referred to as "recruited basal area") of species $i$ at site $t$ across the available samples at that site.
+Furthermore, regarding the share in the recruited  basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the recruited basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total recruited basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the recruited basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero.
 
@@
 
-Simulated regeneration levels varied strongly across the 15 models and typically did not match the levels found in the forest reserve data (Figure \@ref(fig:over)).
-Regeneration was overestimated in most models for both the 7 and 10 cm diameter thresholds, with the exception of the empirical stand model SIBYLA, the landscape model Landis-II and the global model aDGVM2, which estimated regeneration levels at the lower end of the plausibility interval of the observed data (Figure \@ref(fig:over)).
+Simulated levels of recruitment varied strongly across the 15 models and typically did not match the levels found in the forest reserve data (Figure \@ref(fig:over)).
+Recruitment was overestimated in most models for both the 7 and 10 cm diameter thresholds, with the exception of the empirical stand model SIBYLA, the landscape model Landis-II and the global model aDGVM2, which estimated recruited levels at the lower end of the plausibility interval of the observed data (Figure \@ref(fig:over)).
 The models with the largest overestimation were the stand model PICUS and the landscape model TreeMig.
-For most models, the variability of simulated regeneration levels across the 200 sites (visible from the interquartile range in the box plots of Figure \@ref(fig:over)) was similar to or smaller than observed, with the notable exception of PICUS, where simulated regeneration variability was much larger.
+For most models, the variability of simulated regeneration levels across the 200 sites (visible from the interquartile range in the box plots of Figure \@ref(fig:over)) was similar to or smaller than observed, with the notable exception of PICUS, where simulated recruitment variability was much larger.
 
-Both the observations and the simulated data had no regeneration in some samples and at some sites (for details, cf.
+Both the observations and the simulated data had no recruitment in some samples and at some sites (for details, cf.
 Table S1).
-The observed data had 4% of the samples with no regeneration.
-Only three models had a larger proportion of no regeneration (4C, Landis-II and aDVM2).
-Two models always simulated regeneration for both the 7 and 10 cm threshold (xComp and TreeMig), i.e., they did not feature any zero values.
-The other ten models had a very low percentage of samples with no regeneration (0.01% - 2.39%), i.e., they had distinctly fewer occurrences of zeros compared to the observations.
+The observed data had 4% of the samples with no recruitment.
+Only three models had a larger proportion of no recruitment (4C, Landis-II and aDVM2).
+Two models always simulated recruitment for both the 7 and 10 cm threshold (xComp and TreeMig), i.e., they did not feature any zero values.
+The other ten models had a very low percentage of samples with no recruitment (0.01% - 2.39%), i.e., they had distinctly fewer occurrences of zeros compared to the observations.
 
@@
 Most models matched the level of diversity of the observed data quite well (Figure \@ref(fig:H710)).
-Five models overestimated regeneration diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter particularly for the 7 cm diameter threshold.
+Five models overestimated recruitment diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter particularly for the 7 cm diameter threshold.
 The model 4C is a special case, as it simulated five species only, i.e., its diversity values are not directly comparable to those of the other models, nor to the observations.
-Only one model, Landis-II, consistently underestimated regeneration diversity.
+Only one model, Landis-II, consistently underestimated recruited diversity.
 
-In most models, there were only small but significant differences in the species diversity of regeneration between 7 and 10 cm across sites (Tables S2).
-Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained the regeneration diversity between the 7 and 10 cm thresholds (i.e., the differences between them were not significant, see Table S2), and the same was evident from the observed data.
+In most models, there were only small but significant differences in the species diversity of recruitment between 7 and 10 cm across sites (Tables S2).
+Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained the recruitment diversity between the 7 and 10 cm thresholds (i.e., the differences between them were not significant, see Table S2), and the same was evident from the observed data.
 
-For both the observed and the simulated data, and both regeneration thresholds, species diversity in the regeneration compared to stand-level diversity did not feature clear patterns across the models (Figures S1 and S2).
+For both the observed and the simulated data, and both recruitment thresholds, species diversity in the recruitment compared to stand-level diversity did not feature clear patterns across the models (Figures S1 and S2).
 Most models captured reasonably well or overpredicted species diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left and center; cf.
@@
 
-The majority of the models overpredicted regeneration diversity; as already observed in Figure \@ref(fig:H710), only two of them underestimated it (4C and Landis-II) (Figure \@ref(fig:combinedRichness)B and Figure S4).
-Both the models with feedback and those without feedback from the adult trees to regeneration (via seed production) overestimated, underestimated or captured reasonably well regeneration species diversity compared to the observations.
-Again, simulated diversity had considerably lower variability than observed diversity in the regeneration.
+The majority of the models overpredicted recruitment diversity; as already observed in Figure \@ref(fig:H710), only two of them underestimated it (4C and Landis-II) (Figure \@ref(fig:combinedRichness)B and Figure S4).
+Both the models with feedback and those without feedback from the adult trees to regeneration (via seed production) overestimated, underestimated or captured reasonably well recruitment species diversity compared to the observations.
+Again, simulated diversity had considerably lower variability than observed diversity in the recruitment.
 
@@
 There were strong differences among the models regarding the mortality rate between 7 and 10 cm (Figure \@ref(fig:mortality11)).
-Most of the models featured a mortality rate significantly larger than the observed data, thus at least partially compensating for the general overestimation of regeneration levels (Figure \@ref(fig:over) and Table S3); six models did not feature a significant difference compared to the observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
-Four models (xComp, PICUS, LandClim and TreeMig) featured very high mortality rates (i.e., well above the Reineke self-thinning line), which compensated for the strong initial overestimation of regeneration (cf.
+Most of the models featured a mortality rate significantly larger than the observed data, thus at least partially compensating for the general overestimation of recruitment levels (Figure \@ref(fig:over) and Table S3); six models did not feature a significant difference compared to the observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
+Four models (xComp, PICUS, LandClim and TreeMig) featured very high mortality rates (i.e., well above the Reineke self-thinning line), which compensated for the strong initial overestimation of recruitment (cf.
 Figure \@ref(fig:over)).
-Interestingly, two models that underestimated overall regeneration levels (SIBYLA and Landis-II; cf.
+Interestingly, two models that underestimated overall recruitment levels (SIBYLA and Landis-II; cf.
 Figure \@ref(fig:over)) featured mortality rates that were close to but still above the self-thinning line (Figure \@ref(fig:mortality11)).
@@
 The fact that some models (and observed data) featured lower mortality than expected by Reineke's self-thinning rule may point to facilitation, or simply a higher mortality rate before the trees had reached 7 cm diameter.
-Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher regeneration mortality compared to even-aged, single-species forests.
-Some models yielded a mortality rate of regeneration that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).
+Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher recruitment mortality compared to even-aged, single-species forests.
+Some models yielded a mortality rate of recruitment that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).
 
-The three models with the largest overestimation of the proportion of regeneration at the 7 cm threshold featured the highest mortality between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A).
-The other models that overestimated regeneration had mortalities lower than the expected self-thinning ratio, with the exception of LandClim (Figure \@ref(fig:mortality13)B).
-It is noteworthy that most of the models that featured a low ratio of regeneration between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small overestimation of regeneration at the 7 cm threshold (Figure \@ref(fig:mortality11)).
-aDGVM2 was the only model that underestimated regeneration at the 7 cm threshold while featuring a low mortality between the two thresholds (Figure \@ref(fig:mortality13)C), and only a few models were close to the correct regeneration levels at 7 cm while also being close to the Reineke line (SIBYLA, Landis-II; cf.
+The three models with the largest overestimation of the proportion of recruitment at the 7 cm threshold featured the highest mortality between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A).
+The other models that overestimated recruitment had mortalities lower than the expected self-thinning ratio, with the exception of LandClim (Figure \@ref(fig:mortality13)B).
+It is noteworthy that most of the models that featured a low ratio of regeneration between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small overestimation of recruitment at the 7 cm threshold (Figure \@ref(fig:mortality11)).
+aDGVM2 was the only model that underestimated recruitment at the 7 cm threshold while featuring a low mortality between the two thresholds (Figure \@ref(fig:mortality13)C), and only a few models were close to the correct recruitment levels at 7 cm while also being close to the Reineke line (SIBYLA, Landis-II; cf.
 Figure \@ref(fig:mortality13)D; and 4C as well as LandClim).
 
-Lastly, some models that had a low number of individuals at the 7 cm threshold also had a smaller decrease of regeneration between 7 and 10 cm (ratio 7-10 cm) (Figure S5).
+Lastly, some models that had a low number of individuals at the 7 cm threshold also had a smaller decrease of recruitment between 7 and 10 cm (ratio 7-10 cm) (Figure S5).
 In these models, competition and self-thinning are either not pronounced or must have occurred before the trees had reached 7 cm.
 However, this pattern was not consistent across models (Figure S5).
-It is noteworthy that there is a relationship between the regeneration levels at 7 cm and the mortality rate (Table S4): most of the models showed a positive effect (higher regeneration at 7 cm is coupled to an increase of the mortality between 7 and 10 cm).
+It is noteworthy that there is a relationship between the recruitment levels at 7 cm and the mortality rate (Table S4): most of the models showed a positive effect (higher regeneration at 7 cm is coupled to an increase of the mortality between 7 and 10 cm).
 While some models showed a negative effect, the linear trend was not pronounced and not always significant regarding the slope.
@@
 Therefore it was not possible to differentiate in terms of performance between models type (empirical or process based), or their scales (stand, landscape or global) (Table S5).
-The values of regeneration species diversity simulated by each model were also assessed against the observed data to see if there was a difference between models that include a feedback compared to those that did not; all the models except ForCEEPs(f) and FORMIND had significant differences (Table S6).
+The values of recruited species diversity simulated by each model were also assessed against the observed data to see if there was a difference between models that include a feedback compared to those that did not; all the models except ForCEEPs(f) and FORMIND had significant differences (Table S6).
 It is noteworthy that for the model ForCEEPS the pattern changed from the overprediction of species diversity in the regeneration to a diversity level that is closer to observations when the recruitment module included feedback (ForCEEPs(f)).
@@
 
-When evaluating regeneration levels along key gradients of light availability (basal area), temperature (degree-days) and soil moisture (climatic water balance), distinct features emerged: the models reproduced the effect of basal area in both its magnitude and patterns (Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better than the effects of the climatic gradients, where they featured varying patterns.
+When evaluating recruited levels along key gradients of light availability (basal area), temperature (degree-days) and soil moisture (climatic water balance), distinct features emerged: the models reproduced the effect of basal area in both its magnitude and patterns (Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better than the effects of the climatic gradients, where they featured varying patterns.
 
-In the observed data, regeneration levels decreased clearly with increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a).
+In the observed data, recruited levels decreased clearly with increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a).
 This trend was captured by the Landis-II model only, albeit at much lower values of basal area than in the observations.
-The other models featured distinctly different trends, such as (1) an increase in regeneration levels with increasing total basal area followed by the absence of regeneration at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant regeneration levels with basal area (ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of regeneration up to a certain value of stand basal area followed by a decrease at even higher values of stand basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).
+The other models featured distinctly different trends, such as (1) an increase in recruited levels with increasing total basal area followed by the absence of regeneration at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant recruitment levels with basal area (ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of recruitment up to a certain value of stand basal area followed by a decrease at even higher values of stand basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).
 
-The observed regeneration did not change much across the climatic water balance gradient, and showed a slight increase with temperature.
-The models did not match the observed regeneration patterns across the gradients of soil moisture and temperature (Figure \@ref(fig:nvTrend710BAlightTemp)b,c), and neither was there a clear pattern across models.
-Regarding the soil moisture gradient, a group of models featured decreasing regeneration with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS).
-Across the temperature gradient, all the models that featured a decrease in regeneration levels with increasing soil moisture showed the opposite trend (i.e., more regeneration with increasing temperature), except for 4C.
+The observed recruitment did not change much across the climatic water balance gradient, and showed a slight increase with temperature.
+The models did not match the observed reruitment patterns across the gradients of soil moisture and temperature (Figure \@ref(fig:nvTrend710BAlightTemp)b,c), and neither was there a clear pattern across models.
+Regarding the soil moisture gradient, a group of models featured decreasing recruitment with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS).
+Across the temperature gradient, all the models that featured a decrease in recruitment levels with increasing soil moisture showed the opposite trend (i.e., more recruitment with increasing temperature), except for 4C.
 The other models did not feature a clear pattern.
@@
 Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus sylvestris*) and oak (*Quercus* spp.) were the species observed most frequently in the EuFoRIa data.
-Most models captured well the share of basal area in the regeneration of these five species compared to the observed data along the two major climatic gradients (Figure \@ref(fig:diff), cf.
+Most models captured well the share of basal area in the recruitment of these five species compared to the observed data along the two major climatic gradients (Figure \@ref(fig:diff), cf.
 Figure S8).
@@
 
-First, some models overestimated regeneration across the entire gradients of at least one of the five main tree species (Figure \@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated regeneration of more than one species for a large part of the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
+First, some models overestimated recruitment across the entire gradients of at least one of the five main tree species (Figure \@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated recruitment of more than one species for a large part of the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
 red colors in Figure \@ref(fig:diff)).
-These trends were consistent for both regeneration thresholds, although the exact changes in the share of the regeneration basal area were different (Figure S8).
+These trends were consistent for both recruitment thresholds, although the exact changes in the share of the recruited basal area were different (Figure S8).
 
 Second, there was some congruence in the simulation results by species across the models.
-The share in the regeneration of Silver fir (*Abies alba*) was captured well across the gradient by most models except for xComp, PICUS, and LandClim, which overestimated the regeneration share of this species.
+The share in the recruitment of Silver fir (*Abies alba*) was captured well across the gradient by most models except for xComp, PICUS, and LandClim, which overestimated the recruitment share of this species.
 There were also some peculiarities evident for some models, such as 4C that did not simulate *Abies alba*.
-The patterns across models for *Fagus sylvatica* regeneration were more complex, as its regeneration was underestimated by many models across the environmental gradients featuring negligible regeneration at most sites, while others overestimated it in different parts of the environmental space.
-Most models overestimated the share of *Picea abies* regeneration in the cold-wet part of the gradients.
+The patterns across models for *Fagus sylvatica* recruitment were more complex, as its recruitment was underestimated by many models across the environmental gradients featuring negligible regeneration at most sites, while others overestimated it in different parts of the environmental space.
+Most models overestimated the share of *Picea abies* recruitment in the cold-wet part of the gradients.
 *Pinus sylvestris* and *Quercus* spp.
-represented a small share of the regeneration in the observed data mostly at warm-dry sites.
-Many models, however, erroneously featured regeneration for these two species along most of the climatic gradients, although with a low share of basal area.
-Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had almost no regeneration of *Pinus sylvestris* and *Quercus* spp., or did not feature any regeneration of these species at all.
+represented a small share of the recruitment in the observed data mostly at warm-dry sites.
+Many models, however, erroneously featured recruitment for these two species along most of the climatic gradients, although with a low share of basal area.
+Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had almost no recruitment of *Pinus sylvestris* and *Quercus* spp., or did not feature any recruitment of these species at all.
 
-Lastly, there is no model that performed well across the five species regarding the species-specific regeneration levels in the environmental space (Figure \@ref(fig:diff)).
-Some models (e.g., ForClim11, PICUS) tended to consistently overestimate the regeneration share of all five species, but most models overestimated the regeneration share of some species while underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
+Lastly, there is no model that performed well across the five species regarding the species-specific recruitment levels in the environmental space (Figure \@ref(fig:diff)).
+Some models (e.g., ForClim11, PICUS) tended to consistently overestimate the recruitment share of all five species, but most models overestimated the recruitment share of some species while underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
 
@@
 
-The EuFoRIa data [@käber2023] are exceptional, particularly with respect to the number of records (number of sites and repeated measurements), which is essential for capturing a highly 'noisy' process such as tree regeneration.
+The EuFoRIa data [@käber2023] are exceptional, particularly with respect to the number of records (number of sites and repeated measurements on unmanaged forests), which is essential for capturing a highly 'noisy' process such as tree regeneration.
 The use of this dataset for model benchmarking provided novel insights on the ability of state-of-the-art models to accurately simulate regeneration levels, species composition, and mortality in an early stage of tree life, i.e., between a tree diameter of 7 and 10 cm.
@@
 
-Most of the models overestimated tree regeneration levels.
+Most of the models overestimated tree recruitment levels.
 This has potentially far-reaching implications, e.g., regarding biomass (and thus carbon) turnover, with a potential overestimation of the capacity for forest carbon sequestration [@pugh2019; @pan2011].
-Yet, we focused on one specific stage, i.e., regeneration into the 7 and 10 cm diameter classes.
+Yet, we focused on one specific stage, i.e., recruitment into the 7 and 10 cm diameter classes.
 Trees at this size contribute little to carbon sequestration, and if the excess regeneration at this stage is compensated -- in the models -- soon thereafter by higher mortality, overall simulated regeneration may still yield trustworthy results at the level of stand structure and related forest attributes such as basal area or biomass.
@@
 above), their performance is better when it comes to matching the composition of regeneration as a function of abiotic and biotic conditions.
-Most models maintained or even decreased regeneration diversity between the 7 and 10 cm thresholds, and the same is visible from the empirical data, but the differences were not significant.
+Most models maintained or even decreased recruitment diversity between the 7 and 10 cm thresholds, and the same is visible from the empirical data, but the differences were not significant.
 Diversity variations may be explained by the way the individual models consider regeneration processes [@konig_tree_2022], e.g., without considering the feedback from adults, regeneration can be linked to either seed rain from dispersal or a seed/seedling bank; neither process is incorporated across many models studied here, however.
 
-The species diversity of the entire stand was captured better by the models than the diversity of tree regeneration, and only a few models overpredicted stand-level diversity.
+The species diversity of the entire stand was captured better by the models than the diversity of tree recruitment, and only a few models overpredicted stand-level diversity.
 Defining regeneration composition as being proportional to adult abundance, regardless of productivity, might be a simple and conservative assumption to safely maintain relative species abundances [@hanbury-brown2022], although this may be simplistic.
-Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the regeneration than those without feedback, with the exception of ForCEEPS that significantly improved regeneration species diversity with the model version that included feedback (ForCEEPS(f)).
+Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the recruitment than those without feedback, with the exception of ForCEEPS that significantly improved regeneration species diversity with the model version that included feedback (ForCEEPS(f)).
 The similar performance of models with and without feedback is likely because the models put more weight on the regeneration niche arising from abiotic and biotic filters, than from the habitat niche of the adult trees [@grubb1977].
@@
 
-In spite of the critical considerations above, it is remarkable that most models did not deviate exceedingly from the observations with respect to simulated regeneration levels -- after all, few if any of the models' regeneration routines are sufficiently constrained by data, with the exception of the empirical models xComp and SIBYLA.
+In spite of the critical considerations above, it is remarkable that most models did not deviate exceedingly from the observations with respect to simulated recruited levels -- after all, few if any of the models' regeneration routines are sufficiently constrained by data, with the exception of the empirical models xComp and SIBYLA.
 It is noteworthy that in spite of their empirical basis, these two models did not match empirical regeneration levels, in a similar magnitude as the other models and even in opposite directions (SIBYLA: underestimation; xComp: overestimation).
@@
 At the level of the simulated one-hectare samples, average basal area is typically not very low as long as no larger disturbances occur, which was explicitly excluded in our simulation protocol.
-Towards the other end of the spectrum, i.e., with increasing stand basal area, it would be reasonable to expect that the regeneration of the different species would become sparser and drop out entirely under low-light conditions [@klopcic2012; @zell2019].
+Towards the other end of the spectrum, i.e., with increasing stand basal area, it would be reasonable to expect that the recruitment of the different species would become sparser and drop out entirely under low-light conditions [@klopcic2012; @zell2019].
 However, few models showed this trend, thus indicating that the relation between regeneration and light availability is not yet captured correctly in most models.
@@
 
-The EuFoRIa data as used here are unique and well suited to better understand tree regeneration.
+The EuFoRIa data as used here are well suited to better understand tree regeneration.
 However, three aspects of these data may represent considerable limitations.
@@
 Morin, W.
-Rammer and D.
-Scherrer H.
+Rammer, D.
+Scherrer and H.
 Bugmann.
@@
 
-**Table 1.** Models included, their characteristics and simulation strategies used in this study.
-Where Scale refers to the models scale, Type to the models type, Population structure to the model population approach, Feedback to the inclusion or not of adult tree feedback to the level and species composition of regeneration, Approach to the model regeneration module approach, Start from to the stage from which trees are recruited, Complexity, to the mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), Species to which species from the original eleven species were simulated , Runtime (spin up) [years] to the number of years used per model in their simulations for the spin up period and regeneration sampling; Climate data type (use of years) to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details.
+**Table 1.** Models included in this study, their characteristics and simulation strategies used.
+Scale refers to the models scale, Type to the models type, Population structure to the model population approach, Feedback to the inclusion or not of adult tree feedback to the level and species composition of regeneration, Approach to the model regeneration module approach, Start from to the stage from which trees are recruited, Complexity, to the mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), Species to which species from the original eleven species were simulated, Runtime (spin up) [years] to the number of years used per model in their simulations for the spin up period and regeneration sampling; Climate data type (use of years) to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details.
 Further information can be found on the individual model reports available in the supplementary materials.

### 2023-10-10 manuscript/review/authorsResponse.qmd
more updates in the revision before sending to co-authors

@@
 
+Thank you for taking the time to review our manuscript and for your comments.
 Our main take-away from this manuscript is that while it would certainly be possible to calibrate all 15 models against the EuFoRIa data, it is important to ensure that this aligns with the study's goals.
@@
 
-Our general comments to reviewer 1:
+**Our general comments to reviewer 1:**
 
@@
 
-[PENDING: I am going to do this.]{.underline}
+[Pending]{.underline}
 
@@
 
-[Clean together with comment 2]{.underline}
+[Clean together with comment 3]{.underline}
 
@@
 
-In this first assessment we were not aiming at identify what specific subcomponents may hold the key and therefore dased on the current approach we could not provide more concrete recommendation for model development which most likely be specific for each of the 15 models.
+In this first assessment we were not aiming at identify what specific subcomponents may hold the key and therefore based on the current approach we could not provide more concrete recommendation for model development which most likely be specific for each of the 15 models.
 Maybe we need more such studies, including calibration as a first step, and then application against a withheld part of the data; or application (and comparison) under climate change scenarios to evaluate the robustness of the signal under future conditions.
@@
 
-### [COMMENT 13]{.underline}
+### COMMENT 13
 
@@
 
-### **COMMENT 18**
+### COMMENT 18
 
@@
 
-### **COMMENT 19**
+### COMMENT 19
 
@@
 
-### [**COMMENT 20**]{.underline}
+### [COMMENT 20]{.underline}
 
@@
 
-Response to comment 22:
+**Response to comment 22:**
 
@@
 
-### **COMMENT 23**
+### COMMENT 21
 
@@
 
-Response to comment 22:
+Response to comment 21:
 
@@
 
-### **COMMENT 24**
+### COMMENT 22
 
@@
 
-Response to comment 24:
+**Response to comment 22:**
 
@@
 
-### **COMMENT 26**
+### COMMENT 23
 
@@
 
-Response to comment 26:
+**Response to comment 23:**
 
@@
 
-### **COMMENT 27**
+### COMMENT 24
 
@@
 
-Response to comment 27:
+**Response to comment 24:**
 
@@
 
-### **COMMENT 28**
+### COMMENT 25
 
@@
 
-Response to comment 28:
+**Response to comment 25:**
 
@@
 
-### **COMMENT 29**
+### COMMENT 26
 
@@
 
-Response to comment 29:
+**Response to comment 26:**
 
@@
 
-### **COMMENT 30**
+### COMMENT 27
 
@@
 
-Response to comment 30:
+**Response to comment 27:**
 
@@
 
-### **COMMENT 31**
+### COMMENT 28
 
@@
 
-Response to comment 31:
-
-Please, see our reply on comment 25
-
-### **COMMENT 32**
+**Response to comment 28:**
+
+We have changed this text L\[XX-XX\]
+
+### COMMENT 29
 
@@
 
-Response to comment 32:
+**Response to comment 32:**
 
@@
 
-### **COMMENT 33**
+### COMMENT 30
 
@@
 
-Response to comment 33:
+**Response to comment 30:**
 
@@
 
-### [**COMMENT 34**]{.underline}
+### COMMENT 31
 
@@
 
-Response to comment 34:
+**Response to comment 31:**
 
@@
 
-### **COMMENT 35**
+### COMMENT 32
 
@@
 
-Response to comment 35:
+**Response to comment 32:**
 
@@
 
-### **COMMENT 36**
+### COMMENT 33
 
@@
 
-Response to comment 36:
+**Response to comment 33:**
 
@@
 
-### **COMMENT 37**
+### COMMENT 34
 
@@
 
-Response to comment 37:
+**Response to comment 34:**
 
@@
 
-### **COMMENT 38**
+### COMMENT 35
 
@@
 
-Response to comment 38:
+**Response to comment 35:**
 
@@
 
-**COMMENT 39**
+### COMMENT 36
 
@@
 
-Response to comment 39:
+**Response to comment 36:**
 
@@
 
-### **COMMENT 40**
+### **CO**MMENT 37
 
@@
 
-Response to comment 40:
+**Response to comment 37:**
 
@@
 
-### **COMMENT 41**
+### COMMENT 38
 
@@
 
-### **COMMENT 42**
+### COMMENT 39
 
@@
 
-### **COMMENT 43**
+### COMMENT 40
 
@@
 
-### **COMMENT 44**
+### COMMENT 41
 
@@
 
-Response to comment 44:
+Response to comment 41:
 
@@
 
-### **COMMENT 45**
+### COMMENT 42
 
@@
 
-Response to comment 45:
+**Response to comment 42:**
 
@@
 
-### **COMMENT 46**
+### COMMENT 43
 
@@
 
-Response to comment 46:
+**Response to comment 43:**
 
@@
 We have also reported later on in this section (but before this line) that each of the models reported the number of trees crossing the two diameter thresholds.
-
-### COMMENT 47
+L\[XX-XX\]
+
+### COMMENT 44
 
@@
 
-Response to comment 47:
-
-Following the reviewer recomendation we have changed the equation formulation to better represent the Reineke relationship.
+**Response to comment 44:**
+
+Following the reviewer recommendation we have changed the equation formulation to better represent the Reineke relationship.
 The reviewer is correct that the Reineke is applied to all stems in the stand, in our case the N value is only changing due to the available regeneration at the two diameter thresholds as the adult trees number is the same.
@@
 
-### COMMENT 50
+### COMMENT 45
 
@@
 
-Response to comment 50:
+**Response to comment 45:**
 
@@
 
+### COMMENT 46
+
+L319-321 - overly complex sentence structure
+
+**Response to comment 46:**
+
+We have improved this sentence L\[XX-XX\]
+
+"We tested for significant differences between simulated and observed values of the recruitment levels and species diversity by using paired t-tests."
+
+### COMMENT 47
+
+L323-324 presumably, recruits make up a very small proportion of stand basal if the models have been run to equilibrium?
+
+**Response to comment 47:**
+
+Yes the reviewer is correct, the recruits represent a small proportion of the total stand basal area.
+
+### COMMENT 48
+
+L327 of each species?
+
+**Response to comment 48:**
+
+Added as suggested.
+
+### COMMENT 49
+
+L332-335 This give a matric of species-level variation in recruitment across sites, but how does this relate to the regeneration niche concept of Grubb 1977?
+
+**Response to comment 49:**
+
+We have changed the text to clarify this point. L\[XX-XX\]
+
+"For these species, we calculated the share in the recruited basal area per site, where each site represents a part of the environmental space, as the mean across the available samples per site:"
+
+### COMMENT 50
+
+L496 "confronted with a unique dataset from unmanaged"
+
+L498 "The EuFoRIa data (Käber et al. 2023) are exceptional, particularly"
+
+I think there are plenty of other datasets that could be used and have better information on smaller trees than EuFoRIa,  so suggest a more nuanced sales pitch here.
+
+**Response to comment 50:**
+
+We have changed the text L\[XX-XX\]
+
+###
+
 ### COMMENT 51
-
-L496 "confronted with a unique dataset from unmanaged"
-
-L498 "The EuFoRIa data (Käber et al. 2023) are exceptional, particularly"
-
-I think there are plenty of other datasets that could be used and have better information on smaller trees than EuFoRIa,  so suggest a more nuanced sales pitch here.
-
-Response to comment 51:
-
-We have changed the text.
-
-###
-
-### COMMENT 52
 
@@
 
-Response to comment 52:
+**Response to comment 51:**
 
@@
 
-### COMMENT 53
+### COMMENT 52
 
@@
 
-Response to comment 53:
+Response to comment 52:
 
@@
 
+### COMMENT 53
+
+L552 "there is no evidence that models with feedback from the canopy captured better the species" could you explain what you mean by "feedback" here?
+
+**Response to comment 53:**
+
+We have modified this sentence to clarify what do we mean here with feedback: " Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the regeneration than those without feedback, "
+
+This had also been defined in the material and methods where we stated that "i.e., where the level and species composition of regeneration is influenced by the existing adult tree community via the production of seeds, seedlings or saplings, from models that do not contain such a feedback."
+
+###
+
 ### COMMENT 54
-
-L552 "there is no evidence that models with feedback from the canopy captured better the species" could you explain what you mean by "feedback" here?
-
-Response to comment 54:
-
-We have modified this sentence to clarify what do we mean here with feedback: " Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the regeneration than those without feedback, "
-
-This had also been defined in the material and methods where we stated that "i.e., where the level and species composition of regeneration is influenced by the existing adult tree community via the production of seeds, seedlings or saplings, from models that do not contain such a feedback."
-
-###
-
-### COMMENT 55
 
@@
 
-Response to comment 55:
+**Response to comment 54:**
 
@@
 
-### COMMENT 56
+### COMMENT 55
 
@@
 
-Response to comment 56:
+**Response to comment 55:**
 
@@
 
-### [COMMENT 57]{.underline}
+### COMMENT 56
 
@@
 
+Response to comment 56:
+
+We have decided not to include this aspect in the abstract due to lack of space.
+
+###
+
+### COMMENT 57
+
+L656 "Yet, the global models should not be at a disadvantage due to the limited spatial consideration" Don't quite follow your point here. what does "at a disadvantage" mean?
+
 Response to comment 57:
 
-[Should we add here:]{.underline}
-
-["Using 15 models built to capture long-term forest dynamics at the stand, landscape, and global levels, we simulate tree regeneration at 200 sites representing large environmental gradients across Central Europe WITHOUT DISTURBANCES CONSIDERATION" ?
-or something like this?]{.underline}
-
-###
+In this sentence we are discussing how the sampling strategy presented in the study protocol might have represented a disadvantage for landscape level models due to the limited spatial scale, but how this might not be the same for Global level models as they usually lack dispersal between cells.
 
 ### COMMENT 58
-
-L656 "Yet, the global models should not be at a disadvantage due to the limited spatial consideration" Don't quite follow your point here. what does "at a disadvantage" mean?
-
-Response to comment 58:
-
-In this sentence we are discussing how the sampling strategy presented in the study protocol might have represented a disadvantage for landscape level models due to the limited spatial scale, but how this might not be the same for Global level models as they usually lack dispersal between cells.
-
-### COMMENT 59
 
@@
 
-### COMMENT 60
+### COMMENT 59
 
@@
 
+**Response to comment 59:**
+
+We agree with the reviewer that a more systematic kind of inventory could be useful to detect regeneration patterns under management or disturbances impacts.
+
+###
+
+### COMMENT 60
+
+L696 "Exercises like the one presented here, where the models are operated in"blind flight" mode." Definition of blind flight "To do something based on guesswork, intuition, or without any help or instructions"....is that what you mean?
+
 Response to comment 60:
 
-We agree with the reviewer that a more systematic kind of inventory could be useful to detect regeneration patterns under management or disturbances impacts.
-
-###
+With blind flight we refer to doing the simulations without having important information about the expected regeneration values.
 
 ### COMMENT 61
-
-L696 "Exercises like the one presented here, where the models are operated in"blind flight" mode." Definition of blind flight "To do something based on guesswork, intuition, or without any help or instructions"....is that what you mean?
-
-Response to comment 60:
-
-With blind flight we refer to doing the simulations without having important information about the expected regeneration values.
-
-### COMMENT 62
 
@@
 
-Response to comment 62:
+**Response to comment 61:**
 
@@
 
+### COMMENT 62
+
+L831 Add publisher
+
+L939 Correct doi
+
+**Response to comment 62:**
+
+Changed as suggested.
+
 ### COMMENT 63
 
-L831 Add publisher
-
-L939 Correct doi
-
-Response to comment 63:
+L1041 suggest sticking with "ingrowth" (as used in the figure) or "recruitment"
+
+**Response to comment 63:**
+
+We have changed the terminology.
+
+### COMMENT 64
+
+L1050 Overperdicted
+
+**Response to comment 64:**
 
@@
 
-### COMMENT 64
-
-L1041 suggest sticking with "ingrowth" (as used in the figure) or "recruitment"
-
-Response to comment 65:
-
-We have changed the terminology.
-
-### COMMENT 65
-
-L1050 Overperdicted
-
-Response to comment 65:
-
-Changed as suggested.
-
 ​

### 2023-10-10 code/figures.R

@@ -8,7 +8,7 @@
 
 # Load plotting styles 
 
-    source("code/plottingManuscript.R")
+overDTsim    source("code/plottingManuscript.R")
 
 # Load data ----
 
@@ -220,7 +220,7 @@
     
     sigH7_10 <- ggpubr::compare_means(ShannonIndexRecruit ~ dbh, 
                                       data = dat710[!dat710$model == "aDGVM2", ],
-                          group.by = "model", method = "t.test")
+                                      group.by = "model", method = "t.test")
     
     write.csv(sigH7_10,"figures/sigH7_10.csv", row.names = FALSE)
     
@@ -454,7 +454,7 @@
                                                   label = label)) +
         ggplot2::scale_color_manual(values = c("7" = "#FDAE61", "10" = "grey"),
                                     name = "Diameter threshold",
-                                    labels = c("7" = '7 cm', "10" = '10 cm'))+
+                                    labels = c("7" = '7 cm', "10" = '10 cm')) +
       ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 10)))
     
     ggplot2::ggsave("figures/allAgesRichness.png",
@@ -505,7 +505,8 @@
                            show.legend = FALSE) +
         ggplot2::scale_color_manual(values = c("7" = "#FDAE61", "10" = "grey"),
                                     name = "Diameter threshold",
-                                    labels = c("7" = '7 cm', "10" = '10 cm')) 
+                                    labels = c("7" = '7 cm', "10" = '10 cm')) +
+      ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 10)))
 
     
     ggplot2::ggsave("figures/recruitmentRichness.png",
@@ -821,7 +822,7 @@
                                     ref.group = "Observed", method = "t.test")
     write.csv(rtreesDiff ,"figures/rtreesDiff.csv", row.names = FALSE)
     
-    
+   
 ### feedback / not feedback H simulated vs observed -----  
 
     dat710 <- ShannonIndex |>  
@@ -835,6 +836,10 @@
                                         data = dat7H[!dat7H$model == "aDGVM2", ], 
                                         ref.group = "Observed", method = "t.test")
     write.csv(rtreesDiff, "figures/sigH7_SimObs.csv", row.names = FALSE)  
+    
+    
+    
+    
     
     
 # Recruitment niche-----

### 2023-10-19 manuscript/manuscriptRW.Rmd
changes in manuscript before removing figures

@@
       $^2$Senckenberg Biodiversity and Climate Research Centre (SBiK-F), 60325 Frankfurt am Main, Germany. \newline
-      $^3$Department of Computational Hydrosystem, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline
+      $^3$Department of Ecological Modelling, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline
       $^4$Ecosystem Dynamics and Forest Management Group, TUM School of Life Sciences, Technical University of Munich, 85354 Freising, Germany. \newline
@@
        $^9$Department of Soil Sciences, of Plants and Food, University of Bari Aldo Moro, Italy. \newline
-       $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen \newline
+       $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen, Slovakia \newline
        $^{11}$Czech University of Life Sciences Prague, Faculty of Forestry and Wood Sciences, Kamýcká 129, 16500 Praha 6 – Suchdol, Czech Republic. \newline
        $^{12}$Department of Biodiversity of Ecosystems and Landscape, Slovak Academy of Sciences, Štefánikova 3, P.O.Box 25, 81499 Bratislava, Slovakia. \newline 
-       $^{13}$Dpt. of Soil and Climate, Bavarian State Institute of Forestry. \newline 
+       $^{13}$Dpt. of Soil and Climate, Bavarian State Institute of Forestry, Germany. \newline 
        $^{14}$Institute for Alpine Environment, Eurac Research, Bolzano/Bozen, Italy. \newline 
@@
     keep_tex: yes
-    number_sections: yes
+    number_sections: no
     toc: no
@@
 
-**OPEN RESEARCH STATEMENT:** This submission uses novel code, both the data and code are provided in an external repository to be evaluated during the peer review process.
-The repository can be downloaded from [this figshare private link: https://figshare.com/s/100857ba67d2351dcb15](https://figshare.com/s/100857ba67d2351dcb15).
-The protocol is available at @díaz-yáñez2022 ([with link: https://osf.io/czdxp/](https://osf.io/czdxp/)).
-We will make publicly available the code and data if the manuscript is accepted.
-
-**Keywords:** tree regeneration \| tree recruitment \| tree establishment \| forest dynamics \| models \| landscape \| stand \| global
+**OPEN RESEARCH STATEMENT:** This submission uses novel code, both the
+data and code are provided in an external repository to be evaluated
+during the peer review process. The repository can be downloaded from
+[this figshare private link:
+https://figshare.com/s/100857ba67d2351dcb15](https://figshare.com/s/100857ba67d2351dcb15).
+The protocol is available at @díaz-yáñez2022 ([with link:
+https://osf.io/czdxp/](https://osf.io/czdxp/)). We will make publicly
+available the code and data if the manuscript is accepted.
+
+**Keywords:** tree regeneration \| tree recruitment \| tree
+establishment \| forest dynamics \| models \| landscape \| stand \|
+global
 
@@
 
-Tree regeneration is a key process in forest dynamics, particularly in the context of forest resilience and climate change.
-Models are pivotal for assessing long-term forest dynamics, and they have been in use for more than 50 years.
-However, there is a need to evaluate their capacity to accurately represent tree regeneration.
-We assess how well current models capture the overall abundance, species composition, and mortality of tree regeneration.
-
-Using 15 models built to capture long-term forest dynamics at the stand, landscape, and global levels, we simulate tree regeneration at 200 sites representing large environmental gradients across Central Europe.
-The results are evaluated against extensive data from unmanaged forests.
-
-Most of the models overestimate recruitment levels, which is only compensated in some models by high simulated mortality rates in the early stages of individual trees dynamics.
-Simulated species diversity of recruitment matches the observed ranges.
-Models simulating higher species diversity at the stand level do not present higher recruitment diversity.
-The effect of light availability on recruitment levels is captured better than the effect of temperature and soil moisture, but patterns are not consistent across models.
-Increasing complexity in the tree regeneration modules of the models is not related to higher accuracy of simulated tree recruitment.
-Furthermore, individual model design is more important than scale (stand, landscape, global) and approach (empirical, process-based) for accurately capturing tree regeneration.
-
-Despite the mismatches between simulation results and data, it is remarkable that most models capture the essential features of the highly complex process of tree regeneration, while not having been parameterized with such data.
-We conclude that much can be gained by evaluating and refining the modeling of regeneration processes.
-This has the potential to render long-term projections of forest dynamics under changing environmental conditions that are much more robust.
+Tree regeneration is a key process in forest dynamics, particularly in
+the context of forest resilience and climate change. Models are pivotal
+for assessing long-term forest dynamics, and they have been in use for
+more than 50 years. However, there is a need to evaluate their capacity
+to accurately represent tree regeneration. We assess how well current
+models capture the overall abundance, species composition, and mortality
+of tree regeneration.
+
+Using 15 models built to capture long-term forest dynamics at the stand,
+landscape, and global levels, we simulate tree regeneration at 200 sites
+representing large environmental gradients across Central Europe. The
+results are evaluated against extensive data from unmanaged forests.
+
+Most of the models overestimate recruitment levels, which is only
+compensated in some models by high simulated mortality rates in the
+early stages of individual trees dynamics. Simulated species diversity
+of recruitment matches the observed ranges. Models simulating higher
+species diversity at the stand level do not present higher species
+diversity in the recruitment layer. The effect of light availability on
+recruitment levels is captured better than the effect of temperature and
+soil moisture, but patterns are not consistent across models. Increasing
+complexity in the tree regeneration modules of the models is not related
+to higher accuracy of simulated tree recruitment. Furthermore,
+individual model design is more important than scale (stand, landscape,
+global) and approach (empirical, process-based) for accurately capturing
+tree regeneration.
+
+Despite the mismatches between simulation results and data, it is
+remarkable that most models capture the essential features of the highly
+complex process of tree regeneration, while not having been
+parameterized with such data. We conclude that much can be gained by
+evaluating and refining the modeling of regeneration processes. This has
+the potential to render long-term projections of forest dynamics under
+changing environmental conditions that are much more robust.
 
@@
 
-Forests provide a wide range of products and services of vital importance to humankind [@foresteurope2020].
-Under the pressure of climate change, increasing disturbance impacts and changing societal demands on forest ecosystem services, it is becoming ever more important to understand how forest structure, composition and function will change, and to evaluate forest capacity to adapt to or be resilient after disturbance [@lindner2010; @seidl2022].
-A wide range of models of forest dynamics have been developed over the past decades considering the impacts of climate [@bugmann2022].
-From these studies, it is evident that we have a reasonably good understanding how to model tree growth [@vanclay1997; @bugmann1996], and substantial efforts have been dedicated to improving the representation of tree mortality [@bugmann2019; @cailleret2017].
-In contrast, tree regeneration is a much less studied process that is often represented rather coarsely in models [@walck2011; @leishman1992; @price2001].
-This presents an important research gap, particularly in the context of climate-induced forest disturbances and forest resilience.
-
-Tree regeneration arises from multiple processes including pollination, fruit maturation, seed production, dispersal, germination, juvenile growth and survival [@price2001; @vacchiano2018].
-All these processes are difficult to assess, and some of them are scarcely understood and thus appear highly stochastic [@bogdziewicz2021].
-Identifying the appropriate level of complexity for the mathematical formulation of the key factors that are leading to successful tree regeneration is challenging.
-Currently, tree regeneration processes in dynamic forest models are handled in a multitude of ways [@konig_tree_2022; @bugmann2022]: from 1) entirely ignoring them [as done in classical forest growth models, e.g., Pretzsch et al.
--@pretzsch2002], across 2) the use of a few simple environmental filters, as done in most forest "gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g., @hickler2012; @smith2001], to 3) complex approaches that incorporate local feedback from the canopy, multiple ecological processes and often also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based statistical parameterizations, which however are not easy to extrapolate in space and time [e.g., @ribbens1994].
-
-Overall, models are needed to (1) synthesize existing empirical data and explore their relationships, (2) assess future tree regeneration e.g.
-in the context of global change scenarios, and (3) identify the most important processes that are shaping ecological patterns.
-Given the current strategies that are used in models of forest dynamics to represent tree regeneration, model behavior often is prone to problems, such as very high levels of tree regeneration that necessitate excess mortality at early stages of tree life to simulate correct stand structure and composition.
-Also, inadequately high species diversity in tree regeneration may be simulated, which is characteristic of "classical" forest gap models [@gutiérrez2016], at least as long as the simulation set-up comprises a multitude of species.
-Some models use calibration against local canopy-level data to constrain simulated regeneration levels, which is likely to hamper the general applicability of these models e.g.
-under scenarios of climate change.
-Furthermore, correctly capturing the species composition in tree regeneration as a function of the presence of seed trees in the canopy is often a particular challenge, potentially leading to unrealistic successional drift in the model, which must be corrected by factors that are hard to parameterize [e.g., @lischke2006].
-A related issue is the excessive reduction of species diversity due to positive feedback effects, such that eventually just single-species stands remain.
-This is sometimes corrected by the incorporation of a low level of seed influx of all species at all times [@schumacher2006] or by restricting the number of seeds per species in the seedbank [@lischke2006].
-However, simulated species composition is usually exceedingly sensitive to assumptions about seed availability, whereas the parameters of such functions are poorly constrained by field data.
-Lastly, there is often a problem with insufficient observational constraints on parameter values for models that start from very small tree sizes (e.g., 10 cm of height or even seeds) and track tree development in a process-oriented manner by considering a multitude of ecological influences, rather than emphasizing tree regeneration into a larger size class [e.g., @wehrli2006].
-
-Modeling tree regeneration processes is challenging, and even in empirical ecology, it has not received much attention [@hanbury-brown2022], although many sub-processes have been studied in detail [@miina2006].
-Yet, little data are available that covers all the processes within one species along environmental gradients, let alone for a vast suite of species.
-Data on forest regeneration data are often fragmented, which constitutes a major problem for model building [@clark1999a].
-For example, monitoring on permanent plots (such as National Forest Inventories) often measures tree regeneration [@zell2019], however with a design that captures the rate after the stand initiation phase [@hallsby2015].
-Data from permanent plots, although available across regions, are also highly heterogeneous, rendering their use difficult in a modeling context [@käber2021], and targeted studies to measure tree regeneration on experimental sites are often limited in spatial extent, species studied, or the sub-set of processes that are targeted [@berdanier2016; @collet2006].
-
-Thus, a focus on the modeling of tree regeneration processes is sorely needed and overdue [@price2001; @walck2011] if we are to make reliable projections of future forest dynamics, i.e., when the models need to be operated in extrapolation mode, as well as from a fundamental ecological point of view for increasing systems understanding.
-In the present study, for the first time, a large number of forest models commonly used to assess forest dynamics under climate change are evaluated against a continental-scale, multi-species harmonized dataset on tree recruitment [@käber2023].
-By tree recruitment, we refer to the passing of trees across a specific diameter threshold ("ingrowth").
-We included models that are based on a range of "philosophies", from models operating at the stand to the global scale as well as the range of models from empirically derived to "process-based" [e.g., @bugmann1996; @fabrika2012; @lexer2001; @reyer2014].
-Due to the large variability in tree regeneration patterns in nature and the large number of factors driving and constraining this process --- including some that are not incorporated explicitly in most models, such as deer browsing --- we do not aim for a detailed statistical evaluation of each model.
-Instead, we aim to evaluate the general regeneration patterns and magnitudes simulated by the models, and benchmark the simulated regeneration niche of multiple species against empirical data along a wide environmental gradient of temperature, moisture, and light availability [@grubb1977].
-
-More specifically, we aim to answer the following questions: (1) Are models of forest dynamics capturing accurately tree recruitment levels, initial tree species diversity and mortality in the recruitment?
-(2) Do model traits explain differences in model performance?
-(3) How well do the models capture total recruitment and individual species regeneration niches across environmental gradients of light availability, temperature, and soil moisture?
-We evaluate the performance of regeneration models by analyzing simulated data alongside observed data.
-Specifically, we examine tree species diversity, mortality rates in tree recruitment, and recruitment levels.
-Then, we link the model's traits and performance to define regeneration and we assess the recruitment gradients on regeneration niches.
-Finally, we provide research recommendations for next steps towards improving tree regeneration modules.
+Forests provide a wide range of products and services of vital
+importance to humankind [@foresteurope2020]. Under the pressure of
+climate change, increasing disturbance impacts and changing societal
+demands on forest ecosystem services, it is becoming ever more important
+to understand how forest structure, composition and function will
+change, and to evaluate forest capacity to adapt to or be resilient
+after disturbance [@lindner2010; @seidl2022]. A wide range of models of
+forest dynamics have been developed over the past decades considering
+the impacts of climate [@bugmann2022]. From these studies, it is evident
+that we have a reasonably good understanding how to model tree growth
+[@vanclay1997; @bugmann1996], and substantial efforts have been
+dedicated to improving the representation of tree mortality
+[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is a much
+less studied process that is often represented rather coarsely in models
+[@walck2011; @leishman1992; @price2001]. This presents an important
+research gap, particularly in the context of climate-induced forest
+disturbances and forest resilience.
+
+Tree regeneration arises from multiple processes including pollination,
+fruit maturation, seed production, dispersal, germination, juvenile
+growth and survival [@price2001; @vacchiano2018]. All these processes
+are difficult to assess, and some of them are scarcely understood and
+thus appear highly stochastic [@bogdziewicz2021]. Identifying the
+appropriate level of complexity for the mathematical formulation of the
+key factors that are leading to successful tree regeneration is
+challenging. Currently, tree regeneration processes in dynamic forest
+models are handled in a multitude of ways [@konig_tree_2022;
+@bugmann2022]: from 1) entirely ignoring them [as done in classical
+forest growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2)
+the use of a few simple environmental filters, as done in most forest
+"gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
+@hickler2012; @smith2001], to 3) complex approaches that incorporate
+local feedback from the canopy, multiple ecological processes and often
+also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
+statistical parameterizations, which however are not easy to extrapolate
+in space and time [e.g., @ribbens1994].
+
+Overall, models are needed to (1) synthesize existing empirical data and
+explore their relationships, (2) assess future tree regeneration e.g. in
+the context of global change scenarios, and (3) identify the most
+important processes that are shaping ecological patterns. Given the
+current strategies that are used in models of forest dynamics to
+represent tree regeneration, model behavior often is prone to problems,
+such as very high levels of tree regeneration that necessitate excess
+mortality at early stages of tree life to simulate correct stand
+structure and composition [@kroiss2015]. Also, inadequately high species
+diversity in tree regeneration may be simulated, which is characteristic
+of "classical" forest gap models [@gutiérrez2016], at least as long as
+the simulation set-up comprises a multitude of species. Some models use
+calibration against local canopy-level data to constrain simulated
+regeneration levels, which is likely to hamper the general applicability
+of these models e.g. under scenarios of climate change. Furthermore,
+correctly capturing the species composition in tree regeneration as a
+function of the presence of seed trees in the canopy is often a
+particular challenge, potentially leading to unrealistic successional
+drift in the model, which must be corrected by factors that are hard to
+parameterize [e.g., @lischke2006]. A related issue is the excessive
+reduction of species diversity due to positive feedback effects, such
+that eventually just single-species stands remain [@meier2011]. This is
+sometimes corrected by the incorporation of a low level of seed influx
+of all species at all times [@schumacher2006] or by restricting the
+number of seeds per species in the seedbank [@lischke2006]. However,
+simulated species composition is usually exceedingly sensitive to
+assumptions about seed availability, whereas the parameters of such
+functions are poorly constrained by field data. Lastly, there is often a
+problem with insufficient observational constraints on parameter values
+for models that start from very small tree sizes (e.g., 10 cm in height
+or even seeds) and track tree development in a process-oriented manner
+by considering a multitude of ecological influences, rather than
+emphasizing tree regeneration into a larger size class [e.g.,
+@wehrli2006].
+
+Modeling tree regeneration processes is challenging, and even in
+empirical ecology, it has not received much attention
+[@hanbury-brown2022], although many sub-processes have been studied in
+detail [@miina2006]. Yet, little data are available that cover all the
+processes within one species along environmental gradients, let alone
+for a vast suite of species. Data on forest regeneration are often
+fragmented, which constitutes a major problem for model building
+[@clark1999a]. For example, monitoring on permanent plots (such as
+National Forest Inventories) often measures tree regeneration
+[@zell2019], however with a design that captures the rate after the
+stand initiation phase [@hallsby2015]. Data from permanent plots,
+although available across regions, are also highly heterogeneous,
+rendering their use difficult in a modeling context [@käber2021], and
+targeted studies to measure tree regeneration on experimental sites are
+often limited in spatial extent, species studied, or the sub-set of
+processes that are targeted [@berdanier2016; @collet2006].
+
+Thus, a focus on the modeling of tree regeneration processes is sorely
+needed and overdue [@price2001; @walck2011] if we are to make reliable
+projections of future forest dynamics, i.e., when the models need to be
+operated in extrapolation mode, as well as from a fundamental ecological
+point of view for increasing systems understanding. In the present
+study, for the first time, a large number of forest models commonly used
+to assess forest dynamics under climate change are evaluated against a
+continental-scale, multi-species harmonized dataset on tree recruitment
+[@käber2023]. By tree recruitment, we refer to the passing of trees
+across a specific diameter threshold ("ingrowth"). We included models
+that are based on a range of "philosophies", from models operating at
+the stand to the global scale as well as the range of models from
+empirically derived to "process-based" [e.g., @bugmann1996;
+@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
+tree regeneration patterns in nature and the large number of factors
+driving and constraining this process --- including some that are not
+incorporated explicitly in most models, such as deer browsing --- we do
+not aim for a detailed statistical evaluation of each model. Instead, we
+aim to evaluate the general regeneration patterns and magnitudes
+simulated by the models, and benchmark the simulated regeneration niche
+of multiple species against empirical data along a wide environmental
+gradient of temperature, moisture, and light availability [@grubb1977].
+
+More specifically, we aim to answer the following questions: (1) Are
+models of forest dynamics capturing accurately tree recruitment levels,
+initial tree species diversity and mortality in the recruitment? (2) Do
+model traits explain differences in model performance? (3) How well do
+the models capture total recruitment and individual species regeneration
+niches across environmental gradients of light availability,
+temperature, and soil moisture? We evaluate the performance of
+regeneration models by analyzing simulated data alongside observed data.
+Specifically, we examine tree species diversity, mortality rates in tree
+recruitment, and recruitment levels. Then, we link model traits (e.g.,
+complexity, scale) to performance, and we assess whether modeled total
+recruitment or individual species regeneration niches (i.e.,
+proportional recruitment along an environmental gradient) align with
+observational data.
 
@@
 
-Fifteen models of forest dynamics (including two models featuring model variants) were used to simulate forest dynamics at the stand, landscape or global scale (Table \@ref(tab:tableModels)).
-The approaches used for model construction and their origin differ strongly, with most of the models featuring a largely "process-based" approach, whereas two models are based on formulations derived from the statistical analysis of inventory data (SIBYLA and xComp).
-Some of the models largely rely on the original approach underlying forest gap models (e.g., ForClim 1), whereas some are based on plant ecophysiological processes (e.g., FORMIND and iLand).
-The different approaches underlying the models have strongly influenced the formulation of tree regeneration processes.
-
-The regeneration models can be differentiated into 'regeneration' and 'recruitment' approaches [@vanclay1994; @konig_tree_2022].
-*Regeneration models* include processes such as flowering and pollination, seed production, seed dispersal, germination, and seedling growth, which ultimately lead to the simulated number of established trees.
-*Recruitment models*, in contrast, introduce a number of new trees with certain characteristics such as biomass or diameter, without explicitly considering earlier development processes.
-We can further distinguish models that feature a feedback in the regeneration, i.e., where the level and species composition of regeneration is influenced by the existing adult tree community via the production of seeds, seedlings or saplings, from models that do not contain such a feedback.
-
-The starting point for tree regeneration in the models differs as well, ranging from seedbank, seed or seedling to sapling (i.e., trees much larger than 10 cm height, often ca.
-2-3 m tall).
-Models that start from seed need to include a larger number of ecological processes such as germination and survival of young seedlings, whereas models that start from saplings have to aggregate via parameterizations several ecological processes that are not treated explicitly.
-This latter approach reduces model complexity but comes at the cost of blurring process representation.
-
-In the models used here, the overall complexity in the regeneration modules varies considerably.
-Following @bugmann2022, we can classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and TreeMig) as having rather high complexity in their regeneration modules (mean regeneration formulation complexity across all processes \>0, Table \@ref(tab:tableModels)).
-Two models, ForClim and ForCEEPS, were used here with two alternative variants of regeneration.
-ForClim variant 1 [@bugmann1996] is based on a regeneration module that adheres closely to the concept introduced by @botkin1972 which is an environmental filter considering species specific thresholds of light availability and climatic variables.
-Whereas ForClim variant 11 is adopting a slightly more complex approach where individual species properties and their relation to the environment are represented gradually along with the relative suitability of different species (i.e., competition among tree species) [@huber2020].
-These two model variants allow us to evaluate a more process-based and complex module (variant 11) against a simple module (variant 1), while the rest of the model structure is identical.
-Similarly, the two ForCEEPS variants allow us to isolate the importance of the canopy feedback (i.e., simulated actual composition and relative abundance of species in the plot) via seed trees for the quantity and quality (e.g., diversity and composition) of simulated regeneration, as one variant includes this feedback whereas the other does not.
+Fifteen models of forest dynamics (including two models featuring model
+variants) were used to simulate forest dynamics at the stand, landscape
+or global scale (Table \@ref(tab:tableModels)). The approaches used for
+model construction and their origin differ strongly, with most of the
+models featuring a largely "process-based" approach, whereas two models
+are based on formulations derived from the statistical analysis of
+inventory data (SIBYLA and xComp). Some of the models largely rely on
+the original approach underlying forest gap models (e.g., ForClim 1),
+whereas some are based on plant ecophysiological processes (e.g.,
+FORMIND and iLand). The different approaches underlying the models have
+strongly influenced the formulation of tree regeneration processes.
+
+The regeneration models can be differentiated into 'regeneration' and
+'recruitment' approaches [@vanclay1994; @konig_tree_2022]. *Regeneration
+models* include processes such as flowering and pollination, seed
+production, seed dispersal, germination, and seedling growth, which
+ultimately lead to the simulated number of established trees.
+*Recruitment models*, in contrast, introduce a number of new trees with
+certain characteristics such as biomass or diameter, without explicitly
+considering earlier development processes. We can further distinguish
+models that feature a feedback in the regeneration, i.e., where the
+level and species composition of regeneration is influenced by the
+existing adult tree community via the production of seeds, seedlings or
+saplings, from models that do not contain such a feedback.
+
+The starting point for tree regeneration in the models differs as well,
+ranging from seedbank, seed or seedling to sapling (i.e., trees much
+larger than 10 cm height, often ca. 2-3 m tall). Models that start from
+seed need to include a larger number of ecological processes such as
+germination and survival of young seedlings, whereas models that start
+from saplings have to aggregate via parameterizations several ecological
+processes that are not treated explicitly. This latter approach reduces
+model complexity but comes at the cost of blurring process
+representation.
+
+In the models used here, the overall complexity in the regeneration
+modules varies considerably. Following @bugmann2022, we can classify
+seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and
+TreeMig) as having rather high complexity in their regeneration modules
+(mean regeneration formulation complexity across all processes \>0,
+Table \@ref(tab:tableModels)). Two models, ForClim and ForCEEPS, were
+used here with two alternative variants of regeneration. ForClim variant
+1 [@bugmann1996] is based on a regeneration module that adheres closely
+to the concept introduced by @botkin1972 which is an environmental
+filter considering species specific thresholds of light availability and
+climatic variables. Whereas ForClim variant 11 is adopting a slightly
+more complex approach where individual species properties and their
+relation to the environment are represented gradually along with the
+relative suitability of different species (i.e., competition among tree
+species) [@huber2020]. These two model variants allow us to evaluate a
+more process-based and complex module (variant 11) against a simple
+module (variant 1), while the rest of the model structure is identical.
+Similarly, the two ForCEEPS variants allow us to isolate the importance
+of the canopy feedback (i.e., simulated actual composition and relative
+abundance of species in the plot) via seed trees for the quantity and
+quality (e.g., diversity and composition) of simulated regeneration, as
+one variant includes this feedback whereas the other does not.
 
@@
 
-Recruitment data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset covering such gradients over a large number of sites, to evaluate how well regeneration is captured.
-The observations used here are derived from a novel network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network [@euforia2019;@käber2023] (Figure \@ref(fig:map)).
-The environmental coverage and the variety of forests represented by this data is unprecedented within Europe.
-These forests have been unmanaged for at least 10 years prior to their designation as reserves, and most of them provide time series of natural forest dynamics over multiple decades and up to ca.
-80 years.
-The census periods range from 3 to 37 years, with an average of 14 years.
-The data provide information at the tree level, thus allowing for the sequential comparison of processes such as individual-tree regeneration and death between the measurements.
-
-We selected 200 sites from this network as the benchmarking dataset for the simulation to be representative of the environmental variation contained in the data.
-This was achieved by applying k-means clustering to define 200 clusters of plots from the original set of 869 plots along the environmental dimensions of temperature, climatic water balance, soil quality, slope and aspect.
-Each of these sites featured at least two consecutive measurements.
-Recruitment minimum size thresholds for these sites differed between stem diameters of 0 and 10 cm.
-For our study, we defined two datasets where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds from 10 cm.
-Therefore we had in total 200 sites with observations of newly established trees above 10 cm and 165 sites above 7cm.
-In the results section, any result from observed data at 7 cm threshold is originated from the 165 sites.
-
-Plot size ranged from 0.02 ha to 5.52 ha, and the data were further processed and aggregated following @käber2023 to provide recruitment rates per ha and per decade.
-The observations used in this study featured 30,900 newly established trees.
-Recruitment rates per site, sample and decade ranged from zero to 1246 trees, with a mean of 56 trees.
-Adult species composition was also available for each plot.
-For more details on this dataset and the detailed regeneration information, cf.
-@käber2023.
+Recruitment data covering a wide range of environmental conditions are
+hard to obtain, and this is one of the reasons why most models of forest
+dynamics have never been confronted with a dataset covering such
+gradients over a large number of sites, to evaluate how well
+regeneration is captured. The observations used here are derived from a
+novel network of sites in forest reserves that represent the range of
+environmental gradients in temperature and precipitation in Central
+Europe as compiled in the framework of the EuFoRIa network
+[@euforia2019; @käber2023] (Figure \@ref(fig:map)). The environmental
+coverage and the variety of forests represented by this data is
+unprecedented within Europe. These forests have been unmanaged for at
+least 10 years prior to their designation as reserves, and most of them
+provide time series of natural forest dynamics over multiple decades and
+up to ca. 80 years. The census periods range from 3 to 37 years, with an
+average of 14 years. The data provide information at the tree level,
+thus allowing for the sequential comparison of processes such as
+individual-tree regeneration and death between the measurements.
+
+We selected 200 sites from this network as the benchmarking dataset for
+the simulation to be representative of the environmental variation
+contained in the data. This was achieved by applying k-means clustering
+to define 200 clusters of plots from the original set of 869 plots along
+the environmental dimensions of temperature, climatic water balance,
+soil quality, slope and aspect. Each of these sites featured at least
+two consecutive measurements. Recruitment minimum size thresholds for
+these sites differed between stem diameters of 0 and 10 cm. For our
+study, we defined two datasets where one included 165 sites with a
+diameter threshold of 7 cm or lower, and the other included another 35
+sites with diameter thresholds from 10 cm. Therefore we had in total 200
+sites with observations of newly established trees above 10 cm and 165
+sites above 7cm. In the results section, any result from observed data
+at 7 cm threshold originated from the 165 sites.
+
+Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
+processed and aggregated following @käber2023 to provide recruitment
+rates per ha and per decade. The observations used in this study
+featured 30,900 newly established trees. Recruitment rates per site,
+sample and decade ranged from zero to 1246 trees, with a mean of 56
+trees. Adult species composition was also available for each plot. For
+more details on this dataset and the detailed regeneration information,
+cf. @käber2023.
 
@@
 
-The overarching goal of the simulation experiments was to assess tree recruitment as it arises from empirical data against its representation in a wide range of models of forest dynamics.
-We define tree recruitment as the passing of a breast height diameter threshold of 7 and 10 cm (synonym: ingrowth).
-To this end, each modeling group was provided with a detailed protocol [@díaz-yáñez2022] with instructions how to perform the simulations, input variables on climate and soil conditions, and the list of expected output variables.
-Neither were further site information (except for the data specified below) nor any data on tree recruitment or forest stand features provided prior to the simulation.
-
-The input variables were collected from different data sources and aggregated to be adapted to the needs of the different models.
-Time series of climatic variables from 1981 to 2018 were provided in hourly [Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly resolution [CHELSA data, Karger et al.
--@karger2021].
-Some variables required for some models, such as relative humidity or vapor pressure deficit, were calculated from these variables.
-The final instructions for using the climate data were slightly different in each model, based on the approach that best suited the model (Table \@ref(tab:tableModels)).
-Soil quality data were provided as continuous values between 1 and 5 [Soilgrids dataset, Hengl et al.
--@hengl2017].
-The protocol also provided the elevation, slope and aspect for each of the 200 sites [@asterscienceteam2019], but no other spatial information such as coordinates, with the exception of iLand and aDGVM2, which required blurred coordinates to derive highly detailed soil data.
-
-The simulations were run in the absence of natural disturbances.
-Only the model LPJ-GUESS had to include a background disturbance to increase chances of shade-intolerant species to establish.
-The simulations were set up to sample species-specific recruitment rates per decade and per ha in the equilibrium state of the model, typically entailing a "spin-up" run (as we did not provide any forest data).
-The modeling teams decided on the simulated area and how they derived these samples.
-The exact length of the simulation was also decided by the modeling teams (Table \@ref(tab:tableModels)).
-Further details on how each modeling team prepared the simulations and the outputs are available in Supplementary Material 1.
-
-The simulations were run in the absence of management to a simulated equilibrium ("Potential Natural Vegetation") with the current climate, as the ultimate goal was to evaluate tree recruitment under comparable and near-equilibrium conditions.
-This entails the assumption that (1) the observations from the forest reserves reflect no traces of forest management, and (2) there is an equilibrium between forest dynamics and climate.
-While the former might be starting to be visible in many of the EuFoRIA reserves, the latter may be more debatable.
-However, in the absence of detailed data on the history of each plot in the EuFoRIa network, some broad assumptions had to be made.
-Both the width of the regeneration niche (i.e., in environmental space) as well as the intensity of the regeneration process (i.e., the number of recruited trees per area and per unit of time) were of interest.
-
-The simulations were run for mixed-species forests (not multiple single-species simulations) using mixtures of eleven species or genera for which regeneration data of sufficient quality were available from EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp.
-L., Fraxinus excelsior L., Quercus spp.
-L., Alnus glutinosa L., Pinus sylvestris L.* The same set of species was used at all 200 sites.
-Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species simulated beyond the eleven species were lumped in a category "others".
-In three models, fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)).
-In 4C only *Fagus sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp.
-and *Pinus sylvestris* are parameterised.
-xComp simulations did not consider *Fraxinus excelsior* due to a depreciated species parameterisation.
-In LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not parameterized and therefore these species could not be included in the simulation.
-Finally, the model aDGVM2 does not simulate individual species.
-Rather, community assembly processes and trait filtering generate plant communities that are adjusted to the biotic and abiotic conditions, and the simulated plants can be classified into ecological strategies based on their trait values in a post-processing step.
-
-Each of the models reported the number of trees crossing the two diameter thresholds by sampling 200 times in a 10-year interval for each species and per ha for each of the 200 sites.
-Multiple samples per site were used to better understand the simulated variation within each site.
-This was done using different strategies, depending on the model, including (1) sampling simulated data from the same 1 ha plot in the equilibrium over time, (2) sampling several 1 ha plots from the simulated forest at one specific point in time (in the equilibrium) or (3) a combination of (1) and (2).
-This resulted in 880,000 observations per model that simulated the 11 species included in the protocol (200 sites, 200 samples per site, 11 species and two diameter thresholds).
-For the models that simulated additional species, their regeneration rates were aggregated as "others".
-
-Two models did not provide results from all the simulated samples or sites to avoid unrealistic results, as follows.
-In the model 4C, a threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid unrealistic stand basal area data, and therefore not all the samples and sites were reported.
-The reason for this is that 4C is not suitable for long-term simulations without management, due to misrepresentations in density-dependent mortality processes in long-term simulations and assumptions of tree geometry that lead to unrealistic single tree dimensions of very old individuals.
-LPJ-GUESS had 2% of the sample outputs without tree recruitment, and these were considered as zero stand basal area and zero regeneration for all the tree species simulated; one site produced grassland rather than a forest and was not included in the results.
+The overarching goal of the simulation experiments was to assess tree
+recruitment as it arises from empirical data against its representation
+in a wide range of models of forest dynamics. We define tree recruitment
+as the passing of a breast height diameter threshold of 7 and 10 cm
+(synonym: ingrowth). To this end, each modeling group was provided with
+a detailed protocol [@díaz-yáñez2022] with instructions how to perform
+the simulations, input variables on climate and soil conditions, and the
+list of expected output variables. Neither were further site information
+(except for the data specified below) nor any data on tree recruitment
+or forest stand features provided prior to the simulation.
+
+The input variables were collected from different data sources and
+aggregated to be adapted to the needs of the different models. Time
+series of climatic variables from 1981 to 2018 were provided in hourly
+[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
+resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
+required for some models, such as relative humidity or vapor pressure
+deficit, were calculated from these variables. The final instructions
+for using the climate data were slightly different in each model, based
+on the approach that best suited the model (Table
+\@ref(tab:tableModels)). Soil quality data were provided as continuous
+values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
+The protocol also provided the elevation, slope and aspect for each of
+the 200 sites [@asterscienceteam2019], but no other spatial information
+such as coordinates, with the exception of iLand and aDGVM2, which
+required blurred coordinates to derive highly detailed soil data.
+
+The simulations were run in the absence of natural disturbances. Only
+the model LPJ-GUESS had to include a background disturbance to increase
+chances of shade-intolerant species to establish. The simulations were
+set up to sample species-specific recruitment rates per decade and per
+ha in the equilibrium state of the model, typically entailing a
+"spin-up" run (as we did not provide any forest data). The modeling
+teams decided on the simulated area and how they derived these samples.
+The exact length of the simulation was also decided by the modeling
+teams (Table \@ref(tab:tableModels)). Further details on how each
+modeling team prepared the simulations and the outputs are available in
+Supplementary Material 1.
+
+The simulations were run in the absence of management to a simulated
+equilibrium ("Potential Natural Vegetation") with the current climate,
+as the ultimate goal was to evaluate tree recruitment under comparable
+and near-equilibrium conditions. This entails the assumption that (1)
+the observations from the forest reserves reflect no traces of forest
+management, and (2) there is an equilibrium between forest dynamics and
+climate. While the former might be starting to be visible in many of the
+EuFoRIA reserves, the latter may be more debatable. However, in the
+absence of detailed data on the history of each plot in the EuFoRIa
+network, some broad assumptions had to be made. Both the width of the
+regeneration niche (i.e., in environmental space) as well as the
+intensity of the regeneration process (i.e., the number of recruited
+trees per area and per unit of time) were of interest.
+
+The simulations were run for mixed-species forests (not multiple
+single-species simulations) using mixtures of eleven species or genera
+for which regeneration data of sufficient quality were available from
+EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
+betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
+Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus
+sylvestris L.* The same set of species was used at all 200 sites. Two
+models included their standard set of species for the simulations, which
+is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species
+simulated beyond the eleven species were lumped in a category "others".
+In three models, fewer than these eleven species were simulated (4C,
+xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). In 4C only *Fagus
+sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp. and *Pinus
+sylvestris* are parameterised. xComp simulations did not consider
+*Fraxinus excelsior* due to a deprecated species parameterisation. In
+LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not
+parameterized and therefore these species could not be included in the
+simulation. Finally, the model aDGVM2 does not simulate individual
+species. Rather, community assembly processes and trait filtering
+generate plant communities that are adjusted to the biotic and abiotic
+conditions, and the simulated plants can be classified into ecological
+strategies based on their trait values in a post-processing step.
+
+Each of the models reported the number of trees crossing the two
+diameter thresholds by sampling 200 times in a 10-year interval for each
+species and per ha for each of the 200 sites. Multiple samples per site
+were used to better understand the simulated variation within each site.
+This was done using different strategies, depending on the model,
+including (1) sampling simulated data from the same 1 ha plot in the
+equilibrium over time, (2) sampling several 1 ha plots from the
+simulated forest at one specific point in time (in the equilibrium) or
+(3) a combination of (1) and (2). This resulted in 880,000 observations
+per model that simulated the 11 species included in the protocol (200
+sites, 200 samples per site, 11 species and two diameter thresholds).
+For the models that simulated additional species, their regeneration
+rates were aggregated as "others".
+
+Two models did not provide results from all the simulated samples or
+sites to avoid unrealistic results, as follows. In the model 4C, a
+threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid
+unrealistic stand basal area data, and therefore not all the samples and
+sites were reported. The reason for this is that 4C is not suitable for
+long-term simulations without management, due to misrepresentations in
+density-dependent mortality processes in long-term simulations and
+assumptions of tree geometry that lead to unrealistic single tree
+dimensions of very old individuals. LPJ-GUESS had 2% of the sample
+outputs without tree recruitment, and these were considered as zero
+stand basal area and zero regeneration for all the tree species
+simulated; one site produced grassland rather than a forest and was not
+included in the results.
 
@@
 
-The simulation results were analyzed regarding (1) recruitment levels (i.e., ingrowth number per unit time and space), (2) recruitment species diversity, (3) recruitment mortality, (4) the relationship between model performance and model traits, and (5) ingrowth gradients along the regeneration niches.
-We evaluated species diversity across the models and in relation to the observed data by calculating the Shannon index $H$ based on the relative proportion of the species in terms of basal area.
-It was calculated for the regeneration ($H_{R_{n}}$) (Equation \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation \ref{eqpiShannon}).
-The higher the value of the index, the higher the species diversity at a particular site and sample.
+The simulation results were analyzed regarding (1) recruitment levels
+(i.e., ingrowth number per unit time and space), (2) recruitment species
+diversity, (3) recruitment mortality, (4) the relationship between model
+performance and model traits, and (5) ingrowth gradients along the
+regeneration niches. We evaluated species diversity across the models
+and in relation to the observed data by calculating the Shannon index
+$H$ based on the relative proportion of the species in terms of basal
+area. It was calculated for the regeneration ($H_{R_{n}}$) (Equation
+\ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation
+\ref{eqpiShannon}). The higher the value of the index, the higher the
+species diversity at a particular site and sample.
 
@@
 ```
-where $s$ is the total number of species present and that have a basal area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of species $i$ in sample $n$ calculated as the regeneration basal area ($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the proportion of species $i$ calculated as the basal area of all trees ($BA_{n,i}$) of that species $i$ relative to total basal area ($TotalR_{BA_{n}}$) of the sample $n$.
-Species diversity was not assessed for aDGVM2 as this model does not simulate individual species.
-
-Mortality in tree recruitment was assessed based on the ratio of recruitment between the 7 and 10 cm diameter thresholds.
-We used the Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference to estimate whether the ratio of recruitment between the 7 and 10 cm diameter thresholds were above or below the expected theoretical rate.
-The Reineke self-thinning rule is usually calculated for even-aged, single species stands and assumes a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands.
-The value used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm to be 77% higher than at 10 cm), calculated using Equation \ref{eqReineke}.
+where $s$ is the total number of species present that have a basal area
+larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
+species $i$ in sample $n$ calculated as the regeneration basal area
+($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal
+area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
+proportion of species $i$ calculated as the basal area of all trees
+($BA_{n,i}$) of that species $i$ relative to total basal area
+($Total_{BA_{n}}$) of the sample $n$. Species diversity was not assessed
+for aDGVM2 as this model does not simulate individual species.
+
+Mortality in tree recruitment was assessed based on the ratio of
+recruitment between the 7 and 10 cm diameter thresholds. We used the
+Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference
+to estimate whether the ratio of recruitment between the 7 and 10 cm
+diameter thresholds were above or below the expected theoretical rate.
+The Reineke self-thinning rule is usually calculated for even-aged,
+single species stands and assumes a fixed relationship between the
+number of stems and the quadratic mean diameter in fully stocked pure
+stands. The value used in our comparisons was 1.77 (i.e., we expect stem
+numbers at 7 cm to be 77% higher than at 10 cm), calculated using
+Equation \ref{eqReineke}.
 
@@
 ```
-We assessed model performance in relation to model traits focusing on (1) model complexity as defined by @bugmann2022, (2) model type (empirical or process based), (3) the presence or absence of a canopy feedback for regeneration, and (4) the scale of application of the model (stand, landscape or global).
-We tested for significant differences between simulated and observed values of the recruitment levels and species diversity by using paired t-tests.
-
-Total recruitment and the regeneration niches of the individual species were evaluated across the environmental gradients of light availability, temperature, and soil moisture.
-Stand basal area (the basal area of all the trees in each sample including tree regeneration) was used as a proxy for light availability at the forest floor, the annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing season warmth, and the climatic water balance as a proxy for soil moisture [@speich2019].
-Recruitment values of each species were calculated as the mean across the 200 samples per site.
-The observed data were modeled using a Generalized Additive Model [@wood2011] with a negative binomial distribution and restricted maximum likelihood to better understand the relationship between the environmental gradients and the levels of observed regeneration, relative to the simulation results per model.
-
-In order to analyze the regeneration niches across the climatic gradients, we focused on five common tree species or genera: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and *Quercus* spp.
-For these species, we calculated the share in the recruited basal area per site, where each site represents a part of the environmental space, as the mean across the available samples per site:
+We assessed model performance in relation to model traits focusing on
+(1) model complexity as defined by @bugmann2022, (2) model type
+(empirical or process based), (3) the presence or absence of a canopy
+feedback for regeneration, and (4) the scale of application of the model
+(stand, landscape or global). We tested for significant differences
+between simulated and observed values of the recruitment levels and
+species diversity by using paired t-tests.
+
+Total recruitment and the regeneration niches of the individual species
+were evaluated across the environmental gradients of light availability,
+temperature, and soil moisture. Stand basal area (the basal area of all
+the trees in each sample including tree regeneration) was used as a
+proxy for light availability at the forest floor, the annual degree-day
+sum [@allen1976; @fischlin1995] as a proxy for growing season warmth,
+and the climatic water balance as a proxy for soil moisture
+[@speich2019]. Recruitment values of each species were calculated as the
+mean across the 200 samples per site. The observed data were modeled
+using a Generalized Additive Model [@wood2011] with a negative binomial
+distribution and restricted maximum likelihood to better understand the
+relationship between the environmental gradients and the levels of
+observed regeneration, relative to the simulation results per model.
+
+In order to analyze the regeneration niches across the climatic
+gradients, we focused on five common tree species or genera: *Abies
+alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
+*Quercus* spp. For these species, we calculated the share in the
+recruited basal area per site, where each site represents a part of the
+environmental space, as the mean across the available samples per site:
 
@@
 ```
-where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$ is the mean basal area in the recruitment (subsequently referred to as "recruited basal area") of species $i$ at site $t$ across the available samples at that site.
-Furthermore, regarding the share in the recruited  basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the recruited basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total recruited basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the recruited basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero.
+where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
+is the mean basal area in the recruitment (subsequently referred to as
+"recruited basal area") of species $i$ at site $t$ across the available
+samples at that site. Furthermore, regarding the share in the recruited
+basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this
+as zero when the recruited basal area of that species
+($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total recruited
+basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the recruited
+basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero.
 
@@
 
-Simulated levels of recruitment varied strongly across the 15 models and typically did not match the levels found in the forest reserve data (Figure \@ref(fig:over)).
-Recruitment was overestimated in most models for both the 7 and 10 cm diameter thresholds, with the exception of the empirical stand model SIBYLA, the landscape model Landis-II and the global model aDGVM2, which estimated recruited levels at the lower end of the plausibility interval of the observed data (Figure \@ref(fig:over)).
-The models with the largest overestimation were the stand model PICUS and the landscape model TreeMig.
-For most models, the variability of simulated regeneration levels across the 200 sites (visible from the interquartile range in the box plots of Figure \@ref(fig:over)) was similar to or smaller than observed, with the notable exception of PICUS, where simulated recruitment variability was much larger.
-
-Both the observations and the simulated data had no recruitment in some samples and at some sites (for details, cf.
-Table S1).
-The observed data had 4% of the samples with no recruitment.
-Only three models had a larger proportion of no recruitment (4C, Landis-II and aDVM2).
-Two models always simulated recruitment for both the 7 and 10 cm threshold (xComp and TreeMig), i.e., they did not feature any zero values.
-The other ten models had a very low percentage of samples with no recruitment (0.01% - 2.39%), i.e., they had distinctly fewer occurrences of zeros compared to the observations.
+Simulated levels of recruitment varied strongly across the 15 models and
+typically did not match the levels found in the forest reserve data
+(Figure \@ref(fig:over)). Recruitment was overestimated in most models
+for both the 7 and 10 cm diameter thresholds, with the exception of the
+empirical stand model SIBYLA, the landscape model Landis-II and the
+global model aDGVM2, which estimated recruited levels at the lower end
+of the plausibility interval of the observed data (Figure
+\@ref(fig:over)). The models with the largest overestimation were the
+stand model PICUS and the landscape model TreeMig. For most models, the
+variability of simulated recruitment levels across the 200 sites
+(visible from the interquartile range in the box plots of Figure
+\@ref(fig:over)) was similar to or smaller than observed, with the
+notable exception of PICUS, where simulated recruitment variability was
+much larger.
+
+Both the observations and the simulated data had no recruitment in some
+samples and at some sites (for details, cf. Table S1). The observed data
+had 4% of the samples with no recruitment. Only three models had a
+larger proportion of no recruitment (4C, Landis-II and aDVM2). Two
+models always simulated recruitment for both the 7 and 10 cm threshold
+(xComp and TreeMig), i.e., they did not feature any zero values. The
+other ten models had a very low percentage of samples with no
+recruitment (0.01% - 2.39%), i.e., they had distinctly fewer occurrences
+of zeros compared to the observations.
 
@@
 
-Most models matched the level of diversity of the observed data quite well (Figure \@ref(fig:H710)).
-Five models overestimated recruitment diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter particularly for the 7 cm diameter threshold.
-The model 4C is a special case, as it simulated five species only, i.e., its diversity values are not directly comparable to those of the other models, nor to the observations.
-Only one model, Landis-II, consistently underestimated recruited diversity.
-
-In most models, there were only small but significant differences in the species diversity of recruitment between 7 and 10 cm across sites (Tables S2).
-Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained the recruitment diversity between the 7 and 10 cm thresholds (i.e., the differences between them were not significant, see Table S2), and the same was evident from the observed data.
-
-For both the observed and the simulated data, and both recruitment thresholds, species diversity in the recruitment compared to stand-level diversity did not feature clear patterns across the models (Figures S1 and S2).
-Most models captured reasonably well or overpredicted species diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left and center; cf.
-Figure S3), but a characteristic feature was that the simulations had a much lower variability of diversity compared to observations.
-Models overpredicting species diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A, left).
-Several models underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).
-
-The majority of the models overpredicted recruitment diversity; as already observed in Figure \@ref(fig:H710), only two of them underestimated it (4C and Landis-II) (Figure \@ref(fig:combinedRichness)B and Figure S4).
-Both the models with feedback and those without feedback from the adult trees to regeneration (via seed production) overestimated, underestimated or captured reasonably well recruitment species diversity compared to the observations.
-Again, simulated diversity had considerably lower variability than observed diversity in the recruitment.
+Most models matched the level of diversity of the observed data quite
+well (Figure \@ref(fig:H710)). Five models overestimated recruitment
+diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
+particularly for the 7 cm diameter threshold. The model 4C is a special
+case, as it simulated five species only, i.e., its diversity values are
+not directly comparable to those of the other models, nor to the
+observations. Only one model, Landis-II, consistently underestimated
+recruited diversity.
+
+In most models, there were only small but significant differences in the
+species diversity of recruitment between 7 and 10 cm across sites
+(Tables S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS)
+maintained the recruitment diversity between the 7 and 10 cm thresholds
+(i.e., the differences between them were not significant, see Table S2),
+and the same was evident from the observed data.
+
+For both the observed and the simulated data, and both recruitment
+thresholds, species diversity in the recruitment compared to stand-level
+diversity did not feature clear patterns across the models (Figures S1
+and S2). Most models captured reasonably well or overpredicted species
+diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left
+and center; cf. Figure S3), but a characteristic feature was that the
+simulations had a much lower variability of diversity compared to
+observations. Models overpredicting species diversity included ForClim1,
+ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A,
+left). Several models underpredicted species diversity, i.e., LPJ-GUESS,
+ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).
+
+The majority of the models overpredicted recruitment diversity; as
+already observed in Figure \@ref(fig:H710), only two of them
+underestimated it (4C and Landis-II) (Figure
+\@ref(fig:combinedRichness)B and Figure S4). Both the models with
+feedback and those without feedback from the adult trees to regeneration
+(via seed production) overestimated, underestimated or captured
+reasonably well recruitment species diversity compared to the
+observations. Again, simulated diversity had considerably lower
+variability than observed diversity in the recruitment.
 
@@
 
-There were strong differences among the models regarding the mortality rate between 7 and 10 cm (Figure \@ref(fig:mortality11)).
-Most of the models featured a mortality rate significantly larger than the observed data, thus at least partially compensating for the general overestimation of recruitment levels (Figure \@ref(fig:over) and Table S3); six models did not feature a significant difference compared to the observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
-Four models (xComp, PICUS, LandClim and TreeMig) featured very high mortality rates (i.e., well above the Reineke self-thinning line), which compensated for the strong initial overestimation of recruitment (cf.
-Figure \@ref(fig:over)).
-Interestingly, two models that underestimated overall recruitment levels (SIBYLA and Landis-II; cf.
-Figure \@ref(fig:over)) featured mortality rates that were close to but still above the self-thinning line (Figure \@ref(fig:mortality11)).
-
-The fact that some models (and observed data) featured lower mortality than expected by Reineke's self-thinning rule may point to facilitation, or simply a higher mortality rate before the trees had reached 7 cm diameter.
-Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher recruitment mortality compared to even-aged, single-species forests.
-Some models yielded a mortality rate of recruitment that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).
-
-The three models with the largest overestimation of the proportion of recruitment at the 7 cm threshold featured the highest mortality between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A).
-The other models that overestimated recruitment had mortalities lower than the expected self-thinning ratio, with the exception of LandClim (Figure \@ref(fig:mortality13)B).
-It is noteworthy that most of the models that featured a low ratio of regeneration between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small overestimation of recruitment at the 7 cm threshold (Figure \@ref(fig:mortality11)).
-aDGVM2 was the only model that underestimated recruitment at the 7 cm threshold while featuring a low mortality between the two thresholds (Figure \@ref(fig:mortality13)C), and only a few models were close to the correct recruitment levels at 7 cm while also being close to the Reineke line (SIBYLA, Landis-II; cf.
-Figure \@ref(fig:mortality13)D; and 4C as well as LandClim).
-
-Lastly, some models that had a low number of individuals at the 7 cm threshold also had a smaller decrease of recruitment between 7 and 10 cm (ratio 7-10 cm) (Figure S5).
-In these models, competition and self-thinning are either not pronounced or must have occurred before the trees had reached 7 cm.
-However, this pattern was not consistent across models (Figure S5).
-It is noteworthy that there is a relationship between the recruitment levels at 7 cm and the mortality rate (Table S4): most of the models showed a positive effect (higher regeneration at 7 cm is coupled to an increase of the mortality between 7 and 10 cm).
-While some models showed a negative effect, the linear trend was not pronounced and not always significant regarding the slope.
+There were strong differences among the models regarding the mortality
+rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the
+models featured a mortality rate significantly larger than the observed
+data, thus at least partially compensating for the general
+overestimation of recruitment levels (Figure \@ref(fig:over) and Table
+S3); six models did not feature a significant difference compared to the
+observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
+Four models (xComp, PICUS, LandClim and TreeMig) featured very high
+mortality rates (i.e., well above the Reineke self-thinning line), which
+compensated for the strong initial overestimation of recruitment (cf.
+Figure \@ref(fig:over)). Interestingly, two models that underestimated
+overall recruitment levels (SIBYLA and Landis-II; cf. Figure
+\@ref(fig:over)) featured mortality rates that were close to but still
+above the self-thinning line (Figure \@ref(fig:mortality11)).
+
+The fact that some models (and observed data) featured lower mortality
+than expected by Reineke's self-thinning rule may point to facilitation,
+or simply a higher mortality rate before the trees had reached 7 cm
+diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
+whose mortality was well above the self-thinning line, indicates that
+they feature higher recruitment mortality compared to even-aged,
+single-species forests. Some models yielded a mortality rate of
+recruitment that is broadly compatible with the self-thinning rule
+(e.g., 4C and LandClim).
+
+The three models with the largest overestimation of the proportion of
+recruitment at the 7 cm threshold featured the highest mortality between
+the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The other
+models that overestimated recruitment had mortalities lower than the
+expected self-thinning ratio, with the exception of LandClim (Figure
+\@ref(fig:mortality13)B). It is noteworthy that most of the models that
+featured a low ratio of regeneration between 7 and 10 cm (i.e.,
+ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small
+overestimation of recruitment at the 7 cm threshold (Figure
+\@ref(fig:mortality11)). aDGVM2 was the only model that underestimated
+recruitment at the 7 cm threshold while featuring a low mortality
+between the two thresholds (Figure \@ref(fig:mortality13)C), and only a
+few models were close to the correct recruitment levels at 7 cm while
+also being close to the Reineke line (SIBYLA, Landis-II; cf. Figure
+\@ref(fig:mortality13)D; and 4C as well as LandClim).
+
+Lastly, some models that had a low number of individuals at the 7 cm
+threshold also had a smaller decrease of recruitment between 7 and 10 cm
+(ratio 7-10 cm) (Figure S5). In these models, competition and
+self-thinning are either not pronounced or must have occurred before the
+trees had reached 7 cm. However, this pattern was not consistent across
+models (Figure S5). It is noteworthy that there is a relationship
+between the recruitment levels at 7 cm and the mortality rate (Table
+S4): most of the models showed a positive effect (higher regeneration at
+7 cm is coupled to an increase of the mortality between 7 and 10 cm).
+While some models showed a negative effect, the linear trend was not
+pronounced and not always significant regarding the slope.
 
@@
 
-There was no significant relationship between the mean complexity of the regeneration module (Table \@ref(tab:tableModels)) and the overestimation proportion at a diameter threshold of 7 cm (Figure S6).
-All the models except 4C had significant differences between the observed and simulated mean regeneration values.
-Therefore it was not possible to differentiate in terms of performance between models type (empirical or process based), or their scales (stand, landscape or global) (Table S5).
-The values of recruited species diversity simulated by each model were also assessed against the observed data to see if there was a difference between models that include a feedback compared to those that did not; all the models except ForCEEPs(f) and FORMIND had significant differences (Table S6).
-It is noteworthy that for the model ForCEEPS the pattern changed from the overprediction of species diversity in the regeneration to a diversity level that is closer to observations when the recruitment module included feedback (ForCEEPs(f)).
+There was no significant relationship between the mean complexity of the
+regeneration module (Table \@ref(tab:tableModels)) and the
+overestimation proportion at a diameter threshold of 7 cm (Figure S6).
+All the models except 4C had significant differences between the
+observed and simulated mean recruitment values. Therefore it was not
+possible to differentiate in terms of performance between models type
+(empirical or process based), or their scales (stand, landscape or
+global) (Table S5). The values of recruited species diversity simulated
+by each model were also assessed against the observed data to see if
+there was a difference between models that include a feedback compared
+to those that did not; all the models except ForCEEPs(f) and FORMIND had
+significant differences (Table S6). It is noteworthy that for the model
+ForCEEPS the pattern changed from the overprediction of species
+diversity in the regeneration to a diversity level that is closer to
+observations when the recruitment module included feedback
+(ForCEEPs(f)).
 
@@
 
-When evaluating recruited levels along key gradients of light availability (basal area), temperature (degree-days) and soil moisture (climatic water balance), distinct features emerged: the models reproduced the effect of basal area in both its magnitude and patterns (Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better than the effects of the climatic gradients, where they featured varying patterns.
-
-In the observed data, recruited levels decreased clearly with increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a).
-This trend was captured by the Landis-II model only, albeit at much lower values of basal area than in the observations.
-The other models featured distinctly different trends, such as (1) an increase in recruited levels with increasing total basal area followed by the absence of regeneration at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant recruitment levels with basal area (ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of recruitment up to a certain value of stand basal area followed by a decrease at even higher values of stand basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).
-
-The observed recruitment did not change much across the climatic water balance gradient, and showed a slight increase with temperature.
-The models did not match the observed reruitment patterns across the gradients of soil moisture and temperature (Figure \@ref(fig:nvTrend710BAlightTemp)b,c), and neither was there a clear pattern across models.
-Regarding the soil moisture gradient, a group of models featured decreasing recruitment with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS).
-Across the temperature gradient, all the models that featured a decrease in recruitment levels with increasing soil moisture showed the opposite trend (i.e., more recruitment with increasing temperature), except for 4C.
-The other models did not feature a clear pattern.
-
-Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus sylvestris*) and oak (*Quercus* spp.) were the species observed most frequently in the EuFoRIa data.
-Most models captured well the share of basal area in the recruitment of these five species compared to the observed data along the two major climatic gradients (Figure \@ref(fig:diff), cf.
-Figure S8).
-However, the models differed considerably in the way in which they simulated these climatic niches.
-
-First, some models overestimated recruitment across the entire gradients of at least one of the five main tree species (Figure \@ref(fig:diff), e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated recruitment of more than one species for a large part of the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
-red colors in Figure \@ref(fig:diff)).
-These trends were consistent for both recruitment thresholds, although the exact changes in the share of the recruited basal area were different (Figure S8).
-
-Second, there was some congruence in the simulation results by species across the models.
-The share in the recruitment of Silver fir (*Abies alba*) was captured well across the gradient by most models except for xComp, PICUS, and LandClim, which overestimated the recruitment share of this species.
-There were also some peculiarities evident for some models, such as 4C that did not simulate *Abies alba*.
-The patterns across models for *Fagus sylvatica* recruitment were more complex, as its recruitment was underestimated by many models across the environmental gradients featuring negligible regeneration at most sites, while others overestimated it in different parts of the environmental space.
-Most models overestimated the share of *Picea abies* recruitment in the cold-wet part of the gradients.
-*Pinus sylvestris* and *Quercus* spp.
-represented a small share of the recruitment in the observed data mostly at warm-dry sites.
-Many models, however, erroneously featured recruitment for these two species along most of the climatic gradients, although with a low share of basal area.
-Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had almost no recruitment of *Pinus sylvestris* and *Quercus* spp., or did not feature any recruitment of these species at all.
-
-Lastly, there is no model that performed well across the five species regarding the species-specific recruitment levels in the environmental space (Figure \@ref(fig:diff)).
-Some models (e.g., ForClim11, PICUS) tended to consistently overestimate the recruitment share of all five species, but most models overestimated the recruitment share of some species while underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
+When evaluating total recruited levels along key gradients of light
+availability (basal area), temperature (degree-days) and soil moisture
+(climatic water balance), distinct features emerged: the models
+reproduced the effect of basal area in both its magnitude and patterns
+(Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better
+than the effects of the climatic gradients, where they featured varying
+patterns.
+
+In the observed data, recruited levels decreased clearly with increasing
+stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a). This trend
+was captured by the Landis-II model only, albeit at much lower values of
+basal area than in the observations. The other models featured
+distinctly different trends, such as (1) an increase in recruited levels
+with increasing total basal area followed by the absence of regeneration
+at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig
+and LPJ-GUESS), (2) almost constant recruitment levels with basal area
+(ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of
+recruitment up to a certain value of stand basal area followed by a
+decrease at even higher values of stand basal area, with model-specific
+thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).
+
+The observed recruitment did not change much across the climatic water
+balance gradient, and showed a slight increase with temperature. The
+models did not match the observed recruitment patterns across the
+gradients of soil moisture and temperature (Figure
+\@ref(fig:nvTrend710BAlightTemp)b,c), and neither was there a clear
+pattern across models. Regarding the soil moisture gradient, a group of
+models featured decreasing recruitment with increasing soil moisture
+(4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and
+iLand), while a few models showed the opposite trend (ForCEEPS(f),
+FORMIND and LPJ-GUESS). Across the temperature gradient, all the models
+that featured a decrease in recruitment levels with increasing soil
+moisture showed the opposite trend (i.e., more recruitment with
+increasing temperature), except for 4C. The other models did not feature
+a clear pattern.
+
+Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
+sylvestris*) and oak (*Quercus* spp.) were the species observed most
+frequently in the EuFoRIa data. Most models captured well the share of
+basal area in the recruitment of these five species compared to the
+observed data along the two major climatic gradients (Figure
+\@ref(fig:diff), cf. Figure S8). However, the models differed
+considerably in the way in which they simulated these climatic niches.
+
+First, some models overestimated recruitment across the entire gradients
+of at least one of the five main tree species (Figure \@ref(fig:diff),
+e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models
+overestimated recruitment of more than one species for a large part of
+the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
+red colors in Figure \@ref(fig:diff)). These trends were consistent for
+both recruitment thresholds, although the exact changes in the share of
+the recruited basal area were different (Figure S8).
+
+Second, there was some congruence in the simulation results by species
+across the models. The share in the recruitment of Silver fir (*Abies
+alba*) was captured well across the gradient by most models except for
+xComp, PICUS, and LandClim, which overestimated the recruitment share of
+this species. There were also some peculiarities evident for some
+models, such as 4C that did not simulate *Abies alba*. The patterns
+across models for *Fagus sylvatica* recruitment were more complex, as
+its recruitment was underestimated by many models across the
+environmental gradients featuring negligible regeneration at most sites,
+while others overestimated it in different parts of the environmental
+space. Most models overestimated the share of *Picea abies* recruitment
+in the cold-wet part of the gradients. *Pinus sylvestris* and *Quercus*
+spp. represented a small share of the recruitment in the observed data
+mostly at warm-dry sites. Many models, however, erroneously featured
+recruitment for these two species along most of the climatic gradients,
+although with a low share of basal area. Five models (4C, FORMIND,
+SIBYLA, xComp, and LandClim) had almost no recruitment of *Pinus
+sylvestris* and *Quercus* spp., or did not feature any recruitment of
+these species at all.
+
+Lastly, there is no model that performed well across the five species
+regarding the species-specific recruitment levels in the environmental
+space (Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
+tended to consistently overestimate the recruitment share of all five
+species, but most models overestimated the recruitment share of some
+species while underestimating the share of others (e.g., ForCEEPS,
+LPJ-GUESS).
 
@@
 
-Tree regeneration is a fundamental process in forest dynamics.
-Correctly capturing tree regeneration in dynamic models is fundamental to, for example, evaluate post-disturbance dynamics and potential long-term recovery trajectories as it will define the forest state in the first decades [@seidl2022].
-If this initial phase is not captured well, we can not properly assess aspects such as forest resilience or the timing, magnitude and progression of carbon sequestration.
-In this study, for the first time, projections of tree regeneration from multiple models of forest dynamics were confronted with a unique dataset from unmanaged forest reserves across a large environmental gradient in Europe.
-
-The EuFoRIa data [@käber2023] are exceptional, particularly with respect to the number of records (number of sites and repeated measurements on unmanaged forests), which is essential for capturing a highly 'noisy' process such as tree regeneration.
-The use of this dataset for model benchmarking provided novel insights on the ability of state-of-the-art models to accurately simulate regeneration levels, species composition, and mortality in an early stage of tree life, i.e., between a tree diameter of 7 and 10 cm.
-Overall, by adopting this approach a much broader understanding resulted than if we had used these data for model calibration: it is primarily from the shortcomings of the models that we can gain ecological insights [cf.
-@trugman2022].
+Tree regeneration is a fundamental process in forest dynamics. Correctly
+capturing tree regeneration in dynamic models is fundamental to, for
+example, evaluate post-disturbance dynamics and potential long-term
+recovery trajectories as it will define the forest state in the first
+decades [@seidl2022]. If this initial phase is not captured well, we can
+not properly assess aspects such as forest resilience or the timing,
+magnitude and progression of carbon sequestration. In this study, for
+the first time, projections of tree regeneration from multiple models of
+forest dynamics were confronted with a unique dataset from unmanaged
+forest reserves across a large environmental gradient in Europe.
+
+The EuFoRIa data [@käber2023] are exceptional, particularly with respect
+to the number of records (number of sites and repeated measurements on
+unmanaged forests), which is essential for capturing a highly 'noisy'
+process such as tree regeneration. The use of this dataset for model
+benchmarking provided novel insights on the ability of state-of-the-art
+models to accurately simulate regeneration levels, species composition,
+and mortality in an early stage of tree life, i.e., between tree
+diameters of 7 and 10 cm. Overall, by adopting this approach a much
+broader understanding resulted than if we had used these data for model
+calibration: it is primarily from the shortcomings of the models that we
+can gain ecological insights [cf. @trugman2022].
 
@@
 
-Most of the models overestimated tree recruitment levels.
-This has potentially far-reaching implications, e.g., regarding biomass (and thus carbon) turnover, with a potential overestimation of the capacity for forest carbon sequestration [@pugh2019; @pan2011].
-Yet, we focused on one specific stage, i.e., recruitment into the 7 and 10 cm diameter classes.
-Trees at this size contribute little to carbon sequestration, and if the excess regeneration at this stage is compensated -- in the models -- soon thereafter by higher mortality, overall simulated regeneration may still yield trustworthy results at the level of stand structure and related forest attributes such as basal area or biomass.
-
-From a modeling point of view, excess regeneration inevitably requires excess mortality rates in a later stage, either -- as observed for some models -- between the 7 and 10 cm diameter thresholds, or soon after the 10 cm threshold has been crossed.
-In any case, correcting at early stages for the expected forest densities at later stages is equivalent to compensating for a first error (excess regeneration) by a second error (excess mortality).
-It is highly likely that biased projections will result, because the two errors are unlikely to be perfectly linked and thus will not always compensate each other.
-Hence, this structural problem of most models investigated here is problematic particularly if the models are to be used under novel conditions such as under climate change [e.g., @huber2021] or in a decision support context [e.g., @thrippleton2021].
-
-Regeneration levels define the structure and composition of future forests, but it is equally important to correctly identify areas where regeneration is lacking [@rammer2021].
-There are multiple constraints to the regeneration niche of tree species [@price2001], and therefore the absence of regeneration is likely to be common [@fortin2007], even over larger areas such as the one-hectare plots used here.
-Tree regeneration data are characterized by zero-inflation.
-This was clearly evident from the EuFoRIa dataset, but some of the models did not produce zeros at all, or featured a very low proportion of zero data.
-This substantial difference may be due to the fact that the simulation results were drawn from equilibrium forests, whereas in reality many of the forest reserves are recovering from past management activities and have become denser over the past decades [e.g., @heiri2009], leading to less regeneration than in an equilibrium situation.
-Another possible reason for this difference may be the exclusion of factors like deer browsing or failure to accurately measure establishment filters, such as those to site limitants or competition with herb layers.
+Most of the models overestimated tree recruitment levels. This has
+potentially far-reaching implications, e.g., regarding biomass (and thus
+carbon) turnover, with a potential overestimation of the capacity for
+forest carbon sequestration [@pugh2019; @pan2011]. Yet, we focused on
+one specific stage, i.e., recruitment into the 7 and 10 cm diameter
+classes. Trees at this size contribute little to carbon sequestration,
+and if the excess regeneration at this stage is compensated -- in the
+models -- soon thereafter by higher mortality, overall simulated
+regeneration may still yield trustworthy results at the level of stand
+structure and related forest attributes such as basal area or biomass.
+
+From a modeling point of view, excess regeneration inevitably requires
+excess mortality rates in a later stage, either -- as observed for some
+models -- between the 7 and 10 cm diameter thresholds, or soon after the
+10 cm threshold has been crossed. In any case, correcting at early
+stages for the expected forest densities at later stages is equivalent
+to compensating for a first error (excess regeneration) by a second
+error (excess mortality). It is highly likely that biased projections
+will result, because the two errors are unlikely to be perfectly linked
+and thus will not always compensate each other. Hence, this structural
+problem of most models investigated here is problematic particularly if
+the models are to be used under novel conditions such as under climate
+change [e.g., @huber2021] or in a decision support context [e.g.,
+@thrippleton2021].
+
+Regeneration levels define the structure and composition of future
+forests, but it is equally important to correctly identify areas where
+regeneration is lacking [@rammer2021]. There are multiple constraints to
+the regeneration niche of tree species [@price2001], and therefore the
+absence of regeneration is likely to be common [@fortin2007], even over
+larger areas such as the one-hectare plots used here. Tree regeneration
+data are characterized by zero-inflation. This was clearly evident from
+the EuFoRIa dataset, but some of the models did not produce zeros at
+all, or featured a very low proportion of zero data. This substantial
+difference may be due to the fact that the simulation results were drawn
+from equilibrium forests, whereas in reality many of the forest reserves
+are recovering from past management activities and have become denser
+over the past decades [e.g., @heiri2009], leading to less regeneration
+than in an equilibrium situation. Another possible reason for this
+difference may be the exclusion of factors like deer browsing or failure
+to accurately measure establishment filters, such as competition with
+herb layers, site or environmental limitants.
 
@@
 
-Correctly capturing species composition in tree regeneration is important to assess the future functional diversity of a forest, e.g., its sensitivity to drought or resilience to disturbances [@redmond2015; @seidl2022].
-In the simulations, overall species diversity levels in the regeneration were well within the observed range for half of the models (7 out of 14).
-Thus, while most of the models are facing difficulties to quantitatively match regeneration levels (cf.
-above), their performance is better when it comes to matching the composition of regeneration as a function of abiotic and biotic conditions.
-Most models maintained or even decreased recruitment diversity between the 7 and 10 cm thresholds, and the same is visible from the empirical data, but the differences were not significant.
-Diversity variations may be explained by the way the individual models consider regeneration processes [@konig_tree_2022], e.g., without considering the feedback from adults, regeneration can be linked to either seed rain from dispersal or a seed/seedling bank; neither process is incorporated across many models studied here, however.
-
-The species diversity of the entire stand was captured better by the models than the diversity of tree recruitment, and only a few models overpredicted stand-level diversity.
-Defining regeneration composition as being proportional to adult abundance, regardless of productivity, might be a simple and conservative assumption to safely maintain relative species abundances [@hanbury-brown2022], although this may be simplistic.
-Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the recruitment than those without feedback, with the exception of ForCEEPS that significantly improved regeneration species diversity with the model version that included feedback (ForCEEPS(f)).
-The similar performance of models with and without feedback is likely because the models put more weight on the regeneration niche arising from abiotic and biotic filters, than from the habitat niche of the adult trees [@grubb1977].
+Correctly capturing species composition in tree regeneration is
+important to assess the future functional diversity of a forest, e.g.,
+its sensitivity to drought or resilience to disturbances [@redmond2015;
+@seidl2022]. In the simulations, overall species diversity levels in the
+regeneration were well within the observed range for half of the models
+(7 out of 14). Thus, while most of the models are facing difficulties to
+quantitatively match regeneration levels (cf. above), their performance
+is better when it comes to matching the composition of regeneration as a
+function of abiotic and biotic conditions. Most models maintained or
+even decreased recruitment diversity between the 7 and 10 cm thresholds,
+and the same is visible from the empirical data, but the differences
+were not significant. Diversity variations may be explained by the way
+the individual models consider regeneration processes
+[@konig_tree_2022], e.g., without considering the feedback from adults,
+regeneration can be linked to either seed rain from dispersal or a
+seed/seedling bank; neither process is incorporated across many models
+studied here, however.
+
+The species diversity of the entire stand was captured better by the
+models than the diversity of tree recruitment, and only a few models
+overpredicted stand-level diversity. Defining regeneration composition
+as being proportional to adult abundance, regardless of productivity,
+might be a simple and conservative assumption to safely maintain
+relative species abundances [@hanbury-brown2022], although this may be
+simplistic. Furthermore, based on the simulation results, there is no
+evidence that models with feedback from the canopy (in terms of species
+composition of regeneration) better captured the species diversity of
+the recruitment than those without feedback, with the exception of
+ForCEEPS that significantly improved regeneration species diversity with
+the model version that included feedback (ForCEEPS(f)). The similar
+performance of models with and without feedback is likely because the
+models put more weight on the regeneration niche arising from abiotic
+and biotic filters, than from the habitat niche of the adult trees
+[@grubb1977].
 
@@
 
-There are several factors that lead to mortality during the regeneration phase, such as competition [@casper1997] and multiple abiotic factors [@schmid2021; @cunningham2006].
-As mentioned above, it is reasonable to expect that models that overestimate regeneration may have a particularly high mortality between the 7 and 10 cm thresholds.
-However, this was not consistently evident from the simulations.
-This implies that these models must have an excess mortality in later stages, if we assume that all models were able to capture the structure and composition of the adult stands along the EuFoRIa gradient; this however was not tested here.
-
-Higher mortality towards the adult phase has important implications for forest dynamics and the goods and services provided by forests.
-On the one hand, mortality in later stages may erroneously enhance the share of less shade tolerant species in the models [@klopcic2015], therefore shifting the species composition.
-Unrealistic high stem densities for a longer period of time may overestimate the role of tree regrowth in carbon sink dynamics [@pugh2019].
-Given our set of simulations and analyses, we cannot conclusively assess what is happening in the models, and further studies are required that focus on a wider range of tree sizes and the fate of tree regeneration along such a size continuum.
+There are several factors that lead to mortality during the regeneration
+phase, such as competition [@casper1997] and multiple abiotic factors
+[@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
+expect that models that overestimate regeneration may have a
+particularly high mortality between the 7 and 10 cm thresholds. However,
+this was not consistently evident from the simulations. This implies
+that these models must have an excess mortality in later stages, if we
+assume that all models were able to capture the structure and
+composition of the adult stands along the EuFoRIa gradient; this however
+was not tested here.
+
+Higher mortality towards the adult phase has important implications for
+forest dynamics and the goods and services provided by forests. On the
+one hand, mortality in later stages may erroneously enhance the share of
+less shade tolerant species in the models [@klopcic2015], therefore
+shifting the species composition. Unrealistic high stem densities for a
+longer period of time may overestimate the role of tree regrowth in
+carbon sink dynamics [@pugh2019]. Given our set of simulations and
+analyses, we cannot conclusively assess what is happening in the models,
+and further studies are required that focus on a wider range of tree
+sizes and the fate of tree regeneration along such a size continuum.
 
@@
 
-In spite of the critical considerations above, it is remarkable that most models did not deviate exceedingly from the observations with respect to simulated recruited levels -- after all, few if any of the models' regeneration routines are sufficiently constrained by data, with the exception of the empirical models xComp and SIBYLA.
-It is noteworthy that in spite of their empirical basis, these two models did not match empirical regeneration levels, in a similar magnitude as the other models and even in opposite directions (SIBYLA: underestimation; xComp: overestimation).
-It appears that using empirical data of limited geographical scope and constrained to managed forests, as done in these models, leads to extrapolation problems already under current climatic conditions.
-For the other models, whose regeneration modules are not strongly constrained by empirical data, the multiple strategies that are available for modeling regeneration [@konig_tree_2022] could have implied that model performance would be much worse than what we found.
-
-Our study showed that increasing complexity in the regeneration modules is not linked with a higher accuracy of the projections of regeneration levels, species composition or mortality at early tree stages, as there was no significant relationship between model performance and model complexity.
-Increasing complexity in regeneration modules has been motivated by better process understanding and enhancing of model accuracy [@bugmann2022].
-However, more complex models do not necessarily lead to better projections and rather to a reduced transparency and lower predictive power [@franklin2020].
-Thus, the question of what level of detail is appropriate and parsimonious for modeling tree regeneration processes remains open [@konig_tree_2022; @bugmann2022].
+In spite of the critical considerations above, it is remarkable that
+most models did not deviate exceedingly from the observations with
+respect to simulated recruited levels -- after all, few if any of the
+models' regeneration routines are sufficiently constrained by data, with
+the exception of the empirical models xComp and SIBYLA. It is noteworthy
+that in spite of their empirical basis, these two models did not match
+empirical regeneration levels, in a similar magnitude as the other
+models and even in opposite directions (SIBYLA: underestimation; xComp:
+overestimation). It appears that using empirical data of limited
+geographical scope and constrained to managed forests, as done in these
+models, leads to extrapolation problems already under current climatic
+conditions. For the other models, whose regeneration modules are not
+strongly constrained by empirical data, the multiple strategies that are
+available for modeling regeneration [@konig_tree_2022] could have
+implied that model performance would be much worse than what we found.
+
+Our study showed that increasing complexity in the regeneration modules
+is not linked with a higher accuracy of the projections of regeneration
+levels, species composition or mortality at early tree stages, as there
+was no significant relationship between model performance and model
+complexity. Increasing complexity in regeneration modules has been
+motivated by better process understanding and enhancing of model
+accuracy [@bugmann2022]. However, more complex models do not necessarily
+lead to better projections and rather to a reduced transparency and
+lower predictive power [@franklin2020]. Thus, the question of what level
+of detail is appropriate and parsimonious for modeling tree regeneration
+processes remains open [@konig_tree_2022; @bugmann2022].
 
@@
 
-Competition for light as a strong filter for tree regeneration has been widely documented [@messier1999; @collet2006; @berdanier2016], but the models examined here did not reproduce this expectation (i.e., decreasing regeneration levels with increasing stand basal area).
-However, it is difficult to measure light availability at large spatial and temporal scales.
-We used total stand basal area as a proxy for light availability [cf.
-@schmid2021].
-However, we were unable to consider light availability restrictions caused by ground vegetation, which may be an important filter for some EuForIa stands [@woltjer2008].
-We found pronounced differences in the stand basal area ranges simulated by the models, but also between models and observations.
-This made it impossible to evaluate the regeneration for the extremes of the stand density ranges in some models.
-For example, regeneration levels at low stand densities are relevant to assess how well forests are recovering e.g.
-after gap creation due to disturbance [@seidl2022; @grubb1977].
-At the level of the simulated one-hectare samples, average basal area is typically not very low as long as no larger disturbances occur, which was explicitly excluded in our simulation protocol.
-Towards the other end of the spectrum, i.e., with increasing stand basal area, it would be reasonable to expect that the recruitment of the different species would become sparser and drop out entirely under low-light conditions [@klopcic2012; @zell2019].
-However, few models showed this trend, thus indicating that the relation between regeneration and light availability is not yet captured correctly in most models.
-Yet, several of the models that did feature an increase of regeneration with increasing basal area include a feedback between seed production of mature trees and regeneration.
-Thus, it seems that higher light competition does not sufficiently compensate for increased seed availability with higher basal area due to a higher abundance of mature trees.
-
-There were pronounced differences how the main tree species were represented by the models along the environmental gradients, in particular the dominant species *Abies alba*, *Fagus sylvatica* and *Picea abies*.
-The regeneration levels were sampled from simulations in the equilibrium, and in this state it is expected that non-climax species such as *Pinus sylvestris* or *Quercus* would be of minor importance, or absent [@klopcic2015].
-Most models captured this low abundance, which is also found in the empirical data.
-Thus, although the broad patterns are matched by many models, improvements in the quantification of the regeneration niche of the species are needed, but this cannot be done in the absence of robust datasets across multiple species.
+Competition for light as a strong filter for tree regeneration has been
+widely documented [@messier1999; @collet2006; @berdanier2016], but the
+models examined here did not reproduce this expectation (i.e.,
+decreasing regeneration levels with increasing stand basal area).
+However, it is difficult to measure light availability at large spatial
+and temporal scales. We used total stand basal area as a proxy for light
+availability [cf. @schmid2021]. However, we were unable to consider
+light availability restrictions caused by ground vegetation, which may
+be an important filter for some EuForIa stands [@woltjer2008]. We found
+pronounced differences in the stand basal area ranges simulated by the
+models, but also between models and observations. This made it
+impossible to evaluate the regeneration for the extremes of the stand
+density ranges in some models. For example, regeneration levels at low
+stand densities are relevant to assess how well forests are recovering
+e.g. after gap creation due to disturbance [@seidl2022; @grubb1977]. At
+the level of the simulated one-hectare samples, average basal area is
+typically not very low as long as no larger disturbances occur, which
+was explicitly excluded in our simulation protocol. Towards the other
+end of the spectrum, i.e., with increasing stand basal area, it would be
+reasonable to expect that the recruitment of the different species would
+become sparser and drop out entirely under low-light conditions
+[@klopcic2012; @zell2019]. However, few models showed this trend, thus
+indicating that the relation between regeneration and light availability
+is not yet captured correctly in most models. Yet, several of the models
+that did feature an increase of regeneration with increasing basal area
+include a feedback between seed production of mature trees and
+regeneration. Thus, it seems that higher light competition does not
+sufficiently compensate for increased seed availability with higher
+basal area due to a higher abundance of mature trees.
+
+There were pronounced differences in how the main tree species were
+represented by the models along the environmental gradients, in
+particular the dominant species *Abies alba*, *Fagus sylvatica* and
+*Picea abies*. The regeneration levels were sampled from simulations in
+the equilibrium, and in this state it is expected that non-climax
+species such as *Pinus sylvestris* or *Quercus* would be of minor
+importance, or absent [@klopcic2015]. Most models captured this low
+abundance, which is also found in the empirical data. Thus, although the
+broad patterns are matched by many models, improvements in the
+quantification of the regeneration niche of the species are needed, but
+this cannot be done in the absence of robust datasets across multiple
+species.
 
@@
 
-The EuFoRIa data as used here are well suited to better understand tree regeneration.
-However, three aspects of these data may represent considerable limitations.
-First, we made a comparison of tree regeneration in an equilibrium state, but we cannot assess how close the forests included in the EuFoRIa dataset were to such a state.
-The data were collected in forest reserves where no management has taken place for long periods of time.
-This makes our assumption of an equilibrium between forest properties and environmental drivers more reasonable than it might appear at first sight.
-In an analysis of primeval forests at demographic equilibrium, @brzeziecki2021 found higher regeneration rates than the ones observed in the EuFoRia data.
-Thus, the overestimation of regeneration rates by the models may not be so problematic.
-Second, the data were collected for regeneration above a 7 cm threshold, thus limiting the assessment of tree regeneration to a holistic viewpoint.
-This constituted a hard limit based on which we can understand only some aspects of tree regeneration, which in its entirety often comprises a rather long period since seed production [@price2001].
-In reality, many environmental constraints are acting on young trees [@käber2021] that we were unable to assess.
-Yet, regeneration data with lower thresholds are simply not available in a harmonized manner across large environmental gradients.
-Third, the empirical data were collected from rather small plots, while we sampled simulated regeneration levels from 1 ha areas, which may lead to an incorrect representation of space.
-Even though the strategy we adopted is not ideal, it represents a common challenge when harmonizing diverse data sources originating from varied sampling strategies [@portier2022].
-It would have been extremely challenging for such a variety of models to follow a protocol where the spatial sampling size was different at each of the 200 sites, and it would have introduced additional uncertainty in the results.
-
-The design of our sampling protocol did not include spatial aspects such as seed dispersal or detailed soil data.
-While we considered a wide range of models of forest dynamics, from stand to global scales, the simulation setup was limited to 1 ha regeneration values.
-This lack of consideration of spatial scale is appropriate for stand-scale models, but it potentially puts the landscape models at a disadvantage, as they have been built to be accurate at the landscape level.
-Without spatial context, we are limiting regeneration to the seed influx from the stand itself, unless the model has a background seed input.
-Yet, the global models should not be at a disadvantage due to the limited spatial considerations [but cf.
-Snell et al.
--@snell2018; or Lehsten et al.
--@lehsten2019], as they are usually lacking dispersal between cells and are based on a strong abstraction of horizontal space [@hanbury-brown2022].
-Lastly, detailed data on soil conditions were not available from the observed data, and independent, admittedly rather coarse data for soil and climatic water balance had to be used instead.
-It is noteworthy that many models represent drought based on detailed indicators based e.g.
-on soil water holding capacity, which had to be derived from a rough soil quality measure.
-This may at least partially explain the unsatisfactory performance of many models along the drought axis (i.e., climatic water balance).
+The EuFoRIa data as used here are well suited to better understand tree
+regeneration. However, three aspects of these data may represent
+considerable limitations. First, we made a comparison of tree
+regeneration in an equilibrium state, but we cannot assess how close the
+forests included in the EuFoRIa dataset were to such a state. The data
+were collected in forest reserves where no management has taken place
+for long periods of time. This makes our assumption of an equilibrium
+between forest properties and environmental drivers more reasonable than
+it might appear at first sight. In an analysis of primeval forests at
+demographic equilibrium, @brzeziecki2021 found higher regeneration rates
+than the ones observed in the EuFoRia data. Thus, the overestimation of
+regeneration rates by the models may not be so problematic. Second, the
+data were collected for regeneration above a 7 cm threshold, thus
+limiting the assessment of tree regeneration to a holistic viewpoint.
+This constituted a hard limit based on which we can understand only some
+aspects of tree regeneration, which in its entirety often comprises a
+rather long period since seed production [@price2001]. In reality, many
+environmental constraints are acting on young trees [@käber2021] that we
+were unable to assess. Yet, regeneration data with lower thresholds are
+simply not available in a harmonized manner across large environmental
+gradients. Third, the empirical data were collected from rather small
+plots, while we sampled simulated regeneration levels from 1 ha areas,
+which may lead to an incorrect representation of space. Even though the
+strategy we adopted is not ideal, it represents a common challenge when
+harmonizing diverse data sources originating from varied sampling
+strategies [@portier2022]. It would have been extremely challenging for
+such a variety of models to follow a protocol where the spatial sampling
+size was different at each of the 200 sites, and it would have
+introduced additional uncertainty in the results.
+
+The design of our sampling protocol did not include spatial aspects such
+as seed dispersal or detailed soil data. While we considered a wide
+range of models of forest dynamics, from stand to global scales, the
+simulation setup was limited to 1 ha regeneration values. This lack of
+consideration of spatial scale is appropriate for stand-scale models,
+but it potentially puts the landscape models at a disadvantage, as they
+have been built to be accurate at the landscape level. Without spatial
+context, we are limiting regeneration to the seed influx from the stand
+itself, unless the model has a background seed input. Yet, the global
+models should not be at a disadvantage due to the limited spatial
+considerations [but cf. Snell et al. -@snell2018; or Lehsten et al.
+-@lehsten2019], as they are usually lacking dispersal between cells and
+are based on a strong abstraction of horizontal space
+[@hanbury-brown2022]. Lastly, detailed data on soil conditions were not
+available from the observed data, and independent, admittedly rather
+coarse data for soil and climatic water balance had to be used instead.
+It is noteworthy that many models represent drought based on detailed
+indicators based e.g. on soil water holding capacity, which had to be
+derived from a rough soil quality measure. This may at least partially
+explain the unsatisfactory performance of many models along the drought
+axis (i.e., climatic water balance).
 
@@
 
-With our study, we have demonstrated that models of forest dynamics need a focus on tree regeneration modules to make them more robust.
-It remains uncertain what level of detail is required to model regeneration processes, and this must be addressed in future research.
-We recommend that the improvement of the regeneration modules is implemented as additional features that can be traced back, as done here for the variants of ForClim and ForCEEPS, and that model complexity and structure must always be connected with modeling objectives [@albrich2020].
-If it should be necessary to include more detail in the regeneration models, this will come with higher parameterisation efforts.
-This will most likely lead to lower generalization because the required data will have to be collected from specific locations, as currently there is no general, comprehensive regeneration dataset available.
-
-Therefore, we further recommend that more effort should be invested into collecting harmonized datasets in a site-specific manner covering the different aspects leading to tree regeneration.
-We emphasize that datasets such as EuFoRIa are invaluable and should be expanded in both their spatial extent (e.g., towards boreal and Mediterranean conditions) as well as in time (e.g., continuing the monitoring into the future).
-Such data will allow for a better evaluation of forest models and help to reduce the uncertainty in their projections, which is crucial when they are used as tools for predicting e.g.
-the impacts of anthropogenic climate change.
-
-In the present study, we have considered tree regeneration in the equilibrium state only.
-It is equally important to understand how these models project tree regeneration after changes in forest structure by disturbances [@seidl2022], or under different management strategies [@lindner2000].
-However, this will require an entirely different set of observed data, and potentially not all models of forest dynamics would be able to assess the relationship of these aspects on tree regeneration, e.g.
-due to the lack of disturbance or appropriate management modules.
-
-We recommend to investigate in detail the implications of the current modeling strategies for tree regeneration and, ultimately, simulated forest stand structure.
-This applies particularly to the erroneous patterns of excess tree regeneration and later excess mortality, by focusing on a wider range of tree sizes and the related regeneration dynamics.
-We also recommend, especially for the landscape-level models, the inclusion of explicit spatial considerations regarding tree regeneration [@beckage2003]; this, however, is a serious challenge regarding appropriate datasets, as inventory data have wide coverage, but by definition do not allow for the assessment of spatial interaction effects.
-
-Exercises like the one presented here, where the models are operated in "blind flight" mode, i.e., without the possibility of being tuned towards capturing the expected patterns, should be repeated.
-Such benchmarking exercises should next focus on aspects such as specific model traits and the ecological formulations of particular (sub-)processes, to better understand the implications of the assumptions on which the models are based.
-Furthermore, the inclusion of a wide range of models with different scales, type, population structures, tree regeneration modules and complexity will ensure a wider benefit to the whole modeling community.
+With our study, we have demonstrated that models of forest dynamics need
+a focus on tree regeneration modules to make them more robust. It
+remains uncertain what level of detail is required to model regeneration
+processes, and this must be addressed in future research. We recommend
+that the improvement of the regeneration modules is implemented as
+additional features that can be traced back, as done here for the
+variants of ForClim and ForCEEPS, and that model complexity and
+structure must always be connected with modeling objectives
+[@albrich2020]. If it should be necessary to include more detail in the
+regeneration models, this will come with higher parameterisation
+efforts. This will most likely lead to lower generalization because the
+required data will have to be collected from specific locations, as
+currently there is no general, comprehensive regeneration dataset
+available.
+
+Therefore, we further recommend that more effort should be invested into
+collecting harmonized datasets in a site-specific manner covering the
+different aspects leading to tree regeneration. We emphasize that
+datasets such as EuFoRIa are invaluable and should be expanded in both
+their spatial extent (e.g., towards boreal and Mediterranean conditions)
+as well as in time (e.g., continuing the monitoring into the future).
+Such data will allow for a better evaluation of forest models and help
+to reduce the uncertainty in their projections, which is crucial when
+they are used as tools for predicting e.g. the impacts of anthropogenic
+climate change.
+
+In the present study, we have considered tree regeneration in the
+equilibrium state only. It is equally important to understand how these
+models project tree regeneration after changes in forest structure by
+disturbances [@seidl2022], or under different management strategies
+[@lindner2000]. However, this will require an entirely different set of
+observed data, and potentially not all models of forest dynamics would
+be able to assess the relationship of these aspects on tree
+regeneration, e.g. due to the lack of disturbance or appropriate
+management modules.
+
+We recommend to investigate in detail the implications of the current
+modeling strategies for tree regeneration and, ultimately, simulated
+forest stand structure. This applies particularly to the erroneous
+patterns of excess tree regeneration and later excess mortality, by
+focusing on a wider range of tree sizes and the related regeneration
+dynamics. We also recommend, especially for the landscape-level models,
+the inclusion of explicit spatial considerations regarding tree
+regeneration [@beckage2003]; this, however, is a serious challenge
+regarding appropriate datasets, as inventory data have wide coverage,
+but by definition do not allow for the assessment of spatial interaction
+effects.
+
+Exercises like the one presented here, where the models are operated in
+"blind flight" mode, i.e., without the possibility of being tuned
+towards capturing the expected patterns, should be repeated. Such
+benchmarking exercises should next focus on aspects such as specific
+model traits and the ecological formulations of particular
+(sub-)processes, to better understand the implications of the
+assumptions on which the models are based. Furthermore, the inclusion of
+a wide range of models with different scales, type, population
+structures, tree regeneration modules and complexity will ensure a wider
+benefit to the whole modeling community.
 
@@
 
-Models of forest dynamics are important tools in science and decision support, and the formulation of tree regeneration has strong implications for simulated forest properties.
-The 15 models and variants used here are facing similar challenges in their representation of tree regeneration: they generally overestimate tree regeneration levels, and the simulated regeneration niche is not always captured accurately as a function of biotic (light) and abiotic (temperature and moisture) factors.
-
-However, most models properly capture the diversity of the initial tree community, and differences between model formulations, e.g., the presence or absence of feedback from the adult trees, did not have a strong effect for capturing the species composition of regeneration.
-
-Regarding mortality in the early phase of tree life, many models that feature a particularly high overestimation of regeneration levels are compensating for this by a larger tree mortality.
-Often, this compensation is not sufficient to reduce the high regeneration levels to realistic values.
-Overall, there is no clear mortality pattern across all models.
-
-When capturing tree regeneration, the specific design decisions taken in the development of any model are more important for its behavior (accuracy) than scale (stand, landscape, global), modelling approach (empirical vs.
-process-based), and complexity.
-Having both empirical and process-based models in our set, the empirically-based models could have been expected to have a better performance, as they were calibrated with inventory data, but this was not the case.
-Similarly, higher model complexity does not represent an improvement for capturing tree regeneration.
-
-Even though the regeneration routines of most of the models investigated here have never been sufficiently constrained by robust data, the regeneration projections are not overly off.
-This indicates that a lot can be gained by a focus on the modeling of regeneration processes.
-The representation of forest dynamics in these models would become much more robust particularly in the face of climate change and post-disturbance dynamics, thus strongly reducing the uncertainty in long-term projections of future forest dynamics.
+Models of forest dynamics are important tools in science and decision
+support, and the formulation of tree regeneration has strong
+implications for simulated forest properties. The 15 models and variants
+used here are facing similar challenges in their representation of tree
+regeneration: they generally overestimate tree regeneration levels, and
+the simulated regeneration niche is not always captured accurately as a
+function of biotic (light) and abiotic (temperature and moisture)
+factors.
+
+However, most models properly capture the diversity of the initial tree
+community, and differences between model formulations, e.g., the
+presence or absence of feedback from the adult trees did not have a
+strong effect on capturing the species composition of regeneration.
+
+Regarding mortality in the early phase of tree life, many models that
+feature a particularly high overestimation of regeneration levels are
+compensating for this by a larger tree mortality. Often, this
+compensation is not sufficient to reduce the high regeneration levels to
+realistic values. Overall, there is no clear mortality pattern across
+all models.
+
+When capturing tree regeneration, the specific design decisions taken in
+the development of any model are more important for its behavior
+(accuracy) than scale (stand, landscape, global), modelling approach
+(empirical vs. process-based), and complexity. Having both empirical and
+process-based models in our set, the empirically-based models could have
+been expected to have a better performance, as they were calibrated with
+inventory data, but this was not the case. Similarly, higher model
+complexity does not represent an improvement for capturing tree
+regeneration.
+
+Even though the regeneration routines of most of the models investigated
+here have never been sufficiently constrained by robust data, the
+regeneration projections are not overly off. This indicates that a lot
+can be gained by a focus on the modeling of regeneration processes. The
+representation of forest dynamics in these models would become much more
+robust particularly in the face of climate change and post-disturbance
+dynamics, thus strongly reducing the uncertainty in long-term
+projections of future forest dynamics.
 
@@
 
-O.
-Díaz-Yáñez and Y.
-Käber contributed equally and share the first authorship.
-H.
-Bugmann, Y.
-Käber, and O.
-Díaz-Yáñez developed the study design.
-All authors contributed to preparing the simulations with their respective models.
-O.
-Díaz-Yáñez led the data analysis with contributions during the workshop from Y.Käber, T.
-Anders, K.
-H.
-Braziunas, J.
-Brůna, S.
-M.
-Fischer, J.
-Hetzer, T.
-Hickler, H.
-Lischke, M.Mahnken, P.
-Mairota, K.
-Merganičová, T.
-Mette, X.
-Morin, W.
-Rammer, D.
-Scherrer and H.
-Bugmann.
-O.
-Díaz-Yáñez and H.
-Bugmann led the writing of the manuscript.
-All authors participated in the revision of the manuscript and approved its submission.
+O. Díaz-Yáñez and Y. Käber contributed equally and share the first
+authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study
+design. All authors contributed to preparing the simulations with their
+respective models. O. Díaz-Yáñez led the data analysis with
+contributions during the workshop from Y.Käber, T. Anders, K. H.
+Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T. Hickler, H. Lischke,
+M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X. Morin, W. Rammer, D.
+Scherrer and H. Bugmann. O. Díaz-Yáñez and H. Bugmann led the writing of
+the manuscript. All authors participated in the revision of the
+manuscript and approved its submission.
 
@@
 
-This paper arises from a workshop held in Davos (Switzerland) in June 2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
-Funding for the workshop was provided by COST Action CA19139 PROCLIAS (PROcess-based models for CLimate Impact Attribution across Sectors), supported by COST (European Cooperation in Science and Technology; [https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research Grant ETH-35 18-1).
-ODY was funded by the Swiss National Science Foundation project 200020_188882 / 1.
-KB received support from the European Research Council under the European Union's Horizon 2020 research and innovation program (Grant Agreement 101001905).
-JB acknowledges funding from the long-term research development project no.
-RVO 67985939 from the Czech Academy of Sciences.
-MM acknowledges financial support from I-Maestro (Innovative forest management strategies for a resilient bioeconomy under climate change and disturbances, grant no.
-22035418, 2019--2022) funded by the ERA-NET Cofund ForestValue.
-MMina acknowledges funding from the H2020-MSCA-IF project REINFORCE (Grant No.
-891671).
-KM received funding from the Project No.
-CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant no.
-ITMS2014+313011W580 supported by the Integrated Infrastructure Operational Programme funded by the ERDF.
-CH acknowledges financial support from Waldfonds WASIM2100 project.
-We would also like to thank all the people that helped to make this study possible with contributions at different stages such as Martin Gutsch and Giorgio Vacchiano.
+This paper arises from a workshop held in Davos (Switzerland) in June
+2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
+Funding for the workshop was provided by COST Action CA19139 PROCLIAS
+(PROcess-based models for CLimate Impact Attribution across Sectors),
+supported by COST (European Cooperation in Science and Technology;
+[https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research
+Grant ETH-35 18-1). ODY was funded by the Swiss National Science
+Foundation project 200020_188882 / 1. KB received support from the
+European Research Council under the European Union's Horizon 2020
+research and innovation program (Grant Agreement 101001905). JB
+acknowledges funding from the long-term research development project no.
+RVO 67985939 from the Czech Academy of Sciences. MM acknowledges
+financial support from I-Maestro (Innovative forest management
+strategies for a resilient bioeconomy under climate change and
+disturbances, grant no. 22035418, 2019--2022) funded by the ERA-NET
+Cofund ForestValue. MMina acknowledges funding from the H2020-MSCA-IF
+project REINFORCE (Grant No. 891671). KM received funding from the
+Project No. CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant
+no. ITMS2014+313011W580 supported by the Integrated Infrastructure
+Operational Programme funded by the ERDF. CH acknowledges financial
+support from Waldfonds WASIM2100 project. We would also like to thank
+all the people that helped to make this study possible with
+contributions at different stages such as Martin Gutsch and Giorgio
+Vacchiano.
 
@@
 
-**Table 1.** Models included in this study, their characteristics and simulation strategies used.
-Scale refers to the models scale, Type to the models type, Population structure to the model population approach, Feedback to the inclusion or not of adult tree feedback to the level and species composition of regeneration, Approach to the model regeneration module approach, Start from to the stage from which trees are recruited, Complexity, to the mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), Species to which species from the original eleven species were simulated, Runtime (spin up) [years] to the number of years used per model in their simulations for the spin up period and regeneration sampling; Climate data type (use of years) to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details.
-Further information can be found on the individual model reports available in the supplementary materials.
-
-**Figure 1.** The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed regeneration at the 10 cm threshold.
-
-**Figure 2.** Mean regeneration levels across all samples per site, plotted for the 200 sites and for each model.
-The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
-There are two boxplots for each model where the left lighter boxplot corresponds to 7 cm and the right darker boxplot to 10 cm
-
-**Figure 3.** Shannon index of tree regeneration, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
-The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
-
-**Figure 4.** Mean Shannon index across all samples per site for observed and simulated data.
-Each plot shows one pattern represented by one exemplary model of each category (overpredicted, intermediate and underpredicted).
-The full data with the grouping of the models are shown in Figures S3 and S4.
-n indicates the number of models falling in each group.
-A) Examples for the three trends across models for species diversity at the stand level.
-B) Examples for the three trends across models for regeneration at the 7 cm threshold (165 sites).
-
-**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm thresholds.
-Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree regeneration between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.
-
-**Figure 6.** Ratio of tree regeneration between the 7 and 10 cm thresholds (regeneration 7 cm / regeneration 10 cm) and the overestimation proportion at 7 cm ([simulated - observed] / observed) for the mean regeneration per model across sites and samples.
-The horizontal dashed lines show a ratio equal to 1, indicating no decrease of regeneration between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
-The vertical line indicates zero overestimation.
-
-**Figure 7.** Mean regeneration values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum.
-The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
-Y-axis are not equally scaled.
-
-**Figure 8.** Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold.
-The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient.
-The sizes of the circles represent the ratio between the regeneration basal area of the species and the total regeneration basal area for all species.
-The absence of a circle indicates a zero basal area share in the regeneration, or the absence of regeneration altogether.
-The color gradient (for the models only) shows the difference between the simulated and observed ratio of regeneration basal area of the species and the total regeneration basal area for all the species.
+**Table 1.** Models included in this study, their characteristics and
+simulation strategies used. *Scale* refers to the models scale, *Type*
+to the models type, *Population structure* to the model population
+approach, *Feedback* to the inclusion or not of adult tree feedback to
+the level and species composition of regeneration, *Approach* to the
+model regeneration module approach, *Start from* to the stage from which
+trees are recruited, *Complexity*, to the mean regeneration formulation
+as estimated in (Bugmann & Seidl, 2022), *Species* to which species from
+the original eleven species were simulated, *Runtime* (spin up) [years]
+to the number of years used per model in their simulations for the spin
+up period and regeneration sampling; *Climate data type (use of years)*
+to the type of data used in the simulations and how it was used, and
+*Reference* provides the link to the publication with more model
+details. Further information can be found on the individual model
+reports available in the supplementary materials.
+
+**Figure 1.** The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used
+in the present study, the colors show the dominant tree species per site
+in the observed regeneration at the 10 cm threshold.
+
+**Figure 2.** Mean regeneration levels across all samples per site,
+plotted for the 200 sites and for each model. The red dashed lines shows
+the 25th and 75th percentiles for the 7 cm diameter threshold in the
+observed data. There are two boxplots for each model where the left
+lighter boxplot corresponds to 7 cm and the right darker boxplot to 10
+cm
+
+**Figure 3.** Shannon index of tree regeneration, calculated by basal
+area, as the mean value across the 200 samples per site for the diameter
+thresholds of 7 and 10 cm, respectively. The red dashed lines shows the
+25th and 75th percentiles for the 7 cm diameter threshold in the
+observed data.
+
+**Figure 4.** Mean Shannon index across all samples per site for
+observed and simulated data. Each plot shows one pattern represented by
+one exemplary model of each category (overpredicted, intermediate and
+underpredicted). The full data with the grouping of the models are shown
+in Figures S3 and S4. n indicates the number of models falling in each
+group. A) Examples for the three trends across models for species
+diversity at the stand level. B) Examples for the three trends across
+models for regeneration at the 7 cm threshold (165 sites).
+
+**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm
+thresholds. Dashed blue lines mark a ratio equal to 1, indicating no
+decrease in tree regeneration between 7 and 10 cm, and a ratio equal to
+1.77, corresponding to the Reineke self-thinning ratio under even-aged
+conditions.
+
+**Figure 6.** Ratio of tree regeneration between the 7 and 10 cm
+thresholds (regeneration 7 cm / regeneration 10 cm) and the
+overestimation proportion at 7 cm ([simulated - observed] / observed)
+for the mean regeneration per model across sites and samples. The
+horizontal dashed lines show a ratio equal to 1, indicating no decrease
+of regeneration between 7 and 10 cm, and a ratio equal to 1.77
+corresponding to the Reineke self-thinning line under even-aged
+conditions. The vertical line indicates zero overestimation.
+
+**Figure 7.** Mean recruitment across the 200 samples per site with the
+y-axis scaled differently by model, for the 200 sites against gradients
+of (a) total basal area; (b) climatic water balance; (c) seasonal
+degree-day sum. The values were split into ten bins; the red lines
+represent a GAM model showing the trend in the observed data. Y-axis are
+not equally scaled.
+
+**Figure 8.** Regeneration levels of the five main species in terms of
+their basal area share in the observations (top row) and the performance
+of each model across the environmental gradients (other rows) for the 7
+cm regeneration threshold. The values shown are the mean of the 200
+samples per site and across the sites in each bin (tile), with ten bins
+per gradient. The sizes of the circles represent the ratio between the
+regeneration basal area of the species and the total regeneration basal
+area for all species. The absence of a circle indicates a zero basal
+area share in the regeneration, or the absence of regeneration
+altogether. The color gradient (for the models only) shows the
+difference between the simulated and observed ratio of regeneration
+basal area of the species and the total regeneration basal area for all
+the species.
 

### 2023-10-19 manuscript/manuscriptRW.Rmd
change figures names and remove them from manuscript

@@
 Europe as compiled in the framework of the EuFoRIa network
-[@euforia2019; @käber2023] (Figure \@ref(fig:map)). The environmental
-coverage and the variety of forests represented by this data is
-unprecedented within Europe. These forests have been unmanaged for at
-least 10 years prior to their designation as reserves, and most of them
-provide time series of natural forest dynamics over multiple decades and
-up to ca. 80 years. The census periods range from 3 to 37 years, with an
-average of 14 years. The data provide information at the tree level,
-thus allowing for the sequential comparison of processes such as
-individual-tree regeneration and death between the measurements.
+[@euforia2019; @käber2023] (Figure 1). The environmental coverage and
+the variety of forests represented by this data is unprecedented within
+Europe. These forests have been unmanaged for at least 10 years prior to
+their designation as reserves, and most of them provide time series of
+natural forest dynamics over multiple decades and up to ca. 80 years.
+The census periods range from 3 to 37 years, with an average of 14
+years. The data provide information at the tree level, thus allowing for
+the sequential comparison of processes such as individual-tree
+regeneration and death between the measurements.
 
@@
 typically did not match the levels found in the forest reserve data
-(Figure \@ref(fig:over)). Recruitment was overestimated in most models
-for both the 7 and 10 cm diameter thresholds, with the exception of the
-empirical stand model SIBYLA, the landscape model Landis-II and the
-global model aDGVM2, which estimated recruited levels at the lower end
-of the plausibility interval of the observed data (Figure
-\@ref(fig:over)). The models with the largest overestimation were the
-stand model PICUS and the landscape model TreeMig. For most models, the
-variability of simulated recruitment levels across the 200 sites
-(visible from the interquartile range in the box plots of Figure
-\@ref(fig:over)) was similar to or smaller than observed, with the
+(Figure 2). Recruitment was overestimated in most models for both the 7
+and 10 cm diameter thresholds, with the exception of the empirical stand
+model SIBYLA, the landscape model Landis-II and the global model aDGVM2,
+which estimated recruited levels at the lower end of the plausibility
+interval of the observed data (Figure 2). The models with the largest
+overestimation were the stand model PICUS and the landscape model
+TreeMig. For most models, the variability of simulated recruitment
+levels across the 200 sites (visible from the interquartile range in the
+box plots of Figure 2) was similar to or smaller than observed, with the
 notable exception of PICUS, where simulated recruitment variability was
@@
 Most models matched the level of diversity of the observed data quite
-well (Figure \@ref(fig:H710)). Five models overestimated recruitment
-diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
+well (Figure 3). Five models overestimated recruitment diversity:
+ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
 particularly for the 7 cm diameter threshold. The model 4C is a special
@@
 and S2). Most models captured reasonably well or overpredicted species
-diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left
-and center; cf. Figure S3), but a characteristic feature was that the
-simulations had a much lower variability of diversity compared to
-observations. Models overpredicting species diversity included ForClim1,
-ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A,
-left). Several models underpredicted species diversity, i.e., LPJ-GUESS,
-ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).
+diversity at the stand level (Figure 4A, left and center; cf. Figure
+S3), but a characteristic feature was that the simulations had a much
+lower variability of diversity compared to observations. Models
+overpredicting species diversity included ForClim1, ForClim 11, FORMIND
+and TreeMig (Figure 4A, left). Several models underpredicted species
+diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure 4A, right).
 
 The majority of the models overpredicted recruitment diversity; as
-already observed in Figure \@ref(fig:H710), only two of them
-underestimated it (4C and Landis-II) (Figure
-\@ref(fig:combinedRichness)B and Figure S4). Both the models with
-feedback and those without feedback from the adult trees to regeneration
-(via seed production) overestimated, underestimated or captured
-reasonably well recruitment species diversity compared to the
-observations. Again, simulated diversity had considerably lower
-variability than observed diversity in the recruitment.
+already observed in Figure 3, only two of them underestimated it (4C and
+Landis-II) (Figure 4B and Figure S4). Both the models with feedback and
+those without feedback from the adult trees to regeneration (via seed
+production) overestimated, underestimated or captured reasonably well
+recruitment species diversity compared to the observations. Again,
+simulated diversity had considerably lower variability than observed
+diversity in the recruitment.
 
@@
 There were strong differences among the models regarding the mortality
-rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the
-models featured a mortality rate significantly larger than the observed
-data, thus at least partially compensating for the general
-overestimation of recruitment levels (Figure \@ref(fig:over) and Table
-S3); six models did not feature a significant difference compared to the
-observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
-Four models (xComp, PICUS, LandClim and TreeMig) featured very high
-mortality rates (i.e., well above the Reineke self-thinning line), which
-compensated for the strong initial overestimation of recruitment (cf.
-Figure \@ref(fig:over)). Interestingly, two models that underestimated
-overall recruitment levels (SIBYLA and Landis-II; cf. Figure
-\@ref(fig:over)) featured mortality rates that were close to but still
-above the self-thinning line (Figure \@ref(fig:mortality11)).
+rate between 7 and 10 cm (Figure 5). Most of the models featured a
+mortality rate significantly larger than the observed data, thus at
+least partially compensating for the general overestimation of
+recruitment levels (Figure 2 and Table S3); six models did not feature a
+significant difference compared to the observed data (ForCEEPS, FORMIND,
+ForClim 1 and 11, iLand and aDGVM2). Four models (xComp, PICUS, LandClim
+and TreeMig) featured very high mortality rates (i.e., well above the
+Reineke self-thinning line), which compensated for the strong initial
+overestimation of recruitment (cf. Figure 2). Interestingly, two models
+that underestimated overall recruitment levels (SIBYLA and Landis-II;
+cf. Figure 2) featured mortality rates that were close to but still
+above the self-thinning line (Figure 5).
 
@@
 recruitment at the 7 cm threshold featured the highest mortality between
-the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The other
-models that overestimated recruitment had mortalities lower than the
-expected self-thinning ratio, with the exception of LandClim (Figure
-\@ref(fig:mortality13)B). It is noteworthy that most of the models that
-featured a low ratio of regeneration between 7 and 10 cm (i.e.,
-ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small
-overestimation of recruitment at the 7 cm threshold (Figure
-\@ref(fig:mortality11)). aDGVM2 was the only model that underestimated
+the 7 and 10 cm thresholds (Figure 6A). The other models that
+overestimated recruitment had mortalities lower than the expected
+self-thinning ratio, with the exception of LandClim (Figure 6B). It is
+noteworthy that most of the models that featured a low ratio of
+regeneration between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11
+and iLand) had only a small overestimation of recruitment at the 7 cm
+threshold (Figure 5). aDGVM2 was the only model that underestimated
 recruitment at the 7 cm threshold while featuring a low mortality
-between the two thresholds (Figure \@ref(fig:mortality13)C), and only a
-few models were close to the correct recruitment levels at 7 cm while
-also being close to the Reineke line (SIBYLA, Landis-II; cf. Figure
-\@ref(fig:mortality13)D; and 4C as well as LandClim).
+between the two thresholds (Figure 6C), and only a few models were close
+to the correct recruitment levels at 7 cm while also being close to the
+Reineke line (SIBYLA, Landis-II; cf. Figure 6D; and 4C as well as
+LandClim).
 
@@
 reproduced the effect of basal area in both its magnitude and patterns
-(Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better
-than the effects of the climatic gradients, where they featured varying
-patterns.
+(Figure 7 and S7) considerably better than the effects of the climatic
+gradients, where they featured varying patterns.
 
 In the observed data, recruited levels decreased clearly with increasing
-stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a). This trend
-was captured by the Landis-II model only, albeit at much lower values of
-basal area than in the observations. The other models featured
-distinctly different trends, such as (1) an increase in recruited levels
-with increasing total basal area followed by the absence of regeneration
-at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig
-and LPJ-GUESS), (2) almost constant recruitment levels with basal area
-(ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of
-recruitment up to a certain value of stand basal area followed by a
-decrease at even higher values of stand basal area, with model-specific
-thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).
+stand basal area (Figure 7a). This trend was captured by the Landis-II
+model only, albeit at much lower values of basal area than in the
+observations. The other models featured distinctly different trends,
+such as (1) an increase in recruited levels with increasing total basal
+area followed by the absence of regeneration at high values of basal
+area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost
+constant recruitment levels with basal area (ForClim 11, xComp, LandClim
+and aDGVM2), or (3) an increase of recruitment up to a certain value of
+stand basal area followed by a decrease at even higher values of stand
+basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA
+and iLand).
 
@@
 models did not match the observed recruitment patterns across the
-gradients of soil moisture and temperature (Figure
-\@ref(fig:nvTrend710BAlightTemp)b,c), and neither was there a clear
-pattern across models. Regarding the soil moisture gradient, a group of
-models featured decreasing recruitment with increasing soil moisture
-(4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and
-iLand), while a few models showed the opposite trend (ForCEEPS(f),
-FORMIND and LPJ-GUESS). Across the temperature gradient, all the models
-that featured a decrease in recruitment levels with increasing soil
-moisture showed the opposite trend (i.e., more recruitment with
-increasing temperature), except for 4C. The other models did not feature
-a clear pattern.
+gradients of soil moisture and temperature (Figure 7b,c), and neither
+was there a clear pattern across models. Regarding the soil moisture
+gradient, a group of models featured decreasing recruitment with
+increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp,
+PICUS, LandClim, TreeMig and iLand), while a few models showed the
+opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS). Across the
+temperature gradient, all the models that featured a decrease in
+recruitment levels with increasing soil moisture showed the opposite
+trend (i.e., more recruitment with increasing temperature), except for
+4C. The other models did not feature a clear pattern.
 
@@
 basal area in the recruitment of these five species compared to the
-observed data along the two major climatic gradients (Figure
-\@ref(fig:diff), cf. Figure S8). However, the models differed
-considerably in the way in which they simulated these climatic niches.
+observed data along the two major climatic gradients (Figure 8, cf.
+Figure S8). However, the models differed considerably in the way in
+which they simulated these climatic niches.
 
 First, some models overestimated recruitment across the entire gradients
-of at least one of the five main tree species (Figure \@ref(fig:diff),
-e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models
-overestimated recruitment of more than one species for a large part of
-the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
-red colors in Figure \@ref(fig:diff)). These trends were consistent for
-both recruitment thresholds, although the exact changes in the share of
-the recruited basal area were different (Figure S8).
+of at least one of the five main tree species (Figure 8, e.g., ForCEEPS,
+ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated
+recruitment of more than one species for a large part of the climate
+gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf. red colors in
+Figure 8). These trends were consistent for both recruitment thresholds,
+although the exact changes in the share of the recruited basal area were
+different (Figure S8).
 
@@
 regarding the species-specific recruitment levels in the environmental
-space (Figure \@ref(fig:diff)). Some models (e.g., ForClim11, PICUS)
-tended to consistently overestimate the recruitment share of all five
-species, but most models overestimated the recruitment share of some
-species while underestimating the share of others (e.g., ForCEEPS,
-LPJ-GUESS).
+space (Figure 8). Some models (e.g., ForClim11, PICUS) tended to
+consistently overestimate the recruitment share of all five species, but
+most models overestimated the recruitment share of some species while
+underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
 
@@
 
-```{r map, echo=FALSE, fig.cap=".", out.width='80%'}
-knitr::include_graphics(here::here("figures", "theMap.png"))
-```
-
-\newpage
-
-```{r over, echo=FALSE, fig.cap=".", out.width='100%'}
-        knitr::include_graphics(here::here("figures", "recOverAll710.png"))
-```
-
-\newpage
-
-```{r H710, echo=FALSE, fig.cap=".", out.width='100%'}
-    knitr::include_graphics(here::here("figures", "H7_10.png"))
-```
-
-\newpage
-
-```{r combinedRichness, echo=FALSE, fig.cap=".", out.width='75%'}
-knitr::include_graphics(here::here("figures", "combined_richness.png"))
-```
-
-\newpage
-
-```{r mortality11, echo=FALSE, fig.cap=".", out.width='75%'}
-knitr::include_graphics(here::here("figures", "ratio7_10.png"))
-```
-
-\newpage
-
-```{r mortality13, echo=FALSE, fig.cap=".", out.width='80%'}
-knitr::include_graphics(here::here("figures", "mort7_10_medianOver.png"))
-```
-
-\newpage
-
-```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap=".", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}
-
-   knitr::include_graphics(c(here::here("figures", "envTrend7Totba.png"),
-                              here::here("figures", "envTrend7WB.png"),
-                              here::here("figures", "envTrend7DDS.png")))
-```
-
-\newpage
-
 ```{r diff, echo=FALSE, fig.cap=".", out.width='100%', out.height= '85%'}

### 2023-10-19 code/figures.R

@@ -8,7 +8,7 @@
 
 # Load plotting styles 
 
-overDTsim    source("code/plottingManuscript.R")
+    source("code/plottingManuscript.R")
 
 # Load data ----
 
@@ -223,6 +223,9 @@ overDTsim    source("code/plottingManuscript.R")
                                       group.by = "model", method = "t.test")
     
     write.csv(sigH7_10,"figures/sigH7_10.csv", row.names = FALSE)
+    
+    
+    # GLME 
     
    
     ### Fig. H obs. VS sim. in stand and recruitment  ----
@@ -818,9 +821,18 @@ overDTsim    source("code/plottingManuscript.R")
     
   
     rtreesDiff <- ggpubr::compare_means(r.trees ~ model,  
-                                    data = overDTsim, 
+                                    data = overDTsim[overDTsim$dbh == 7,], 
                                     ref.group = "Observed", method = "t.test")
-    write.csv(rtreesDiff ,"figures/rtreesDiff.csv", row.names = FALSE)
+    write.csv(rtreesDiff, "figures/rtreesDiff.csv", row.names = FALSE)
+    
+    #### GLME -----
+    
+    rtreesDiffGLME <- lme4::glmer(r.trees ~ model + (1 | site), 
+                                  family = poisson(link = "log"), 
+                                  data = overDTsim[overDTsim$dbh == 7,],
+                                  verbose = 1, optimizer="nloptwrap")
+    
+  
     
    
 ### feedback / not feedback H simulated vs observed -----  

### 2023-10-19 manuscript/review/authorsResponse.qmd
last changes before sent for internal round

@@
   - name: Tobias Mette,
-  - name: Marco Mina$,
+  - name: Marco Mina,
   - name: Xavier Morin,
@@
 
-# Internal instructions
+# Internal instructions \[this section will be deleted\]
 
 -   underlined comments are comments that still need attention
-
-# Journal formatting (ToDo list)
+-   I will delete most of the quoted text and add Line numbers, all the L\[XX-XX\] will be added in the last round after all the changes are done
+
+## Journal formatting (ToDo list)
 
@@
 Thank you for taking the time to review our manuscript.
-We appreciate your thoughtful comments and have carefully considered each of them.
 We have made changes to the manuscript based on your feedback and believe that these changes have improved the overall quality of our manuscript.
@@
 
-[Pending]{.underline}
+The objective of running the different t-tests were to allow us for discussing irrefutable differences between models and observations.
+The t-test, as employed in our study, is not intended to serve as definitive proof of our findings.
+Rather, it serves as a practical tool for managing the extensive number of comparisons in our study.
+As such, we intend to include a statement in the manuscript:
+
+"Note that the t-test does not present conclusive evidence for a specific hypotheses, but rather facilitates managing the extensive number of comparisons discussed in our study."
+
+Regarding the other alternative approaches suggested by the reviewer: We can not do a paired t-test as the for paired samples it is necessary that you have the same number of subjects in both groups (observed and simulated) which is not the case.
+We tested the possibility of creating a GLME using as random effect the site and as a explanatory variables all the variables that we were interested in (model complexity, feedback, model type and scale).
+This model approach did not work as the combination as including some models having the same values in the combination of type and scale (factors did not have 2 or more levels).
+
+-   "This more accurately answer your question about if any of these fixed explanatory variables can predict over or under representation of recruitment. " how to respond to this?
+
+Account for Type I error rates and adjust your p-values accordingly
 
@@
 
-Response to comment 20:
+**Response to comment 20:**
 
@@
 
-Response to comment 21:
+**Response to comment 21:**
 
@@
 
-Response to comment 24:
+**Response to comment 24:**
 
@@
 
-We would like to thank Reviewer #2 for this positive overall assessment and the level of detail of the review.
-We have taken care to go through all suggestions made in the PDF, and have changed/improved the manuscript in most instances; however these suggestions are too numerous to be dealt with on an individual basis here.
+We would like to thank Reviewer 2 for this positive overall assessment and the level of detail of the review.
+We have taken care to go through all suggestions made in the PDF, and have changed/improved the manuscript in most instances.
+Our responses to each of the comments and the concrete actions we have taken are detailed below.
 
@@
 
-Response to comment 5:
+**Response to comment 5:**
 
@@
 
-Response to comment 6:
+**Response to comment 6:**
 
@@
 
-Response to comment 14:
-
-We are not complete sure of the reviewers point here.
+**Response to comment 14:**
+
+We are not completely sure of the reviewers point here.
 If the reviewer is referring to the fact that the most important processes will be those included in the model, it is true that you can only asses those processes included or how including or not another processes impact regeneration.
@@
 
-### [COMMENT 15]{.underline}
+### COMMENT 15
 
@@
 
-[ToDo find a reference! Any suggestions?]{.underline}
-
-### [COMMENT 16]{.underline}
+We have added a reference here.
+
+### COMMENT 16
 
@@
 
-[ToDo find a reference!Any suggestions?]{.underline}
+We have added a reference here.
 
@@
 
-Response to comment 19:
+**Response to comment 19:**
 
@@
 
-### [COMMENT 20]{.underline}
+### COMMENT 20
 
@@
 
-[Is there a way to explain this shortly \@ Yannek ??]{.underline}
+We have added information to briefly explain the general approach.
 
@@
 
-Response to comment 21:
+**Response to comment 21:**
 
 We agree with the reviewer that permanent plots such as those present in National Forest Inventories could potentially be a good source of information to obtain regeneration data.
-However these often are collected from managed sites, do not capture ingrowth below 10cm, and are hard to harmonized between different inventory strategies.
+However these often are collected from managed sites, do not capture ingrowth below 10cm, and are hard to harmonize between different inventory strategies.
 
@@
 
-Response to comment 41:
+Response to comment 38:
 
@@
 
-Response to comment 42:
+Response to comment 39:
 
@@
 
-Response to comment 43:
-
-We have change the terminology.
+Response to comment 40:
+
+We have changed the terminology.
 
@@
 
-Response to comment 41:
+**Response to comment 41:**
 
@@
 We agree with the reviewer that the Shannon index is very well known index and its equation as well.
-However we think is still important that we define how we did calculate the Shannon index based on the available data we had as we had some internal discussion on how to do this and the meaning of this index in this context, we though the reader will benefit from a more detailed explanation.
+However we think is still important that we define how we did calculate the Shannon index based on the available data we had as we had some internal discussion on how to do this and the meaning of this index in this context, we thought the reader will benefit from a more detailed explanation.
 In any case the exact details of how we did calculate this index based on the simulation outputs are also available in the software supplementary materials, in the script figures.R L42-78.
@@
 
-###
-
 ### COMMENT 51
@@
 
-Response to comment 52:
+**Response to comment 52:**
 
@@
 This had also been defined in the material and methods where we stated that "i.e., where the level and species composition of regeneration is influenced by the existing adult tree community via the production of seeds, seedlings or saplings, from models that do not contain such a feedback."
-
-###
 
@@
 We agree with the reviewers statement that we need the "right amount" of complexity which considers the most relevant processes.
-however there are limited studies looking at this aspect in models of forest dynamics and even more limited looking at this aspect in the regeneration related processes in models of forest dynamics.
+However there are limited studies looking at this aspect in models of forest dynamics and even more limited looking at this aspect in the regeneration related processes in models of forest dynamics.
 We still think it is important to discuss this issue in the context of our findings even though this was not the solely focus of our research.
-
-###
 
@@
 
-Response to comment 56:
+**Response to comment 56:**
 
@@
 
-###
-
 ### COMMENT 57
@@
 
-Response to comment 57:
-
-In this sentence we are discussing how the sampling strategy presented in the study protocol might have represented a disadvantage for landscape level models due to the limited spatial scale, but how this might not be the same for Global level models as they usually lack dispersal between cells.
+**Response to comment 57:**
+
+In this sentence we are discussing how the sampling strategy presented in the study protocol might have represented a disadvantage for landscape level models due to the limited spatial scale, but how this might not be the same for global level models as they usually lack dispersal between cells.
 
@@
 
-Response to comment 58:
+**Response to comment 58:**
 
@@
 
-###
-
 ### COMMENT 60
@@
 
-Response to comment 60:
+**Response to comment 60:**
 
@@
 "Models included, their characteristics and simulation strategies used in this study.
-Where Scale refers to the models scale, *Type* to the models type, *Population structure* to the model population approach, *Feedback* to the inclusion or not of adult tree feedback to the level and species composition of regeneration, *Approach* to the model regeneration module approach, *Start from* to the stage from which trees are recruited, *Complexity,* to *the* mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), *Species* to which species from the original eleven species were simulated , *Runtime (spin up) \[years\]* to the number of years used per model in their simulations for the spin up period and regeneration sampling; *Climate data type (use of years)* to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details.
+Where *Scale* refers to the models scale, *Type* to the models type, *Population structure* to the model population approach, *Feedback* to the inclusion or not of adult tree feedback to the level and species composition of regeneration, *Approach* to the model regeneration module approach, *Start from* to the stage from which trees are recruited, *Complexity,* to *t*he mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), *Species* to which species from the original eleven species were simulated, *Runtime (spin up) \[years\]* to the number of years used per model in their simulations for the spin up period and regeneration sampling; *Climate data type (use of years)* to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details.
 Further information can be found on the individual model reports available in the supplementary materials."

### 2023-10-19 manuscript/supplementaryRW.Rmd
last changes before sent for internal round

@@
 TreeMig simulates directly the mean dynamics of entire stands, but based
-on the variability of light conditions within the stand [distribution
-based approach, Lischke, 1998, aggregation. To come up with the required
+on the variability of light conditions within the stand (distribution based approach, cf. Lischke et al. 1998). To come up with the required
 200 random patches of 1 ha, temporal and spatial variability intrinsic

### 2023-10-19 code/figures.R

@@ -159,11 +159,11 @@
                                                                na.rm = TRUE)),
                             colour = "#990000", linetype = "dashed") 
     
-    ggplot2::ggsave("figures/recOverAll710.png",
+    ggplot2::ggsave("figures/Figure2.jpg", 
                     plot =  recOverAll,
                     width = 21, height = 12, scale = 0.9,
                     dpi = 300, units = "cm", 
-                    device = 'png') 
+                    device = 'jpg') 
     
 # Diversity in recruitment ----
 
@@ -210,10 +210,10 @@
                                                                c(0.75))),
                             colour = "#990000", linetype = "dashed") 
     
-    ggplot2::ggsave("figures/H7_10.png",
+    ggplot2::ggsave("figures/Figure3.jpg",
                     plot =   H7_10,
                     width = 21, height = 12, scale = 0.9,
-                    dpi = 300, units = "cm", device = 'png')
+                    dpi = 300, units = "cm", device = 'jpg')
     
     
     # Significant difference per model between 7 and 10 diameter threshold
@@ -351,11 +351,11 @@
                                   bottom = grid::grid.text("Observed", 
                                                            gp = grid::gpar(cex = 1.3)))
     
-    ggplot2::ggsave("figures/combined_richness.png",
+    ggplot2::ggsave("figures/Figure4.jpg",
                     plot = xx,
                     dpi = 300, 
                     width = 20, height = 18, scale = 0.9,
-                    units = "cm", device = 'png')
+                    units = "cm", device = 'jpg')
     
     
     #### FigS. rec vs all 7cm ----
@@ -568,10 +568,10 @@
                                  values =  values_color[names(values_color) %in% unique(mortAll$model)],
                                  guide = "none")
     
-    ggplot2::ggsave("figures/ratio7_10.png",
+    ggplot2::ggsave("figures/Figure5.jpg",
                     plot = ratio7_10,
                     width = 21, height = 12, scale = 0.9,
-                    dpi = 300, units = "cm", device = 'png') 
+                    dpi = 300, units = "cm", device = 'jpg') 
     
     
     mortAll$nn710 <-  as.numeric(as.character(mortAll$nn710))
@@ -649,10 +649,10 @@
[…]

### 2023-10-24 manuscript/significTest.qmd
first analysis

---
title: "signifTest"
format: pdf
editor: visual
---

## Test of significant differences between observed and models simulations and between 7 and 10 cm

### Summary (potential response to reviewer)
The objective of running the different t-tests were to allow us for discussing irrefutable differences between mean values per site between models and observations and between 7 and 10cm diameter thresholds. The t-test, as employed in our study, is not intended to serve as definitive proof of our findings and we have deliveratively avoid over-interpretation of their outcomes. Rather, it serves as a practical tool for managing the extensive number of comparisons in our study. As such, we intend to include a statement in the manuscript:

"Note that the t-test does not present conclusive evidence for a specific hypotheses, but rather facilitates managing the extensive number of comparisons discussed in our study."

Following the reviewers comment we included a correction for multiple testing although we are aware that all corrections have pros and cons. We use the Bonferroni correction to make sure that the actual Type I error probability for the combined H0 is smaller or equal than the nominal test level.

Regarding the other alternative approaches suggested by the reviewer: We can not do a paired t-test as the for paired samples it is necessary that you have the same number of subjects in both groups (observed and simulated) which is not the case. We tested the possibility of creating a GLME using as random effect the site and as a explanatory variables all the variables that we were interested in (model complexity, feedback, model type and scale) as suggested. However this approach will not give us the observed-model comparison that we wanted but a if any of these fixed explanatory variables can predict over or under representation of recruitment, which was not our intended goal.

Following the reviewers comments we also did further testing to see if with other approaches we could better explain differences between observed and simulated values. We tested a mixed model to account for the randomness induce due to site (plot) differences. For example a glme model with the recruitment values as the response variable and models as the explanatory variable with observed data as the reference (with poisson distribution-count data). This worked well and the patterns were the same as in the t-test (expect for the model 4C). We also considered doing the same approach the shannon index and mortality differences but in this case we think it is more interesting to understand the mean differences across sites and therefore a t-test is a valid approach. We also tested a linear mixed model to assess differences between 7 and 10cm on the Shannon index creating a new parameter where the explanatory variable was a combination of the model and the diameter threshold, then we run a pairwise multiple comparisons to determine which means differ. The patterns here were also similar to the ones observed in the t-test except for the three landscape level models that were significant in this test contrary to the t-test.

We have included here

## Prepare the data for analysis

```{r dataPrep}
source("../code/plottingManuscript.R")

# Recruitment data 
outputsDF <- data.table::fread("../data/dataOutputs.csv")
outputsDF$model[outputsDF$model == "Empirical"] <- "Observed"

# Assign models and dbh to levels

outputsDF$model <- factor(outputsDF$model, levels = modelsOrder)
outputsDF$dbh <- factor(outputsDF$dbh, levels = c("7", "10"))

# Remove the 35 sites from the models data for threshold 7cm 
# as they do not have observations for those sites

sitesWoObs <- outputsDF[is.na(outputsDF$r.trees), ]
outputsDF <- outputsDF[!(outputsDF$dbh == 7 & outputsDF$site %in% unique(sitesWoObs$site)), ]

# Aggegate data based on site, sample, dbh and model/obs
overDTsim <- outputsDF |> 
      dplyr::group_by(site, sample, dbh, model) |>
      dplyr::summarise(r.trees = sum(r.trees),
                       r.ba = sum(r.ba),
                       Totba = unique(Totba)) |>
      dplyr::group_by(site, dbh, model) |>
      dplyr::summarise(r.trees = mean(r.trees),
                       r.ba = mean(r.ba),
                       Totba = mean(Totba))

# Shannon index data 
  ## Recruitment Shannon index  by basal area ----
    ShannonDF <- outputsDF 
    ShannonDF$giDIVG <- ShannonDF$r.ba / ShannonDF$Totr.ba
    ShannonDF$lngiDIVG <- log(ShannonDF$r.ba / ShannonDF$Totr.ba)
    ShannonDF$mult <- round(ShannonDF$giDIVG * ShannonDF$lngiDIVG, 2)

    # Sum across species in each observation
    ShannonIndex <- data.frame(cbind(aggregate(mult ~ site + sample + model + dbh,
                                               sum, data = ShannonDF)))

    ShannonIndex$mult <- (-1) * (ShannonIndex$mult)
    colnames(ShannonIndex) <- c("site", "sample", "model", "dbh",
                                "ShannonIndexRecruit")

    ## Stand Shannon index ----
    ShannonDFAll <- outputsDF 
    ShannonDFAll$giDIVG <- ShannonDFAll$ba / ShannonDFAll$Totba
    ShannonDFAll$lngiDIVG <- log(ShannonDFAll$ba / ShannonDFAll$Totba)
    ShannonDFAll$mult <- round(ShannonDFAll$giDIVG * ShannonDFAll$lngiDIVG, 2)

    # Sum across species per site and sample
    ShannonIndexAll <- data.frame(cbind(aggregate(mult ~ site + sample + model + dbh,
                                                  sum, data = ShannonDFAll)))

    ShannonIndexAll$mult <- (-1) * (ShannonIndexAll$mult)
    colnames(ShannonIndexAll) <- c("site", "sample", "model",
                                   "dbh", "ShannonIndexAllAges")

    ShannonIndex <- merge(ShannonIndex, ShannonIndexAll, by = c("site", "sample",
                                                                "model","dbh"))

    ## Relative Shannon index ----

    ShannonIndex$relShannon <- ShannonIndex$ShannonIndexRecruit / ShannonIndex$ShannonIndexAllAges

     dat710 <- ShannonIndex |>  
      dplyr::group_by(model, site, dbh) |> 
      dplyr::summarise(ShannonIndexRecruit = mean(ShannonIndexRecruit))
      dat710$dbh <- ordered(dat710$dbh, levels = c("7", "10"))

    dat7H <- dat710[dat710$dbh == "7", ]

# data preparation mortality 
    # Data preparation    
    dataMort <- outputsDF[outputsDF$species %in% c(selSpecies, aDVGMSpecies), ]

    treesTot <- dataMort |>
      dplyr::group_by(model, site, sample, dbh) |>
      dplyr::summarise(Tot.rtrees = sum(r.trees),
                       dds = unique(dds),      
                       wb = unique(wb))
    meanTrees <- treesTot |>
      dplyr::group_by(model, site, dbh) |>
      dplyr::summarise(mean.rtrees = mean(Tot.rtrees),
                       dds = unique(dds),  
                       wb = unique(wb))

    meanTrees7 <- meanTrees[meanTrees$dbh == 7,]
    colnames(meanTrees7) <- c("model", "site", "dbh", "nn7", "dds", "wb")
    meanTrees7 <- meanTrees7[, c("model", "site", "nn7")]
    meanTrees10 <- meanTrees[meanTrees$dbh == 10,]
    colnames(meanTrees10) <- c("model", "site", "dbh", "nn10", "dds", "wb")
    meanTrees10 <- meanTrees10[, c("model", "site", "nn10",  "dds", "wb")]

    mortAll <- merge(meanTrees7, meanTrees10, by = c("model","site"))
    mortAll$nn710 <- mortAll$nn7 / mortAll$nn10

    # the plot 
    ratio7_10 <- ggplot2::ggplot(mortAll, ggplot2::aes(y = nn710, x = model, 
                                                     fill = model)) + 
      ggplot2::geom_boxplot() +
      ggplot2::geom_hline(ggplot2::aes(yintercept = 1), colour = "darkblue", 
                          linetype = "dashed") +
      ggplot2::geom_hline(ggplot2::aes(yintercept = 1.77), colour = "darkblue",
                          linetype = "dashed") +
      ggplot2::ylim(c(0,4)) +
      ggplot2::xlab(label = "") +
      ggplot2::ylab(label = "Ratio of recruitment (7 and 10 cm)") +
                    # bquote(bar(R) * "(7cm) / " * bar(R) * "(10cm)")) +
      ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90),
                     strip.text.y = ggplot2::element_text(angle = 0),
                     legend.position = "none",
                     panel.background = ggplot2::element_blank(), 
                     axis.line = ggplot2::element_line(colour = "black"),
                     legend.key =  ggplot2::element_blank(), 
                     legend.title = ggplot2::element_blank(),
                     strip.background =  ggplot2::element_blank()) +
      ggplot2::scale_fill_manual(labels = labels[names(labels) %in%  unique(mortAll$model)],
                                 values =  values_color[names(values_color) %in% unique(mortAll$model)],
                                 guide = "none")

    ggplot2::ggsave("figures/Figure5.jpg",
                    plot = ratio7_10,
                    width = 21, height = 12, scale = 0.9,
                    dpi = 300, units = "cm", device = 'jpg')

    mortAll$nn710 <-  as.numeric(as.character(mortAll$nn710))

```

## Tree recruitment

### T-test

With this tests we wanted to see if the recruitment levels simulated per each model were different from the observations. We have included the bonferroni correction due to multiple testing.

```{r t-testR7, warning=FALSE}
rtreesDiff <- ggpubr::compare_means(r.trees ~ model,  
                                    data = overDTsim[overDTsim$dbh == 7,], 
                                    ref.group = "Observed",
                                    method = "t.test",
                                    p.adjust.method = "bonferroni")

# Clean the table 
rtreesDiff$method <- NULL
rtreesDiff$p.adj <- NULL
rtreesDiff$p <- NULL
```

```{r t-testR7T, warning=FALSE, echo=FALSE}
library(knitr)
kable(rtreesDiff)
```

The conclusion here is, all the models except 4C have significantly different mean values per site of recruitment levels when compared with the observed data.

### GLME

Here we try to also check if the recruitment levels simulated per each model were different from the observations but taking into consideration the random effects caused by the simulations being done at different sites. As here we are dealing with count data we use the poisson distribution.

```{r glmeRtrees7,  warning=FALSE}
  nomeansdata <- outputsDF |> 
     dplyr::group_by(site, sample, dbh, model) |>
     dplyr::summarise(r.trees = sum(r.trees),
                      r.ba = sum(r.ba),
                      Totba = unique(Totba))

   dataGLMrtrees2 <- nomeansdata[nomeansdata$dbh == 7,]
   dataGLMrtrees2$r.trees <- as.integer(dataGLMrtrees2$r.trees) 
   rtreesDiffGLME2 <- lme4::glmer(r.trees ~ model + (1 | site),
                                  family = poisson(link = "log"),
                                  data =  dataGLMrtrees2)

   summary(rtreesDiffGLME2)

```

Contrary to the t-test, here all the models are significantly different to the observed data, so when taking into consideration the variability per site and not only the mean value per site, then also 4C is significantly different.

## Shannons index -Models with feedback and wo feedback and observations

### T-test

With this tests we wanted to see if the mean Shannon index values per site simulated per each model were different from the observations.
```{r t-testH7, warning=FALSE}
HtreesDiff <- ggpubr::compare_means(ShannonIndexRecruit ~ model, 
                                    data = dat7H[!dat7H$model == "aDGVM2", ], 
                                    ref.group = "Observed", method = "t.test",
                                    p.adjust.method = "bonferroni")

# Clean the table 
HtreesDiff$method <- NULL
HtreesDiff$p.adj <- NULL
HtreesDiff$p <- NULL
```

```{r t-testH7T, warning=FALSE, echo=FALSE}
library(knitr)
kable(HtreesDiff)
```

The conclusion here is, that when looking at mean values per site, few models match the observed species diversity in recruitment (not significantly different).

### LMM
Here we try to also check if the Shannon index obtained from the simulations in each model were different from the observations but taking into consideration the random effects caused by the simulations being done at different sites. As here we are dealing with and index that starts from 0, we assumed a Gaussian distribution.

```{r ,  warning=FALSE, message = FALSE}
  ShannonIndex7 <- ShannonIndex[ShannonIndex$dbh == "7", ]
    ShannonIndex7$ShannonIndexRecruit[ShannonIndex7$ShannonIndexRecruit < 0] <- 0
    sigH7_SimObsGLME2 <- lme4::lmer(ShannonIndexRecruit ~ model + (1 | site),
                                    data = ShannonIndex7[!ShannonIndex7$model == "aDGVM2", ])
    summary(sigH7_SimObsGLME2)
```

No significant differences of the models with the observed data, because here we are taking into consideration the variability at the different sites which is large. This answer might be also interesting but not the intended analysis I think. The conclusion here would be that none of the models matched species diversity in recruitment when accounting for site variability.

## Mortality

### 2023-10-24 code/figures.R
additions analysis further tests models GLME and LMM

@@ -216,16 +216,82 @@
                     dpi = 300, units = "cm", device = 'jpg')
     
     
-    # Significant difference per model between 7 and 10 diameter threshold
+    # Significant difference per model between 7 and 10 diameter threshold-----
     
     sigH7_10 <- ggpubr::compare_means(ShannonIndexRecruit ~ dbh, 
                                       data = dat710[!dat710$model == "aDGVM2", ],
-                                      group.by = "model", method = "t.test")
+                                      group.by = "model", method = "t.test",
+                                      p.adjust.method = "bonferroni")
     
-    write.csv(sigH7_10,"figures/sigH7_10.csv", row.names = FALSE)
+    # Clean table
+    sigH7_10$p.adf <- NULL
+    sigH7_10$p.format <- NULL
+    sigH7_10$method <- NULL
     
+     write.csv(sigH7_10,"figures/sigH7_10.csv", row.names = FALSE)
     
-    # GLME 
+     #####  GLME approach -----
+     # briefly thought about how to analyse your data (Shannon index and mortality),
+     #  when you want to focus on the differences between the 7 and 10 cm thresholds 
+     #  within the same model. We actually use a function in our applied stats 
+     #  course that we apply to anova, but I didn’t know whether this will 
+     #  work with random effects as well.
+     # 
+     # You can try to use the function "glht" in the R library "multcomp",
+     #  which allows to define specific contrasts that you want to test. 
+     #  The function can also be applied to mixed effects models. 
+     #  You can try whether the following works: Calculate a repeated-measures ANOVA, 
+     #  i.e. a linear mixed-effects model, e.g. lmer(y ~ model.thr + (1 | sample)), 
+     #  where y is the response variable and model.thr a factor, which combines 
+     #  the model (e.g. LandClim) and the threshold (e.g. 7 cm), e.g. LandClim.7, 
+     #  LandClim.10. For the Shannon index, you can probably assume a linear
+     #   mixed-effects model with log(y) as response. For the mortality data, 
+     #   the best would be to use a binomial distribution, because you know the 
[…]

### 2023-10-24 manuscript/manuscriptRW.Rmd
last changes before sent for internal round
add revisions from Harald

@@
 - Heike	Lischke$^7$,
-- Mats Mahnken$^8$,
-- Paola	Mairota$^9$,
-- Ján Merganič$^{10}$,
-- Katarina Merganičová$^{11,12}$,
-- Tobias Mette$^{13}$,
-- Marco Mina$^{14}$,
-- Xavier Morin$^{15}$
+- Paola	Mairota$^8$,
+- Ján Merganič$^{9}$,
+- Katarina Merganičová$^{10,11}$,
+- Tobias Mette$^{12}$,
+- Marco Mina$^{13}$,
+- Xavier Morin$^{14}$
+- Mats Nieberg,$^{15}$,
 - Werner Rammer$^4$,
@@
 
-Most of the models overestimate recruitment levels, which is only
-compensated in some models by high simulated mortality rates in the
-early stages of individual trees dynamics. Simulated species diversity
-of recruitment matches the observed ranges. Models simulating higher
-species diversity at the stand level do not present higher species
-diversity in the recruitment layer. The effect of light availability on
-recruitment levels is captured better than the effect of temperature and
-soil moisture, but patterns are not consistent across models. Increasing
-complexity in the tree regeneration modules of the models is not related
-to higher accuracy of simulated tree recruitment. Furthermore,
-individual model design is more important than scale (stand, landscape,
-global) and approach (empirical, process-based) for accurately capturing
-tree regeneration.
+Most of the models overestimate recruitment levels, which is compensated
+only in some models by high simulated mortality rates in the early
+stages of individual tree dynamics. Simulated species diversity of
+recruitment generally matches observed ranges. Models simulating higher
+stand-level species diversity do not feature higher species diversity in
+the recruitment layer. The effect of light availability on recruitment
+levels is captured better than the effects of temperature and soil
+moisture, but patterns are not consistent across models. Increasing
+complexity in the tree regeneration modules is not related to higher
+accuracy of simulated tree recruitment. Furthermore, individual model
+design is more important than scale (stand, landscape, global) and
+approach (empirical, process-based) for accurately capturing tree
+regeneration.
 
@@
 parameterized with such data. We conclude that much can be gained by
-evaluating and refining the modeling of regeneration processes. This has
-the potential to render long-term projections of forest dynamics under
-changing environmental conditions that are much more robust.
+evaluating and refining the modeling of tree regeneration processes.
+This has the potential to render long-term projections of forest
+dynamics under changing environmental conditions much more robust.
 
@@
 dedicated to improving the representation of tree mortality
-[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is a much
-less studied process that is often represented rather coarsely in models
+[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is much
+less studied and is often represented rather coarsely in models
 [@walck2011; @leishman1992; @price2001]. This presents an important
@@
 the context of global change scenarios, and (3) identify the most
-important processes that are shaping ecological patterns. Given the
-current strategies that are used in models of forest dynamics to
-represent tree regeneration, model behavior often is prone to problems,
-such as very high levels of tree regeneration that necessitate excess
-mortality at early stages of tree life to simulate correct stand
-structure and composition [@kroiss2015]. Also, inadequately high species
-diversity in tree regeneration may be simulated, which is characteristic
-of "classical" forest gap models [@gutiérrez2016], at least as long as
-the simulation set-up comprises a multitude of species. Some models use
-calibration against local canopy-level data to constrain simulated
-regeneration levels, which is likely to hamper the general applicability
-of these models e.g. under scenarios of climate change. Furthermore,
-correctly capturing the species composition in tree regeneration as a
+important processes that are shaping vegetation patterns. Given the
+strategies that are used in models of forest dynamics to represent tree
+regeneration, their behavior often is prone to problems, such as very
+high levels of tree regeneration that necessitate excess mortality at
+early stages of tree life to simulate correct stand structure and
+composition [@kroiss2015]. Also, inadequately high species diversity in
+tree regeneration may be simulated, which is characteristic of
+"classical" forest gap models [@gutiérrez2016], at least as long as the
+simulation set-up comprises a multitude of species. Some models use
+calibration against local canopy-level data to constrain simulated tree
+regeneration, which is likely to hamper the general applicability of
+these models e.g. under scenarios of climate change. Furthermore,
+correctly capturing the species composition of tree regeneration as a
 function of the presence of seed trees in the canopy is often a
@@
 problem with insufficient observational constraints on parameter values
-for models that start from very small tree sizes (e.g., 10 cm in height
-or even seeds) and track tree development in a process-oriented manner
-by considering a multitude of ecological influences, rather than
-emphasizing tree regeneration into a larger size class [e.g.,
+for models that start from very small tree sizes (e.g., 10 cm in height,
+or even from seed) and track tree development in a process-oriented
+manner by considering a multitude of ecological influences, rather than
+emphasizing tree recruitment into a larger size class [e.g.,
 @wehrli2006].
@@
 [@clark1999a]. For example, monitoring on permanent plots (such as
-National Forest Inventories) often measures tree regeneration
-[@zell2019], however with a design that captures the rate after the
-stand initiation phase [@hallsby2015]. Data from permanent plots,
-although available across regions, are also highly heterogeneous,
-rendering their use difficult in a modeling context [@käber2021], and
-targeted studies to measure tree regeneration on experimental sites are
-often limited in spatial extent, species studied, or the sub-set of
-processes that are targeted [@berdanier2016; @collet2006].
+National Forest Inventories) often measures the ingrowth of new trees
+into a specific size class [@zell2019], however with a design that
+captures the rate after the stand initiation phase [@hallsby2015]. Data
+from permanent plots, although available across regions, are also highly
+heterogeneous, rendering their use difficult in a modeling context
+[@käber2021]. Lastly, targeted studies to measure tree regeneration on
+experimental sites are often limited in spatial extent, species studied,
+or the sub-set of processes that are investigated [@berdanier2016;
+@collet2006].
 
@@
 not aim for a detailed statistical evaluation of each model. Instead, we
-aim to evaluate the general regeneration patterns and magnitudes
+aim to evaluate the general recruitment patterns and magnitudes
 simulated by the models, and benchmark the simulated regeneration niche
@@
 model traits explain differences in model performance? (3) How well do
-the models capture total recruitment and individual species regeneration
-niches across environmental gradients of light availability,
-temperature, and soil moisture? We evaluate the performance of
-regeneration models by analyzing simulated data alongside observed data.
-Specifically, we examine tree species diversity, mortality rates in tree
-recruitment, and recruitment levels. Then, we link model traits (e.g.,
-complexity, scale) to performance, and we assess whether modeled total
-recruitment or individual species regeneration niches (i.e.,
-proportional recruitment along an environmental gradient) align with
-observational data.
+the models capture total recruitment and the regeneration niches of
+individual species across environmental gradients of light availability,
+temperature, and soil moisture?
+
+We evaluate the performance of the models by analyzing simulated data
+alongside observed data. Specifically, we examine tree species
+diversity, mortality rates in tree recruitment, and recruitment levels.
+Then, we link model traits (e.g., complexity, scale) to performance, and
+we assess whether modeled total recruitment or individual species
+regeneration niches (i.e., proportional recruitment along an
+environmental gradient) align with observational data.
 
@@
 Fifteen models of forest dynamics (including two models featuring model
-variants) were used to simulate forest dynamics at the stand, landscape
-or global scale (Table \@ref(tab:tableModels)). The approaches used for
-model construction and their origin differ strongly, with most of the
-models featuring a largely "process-based" approach, whereas two models
-are based on formulations derived from the statistical analysis of
-inventory data (SIBYLA and xComp). Some of the models largely rely on
-the original approach underlying forest gap models (e.g., ForClim 1),
-whereas some are based on plant ecophysiological processes (e.g.,
-FORMIND and iLand). The different approaches underlying the models have
-strongly influenced the formulation of tree regeneration processes.
-
-The regeneration models can be differentiated into 'regeneration' and
-'recruitment' approaches [@vanclay1994; @konig_tree_2022]. *Regeneration
-models* include processes such as flowering and pollination, seed
-production, seed dispersal, germination, and seedling growth, which
-ultimately lead to the simulated number of established trees.
-*Recruitment models*, in contrast, introduce a number of new trees with
-certain characteristics such as biomass or diameter, without explicitly
-considering earlier development processes. We can further distinguish
-models that feature a feedback in the regeneration, i.e., where the
-level and species composition of regeneration is influenced by the
-existing adult tree community via the production of seeds, seedlings or
-saplings, from models that do not contain such a feedback.
+variants) developed for the stand, landscape or global scale were used
+to simulate forest dynamics (Table \@ref(tab:tableModels)). The
+approaches used for model construction and their origin differ strongly,
+with most of the models featuring a largely "process-based" approach,
+whereas two models are based on formulations derived from the
+statistical analysis of inventory data (SIBYLA and xComp). Some of the
+models largely rely on the original approach underlying forest gap
+models (e.g., ForClim 1), whereas some are based on plant
+ecophysiological processes (e.g., FORMIND and iLand). The different
+approaches underlying the models have strongly influenced the
+formulation of tree regeneration processes.
+
+The models can be differentiated into 'regeneration' and 'recruitment'
+approaches [@vanclay1994; @konig_tree_2022]. *Regeneration models*
+include processes such as flowering and pollination, seed production,
+seed dispersal, germination, and seedling growth, which ultimately lead
+to the simulated number of established trees. *Recruitment models*, in
+contrast, introduce a number of new trees with certain characteristics
+such as biomass or diameter, without explicitly considering earlier
+development processes. We can further distinguish models that feature a
+feedback from stand properties to simulated recruitment, i.e., where the
+level and species composition of recruitment is influenced by the adult
+tree community via the production of seeds (or seedlings/saplings), from
+models that do not contain such feedback.
 
@@
 processes that are not treated explicitly. This latter approach reduces
-model complexity but comes at the cost of blurring process
+model complexity but comes at the cost of blurred process
 representation.
@@
 (mean regeneration formulation complexity across all processes \>0,
-Table \@ref(tab:tableModels)). Two models, ForClim and ForCEEPS, were
-used here with two alternative variants of regeneration. ForClim variant
-1 [@bugmann1996] is based on a regeneration module that adheres closely
-to the concept introduced by @botkin1972 which is an environmental
-filter considering species specific thresholds of light availability and
-climatic variables. Whereas ForClim variant 11 is adopting a slightly
-more complex approach where individual species properties and their
-relation to the environment are represented gradually along with the
-relative suitability of different species (i.e., competition among tree
-species) [@huber2020]. These two model variants allow us to evaluate a
-more process-based and complex module (variant 11) against a simple
-module (variant 1), while the rest of the model structure is identical.
-Similarly, the two ForCEEPS variants allow us to isolate the importance
-of the canopy feedback (i.e., simulated actual composition and relative
-abundance of species in the plot) via seed trees for the quantity and
-quality (e.g., diversity and composition) of simulated regeneration, as
-one variant includes this feedback whereas the other does not.
+Table \@ref(tab:tableModels)), whereas the other eight models feture
+relatively low complexity. Two models, ForClim and ForCEEPS, were used
+here with two alternative variants of regeneration. ForClim variant 1
+[@bugmann1996] is based on a regeneration module that adheres closely to
+the concept introduced by @botkin1972 which is based on environmental
+filters considering species-specific thresholds of light availability
+and climatic variables. In contrast, ForClim variant 11 is adopting a
+slightly more complex approach where individual species properties and
+their relation to the environment are represented gradually along with
+the relative suitability of different species (i.e., competition among
+tree species) [@huber2020]. These two model variants allow us to
+evaluate a more process-based and complex module (variant 11) against a
+simple module (variant 1), while the rest of the model structure is
+identical. Similarly, the two ForCEEPS variants allow us to isolate the
+importance of the canopy feedback (i.e., simulated actual composition
+and relative abundance of species in the plot) via seed trees for the
+quantity and quality (e.g., diversity and composition) of simulated
+regeneration, as one variant includes this feedback whereas the other
+does not.
 
@@
 environmental gradients in temperature and precipitation in Central
-Europe as compiled in the framework of the EuFoRIa network
-[@euforia2019; @käber2023] (Figure 1). The environmental coverage and
-the variety of forests represented by this data is unprecedented within
-Europe. These forests have been unmanaged for at least 10 years prior to
-their designation as reserves, and most of them provide time series of
-natural forest dynamics over multiple decades and up to ca. 80 years.
-The census periods range from 3 to 37 years, with an average of 14
-years. The data provide information at the tree level, thus allowing for
-the sequential comparison of processes such as individual-tree
-regeneration and death between the measurements.
+Europe as compiled in the EuFoRIa network [@euforia2019; @käber2023]
+(Figure 1). The environmental coverage and the variety of forests
+represented by this data is unprecedented within Europe. These forests
+have been unmanaged for at least 10 years prior to their designation as
+reserves, and most of them provide time series of natural forest
+dynamics over multiple decades and up to ca. 85 years. The census
+periods range from 3 to 37 years, with an average of 14 years. The data
+provide information at the tree level, thus allowing for the sequential
+comparison of processes such as individual-tree recruitment and death
+between the measurements.
 
 We selected 200 sites from this network as the benchmarking dataset for
-the simulation to be representative of the environmental variation
-contained in the data. This was achieved by applying k-means clustering
-to define 200 clusters of plots from the original set of 869 plots along
-the environmental dimensions of temperature, climatic water balance,
-soil quality, slope and aspect. Each of these sites featured at least
-two consecutive measurements. Recruitment minimum size thresholds for
-these sites differed between stem diameters of 0 and 10 cm. For our
-study, we defined two datasets where one included 165 sites with a
+the simulation, so as to be representative of the environmental
+variation contained in the data. This was achieved by applying k-means
+clustering to define 200 clusters of plots from the original set of 869
+plots along the environmental dimensions of temperature, climatic water
+balance, soil quality, slope and aspect. Each of these sites featured at
+least two consecutive measurements. Recruitment size thresholds ranged
+from stem diameters at breast height (dbh) from 0 to 10 cm. For our
+study, we defined two datasets, where one included 165 sites with a
 diameter threshold of 7 cm or lower, and the other included another 35
-sites with diameter thresholds from 10 cm. Therefore we had in total 200
-sites with observations of newly established trees above 10 cm and 165
-sites above 7cm. In the results section, any result from observed data
-at 7 cm threshold originated from the 165 sites.
+sites with diameter thresholds between 7 and 10 cm. Therefore we had in
+total 200 sites with observations of newly recruited trees above 10 cm
+and a subset of 165 sites with recruitment data above 7 cm. In the
+results section, any observations from the data at the 7 cm threshold
+originated from 165 sites.
 
@@
 featured 30,900 newly established trees. Recruitment rates per site,
-sample and decade ranged from zero to 1246 trees, with a mean of 56
+sample and decade ranged from zero to 1,246 trees, with a mean of 56
 trees. Adult species composition was also available for each plot. For
-more details on this dataset and the detailed regeneration information,
+more details on this dataset and the detailed recruitment information,
 cf. @käber2023.
@@
 The input variables were collected from different data sources and
-aggregated to be adapted to the needs of the different models. Time
+aggregated to be suitable for the needs of the different models. Time
 series of climatic variables from 1981 to 2018 were provided in hourly
@@
 The simulations were run in the absence of natural disturbances. Only
-the model LPJ-GUESS had to include a background disturbance to increase
-chances of shade-intolerant species to establish. The simulations were
-set up to sample species-specific recruitment rates per decade and per
-ha in the equilibrium state of the model, typically entailing a
-"spin-up" run (as we did not provide any forest data). The modeling
-teams decided on the simulated area and how they derived these samples.
-The exact length of the simulation was also decided by the modeling
-teams (Table \@ref(tab:tableModels)). Further details on how each
-modeling team prepared the simulations and the outputs are available in
-Supplementary Material 1.
+the model LPJ-GUESS had to include a "background" disturbance
+probability to increase chances of shade-intolerant species to
+establish. The simulations were set up to sample species-specific
+recruitment rates per decade and per ha in the equilibrium state of the
+model, typically entailing a "spin-up" run (as we did not provide any
+forest data). The modeling teams decided on the simulated area and how
+they derived these samples. The exact length of the simulation was also
+decided by the modeling teams (Table \@ref(tab:tableModels)). Further
+details on how each modeling team prepared the simulations and the
+outputs are available in Supplementary Material 1.
 
@@
 regeneration niche (i.e., in environmental space) as well as the
-intensity of the regeneration process (i.e., the number of recruited
-trees per area and per unit of time) were of interest.
+intensity of recruitment (i.e., the number of recruited trees per area
+and per unit of time) were of interest.
 
@@
 betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
-Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus
+Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L.,* and *Pinus
 sylvestris L.* The same set of species was used at all 200 sites. Two
 models included their standard set of species for the simulations, which
-is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species
+is much larger (i.e., ForClim 1, ForClim 11, TreeMig), and the species
 simulated beyond the eleven species were lumped in a category "others".
@@
 species. Rather, community assembly processes and trait filtering
-generate plant communities that are adjusted to the biotic and abiotic
-conditions, and the simulated plants can be classified into ecological
-strategies based on their trait values in a post-processing step.
+generate plant communities that are adjusted to the local biotic and
+abiotic conditions, and the simulated plants can be classified into
+ecological strategies based on their trait values in a post-processing
+step.
 
@@
 Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference
-to estimate whether the ratio of recruitment between the 7 and 10 cm
-diameter thresholds were above or below the expected theoretical rate.
-The Reineke self-thinning rule is usually calculated for even-aged,
-single species stands and assumes a fixed relationship between the
-number of stems and the quadratic mean diameter in fully stocked pure
-stands. The value used in our comparisons was 1.77 (i.e., we expect stem
-numbers at 7 cm to be 77% higher than at 10 cm), calculated using
-Equation \ref{eqReineke}.
+to estimate whether the ratio of recruitment between these two
+thresholds was above or below the expected theoretical rate. The Reineke
+self-thinning rule is usually calculated for even-aged, single species
+stands and is based on a fixed relationship between the number of stems
+and the quadratic mean diameter in fully stocked pure stands. The value
+used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm
+to be 77% higher than at 10 cm), calculated using Equation
+\ref{eqReineke}.
 
@@
 (1) model complexity as defined by @bugmann2022, (2) model type
-(empirical or process based), (3) the presence or absence of a canopy
+(empirical or process-based), (3) the presence or absence of a canopy
 feedback for regeneration, and (4) the scale of application of the model
@@
 Total recruitment and the regeneration niches of the individual species
-were evaluated across the environmental gradients of light availability,
-temperature, and soil moisture. Stand basal area (the basal area of all
-the trees in each sample including tree regeneration) was used as a
-proxy for light availability at the forest floor, the annual degree-day
-sum [@allen1976; @fischlin1995] as a proxy for growing season warmth,
-and the climatic water balance as a proxy for soil moisture
-[@speich2019]. Recruitment values of each species were calculated as the
-mean across the 200 samples per site. The observed data were modeled
-using a Generalized Additive Model [@wood2011] with a negative binomial
-distribution and restricted maximum likelihood to better understand the
-relationship between the environmental gradients and the levels of
-observed regeneration, relative to the simulation results per model.
+were evaluated across the gradients of light availability, temperature,
+and soil moisture as captured in the data. Stand basal area (the basal
+area of all the trees in each sample above the recruitment threshold)
+was used as a proxy for light availability at the forest floor, the
+annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing
+season warmth, and the climatic water balance as a proxy for soil
+moisture [@speich2019]. Recruitment values of each species were
+calculated as the mean across the 200 samples per site. The observed
+data were modeled using a Generalized Additive Model [@wood2011] with a
+negative binomial distribution and restricted maximum likelihood to
+better understand the relationship between the environmental gradients
+and the levels of observed regeneration, relative to the simulation
+results per model.
 
@@
 
-## Regeneration levels
+## Recruitment levels
 
@@
 recruitment (0.01% - 2.39%), i.e., they had distinctly fewer occurrences
-of zeros compared to the observations.
-
-## Tree species diversity of regeneration
+of zeros than the observations.
+
+## Tree species diversity of recruitment
 
@@
 lower variability of diversity compared to observations. Models
-overpredicting species diversity included ForClim1, ForClim 11, FORMIND
-and TreeMig (Figure 4A, left). Several models underpredicted species
-diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure 4A, right).
+overpredicting stand-level species diversity included ForClim1, ForClim
+11, FORMIND and TreeMig (Figure 4A, left). Several models underpredicted
+species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure 4A,
+right).
 
@@
 already observed in Figure 3, only two of them underestimated it (4C and
-Landis-II) (Figure 4B and Figure S4). Both the models with feedback and
-those without feedback from the adult trees to regeneration (via seed
-production) overestimated, underestimated or captured reasonably well
-recruitment species diversity compared to the observations. Again,
-simulated diversity had considerably lower variability than observed
-diversity in the recruitment.
-
-## Mortality in tree regeneration
-
-There were strong differences among the models regarding the mortality
-rate between 7 and 10 cm (Figure 5). Most of the models featured a
-mortality rate significantly larger than the observed data, thus at
-least partially compensating for the general overestimation of
-recruitment levels (Figure 2 and Table S3); six models did not feature a
+Landis-II) (Figure 4B and Figure S4). Both the models with and without
+feedback from the adult trees to regeneration (via seed production)
+overestimated, underestimated or captured reasonably well recruitment
+species diversity compared to the observations. Again, simulated
+diversity had considerably lower variability than observed diversity in
+the recruitment.
+
+## Recruitment mortality
+
+There were strong differences among models regarding the mortality rate
+between 7 and 10 cm (Figure 5). Most of the models featured a mortality
+rate significantly larger than the observed data, thus at least
+partially compensating for the general overestimation of recruitment
+levels at 7 cm (Figure 2 and Table S3); six models did not feature a
 significant difference compared to the observed data (ForCEEPS, FORMIND,
@@
 noteworthy that most of the models that featured a low ratio of
-regeneration between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11
+recruitment between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11
 and iLand) had only a small overestimation of recruitment at the 7 cm
@@
 
-Lastly, some models that had a low number of individuals at the 7 cm
-threshold also had a smaller decrease of recruitment between 7 and 10 cm
+Lastly, some models that had a low number of recruits at the 7 cm
+threshold also had a small decrease of recruitment between 7 and 10 cm
 (ratio 7-10 cm) (Figure S5). In these models, competition and
@@
 While some models showed a negative effect, the linear trend was not
-pronounced and not always significant regarding the slope.
+pronounced and not always significant regarding its slope.
 
@@
 overestimation proportion at a diameter threshold of 7 cm (Figure S6).
-All the models except 4C had significant differences between the
-observed and simulated mean recruitment values. Therefore it was not
-possible to differentiate in terms of performance between models type
-(empirical or process based), or their scales (stand, landscape or
-global) (Table S5). The values of recruited species diversity simulated
-by each model were also assessed against the observed data to see if
-there was a difference between models that include a feedback compared
-to those that did not; all the models except ForCEEPs(f) and FORMIND had
-significant differences (Table S6). It is noteworthy that for the model
-ForCEEPS the pattern changed from the overprediction of species
-diversity in the regeneration to a diversity level that is closer to
-observations when the recruitment module included feedback
-(ForCEEPs(f)).
-
-## Regeneration gradients on regeneration niches
-
-When evaluating total recruited levels along key gradients of light
-availability (basal area), temperature (degree-days) and soil moisture
-(climatic water balance), distinct features emerged: the models
+All models except 4C had significant differences between the observed
+and simulated mean recruitment values. Therefore it was not possible to
+differentiate in terms of performance between models type (empirical or
+process based), or their scales (stand, landscape or global) (Table S5).
+The values of species diversity in the recruitment simulated by each
+model were also assessed against the observed data to see if there was a
+difference between models that include a feedback compared to those that
+did not; all the models except ForCEEPs(f) and FORMIND had significant
+differences (Table S6). It is noteworthy that for the model ForCEEPS the
+pattern changed from the overprediction of species diversity in the
+regeneration to a diversity level that is closer to observations when
+the recruitment module included feedback (ForCEEPs(f)).
+
+## Regeneration gradients and regeneration niches
+
+When evaluating total tree recruitment levels along key gradients of
+light availability (basal area), temperature (degree-days) and soil
+moisture (climatic water balance), distinct features emerged: the models
 reproduced the effect of basal area in both its magnitude and patterns
@@
 
-In the observed data, recruited levels decreased clearly with increasing
-stand basal area (Figure 7a). This trend was captured by the Landis-II
-model only, albeit at much lower values of basal area than in the
-observations. The other models featured distinctly different trends,
-such as (1) an increase in recruited levels with increasing total basal
-area followed by the absence of regeneration at high values of basal
-area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost
-constant recruitment levels with basal area (ForClim 11, xComp, LandClim
-and aDGVM2), or (3) an increase of recruitment up to a certain value of
-stand basal area followed by a decrease at even higher values of stand
-basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA
-and iLand).
+In the observed data, recruitment levels decreased clearly with
+increasing stand basal area (Figure 7a). This trend was captured by the
+Landis-II model only, albeit at much lower values of basal area than in
+the observations. The other models featured distinctly different trends,
+such as (1) an increase in recruitment levels with increasing total
+basal area followed by the absence of recruitment at high values of
+basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2)
+almost constant recruitment levels with basal area (ForClim 11, xComp,
+LandClim and aDGVM2), or (3) an increase of recruitment up to a certain
+value of stand basal area followed by a decrease at even higher values
+of stand basal area, with model-specific thresholds (4C, ForClim 1,
+PICUS, SIBYLA and iLand).
 
 The observed recruitment did not change much across the climatic water
-balance gradient, and showed a slight increase with temperature. The
-models did not match the observed recruitment patterns across the
-gradients of soil moisture and temperature (Figure 7b,c), and neither
-was there a clear pattern across models. Regarding the soil moisture
-gradient, a group of models featured decreasing recruitment with
-increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp,
-PICUS, LandClim, TreeMig and iLand), while a few models showed the
-opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS). Across the
-temperature gradient, all the models that featured a decrease in
-recruitment levels with increasing soil moisture showed the opposite
-trend (i.e., more recruitment with increasing temperature), except for
-4C. The other models did not feature a clear pattern.
+balance gradient, and showed a slight increase with temperature. Neither
+did the models generally match the observed recruitment patterns across
+these gradients (Figure 7b,c), nor was there a clear pattern across
+models. Regarding the soil moisture gradient, a group of models featured
+decreasing recruitment with increasing soil moisture (4C, ForClim 1,
+ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a
+few models showed the opposite trend (ForCEEPS(f), FORMIND and
+LPJ-GUESS). Across the temperature gradient, all the models that
+featured a decrease in recruitment levels with increasing soil moisture
+showed the opposite trend (i.e., more recruitment with increasing
+temperature), except for 4C. The other models did not feature a clear
+pattern.
 
@@
 although with a low share of basal area. Five models (4C, FORMIND,
-SIBYLA, xComp, and LandClim) had almost no recruitment of *Pinus
-sylvestris* and *Quercus* spp., or did not feature any recruitment of
-these species at all.
+SIBYLA, xComp, and LandClim) had very small amounts of recruitment of
+*Pinus sylvestris* and *Quercus* spp., or did not feature any
+recruitment of these species at all.
 
@@
 Tree regeneration is a fundamental process in forest dynamics. Correctly
-capturing tree regeneration in dynamic models is fundamental to, for
-example, evaluate post-disturbance dynamics and potential long-term
-recovery trajectories as it will define the forest state in the first
-decades [@seidl2022]. If this initial phase is not captured well, we can
-not properly assess aspects such as forest resilience or the timing,
-magnitude and progression of carbon sequestration. In this study, for
-the first time, projections of tree regeneration from multiple models of
-forest dynamics were confronted with a unique dataset from unmanaged
-forest reserves across a large environmental gradient in Europe.
+capturing it in dynamic models is fundamental to, for example, evaluate
+post-disturbance dynamics and potential long-term recovery trajectories
+as it will define the forest state in the first decades [@seidl2022]. If
+this initial phase is not captured well, we can not properly assess
+aspects such as forest resilience or the timing, magnitude and
+progression of carbon sequestration. In this study, for the first time,
+projections of tree recruitment from multiple models of forest dynamics
+were confronted with a unique dataset from unmanaged forest reserves
+across a large environmental gradient in Europe.
 
@@
 benchmarking provided novel insights on the ability of state-of-the-art
-models to accurately simulate regeneration levels, species composition,
-and mortality in an early stage of tree life, i.e., between tree
-diameters of 7 and 10 cm. Overall, by adopting this approach a much
+models to accurately simulate recruitment levels, its species
+composition and mortality in an early stage of tree life, i.e., between
+tree diameters of 7 and 10 cm. Overall, by adopting this approach a much
 broader understanding resulted than if we had used these data for model
@@
 and if the excess regeneration at this stage is compensated -- in the
-models -- soon thereafter by higher mortality, overall simulated
-regeneration may still yield trustworthy results at the level of stand
-structure and related forest attributes such as basal area or biomass.
-
-From a modeling point of view, excess regeneration inevitably requires
-excess mortality rates in a later stage, either -- as observed for some
-models -- between the 7 and 10 cm diameter thresholds, or soon after the
-10 cm threshold has been crossed. In any case, correcting at early
-stages for the expected forest densities at later stages is equivalent
-to compensating for a first error (excess regeneration) by a second
-error (excess mortality). It is highly likely that biased projections
-will result, because the two errors are unlikely to be perfectly linked
-and thus will not always compensate each other. Hence, this structural
-problem of most models investigated here is problematic particularly if
-the models are to be used under novel conditions such as under climate
-change [e.g., @huber2021] or in a decision support context [e.g.,
-@thrippleton2021].
-
-Regeneration levels define the structure and composition of future
+models -- soon thereafter by higher mortality, overall simulated tree
+population dynamics may still be trustworthy at the level of stand
+structure and attributes such as basal area or biomass.
+
+From a modeling point of view, excess tree recruitment inevitably
+requires excess mortality rates in a later stage, either -- as observed
+for some models -- between the 7 and 10 cm diameter thresholds, or soon
+after the 10 cm threshold has been crossed. In any case, correcting at
+early stages for the expected forest densities at later stages is
+equivalent to compensating for a first error (excess recruitment) by a
+second error (excess mortality). It is highly likely that biased
+projections will result, because the two errors are unlikely to be
+perfectly linked and thus will not always compensate each other. Hence,
+this structural problem of most models investigated here is problematic
+particularly if the models are to be used under novel conditions such as
+under climate change [e.g., @huber2021] or in a decision support context
+[e.g., @thrippleton2021].
+
+Recruitment levels define the structure and composition of future
 forests, but it is equally important to correctly identify areas where
@@
 to accurately measure establishment filters, such as competition with
-herb layers, site or environmental limitants.
-
-## Species composition of regeneration
-
-Correctly capturing species composition in tree regeneration is
-important to assess the future functional diversity of a forest, e.g.,
-its sensitivity to drought or resilience to disturbances [@redmond2015;
+the herb layer, site-specific or environmental limitations.
+
+## Species composition of recruitment
+
+Correctly capturing species composition in tree recruitment is important
+to assess the future functional diversity of a forest, e.g., its
+sensitivity to drought or resilience to disturbances [@redmond2015;
 @seidl2022]. In the simulations, overall species diversity levels in the
-regeneration were well within the observed range for half of the models
-(7 out of 14). Thus, while most of the models are facing difficulties to
-quantitatively match regeneration levels (cf. above), their performance
-is better when it comes to matching the composition of regeneration as a
-function of abiotic and biotic conditions. Most models maintained or
-even decreased recruitment diversity between the 7 and 10 cm thresholds,
-and the same is visible from the empirical data, but the differences
-were not significant. Diversity variations may be explained by the way
-the individual models consider regeneration processes
-[@konig_tree_2022], e.g., without considering the feedback from adults,
-regeneration can be linked to either seed rain from dispersal or a
-seed/seedling bank; neither process is incorporated across many models
-studied here, however.
+recruitment were well within the observed range for half of the models
+(7 out of 14). Thus, while most models are facing difficulties to
+quantitatively match recruitment levels (cf. above), their performance
+is better regarding the composition of recruitment as a function of
+abiotic and biotic conditions. Most models maintained or even decreased
+recruitment diversity between the 7 and 10 cm thresholds, and the same
+is visible from the empirical data, but the differences between the 7
+and 10 cm thresholds were not significant. Diversity variations may be
+explained by the way the individual models consider regeneration
+processes [@konig_tree_2022], e.g., regeneration can be linked to the
+seed/seedling bank, seed rain from dispersal or from a pool of available
+species.
 
@@
 models than the diversity of tree recruitment, and only a few models
-overpredicted stand-level diversity. Defining regeneration composition
-as being proportional to adult abundance, regardless of productivity,
-might be a simple and conservative assumption to safely maintain
-relative species abundances [@hanbury-brown2022], although this may be
+overpredicted stand-level diversity. Defining recruitment composition as
+being proportional to local adult abundance, regardless of productivity,
+might be a simple and conservative assumption to maintain relative
+species abundances [@hanbury-brown2022], although this may be
 simplistic. Furthermore, based on the simulation results, there is no
 evidence that models with feedback from the canopy (in terms of species
-composition of regeneration) better captured the species diversity of
-the recruitment than those without feedback, with the exception of
-ForCEEPS that significantly improved regeneration species diversity with
-the model version that included feedback (ForCEEPS(f)). The similar
-performance of models with and without feedback is likely because the
-models put more weight on the regeneration niche arising from abiotic
-and biotic filters, than from the habitat niche of the adult trees
-[@grubb1977].
-
-## Regeneration mortality
+composition of regeneration) captured the species diversity of the
+recruitment better than those without feedback, with the exception of
+ForCEEPS that featured significantly more accurate recruitment species
+diversity with the model version that included feedback (ForCEEPS(f)).
+The similar performance of models with and without feedback is likely
+because the models put more weight on the regeneration niche arising
+from abiotic and biotic filters, than from the habitat niche of the
+adult trees [@grubb1977].
+
+## Recruitment mortality
 
@@
 [@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
-expect that models that overestimate regeneration may have a
+expect that models that overestimate recruitment at 7 cm may have a
 particularly high mortality between the 7 and 10 cm thresholds. However,
@@
 respect to simulated recruited levels -- after all, few if any of the
-models' regeneration routines are sufficiently constrained by data, with
-the exception of the empirical models xComp and SIBYLA. It is noteworthy
+models' regeneration routines are well constrained by data, with the
+exception of the empirical models xComp and SIBYLA. It is noteworthy
 that in spite of their empirical basis, these two models did not match
-empirical regeneration levels, in a similar magnitude as the other
-models and even in opposite directions (SIBYLA: underestimation; xComp:
+empirical recruitment levels, in a similar magnitude as the other models
+and even in opposite directions (SIBYLA: underestimation; xComp:
 overestimation). It appears that using empirical data of limited
-geographical scope and constrained to managed forests, as done in these
-models, leads to extrapolation problems already under current climatic
-conditions. For the other models, whose regeneration modules are not
-strongly constrained by empirical data, the multiple strategies that are
-available for modeling regeneration [@konig_tree_2022] could have
-implied that model performance would be much worse than what we found.
+geographical scope that are constrained to managed forests, as done in
+these models, leads to extrapolation problems already under current
+climatic conditions. For the other models, whose regeneration modules
+are not strongly constrained by empirical data, the multiple strategies
+that are available for modeling regeneration processes
+[@konig_tree_2022] could have implied that model performance would be
+much worse than what we found.
 
 Our study showed that increasing complexity in the regeneration modules
-is not linked with a higher accuracy of the projections of regeneration
+is not linked to a higher accuracy of the projections of recruitment
 levels, species composition or mortality at early tree stages, as there
@@
 complexity. Increasing complexity in regeneration modules has been
-motivated by better process understanding and enhancing of model
-accuracy [@bugmann2022]. However, more complex models do not necessarily
-lead to better projections and rather to a reduced transparency and
-lower predictive power [@franklin2020]. Thus, the question of what level
-of detail is appropriate and parsimonious for modeling tree regeneration
-processes remains open [@konig_tree_2022; @bugmann2022].
+called for a long time ago (e.g., @price2001) motivated by better
+process understanding and with the goal of enhancing model accuracy
+[@bugmann2022]. However, more complex models do may not necessarily lead
+to better projections but rather to reduced transparency and lower
+predictive power [@franklin2020]. Thus, the question regarding the level
+of detail that is appropriate and parsimonious for modeling tree
+regeneration remains open [@konig_tree_2022; @bugmann2022].
 
@@
 models examined here did not reproduce this expectation (i.e.,
-decreasing regeneration levels with increasing stand basal area).
+decreasing recruitment levels with increasing stand basal area).
 However, it is difficult to measure light availability at large spatial
 and temporal scales. We used total stand basal area as a proxy for light
-availability [cf. @schmid2021]. However, we were unable to consider
-light availability restrictions caused by ground vegetation, which may
-be an important filter for some EuForIa stands [@woltjer2008]. We found
-pronounced differences in the stand basal area ranges simulated by the
-models, but also between models and observations. This made it
-impossible to evaluate the regeneration for the extremes of the stand
-density ranges in some models. For example, regeneration levels at low
-stand densities are relevant to assess how well forests are recovering
-e.g. after gap creation due to disturbance [@seidl2022; @grubb1977]. At
-the level of the simulated one-hectare samples, average basal area is
-typically not very low as long as no larger disturbances occur, which
-was explicitly excluded in our simulation protocol. Towards the other
-end of the spectrum, i.e., with increasing stand basal area, it would be
-reasonable to expect that the recruitment of the different species would
-become sparser and drop out entirely under low-light conditions
-[@klopcic2012; @zell2019]. However, few models showed this trend, thus
-indicating that the relation between regeneration and light availability
-is not yet captured correctly in most models. Yet, several of the models
-that did feature an increase of regeneration with increasing basal area
-include a feedback between seed production of mature trees and
-regeneration. Thus, it seems that higher light competition does not
-sufficiently compensate for increased seed availability with higher
-basal area due to a higher abundance of mature trees.
-
-There were pronounced differences in how the main tree species were
+availability [cf. @schmid2021]. Yet, we were unable to consider light
+availability restrictions caused by ground vegetation, which may be an
+important filter for forest dynamics at least in some EuFoRIa stands
+[@woltjer2008]. We found pronounced differences in the ranges of stand
+basal area simulated by the models, but also between models and
+observations. This made it impossible to evaluate tree recruitment for
+the extremes of the stand density ranges in some models. This is
+unfortunate because recruitment levels at low stand density are relevant
+to assess how well forests are recovering e.g. after gap creation due to
+disturbance [@seidl2022; @grubb1977]. At the level of the simulated
+one-hectare samples, average basal area is typically not very low as
+long as no larger disturbances occur, which was explicitly excluded in
+our simulation protocol. Towards the other end of the spectrum, i.e.,
+with increasing stand basal area, it would be reasonable to expect that
+the recruitment of the different species would become sparser and drop
+out entirely under low-light conditions [@klopcic2012; @zell2019].
+However, few models showed this trend, thus indicating that the relation
+between regeneration and light availability is not yet captured
+correctly in most models. Yet, several of the models that did feature an
+increase of regeneration with increasing basal area include a feedback
+between seed production of mature trees and regeneration. Thus, it seems
+that in these models higher light competition does not sufficiently
+compensate for increased seed availability with higher basal area due to
+a higher abundance of mature trees.
+
+Lastly, there were pronounced differences how the main tree species were
 represented by the models along the environmental gradients, in
 particular the dominant species *Abies alba*, *Fagus sylvatica* and
-*Picea abies*. The regeneration levels were sampled from simulations in
+*Picea abies*. The recruitment levels were sampled from simulations in
 the equilibrium, and in this state it is expected that non-climax
@@
 it might appear at first sight. In an analysis of primeval forests at
-demographic equilibrium, @brzeziecki2021 found higher regeneration rates
+demographic equilibrium, @brzeziecki2021 found higher recruitment rates
 than the ones observed in the EuFoRia data. Thus, the overestimation of
-regeneration rates by the models may not be so problematic. Second, the
-data were collected for regeneration above a 7 cm threshold, thus
-limiting the assessment of tree regeneration to a holistic viewpoint.
-This constituted a hard limit based on which we can understand only some
-aspects of tree regeneration, which in its entirety often comprises a
-rather long period since seed production [@price2001]. In reality, many
-environmental constraints are acting on young trees [@käber2021] that we
-were unable to assess. Yet, regeneration data with lower thresholds are
-simply not available in a harmonized manner across large environmental
-gradients. Third, the empirical data were collected from rather small
-plots, while we sampled simulated regeneration levels from 1 ha areas,
-which may lead to an incorrect representation of space. Even though the
-strategy we adopted is not ideal, it represents a common challenge when
-harmonizing diverse data sources originating from varied sampling
-strategies [@portier2022]. It would have been extremely challenging for
-such a variety of models to follow a protocol where the spatial sampling
-size was different at each of the 200 sites, and it would have
-introduced additional uncertainty in the results.
+recruitment rates by the models may not be so problematic. Second, the
+data were collected for recruitment above a 7 cm threshold, thus
+limiting the assessment of tree regeneration to a specific point of
+stand dynamics. This constituted a hard limit based on which we can
+understand only some aspects of tree regeneration, which in its entirety
+often comprises a rather long period since seed production [@price2001].
+In reality, many environmental constraints are acting on young trees
+[@käber2021] that we were unable to assess. Yet, recruitment data with
+lower thresholds are simply not available in a harmonized manner across
+large environmental gradients. Third, the empirical data were collected
+from rather small plots, while we sampled simulated recruitment levels
+from 1 ha areas, which may lead to an incorrect representation of space.
+Even though the strategy we adopted is not ideal, it represents a common
+challenge when harmonizing diverse data sources originating from varied
+sampling strategies [@portier2022]. It would have been extremely
+challenging for such a variety of models to follow a protocol where the
+spatial sampling size was different at each of the 200 sites, and it
+would have introduced additional uncertainty in the results.
 
@@
 range of models of forest dynamics, from stand to global scales, the
-simulation setup was limited to 1 ha regeneration values. This lack of
-consideration of spatial scale is appropriate for stand-scale models,
-but it potentially puts the landscape models at a disadvantage, as they
-have been built to be accurate at the landscape level. Without spatial
-context, we are limiting regeneration to the seed influx from the stand
-itself, unless the model has a background seed input. Yet, the global
-models should not be at a disadvantage due to the limited spatial
-considerations [but cf. Snell et al. -@snell2018; or Lehsten et al.
--@lehsten2019], as they are usually lacking dispersal between cells and
-are based on a strong abstraction of horizontal space
-[@hanbury-brown2022]. Lastly, detailed data on soil conditions were not
-available from the observed data, and independent, admittedly rather
-coarse data for soil and climatic water balance had to be used instead.
-It is noteworthy that many models represent drought based on detailed
-indicators based e.g. on soil water holding capacity, which had to be
-derived from a rough soil quality measure. This may at least partially
-explain the unsatisfactory performance of many models along the drought
-axis (i.e., climatic water balance).
+simulation setup was limited a spatial scale of 1 ha. This lack of
+consideration of spatial scale relationships is appropriate for
+stand-scale models, but it potentially puts the landscape models at a
+disadvantage, as they have been built to be accurate at the landscape
+level: without any spatial context, we are limiting tree regeneration to
+the seed influx from the stand itself, unless the model has a background
+seed input. Yet, the global models should not be at a disadvantage here
+due to their inherently limited spatial considerations [but cf. Snell et
+al. -@snell2018; or Lehsten et al. -@lehsten2019], as they are usually
+lacking dispersal between cells and are based on a strong abstraction of
+horizontal space [@hanbury-brown2022]. Lastly, detailed data on soil
+conditions were not available from the observed data, and independent,
+admittedly coarse data for soil properties and the climatic water
+balance had to be used instead. It is noteworthy that many models
+represent drought using detailed indicators based e.g. on soil water
+holding capacity, which had to be derived from a rough soil quality
+measure. This may at least partially explain the unsatisfactory
+performance of many models along the drought axis (i.e., climatic water
+balance).
 
@@
 With our study, we have demonstrated that models of forest dynamics need
-a focus on tree regeneration modules to make them more robust. It
-remains uncertain what level of detail is required to model regeneration
-processes, and this must be addressed in future research. We recommend
-that the improvement of the regeneration modules is implemented as
-additional features that can be traced back, as done here for the
-variants of ForClim and ForCEEPS, and that model complexity and
+a focus on their regeneration modules to make them more robust. It
+remains uncertain what level of detail is required to model tree
+regeneration, and this must be addressed in future research. We
+recommend that the improvement of the regeneration modules is
+implemented as additional features that can be traced back, as done here
+for the variants of ForClim and ForCEEPS, and that model complexity and
 structure must always be connected with modeling objectives
@@
 [@lindner2000]. However, this will require an entirely different set of
-observed data, and potentially not all models of forest dynamics would
-be able to assess the relationship of these aspects on tree
-regeneration, e.g. due to the lack of disturbance or appropriate
-management modules.
+observed data, and potentially not all forest models would be able to
+assess the relationship of these aspects on tree regeneration, e.g. due
+to the lack of disturbance or appropriate management modules.
 
@@
 regeneration [@beckage2003]; this, however, is a serious challenge
-regarding appropriate datasets, as inventory data have wide coverage,
-but by definition do not allow for the assessment of spatial interaction
-effects.
+regarding appropriate datasets, as large-scale inventory data have a
+wide coverage, but by definition do not allow for the assessment of
+spatial interactions.
 
@@
 assumptions on which the models are based. Furthermore, the inclusion of
-a wide range of models with different scales, type, population
-structures, tree regeneration modules and complexity will ensure a wider
-benefit to the whole modeling community.
+a wide range of models with different (1) scales, (2) approaches for
+capturing population structure, (3) tree regeneration modules and (4)
+complexity of their formulations will ensure a large benefit to the
+entire modeling community and beyond.
 
@@
 used here are facing similar challenges in their representation of tree
-regeneration: they generally overestimate tree regeneration levels, and
+regeneration: they generally overestimate tree recruitment levels, and
 the simulated regeneration niche is not always captured accurately as a
@@
 Regarding mortality in the early phase of tree life, many models that
-feature a particularly high overestimation of regeneration levels are
+feature a particularly high overestimation of recruitment levels are
 compensating for this by a larger tree mortality. Often, this
-compensation is not sufficient to reduce the high regeneration levels to
+compensation is not sufficient to reduce the high recruitment levels to
 realistic values. Overall, there is no clear mortality pattern across
@@
 Even though the regeneration routines of most of the models investigated
-here have never been sufficiently constrained by robust data, the
-regeneration projections are not overly off. This indicates that a lot
-can be gained by a focus on the modeling of regeneration processes. The
+here have never been constrained well by robust data, their projections
+of recruitment are not overly off. This indicates that a lot can be
+gained by a focus on the modeling of regeneration processes. The
 representation of forest dynamics in these models would become much more
@@
 authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study
-design. All authors contributed to preparing the simulations with their
-respective models. O. Díaz-Yáñez led the data analysis with
-contributions during the workshop from Y.Käber, T. Anders, K. H.
-Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T. Hickler, H. Lischke,
-M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X. Morin, W. Rammer, D.
-Scherrer and H. Bugmann. O. Díaz-Yáñez and H. Bugmann led the writing of
-the manuscript. All authors participated in the revision of the
-manuscript and approved its submission.
+design. Y. Käber and O. Díaz-Yáñez prepared the input data used by all
+modeling teams. All authors contributed by conducting the simulations
+with their respective models. O. Díaz-Yáñez led the analysis of the
+simulation results with contributions during the workshop from Y.Käber,
+T. Anders, K. H. Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T.
+Hickler, H. Lischke, M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X.
+Morin, W. Rammer, and D. Scherrer. O. Díaz-Yáñez and H. Bugmann led the
+writing of the manuscript and provided the first draft. All authors
+participated in the revision of the manuscript and approved its
+submission.
 
@@
 support from Waldfonds WASIM2100 project. We would also like to thank
-all the people that helped to make this study possible with
-contributions at different stages such as Martin Gutsch and Giorgio
-Vacchiano.
+all the people who helped to make this study possible with contributions
+at different stages particularly Martin Gutsch and Giorgio Vacchiano.
 
@@
 \newpage
-
-# Figures
-
-```{r diff, echo=FALSE, fig.cap=".", out.width='100%', out.height= '85%'}
-knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
-```
-
-\newpage

### 2023-10-30 code/figures.R
last figures changes

@@ -176,6 +176,7 @@
         dplyr::summarise(ShannonIndexRecruit = mean(ShannonIndexRecruit))
     dat710$dbh <- ordered(dat710$dbh, levels = c("7", "10"))
     modelsSel <-  modelsOrder[!modelsOrder  == "aDGVM2"]
+    dat7 <- dat710[dat710$dbh == "7",]
     
     #### Plot
     H7_10 <- ggplot2::ggplot(dat710[!dat710$model == "aDGVM2", ],
@@ -928,6 +929,9 @@
                                     data = overDTsim[overDTsim$dbh == 7,], 
                                     ref.group = "Observed", method = "t.test",
                                     p.adjust.method = "bonferroni")
+    pp <- overDTsim[overDTsim$dbh == 7,]
+    pp <-  pp [!pp $model %in% c("4C", "LPJ-GUESS"  ),]
+    
     # Clean the table 
     rtreesDiff$method <- NULL
     rtreesDiff$p.adj <- NULL
@@ -975,7 +979,7 @@
     
     dat7H <- dat710[dat710$dbh == "7", ]
     
-    HtreesDiff <- ggpubr::compare_means(ShannonIndexRecruit ~ model,  
+    HtreesDiff <- ggpubr::compare_means(ShannonIndexRecruit ~ model,
                                         data = dat7H[!dat7H$model == "aDGVM2", ], 
                                         ref.group = "Observed", method = "t.test",
                                         p.adjust.method = "bonferroni")

### 2023-10-30 manuscript/manuscriptRW.Rmd
small correction in the manuscript, last version before diff

@@
 subtitle: "\\fontsize{16pt}{3pt}\\selectfont \\textbf{\\textit{Tree regeneration in models of forest dynamics: a key priority for further research}} "
-title: "\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to Ecosphere. Manuscript type: Article}}"
+title: "\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Journal: Ecosphere. Manuscript ID: ECS23-0342. Manuscript type: Article}}"
 author: 
@@
 - Werner Rammer$^4$,
-- Christopher P.O. Reyer$^8$,
+- Christopher P.O. Reyer$^{15}$,
 - Simon	Scheiter$^2$,
@@
        $^5$Department of Geoecology, Institute of Botany of the Czech Academy of Sciences, Průhonice, Czech Republic. \newline
-       $^6$Institute of Silviculture, Department of Forest and Soil Sciences, University of Natural Resources and Life Sciences, Peter Jordan Strasse 82, A-1190 Vienna, Austria \newline
+       $^6$Institute of Silviculture, Department of Forest and Soil Sciences, University of Natural Resources and Life Sciences, Peter Jordan Strasse 82, A-1190 Vienna, Austria. \newline
        $^7$Dynamic Macroecology, Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland. \newline
@@
        $^9$Department of Soil Sciences, of Plants and Food, University of Bari Aldo Moro, Italy. \newline
-       $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen, Slovakia \newline
+       $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen, Slovakia. \newline
        $^{11}$Czech University of Life Sciences Prague, Faculty of Forestry and Wood Sciences, Kamýcká 129, 16500 Praha 6 – Suchdol, Czech Republic. \newline
@@
 
-Each of the models reported the number of trees crossing the two
+Each of the models reported the number of new trees crossing the two
 diameter thresholds by sampling 200 times in a 10-year interval for each
@@
 between simulated and observed values of the recruitment levels and
-species diversity by using paired t-tests.
+species diversity by using t-tests with a correction for multiple
+testing. Note that the t-tests does not present conclusive evidence for
+a specific hypotheses, but rather facilitates managing the extensive
+number of comparisons discussed in our study.
 
@@
 
-# Author contributions
-
-O. Díaz-Yáñez and Y. Käber contributed equally and share the first
-authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study
-design. Y. Käber and O. Díaz-Yáñez prepared the input data used by all
-modeling teams. All authors contributed by conducting the simulations
-with their respective models. O. Díaz-Yáñez led the analysis of the
-simulation results with contributions during the workshop from Y.Käber,
-T. Anders, K. H. Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T.
-Hickler, H. Lischke, M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X.
-Morin, W. Rammer, and D. Scherrer. O. Díaz-Yáñez and H. Bugmann led the
-writing of the manuscript and provided the first draft. All authors
-participated in the revision of the manuscript and approved its
-submission.
-
 # Acknowledgements
@@
 
+# Author contributions
+
+O. Díaz-Yáñez and Y. Käber contributed equally and share the first
+authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study
+design. Y. Käber and O. Díaz-Yáñez prepared the input data used by all
+modeling teams. All authors contributed by conducting the simulations
+with their respective models. O. Díaz-Yáñez led the analysis of the
+simulation results with contributions during the workshop from Y. Käber,
+T. Anders, K. H. Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T.
+Hickler, H. Lischke, M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X.
+Morin, W. Rammer, and D. Scherrer. O. Díaz-Yáñez and H. Bugmann led the
+writing of the manuscript and provided the first draft. All authors
+participated in the revision of the manuscript and approved its
+submission.
+
 # Conflict of interest statement
@@
 
+**Table 1.** Models included in this study, their characteristics and
+simulation strategies used. Scale refers to the models scale, Type to
+the models type, Population structure to the model population approach,
+Feedback to the inclusion or not of adult tree feedback to the level and
+species composition of regeneration, Approach to the model regeneration
+module approach, Start from to the stage from which trees are recruited,
+Complexity, to the mean regeneration formulation as estimated in
+(Bugmann & Seidl, 2022), Species to which species from the original
+eleven species were simulated, Runtime (spin up) [years] to the number
+of years used per model in their simulations for the spin up period and
+regeneration sampling; Climate data type (use of years) to the type of
+data used in the simulations and how it was used, and Reference provides
+the link to the publication with more model details. Further information
+can be found on the individual model reports available in the
+supplementary materials.
+
 ```{r tableModels, echo=FALSE,  message=FALSE, warning=TRUE}
@@
 modelDF$Sampling <- NULL
+captionText <- ""
 
@@
       format = "latex", booktabs = TRUE,
-      caption = "") |>
+      caption = captionText) |>
         kableExtra::landscape() |>
@@
 
-# Table and Figure captions
-
-**Table 1.** Models included in this study, their characteristics and
-simulation strategies used. *Scale* refers to the models scale, *Type*
-to the models type, *Population structure* to the model population
-approach, *Feedback* to the inclusion or not of adult tree feedback to
-the level and species composition of regeneration, *Approach* to the
-model regeneration module approach, *Start from* to the stage from which
-trees are recruited, *Complexity*, to the mean regeneration formulation
-as estimated in (Bugmann & Seidl, 2022), *Species* to which species from
-the original eleven species were simulated, *Runtime* (spin up) [years]
-to the number of years used per model in their simulations for the spin
-up period and regeneration sampling; *Climate data type (use of years)*
-to the type of data used in the simulations and how it was used, and
-*Reference* provides the link to the publication with more model
-details. Further information can be found on the individual model
-reports available in the supplementary materials.
+# Figure captions
 

### 2023-10-30 manuscript/supplementaryRW.Rmd
corrections in suplementary doc

@@
 title: "Appendix"
-subtitle: "Tree regeneration in models of forest dynamics: a key priority for further research"
+subtitle: "Tree regeneration in models of forest dynamics: a key priority for further research  ECOSPHERE"
 author: 
@@
 - Heike	Lischke,
-- Mats Mahnken,
 - Paola	Mairota,
@@
 - Xavier Morin,
+- Mats Nieberg,
 - Werner Rammer,
@@
 - \usepackage[left]{lineno}
-- \linenumbers
 - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
@@
 
-# Appendix 1: Model reports
+# Model reports
 
@@
 
-Mahnken M., Gutsch M., Reyer C.P.O., Lasch-Born P.
+Nieberg M., Gutsch M., Reyer C.P.O., Lasch-Born P.
 
@@
 
-All species required were simulated: Fagus sylvatica, Picea abies, Abies
-alba, Carpinus betulus, Tilia cordata, Acer pseudoplatanus, Betula spp.,
-Fraxinus excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris
-
-In addition the following species were simulated: Acer campestre, Acer
+All species required were simulated: *Fagus sylvatica, Picea abies,
+Abies alba, Carpinus betulus, Tilia cordata, Acer pseudoplatanus, Betula
+spp., Fraxinus excelsior, Quercus spp., Alnus glutinosa, Pinus
+sylvestris.*
+
+In addition the following species were simulated: *Acer campestre, Acer
 platanoides, Alnus incana, Alnus viridis, Castanea sativa, Corylus
@@
 Salix alba, Sorbus aria, Sorbus aucuparia, Taxus baccata, Tilia
-platyphyllos, Ulmus glabra, Pinus cembra
+platyphyllos, Ulmus glabra, Pinus cembra*
 
@@
 simulator. Acta Facultatis Forestalis Zvolen, 51(3), p. 125-137, ISSN:
-0231-5785 Fabrika, M., (2005): Simulátor biodynamiky lesa SIBYLA,
-koncepcia, konštrukcia a programové riešenie. Habilitačná práca.
-Technická univerzita vo
-
-Zvolene, 238 p. Kahn, M., (1994): Modellierung der Höhenentwicklung
-ausgewählter Baumarten in Abhängigkeit vom Standort. Forstliche
-Forschungsber. München, Vol. 141, 221 p. Keeling C. D., R. B. Bacastow,
-A. E. Bainbridge, C. A. Ekdahl Jr., P. R. Guenther, L. S. Waterman, J.
-F. S. Chin 1976: Atmospheric carbon dioxide variations at Mauna Loa
-Observatory, Hawaii. Tellus. XXVIII, 6: 538-551,
+0231-5785
+
+Fabrika, M., (2005): Simulátor biodynamiky lesa SIBYLA, koncepcia,
+konštrukcia a programové riešenie. Habilitačná práca. Technická
+univerzita vo Zvolene, 238 p.
+
+Kahn, M., (1994): Modellierung der Höhenentwicklung ausgewählter
+Baumarten in Abhängigkeit vom Standort. Forstliche Forschungsber.
+München, Vol. 141, 221 p.
+
+Keeling C. D., R. B. Bacastow, A. E. Bainbridge, C. A. Ekdahl Jr., P. R.
+Guenther, L. S. Waterman, J. F. S. Chin 1976: Atmospheric carbon dioxide
+variations at Mauna Loa Observatory, Hawaii. Tellus. XXVIII, 6: 538-551,
 <https://doi.org/10.1111/j.2153-3490.1976.tb00701.x>
@@
 Rein- und Mischbestände. Forstliche Forschungsberichte München, Nr.115,
-358 p. Valent, P., Merganič J., Výbošťok J. 2018: Optimus -- Software
-for Multi-criteria analysis. TU Zvolen.
+358 p.
+
+Valent, P., Merganič J., Výbošťok J. 2018: Optimus -- Software for
+Multi-criteria analysis. TU Zvolen.
 
@@
 TreeMig simulates directly the mean dynamics of entire stands, but based
-on the variability of light conditions within the stand (distribution based approach, cf. Lischke et al. 1998). To come up with the required
+on the variability of light conditions within the stand (distribution
+based approach, cf. Lischke et al. 1998). To come up with the required
 200 random patches of 1 ha, temporal and spatial variability intrinsic
@@
 ```
-# Appendix 2: Supplementary figures and tables
+# Supplementary figures and tables
 
@@
 
-kable(zeroInflated,  format = "latex", caption = "Number of total samples, samples with zero recruitment and the proportion of zero recruitment per model and observed data.")|>
-        kableExtra::kable_styling(latex_options = "HOLD_position")|>
-        #kableExtra::landscape() |>
+kable(zeroInflated,  format = "latex", caption = "Number of total samples, samples with zero recruitment and the proportion of zero recruitment per model and observed data.") |>
+        kableExtra::kable_styling(latex_options = "HOLD_position") |>
         kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
@@
 sigDiff <- readr::read_csv(here::here("figures", "sigH7_10.csv"))
+sigDiff$p.adj <- NULL
+colnames(sigDiff) <- c("model", "y", "Group 1", "Group 2", "p", "p.signif")
 
 kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of Shannon index in each model for 7 and 10 cm threshold.") |>
-        kableExtra::kable_styling(latex_options = "HOLD_position")|>
-        #kableExtra::landscape() |>
-        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+        kableExtra::kable_styling(latex_options = "HOLD_position")
 ```
@@
 ```{r recrStand7, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 7cm.
-The black line marks where the mean Shannon index per site at recruitment and at the stand level would be equal", out.width='100%', fig.pos='H'}
+The black line marks where the mean Shannon index per site at recruitment and at the stand level would be equal.", out.width='100%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "recruitmentAdultRichness7.png"))
@@
 
-```{r recrStand10, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 10cm", out.width='100%', fig.pos='H'}
+```{r recrStand10, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 10cm.", out.width='100%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "recruitmentAdultRichness10.png"))
@@
 A1) groups models where the simulated species composition in adults is higher than the observed, A2) groups models where the simulated species composition in adults is similar to the observed, A3) groups models where the simulated species composition in adults is lower than the observed.
-NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%', fig.pos='H'}
+NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA.", out.width='100%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "allAgesRichness.png"))
@@
 B1) groups models where the simulated species composition in recruitment is higher than the observed, A2) groups models where the simulated species composition in recruitment is similar to the observed, A3) groups models where the simulated species composition in recruitment is lower than the observed.
-NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%', fig.pos='H'}
+NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA.", out.width='100%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "recruitmentRichness.png"))
@@
 kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold over 10 cm threshold between the observed values and the simulated values per model.") |>
-    kableExtra::kable_styling(latex_options = "HOLD_position")|>
-#kableExtra::landscape()
-kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+    kableExtra::kable_styling(latex_options = "HOLD_position") 
 ```
@@
 tabletrends <- readr::read_csv(here::here("figures", "mort7_10_trends.csv"))
-kable(tabletrends,  format = "latex", caption = "Models trends in mortality between 7 and 10cm, and the ingrowth at 7cm, based in the mean oer site .")|>
+tabletrends$Significance <- round(tabletrends$Significance, 4)
+tabletrends$Slope <- round(tabletrends$Slope, 4)
+kable(tabletrends,  format = "latex", caption = "Models trends in mortality between 7 and 10cm (represented in Figure S5), and the ingrowth at 7cm, based in the mean per site.
+The slope is the rate of change from the linear model between recruitment ratio 7-10 cm and the initial recruitment tree number.") |>
 kableExtra::kable_styling(latex_options = "HOLD_position")
@@
 ```{r mortality3 , echo=FALSE, fig.cap="Recruitment ratio between 7 and 10cm and the initial recruitment tree number at 7cm.
-Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke self-thinning ratio under evenaged conditions", out.width='70%', fig.pos='H'}
+Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke self-thinning ratio under evenaged conditions.", out.width='70%', fig.pos='H'}
 knitr::include_graphics(here::here("figures", "mort7_10_trends.png"))
@@
 
-```{r plotComp, echo=FALSE, warning=FALSE, fig.cap="Plot and summary statistics of the linear model between the overestimation proportion in ingrowth levels and the mean complexity value of the establisment modules", fig.show = 'hold', out.width = '70%', fig.pos='H'}
+```{r plotComp, echo=FALSE, warning=FALSE, fig.cap="Plot and summary statistics of the linear model between the overestimation proportion in ingrowth levels and the mean complexity value of the establisment modules.", fig.show = 'hold', out.width = '70%', fig.pos='H'}
 
@@
 kable(rtreesDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold of the observed values and the simulated values per model.") |>
-    kableExtra::kable_styling(latex_options = "HOLD_position") |>
-    kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
+    kableExtra::kable_styling(latex_options = "HOLD_position") 
 ```

### 2023-10-30 manuscript/review/authorsResponse.qmd
last author responses changes

@@
   - name: Heike	Lischke,
-  - name: Mats Mahnken,
   - name: Paola	Mairota,
@@
   - name: Xavier Morin,
+  - name: Mats Nieberg,
   - name: Werner Rammer,
@@
 
-# Internal instructions \[this section will be deleted\]
-
--   underlined comments are comments that still need attention
--   I will delete most of the quoted text and add Line numbers, all the L\[XX-XX\] will be added in the last round after all the changes are done
-
-## Journal formatting (ToDo list)
-
--   Respond to the review comments in the space provided.
-You can use this space to document any changes you make to the original manuscript.
-In order to expedite the processing of the resubmitted manuscript, please be as specific as possible in your response to the review comments.
-If you disagree with a point, please explain why.
-Please note that the field in ScholarOne does not retain type formats such as italics, boldface, or colors, so please format the responses accordingly.
-We suggest you upload a separate file for your "Author's Response to Reviewer Comments" document and use our template: <https://www.esa.org/wp-content/uploads/2021/04/Author-Response-to-Reviewers-Template.docx>.
-
--   A version uploaded in the file list should be given the designation "Additional File for Review but NOT for publication".
-
--   Please also include a "track changes" version of your file, along with a clean copy as your Main Document.
-This "track changes" version should be uploaded in the file list list and given the designation "Additional File for Review but NOT for publication".
-
--   IMPORTANT: Your original files are available to you when you upload your revised manuscript.
-Please replace files of the earlier version and delete any redundant files before completing the submission.
-
--   To submit your revised manuscript, log into <https://mc.manuscriptcentral.com/ecosphere> and enter your Author Center.
-You will find your manuscript title listed under "Manuscripts with Decisions." Under "Actions," click on "Create a Revision." Your manuscript number has been appended to denote a revision.
-Please DO NOT upload your revised manuscript as a new submission.
-
--   DEADLINE: We encourage you to submit your revised manuscript within three weeks from today's date. The final due date for submitting your manuscript with the same manuscript number is 14-Dec-2023.
-
--   If you are unable to meet the final deadline, you can request an extension by contacting the the journal's editorial staff ([esajournals\@esa.org](mailto:esajournals@esa.org)).
-Please note that we can only consider extensions of a few weeks.
-If you cannot submit by the final due date, you must create a new submission for your manuscript.
-If creating a new submission, be sure to note the previous manuscript ID (ECS23-0342) in the online submission form and include an "Author's Response to Reviewer Comments" document (using the template linked to earlier in this message) in the file list.
-
--   Editorial staff note: In addition to revisions according to the review comments below, please make revisions to your paper to conform to the journal style at this stage, especially the tables, figures, and supporting information, anticipating the possibility that your paper may be accepted for publication.
-
--   Please provide the main manuscript file composed in Word, taking care to use the Table function for the tables.
-However, if you created your manuscript in LaTex, please provide only the .tex file of the main manuscript, a .pdf file of the main manuscript (that is the exact same in content as the .tex file), and a .bib file.
-
--   Ecosphere page style allows for five levels of headings.
-Please use different type styles (italics, boldface, boldface italics, etc.) consistently to distinguish between the levels of headings in the article and delete the section numbers in the headings.
-
--   Literature references: Please check that each citation has complete information, and check that all references in text have a matching citation in the References section and vice versa.
-
--   Table and figure references: Please be sure that all tables and all figures have in-text citations in the main body of the manuscript, and that tables and figures are called out in numerical order (i.e., Figure 1 is called out before Figure 2).
-
--   Figures: Each figure should be provided in an individual, complete file (all panels in one image, identified by figure number in the file name), preferably as a high-resolution (300-600 dpi) image file or high-quality PDF.
-Preferred file formats include TIF, EPS, PDF, or AI at 600 ppi, while JPEG, PPT/PPTX, or DOC/DOCX are acceptable if the resolution is sufficient.
-Only one version of each figure should be provided.
-Please delete figure images from the main manuscript file.
-Please check for misspellings and be sure that all parts, including all lettering, symbols, labels, and scale numbers, are discernible, clear, and readable when the figures are presented at the maximum allowable size (18 cm wide by 22 cm tall) in the typeset and paged PDF version of the article.
-See additional instructions for preparation in the Manuscript Preparation Guide.
-
--   Figure captions: Please be sure to include a list of figure captions in the main manuscript file and ensure that any abbreviation that appears in a figure is identified in the caption.
-
--   Supporting Information: Each appendix must be assembled exactly as you intend it to appear online.
-(Appendices are posted "as is"; ESA and Wiley staff do not copy edit or assemble appendices.) Note also that the Ecosphere naming convention for appendices is of the style "Appendix S1," "Appendix S2," etc., and for tables and figures within appendices is "Table S1," "Table S2," etc.
-or "Fig.
-S1," "Fig.
-S2," etc., and requires that the numbering begins at "S1" in each appendix (and revise references to material in other appendices as necessary).
-Please be sure to eliminate line numbering and to include the title of the journal, the title of the article, and all of the authors' names on the first page of each appendix.
-In addition, if there are literature references within an appendix, the appendix should include a list of literature cited.
-
--   If the journal style is not followed at submission of the revision, we may return the manuscript to you for further revision before sending it along to the Subject-matter Editor.
-Closely following our manuscript preparation guide at this stage would expedite the production of your paper for publication, should it be accepted.
-You can download our guide here: <https://www.esa.org/wp-content/uploads/2022/05/ESA-Manuscript-Preparation-Guide.pdf>
-
 # Response to the Editor
@@
 We have followed your suggestion and have replaced the erroneous use of 'regeneration' by 'recruitment' where we refer to ingrowth into a certain DBH class.
-However, in some places the term 'regeneration' is more accurate, especially in many parts of the Introduction and Discussion sections.
+However, in some places the term 'regeneration' is more accurate, especially in many parts of the introduction and discussion sections.
 'Regeneration' is a broader term that starts from flowering and ends with ingrowth into a given size class; that is, recruitment is the result of successful regeneration.
@@
 Thus, the 7 cm dataset (*n* = 165) actually is a subset of the 10 cm dataset (*n* = 200), as all data from the 7-cm threshold could be re-calculated for the 10-cm threshold.
-Thus, there was no 'like-to-like' comparison for any other diameter threshold, only for 7 and 10cm.
+Thus, there was no 'like-to-like' comparison for any other diameter threshold, only for 7 and 10 cm.
 
 When examining the change in mortality between 7 and 10 cm, or the difference in species diversity between observed and simulated data (Figure 4), we could only use the 165 sites that contained both pieces of information at the 7 and 10 cm thresholds in the observed data.
-This is reported in the supplementary materials software, script \`\`\`Figures.R\`\`\` line 536 and L284-349 and now better explained in the manuscript:
-
-1\) We have better explained the difference on the observed data in the section 3.2 Observed data L\[XX-XX\]
-
-"For our study, we defined two datasets where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds from 10 cm.
-Therefore we had in total 200 sites with observations of newly established trees above 10 cm and 165 sites above 7cm.
-In the results section any result from observed data at 7 cm threshold is originated from the 165 sites."
-
-2\) We have added a new text in the caption of figure 4 L\[XX-XX\]
-
-"B) Examples for the three trends across models for regeneration at the 7 cm threshold (165 sites)."
-
-### [COMMENT 3]{.underline}
+This is reported in the supplementary materials software, script \`\`\`Figures.R\`\`\` line 536 and L284-349 and now better explained in the manuscript: we have better explained the difference on the observed data in the section 3.2 Observed data L\[225-230\] and we have added a new text in the caption of figure 4 L\[1085-1086\].
+
+### COMMENT 3
 
@@
 
-The objective of running the different t-tests were to allow us for discussing irrefutable differences between models and observations.
-The t-test, as employed in our study, is not intended to serve as definitive proof of our findings.
+The objective of running the different t-tests were to allow us for discussing differences between mean values per site between models and observations and between 7 and 10 cm diameter thresholds.
+The t-test, as employed in our study, is not intended to serve as definitive proof of our findings and we have deliberately avoid over-interpretation of their outcomes.
 Rather, it serves as a practical tool for managing the extensive number of comparisons in our study.
-As such, we intend to include a statement in the manuscript:
-
-"Note that the t-test does not present conclusive evidence for a specific hypotheses, but rather facilitates managing the extensive number of comparisons discussed in our study."
-
-Regarding the other alternative approaches suggested by the reviewer: We can not do a paired t-test as the for paired samples it is necessary that you have the same number of subjects in both groups (observed and simulated) which is not the case.
-We tested the possibility of creating a GLME using as random effect the site and as a explanatory variables all the variables that we were interested in (model complexity, feedback, model type and scale).
-This model approach did not work as the combination as including some models having the same values in the combination of type and scale (factors did not have 2 or more levels).
-
--   "This more accurately answer your question about if any of these fixed explanatory variables can predict over or under representation of recruitment. " how to respond to this?
-
-Account for Type I error rates and adjust your p-values accordingly
+As such, we intend to include a statement in the manuscript to clarify this aspect, please see L\[336-337\].
+
+Following the reviewers comment we included a correction for multiple testing although we are aware that all corrections have pros and cons.
+In this case we used the Bonferroni correction to make sure that the actual Type I error probability for the combined H0 is smaller or equal than the nominal test level.
+
+We also look into the alternative approaches suggested by the reviewer.
+We can not do a paired t-test as the for paired samples it is necessary that you have the same number of subjects in both groups (observed and simulated) which is not the case as for example some models did not simulate all the sites.
+We tested the possibility of creating a GLME using as random effect the site and as a explanatory variables all the variables that we were interested in (model complexity, feedback, model type and scale) as suggested.
+However this approach will not give us the observed-model comparison that we are interested in.
+Following the reviewers comments we also did further testing to see if with other approaches we could better explain differences between observed and simulated values.
+We tested a mixed model to account for the randomness induce due to site (plot) differences.
+For example a glme model with the recruitment values as the response variable and models as the explanatory variable with observed data as the reference (with poisson distribution-count data).
+This worked well and the patterns were the same as in the t-test (expect for the model 4C).
+We also considered doing the same approach with the shannon index and mortality differences but in this case we think it is more interesting to understand the mean differences across sites and therefore a t-test is a valid approach.
+We also tested a linear mixed model to assess differences between 7 and 10cm on the Shannon index creating a new parameter where the explanatory variable was a combination of the model and the diameter threshold, then we run a pairwise multiple comparisons to determine which means differ.
+The patterns here were also similar to the ones observed in the t-test except for two landscape level models (TreeMig and LandisII) that were significant in this test contrary to the t-test.
+
+After considering all the options that the reviewer suggested and testing others, we made some adjustments to the t-tests analysis, with the correction for multiple testing and more clean output tables as suggested by the reviewer. In addition we explain the purpose and interpretation of these tests better in line \[336-337\].
 
@@
 
-We have added a new column classifying the models into recruitment or regeneration models; and the letters a,b,c, d,e and f in the table are now better described.
+We have added a new column classifying the models into recruitment or regeneration models; and the letters a, b, c, d, e and f in the table are now better described.
 Regarding the suggestion of adding the size of the saplings or ingrowth: we are concerned that this would be more confusing than helpful.
@@
 These were counted when we calculated the diversity index as well as for total recruitment values.
-We have added this information in the simulation protocol section L\[XX-XX\].
-
-"Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species simulated beyond the eleven species were lumped in a category"others""
+We have added this information in the simulation protocol section L\[278-281\].
 
@@
 
-Following the reviewers comment we have modified the text L\[XX-XX\]
-
-"Furthermore, regarding the share in the regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero."
+Following the reviewers comment we have modified the text L\[354-357\].
 
@@
 
-### [COMMENT 11]{.underline}
+### COMMENT 11
 
@@
 
-[To be addressed in combination with COMMENT 3]{.underline}
+We have cleaned all the output tables.
 
@@
 
-We have increased the font size of the legend, and now the difference between observed and simulated is visible. We have also explained the meaning of the black line in the caption L\[XX-XX\].
-
-"Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 7cm.
-The black line marks where the mean Shannon index per site at recruitment and at the stand level would be equal"
+We have increased the font size of the legend, and now the difference between observed and simulated is visible. We have also explained the meaning of the black line in the caption.
 
@@
 Figure 4 shows three representative patterns observed across models using 6 of the models as an example, and Figure S3 provides the details for each model.
-We believe it is not very informative to color the points from the models with feedback in Figure 4 as this will only color all the points from some of the panels and it is possible to identify the models with feedback in table 1 We have addressed this point in the discussion section L\[XX-XX\].
-
-"Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the regeneration than those without feedback, with the exception of ForCEEPS that significantly improved regeneration species diversity with the model version that included feedback (ForCEEPS(f)).
-The similar performance of models with and without feedback is likely because the models put more weight on the regeneration niche arising from abiotic and biotic filters, than from the habitat niche of the adult trees [@grubb1977]."
+We believe it is not very informative to color the points from the models with feedback in Figure 4 as this will only color all the points from some of the panels and it is possible to identify the models with feedback in table 1 We have addressed this point in the discussion section L\[563-570\].
 
@@
 
-Done.
+We have corrected these figures.
 
@@
 
-This paragraph is addressing the comparison with the theoretical Reineke value which is usually calculated for even-aged, single species stands. We have change this text to clarify this aspect.
+This paragraph is addressing the comparison with the theoretical Reineke value which is usually calculated for even-aged, single species stands.
+We have change this text to clarify this aspect L\[324-330\].
 
@@
 
-### [COMMENT 17]{.underline}
+### COMMENT 17
 
@@
 
-[Clean together with comment 3]{.underline}
+We have cleaned this table and changed the table caption including a reference to Figure S5 as suggested by the reviewer.
 
@@
 
-With this statement, we were referring to the multiple potential combinations of approaches of modeling tree regeneration. We have modified this for clarity L\[XX-XX\]
+With this statement, we were referring to the multiple potential combinations of approaches of modeling tree regeneration. We have modified this for clarity L\[596-598\]
 
@@
 
-This was a glitch, which now is corrected L\[ XX-XX\]
+This was a glitch, which now is corrected L\[636-638\]
 
@@
 
-We have changed this text L\[XX-XX\].
-Although there are multiple data sources (such as MASTIF, MASTREE, National Forest Inventories, Forest Reserve plots, etc.), usually these data do not refer to the same locations and thus cannot be used in a site-specific manner for calibrating or testing models that intend to follow the ‚regeneration' chain of processes.
-
-"Therefore, we further recommend that more effort should be invested into collecting harmonized datasets in a site-specific manner covering the different aspects leading to tree regeneration."
+We have changed this text L\[684-688\].
+Although there are multiple data sources (such as MASTIF, MASTREE, National Forest Inventories, Forest Reserve plots, etc.), usually these data do not refer to the same locations and thus cannot be used in a site-specific manner for calibrating or testing models that intend to follow the "regeneration" chain of processes.
 
@@
 
-We have added a "road map" at the end of the introduction.
+We have added a "road map" at the end of the introduction L\[159-163\].
 
@@
 
-We have modified this sentence L\[XX-XX\]
-
-"However, there is a need to evaluate their capacity to accurately represent tree regeneration"
+We have modified this sentence L\[52-53\]
 
@@
 
-We have changed the word "comprehensive" to "extensive" to express that we used a wide-range data that considers many important elements of regeneration.
+We have changed the word "comprehensive" to "extensive" to express that we used a wide-range data that considers many important elements of regeneration L\[57\]
 
@@
 
-We have changed "feature" with "present".
+We have kept "feature" in this case.
 
@@
 
-We have changed the word "were" with "have been".
+We have changed the word "were" with "have been" L\[78\].
 
@@
 
-We have corrected the reference style.
+We have corrected the reference style L\[80\].
 
@@
 
-We have changed this sentence to: "This presents an important research gap, particularly in the context of climate-induced forest disturbances and forest resilience."
+We have changed this sentence L\[84-85\].
 
@@
 
-We have changed "such as" with "including"
+We have changed "such as" with "including" L\[86\].
 
@@
 
-We have changed "it" with "them"
+We have changed "it" with "them" L\[92\].
 
@@
 
-We have added a reference here.
+We have added a reference here L\[105\].
 
@@
 
-We have added information to briefly explain the general approach.
+We have added information to briefly explain the general approach L\[196-197\].
 
@@
 
-Yet we acknowledge that our statement needs to be put in a European context much clearer. We hope that our adjustments to the text resolve the exaggeration of our initial formulation.
+Yet we acknowledge that our statement needs to be put in a European context much clearer. We hope that our adjustments to the text resolve the exaggeration of our initial formulation L\[211-213\].
 
@@
 
-We have changed "and exposition (i.e.,slope and aspect)." with ", slope and aspect."
+We have changed "and exposition (i.e.,slope and aspect)." with ", slope and aspect." L\[223\].
 
@@
 
-Following the reviewer's comment we have changed this sentence to: "Regeneration minimum size thresholds for these sites differed between stem diameters of 0 and 10 cm"
+Following the reviewer's comment we have changed this sentence L\[224-225\].
 
@@
 
+L218 "featured 30,900 newly established trees" so they are trees....
+
 **Response to comment 26:**
@@
 
-L215 "another 35 sites with diameter thresholds between 7 and 10 cm." OK, so we're talking about v large trees here
-
-L218 "featured 30,900 newly established trees" so they are trees....
+L215 "on this unique dataset" not so unique
 
@@
 
-The minimum threshold that could be standardized across reserves was of at least 7cm.
+We have changed this text L\[235\].
 
@@
 
-L215 "on this unique dataset" not so unique
+L221 "nformation, cf. Käber et al. (2023)." in review and not available
 
@@
 
-We have changed this text L\[XX-XX\]
+This paper have recently being published and the reference has been updated accordingly.
 
@@
 
-L221 "nformation, cf. Käber et al. (2023)." in review and not available
+L223 "The overarching goal of the experiments was to assess the tree regeneration as it arises" simulation experiments and remove "the"
+
+**Response to comment 29:**
+
+Changed as suggested L\[238-239\].
+
+### COMMENT 30
+
+L225 "tree regeneration as the passing of a breast height diameter threshold of 7 or 10 cm, respectively" I suggest "tree recruitment" not "tree regeneration" here. respectively??
+
+**Response to comment 30:**
+
+We have removed "respectively" L\[240\].
+
+We have re-assessed the terms used across this manuscript and change it across the text accordingly.
+
+### COMMENT 31
+
+L 227-228 "simulations, providing input variables" provided with
+
+**Response to comment 31:**
+
+We have removed "providing" L\[242\].
+
+### COMMENT 32
+
+L229 "Neither were further site information (except for the data specified below) nor any data" These were "blind trials":  modelling group were not provided with site information \...
 
 **Response to comment 32:**
-
-This paper have recently being published and the reference has been updated accordingly.
-
-### COMMENT 30
-
-L223 "The overarching goal of the experiments was to assess the tree regeneration as it arises" simulation experiments and remove "the"
-
-**Response to comment 30:**
-
-Changed as suggested
-
-### COMMENT 31
-
-L225 "tree regeneration as the passing of a breast height diameter threshold of 7 or 10 cm, respectively" I suggest "tree recruitment" not "tree regeneration" here. respectively??
-
-**Response to comment 31:**
-
-We have removed "respectively".
-
-We have re-assessed the terms used across this manuscript and change it across the text accordingly.
-
-### COMMENT 32
-
-L 227-228 "simulations, providing input variables" provided with
-
-**Response to comment 32:**
-
-We have removed "providing".
-
-### COMMENT 33
-
-L229 "Neither were further site information (except for the data specified below) nor any data" These were "blind trials":  modelling group were not provided with site information \...
-
-**Response to comment 33:**
 
@@
 
+### COMMENT 33
+
+L230-231 "That is, the models were run in"blind flight" mode." Remove
+
+**Response to comment 33:**
+
+Changed as suggested.
+
 ### COMMENT 34
 
-L230-231 "That is, the models were run in"blind flight" mode." Remove
+L237-238 "Soil quality data were provided as continuous values between 1 and 5" ?hydrological quality?
 
 **Response to comment 34:**
-
-Changed as suggested.
-
-### COMMENT 35
-
-L237-238 "Soil quality data were provided as continuous values between 1 and 5" ?hydrological quality?
-
-**Response to comment 35:**
 
@@
 
-### COMMENT 36
+### COMMENT 35
 
@@
 
+**Response to comment 35:**
+
+We understand the reviewers concern on not constraining the length of the simulations, but we decided to leave it to the assessment of the individual modelers, as we wanted to obtain simulation that were run from bare ground in the absence of management to a simulated equilibrium ("potential natural vegetation") with current climate.
+Each modeler know best how to obtain this in each model and different model uses its own logic to get to the equilibriium that can require different lengths.
+
+### **CO**MMENT 36
+
+L250 "The simulations were run in the absence of management to a simulated equilibrium" The absence of disturbance and running to equilibrium set quite specific conditions; suggest this is emphasised more in abstract / intro / discussion.
+
 **Response to comment 36:**
-
-We understand the reviewers concern on not constraining the length of the simulations, but we decided to leave it to the assessment of the individual modelers, as we wanted to obtain simulation that were run from from bare ground in the absence of management to a simulated equilibrium ("potential natural vegetation") with current climate, and each modeler know best how to obtain this in each model.
-This information is currently available in the manuscript and in the protocol of this study.
-
-### **CO**MMENT 37
-
-L250 "The simulations were run in the absence of management to a simulated equilibrium" The absence of disturbance and running to equilibrium set quite specific conditions; suggest this is emphasised more in abstract / intro / discussion.
-
-**Response to comment 37:**
 
@@
 
-### COMMENT 38
+### COMMENT 37
 
@@
 
-Response to comment 38:
+Response to comment 37:
 
@@
 The data were collected in forest reserves where no management has taken place for long periods of time which makes our assumption of an equilibrium between forest properties and environmental drivers reasonable.
-This is the best data we have and we have address its limitations, as this one, in the discussion because we are also aware that is not ideal.
+This is the best data we have and we have addressed its limitations, as this one, in the discussion because we are also aware that is not ideal.
+
+### COMMENT 38
+
+L275 "Each of the models reported the regeneration number by sampling 200 times in a 10-year interval" regeneration number = number of "recruits" entering the 7-cm size class?
+
+Response to comment 38:
+
+Following the reviewer's comment we have changed this sentence L\[290-291\]
 
@@
 
-L275 "Each of the models reported the regeneration number by sampling 200 times in a 10-year interval" regeneration number = number of "recruits" entering the 7-cm size class?
+L295 "(2) regeneration species diversity" species diversity of recruitment
+
+L295 "(3) regeneration mortality" mortality of recruits
 
@@
 
-Following the reviewer's comment we have changed "regeneration number" with "number of trees crossing the two diameter thresholds"
+We have changed the terminology.
 
@@
 
-L295 "(2) regeneration species diversity" species diversity of recruitment
-
-L295 "(3) regeneration mortality" mortality of recruits
-
-Response to comment 40:
-
-We have changed the terminology.
-
-### COMMENT 41
-
 L296 "ingrowth gradients along the regeneration niches." ?? not clear to me. How is regeneration niche defined??
 
-**Response to comment 41:**
+**Response to comment 40:**
 
@@
 
-### COMMENT 42
+### COMMENT 41
 
@@
 
-**Response to comment 42:**
+**Response to comment 41:**
 
@@
 
+### COMMENT 42
+
+L308 "Mortality in tree regeneration was assessed based on the ratio of regeneration between the 7 and 10" [of recent recruits.]{.underline} Now I understand the 7 and 10 cm thresholds mentioned earlier;  please clarify earlier mention.
+
+**Response to comment 42:**
+
+Following the reviewer's comment on the lack of clarity of the selected thresholds explanation we have changed at the beginning of the section simulation protocol the first sentence where we explain that we considered tree regeneration as the passing of a breast height diameter threshold of 7 and 10 cm.
+We have also reported later on in this section (but before this line) that each of the models reported the number of trees crossing the two diameter thresholds L\[290-291\].
+
 ### COMMENT 43
-
-L308 "Mortality in tree regeneration was assessed based on the ratio of regeneration between the 7 and 10" [of recent recruits.]{.underline} Now I understand the 7 and 10 cm thresholds mentioned earlier;  please clarify earlier mention.
-
-**Response to comment 43:**
-
-Following the reviewer's comment on the lack of clarity of the selected thresholds explanation we have changed at the beginning of the section simulation protocol the first sentence where we explain that we considered tree regeneration as the passing of a breast height diameter threshold of 7 and 10 cm.
-We have also reported later on in this section (but before this line) that each of the models reported the number of trees crossing the two diameter thresholds.
-L\[XX-XX\]
-
-### COMMENT 44
 
@@
 
-**Response to comment 44:**
+**Response to comment 43:**
 
@@
 
-### COMMENT 45
+### COMMENT 44
 
@@
 
-**Response to comment 45:**
+**Response to comment 44:**
 
@@
 
+### COMMENT 45
+
+L319-321 - overly complex sentence structure
+
+**Response to comment 45:**
+
+We have improved this sentence L\[334-337\].
+
 ### COMMENT 46
 
-L319-321 - overly complex sentence structure
+L323-324 presumably, recruits make up a very small proportion of stand basal if the models have been run to equilibrium?
 
@@
 
-We have improved this sentence L\[XX-XX\]
-
-"We tested for significant differences between simulated and observed values of the recruitment levels and species diversity by using paired t-tests."
+Yes the reviewer is correct, the recruits represent a small proportion of the total stand basal area.
 
@@
 
-L323-324 presumably, recruits make up a very small proportion of stand basal if the models have been run to equilibrium?
+L327 of each species?
 
@@
 
-Yes the reviewer is correct, the recruits represent a small proportion of the total stand basal area.
+Added as suggested L\[343\].
 
@@
 
-L327 of each species?
+L332-335 This give a matric of species-level variation in recruitment across sites, but how does this relate to the regeneration niche concept of Grubb 1977?
 
@@
 
-Added as suggested.
+We have changed the text to clarify this point. L\[350-351\].
 
@@
 
-L332-335 This give a matric of species-level variation in recruitment across sites, but how does this relate to the regeneration niche concept of Grubb 1977?
+L496 "confronted with a unique dataset from unmanaged"
+
+L498 "The EuFoRIa data (Käber et al. 2023) are exceptional, particularly"
+
+I think there are plenty of other datasets that could be used and have better information on smaller trees than EuFoRIa,  so suggest a more nuanced sales pitch here.
 
@@
 
-We have changed the text to clarify this point. L\[XX-XX\]
-
-"For these species, we calculated the share in the recruited basal area per site, where each site represents a part of the environmental space, as the mean across the available samples per site:"
+We agree with the reviewer's comment that there are other forest inventory data networks outside Central Europe even on forest reserves that could provide this sort of data.
+However, within Europe no such data existed before, e.g.
+from unmanaged forests and collecting such a large number of records.
 
 ### COMMENT 50
-
-L496 "confronted with a unique dataset from unmanaged"
-
-L498 "The EuFoRIa data (Käber et al. 2023) are exceptional, particularly"
-
-I think there are plenty of other datasets that could be used and have better information on smaller trees than EuFoRIa,  so suggest a more nuanced sales pitch here.
-
-**Response to comment 50:**
-
-We have changed the text L\[XX-XX\]
-
-### COMMENT 51
 
@@
 
-**Response to comment 51:**
+**Response to comment 50:**
 
 Thank you for this remark, we agree with the reviewer that this is an important point and specially how different models are capturing or missing to represent the absence of regeneration.
-We have now captured this aspect also in the introduction in the sentence L\[XX-XX\]
-
-"Due to the large variability in tree regeneration patterns in nature and the large number of factors driving and constraining this process" on top of the current whole paragraph on this issue present in the discussion.
-
-### COMMENT 52
+We have now captured this aspect also in the introduction in the sentence, L\[147-148\], on top of the current whole paragraph on this issue present in the discussion.
+
+### COMMENT 51
 
@@
 
+**Response to comment 51:**
+
+We have included this point by adding into this paragraph a new sentence: L\[544-546\].
+
+### COMMENT 52
+
+L552 "there is no evidence that models with feedback from the canopy captured better the species" could you explain what you mean by "feedback" here?
+
 **Response to comment 52:**
 
-We have included this point by adding into this paragraph: "Another possible reason for this difference may be the exclusion of factors like deer browsing or failure to accurately measure establishment filters, such as those to site limitants or competition with herb layers."
+We have modified this sentence to clarify what do we mean here with feedback L\[563-567\]
+
+This had also been defined in the material and methods where we stated L\[180-182\].
 
@@
 
-L552 "there is no evidence that models with feedback from the canopy captured better the species" could you explain what you mean by "feedback" here?
+L586-587 "Our study showed that increasing complexity in the regeneration modules is not linked with a higher accuracy of the projections of regeneration levels," Surely more complex models WOULD be better if they captured the relevant processes!  Arguing for simple vs complex models is a distraction here\...
+we need models that are as simple as possible but not too simple..
 
 **Response to comment 53:**
-
-We have modified this sentence to clarify what do we mean here with feedback: " Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the regeneration than those without feedback, "
-
-This had also been defined in the material and methods where we stated that "i.e., where the level and species composition of regeneration is influenced by the existing adult tree community via the production of seeds, seedlings or saplings, from models that do not contain such a feedback."
-
-### COMMENT 54
-
-L586-587 "Our study showed that increasing complexity in the regenerationmodules is not linkedwith a higher accuracy of the projections of regeneration levels," Surely more complex models WOULD be better if they captured the relevant processes!  Arguing for simple vs complex models is a distraction here\...
-we need models that are as simple as possible but not too simple..
-
-**Response to comment 54:**
 
@@
 
-### COMMENT 55
+### COMMENT 54
 
@@
 
-**Response to comment 55:**
-
-Following the reviewers comment we have added the following sentence in this part of the text:
-
-"(i.e., decreasing regeneration levels with increasing stand basal area)"
-
-### COMMENT 56
+**Response to comment 54:**
+
+Following the reviewers comment we have added the a sentence in this part of the text L\[611-612\].
+
+### COMMENT 55
 
@@
 
+**Response to comment 55:**
+
+We have decided not to include this aspect in the abstract due to lack of space.
+
+### COMMENT 56
+
+L656 "Yet, the global models should not be at a disadvantage due to the limited spatial consideration" Don't quite follow your point here. what does "at a disadvantage" mean?
+
 **Response to comment 56:**
 
-We have decided not to include this aspect in the abstract due to lack of space.
+In this sentence we are discussing how the sampling strategy presented in the study protocol might have represented a disadvantage for landscape level models due to the limited spatial scale, but how this might not be the same for global level models as they usually lack dispersal between cells.
 
 ### COMMENT 57
-
-L656 "Yet, the global models should not be at a disadvantage due to the limited spatial consideration" Don't quite follow your point here. what does "at a disadvantage" mean?
-
-**Response to comment 57:**
-
-In this sentence we are discussing how the sampling strategy presented in the study protocol might have represented a disadvantage for landscape level models due to the limited spatial scale, but how this might not be the same for global level models as they usually lack dispersal between cells.
-
-### COMMENT 58
 
@@
 
-**Response to comment 58:**
-
-We have changed this text L\[XX-XX\]
-
-### COMMENT 59
+**Response to comment 57:**
+
+We have changed this text L\[687-688\].
+
+### COMMENT 58
 
@@
 
+**Response to comment 58:**
+
+We agree with the reviewer that a more systematic kind of inventory could be useful to detect regeneration patterns under management or disturbances impacts.
+
+### COMMENT 59
+
+L696 "Exercises like the one presented here, where the models are operated in"blind flight" mode." Definition of blind flight "To do something based on guesswork, intuition, or without any help or instructions"....is that what you mean?
+
 **Response to comment 59:**
 
-We agree with the reviewer that a more systematic kind of inventory could be useful to detect regeneration patterns under management or disturbances impacts.
+With blind flight we refer to doing the simulations without having important information about the expected regeneration values.
 
 ### COMMENT 60
-
-L696 "Exercises like the one presented here, where the models are operated in"blind flight" mode." Definition of blind flight "To do something based on guesswork, intuition, or without any help or instructions"....is that what you mean?
-
-**Response to comment 60:**
-
-With blind flight we refer to doing the simulations without having important information about the expected regeneration values.
-
-### COMMENT 61
 
@@
 
+**Response to comment 60:**
+
+Following the reviewers comment we have expanded the caption of table 1.
+
+### COMMENT 61
+
+L831 Add publisher
+
+L939 Correct doi
+
 **Response to comment 61:**
 
-Following the reviewers comment we have expanded the caption of table 1 with the following text:
-
-"Models included, their characteristics and simulation strategies used in this study.
-Where *Scale* refers to the models scale, *Type* to the models type, *Population structure* to the model population approach, *Feedback* to the inclusion or not of adult tree feedback to the level and species composition of regeneration, *Approach* to the model regeneration module approach, *Start from* to the stage from which trees are recruited, *Complexity,* to *t*he mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), *Species* to which species from the original eleven species were simulated, *Runtime (spin up) \[years\]* to the number of years used per model in their simulations for the spin up period and regeneration sampling; *Climate data type (use of years)* to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details.
-Further information can be found on the individual model reports available in the supplementary materials."
-
-### COMMENT 62
-
-L831 Add publisher
-
-L939 Correct doi
-
-**Response to comment 62:**
-
 Changed as suggested.

### 2023-10-30 manuscript/review/journalIntructionsqmd.qmd
journal instructions for revision

---
title: "Untitled"
format: pdf
editor: visual
---

## Internal instructions \[this section will be deleted\]

-   underlined comments are comments that still need attention
-   I will delete most of the quoted text and add Line numbers, all the L\[XX-XX\] will be added in the last round after all the changes are done

## Journal formatting (ToDo list)

-   Respond to the review comments in the space provided. You can use this space to document any changes you make to the original manuscript. In order to expedite the processing of the resubmitted manuscript, please be as specific as possible in your response to the review comments. If you disagree with a point, please explain why. Please note that the field in ScholarOne does not retain type formats such as italics, boldface, or colors, so please format the responses accordingly. We suggest you upload a separate file for your "Author's Response to Reviewer Comments" document and use our template: <https://www.esa.org/wp-content/uploads/2021/04/Author-Response-to-Reviewers-Template.docx>.

-   A version uploaded in the file list should be given the designation "Additional File for Review but NOT for publication".

-   Please also include a "track changes" version of your file, along with a clean copy as your Main Document. This "track changes" version should be uploaded in the file list list and given the designation "Additional File for Review but NOT for publication".

-   IMPORTANT: Your original files are available to you when you upload your revised manuscript. Please replace files of the earlier version and delete any redundant files before completing the submission.

-   To submit your revised manuscript, log into <https://mc.manuscriptcentral.com/ecosphere> and enter your Author Center. You will find your manuscript title listed under "Manuscripts with Decisions." Under "Actions," click on "Create a Revision." Your manuscript number has been appended to denote a revision. Please DO NOT upload your revised manuscript as a new submission.

-   DEADLINE: We encourage you to submit your revised manuscript within three weeks from today's date. The final due date for submitting your manuscript with the same manuscript number is 14-Dec-2023.

-   If you are unable to meet the final deadline, you can request an extension by contacting the the journal's editorial staff ([esajournals\@esa.org](mailto:esajournals@esa.org)). Please note that we can only consider extensions of a few weeks. If you cannot submit by the final due date, you must create a new submission for your manuscript. If creating a new submission, be sure to note the previous manuscript ID (ECS23-0342) in the online submission form and include an "Author's Response to Reviewer Comments" document (using the template linked to earlier in this message) in the file list.

-   Editorial staff note: In addition to revisions according to the review comments below, please make revisions to your paper to conform to the journal style at this stage, especially the tables, figures, and supporting information, anticipating the possibility that your paper may be accepted for publication.

-   Please provide the main manuscript file composed in Word, taking care to use the Table function for the tables. However, if you created your manuscript in LaTex, please provide only the .tex file of the main manuscript, a .pdf file of the main manuscript (that is the exact same in content as the .tex file), and a .bib file.

-   Ecosphere page style allows for five levels of headings. Please use different type styles (italics, boldface, boldface italics, etc.) consistently to distinguish between the levels of headings in the article and delete the section numbers in the headings.

-   Literature references: Please check that each citation has complete information, and check that all references in text have a matching citation in the References section and vice versa.

-   Table and figure references: Please be sure that all tables and all figures have in-text citations in the main body of the manuscript, and that tables and figures are called out in numerical order (i.e., Figure 1 is called out before Figure 2).

-   Figures: Each figure should be provided in an individual, complete file (all panels in one image, identified by figure number in the file name), preferably as a high-resolution (300-600 dpi) image file or high-quality PDF. Preferred file formats include TIF, EPS, PDF, or AI at 600 ppi, while JPEG, PPT/PPTX, or DOC/DOCX are acceptable if the resolution is sufficient. Only one version of each figure should be provided. Please delete figure images from the main manuscript file. Please check for misspellings and be sure that all parts, including all lettering, symbols, labels, and scale numbers, are discernible, clear, and readable when the figures are presented at the maximum allowable size (18 cm wide by 22 cm tall) in the typeset and paged PDF version of the article. See additional instructions for preparation in the Manuscript Preparation Guide.

-   Figure captions: Please be sure to include a list of figure captions in the main manuscript file and ensure that any abbreviation that appears in a figure is identified in the caption.

-   Supporting Information: Each appendix must be assembled exactly as you intend it to appear online. (Appendices are posted "as is"; ESA and Wiley staff do not copy edit or assemble appendices.) Note also that the Ecosphere naming convention for appendices is of the style "Appendix S1," "Appendix S2," etc., and for tables and figures within appendices is "Table S1," "Table S2," etc. or "Fig. S1," "Fig. S2," etc., and requires that the numbering begins at "S1" in each appendix (and revise references to material in other appendices as necessary). Please be sure to eliminate line numbering and to include the title of the journal, the title of the article, and all of the authors' names on the first page of each appendix. In addition, if there are literature references within an appendix, the appendix should include a list of literature cited.

-   If the journal style is not followed at submission of the revision, we may return the manuscript to you for further revision before sending it along to the Subject-matter Editor. Closely following our manuscript preparation guide at this stage would expedite the production of your paper for publication, should it be accepted. You can download our guide here: <https://www.esa.org/wp-content/uploads/2022/05/ESA-Manuscript-Preparation-Guide.pdf>

### 2023-10-30 code/figures.R
delete all the code from GLME and LMM tests

@@ -231,70 +231,6 @@
     
      write.csv(sigH7_10,"figures/sigH7_10.csv", row.names = FALSE)
     
-     #####  GLME approach -----
-     # briefly thought about how to analyse your data (Shannon index and mortality),
-     #  when you want to focus on the differences between the 7 and 10 cm thresholds 
-     #  within the same model. We actually use a function in our applied stats 
-     #  course that we apply to anova, but I didn’t know whether this will 
-     #  work with random effects as well.
-     # 
-     # You can try to use the function "glht" in the R library "multcomp",
-     #  which allows to define specific contrasts that you want to test. 
-     #  The function can also be applied to mixed effects models. 
-     #  You can try whether the following works: Calculate a repeated-measures ANOVA, 
-     #  i.e. a linear mixed-effects model, e.g. lmer(y ~ model.thr + (1 | sample)), 
-     #  where y is the response variable and model.thr a factor, which combines 
-     #  the model (e.g. LandClim) and the threshold (e.g. 7 cm), e.g. LandClim.7, 
-     #  LandClim.10. For the Shannon index, you can probably assume a linear
-     #   mixed-effects model with log(y) as response. For the mortality data, 
-     #   the best would be to use a binomial distribution, because you know the 
-     #   number of dead trees and the number of surviving trees, i.e. use 
-     #   glmer(cbind(dead, alive) ~ model.thr + (1 | sample), family=binomial(link=“logit”)).
-     # 
-     # The reference level for model.thr could be the observations. 
-     # Then you apply the function "glht" on this model. You will have to define
-     #  in the argument "linfct" the contrasts that you want to test, 
-     #  i.e. you only test between the data of the 7 and 10 cm threshold 
-     #  of the same model. This will require some time to figure out, 
-     #  how this should be done, but you can find some examples in the R-help or online.
-     #   If my explanation was very unclear we have to discuss this in person.
-     # 
-     
-     
[…]

### 2023-10-31 manuscript/manuscriptRW.Rmd
change supplementary wiyh appendix

@@
 details on how each modeling team prepared the simulations and the
-outputs are available in Supplementary Material 1.
+outputs are available in Appendix S1: Model reports.
 
@@
 between simulated and observed values of the recruitment levels and
-species diversity by using t-tests with a correction for multiple
-testing. Note that the t-tests does not present conclusive evidence for
-a specific hypotheses, but rather facilitates managing the extensive
-number of comparisons discussed in our study.
+species diversity by using t-tests with the Bonferroni correction for
+multiple testing. Note that the t-tests does not present conclusive
+evidence for a specific hypotheses, but rather facilitates managing the
+extensive number of comparisons discussed in our study.
 
@@
 species diversity of recruitment between 7 and 10 cm across sites
-(Tables S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS)
-maintained the recruitment diversity between the 7 and 10 cm thresholds
-(i.e., the differences between them were not significant, see Table S2),
-and the same was evident from the observed data.
+(Appendix S1: Table S2). Four models (iLand, Landis-II, TreeMig and
+LPJ-GUESS) maintained the recruitment diversity between the 7 and 10 cm
+thresholds (i.e., the differences between them were not significant, see
+Appendix S1: Table S2), and the same was evident from the observed data.
 
@@
 thresholds, species diversity in the recruitment compared to stand-level
-diversity did not feature clear patterns across the models (Figures S1
-and S2). Most models captured reasonably well or overpredicted species
-diversity at the stand level (Figure 4A, left and center; cf. Figure
-S3), but a characteristic feature was that the simulations had a much
-lower variability of diversity compared to observations. Models
-overpredicting stand-level species diversity included ForClim1, ForClim
-11, FORMIND and TreeMig (Figure 4A, left). Several models underpredicted
-species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure 4A,
-right).
+diversity did not feature clear patterns across the models (Appendix S1:
+Figures S1 and S2). Most models captured reasonably well or
+overpredicted species diversity at the stand level (Figure 4A, left and
+center; cf. Appendix S1: Figure S3), but a characteristic feature was
+that the simulations had a much lower variability of diversity compared
+to observations. Models overpredicting stand-level species diversity
+included ForClim1, ForClim 11, FORMIND and TreeMig (Figure 4A, left).
+Several models underpredicted species diversity, i.e., LPJ-GUESS,
+ForCEEPS(f) and 4C (Figure 4A, right).
 
@@
 already observed in Figure 3, only two of them underestimated it (4C and
-Landis-II) (Figure 4B and Figure S4). Both the models with and without
-feedback from the adult trees to regeneration (via seed production)
-overestimated, underestimated or captured reasonably well recruitment
-species diversity compared to the observations. Again, simulated
-diversity had considerably lower variability than observed diversity in
-the recruitment.
+Landis-II) (Figure 4B and Appendix S1: Figure S4). Both the models with
+and without feedback from the adult trees to regeneration (via seed
+production) overestimated, underestimated or captured reasonably well
+recruitment species diversity compared to the observations. Again,
+simulated diversity had considerably lower variability than observed
+diversity in the recruitment.
 
@@
 partially compensating for the general overestimation of recruitment
-levels at 7 cm (Figure 2 and Table S3); six models did not feature a
-significant difference compared to the observed data (ForCEEPS, FORMIND,
-ForClim 1 and 11, iLand and aDGVM2). Four models (xComp, PICUS, LandClim
-and TreeMig) featured very high mortality rates (i.e., well above the
-Reineke self-thinning line), which compensated for the strong initial
-overestimation of recruitment (cf. Figure 2). Interestingly, two models
-that underestimated overall recruitment levels (SIBYLA and Landis-II;
-cf. Figure 2) featured mortality rates that were close to but still
-above the self-thinning line (Figure 5).
+levels at 7 cm (Figure 2 and Appendix S1: Table S3); six models did not
+feature a significant difference compared to the observed data
+(ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2). Four models
+(xComp, PICUS, LandClim and TreeMig) featured very high mortality rates
+(i.e., well above the Reineke self-thinning line), which compensated for
+the strong initial overestimation of recruitment (cf. Figure 2).
+Interestingly, two models that underestimated overall recruitment levels
+(SIBYLA and Landis-II; cf. Figure 2) featured mortality rates that were
+close to but still above the self-thinning line (Figure 5).
 
@@
 threshold also had a small decrease of recruitment between 7 and 10 cm
-(ratio 7-10 cm) (Figure S5). In these models, competition and
-self-thinning are either not pronounced or must have occurred before the
-trees had reached 7 cm. However, this pattern was not consistent across
-models (Figure S5). It is noteworthy that there is a relationship
-between the recruitment levels at 7 cm and the mortality rate (Table
-S4): most of the models showed a positive effect (higher regeneration at
-7 cm is coupled to an increase of the mortality between 7 and 10 cm).
-While some models showed a negative effect, the linear trend was not
-pronounced and not always significant regarding its slope.
+(ratio 7-10 cm) (Appendix S1: Figure S5). In these models, competition
+and self-thinning are either not pronounced or must have occurred before
+the trees had reached 7 cm. However, this pattern was not consistent
+across models (Appendix S1: Figure S5). It is noteworthy that there is a
+relationship between the recruitment levels at 7 cm and the mortality
+rate (Appendix S1: Table S4): most of the models showed a positive
+effect (higher regeneration at 7 cm is coupled to an increase of the
+mortality between 7 and 10 cm). While some models showed a negative
+effect, the linear trend was not pronounced and not always significant
+regarding its slope.
 
@@
 regeneration module (Table \@ref(tab:tableModels)) and the
-overestimation proportion at a diameter threshold of 7 cm (Figure S6).
-All models except 4C had significant differences between the observed
-and simulated mean recruitment values. Therefore it was not possible to
-differentiate in terms of performance between models type (empirical or
-process based), or their scales (stand, landscape or global) (Table S5).
-The values of species diversity in the recruitment simulated by each
-model were also assessed against the observed data to see if there was a
-difference between models that include a feedback compared to those that
-did not; all the models except ForCEEPs(f) and FORMIND had significant
-differences (Table S6). It is noteworthy that for the model ForCEEPS the
-pattern changed from the overprediction of species diversity in the
-regeneration to a diversity level that is closer to observations when
-the recruitment module included feedback (ForCEEPs(f)).
+overestimation proportion at a diameter threshold of 7 cm (Appendix S1:
+Figure S6). All models except 4C had significant differences between the
+observed and simulated mean recruitment values. Therefore it was not
+possible to differentiate in terms of performance between models type
+(empirical or process based), or their scales (stand, landscape or
+global) (Appendix S1: Table S5). The values of species diversity in the
+recruitment simulated by each model were also assessed against the
+observed data to see if there was a difference between models that
+include a feedback compared to those that did not; all the models except
+ForCEEPs(f) and FORMIND had significant differences (Appendix S1: Table
+S6). It is noteworthy that for the model ForCEEPS the pattern changed
+from the overprediction of species diversity in the regeneration to a
+diversity level that is closer to observations when the recruitment
+module included feedback (ForCEEPs(f)).
 
@@
 reproduced the effect of basal area in both its magnitude and patterns
-(Figure 7 and S7) considerably better than the effects of the climatic
-gradients, where they featured varying patterns.
+(Figure 7 and Appendix S1: Figure S7) considerably better than the
+effects of the climatic gradients, where they featured varying patterns.
 
@@
 observed data along the two major climatic gradients (Figure 8, cf.
-Figure S8). However, the models differed considerably in the way in
-which they simulated these climatic niches.
+Appendix S1: Figure S8). However, the models differed considerably in
+the way in which they simulated these climatic niches.
 
@@
 although the exact changes in the share of the recruited basal area were
-different (Figure S8).
+different (Appendix S1: Figure S8).
 
@@
 the link to the publication with more model details. Further information
-can be found on the individual model reports available in the
-supplementary materials.
+can be found on the individual model reports available in the Appendix
+S1: Model reports.
 

### 2023-10-31 manuscript/supplementaryRW.Rmd
changing appendix name

@@
 ---
-title: "Appendix"
+title: "Appendix S1"
 subtitle: "Tree regeneration in models of forest dynamics: a key priority for further research  ECOSPHERE"
@@
 sigDiff$p.adj <- NULL
+sigDiff$p <- NULL
+sigDiff$method <- NULL
 colnames(sigDiff) <- c("model", "y", "Group 1", "Group 2", "p", "p.signif")

### 2023-10-31 manuscript/review/authorsResponse.qmd
small corrections in the response based on Haralds comments

@@
 
-The objective of running the different t-tests were to allow us for discussing differences between mean values per site between models and observations and between 7 and 10 cm diameter thresholds.
-The t-test, as employed in our study, is not intended to serve as definitive proof of our findings and we have deliberately avoid over-interpretation of their outcomes.
-Rather, it serves as a practical tool for managing the extensive number of comparisons in our study.
-As such, we intend to include a statement in the manuscript to clarify this aspect, please see L\[336-337\].
-
-Following the reviewers comment we included a correction for multiple testing although we are aware that all corrections have pros and cons.
-In this case we used the Bonferroni correction to make sure that the actual Type I error probability for the combined H0 is smaller or equal than the nominal test level.
-
-We also look into the alternative approaches suggested by the reviewer.
-We can not do a paired t-test as the for paired samples it is necessary that you have the same number of subjects in both groups (observed and simulated) which is not the case as for example some models did not simulate all the sites.
-We tested the possibility of creating a GLME using as random effect the site and as a explanatory variables all the variables that we were interested in (model complexity, feedback, model type and scale) as suggested.
-However this approach will not give us the observed-model comparison that we are interested in.
-Following the reviewers comments we also did further testing to see if with other approaches we could better explain differences between observed and simulated values.
-We tested a mixed model to account for the randomness induce due to site (plot) differences.
-For example a glme model with the recruitment values as the response variable and models as the explanatory variable with observed data as the reference (with poisson distribution-count data).
-This worked well and the patterns were the same as in the t-test (expect for the model 4C).
-We also considered doing the same approach with the shannon index and mortality differences but in this case we think it is more interesting to understand the mean differences across sites and therefore a t-test is a valid approach.
-We also tested a linear mixed model to assess differences between 7 and 10cm on the Shannon index creating a new parameter where the explanatory variable was a combination of the model and the diameter threshold, then we run a pairwise multiple comparisons to determine which means differ.
-The patterns here were also similar to the ones observed in the t-test except for two landscape level models (TreeMig and LandisII) that were significant in this test contrary to the t-test.
-
-After considering all the options that the reviewer suggested and testing others, we made some adjustments to the t-tests analysis, with the correction for multiple testing and more clean output tables as suggested by the reviewer. In addition we explain the purpose and interpretation of these tests better in line \[336-337\].
+The objective of running the t-tests was to allow us to discuss differences between mean values per site between models, and observations and simulations between the 7 and 10 cm diameter thresholds.
+The t-test, as employed in our study, is not intended to serve as \'proof\' of our findings, and we have carefully avoided an over-interpretation of their outcomes.
+Rather, they serve as a practical tool for handling the extensive number of comparisons in our study.
+To clarify this, we have included a statement in the manuscript, cf.
+L\[336-337\].
+
+Following the reviewer\'s comment, we have included a correction for multiple tests, although we are aware that all corrections have pros and cons.
+We used the Bonferroni correction to make sure that the actual Type I error probability for the combined H0 is smaller than or equal to the nominal test level.
+
+We have also looked into the alternative approaches suggested by the reviewer.
+We cannot do a paired t-test as for paired samples it is necessary to have the same number of subjects in both groups (observed and simulated), which is not the case here, as for example some models did not simulate all the sites.
+Yet, we tested the possibility of creating a GLME using as random effect the site and as explanatory variables all the features that we were interested in (model complexity, feedback, model type and scale), as suggested by the reviewer.
+However, this approach does not provide the comparison of observed vs.
+simulated data that we are interested in.
+Following the reviewer\'s comments, we conducted further tests to see if we could better explain differences between observed and simulated values using other approaches.
+We tested a mixed model to account for the randomness induced by to site (plot) differences.
+For example, a GLME with recruitment values as the response variable and models as the explanatory variable, with observed data as the reference (with a Poisson distribution because we are dealing with count data).
+This worked well, and the patterns were the same as in the t-test (except for the model 4C).
+We also considered pursuing the same approach with the Shannon index and the differences in mortality, but in this case we think it is more relevant to understand the mean differences across sites, and therefore a t-test is more suitable.
+We furthermore tested a linear mixed model to assess differences between the 7 and 10 cm thresholds regarding the Shannon index, by creating a new parameter where the explanatory variable was a combination of the model and the diameter threshold.
+We then ran pairwise multiple comparisons to determine which means are different.
+Also here, the patterns were quite similar to the ones resulting from the t-test except for two landscape-level models (TreeMig and Landis-II) that were significant in this test, contrary to the t-test.
+
+After considering all the options that the reviewer suggested and testing others, we have made some adjustments to the t-test analysis, with the correction for multiple testing and cleaner output tables, as suggested by the reviewer.
+In addition, we have explained the purpose and interpretation of these tests better on lines \[336-337\].
 
@@
 
-In this first assessment we were not aiming at identify what specific subcomponents may hold the key and therefore based on the current approach we could not provide more concrete recommendation for model development which most likely be specific for each of the 15 models.
-Maybe we need more such studies, including calibration as a first step, and then application against a withheld part of the data; or application (and comparison) under climate change scenarios to evaluate the robustness of the signal under future conditions.
+In this first assessment, we were not aiming to identify what specific subcomponents may hold the key.
+Therefore, based on the current approach, we could not provide more concrete recommendations for model development, which would most likely be specific for each of the 15 models.
+Maybe we need more such studies, including calibration as a first step, and then application against a withheld part of the data or application (and comparison) under climate change scenarios to evaluate the robustness of the signal under future conditions.
+We agree with the reviewer that we need more insights into how we can better simulate regeneration processes in each of these models, but this was beyond the scope of this paper.
 
@@
 The reviewer is correct that we did not provide e.g.
-site coordinates (with the exception of iLand that needed this for a model specific soil input, this is described in the manuscript), but we did provide topography, climate and soil site specific variables.
+site coordinates (with the exception of iLand and aDGVM2 that needed this for a model specific soil input, this is described in the manuscript), but we did provide topography, climate and soil site specific variables.
 
@@
 Following the reviewers comment we have expanded the caption of table 1.
+We have decided not to include further variables on the model assumptions as we believe the current information provides a good general overview of the models approach and further details are included in the appendix.
 

### 2023-10-31 manuscript/review/editorsLetter.qmd
last changes before sent for internal round
editors latter notes

@@
 
-I want to express my gratitude for taking the time to review our manuscript. Your insightful thoughts on our findings were very helpful.
+I would like to express my sincere gratitude for taking the time to review our manuscript. Your insightful feedback on our findings was immensely helpful.
 
-We believe that the changes we made have significantly improved the manuscript.
+We have taken all the reviewer\'s comments and suggestions into consideration and have made significant changes to the manuscript.
+Please find enclosed a detailed report outlining the actions we have taken in response to each of the reviewer's comments.
 
-Once again, I want to thank you for the time and effort you dedicated to reviewing our work. We look forward to hearing from you again.
+Thank you again for your time and effort. We look forward to hearing from you soon.
 
@@
 
-Olalla Díaz-Yáñez on behalf of all the co-authors
-
-##
-
-## Reviewer 1
-
-Dear Reviewer 1,
-
-Thank you for taking the time to review our manuscript.
-We appreciate your thoughtful comments and have carefully considered each of them.
-We have made changes to the manuscript based on your feedback and believe that these changes have improved the overall quality of the work.
-
-Please find below our responses to each of your comments and the concrete actions we have taken in response.
-
-Thank you again for your valuable feedback.
+Olalla Díaz-Yáñez, on behalf of all the co-authors

### 2023-10-31 manuscript/review/authorsResponse.qmd
comment 3

@@
 The objective of running the t-tests was to allow us to discuss differences between mean values per site between models, and observations and simulations between the 7 and 10 cm diameter thresholds.
-The t-test, as employed in our study, is not intended to serve as \'proof\' of our findings, and we have carefully avoided an over-interpretation of their outcomes.
+The t-test, as employed in our study, is not intended to serve as 'proof' of our findings, and we have carefully avoided an over-interpretation of their outcomes.
 Rather, they serve as a practical tool for handling the extensive number of comparisons in our study.
@@
 
-Following the reviewer\'s comment, we have included a correction for multiple tests, although we are aware that all corrections have pros and cons.
+Following the reviewer's comment, we have included a correction for multiple tests, although we are aware that all corrections have pros and cons.
 We used the Bonferroni correction to make sure that the actual Type I error probability for the combined H0 is smaller than or equal to the nominal test level.
@@
 simulated data that we are interested in.
-Following the reviewer\'s comments, we conducted further tests to see if we could better explain differences between observed and simulated values using other approaches.
+Following the reviewer's comments, we conducted further tests to see if we could better explain differences between observed and simulated values using other approaches.
 We tested a mixed model to account for the randomness induced by to site (plot) differences.

### 2023-10-31 manuscript/manuscriptRW.Rmd
authors afiliation

@@
 - Josef	Brůna$^5$,
-- Rico Fischer$^3$,
-- Samuel M. Fischer$^3$,
+- Rico Fischer$^6$,
+- Samuel M. Fischer$^6$,
 - Jessica Hetzer$^2$,
 - Thomas Hickler$^2$,
-- Christian	Hochauer$^6$,
-- Manfred J. Lexer$^6$,
-- Heike	Lischke$^7$,
-- Paola	Mairota$^8$,
-- Ján Merganič$^{9}$,
-- Katarina Merganičová$^{10,11}$,
-- Tobias Mette$^{12}$,
-- Marco Mina$^{13}$,
-- Xavier Morin$^{14}$
-- Mats Nieberg,$^{15}$,
+- Christian	Hochauer$^7$,
+- Manfred J. Lexer$^7$,
+- Heike	Lischke$^8$,
+- Paola	Mairota$^9$,
+- Ján Merganič$^{10}$,
+- Katarina Merganičová$^{11,12}$,
+- Tobias Mette$^{13}$,
+- Marco Mina$^{14}$,
+- Xavier Morin$^{15}$
+- Mats Nieberg,$^{16}$,
 - Werner Rammer$^4$,
-- Christopher P.O. Reyer$^{15}$,
+- Christopher P.O. Reyer$^{16}$,
 - Simon	Scheiter$^2$,
@@
       $^2$Senckenberg Biodiversity and Climate Research Centre (SBiK-F), 60325 Frankfurt am Main, Germany. \newline
-      $^3$Department of Ecological Modelling, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline
+      $^3$Department of Computational Hydrosystem, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline
       $^4$Ecosystem Dynamics and Forest Management Group, TUM School of Life Sciences, Technical University of Munich, 85354 Freising, Germany. \newline
        $^5$Department of Geoecology, Institute of Botany of the Czech Academy of Sciences, Průhonice, Czech Republic. \newline
-       $^6$Institute of Silviculture, Department of Forest and Soil Sciences, University of Natural Resources and Life Sciences, Peter Jordan Strasse 82, A-1190 Vienna, Austria. \newline
-       $^7$Dynamic Macroecology, Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland. \newline
-       $^8$Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, 14473 Potsdam, Germany. \newline
+       $^6 Department of Ecological Modelling, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline
+       $^7$Institute of Silviculture, Department of Forest and Soil Sciences, University of Natural Resources and Life Sciences, Peter Jordan Strasse 82, A-1190 Vienna, Austria. \newline
+       $^8$Dynamic Macroecology, Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland. \newline
        $^9$Department of Soil Sciences, of Plants and Food, University of Bari Aldo Moro, Italy. \newline
@@
        $^{14}$Institute for Alpine Environment, Eurac Research, Bolzano/Bozen, Italy. \newline 
-       $^{15}$CEFE, CNRS Campus du CNRS 1919, route de Mende 34293 Montpellier cedex, France. \newline 
-       $^{16}$Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland.  \newline
+       $^{15}$CEFE, CNRS Campus du CNRS 1919, route de Mende 34293 Montpellier cedex, France.
+       $^{16}$Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, 14473 Potsdam, Germany. \newline
+       $^{17}$Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland.  \newline
        $^*$Corresponding author (olalla.diaz@usys.ethz.ch)

### 2023-10-31 manuscript/review/editorsLetter.qmd
update gitignore

@@
 
-I would like to express my sincere gratitude for taking the time to review our manuscript. Your insightful feedback on our findings was immensely helpful.
+Thank you for taking the time to review our manuscript and for your valuable feedback.
 
-We have taken all the reviewer\'s comments and suggestions into consideration and have made significant changes to the manuscript.
-Please find enclosed a detailed report outlining the actions we have taken in response to each of the reviewer's comments.
+In this revised version, we have taken all the reviewer\'s comments and suggestions into consideration and have made significant changes to the manuscript accordingly.
+We have submitted a detailed report outlining the actions we have taken in response to each of the reviewer's comments.
+We believe that the changes we have made have significantly improved the quality of our manuscript.
 
-Thank you again for your time and effort. We look forward to hearing from you soon.
+Thank you again for your time. We look forward to hearing from you soon.
 

### 2023-10-31 code/figures.R

@@ -1456,7 +1456,7 @@
     
     ###### DO NOT RUN #########
 
-    # #Add location
+    # # #Add location
     # location <- data.table::fread("data/coords_blurred_dt4326.csv")
     # 
     # domSpp7 <- merge(standSpp, recruitSpp7, by = "site")
@@ -1509,5 +1509,5 @@
     # ggplot2::ggsave("figures/Figure1.jpg",
     #                 plot =   plotsMap,
     #                 width = 21, height = 12, scale = 0.9,
-    #                 dpi = 300, units = "cm", device = 'png')
-    # 
+    #                 dpi = 300, units = "cm", device = 'jpg')
+

### 2024-02-12 manuscript/manuscriptRW.Rmd
update gitignore
change Paola's affiliation

@@
 - Simon	Scheiter$^2$,
-- Daniel Scherrer$^{16}$,
+- Daniel Scherrer$^{17}$,
 - Harald Bugmann$^1$
@@
        $^5$Department of Geoecology, Institute of Botany of the Czech Academy of Sciences, Průhonice, Czech Republic. \newline
-       $^6 Department of Ecological Modelling, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline
+       $^6$Department of Ecological Modelling, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline
        $^7$Institute of Silviculture, Department of Forest and Soil Sciences, University of Natural Resources and Life Sciences, Peter Jordan Strasse 82, A-1190 Vienna, Austria. \newline
        $^8$Dynamic Macroecology, Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland. \newline
-       $^9$Department of Soil Sciences, of Plants and Food, University of Bari Aldo Moro, Italy. \newline
+       $^9$Dipartimento di Scienze del suolo, della pianta e degli alimenti Università degli Studi di Bari Aldo Moro via Amendola 165/A, 70126 Bari, Italy \newline
        $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen, Slovakia. \newline
@@
 
-Tree regeneration is a key process in forest dynamics, particularly in
-the context of forest resilience and climate change. Models are pivotal
-for assessing long-term forest dynamics, and they have been in use for
-more than 50 years. However, there is a need to evaluate their capacity
-to accurately represent tree regeneration. We assess how well current
-models capture the overall abundance, species composition, and mortality
-of tree regeneration.
-
-Using 15 models built to capture long-term forest dynamics at the stand,
-landscape, and global levels, we simulate tree regeneration at 200 sites
-representing large environmental gradients across Central Europe. The
-results are evaluated against extensive data from unmanaged forests.
-
-Most of the models overestimate recruitment levels, which is compensated
-only in some models by high simulated mortality rates in the early
-stages of individual tree dynamics. Simulated species diversity of
-recruitment generally matches observed ranges. Models simulating higher
-stand-level species diversity do not feature higher species diversity in
-the recruitment layer. The effect of light availability on recruitment
-levels is captured better than the effects of temperature and soil
-moisture, but patterns are not consistent across models. Increasing
-complexity in the tree regeneration modules is not related to higher
-accuracy of simulated tree recruitment. Furthermore, individual model
-design is more important than scale (stand, landscape, global) and
-approach (empirical, process-based) for accurately capturing tree
-regeneration.
-
-Despite the mismatches between simulation results and data, it is
-remarkable that most models capture the essential features of the highly
-complex process of tree regeneration, while not having been
-parameterized with such data. We conclude that much can be gained by
-evaluating and refining the modeling of tree regeneration processes.
-This has the potential to render long-term projections of forest
-dynamics under changing environmental conditions much more robust.
+Tree regeneration is a key process in forest dynamics, particularly in the context of forest resilience and climate change.
+Models are pivotal for assessing long-term forest dynamics, and they have been in use for more than 50 years.
+However, there is a need to evaluate their capacity to accurately represent tree regeneration.
+We assess how well current models capture the overall abundance, species composition, and mortality of tree regeneration.
+
+Using 15 models built to capture long-term forest dynamics at the stand, landscape, and global levels, we simulate tree regeneration at 200 sites representing large environmental gradients across Central Europe.
+The results are evaluated against extensive data from unmanaged forests.
+
+Most of the models overestimate recruitment levels, which is compensated only in some models by high simulated mortality rates in the early stages of individual tree dynamics.
+Simulated species diversity of recruitment generally matches observed ranges.
+Models simulating higher stand-level species diversity do not feature higher species diversity in the recruitment layer.
+The effect of light availability on recruitment levels is captured better than the effects of temperature and soil moisture, but patterns are not consistent across models.
+Increasing complexity in the tree regeneration modules is not related to higher accuracy of simulated tree recruitment.
+Furthermore, individual model design is more important than scale (stand, landscape, global) and approach (empirical, process-based) for accurately capturing tree regeneration.
+
+Despite the mismatches between simulation results and data, it is remarkable that most models capture the essential features of the highly complex process of tree regeneration, while not having been parameterized with such data.
+We conclude that much can be gained by evaluating and refining the modeling of tree regeneration processes.
+This has the potential to render long-term projections of forest dynamics under changing environmental conditions much more robust.
 
@@
 
-Forests provide a wide range of products and services of vital
-importance to humankind [@foresteurope2020]. Under the pressure of
-climate change, increasing disturbance impacts and changing societal
-demands on forest ecosystem services, it is becoming ever more important
-to understand how forest structure, composition and function will
-change, and to evaluate forest capacity to adapt to or be resilient
-after disturbance [@lindner2010; @seidl2022]. A wide range of models of
-forest dynamics have been developed over the past decades considering
-the impacts of climate [@bugmann2022]. From these studies, it is evident
-that we have a reasonably good understanding how to model tree growth
-[@vanclay1997; @bugmann1996], and substantial efforts have been
-dedicated to improving the representation of tree mortality
-[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is much
-less studied and is often represented rather coarsely in models
-[@walck2011; @leishman1992; @price2001]. This presents an important
-research gap, particularly in the context of climate-induced forest
-disturbances and forest resilience.
-
-Tree regeneration arises from multiple processes including pollination,
-fruit maturation, seed production, dispersal, germination, juvenile
-growth and survival [@price2001; @vacchiano2018]. All these processes
-are difficult to assess, and some of them are scarcely understood and
-thus appear highly stochastic [@bogdziewicz2021]. Identifying the
-appropriate level of complexity for the mathematical formulation of the
-key factors that are leading to successful tree regeneration is
-challenging. Currently, tree regeneration processes in dynamic forest
-models are handled in a multitude of ways [@konig_tree_2022;
-@bugmann2022]: from 1) entirely ignoring them [as done in classical
-forest growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2)
-the use of a few simple environmental filters, as done in most forest
-"gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
-@hickler2012; @smith2001], to 3) complex approaches that incorporate
-local feedback from the canopy, multiple ecological processes and often
-also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
-statistical parameterizations, which however are not easy to extrapolate
-in space and time [e.g., @ribbens1994].
-
-Overall, models are needed to (1) synthesize existing empirical data and
-explore their relationships, (2) assess future tree regeneration e.g. in
-the context of global change scenarios, and (3) identify the most
-important processes that are shaping vegetation patterns. Given the
-strategies that are used in models of forest dynamics to represent tree
-regeneration, their behavior often is prone to problems, such as very
-high levels of tree regeneration that necessitate excess mortality at
-early stages of tree life to simulate correct stand structure and
-composition [@kroiss2015]. Also, inadequately high species diversity in
-tree regeneration may be simulated, which is characteristic of
-"classical" forest gap models [@gutiérrez2016], at least as long as the
-simulation set-up comprises a multitude of species. Some models use
-calibration against local canopy-level data to constrain simulated tree
-regeneration, which is likely to hamper the general applicability of
-these models e.g. under scenarios of climate change. Furthermore,
-correctly capturing the species composition of tree regeneration as a
-function of the presence of seed trees in the canopy is often a
-particular challenge, potentially leading to unrealistic successional
-drift in the model, which must be corrected by factors that are hard to
-parameterize [e.g., @lischke2006]. A related issue is the excessive
-reduction of species diversity due to positive feedback effects, such
-that eventually just single-species stands remain [@meier2011]. This is
-sometimes corrected by the incorporation of a low level of seed influx
-of all species at all times [@schumacher2006] or by restricting the
-number of seeds per species in the seedbank [@lischke2006]. However,
-simulated species composition is usually exceedingly sensitive to
-assumptions about seed availability, whereas the parameters of such
-functions are poorly constrained by field data. Lastly, there is often a
-problem with insufficient observational constraints on parameter values
-for models that start from very small tree sizes (e.g., 10 cm in height,
-or even from seed) and track tree development in a process-oriented
-manner by considering a multitude of ecological influences, rather than
-emphasizing tree recruitment into a larger size class [e.g.,
-@wehrli2006].
-
-Modeling tree regeneration processes is challenging, and even in
-empirical ecology, it has not received much attention
-[@hanbury-brown2022], although many sub-processes have been studied in
-detail [@miina2006]. Yet, little data are available that cover all the
-processes within one species along environmental gradients, let alone
-for a vast suite of species. Data on forest regeneration are often
-fragmented, which constitutes a major problem for model building
-[@clark1999a]. For example, monitoring on permanent plots (such as
-National Forest Inventories) often measures the ingrowth of new trees
-into a specific size class [@zell2019], however with a design that
-captures the rate after the stand initiation phase [@hallsby2015]. Data
-from permanent plots, although available across regions, are also highly
-heterogeneous, rendering their use difficult in a modeling context
-[@käber2021]. Lastly, targeted studies to measure tree regeneration on
-experimental sites are often limited in spatial extent, species studied,
-or the sub-set of processes that are investigated [@berdanier2016;
-@collet2006].
-
-Thus, a focus on the modeling of tree regeneration processes is sorely
-needed and overdue [@price2001; @walck2011] if we are to make reliable
-projections of future forest dynamics, i.e., when the models need to be
-operated in extrapolation mode, as well as from a fundamental ecological
-point of view for increasing systems understanding. In the present
-study, for the first time, a large number of forest models commonly used
-to assess forest dynamics under climate change are evaluated against a
-continental-scale, multi-species harmonized dataset on tree recruitment
-[@käber2023]. By tree recruitment, we refer to the passing of trees
-across a specific diameter threshold ("ingrowth"). We included models
-that are based on a range of "philosophies", from models operating at
-the stand to the global scale as well as the range of models from
-empirically derived to "process-based" [e.g., @bugmann1996;
-@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
-tree regeneration patterns in nature and the large number of factors
-driving and constraining this process --- including some that are not
-incorporated explicitly in most models, such as deer browsing --- we do
-not aim for a detailed statistical evaluation of each model. Instead, we
-aim to evaluate the general recruitment patterns and magnitudes
-simulated by the models, and benchmark the simulated regeneration niche
-of multiple species against empirical data along a wide environmental
-gradient of temperature, moisture, and light availability [@grubb1977].
-
-More specifically, we aim to answer the following questions: (1) Are
-models of forest dynamics capturing accurately tree recruitment levels,
-initial tree species diversity and mortality in the recruitment? (2) Do
-model traits explain differences in model performance? (3) How well do
-the models capture total recruitment and the regeneration niches of
-individual species across environmental gradients of light availability,
-temperature, and soil moisture?
-
-We evaluate the performance of the models by analyzing simulated data
-alongside observed data. Specifically, we examine tree species
-diversity, mortality rates in tree recruitment, and recruitment levels.
-Then, we link model traits (e.g., complexity, scale) to performance, and
-we assess whether modeled total recruitment or individual species
-regeneration niches (i.e., proportional recruitment along an
-environmental gradient) align with observational data.
+Forests provide a wide range of products and services of vital importance to humankind [@foresteurope2020].
+Under the pressure of climate change, increasing disturbance impacts and changing societal demands on forest ecosystem services, it is becoming ever more important to understand how forest structure, composition and function will change, and to evaluate forest capacity to adapt to or be resilient after disturbance [@lindner2010; @seidl2022].
+A wide range of models of forest dynamics have been developed over the past decades considering the impacts of climate [@bugmann2022].
+From these studies, it is evident that we have a reasonably good understanding how to model tree growth [@vanclay1997; @bugmann1996], and substantial efforts have been dedicated to improving the representation of tree mortality [@bugmann2019; @cailleret2017].
+In contrast, tree regeneration is much less studied and is often represented rather coarsely in models [@walck2011; @leishman1992; @price2001].
+This presents an important research gap, particularly in the context of climate-induced forest disturbances and forest resilience.
+
+Tree regeneration arises from multiple processes including pollination, fruit maturation, seed production, dispersal, germination, juvenile growth and survival [@price2001; @vacchiano2018].
+All these processes are difficult to assess, and some of them are scarcely understood and thus appear highly stochastic [@bogdziewicz2021].
+Identifying the appropriate level of complexity for the mathematical formulation of the key factors that are leading to successful tree regeneration is challenging.
+Currently, tree regeneration processes in dynamic forest models are handled in a multitude of ways [@konig_tree_2022; @bugmann2022]: from 1) entirely ignoring them [as done in classical forest growth models, e.g., Pretzsch et al.
+-@pretzsch2002], across 2) the use of a few simple environmental filters, as done in most forest "gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g., @hickler2012; @smith2001], to 3) complex approaches that incorporate local feedback from the canopy, multiple ecological processes and often also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based statistical parameterizations, which however are not easy to extrapolate in space and time [e.g., @ribbens1994].
+
+Overall, models are needed to (1) synthesize existing empirical data and explore their relationships, (2) assess future tree regeneration e.g.
+in the context of global change scenarios, and (3) identify the most important processes that are shaping vegetation patterns.
+Given the strategies that are used in models of forest dynamics to represent tree regeneration, their behavior often is prone to problems, such as very high levels of tree regeneration that necessitate excess mortality at early stages of tree life to simulate correct stand structure and composition [@kroiss2015].
+Also, inadequately high species diversity in tree regeneration may be simulated, which is characteristic of "classical" forest gap models [@gutiérrez2016], at least as long as the simulation set-up comprises a multitude of species.
+Some models use calibration against local canopy-level data to constrain simulated tree regeneration, which is likely to hamper the general applicability of these models e.g.
+under scenarios of climate change.
+Furthermore, correctly capturing the species composition of tree regeneration as a function of the presence of seed trees in the canopy is often a particular challenge, potentially leading to unrealistic successional drift in the model, which must be corrected by factors that are hard to parameterize [e.g., @lischke2006].
+A related issue is the excessive reduction of species diversity due to positive feedback effects, such that eventually just single-species stands remain [@meier2011].
+This is sometimes corrected by the incorporation of a low level of seed influx of all species at all times [@schumacher2006] or by restricting the number of seeds per species in the seedbank [@lischke2006].
+However, simulated species composition is usually exceedingly sensitive to assumptions about seed availability, whereas the parameters of such functions are poorly constrained by field data.
+Lastly, there is often a problem with insufficient observational constraints on parameter values for models that start from very small tree sizes (e.g., 10 cm in height, or even from seed) and track tree development in a process-oriented manner by considering a multitude of ecological influences, rather than emphasizing tree recruitment into a larger size class [e.g., @wehrli2006].
+
+Modeling tree regeneration processes is challenging, and even in empirical ecology, it has not received much attention [@hanbury-brown2022], although many sub-processes have been studied in detail [@miina2006].
+Yet, little data are available that cover all the processes within one species along environmental gradients, let alone for a vast suite of species.
+Data on forest regeneration are often fragmented, which constitutes a major problem for model building [@clark1999a].
+For example, monitoring on permanent plots (such as National Forest Inventories) often measures the ingrowth of new trees into a specific size class [@zell2019], however with a design that captures the rate after the stand initiation phase [@hallsby2015].
+Data from permanent plots, although available across regions, are also highly heterogeneous, rendering their use difficult in a modeling context [@käber2021].
+Lastly, targeted studies to measure tree regeneration on experimental sites are often limited in spatial extent, species studied, or the sub-set of processes that are investigated [@berdanier2016; @collet2006].
+
+Thus, a focus on the modeling of tree regeneration processes is sorely needed and overdue [@price2001; @walck2011] if we are to make reliable projections of future forest dynamics, i.e., when the models need to be operated in extrapolation mode, as well as from a fundamental ecological point of view for increasing systems understanding.
+In the present study, for the first time, a large number of forest models commonly used to assess forest dynamics under climate change are evaluated against a continental-scale, multi-species harmonized dataset on tree recruitment [@käber2023].
+By tree recruitment, we refer to the passing of trees across a specific diameter threshold ("ingrowth").
+We included models that are based on a range of "philosophies", from models operating at the stand to the global scale as well as the range of models from empirically derived to "process-based" [e.g., @bugmann1996; @fabrika2012; @lexer2001; @reyer2014].
+Due to the large variability in tree regeneration patterns in nature and the large number of factors driving and constraining this process --- including some that are not incorporated explicitly in most models, such as deer browsing --- we do not aim for a detailed statistical evaluation of each model.
+Instead, we aim to evaluate the general recruitment patterns and magnitudes simulated by the models, and benchmark the simulated regeneration niche of multiple species against empirical data along a wide environmental gradient of temperature, moisture, and light availability [@grubb1977].
+
+More specifically, we aim to answer the following questions: (1) Are models of forest dynamics capturing accurately tree recruitment levels, initial tree species diversity and mortality in the recruitment?
+(2) Do model traits explain differences in model performance?
+(3) How well do the models capture total recruitment and the regeneration niches of individual species across environmental gradients of light availability, temperature, and soil moisture?
+
+We evaluate the performance of the models by analyzing simulated data alongside observed data.
+Specifically, we examine tree species diversity, mortality rates in tree recruitment, and recruitment levels.
+Then, we link model traits (e.g., complexity, scale) to performance, and we assess whether modeled total recruitment or individual species regeneration niches (i.e., proportional recruitment along an environmental gradient) align with observational data.
 
@@
 
-Fifteen models of forest dynamics (including two models featuring model
-variants) developed for the stand, landscape or global scale were used
-to simulate forest dynamics (Table \@ref(tab:tableModels)). The
-approaches used for model construction and their origin differ strongly,
-with most of the models featuring a largely "process-based" approach,
-whereas two models are based on formulations derived from the
-statistical analysis of inventory data (SIBYLA and xComp). Some of the
-models largely rely on the original approach underlying forest gap
-models (e.g., ForClim 1), whereas some are based on plant
-ecophysiological processes (e.g., FORMIND and iLand). The different
-approaches underlying the models have strongly influenced the
-formulation of tree regeneration processes.
-
-The models can be differentiated into 'regeneration' and 'recruitment'
-approaches [@vanclay1994; @konig_tree_2022]. *Regeneration models*
-include processes such as flowering and pollination, seed production,
-seed dispersal, germination, and seedling growth, which ultimately lead
-to the simulated number of established trees. *Recruitment models*, in
-contrast, introduce a number of new trees with certain characteristics
-such as biomass or diameter, without explicitly considering earlier
-development processes. We can further distinguish models that feature a
-feedback from stand properties to simulated recruitment, i.e., where the
-level and species composition of recruitment is influenced by the adult
-tree community via the production of seeds (or seedlings/saplings), from
-models that do not contain such feedback.
-
-The starting point for tree regeneration in the models differs as well,
-ranging from seedbank, seed or seedling to sapling (i.e., trees much
-larger than 10 cm height, often ca. 2-3 m tall). Models that start from
-seed need to include a larger number of ecological processes such as
-germination and survival of young seedlings, whereas models that start
-from saplings have to aggregate via parameterizations several ecological
-processes that are not treated explicitly. This latter approach reduces
-model complexity but comes at the cost of blurred process
-representation.
-
-In the models used here, the overall complexity in the regeneration
-modules varies considerably. Following @bugmann2022, we can classify
-seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and
-TreeMig) as having rather high complexity in their regeneration modules
-(mean regeneration formulation complexity across all processes \>0,
-Table \@ref(tab:tableModels)), whereas the other eight models feture
-relatively low complexity. Two models, ForClim and ForCEEPS, were used
-here with two alternative variants of regeneration. ForClim variant 1
-[@bugmann1996] is based on a regeneration module that adheres closely to
-the concept introduced by @botkin1972 which is based on environmental
-filters considering species-specific thresholds of light availability
-and climatic variables. In contrast, ForClim variant 11 is adopting a
-slightly more complex approach where individual species properties and
-their relation to the environment are represented gradually along with
-the relative suitability of different species (i.e., competition among
-tree species) [@huber2020]. These two model variants allow us to
-evaluate a more process-based and complex module (variant 11) against a
-simple module (variant 1), while the rest of the model structure is
-identical. Similarly, the two ForCEEPS variants allow us to isolate the
-importance of the canopy feedback (i.e., simulated actual composition
-and relative abundance of species in the plot) via seed trees for the
-quantity and quality (e.g., diversity and composition) of simulated
-regeneration, as one variant includes this feedback whereas the other
-does not.
+Fifteen models of forest dynamics (including two models featuring model variants) developed for the stand, landscape or global scale were used to simulate forest dynamics (Table \@ref(tab:tableModels)).
+The approaches used for model construction and their origin differ strongly, with most of the models featuring a largely "process-based" approach, whereas two models are based on formulations derived from the statistical analysis of inventory data (SIBYLA and xComp).
+Some of the models largely rely on the original approach underlying forest gap models (e.g., ForClim 1), whereas some are based on plant ecophysiological processes (e.g., FORMIND and iLand).
+The different approaches underlying the models have strongly influenced the formulation of tree regeneration processes.
+
+The models can be differentiated into 'regeneration' and 'recruitment' approaches [@vanclay1994; @konig_tree_2022].
+*Regeneration models* include processes such as flowering and pollination, seed production, seed dispersal, germination, and seedling growth, which ultimately lead to the simulated number of established trees.
+*Recruitment models*, in contrast, introduce a number of new trees with certain characteristics such as biomass or diameter, without explicitly considering earlier development processes.
+We can further distinguish models that feature a feedback from stand properties to simulated recruitment, i.e., where the level and species composition of recruitment is influenced by the adult tree community via the production of seeds (or seedlings/saplings), from models that do not contain such feedback.
+
+The starting point for tree regeneration in the models differs as well, ranging from seedbank, seed or seedling to sapling (i.e., trees much larger than 10 cm height, often ca.
+2-3 m tall).
+Models that start from seed need to include a larger number of ecological processes such as germination and survival of young seedlings, whereas models that start from saplings have to aggregate via parameterizations several ecological processes that are not treated explicitly.
+This latter approach reduces model complexity but comes at the cost of blurred process representation.
+
+In the models used here, the overall complexity in the regeneration modules varies considerably.
+Following @bugmann2022, we can classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and TreeMig) as having rather high complexity in their regeneration modules (mean regeneration formulation complexity across all processes \>0, Table \@ref(tab:tableModels)), whereas the other eight models feture relatively low complexity.
+Two models, ForClim and ForCEEPS, were used here with two alternative variants of regeneration.
+ForClim variant 1 [@bugmann1996] is based on a regeneration module that adheres closely to the concept introduced by @botkin1972 which is based on environmental filters considering species-specific thresholds of light availability and climatic variables.
+In contrast, ForClim variant 11 is adopting a slightly more complex approach where individual species properties and their relation to the environment are represented gradually along with the relative suitability of different species (i.e., competition among tree species) [@huber2020].
+These two model variants allow us to evaluate a more process-based and complex module (variant 11) against a simple module (variant 1), while the rest of the model structure is identical.
+Similarly, the two ForCEEPS variants allow us to isolate the importance of the canopy feedback (i.e., simulated actual composition and relative abundance of species in the plot) via seed trees for the quantity and quality (e.g., diversity and composition) of simulated regeneration, as one variant includes this feedback whereas the other does not.
 
@@
 
-Recruitment data covering a wide range of environmental conditions are
-hard to obtain, and this is one of the reasons why most models of forest
-dynamics have never been confronted with a dataset covering such
-gradients over a large number of sites, to evaluate how well
-regeneration is captured. The observations used here are derived from a
-novel network of sites in forest reserves that represent the range of
-environmental gradients in temperature and precipitation in Central
-Europe as compiled in the EuFoRIa network [@euforia2019; @käber2023]
-(Figure 1). The environmental coverage and the variety of forests
-represented by this data is unprecedented within Europe. These forests
-have been unmanaged for at least 10 years prior to their designation as
-reserves, and most of them provide time series of natural forest
-dynamics over multiple decades and up to ca. 85 years. The census
-periods range from 3 to 37 years, with an average of 14 years. The data
-provide information at the tree level, thus allowing for the sequential
-comparison of processes such as individual-tree recruitment and death
-between the measurements.
-
-We selected 200 sites from this network as the benchmarking dataset for
-the simulation, so as to be representative of the environmental
-variation contained in the data. This was achieved by applying k-means
-clustering to define 200 clusters of plots from the original set of 869
-plots along the environmental dimensions of temperature, climatic water
-balance, soil quality, slope and aspect. Each of these sites featured at
-least two consecutive measurements. Recruitment size thresholds ranged
-from stem diameters at breast height (dbh) from 0 to 10 cm. For our
-study, we defined two datasets, where one included 165 sites with a
-diameter threshold of 7 cm or lower, and the other included another 35
-sites with diameter thresholds between 7 and 10 cm. Therefore we had in
-total 200 sites with observations of newly recruited trees above 10 cm
-and a subset of 165 sites with recruitment data above 7 cm. In the
-results section, any observations from the data at the 7 cm threshold
-originated from 165 sites.
-
-Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
-processed and aggregated following @käber2023 to provide recruitment
-rates per ha and per decade. The observations used in this study
-featured 30,900 newly established trees. Recruitment rates per site,
-sample and decade ranged from zero to 1,246 trees, with a mean of 56
-trees. Adult species composition was also available for each plot. For
-more details on this dataset and the detailed recruitment information,
-cf. @käber2023.
+Recruitment data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset covering such gradients over a large number of sites, to evaluate how well regeneration is captured.
+The observations used here are derived from a novel network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the EuFoRIa network [@euforia2019; @käber2023] (Figure 1).
+The environmental coverage and the variety of forests represented by this data is unprecedented within Europe.
+These forests have been unmanaged for at least 10 years prior to their designation as reserves, and most of them provide time series of natural forest dynamics over multiple decades and up to ca.
+85 years.
+The census periods range from 3 to 37 years, with an average of 14 years.
+The data provide information at the tree level, thus allowing for the sequential comparison of processes such as individual-tree recruitment and death between the measurements.
+
+We selected 200 sites from this network as the benchmarking dataset for the simulation, so as to be representative of the environmental variation contained in the data.
+This was achieved by applying k-means clustering to define 200 clusters of plots from the original set of 869 plots along the environmental dimensions of temperature, climatic water balance, soil quality, slope and aspect.
+Each of these sites featured at least two consecutive measurements.
+Recruitment size thresholds ranged from stem diameters at breast height (dbh) from 0 to 10 cm.
+For our study, we defined two datasets, where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds between 7 and 10 cm.
+Therefore we had in total 200 sites with observations of newly recruited trees above 10 cm and a subset of 165 sites with recruitment data above 7 cm.
+In the results section, any observations from the data at the 7 cm threshold originated from 165 sites.
+
+Plot size ranged from 0.02 ha to 5.52 ha, and the data were further processed and aggregated following @käber2023 to provide recruitment rates per ha and per decade.
+The observations used in this study featured 30,900 newly established trees.
+Recruitment rates per site, sample and decade ranged from zero to 1,246 trees, with a mean of 56 trees.
+Adult species composition was also available for each plot.
+For more details on this dataset and the detailed recruitment information, cf.
+@käber2023.
 
@@
 
-The overarching goal of the simulation experiments was to assess tree
-recruitment as it arises from empirical data against its representation
-in a wide range of models of forest dynamics. We define tree recruitment
-as the passing of a breast height diameter threshold of 7 and 10 cm
-(synonym: ingrowth). To this end, each modeling group was provided with
-a detailed protocol [@díaz-yáñez2022] with instructions how to perform
-the simulations, input variables on climate and soil conditions, and the
-list of expected output variables. Neither were further site information
-(except for the data specified below) nor any data on tree recruitment
-or forest stand features provided prior to the simulation.
-
-The input variables were collected from different data sources and
-aggregated to be suitable for the needs of the different models. Time
-series of climatic variables from 1981 to 2018 were provided in hourly
-[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
-resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
-required for some models, such as relative humidity or vapor pressure
-deficit, were calculated from these variables. The final instructions
-for using the climate data were slightly different in each model, based
-on the approach that best suited the model (Table
-\@ref(tab:tableModels)). Soil quality data were provided as continuous
-values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
-The protocol also provided the elevation, slope and aspect for each of
-the 200 sites [@asterscienceteam2019], but no other spatial information
-such as coordinates, with the exception of iLand and aDGVM2, which
-required blurred coordinates to derive highly detailed soil data.
-
-The simulations were run in the absence of natural disturbances. Only
-the model LPJ-GUESS had to include a "background" disturbance
-probability to increase chances of shade-intolerant species to
-establish. The simulations were set up to sample species-specific
-recruitment rates per decade and per ha in the equilibrium state of the
-model, typically entailing a "spin-up" run (as we did not provide any
-forest data). The modeling teams decided on the simulated area and how
-they derived these samples. The exact length of the simulation was also
-decided by the modeling teams (Table \@ref(tab:tableModels)). Further
-details on how each modeling team prepared the simulations and the
-outputs are available in Appendix S1: Model reports.
-
-The simulations were run in the absence of management to a simulated
-equilibrium ("Potential Natural Vegetation") with the current climate,
-as the ultimate goal was to evaluate tree recruitment under comparable
-and near-equilibrium conditions. This entails the assumption that (1)
-the observations from the forest reserves reflect no traces of forest
-management, and (2) there is an equilibrium between forest dynamics and
-climate. While the former might be starting to be visible in many of the
-EuFoRIA reserves, the latter may be more debatable. However, in the
-absence of detailed data on the history of each plot in the EuFoRIa
-network, some broad assumptions had to be made. Both the width of the
-regeneration niche (i.e., in environmental space) as well as the
-intensity of recruitment (i.e., the number of recruited trees per area
-and per unit of time) were of interest.
-
-The simulations were run for mixed-species forests (not multiple
-single-species simulations) using mixtures of eleven species or genera
-for which regeneration data of sufficient quality were available from
-EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
-betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
-Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L.,* and *Pinus
-sylvestris L.* The same set of species was used at all 200 sites. Two
-models included their standard set of species for the simulations, which
-is much larger (i.e., ForClim 1, ForClim 11, TreeMig), and the species
-simulated beyond the eleven species were lumped in a category "others".
-In three models, fewer than these eleven species were simulated (4C,
-xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). In 4C only *Fagus
-sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp. and *Pinus
-sylvestris* are parameterised. xComp simulations did not consider
-*Fraxinus excelsior* due to a deprecated species parameterisation. In
-LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not
-parameterized and therefore these species could not be included in the
-simulation. Finally, the model aDGVM2 does not simulate individual
-species. Rather, community assembly processes and trait filtering
-generate plant communities that are adjusted to the local biotic and
-abiotic conditions, and the simulated plants can be classified into
-ecological strategies based on their trait values in a post-processing
-step.
-
-Each of the models reported the number of new trees crossing the two
-diameter thresholds by sampling 200 times in a 10-year interval for each
-species and per ha for each of the 200 sites. Multiple samples per site
-were used to better understand the simulated variation within each site.
-This was done using different strategies, depending on the model,
-including (1) sampling simulated data from the same 1 ha plot in the
-equilibrium over time, (2) sampling several 1 ha plots from the
-simulated forest at one specific point in time (in the equilibrium) or
-(3) a combination of (1) and (2). This resulted in 880,000 observations
-per model that simulated the 11 species included in the protocol (200
-sites, 200 samples per site, 11 species and two diameter thresholds).
-For the models that simulated additional species, their regeneration
-rates were aggregated as "others".
-
-Two models did not provide results from all the simulated samples or
-sites to avoid unrealistic results, as follows. In the model 4C, a
-threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid
-unrealistic stand basal area data, and therefore not all the samples and
-sites were reported. The reason for this is that 4C is not suitable for
-long-term simulations without management, due to misrepresentations in
-density-dependent mortality processes in long-term simulations and
-assumptions of tree geometry that lead to unrealistic single tree
-dimensions of very old individuals. LPJ-GUESS had 2% of the sample
-outputs without tree recruitment, and these were considered as zero
-stand basal area and zero regeneration for all the tree species
-simulated; one site produced grassland rather than a forest and was not
-included in the results.
+The overarching goal of the simulation experiments was to assess tree recruitment as it arises from empirical data against its representation in a wide range of models of forest dynamics.
+We define tree recruitment as the passing of a breast height diameter threshold of 7 and 10 cm (synonym: ingrowth).
+To this end, each modeling group was provided with a detailed protocol [@díaz-yáñez2022] with instructions how to perform the simulations, input variables on climate and soil conditions, and the list of expected output variables.
+Neither were further site information (except for the data specified below) nor any data on tree recruitment or forest stand features provided prior to the simulation.
+
+The input variables were collected from different data sources and aggregated to be suitable for the needs of the different models.
+Time series of climatic variables from 1981 to 2018 were provided in hourly [Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly resolution [CHELSA data, Karger et al.
+-@karger2021].
+Some variables required for some models, such as relative humidity or vapor pressure deficit, were calculated from these variables.
+The final instructions for using the climate data were slightly different in each model, based on the approach that best suited the model (Table \@ref(tab:tableModels)).
+Soil quality data were provided as continuous values between 1 and 5 [Soilgrids dataset, Hengl et al.
+-@hengl2017].
+The protocol also provided the elevation, slope and aspect for each of the 200 sites [@asterscienceteam2019], but no other spatial information such as coordinates, with the exception of iLand and aDGVM2, which required blurred coordinates to derive highly detailed soil data.
+
+The simulations were run in the absence of natural disturbances.
+Only the model LPJ-GUESS had to include a "background" disturbance probability to increase chances of shade-intolerant species to establish.
+The simulations were set up to sample species-specific recruitment rates per decade and per ha in the equilibrium state of the model, typically entailing a "spin-up" run (as we did not provide any forest data).
+The modeling teams decided on the simulated area and how they derived these samples.
+The exact length of the simulation was also decided by the modeling teams (Table \@ref(tab:tableModels)).
+Further details on how each modeling team prepared the simulations and the outputs are available in Appendix S1: Model reports.
+
+The simulations were run in the absence of management to a simulated equilibrium ("Potential Natural Vegetation") with the current climate, as the ultimate goal was to evaluate tree recruitment under comparable and near-equilibrium conditions.
+This entails the assumption that (1) the observations from the forest reserves reflect no traces of forest management, and (2) there is an equilibrium between forest dynamics and climate.
+While the former might be starting to be visible in many of the EuFoRIA reserves, the latter may be more debatable.
+However, in the absence of detailed data on the history of each plot in the EuFoRIa network, some broad assumptions had to be made.
+Both the width of the regeneration niche (i.e., in environmental space) as well as the intensity of recruitment (i.e., the number of recruited trees per area and per unit of time) were of interest.
+
+The simulations were run for mixed-species forests (not multiple single-species simulations) using mixtures of eleven species or genera for which regeneration data of sufficient quality were available from EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp.
+L., Fraxinus excelsior L., Quercus spp.
+L., Alnus glutinosa L.,* and *Pinus sylvestris L.* The same set of species was used at all 200 sites.
+Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig), and the species simulated beyond the eleven species were lumped in a category "others".
+In three models, fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)).
+In 4C only *Fagus sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp.
+and *Pinus sylvestris* are parameterised.
+xComp simulations did not consider *Fraxinus excelsior* due to a deprecated species parameterisation.
+In LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not parameterized and therefore these species could not be included in the simulation.
+Finally, the model aDGVM2 does not simulate individual species.
+Rather, community assembly processes and trait filtering generate plant communities that are adjusted to the local biotic and abiotic conditions, and the simulated plants can be classified into ecological strategies based on their trait values in a post-processing step.
+
+Each of the models reported the number of new trees crossing the two diameter thresholds by sampling 200 times in a 10-year interval for each species and per ha for each of the 200 sites.
+Multiple samples per site were used to better understand the simulated variation within each site.
+This was done using different strategies, depending on the model, including (1) sampling simulated data from the same 1 ha plot in the equilibrium over time, (2) sampling several 1 ha plots from the simulated forest at one specific point in time (in the equilibrium) or (3) a combination of (1) and (2).
+This resulted in 880,000 observations per model that simulated the 11 species included in the protocol (200 sites, 200 samples per site, 11 species and two diameter thresholds).
+For the models that simulated additional species, their regeneration rates were aggregated as "others".
+
+Two models did not provide results from all the simulated samples or sites to avoid unrealistic results, as follows.
+In the model 4C, a threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid unrealistic stand basal area data, and therefore not all the samples and sites were reported.
+The reason for this is that 4C is not suitable for long-term simulations without management, due to misrepresentations in density-dependent mortality processes in long-term simulations and assumptions of tree geometry that lead to unrealistic single tree dimensions of very old individuals.
+LPJ-GUESS had 2% of the sample outputs without tree recruitment, and these were considered as zero stand basal area and zero regeneration for all the tree species simulated; one site produced grassland rather than a forest and was not included in the results.
 
@@
 
-The simulation results were analyzed regarding (1) recruitment levels
-(i.e., ingrowth number per unit time and space), (2) recruitment species
-diversity, (3) recruitment mortality, (4) the relationship between model
-performance and model traits, and (5) ingrowth gradients along the
-regeneration niches. We evaluated species diversity across the models
-and in relation to the observed data by calculating the Shannon index
-$H$ based on the relative proportion of the species in terms of basal
-area. It was calculated for the regeneration ($H_{R_{n}}$) (Equation
-\ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation
-\ref{eqpiShannon}). The higher the value of the index, the higher the
-species diversity at a particular site and sample.
+The simulation results were analyzed regarding (1) recruitment levels (i.e., ingrowth number per unit time and space), (2) recruitment species diversity, (3) recruitment mortality, (4) the relationship between model performance and model traits, and (5) ingrowth gradients along the regeneration niches.
+We evaluated species diversity across the models and in relation to the observed data by calculating the Shannon index $H$ based on the relative proportion of the species in terms of basal area.
+It was calculated for the regeneration ($H_{R_{n}}$) (Equation \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation \ref{eqpiShannon}).
+The higher the value of the index, the higher the species diversity at a particular site and sample.
 
@@
 ```
-where $s$ is the total number of species present that have a basal area
-larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
-species $i$ in sample $n$ calculated as the regeneration basal area
-($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal
-area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
-proportion of species $i$ calculated as the basal area of all trees
-($BA_{n,i}$) of that species $i$ relative to total basal area
-($Total_{BA_{n}}$) of the sample $n$. Species diversity was not assessed
-for aDGVM2 as this model does not simulate individual species.
-
-Mortality in tree recruitment was assessed based on the ratio of
-recruitment between the 7 and 10 cm diameter thresholds. We used the
-Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference
-to estimate whether the ratio of recruitment between these two
-thresholds was above or below the expected theoretical rate. The Reineke
-self-thinning rule is usually calculated for even-aged, single species
-stands and is based on a fixed relationship between the number of stems
-and the quadratic mean diameter in fully stocked pure stands. The value
-used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm
-to be 77% higher than at 10 cm), calculated using Equation
-\ref{eqReineke}.
+where $s$ is the total number of species present that have a basal area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of species $i$ in sample $n$ calculated as the regeneration basal area ($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the proportion of species $i$ calculated as the basal area of all trees ($BA_{n,i}$) of that species $i$ relative to total basal area ($Total_{BA_{n}}$) of the sample $n$.
+Species diversity was not assessed for aDGVM2 as this model does not simulate individual species.
+
+Mortality in tree recruitment was assessed based on the ratio of recruitment between the 7 and 10 cm diameter thresholds.
+We used the Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference to estimate whether the ratio of recruitment between these two thresholds was above or below the expected theoretical rate.
+The Reineke self-thinning rule is usually calculated for even-aged, single species stands and is based on a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands.
+The value used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm to be 77% higher than at 10 cm), calculated using Equation \ref{eqReineke}.
 
@@
 ```
-We assessed model performance in relation to model traits focusing on
-(1) model complexity as defined by @bugmann2022, (2) model type
-(empirical or process-based), (3) the presence or absence of a canopy
-feedback for regeneration, and (4) the scale of application of the model
-(stand, landscape or global). We tested for significant differences
-between simulated and observed values of the recruitment levels and
-species diversity by using t-tests with the Bonferroni correction for
-multiple testing. Note that the t-tests does not present conclusive
-evidence for a specific hypotheses, but rather facilitates managing the
-extensive number of comparisons discussed in our study.
-
-Total recruitment and the regeneration niches of the individual species
-were evaluated across the gradients of light availability, temperature,
-and soil moisture as captured in the data. Stand basal area (the basal
-area of all the trees in each sample above the recruitment threshold)
-was used as a proxy for light availability at the forest floor, the
-annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing
-season warmth, and the climatic water balance as a proxy for soil
-moisture [@speich2019]. Recruitment values of each species were
-calculated as the mean across the 200 samples per site. The observed
-data were modeled using a Generalized Additive Model [@wood2011] with a
-negative binomial distribution and restricted maximum likelihood to
-better understand the relationship between the environmental gradients
-and the levels of observed regeneration, relative to the simulation
-results per model.
-
-In order to analyze the regeneration niches across the climatic
-gradients, we focused on five common tree species or genera: *Abies
-alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
-*Quercus* spp. For these species, we calculated the share in the
-recruited basal area per site, where each site represents a part of the
-environmental space, as the mean across the available samples per site:
+We assessed model performance in relation to model traits focusing on (1) model complexity as defined by @bugmann2022, (2) model type (empirical or process-based), (3) the presence or absence of a canopy feedback for regeneration, and (4) the scale of application of the model (stand, landscape or global).
+We tested for significant differences between simulated and observed values of the recruitment levels and species diversity by using t-tests with the Bonferroni correction for multiple testing.
+Note that the t-tests does not present conclusive evidence for a specific hypotheses, but rather facilitates managing the extensive number of comparisons discussed in our study.
+
+Total recruitment and the regeneration niches of the individual species were evaluated across the gradients of light availability, temperature, and soil moisture as captured in the data.
+Stand basal area (the basal area of all the trees in each sample above the recruitment threshold) was used as a proxy for light availability at the forest floor, the annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing season warmth, and the climatic water balance as a proxy for soil moisture [@speich2019].
+Recruitment values of each species were calculated as the mean across the 200 samples per site.
+The observed data were modeled using a Generalized Additive Model [@wood2011] with a negative binomial distribution and restricted maximum likelihood to better understand the relationship between the environmental gradients and the levels of observed regeneration, relative to the simulation results per model.
+
+In order to analyze the regeneration niches across the climatic gradients, we focused on five common tree species or genera: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and *Quercus* spp.
+For these species, we calculated the share in the recruited basal area per site, where each site represents a part of the environmental space, as the mean across the available samples per site:
 
@@
 ```
-where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
-is the mean basal area in the recruitment (subsequently referred to as
-"recruited basal area") of species $i$ at site $t$ across the available
-samples at that site. Furthermore, regarding the share in the recruited
-basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this
-as zero when the recruited basal area of that species
-($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total recruited
-basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the recruited
-basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero.
+where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$ is the mean basal area in the recruitment (subsequently referred to as "recruited basal area") of species $i$ at site $t$ across the available samples at that site.
+Furthermore, regarding the share in the recruited basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the recruited basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total recruited basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the recruited basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero.
 
@@
 
-Simulated levels of recruitment varied strongly across the 15 models and
-typically did not match the levels found in the forest reserve data
-(Figure 2). Recruitment was overestimated in most models for both the 7
-and 10 cm diameter thresholds, with the exception of the empirical stand
-model SIBYLA, the landscape model Landis-II and the global model aDGVM2,
-which estimated recruited levels at the lower end of the plausibility
-interval of the observed data (Figure 2). The models with the largest
-overestimation were the stand model PICUS and the landscape model
-TreeMig. For most models, the variability of simulated recruitment
-levels across the 200 sites (visible from the interquartile range in the
-box plots of Figure 2) was similar to or smaller than observed, with the
-notable exception of PICUS, where simulated recruitment variability was
-much larger.
-
-Both the observations and the simulated data had no recruitment in some
-samples and at some sites (for details, cf. Table S1). The observed data
-had 4% of the samples with no recruitment. Only three models had a
-larger proportion of no recruitment (4C, Landis-II and aDVM2). Two
-models always simulated recruitment for both the 7 and 10 cm threshold
-(xComp and TreeMig), i.e., they did not feature any zero values. The
-other ten models had a very low percentage of samples with no
-recruitment (0.01% - 2.39%), i.e., they had distinctly fewer occurrences
-of zeros than the observations.
+Simulated levels of recruitment varied strongly across the 15 models and typically did not match the levels found in the forest reserve data (Figure 2).
+Recruitment was overestimated in most models for both the 7 and 10 cm diameter thresholds, with the exception of the empirical stand model SIBYLA, the landscape model Landis-II and the global model aDGVM2, which estimated recruited levels at the lower end of the plausibility interval of the observed data (Figure 2).
+The models with the largest overestimation were the stand model PICUS and the landscape model TreeMig.
+For most models, the variability of simulated recruitment levels across the 200 sites (visible from the interquartile range in the box plots of Figure 2) was similar to or smaller than observed, with the notable exception of PICUS, where simulated recruitment variability was much larger.
+
+Both the observations and the simulated data had no recruitment in some samples and at some sites (for details, cf.
+Table S1).
+The observed data had 4% of the samples with no recruitment.
+Only three models had a larger proportion of no recruitment (4C, Landis-II and aDVM2).
+Two models always simulated recruitment for both the 7 and 10 cm threshold (xComp and TreeMig), i.e., they did not feature any zero values.
+The other ten models had a very low percentage of samples with no recruitment (0.01% - 2.39%), i.e., they had distinctly fewer occurrences of zeros than the observations.
 
@@
 
-Most models matched the level of diversity of the observed data quite
-well (Figure 3). Five models overestimated recruitment diversity:
-ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
-particularly for the 7 cm diameter threshold. The model 4C is a special
-case, as it simulated five species only, i.e., its diversity values are
-not directly comparable to those of the other models, nor to the
-observations. Only one model, Landis-II, consistently underestimated
-recruited diversity.
-
-In most models, there were only small but significant differences in the
-species diversity of recruitment between 7 and 10 cm across sites
-(Appendix S1: Table S2). Four models (iLand, Landis-II, TreeMig and
-LPJ-GUESS) maintained the recruitment diversity between the 7 and 10 cm
-thresholds (i.e., the differences between them were not significant, see
-Appendix S1: Table S2), and the same was evident from the observed data.
-
-For both the observed and the simulated data, and both recruitment
-thresholds, species diversity in the recruitment compared to stand-level
-diversity did not feature clear patterns across the models (Appendix S1:
-Figures S1 and S2). Most models captured reasonably well or
-overpredicted species diversity at the stand level (Figure 4A, left and
-center; cf. Appendix S1: Figure S3), but a characteristic feature was
-that the simulations had a much lower variability of diversity compared
-to observations. Models overpredicting stand-level species diversity
-included ForClim1, ForClim 11, FORMIND and TreeMig (Figure 4A, left).
-Several models underpredicted species diversity, i.e., LPJ-GUESS,
-ForCEEPS(f) and 4C (Figure 4A, right).
-
-The majority of the models overpredicted recruitment diversity; as
-already observed in Figure 3, only two of them underestimated it (4C and
-Landis-II) (Figure 4B and Appendix S1: Figure S4). Both the models with
-and without feedback from the adult trees to regeneration (via seed
-production) overestimated, underestimated or captured reasonably well
-recruitment species diversity compared to the observations. Again,
-simulated diversity had considerably lower variability than observed
-diversity in the recruitment.
+Most models matched the level of diversity of the observed data quite well (Figure 3).
+Five models overestimated recruitment diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter particularly for the 7 cm diameter threshold.
+The model 4C is a special case, as it simulated five species only, i.e., its diversity values are not directly comparable to those of the other models, nor to the observations.
+Only one model, Landis-II, consistently underestimated recruited diversity.
+
+In most models, there were only small but significant differences in the species diversity of recruitment between 7 and 10 cm across sites (Appendix S1: Table S2).
+Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained the recruitment diversity between the 7 and 10 cm thresholds (i.e., the differences between them were not significant, see Appendix S1: Table S2), and the same was evident from the observed data.
+
+For both the observed and the simulated data, and both recruitment thresholds, species diversity in the recruitment compared to stand-level diversity did not feature clear patterns across the models (Appendix S1: Figures S1 and S2).
+Most models captured reasonably well or overpredicted species diversity at the stand level (Figure 4A, left and center; cf.
+Appendix S1: Figure S3), but a characteristic feature was that the simulations had a much lower variability of diversity compared to observations.
+Models overpredicting stand-level species diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure 4A, left).
+Several models underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure 4A, right).
+
+The majority of the models overpredicted recruitment diversity; as already observed in Figure 3, only two of them underestimated it (4C and Landis-II) (Figure 4B and Appendix S1: Figure S4).
+Both the models with and without feedback from the adult trees to regeneration (via seed production) overestimated, underestimated or captured reasonably well recruitment species diversity compared to the observations.
+Again, simulated diversity had considerably lower variability than observed diversity in the recruitment.
 
@@
 
-There were strong differences among models regarding the mortality rate
-between 7 and 10 cm (Figure 5). Most of the models featured a mortality
-rate significantly larger than the observed data, thus at least
-partially compensating for the general overestimation of recruitment
-levels at 7 cm (Figure 2 and Appendix S1: Table S3); six models did not
-feature a significant difference compared to the observed data
-(ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2). Four models
-(xComp, PICUS, LandClim and TreeMig) featured very high mortality rates
-(i.e., well above the Reineke self-thinning line), which compensated for
-the strong initial overestimation of recruitment (cf. Figure 2).
-Interestingly, two models that underestimated overall recruitment levels
-(SIBYLA and Landis-II; cf. Figure 2) featured mortality rates that were
-close to but still above the self-thinning line (Figure 5).
-
-The fact that some models (and observed data) featured lower mortality
-than expected by Reineke's self-thinning rule may point to facilitation,
-or simply a higher mortality rate before the trees had reached 7 cm
-diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
-whose mortality was well above the self-thinning line, indicates that
-they feature higher recruitment mortality compared to even-aged,
-single-species forests. Some models yielded a mortality rate of
-recruitment that is broadly compatible with the self-thinning rule
-(e.g., 4C and LandClim).
-
-The three models with the largest overestimation of the proportion of
-recruitment at the 7 cm threshold featured the highest mortality between
-the 7 and 10 cm thresholds (Figure 6A). The other models that
-overestimated recruitment had mortalities lower than the expected
-self-thinning ratio, with the exception of LandClim (Figure 6B). It is
-noteworthy that most of the models that featured a low ratio of
-recruitment between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11
-and iLand) had only a small overestimation of recruitment at the 7 cm
-threshold (Figure 5). aDGVM2 was the only model that underestimated
-recruitment at the 7 cm threshold while featuring a low mortality
-between the two thresholds (Figure 6C), and only a few models were close
-to the correct recruitment levels at 7 cm while also being close to the
-Reineke line (SIBYLA, Landis-II; cf. Figure 6D; and 4C as well as
-LandClim).
-
-Lastly, some models that had a low number of recruits at the 7 cm
-threshold also had a small decrease of recruitment between 7 and 10 cm
-(ratio 7-10 cm) (Appendix S1: Figure S5). In these models, competition
-and self-thinning are either not pronounced or must have occurred before
-the trees had reached 7 cm. However, this pattern was not consistent
-across models (Appendix S1: Figure S5). It is noteworthy that there is a
-relationship between the recruitment levels at 7 cm and the mortality
-rate (Appendix S1: Table S4): most of the models showed a positive
-effect (higher regeneration at 7 cm is coupled to an increase of the
-mortality between 7 and 10 cm). While some models showed a negative
-effect, the linear trend was not pronounced and not always significant
-regarding its slope.
+There were strong differences among models regarding the mortality rate between 7 and 10 cm (Figure 5).
+Most of the models featured a mortality rate significantly larger than the observed data, thus at least partially compensating for the general overestimation of recruitment levels at 7 cm (Figure 2 and Appendix S1: Table S3); six models did not feature a significant difference compared to the observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
+Four models (xComp, PICUS, LandClim and TreeMig) featured very high mortality rates (i.e., well above the Reineke self-thinning line), which compensated for the strong initial overestimation of recruitment (cf.
+Figure 2).
+Interestingly, two models that underestimated overall recruitment levels (SIBYLA and Landis-II; cf.
+Figure 2) featured mortality rates that were close to but still above the self-thinning line (Figure 5).
+
+The fact that some models (and observed data) featured lower mortality than expected by Reineke's self-thinning rule may point to facilitation, or simply a higher mortality rate before the trees had reached 7 cm diameter.
+Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher recruitment mortality compared to even-aged, single-species forests.
+Some models yielded a mortality rate of recruitment that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).
+
+The three models with the largest overestimation of the proportion of recruitment at the 7 cm threshold featured the highest mortality between the 7 and 10 cm thresholds (Figure 6A).
+The other models that overestimated recruitment had mortalities lower than the expected self-thinning ratio, with the exception of LandClim (Figure 6B).
+It is noteworthy that most of the models that featured a low ratio of recruitment between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small overestimation of recruitment at the 7 cm threshold (Figure 5).
+aDGVM2 was the only model that underestimated recruitment at the 7 cm threshold while featuring a low mortality between the two thresholds (Figure 6C), and only a few models were close to the correct recruitment levels at 7 cm while also being close to the Reineke line (SIBYLA, Landis-II; cf.
+Figure 6D; and 4C as well as LandClim).
+
+Lastly, some models that had a low number of recruits at the 7 cm threshold also had a small decrease of recruitment between 7 and 10 cm (ratio 7-10 cm) (Appendix S1: Figure S5).
+In these models, competition and self-thinning are either not pronounced or must have occurred before the trees had reached 7 cm.
+However, this pattern was not consistent across models (Appendix S1: Figure S5).
+It is noteworthy that there is a relationship between the recruitment levels at 7 cm and the mortality rate (Appendix S1: Table S4): most of the models showed a positive effect (higher regeneration at 7 cm is coupled to an increase of the mortality between 7 and 10 cm).
+While some models showed a negative effect, the linear trend was not pronounced and not always significant regarding its slope.
 
@@
 
-There was no significant relationship between the mean complexity of the
-regeneration module (Table \@ref(tab:tableModels)) and the
-overestimation proportion at a diameter threshold of 7 cm (Appendix S1:
-Figure S6). All models except 4C had significant differences between the
-observed and simulated mean recruitment values. Therefore it was not
-possible to differentiate in terms of performance between models type
-(empirical or process based), or their scales (stand, landscape or
-global) (Appendix S1: Table S5). The values of species diversity in the
-recruitment simulated by each model were also assessed against the
-observed data to see if there was a difference between models that
-include a feedback compared to those that did not; all the models except
-ForCEEPs(f) and FORMIND had significant differences (Appendix S1: Table
-S6). It is noteworthy that for the model ForCEEPS the pattern changed
-from the overprediction of species diversity in the regeneration to a
-diversity level that is closer to observations when the recruitment
-module included feedback (ForCEEPs(f)).
+There was no significant relationship between the mean complexity of the regeneration module (Table \@ref(tab:tableModels)) and the overestimation proportion at a diameter threshold of 7 cm (Appendix S1: Figure S6).
+All models except 4C had significant differences between the observed and simulated mean recruitment values.
+Therefore it was not possible to differentiate in terms of performance between models type (empirical or process based), or their scales (stand, landscape or global) (Appendix S1: Table S5).
+The values of species diversity in the recruitment simulated by each model were also assessed against the observed data to see if there was a difference between models that include a feedback compared to those that did not; all the models except ForCEEPs(f) and FORMIND had significant differences (Appendix S1: Table S6).
+It is noteworthy that for the model ForCEEPS the pattern changed from the overprediction of species diversity in the regeneration to a diversity level that is closer to observations when the recruitment module included feedback (ForCEEPs(f)).
 
@@
 
-When evaluating total tree recruitment levels along key gradients of
-light availability (basal area), temperature (degree-days) and soil
-moisture (climatic water balance), distinct features emerged: the models
-reproduced the effect of basal area in both its magnitude and patterns
-(Figure 7 and Appendix S1: Figure S7) considerably better than the
-effects of the climatic gradients, where they featured varying patterns.
-
-In the observed data, recruitment levels decreased clearly with
-increasing stand basal area (Figure 7a). This trend was captured by the
-Landis-II model only, albeit at much lower values of basal area than in
-the observations. The other models featured distinctly different trends,
-such as (1) an increase in recruitment levels with increasing total
-basal area followed by the absence of recruitment at high values of
-basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2)
-almost constant recruitment levels with basal area (ForClim 11, xComp,
-LandClim and aDGVM2), or (3) an increase of recruitment up to a certain
-value of stand basal area followed by a decrease at even higher values
-of stand basal area, with model-specific thresholds (4C, ForClim 1,
-PICUS, SIBYLA and iLand).
-
-The observed recruitment did not change much across the climatic water
-balance gradient, and showed a slight increase with temperature. Neither
-did the models generally match the observed recruitment patterns across
-these gradients (Figure 7b,c), nor was there a clear pattern across
-models. Regarding the soil moisture gradient, a group of models featured
-decreasing recruitment with increasing soil moisture (4C, ForClim 1,
-ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a
-few models showed the opposite trend (ForCEEPS(f), FORMIND and
-LPJ-GUESS). Across the temperature gradient, all the models that
-featured a decrease in recruitment levels with increasing soil moisture
-showed the opposite trend (i.e., more recruitment with increasing
-temperature), except for 4C. The other models did not feature a clear
-pattern.
-
-Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
-sylvestris*) and oak (*Quercus* spp.) were the species observed most
-frequently in the EuFoRIa data. Most models captured well the share of
-basal area in the recruitment of these five species compared to the
-observed data along the two major climatic gradients (Figure 8, cf.
-Appendix S1: Figure S8). However, the models differed considerably in
-the way in which they simulated these climatic niches.
-
-First, some models overestimated recruitment across the entire gradients
-of at least one of the five main tree species (Figure 8, e.g., ForCEEPS,
-ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated
-recruitment of more than one species for a large part of the climate
-gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf. red colors in
-Figure 8). These trends were consistent for both recruitment thresholds,
-although the exact changes in the share of the recruited basal area were
-different (Appendix S1: Figure S8).
-
-Second, there was some congruence in the simulation results by species
-across the models. The share in the recruitment of Silver fir (*Abies
-alba*) was captured well across the gradient by most models except for
-xComp, PICUS, and LandClim, which overestimated the recruitment share of
-this species. There were also some peculiarities evident for some
-models, such as 4C that did not simulate *Abies alba*. The patterns
-across models for *Fagus sylvatica* recruitment were more complex, as
-its recruitment was underestimated by many models across the
-environmental gradients featuring negligible regeneration at most sites,
-while others overestimated it in different parts of the environmental
-space. Most models overestimated the share of *Picea abies* recruitment
-in the cold-wet part of the gradients. *Pinus sylvestris* and *Quercus*
-spp. represented a small share of the recruitment in the observed data
-mostly at warm-dry sites. Many models, however, erroneously featured
-recruitment for these two species along most of the climatic gradients,
-although with a low share of basal area. Five models (4C, FORMIND,
-SIBYLA, xComp, and LandClim) had very small amounts of recruitment of
-*Pinus sylvestris* and *Quercus* spp., or did not feature any
-recruitment of these species at all.
-
-Lastly, there is no model that performed well across the five species
-regarding the species-specific recruitment levels in the environmental
-space (Figure 8). Some models (e.g., ForClim11, PICUS) tended to
-consistently overestimate the recruitment share of all five species, but
-most models overestimated the recruitment share of some species while
-underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
+When evaluating total tree recruitment levels along key gradients of light availability (basal area), temperature (degree-days) and soil moisture (climatic water balance), distinct features emerged: the models reproduced the effect of basal area in both its magnitude and patterns (Figure 7 and Appendix S1: Figure S7) considerably better than the effects of the climatic gradients, where they featured varying patterns.
+
+In the observed data, recruitment levels decreased clearly with increasing stand basal area (Figure 7a).
+This trend was captured by the Landis-II model only, albeit at much lower values of basal area than in the observations.
+The other models featured distinctly different trends, such as (1) an increase in recruitment levels with increasing total basal area followed by the absence of recruitment at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant recruitment levels with basal area (ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of recruitment up to a certain value of stand basal area followed by a decrease at even higher values of stand basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).
+
+The observed recruitment did not change much across the climatic water balance gradient, and showed a slight increase with temperature.
+Neither did the models generally match the observed recruitment patterns across these gradients (Figure 7b,c), nor was there a clear pattern across models.
+Regarding the soil moisture gradient, a group of models featured decreasing recruitment with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS).
+Across the temperature gradient, all the models that featured a decrease in recruitment levels with increasing soil moisture showed the opposite trend (i.e., more recruitment with increasing temperature), except for 4C.
+The other models did not feature a clear pattern.
+
+Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus sylvestris*) and oak (*Quercus* spp.) were the species observed most frequently in the EuFoRIa data.
+Most models captured well the share of basal area in the recruitment of these five species compared to the observed data along the two major climatic gradients (Figure 8, cf.
+Appendix S1: Figure S8).
+However, the models differed considerably in the way in which they simulated these climatic niches.
+
+First, some models overestimated recruitment across the entire gradients of at least one of the five main tree species (Figure 8, e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated recruitment of more than one species for a large part of the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
+red colors in Figure 8).
+These trends were consistent for both recruitment thresholds, although the exact changes in the share of the recruited basal area were different (Appendix S1: Figure S8).
+
+Second, there was some congruence in the simulation results by species across the models.
+The share in the recruitment of Silver fir (*Abies alba*) was captured well across the gradient by most models except for xComp, PICUS, and LandClim, which overestimated the recruitment share of this species.
+There were also some peculiarities evident for some models, such as 4C that did not simulate *Abies alba*.
+The patterns across models for *Fagus sylvatica* recruitment were more complex, as its recruitment was underestimated by many models across the environmental gradients featuring negligible regeneration at most sites, while others overestimated it in different parts of the environmental space.
+Most models overestimated the share of *Picea abies* recruitment in the cold-wet part of the gradients.
+*Pinus sylvestris* and *Quercus* spp.
+represented a small share of the recruitment in the observed data mostly at warm-dry sites.
+Many models, however, erroneously featured recruitment for these two species along most of the climatic gradients, although with a low share of basal area.
+Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had very small amounts of recruitment of *Pinus sylvestris* and *Quercus* spp., or did not feature any recruitment of these species at all.
+
+Lastly, there is no model that performed well across the five species regarding the species-specific recruitment levels in the environmental space (Figure 8).
+Some models (e.g., ForClim11, PICUS) tended to consistently overestimate the recruitment share of all five species, but most models overestimated the recruitment share of some species while underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
 
@@
 
-Tree regeneration is a fundamental process in forest dynamics. Correctly
-capturing it in dynamic models is fundamental to, for example, evaluate
-post-disturbance dynamics and potential long-term recovery trajectories
-as it will define the forest state in the first decades [@seidl2022]. If
-this initial phase is not captured well, we can not properly assess
-aspects such as forest resilience or the timing, magnitude and
-progression of carbon sequestration. In this study, for the first time,
-projections of tree recruitment from multiple models of forest dynamics
-were confronted with a unique dataset from unmanaged forest reserves
-across a large environmental gradient in Europe.
-
-The EuFoRIa data [@käber2023] are exceptional, particularly with respect
-to the number of records (number of sites and repeated measurements on
-unmanaged forests), which is essential for capturing a highly 'noisy'
-process such as tree regeneration. The use of this dataset for model
-benchmarking provided novel insights on the ability of state-of-the-art
-models to accurately simulate recruitment levels, its species
-composition and mortality in an early stage of tree life, i.e., between
-tree diameters of 7 and 10 cm. Overall, by adopting this approach a much
-broader understanding resulted than if we had used these data for model
-calibration: it is primarily from the shortcomings of the models that we
-can gain ecological insights [cf. @trugman2022].
+Tree regeneration is a fundamental process in forest dynamics.
+Correctly capturing it in dynamic models is fundamental to, for example, evaluate post-disturbance dynamics and potential long-term recovery trajectories as it will define the forest state in the first decades [@seidl2022].
+If this initial phase is not captured well, we can not properly assess aspects such as forest resilience or the timing, magnitude and progression of carbon sequestration.
+In this study, for the first time, projections of tree recruitment from multiple models of forest dynamics were confronted with a unique dataset from unmanaged forest reserves across a large environmental gradient in Europe.
+
+The EuFoRIa data [@käber2023] are exceptional, particularly with respect to the number of records (number of sites and repeated measurements on unmanaged forests), which is essential for capturing a highly 'noisy' process such as tree regeneration.
+The use of this dataset for model benchmarking provided novel insights on the ability of state-of-the-art models to accurately simulate recruitment levels, its species composition and mortality in an early stage of tree life, i.e., between tree diameters of 7 and 10 cm.
+Overall, by adopting this approach a much broader understanding resulted than if we had used these data for model calibration: it is primarily from the shortcomings of the models that we can gain ecological insights [cf.
+@trugman2022].
 
@@
 
-Most of the models overestimated tree recruitment levels. This has
-potentially far-reaching implications, e.g., regarding biomass (and thus
-carbon) turnover, with a potential overestimation of the capacity for
-forest carbon sequestration [@pugh2019; @pan2011]. Yet, we focused on
-one specific stage, i.e., recruitment into the 7 and 10 cm diameter
-classes. Trees at this size contribute little to carbon sequestration,
-and if the excess regeneration at this stage is compensated -- in the
-models -- soon thereafter by higher mortality, overall simulated tree
-population dynamics may still be trustworthy at the level of stand
-structure and attributes such as basal area or biomass.
-
-From a modeling point of view, excess tree recruitment inevitably
-requires excess mortality rates in a later stage, either -- as observed
-for some models -- between the 7 and 10 cm diameter thresholds, or soon
-after the 10 cm threshold has been crossed. In any case, correcting at
-early stages for the expected forest densities at later stages is
-equivalent to compensating for a first error (excess recruitment) by a
-second error (excess mortality). It is highly likely that biased
-projections will result, because the two errors are unlikely to be
-perfectly linked and thus will not always compensate each other. Hence,
-this structural problem of most models investigated here is problematic
-particularly if the models are to be used under novel conditions such as
-under climate change [e.g., @huber2021] or in a decision support context
-[e.g., @thrippleton2021].
-
-Recruitment levels define the structure and composition of future
-forests, but it is equally important to correctly identify areas where
-regeneration is lacking [@rammer2021]. There are multiple constraints to
-the regeneration niche of tree species [@price2001], and therefore the
-absence of regeneration is likely to be common [@fortin2007], even over
-larger areas such as the one-hectare plots used here. Tree regeneration
-data are characterized by zero-inflation. This was clearly evident from
-the EuFoRIa dataset, but some of the models did not produce zeros at
-all, or featured a very low proportion of zero data. This substantial
-difference may be due to the fact that the simulation results were drawn
-from equilibrium forests, whereas in reality many of the forest reserves
-are recovering from past management activities and have become denser
-over the past decades [e.g., @heiri2009], leading to less regeneration
-than in an equilibrium situation. Another possible reason for this
-difference may be the exclusion of factors like deer browsing or failure
-to accurately measure establishment filters, such as competition with
-the herb layer, site-specific or environmental limitations.
+Most of the models overestimated tree recruitment levels.
+This has potentially far-reaching implications, e.g., regarding biomass (and thus carbon) turnover, with a potential overestimation of the capacity for forest carbon sequestration [@pugh2019; @pan2011].
+Yet, we focused on one specific stage, i.e., recruitment into the 7 and 10 cm diameter classes.
+Trees at this size contribute little to carbon sequestration, and if the excess regeneration at this stage is compensated -- in the models -- soon thereafter by higher mortality, overall simulated tree population dynamics may still be trustworthy at the level of stand structure and attributes such as basal area or biomass.
+
+From a modeling point of view, excess tree recruitment inevitably requires excess mortality rates in a later stage, either -- as observed for some models -- between the 7 and 10 cm diameter thresholds, or soon after the 10 cm threshold has been crossed.
+In any case, correcting at early stages for the expected forest densities at later stages is equivalent to compensating for a first error (excess recruitment) by a second error (excess mortality).
+It is highly likely that biased projections will result, because the two errors are unlikely to be perfectly linked and thus will not always compensate each other.
+Hence, this structural problem of most models investigated here is problematic particularly if the models are to be used under novel conditions such as under climate change [e.g., @huber2021] or in a decision support context [e.g., @thrippleton2021].
+
+Recruitment levels define the structure and composition of future forests, but it is equally important to correctly identify areas where regeneration is lacking [@rammer2021].
+There are multiple constraints to the regeneration niche of tree species [@price2001], and therefore the absence of regeneration is likely to be common [@fortin2007], even over larger areas such as the one-hectare plots used here.
+Tree regeneration data are characterized by zero-inflation.
+This was clearly evident from the EuFoRIa dataset, but some of the models did not produce zeros at all, or featured a very low proportion of zero data.
+This substantial difference may be due to the fact that the simulation results were drawn from equilibrium forests, whereas in reality many of the forest reserves are recovering from past management activities and have become denser over the past decades [e.g., @heiri2009], leading to less regeneration than in an equilibrium situation.
+Another possible reason for this difference may be the exclusion of factors like deer browsing or failure to accurately measure establishment filters, such as competition with the herb layer, site-specific or environmental limitations.
 
@@
 
-Correctly capturing species composition in tree recruitment is important
-to assess the future functional diversity of a forest, e.g., its
-sensitivity to drought or resilience to disturbances [@redmond2015;
-@seidl2022]. In the simulations, overall species diversity levels in the
-recruitment were well within the observed range for half of the models
-(7 out of 14). Thus, while most models are facing difficulties to
-quantitatively match recruitment levels (cf. above), their performance
-is better regarding the composition of recruitment as a function of
-abiotic and biotic conditions. Most models maintained or even decreased
-recruitment diversity between the 7 and 10 cm thresholds, and the same
-is visible from the empirical data, but the differences between the 7
-and 10 cm thresholds were not significant. Diversity variations may be
-explained by the way the individual models consider regeneration
-processes [@konig_tree_2022], e.g., regeneration can be linked to the
-seed/seedling bank, seed rain from dispersal or from a pool of available
-species.
-
-The species diversity of the entire stand was captured better by the
-models than the diversity of tree recruitment, and only a few models
-overpredicted stand-level diversity. Defining recruitment composition as
-being proportional to local adult abundance, regardless of productivity,
-might be a simple and conservative assumption to maintain relative
-species abundances [@hanbury-brown2022], although this may be
-simplistic. Furthermore, based on the simulation results, there is no
-evidence that models with feedback from the canopy (in terms of species
-composition of regeneration) captured the species diversity of the
-recruitment better than those without feedback, with the exception of
-ForCEEPS that featured significantly more accurate recruitment species
-diversity with the model version that included feedback (ForCEEPS(f)).
-The similar performance of models with and without feedback is likely
-because the models put more weight on the regeneration niche arising
-from abiotic and biotic filters, than from the habitat niche of the
-adult trees [@grubb1977].
+Correctly capturing species composition in tree recruitment is important to assess the future functional diversity of a forest, e.g., its sensitivity to drought or resilience to disturbances [@redmond2015; @seidl2022].
+In the simulations, overall species diversity levels in the recruitment were well within the observed range for half of the models (7 out of 14).
+Thus, while most models are facing difficulties to quantitatively match recruitment levels (cf.
+above), their performance is better regarding the composition of recruitment as a function of abiotic and biotic conditions.
+Most models maintained or even decreased recruitment diversity between the 7 and 10 cm thresholds, and the same is visible from the empirical data, but the differences between the 7 and 10 cm thresholds were not significant.
+Diversity variations may be explained by the way the individual models consider regeneration processes [@konig_tree_2022], e.g., regeneration can be linked to the seed/seedling bank, seed rain from dispersal or from a pool of available species.
+
+The species diversity of the entire stand was captured better by the models than the diversity of tree recruitment, and only a few models overpredicted stand-level diversity.
+Defining recruitment composition as being proportional to local adult abundance, regardless of productivity, might be a simple and conservative assumption to maintain relative species abundances [@hanbury-brown2022], although this may be simplistic.
+Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured the species diversity of the recruitment better than those without feedback, with the exception of ForCEEPS that featured significantly more accurate recruitment species diversity with the model version that included feedback (ForCEEPS(f)).
+The similar performance of models with and without feedback is likely because the models put more weight on the regeneration niche arising from abiotic and biotic filters, than from the habitat niche of the adult trees [@grubb1977].
 
@@
 
-There are several factors that lead to mortality during the regeneration
-phase, such as competition [@casper1997] and multiple abiotic factors
-[@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
-expect that models that overestimate recruitment at 7 cm may have a
-particularly high mortality between the 7 and 10 cm thresholds. However,
-this was not consistently evident from the simulations. This implies
-that these models must have an excess mortality in later stages, if we
-assume that all models were able to capture the structure and
-composition of the adult stands along the EuFoRIa gradient; this however
-was not tested here.
-
-Higher mortality towards the adult phase has important implications for
-forest dynamics and the goods and services provided by forests. On the
-one hand, mortality in later stages may erroneously enhance the share of
-less shade tolerant species in the models [@klopcic2015], therefore
-shifting the species composition. Unrealistic high stem densities for a
-longer period of time may overestimate the role of tree regrowth in
-carbon sink dynamics [@pugh2019]. Given our set of simulations and
-analyses, we cannot conclusively assess what is happening in the models,
-and further studies are required that focus on a wider range of tree
-sizes and the fate of tree regeneration along such a size continuum.
+There are several factors that lead to mortality during the regeneration phase, such as competition [@casper1997] and multiple abiotic factors [@schmid2021; @cunningham2006].
+As mentioned above, it is reasonable to expect that models that overestimate recruitment at 7 cm may have a particularly high mortality between the 7 and 10 cm thresholds.
+However, this was not consistently evident from the simulations.
+This implies that these models must have an excess mortality in later stages, if we assume that all models were able to capture the structure and composition of the adult stands along the EuFoRIa gradient; this however was not tested here.
+
+Higher mortality towards the adult phase has important implications for forest dynamics and the goods and services provided by forests.
+On the one hand, mortality in later stages may erroneously enhance the share of less shade tolerant species in the models [@klopcic2015], therefore shifting the species composition.
+Unrealistic high stem densities for a longer period of time may overestimate the role of tree regrowth in carbon sink dynamics [@pugh2019].
+Given our set of simulations and analyses, we cannot conclusively assess what is happening in the models, and further studies are required that focus on a wider range of tree sizes and the fate of tree regeneration along such a size continuum.
 
@@
 
-In spite of the critical considerations above, it is remarkable that
-most models did not deviate exceedingly from the observations with
-respect to simulated recruited levels -- after all, few if any of the
-models' regeneration routines are well constrained by data, with the
-exception of the empirical models xComp and SIBYLA. It is noteworthy
-that in spite of their empirical basis, these two models did not match
-empirical recruitment levels, in a similar magnitude as the other models
-and even in opposite directions (SIBYLA: underestimation; xComp:
-overestimation). It appears that using empirical data of limited
-geographical scope that are constrained to managed forests, as done in
-these models, leads to extrapolation problems already under current
-climatic conditions. For the other models, whose regeneration modules
-are not strongly constrained by empirical data, the multiple strategies
-that are available for modeling regeneration processes
-[@konig_tree_2022] could have implied that model performance would be
-much worse than what we found.
-
-Our study showed that increasing complexity in the regeneration modules
-is not linked to a higher accuracy of the projections of recruitment
-levels, species composition or mortality at early tree stages, as there
-was no significant relationship between model performance and model
-complexity. Increasing complexity in regeneration modules has been
-called for a long time ago (e.g., @price2001) motivated by better
-process understanding and with the goal of enhancing model accuracy
-[@bugmann2022]. However, more complex models do may not necessarily lead
-to better projections but rather to reduced transparency and lower
-predictive power [@franklin2020]. Thus, the question regarding the level
-of detail that is appropriate and parsimonious for modeling tree
-regeneration remains open [@konig_tree_2022; @bugmann2022].
+In spite of the critical considerations above, it is remarkable that most models did not deviate exceedingly from the observations with respect to simulated recruited levels -- after all, few if any of the models' regeneration routines are well constrained by data, with the exception of the empirical models xComp and SIBYLA.
+It is noteworthy that in spite of their empirical basis, these two models did not match empirical recruitment levels, in a similar magnitude as the other models and even in opposite directions (SIBYLA: underestimation; xComp: overestimation).
+It appears that using empirical data of limited geographical scope that are constrained to managed forests, as done in these models, leads to extrapolation problems already under current climatic conditions.
+For the other models, whose regeneration modules are not strongly constrained by empirical data, the multiple strategies that are available for modeling regeneration processes [@konig_tree_2022] could have implied that model performance would be much worse than what we found.
+
+Our study showed that increasing complexity in the regeneration modules is not linked to a higher accuracy of the projections of recruitment levels, species composition or mortality at early tree stages, as there was no significant relationship between model performance and model complexity.
+Increasing complexity in regeneration modules has been called for a long time ago (e.g., @price2001) motivated by better process understanding and with the goal of enhancing model accuracy [@bugmann2022].
+However, more complex models do may not necessarily lead to better projections but rather to reduced transparency and lower predictive power [@franklin2020].
+Thus, the question regarding the level of detail that is appropriate and parsimonious for modeling tree regeneration remains open [@konig_tree_2022; @bugmann2022].
 
@@
 
-Competition for light as a strong filter for tree regeneration has been
-widely documented [@messier1999; @collet2006; @berdanier2016], but the
-models examined here did not reproduce this expectation (i.e.,
-decreasing recruitment levels with increasing stand basal area).
-However, it is difficult to measure light availability at large spatial
-and temporal scales. We used total stand basal area as a proxy for light
-availability [cf. @schmid2021]. Yet, we were unable to consider light
-availability restrictions caused by ground vegetation, which may be an
-important filter for forest dynamics at least in some EuFoRIa stands
-[@woltjer2008]. We found pronounced differences in the ranges of stand
-basal area simulated by the models, but also between models and
-observations. This made it impossible to evaluate tree recruitment for
-the extremes of the stand density ranges in some models. This is
-unfortunate because recruitment levels at low stand density are relevant
-to assess how well forests are recovering e.g. after gap creation due to
-disturbance [@seidl2022; @grubb1977]. At the level of the simulated
-one-hectare samples, average basal area is typically not very low as
-long as no larger disturbances occur, which was explicitly excluded in
-our simulation protocol. Towards the other end of the spectrum, i.e.,
-with increasing stand basal area, it would be reasonable to expect that
-the recruitment of the different species would become sparser and drop
-out entirely under low-light conditions [@klopcic2012; @zell2019].
-However, few models showed this trend, thus indicating that the relation
-between regeneration and light availability is not yet captured
-correctly in most models. Yet, several of the models that did feature an
-increase of regeneration with increasing basal area include a feedback
-between seed production of mature trees and regeneration. Thus, it seems
-that in these models higher light competition does not sufficiently
-compensate for increased seed availability with higher basal area due to
-a higher abundance of mature trees.
-
-Lastly, there were pronounced differences how the main tree species were
-represented by the models along the environmental gradients, in
-particular the dominant species *Abies alba*, *Fagus sylvatica* and
-*Picea abies*. The recruitment levels were sampled from simulations in
-the equilibrium, and in this state it is expected that non-climax
-species such as *Pinus sylvestris* or *Quercus* would be of minor
-importance, or absent [@klopcic2015]. Most models captured this low
-abundance, which is also found in the empirical data. Thus, although the
-broad patterns are matched by many models, improvements in the
-quantification of the regeneration niche of the species are needed, but
-this cannot be done in the absence of robust datasets across multiple
-species.
+Competition for light as a strong filter for tree regeneration has been widely documented [@messier1999; @collet2006; @berdanier2016], but the models examined here did not reproduce this expectation (i.e., decreasing recruitment levels with increasing stand basal area).
+However, it is difficult to measure light availability at large spatial and temporal scales.
+We used total stand basal area as a proxy for light availability [cf.
+@schmid2021].
+Yet, we were unable to consider light availability restrictions caused by ground vegetation, which may be an important filter for forest dynamics at least in some EuFoRIa stands [@woltjer2008].
+We found pronounced differences in the ranges of stand basal area simulated by the models, but also between models and observations.
+This made it impossible to evaluate tree recruitment for the extremes of the stand density ranges in some models.
+This is unfortunate because recruitment levels at low stand density are relevant to assess how well forests are recovering e.g.
+after gap creation due to disturbance [@seidl2022; @grubb1977].
+At the level of the simulated one-hectare samples, average basal area is typically not very low as long as no larger disturbances occur, which was explicitly excluded in our simulation protocol.
+Towards the other end of the spectrum, i.e., with increasing stand basal area, it would be reasonable to expect that the recruitment of the different species would become sparser and drop out entirely under low-light conditions [@klopcic2012; @zell2019].
+However, few models showed this trend, thus indicating that the relation between regeneration and light availability is not yet captured correctly in most models.
+Yet, several of the models that did feature an increase of regeneration with increasing basal area include a feedback between seed production of mature trees and regeneration.
+Thus, it seems that in these models higher light competition does not sufficiently compensate for increased seed availability with higher basal area due to a higher abundance of mature trees.
+
+Lastly, there were pronounced differences how the main tree species were represented by the models along the environmental gradients, in particular the dominant species *Abies alba*, *Fagus sylvatica* and *Picea abies*.
+The recruitment levels were sampled from simulations in the equilibrium, and in this state it is expected that non-climax species such as *Pinus sylvestris* or *Quercus* would be of minor importance, or absent [@klopcic2015].
+Most models captured this low abundance, which is also found in the empirical data.
+Thus, although the broad patterns are matched by many models, improvements in the quantification of the regeneration niche of the species are needed, but this cannot be done in the absence of robust datasets across multiple species.
 
@@
 
-The EuFoRIa data as used here are well suited to better understand tree
-regeneration. However, three aspects of these data may represent
-considerable limitations. First, we made a comparison of tree
-regeneration in an equilibrium state, but we cannot assess how close the
-forests included in the EuFoRIa dataset were to such a state. The data
-were collected in forest reserves where no management has taken place
-for long periods of time. This makes our assumption of an equilibrium
-between forest properties and environmental drivers more reasonable than
-it might appear at first sight. In an analysis of primeval forests at
-demographic equilibrium, @brzeziecki2021 found higher recruitment rates
-than the ones observed in the EuFoRia data. Thus, the overestimation of
-recruitment rates by the models may not be so problematic. Second, the
-data were collected for recruitment above a 7 cm threshold, thus
-limiting the assessment of tree regeneration to a specific point of
-stand dynamics. This constituted a hard limit based on which we can
-understand only some aspects of tree regeneration, which in its entirety
-often comprises a rather long period since seed production [@price2001].
-In reality, many environmental constraints are acting on young trees
-[@käber2021] that we were unable to assess. Yet, recruitment data with
-lower thresholds are simply not available in a harmonized manner across
-large environmental gradients. Third, the empirical data were collected
-from rather small plots, while we sampled simulated recruitment levels
-from 1 ha areas, which may lead to an incorrect representation of space.
-Even though the strategy we adopted is not ideal, it represents a common
-challenge when harmonizing diverse data sources originating from varied
-sampling strategies [@portier2022]. It would have been extremely
-challenging for such a variety of models to follow a protocol where the
-spatial sampling size was different at each of the 200 sites, and it
-would have introduced additional uncertainty in the results.
-
-The design of our sampling protocol did not include spatial aspects such
-as seed dispersal or detailed soil data. While we considered a wide
-range of models of forest dynamics, from stand to global scales, the
-simulation setup was limited a spatial scale of 1 ha. This lack of
-consideration of spatial scale relationships is appropriate for
-stand-scale models, but it potentially puts the landscape models at a
-disadvantage, as they have been built to be accurate at the landscape
-level: without any spatial context, we are limiting tree regeneration to
-the seed influx from the stand itself, unless the model has a background
-seed input. Yet, the global models should not be at a disadvantage here
-due to their inherently limited spatial considerations [but cf. Snell et
-al. -@snell2018; or Lehsten et al. -@lehsten2019], as they are usually
-lacking dispersal between cells and are based on a strong abstraction of
-horizontal space [@hanbury-brown2022]. Lastly, detailed data on soil
-conditions were not available from the observed data, and independent,
-admittedly coarse data for soil properties and the climatic water
-balance had to be used instead. It is noteworthy that many models
-represent drought using detailed indicators based e.g. on soil water
-holding capacity, which had to be derived from a rough soil quality
-measure. This may at least partially explain the unsatisfactory
-performance of many models along the drought axis (i.e., climatic water
-balance).
+The EuFoRIa data as used here are well suited to better understand tree regeneration.
+However, three aspects of these data may represent considerable limitations.
+First, we made a comparison of tree regeneration in an equilibrium state, but we cannot assess how close the forests included in the EuFoRIa dataset were to such a state.
+The data were collected in forest reserves where no management has taken place for long periods of time.
+This makes our assumption of an equilibrium between forest properties and environmental drivers more reasonable than it might appear at first sight.
+In an analysis of primeval forests at demographic equilibrium, @brzeziecki2021 found higher recruitment rates than the ones observed in the EuFoRia data.
+Thus, the overestimation of recruitment rates by the models may not be so problematic.
+Second, the data were collected for recruitment above a 7 cm threshold, thus limiting the assessment of tree regeneration to a specific point of stand dynamics.
+This constituted a hard limit based on which we can understand only some aspects of tree regeneration, which in its entirety often comprises a rather long period since seed production [@price2001].
+In reality, many environmental constraints are acting on young trees [@käber2021] that we were unable to assess.
+Yet, recruitment data with lower thresholds are simply not available in a harmonized manner across large environmental gradients.
+Third, the empirical data were collected from rather small plots, while we sampled simulated recruitment levels from 1 ha areas, which may lead to an incorrect representation of space.
+Even though the strategy we adopted is not ideal, it represents a common challenge when harmonizing diverse data sources originating from varied sampling strategies [@portier2022].
+It would have been extremely challenging for such a variety of models to follow a protocol where the spatial sampling size was different at each of the 200 sites, and it would have introduced additional uncertainty in the results.
+
+The design of our sampling protocol did not include spatial aspects such as seed dispersal or detailed soil data.
+While we considered a wide range of models of forest dynamics, from stand to global scales, the simulation setup was limited a spatial scale of 1 ha.
+This lack of consideration of spatial scale relationships is appropriate for stand-scale models, but it potentially puts the landscape models at a disadvantage, as they have been built to be accurate at the landscape level: without any spatial context, we are limiting tree regeneration to the seed influx from the stand itself, unless the model has a background seed input.
+Yet, the global models should not be at a disadvantage here due to their inherently limited spatial considerations [but cf.
+Snell et al.
+-@snell2018; or Lehsten et al.
+-@lehsten2019], as they are usually lacking dispersal between cells and are based on a strong abstraction of horizontal space [@hanbury-brown2022].
+Lastly, detailed data on soil conditions were not available from the observed data, and independent, admittedly coarse data for soil properties and the climatic water balance had to be used instead.
+It is noteworthy that many models represent drought using detailed indicators based e.g.
+on soil water holding capacity, which had to be derived from a rough soil quality measure.
+This may at least partially explain the unsatisfactory performance of many models along the drought axis (i.e., climatic water balance).
 
@@
 
-With our study, we have demonstrated that models of forest dynamics need
-a focus on their regeneration modules to make them more robust. It
-remains uncertain what level of detail is required to model tree
-regeneration, and this must be addressed in future research. We
-recommend that the improvement of the regeneration modules is
-implemented as additional features that can be traced back, as done here
-for the variants of ForClim and ForCEEPS, and that model complexity and
-structure must always be connected with modeling objectives
-[@albrich2020]. If it should be necessary to include more detail in the
-regeneration models, this will come with higher parameterisation
-efforts. This will most likely lead to lower generalization because the
-required data will have to be collected from specific locations, as
-currently there is no general, comprehensive regeneration dataset
-available.
-
-Therefore, we further recommend that more effort should be invested into
-collecting harmonized datasets in a site-specific manner covering the
-different aspects leading to tree regeneration. We emphasize that
-datasets such as EuFoRIa are invaluable and should be expanded in both
-their spatial extent (e.g., towards boreal and Mediterranean conditions)
-as well as in time (e.g., continuing the monitoring into the future).
-Such data will allow for a better evaluation of forest models and help
-to reduce the uncertainty in their projections, which is crucial when
-they are used as tools for predicting e.g. the impacts of anthropogenic
-climate change.
-
-In the present study, we have considered tree regeneration in the
-equilibrium state only. It is equally important to understand how these
-models project tree regeneration after changes in forest structure by
-disturbances [@seidl2022], or under different management strategies
-[@lindner2000]. However, this will require an entirely different set of
-observed data, and potentially not all forest models would be able to
-assess the relationship of these aspects on tree regeneration, e.g. due
-to the lack of disturbance or appropriate management modules.
-
-We recommend to investigate in detail the implications of the current
-modeling strategies for tree regeneration and, ultimately, simulated
-forest stand structure. This applies particularly to the erroneous
-patterns of excess tree regeneration and later excess mortality, by
-focusing on a wider range of tree sizes and the related regeneration
-dynamics. We also recommend, especially for the landscape-level models,
-the inclusion of explicit spatial considerations regarding tree
-regeneration [@beckage2003]; this, however, is a serious challenge
-regarding appropriate datasets, as large-scale inventory data have a
-wide coverage, but by definition do not allow for the assessment of
-spatial interactions.
-
-Exercises like the one presented here, where the models are operated in
-"blind flight" mode, i.e., without the possibility of being tuned
-towards capturing the expected patterns, should be repeated. Such
-benchmarking exercises should next focus on aspects such as specific
-model traits and the ecological formulations of particular
-(sub-)processes, to better understand the implications of the
-assumptions on which the models are based. Furthermore, the inclusion of
-a wide range of models with different (1) scales, (2) approaches for
-capturing population structure, (3) tree regeneration modules and (4)
-complexity of their formulations will ensure a large benefit to the
-entire modeling community and beyond.
+With our study, we have demonstrated that models of forest dynamics need a focus on their regeneration modules to make them more robust.
+It remains uncertain what level of detail is required to model tree regeneration, and this must be addressed in future research.
+We recommend that the improvement of the regeneration modules is implemented as additional features that can be traced back, as done here for the variants of ForClim and ForCEEPS, and that model complexity and structure must always be connected with modeling objectives [@albrich2020].
+If it should be necessary to include more detail in the regeneration models, this will come with higher parameterisation efforts.
+This will most likely lead to lower generalization because the required data will have to be collected from specific locations, as currently there is no general, comprehensive regeneration dataset available.
+
+Therefore, we further recommend that more effort should be invested into collecting harmonized datasets in a site-specific manner covering the different aspects leading to tree regeneration.
+We emphasize that datasets such as EuFoRIa are invaluable and should be expanded in both their spatial extent (e.g., towards boreal and Mediterranean conditions) as well as in time (e.g., continuing the monitoring into the future).
+Such data will allow for a better evaluation of forest models and help to reduce the uncertainty in their projections, which is crucial when they are used as tools for predicting e.g.
+the impacts of anthropogenic climate change.
+
+In the present study, we have considered tree regeneration in the equilibrium state only.
+It is equally important to understand how these models project tree regeneration after changes in forest structure by disturbances [@seidl2022], or under different management strategies [@lindner2000].
+However, this will require an entirely different set of observed data, and potentially not all forest models would be able to assess the relationship of these aspects on tree regeneration, e.g.
+due to the lack of disturbance or appropriate management modules.
+
+We recommend to investigate in detail the implications of the current modeling strategies for tree regeneration and, ultimately, simulated forest stand structure.
+This applies particularly to the erroneous patterns of excess tree regeneration and later excess mortality, by focusing on a wider range of tree sizes and the related regeneration dynamics.
+We also recommend, especially for the landscape-level models, the inclusion of explicit spatial considerations regarding tree regeneration [@beckage2003]; this, however, is a serious challenge regarding appropriate datasets, as large-scale inventory data have a wide coverage, but by definition do not allow for the assessment of spatial interactions.
+
+Exercises like the one presented here, where the models are operated in "blind flight" mode, i.e., without the possibility of being tuned towards capturing the expected patterns, should be repeated.
+Such benchmarking exercises should next focus on aspects such as specific model traits and the ecological formulations of particular (sub-)processes, to better understand the implications of the assumptions on which the models are based.
+Furthermore, the inclusion of a wide range of models with different (1) scales, (2) approaches for capturing population structure, (3) tree regeneration modules and (4) complexity of their formulations will ensure a large benefit to the entire modeling community and beyond.
 
@@
 
-Models of forest dynamics are important tools in science and decision
-support, and the formulation of tree regeneration has strong
-implications for simulated forest properties. The 15 models and variants
-used here are facing similar challenges in their representation of tree
-regeneration: they generally overestimate tree recruitment levels, and
-the simulated regeneration niche is not always captured accurately as a
-function of biotic (light) and abiotic (temperature and moisture)
-factors.
-
-However, most models properly capture the diversity of the initial tree
-community, and differences between model formulations, e.g., the
-presence or absence of feedback from the adult trees did not have a
-strong effect on capturing the species composition of regeneration.
-
-Regarding mortality in the early phase of tree life, many models that
-feature a particularly high overestimation of recruitment levels are
-compensating for this by a larger tree mortality. Often, this
-compensation is not sufficient to reduce the high recruitment levels to
-realistic values. Overall, there is no clear mortality pattern across
-all models.
-
-When capturing tree regeneration, the specific design decisions taken in
-the development of any model are more important for its behavior
-(accuracy) than scale (stand, landscape, global), modelling approach
-(empirical vs. process-based), and complexity. Having both empirical and
-process-based models in our set, the empirically-based models could have
-been expected to have a better performance, as they were calibrated with
-inventory data, but this was not the case. Similarly, higher model
-complexity does not represent an improvement for capturing tree
-regeneration.
-
-Even though the regeneration routines of most of the models investigated
-here have never been constrained well by robust data, their projections
-of recruitment are not overly off. This indicates that a lot can be
-gained by a focus on the modeling of regeneration processes. The
-representation of forest dynamics in these models would become much more
-robust particularly in the face of climate change and post-disturbance
-dynamics, thus strongly reducing the uncertainty in long-term
-projections of future forest dynamics.
+Models of forest dynamics are important tools in science and decision support, and the formulation of tree regeneration has strong implications for simulated forest properties.
+The 15 models and variants used here are facing similar challenges in their representation of tree regeneration: they generally overestimate tree recruitment levels, and the simulated regeneration niche is not always captured accurately as a function of biotic (light) and abiotic (temperature and moisture) factors.
+
+However, most models properly capture the diversity of the initial tree community, and differences between model formulations, e.g., the presence or absence of feedback from the adult trees did not have a strong effect on capturing the species composition of regeneration.
+
+Regarding mortality in the early phase of tree life, many models that feature a particularly high overestimation of recruitment levels are compensating for this by a larger tree mortality.
+Often, this compensation is not sufficient to reduce the high recruitment levels to realistic values.
+Overall, there is no clear mortality pattern across all models.
+
+When capturing tree regeneration, the specific design decisions taken in the development of any model are more important for its behavior (accuracy) than scale (stand, landscape, global), modelling approach (empirical vs.
+process-based), and complexity.
+Having both empirical and process-based models in our set, the empirically-based models could have been expected to have a better performance, as they were calibrated with inventory data, but this was not the case.
+Similarly, higher model complexity does not represent an improvement for capturing tree regeneration.
+
+Even though the regeneration routines of most of the models investigated here have never been constrained well by robust data, their projections of recruitment are not overly off.
+This indicates that a lot can be gained by a focus on the modeling of regeneration processes.
+The representation of forest dynamics in these models would become much more robust particularly in the face of climate change and post-disturbance dynamics, thus strongly reducing the uncertainty in long-term projections of future forest dynamics.
 
@@
 
-This paper arises from a workshop held in Davos (Switzerland) in June
-2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
-Funding for the workshop was provided by COST Action CA19139 PROCLIAS
-(PROcess-based models for CLimate Impact Attribution across Sectors),
-supported by COST (European Cooperation in Science and Technology;
-[https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research
-Grant ETH-35 18-1). ODY was funded by the Swiss National Science
-Foundation project 200020_188882 / 1. KB received support from the
-European Research Council under the European Union's Horizon 2020
-research and innovation program (Grant Agreement 101001905). JB
-acknowledges funding from the long-term research development project no.
-RVO 67985939 from the Czech Academy of Sciences. MM acknowledges
-financial support from I-Maestro (Innovative forest management
-strategies for a resilient bioeconomy under climate change and
-disturbances, grant no. 22035418, 2019--2022) funded by the ERA-NET
-Cofund ForestValue. MMina acknowledges funding from the H2020-MSCA-IF
-project REINFORCE (Grant No. 891671). KM received funding from the
-Project No. CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant
-no. ITMS2014+313011W580 supported by the Integrated Infrastructure
-Operational Programme funded by the ERDF. CH acknowledges financial
-support from Waldfonds WASIM2100 project. We would also like to thank
-all the people who helped to make this study possible with contributions
-at different stages particularly Martin Gutsch and Giorgio Vacchiano.
+This paper arises from a workshop held in Davos (Switzerland) in June 2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
+Funding for the workshop was provided by COST Action CA19139 PROCLIAS (PROcess-based models for CLimate Impact Attribution across Sectors), supported by COST (European Cooperation in Science and Technology; [https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research Grant ETH-35 18-1).
+ODY was funded by the Swiss National Science Foundation project 200020_188882 / 1.
+KB received support from the European Research Council under the European Union's Horizon 2020 research and innovation program (Grant Agreement 101001905).
+JB acknowledges funding from the long-term research development project no.
+RVO 67985939 from the Czech Academy of Sciences.
+MM acknowledges financial support from I-Maestro (Innovative forest management strategies for a resilient bioeconomy under climate change and disturbances, grant no.
+22035418, 2019--2022) funded by the ERA-NET Cofund ForestValue.
+MMina acknowledges funding from the H2020-MSCA-IF project REINFORCE (Grant No.
+891671).
+KM received funding from the Project No.
+CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant no.
+ITMS2014+313011W580 supported by the Integrated Infrastructure Operational Programme funded by the ERDF.
+CH acknowledges financial support from Waldfonds WASIM2100 project.
+We would also like to thank all the people who helped to make this study possible with contributions at different stages particularly Martin Gutsch and Giorgio Vacchiano.
 
@@
 
-O. Díaz-Yáñez and Y. Käber contributed equally and share the first
-authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study
-design. Y. Käber and O. Díaz-Yáñez prepared the input data used by all
-modeling teams. All authors contributed by conducting the simulations
-with their respective models. O. Díaz-Yáñez led the analysis of the
-simulation results with contributions during the workshop from Y. Käber,
-T. Anders, K. H. Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T.
-Hickler, H. Lischke, M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X.
-Morin, W. Rammer, and D. Scherrer. O. Díaz-Yáñez and H. Bugmann led the
-writing of the manuscript and provided the first draft. All authors
-participated in the revision of the manuscript and approved its
-submission.
+O.
+Díaz-Yáñez and Y.
+Käber contributed equally and share the first authorship.
+H.
+Bugmann, Y.
+Käber, and O.
+Díaz-Yáñez developed the study design.
+Y.
+Käber and O.
+Díaz-Yáñez prepared the input data used by all modeling teams.
+All authors contributed by conducting the simulations with their respective models.
+O.
+Díaz-Yáñez led the analysis of the simulation results with contributions during the workshop from Y.
+Käber, T.
+Anders, K.
+H.
+Braziunas, J.
+Brůna, S.
+M.
+Fischer, J.
+Hetzer, T.
+Hickler, H.
+Lischke, M.Mahnken, P.
+Mairota, K.
+Merganičová, T.
+Mette, X.
+Morin, W.
+Rammer, and D.
+Scherrer.
+O.
+Díaz-Yáñez and H.
+Bugmann led the writing of the manuscript and provided the first draft.
+All authors participated in the revision of the manuscript and approved its submission.
 
@@
 
-**Table 1.** Models included in this study, their characteristics and
-simulation strategies used. Scale refers to the models scale, Type to
-the models type, Population structure to the model population approach,
-Feedback to the inclusion or not of adult tree feedback to the level and
-species composition of regeneration, Approach to the model regeneration
-module approach, Start from to the stage from which trees are recruited,
-Complexity, to the mean regeneration formulation as estimated in
-(Bugmann & Seidl, 2022), Species to which species from the original
-eleven species were simulated, Runtime (spin up) [years] to the number
-of years used per model in their simulations for the spin up period and
-regeneration sampling; Climate data type (use of years) to the type of
-data used in the simulations and how it was used, and Reference provides
-the link to the publication with more model details. Further information
-can be found on the individual model reports available in the Appendix
-S1: Model reports.
+**Table 1.** Models included in this study, their characteristics and simulation strategies used.
+Scale refers to the models scale, Type to the models type, Population structure to the model population approach, Feedback to the inclusion or not of adult tree feedback to the level and species composition of regeneration, Approach to the model regeneration module approach, Start from to the stage from which trees are recruited, Complexity, to the mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), Species to which species from the original eleven species were simulated, Runtime (spin up) [years] to the number of years used per model in their simulations for the spin up period and regeneration sampling; Climate data type (use of years) to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details.
+Further information can be found on the individual model reports available in the Appendix S1: Model reports.
 
@@
 
-**Figure 1.** The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used
-in the present study, the colors show the dominant tree species per site
-in the observed regeneration at the 10 cm threshold.
-
-**Figure 2.** Mean regeneration levels across all samples per site,
-plotted for the 200 sites and for each model. The red dashed lines shows
-the 25th and 75th percentiles for the 7 cm diameter threshold in the
-observed data. There are two boxplots for each model where the left
-lighter boxplot corresponds to 7 cm and the right darker boxplot to 10
-cm
-
-**Figure 3.** Shannon index of tree regeneration, calculated by basal
-area, as the mean value across the 200 samples per site for the diameter
-thresholds of 7 and 10 cm, respectively. The red dashed lines shows the
-25th and 75th percentiles for the 7 cm diameter threshold in the
-observed data.
-
-**Figure 4.** Mean Shannon index across all samples per site for
-observed and simulated data. Each plot shows one pattern represented by
-one exemplary model of each category (overpredicted, intermediate and
-underpredicted). The full data with the grouping of the models are shown
-in Figures S3 and S4. n indicates the number of models falling in each
-group. A) Examples for the three trends across models for species
-diversity at the stand level. B) Examples for the three trends across
-models for regeneration at the 7 cm threshold (165 sites).
-
-**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm
-thresholds. Dashed blue lines mark a ratio equal to 1, indicating no
-decrease in tree regeneration between 7 and 10 cm, and a ratio equal to
-1.77, corresponding to the Reineke self-thinning ratio under even-aged
-conditions.
-
-**Figure 6.** Ratio of tree regeneration between the 7 and 10 cm
-thresholds (regeneration 7 cm / regeneration 10 cm) and the
-overestimation proportion at 7 cm ([simulated - observed] / observed)
-for the mean regeneration per model across sites and samples. The
-horizontal dashed lines show a ratio equal to 1, indicating no decrease
-of regeneration between 7 and 10 cm, and a ratio equal to 1.77
-corresponding to the Reineke self-thinning line under even-aged
-conditions. The vertical line indicates zero overestimation.
-
-**Figure 7.** Mean recruitment across the 200 samples per site with the
-y-axis scaled differently by model, for the 200 sites against gradients
-of (a) total basal area; (b) climatic water balance; (c) seasonal
-degree-day sum. The values were split into ten bins; the red lines
-represent a GAM model showing the trend in the observed data. Y-axis are
-not equally scaled.
-
-**Figure 8.** Regeneration levels of the five main species in terms of
-their basal area share in the observations (top row) and the performance
-of each model across the environmental gradients (other rows) for the 7
-cm regeneration threshold. The values shown are the mean of the 200
-samples per site and across the sites in each bin (tile), with ten bins
-per gradient. The sizes of the circles represent the ratio between the
-regeneration basal area of the species and the total regeneration basal
-area for all species. The absence of a circle indicates a zero basal
-area share in the regeneration, or the absence of regeneration
-altogether. The color gradient (for the models only) shows the
-difference between the simulated and observed ratio of regeneration
-basal area of the species and the total regeneration basal area for all
-the species.
+**Figure 1.** The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed regeneration at the 10 cm threshold.
+
+**Figure 2.** Mean regeneration levels across all samples per site, plotted for the 200 sites and for each model.
+The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
+There are two boxplots for each model where the left lighter boxplot corresponds to 7 cm and the right darker boxplot to 10 cm
+
+**Figure 3.** Shannon index of tree regeneration, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
+The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
+
+**Figure 4.** Mean Shannon index across all samples per site for observed and simulated data.
+Each plot shows one pattern represented by one exemplary model of each category (overpredicted, intermediate and underpredicted).
+The full data with the grouping of the models are shown in Figures S3 and S4.
+n indicates the number of models falling in each group.
+A) Examples for the three trends across models for species diversity at the stand level.
+B) Examples for the three trends across models for regeneration at the 7 cm threshold (165 sites).
+
+**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm thresholds.
+Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree regeneration between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.
+
+**Figure 6.** Ratio of tree regeneration between the 7 and 10 cm thresholds (regeneration 7 cm / regeneration 10 cm) and the overestimation proportion at 7 cm ([simulated - observed] / observed) for the mean regeneration per model across sites and samples.
+The horizontal dashed lines show a ratio equal to 1, indicating no decrease of regeneration between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
+The vertical line indicates zero overestimation.
+
+**Figure 7.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum.
+The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
+Y-axis are not equally scaled.
+
+**Figure 8.** Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold.
+The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient.
+The sizes of the circles represent the ratio between the regeneration basal area of the species and the total regeneration basal area for all species.
+The absence of a circle indicates a zero basal area share in the regeneration, or the absence of regeneration altogether.
+The color gradient (for the models only) shows the difference between the simulated and observed ratio of regeneration basal area of the species and the total regeneration basal area for all the species.
 

### 2024-02-12 manuscript/review/editorsLetter.qmd
last comment not pushc from the editors revision

@@
 
-In this revised version, we have taken all the reviewer\'s comments and suggestions into consideration and have made significant changes to the manuscript accordingly.
+In this revised version, we have taken all the reviewer's comments and suggestions into consideration and have made significant changes to the manuscript accordingly.
 We have submitted a detailed report outlining the actions we have taken in response to each of the reviewer's comments.

### 2024-02-12 manuscript/review/editorsLetter.qmd

--- 2023-10-31 manuscript/review/editorsLetter.qmd
@@
 
-I would like to express my sincere gratitude for taking the time to review our manuscript. Your insightful feedback on our findings was immensely helpful.
+Thank you for taking the time to review our manuscript and for your valuable feedback.
 
@@
 
-Thank you again for your time and effort. We look forward to hearing from you soon.
+Thank you again for your time. We look forward to hearing from you soon.
 

### 2024-02-12 code/figures.R
comment5- use black outlines for the underlying map

@@ -42,7 +42,7 @@
 
 # Data preparation ----    
     
-    ## Recruitment Shannon index  by basal area ----
+    ## Recruitment Shannon index by basal area ----
     ShannonDF <- outputsDF 
     ShannonDF$giDIVG <- ShannonDF$r.ba / ShannonDF$Totr.ba
     ShannonDF$lngiDIVG <- log(ShannonDF$r.ba / ShannonDF$Totr.ba)
@@ -1479,7 +1479,7 @@
     # Europe <- world[which(world$continent == "Europe"), ]
     # plotsMap <- ggplot(Europe) +
     #   ggthemes::theme_map() +
-    #   geom_sf(fill = "transparent" , colour = "lightgrey",  lwd = 0.25) +
+    #   geom_sf(fill = "transparent" , colour = "black",  lwd = 0.25) +
     #   coord_sf(xlim = c(-15, 30), ylim = c(40,55), expand = FALSE) +
     #   geom_point(data = domSpp10, shape = 20, stroke = FALSE,
     #              mapping = aes(x = X, y = Y, color = recruitment),
@@ -1510,4 +1510,4 @@
     #                 plot =   plotsMap,
     #                 width = 21, height = 12, scale = 0.9,
     #                 dpi = 300, units = "cm", device = 'jpg')
-
+    # 

### 2024-02-12 manuscript/manuscriptRW.Rmd
comment 6. include explanation of the boxplot fetures in figure 2,3, and 5

@@
 The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
-There are two boxplots for each model where the left lighter boxplot corresponds to 7 cm and the right darker boxplot to 10 cm
+There are two boxplots for each model where the left lighter boxplot corresponds to 7 cm and the right darker boxplot to 10 cm.
+The boxplot shows the median (midline), interquartile range (box limits), largest and smallest value within 1.5 times interquartile range (whiskers), larger values are represented as outliers.
 
@@
 The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
+Boxplot features are the same as in Figure 2.
 
@@
 **Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm thresholds.
-Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree regeneration between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.
+Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree regeneration between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.Boxplot features are the same as in Figure 2.
 

### 2024-02-12 manuscript/manuscriptRW.Rmd
comment 7: figure 7 is too big, create separate figures for each pannel and change references accordingly in the manuscript

@@
 
-In the observed data, recruitment levels decreased clearly with increasing stand basal area (Figure 7a).
+In the observed data, recruitment levels decreased clearly with increasing stand basal area (Figure 7).
 This trend was captured by the Landis-II model only, albeit at much lower values of basal area than in the observations.
@@
 The observed recruitment did not change much across the climatic water balance gradient, and showed a slight increase with temperature.
-Neither did the models generally match the observed recruitment patterns across these gradients (Figure 7b,c), nor was there a clear pattern across models.
+Neither did the models generally match the observed recruitment patterns across these gradients (Figure 8,9), nor was there a clear pattern across models.
 Regarding the soil moisture gradient, a group of models featured decreasing recruitment with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS).
@@
 
-**Figure 7.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum.
+**Figure 7.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of total basal area.
 The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
-Y-axis are not equally scaled.
 
-**Figure 8.** Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold.
+**Figure 8.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of climatic water balance.
+The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
+
+**Figure 9.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of seasonal degree-day sum.
+The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
+
+**Figure 10.** Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold.
 The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient.

### 2024-02-12 code/figures.R

@@ -947,7 +947,7 @@
     
     
     #### Fig. R VS BA----
-    png(file = "figures/envTrend7Totba.png", width = 16, height = 12, 
+    jpeg(file = "figures/Figure7.jpg", width = 16, height = 12, 
         units = "cm", res = 300)
     par(mfrow = c(4, 4), mar = c(2, 3, 1, 1), oma = c(2, 2, 1, 1))
     
@@ -1074,7 +1074,7 @@
     
     #### FigS. R VS WB -----
     
-    png(file = "figures/envTrend7WB.png", width = 16, height = 12, 
+    jpeg(file = "figures/Figure8.jpg", width = 16, height = 12, 
         units = "cm", res = 300)
     par(mfrow = c(4, 4), mar = c(2, 3, 1, 1), oma = c(2, 2, 1, 1))
     
@@ -1178,7 +1178,7 @@
     
     #### FigS. R VS. DDS -----
     
-    png(file = "figures/envTrend7DDS.png", width = 16, height = 12, 
+    jpeg(file = "figures/Figure9.jpg", width = 16, height = 12, 
         units = "cm", res = 300)
     par(mfrow = c(4, 4), mar = c(2, 3, 1, 1), oma = c(2, 2, 1, 1))
     
@@ -1199,6 +1199,10 @@
     
     for (modelSel in unique(envTrend$model)) {
       ienvTrendModel <- envTrend[envTrend$model == modelSel,]
+      
+      axisPlot <- ifelse(modelSel %in% c("Landis II", "TreeMig", "LPJ-GUESS",
+                                         "aDGVM2"), "s", "n")
+      
       dataSel <- ienvTrendModel[ienvTrendModel$dbh == 7 & ienvTrendModel$model == modelSel, ]
       boxplot(r.trees ~ dds_cut, 
               data = dataSel, 
@@ -1224,6 +1228,7 @@
     
     dev.off()
     
+
     
     png(file = "figures/envTrend7DDS_500.png", width = 16, height = 12, 
         units = "cm", res = 300)
@@ -1401,7 +1406,7 @@
                       dpi = 300, units = "cm", device = 'png') 
       
       if(dbhSel == 7) {
-        ggplot2::ggsave("figures/Figures8.jpg",
+        ggplot2::ggsave("figures/Figure10.jpg",
                         plot = heatCircle,
                         width = 27, height = 60, scale = 0.9,
                         dpi = 300, units = "cm", device = 'jpg') 

### 2024-02-12 manuscript/manuscriptRW.Rmd
comment8: letteri is too small, revise the gfigure

@@
 
-First, some models overestimated recruitment across the entire gradients of at least one of the five main tree species (Figure 8, e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated recruitment of more than one species for a large part of the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
-red colors in Figure 8).
+First, some models overestimated recruitment across the entire gradients of at least one of the five main tree species (Figure 10, e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated recruitment of more than one species for a large part of the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
+red colors in Figure 10).
 These trends were consistent for both recruitment thresholds, although the exact changes in the share of the recruited basal area were different (Appendix S1: Figure S8).
@@
 
-Lastly, there is no model that performed well across the five species regarding the species-specific recruitment levels in the environmental space (Figure 8).
+Lastly, there is no model that performed well across the five species regarding the species-specific recruitment levels in the environmental space (Figure 10).
 Some models (e.g., ForClim11, PICUS) tended to consistently overestimate the recruitment share of all five species, but most models overestimated the recruitment share of some species while underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).

### 2024-02-12 code/figures.R

@@ -1400,15 +1400,141 @@
         ggplot2::labs(size = expression(bar(R) * " BA share")) 
       
       
-      ggplot2::ggsave(paste0("figures/HeatEnvGrad_", binSize - 1, "circles_",dbhSel, ".png"),
+      ggplot2::ggsave(paste0("figures/HeatEnvGrad_", binSize - 1, "circles_", 
+                             dbhSel, ".png"),
                       plot = heatCircle,
                       width = 27, height = 60, scale = 0.9,
                       dpi = 300, units = "cm", device = 'png') 
       
+      #Divide the figure in two panels for publication
+      panel1models <-  c( "Observed", "4C", "ForCEEPS", "ForCEEPS(f)",  "FORMIND", 
+                          "ForClim 1",  "ForClim 11", "SIBYLA")
+      panel2models <-  c("Observed","xComp", "PICUS", "iLand", "LandClim", "Landis II", 
+                         "TreeMig", "LPJ-GUESS") 
+      
+      heatCircle1 <- ggplot2::ggplot(data = mainSppBAshareBinsMeanDiff[mainSppBAshareBinsMeanDiff$model %in% panel1models,],
+                                    mapping = ggplot2::aes(x = dds_bin,
+                                                           y = wb_bin,
+                                                           fill = Diff)) +
+        ggplot2::scale_fill_gradient2(low = "blue", mid = 'white', high = "red",
+                                      breaks = c(-1, -0.5, 0, 0.5, 1),
+                                      midpoint = 0, 
+                                      name = expression(bar(R) * " BA share difference")) +
+        
+        ggplot2::geom_tile() +
+        ggplot2::xlab(label = "") +
+        ggplot2::ylab(label = "") +
+        ggplot2::facet_grid(model ~ species,  
+                            labeller = labeller(species = labelsSpecies)) +
+        ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90),
+                       strip.text.y = ggplot2::element_text(angle = 0),
+                       strip.text.x = ggplot2::element_text(size = 6), 
+                       panel.grid.major = ggplot2::element_blank(), 
+                       panel.grid.minor = ggplot2::element_blank(),
+                       panel.background = ggplot2::element_rect(fill = "lightgrey",
+                                                                colour = "lightblue"),
+                       axis.line = ggplot2::element_line(colour = "black"),
+                       strip.background =  ggplot2::element_blank(),
[…]

### 2024-02-12 code/plottingManuscript.R

@@ -62,6 +62,20 @@
                 "Treemig" = "TreeMig",
                 "aDGVM2" = "aDGVM2", 
                 "LPJ-GUESS" = "LPJ-GUESS")
+    
+# Labels per species -----
+
+    labelsSpecies <- c("Abies alba" = "A. alba" ,
+                       "Acer pseudoplatanus" = "A. pseudoplatanus",
+                       "Alnus glutinosa" = "A. glutinosa",
+                       "Betula spp." = "Betula spp.",
+                       "Carpinus betulus" = "C. betulus",
+                       "Fagus sylvatica" = "F. sylvatica",
+                       "Fraxinus excelsior" = "F. excelsior",
+                       "Picea abies" = "P. abies",
+                       "Pinus sylvestris" = "P. sylvestris",
+                       "Quercus spp." =  "Quercus spp.",
+                       "Tilia cordata" = "T.cordata")    
 
 # Labels figures  ----
 

### 2024-02-12 manuscript/manuscriptRW.Rmd
comment9: add to figure captions all abreviations from figures

@@
 
-Tree regeneration is a key process in forest dynamics, particularly in the context of forest resilience and climate change.
-Models are pivotal for assessing long-term forest dynamics, and they have been in use for more than 50 years.
-However, there is a need to evaluate their capacity to accurately represent tree regeneration.
-We assess how well current models capture the overall abundance, species composition, and mortality of tree regeneration.
-
-Using 15 models built to capture long-term forest dynamics at the stand, landscape, and global levels, we simulate tree regeneration at 200 sites representing large environmental gradients across Central Europe.
-The results are evaluated against extensive data from unmanaged forests.
-
-Most of the models overestimate recruitment levels, which is compensated only in some models by high simulated mortality rates in the early stages of individual tree dynamics.
-Simulated species diversity of recruitment generally matches observed ranges.
-Models simulating higher stand-level species diversity do not feature higher species diversity in the recruitment layer.
-The effect of light availability on recruitment levels is captured better than the effects of temperature and soil moisture, but patterns are not consistent across models.
-Increasing complexity in the tree regeneration modules is not related to higher accuracy of simulated tree recruitment.
-Furthermore, individual model design is more important than scale (stand, landscape, global) and approach (empirical, process-based) for accurately capturing tree regeneration.
-
-Despite the mismatches between simulation results and data, it is remarkable that most models capture the essential features of the highly complex process of tree regeneration, while not having been parameterized with such data.
-We conclude that much can be gained by evaluating and refining the modeling of tree regeneration processes.
-This has the potential to render long-term projections of forest dynamics under changing environmental conditions much more robust.
+Tree regeneration is a key process in forest dynamics, particularly in
+the context of forest resilience and climate change. Models are pivotal
+for assessing long-term forest dynamics, and they have been in use for
+more than 50 years. However, there is a need to evaluate their capacity
+to accurately represent tree regeneration. We assess how well current
+models capture the overall abundance, species composition, and mortality
+of tree regeneration.
+
+Using 15 models built to capture long-term forest dynamics at the stand,
+landscape, and global levels, we simulate tree regeneration at 200 sites
+representing large environmental gradients across Central Europe. The
+results are evaluated against extensive data from unmanaged forests.
+
+Most of the models overestimate recruitment levels, which is compensated
+only in some models by high simulated mortality rates in the early
+stages of individual tree dynamics. Simulated species diversity of
+recruitment generally matches observed ranges. Models simulating higher
+stand-level species diversity do not feature higher species diversity in
+the recruitment layer. The effect of light availability on recruitment
+levels is captured better than the effects of temperature and soil
+moisture, but patterns are not consistent across models. Increasing
+complexity in the tree regeneration modules is not related to higher
+accuracy of simulated tree recruitment. Furthermore, individual model
+design is more important than scale (stand, landscape, global) and
+approach (empirical, process-based) for accurately capturing tree
+regeneration.
+
+Despite the mismatches between simulation results and data, it is
+remarkable that most models capture the essential features of the highly
+complex process of tree regeneration, while not having been
+parameterized with such data. We conclude that much can be gained by
+evaluating and refining the modeling of tree regeneration processes.
+This has the potential to render long-term projections of forest
+dynamics under changing environmental conditions much more robust.
 
@@
 
-Forests provide a wide range of products and services of vital importance to humankind [@foresteurope2020].
-Under the pressure of climate change, increasing disturbance impacts and changing societal demands on forest ecosystem services, it is becoming ever more important to understand how forest structure, composition and function will change, and to evaluate forest capacity to adapt to or be resilient after disturbance [@lindner2010; @seidl2022].
-A wide range of models of forest dynamics have been developed over the past decades considering the impacts of climate [@bugmann2022].
-From these studies, it is evident that we have a reasonably good understanding how to model tree growth [@vanclay1997; @bugmann1996], and substantial efforts have been dedicated to improving the representation of tree mortality [@bugmann2019; @cailleret2017].
-In contrast, tree regeneration is much less studied and is often represented rather coarsely in models [@walck2011; @leishman1992; @price2001].
-This presents an important research gap, particularly in the context of climate-induced forest disturbances and forest resilience.
-
-Tree regeneration arises from multiple processes including pollination, fruit maturation, seed production, dispersal, germination, juvenile growth and survival [@price2001; @vacchiano2018].
-All these processes are difficult to assess, and some of them are scarcely understood and thus appear highly stochastic [@bogdziewicz2021].
-Identifying the appropriate level of complexity for the mathematical formulation of the key factors that are leading to successful tree regeneration is challenging.
-Currently, tree regeneration processes in dynamic forest models are handled in a multitude of ways [@konig_tree_2022; @bugmann2022]: from 1) entirely ignoring them [as done in classical forest growth models, e.g., Pretzsch et al.
--@pretzsch2002], across 2) the use of a few simple environmental filters, as done in most forest "gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g., @hickler2012; @smith2001], to 3) complex approaches that incorporate local feedback from the canopy, multiple ecological processes and often also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based statistical parameterizations, which however are not easy to extrapolate in space and time [e.g., @ribbens1994].
-
-Overall, models are needed to (1) synthesize existing empirical data and explore their relationships, (2) assess future tree regeneration e.g.
-in the context of global change scenarios, and (3) identify the most important processes that are shaping vegetation patterns.
-Given the strategies that are used in models of forest dynamics to represent tree regeneration, their behavior often is prone to problems, such as very high levels of tree regeneration that necessitate excess mortality at early stages of tree life to simulate correct stand structure and composition [@kroiss2015].
-Also, inadequately high species diversity in tree regeneration may be simulated, which is characteristic of "classical" forest gap models [@gutiérrez2016], at least as long as the simulation set-up comprises a multitude of species.
-Some models use calibration against local canopy-level data to constrain simulated tree regeneration, which is likely to hamper the general applicability of these models e.g.
-under scenarios of climate change.
-Furthermore, correctly capturing the species composition of tree regeneration as a function of the presence of seed trees in the canopy is often a particular challenge, potentially leading to unrealistic successional drift in the model, which must be corrected by factors that are hard to parameterize [e.g., @lischke2006].
-A related issue is the excessive reduction of species diversity due to positive feedback effects, such that eventually just single-species stands remain [@meier2011].
-This is sometimes corrected by the incorporation of a low level of seed influx of all species at all times [@schumacher2006] or by restricting the number of seeds per species in the seedbank [@lischke2006].
-However, simulated species composition is usually exceedingly sensitive to assumptions about seed availability, whereas the parameters of such functions are poorly constrained by field data.
-Lastly, there is often a problem with insufficient observational constraints on parameter values for models that start from very small tree sizes (e.g., 10 cm in height, or even from seed) and track tree development in a process-oriented manner by considering a multitude of ecological influences, rather than emphasizing tree recruitment into a larger size class [e.g., @wehrli2006].
-
-Modeling tree regeneration processes is challenging, and even in empirical ecology, it has not received much attention [@hanbury-brown2022], although many sub-processes have been studied in detail [@miina2006].
-Yet, little data are available that cover all the processes within one species along environmental gradients, let alone for a vast suite of species.
-Data on forest regeneration are often fragmented, which constitutes a major problem for model building [@clark1999a].
-For example, monitoring on permanent plots (such as National Forest Inventories) often measures the ingrowth of new trees into a specific size class [@zell2019], however with a design that captures the rate after the stand initiation phase [@hallsby2015].
-Data from permanent plots, although available across regions, are also highly heterogeneous, rendering their use difficult in a modeling context [@käber2021].
-Lastly, targeted studies to measure tree regeneration on experimental sites are often limited in spatial extent, species studied, or the sub-set of processes that are investigated [@berdanier2016; @collet2006].
-
-Thus, a focus on the modeling of tree regeneration processes is sorely needed and overdue [@price2001; @walck2011] if we are to make reliable projections of future forest dynamics, i.e., when the models need to be operated in extrapolation mode, as well as from a fundamental ecological point of view for increasing systems understanding.
-In the present study, for the first time, a large number of forest models commonly used to assess forest dynamics under climate change are evaluated against a continental-scale, multi-species harmonized dataset on tree recruitment [@käber2023].
-By tree recruitment, we refer to the passing of trees across a specific diameter threshold ("ingrowth").
-We included models that are based on a range of "philosophies", from models operating at the stand to the global scale as well as the range of models from empirically derived to "process-based" [e.g., @bugmann1996; @fabrika2012; @lexer2001; @reyer2014].
-Due to the large variability in tree regeneration patterns in nature and the large number of factors driving and constraining this process --- including some that are not incorporated explicitly in most models, such as deer browsing --- we do not aim for a detailed statistical evaluation of each model.
-Instead, we aim to evaluate the general recruitment patterns and magnitudes simulated by the models, and benchmark the simulated regeneration niche of multiple species against empirical data along a wide environmental gradient of temperature, moisture, and light availability [@grubb1977].
-
-More specifically, we aim to answer the following questions: (1) Are models of forest dynamics capturing accurately tree recruitment levels, initial tree species diversity and mortality in the recruitment?
-(2) Do model traits explain differences in model performance?
-(3) How well do the models capture total recruitment and the regeneration niches of individual species across environmental gradients of light availability, temperature, and soil moisture?
-
-We evaluate the performance of the models by analyzing simulated data alongside observed data.
-Specifically, we examine tree species diversity, mortality rates in tree recruitment, and recruitment levels.
-Then, we link model traits (e.g., complexity, scale) to performance, and we assess whether modeled total recruitment or individual species regeneration niches (i.e., proportional recruitment along an environmental gradient) align with observational data.
+Forests provide a wide range of products and services of vital
+importance to humankind [@foresteurope2020]. Under the pressure of
+climate change, increasing disturbance impacts and changing societal
+demands on forest ecosystem services, it is becoming ever more important
+to understand how forest structure, composition and function will
+change, and to evaluate forest capacity to adapt to or be resilient
+after disturbance [@lindner2010; @seidl2022]. A wide range of models of
+forest dynamics have been developed over the past decades considering
+the impacts of climate [@bugmann2022]. From these studies, it is evident
+that we have a reasonably good understanding how to model tree growth
+[@vanclay1997; @bugmann1996], and substantial efforts have been
+dedicated to improving the representation of tree mortality
+[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is much
+less studied and is often represented rather coarsely in models
+[@walck2011; @leishman1992; @price2001]. This presents an important
+research gap, particularly in the context of climate-induced forest
+disturbances and forest resilience.
+
+Tree regeneration arises from multiple processes including pollination,
+fruit maturation, seed production, dispersal, germination, juvenile
+growth and survival [@price2001; @vacchiano2018]. All these processes
+are difficult to assess, and some of them are scarcely understood and
+thus appear highly stochastic [@bogdziewicz2021]. Identifying the
+appropriate level of complexity for the mathematical formulation of the
+key factors that are leading to successful tree regeneration is
+challenging. Currently, tree regeneration processes in dynamic forest
+models are handled in a multitude of ways [@konig_tree_2022;
+@bugmann2022]: from 1) entirely ignoring them [as done in classical
+forest growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2)
+the use of a few simple environmental filters, as done in most forest
+"gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
+@hickler2012; @smith2001], to 3) complex approaches that incorporate
+local feedback from the canopy, multiple ecological processes and often
+also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
+statistical parameterizations, which however are not easy to extrapolate
+in space and time [e.g., @ribbens1994].
+
+Overall, models are needed to (1) synthesize existing empirical data and
+explore their relationships, (2) assess future tree regeneration e.g. in
+the context of global change scenarios, and (3) identify the most
+important processes that are shaping vegetation patterns. Given the
+strategies that are used in models of forest dynamics to represent tree
+regeneration, their behavior often is prone to problems, such as very
+high levels of tree regeneration that necessitate excess mortality at
+early stages of tree life to simulate correct stand structure and
+composition [@kroiss2015]. Also, inadequately high species diversity in
+tree regeneration may be simulated, which is characteristic of
+"classical" forest gap models [@gutiérrez2016], at least as long as the
+simulation set-up comprises a multitude of species. Some models use
+calibration against local canopy-level data to constrain simulated tree
+regeneration, which is likely to hamper the general applicability of
+these models e.g. under scenarios of climate change. Furthermore,
+correctly capturing the species composition of tree regeneration as a
+function of the presence of seed trees in the canopy is often a
+particular challenge, potentially leading to unrealistic successional
+drift in the model, which must be corrected by factors that are hard to
+parameterize [e.g., @lischke2006]. A related issue is the excessive
+reduction of species diversity due to positive feedback effects, such
+that eventually just single-species stands remain [@meier2011]. This is
+sometimes corrected by the incorporation of a low level of seed influx
+of all species at all times [@schumacher2006] or by restricting the
+number of seeds per species in the seedbank [@lischke2006]. However,
+simulated species composition is usually exceedingly sensitive to
+assumptions about seed availability, whereas the parameters of such
+functions are poorly constrained by field data. Lastly, there is often a
+problem with insufficient observational constraints on parameter values
+for models that start from very small tree sizes (e.g., 10 cm in height,
+or even from seed) and track tree development in a process-oriented
+manner by considering a multitude of ecological influences, rather than
+emphasizing tree recruitment into a larger size class [e.g.,
+@wehrli2006].
+
+Modeling tree regeneration processes is challenging, and even in
+empirical ecology, it has not received much attention
+[@hanbury-brown2022], although many sub-processes have been studied in
+detail [@miina2006]. Yet, little data are available that cover all the
+processes within one species along environmental gradients, let alone
+for a vast suite of species. Data on forest regeneration are often
+fragmented, which constitutes a major problem for model building
+[@clark1999a]. For example, monitoring on permanent plots (such as
+National Forest Inventories) often measures the ingrowth of new trees
+into a specific size class [@zell2019], however with a design that
+captures the rate after the stand initiation phase [@hallsby2015]. Data
+from permanent plots, although available across regions, are also highly
+heterogeneous, rendering their use difficult in a modeling context
+[@käber2021]. Lastly, targeted studies to measure tree regeneration on
+experimental sites are often limited in spatial extent, species studied,
+or the sub-set of processes that are investigated [@berdanier2016;
+@collet2006].
+
+Thus, a focus on the modeling of tree regeneration processes is sorely
+needed and overdue [@price2001; @walck2011] if we are to make reliable
+projections of future forest dynamics, i.e., when the models need to be
+operated in extrapolation mode, as well as from a fundamental ecological
+point of view for increasing systems understanding. In the present
+study, for the first time, a large number of forest models commonly used
+to assess forest dynamics under climate change are evaluated against a
+continental-scale, multi-species harmonized dataset on tree recruitment
+[@käber2023]. By tree recruitment, we refer to the passing of trees
+across a specific diameter threshold ("ingrowth"). We included models
+that are based on a range of "philosophies", from models operating at
+the stand to the global scale as well as the range of models from
+empirically derived to "process-based" [e.g., @bugmann1996;
+@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
+tree regeneration patterns in nature and the large number of factors
+driving and constraining this process --- including some that are not
+incorporated explicitly in most models, such as deer browsing --- we do
+not aim for a detailed statistical evaluation of each model. Instead, we
+aim to evaluate the general recruitment patterns and magnitudes
+simulated by the models, and benchmark the simulated regeneration niche
+of multiple species against empirical data along a wide environmental
+gradient of temperature, moisture, and light availability [@grubb1977].
+
+More specifically, we aim to answer the following questions: (1) Are
+models of forest dynamics capturing accurately tree recruitment levels,
+initial tree species diversity and mortality in the recruitment? (2) Do
+model traits explain differences in model performance? (3) How well do
+the models capture total recruitment and the regeneration niches of
+individual species across environmental gradients of light availability,
+temperature, and soil moisture?
+
+We evaluate the performance of the models by analyzing simulated data
+alongside observed data. Specifically, we examine tree species
+diversity, mortality rates in tree recruitment, and recruitment levels.
+Then, we link model traits (e.g., complexity, scale) to performance, and
+we assess whether modeled total recruitment or individual species
+regeneration niches (i.e., proportional recruitment along an
+environmental gradient) align with observational data.
 
@@
 
-Fifteen models of forest dynamics (including two models featuring model variants) developed for the stand, landscape or global scale were used to simulate forest dynamics (Table \@ref(tab:tableModels)).
-The approaches used for model construction and their origin differ strongly, with most of the models featuring a largely "process-based" approach, whereas two models are based on formulations derived from the statistical analysis of inventory data (SIBYLA and xComp).
-Some of the models largely rely on the original approach underlying forest gap models (e.g., ForClim 1), whereas some are based on plant ecophysiological processes (e.g., FORMIND and iLand).
-The different approaches underlying the models have strongly influenced the formulation of tree regeneration processes.
-
-The models can be differentiated into 'regeneration' and 'recruitment' approaches [@vanclay1994; @konig_tree_2022].
-*Regeneration models* include processes such as flowering and pollination, seed production, seed dispersal, germination, and seedling growth, which ultimately lead to the simulated number of established trees.
-*Recruitment models*, in contrast, introduce a number of new trees with certain characteristics such as biomass or diameter, without explicitly considering earlier development processes.
-We can further distinguish models that feature a feedback from stand properties to simulated recruitment, i.e., where the level and species composition of recruitment is influenced by the adult tree community via the production of seeds (or seedlings/saplings), from models that do not contain such feedback.
-
-The starting point for tree regeneration in the models differs as well, ranging from seedbank, seed or seedling to sapling (i.e., trees much larger than 10 cm height, often ca.
-2-3 m tall).
-Models that start from seed need to include a larger number of ecological processes such as germination and survival of young seedlings, whereas models that start from saplings have to aggregate via parameterizations several ecological processes that are not treated explicitly.
-This latter approach reduces model complexity but comes at the cost of blurred process representation.
-
-In the models used here, the overall complexity in the regeneration modules varies considerably.
-Following @bugmann2022, we can classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and TreeMig) as having rather high complexity in their regeneration modules (mean regeneration formulation complexity across all processes \>0, Table \@ref(tab:tableModels)), whereas the other eight models feture relatively low complexity.
-Two models, ForClim and ForCEEPS, were used here with two alternative variants of regeneration.
-ForClim variant 1 [@bugmann1996] is based on a regeneration module that adheres closely to the concept introduced by @botkin1972 which is based on environmental filters considering species-specific thresholds of light availability and climatic variables.
-In contrast, ForClim variant 11 is adopting a slightly more complex approach where individual species properties and their relation to the environment are represented gradually along with the relative suitability of different species (i.e., competition among tree species) [@huber2020].
-These two model variants allow us to evaluate a more process-based and complex module (variant 11) against a simple module (variant 1), while the rest of the model structure is identical.
-Similarly, the two ForCEEPS variants allow us to isolate the importance of the canopy feedback (i.e., simulated actual composition and relative abundance of species in the plot) via seed trees for the quantity and quality (e.g., diversity and composition) of simulated regeneration, as one variant includes this feedback whereas the other does not.
+Fifteen models of forest dynamics (including two models featuring model
+variants) developed for the stand, landscape or global scale were used
+to simulate forest dynamics (Table \@ref(tab:tableModels)). The
+approaches used for model construction and their origin differ strongly,
+with most of the models featuring a largely "process-based" approach,
+whereas two models are based on formulations derived from the
+statistical analysis of inventory data (SIBYLA and xComp). Some of the
+models largely rely on the original approach underlying forest gap
+models (e.g., ForClim 1), whereas some are based on plant
+ecophysiological processes (e.g., FORMIND and iLand). The different
+approaches underlying the models have strongly influenced the
+formulation of tree regeneration processes.
+
+The models can be differentiated into 'regeneration' and 'recruitment'
+approaches [@vanclay1994; @konig_tree_2022]. *Regeneration models*
+include processes such as flowering and pollination, seed production,
+seed dispersal, germination, and seedling growth, which ultimately lead
+to the simulated number of established trees. *Recruitment models*, in
+contrast, introduce a number of new trees with certain characteristics
+such as biomass or diameter, without explicitly considering earlier
+development processes. We can further distinguish models that feature a
+feedback from stand properties to simulated recruitment, i.e., where the
+level and species composition of recruitment is influenced by the adult
+tree community via the production of seeds (or seedlings/saplings), from
+models that do not contain such feedback.
+
+The starting point for tree regeneration in the models differs as well,
+ranging from seedbank, seed or seedling to sapling (i.e., trees much
+larger than 10 cm height, often ca. 2-3 m tall). Models that start from
+seed need to include a larger number of ecological processes such as
+germination and survival of young seedlings, whereas models that start
+from saplings have to aggregate via parameterizations several ecological
+processes that are not treated explicitly. This latter approach reduces
+model complexity but comes at the cost of blurred process
+representation.
+
+In the models used here, the overall complexity in the regeneration
+modules varies considerably. Following @bugmann2022, we can classify
+seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and
+TreeMig) as having rather high complexity in their regeneration modules
+(mean regeneration formulation complexity across all processes \>0,
+Table \@ref(tab:tableModels)), whereas the other eight models feture
+relatively low complexity. Two models, ForClim and ForCEEPS, were used
+here with two alternative variants of regeneration. ForClim variant 1
+[@bugmann1996] is based on a regeneration module that adheres closely to
+the concept introduced by @botkin1972 which is based on environmental
+filters considering species-specific thresholds of light availability
+and climatic variables. In contrast, ForClim variant 11 is adopting a
+slightly more complex approach where individual species properties and
+their relation to the environment are represented gradually along with
+the relative suitability of different species (i.e., competition among
+tree species) [@huber2020]. These two model variants allow us to
+evaluate a more process-based and complex module (variant 11) against a
+simple module (variant 1), while the rest of the model structure is
+identical. Similarly, the two ForCEEPS variants allow us to isolate the
+importance of the canopy feedback (i.e., simulated actual composition
+and relative abundance of species in the plot) via seed trees for the
+quantity and quality (e.g., diversity and composition) of simulated
+regeneration, as one variant includes this feedback whereas the other
+does not.
 
@@
 
-Recruitment data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset covering such gradients over a large number of sites, to evaluate how well regeneration is captured.
-The observations used here are derived from a novel network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the EuFoRIa network [@euforia2019; @käber2023] (Figure 1).
-The environmental coverage and the variety of forests represented by this data is unprecedented within Europe.
-These forests have been unmanaged for at least 10 years prior to their designation as reserves, and most of them provide time series of natural forest dynamics over multiple decades and up to ca.
-85 years.
-The census periods range from 3 to 37 years, with an average of 14 years.
-The data provide information at the tree level, thus allowing for the sequential comparison of processes such as individual-tree recruitment and death between the measurements.
-
-We selected 200 sites from this network as the benchmarking dataset for the simulation, so as to be representative of the environmental variation contained in the data.
-This was achieved by applying k-means clustering to define 200 clusters of plots from the original set of 869 plots along the environmental dimensions of temperature, climatic water balance, soil quality, slope and aspect.
-Each of these sites featured at least two consecutive measurements.
-Recruitment size thresholds ranged from stem diameters at breast height (dbh) from 0 to 10 cm.
-For our study, we defined two datasets, where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds between 7 and 10 cm.
-Therefore we had in total 200 sites with observations of newly recruited trees above 10 cm and a subset of 165 sites with recruitment data above 7 cm.
-In the results section, any observations from the data at the 7 cm threshold originated from 165 sites.
-
-Plot size ranged from 0.02 ha to 5.52 ha, and the data were further processed and aggregated following @käber2023 to provide recruitment rates per ha and per decade.
-The observations used in this study featured 30,900 newly established trees.
-Recruitment rates per site, sample and decade ranged from zero to 1,246 trees, with a mean of 56 trees.
-Adult species composition was also available for each plot.
-For more details on this dataset and the detailed recruitment information, cf.
-@käber2023.
+Recruitment data covering a wide range of environmental conditions are
+hard to obtain, and this is one of the reasons why most models of forest
+dynamics have never been confronted with a dataset covering such
+gradients over a large number of sites, to evaluate how well
+regeneration is captured. The observations used here are derived from a
+novel network of sites in forest reserves that represent the range of
+environmental gradients in temperature and precipitation in Central
+Europe as compiled in the EuFoRIa network [@euforia2019; @käber2023]
+(Figure 1). The environmental coverage and the variety of forests
+represented by this data is unprecedented within Europe. These forests
+have been unmanaged for at least 10 years prior to their designation as
+reserves, and most of them provide time series of natural forest
+dynamics over multiple decades and up to ca. 85 years. The census
+periods range from 3 to 37 years, with an average of 14 years. The data
+provide information at the tree level, thus allowing for the sequential
+comparison of processes such as individual-tree recruitment and death
+between the measurements.
+
+We selected 200 sites from this network as the benchmarking dataset for
+the simulation, so as to be representative of the environmental
+variation contained in the data. This was achieved by applying k-means
+clustering to define 200 clusters of plots from the original set of 869
+plots along the environmental dimensions of temperature, climatic water
+balance, soil quality, slope and aspect. Each of these sites featured at
+least two consecutive measurements. Recruitment size thresholds ranged
+from stem diameters at breast height (dbh) from 0 to 10 cm. For our
+study, we defined two datasets, where one included 165 sites with a
+diameter threshold of 7 cm or lower, and the other included another 35
+sites with diameter thresholds between 7 and 10 cm. Therefore we had in
+total 200 sites with observations of newly recruited trees above 10 cm
+and a subset of 165 sites with recruitment data above 7 cm. In the
+results section, any observations from the data at the 7 cm threshold
+originated from 165 sites.
+
+Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
+processed and aggregated following @käber2023 to provide recruitment
+rates per ha and per decade. The observations used in this study
+featured 30,900 newly established trees. Recruitment rates per site,
+sample and decade ranged from zero to 1,246 trees, with a mean of 56
+trees. Adult species composition was also available for each plot. For
+more details on this dataset and the detailed recruitment information,
+cf. @käber2023.
 
@@
 
-The overarching goal of the simulation experiments was to assess tree recruitment as it arises from empirical data against its representation in a wide range of models of forest dynamics.
-We define tree recruitment as the passing of a breast height diameter threshold of 7 and 10 cm (synonym: ingrowth).
-To this end, each modeling group was provided with a detailed protocol [@díaz-yáñez2022] with instructions how to perform the simulations, input variables on climate and soil conditions, and the list of expected output variables.
-Neither were further site information (except for the data specified below) nor any data on tree recruitment or forest stand features provided prior to the simulation.
-
-The input variables were collected from different data sources and aggregated to be suitable for the needs of the different models.
-Time series of climatic variables from 1981 to 2018 were provided in hourly [Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly resolution [CHELSA data, Karger et al.
--@karger2021].
-Some variables required for some models, such as relative humidity or vapor pressure deficit, were calculated from these variables.
-The final instructions for using the climate data were slightly different in each model, based on the approach that best suited the model (Table \@ref(tab:tableModels)).
-Soil quality data were provided as continuous values between 1 and 5 [Soilgrids dataset, Hengl et al.
--@hengl2017].
-The protocol also provided the elevation, slope and aspect for each of the 200 sites [@asterscienceteam2019], but no other spatial information such as coordinates, with the exception of iLand and aDGVM2, which required blurred coordinates to derive highly detailed soil data.
-
-The simulations were run in the absence of natural disturbances.
-Only the model LPJ-GUESS had to include a "background" disturbance probability to increase chances of shade-intolerant species to establish.
-The simulations were set up to sample species-specific recruitment rates per decade and per ha in the equilibrium state of the model, typically entailing a "spin-up" run (as we did not provide any forest data).
-The modeling teams decided on the simulated area and how they derived these samples.
-The exact length of the simulation was also decided by the modeling teams (Table \@ref(tab:tableModels)).
-Further details on how each modeling team prepared the simulations and the outputs are available in Appendix S1: Model reports.
-
-The simulations were run in the absence of management to a simulated equilibrium ("Potential Natural Vegetation") with the current climate, as the ultimate goal was to evaluate tree recruitment under comparable and near-equilibrium conditions.
-This entails the assumption that (1) the observations from the forest reserves reflect no traces of forest management, and (2) there is an equilibrium between forest dynamics and climate.
-While the former might be starting to be visible in many of the EuFoRIA reserves, the latter may be more debatable.
-However, in the absence of detailed data on the history of each plot in the EuFoRIa network, some broad assumptions had to be made.
-Both the width of the regeneration niche (i.e., in environmental space) as well as the intensity of recruitment (i.e., the number of recruited trees per area and per unit of time) were of interest.
-
-The simulations were run for mixed-species forests (not multiple single-species simulations) using mixtures of eleven species or genera for which regeneration data of sufficient quality were available from EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp.
-L., Fraxinus excelsior L., Quercus spp.
-L., Alnus glutinosa L.,* and *Pinus sylvestris L.* The same set of species was used at all 200 sites.
-Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig), and the species simulated beyond the eleven species were lumped in a category "others".
-In three models, fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)).
-In 4C only *Fagus sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp.
-and *Pinus sylvestris* are parameterised.
-xComp simulations did not consider *Fraxinus excelsior* due to a deprecated species parameterisation.
-In LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not parameterized and therefore these species could not be included in the simulation.
-Finally, the model aDGVM2 does not simulate individual species.
-Rather, community assembly processes and trait filtering generate plant communities that are adjusted to the local biotic and abiotic conditions, and the simulated plants can be classified into ecological strategies based on their trait values in a post-processing step.
-
-Each of the models reported the number of new trees crossing the two diameter thresholds by sampling 200 times in a 10-year interval for each species and per ha for each of the 200 sites.
-Multiple samples per site were used to better understand the simulated variation within each site.
-This was done using different strategies, depending on the model, including (1) sampling simulated data from the same 1 ha plot in the equilibrium over time, (2) sampling several 1 ha plots from the simulated forest at one specific point in time (in the equilibrium) or (3) a combination of (1) and (2).
-This resulted in 880,000 observations per model that simulated the 11 species included in the protocol (200 sites, 200 samples per site, 11 species and two diameter thresholds).
-For the models that simulated additional species, their regeneration rates were aggregated as "others".
-
-Two models did not provide results from all the simulated samples or sites to avoid unrealistic results, as follows.
-In the model 4C, a threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid unrealistic stand basal area data, and therefore not all the samples and sites were reported.
-The reason for this is that 4C is not suitable for long-term simulations without management, due to misrepresentations in density-dependent mortality processes in long-term simulations and assumptions of tree geometry that lead to unrealistic single tree dimensions of very old individuals.
-LPJ-GUESS had 2% of the sample outputs without tree recruitment, and these were considered as zero stand basal area and zero regeneration for all the tree species simulated; one site produced grassland rather than a forest and was not included in the results.
+The overarching goal of the simulation experiments was to assess tree
+recruitment as it arises from empirical data against its representation
+in a wide range of models of forest dynamics. We define tree recruitment
+as the passing of a breast height diameter threshold of 7 and 10 cm
+(synonym: ingrowth). To this end, each modeling group was provided with
+a detailed protocol [@díaz-yáñez2022] with instructions how to perform
+the simulations, input variables on climate and soil conditions, and the
+list of expected output variables. Neither were further site information
+(except for the data specified below) nor any data on tree recruitment
+or forest stand features provided prior to the simulation.
+
+The input variables were collected from different data sources and
+aggregated to be suitable for the needs of the different models. Time
+series of climatic variables from 1981 to 2018 were provided in hourly
+[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
+resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
+required for some models, such as relative humidity or vapor pressure
+deficit, were calculated from these variables. The final instructions
+for using the climate data were slightly different in each model, based
+on the approach that best suited the model (Table
+\@ref(tab:tableModels)). Soil quality data were provided as continuous
+values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
+The protocol also provided the elevation, slope and aspect for each of
+the 200 sites [@asterscienceteam2019], but no other spatial information
+such as coordinates, with the exception of iLand and aDGVM2, which
+required blurred coordinates to derive highly detailed soil data.
+
+The simulations were run in the absence of natural disturbances. Only
+the model LPJ-GUESS had to include a "background" disturbance
+probability to increase chances of shade-intolerant species to
+establish. The simulations were set up to sample species-specific
+recruitment rates per decade and per ha in the equilibrium state of the
+model, typically entailing a "spin-up" run (as we did not provide any
+forest data). The modeling teams decided on the simulated area and how
+they derived these samples. The exact length of the simulation was also
+decided by the modeling teams (Table \@ref(tab:tableModels)). Further
+details on how each modeling team prepared the simulations and the
+outputs are available in Appendix S1: Model reports.
+
+The simulations were run in the absence of management to a simulated
+equilibrium ("Potential Natural Vegetation") with the current climate,
+as the ultimate goal was to evaluate tree recruitment under comparable
+and near-equilibrium conditions. This entails the assumption that (1)
+the observations from the forest reserves reflect no traces of forest
+management, and (2) there is an equilibrium between forest dynamics and
+climate. While the former might be starting to be visible in many of the
+EuFoRIA reserves, the latter may be more debatable. However, in the
+absence of detailed data on the history of each plot in the EuFoRIa
+network, some broad assumptions had to be made. Both the width of the
+regeneration niche (i.e., in environmental space) as well as the
+intensity of recruitment (i.e., the number of recruited trees per area
+and per unit of time) were of interest.
+
+The simulations were run for mixed-species forests (not multiple
+single-species simulations) using mixtures of eleven species or genera
+for which regeneration data of sufficient quality were available from
+EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
+betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
+Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L.,* and *Pinus
+sylvestris L.* The same set of species was used at all 200 sites. Two
+models included their standard set of species for the simulations, which
+is much larger (i.e., ForClim 1, ForClim 11, TreeMig), and the species
+simulated beyond the eleven species were lumped in a category "others".
+In three models, fewer than these eleven species were simulated (4C,
+xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). In 4C only *Fagus
+sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp. and *Pinus
+sylvestris* are parameterised. xComp simulations did not consider
+*Fraxinus excelsior* due to a deprecated species parameterisation. In
+LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not
+parameterized and therefore these species could not be included in the
+simulation. Finally, the model aDGVM2 does not simulate individual
+species. Rather, community assembly processes and trait filtering
+generate plant communities that are adjusted to the local biotic and
+abiotic conditions, and the simulated plants can be classified into
+ecological strategies based on their trait values in a post-processing
+step.
+
+Each of the models reported the number of new trees crossing the two
+diameter thresholds by sampling 200 times in a 10-year interval for each
+species and per ha for each of the 200 sites. Multiple samples per site
+were used to better understand the simulated variation within each site.
+This was done using different strategies, depending on the model,
+including (1) sampling simulated data from the same 1 ha plot in the
+equilibrium over time, (2) sampling several 1 ha plots from the
+simulated forest at one specific point in time (in the equilibrium) or
+(3) a combination of (1) and (2). This resulted in 880,000 observations
+per model that simulated the 11 species included in the protocol (200
+sites, 200 samples per site, 11 species and two diameter thresholds).
+For the models that simulated additional species, their regeneration
+rates were aggregated as "others".
+
+Two models did not provide results from all the simulated samples or
+sites to avoid unrealistic results, as follows. In the model 4C, a
+threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid
+unrealistic stand basal area data, and therefore not all the samples and
+sites were reported. The reason for this is that 4C is not suitable for
+long-term simulations without management, due to misrepresentations in
+density-dependent mortality processes in long-term simulations and
+assumptions of tree geometry that lead to unrealistic single tree
+dimensions of very old individuals. LPJ-GUESS had 2% of the sample
+outputs without tree recruitment, and these were considered as zero
+stand basal area and zero regeneration for all the tree species
+simulated; one site produced grassland rather than a forest and was not
+included in the results.
 
@@
 ```
-where $s$ is the total number of species present that have a basal area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of species $i$ in sample $n$ calculated as the regeneration basal area ($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the proportion of species $i$ calculated as the basal area of all trees ($BA_{n,i}$) of that species $i$ relative to total basal area ($Total_{BA_{n}}$) of the sample $n$.
-Species diversity was not assessed for aDGVM2 as this model does not simulate individual species.
-
-Mortality in tree recruitment was assessed based on the ratio of recruitment between the 7 and 10 cm diameter thresholds.
-We used the Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference to estimate whether the ratio of recruitment between these two thresholds was above or below the expected theoretical rate.
-The Reineke self-thinning rule is usually calculated for even-aged, single species stands and is based on a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands.
-The value used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm to be 77% higher than at 10 cm), calculated using Equation \ref{eqReineke}.
+where $s$ is the total number of species present that have a basal area
+larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
+species $i$ in sample $n$ calculated as the regeneration basal area
+($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal
+area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
+proportion of species $i$ calculated as the basal area of all trees
+($BA_{n,i}$) of that species $i$ relative to total basal area
+($Total_{BA_{n}}$) of the sample $n$. Species diversity was not assessed
+for aDGVM2 as this model does not simulate individual species.
+
+Mortality in tree recruitment was assessed based on the ratio of
+recruitment between the 7 and 10 cm diameter thresholds. We used the
+Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference
+to estimate whether the ratio of recruitment between these two
+thresholds was above or below the expected theoretical rate. The Reineke
+self-thinning rule is usually calculated for even-aged, single species
+stands and is based on a fixed relationship between the number of stems
+and the quadratic mean diameter in fully stocked pure stands. The value
+used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm
+to be 77% higher than at 10 cm), calculated using Equation
+\ref{eqReineke}.
 
@@
 ```
-We assessed model performance in relation to model traits focusing on (1) model complexity as defined by @bugmann2022, (2) model type (empirical or process-based), (3) the presence or absence of a canopy feedback for regeneration, and (4) the scale of application of the model (stand, landscape or global).
-We tested for significant differences between simulated and observed values of the recruitment levels and species diversity by using t-tests with the Bonferroni correction for multiple testing.
-Note that the t-tests does not present conclusive evidence for a specific hypotheses, but rather facilitates managing the extensive number of comparisons discussed in our study.
-
-Total recruitment and the regeneration niches of the individual species were evaluated across the gradients of light availability, temperature, and soil moisture as captured in the data.
-Stand basal area (the basal area of all the trees in each sample above the recruitment threshold) was used as a proxy for light availability at the forest floor, the annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing season warmth, and the climatic water balance as a proxy for soil moisture [@speich2019].
-Recruitment values of each species were calculated as the mean across the 200 samples per site.
-The observed data were modeled using a Generalized Additive Model [@wood2011] with a negative binomial distribution and restricted maximum likelihood to better understand the relationship between the environmental gradients and the levels of observed regeneration, relative to the simulation results per model.
-
-In order to analyze the regeneration niches across the climatic gradients, we focused on five common tree species or genera: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and *Quercus* spp.
-For these species, we calculated the share in the recruited basal area per site, where each site represents a part of the environmental space, as the mean across the available samples per site:
+We assessed model performance in relation to model traits focusing on
+(1) model complexity as defined by @bugmann2022, (2) model type
+(empirical or process-based), (3) the presence or absence of a canopy
+feedback for regeneration, and (4) the scale of application of the model
+(stand, landscape or global). We tested for significant differences
+between simulated and observed values of the recruitment levels and
+species diversity by using t-tests with the Bonferroni correction for
+multiple testing. Note that the t-tests does not present conclusive
+evidence for a specific hypotheses, but rather facilitates managing the
+extensive number of comparisons discussed in our study.
+
+Total recruitment and the regeneration niches of the individual species
+were evaluated across the gradients of light availability, temperature,
+and soil moisture as captured in the data. Stand basal area (the basal
+area of all the trees in each sample above the recruitment threshold)
+was used as a proxy for light availability at the forest floor, the
+annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing
+season warmth, and the climatic water balance as a proxy for soil
+moisture [@speich2019]. Recruitment values of each species were
+calculated as the mean across the 200 samples per site. The observed
+data were modeled using a Generalized Additive Model [@wood2011] with a
+negative binomial distribution and restricted maximum likelihood to
+better understand the relationship between the environmental gradients
+and the levels of observed regeneration, relative to the simulation
+results per model.
+
+In order to analyze the regeneration niches across the climatic
+gradients, we focused on five common tree species or genera: *Abies
+alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
+*Quercus* spp. For these species, we calculated the share in the
+recruited basal area per site, where each site represents a part of the
+environmental space, as the mean across the available samples per site:
 
@@
 ```
-where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$ is the mean basal area in the recruitment (subsequently referred to as "recruited basal area") of species $i$ at site $t$ across the available samples at that site.
-Furthermore, regarding the share in the recruited basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the recruited basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total recruited basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the recruited basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero.
+where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$
+is the mean basal area in the recruitment (subsequently referred to as
+"recruited basal area") of species $i$ at site $t$ across the available
+samples at that site. Furthermore, regarding the share in the recruited
+basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this
+as zero when the recruited basal area of that species
+($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total recruited
+basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the recruited
+basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero.
 
@@
 
-Simulated levels of recruitment varied strongly across the 15 models and typically did not match the levels found in the forest reserve data (Figure 2).
-Recruitment was overestimated in most models for both the 7 and 10 cm diameter thresholds, with the exception of the empirical stand model SIBYLA, the landscape model Landis-II and the global model aDGVM2, which estimated recruited levels at the lower end of the plausibility interval of the observed data (Figure 2).
-The models with the largest overestimation were the stand model PICUS and the landscape model TreeMig.
-For most models, the variability of simulated recruitment levels across the 200 sites (visible from the interquartile range in the box plots of Figure 2) was similar to or smaller than observed, with the notable exception of PICUS, where simulated recruitment variability was much larger.
-
-Both the observations and the simulated data had no recruitment in some samples and at some sites (for details, cf.
-Table S1).
-The observed data had 4% of the samples with no recruitment.
-Only three models had a larger proportion of no recruitment (4C, Landis-II and aDVM2).
-Two models always simulated recruitment for both the 7 and 10 cm threshold (xComp and TreeMig), i.e., they did not feature any zero values.
-The other ten models had a very low percentage of samples with no recruitment (0.01% - 2.39%), i.e., they had distinctly fewer occurrences of zeros than the observations.
+Simulated levels of recruitment varied strongly across the 15 models and
+typically did not match the levels found in the forest reserve data
+(Figure 2). Recruitment was overestimated in most models for both the 7
+and 10 cm diameter thresholds, with the exception of the empirical stand
+model SIBYLA, the landscape model Landis-II and the global model aDGVM2,
+which estimated recruited levels at the lower end of the plausibility
+interval of the observed data (Figure 2). The models with the largest
+overestimation were the stand model PICUS and the landscape model
+TreeMig. For most models, the variability of simulated recruitment
+levels across the 200 sites (visible from the interquartile range in the
+box plots of Figure 2) was similar to or smaller than observed, with the
+notable exception of PICUS, where simulated recruitment variability was
+much larger.
+
+Both the observations and the simulated data had no recruitment in some
+samples and at some sites (for details, cf. Table S1). The observed data
+had 4% of the samples with no recruitment. Only three models had a
+larger proportion of no recruitment (4C, Landis-II and aDVM2). Two
+models always simulated recruitment for both the 7 and 10 cm threshold
+(xComp and TreeMig), i.e., they did not feature any zero values. The
+other ten models had a very low percentage of samples with no
+recruitment (0.01% - 2.39%), i.e., they had distinctly fewer occurrences
+of zeros than the observations.
 
@@
 
-Most models matched the level of diversity of the observed data quite well (Figure 3).
-Five models overestimated recruitment diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter particularly for the 7 cm diameter threshold.
-The model 4C is a special case, as it simulated five species only, i.e., its diversity values are not directly comparable to those of the other models, nor to the observations.
-Only one model, Landis-II, consistently underestimated recruited diversity.
-
-In most models, there were only small but significant differences in the species diversity of recruitment between 7 and 10 cm across sites (Appendix S1: Table S2).
-Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained the recruitment diversity between the 7 and 10 cm thresholds (i.e., the differences between them were not significant, see Appendix S1: Table S2), and the same was evident from the observed data.
-
-For both the observed and the simulated data, and both recruitment thresholds, species diversity in the recruitment compared to stand-level diversity did not feature clear patterns across the models (Appendix S1: Figures S1 and S2).
-Most models captured reasonably well or overpredicted species diversity at the stand level (Figure 4A, left and center; cf.
-Appendix S1: Figure S3), but a characteristic feature was that the simulations had a much lower variability of diversity compared to observations.
-Models overpredicting stand-level species diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure 4A, left).
-Several models underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure 4A, right).
-
-The majority of the models overpredicted recruitment diversity; as already observed in Figure 3, only two of them underestimated it (4C and Landis-II) (Figure 4B and Appendix S1: Figure S4).
-Both the models with and without feedback from the adult trees to regeneration (via seed production) overestimated, underestimated or captured reasonably well recruitment species diversity compared to the observations.
-Again, simulated diversity had considerably lower variability than observed diversity in the recruitment.
+Most models matched the level of diversity of the observed data quite
+well (Figure 3). Five models overestimated recruitment diversity:
+ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter
+particularly for the 7 cm diameter threshold. The model 4C is a special
+case, as it simulated five species only, i.e., its diversity values are
+not directly comparable to those of the other models, nor to the
+observations. Only one model, Landis-II, consistently underestimated
+recruited diversity.
+
+In most models, there were only small but significant differences in the
+species diversity of recruitment between 7 and 10 cm across sites
+(Appendix S1: Table S2). Four models (iLand, Landis-II, TreeMig and
+LPJ-GUESS) maintained the recruitment diversity between the 7 and 10 cm
+thresholds (i.e., the differences between them were not significant, see
+Appendix S1: Table S2), and the same was evident from the observed data.
+
+For both the observed and the simulated data, and both recruitment
+thresholds, species diversity in the recruitment compared to stand-level
+diversity did not feature clear patterns across the models (Appendix S1:
+Figures S1 and S2). Most models captured reasonably well or
+overpredicted species diversity at the stand level (Figure 4A, left and
+center; cf. Appendix S1: Figure S3), but a characteristic feature was
+that the simulations had a much lower variability of diversity compared
+to observations. Models overpredicting stand-level species diversity
+included ForClim1, ForClim 11, FORMIND and TreeMig (Figure 4A, left).
+Several models underpredicted species diversity, i.e., LPJ-GUESS,
+ForCEEPS(f) and 4C (Figure 4A, right).
+
+The majority of the models overpredicted recruitment diversity; as
+already observed in Figure 3, only two of them underestimated it (4C and
+Landis-II) (Figure 4B and Appendix S1: Figure S4). Both the models with
+and without feedback from the adult trees to regeneration (via seed
+production) overestimated, underestimated or captured reasonably well
+recruitment species diversity compared to the observations. Again,
+simulated diversity had considerably lower variability than observed
+diversity in the recruitment.
 
@@
 
-There were strong differences among models regarding the mortality rate between 7 and 10 cm (Figure 5).
-Most of the models featured a mortality rate significantly larger than the observed data, thus at least partially compensating for the general overestimation of recruitment levels at 7 cm (Figure 2 and Appendix S1: Table S3); six models did not feature a significant difference compared to the observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2).
-Four models (xComp, PICUS, LandClim and TreeMig) featured very high mortality rates (i.e., well above the Reineke self-thinning line), which compensated for the strong initial overestimation of recruitment (cf.
-Figure 2).
-Interestingly, two models that underestimated overall recruitment levels (SIBYLA and Landis-II; cf.
-Figure 2) featured mortality rates that were close to but still above the self-thinning line (Figure 5).
-
-The fact that some models (and observed data) featured lower mortality than expected by Reineke's self-thinning rule may point to facilitation, or simply a higher mortality rate before the trees had reached 7 cm diameter.
-Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher recruitment mortality compared to even-aged, single-species forests.
-Some models yielded a mortality rate of recruitment that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).
-
-The three models with the largest overestimation of the proportion of recruitment at the 7 cm threshold featured the highest mortality between the 7 and 10 cm thresholds (Figure 6A).
-The other models that overestimated recruitment had mortalities lower than the expected self-thinning ratio, with the exception of LandClim (Figure 6B).
-It is noteworthy that most of the models that featured a low ratio of recruitment between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small overestimation of recruitment at the 7 cm threshold (Figure 5).
-aDGVM2 was the only model that underestimated recruitment at the 7 cm threshold while featuring a low mortality between the two thresholds (Figure 6C), and only a few models were close to the correct recruitment levels at 7 cm while also being close to the Reineke line (SIBYLA, Landis-II; cf.
-Figure 6D; and 4C as well as LandClim).
-
-Lastly, some models that had a low number of recruits at the 7 cm threshold also had a small decrease of recruitment between 7 and 10 cm (ratio 7-10 cm) (Appendix S1: Figure S5).
-In these models, competition and self-thinning are either not pronounced or must have occurred before the trees had reached 7 cm.
-However, this pattern was not consistent across models (Appendix S1: Figure S5).
-It is noteworthy that there is a relationship between the recruitment levels at 7 cm and the mortality rate (Appendix S1: Table S4): most of the models showed a positive effect (higher regeneration at 7 cm is coupled to an increase of the mortality between 7 and 10 cm).
-While some models showed a negative effect, the linear trend was not pronounced and not always significant regarding its slope.
+There were strong differences among models regarding the mortality rate
+between 7 and 10 cm (Figure 5). Most of the models featured a mortality
+rate significantly larger than the observed data, thus at least
+partially compensating for the general overestimation of recruitment
+levels at 7 cm (Figure 2 and Appendix S1: Table S3); six models did not
+feature a significant difference compared to the observed data
+(ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2). Four models
+(xComp, PICUS, LandClim and TreeMig) featured very high mortality rates
+(i.e., well above the Reineke self-thinning line), which compensated for
+the strong initial overestimation of recruitment (cf. Figure 2).
+Interestingly, two models that underestimated overall recruitment levels
+(SIBYLA and Landis-II; cf. Figure 2) featured mortality rates that were
+close to but still above the self-thinning line (Figure 5).
+
+The fact that some models (and observed data) featured lower mortality
+than expected by Reineke's self-thinning rule may point to facilitation,
+or simply a higher mortality rate before the trees had reached 7 cm
+diameter. Yet, the case of models such as xComp, PICUS and TreeMig,
+whose mortality was well above the self-thinning line, indicates that
+they feature higher recruitment mortality compared to even-aged,
+single-species forests. Some models yielded a mortality rate of
+recruitment that is broadly compatible with the self-thinning rule
+(e.g., 4C and LandClim).
+
+The three models with the largest overestimation of the proportion of
+recruitment at the 7 cm threshold featured the highest mortality between
+the 7 and 10 cm thresholds (Figure 6A). The other models that
+overestimated recruitment had mortalities lower than the expected
+self-thinning ratio, with the exception of LandClim (Figure 6B). It is
+noteworthy that most of the models that featured a low ratio of
+recruitment between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11
+and iLand) had only a small overestimation of recruitment at the 7 cm
+threshold (Figure 5). aDGVM2 was the only model that underestimated
+recruitment at the 7 cm threshold while featuring a low mortality
+between the two thresholds (Figure 6C), and only a few models were close
+to the correct recruitment levels at 7 cm while also being close to the
+Reineke line (SIBYLA, Landis-II; cf. Figure 6D; and 4C as well as
+LandClim).
+
+Lastly, some models that had a low number of recruits at the 7 cm
+threshold also had a small decrease of recruitment between 7 and 10 cm
+(ratio 7-10 cm) (Appendix S1: Figure S5). In these models, competition
+and self-thinning are either not pronounced or must have occurred before
+the trees had reached 7 cm. However, this pattern was not consistent
+across models (Appendix S1: Figure S5). It is noteworthy that there is a
+relationship between the recruitment levels at 7 cm and the mortality
+rate (Appendix S1: Table S4): most of the models showed a positive
+effect (higher regeneration at 7 cm is coupled to an increase of the
+mortality between 7 and 10 cm). While some models showed a negative
+effect, the linear trend was not pronounced and not always significant
+regarding its slope.
 
@@
 
-There was no significant relationship between the mean complexity of the regeneration module (Table \@ref(tab:tableModels)) and the overestimation proportion at a diameter threshold of 7 cm (Appendix S1: Figure S6).
-All models except 4C had significant differences between the observed and simulated mean recruitment values.
-Therefore it was not possible to differentiate in terms of performance between models type (empirical or process based), or their scales (stand, landscape or global) (Appendix S1: Table S5).
-The values of species diversity in the recruitment simulated by each model were also assessed against the observed data to see if there was a difference between models that include a feedback compared to those that did not; all the models except ForCEEPs(f) and FORMIND had significant differences (Appendix S1: Table S6).
-It is noteworthy that for the model ForCEEPS the pattern changed from the overprediction of species diversity in the regeneration to a diversity level that is closer to observations when the recruitment module included feedback (ForCEEPs(f)).
+There was no significant relationship between the mean complexity of the
+regeneration module (Table \@ref(tab:tableModels)) and the
+overestimation proportion at a diameter threshold of 7 cm (Appendix S1:
+Figure S6). All models except 4C had significant differences between the
+observed and simulated mean recruitment values. Therefore it was not
+possible to differentiate in terms of performance between models type
+(empirical or process based), or their scales (stand, landscape or
+global) (Appendix S1: Table S5). The values of species diversity in the
+recruitment simulated by each model were also assessed against the
+observed data to see if there was a difference between models that
+include a feedback compared to those that did not; all the models except
+ForCEEPs(f) and FORMIND had significant differences (Appendix S1: Table
+S6). It is noteworthy that for the model ForCEEPS the pattern changed
+from the overprediction of species diversity in the regeneration to a
+diversity level that is closer to observations when the recruitment
+module included feedback (ForCEEPs(f)).
 
@@
 
-When evaluating total tree recruitment levels along key gradients of light availability (basal area), temperature (degree-days) and soil moisture (climatic water balance), distinct features emerged: the models reproduced the effect of basal area in both its magnitude and patterns (Figure 7 and Appendix S1: Figure S7) considerably better than the effects of the climatic gradients, where they featured varying patterns.
-
-In the observed data, recruitment levels decreased clearly with increasing stand basal area (Figure 7).
-This trend was captured by the Landis-II model only, albeit at much lower values of basal area than in the observations.
-The other models featured distinctly different trends, such as (1) an increase in recruitment levels with increasing total basal area followed by the absence of recruitment at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant recruitment levels with basal area (ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of recruitment up to a certain value of stand basal area followed by a decrease at even higher values of stand basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).
-
-The observed recruitment did not change much across the climatic water balance gradient, and showed a slight increase with temperature.
-Neither did the models generally match the observed recruitment patterns across these gradients (Figure 8,9), nor was there a clear pattern across models.
-Regarding the soil moisture gradient, a group of models featured decreasing recruitment with increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a few models showed the opposite trend (ForCEEPS(f), FORMIND and LPJ-GUESS).
-Across the temperature gradient, all the models that featured a decrease in recruitment levels with increasing soil moisture showed the opposite trend (i.e., more recruitment with increasing temperature), except for 4C.
-The other models did not feature a clear pattern.
-
-Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus sylvestris*) and oak (*Quercus* spp.) were the species observed most frequently in the EuFoRIa data.
-Most models captured well the share of basal area in the recruitment of these five species compared to the observed data along the two major climatic gradients (Figure 8, cf.
-Appendix S1: Figure S8).
-However, the models differed considerably in the way in which they simulated these climatic niches.
-
-First, some models overestimated recruitment across the entire gradients of at least one of the five main tree species (Figure 10, e.g., ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated recruitment of more than one species for a large part of the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
-red colors in Figure 10).
-These trends were consistent for both recruitment thresholds, although the exact changes in the share of the recruited basal area were different (Appendix S1: Figure S8).
-
-Second, there was some congruence in the simulation results by species across the models.
-The share in the recruitment of Silver fir (*Abies alba*) was captured well across the gradient by most models except for xComp, PICUS, and LandClim, which overestimated the recruitment share of this species.
-There were also some peculiarities evident for some models, such as 4C that did not simulate *Abies alba*.
-The patterns across models for *Fagus sylvatica* recruitment were more complex, as its recruitment was underestimated by many models across the environmental gradients featuring negligible regeneration at most sites, while others overestimated it in different parts of the environmental space.
-Most models overestimated the share of *Picea abies* recruitment in the cold-wet part of the gradients.
-*Pinus sylvestris* and *Quercus* spp.
-represented a small share of the recruitment in the observed data mostly at warm-dry sites.
-Many models, however, erroneously featured recruitment for these two species along most of the climatic gradients, although with a low share of basal area.
-Five models (4C, FORMIND, SIBYLA, xComp, and LandClim) had very small amounts of recruitment of *Pinus sylvestris* and *Quercus* spp., or did not feature any recruitment of these species at all.
-
-Lastly, there is no model that performed well across the five species regarding the species-specific recruitment levels in the environmental space (Figure 10).
-Some models (e.g., ForClim11, PICUS) tended to consistently overestimate the recruitment share of all five species, but most models overestimated the recruitment share of some species while underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
+When evaluating total tree recruitment levels along key gradients of
+light availability (basal area), temperature (degree-days) and soil
+moisture (climatic water balance), distinct features emerged: the models
+reproduced the effect of basal area in both its magnitude and patterns
+(Figure 7 and Appendix S1: Figure S7) considerably better than the
+effects of the climatic gradients, where they featured varying patterns.
+
+In the observed data, recruitment levels decreased clearly with
+increasing stand basal area (Figure 7). This trend was captured by the
+Landis-II model only, albeit at much lower values of basal area than in
+the observations. The other models featured distinctly different trends,
+such as (1) an increase in recruitment levels with increasing total
+basal area followed by the absence of recruitment at high values of
+basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2)
+almost constant recruitment levels with basal area (ForClim 11, xComp,
+LandClim and aDGVM2), or (3) an increase of recruitment up to a certain
+value of stand basal area followed by a decrease at even higher values
+of stand basal area, with model-specific thresholds (4C, ForClim 1,
+PICUS, SIBYLA and iLand).
+
+The observed recruitment did not change much across the climatic water
+balance gradient, and showed a slight increase with temperature. Neither
+did the models generally match the observed recruitment patterns across
+these gradients (Figure 8,9), nor was there a clear pattern across
+models. Regarding the soil moisture gradient, a group of models featured
+decreasing recruitment with increasing soil moisture (4C, ForClim 1,
+ForClim11, SIBYLA, xComp, PICUS, LandClim, TreeMig and iLand), while a
+few models showed the opposite trend (ForCEEPS(f), FORMIND and
+LPJ-GUESS). Across the temperature gradient, all the models that
+featured a decrease in recruitment levels with increasing soil moisture
+showed the opposite trend (i.e., more recruitment with increasing
+temperature), except for 4C. The other models did not feature a clear
+pattern.
+
+Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
+sylvestris*) and oak (*Quercus* spp.) were the species observed most
+frequently in the EuFoRIa data. Most models captured well the share of
+basal area in the recruitment of these five species compared to the
+observed data along the two major climatic gradients (Figure 8, cf.
+Appendix S1: Figure S8). However, the models differed considerably in
+the way in which they simulated these climatic niches.
+
+First, some models overestimated recruitment across the entire gradients
+of at least one of the five main tree species (Figure 10, e.g.,
+ForCEEPS, ForCEEPS(f), LandClim or LPJ-GUESS), while other models
+overestimated recruitment of more than one species for a large part of
+the climate gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf.
+red colors in Figure 10). These trends were consistent for both
+recruitment thresholds, although the exact changes in the share of the
+recruited basal area were different (Appendix S1: Figure S8).
+
+Second, there was some congruence in the simulation results by species
+across the models. The share in the recruitment of Silver fir (*Abies
+alba*) was captured well across the gradient by most models except for
+xComp, PICUS, and LandClim, which overestimated the recruitment share of
+this species. There were also some peculiarities evident for some
+models, such as 4C that did not simulate *Abies alba*. The patterns
+across models for *Fagus sylvatica* recruitment were more complex, as
+its recruitment was underestimated by many models across the
+environmental gradients featuring negligible regeneration at most sites,
+while others overestimated it in different parts of the environmental
+space. Most models overestimated the share of *Picea abies* recruitment
+in the cold-wet part of the gradients. *Pinus sylvestris* and *Quercus*
+spp. represented a small share of the recruitment in the observed data
+mostly at warm-dry sites. Many models, however, erroneously featured
+recruitment for these two species along most of the climatic gradients,
+although with a low share of basal area. Five models (4C, FORMIND,
+SIBYLA, xComp, and LandClim) had very small amounts of recruitment of
+*Pinus sylvestris* and *Quercus* spp., or did not feature any
+recruitment of these species at all.
+
+Lastly, there is no model that performed well across the five species
+regarding the species-specific recruitment levels in the environmental
+space (Figure 10). Some models (e.g., ForClim11, PICUS) tended to
+consistently overestimate the recruitment share of all five species, but
+most models overestimated the recruitment share of some species while
+underestimating the share of others (e.g., ForCEEPS, LPJ-GUESS).
 
@@
 
-Tree regeneration is a fundamental process in forest dynamics.
-Correctly capturing it in dynamic models is fundamental to, for example, evaluate post-disturbance dynamics and potential long-term recovery trajectories as it will define the forest state in the first decades [@seidl2022].
-If this initial phase is not captured well, we can not properly assess aspects such as forest resilience or the timing, magnitude and progression of carbon sequestration.
-In this study, for the first time, projections of tree recruitment from multiple models of forest dynamics were confronted with a unique dataset from unmanaged forest reserves across a large environmental gradient in Europe.
-
-The EuFoRIa data [@käber2023] are exceptional, particularly with respect to the number of records (number of sites and repeated measurements on unmanaged forests), which is essential for capturing a highly 'noisy' process such as tree regeneration.
-The use of this dataset for model benchmarking provided novel insights on the ability of state-of-the-art models to accurately simulate recruitment levels, its species composition and mortality in an early stage of tree life, i.e., between tree diameters of 7 and 10 cm.
-Overall, by adopting this approach a much broader understanding resulted than if we had used these data for model calibration: it is primarily from the shortcomings of the models that we can gain ecological insights [cf.
-@trugman2022].
+Tree regeneration is a fundamental process in forest dynamics. Correctly
+capturing it in dynamic models is fundamental to, for example, evaluate
+post-disturbance dynamics and potential long-term recovery trajectories
+as it will define the forest state in the first decades [@seidl2022]. If
+this initial phase is not captured well, we can not properly assess
+aspects such as forest resilience or the timing, magnitude and
+progression of carbon sequestration. In this study, for the first time,
+projections of tree recruitment from multiple models of forest dynamics
+were confronted with a unique dataset from unmanaged forest reserves
+across a large environmental gradient in Europe.
+
+The EuFoRIa data [@käber2023] are exceptional, particularly with respect
+to the number of records (number of sites and repeated measurements on
+unmanaged forests), which is essential for capturing a highly 'noisy'
+process such as tree regeneration. The use of this dataset for model
+benchmarking provided novel insights on the ability of state-of-the-art
+models to accurately simulate recruitment levels, its species
+composition and mortality in an early stage of tree life, i.e., between
+tree diameters of 7 and 10 cm. Overall, by adopting this approach a much
+broader understanding resulted than if we had used these data for model
+calibration: it is primarily from the shortcomings of the models that we
+can gain ecological insights [cf. @trugman2022].
 
@@
 
-Most of the models overestimated tree recruitment levels.
-This has potentially far-reaching implications, e.g., regarding biomass (and thus carbon) turnover, with a potential overestimation of the capacity for forest carbon sequestration [@pugh2019; @pan2011].
-Yet, we focused on one specific stage, i.e., recruitment into the 7 and 10 cm diameter classes.
-Trees at this size contribute little to carbon sequestration, and if the excess regeneration at this stage is compensated -- in the models -- soon thereafter by higher mortality, overall simulated tree population dynamics may still be trustworthy at the level of stand structure and attributes such as basal area or biomass.
-
-From a modeling point of view, excess tree recruitment inevitably requires excess mortality rates in a later stage, either -- as observed for some models -- between the 7 and 10 cm diameter thresholds, or soon after the 10 cm threshold has been crossed.
-In any case, correcting at early stages for the expected forest densities at later stages is equivalent to compensating for a first error (excess recruitment) by a second error (excess mortality).
-It is highly likely that biased projections will result, because the two errors are unlikely to be perfectly linked and thus will not always compensate each other.
-Hence, this structural problem of most models investigated here is problematic particularly if the models are to be used under novel conditions such as under climate change [e.g., @huber2021] or in a decision support context [e.g., @thrippleton2021].
-
-Recruitment levels define the structure and composition of future forests, but it is equally important to correctly identify areas where regeneration is lacking [@rammer2021].
-There are multiple constraints to the regeneration niche of tree species [@price2001], and therefore the absence of regeneration is likely to be common [@fortin2007], even over larger areas such as the one-hectare plots used here.
-Tree regeneration data are characterized by zero-inflation.
-This was clearly evident from the EuFoRIa dataset, but some of the models did not produce zeros at all, or featured a very low proportion of zero data.
-This substantial difference may be due to the fact that the simulation results were drawn from equilibrium forests, whereas in reality many of the forest reserves are recovering from past management activities and have become denser over the past decades [e.g., @heiri2009], leading to less regeneration than in an equilibrium situation.
-Another possible reason for this difference may be the exclusion of factors like deer browsing or failure to accurately measure establishment filters, such as competition with the herb layer, site-specific or environmental limitations.
+Most of the models overestimated tree recruitment levels. This has
+potentially far-reaching implications, e.g., regarding biomass (and thus
+carbon) turnover, with a potential overestimation of the capacity for
+forest carbon sequestration [@pugh2019; @pan2011]. Yet, we focused on
+one specific stage, i.e., recruitment into the 7 and 10 cm diameter
+classes. Trees at this size contribute little to carbon sequestration,
+and if the excess regeneration at this stage is compensated -- in the
+models -- soon thereafter by higher mortality, overall simulated tree
+population dynamics may still be trustworthy at the level of stand
+structure and attributes such as basal area or biomass.
+
+From a modeling point of view, excess tree recruitment inevitably
+requires excess mortality rates in a later stage, either -- as observed
+for some models -- between the 7 and 10 cm diameter thresholds, or soon
+after the 10 cm threshold has been crossed. In any case, correcting at
+early stages for the expected forest densities at later stages is
+equivalent to compensating for a first error (excess recruitment) by a
+second error (excess mortality). It is highly likely that biased
+projections will result, because the two errors are unlikely to be
+perfectly linked and thus will not always compensate each other. Hence,
+this structural problem of most models investigated here is problematic
+particularly if the models are to be used under novel conditions such as
+under climate change [e.g., @huber2021] or in a decision support context
+[e.g., @thrippleton2021].
+
+Recruitment levels define the structure and composition of future
+forests, but it is equally important to correctly identify areas where
+regeneration is lacking [@rammer2021]. There are multiple constraints to
+the regeneration niche of tree species [@price2001], and therefore the
+absence of regeneration is likely to be common [@fortin2007], even over
+larger areas such as the one-hectare plots used here. Tree regeneration
+data are characterized by zero-inflation. This was clearly evident from
+the EuFoRIa dataset, but some of the models did not produce zeros at
+all, or featured a very low proportion of zero data. This substantial
+difference may be due to the fact that the simulation results were drawn
+from equilibrium forests, whereas in reality many of the forest reserves
+are recovering from past management activities and have become denser
+over the past decades [e.g., @heiri2009], leading to less regeneration
+than in an equilibrium situation. Another possible reason for this
+difference may be the exclusion of factors like deer browsing or failure
+to accurately measure establishment filters, such as competition with
+the herb layer, site-specific or environmental limitations.
 
@@
 
-Correctly capturing species composition in tree recruitment is important to assess the future functional diversity of a forest, e.g., its sensitivity to drought or resilience to disturbances [@redmond2015; @seidl2022].
-In the simulations, overall species diversity levels in the recruitment were well within the observed range for half of the models (7 out of 14).
-Thus, while most models are facing difficulties to quantitatively match recruitment levels (cf.
-above), their performance is better regarding the composition of recruitment as a function of abiotic and biotic conditions.
-Most models maintained or even decreased recruitment diversity between the 7 and 10 cm thresholds, and the same is visible from the empirical data, but the differences between the 7 and 10 cm thresholds were not significant.
-Diversity variations may be explained by the way the individual models consider regeneration processes [@konig_tree_2022], e.g., regeneration can be linked to the seed/seedling bank, seed rain from dispersal or from a pool of available species.
-
-The species diversity of the entire stand was captured better by the models than the diversity of tree recruitment, and only a few models overpredicted stand-level diversity.
-Defining recruitment composition as being proportional to local adult abundance, regardless of productivity, might be a simple and conservative assumption to maintain relative species abundances [@hanbury-brown2022], although this may be simplistic.
-Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured the species diversity of the recruitment better than those without feedback, with the exception of ForCEEPS that featured significantly more accurate recruitment species diversity with the model version that included feedback (ForCEEPS(f)).
-The similar performance of models with and without feedback is likely because the models put more weight on the regeneration niche arising from abiotic and biotic filters, than from the habitat niche of the adult trees [@grubb1977].
+Correctly capturing species composition in tree recruitment is important
+to assess the future functional diversity of a forest, e.g., its
+sensitivity to drought or resilience to disturbances [@redmond2015;
+@seidl2022]. In the simulations, overall species diversity levels in the
+recruitment were well within the observed range for half of the models
+(7 out of 14). Thus, while most models are facing difficulties to
+quantitatively match recruitment levels (cf. above), their performance
+is better regarding the composition of recruitment as a function of
+abiotic and biotic conditions. Most models maintained or even decreased
+recruitment diversity between the 7 and 10 cm thresholds, and the same
+is visible from the empirical data, but the differences between the 7
+and 10 cm thresholds were not significant. Diversity variations may be
+explained by the way the individual models consider regeneration
+processes [@konig_tree_2022], e.g., regeneration can be linked to the
+seed/seedling bank, seed rain from dispersal or from a pool of available
+species.
+
+The species diversity of the entire stand was captured better by the
+models than the diversity of tree recruitment, and only a few models
+overpredicted stand-level diversity. Defining recruitment composition as
+being proportional to local adult abundance, regardless of productivity,
+might be a simple and conservative assumption to maintain relative
+species abundances [@hanbury-brown2022], although this may be
+simplistic. Furthermore, based on the simulation results, there is no
+evidence that models with feedback from the canopy (in terms of species
+composition of regeneration) captured the species diversity of the
+recruitment better than those without feedback, with the exception of
+ForCEEPS that featured significantly more accurate recruitment species
+diversity with the model version that included feedback (ForCEEPS(f)).
+The similar performance of models with and without feedback is likely
+because the models put more weight on the regeneration niche arising
+from abiotic and biotic filters, than from the habitat niche of the
+adult trees [@grubb1977].
 
@@
 
-There are several factors that lead to mortality during the regeneration phase, such as competition [@casper1997] and multiple abiotic factors [@schmid2021; @cunningham2006].
-As mentioned above, it is reasonable to expect that models that overestimate recruitment at 7 cm may have a particularly high mortality between the 7 and 10 cm thresholds.
-However, this was not consistently evident from the simulations.
-This implies that these models must have an excess mortality in later stages, if we assume that all models were able to capture the structure and composition of the adult stands along the EuFoRIa gradient; this however was not tested here.
-
-Higher mortality towards the adult phase has important implications for forest dynamics and the goods and services provided by forests.
-On the one hand, mortality in later stages may erroneously enhance the share of less shade tolerant species in the models [@klopcic2015], therefore shifting the species composition.
-Unrealistic high stem densities for a longer period of time may overestimate the role of tree regrowth in carbon sink dynamics [@pugh2019].
-Given our set of simulations and analyses, we cannot conclusively assess what is happening in the models, and further studies are required that focus on a wider range of tree sizes and the fate of tree regeneration along such a size continuum.
+There are several factors that lead to mortality during the regeneration
+phase, such as competition [@casper1997] and multiple abiotic factors
+[@schmid2021; @cunningham2006]. As mentioned above, it is reasonable to
+expect that models that overestimate recruitment at 7 cm may have a
+particularly high mortality between the 7 and 10 cm thresholds. However,
+this was not consistently evident from the simulations. This implies
+that these models must have an excess mortality in later stages, if we
+assume that all models were able to capture the structure and
+composition of the adult stands along the EuFoRIa gradient; this however
+was not tested here.
+
+Higher mortality towards the adult phase has important implications for
+forest dynamics and the goods and services provided by forests. On the
+one hand, mortality in later stages may erroneously enhance the share of
+less shade tolerant species in the models [@klopcic2015], therefore
+shifting the species composition. Unrealistic high stem densities for a
+longer period of time may overestimate the role of tree regrowth in
+carbon sink dynamics [@pugh2019]. Given our set of simulations and
+analyses, we cannot conclusively assess what is happening in the models,
+and further studies are required that focus on a wider range of tree
+sizes and the fate of tree regeneration along such a size continuum.
 
@@
 
-In spite of the critical considerations above, it is remarkable that most models did not deviate exceedingly from the observations with respect to simulated recruited levels -- after all, few if any of the models' regeneration routines are well constrained by data, with the exception of the empirical models xComp and SIBYLA.
-It is noteworthy that in spite of their empirical basis, these two models did not match empirical recruitment levels, in a similar magnitude as the other models and even in opposite directions (SIBYLA: underestimation; xComp: overestimation).
-It appears that using empirical data of limited geographical scope that are constrained to managed forests, as done in these models, leads to extrapolation problems already under current climatic conditions.
-For the other models, whose regeneration modules are not strongly constrained by empirical data, the multiple strategies that are available for modeling regeneration processes [@konig_tree_2022] could have implied that model performance would be much worse than what we found.
-
-Our study showed that increasing complexity in the regeneration modules is not linked to a higher accuracy of the projections of recruitment levels, species composition or mortality at early tree stages, as there was no significant relationship between model performance and model complexity.
-Increasing complexity in regeneration modules has been called for a long time ago (e.g., @price2001) motivated by better process understanding and with the goal of enhancing model accuracy [@bugmann2022].
-However, more complex models do may not necessarily lead to better projections but rather to reduced transparency and lower predictive power [@franklin2020].
-Thus, the question regarding the level of detail that is appropriate and parsimonious for modeling tree regeneration remains open [@konig_tree_2022; @bugmann2022].
+In spite of the critical considerations above, it is remarkable that
+most models did not deviate exceedingly from the observations with
+respect to simulated recruited levels -- after all, few if any of the
+models' regeneration routines are well constrained by data, with the
+exception of the empirical models xComp and SIBYLA. It is noteworthy
+that in spite of their empirical basis, these two models did not match
+empirical recruitment levels, in a similar magnitude as the other models
+and even in opposite directions (SIBYLA: underestimation; xComp:
+overestimation). It appears that using empirical data of limited
+geographical scope that are constrained to managed forests, as done in
+these models, leads to extrapolation problems already under current
+climatic conditions. For the other models, whose regeneration modules
+are not strongly constrained by empirical data, the multiple strategies
+that are available for modeling regeneration processes
+[@konig_tree_2022] could have implied that model performance would be
+much worse than what we found.
+
+Our study showed that increasing complexity in the regeneration modules
+is not linked to a higher accuracy of the projections of recruitment
+levels, species composition or mortality at early tree stages, as there
+was no significant relationship between model performance and model
+complexity. Increasing complexity in regeneration modules has been
+called for a long time ago (e.g., @price2001) motivated by better
+process understanding and with the goal of enhancing model accuracy
+[@bugmann2022]. However, more complex models do may not necessarily lead
+to better projections but rather to reduced transparency and lower
+predictive power [@franklin2020]. Thus, the question regarding the level
+of detail that is appropriate and parsimonious for modeling tree
+regeneration remains open [@konig_tree_2022; @bugmann2022].
 
@@
 
-Competition for light as a strong filter for tree regeneration has been widely documented [@messier1999; @collet2006; @berdanier2016], but the models examined here did not reproduce this expectation (i.e., decreasing recruitment levels with increasing stand basal area).
-However, it is difficult to measure light availability at large spatial and temporal scales.
-We used total stand basal area as a proxy for light availability [cf.
-@schmid2021].
-Yet, we were unable to consider light availability restrictions caused by ground vegetation, which may be an important filter for forest dynamics at least in some EuFoRIa stands [@woltjer2008].
-We found pronounced differences in the ranges of stand basal area simulated by the models, but also between models and observations.
-This made it impossible to evaluate tree recruitment for the extremes of the stand density ranges in some models.
-This is unfortunate because recruitment levels at low stand density are relevant to assess how well forests are recovering e.g.
-after gap creation due to disturbance [@seidl2022; @grubb1977].
-At the level of the simulated one-hectare samples, average basal area is typically not very low as long as no larger disturbances occur, which was explicitly excluded in our simulation protocol.
-Towards the other end of the spectrum, i.e., with increasing stand basal area, it would be reasonable to expect that the recruitment of the different species would become sparser and drop out entirely under low-light conditions [@klopcic2012; @zell2019].
-However, few models showed this trend, thus indicating that the relation between regeneration and light availability is not yet captured correctly in most models.
-Yet, several of the models that did feature an increase of regeneration with increasing basal area include a feedback between seed production of mature trees and regeneration.
-Thus, it seems that in these models higher light competition does not sufficiently compensate for increased seed availability with higher basal area due to a higher abundance of mature trees.
-
-Lastly, there were pronounced differences how the main tree species were represented by the models along the environmental gradients, in particular the dominant species *Abies alba*, *Fagus sylvatica* and *Picea abies*.
-The recruitment levels were sampled from simulations in the equilibrium, and in this state it is expected that non-climax species such as *Pinus sylvestris* or *Quercus* would be of minor importance, or absent [@klopcic2015].
-Most models captured this low abundance, which is also found in the empirical data.
-Thus, although the broad patterns are matched by many models, improvements in the quantification of the regeneration niche of the species are needed, but this cannot be done in the absence of robust datasets across multiple species.
+Competition for light as a strong filter for tree regeneration has been
+widely documented [@messier1999; @collet2006; @berdanier2016], but the
+models examined here did not reproduce this expectation (i.e.,
+decreasing recruitment levels with increasing stand basal area).
+However, it is difficult to measure light availability at large spatial
+and temporal scales. We used total stand basal area as a proxy for light
+availability [cf. @schmid2021]. Yet, we were unable to consider light
+availability restrictions caused by ground vegetation, which may be an
+important filter for forest dynamics at least in some EuFoRIa stands
+[@woltjer2008]. We found pronounced differences in the ranges of stand
+basal area simulated by the models, but also between models and
+observations. This made it impossible to evaluate tree recruitment for
+the extremes of the stand density ranges in some models. This is
+unfortunate because recruitment levels at low stand density are relevant
+to assess how well forests are recovering e.g. after gap creation due to
+disturbance [@seidl2022; @grubb1977]. At the level of the simulated
+one-hectare samples, average basal area is typically not very low as
+long as no larger disturbances occur, which was explicitly excluded in
+our simulation protocol. Towards the other end of the spectrum, i.e.,
+with increasing stand basal area, it would be reasonable to expect that
+the recruitment of the different species would become sparser and drop
+out entirely under low-light conditions [@klopcic2012; @zell2019].
+However, few models showed this trend, thus indicating that the relation
+between regeneration and light availability is not yet captured
+correctly in most models. Yet, several of the models that did feature an
+increase of regeneration with increasing basal area include a feedback
+between seed production of mature trees and regeneration. Thus, it seems
+that in these models higher light competition does not sufficiently
+compensate for increased seed availability with higher basal area due to
+a higher abundance of mature trees.
+
+Lastly, there were pronounced differences how the main tree species were
+represented by the models along the environmental gradients, in
+particular the dominant species *Abies alba*, *Fagus sylvatica* and
+*Picea abies*. The recruitment levels were sampled from simulations in
+the equilibrium, and in this state it is expected that non-climax
+species such as *Pinus sylvestris* or *Quercus* would be of minor
+importance, or absent [@klopcic2015]. Most models captured this low
+abundance, which is also found in the empirical data. Thus, although the
+broad patterns are matched by many models, improvements in the
+quantification of the regeneration niche of the species are needed, but
+this cannot be done in the absence of robust datasets across multiple
+species.
 
@@
 
-The EuFoRIa data as used here are well suited to better understand tree regeneration.
-However, three aspects of these data may represent considerable limitations.
-First, we made a comparison of tree regeneration in an equilibrium state, but we cannot assess how close the forests included in the EuFoRIa dataset were to such a state.
-The data were collected in forest reserves where no management has taken place for long periods of time.
-This makes our assumption of an equilibrium between forest properties and environmental drivers more reasonable than it might appear at first sight.
-In an analysis of primeval forests at demographic equilibrium, @brzeziecki2021 found higher recruitment rates than the ones observed in the EuFoRia data.
-Thus, the overestimation of recruitment rates by the models may not be so problematic.
-Second, the data were collected for recruitment above a 7 cm threshold, thus limiting the assessment of tree regeneration to a specific point of stand dynamics.
-This constituted a hard limit based on which we can understand only some aspects of tree regeneration, which in its entirety often comprises a rather long period since seed production [@price2001].
-In reality, many environmental constraints are acting on young trees [@käber2021] that we were unable to assess.
-Yet, recruitment data with lower thresholds are simply not available in a harmonized manner across large environmental gradients.
-Third, the empirical data were collected from rather small plots, while we sampled simulated recruitment levels from 1 ha areas, which may lead to an incorrect representation of space.
-Even though the strategy we adopted is not ideal, it represents a common challenge when harmonizing diverse data sources originating from varied sampling strategies [@portier2022].
-It would have been extremely challenging for such a variety of models to follow a protocol where the spatial sampling size was different at each of the 200 sites, and it would have introduced additional uncertainty in the results.
-
-The design of our sampling protocol did not include spatial aspects such as seed dispersal or detailed soil data.
-While we considered a wide range of models of forest dynamics, from stand to global scales, the simulation setup was limited a spatial scale of 1 ha.
-This lack of consideration of spatial scale relationships is appropriate for stand-scale models, but it potentially puts the landscape models at a disadvantage, as they have been built to be accurate at the landscape level: without any spatial context, we are limiting tree regeneration to the seed influx from the stand itself, unless the model has a background seed input.
-Yet, the global models should not be at a disadvantage here due to their inherently limited spatial considerations [but cf.
-Snell et al.
--@snell2018; or Lehsten et al.
--@lehsten2019], as they are usually lacking dispersal between cells and are based on a strong abstraction of horizontal space [@hanbury-brown2022].
-Lastly, detailed data on soil conditions were not available from the observed data, and independent, admittedly coarse data for soil properties and the climatic water balance had to be used instead.
-It is noteworthy that many models represent drought using detailed indicators based e.g.
-on soil water holding capacity, which had to be derived from a rough soil quality measure.
-This may at least partially explain the unsatisfactory performance of many models along the drought axis (i.e., climatic water balance).
+The EuFoRIa data as used here are well suited to better understand tree
+regeneration. However, three aspects of these data may represent
+considerable limitations. First, we made a comparison of tree
+regeneration in an equilibrium state, but we cannot assess how close the
+forests included in the EuFoRIa dataset were to such a state. The data
+were collected in forest reserves where no management has taken place
+for long periods of time. This makes our assumption of an equilibrium
+between forest properties and environmental drivers more reasonable than
+it might appear at first sight. In an analysis of primeval forests at
+demographic equilibrium, @brzeziecki2021 found higher recruitment rates
+than the ones observed in the EuFoRia data. Thus, the overestimation of
+recruitment rates by the models may not be so problematic. Second, the
+data were collected for recruitment above a 7 cm threshold, thus
+limiting the assessment of tree regeneration to a specific point of
+stand dynamics. This constituted a hard limit based on which we can
+understand only some aspects of tree regeneration, which in its entirety
+often comprises a rather long period since seed production [@price2001].
+In reality, many environmental constraints are acting on young trees
+[@käber2021] that we were unable to assess. Yet, recruitment data with
+lower thresholds are simply not available in a harmonized manner across
+large environmental gradients. Third, the empirical data were collected
+from rather small plots, while we sampled simulated recruitment levels
+from 1 ha areas, which may lead to an incorrect representation of space.
+Even though the strategy we adopted is not ideal, it represents a common
+challenge when harmonizing diverse data sources originating from varied
+sampling strategies [@portier2022]. It would have been extremely
+challenging for such a variety of models to follow a protocol where the
+spatial sampling size was different at each of the 200 sites, and it
+would have introduced additional uncertainty in the results.
+
+The design of our sampling protocol did not include spatial aspects such
+as seed dispersal or detailed soil data. While we considered a wide
+range of models of forest dynamics, from stand to global scales, the
+simulation setup was limited a spatial scale of 1 ha. This lack of
+consideration of spatial scale relationships is appropriate for
+stand-scale models, but it potentially puts the landscape models at a
+disadvantage, as they have been built to be accurate at the landscape
+level: without any spatial context, we are limiting tree regeneration to
+the seed influx from the stand itself, unless the model has a background
+seed input. Yet, the global models should not be at a disadvantage here
+due to their inherently limited spatial considerations [but cf. Snell et
+al. -@snell2018; or Lehsten et al. -@lehsten2019], as they are usually
+lacking dispersal between cells and are based on a strong abstraction of
+horizontal space [@hanbury-brown2022]. Lastly, detailed data on soil
+conditions were not available from the observed data, and independent,
+admittedly coarse data for soil properties and the climatic water
+balance had to be used instead. It is noteworthy that many models
+represent drought using detailed indicators based e.g. on soil water
+holding capacity, which had to be derived from a rough soil quality
+measure. This may at least partially explain the unsatisfactory
+performance of many models along the drought axis (i.e., climatic water
+balance).
 
@@
 
-With our study, we have demonstrated that models of forest dynamics need a focus on their regeneration modules to make them more robust.
-It remains uncertain what level of detail is required to model tree regeneration, and this must be addressed in future research.
-We recommend that the improvement of the regeneration modules is implemented as additional features that can be traced back, as done here for the variants of ForClim and ForCEEPS, and that model complexity and structure must always be connected with modeling objectives [@albrich2020].
-If it should be necessary to include more detail in the regeneration models, this will come with higher parameterisation efforts.
-This will most likely lead to lower generalization because the required data will have to be collected from specific locations, as currently there is no general, comprehensive regeneration dataset available.
-
-Therefore, we further recommend that more effort should be invested into collecting harmonized datasets in a site-specific manner covering the different aspects leading to tree regeneration.
-We emphasize that datasets such as EuFoRIa are invaluable and should be expanded in both their spatial extent (e.g., towards boreal and Mediterranean conditions) as well as in time (e.g., continuing the monitoring into the future).
-Such data will allow for a better evaluation of forest models and help to reduce the uncertainty in their projections, which is crucial when they are used as tools for predicting e.g.
-the impacts of anthropogenic climate change.
-
-In the present study, we have considered tree regeneration in the equilibrium state only.
-It is equally important to understand how these models project tree regeneration after changes in forest structure by disturbances [@seidl2022], or under different management strategies [@lindner2000].
-However, this will require an entirely different set of observed data, and potentially not all forest models would be able to assess the relationship of these aspects on tree regeneration, e.g.
-due to the lack of disturbance or appropriate management modules.
-
-We recommend to investigate in detail the implications of the current modeling strategies for tree regeneration and, ultimately, simulated forest stand structure.
-This applies particularly to the erroneous patterns of excess tree regeneration and later excess mortality, by focusing on a wider range of tree sizes and the related regeneration dynamics.
-We also recommend, especially for the landscape-level models, the inclusion of explicit spatial considerations regarding tree regeneration [@beckage2003]; this, however, is a serious challenge regarding appropriate datasets, as large-scale inventory data have a wide coverage, but by definition do not allow for the assessment of spatial interactions.
-
-Exercises like the one presented here, where the models are operated in "blind flight" mode, i.e., without the possibility of being tuned towards capturing the expected patterns, should be repeated.
-Such benchmarking exercises should next focus on aspects such as specific model traits and the ecological formulations of particular (sub-)processes, to better understand the implications of the assumptions on which the models are based.
-Furthermore, the inclusion of a wide range of models with different (1) scales, (2) approaches for capturing population structure, (3) tree regeneration modules and (4) complexity of their formulations will ensure a large benefit to the entire modeling community and beyond.
+With our study, we have demonstrated that models of forest dynamics need
+a focus on their regeneration modules to make them more robust. It
+remains uncertain what level of detail is required to model tree
+regeneration, and this must be addressed in future research. We
+recommend that the improvement of the regeneration modules is
+implemented as additional features that can be traced back, as done here
+for the variants of ForClim and ForCEEPS, and that model complexity and
+structure must always be connected with modeling objectives
+[@albrich2020]. If it should be necessary to include more detail in the
+regeneration models, this will come with higher parameterisation
+efforts. This will most likely lead to lower generalization because the
+required data will have to be collected from specific locations, as
+currently there is no general, comprehensive regeneration dataset
+available.
+
+Therefore, we further recommend that more effort should be invested into
+collecting harmonized datasets in a site-specific manner covering the
+different aspects leading to tree regeneration. We emphasize that
+datasets such as EuFoRIa are invaluable and should be expanded in both
+their spatial extent (e.g., towards boreal and Mediterranean conditions)
+as well as in time (e.g., continuing the monitoring into the future).
+Such data will allow for a better evaluation of forest models and help
+to reduce the uncertainty in their projections, which is crucial when
+they are used as tools for predicting e.g. the impacts of anthropogenic
+climate change.
+
+In the present study, we have considered tree regeneration in the
+equilibrium state only. It is equally important to understand how these
+models project tree regeneration after changes in forest structure by
+disturbances [@seidl2022], or under different management strategies
+[@lindner2000]. However, this will require an entirely different set of
+observed data, and potentially not all forest models would be able to
+assess the relationship of these aspects on tree regeneration, e.g. due
+to the lack of disturbance or appropriate management modules.
+
+We recommend to investigate in detail the implications of the current
+modeling strategies for tree regeneration and, ultimately, simulated
+forest stand structure. This applies particularly to the erroneous
+patterns of excess tree regeneration and later excess mortality, by
+focusing on a wider range of tree sizes and the related regeneration
+dynamics. We also recommend, especially for the landscape-level models,
+the inclusion of explicit spatial considerations regarding tree
+regeneration [@beckage2003]; this, however, is a serious challenge
+regarding appropriate datasets, as large-scale inventory data have a
+wide coverage, but by definition do not allow for the assessment of
+spatial interactions.
+
+Exercises like the one presented here, where the models are operated in
+"blind flight" mode, i.e., without the possibility of being tuned
+towards capturing the expected patterns, should be repeated. Such
+benchmarking exercises should next focus on aspects such as specific
+model traits and the ecological formulations of particular
+(sub-)processes, to better understand the implications of the
+assumptions on which the models are based. Furthermore, the inclusion of
+a wide range of models with different (1) scales, (2) approaches for
+capturing population structure, (3) tree regeneration modules and (4)
+complexity of their formulations will ensure a large benefit to the
+entire modeling community and beyond.
 
@@
 
-Models of forest dynamics are important tools in science and decision support, and the formulation of tree regeneration has strong implications for simulated forest properties.
-The 15 models and variants used here are facing similar challenges in their representation of tree regeneration: they generally overestimate tree recruitment levels, and the simulated regeneration niche is not always captured accurately as a function of biotic (light) and abiotic (temperature and moisture) factors.
-
-However, most models properly capture the diversity of the initial tree community, and differences between model formulations, e.g., the presence or absence of feedback from the adult trees did not have a strong effect on capturing the species composition of regeneration.
-
-Regarding mortality in the early phase of tree life, many models that feature a particularly high overestimation of recruitment levels are compensating for this by a larger tree mortality.
-Often, this compensation is not sufficient to reduce the high recruitment levels to realistic values.
-Overall, there is no clear mortality pattern across all models.
-
-When capturing tree regeneration, the specific design decisions taken in the development of any model are more important for its behavior (accuracy) than scale (stand, landscape, global), modelling approach (empirical vs.
-process-based), and complexity.
-Having both empirical and process-based models in our set, the empirically-based models could have been expected to have a better performance, as they were calibrated with inventory data, but this was not the case.
-Similarly, higher model complexity does not represent an improvement for capturing tree regeneration.
-
-Even though the regeneration routines of most of the models investigated here have never been constrained well by robust data, their projections of recruitment are not overly off.
-This indicates that a lot can be gained by a focus on the modeling of regeneration processes.
-The representation of forest dynamics in these models would become much more robust particularly in the face of climate change and post-disturbance dynamics, thus strongly reducing the uncertainty in long-term projections of future forest dynamics.
+Models of forest dynamics are important tools in science and decision
+support, and the formulation of tree regeneration has strong
+implications for simulated forest properties. The 15 models and variants
+used here are facing similar challenges in their representation of tree
+regeneration: they generally overestimate tree recruitment levels, and
+the simulated regeneration niche is not always captured accurately as a
+function of biotic (light) and abiotic (temperature and moisture)
+factors.
+
+However, most models properly capture the diversity of the initial tree
+community, and differences between model formulations, e.g., the
+presence or absence of feedback from the adult trees did not have a
+strong effect on capturing the species composition of regeneration.
+
+Regarding mortality in the early phase of tree life, many models that
+feature a particularly high overestimation of recruitment levels are
+compensating for this by a larger tree mortality. Often, this
+compensation is not sufficient to reduce the high recruitment levels to
+realistic values. Overall, there is no clear mortality pattern across
+all models.
+
+When capturing tree regeneration, the specific design decisions taken in
+the development of any model are more important for its behavior
+(accuracy) than scale (stand, landscape, global), modelling approach
+(empirical vs. process-based), and complexity. Having both empirical and
+process-based models in our set, the empirically-based models could have
+been expected to have a better performance, as they were calibrated with
+inventory data, but this was not the case. Similarly, higher model
+complexity does not represent an improvement for capturing tree
+regeneration.
+
+Even though the regeneration routines of most of the models investigated
+here have never been constrained well by robust data, their projections
+of recruitment are not overly off. This indicates that a lot can be
+gained by a focus on the modeling of regeneration processes. The
+representation of forest dynamics in these models would become much more
+robust particularly in the face of climate change and post-disturbance
+dynamics, thus strongly reducing the uncertainty in long-term
+projections of future forest dynamics.
 
@@
 
-This paper arises from a workshop held in Davos (Switzerland) in June 2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
-Funding for the workshop was provided by COST Action CA19139 PROCLIAS (PROcess-based models for CLimate Impact Attribution across Sectors), supported by COST (European Cooperation in Science and Technology; [https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research Grant ETH-35 18-1).
-ODY was funded by the Swiss National Science Foundation project 200020_188882 / 1.
-KB received support from the European Research Council under the European Union's Horizon 2020 research and innovation program (Grant Agreement 101001905).
-JB acknowledges funding from the long-term research development project no.
-RVO 67985939 from the Czech Academy of Sciences.
-MM acknowledges financial support from I-Maestro (Innovative forest management strategies for a resilient bioeconomy under climate change and disturbances, grant no.
-22035418, 2019--2022) funded by the ERA-NET Cofund ForestValue.
-MMina acknowledges funding from the H2020-MSCA-IF project REINFORCE (Grant No.
-891671).
-KM received funding from the Project No.
-CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant no.
-ITMS2014+313011W580 supported by the Integrated Infrastructure Operational Programme funded by the ERDF.
-CH acknowledges financial support from Waldfonds WASIM2100 project.
-We would also like to thank all the people who helped to make this study possible with contributions at different stages particularly Martin Gutsch and Giorgio Vacchiano.
+This paper arises from a workshop held in Davos (Switzerland) in June
+2022, co-led by Olalla Díaz-Yáñez, Yannek Käber and Harald Bugmann.
+Funding for the workshop was provided by COST Action CA19139 PROCLIAS
+(PROcess-based models for CLimate Impact Attribution across Sectors),
+supported by COST (European Cooperation in Science and Technology;
+[https://www.cost.eu](https://www.cost.eu/)) and ETH Zurich (Research
+Grant ETH-35 18-1). ODY was funded by the Swiss National Science
+Foundation project 200020_188882 / 1. KB received support from the
+European Research Council under the European Union's Horizon 2020
+research and innovation program (Grant Agreement 101001905). JB
+acknowledges funding from the long-term research development project no.
+RVO 67985939 from the Czech Academy of Sciences. MM acknowledges
+financial support from I-Maestro (Innovative forest management
+strategies for a resilient bioeconomy under climate change and
+disturbances, grant no. 22035418, 2019--2022) funded by the ERA-NET
+Cofund ForestValue. MMina acknowledges funding from the H2020-MSCA-IF
+project REINFORCE (Grant No. 891671). KM received funding from the
+Project No. CZ.02.1.01/0.0/0.0/16_019/0000803 financed by OP RDE, grant
+no. ITMS2014+313011W580 supported by the Integrated Infrastructure
+Operational Programme funded by the ERDF. CH acknowledges financial
+support from Waldfonds WASIM2100 project. We would also like to thank
+all the people who helped to make this study possible with contributions
+at different stages particularly Martin Gutsch and Giorgio Vacchiano.
 
@@
 
-O.
-Díaz-Yáñez and Y.
-Käber contributed equally and share the first authorship.
-H.
-Bugmann, Y.
-Käber, and O.
-Díaz-Yáñez developed the study design.
-Y.
-Käber and O.
-Díaz-Yáñez prepared the input data used by all modeling teams.
-All authors contributed by conducting the simulations with their respective models.
-O.
-Díaz-Yáñez led the analysis of the simulation results with contributions during the workshop from Y.
-Käber, T.
-Anders, K.
-H.
-Braziunas, J.
-Brůna, S.
-M.
-Fischer, J.
-Hetzer, T.
-Hickler, H.
-Lischke, M.Mahnken, P.
-Mairota, K.
-Merganičová, T.
-Mette, X.
-Morin, W.
-Rammer, and D.
-Scherrer.
-O.
-Díaz-Yáñez and H.
-Bugmann led the writing of the manuscript and provided the first draft.
-All authors participated in the revision of the manuscript and approved its submission.
+O. Díaz-Yáñez and Y. Käber contributed equally and share the first
+authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study
+design. Y. Käber and O. Díaz-Yáñez prepared the input data used by all
+modeling teams. All authors contributed by conducting the simulations
+with their respective models. O. Díaz-Yáñez led the analysis of the
+simulation results with contributions during the workshop from Y. Käber,
+T. Anders, K. H. Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T.
+Hickler, H. Lischke, M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X.
+Morin, W. Rammer, and D. Scherrer. O. Díaz-Yáñez and H. Bugmann led the
+writing of the manuscript and provided the first draft. All authors
+participated in the revision of the manuscript and approved its
+submission.
 
@@
 
-**Table 1.** Models included in this study, their characteristics and simulation strategies used.
-Scale refers to the models scale, Type to the models type, Population structure to the model population approach, Feedback to the inclusion or not of adult tree feedback to the level and species composition of regeneration, Approach to the model regeneration module approach, Start from to the stage from which trees are recruited, Complexity, to the mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), Species to which species from the original eleven species were simulated, Runtime (spin up) [years] to the number of years used per model in their simulations for the spin up period and regeneration sampling; Climate data type (use of years) to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details.
-Further information can be found on the individual model reports available in the Appendix S1: Model reports.
+**Table 1.** Models included in this study, their characteristics and
+simulation strategies used. Scale refers to the models scale, Type to
+the models type, Population structure to the model population approach,
+Feedback to the inclusion or not of adult tree feedback to the level and
+species composition of regeneration, Approach to the model regeneration
+module approach, Start from to the stage from which trees are recruited,
+Complexity, to the mean regeneration formulation as estimated in
+(Bugmann & Seidl, 2022), Species to which species from the original
+eleven species were simulated, Runtime (spin up) [years] to the number
+of years used per model in their simulations for the spin up period and
+regeneration sampling; Climate data type (use of years) to the type of
+data used in the simulations and how it was used, and Reference provides
+the link to the publication with more model details. Further information
+can be found on the individual model reports available in the Appendix
+S1: Model reports.
 
@@
 
-**Figure 1.** The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed regeneration at the 10 cm threshold.
-
-**Figure 2.** Mean regeneration levels across all samples per site, plotted for the 200 sites and for each model.
-The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
-There are two boxplots for each model where the left lighter boxplot corresponds to 7 cm and the right darker boxplot to 10 cm.
-The boxplot shows the median (midline), interquartile range (box limits), largest and smallest value within 1.5 times interquartile range (whiskers), larger values are represented as outliers.
-
-**Figure 3.** Shannon index of tree regeneration, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively.
-The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.
-Boxplot features are the same as in Figure 2.
-
-**Figure 4.** Mean Shannon index across all samples per site for observed and simulated data.
-Each plot shows one pattern represented by one exemplary model of each category (overpredicted, intermediate and underpredicted).
-The full data with the grouping of the models are shown in Figures S3 and S4.
-n indicates the number of models falling in each group.
-A) Examples for the three trends across models for species diversity at the stand level.
-B) Examples for the three trends across models for regeneration at the 7 cm threshold (165 sites).
-
-**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm thresholds.
-Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree regeneration between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.Boxplot features are the same as in Figure 2.
-
-**Figure 6.** Ratio of tree regeneration between the 7 and 10 cm thresholds (regeneration 7 cm / regeneration 10 cm) and the overestimation proportion at 7 cm ([simulated - observed] / observed) for the mean regeneration per model across sites and samples.
-The horizontal dashed lines show a ratio equal to 1, indicating no decrease of regeneration between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions.
-The vertical line indicates zero overestimation.
-
-**Figure 7.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of total basal area.
-The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
-
-**Figure 8.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of climatic water balance.
-The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
-
-**Figure 9.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of seasonal degree-day sum.
-The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.
-
-**Figure 10.** Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold.
-The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient.
-The sizes of the circles represent the ratio between the regeneration basal area of the species and the total regeneration basal area for all species.
-The absence of a circle indicates a zero basal area share in the regeneration, or the absence of regeneration altogether.
-The color gradient (for the models only) shows the difference between the simulated and observed ratio of regeneration basal area of the species and the total regeneration basal area for all the species.
+**Figure 1.** The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used
+in the present study, the colors show the dominant tree species per site
+in the observed regeneration at the 10 cm threshold.
+
+**Figure 2.** Mean regeneration levels ($\overline{R}$) across all samples per
+site, plotted for the 200 sites and for each model. The red dashed lines
+shows the 25th and 75th percentiles for the 7 cm diameter threshold in
+the observed data. There are two boxplots for each model where the left
+lighter boxplot corresponds to 7 cm and the right darker boxplot to 10
+cm. The boxplot shows the median (midline), interquartile range (box
+limits), largest and smallest value within 1.5 times interquartile range
+(whiskers), larger values are represented as outliers.
+
+**Figure 3.** Mean Shannon index ($\overline{H}$) of tree regeneration, calculated by basal
+area, as the mean value across the 200 samples per site for the diameter
+thresholds of 7 and 10 cm, respectively. The red dashed lines shows the
+25th and 75th percentiles for the 7 cm diameter threshold in the
+observed data. Boxplot features are the same as in Figure 2.
+
+**Figure 4.** Mean Shannon index ($\overline{H}$) across all samples per site for
+observed and simulated data. Each plot shows one pattern represented by
+one exemplary model of each category (overpredicted, intermediate and
+underpredicted). The full data with the grouping of the models are shown
+in Figures S3 and S4. n indicates the number of models falling in each
+group. A) Examples for the three trends across models for species
+diversity at the stand level. B) Examples for the three trends across
+models for regeneration at the 7 cm threshold (165 sites).
+
+**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm
+thresholds. Dashed blue lines mark a ratio equal to 1, indicating no
+decrease in tree regeneration between 7 and 10 cm, and a ratio equal to
+1.77, corresponding to the Reineke self-thinning ratio under even-aged
+conditions. Boxplot features are the same as in Figure 2.
+
+**Figure 6.** Ratio of tree regeneration between the 7 and 10 cm
+thresholds (regeneration 7 cm / regeneration 10 cm) and the
+overestimation proportion at 7 cm ([simulated - observed] / observed)
+for the mean regeneration per model across sites and samples. The
+horizontal dashed lines show a ratio equal to 1, indicating no decrease
+of regeneration between 7 and 10 cm, and a ratio equal to 1.77
+corresponding to the Reineke self-thinning line under even-aged
+conditions. The vertical line indicates zero overestimation.
+
+**Figure 7.** Mean recruitment ($\overline{R}$) across the 200 samples per site with the
+y-axis scaled differently by model, for the 200 sites against gradients
+of total basal area. The values were split into ten bins; the red lines
+represent a GAM model showing the trend in the observed data.
+
+**Figure 8.** Mean recruitment ($\overline{R}$) across the 200 samples per site with the
+y-axis scaled differently by model, for the 200 sites against gradients
+of climatic water balance. The values were split into ten bins; the red
+lines represent a GAM model showing the trend in the observed data.
+
+**Figure 9.** Mean recruitment ($\overline{R}$) across the 200 samples per site with the
+y-axis scaled differently by model, for the 200 sites against gradients
+of seasonal degree-day sum. The values were split into ten bins; the red
+lines represent a GAM model showing the trend in the observed data.
+
+**Figure 10.** Mean recruitment levels ($\overline{R}$) of the five main species in terms of
+their basal area share in the observations (top row) and the performance
+of each model across the environmental gradients (other rows) for the 7
+cm regeneration threshold. The values shown are the mean of the 200
+samples per site and across the sites in each bin (tile), with ten bins
+per gradient. The sizes of the circles represent the ratio between the
+regeneration basal area of the species and the total regeneration basal
+area for all species. The absence of a circle indicates a zero basal
+area share in the regeneration, or the absence of regeneration
+altogether. The color gradient (for the models only) shows the
+difference between the simulated and observed ratio of regeneration
+basal area of the species and the total regeneration basal area for all
+the species.
 

### 2024-02-13 code/dataPreparation.R
add instructions to download data needed from the data portal

@@ -1,49 +1,23 @@
 ################################################################################
 # Title: Data preparation from model simulations and observations
 # Aim: Prepare and merge data from simulations and observations
+#      produces dataOutputs.csv file needed in code/figures.R
 # Authors: Olalla Díaz-Yáñez | @olalla | olalladiaz.net
 #          and Yannek Käber 
 ################################################################################ 
 
-# Load observations -----
+# Download the data 
+# IMPORTANT! this data is needed to run this script
 
-    linkObs <- "https://www.research-collection.ethz.ch/bitstream/handle/20.500.11850/643426/observations.zip"
-   
-    
-     download.file(url = linkObs, destfile = here::here("data/observations.zip"),  mode="wb")
-    downloader::download("https://www.research-collection.ethz.ch/bitstream/handle/20.500.11850/643426/observations.zip",
-                         destfile = here::here("data/observations.csv"))
-    
-    temp <- tempfile()
-    download.file(linkObs, "data/observations.zip", mode="wb")
-    data <- read.table(unz(temp, here::here("data/observations.csv")))
-    unlink(temp)
-   
-     download.file(myURL, dir, mode="wb")
-    observationsData <- data.table::fread("data/observations/observations.csv")
-    
-    
-    
-    
-    
-    url <- "https://www.research-collection.ethz.ch/bitstream/handle/20.500.11850/643426/observations.zip"
-    
-    # Local folder to save the downloaded zip file
-    local_folder <-  here::here("data")
-    
-    # File path for the downloaded zip file
-    zip_file <- file.path(local_folder, "observations.zip")
-    
-    # Download the zip file
-    download.file(url, destfile = zip_file, mode = "wb")# wget or curl
-    
-    # Unzip the file
-    unzip(zip_file, exdir = local_folder, list = TRUE)
+# this section requires to download the study data
+# the data is available in: 
+# https://doi.org/10.3929/ethz-b-000643426
+# via this link you can download the observations and simulation_results data
+# these two folders should be placed under the folder data 
+# data/observations/
+# data/simulations_results/
 
-    
-    
-    
-    
+
 
 # Load model simulations ----
     

### 2024-02-13 code/figures.R

@@ -2,9 +2,13 @@
 # Title: Figures
 # Aim: Create figures and analysis for the manuscript
 # Author: Olalla Díaz-Yáñez | @olalla | olalladiaz.net &
-#         some part of the figures code were based in the work of the 
+#         some parts of the figures code were based in the work of the 
 #         regeneration workshop attendees
 ################################################################################
+
+# Note:
+# code/dataPreparation.R script should be run before this script
+# to produce the file data/dataOutputs.csv needed for this script
 
 # Load plotting styles 
 

### 2024-02-13 manuscript/manuscriptRW.Rmd
correct the open research statement

@@
 - Harald Bugmann$^1$
-date: $^1$ETH Zurich. Department of Environmental Systems Science. Institute of Terrestrial Ecosystems. Universitätstrasse 16, 8092 Zürich, Switzerland. \newline
+date: $^1$ETH Zurich. Department of Environmental Systems Science. Institute of Terrestrial Ecosystems. Forest Ecology. Universitätstrasse 16, 8092 Zürich, Switzerland. \newline
       $^2$Senckenberg Biodiversity and Climate Research Centre (SBiK-F), 60325 Frankfurt am Main, Germany. \newline
@@
 
-**OPEN RESEARCH STATEMENT:** This submission uses novel code, both the
-data and code are provided in an external repository to be evaluated
-during the peer review process. The repository can be downloaded from
-[this figshare private link:
-https://figshare.com/s/100857ba67d2351dcb15](https://figshare.com/s/100857ba67d2351dcb15).
-The protocol is available at @díaz-yáñez2022 ([with link:
-https://osf.io/czdxp/](https://osf.io/czdxp/)). We will make publicly
-available the code and data if the manuscript is accepted.
+**OPEN RESEARCH STATEMENT:** This paper uses novel protocol, data and
+code, alle are available open access. The protocol and the protocol data
+is available at @díaz-yáñez2022. The data is available at
+@diaz-yanez2023. The repository containing the code used in the
+analysis, figure creation and manuscript writing can be downloaded from
+XXXXXXX.
 
@@
 
-**Figure 2.** Mean regeneration levels ($\overline{R}$) across all samples per
-site, plotted for the 200 sites and for each model. The red dashed lines
-shows the 25th and 75th percentiles for the 7 cm diameter threshold in
-the observed data. There are two boxplots for each model where the left
-lighter boxplot corresponds to 7 cm and the right darker boxplot to 10
-cm. The boxplot shows the median (midline), interquartile range (box
-limits), largest and smallest value within 1.5 times interquartile range
-(whiskers), larger values are represented as outliers.
-
-**Figure 3.** Mean Shannon index ($\overline{H}$) of tree regeneration, calculated by basal
-area, as the mean value across the 200 samples per site for the diameter
-thresholds of 7 and 10 cm, respectively. The red dashed lines shows the
-25th and 75th percentiles for the 7 cm diameter threshold in the
-observed data. Boxplot features are the same as in Figure 2.
-
-**Figure 4.** Mean Shannon index ($\overline{H}$) across all samples per site for
-observed and simulated data. Each plot shows one pattern represented by
-one exemplary model of each category (overpredicted, intermediate and
-underpredicted). The full data with the grouping of the models are shown
-in Figures S3 and S4. n indicates the number of models falling in each
-group. A) Examples for the three trends across models for species
-diversity at the stand level. B) Examples for the three trends across
-models for regeneration at the 7 cm threshold (165 sites).
+**Figure 2.** Mean regeneration levels ($\overline{R}$) across all
+samples per site, plotted for the 200 sites and for each model. The red
+dashed lines shows the 25th and 75th percentiles for the 7 cm diameter
+threshold in the observed data. There are two boxplots for each model
+where the left lighter boxplot corresponds to 7 cm and the right darker
+boxplot to 10 cm. The boxplot shows the median (midline), interquartile
+range (box limits), largest and smallest value within 1.5 times
+interquartile range (whiskers), larger values are represented as
+outliers.
+
+**Figure 3.** Mean Shannon index ($\overline{H}$) of tree regeneration,
+calculated by basal area, as the mean value across the 200 samples per
+site for the diameter thresholds of 7 and 10 cm, respectively. The red
+dashed lines shows the 25th and 75th percentiles for the 7 cm diameter
+threshold in the observed data. Boxplot features are the same as in
+Figure 2.
+
+**Figure 4.** Mean Shannon index ($\overline{H}$) across all samples per
+site for observed and simulated data. Each plot shows one pattern
+represented by one exemplary model of each category (overpredicted,
+intermediate and underpredicted). The full data with the grouping of the
+models are shown in Figures S3 and S4. n indicates the number of models
+falling in each group. A) Examples for the three trends across models
+for species diversity at the stand level. B) Examples for the three
+trends across models for regeneration at the 7 cm threshold (165 sites).
 
@@
 
-**Figure 7.** Mean recruitment ($\overline{R}$) across the 200 samples per site with the
-y-axis scaled differently by model, for the 200 sites against gradients
-of total basal area. The values were split into ten bins; the red lines
-represent a GAM model showing the trend in the observed data.
-
-**Figure 8.** Mean recruitment ($\overline{R}$) across the 200 samples per site with the
-y-axis scaled differently by model, for the 200 sites against gradients
-of climatic water balance. The values were split into ten bins; the red
-lines represent a GAM model showing the trend in the observed data.
-
-**Figure 9.** Mean recruitment ($\overline{R}$) across the 200 samples per site with the
-y-axis scaled differently by model, for the 200 sites against gradients
-of seasonal degree-day sum. The values were split into ten bins; the red
-lines represent a GAM model showing the trend in the observed data.
-
-**Figure 10.** Mean recruitment levels ($\overline{R}$) of the five main species in terms of
-their basal area share in the observations (top row) and the performance
-of each model across the environmental gradients (other rows) for the 7
-cm regeneration threshold. The values shown are the mean of the 200
-samples per site and across the sites in each bin (tile), with ten bins
-per gradient. The sizes of the circles represent the ratio between the
-regeneration basal area of the species and the total regeneration basal
-area for all species. The absence of a circle indicates a zero basal
-area share in the regeneration, or the absence of regeneration
-altogether. The color gradient (for the models only) shows the
-difference between the simulated and observed ratio of regeneration
+**Figure 7.** Mean recruitment ($\overline{R}$) across the 200 samples
+per site with the y-axis scaled differently by model, for the 200 sites
+against gradients of total basal area. The values were split into ten
+bins; the red lines represent a GAM model showing the trend in the
+observed data.
+
+**Figure 8.** Mean recruitment ($\overline{R}$) across the 200 samples
+per site with the y-axis scaled differently by model, for the 200 sites
+against gradients of climatic water balance. The values were split into
+ten bins; the red lines represent a GAM model showing the trend in the
+observed data.
+
+**Figure 9.** Mean recruitment ($\overline{R}$) across the 200 samples
+per site with the y-axis scaled differently by model, for the 200 sites
+against gradients of seasonal degree-day sum. The values were split into
+ten bins; the red lines represent a GAM model showing the trend in the
+observed data.
+
+**Figure 10.** Mean recruitment levels ($\overline{R}$) of the five main
+species in terms of their basal area share in the observations (top row)
+and the performance of each model across the environmental gradients
+(other rows) for the 7 cm regeneration threshold. The values shown are
+the mean of the 200 samples per site and across the sites in each bin
+(tile), with ten bins per gradient. The sizes of the circles represent
+the ratio between the regeneration basal area of the species and the
+total regeneration basal area for all species. The absence of a circle
+indicates a zero basal area share in the regeneration, or the absence of
+regeneration altogether. The color gradient (for the models only) shows
+the difference between the simulated and observed ratio of regeneration
 basal area of the species and the total regeneration basal area for all

### 2024-02-13 manuscript/supplementaryRW.Rmd
small correction in appendix doc

@@
 title: "Appendix S1"
-subtitle: "Tree regeneration in models of forest dynamics: a key priority for further research  ECOSPHERE"
+subtitle: "Tree regeneration in models of forest dynamics: a key priority for further research"
 author: 

### 2024-02-14 code/figures.R
correct legend figure 1

@@ -1632,14 +1632,13 @@
     #                      labels = c("Picea abies" = expression(italic("Picea abies")),
     #                                    "Fagus sylvatica" = expression(italic("Fagus sylvatica")),
     #                                    "Abies alba" =  expression(italic("Abies alba")),
-    #                                    "Quercus spp." = expression(italic("Quercus"), atop(" spp."),
-    #                                    "Other spp." = expression(atop("Other spp.")),
+    #                                    "Quercus spp." = expression(italic("Quercus spp.")),
+    #                                    "Other spp." = "Other spp.",
     #                                    "Pinus sylvestris" = expression(italic("Pinus sylvestris")),
     #                                    "Fraxinus excelsior" = expression(italic("Fraxinus excelsior")),
     #                                    "Alnus glutinosa" = expression(italic("Alnus glutinosa")),
-    #                                    "Tilia cordata" = expression(italic("Tilia cordata")))
-    # 
-    #                                 )) + theme(legend.text.align = 0)
+    #                                    "Tilia cordata" = expression(italic("Tilia cordata")))) +
+    #          theme(legend.text.align = 0)
     # 
     # ggplot2::ggsave("figures/Figure1.jpg",
     #                 plot =   plotsMap,

### 2024-02-15 code/dataPreparation.R
add data prep line

@@ -17,7 +17,9 @@
 # data/observations/
 # data/simulations_results/
 
+# Load observations -----
 
+    observationsData <- data.table::fread("data/observations/observations.csv")
 
 # Load model simulations ----
     

### 2024-02-15 manuscript/manuscriptRW.Rmd
add the reference to the published code

@@
 
-**OPEN RESEARCH STATEMENT:** This paper uses novel protocol, data and
-code, alle are available open access. The protocol and the protocol data
+**OPEN RESEARCH STATEMENT:** This paper uses a novel protocol, data and
+code, all are available open access. The protocol and the protocol data
 is available at @díaz-yáñez2022. The data is available at
@@
 analysis, figure creation and manuscript writing can be downloaded from
-XXXXXXX.
+@díaz-yáñez2024.
 
